### 2019-05-24 manuscript/manuscript draft.Rmd

---
title: "Untitled"
output: html_document
bibliography: library.bib
---

### Methods
1.	ForC database [@Anderson-Teixeira2018]
+ Used data compiled from existing literature
2.	Climate data
+	WorldClim [@Hijmans2005]; WorldClim 2: vapour pressure and solar radiation [@Fick2017]
+	CRU: cloud cover; annual frost days; potential evapotranspiration; annual wet days [@Harris2014]
+	Global Aridity Index and Potential Evapotranspiration Climate Database [@Trabucco2018]
+	TerraClimate: vapour pressure deficit [@Abatzoglou2018]
3.	Run models in R

### References

### 2019-05-28 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@
 + Used data compiled from existing literature
++ 
 2.	Climate data
@@
 3.	Run models in R
++ Model specification: mixed model using the lme4 package in R. 
++ Model = response_variable ~ fixed_variable + altitude + (1/geographic_area/plot_name), weighted by koeppen climate area
+
+#### Database  
+Analyses were conducted on data contained in the open-access ForC database [@Anderson-Teixeira2016; @Anderson-Teixeira2018].
+This database contains records of field-based measurements of forest carbon stocks and annual fluxes, compiled from previously published studies and existing databases (e.g.
+ORNL DAAC; Luyssaert etc?).
+*Data selection.* ForC contains measures of carbon stocks and fluxes; for the purposes of this analysis only measures of primary productivity were selected from the database.
+The variables selected were: GPP, NPP, BNPP (total root and fine root), ANPP (total ANPP, foliage, wood and woody stem).
+Sites were excluded from analysis if they were managed, defined as plots that were planted, managed as plantations, irrigated, fertilised or including the term "managed" in their site description.
+Sites that had experienced significant disturbance were also excluded.
+Disturbances that justified site exclusion were major cutting or harvesting, and/or burning, flooding, drought and storm events with site mortality >10% of trees.
+There is evidence that stand age influences patterns of primary productivity and carbon allocation in forest ecosystems [@DeLucia2007; @Gillman2015].
+To reduce any biasing effects of stand age, stands under 100 years of age were excluded from analysis.
 

### 2019-05-28 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@
 
+*Methodological consistency* The data in ForC is derived from a range of studies, often employing different methods.
+For this reason, criteria were introduced to standardise for differences in methodology.
+Where data was based on forest plot census measurements, studies which uses a minimum dbh measure of 10cm or greater were excluded from analysis.
+Estimates of total ecosystem primary productivity are based on measurements of the component parts of forest productivity.
+Since the components included in productivity estimates vary between studies, estimates of productivity were classified within the ForC database according to their components, and then filtered for analysis.
+Estimates of NPP were selected if they included foliage, branch, stem, coarse root, fine root and understory.
+Measures of NPP which included additional components were excluded.
+Estimates of ANPP were selected if they included foliage, stem growth and branch turnover.
+Any measures of primary productivity where components were unknown were excluded from analysis.
+
 ### References

### 2019-05-28 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@
 ORNL DAAC; Luyssaert etc?).
+
 *Data selection.* ForC contains measures of carbon stocks and fluxes; for the purposes of this analysis only measures of primary productivity were selected from the database.
@@
 
-*Methodological consistency* The data in ForC is derived from a range of studies, often employing different methods.
+*Methodological consistency.* The data in ForC is derived from a range of studies, often employing different methods.
 For this reason, criteria were introduced to standardise for differences in methodology.
@@
 
+*Climate datasets.* Climate data was extracted from four open-access climate datasets.
+
+*Model specification.*
+
 ### References

### 2019-05-28 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@
 ORNL DAAC; Luyssaert etc?).
+Additional targeted literature searches were conducted to identify any further available data on primary productivity, with particular focus on old-growth forests in temperate and boreal regions.
+
+**Fig. 1** - Map showing all data used in the anlysis, coded by variable.
+![Distribution of carbon data](https://github.com/forc-db/Global_Productivity/blob/master/results/figures/final_figures/maps/distribution_all_variables.png)
 
@@
 The variables selected were: GPP, NPP, BNPP (total root and fine root), ANPP (total ANPP, foliage, wood and woody stem).
+
+*[insert table here similar to in trop-forC paper, including information on productivity measures, methods used to estimate, components included etc]*
+
 Sites were excluded from analysis if they were managed, defined as plots that were planted, managed as plantations, irrigated, fertilised or including the term "managed" in their site description.
@@
 
-*Climate datasets.* Climate data was extracted from four open-access climate datasets.
+*Climate datasets.* Where site-level data on mean annual temperature, mean annual precipitation, and latitude were available in the primary literature, this data was compiled and entered directly into the ForC database.
+In addition to this data, climate data for each site was extracted from five open-access climate datasets.
 
-*Model specification.*
+\begin{table}
+\centering
+\caption{}
+\label{table-paramvalues}
+\begin{tabular}{ p{4cm} p{4cm} p{4cm} }
+\hline \\ [-1.5ex]
+Database &amp; Variables downloaded &amp; Citation \\ [1ex]
+\hline \\ [-1.5ex]
+WorldClim & mean annual temperature; temperature seasonality; annual temperature range; mean annual precipitation & [@Hijmans2005] \\ [1ex]
+WorldClim2 & Vapour pressure; solar radiation & [@Fick2017] \\ [1ex]
+Climate Research Unit (CRU) time-series dataset v 4.03 & Cloud cover; annual frost days; annual wet days; potential evapotranspiration & [@Harris2014] \\ [1ex]
+Global Aridity Index and Potential Evapotranspiration Climate Database & Aridity; potential evapotranspiration & [@Trabucco2018] \\ [1ex]
+TerraClimate & Vapour pressure deficit & [@Abatzoglou2018] \\ [1ex]
+\hline
+\end{tabular}
+\end{table}
+
+| Database | Variables downloaded | Citation |
+|---------------|-------------------------------------------------|-------------------|
+| WorldClim | Mean annual temperature; temperature seasonality; annual temperature range; mean annual precipitation | [@Hijmans2005] |
+| WorldClim2 | Vapour pressure; solar radiation | [@Fick2017] |
+| Climate Research Unit (CRU) time-series dataset v 4.03 | Cloud cover; annual frost days; annual wet days; potential evapotranspiration | [@Harris2014] |
+| Global Aridity Index and Potential Evapotranspiration Climate Database | Aridity; potential evapotranspiration | [@Trabucco2018] |
+| TerraClimate | Vapour pressure deficit | [@Abatzoglou2018] |
+
+*Model specification.* The effects of climate and latitude on primary productivity were analysed using mixed effects models using the package 'lme4' in R v. 1.1.463 (cite).
 

### 2019-05-28 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@
 
-**Fig. 1** - Map showing all data used in the anlysis, coded by variable.
-![Distribution of carbon data](https://github.com/forc-db/Global_Productivity/blob/master/results/figures/final_figures/maps/distribution_all_variables.png)
+---
+output:
+  pdf_document:
+    fig_caption: yes
+---
+
+![Figure 1: Map showing all data used in the anlysis, coded by variable \label{figurelabel}](distribution_all_variables.png)
 
@@
 
-\begin{table}
-\centering
-\caption{}
-\label{table-paramvalues}
-\begin{tabular}{ p{4cm} p{4cm} p{4cm} }
-\hline \\ [-1.5ex]
-Database &amp; Variables downloaded &amp; Citation \\ [1ex]
-\hline \\ [-1.5ex]
-WorldClim & mean annual temperature; temperature seasonality; annual temperature range; mean annual precipitation & [@Hijmans2005] \\ [1ex]
-WorldClim2 & Vapour pressure; solar radiation & [@Fick2017] \\ [1ex]
-Climate Research Unit (CRU) time-series dataset v 4.03 & Cloud cover; annual frost days; annual wet days; potential evapotranspiration & [@Harris2014] \\ [1ex]
-Global Aridity Index and Potential Evapotranspiration Climate Database & Aridity; potential evapotranspiration & [@Trabucco2018] \\ [1ex]
-TerraClimate & Vapour pressure deficit & [@Abatzoglou2018] \\ [1ex]
-\hline
-\end{tabular}
-\end{table}
-
 | Database | Variables downloaded | Citation |

### 2019-05-28 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
 
-*Model specification.* The effects of climate and latitude on primary productivity were analysed using mixed effects models using the package 'lme4' in R v. 1.1.463 (cite).
+*Model specification.* The effects of climate and latitude on primary productivity were analysed using mixed effects models using the package 'lme4' in R v.
+1.1.463 (cite).
+The effect of each extracted climate variable on each measure of primary productivity was modelled by specifying the climate variable as a fixed effect.
+Site altitude was also included as a fixed effect.
+Random effect was stand nested within area.
+Data from the temperate regions was heavily skewed towards studies from the old-growth forests of the Pacific Northwest.
+These forests have very high productivity, and so to reduce any bias from over-sampling of this region, the models were weighted according to the proportion of forest cover by Koeppen climate zone.
 

### 2019-05-29 manuscript/manuscript draft.Rmd
manuscript draft

@@
 ---
-title: "Untitled"
+title: "Manuscript draft"
 output: html_document
@@
 
+### Introduction
+
 ### Methods
-1.	ForC database [@Anderson-Teixeira2018]
-+ Used data compiled from existing literature
-+ 
-2.	Climate data
-+	WorldClim [@Hijmans2005]; WorldClim 2: vapour pressure and solar radiation [@Fick2017]
-+	CRU: cloud cover; annual frost days; potential evapotranspiration; annual wet days [@Harris2014]
-+	Global Aridity Index and Potential Evapotranspiration Climate Database [@Trabucco2018]
-+	TerraClimate: vapour pressure deficit [@Abatzoglou2018]
-3.	Run models in R
-+ Model specification: mixed model using the lme4 package in R. 
-+ Model = response_variable ~ fixed_variable + altitude + (1/geographic_area/plot_name), weighted by koeppen climate area
 
-#### Database  
 Analyses were conducted on data contained in the open-access ForC database [@Anderson-Teixeira2016; @Anderson-Teixeira2018].
@@
 These forests have very high productivity, and so to reduce any bias from over-sampling of this region, the models were weighted according to the proportion of forest cover by Koeppen climate zone.
+ranch turnover. Any measures of primary productivity where components were unknown were excluded from analysis.
 

### 2019-05-30 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
 ### Introduction
+
+Globally, forests play an important role in the carbon cycle,
+
+Forests are estimated to sequester [x] carbon annually, making them a significant carbon sink, and an important component of global carbon dioxide budgets.
+
+Understanding global patterns of net primary productivity in forest ecosystems is essential for understanding the global carbon cycle.
+
+The relationship between primary productivity and latitude is unclear.
+
+*Understanding components of primary productivity.* Gross primary productivity (GPP) represents the gross uptake of carbon dioxide by plants in an ecosystem, to be used in photosynthesis and energy production.
+Only a fraction of this carbon is assimilated into plant tissues; the rest is used in autotrophic respiration.
+The component of GPP that is stored as plant material is the net primary productivity (NPP) of an ecosystem; thus GPP can be described as the sum of NPP and ecosystem respiration.
+
+Currently, GPP cannot be measured directly by observing total ecosystem photosynthesis.
+Instead field estimates of GPP have to be derived based on modelling and extrapolation of eddy-covariance studies and measurements of net ecosystem exchange (NEE) [@Clark2017].
+
+NPP is typically measured through direct measurements of the productivity of its components, which are then summed to generate a whole-ecosystem estimate.
+These components can be broadly divided into aboveground NPP (ANPP) and belowground NPP (BNPP).
+Measurement of aboveground NPP - primarily consisting of woody stem, woody branch and canopy/foliage primary productivity, but also sometimes including estimates of productivity of reproductive and understory components, and losses due to herbivory - is most commonly estimated through forest plot censuses.
+Belowground NPP is significantly more challenging to measure, meaning that estimates of belowground productivity are less reliable than those of aboveground productivity.
+Coarse root productivity is typically determined by using allometric equations relating root mass to stem diameter.
+This relies on extrapolation of ANPP measures, rather than being based on direct measures of root productivity.
+Fine root productivity is estimated using a variety of methods, such as taking soil cores and minirhizotrons, making it typically more reliable than measures of coarse root productivity.
+
+*GPP* Gross primary productivity is generally thought to be highest in the tropics.
+Modelling of global terrestrial ecosystem GPP though upscaling and calibration of eddy flux measurements indicates a peak in GPP in the tropics, with the highest levels in tropical forests [@Beer2010; @Jung2011].
+This is corroborated by analysis of site-level GPP measurements, which show a strong positive correlation between GPP and mean annual temperature (MAT) and mean annual precipitation (MAP) [@Luyssaert2007], with the highest GPP values reported in tropical forests.
+There is evidence that this increase is not linear: increases in GPP have been reported to saturate above 1500mm MAP [@Luyssaert2007] and 25$^\circ$C [@Larjavaara2012].
 

### 2019-05-30 manuscript/manuscript draft.Rmd
update manuscript draft

@@
 This is corroborated by analysis of site-level GPP measurements, which show a strong positive correlation between GPP and mean annual temperature (MAT) and mean annual precipitation (MAP) [@Luyssaert2007], with the highest GPP values reported in tropical forests.
-There is evidence that this increase is not linear: increases in GPP have been reported to saturate above 1500mm MAP [@Luyssaert2007] and 25$^\circ$C [@Larjavaara2012].
+There is evidence that this increase is not linear: increases in GPP have been reported to saturate above 1500mm MAP [@Luyssaert2007] and 25$^\circ$C MAT [@Larjavaara2012].
+
+*NPP* The influence of latitude on global patterns of NPP is less clear than that of GPP.
+Simova and Storch (2017) found that, as with GPP, NPP decreases with latitude, peaking in the tropics.
+However, other studies have found the highest values of NPP in temperate forests, with increases in NPP saturating at 1500mm MAP and 10$^\circ$C MAT [@Luyssaert2007].
+
+Primary productivity of forests is influenced by a range of factors.
+On a local scale, stand age [@Litton2007; @Gillman2015], management [@Simova2017]; nutrient availability [cite]; and altitude [@Girardin2010; @Malhi2017] all impact stand productivity.
+On a global scale, changes are likely to be influenced by climatic variables.
+The majority of studies to date have focused on the influence of MAP and MAT on productivity.
+While MAP and MAT have been argued to be significant predictors of productivity [@Chu2016], other studies have found that the correlation between productivity and MAT is a factor of the relationship between productivity and growing season length [@Kerkhoff2005; @Malhi2012; @Michaletz2014; @Michaletz2018].
 

### 2019-05-30 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
 Simova and Storch (2017) found that, as with GPP, NPP decreases with latitude, peaking in the tropics.
-However, other studies have found the highest values of NPP in temperate forests, with increases in NPP saturating at 1500mm MAP and 10$^\circ$C MAT [@Luyssaert2007].
+However, other studies have found the highest values of NPP in temperate forests [@Huston2009], with increases in NPP saturating at 1500mm MAP and 10$^\circ$C MAT [@Luyssaert2007].
+
+Because of the challenges in accurately obtaining unbiased measures of belowground productivity, many studies focus on ANPP in preference to measures of NPP.
+Studies on global patterns of ANPP are equally inconclusive: Gillman et al.
+(2015) found a weak negative relationship between ANPP and latitude, with the relationship becoming stronger in older forest stands.
+These findings were echoed in other studies, which have found weak or no relationships between ANPP and latitude [@Huston2009].
 
@@
 
-*[insert table here similar to in trop-forC paper, including information on productivity measures, methods used to estimate, components included etc]*
+| Variable | Definition | Components included | Methodologies used|
+|-------------|-------------------|-------------|------------------------|
+| GPP | Annual gross primary production; annual uptake of carbon dioxide by an ecosystem | NA | Flux partitioning of eddy covariance |
+| NPP | Annual net primary production; the component of GPP that is stored in plant tissue; GPP minus ecosystem respiration | Foliage, branch, stem, coarse root, fine root, and optionally understory | Direct measurement of annual increments of components |
+| ANPP | Aboveground net primary production | Foliage, stem, and optionally branch | Direct measurement of annual increments of components | 
+| ANPP_foliage | Net primary production of foliage | Foliage | Direct measurement of litterfall, correcting for changes in leaf biomass when measured |
+| ANPP_woody | Net primary production of woody components | Stems and branches | Direct measurement of stem growth and branch fall |
+| ANPP_woody_stem | Net primary production of woody stems | Woody stems | Direct measurement of stem growth increment |
+| BNPP_root | Belowground net primary production | Coarse and fine roots | Direct measurement of one or more of: fine root turnover, soil cores, root ingrowth cores, minirhizotrons; indirect estimates of coarse roots using allometries based on aboveground stem increment measures |
+| BNPP_root_fine | Net primary production of fine roots | Fine roots | Direct measurement of one or more of: minirhizotrons, fine root turnover, soil cores, root ingrowth cores |
 
@@
 Disturbances that justified site exclusion were major cutting or harvesting, and/or burning, flooding, drought and storm events with site mortality >10% of trees.
-There is evidence that stand age influences patterns of primary productivity and carbon allocation in forest ecosystems [@DeLucia2007; @Gillman2015].
+There is evidence that stand age influences patterns of primary productivity and carbon allocation in forest ecosystems, and can confound relationships between latitude and primary productivity [@DeLucia2007; @Gillman2015].
 To reduce any biasing effects of stand age, stands under 100 years of age were excluded from analysis.

### 2019-05-31 manuscript/manuscript draft.Rmd
updated manuscript

@@
 
-Globally, forests play an important role in the carbon cycle,
+Globally, forests play an important role in the carbon cycle.
 
 Forests are estimated to sequester [x] carbon annually, making them a significant carbon sink, and an important component of global carbon dioxide budgets.
+
+The productivity of forests shows variation with latitude, however the exact nature of this pattern and the mechanisms that drive it are poorly understood.
+On a global scale variation in primary productivity is likely to be climate-driven, meaning that achieving a better understanding of the climatic controls on forest productivity is essential for understanding and predicting the reponses and feedbacks of forests to climate change.
 
@@
 
-*Understanding components of primary productivity.* Gross primary productivity (GPP) represents the gross uptake of carbon dioxide by plants in an ecosystem, to be used in photosynthesis and energy production.
+In order to better understand and predict responses and feedbacks to climate change, we need a better fundamental understand of how forest carbon cycling varies with respect to climate globally.
+
+*Patterns of primary productivity.* Understanding variation in primary productivity is complex.
+Firstly, there are many factors which can influence productivity on a range of scales, many of which show interactive effects with each other.
+In addition, the primary productivity of a forest is the sum of the primary productivity of a number of individual components, which may all show individual responses to drivers of productivity.
+Understanding these relationships and interdependencies is complex.
+
+Primary productivity of forests is influenced by a range of factors.
+On a local scale, stand age [@Litton2007; @Gillman2015], management [@Simova2017]; nutrient availability [cite]; and altitude [@Girardin2010; @Malhi2017] all impact stand productivity.
+On a global scale, changes in primary productivity are influenced by climatic variables and abiotic gradients, such as the length of growing season [@Michaletz2014].
+The majority of studies to date have focused on the influence of MAP and MAT on productivity.
+While MAP and MAT have been argued to be significant predictors of productivity [@Chu2016], other studies have found that the correlation between productivity and MAT is a factor of the relationship between productivity and growing season length [@Kerkhoff2005; @Malhi2012; @Michaletz2014; @Michaletz2018].
+The influence of other climate variables on global patterns of productivity remains under-explored.
+
+*Understanding components of primary productivity.* There are several different measures of primary productivity.
+Gross primary productivity (GPP) represents the gross uptake of carbon dioxide by plants in an ecosystem, to be used in photosynthesis and energy production.
 Only a fraction of this carbon is assimilated into plant tissues; the rest is used in autotrophic respiration.
@@
 
-*GPP* Gross primary productivity is generally thought to be highest in the tropics.
+Current knowledge indicates that these different components of productivity show different latitudinal patterns.
+Gross primary productivity is generally thought to be highest in the tropics.
 Modelling of global terrestrial ecosystem GPP though upscaling and calibration of eddy flux measurements indicates a peak in GPP in the tropics, with the highest levels in tropical forests [@Beer2010; @Jung2011].
 This is corroborated by analysis of site-level GPP measurements, which show a strong positive correlation between GPP and mean annual temperature (MAT) and mean annual precipitation (MAP) [@Luyssaert2007], with the highest GPP values reported in tropical forests.
-There is evidence that this increase is not linear: increases in GPP have been reported to saturate above 1500mm MAP [@Luyssaert2007] and 25$^\circ$C MAT [@Larjavaara2012].
-
-*NPP* The influence of latitude on global patterns of NPP is less clear than that of GPP.
+The influence of latitude on global patterns of NPP is less clear than that of GPP.
 Simova and Storch (2017) found that, as with GPP, NPP decreases with latitude, peaking in the tropics.
-However, other studies have found the highest values of NPP in temperate forests [@Huston2009], with increases in NPP saturating at 1500mm MAP and 10$^\circ$C MAT [@Luyssaert2007].
-
+However, other studies have found the highest values of NPP in temperate forests [@Luyssaert2007; @Huston2009].
 Because of the challenges in accurately obtaining unbiased measures of belowground productivity, many studies focus on ANPP in preference to measures of NPP.
@@
 
-Primary productivity of forests is influenced by a range of factors.
-On a local scale, stand age [@Litton2007; @Gillman2015], management [@Simova2017]; nutrient availability [cite]; and altitude [@Girardin2010; @Malhi2017] all impact stand productivity.
-On a global scale, changes are likely to be influenced by climatic variables.
-The majority of studies to date have focused on the influence of MAP and MAT on productivity.
-While MAP and MAT have been argued to be significant predictors of productivity [@Chu2016], other studies have found that the correlation between productivity and MAT is a factor of the relationship between productivity and growing season length [@Kerkhoff2005; @Malhi2012; @Michaletz2014; @Michaletz2018].
+These variations in the responses of different components of productivity to changes in latitude suggest that components may vary in the strength of their response to climatic drivers.
+There is some empirical evidence for this: increases in GPP are reported to saturate above 25$^\circ$C MAT [@Larjavaara2012], while increases in NPP are recorded to saturate above 10$^\circ$C MAT [@Luyssaert2007].
+Similarly, allocation to different components of ANPP varies with climate.
+Within the tropics, allocation to canopy NPP appears fairly consistent, with significantly greater variation in allocation to woody and belowground NPP [@Litton2007; @Malhi2011].
+Allocation to these structural biomass components has been shown to increase with water availability [@Litton2007; @Bloom2016], and is highest in the wet tropics, indicating that control of woody productivity by MAP may be more significant than control of foliar productivity.
+However, these studies are regional, meaning that our understanding of variation in allocation and its relationship to climate on a global scale remains limited.
+
+Allocation is also influenced by stand age [@DeLucia2007], nutrient availability [@Litton2007] and forest structure [@Taylor2019], which can make it challenging to disentangle the effects of climate.
 

### 2019-05-31 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
 Globally, forests play an important role in the carbon cycle.
+Old growth forests alone are estimated to sequester up to 1.4 GtCyr superscript^-1^ [@Malhi1999], making them a significant carbon sink, and am important component of global carbon dioxide budgets [@Luyssaert2008].
+Understanding the patterns and mechanisms of forest carbon cycling on a global scale is therefore of high importance in understanding the potential role of forests in mitigating climate change.
 
-Forests are estimated to sequester [x] carbon annually, making them a significant carbon sink, and an important component of global carbon dioxide budgets.
-
-The productivity of forests shows variation with latitude, however the exact nature of this pattern and the mechanisms that drive it are poorly understood.
-On a global scale variation in primary productivity is likely to be climate-driven, meaning that achieving a better understanding of the climatic controls on forest productivity is essential for understanding and predicting the reponses and feedbacks of forests to climate change.
-
-Understanding global patterns of net primary productivity in forest ecosystems is essential for understanding the global carbon cycle.
-
-The relationship between primary productivity and latitude is unclear.
+It is understood that the productivity of forests shows variation with latitude, with a general trend of increasing productivity towards the tropics, however the exact nature of this pattern and the mechanisms that drive it are poorly understood.
+Climate can be an important driver of productivity, particularly on a global scale; achieving a better understanding of the climatic controls on forest productivity will help to illuminate potential responses and feedbacks of forests to climate change.
 
@@
 Allocation is also influenced by stand age [@DeLucia2007], nutrient availability [@Litton2007] and forest structure [@Taylor2019], which can make it challenging to disentangle the effects of climate.
+
+Data that control for stand age and standardize methodologies are required to resolve this question. Here, we use a comprehensive global database to explore global patterns in productivity.
 

### 2019-06-03 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
 title: "Manuscript draft"
-output: html_document
+output:
+  pdf_document:
+    citation_package: natbib
 bibliography: library.bib
+csl: apa.csl
 ---
@@
 Globally, forests play an important role in the carbon cycle.
-Old growth forests alone are estimated to sequester up to 1.4 GtCyr superscript^-1^ [@Malhi1999], making them a significant carbon sink, and am important component of global carbon dioxide budgets [@Luyssaert2008].
+Old growth forests alone are estimated to sequester up to 1.4 GtCyr ^-1^ [@Malhi1999], making them a significant carbon sink, and an important component of global carbon dioxide budgets [@Luyssaert2008].
 Understanding the patterns and mechanisms of forest carbon cycling on a global scale is therefore of high importance in understanding the potential role of forests in mitigating climate change.
 
-It is understood that the productivity of forests shows variation with latitude, with a general trend of increasing productivity towards the tropics, however the exact nature of this pattern and the mechanisms that drive it are poorly understood.
+It is understood that the productivity of forests shows variation with latitude, with a general trend of increasing productivity towards the tropics; however the exact nature of this pattern and the mechanisms that drive it are poorly understood.
 Climate can be an important driver of productivity, particularly on a global scale; achieving a better understanding of the climatic controls on forest productivity will help to illuminate potential responses and feedbacks of forests to climate change.
@@
 Studies on global patterns of ANPP are equally inconclusive: Gillman et al.
-(2015) found a weak negative relationship between ANPP and latitude, with the relationship becoming stronger in older forest stands.
+[-@Gillman2015] found a weak negative relationship between ANPP and latitude, with the relationship becoming stronger in older forest stands.
 These findings were echoed in other studies, which have found weak or no relationships between ANPP and latitude [@Huston2009].

### 2019-06-03 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
 
-It is understood that the productivity of forests shows variation with latitude, with a general trend of increasing productivity towards the tropics; however the exact nature of this pattern and the mechanisms that drive it are poorly understood.
+The productivity of forests shows variation with latitude, with a general trend of increasing productivity towards the tropics; however the exact nature of this pattern and the mechanisms that drive it are poorly understood.
 Climate can be an important driver of productivity, particularly on a global scale; achieving a better understanding of the climatic controls on forest productivity will help to illuminate potential responses and feedbacks of forests to climate change.

### 2019-06-03 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
 
+Little is known about the main environmental drivers of these patterns.
+
 The productivity of forests shows variation with latitude, with a general trend of increasing productivity towards the tropics; however the exact nature of this pattern and the mechanisms that drive it are poorly understood.
@@
 Primary productivity of forests is influenced by a range of factors.
-On a local scale, stand age [@Litton2007; @Gillman2015], management [@Simova2017]; nutrient availability [cite]; and altitude [@Girardin2010; @Malhi2017] all impact stand productivity.
+On a local scale, stand age [@Litton2007; @Gillman2015], management [@Simova2017]; nutrient availability [@Aragao2009]; and altitude [@Girardin2010; @Malhi2017] all impact stand productivity.
 On a global scale, changes in primary productivity are influenced by climatic variables and abiotic gradients, such as the length of growing season [@Michaletz2014].

### 2019-06-04 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd
updated manuscript draft

@@
 
+add hypotheses, can number/bullet point them
+redo first couple paragraphs
+look again at breakdown of primary productivity components --> add discussion on understanding belowground productivity
+
 ### Introduction
 
-Globally, forests play an important role in the carbon cycle.
-Old growth forests alone are estimated to sequester up to 1.4 GtCyr ^-1^ [@Malhi1999], making them a significant carbon sink, and an important component of global carbon dioxide budgets [@Luyssaert2008].
-Understanding the patterns and mechanisms of forest carbon cycling on a global scale is therefore of high importance in understanding the potential role of forests in mitigating climate change.
+Globally, forests play an important role in the carbon cycle and are an important component of global carbon dioxide budgets [@Luyssaert2008].
+They show higher levels of productivity than non-forest  terrestrial ecosystems [@DelGrosso2008], and as a result achieve significant carbon sequestration and storage.
+Estimating the total role of forests in the carbon cycle is challenging, but studies indicate that old growth forests alone sequester up to 1.4 GtCyr^-1^ [@Malhi1999], while the total sequestration of carbon by established forests globally could be up to 2.4 GtCyr^-1^, with the largest sinks being in old-growth tropical forests [@Pan2011].
+As atmospheric carbon dioxide levels continue to rise, with consequences for global climate, there is increasing recognition that proper protection and management of forest resources will have an important role to play in mitigating climate change.
+Understanding the patterns of forest productivity on a global scale, and the drivers behind them, is therefore a priority in forest research.
 
-Little is known about the main environmental drivers of these patterns.
-
-The productivity of forests shows variation with latitude, with a general trend of increasing productivity towards the tropics; however the exact nature of this pattern and the mechanisms that drive it are poorly understood.
-Climate can be an important driver of productivity, particularly on a global scale; achieving a better understanding of the climatic controls on forest productivity will help to illuminate potential responses and feedbacks of forests to climate change.
-
-In order to better understand and predict responses and feedbacks to climate change, we need a better fundamental understand of how forest carbon cycling varies with respect to climate globally.
+There are two major questions to understand: firstly, how forest productivity varies globally, and specifically which areas show the greatest peaks in productivity; and secondly, which climate variables drive this variation.
+On a global scale, the productivity of forests varies with latitude, with a general trend of increasing productivity towards the tropics [@Beer2010; @Jung2011]; however the exact nature of this pattern, and how it varies by component of productivity, is poorly understood.
+This latitudinal gradient is most likely to be explained by climatic gradients in temperature, precipitation, length of growing season, and combinations of the above.
+Productivity is influenced by a range of climatic drivers, including mean annual temperature (MAT) and mean annual precipitation (MAP) [@DelGrosso2008], but doesn't necessarily respond linearly to these drivers.
+Disentangling the shape of productivity responses to climate drivers will enable better predictions of future responses under climate change.
 
@@
 The component of GPP that is stored as plant material is the net primary productivity (NPP) of an ecosystem; thus GPP can be described as the sum of NPP and ecosystem respiration.
-
 Currently, GPP cannot be measured directly by observing total ecosystem photosynthesis.
@@
 Fine root productivity is estimated using a variety of methods, such as taking soil cores and minirhizotrons, making it typically more reliable than measures of coarse root productivity.
+
+[[Belowground productivity is a potentially often underestimated component of productivity, but has been shown to be significant, particularly in higher latitude forests [@Pan2011].]]
 
@@
 
-Data that control for stand age and standardize methodologies are required to resolve this question. Here, we use a comprehensive global database to explore global patterns in productivity.
+Data that control for stand age and standardize methodologies are required to resolve this question.
+Here, we use a comprehensive global database to explore global patterns in productivity.
+We explore three questions:
+1. Which climatic variables are the most important drivers behind the latitudinal pattern in primary productivity?
+2. Do the components of primary productivity show variation in their responses to these climatic drivers? Is carbon allocation climate invariant?
+3.
+Does climate explain the same proportion of variation in different components of primary productivity?
+Is the relative importance of climatic drivers the same across components of primary productivity?
 

### 2019-06-05 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
 
-add hypotheses, can number/bullet point them
-redo first couple paragraphs
-look again at breakdown of primary productivity components --> add discussion on understanding belowground productivity
+, fig.cap = "Map showing all data used in the analysis, coded by variable"
 
@@
 
-There are two major questions to understand: firstly, how forest productivity varies globally, and specifically which areas show the greatest peaks in productivity; and secondly, which climate variables drive this variation.
+There are two major questions to understand: firstly, how forest productivity varies globally - and specifically which areas show the greatest peaks in productivity -; and secondly, which climate variables drive this variation.
 On a global scale, the productivity of forests varies with latitude, with a general trend of increasing productivity towards the tropics [@Beer2010; @Jung2011]; however the exact nature of this pattern, and how it varies by component of productivity, is poorly understood.
@@
 
-*Patterns of primary productivity.* Understanding variation in primary productivity is complex.
-Firstly, there are many factors which can influence productivity on a range of scales, many of which show interactive effects with each other.
-In addition, the primary productivity of a forest is the sum of the primary productivity of a number of individual components, which may all show individual responses to drivers of productivity.
-Understanding these relationships and interdependencies is complex.
+*What is primary productivity?* During photosynthesis, plants capture carbon dioxide from the atmosphere.
+The gross primary productivity (GPP) of an ecosystem is the gross uptake, via photosynthesis, of carbon dioxide by plants in that ecosystem.
+Only a fraction of the carbon captured is assimilated into plant tissue; the rest is used in autotrophic respiration (R~a~).
+The component of GPP that is stored as plant material is the net primary productivity (NPP) of an ecosystem.
+Net primary productivity can therefore be expressed as:
 
-Primary productivity of forests is influenced by a range of factors.
-On a local scale, stand age [@Litton2007; @Gillman2015], management [@Simova2017]; nutrient availability [@Aragao2009]; and altitude [@Girardin2010; @Malhi2017] all impact stand productivity.
-On a global scale, changes in primary productivity are influenced by climatic variables and abiotic gradients, such as the length of growing season [@Michaletz2014].
-The majority of studies to date have focused on the influence of MAP and MAT on productivity.
-While MAP and MAT have been argued to be significant predictors of productivity [@Chu2016], other studies have found that the correlation between productivity and MAT is a factor of the relationship between productivity and growing season length [@Kerkhoff2005; @Malhi2012; @Michaletz2014; @Michaletz2018].
-The influence of other climate variables on global patterns of productivity remains under-explored.
+$$ NPP = GPP + R~a~ $$
 
-*Understanding components of primary productivity.* There are several different measures of primary productivity.
-Gross primary productivity (GPP) represents the gross uptake of carbon dioxide by plants in an ecosystem, to be used in photosynthesis and energy production.
-Only a fraction of this carbon is assimilated into plant tissues; the rest is used in autotrophic respiration.
-The component of GPP that is stored as plant material is the net primary productivity (NPP) of an ecosystem; thus GPP can be described as the sum of NPP and ecosystem respiration.
 Currently, GPP cannot be measured directly by observing total ecosystem photosynthesis.
@@
 
-NPP is typically measured through direct measurements of the productivity of its components, which are then summed to generate a whole-ecosystem estimate.
-These components can be broadly divided into aboveground NPP (ANPP) and belowground NPP (BNPP).
-Measurement of aboveground NPP - primarily consisting of woody stem, woody branch and canopy/foliage primary productivity, but also sometimes including estimates of productivity of reproductive and understory components, and losses due to herbivory - is most commonly estimated through forest plot censuses.
-Belowground NPP is significantly more challenging to measure, meaning that estimates of belowground productivity are less reliable than those of aboveground productivity.
-Coarse root productivity is typically determined by using allometric equations relating root mass to stem diameter.
-This relies on extrapolation of ANPP measures, rather than being based on direct measures of root productivity.
-Fine root productivity is estimated using a variety of methods, such as taking soil cores and minirhizotrons, making it typically more reliable than measures of coarse root productivity.
+In contrast, NPP can be calculated through direct field observations.
+In order to achieve greater accuracy in estimating NPP, NPP is often broken down into its component parts, with aboveground NPP (ANPP) and belowground NPP (BNPP) considered separately to each other.
+The components included when estimating ANPP and BNPP often vary between studies, depending on the intensity of fieldwork effort.
+At its most basic level, ANPP can be expressed as:
 
-[[Belowground productivity is a potentially often underestimated component of productivity, but has been shown to be significant, particularly in higher latitude forests [@Pan2011].]]
+$$ ANPP = NPP~stem~ + NPP~branch~ + NPP~canopy~ $$
+where NPP~stem~ is the annual woody increment of all stems above a specified diameter at breast height (DBH), NPP~branch~ is annual woody branch turnover, and NPP~canopy~ is annual foliage production, including leaves, twigs, and reproductive structures.
+ANPP may also include NPP~VOC~, the annual emission of volatile organic compounds, and NPP~herbivory~, the annual consumption of plant matter by herbivores, but these components are often excluded from field observations as they are much harder to quantify.
+Other components of aboveground productivity that remain largely unquantified include epiphytes, hemiepiphytes, and understory plants [@Clark2017].
+All current ANPP estimates are based on the assumption that the contribution of these components to overall NPP is insignificant.
 
-Current knowledge indicates that these different components of productivity show different latitudinal patterns.
+There are two major subcomponents of BNPP, which can be expressed as:
+
+$$ BNPP = NPP~coarse root~ + NPP~fine root~ $$
+where NPP~coarse root~ is the annual production of coarse roots (typically roots >2mm diameter), and NPP~fine root~ is the annual production of fine roots (typically roots <2mm diameter) [@Aragao2009].
+Calculations of BNPP may also include NPP~exudation~, a measure of annual carbon losses through root exudation, and NPP~symbionts~, the annual carbon allocation to mycorrhizae and legumes, but, as before, this is challenging to quantify and is often excluded from field observations.
+
+BNPP is a poorly understood component of total ecosystem productivity, primarily because of the challenges in gaining accurate field measurements.
+Coarse root productivity is often estimated via extrapolation of NPP~stem~ estimates using allometries that may not have been empirically verified [@Clark2001].
+NPP~fine root~ is easier to quantify through soil cores and minirhizotrons, however, sampling tends to be limited to the surface soils, with very few studies sampling to depths below 3 metres [cite].
+As a result, it is possible that BNPP is currently significantly underestimated, despite being a hugely significant component of total ecosystem productivity [@Pan2011].
+
+*Which factors influence primary productivity?* Primary productivity is influenced by many factors, which often act across a range of scales, and may show interactive effects with each other.
+On a local scale, stand age [@Litton2007; @Gillman2015], management [@Simova2017]; nutrient availability [@Aragao2009]; and altitude [@Girardin2010; @Malhi2017] all impact forest productivity.
+On a global scale, changes in primary productivity are influenced by climatic variables and abiotic gradients, such as the length of growing season [@Michaletz2014].
+There is some debate over the precise relationship between these drivers and productivity; While mean annual precipitation (MAP) and mean annual temperature (MAT) have been argued to be significant predictors of productivity [@Chu2016], other studies have found that the correlation between productivity and MAT is a factor of the relationship between productivity and growing season length [@Kerkhoff2005; @Malhi2012; @Michaletz2014; @Michaletz2018].
+Improving understanding of how these factors interact to control global patterns in primary productivity is essential to understanding the global carbon cycle.
+
+Current research into how primary productivity varies with latitude is inconclusive, and - though it has primarily focussed on patterns of GPP, NPP, and ANPP - indicates that different components of productivity may show different relationships to latitude.
 Gross primary productivity is generally thought to be highest in the tropics.
 Modelling of global terrestrial ecosystem GPP though upscaling and calibration of eddy flux measurements indicates a peak in GPP in the tropics, with the highest levels in tropical forests [@Beer2010; @Jung2011].
-This is corroborated by analysis of site-level GPP measurements, which show a strong positive correlation between GPP and mean annual temperature (MAT) and mean annual precipitation (MAP) [@Luyssaert2007], with the highest GPP values reported in tropical forests.
+This is corroborated by analysis of site-level GPP measurements, which show a strong positive correlation between GPP and MAT and MAP [@Luyssaert2007], with the highest GPP values reported in tropical forests.
 The influence of latitude on global patterns of NPP is less clear than that of GPP.
@@
 
-These variations in the responses of different components of productivity to changes in latitude suggest that components may vary in the strength of their response to climatic drivers.
-There is some empirical evidence for this: increases in GPP are reported to saturate above 25$^\circ$C MAT [@Larjavaara2012], while increases in NPP are recorded to saturate above 10$^\circ$C MAT [@Luyssaert2007].
+Furthermore, there is evidence that different components of productivity show individual responses to drivers of productivity.
+For example, increases in GPP have been reported to saturate above 25$^\circ$C MAT [@Larjavaara2012], while increases in NPP are recorded to saturate above 10$^\circ$C MAT [@Luyssaert2007].
 Similarly, allocation to different components of ANPP varies with climate.
@@
 However, these studies are regional, meaning that our understanding of variation in allocation and its relationship to climate on a global scale remains limited.
-
-Allocation is also influenced by stand age [@DeLucia2007], nutrient availability [@Litton2007] and forest structure [@Taylor2019], which can make it challenging to disentangle the effects of climate.
+In addition, allocation is also influenced by stand age [@DeLucia2007], nutrient availability [@Litton2007] and forest structure [@Taylor2019], which can make it challenging to disentangle the effects of climate.
 
@@
 We explore three questions:
-1. Which climatic variables are the most important drivers behind the latitudinal pattern in primary productivity?
-2. Do the components of primary productivity show variation in their responses to these climatic drivers? Is carbon allocation climate invariant?
+1.
+**Which climatic variables are the most important drivers behind the latitudinal pattern in primary productivity?** To date, the majority of studies have focussed on productivity responses to precipitation and temperature, while the influence of other climate variables on productivity remains under-explored.
+We utilise global climate datasets to investigate the relationships between productivity and a range of climate variables.
+2.
+**Do the components of primary productivity show variation in their responses to these climatic drivers?** Allocation to components of primary productivity appears to show some variation across climate and latitudinal gradients.
+We use global datasets of a range of primary productivity components to explore how significant this variation in allocation is.
 3.
-Does climate explain the same proportion of variation in different components of primary productivity?
-Is the relative importance of climatic drivers the same across components of primary productivity?
+**Does climate explain the same proportion of variation in different components of primary productivity?** Review of current research shows that, while the relationship between GPP and latitude is consistent, NPP and ANPP show weaker relationships with latitude.
+This may be a result of varying sensitivities to climate drivers.
+We use mixed models to estimate the proportion of variation explained by climate for different components of productivity, to investigate how the relative importance of climatic drivers varies across components.
 
@@
 
----
-output:
-  pdf_document:
-    fig_caption: yes
----
-
-![Figure 1: Map showing all data used in the anlysis, coded by variable \label{figurelabel}](distribution_all_variables.png)
+  ```{r echo=FALSE, out.width='100%'}
+  knitr::include_graphics("distribution_all_variables.png")
+  ```
 
@@
 ranch turnover. Any measures of primary productivity where components were unknown were excluded from analysis.
-
-### References

### 2019-06-07 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
     citation_package: natbib
+    fig_caption: yes
 bibliography: library.bib
 csl: apa.csl
+header-includes:  \usepackage{float}
 ---
 
-, fig.cap = "Map showing all data used in the analysis, coded by variable"
+```{r, include=FALSE}
+options(tinytex.verbose = TRUE)
+```
 
@@
 Analyses were conducted on data contained in the open-access ForC database [@Anderson-Teixeira2016; @Anderson-Teixeira2018].
-This database contains records of field-based measurements of forest carbon stocks and annual fluxes, compiled from previously published studies and existing databases (e.g.
-ORNL DAAC; Luyssaert etc?).
+This database contains records of field-based measurements of forest carbon stocks and annual fluxes, compiled from original publications and existing data compilations and databases.
+Associated data, such as stand age, measurement methodologies, and disturbance history, are also included.
+For each site, site geographic co-ordinates were used to extract the Koeppen-Geiger zone from the ESRI Koeppen-Geiger map [cite], and the FAO Global Ecological Zone from the ???
 Additional targeted literature searches were conducted to identify any further available data on primary productivity, with particular focus on old-growth forests in temperate and boreal regions.
+ForC currently contains [] records from [] plots, representing [] distinct geographic areas across all forested biogeographic and climate zones.
 
-  ```{r echo=FALSE, out.width='100%'}
+  ```{r echo=FALSE, out.width='100%', fig.cap = "Map showing all data used in the analysis, coded by variable", fig.pos='H'}
   knitr::include_graphics("distribution_all_variables.png")
   ```
-
-*Data selection.* ForC contains measures of carbon stocks and fluxes; for the purposes of this analysis only measures of primary productivity were selected from the database.
-The variables selected were: GPP, NPP, BNPP (total root and fine root), ANPP (total ANPP, foliage, wood and woody stem).
+*Data selection.* Over 50 variables of forest carbon stocks and annual fluxes are represented in the ForC database; this analysis focussed on measures of primary productivity.
+Table 1 contains details of the variables selected for analysis.
 
 | Variable | Definition | Components included | Methodologies used|
-|-------------|-------------------|-------------|------------------------|
+|--------------|-------------------|-------------|------------------------|
 | GPP | Annual gross primary production; annual uptake of carbon dioxide by an ecosystem | NA | Flux partitioning of eddy covariance |
@@
 
-Sites were excluded from analysis if they were managed, defined as plots that were planted, managed as plantations, irrigated, fertilised or including the term "managed" in their site description.
+Table: Definitions of variables used in analysis
+
+A subset of the ForC database was generated for the purposes of this analysis, in order to control for data quality and remove biasing factors.
+Since management can alter observed patterns of primary productivity [@Simova2017], sites were excluded from analysis if they were managed, defined as plots that were planted, managed as plantations, irrigated, fertilised or including the term "managed" in their site description.
 Sites that had experienced significant disturbance were also excluded.
@@
 To reduce any biasing effects of stand age, stands under 100 years of age were excluded from analysis.
+Sites for which stand age was unknown were excluded from analysis.
 
@@
 For this reason, criteria were introduced to standardise for differences in methodology.
-Where data was based on forest plot census measurements, studies which uses a minimum dbh measure of 10cm or greater were excluded from analysis.
+Where data was based on forest plot census measurements, studies which used a minimum diameter at breast height (DBH) measure of 10cm or greater were excluded from analysis.
 Estimates of total ecosystem primary productivity are based on measurements of the component parts of forest productivity.
@@
 *Climate datasets.* Where site-level data on mean annual temperature, mean annual precipitation, and latitude were available in the primary literature, this data was compiled and entered directly into the ForC database.
-In addition to this data, climate data for each site was extracted from five open-access climate datasets.
+In addition to this data, climate data for each site was extracted from five open-access climate datasets based on site geographic co-ordinates.
 
@@
 
+Table: Sources of climate data
+
 *Model specification.* The effects of climate and latitude on primary productivity were analysed using mixed effects models using the package 'lme4' in R v.

### 2019-06-07 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
   pdf_document:
-    citation_package: natbib
     fig_caption: yes
 bibliography: library.bib
-csl: apa.csl
+csl: global-change-biology.csl
 header-includes:  \usepackage{float}
@@
 Associated data, such as stand age, measurement methodologies, and disturbance history, are also included.
-For each site, site geographic co-ordinates were used to extract the Koeppen-Geiger zone from the ESRI Koeppen-Geiger map [cite], and the FAO Global Ecological Zone from the ???
+For each site, site geographic co-ordinates were used to extract the Koeppen-Geiger zone from the ESRI Koeppen-Geiger map [cite], and the FAO Global Ecological Zone from the ???.
 Additional targeted literature searches were conducted to identify any further available data on primary productivity, with particular focus on old-growth forests in temperate and boreal regions.
@@
 Disturbances that justified site exclusion were major cutting or harvesting, and/or burning, flooding, drought and storm events with site mortality >10% of trees.
+Grazed sites were retained.
 There is evidence that stand age influences patterns of primary productivity and carbon allocation in forest ecosystems, and can confound relationships between latitude and primary productivity [@DeLucia2007; @Gillman2015].
@@
 Where data was based on forest plot census measurements, studies which used a minimum diameter at breast height (DBH) measure of 10cm or greater were excluded from analysis.
-Estimates of total ecosystem primary productivity are based on measurements of the component parts of forest productivity.
+It would be preferable to standardise by minimum area sampled; however x% of plots in the database are 1 ha or under in size; excluding these plots would place significant constraints on sample size.
+
+As discussed above, estimates of NPP, ANPP, and BNPP are generated through summing measurements of their component parts.
 Since the components included in productivity estimates vary between studies, estimates of productivity were classified within the ForC database according to their components, and then filtered for analysis.
-Estimates of NPP were selected if they included foliage, branch, stem, coarse root, fine root and understory.
+Estimates of NPP were selected if they included foliage, branch, stem, coarse root, fine root and optionally understory.
 Measures of NPP which included additional components were excluded.
@@
 
-*Model specification.* The effects of climate and latitude on primary productivity were analysed using mixed effects models using the package 'lme4' in R v.
+*Model specification.* The effects of climate and latitude on primary productivity were analysed using mixed effects models using the package 'lme4' in RStudio v.
 1.1.463 (cite).
@@
 Data from the temperate regions was heavily skewed towards studies from the old-growth forests of the Pacific Northwest.
-These forests have very high productivity, and so to reduce any bias from over-sampling of this region, the models were weighted according to the proportion of forest cover by Koeppen climate zone.
-ranch turnover. Any measures of primary productivity where components were unknown were excluded from analysis.
+These forests have very high productivity, and so to reduce any bias from over-sampling of this region, the proportion of global forest cover contributed by each Koeppen climate zone was calculated, and the models weighted according these proportions.
+
+Two models were specified: one with the climate variable as a linear term, and one with the climate variable as a polynomial term.
+AIC values were calculated for the models and used to select the best model.
+R-squared values were calculated for the best model.
+In addition, slope values were calculated for the linear models.

### 2019-06-07 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
 | Database | Variables downloaded | Citation |
-|---------------|-------------------------------------------------|-------------------|
+|-----------------------|--------------------------------------------|---------------------|
 | WorldClim | Mean annual temperature; temperature seasonality; annual temperature range; mean annual precipitation | [@Hijmans2005] |
@@
 In addition, slope values were calculated for the linear models.
+
+###References

### 2019-06-10 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
   pdf_document:
+    citation_package: natbib
     fig_caption: yes
 bibliography: library.bib
-csl: global-change-biology.csl
+biblio-style: apalike
+csl: apa.csl
 header-includes:  \usepackage{float}
+
 ---
@@
 
-Globally, forests play an important role in the carbon cycle and are an important component of global carbon dioxide budgets [@Luyssaert2008].
-They show higher levels of productivity than non-forest  terrestrial ecosystems [@DelGrosso2008], and as a result achieve significant carbon sequestration and storage.
-Estimating the total role of forests in the carbon cycle is challenging, but studies indicate that old growth forests alone sequester up to 1.4 GtCyr^-1^ [@Malhi1999], while the total sequestration of carbon by established forests globally could be up to 2.4 GtCyr^-1^, with the largest sinks being in old-growth tropical forests [@Pan2011].
+Globally, forests play an important role in the carbon cycle and are an important component of global carbon dioxide budgets [@luyssaert_old-growth_2008].
+They show higher levels of productivity than non-forest  terrestrial ecosystems [@del_grosso_global_2008], and as a result achieve significant carbon sequestration and storage.
+Estimating the total role of forests in the carbon cycle is challenging, but studies indicate that old growth forests alone sequester up to 1.4 GtCyr^-1^ [@malhi_carbon_1999], while the total sequestration of carbon by established forests globally could be up to 2.4 GtCyr^-1^, with the largest sinks being in old-growth tropical forests [@pan_large_2011].
 As atmospheric carbon dioxide levels continue to rise, with consequences for global climate, there is increasing recognition that proper protection and management of forest resources will have an important role to play in mitigating climate change.
@@
 There are two major questions to understand: firstly, how forest productivity varies globally - and specifically which areas show the greatest peaks in productivity -; and secondly, which climate variables drive this variation.
-On a global scale, the productivity of forests varies with latitude, with a general trend of increasing productivity towards the tropics [@Beer2010; @Jung2011]; however the exact nature of this pattern, and how it varies by component of productivity, is poorly understood.
+On a global scale, the productivity of forests varies with latitude, with a general trend of increasing productivity towards the tropics [@beer_terrestrial_2010; @jung_global_2011]; however the exact nature of this pattern, and how it varies by component of productivity, is poorly understood.
 This latitudinal gradient is most likely to be explained by climatic gradients in temperature, precipitation, length of growing season, and combinations of the above.
-Productivity is influenced by a range of climatic drivers, including mean annual temperature (MAT) and mean annual precipitation (MAP) [@DelGrosso2008], but doesn't necessarily respond linearly to these drivers.
+Productivity is influenced by a range of climatic drivers, including mean annual temperature (MAT) and mean annual precipitation (MAP) [@del_grosso_global_2008], but doesn't necessarily respond linearly to these drivers.
 Disentangling the shape of productivity responses to climate drivers will enable better predictions of future responses under climate change.
@@
 Currently, GPP cannot be measured directly by observing total ecosystem photosynthesis.
-Instead field estimates of GPP have to be derived based on modelling and extrapolation of eddy-covariance studies and measurements of net ecosystem exchange (NEE) [@Clark2017].
+Instead field estimates of GPP have to be derived based on modelling and extrapolation of eddy-covariance studies and measurements of net ecosystem exchange (NEE) [@clark_reviews_2017].
 
@@
 ANPP may also include NPP~VOC~, the annual emission of volatile organic compounds, and NPP~herbivory~, the annual consumption of plant matter by herbivores, but these components are often excluded from field observations as they are much harder to quantify.
-Other components of aboveground productivity that remain largely unquantified include epiphytes, hemiepiphytes, and understory plants [@Clark2017].
+Other components of aboveground productivity that remain largely unquantified include epiphytes, hemiepiphytes, and understory plants [@clark_reviews_2017].
 All current ANPP estimates are based on the assumption that the contribution of these components to overall NPP is insignificant.
@@
 $$ BNPP = NPP~coarse root~ + NPP~fine root~ $$
-where NPP~coarse root~ is the annual production of coarse roots (typically roots >2mm diameter), and NPP~fine root~ is the annual production of fine roots (typically roots <2mm diameter) [@Aragao2009].
+where NPP~coarse root~ is the annual production of coarse roots (typically roots >2mm diameter), and NPP~fine root~ is the annual production of fine roots (typically roots <2mm diameter) [@aragao_above-_2009].
 Calculations of BNPP may also include NPP~exudation~, a measure of annual carbon losses through root exudation, and NPP~symbionts~, the annual carbon allocation to mycorrhizae and legumes, but, as before, this is challenging to quantify and is often excluded from field observations.
@@
 BNPP is a poorly understood component of total ecosystem productivity, primarily because of the challenges in gaining accurate field measurements.
-Coarse root productivity is often estimated via extrapolation of NPP~stem~ estimates using allometries that may not have been empirically verified [@Clark2001].
+Coarse root productivity is often estimated via extrapolation of NPP~stem~ estimates using allometries that may not have been empirically verified [@clark_net_2001].
 NPP~fine root~ is easier to quantify through soil cores and minirhizotrons, however, sampling tends to be limited to the surface soils, with very few studies sampling to depths below 3 metres [cite].
-As a result, it is possible that BNPP is currently significantly underestimated, despite being a hugely significant component of total ecosystem productivity [@Pan2011].
+As a result, it is possible that BNPP is currently significantly underestimated, despite being a hugely significant component of total ecosystem productivity [@pan_large_2011].
 
 *Which factors influence primary productivity?* Primary productivity is influenced by many factors, which often act across a range of scales, and may show interactive effects with each other.
-On a local scale, stand age [@Litton2007; @Gillman2015], management [@Simova2017]; nutrient availability [@Aragao2009]; and altitude [@Girardin2010; @Malhi2017] all impact forest productivity.
-On a global scale, changes in primary productivity are influenced by climatic variables and abiotic gradients, such as the length of growing season [@Michaletz2014].
-There is some debate over the precise relationship between these drivers and productivity; While mean annual precipitation (MAP) and mean annual temperature (MAT) have been argued to be significant predictors of productivity [@Chu2016], other studies have found that the correlation between productivity and MAT is a factor of the relationship between productivity and growing season length [@Kerkhoff2005; @Malhi2012; @Michaletz2014; @Michaletz2018].
+On a local scale, stand age [@litton_carbon_2007; @gillman_latitude_2015], management [@simova_enigma_2017]; nutrient availability [@aragao_above-_2009]; and altitude [@girardin_net_2010; @malhi_variation_2017] all impact forest productivity.
+On a global scale, changes in primary productivity are influenced by climatic variables and abiotic gradients, such as the length of growing season [@michaletz_convergence_2014].
+There is some debate over the precise relationship between these drivers and productivity; While mean annual precipitation (MAP) and mean annual temperature (MAT) have been argued to be significant predictors of productivity [@chu_does_2016], other studies have found that the correlation between productivity and MAT is a factor of the relationship between productivity and growing season length [@kerkhoff_plant_2005; @malhi_productivity_2012; @michaletz_convergence_2014; @michaletz_drivers_2018].
 Improving understanding of how these factors interact to control global patterns in primary productivity is essential to understanding the global carbon cycle.
@@
 Gross primary productivity is generally thought to be highest in the tropics.
-Modelling of global terrestrial ecosystem GPP though upscaling and calibration of eddy flux measurements indicates a peak in GPP in the tropics, with the highest levels in tropical forests [@Beer2010; @Jung2011].
-This is corroborated by analysis of site-level GPP measurements, which show a strong positive correlation between GPP and MAT and MAP [@Luyssaert2007], with the highest GPP values reported in tropical forests.
+Modelling of global terrestrial ecosystem GPP though upscaling and calibration of eddy flux measurements indicates a peak in GPP in the tropics, with the highest levels in tropical forests [@beer_terrestrial_2010; @jung_global_2011].
+This is corroborated by analysis of site-level GPP measurements, which show a strong positive correlation between GPP and MAT and MAP [@luyssaert_co2_2007], with the highest GPP values reported in tropical forests.
 The influence of latitude on global patterns of NPP is less clear than that of GPP.
 Simova and Storch (2017) found that, as with GPP, NPP decreases with latitude, peaking in the tropics.
-However, other studies have found the highest values of NPP in temperate forests [@Luyssaert2007; @Huston2009].
+However, other studies have found the highest values of NPP in temperate forests [@luyssaert_co2_2007; @huston_global_2009].
 Because of the challenges in accurately obtaining unbiased measures of belowground productivity, many studies focus on ANPP in preference to measures of NPP.
 Studies on global patterns of ANPP are equally inconclusive: Gillman et al.
-[-@Gillman2015] found a weak negative relationship between ANPP and latitude, with the relationship becoming stronger in older forest stands.
-These findings were echoed in other studies, which have found weak or no relationships between ANPP and latitude [@Huston2009].
+(2015) found a weak negative relationship between ANPP and latitude, with the relationship becoming stronger in older forest stands.
+These findings were echoed in other studies, which have found weak or no relationships between ANPP and latitude [@huston_global_2009].
 
 Furthermore, there is evidence that different components of productivity show individual responses to drivers of productivity.
-For example, increases in GPP have been reported to saturate above 25$^\circ$C MAT [@Larjavaara2012], while increases in NPP are recorded to saturate above 10$^\circ$C MAT [@Luyssaert2007].
+For example, increases in GPP have been reported to saturate above 25$^\circ$C MAT [@larjavaara_temperature_2012], while increases in NPP are recorded to saturate above 10$^\circ$C MAT [@luyssaert_co2_2007].
 Similarly, allocation to different components of ANPP varies with climate.
-Within the tropics, allocation to canopy NPP appears fairly consistent, with significantly greater variation in allocation to woody and belowground NPP [@Litton2007; @Malhi2011].
-Allocation to these structural biomass components has been shown to increase with water availability [@Litton2007; @Bloom2016], and is highest in the wet tropics, indicating that control of woody productivity by MAP may be more significant than control of foliar productivity.
+Within the tropics, allocation to canopy NPP appears fairly consistent, with significantly greater variation in allocation to woody and belowground NPP [@litton_carbon_2007; @malhi_allocation_2011].
+Allocation to these structural biomass components has been shown to increase with water availability [@litton_carbon_2007; @bloom_decadal_2016], and is highest in the wet tropics, indicating that control of woody productivity by MAP may be more significant than control of foliar productivity.
 However, these studies are regional, meaning that our understanding of variation in allocation and its relationship to climate on a global scale remains limited.
-In addition, allocation is also influenced by stand age [@DeLucia2007], nutrient availability [@Litton2007] and forest structure [@Taylor2019], which can make it challenging to disentangle the effects of climate.
+In addition, allocation is also influenced by stand age [@de_lucia_forest_2007], nutrient availability [@litton_carbon_2007] and forest structure [@taylor_greater_2019], which can make it challenging to disentangle the effects of climate.
 
@@
 
-Analyses were conducted on data contained in the open-access ForC database [@Anderson-Teixeira2016; @Anderson-Teixeira2018].
+Analyses were conducted on data contained in the open-access ForC database [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018].
 This database contains records of field-based measurements of forest carbon stocks and annual fluxes, compiled from original publications and existing data compilations and databases.
@@
 A subset of the ForC database was generated for the purposes of this analysis, in order to control for data quality and remove biasing factors.
-Since management can alter observed patterns of primary productivity [@Simova2017], sites were excluded from analysis if they were managed, defined as plots that were planted, managed as plantations, irrigated, fertilised or including the term "managed" in their site description.
+Since management can alter observed patterns of primary productivity [@simova_enigma_2017], sites were excluded from analysis if they were managed, defined as plots that were planted, managed as plantations, irrigated, fertilised or including the term "managed" in their site description.
 Sites that had experienced significant disturbance were also excluded.
@@
 Grazed sites were retained.
-There is evidence that stand age influences patterns of primary productivity and carbon allocation in forest ecosystems, and can confound relationships between latitude and primary productivity [@DeLucia2007; @Gillman2015].
+There is evidence that stand age influences patterns of primary productivity and carbon allocation in forest ecosystems, and can confound relationships between latitude and primary productivity [@de_lucia_forest_2007; @gillman_latitude_2015].
 To reduce any biasing effects of stand age, stands under 100 years of age were excluded from analysis.
@@
 |-----------------------|--------------------------------------------|---------------------|
-| WorldClim | Mean annual temperature; temperature seasonality; annual temperature range; mean annual precipitation | [@Hijmans2005] |
-| WorldClim2 | Vapour pressure; solar radiation | [@Fick2017] |
-| Climate Research Unit (CRU) time-series dataset v 4.03 | Cloud cover; annual frost days; annual wet days; potential evapotranspiration | [@Harris2014] |
-| Global Aridity Index and Potential Evapotranspiration Climate Database | Aridity; potential evapotranspiration | [@Trabucco2018] |
-| TerraClimate | Vapour pressure deficit | [@Abatzoglou2018] |
+| WorldClim | Mean annual temperature; temperature seasonality; annual temperature range; mean annual precipitation | [@hijmans_very_2005] |
+| WorldClim2 | Vapour pressure; solar radiation | [@fick_worldclim_2017] |
+| Climate Research Unit (CRU) time-series dataset v 4.03 | Cloud cover; annual frost days; annual wet days; potential evapotranspiration | [@harris_updated_2014] |
+| Global Aridity Index and Potential Evapotranspiration Climate Database | Aridity; potential evapotranspiration | [@trabucco_global_2018] |
+| TerraClimate | Vapour pressure deficit | [@abatzoglou_terraclimate_2018] |
 
@@
 
-*Model specification.* The effects of climate and latitude on primary productivity were analysed using mixed effects models using the package 'lme4' in RStudio v.
-1.1.463 (cite).
+*Model specification.* The effects of climate and latitude on primary productivity were analysed using mixed effects models using the package 'lme4' [@bates_fitting_2015] in R v.3.5.1 [@r_core_team_r_2018].
 The effect of each extracted climate variable on each measure of primary productivity was modelled by specifying the climate variable as a fixed effect.

### 2019-06-12 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
 
+###Results
+
+*How do carbon fluxes vary with latitude?* All major carbon fluxes increase with decreasing latitude.
+All models were improved by the inclusion of altitude as a fixed effect, with carbon fluxes showing a significant decrease with increasing altitude.
+
+*Which climatic variables are the most important drivers behind the latitudinal pattern in primary productivity?* All major carbon fluxes show similar responses to climate variables.
+Of the climate variables tested, annual wet days, aridity, cloud cover, mean diurnal temperature range and precipitation seasonality were poor explainers of variation in primary productivity.
+Carbon fluxes show a linear increase with increasing mean annual temperature.
+However, they show a saturation point with mean annual precipitation, potential evapotranspiration, and vapour pressure deficit.
+Generally, the climate variables which explain the highest amount of variation in carbon fluxes are temperature seasonality, mean annual temperature, annual temperature range, and vapour pressure.
+For measures of aboveground woody productivity (ANPP_woody and ANPP_woody_stem), vapour pressure deficit,  potential evapotranspiration, and mean annual temperature are the strongest predictors.
+
 ###References

### 2019-06-14 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd
updated manuscript draft

@@
 ```
+
+###Abstract
+
+Globally, forests are a significant carbon sink, showing high levels of productivity and representing an important component of the carbon cycle.
+Current knowledge indicates that forest productivity decreases with latitude; however, there is evidence that this relationship varies across different fluxes.
+Furthermore, the climate variables that drive this relationship are poorly understood.
+Here, we use a comprehensive global database of forest carbon fluxes to show that, while all major carbon fluxes decrease with latitude, allocation to woody productivity increases with latitude, and allocation to root productivity declines.
+The best predictors of global patterns of productivity are temperature variables, including mean annual temperature, annual temperature range, and temperature seasonality, with the exception of woody productivity, which is most strongly influenced by potential evapotranspiration and vapour pressure deficit.
+Climate variables explain a large proportion of the variation in major carbon fluxes, but are a less significant predictor of subsidiary components.
+Our results illustrate the strong influence of climate on primary productivity, and particularly the importance of temperature in determining forest productivity.
+However, they also indicate that effects of climate are complex: fluxes show non-linear responses to climate variables, and many factors other than climate interact to determine allocation to subsidiary components.
+Knowledge of the relative influence of climate variables on forest carbon fluxes is an important component of understanding the responses of primary productivity to climate change.
 
@@
 Data that control for stand age and standardize methodologies are required to resolve this question.
-Here, we use a comprehensive global database to explore global patterns in productivity.
+Here, we use a comprehensive global database of forest carbon fluxes to explore global patterns in productivity.
 We explore three questions:
@@
 
-###References
+*Does climate explain the same proportion of variation in different components of primary productivity?*

### 2019-06-14 manuscript/manuscript draft.Rmd
update manuscript draft

@@
 
-Analyses were conducted on data contained in the open-access ForC database [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018].
+Analyses were conducted on data contained in the open-access ForC database [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc:_2018].
 This database contains records of field-based measurements of forest carbon stocks and annual fluxes, compiled from original publications and existing data compilations and databases.
@@
 
-*Model specification.* The effects of climate and latitude on primary productivity were analysed using mixed effects models using the package 'lme4' [@bates_fitting_2015] in R v.3.5.1 [@r_core_team_r_2018].
+*Model specification.* The effects of climate and latitude on primary productivity were analysed using mixed effects models using the package 'lme4' [@bates_fitting_2015] in R v.3.5.1 [@r_core_team_r:_2018].
 The effect of each extracted climate variable on each measure of primary productivity was modelled by specifying the climate variable as a fixed effect.

### 2019-06-17 scripts/Maps_script.R
Update Maps_script.R

@@ -109,7 +109,7 @@ fixed.variables <- c("AnnualMeanTemp")
 
 all.results <- NULL
 
-ForC_simplified$variable.name <- gsub("(_0|_1|_2)", "", ForC_simplified$variable.name)
+ForC_simplified$variable.name <- gsub("(_1|_2)", "", ForC_simplified$variable.name)
 
 ForC_simplified <- ForC_simplified[ForC_simplified$variable.name %in% c("GPP", "NPP", "BNPP_root", "BNPP_root_fine", "ANPP", "ANPP_foliage", "ANPP_woody", "ANPP_woody_stem"),]
 

### 2019-06-17 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
 Other components of aboveground productivity that remain largely unquantified include epiphytes, hemiepiphytes, and understory plants [@clark_reviews_2017].
-All current ANPP estimates are based on the assumption that the contribution of these components to overall NPP is insignificant.
+The majority of current ANPP estimates are based on the assumption that the contribution of these components to overall NPP is insignificant.
 
@@
 Estimates of NPP were selected if they included foliage, branch, stem, coarse root, fine root and optionally understory.
-Measures of NPP which included additional components were excluded.
+Measures of NPP which included additional components, including volatile organic compounds (VOCs), exudates, estimates of NPP lost to herbivory, and the NPP of reproductive structures, were excluded.
 Estimates of ANPP were selected if they included foliage, stem growth and branch turnover.
@@
 *Does climate explain the same proportion of variation in different components of primary productivity?*
+
+###Discussion
+
+*Challenges in estimating NPP*  
+The impact of excluding certain components of NPP on this analysis is unclear, primarily because these components are poorly understood and rarely measured [@clark_net_2001].
+Do we want to include a section assessing how much this may or may not impact results?
+1. Hard to measure BNPP because it often isn't sampled deep enough; impact on ratio tests?  
+2. Unclear if there are latitudinal patterns e.g. of herbivory/VOCs  
+  + Herbivory can be an important component of total NPP so this may be important to account for; likely to be higher in the tropics?

### 2019-06-19 scripts/best_model_scaled.R
Update best_model_scaled.R

@@ -62,7 +62,7 @@ ForC_simplified <- ForC_simplified[dist.to.keep, ]
 ## keep only records with min.dbh <= 10cm
 ForC_simplified$min.dbh <- as.numeric(ForC_simplified$min.dbh)
 # min.dbh.to.keep <- ForC_simplified$min.dbh <= 10 & is.na(ForC_simplified$min.dbh) # this doesn work great, not enough data, but maybe we can just keep the NAs?
-min.dbh.to.remove <- ForC_simplified$min.dbh >= 10 & !is.na(ForC_simplified$min.dbh)
+min.dbh.to.remove <- ForC_simplified$min.dbh > 10 & !is.na(ForC_simplified$min.dbh)
 ForC_simplified <- ForC_simplified[-min.dbh.to.remove, ]
 
 ###keep only altitude below 500 metres
@@ -249,16 +249,6 @@ for(response.variables in response.variables.groups){
         # lowerCI <- boot.ci(lmerboot, type = "bca")$'bca'[4]
         # upperCI <- boot.ci(lmerboot, type = "bca")$'bca'[5]
         
-        # summary <- summary(mod.linear)
-        # CI <- summary$coefficient[,"Std. Error"]*1.96
-        # lowerCI <- summary$coefficient[2,1] - CI[2]
-        # upperCI <- summary$coefficient[2,1] + CI[2]
-        
-        # profileCI <- confint.merMod(mod.linear, method = "profile")[5,]
-        # WaldCI <- confint.merMod(mod.linear, method = "Wald")[5,]
-        # bootCI <- confint.merMod(mod.linear, method = "boot")[5,]
-        # CI <- rbind(profileCI, WaldCI, bootCI)
-        
         r <- round(fixef(mod.linear), 5)
         slope <- r[2]
         
@@ -297,7 +287,7 @@ for(response.variables in response.variables.groups){
 }
 
 
-write.csv(all.results, file = "C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/tables/best_model_outputs/best_model_scaled_with_ci.csv", row.names = F)
+write.csv(all.results, file = "C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/tables/best_model_outputs/best_model_scaled.csv", row.names = F)
 
[…]

### 2019-06-19 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
 For this reason, criteria were introduced to standardise for differences in methodology.
-Where data was based on forest plot census measurements, studies which used a minimum diameter at breast height (DBH) measure of 10cm or greater were excluded from analysis.
+Where data was based on forest plot census measurements, studies which used a minimum diameter at breast height (DBH) measure of >10cm were excluded from analysis.
 It would be preferable to standardise by minimum area sampled; however x% of plots in the database are 1 ha or under in size; excluding these plots would place significant constraints on sample size.

### 2019-06-19 scripts/best_model_scaled.R
Update best_model_scaled.R

@@ -62,8 +62,9 @@ ForC_simplified <- ForC_simplified[dist.to.keep, ]
 ## keep only records with min.dbh <= 10cm
 ForC_simplified$min.dbh <- as.numeric(ForC_simplified$min.dbh)
 # min.dbh.to.keep <- ForC_simplified$min.dbh <= 10 & is.na(ForC_simplified$min.dbh) # this doesn work great, not enough data, but maybe we can just keep the NAs?
-min.dbh.to.remove <- ForC_simplified$min.dbh > 10 & !is.na(ForC_simplified$min.dbh)
-ForC_simplified <- ForC_simplified[-min.dbh.to.remove, ]
+## commented out because no values with min.dbh > 10; comment in with future merges if this changes
+# min.dbh.to.remove <- ForC_simplified$min.dbh > 10 & !is.na(ForC_simplified$min.dbh)
+# ForC_simplified <- ForC_simplified[-min.dbh.to.remove, ]
 
 ###keep only altitude below 500 metres
 # 

### 2019-06-19 scripts/best_model_scaled.R
Update best_model_scaled.R

@@ -59,11 +59,11 @@ ages <- c("age.greater.than.100")
 dist.to.keep <- ForC_simplified$managed %in% 0 & ForC_simplified$disturbed %in% 0
 ForC_simplified <- ForC_simplified[dist.to.keep, ]
 
-## keep only records with min.dbh <= 10cm
-ForC_simplified$min.dbh <- as.numeric(ForC_simplified$min.dbh)
-# min.dbh.to.keep <- ForC_simplified$min.dbh <= 10 & is.na(ForC_simplified$min.dbh) # this doesn work great, not enough data, but maybe we can just keep the NAs?
-## commented out because no values with min.dbh > 10; comment in with future merges if this changes
-# min.dbh.to.remove <- ForC_simplified$min.dbh > 10 & !is.na(ForC_simplified$min.dbh)
+## keep only records with min.dbh <= 10cm ## commented out because no values with min.dbh > 10; comment in with future merges if this changes
+# ForC_simplified$min.dbh <- as.numeric(ForC_simplified$min.dbh)
+# # min.dbh.to.keep <- ForC_simplified$min.dbh <= 10 & is.na(ForC_simplified$min.dbh) # this doesn work great, not enough data, but maybe we can just keep the NAs?
+# 
+# min.dbh.to.remove <- ForC_simplified$min.dbh >= 10 & !is.na(ForC_simplified$min.dbh)
 # ForC_simplified <- ForC_simplified[-min.dbh.to.remove, ]
 
 ###keep only altitude below 500 metres
@@ -99,8 +99,6 @@ all.response.variables <- gsub("(_OM|_C)", "", all.response.variables)
 all.response.variables <- gsub("(_0|_1|_2|_3|_4|_5)", "", all.response.variables)
 all.response.variables <- all.response.variables[all.response.variables %in% ForC_simplified$variable.name]
 all.response.variables <- unique(gsub("_\\d", "", all.response.variables))
-
-all.response.variables[!all.response.variables %in% unlist(response.variables.groups)]
 
 ## fixed variables list ####
 fixed.variables <- c("mat", "map", "lat", "AnnualMeanTemp", "MeanDiurnalRange", "TempSeasonality", "TempRangeAnnual", "AnnualPre", "PreSeasonality", "CloudCover", "AnnualFrostDays", "AnnualWetDays", "VapourPressure", "SolarRadiation", "Aridity", "PotentialEvapotranspiration", "VapourPressureDeficit")

### 2019-06-19 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
 Additional targeted literature searches were conducted to identify any further available data on primary productivity, with particular focus on old-growth forests in temperate and boreal regions.
-ForC currently contains [] records from [] plots, representing [] distinct geographic areas across all forested biogeographic and climate zones.
+ForC currently contains 29768 records from 5227 plots, representing 20 distinct ecozones across all forested biogeographic and climate zones.
 
@@
 
-*How do carbon fluxes vary with latitude?* All major carbon fluxes increase with decreasing latitude.
+*How do carbon fluxes vary with latitude?* All major carbon fluxes increase with decreasing latitude (fig.
+2).
 All models were improved by the inclusion of altitude as a fixed effect, with carbon fluxes showing a significant decrease with increasing altitude.
 
+  ```{r echo=FALSE, out.width='100%', fig.cap = "Graph to show primary productivity $(MgCha^-1^yr^-1^)$ regressed against latitude.
+Lines of best fit are plotted according to the best model selected during analysis.
+All regressions are significant $(p<0.05)$.", fig.pos='H'}
+  knitr::include_graphics("effect_of_latitude.png")
+  ```
+
+Latitude is a strong predictor for many of the carbon fluxes, explaining ~70% of variation in GPP (n = 241, p<0.01), 64% in NPP (n = 109, p<0.01) and 60% in ANPP (n = 252, p<0.01).
+
 *Which climatic variables are the most important drivers behind the latitudinal pattern in primary productivity?* All major carbon fluxes show similar responses to climate variables.
-Of the climate variables tested, annual wet days, aridity, cloud cover, mean diurnal temperature range and precipitation seasonality were poor explainers of variation in primary productivity.
-Carbon fluxes show a linear increase with increasing mean annual temperature.
-However, they show a saturation point with mean annual precipitation, potential evapotranspiration, and vapour pressure deficit.
-Generally, the climate variables which explain the highest amount of variation in carbon fluxes are temperature seasonality, mean annual temperature, annual temperature range, and vapour pressure.
-For measures of aboveground woody productivity (ANPP_woody and ANPP_woody_stem), vapour pressure deficit,  potential evapotranspiration, and mean annual temperature are the strongest predictors.
+Of the climate variables tested, annual wet days, aridity, cloud cover, mean diurnal temperature range and precipitation seasonality were poor explainers of variation in primary productivity, explaining less than 30% of the variation in each of the carbon fluxes.
+We found that both measures of mean annual precipitation used were significant (p<0.05) but poor predictors of productivity, explaining at most 45% of variation in carbon flux.
+Despite being a poor predictor, the results indicate that there is a saturation point at between 3000 and 4000mm annual precipitation, above which productivity no longer increases.
+Looking across all productivity components, the climate variables which explain the highest amount of variation in carbon fluxes are temperature seasonality, mean annual temperature, annual temperature range, and vapour pressure.
+For measures of woody productivity (ANPP~woody~ and ANPP~woody stem~), potential evapotranspiration and vapour pressure deficit were also strong predictors.
+**Temperature seasonality**  
+**Mean annual temperature**  
+**Annual temperature range**  
+**Vapour pressure**
+
+*Do carbon fluxes vary in their responses to climate, particularly with regards to allocation between fluxes?*
 

### 2019-06-19 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
 Looking across all productivity components, the climate variables which explain the highest amount of variation in carbon fluxes are temperature seasonality, mean annual temperature, annual temperature range, and vapour pressure.
-For measures of woody productivity (ANPP~woody~ and ANPP~woody stem~), potential evapotranspiration and vapour pressure deficit were also strong predictors.
+For measures of woody productivity (ANPP~woody~ and ANPP~woodystem~), potential evapotranspiration and vapour pressure deficit were also strong predictors.
+
+  ```{r echo=FALSE, out.width='100%', fig.cap = "Plots of primary productivity $(MgCha^-1^yr^-1^)$ regressed against the climate variables identified as the best predictors.
+Lines of best fit are plotted according to the best model selected during analysis.
+All regressions are significant $(p<0.05)$.", fig.pos='H'}
+  knitr::include_graphics("combined_plots.png")
+  ```
+
 **Temperature seasonality**  

### 2019-06-19 scripts/best_model_scaled.R
Update best_model_scaled.R
updated to add code for creating combined plots

@@ -304,7 +304,7 @@ for (age in ages){
     
     png(file = paste0("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/figures/final_figures/scaled_best_model_with_alt/Effect_of_", fixed.v, "_MATURE_only_poly_all.png"), width = 2255, height = 2000, units = "px", res = 300)
     
-    par(mfrow = c(1,1), mar = c(0,0,0,8), oma = c(5,5,2,0))
+    par(mfrow = c(2,2), mar = c(0,0,0,8), oma = c(5,5,2,0))
     print(fixed.v)
     
     ylim <- range(ForC_simplified[ForC_simplified$variable.name %in% unlist(response.variables),]$mean)
@@ -402,3 +402,130 @@ for (age in ages){
   }
   
 }
+
+  
+fixed.variables <- c("AnnualMeanTemp", "TempSeasonality", "TempRangeAnnual", "VapourPressure")
+  
+response.variables <- c("GPP", "NPP", "BNPP_root", "ANPP", "ANPP_foliage", "ANPP_woody_stem")
+
+all.results = NULL
+
+### mature forests only ####
+for (age in ages){
+  
+  if (age %in% "age.greater.than.100") ages.to.keep <- ForC_simplified$stand.age >= 100 & !is.na(ForC_simplified$stand.age)
+  if (age %in% "age.greater.than.200") ages.to.keep <- ForC_simplified$stand.age >= 200 & !is.na(ForC_simplified$stand.age)
+  
+  png(file = paste0("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/figures/final_figures/scaled_best_model_with_alt/combined_plots.png"), width = 2255, height = 2000, units = "px", res = 300)
+  
+  par(mfrow = c(2,2), mar = c(3,3,3,3))
+  
+  pannel.nb <- 1
+  
+  for(fixed.v in fixed.variables){
+    
+    
+    print(fixed.v)
+    
+    ylim <- range(ForC_simplified[ForC_simplified$variable.name %in% unlist(response.variables),]$mean)
+    
+    
+    ###subset ForC
+    response.variables.col <- c(1, 3, 4, 5, 6, 7)
+    
+    first.plot <- TRUE
+    
+    for (response.v in response.variables){
+      
[…]

### 2019-06-20 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
 Table 1 contains details of the variables selected for analysis.
+For the purposes of analysis, all estimates of NPP_1 and NPP_2 were combined as NPP, and all estimates of ANPP_1 and ANPP_2 were combined as ANPP.
 
@@
 | GPP | Annual gross primary production; annual uptake of carbon dioxide by an ecosystem | NA | Flux partitioning of eddy covariance |
-| NPP | Annual net primary production; the component of GPP that is stored in plant tissue; GPP minus ecosystem respiration | Foliage, branch, stem, coarse root, fine root, and optionally understory | Direct measurement of annual increments of components |
-| ANPP | Aboveground net primary production | Foliage, stem, and optionally branch | Direct measurement of annual increments of components | 
+| NPP_1 | Annual net primary production; the component of GPP that is stored in plant tissue; GPP minus ecosystem respiration | Foliage, branch, stem, coarse root and fine root | Direct measurement of annual increments of components |
+| NPP_2 | Annual net primary production; the component of GPP that is stored in plant tissue; GPP minus ecosystem respiration | Foliage, branch, stem, coarse root, fine root, and understory | Direct measurement of annual increments of components |
+| ANPP_1 | Aboveground net primary production | Foliage and stem | Direct measurement of annual increments of components | 
+| ANPP_2 | Aboveground net primary production | Foliage, stem, and branch | Direct measurement of annual increments of components |
 | ANPP_foliage | Net primary production of foliage | Foliage | Direct measurement of litterfall, correcting for changes in leaf biomass when measured |
@@
 
+*Validating models of component fluxes.* Comparison of component fluxes is based on the assumption that components sum accurately to estimates of larger fluxes.
+To test this, components of larger fluxes were regressed against latitude, and the models used to generate a series of point estimates along lines of best fit for each component.
+The point estimates for smaller component fluxes were summed to generate new "stacked" estimates of larger fluxes, which were then compared against actual measurements of the larger flux.
+Confidence intervals for the larger flux were calculated using the 'bootMer' function from the lme4 package.
+Stacked plots were generated for:
+1. ANPP_2 + BNPP_root = NPP_1  
+2. ANPP_foliage + ANPP_woody_stem = ANPP_1  
+3. NPP_1 + R_auto = GPP
+
 ###Results
@@
 
-  ```{r echo=FALSE, out.width='100%', fig.cap = "Graph to show primary productivity $(MgCha^-1^yr^-1^)$ regressed against latitude.
+  ```{r echo=FALSE, out.width='100%', fig.cap = "Graph to show primary productivity $(MgC~ha^-1~yr^-1)$ regressed against latitude.
 Lines of best fit are plotted according to the best model selected during analysis.
@@
 Latitude is a strong predictor for many of the carbon fluxes, explaining ~70% of variation in GPP (n = 241, p<0.01), 64% in NPP (n = 109, p<0.01) and 60% in ANPP (n = 252, p<0.01).
+
+Analysis of component fluxes indicates that measures of smaller component fluxes do approximately sum to measures of larger fluxes (fig.
+3).
+Modelled estimates of NPP, generated from the sum of ANPP and BNPP~root~, and ANPP, generated from the sum of ANPP~foliage~ and ANPP~woodystem~, fall completely within the confidence intervals of the regressions of field estimates of NPP and ANPP.
+
+  ```{r echo=FALSE, out.width='100%', fig.cap = "Graphs of primary productivity $(MgC~ha^-1~yr^-1)$ regressed against latitude.
+Lines of best fit are plotted according to the best model selected during analysis.
+All regressions are significant $(p<0.05)$.
+Plots show smaller component fluxes (1) and (2); the larger flux (3) defined by the combination of (1) and (2); and a modelled estimate of the sum of (1) and (2).
+95% confidence intervals are plotted for (3)", fig.pos='H'}
+  knitr::include_graphics("combined_stacked.png")
+  ```
 
@@
 
-  ```{r echo=FALSE, out.width='100%', fig.cap = "Plots of primary productivity $(MgCha^-1^yr^-1^)$ regressed against the climate variables identified as the best predictors.
+  ```{r echo=FALSE, out.width='100%', fig.cap = "Plots of primary productivity $(MgC~ha^-1~yr^-1)$ regressed against the climate variables identified as the best predictors.
 Lines of best fit are plotted according to the best model selected during analysis.
@@
 
-**Temperature seasonality**  
-**Mean annual temperature**  
-**Annual temperature range**  
-**Vapour pressure**
+**Mean annual temperature.** All carbon fluxes show a linear increase with temperature.  
+**Temperature seasonality.** Productivity is highest where temperature seasonality = 0.
+GPP, NPP, ANPP, and BNPP~root~ show a polynomial relationship with temperature seasonality.
+As seasonality increases, these variables show a more rapid decrease in productivity.
+ANPP~foliage~ and ANPP~woodystem~ show a linear relationship with temperature seasonality.
+**Annual temperature range.** As with temperature seasonality, productivity is highest at annual temperature ranges of 15$^\circ$C or lower.
+Again, GPP, NPP, ANPP, and BNPP~root~ show a polynomial relationship with temperature seasonality, and ANPP~foliage~ and ANPP~woodystem~ show a linear relationship with temperature seasonality.
+**Vapour pressure.** Productivity increases linearly with vapour pressure for all carbon fluxes, with the exception of ANPP~foliage~, which saturates at around 0.24.
 

### 2019-06-20 scripts/stacked_plots_code.R
Update stacked_plots_code.R

@@ -106,8 +106,10 @@ set1 <- c("ANPP_2")
 set2 <- c("BNPP_root")
 sum <- c("NPP_1")
   
+
+png(file = paste0("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/figures/final_figures/stacked_plots/combined_stacked.png"), width = 2255, height = 2000, units = "px", res = 300)
+
 par(mfrow = c(2,2), mar = c(1,1,1,1), oma = c(3,3,0,0))
- # png(file = paste0("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/figures/final_figures/stacked_plots/combined_stacked.png"), width = 2255, height = 2000, units = "px", res = 300)
 
   ### mature forests only ####
   for (age in ages){
@@ -711,4 +713,4 @@ for (age in ages){
   }
 }
 
-
+dev.off()

### 2019-06-21 scripts/boxplot_comparisons.R
Update boxplot_comparisons.R
added code to create combined box plots

@@ -6,6 +6,8 @@ setwd("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/ForC")
 
 library(ggplot2)
 library(ggpubr)
+library(cowplot)
+library(gridExtra)
 
 ForC_simplified <- read.csv("ForC_simplified/ForC_simplified_WorldClim_CRU_refined.csv", stringsAsFactors = F)
 ForC_simplified$site_plot <- paste0(ForC_simplified$sites.sitename," ", ForC_simplified$plot.name)
@@ -49,12 +51,75 @@ for (i in seq(along = set1)){
       
       my_comparisons <- list( c("boreal", "tropical"), c("boreal", "temperate"), c("tropical", "temperate"))
       
-      tiff(file = paste0("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/figures/final_figures/boxplot_comparisons/boxplot_", set1[[i]], "_", set2[[j]], ".tiff"), width = 2255, height = 2000, units = "px", res = 300)
+      png(file = paste0("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/figures/final_figures/boxplot_comparisons/boxplot_", set1[[i]], "_", set2[[j]], ".png"), width = 2255, height = 2000, units = "px", res = 300)
       p <- ggboxplot(df, x = "biomes", y = "ratio",
                      color = "biomes", palette = "jco", outlier.shape = NA, ylim = c(0, range), order = c("boreal", "temperate", "tropical")) +
-        stat_compare_means(comparisons = my_comparisons, label.y = c(range, range - 0.25 , range - 0.5))
+        stat_compare_means(comparisons = my_comparisons, hide.ns = T, label = "p.signif", label.y = c(range, range - 0.25 , range - 0.5))
       print(p)
       dev.off()
     }
   }
 }
+
+
+#########################
+
+set1 <- "ANPP_foliage"
+set2 <- "ANPP_woody"
+
+
+for (i in seq(along = set1)){
+  for (j in seq(along = set2)){
+    if (i == j){
+      resp1 <- ForC_simplified[ForC_simplified$variable.name %in% set1[[i]],]
+      resp2 <- ForC_simplified[ForC_simplified$variable.name %in% set2[[j]],]
+      
[…]

### 2019-06-21 scripts/boxplot_comparisons.R
updated boxplots with significance codes

@@ -54,7 +54,7 @@ for (i in seq(along = set1)){
       png(file = paste0("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/figures/final_figures/boxplot_comparisons/boxplot_", set1[[i]], "_", set2[[j]], ".png"), width = 2255, height = 2000, units = "px", res = 300)
       p <- ggboxplot(df, x = "biomes", y = "ratio",
                      color = "biomes", palette = "jco", outlier.shape = NA, ylim = c(0, range), order = c("boreal", "temperate", "tropical")) +
-        stat_compare_means(comparisons = my_comparisons, hide.ns = T, label = "p.signif", label.y = c(range, range - 0.25 , range - 0.5))
+        stat_compare_means(comparisons = my_comparisons, hide.ns = T, label = "p.signif", label.y = c(range, range - 0.25 , range - 0.5), tip.length = 0.01)
       print(p)
       dev.off()
     }

### 2019-06-21 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
 Plots show smaller component fluxes (1) and (2); the larger flux (3) defined by the combination of (1) and (2); and a modelled estimate of the sum of (1) and (2).
-95% confidence intervals are plotted for (3)", fig.pos='H'}
+95\\% confidence intervals are plotted for (3).", fig.pos='H'}
+
   knitr::include_graphics("combined_stacked.png")
@@
 
-  ```{r echo=FALSE, out.width='100%', fig.cap = "Plots of primary productivity $(MgC~ha^-1~yr^-1)$ regressed against the climate variables identified as the best predictors.
+```{r echo=FALSE, out.width='100%', fig.cap = "Plots of primary productivity $(MgC~ha^-1~yr^-1)$ regressed against the climate variables identified as the best predictors: mean annual temperature; temperature seasonality; annual temperature range; and vapour pressure.
 Lines of best fit are plotted according to the best model selected during analysis.
@@
 
+```{r echo=FALSE, out.width='100%', fig.cap = "Ratio of ANPP foliage to ANPP woody and ANPP woody stem plotted against major ecozones.
+All pairwise comparisons are significant.
+Significance levels of pairwise comparisons are indicated where * $p \\leq 0.05 $; ** $p \\leq 0.01 $; *** $p \\leq 0.001 $; **** $p \\leq 0.0001 $.", fig.pos='H'}
+  knitr::include_graphics("boxplot_foliage_woody_combined.png")
+```
+
 *Does climate explain the same proportion of variation in different components of primary productivity?*

### 2019-06-21 scripts/boxplot_comparisons.R
Update boxplot_comparisons.R

@@ -31,8 +31,10 @@ ForC_simplified$biomes <- ifelse((grepl("Subtropical", ForC_simplified$FAO.ecozo
 # ANPP_woody_and_foliage <- merge(ANPP_foliage, ANPP_woody[, c("variable.name", "date", "start.date", "end.date", "mean", "citation.ID", "site_plot", "stand.age")], by= c("site_plot", "citation.ID", "stand.age"))
 # ANPP_woody_stem_and_foliage <- merge(ANPP_foliage, ANPP_woody_stem[, c("variable.name", "date", "start.date", "end.date", "mean", "citation.ID", "site_plot", "stand.age")], by= c("site_plot", "citation.ID", "stand.age"))
 
+fixed.variables <- c("mat", "map", "lat", "AnnualMeanTemp", "AnnualPre", "TempSeasonality", "TempRangeAnnual", "VapourPressure", "VapourPressureDeficit")
+
 set1 <- c("GPP", "NPP_1", "ANPP_1", "ANPP_foliage", "ANPP_foliage", "NPP_1", "ANPP_2")
-set2 <- c("NPP_1", "ANPP_1", "BNPP_root", "ANPP_woody", "ANPP_woody_stem", "BNPP_root", "BNPP_root")
+set2 <- c("NPP_1", "ANPP_2", "BNPP_root", "ANPP_woody", "ANPP_woody_stem", "BNPP_root", "BNPP_root")
 
 
 for (i in seq(along = set1)){
@@ -57,11 +59,73 @@ for (i in seq(along = set1)){
         stat_compare_means(comparisons = my_comparisons, hide.ns = T, label = "p.signif", label.y = c(range, range - 0.25 , range - 0.5), tip.length = 0.01)
       print(p)
       dev.off()
+        
+      }
+      
+    }
+  }
+
+set1 <- c("ANPP_foliage", "ANPP_foliage")
+set2 <- c("ANPP_woody", "ANPP_woody_stem")
+
+
+for (i in seq(along = set1)){
+  for (j in seq(along = set2)){
+    if (i == j){
+      resp1 <- ForC_simplified[ForC_simplified$variable.name %in% set1[[i]],]
+      resp2 <- ForC_simplified[ForC_simplified$variable.name %in% set2[[j]],]
+      
[…]

### 2019-06-21 scripts/boxplot_comparisons.R
Update boxplot_comparisons.R

@@ -12,10 +12,13 @@ library(gridExtra)
 ForC_simplified <- read.csv("ForC_simplified/ForC_simplified_WorldClim_CRU_refined.csv", stringsAsFactors = F)
 ForC_simplified$site_plot <- paste0(ForC_simplified$sites.sitename," ", ForC_simplified$plot.name)
 ForC_simplified$stand.age <- as.numeric(ForC_simplified$stand.age)
-ForC_simplified <- ForC_simplified[which(ForC_simplified$stand.age >= 50), ]
-dist.to.keep <- ForC_simplified$managed %in% 0
+ForC_simplified <- ForC_simplified[which(ForC_simplified$stand.age >= 100), ]
+dist.to.keep <- ForC_simplified$managed %in% 0 & ForC_simplified$disturbed %in% 0
 ForC_simplified <- ForC_simplified[which(dist.to.keep),]
 ForC_simplified$lat <- abs(ForC_simplified$lat)
+
+koeppen_prob <- read.csv("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/raw.data/forest_area_koeppen.csv", stringsAsFactors = F)
+ForC_simplified$weight <- koeppen_prob$prob[match(ForC_simplified$Koeppen, koeppen_prob$koeppen)]
 
 
 ForC_simplified$biomes <- NA
@@ -75,7 +78,7 @@ for (i in seq(along = set1)){
       resp1 <- ForC_simplified[ForC_simplified$variable.name %in% set1[[i]],]
       resp2 <- ForC_simplified[ForC_simplified$variable.name %in% set2[[j]],]
       
-      df <- merge(resp1, resp2[, c("variable.name", "date", "start.date", "end.date", "mean", "citation.ID", "site_plot", "stand.age")], by= c("site_plot", "date"))
+      df <- merge(resp1, resp2[, c("variable.name", "date", "start.date", "end.date", "mean", "citation.ID", "site_plot", "stand.age", "weight")], by= c("site_plot", "date"))
       
       df$ratio <- df$mean.x/df$mean.y
       
@@ -84,13 +87,15 @@ for (i in seq(along = set1)){
       for (fixed.variable in fixed.variables){
         df$fixed <- df[, fixed.variable]
         
-        fixed.no.na <- !is.na(df[, fixed.variable])
+        fixed.no.na <- !is.na(df[, fixed.variable]) & !is.na(df[, "masl"])
         
         df <- df[fixed.no.na, ]
         
[…]

### 2019-06-21 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
 
+*Allocation to carbon fluxes along latitudinal gradients* Variation in allocation to component carbon fluxes along latitudinal gradients was explored for a selection of paired carbon fluxes: GPP:NPP; NPP:ANPP; NPP:BNPP; ANPP:BNPP; ANPP~foliage~:ANPP~woody; and ANPP~foliage~:ANPP~woodystem~.
+Three approaches were taken.
+Firstly, for each set of paired fluxes, measurements taken at the same site and plot, and in the same year, were paired together.
+The ratio of each pair of measurements was calculated.
+For each set of paired fluxes, the ratios were regressed against latitude and climate variables, using the linear model specified above.
+Additionally, sites were classified as "boreal", "temperate", or "tropical", using the FAO ecozone classifications from the ForC database.
+Non-parametric Wilcoxon t-tests were carried out to compare the ratio means between each pair of ecozone classifications.
+In addition, each carbon flux was regressed against latitude, and the models used to generate a series of point estimates along lines of best fit for each flux.
+For each pair of fluxes, the ratio of each point estimate was calculated and plotted, to gain an approximation of the variation of the ratio of the paired fluxes with latitude.
+
 ###Results

### 2019-06-24 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
     fig_caption: yes
-bibliography: library.bib
+bibliography: [library.bib,  packages.bib]
 biblio-style: apalike
@@
 options(tinytex.verbose = TRUE)
+```
+
+```{r include=FALSE}
+knitr::write_bib(c(
+  .packages(), 'knitr', 'rmarkdown', 'lme4', 'MuMIn', 'plyr', 'raster', 'ncdf4', 'Hmisc', 'plyr', 'merTools', 'visreg', 'r2glmm', 'nlme', 'viridis', 'ggplot2', 'ggpubr', 'gridExtra'
+), 'packages.bib')
 ```

### 2019-06-25 scripts/final_model_no_weighting.R
Create final_model_no_weighting.R

@@ -1,0 +1,472 @@
+
+# Clean environment ####
+rm(list = ls())
+
+# Set working directory as ForC main folder ####
+setwd("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/ForC")
+
+# Load libaries ####
+library(lme4)
+library(MuMIn)
+library(plyr)
+library(merTools)
+library(visreg)
+library(r2glmm)
+library(nlme)
+library(viridis)
+
+# Load data ####
+ForC_simplified <- read.csv("ForC_simplified/ForC_simplified_WorldClim_CRU_refined.csv", stringsAsFactors = F)
+
+VARIABLES <- read.csv(paste0(dirname(getwd()), "/ForC/data/ForC_variables.csv"), stringsAsFactors = F)
+
+
+na_codes <- c("NA", "NI", "NRA", "NaN", "NAC") 
+my_is.na <- function(x) { is.na(x) | x %in% na_codes}
+my_na.omit <- function(x) { return(x[!my_is.na(x)])}
+
+
+# Prepare data ####
+
+## change to numerics what it needs to
+ForC_simplified$stand.age <- as.numeric(ForC_simplified$stand.age)
+ForC_simplified$min.db <- as.numeric(ForC_simplified$min.db)
+
+## make plot and geographic areas factors
+ForC_simplified$plot.name <- addNA(ForC_simplified$plot.name)
+ForC_simplified$geographic.area <- addNA(ForC_simplified$geographic.area)
+
+## change sign of NEE 
+ForC_simplified[grepl("NEE", ForC_simplified$variable.name,ignore.case = F),]$mean <- -ForC_simplified[grepl("NEE", ForC_simplified$variable.name,ignore.case = F),]$mean
+
+## take absolute value of latitude
+ForC_simplified$lat <- abs(ForC_simplified$lat)
+
+
+# Control for some factors ####
+
+
+## keep all ages except na, 0 and 999
+ages.not.999.nor.0.nor.na <- !ForC_simplified$stand.age %in% 999 &  !ForC_simplified$stand.age %in% 0 & !is.na(ForC_simplified$stand.age)
+
+## Keep only age >=100 (or 999)
+age.greater.than.100 <- ForC_simplified$stand.age >= 100 & !is.na(ForC_simplified$stand.age)
+age.greater.than.200 <- ForC_simplified$stand.age >= 200 & !is.na(ForC_simplified$stand.age)
+ages <- c("age.greater.than.100")
+
[…]

### 2019-06-25 scripts/best_model_scaled.R
Update best_model_scaled.R

@@ -43,6 +43,8 @@ ForC_simplified[grepl("NEE", ForC_simplified$variable.name,ignore.case = F),]$me
 ## take absolute value of latitude
 ForC_simplified$lat <- abs(ForC_simplified$lat)
 
+ForC_simplified$site_plot <- paste0(ForC_simplified$sites.sitename," ", ForC_simplified$plot.name)
+
 
 # Control for some factors ####
 
@@ -264,10 +266,12 @@ for(response.variables in response.variables.groups){
         legend2 <- paste(response.v, "r-squared = ", Rsq[1], "p-value = ", significance)
         mtext(side = 3, line = -which(response.variables %in% response.v), text = legend2, adj = 0.9, col = response.v.color, cex = 0.5)
         
+        number_plots <- length(unique(df$site_plot))
+        
         date = Sys.Date()
         altitude = TRUE
         
-        results <- data.frame(date.run = date, response = response.v, fixed = fixed.v, altitude_included = altitude, random = "geographic.area/plot.name", Age.filter = age, best.model = best.model, significant = significant.effect, p.value = significance, sample.size = sample.size, Rsq = Rsq, delta.aic = delta.aic, linear.slope = slope)
+        results <- data.frame(date.run = date, response = response.v, fixed = fixed.v, altitude_included = altitude, random = "geographic.area/plot.name", Age.filter = age, best.model = best.model, significant = significant.effect, p.value = significance, sample.size = sample.size, Rsq = Rsq, delta.aic = delta.aic, linear.slope = slope, number_plots = number_plots)
         
         all.results <- rbind(all.results, results)
         all.aictab <- rbind(all.aictab, aictab)

### 2019-06-25 manuscript/manuscript draft.Rmd
updated manuscript draft to include table with sample sizes and number of plots

@@
 
-| Variable | Definition | Components included | Methodologies used|
-|--------------|-------------------|-------------|------------------------|
-| GPP | Annual gross primary production; annual uptake of carbon dioxide by an ecosystem | NA | Flux partitioning of eddy covariance |
-| NPP_1 | Annual net primary production; the component of GPP that is stored in plant tissue; GPP minus ecosystem respiration | Foliage, branch, stem, coarse root and fine root | Direct measurement of annual increments of components |
-| NPP_2 | Annual net primary production; the component of GPP that is stored in plant tissue; GPP minus ecosystem respiration | Foliage, branch, stem, coarse root, fine root, and understory | Direct measurement of annual increments of components |
-| ANPP_1 | Aboveground net primary production | Foliage and stem | Direct measurement of annual increments of components | 
-| ANPP_2 | Aboveground net primary production | Foliage, stem, and branch | Direct measurement of annual increments of components |
-| ANPP_foliage | Net primary production of foliage | Foliage | Direct measurement of litterfall, correcting for changes in leaf biomass when measured |
-| ANPP_woody | Net primary production of woody components | Stems and branches | Direct measurement of stem growth and branch fall |
-| ANPP_woody_stem | Net primary production of woody stems | Woody stems | Direct measurement of stem growth increment |
-| BNPP_root | Belowground net primary production | Coarse and fine roots | Direct measurement of one or more of: fine root turnover, soil cores, root ingrowth cores, minirhizotrons; indirect estimates of coarse roots using allometries based on aboveground stem increment measures |
-| BNPP_root_fine | Net primary production of fine roots | Fine roots | Direct measurement of one or more of: minirhizotrons, fine root turnover, soil cores, root ingrowth cores |
+| Variable | Definition | Components included | Methodologies used | Number of records | Number of plots |
+|--------------|-------------------|-------------|------------------------|-------|-------|
+| GPP | Annual gross primary production; annual uptake of carbon dioxide by an ecosystem | NA | Flux partitioning of eddy covariance | 254 | 91 |
+| NPP_1 | Annual net primary production; the component of GPP that is stored in plant tissue; GPP minus ecosystem respiration | Foliage, branch, stem, coarse root and fine root | Direct measurement of annual increments of components | 93 | 79 |
+| NPP_2 | Annual net primary production; the component of GPP that is stored in plant tissue; GPP minus ecosystem respiration | Foliage, branch, stem, coarse root, fine root, and understory | Direct measurement of annual increments of components | 23 | 23 |
+| ANPP_1 | Aboveground net primary production | Foliage and stem | Direct measurement of annual increments of components | 140 | 131 |
+| ANPP_2 | Aboveground net primary production | Foliage, stem, and branch | Direct measurement of annual increments of components | 119 | 77 |
+| ANPP_foliage | Net primary production of foliage | Foliage | Direct measurement of litterfall, correcting for changes in leaf biomass when measured | 100 | 96 |
+| ANPP_woody | Net primary production of woody components | Stems and branches | Direct measurement of stem growth and branch fall | 53 | 50 |
+| ANPP_woody_stem | Net primary production of woody stems | Woody stems | Direct measurement of stem growth increment | 264 | 252 |
+| BNPP_root | Belowground net primary production | Coarse and fine roots | Direct measurement of one or more of: fine root turnover, soil cores, root ingrowth cores, minirhizotrons; indirect estimates of coarse roots using allometries based on aboveground stem increment measures | 105 | 90 |
+| BNPP_root_fine | Net primary production of fine roots | Fine roots | Direct measurement of one or more of: minirhizotrons, fine root turnover, soil cores, root ingrowth cores | 96 | 88 |
 

### 2019-06-27 manuscript/manuscript draft.Rmd
updated manuscript draft

--- 2019-06-24 manuscript/manuscript draft.Rmd
@@
 Table 1 contains details of the variables selected for analysis.
-For the purposes of analysis, all estimates of NPP_1 and NPP_2 were combined as NPP, and all estimates of ANPP_1 and ANPP_2 were combined as ANPP.
 
-| Variable | Definition | Components included | Methodologies used|
-|--------------|-------------------|-------------|------------------------|
-| GPP | Annual gross primary production; annual uptake of carbon dioxide by an ecosystem | NA | Flux partitioning of eddy covariance |
-| NPP_1 | Annual net primary production; the component of GPP that is stored in plant tissue; GPP minus ecosystem respiration | Foliage, branch, stem, coarse root and fine root | Direct measurement of annual increments of components |
-| NPP_2 | Annual net primary production; the component of GPP that is stored in plant tissue; GPP minus ecosystem respiration | Foliage, branch, stem, coarse root, fine root, and understory | Direct measurement of annual increments of components |
-| ANPP_1 | Aboveground net primary production | Foliage and stem | Direct measurement of annual increments of components | 
-| ANPP_2 | Aboveground net primary production | Foliage, stem, and branch | Direct measurement of annual increments of components |
-| ANPP_foliage | Net primary production of foliage | Foliage | Direct measurement of litterfall, correcting for changes in leaf biomass when measured |
-| ANPP_woody | Net primary production of woody components | Stems and branches | Direct measurement of stem growth and branch fall |
-| ANPP_woody_stem | Net primary production of woody stems | Woody stems | Direct measurement of stem growth increment |
-| BNPP_root | Belowground net primary production | Coarse and fine roots | Direct measurement of one or more of: fine root turnover, soil cores, root ingrowth cores, minirhizotrons; indirect estimates of coarse roots using allometries based on aboveground stem increment measures |
-| BNPP_root_fine | Net primary production of fine roots | Fine roots | Direct measurement of one or more of: minirhizotrons, fine root turnover, soil cores, root ingrowth cores |
+| Variable | Definition | Components included | Methodologies used | Number of records | Number of plots |
+|--------------|-------------------|-------------|------------------------|-------|-------|
+| GPP | Annual gross primary production; annual uptake of carbon dioxide by an ecosystem | NA | Flux partitioning of eddy covariance | 254 | 91 |
+| NPP | Annual net primary production; the component of GPP that is stored in plant tissue; GPP minus ecosystem respiration | Foliage, branch, stem, coarse root, fine root and optionally understory | Direct measurement of annual increments of components | 116 | 85 |
+| ANPP | Aboveground net primary production | Foliage, stem, and optionally branch | Direct measurement of annual increments of components | 259 | 170 |
+| ANPP~foliage~ | Net primary production of foliage | Foliage | Direct measurement of litterfall, correcting for changes in leaf biomass when measured | 100 | 96 |
+| ANPP~woodystem~ | Net primary production of woody stems | Woody stems | Direct measurement of stem growth increment | 264 | 252 |
+| BNPP~root~ | Belowground net primary production | Coarse and fine roots | Direct measurement of one or more of: fine root turnover, soil cores, root ingrowth cores, minirhizotrons; indirect estimates of coarse roots using allometries based on aboveground stem increment measures | 105 | 90 |
+| BNPP~rootfine~ | Net primary production of fine roots | Fine roots | Direct measurement of one or more of: minirhizotrons, fine root turnover, soil cores, root ingrowth cores | 96 | 88 |
+| R~auto~ | Annual autotrophic respiration, including above- and belowground components | Foliage, stem, and root | Chamber measurements of component gas exchange | 22 | 22 |
+| R~autoroot~ | Annual root respiration | Coarse and fine roots | Measurement of root gas exchange | 57 | 38 |
 
@@
 The point estimates for smaller component fluxes were summed to generate new "stacked" estimates of larger fluxes, which were then compared against actual measurements of the larger flux.
-Confidence intervals for the larger flux were calculated using the 'bootMer' function from the lme4 package.
+Confidence intervals for the larger flux were calculated using the 'bootMer' function from the lme4 package [@bates_fitting_2015].
 Stacked plots were generated for:
-1. ANPP_2 + BNPP_root = NPP_1  
-2. ANPP_foliage + ANPP_woody_stem = ANPP_1  
-3. NPP_1 + R_auto = GPP
+1. GPP = NPP + R~auto~  
+2. NPP = ANPP + BNPP  
+3. ANPP = ANPP~foliage~ + ANPP~woodystem~  
+4. Total belowground carbon flux = BNPP + R~root~
 

### 2019-06-28 scripts/final_model_no_weighting.R
Update final_model_no_weighting.R

@@ -116,7 +116,7 @@ all.results <- NULL
 all.aictab <- NULL
 all.koeppen <- NULL
 
-fixed.variables <- c("mat", "map", "lat", "AnnualMeanTemp", "TempSeasonality", "TempRangeAnnual", "AnnualPre", "AnnualFrostDays", "AnnualWetDays", "VapourPressure", "Aridity", "PotentialEvapotranspiration", "VapourPressureDeficit", "SolarRadiation")
+fixed.variables <- c("mat", "map", "lat", "AnnualMeanTemp", "TempSeasonality", "TempRangeAnnual", "AnnualPre", "AnnualFrostDays", "AnnualWetDays", "VapourPressure", "Aridity", "PotentialEvapotranspiration", "VapourPressureDeficit", "SolarRadiation", "PreSeasonality")
 
 
 
@@ -393,6 +393,142 @@ for (age in ages){
 }
 
 
+################################
+
+response.variables.groups <- list(c("GPP", "NPP", "ANPP", "BNPP_root", "R_auto"),
+                                  c("ANPP_foliage", "ANPP_woody_stem", "BNPP_root_fine", "R_auto_root"))
+
+all.results = NULL
+
+
+### mature forests only ####
+for (age in ages){
+  
+  if (age %in% "age.greater.than.100") ages.to.keep <- ForC_simplified$stand.age >= 100 & !is.na(ForC_simplified$stand.age)
+  if (age %in% "age.greater.than.200") ages.to.keep <- ForC_simplified$stand.age >= 200 & !is.na(ForC_simplified$stand.age)
+  
+  for(fixed.v in fixed.variables){
+    
+    print(fixed.v)
+   
+    for(response.variables in response.variables.groups){
+      
+      if(response.variables[1] == "GPP") n <- 1
+      if(response.variables[1] == "ANPP_foliage") n <- 2
+      
+      png(file = paste0("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/figures/final_figures/unweighted_model/Effect_of_", fixed.v, n, "_MATURE_only_poly_all.png"), width = 2255, height = 2000, units = "px", res = 300)
+      
+      par(mfrow = c(1,1), mar = c(3,3,3,3))
+      
+      first.plot <- TRUE
+      
+      for (response.v in response.variables){
+        
[…]

### 2019-06-28 scripts/stacked_plots_code.R
Update stacked_plots_code.R

@@ -731,4 +731,212 @@ for (age in ages){
   }
 }
 
+
+set1 <- c("BNPP_root")
+set2 <- c("R_auto_root")
+sum <- c("BNPP_root_fine")
+
+### mature forests only ####
+for (age in ages){
+  
+  if (age %in% "age.greater.than.100") ages.to.keep <- ForC_simplified$stand.age >= 100 & !is.na(ForC_simplified$stand.age)
+  if (age %in% "age.greater.than.200") ages.to.keep <- ForC_simplified$stand.age >= 200 & !is.na(ForC_simplified$stand.age)
+  
+  for(fixed.v in fixed.variables){
+    
+    # par(mfrow = c(1,1), mar = c(0,0,0,0), oma = c(5,5,2,0))
+    print(fixed.v)
+    
+    ###subset ForC
+    response.variables.col <- 1:3
+    
+    first.plot <- TRUE
+    
+    ################################ for response.v.1
+    
+    for(s in sum){
+      
+      sum.response <- ForC_simplified$variable.name %in% s
+      
+      fixed.no.na <- !is.na(ForC_simplified[, fixed.v]) & !is.na(ForC_simplified[, "masl"])
+      
+      df.sum <- ForC_simplified[sum.response & ages.to.keep & fixed.no.na, ]
+      
+      df.sum$masl <- df.sum$masl/1000
+      
+      df.sum$fixed <- df.sum[, fixed.v]
+      # ylim.1 <- range(ForC_simplified[ForC_simplified$variable.name %in% unlist(response.variables.1),]$mean)
+      
+      mod <-  lmer(mean ~ 1 + (1|geographic.area/plot.name), data = df.sum, REML = F)
+      mod.linear <- lmer(mean ~ poly(fixed, 1) + masl + (1|geographic.area/plot.name), data = df.sum, REML = F)
+      mod.poly <- lmer(mean ~ poly(fixed, 2) + masl + (1|geographic.area/plot.name), data = df.sum, REML = F)
+      
+      aictab <- aictab(list(mod.linear = mod.linear, mod.poly = mod.poly), sort = T)
+      
+      best.model <- as.character(aictab(list(mod = mod, mod.linear = mod.linear), sort = T)$Modname[1])
+      delta.aic <- as.numeric(aictab(list(mod.linear = mod.linear), sort = T)$Delta_AICc[2])
+      delta.aic <- signif(delta.aic, digits=4)
+      
[…]

### 2019-06-28 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
 | R~auto~ | Annual autotrophic respiration, including above- and belowground components | Foliage, stem, and root | Chamber measurements of component gas exchange | 22 | 22 |
-| R~autoroot~ | Annual root respiration | Coarse and fine roots | Measurement of root gas exchange | 57 | 38 |
+| R~root~ | Annual root respiration | Coarse and fine roots | Measurement of root gas exchange | 57 | 38 |
 
@@
 
-```{r echo=FALSE, out.width='100%', fig.cap = "Graph to show primary productivity $(MgC~ha^-1~yr^-1)$ regressed against latitude.
+```{r echo=FALSE, out.width='100%', fig.cap = "Graphs to show primary productivity $(MgC~ha^-1~yr^-1)$ regressed against latitude.
 Lines of best fit are plotted according to the best model selected during analysis.
@@
 
-Latitude is a strong predictor for many of the carbon fluxes, explaining ~70% of variation in GPP (n = 241, p<0.01), 64% in NPP (n = 109, p<0.01) and 60% in ANPP (n = 252, p<0.01).
+Latitude is a strong predictor for many of the carbon fluxes, explaining ~65% of variation in GPP (n = 254, p<0.01), 51% in NPP (n = 116, p<0.01) and 50% in ANPP (n = 259, p<0.01).
 
@@
 All regressions are significant $(p<0.05)$.
-Plots show smaller component fluxes (1) and (2); the larger flux (3) defined by the combination of (1) and (2); and a modelled estimate of the sum of (1) and (2).
-95\\% confidence intervals are plotted for (3).", fig.pos='H'}
+Plots 1 - 3 show two component fluxes; a larger flux, defined as the combination of the two component fluxes; and a modelled estimate of the sum of the two component fluxes.
+95\\% confidence intervals are plotted for the larger flux.
+Plot 4 shows three belowground fluxes, and a modelled estimate of the total belowground carbon flux", fig.pos='H'}
 
@@
 
-```{r echo=FALSE, out.width='100%', fig.cap = "Ratio of ANPP foliage to ANPP woody and ANPP woody stem plotted against major ecozones.
+```{r echo=FALSE, out.width='100%', fig.cap = "Ratio of ANPP foliage to ANPP woody stem plotted against major ecozones.
 All pairwise comparisons are significant.
 Significance levels of pairwise comparisons are indicated where * $p \\leq 0.05 $; ** $p \\leq 0.01 $; *** $p \\leq 0.001 $; **** $p \\leq 0.0001 $.", fig.pos='H'}
-  knitr::include_graphics("boxplot_foliage_woody_combined.png")
+  knitr::include_graphics("boxplot_foliage_woodystem.png")
 ```

### 2019-06-28 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
 *Which climatic variables are the most important drivers behind the latitudinal pattern in primary productivity?* All major carbon fluxes show similar responses to climate variables.
-Of the climate variables tested, annual wet days, aridity, cloud cover, mean diurnal temperature range and precipitation seasonality were poor explainers of variation in primary productivity, explaining less than 30% of the variation in each of the carbon fluxes.
-We found that both measures of mean annual precipitation used were significant (p<0.05) but poor predictors of productivity, explaining at most 45% of variation in carbon flux.
+Of the climate variables tested, annual wet days, aridity, cloud cover, mean diurnal temperature range and precipitation seasonality were poor explainers of variation in primary productivity, explaining less than 20% of the variation in each of the carbon fluxes.
+We found that both measures of mean annual precipitation used were significant (p<0.05) but poor predictors of productivity, explaining, with the exception of R~auto~, at most 37% of variation in carbon flux.
 Despite being a poor predictor, the results indicate that there is a saturation point at between 3000 and 4000mm annual precipitation, above which productivity no longer increases.

### 2019-07-02 scripts/final_model_no_weighting.R
Update final_model_no_weighting.R

@@ -14,6 +14,7 @@ library(visreg)
 library(r2glmm)
 library(nlme)
 library(viridis)
+library(AICcmodavg)
 
 # Load data ####
 ForC_simplified <- read.csv("ForC_simplified/ForC_simplified_WorldClim_CRU_refined.csv", stringsAsFactors = F)
@@ -111,7 +112,7 @@ all.results <- NULL
 
 ###comparing models
 
-library(AICcmodavg)
+
 all.results <- NULL
 all.aictab <- NULL
 all.koeppen <- NULL
@@ -283,13 +284,14 @@ for (age in ages){
     print(fixed.v)
     
     
+    
     for(response.variables in response.variables.groups){
       
       if(response.variables[1] == "GPP") n <- 1
       if(response.variables[1] == "ANPP_foliage") n <- 2
     
-    first.plot <- TRUE
-    
+      first.plot <- TRUE
+      
     for (response.v in response.variables){
       
       col.sym <- read.csv("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/raw.data/colsym.csv", stringsAsFactors = F)
@@ -301,9 +303,6 @@ for (age in ages){
       if(response.v %in% "ANPP") responses.to.keep  <- c("ANPP_1", "ANPP_2")
       if(!response.v %in% c("NPP", "ANPP")) responses.to.keep  <- response.v
       
-      
-      response.v.color <- response.variables.col[which(response.variables %in% response.v)]
-      
       rows.with.response <- ForC_simplified$variable.name %in% responses.to.keep
       
       fixed.no.na <- !is.na(ForC_simplified[, fixed.v]) & !is.na(ForC_simplified[, "masl"])
@@ -314,10 +313,19 @@ for (age in ages){
       
       df$fixed <- df[, fixed.v]
       
+      if(response.v == "GPP"){
       a <- ForC_simplified[ForC_simplified$variable.name %in% unlist(response.variables),]
       ylim <- range(tapply(a$mean, a$variable.name, scale))
       ylim[1] <- ylim[1] - 0.25
       ylim[2] <- ylim[2] + 0.25
+      }
+      
+      if(response.v == "GPP"){
+        b <- ForC_simplified[ForC_simplified$variable.name %in% unlist(response.variables.groups),]
[…]

### 2019-07-02 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@
 knitr::write_bib(c(
-  .packages(), 'knitr', 'rmarkdown', 'lme4', 'MuMIn', 'plyr', 'raster', 'ncdf4', 'Hmisc', 'plyr', 'merTools', 'visreg', 'r2glmm', 'nlme', 'viridis', 'ggplot2', 'ggpubr', 'gridExtra'
+  .packages(), 'knitr', 'rmarkdown', 'lme4', 'MuMIn', 'plyr', 'raster', 'ncdf4', 'Hmisc', 'plyr', 'merTools', 'visreg', 'r2glmm', 'nlme', 'viridis', 'ggplot2', 'ggpubr', 'gridExtra', 'AICcmodavg'
 ), 'packages.bib')

### 2019-07-02 scripts/stacked_plots_code.R
Update stacked_plots_code.R

@@ -272,14 +272,14 @@ par(mfrow = c(2,2), mar = c(1,1,1,1), oma = c(3,3,0,0))
           
           # png(file = paste0("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/figures/final_figures/stacked_plots/", set1[[i]], "_to_", set2[[j]],"_", fixed.v, "_stacked.png"), width = 2255, height = 2000, units = "px", res = 300)
           
-          plot(mean ~ fixed, data = df.1, xlab = "", ylab = "", yaxt = "n", ylim = ylim, col = plasma(10)[2], pch = 2)
-          points(mean ~ fixed, data = df.2, ylab = "", col = plasma(10)[8], pch = 8)
+          plot(mean ~ fixed, data = df.1, xlab = "", ylab = "", yaxt = "n", ylim = ylim, col = plasma(10)[3], pch = 3)
+          points(mean ~ fixed, data = df.2, ylab = "", col = plasma(10)[4], pch = 4)
           points(mean ~ fixed, data = df.sum, ylab = "", col = plasma(10)[1], pch = 1)
           
           for(masl in unique(newDat$masl)){
             i <- which(unique(newDat$masl) %in% masl)
-            lines(fit.1 ~ fixed, data = newDat[newDat$masl %in% masl,], lty = ifelse(significant.effect, 1, 2), lwd = i, col = plasma(10)[2])
-            lines(fit.2 ~ fixed, data = newDat[newDat$masl %in% masl,], lty = ifelse(significant.effect, 1, 2), lwd = i, col = plasma(10)[8])
+            lines(fit.1 ~ fixed, data = newDat[newDat$masl %in% masl,], lty = ifelse(significant.effect, 1, 2), lwd = i, col = plasma(10)[3])
+            lines(fit.2 ~ fixed, data = newDat[newDat$masl %in% masl,], lty = ifelse(significant.effect, 1, 2), lwd = i, col = plasma(10)[4])
             lines(fit.3 ~ fixed, data = newDat[newDat$masl %in% masl,], lty = ifelse(significant.effect, 1, 2), lwd = i, col = plasma(10)[1])
             lines(stacked_plot ~ fixed, data = newDat[newDat$masl %in% masl,], lty = ifelse(significant.effect, 1, 2), lwd = i, col = "black")
[…]

### 2019-07-08 scripts/final_model_no_weighting.R
Update final_model_no_weighting.R

@@ -382,7 +382,8 @@ for (age in ages){
     
     if(n == 1) title(paste("Major fluxes"), outer = F, line = 1)
     if(n == 2) title(paste("Subsidiary fluxes"), outer = F, line = 1)
-    mtext(side = 1, line = 2, text = fixed.v, outer = T)
+    xaxis <- ifelse(fixed.v == "lat", "Latitude", paste(fixed.v))
+    mtext(side = 1, line = 2, text = xaxis, outer = T)
     mtext(side = 2, line = 1,  text = expression("Productivity (scaled values)"), outer = T) 
     
     

### 2019-07-08 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
 Data from the temperate regions was heavily skewed towards studies from the old-growth forests of the Pacific Northwest.
-These forests have very high productivity, and so to reduce any bias from over-sampling of this region, the proportion of global forest cover contributed by each Koeppen climate zone was calculated, and the models weighted according these proportions.
+These forests have very high productivity, and so to ensure that results were not unduly influenced by geographic sampling bias, we tried a version of the model where data were weighted according to forested land area within each Koeppen climate zone.
+Results were similar between the weighted and unweighted model, so, to avoid problems of over-fitting, the weighted model was dropped, and results from this are not presented here.
 
-Two models were specified: one with the climate variable as a linear term, and one with the climate variable as a polynomial term.
+For each climate variable, two models were specified: one with the climate variable as a linear term, and one with the climate variable as a polynomial term.
 AIC values were calculated for the models and used to select the best model.
@@
 In addition, slope values were calculated for the linear models.
+
+Because the magnitude of fluxes varies significantly, in order to facilitate comparisons between regression models for each flux, data for each flux was scaled, to give the data a mean of 0 and standard deviation of 1.
+As each data set was scaled separately, this does not allow for statistical comparisons of slope values, but does assist in visualising the data.
 
@@
 
-*Allocation to carbon fluxes along latitudinal gradients* Variation in allocation to component carbon fluxes along latitudinal gradients was explored for a selection of paired carbon fluxes: GPP:NPP; NPP:ANPP; NPP:BNPP; ANPP:BNPP; ANPP~foliage~:ANPP~woody; and ANPP~foliage~:ANPP~woodystem~.
+*Allocation to carbon fluxes along latitudinal gradients* Variation in allocation to component carbon fluxes along latitudinal gradients was explored for a selection of paired carbon fluxes: GPP:NPP; NPP:ANPP; NPP:BNPP; ANPP:BNPP; ANPP~foliage~:ANPP~woody~; and ANPP~foliage~:ANPP~woodystem~.
 Three approaches were taken.
@@
 
-Latitude is a strong predictor for many of the carbon fluxes, explaining ~65% of variation in GPP (n = 254, p<0.01), 51% in NPP (n = 116, p<0.01) and 50% in ANPP (n = 259, p<0.01).
+Latitude is a strong predictor for many of the carbon fluxes, explaining 66% of variation in GPP (n = 254, p<0.0001), 51% in NPP (n = 116, p<0.0001) and 50% in ANPP (n = 259, p<0.0001).
+For all fluxes, their relationship with latitude was best predicted by the linear model.
 
@@
 3).
-Modelled estimates of NPP, generated from the sum of ANPP and BNPP~root~, and ANPP, generated from the sum of ANPP~foliage~ and ANPP~woodystem~, fall completely within the confidence intervals of the regressions of field estimates of NPP and ANPP.
+Modelled estimates of GPP, generated from the sum of NPP and R auto; NPP, generated from the sum of ANPP and BNPP~root~; and ANPP, generated from the sum of ANPP~foliage~ and ANPP~woodystem~, fall completely within the confidence intervals of the regressions of field estimates of GPP, NPP and ANPP respectively.
 
@@
 We found that both measures of mean annual precipitation used were significant (p<0.05) but poor predictors of productivity, explaining, with the exception of R~auto~, at most 37% of variation in carbon flux.
-Despite being a poor predictor, the results indicate that there is a saturation point at between 3000 and 4000mm annual precipitation, above which productivity no longer increases.
-Looking across all productivity components, the climate variables which explain the highest amount of variation in carbon fluxes are temperature seasonality, mean annual temperature, annual temperature range, and vapour pressure.
-For measures of woody productivity (ANPP~woody~ and ANPP~woodystem~), potential evapotranspiration and vapour pressure deficit were also strong predictors.
+
+Looking across all productivity components, the climate variables which explain the highest amount of variation in carbon fluxes are temperature seasonality, mean annual temperature, annual temperature range, and annual frost days.
+Vapour pressure deficit came out as a strong predictor for ANPP~woodystem~, and potential evapotranspiration was a good predictor for BNPP~root~ and ANPP~woodystem~.
 
@@
 **Mean annual temperature.** All carbon fluxes show a linear increase with temperature.  
-**Temperature seasonality.** Productivity is highest where temperature seasonality = 0.
-GPP, NPP, ANPP, and BNPP~root~ show a polynomial relationship with temperature seasonality.
+**Temperature seasonality and annual temperature range.** These two climate variables are strongly correlated, and so only temperature seasonality is presented in figure 4.
+Productivity is highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower.
+GPP, NPP, ANPP, and R~root~ show a polynomial relationship with both variables, and BNPP~root~ shows a polynomial relationship with annual temperature range.
 As seasonality increases, these variables show a more rapid decrease in productivity.
-ANPP~foliage~ and ANPP~woodystem~ show a linear relationship with temperature seasonality.
-**Annual temperature range.** As with temperature seasonality, productivity is highest at annual temperature ranges of 15$^\circ$C or lower.
-Again, GPP, NPP, ANPP, and BNPP~root~ show a polynomial relationship with temperature seasonality, and ANPP~foliage~ and ANPP~woodystem~ show a linear relationship with temperature seasonality.
-**Vapour pressure.** Productivity increases linearly with vapour pressure for all carbon fluxes, with the exception of ANPP~foliage~, which saturates at around 0.24.
+ANPP~foliage~ and ANPP~woodystem~ show a linear relationship with both variables.
+**Mean annual precipitation.** Despite being a poor predictor of productivity, the results indicate that, for the majority of fluxes, productivity increases with precipitation, up until a saturation point at between 3000 and 4000mm annual precipitation, above which productivity starts to decrease.
+The notable exception to this is GPP, which shows a significant (p<0.0001) linear increase with precipitation up to measures of nearly 6000mm annually, although the explanatory power of this model was low (R-squared = 0.15).
+**Potential evapotranspiration.** All fluxes, with the exception of BNPP~rootfine~ and R~root~ show a strong increase with potential evapotranspiration at values below 1000.
+Between 1200 and 1700 productivity peaks, and begins to decrease at values above 1800.
 

### 2019-07-09 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
 
-*Allocation to carbon fluxes along latitudinal gradients* Variation in allocation to component carbon fluxes along latitudinal gradients was explored for a selection of paired carbon fluxes: GPP:NPP; NPP:ANPP; NPP:BNPP; ANPP:BNPP; ANPP~foliage~:ANPP~woody~; and ANPP~foliage~:ANPP~woodystem~.
+*Allocation to carbon fluxes along latitudinal gradients* Variation in allocation to component carbon fluxes along latitudinal gradients was explored for three sets of paired carbon fluxes: GPP:NPP; ANPP:BNPP; and ANPP~foliage~:ANPP~woodystem~.
 Three approaches were taken.
@@
 The notable exception to this is GPP, which shows a significant (p<0.0001) linear increase with precipitation up to measures of nearly 6000mm annually, although the explanatory power of this model was low (R-squared = 0.15).
-**Potential evapotranspiration.** All fluxes, with the exception of BNPP~rootfine~ and R~root~ show a strong increase with potential evapotranspiration at values below 1000.
-Between 1200 and 1700 productivity peaks, and begins to decrease at values above 1800.
+**Potential evapotranspiration.** All fluxes, with the exception of BNPP~rootfine~ and R~root~ show a strong increase with potential evapotranspiration at values below 1000mm.
+Between 1200 and 1700mm productivity peaks, and begins to decrease at values above 1800mm.
 
@@
 
+| Variable | Latitude | Mean annual temperature | Temperature seasonality | Potential evapotranspiration |
+|----------|----------|----------|----------|----------|
+| R~auto~ | 0.64 | 0.60 | 0.54 | 0.63 |
+| GPP | 0.66 | 0.59 | 0.75 | 0.40 |
+| NPP | 0.51 | 0.50 | 0.52 | 0.34 |
+| ANPP | 0.50 | 0.38 | 0.47 | 0.32 |
+| ANPP~foliage~ | 0.55 | 0.60 | 0.54 | 0.40 |
+| ANPP~woodystem~ | 0.25 | 0.22 | 0.20 | 0.26 |
+| BNPP~root~ | 0.31 | 0.32 | 0.31 | 0.32 |
+| BNPP~rootfine~ | 0.14 | 0.16 | 0.16 | 0.11 |
+| R~root~ | 0.25 | 0.16 | 0.29 | 0.04* |
+
+Table: R-squared values presented for four climate variables. Values marked * are not significant.
+
 ###Discussion

### 2019-07-10 scripts/boxplot_comparisons.R
Update boxplot_comparisons.R

@@ -62,80 +62,10 @@ for (i in seq(along = set1)){
     }
   }
 
-set1 <- c("ANPP_foliage", "ANPP_foliage")
-set2 <- c("ANPP_woody", "ANPP_woody_stem")
-
-
-for (i in seq(along = set1)){
-  for (j in seq(along = set2)){
-    if (i == j){
-      resp1 <- ForC_simplified[ForC_simplified$variable.name %in% set1[[i]],]
-      resp2 <- ForC_simplified[ForC_simplified$variable.name %in% set2[[j]],]
-      
-      df <- merge(resp1, resp2[, c("variable.name", "date", "start.date", "end.date", "mean", "citation.ID", "site_plot", "stand.age", "weight")], by= c("site_plot", "date"))
-      
-      df$ratio <- df$mean.x/df$mean.y
-      
-      range <- quantile(df$ratio, 0.99)
-      
-      for (fixed.variable in fixed.variables){
-        df$fixed <- df[, fixed.variable]
-        
-        fixed.no.na <- !is.na(df[, fixed.variable]) & !is.na(df[, "masl"])
-        
-        df <- df[fixed.no.na, ]
-        
-        df$masl <- df$masl/1000
-        
-        mod <-  lmer(ratio ~ 1 + (1|geographic.area/plot.name), data = df, REML = F, weights = weight.x)
-        
-        mod.full <- lmer(ratio ~ fixed + masl + (1|geographic.area/plot.name), data = df, REML = F, weights = weight.x)
-        
-        significant.effect <- anova(mod, mod.full)$"Pr(>Chisq)"[2] < 0.05
-        
-        significance <- anova(mod, mod.full)$"Pr(>Chisq)"[2]
-        significance <- signif(significance, digits=4)
-        legend1 <- paste0("p-value = ", significance)
-        
-        Rsq <- as.data.frame(r.squaredGLMM(mod.full))
-        Rsq <- signif(Rsq, digits=4)
-        legend2 <- paste0("r-squared = ", Rsq[1])
-        
-        sample.size <- length(df$ratio)
-        
-        png(file = paste0("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/figures/final_figures/foliage_woody_ratio/", set1[[i]], "_", set2[[j]], "_", fixed.variable, ".png"), width = 2255, height = 2000, units = "px", res = 300)
-        
[…]

### 2019-07-10 scripts/ratio_regression_script.R
Update ratio_regression_script.R

@@ -61,6 +61,8 @@ ages.to.keep <- ForC_simplified$stand.age >= 100 & !is.na(ForC_simplified$stand.
 dist.to.keep <- ForC_simplified$managed %in% 0 & ForC_simplified$disturbed %in% 0
 ForC_simplified <- ForC_simplified[dist.to.keep, ]
 
+# alt.to.keep <- ForC_simplified$masl <= 2000 & !is.na(ForC_simplified$masl)
+# ForC_simplified <- ForC_simplified[alt.to.keep, ]
 
 ## give leaf type
 broadleaf_codes <- c("2TEB", "2TDB", "2TB")
@@ -99,6 +101,9 @@ fixed.variables <- c("mat", "map", "lat", "AnnualMeanTemp", "TempSeasonality", "
 set1 <- c("GPP", "ANPP_1", "ANPP_foliage")
 set2 <- c("NPP_1", "BNPP_root", "ANPP_woody_stem")
 
+# set1 <- "ANPP_foliage"
+# set2 <- "ANPP_woody_stem"
+
 
 for (i in seq(along = set1)){
   # for (j in seq(along = set2)){
@@ -126,53 +131,70 @@ for (i in seq(along = set1)){
         
         ##remove extreme outliers 
         
-        # outliers <- boxplot(df$ratio, plot=FALSE, range = 6)$out
-        
+        # outliers <- boxplot(df$ratio, plot=FALSE, range = 3)$out
+        # 
         # if(length(outliers > 0)) df <- df[-which(df$ratio %in% outliers),]
         
         
         mod <-  lmer(ratio ~ 1 + (1|geographic.area/plot.name), data = df, REML = F)
         mod.linear <- lmer(ratio ~ poly(fixed, 1, raw = T) + masl + (1|geographic.area/plot.name), data = df, REML = F)
-        mod.poly <- lmer(ratio ~ poly(fixed, 2, raw = T) + masl + (1|geographic.area/plot.name), data = df, REML = F)
         
-        aictab <- aictab(list(mod.linear = mod.linear, mod.poly = mod.poly), sort = T)
+        best.model <- as.character(aictab(list(mod = mod, mod.linear = mod.linear), sort = T)$Modname[1])
         
-        best.model <- as.character(aictab(list(mod = mod, mod.linear = mod.linear, mod.poly = mod.poly), sort = T)$Modname[1])
-        delta.aic <- as.numeric(aictab(list(mod.linear = mod.linear, mod.poly = mod.poly), sort = T)$Delta_AICc[2])
[…]

### 2019-07-10 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
 |----------|----------|----------|----------|----------|
-| R~auto~ | 0.64 | 0.60 | 0.54 | 0.63 |
-| GPP | 0.66 | 0.59 | 0.75 | 0.40 |
-| NPP | 0.51 | 0.50 | 0.52 | 0.34 |
-| ANPP | 0.50 | 0.38 | 0.47 | 0.32 |
-| ANPP~foliage~ | 0.55 | 0.60 | 0.54 | 0.40 |
-| ANPP~woodystem~ | 0.25 | 0.22 | 0.20 | 0.26 |
-| BNPP~root~ | 0.31 | 0.32 | 0.31 | 0.32 |
-| BNPP~rootfine~ | 0.14 | 0.16 | 0.16 | 0.11 |
-| R~root~ | 0.25 | 0.16 | 0.29 | 0.04* |
+| R~auto~ | 0.64 | 0.83 | 0.54 | 0.63 |
+| GPP | 0.66 | 0.61 | 0.75 | 0.40 |
+| NPP | 0.51 | 0.49 | 0.52 | 0.34 |
+| ANPP | 0.50 | 0.45 | 0.47 | 0.32 |
+| ANPP~foliage~ | 0.55 | 0.63 | 0.54 | 0.40 |
+| ANPP~woodystem~ | 0.25 | 0.23 | 0.20 | 0.26 |
+| BNPP~root~ | 0.31 | 0.26 | 0.31 | 0.32 |
+| BNPP~rootfine~ | 0.14 | 0.11 | 0.16 | 0.11 |
+| R~root~ | 0.25 | 0.21 | 0.29 | 0.04* |
 
 Table: R-squared values presented for four climate variables. Values marked * are not significant.
+
+R-squared values are generally highest in the major fluxes, and decrease in subsidiary fluxes.
+Of the major fluxes, R~auto~ and GPP are the most strongly explained by latitude and climate.
+Mean annual temperature explains 83% of variation in R~auto~, while temperature seasonality explains 75% of variation in GPP.
+The proportion of variation explained by climate and latitude decreases in NPP and ANPP.
+The climatic variables with the strongest explanatory power explain around 50% of the variation in these fluxes.
+Of the major fluxes, BNPP~root~ is the least well explained by climate and latitude, with climate explaining at most 32% of variation.
+
+With the exception of ANPP~foliage~, the proportion of variation explained by climate and latitude in subsidiary fluxes is much lower.
+Climate explains at most 26% of variation in ANPP~woodystem~, 17% in BNPP~rootfine~, and 32% in R~root~.
+In contrast, climate strongly explains variation in ANPP~foliage~, with mean annual temperature explaining 63% of variation.
 

### 2019-07-12 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
 
-*Allocation to carbon fluxes along latitudinal gradients* Variation in allocation to component carbon fluxes along latitudinal gradients was explored for three sets of paired carbon fluxes: GPP:NPP; ANPP:BNPP; and ANPP~foliage~:ANPP~woodystem~.
-Three approaches were taken.
-Firstly, for each set of paired fluxes, measurements taken at the same site and plot, and in the same year, were paired together.
-The ratio of each pair of measurements was calculated.
-For each set of paired fluxes, the ratios were regressed against latitude and climate variables, using the linear model specified above.
-Additionally, sites were classified as "boreal", "temperate", or "tropical", using the FAO ecozone classifications from the ForC database.
-Non-parametric Wilcoxon t-tests were carried out to compare the ratio means between each pair of ecozone classifications.
-In addition, each carbon flux was regressed against latitude, and the models used to generate a series of point estimates along lines of best fit for each flux.
-For each pair of fluxes, the ratio of each point estimate was calculated and plotted, to gain an approximation of the variation of the ratio of the paired fluxes with latitude.
+*Allocation to carbon fluxes along latitudinal gradients.* Variation in allocation to component carbon fluxes along latitudinal gradients was explored for a range of pairings: firstly, GPP:NPP, ANPP:BNPP, and ANPP~foliage~:ANPP~woodystem~; and secondly, the ratio of NPP to each of ANPP, BNPP, ANPP~foliage~, and ANPP~woodystem~.
+For each set of paired fluxes, measurements taken at the same site and plot, and in the same year, were paired together, and the ratio of each pair of measurements calculated.
+The ratios were regressed against latitude and climate variables, using the linear model specified above.
+Cook's distance analyses were carried out for each of the models, and indicated that data from a few high-elevation sites were having a disproportionate influence on the regressions.
+To account for this, models were re-run using only data from sites $\leq$ 1000m.
 
@@
 
-*Do carbon fluxes vary in their responses to climate, particularly with regards to allocation between fluxes?*
-
-```{r echo=FALSE, out.width='100%', fig.cap = "Ratio of ANPP foliage to ANPP woody stem plotted against major ecozones.
-All pairwise comparisons are significant.
-Significance levels of pairwise comparisons are indicated where * $p \\leq 0.05 $; ** $p \\leq 0.01 $; *** $p \\leq 0.001 $; **** $p \\leq 0.0001 $.", fig.pos='H'}
-  knitr::include_graphics("boxplot_foliage_woodystem.png")
-```
+*Do carbon fluxes vary in their responses to climate, particularly with regards to allocation between fluxes?* Of all the flux pairings tested, the only ones to show a significant relationship with latitude were ANPP:NPP (p<0.01, r^2^ = 0.44) and ANPP:BNPP (p<0.001, r^2^ = 0.33).
+In both cases, allocation to ANPP increased with latitude, with the lowest allocation to ANPP in the tropical regions.
 

### 2019-07-15 scripts/PCA_analysis_code.R
Update PCA_analysis_code.R

@@ -15,6 +15,7 @@ setwd("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/ForC")
 library(lme4)
 library(devtools)
 library(MuMIn)
+library(factoextra)
 
 # Load data ####
 ForC_simplified <- read.csv("ForC_simplified/ForC_simplified_WorldClim_CRU_refined.csv", stringsAsFactors = F)
@@ -47,7 +48,8 @@ ForC_simplified$lat <- abs(ForC_simplified$lat)
 
 ## keep only stands that are NOT too strongly influence by management/ disturbance
 
-dist.to.keep <- ForC_simplified$managed %in% 0 & ForC_simplified$disturbed %in% 0
+dist.to.keep <- ForC_simplified$managed %in% 0 & ForC_simplified$disturbed %in% 0 
+masl.to.keep <- !is.na(ForC_simplified[, "masl"])
 
 ## keep only records with min.dbh <= 10cm
 min.dbh.to.keep <- ForC_simplified$min.dbh <= 10 ##& !is.na(ForC_simplified$min.dbh) # this doesn work great, not enough data, but maybe we can just keep the NAs?
@@ -74,12 +76,14 @@ ForC_simplified$variable.name <- gsub("(_OM|_C)", "", ForC_simplified$variable.n
 ForC_simplified$variable.name <- gsub("(_0|_1|_2)", "", ForC_simplified$variable.name)
 
 
-ForC_simplified <- ForC_simplified[dist.to.keep & min.dbh.to.keep, ]
+ForC_simplified <- ForC_simplified[dist.to.keep & min.dbh.to.keep & masl.to.keep, ]
+
+ForC_simplified$masl <- ForC_simplified$masl/1000
 ForC_simplified <- ForC_simplified[ForC_simplified$stand.age >= 100 & !is.na(ForC_simplified$stand.age),]
 ForC_simplified <- ForC_simplified[!ForC_simplified$sites.sitename %in% "Paracou B", ]
 ForC_simplified$leaf.type.phenology <- NA
 ForC_simplified$leaf.type.phenology <- paste(ForC_simplified$leaf.type, ForC_simplified$leaf.phenology, sep = "_")
-ForC_simplified <- ForC_simplified[ForC_simplified$variable.name %in% c("GPP", "NPP", "ANPP", "ANPP_foliage", "ANPP_woody", "BNPP_root"),]
[…]

### 2019-07-15 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
 
-*What is primary productivity?* During photosynthesis, plants capture carbon dioxide from the atmosphere.
+*The components of forest* During photosynthesis, plants capture carbon dioxide from the atmosphere.
 The gross primary productivity (GPP) of an ecosystem is the gross uptake, via photosynthesis, of carbon dioxide by plants in that ecosystem.
-Only a fraction of the carbon captured is assimilated into plant tissue; the rest is used in autotrophic respiration (R~a~).
+Only a fraction of the carbon captured is assimilated into plant tissue; the rest is used in autotrophic respiration (R~auto~).
 The component of GPP that is stored as plant material is the net primary productivity (NPP) of an ecosystem.
@@
 
-$$ NPP = GPP + R~a~ $$
+$$ NPP = GPP + R~auto~ $$
 
@@
 | WorldClim | Mean annual temperature; temperature seasonality; annual temperature range; mean annual precipitation | [@hijmans_very_2005] |
-| WorldClim2 | Vapour pressure; solar radiation | [@fick_worldclim_2017] |
+| WorldClim2 | Solar radiation | [@fick_worldclim_2017] |
 | Climate Research Unit (CRU) time-series dataset v 4.03 | Cloud cover; annual frost days; annual wet days; potential evapotranspiration | [@harris_updated_2014] |
@@
 In both cases, allocation to ANPP increased with latitude, with the lowest allocation to ANPP in the tropical regions.
+BNPP:NPP decreased with latitude, though this relationship was non-significant.
 

### 2019-07-17 scripts/Code_for_interactions.R
Update Code_for_interactions.R

@@ -187,8 +187,11 @@ for(response.variables in response.variables.groups){
         if (best.model == "mod.add") mod.full <- lmer(mean ~ fixed1 + fixed2 + masl + (1|geographic.area/plot.name), data = df, REML = T)
         
                
-        r <- round(fixef(mod.full), 2)
+        r <- round(fixef(mod.full), 4)
         equation <-  paste(response.v, "=", r[1], "+", fixed.v1,  "x", r[2])
+        fixed1.coef <- ifelse(best.model %in% c("mod.int", "mod.add"), r[2], NA)
+        fixed2.coef <- ifelse(best.model %in% c("mod.int", "mod.add"), r[3], NA)
+        int.coef <- ifelse(best.model %in% "mod.int", r[5], NA)
        
         significance <- anova(mod, mod.full)$"Pr(>Chisq)"[2]
         significance <- signif(significance, digits=4)
@@ -196,7 +199,7 @@ for(response.variables in response.variables.groups){
         Rsq <- as.data.frame(r.squaredGLMM(mod.full))
         Rsq <- signif(Rsq, digits=4)
         
-        results <- data.frame(response = response.v, fixed1 = fixed.v1, fixed2 = fixed.v2, random = "geographic.area/plot.name", Age.filter = age, best.model = best.model, significant = significant.effect, p.value = significance, sample.size = sample.size, Rsq = Rsq, delta.aic = delta.aic)
+        results <- data.frame(response = response.v, fixed1 = fixed.v1, fixed2 = fixed.v2, random = "geographic.area/plot.name", Age.filter = age, best.model = best.model, significant = significant.effect, p.value = significance, sample.size = sample.size, Rsq = Rsq, delta.aic = delta.aic, fixed1.coef = fixed1.coef, fixed2.coef = fixed2.coef, int.coef = int.coef)
         
         all.results <- rbind(all.results, results)
         

### 2019-07-17 scripts/Code_for_interactions.R
Update Code_for_interactions.R

@@ -160,18 +160,18 @@ for(response.variables in response.variables.groups){
         ylim <- range(ForC_simplified[ForC_simplified$variable.name %in% unlist(response.variables),]$mean)
         
         mod <-  lmer(mean ~ 1 + (1|geographic.area/plot.name), data = df, REML = F)
-        mod.linear <- lmer(mean ~ poly(fixed1, 1) + masl + (1|geographic.area/plot.name), data = df, REML = F)
-        mod.poly <- lmer(mean ~ poly(fixed1, 2) + masl + (1|geographic.area/plot.name), data = df, REML = F)
+        # mod.linear <- lmer(mean ~ poly(fixed1, 1) + masl + (1|geographic.area/plot.name), data = df, REML = F)
+        # mod.poly <- lmer(mean ~ poly(fixed1, 2) + masl + (1|geographic.area/plot.name), data = df, REML = F)
         mod.int <- lmer(mean ~ fixed1*fixed2 + masl + (1|geographic.area/plot.name), data = df, REML = F)
         mod.add <- lmer(mean ~ fixed1 + fixed2 + masl + (1|geographic.area/plot.name), data = df, REML = F)
         
         
-        best.model <- as.character(aictab(list(mod = mod, mod.linear = mod.linear, mod.poly = mod.poly, mod.int = mod.int, mod.add = mod.add), sort = T)$Modname[1])
-        delta.aic <- as.numeric(aictab(list(mod.linear = mod.linear, mod.poly = mod.poly, mod.int = mod.int, mod.add = mod.add), sort = T)$Delta_AICc[2])
+        best.model <- as.character(aictab(list(mod = mod, mod.int = mod.int, mod.add = mod.add), sort = T)$Modname[1])
+        delta.aic <- as.numeric(aictab(list(mod.int = mod.int, mod.add = mod.add), sort = T)$Delta_AICc[2])
         delta.aic <- signif(delta.aic, digits=4)
         
-        if (best.model == "mod.poly") mod.full <- lmer(mean ~ poly(fixed1, 2) + masl + (1|geographic.area/plot.name), data = df, REML = F)
-        if (best.model == "mod.linear") mod.full <- lmer(mean ~ poly(fixed1, 1) + masl + (1|geographic.area/plot.name), data = df, REML = F)
+        # if (best.model == "mod.poly") mod.full <- lmer(mean ~ poly(fixed1, 2) + masl + (1|geographic.area/plot.name), data = df, REML = F)
[…]

### 2019-07-17 scripts/Code_for_interactions.R
Update Code_for_interactions.R

@@ -99,8 +99,8 @@ library(AICcmodavg)
 all.results <- NULL
 all.aictab <- NULL
 
-fixed.variables.set1 <- c("mat", "map", "TempSeasonality", "AnnualWetDays", "SolarRadiation", "Aridity", "PotentialEvapotranspiration", "VapourPressureDeficit")
-fixed.variables.set2 <- c("mat", "map", "TempSeasonality", "AnnualWetDays", "SolarRadiation", "Aridity", "PotentialEvapotranspiration", "VapourPressureDeficit")
+fixed.variables.set1 <- c("mat", "map", "TempSeasonality", "SolarRadiation", "PotentialEvapotranspiration", "VapourPressureDeficit", "PreSeasonality")
+fixed.variables.set2 <- c("mat", "map", "TempSeasonality", "SolarRadiation", "PotentialEvapotranspiration", "VapourPressureDeficit", "PreSeasonality")
 
 
 response.variables.groups <- list(c("GPP", "NPP", "BNPP_root", "BNPP_root_fine"),

### 2019-07-17 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
 Site altitude was also included as a fixed effect.
-Random effect was stand nested within area.
+Within the ForC database, sites within 25km^2^ of each other are clustered into geographic areas.
+To account for correlations in measurements between tightly clustered sites, a random effect was specified as plot nested within geographic area.
 Data from the temperate regions was heavily skewed towards studies from the old-growth forests of the Pacific Northwest.
@@
 
-For each climate variable, two models were specified: one with the climate variable as a linear term, and one with the climate variable as a polynomial term.
+For each climate variable, three models were specified: a null model; a model with the climate variable as a linear term; and a model with the climate variable as a polynomial term.
 AIC values were calculated for the models and used to select the best model.
 R-squared values were calculated for the best model.
-In addition, slope values were calculated for the linear models.
+In addition, slope coefficients were calculated for the linear models.
 
@@
 As each data set was scaled separately, this does not allow for statistical comparisons of slope values, but does assist in visualising the data.
+
+To investigate the potential interactive effects of climate variables on carbon fluxes, multivariate models were also specified.
+Climate variables which were non-significant or poor predictors in the models with a single climate term, or which were strongly correlated with another climate term, were dropped, leaving a subset of 8 climate variables.
+For each possible pairing of climate variables, three models were specified: a null model; a model with the two climate variables showing an additive effect; and a model with the two climate variables showing an interactive effect.
+As before, altitude was included as a fixed effect, and plot nested within geographic area as a random effect.
+AIC values were calculated for the models and used to select the best model.
+R-squared values were calculated for the best model.
+Slope coefficients were calculated for each of the terms, and for the interactive term if relevant.
 

### 2019-07-17 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
 
+```{r eval = TRUE, echo=FALSE, warning=FALSE}
+library(knitr)
+library(kableExtra)
+multivariate_models <- read.csv("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/multivariate_models.csv", stringsAsFactors = FALSE, check.names = FALSE)
+
+kable(multivariate_models, format = "latex", booktabs = TRUE, caption = "Multivariate model summary", escape = FALSE) %>%
+  kable_styling(latex_options = c("scale_down", "hold_position")) %>%
+  column_spec(1:10, width = "2cm") %>%
+  add_header_above(c(" ", " ", "Climate variables" = 7, " "))
+
+```
+
 *Do carbon fluxes vary in their responses to climate, particularly with regards to allocation between fluxes?* Of all the flux pairings tested, the only ones to show a significant relationship with latitude were ANPP:NPP (p<0.01, r^2^ = 0.44) and ANPP:BNPP (p<0.001, r^2^ = 0.33).

### 2019-07-17 manuscript/manuscript draft.Rmd
updated manuscript draft

--- 2019-07-17 manuscript/manuscript draft.Rmd
@@
 **Mean annual precipitation.** Despite being a poor predictor of productivity, the results indicate that, for the majority of fluxes, productivity increases with precipitation, up until a saturation point at between 3000 and 4000mm annual precipitation, above which productivity starts to decrease.
+The notable exception to this is GPP, which shows a significant (p<0.0001) linear increase with precipitation up to measures of nearly 6000mm annually, although the explanatory power of this model was low (r-squared = 0.15).
+**Potential evapotranspiration.** All fluxes, with the exception of BNPP~rootfine~ and R~root~ show a strong increase with potential evapotranspiration at values below 1000mm.
+Between 1200 and 1700mm productivity peaks, and begins to decrease at values above 1800mm.
+
+The best multivariate models explained similar amounts of variation to
+
+```{r echo=FALSE, out.width='100%', fig.cap = "Plots of primary productivity $(MgC~ha^-1~yr^-1)$ regressed against the climate variables identified as the best predictors: mean annual temperature; temperature seasonality; annual temperature range; and vapour pressure.
+Lines of best fit are plotted according to the best model selected during analysis.
+All regressions are significant $(p<0.05)$.", fig.pos='H'}
+  knitr::include_graphics("combined_plots.png")
+```
+
+**Mean annual temperature.** All carbon fluxes show a linear increase with temperature.  
+**Temperature seasonality and annual temperature range.** These two climate variables are strongly correlated, and so only temperature seasonality is presented in figure 4.
+Productivity is highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower.
+GPP, NPP, ANPP, and R~root~ show a polynomial relationship with both variables, and BNPP~root~ shows a polynomial relationship with annual temperature range.
+As seasonality increases, these variables show a more rapid decrease in productivity.
+ANPP~foliage~ and ANPP~woodystem~ show a linear relationship with both variables.
+**Mean annual precipitation.** Despite being a poor predictor of productivity, the results indicate that, for the majority of fluxes, productivity increases with precipitation, up until a saturation point at between 3000 and 4000mm annual precipitation, above which productivity starts to decrease.
 The notable exception to this is GPP, which shows a significant (p<0.0001) linear increase with precipitation up to measures of nearly 6000mm annually, although the explanatory power of this model was low (R-squared = 0.15).
@@
 Between 1200 and 1700mm productivity peaks, and begins to decrease at values above 1800mm.
+
+```{r eval = TRUE, echo=FALSE, warning=FALSE}
+library(knitr)
+library(kableExtra)
+multivariate_models <- read.csv("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/multivariate_models.csv", stringsAsFactors = FALSE, check.names = FALSE)
+rownames(multivariate_models) <- c("GPP", "NPP", "ANPP", "ANPP\\textsubscript{foliage}", "ANPP\\textsubscript{woody stem}", "BNPP", "BNPP\\textsubscript{root}", "R\\textsubscript{root}")
+kable(multivariate_models, format = "latex", booktabs = TRUE, caption = "Multivariate model output summary.
+For each carbon flux, correlation coefficients are presented for the two climate variables which explain the most variation when included in a multivariate model.
+If the model includes an interactive effect, correlation coefficients for the interaction are presented.
+Results for R\\textsubscript{auto} are not presented here due to the small sample size (n = 14).
+", escape = FALSE, row.names = TRUE) %>%
+  kable_styling(latex_options = c("scale_down", "hold_position")) %>%
+  column_spec(1:10, width = "2cm") %>%
+  add_header_above(c(" ", " ", " ", "Correlation coefficients of climate variables" = 6, " "))
+
+```
 

### 2019-07-18 manuscript/manuscript draft.Rmd
update manuscript draft

@@
 ), 'packages.bib')
+```
+
+```{r results="asis", include=FALSE}
+cat("
+<style>
+caption {
+      font-size: 0.75em;
+    }
+</style>
+")
 ```
@@
 
-The best multivariate models explained similar amounts of variation to
-
-```{r echo=FALSE, out.width='100%', fig.cap = "Plots of primary productivity $(MgC~ha^-1~yr^-1)$ regressed against the climate variables identified as the best predictors: mean annual temperature; temperature seasonality; annual temperature range; and vapour pressure.
-Lines of best fit are plotted according to the best model selected during analysis.
-All regressions are significant $(p<0.05)$.", fig.pos='H'}
-  knitr::include_graphics("combined_plots.png")
-```
-
-**Mean annual temperature.** All carbon fluxes show a linear increase with temperature.  
-**Temperature seasonality and annual temperature range.** These two climate variables are strongly correlated, and so only temperature seasonality is presented in figure 4.
-Productivity is highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower.
-GPP, NPP, ANPP, and R~root~ show a polynomial relationship with both variables, and BNPP~root~ shows a polynomial relationship with annual temperature range.
-As seasonality increases, these variables show a more rapid decrease in productivity.
-ANPP~foliage~ and ANPP~woodystem~ show a linear relationship with both variables.
-**Mean annual precipitation.** Despite being a poor predictor of productivity, the results indicate that, for the majority of fluxes, productivity increases with precipitation, up until a saturation point at between 3000 and 4000mm annual precipitation, above which productivity starts to decrease.
-The notable exception to this is GPP, which shows a significant (p<0.0001) linear increase with precipitation up to measures of nearly 6000mm annually, although the explanatory power of this model was low (R-squared = 0.15).
-**Potential evapotranspiration.** All fluxes, with the exception of BNPP~rootfine~ and R~root~ show a strong increase with potential evapotranspiration at values below 1000mm.
-Between 1200 and 1700mm productivity peaks, and begins to decrease at values above 1800mm.
+Including a second climate term in the models didn't tend to significantly increase the explanatory power of the models.
+The best multivariate models for each carbon flux have r-squared values in the range of those recorded for models with a single climate term.
+However, the multivariate models do capture the additive and interactive effects of temperature and water availability.
 
@@
 multivariate_models <- read.csv("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/multivariate_models.csv", stringsAsFactors = FALSE, check.names = FALSE)
-rownames(multivariate_models) <- c("GPP", "NPP", "ANPP", "ANPP\\textsubscript{foliage}", "ANPP\\textsubscript{woody stem}", "BNPP", "BNPP\\textsubscript{root}", "R\\textsubscript{root}")
+rownames(multivariate_models) <- c("GPP", "NPP", "ANPP", "ANPP\\textsubscript{foliage}", "ANPP\\textsubscript{woody stem}", "BNPP", "BNPP\\textsubscript{fine root}", "R\\textsubscript{root}")
 kable(multivariate_models, format = "latex", booktabs = TRUE, caption = "Multivariate model output summary.
@@
 
-| Variable | Latitude | Mean annual temperature | Temperature seasonality | Potential evapotranspiration |
-|----------|----------|----------|----------|----------|
-| R~auto~ | 0.64 | 0.83 | 0.54 | 0.63 |
-| GPP | 0.66 | 0.61 | 0.75 | 0.40 |
-| NPP | 0.51 | 0.49 | 0.52 | 0.34 |
-| ANPP | 0.50 | 0.45 | 0.47 | 0.32 |
-| ANPP~foliage~ | 0.55 | 0.63 | 0.54 | 0.40 |
-| ANPP~woodystem~ | 0.25 | 0.23 | 0.20 | 0.26 |
-| BNPP~root~ | 0.31 | 0.26 | 0.31 | 0.32 |
-| BNPP~rootfine~ | 0.14 | 0.11 | 0.16 | 0.11 |
-| R~root~ | 0.25 | 0.21 | 0.29 | 0.04* |
+| Variable | Latitude | Mean annual temperature | Temperature seasonality | Potential evapotranspiration | Multivariate model |
+|----------|----------|----------|----------|----------|----------|
+| R~auto~ | 0.64 | 0.83 | 0.54 | 0.63 | NA |
+| GPP | 0.66 | 0.61 | 0.75 | 0.40 | 0.74 |
+| NPP | 0.51 | 0.49 | 0.52 | 0.34 | 0.56 |
+| ANPP | 0.50 | 0.45 | 0.47 | 0.32 | 0.49 |
+| ANPP~foliage~ | 0.55 | 0.63 | 0.54 | 0.40 | 0.66 |
+| ANPP~woodystem~ | 0.25 | 0.23 | 0.20 | 0.26 | 0.26 |
+| BNPP~root~ | 0.31 | 0.26 | 0.31 | 0.32 | 0.36 |
+| BNPP~rootfine~ | 0.14 | 0.11 | 0.16 | 0.11 | 0.19 |
+| R~root~ | 0.25 | 0.21 | 0.29 | 0.04* | 0.47 |
 
-Table: R-squared values presented for four climate variables. Values marked * are not significant.
+Table: R-squared values presented for latitude, three climate variables, and the best multivariate model for each carbon flux. Values marked * are not significant.
 
@@
 
+This pattern is also seen in the r-squared values for multivariate models.
+
 ###Discussion

### 2019-07-18 scripts/interactions_all_combinations.R
Create interactions_all_combinations.R

@@ -1,0 +1,176 @@
+
+
+# Clean environment ####
+rm(list = ls())
+
+# Set working directory as ForC main folder ####
+setwd("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/ForC")
+
+# Load libaries ####
+library(lme4)
+library(MuMIn)
+library(AICcmodavg)
+
+# Load data ####
+ForC_simplified <- read.csv("ForC_simplified/ForC_simplified_WorldClim_CRU_refined.csv", stringsAsFactors = F)
+VARIABLES <- read.csv(paste0(dirname(getwd()), "/ForC/data/ForC_variables.csv"), stringsAsFactors = F)
+
+na_codes <- c("NA", "NI", "NRA", "NaN", "NAC") 
+my_is.na <- function(x) { is.na(x) | x %in% na_codes}
+my_na.omit <- function(x) { return(x[!my_is.na(x)])}
+
+
+# Prepare data ####
+
+## change to numerics what it needs to
+ForC_simplified$stand.age <- as.numeric(ForC_simplified$stand.age)
+ForC_simplified$min.db <- as.numeric(ForC_simplified$min.db)
+
+## make plot and geographic areas factors
+ForC_simplified$plot.name <- addNA(ForC_simplified$plot.name)
+ForC_simplified$geographic.area <- addNA(ForC_simplified$geographic.area)
+
+## change sign of NEE 
+ForC_simplified[grepl("NEE", ForC_simplified$variable.name,ignore.case = F),]$mean <- -ForC_simplified[grepl("NEE", ForC_simplified$variable.name,ignore.case = F),]$mean
+
+## take absolute value of latitude
+ForC_simplified$lat <- abs(ForC_simplified$lat)
+
+
+# Control for some factors ####
+
+
+## keep all ages except na, 0 and 999
+ages.not.999.nor.0.nor.na <- !ForC_simplified$stand.age %in% 999 &  !ForC_simplified$stand.age %in% 0 & !is.na(ForC_simplified$stand.age)
+
+## Keep only age >=100 (or 999)
+age.greater.than.100 <- ForC_simplified$stand.age >= 100 & !is.na(ForC_simplified$stand.age)
+age.greater.than.200 <- ForC_simplified$stand.age >= 200 & !is.na(ForC_simplified$stand.age)
+ages <- c("age.greater.than.100")
+
+## keep only stands that are NOT too strongly influence by management/ disturbance
+
[…]

### 2019-07-18 scripts/interactions_all_combinations.R
Update interactions_all_combinations.R

@@ -10,6 +10,7 @@ setwd("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/ForC")
 library(lme4)
 library(MuMIn)
 library(AICcmodavg)
+library(piecewiseSEM)
 
 # Load data ####
 ForC_simplified <- read.csv("ForC_simplified/ForC_simplified_WorldClim_CRU_refined.csv", stringsAsFactors = F)
@@ -170,6 +171,7 @@ response.variables <- "GPP"
           names(lmm_all) <- formula_vec
           
           var_aic <- aictab(lmm_all, second.ord=TRUE, sort=TRUE) #rank based on AICc
+          rsq <- rsquared(lmm_all)
           
         }
 

### 2019-07-23 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
 To investigate the potential interactive effects of climate variables on carbon fluxes, multivariate models were also specified.
-Climate variables which were non-significant or poor predictors in the models with a single climate term, or which were strongly correlated with another climate term, were dropped, leaving a subset of 8 climate variables.
-For each possible pairing of climate variables, three models were specified: a null model; a model with the two climate variables showing an additive effect; and a model with the two climate variables showing an interactive effect.
+Climate variables which were non-significant or poor predictors in the models with a single climate term, or which were strongly correlated with another climate term, were dropped, leaving a subset of 7 climate variables: mean annual temperature; temperature seasonality; mean annual precipitation; precipitation seasonality; potential evapotranspiration; vapour pressure deficit; and solar radiation.
+For each possible pairing of climate variables, two models were specified: a model with the two climate variables showing an additive effect; and a model with the two climate variables showing an interactive effect.
+In addition a null model, and models including each climate term as a individual (univariate) linear term were specified.
 As before, altitude was included as a fixed effect, and plot nested within geographic area as a random effect.
@@
 
-Including a second climate term in the models didn't tend to significantly increase the explanatory power of the models.
-The best multivariate models for each carbon flux have r-squared values in the range of those recorded for models with a single climate term.
-However, the multivariate models do capture the additive and interactive effects of temperature and water availability.
+Including a second climate term in the models tended to have only a small effect on the explanatory power of the models.
+For ANPP~woody stem~ and R~auto~, including two climate terms did not improve the explantory power of the models: mean annual temperature alone was the best climate predictor for both of these fluxes.
+For all the remaining carbon fluxes, including a second climate term did improve the explanatory power of the models.
+ANPP and ANPP~foliage~ were best explained by an additive model, in both cases including mean annual temperature and potential evapotranspiration as the climate terms.
+GPP, NPP, BNPP, BNPP~fine root~, and R~root~ were all best explained by multivariate models which included an interaction term between the two climate terms.
+The best models for GPP, NPP, and R~root~ all included a temperature term and a precipitation term, indicating that although precipitation isn't a strong predictor for these variables, the interaction between temperature and water availability is important.
 
@@
 multivariate_models <- read.csv("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/multivariate_models.csv", stringsAsFactors = FALSE, check.names = FALSE)
-rownames(multivariate_models) <- c("GPP", "NPP", "ANPP", "ANPP\\textsubscript{foliage}", "ANPP\\textsubscript{woody stem}", "BNPP", "BNPP\\textsubscript{fine root}", "R\\textsubscript{root}")
+rownames(multivariate_models) <- c("GPP", "NPP", "ANPP", "ANPP\\textsubscript{foliage}", "ANPP\\textsubscript{woody stem}", "BNPP", "BNPP\\textsubscript{fine root}","R\\textsubscript{auto}", "R\\textsubscript{root}")
 kable(multivariate_models, format = "latex", booktabs = TRUE, caption = "Multivariate model output summary.
@@
 If the model includes an interactive effect, correlation coefficients for the interaction are presented.
-Results for R\\textsubscript{auto} are not presented here due to the small sample size (n = 14).
+For carbon fluxes that were best predicted by a single climate term, correlation coefficients are presented for that term only.
 ", escape = FALSE, row.names = TRUE) %>%
@@
   column_spec(1:10, width = "2cm") %>%
-  add_header_above(c(" ", " ", " ", "Correlation coefficients of climate variables" = 6, " "))
+  add_header_above(c(" ", " ", " ", "Correlation coefficients of climate variables" = 7, " "))
 
@@
 | Variable | Latitude | Mean annual temperature | Temperature seasonality | Potential evapotranspiration | Multivariate model |
-|----------|----------|----------|----------|----------|----------|
+|----------|----------|----------|----------|--------------|----------|
 | R~auto~ | 0.64 | 0.83 | 0.54 | 0.63 | NA |
-| GPP | 0.66 | 0.61 | 0.75 | 0.40 | 0.74 |
-| NPP | 0.51 | 0.49 | 0.52 | 0.34 | 0.56 |
-| ANPP | 0.50 | 0.45 | 0.47 | 0.32 | 0.49 |
+| GPP | 0.66 | 0.61 | 0.75 | 0.40 | 0.73 |
+| NPP | 0.51 | 0.49 | 0.52 | 0.34 | 0.57 |
+| ANPP | 0.50 | 0.45 | 0.47 | 0.32 | 0.50 |
 | ANPP~foliage~ | 0.55 | 0.63 | 0.54 | 0.40 | 0.66 |
-| ANPP~woodystem~ | 0.25 | 0.23 | 0.20 | 0.26 | 0.26 |
-| BNPP~root~ | 0.31 | 0.26 | 0.31 | 0.32 | 0.36 |
+| ANPP~woodystem~ | 0.25 | 0.23 | 0.20 | 0.26 | NA |
+| BNPP~root~ | 0.31 | 0.26 | 0.31 | 0.32 | 0.38 |
 | BNPP~rootfine~ | 0.14 | 0.11 | 0.16 | 0.11 | 0.19 |
-| R~root~ | 0.25 | 0.21 | 0.29 | 0.04* | 0.47 |
+| R~root~ | 0.25 | 0.21 | 0.29 | 0.04* | 0.43 |
 

### 2019-07-23 scripts/interactions_all_combinations.R
Update interactions_all_combinations.R

@@ -37,6 +37,8 @@ ForC_simplified[grepl("NEE", ForC_simplified$variable.name,ignore.case = F),]$me
 ## take absolute value of latitude
 ForC_simplified$lat <- abs(ForC_simplified$lat)
 
+ForC_simplified$SolarRadiation <- ForC_simplified$SolarRadiation/1000
+
 
 # Control for some factors ####
 

### 2019-07-23 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
 | BNPP~root~ | Belowground net primary production | Coarse and fine roots | Direct measurement of one or more of: fine root turnover, soil cores, root ingrowth cores, minirhizotrons; indirect estimates of coarse roots using allometries based on aboveground stem increment measures | 105 | 90 |
-| BNPP~rootfine~ | Net primary production of fine roots | Fine roots | Direct measurement of one or more of: minirhizotrons, fine root turnover, soil cores, root ingrowth cores | 96 | 88 |
+| BNPP~fineroot~ | Net primary production of fine roots | Fine roots | Direct measurement of one or more of: minirhizotrons, fine root turnover, soil cores, root ingrowth cores | 96 | 88 |
 | R~auto~ | Annual autotrophic respiration, including above- and belowground components | Foliage, stem, and root | Chamber measurements of component gas exchange | 22 | 22 |
@@
 2. NPP = ANPP + BNPP  
-3. ANPP = ANPP~foliage~ + ANPP~woodystem~  
+3. ANPP = ANPP~foliage~ + ANPP~woody stem~  
 4. Total belowground carbon flux = BNPP + R~root~
 
-*Allocation to carbon fluxes along latitudinal gradients.* Variation in allocation to component carbon fluxes along latitudinal gradients was explored for a range of pairings: firstly, GPP:NPP, ANPP:BNPP, and ANPP~foliage~:ANPP~woodystem~; and secondly, the ratio of NPP to each of ANPP, BNPP, ANPP~foliage~, and ANPP~woodystem~.
+*Allocation to carbon fluxes along latitudinal gradients.* Variation in allocation to component carbon fluxes along latitudinal gradients was explored for a range of pairings: firstly, GPP:NPP, ANPP:BNPP, and ANPP~foliage~:ANPP~woody stem~; and secondly, the ratio of NPP to each of ANPP, BNPP, ANPP~foliage~, and ANPP~woody stem~.
 For each set of paired fluxes, measurements taken at the same site and plot, and in the same year, were paired together, and the ratio of each pair of measurements calculated.
@@
 3).
-Modelled estimates of GPP, generated from the sum of NPP and R auto; NPP, generated from the sum of ANPP and BNPP~root~; and ANPP, generated from the sum of ANPP~foliage~ and ANPP~woodystem~, fall completely within the confidence intervals of the regressions of field estimates of GPP, NPP and ANPP respectively.
+Modelled estimates of GPP, generated from the sum of NPP and R auto; NPP, generated from the sum of ANPP and BNPP~root~; and ANPP, generated from the sum of ANPP~foliage~ and ANPP~woody stem~, fall completely within the confidence intervals of the regressions of field estimates of GPP, NPP and ANPP respectively.
 
@@
 Looking across all productivity components, the climate variables which explain the highest amount of variation in carbon fluxes are temperature seasonality, mean annual temperature, annual temperature range, and annual frost days.
-Vapour pressure deficit came out as a strong predictor for ANPP~woodystem~, and potential evapotranspiration was a good predictor for BNPP~root~ and ANPP~woodystem~.
+Vapour pressure deficit came out as a strong predictor for ANPP~woody stem~, and potential evapotranspiration was a good predictor for BNPP~root~ and ANPP~woody stem~.
 
@@
 As seasonality increases, these variables show a more rapid decrease in productivity.
-ANPP~foliage~ and ANPP~woodystem~ show a linear relationship with both variables.
+ANPP~foliage~ and ANPP~woody stem~ show a linear relationship with both variables.
 **Mean annual precipitation.** Despite being a poor predictor of productivity, the results indicate that, for the majority of fluxes, productivity increases with precipitation, up until a saturation point at between 3000 and 4000mm annual precipitation, above which productivity starts to decrease.
 The notable exception to this is GPP, which shows a significant (p<0.0001) linear increase with precipitation up to measures of nearly 6000mm annually, although the explanatory power of this model was low (r-squared = 0.15).
-**Potential evapotranspiration.** All fluxes, with the exception of BNPP~rootfine~ and R~root~ show a strong increase with potential evapotranspiration at values below 1000mm.
+**Potential evapotranspiration.** All fluxes, with the exception of BNPP~fine root~ and R~root~ show a strong increase with potential evapotranspiration at values below 1000mm.
 Between 1200 and 1700mm productivity peaks, and begins to decrease at values above 1800mm.
@@
 | BNPP~root~ | 0.31 | 0.26 | 0.31 | 0.32 | 0.38 |
-| BNPP~rootfine~ | 0.14 | 0.11 | 0.16 | 0.11 | 0.19 |
+| BNPP~fineroot~ | 0.14 | 0.11 | 0.16 | 0.11 | 0.19 |
 | R~root~ | 0.25 | 0.21 | 0.29 | 0.04* | 0.43 |
@@
 With the exception of ANPP~foliage~, the proportion of variation explained by climate and latitude in subsidiary fluxes is much lower.
-Climate explains at most 26% of variation in ANPP~woodystem~, 17% in BNPP~rootfine~, and 32% in R~root~.
+Climate explains at most 26% of variation in ANPP~woody stem~, 17% in BNPP~fine root~, and 32% in R~root~.
 In contrast, climate strongly explains variation in ANPP~foliage~, with mean annual temperature explaining 63% of variation.

### 2019-07-29 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
 The effect of each extracted climate variable on each measure of primary productivity was modelled by specifying the climate variable as a fixed effect.
-Site altitude was also included as a fixed effect.
+For each climate variable, three models were specified: a null model; a model with the climate variable as a linear term; and a model with the climate variable as a polynomial term.
+AIC values were calculated for the models and used to select the best model.
+R-squared values were calculated for the best model.
+All R-squared values presented here are marginal r-squared values, and refer to the proportion of variation explained by only the fixed effects, unless otherwise specified.
+In addition, slope coefficients were calculated for the linear models.
+
+Because the magnitude of fluxes varies significantly, in order to facilitate comparisons between regression models for each flux, data for each flux was scaled, to give the data a mean of 0 and standard deviation of 1.
+As each data set was scaled separately, this does not allow for statistical comparisons of slope values, but does assist in visualising the data.
+
+To test for a potential influence of altitude, models were also run with site altitude included as a second fixed effect.
+These models were compared against models with no altitude term, and AIC values calculated to identify whether inclusion of altitude as a term improved the models.
+
 Within the ForC database, sites within 25km^2^ of each other are clustered into geographic areas.
@@
 Results were similar between the weighted and unweighted model, so, to avoid problems of over-fitting, the weighted model was dropped, and results from this are not presented here.
-
-For each climate variable, three models were specified: a null model; a model with the climate variable as a linear term; and a model with the climate variable as a polynomial term.
-AIC values were calculated for the models and used to select the best model.
-R-squared values were calculated for the best model.
-In addition, slope coefficients were calculated for the linear models.
-
-Because the magnitude of fluxes varies significantly, in order to facilitate comparisons between regression models for each flux, data for each flux was scaled, to give the data a mean of 0 and standard deviation of 1.
-As each data set was scaled separately, this does not allow for statistical comparisons of slope values, but does assist in visualising the data.
 
@@
 2).
-All models were improved by the inclusion of altitude as a fixed effect, with carbon fluxes showing a significant decrease with increasing altitude.
+Models showed significant variation in whether they were improved by inclusion of altitude as a fixed effect.
+Including altitude had a very small effect on most models, with the difference in AIC values between models including and excluding altitude often being <2, suggesting the models are very similar in their explanatory power.
+All models presented below do not include altitude, and r-squared values refer to the proportion of variation explained by the climate variable alone.
 

### 2019-07-29 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@
 AIC values were calculated for the models and used to select the best model.
-R-squared values were calculated for the best model.
-All R-squared values presented here are marginal r-squared values, and refer to the proportion of variation explained by only the fixed effects, unless otherwise specified.
+R^2^ values were calculated for the best model.
+All R^2^ values presented here are marginal R^2^ values, and refer to the proportion of variation explained by only the fixed effects, unless otherwise specified.
 In addition, slope coefficients were calculated for the linear models.
@@
 AIC values were calculated for the models and used to select the best model.
-R-squared values were calculated for the best model.
+R^2^ values were calculated for the best model.
 Slope coefficients were calculated for each of the terms, and for the interactive term if relevant.
@@
 Including altitude had a very small effect on most models, with the difference in AIC values between models including and excluding altitude often being <2, suggesting the models are very similar in their explanatory power.
-All models presented below do not include altitude, and r-squared values refer to the proportion of variation explained by the climate variable alone.
+All models presented below do not include altitude, and R^2^ values refer to the proportion of variation explained by the climate variable alone.
 
@@
 
-Latitude is a strong predictor for many of the carbon fluxes, explaining 66% of variation in GPP (n = 254, p<0.0001), 51% in NPP (n = 116, p<0.0001) and 50% in ANPP (n = 259, p<0.0001).
+Latitude is a strong predictor for many of the carbon fluxes, explaining 64% of variation in GPP (n = 254, p<0.0001), 50% in NPP (n = 114, p<0.0001) and 45% in ANPP (n = 259, p<0.0001).
 For all fluxes, their relationship with latitude was best predicted by the linear model.
@@
 **Mean annual precipitation.** Despite being a poor predictor of productivity, the results indicate that, for the majority of fluxes, productivity increases with precipitation, up until a saturation point at between 3000 and 4000mm annual precipitation, above which productivity starts to decrease.
-The notable exception to this is GPP, which shows a significant (p<0.0001) linear increase with precipitation up to measures of nearly 6000mm annually, although the explanatory power of this model was low (r-squared = 0.15).
+The notable exception to this is GPP: the model indicates that GPP continues to increase with precipitation up to measures of at least 5000mm annually (p<0.0001, R^2^ = 0.33.
+Data above this point is not available, but the model trend indicates that the saturation point for this model is around 5000mm mean annual precipitation.
 **Potential evapotranspiration.** All fluxes, with the exception of BNPP~fine root~ and R~root~ show a strong increase with potential evapotranspiration at values below 1000mm.
@@
 
-*Do carbon fluxes vary in their responses to climate, particularly with regards to allocation between fluxes?* Of all the flux pairings tested, the only ones to show a significant relationship with latitude were ANPP:NPP (p<0.01, r^2^ = 0.44) and ANPP:BNPP (p<0.001, r^2^ = 0.33).
+*Do carbon fluxes vary in their responses to climate, particularly with regards to allocation between fluxes?* Of all the flux pairings tested, the only ones to show a significant relationship with latitude were ANPP:NPP (p<0.01, R^2^ = 0.44) and ANPP:BNPP (p<0.001, R^2^ = 0.33).
 In both cases, allocation to ANPP increased with latitude, with the lowest allocation to ANPP in the tropical regions.
@@
 |----------|----------|----------|----------|--------------|----------|
-| R~auto~ | 0.64 | 0.83 | 0.54 | 0.63 | NA |
-| GPP | 0.66 | 0.61 | 0.75 | 0.40 | 0.73 |
-| NPP | 0.51 | 0.49 | 0.52 | 0.34 | 0.57 |
-| ANPP | 0.50 | 0.45 | 0.47 | 0.32 | 0.50 |
-| ANPP~foliage~ | 0.55 | 0.63 | 0.54 | 0.40 | 0.66 |
-| ANPP~woodystem~ | 0.25 | 0.23 | 0.20 | 0.26 | NA |
-| BNPP~root~ | 0.31 | 0.26 | 0.31 | 0.32 | 0.38 |
-| BNPP~fineroot~ | 0.14 | 0.11 | 0.16 | 0.11 | 0.19 |
-| R~root~ | 0.25 | 0.21 | 0.29 | 0.04* | 0.43 |
+| R~auto~ | 0.65 | 0.77 | 0.62 | 0.63 | NA |
+| GPP | 0.64 | 0.63 | 0.72 | 0.36 | 0.73 |
+| NPP | 0.50 | 0.52 | 0.52 | 0.33 | 0.57 |
+| ANPP | 0.45 | 0.46 | 0.42 | 0.28 | 0.50 |
+| ANPP~foliage~ | 0.54 | 0.63 | 0.54 | 0.35 | 0.66 |
+| ANPP~woodystem~ | 0.17 | 0.23 | 0.12 | 0.19 | NA |
+| BNPP~root~ | 0.33 | 0.26 | 0.23 | 0.36 | 0.38 |
+| BNPP~fineroot~ | 0.15 | 0.11 | 0.16 | 0.11 | 0.19 |
+| R~root~ | 0.22 | 0.21 | 0.30 | 0.01* | 0.43 |
 
-Table: R-squared values presented for latitude, three climate variables, and the best multivariate model for each carbon flux. Values marked * are not significant.
+Table: R^2^ values presented for latitude, three climate variables, and the best multivariate model for each carbon flux. Values marked * are not significant.
 
-R-squared values are generally highest in the major fluxes, and decrease in subsidiary fluxes.
+R^2^ values are generally highest in the major fluxes, and decrease in subsidiary fluxes.
 Of the major fluxes, R~auto~ and GPP are the most strongly explained by latitude and climate.
-Mean annual temperature explains 83% of variation in R~auto~, while temperature seasonality explains 75% of variation in GPP.
+Mean annual temperature explains 77% of variation in R~auto~, while temperature seasonality explains 72% of variation in GPP.
 The proportion of variation explained by climate and latitude decreases in NPP and ANPP.
 The climatic variables with the strongest explanatory power explain around 50% of the variation in these fluxes.
-Of the major fluxes, BNPP~root~ is the least well explained by climate and latitude, with climate explaining at most 32% of variation.
+Of the major fluxes, BNPP~root~ is the least well explained by climate and latitude, with climate explaining at most 36% of variation.
 
 With the exception of ANPP~foliage~, the proportion of variation explained by climate and latitude in subsidiary fluxes is much lower.
-Climate explains at most 26% of variation in ANPP~woody stem~, 17% in BNPP~fine root~, and 32% in R~root~.
+Climate explains at most 23% of variation in ANPP~woody stem~, 16% in BNPP~fine root~, and 30% in R~root~.
 In contrast, climate strongly explains variation in ANPP~foliage~, with mean annual temperature explaining 63% of variation.
 
-This pattern is also seen in the r-squared values for multivariate models.
+This pattern is also seen in the R^2^ values for multivariate models.
 

### 2019-07-29 scripts/interactions_all_combinations.R
Update interactions_all_combinations.R
removed altitude term

@@ -105,7 +105,7 @@ all.response.variables[!all.response.variables %in% unlist(response.variables.gr
 all.results <- NULL
 best.results <- NULL
 
-effects <- c("mat", "map", "TempSeasonality", "SolarRadiation", "PotentialEvapotranspiration", "VapourPressureDeficit", "PreSeasonality", "masl", "(1|geographic.area/plot.name)")
+effects <- c("mat", "map", "TempSeasonality", "SolarRadiation", "PotentialEvapotranspiration", "VapourPressureDeficit", "PreSeasonality", "(1|geographic.area/plot.name)")
 
 
 response.variables <- c("GPP", "NPP", "BNPP_root", "BNPP_root_fine", "ANPP", "ANPP_foliage", "ANPP_woody_stem","R_auto", "R_auto_root")
@@ -131,23 +131,22 @@ response.variables <- c("GPP", "NPP", "BNPP_root", "BNPP_root_fine", "ANPP", "AN
           uni_comb <-
             unlist( sapply( seq_len(length(1)), 
                             function(x) {
-                              apply( combn(effects, 3), 2, function(x) paste(x, collapse = "+"))
+                              apply( combn(effects, 2), 2, function(x) paste(x, collapse = "+"))
                             }))
           
           uni_comb <- expand.grid(response, uni_comb) 
           uni_comb <- uni_comb[grepl("1", uni_comb$Var2), ] #keep random effect
-          uni_comb <- uni_comb[grepl("masl", uni_comb$Var2), ] #keep masl
+        
           
           add_comb <- 
             unlist( sapply( seq_len(length(1)), 
                             function(x) {
-                              apply( combn(effects, 4), 2, function(x) paste(x, collapse = "+"))
+                              apply( combn(effects, 3), 2, function(x) paste(x, collapse = "+"))
                             }))
          
           
           add_comb <- expand.grid(response, add_comb) 
           add_comb <- add_comb[grepl("1", add_comb$Var2), ] #keep random effect
-          add_comb <- add_comb[grepl("masl", add_comb$Var2), ] #keep masl
           
           int_comb <-
             unlist( sapply( seq_len(length(1)), 
@@ -155,22 +154,21 @@ response.variables <- c("GPP", "NPP", "BNPP_root", "BNPP_root_fine", "ANPP", "AN
                               apply( combn(effects, 2), 2, function(x) paste(x, collapse = "*"))
                             }))
           int_comb <- expand.grid(response, int_comb) 
[…]

### 2019-08-02 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
 header-includes:  \usepackage{float}
+                  \usepackage{caption}
+                  \captionsetup[table]{font=footnotesize}
+                  \captionsetup[figure]{font=footnotesize}
 
@@
 To investigate the potential interactive effects of climate variables on carbon fluxes, multivariate models were also specified.
-Climate variables which were non-significant or poor predictors in the models with a single climate term, or which were strongly correlated with another climate term, were dropped, leaving a subset of 7 climate variables: mean annual temperature; temperature seasonality; mean annual precipitation; precipitation seasonality; potential evapotranspiration; vapour pressure deficit; and solar radiation.
+To ensure that models were biologically meaningful, the terms included in the models tested built on results from the univariate models.
+Modelling of individual climate variables identified that the best predictors of carbon fluxes were variables related to temperature.
+We therefore decided to include mean annual temperature as a term in all multivariate models.
+We first modelled the interaction effect between mean annual temperature and mean annual precipitation, in order to capture climate variation along the axes of temperature and water availability.
+Models were tested for a significant interactive effect and a significant additive effect.
+We then explored whether any other climate variable, in combination with mean annual temperature, could significantly improve on the combination of mean annual temperature and mean annual precipitation.
+In specifying the range of models to test, climate variables which were strongly correlated with temperature were dropped, in order to capture the greatest range of variation in climate.
 For each possible pairing of climate variables, two models were specified: a model with the two climate variables showing an additive effect; and a model with the two climate variables showing an interactive effect.
-In addition a null model, and models including each climate term as a individual (univariate) linear term were specified.
-As before, altitude was included as a fixed effect, and plot nested within geographic area as a random effect.
-AIC values were calculated for the models and used to select the best model.
-R^2^ values were calculated for the best model.
-Slope coefficients were calculated for each of the terms, and for the interactive term if relevant.
+As before, plot nested within geographic area was included as a random effect.
+Altitude was not considered.
+AIC values were calculated for the models, and used to compare models.
+Models were considered to be significantly better than the baseline MAT*MAP model if:
+  i) the AIC value of the model was smaller than the AIC value of the baseline model by >2  
+  ii) the r-squared value of the model was larger than the r-squared value of the baseline model by >5
 

### 2019-08-07 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
 
-Including a second climate term in the models tended to have only a small effect on the explanatory power of the models.
-For ANPP~woody stem~ and R~auto~, including two climate terms did not improve the explantory power of the models: mean annual temperature alone was the best climate predictor for both of these fluxes.
-For all the remaining carbon fluxes, including a second climate term did improve the explanatory power of the models.
-ANPP and ANPP~foliage~ were best explained by an additive model, in both cases including mean annual temperature and potential evapotranspiration as the climate terms.
-GPP, NPP, BNPP, BNPP~fine root~, and R~root~ were all best explained by multivariate models which included an interaction term between the two climate terms.
-The best models for GPP, NPP, and R~root~ all included a temperature term and a precipitation term, indicating that although precipitation isn't a strong predictor for these variables, the interaction between temperature and water availability is important.
+There was a significant interactive effect between MAT and MAP for GPP, BNPP~root~, BNPP~fine root~, ANPP, ANPP~woody stem~, and R~root~.
+There was a significant additive effect for R~auto~.
+NPP and ANPP~foliage~ showed no significant interactive or additive effect: including MAP as a second variable did not improve on the model including only MAT.
+
+For the variables which showed a significant interactive or additive effect between MAT and MAP, no other climate variable, in combination with MAT, significantly improved on that model.
+For NPP, there was a significant interactive effect between MAT and water stress months, with this model explaining nearly 5% more variation in NPP than MAT alone.
+However, for ANPP~foliage~, no multivariate model improved on the univariate model including only MAT.
 
@@
 
+*The relationship between productivity and latitude*
+
+Past studies have differed in their conclusions regarding the relationship between productivity and latitude.
+Our findings indicate that, for all carbon fluxes studied, productivity peaks in the tropical regions.
+While Huston and Wolverton (2009) found no significant relationship between ANPP and latitude, we find a significant negative relationship between ANPP and latitude, with the highest levels of ANPP in the tropical regions.
+
 *Challenges in estimating NPP*  

### 2019-08-07 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
 
-```{r eval = TRUE, echo=FALSE, warning=FALSE}
-library(knitr)
-library(kableExtra)
-multivariate_models <- read.csv("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/multivariate_models.csv", stringsAsFactors = FALSE, check.names = FALSE)
-rownames(multivariate_models) <- c("GPP", "NPP", "ANPP", "ANPP\\textsubscript{foliage}", "ANPP\\textsubscript{woody stem}", "BNPP", "BNPP\\textsubscript{fine root}","R\\textsubscript{auto}", "R\\textsubscript{root}")
-kable(multivariate_models, format = "latex", booktabs = TRUE, caption = "Multivariate model output summary.
-For each carbon flux, correlation coefficients are presented for the two climate variables which explain the most variation when included in a multivariate model.
-If the model includes an interactive effect, correlation coefficients for the interaction are presented.
-For carbon fluxes that were best predicted by a single climate term, correlation coefficients are presented for that term only.
-", escape = FALSE, row.names = TRUE) %>%
-  kable_styling(latex_options = c("scale_down", "hold_position")) %>%
-  column_spec(1:10, width = "2cm") %>%
-  add_header_above(c(" ", " ", " ", "Correlation coefficients of climate variables" = 7, " "))
-
+```{r echo=FALSE, out.width='100%', fig.cap = "Plots of primary productivity $(MgC~ha^-1~yr^-1)$ regressed against mean annual precipitation.
+Points are grouped into bins of 0 - 1000, 1001 - 2000, 2001 - 3000, and >3000mm mean annual precipitation, and lines of best fit plotted for mean annual precipitation values of 500, 1500, 2500, and 3500mm.
+All regressions are significant $(p<0.05)$.", fig.pos='H'}
+  knitr::include_graphics("mat_map_interaction.png")
 ```

### 2019-08-12 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@
 |--------------|-------------------|-------------|------------------------|-------|-------|
-| GPP | Annual gross primary production; annual uptake of carbon dioxide by an ecosystem | NA | Flux partitioning of eddy covariance | 254 | 91 |
-| NPP | Annual net primary production; the component of GPP that is stored in plant tissue; GPP minus ecosystem respiration | Foliage, branch, stem, coarse root, fine root and optionally understory | Direct measurement of annual increments of components | 116 | 85 |
-| ANPP | Aboveground net primary production | Foliage, stem, and optionally branch | Direct measurement of annual increments of components | 259 | 170 |
-| ANPP~foliage~ | Net primary production of foliage | Foliage | Direct measurement of litterfall, correcting for changes in leaf biomass when measured | 100 | 96 |
-| ANPP~woodystem~ | Net primary production of woody stems | Woody stems | Direct measurement of stem growth increment | 264 | 252 |
-| BNPP~root~ | Belowground net primary production | Coarse and fine roots | Direct measurement of one or more of: fine root turnover, soil cores, root ingrowth cores, minirhizotrons; indirect estimates of coarse roots using allometries based on aboveground stem increment measures | 105 | 90 |
-| BNPP~fineroot~ | Net primary production of fine roots | Fine roots | Direct measurement of one or more of: minirhizotrons, fine root turnover, soil cores, root ingrowth cores | 96 | 88 |
+| GPP | Annual gross primary production; annual uptake of carbon dioxide by an ecosystem | NA | Flux partitioning of eddy covariance | 257 | 80 |
+| NPP | Annual net primary production; the component of GPP that is stored in plant tissue; GPP minus ecosystem respiration | Foliage, branch, stem, coarse root, fine root and optionally understory | Direct measurement of annual increments of components | 93 | 78 |
+| ANPP | Aboveground net primary production | Foliage, stem, and optionally branch | Direct measurement of annual increments of components | 258 | 166 |
+| ANPP~foliage~ | Net primary production of foliage | Foliage | Direct measurement of litterfall, correcting for changes in leaf biomass when measured | 98 | 94 |
+| ANPP~woodystem~ | Net primary production of woody stems | Woody stems | Direct measurement of stem growth increment | 266 | 252 |
+| BNPP~root~ | Belowground net primary production | Coarse and fine roots | Direct measurement of one or more of: fine root turnover, soil cores, root ingrowth cores, minirhizotrons; indirect estimates of coarse roots using allometries based on aboveground stem increment measures | 107 | 87 |
+| BNPP~fineroot~ | Net primary production of fine roots | Fine roots | Direct measurement of one or more of: minirhizotrons, fine root turnover, soil cores, root ingrowth cores | 90 | 77 |
 | R~auto~ | Annual autotrophic respiration, including above- and belowground components | Foliage, stem, and root | Chamber measurements of component gas exchange | 22 | 22 |
-| R~root~ | Annual root respiration | Coarse and fine roots | Measurement of root gas exchange | 57 | 38 |
+| R~root~ | Annual root respiration | Coarse and fine roots | Measurement of root gas exchange | 65 | 43 |
 
@@
 Since the components included in productivity estimates vary between studies, estimates of productivity were classified within the ForC database according to their components, and then filtered for analysis.
-Estimates of NPP were selected if they included foliage, branch, stem, coarse root, fine root and optionally understory.
-Measures of NPP which included additional components, including volatile organic compounds (VOCs), exudates, estimates of NPP lost to herbivory, and the NPP of reproductive structures, were excluded.
-Estimates of ANPP were selected if they included foliage, stem growth and branch turnover.
+Estimates of NPP were selected if they included foliage, branch, stem, coarse root, and fine root.
+Measures of NPP which included additional components, including understorey, volatile organic compounds (VOCs), exudates, estimates of NPP lost to herbivory, and the NPP of reproductive structures, were excluded.
+Estimates of ANPP were selected if they included foliage, stem growth and optionally branch turnover.
 Any measures of primary productivity where components were unknown were excluded from analysis.
@@
 
-| Variable | Latitude | Mean annual temperature | Temperature seasonality | Potential evapotranspiration | Multivariate model |
+| Variable | Latitude | Mean annual temperature | Temperature seasonality | Potential evapotranspiration | MAT*MAP |
 |----------|----------|----------|----------|--------------|----------|
 | R~auto~ | 0.65 | 0.77 | 0.62 | 0.63 | NA |
-| GPP | 0.64 | 0.63 | 0.72 | 0.36 | 0.73 |
-| NPP | 0.50 | 0.52 | 0.52 | 0.33 | 0.57 |
-| ANPP | 0.45 | 0.46 | 0.42 | 0.28 | 0.50 |
-| ANPP~foliage~ | 0.54 | 0.63 | 0.54 | 0.35 | 0.66 |
-| ANPP~woodystem~ | 0.17 | 0.23 | 0.12 | 0.19 | NA |
-| BNPP~root~ | 0.33 | 0.26 | 0.23 | 0.36 | 0.38 |
+| GPP | 0.64 | 0.63 | 0.72 | 0.36 | 0.64 |
+| NPP | 0.50 | 0.52 | 0.52 | 0.33 | 0.56 |
+| ANPP | 0.45 | 0.46 | 0.42 | 0.28 | 0.48 |
+| ANPP~foliage~ | 0.54 | 0.63 | 0.54 | 0.35 | 0.64 |
+| ANPP~woodystem~ | 0.17 | 0.23 | 0.12 | 0.19 | 0.28 |
+| BNPP~root~ | 0.33 | 0.26 | 0.23 | 0.36 | 0.35 |
 | BNPP~fineroot~ | 0.15 | 0.11 | 0.16 | 0.11 | 0.19 |
-| R~root~ | 0.22 | 0.21 | 0.30 | 0.01* | 0.43 |
+| R~root~ | 0.22 | 0.21 | 0.30 | 0.01* | 0.28 |
 

### 2019-08-22 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
 
-Globally, forests play an important role in the carbon cycle and are an important component of global carbon dioxide budgets [@luyssaert_old-growth_2008].
+Globally, forests play an important role in the carbon cycle and are a major component of global carbon dioxide budgets [@luyssaert_old-growth_2008].
 They show higher levels of productivity than non-forest terrestrial ecosystems [@del_grosso_global_2008], and as a result achieve significant carbon sequestration and storage.
@@
 
-Table: Definitions of variables used in analysis
+Table: Definitions and sample sizes of variables used in analysis
 
@@
 In addition to this data, climate data for each site was extracted from five open-access climate datasets based on site geographic co-ordinates.
+Where site-level data was missing for mean annual temperature and/or mean annual precipitation, data was extracted from the WorldClim dataset.
 
@@
 Table: Sources of climate data
+
+Additionally, two climate variables were derived from the above datasets: maximum vapour pressure deficit, defined as the vapour pressure deficit of the month with the largest deficit; and water stress months, defined as the number of months annually where precipitation was lower than potential evapotranspiration.
+
+*Length of growing season.* Growing season months were defined as months with mean minimum temperature > 0.5.
+Growing season months were initially calculated following methods used by Kerkhoff et al.
+(2005), which additionally required that growing season months had a moisture index, defined as (MAT - PET)/PET, > -0.95.
+Michaletz et al.
+(2014) included an equivalent requirement in their calculation of growing season length.
+However, we found that including this requirement had no effect on the estimates of growing season length, and so chose to exclude it.
+
+Monthly data for PET, precipitation, and temperature was downloaded from the Climate Research Unit (CRU) time-series dataset v 4.03 [@harris_updated_2014], and for solar radiation from WorldClim2 [@fick_worldclim_2017], and used to calculate mean monthly PET, precipitation, temperature and solar radiation during the growing season.
+Total growing season precipitation and solar radiation were also calculated.
 
@@
 Results were similar between the weighted and unweighted model, so, to avoid problems of over-fitting, the weighted model was dropped, and results from this are not presented here.
+
+Models were run for total annual productivity against annual climate variables, and for monthly growing season productivity, defined as total productivity/length of growing season, against growing season climate variables.
 

### 2019-08-22 manuscript/manuscript_removed_text.Rmd
Create manuscript_removed_text.Rmd

--- 2019-08-12 manuscript/manuscript draft.Rmd
@@
 ---
-title: "Manuscript draft"
-output:
-  pdf_document:
-    citation_package: natbib
-    fig_caption: yes
-bibliography: [library.bib,  packages.bib]
-biblio-style: apalike
-csl: apa.csl
-header-includes:  \usepackage{float}
-                  \usepackage{caption}
-                  \captionsetup[table]{font=footnotesize}
-                  \captionsetup[figure]{font=footnotesize}
-
+title: "manuscript_removed_text"
+output: pdf_document
 ---
-
-```{r, include=FALSE}
-options(tinytex.verbose = TRUE)
-```
-
-```{r include=FALSE}
-knitr::write_bib(c(
-  .packages(), 'knitr', 'rmarkdown', 'lme4', 'MuMIn', 'plyr', 'raster', 'ncdf4', 'Hmisc', 'plyr', 'merTools', 'visreg', 'r2glmm', 'nlme', 'viridis', 'ggplot2', 'ggpubr', 'gridExtra', 'AICcmodavg'
-), 'packages.bib')
-```
-
-```{r results="asis", include=FALSE}
-cat("
-<style>
-caption {
-      font-size: 0.75em;
-    }
-</style>
-")
-```
-
-###Abstract
-
-Globally, forests are a significant carbon sink, showing high levels of productivity and representing an important component of the carbon cycle.
-Current knowledge indicates that forest productivity decreases with latitude; however, there is evidence that this relationship varies across different fluxes.
-Furthermore, the climate variables that drive this relationship are poorly understood.
-Here, we use a comprehensive global database of forest carbon fluxes to show that, while all major carbon fluxes decrease with latitude, allocation to woody productivity increases with latitude, and allocation to root productivity declines.
-The best predictors of global patterns of productivity are temperature variables, including mean annual temperature, annual temperature range, and temperature seasonality, with the exception of woody productivity, which is most strongly influenced by potential evapotranspiration and vapour pressure deficit.
-Climate variables explain a large proportion of the variation in major carbon fluxes, but are a less significant predictor of subsidiary components.
-Our results illustrate the strong influence of climate on primary productivity, and particularly the importance of temperature in determining forest productivity.
-However, they also indicate that effects of climate are complex: fluxes show non-linear responses to climate variables, and many factors other than climate interact to determine allocation to subsidiary components.
-Knowledge of the relative influence of climate variables on forest carbon fluxes is an important component of understanding the responses of primary productivity to climate change.
-
-### Introduction
-
-Globally, forests play an important role in the carbon cycle and are an important component of global carbon dioxide budgets [@luyssaert_old-growth_2008].
-They show higher levels of productivity than non-forest  terrestrial ecosystems [@del_grosso_global_2008], and as a result achieve significant carbon sequestration and storage.
-Estimating the total role of forests in the carbon cycle is challenging, but studies indicate that old growth forests alone sequester up to 1.4 GtCyr^-1^ [@malhi_carbon_1999], while the total sequestration of carbon by established forests globally could be up to 2.4 GtCyr^-1^, with the largest sinks being in old-growth tropical forests [@pan_large_2011].
-As atmospheric carbon dioxide levels continue to rise, with consequences for global climate, there is increasing recognition that proper protection and management of forest resources will have an important role to play in mitigating climate change.
-Understanding the patterns of forest productivity on a global scale, and the drivers behind them, is therefore a priority in forest research.
-
-There are two major questions to understand: firstly, how forest productivity varies globally - and specifically which areas show the greatest peaks in productivity -; and secondly, which climate variables drive this variation.
-On a global scale, the productivity of forests varies with latitude, with a general trend of increasing productivity towards the tropics [@beer_terrestrial_2010; @jung_global_2011]; however the exact nature of this pattern, and how it varies by component of productivity, is poorly understood.
-This latitudinal gradient is most likely to be explained by climatic gradients in temperature, precipitation, length of growing season, and combinations of the above.
-Productivity is influenced by a range of climatic drivers, including mean annual temperature (MAT) and mean annual precipitation (MAP) [@del_grosso_global_2008], but doesn't necessarily respond linearly to these drivers.
-Disentangling the shape of productivity responses to climate drivers will enable better predictions of future responses under climate change.
-
-*The components of forest* During photosynthesis, plants capture carbon dioxide from the atmosphere.
-The gross primary productivity (GPP) of an ecosystem is the gross uptake, via photosynthesis, of carbon dioxide by plants in that ecosystem.
-Only a fraction of the carbon captured is assimilated into plant tissue; the rest is used in autotrophic respiration (R~auto~).
-The component of GPP that is stored as plant material is the net primary productivity (NPP) of an ecosystem.
-Net primary productivity can therefore be expressed as:
-
-$$ NPP = GPP + R~auto~ $$
-
-Currently, GPP cannot be measured directly by observing total ecosystem photosynthesis.
-Instead field estimates of GPP have to be derived based on modelling and extrapolation of eddy-covariance studies and measurements of net ecosystem exchange (NEE) [@clark_reviews_2017].
-
-In contrast, NPP can be calculated through direct field observations.
-In order to achieve greater accuracy in estimating NPP, NPP is often broken down into its component parts, with aboveground NPP (ANPP) and belowground NPP (BNPP) considered separately to each other.
-The components included when estimating ANPP and BNPP often vary between studies, depending on the intensity of fieldwork effort.
-At its most basic level, ANPP can be expressed as:
-
-$$ ANPP = NPP~stem~ + NPP~branch~ + NPP~canopy~ $$
-where NPP~stem~ is the annual woody increment of all stems above a specified diameter at breast height (DBH), NPP~branch~ is annual woody branch turnover, and NPP~canopy~ is annual foliage production, including leaves, twigs, and reproductive structures.
-ANPP may also include NPP~VOC~, the annual emission of volatile organic compounds, and NPP~herbivory~, the annual consumption of plant matter by herbivores, but these components are often excluded from field observations as they are much harder to quantify.
-Other components of aboveground productivity that remain largely unquantified include epiphytes, hemiepiphytes, and understory plants [@clark_reviews_2017].
-The majority of current ANPP estimates are based on the assumption that the contribution of these components to overall NPP is insignificant.
-
-There are two major subcomponents of BNPP, which can be expressed as:
-
-$$ BNPP = NPP~coarse root~ + NPP~fine root~ $$
-where NPP~coarse root~ is the annual production of coarse roots (typically roots >2mm diameter), and NPP~fine root~ is the annual production of fine roots (typically roots <2mm diameter) [@aragao_above-_2009].
-Calculations of BNPP may also include NPP~exudation~, a measure of annual carbon losses through root exudation, and NPP~symbionts~, the annual carbon allocation to mycorrhizae and legumes, but, as before, this is challenging to quantify and is often excluded from field observations.
-
-BNPP is a poorly understood component of total ecosystem productivity, primarily because of the challenges in gaining accurate field measurements.
-Coarse root productivity is often estimated via extrapolation of NPP~stem~ estimates using allometries that may not have been empirically verified [@clark_net_2001].
-NPP~fine root~ is easier to quantify through soil cores and minirhizotrons, however, sampling tends to be limited to the surface soils, with very few studies sampling to depths below 3 metres [cite].
-As a result, it is possible that BNPP is currently significantly underestimated, despite being a hugely significant component of total ecosystem productivity [@pan_large_2011].
-
-*Which factors influence primary productivity?* Primary productivity is influenced by many factors, which often act across a range of scales, and may show interactive effects with each other.
-On a local scale, stand age [@litton_carbon_2007; @gillman_latitude_2015], management [@simova_enigma_2017]; nutrient availability [@aragao_above-_2009]; and altitude [@girardin_net_2010; @malhi_variation_2017] all impact forest productivity.
-On a global scale, changes in primary productivity are influenced by climatic variables and abiotic gradients, such as the length of growing season [@michaletz_convergence_2014].
-There is some debate over the precise relationship between these drivers and productivity; While mean annual precipitation (MAP) and mean annual temperature (MAT) have been argued to be significant predictors of productivity [@chu_does_2016], other studies have found that the correlation between productivity and MAT is a factor of the relationship between productivity and growing season length [@kerkhoff_plant_2005; @malhi_productivity_2012; @michaletz_convergence_2014; @michaletz_drivers_2018].
-Improving understanding of how these factors interact to control global patterns in primary productivity is essential to understanding the global carbon cycle.
-
-Current research into how primary productivity varies with latitude is inconclusive, and - though it has primarily focussed on patterns of GPP, NPP, and ANPP - indicates that different components of productivity may show different relationships to latitude.
-Gross primary productivity is generally thought to be highest in the tropics.
-Modelling of global terrestrial ecosystem GPP though upscaling and calibration of eddy flux measurements indicates a peak in GPP in the tropics, with the highest levels in tropical forests [@beer_terrestrial_2010; @jung_global_2011].
-This is corroborated by analysis of site-level GPP measurements, which show a strong positive correlation between GPP and MAT and MAP [@luyssaert_co2_2007], with the highest GPP values reported in tropical forests.
-The influence of latitude on global patterns of NPP is less clear than that of GPP.
-Simova and Storch (2017) found that, as with GPP, NPP decreases with latitude, peaking in the tropics.
-However, other studies have found the highest values of NPP in temperate forests [@luyssaert_co2_2007; @huston_global_2009].
-Because of the challenges in accurately obtaining unbiased measures of belowground productivity, many studies focus on ANPP in preference to measures of NPP.
-Studies on global patterns of ANPP are equally inconclusive: Gillman et al.
-(2015) found a weak negative relationship between ANPP and latitude, with the relationship becoming stronger in older forest stands.
-These findings were echoed in other studies, which have found weak or no relationships between ANPP and latitude [@huston_global_2009].
-
-Furthermore, there is evidence that different components of productivity show individual responses to drivers of productivity.
-For example, increases in GPP have been reported to saturate above 25$^\circ$C MAT [@larjavaara_temperature_2012], while increases in NPP are recorded to saturate above 10$^\circ$C MAT [@luyssaert_co2_2007].
-Similarly, allocation to different components of ANPP varies with climate.
-Within the tropics, allocation to canopy NPP appears fairly consistent, with significantly greater variation in allocation to woody and belowground NPP [@litton_carbon_2007; @malhi_allocation_2011].
-Allocation to these structural biomass components has been shown to increase with water availability [@litton_carbon_2007; @bloom_decadal_2016], and is highest in the wet tropics, indicating that control of woody productivity by MAP may be more significant than control of foliar productivity.
-However, these studies are regional, meaning that our understanding of variation in allocation and its relationship to climate on a global scale remains limited.
-In addition, allocation is also influenced by stand age [@de_lucia_forest_2007], nutrient availability [@litton_carbon_2007] and forest structure [@taylor_greater_2019], which can make it challenging to disentangle the effects of climate.
-
-Data that control for stand age and standardize methodologies are required to resolve this question.
-Here, we use a comprehensive global database of forest carbon fluxes to explore global patterns in productivity.
-We explore three questions:
-1.
-**Which climatic variables are the most important drivers behind the latitudinal pattern in primary productivity?** To date, the majority of studies have focussed on productivity responses to precipitation and temperature, while the influence of other climate variables on productivity remains under-explored.
-We utilise global climate datasets to investigate the relationships between productivity and a range of climate variables.
-2.
-**Do the components of primary productivity show variation in their responses to these climatic drivers?** Allocation to components of primary productivity appears to show some variation across climate and latitudinal gradients.
-We use global datasets of a range of primary productivity components to explore how significant this variation in allocation is.
-3.
-**Does climate explain the same proportion of variation in different components of primary productivity?** Review of current research shows that, while the relationship between GPP and latitude is consistent, NPP and ANPP show weaker relationships with latitude.
-This may be a result of varying sensitivities to climate drivers.
-We use mixed models to estimate the proportion of variation explained by climate for different components of productivity, to investigate how the relative importance of climatic drivers varies across components.
-
-### Methods
-
-Analyses were conducted on data contained in the open-access ForC database [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc:_2018].
-This database contains records of field-based measurements of forest carbon stocks and annual fluxes, compiled from original publications and existing data compilations and databases.
-Associated data, such as stand age, measurement methodologies, and disturbance history, are also included.
-For each site, site geographic co-ordinates were used to extract the Koeppen-Geiger zone from the ESRI Koeppen-Geiger map [cite], and the FAO Global Ecological Zone from the ???.
-Additional targeted literature searches were conducted to identify any further available data on primary productivity, with particular focus on old-growth forests in temperate and boreal regions.
-ForC currently contains 29768 records from 5227 plots, representing 20 distinct ecozones across all forested biogeographic and climate zones.
-
-```{r echo=FALSE, out.width='100%', fig.cap = "Map showing all data used in the analysis, coded by variable", fig.pos='H'}
-  knitr::include_graphics("distribution_all_variables.png")
-```
-*Data selection.* Over 50 variables of forest carbon stocks and annual fluxes are represented in the ForC database; this analysis focussed on measures of primary productivity.
-Table 1 contains details of the variables selected for analysis.
-
-| Variable | Definition | Components included | Methodologies used | Number of records | Number of plots |
-|--------------|-------------------|-------------|------------------------|-------|-------|
-| GPP | Annual gross primary production; annual uptake of carbon dioxide by an ecosystem | NA | Flux partitioning of eddy covariance | 257 | 80 |
-| NPP | Annual net primary production; the component of GPP that is stored in plant tissue; GPP minus ecosystem respiration | Foliage, branch, stem, coarse root, fine root and optionally understory | Direct measurement of annual increments of components | 93 | 78 |
-| ANPP | Aboveground net primary production | Foliage, stem, and optionally branch | Direct measurement of annual increments of components | 258 | 166 |
-| ANPP~foliage~ | Net primary production of foliage | Foliage | Direct measurement of litterfall, correcting for changes in leaf biomass when measured | 98 | 94 |
-| ANPP~woodystem~ | Net primary production of woody stems | Woody stems | Direct measurement of stem growth increment | 266 | 252 |
-| BNPP~root~ | Belowground net primary production | Coarse and fine roots | Direct measurement of one or more of: fine root turnover, soil cores, root ingrowth cores, minirhizotrons; indirect estimates of coarse roots using allometries based on aboveground stem increment measures | 107 | 87 |
-| BNPP~fineroot~ | Net primary production of fine roots | Fine roots | Direct measurement of one or more of: minirhizotrons, fine root turnover, soil cores, root ingrowth cores | 90 | 77 |
-| R~auto~ | Annual autotrophic respiration, including above- and belowground components | Foliage, stem, and root | Chamber measurements of component gas exchange | 22 | 22 |
-| R~root~ | Annual root respiration | Coarse and fine roots | Measurement of root gas exchange | 65 | 43 |
-
-Table: Definitions of variables used in analysis
-
-A subset of the ForC database was generated for the purposes of this analysis, in order to control for data quality and remove biasing factors.
-Since management can alter observed patterns of primary productivity [@simova_enigma_2017], sites were excluded from analysis if they were managed, defined as plots that were planted, managed as plantations, irrigated, fertilised or including the term "managed" in their site description.
-Sites that had experienced significant disturbance were also excluded.
-Disturbances that justified site exclusion were major cutting or harvesting, and/or burning, flooding, drought and storm events with site mortality >10% of trees.
-Grazed sites were retained.
-There is evidence that stand age influences patterns of primary productivity and carbon allocation in forest ecosystems, and can confound relationships between latitude and primary productivity [@de_lucia_forest_2007; @gillman_latitude_2015].
-To reduce any biasing effects of stand age, stands under 100 years of age were excluded from analysis.
-Sites for which stand age was unknown were excluded from analysis.
-
-*Methodological consistency.* The data in ForC is derived from a range of studies, often employing different methods.
-For this reason, criteria were introduced to standardise for differences in methodology.
-Where data was based on forest plot census measurements, studies which used a minimum diameter at breast height (DBH) measure of >10cm were excluded from analysis.
-It would be preferable to standardise by minimum area sampled; however x% of plots in the database are 1 ha or under in size; excluding these plots would place significant constraints on sample size.
-
-As discussed above, estimates of NPP, ANPP, and BNPP are generated through summing measurements of their component parts.
-Since the components included in productivity estimates vary between studies, estimates of productivity were classified within the ForC database according to their components, and then filtered for analysis.
-Estimates of NPP were selected if they included foliage, branch, stem, coarse root, and fine root.
-Measures of NPP which included additional components, including understorey, volatile organic compounds (VOCs), exudates, estimates of NPP lost to herbivory, and the NPP of reproductive structures, were excluded.
-Estimates of ANPP were selected if they included foliage, stem growth and optionally branch turnover.
-Any measures of primary productivity where components were unknown were excluded from analysis.
-
-*Climate datasets.* Where site-level data on mean annual temperature, mean annual precipitation, and latitude were available in the primary literature, this data was compiled and entered directly into the ForC database.
-In addition to this data, climate data for each site was extracted from five open-access climate datasets based on site geographic co-ordinates.
-
-| Database | Variables downloaded | Citation |
-|-----------------------|--------------------------------------------|---------------------|
-| WorldClim | Mean annual temperature; temperature seasonality; annual temperature range; mean annual precipitation | [@hijmans_very_2005] |
-| WorldClim2 | Solar radiation | [@fick_worldclim_2017] |
-| Climate Research Unit (CRU) time-series dataset v 4.03 | Cloud cover; annual frost days; annual wet days; potential evapotranspiration | [@harris_updated_2014] |
-| Global Aridity Index and Potential Evapotranspiration Climate Database | Aridity; potential evapotranspiration | [@trabucco_global_2018] |
-| TerraClimate | Vapour pressure deficit | [@abatzoglou_terraclimate_2018] |
-
-Table: Sources of climate data
-
-*Model specification.* The effects of climate and latitude on primary productivity were analysed using mixed effects models using the package 'lme4' [@bates_fitting_2015] in R v.3.5.1 [@r_core_team_r:_2018].
-The effect of each extracted climate variable on each measure of primary productivity was modelled by specifying the climate variable as a fixed effect.
-For each climate variable, three models were specified: a null model; a model with the climate variable as a linear term; and a model with the climate variable as a polynomial term.
-AIC values were calculated for the models and used to select the best model.
-R^2^ values were calculated for the best model.
-All R^2^ values presented here are marginal R^2^ values, and refer to the proportion of variation explained by only the fixed effects, unless otherwise specified.
-In addition, slope coefficients were calculated for the linear models.
-
-Because the magnitude of fluxes varies significantly, in order to facilitate comparisons between regression models for each flux, data for each flux was scaled, to give the data a mean of 0 and standard deviation of 1.
-As each data set was scaled separately, this does not allow for statistical comparisons of slope values, but does assist in visualising the data.
-
-To test for a potential influence of altitude, models were also run with site altitude included as a second fixed effect.
-These models were compared against models with no altitude term, and AIC values calculated to identify whether inclusion of altitude as a term improved the models.
-
-Within the ForC database, sites within 25km^2^ of each other are clustered into geographic areas.
-To account for correlations in measurements between tightly clustered sites, a random effect was specified as plot nested within geographic area.
-Data from the temperate regions was heavily skewed towards studies from the old-growth forests of the Pacific Northwest.
-These forests have very high productivity, and so to ensure that results were not unduly influenced by geographic sampling bias, we tried a version of the model where data were weighted according to forested land area within each Koeppen climate zone.
-Results were similar between the weighted and unweighted model, so, to avoid problems of over-fitting, the weighted model was dropped, and results from this are not presented here.
-
-To investigate the potential interactive effects of climate variables on carbon fluxes, multivariate models were also specified.
-To ensure that models were biologically meaningful, the terms included in the models tested built on results from the univariate models.
-Modelling of individual climate variables identified that the best predictors of carbon fluxes were variables related to temperature.
-We therefore decided to include mean annual temperature as a term in all multivariate models.
-We first modelled the interaction effect between mean annual temperature and mean annual precipitation, in order to capture climate variation along the axes of temperature and water availability.
-Models were tested for a significant interactive effect and a significant additive effect.
-We then explored whether any other climate variable, in combination with mean annual temperature, could significantly improve on the combination of mean annual temperature and mean annual precipitation.
-In specifying the range of models to test, climate variables which were strongly correlated with temperature were dropped, in order to capture the greatest range of variation in climate.
-For each possible pairing of climate variables, two models were specified: a model with the two climate variables showing an additive effect; and a model with the two climate variables showing an interactive effect.
-As before, plot nested within geographic area was included as a random effect.
-Altitude was not considered.
-AIC values were calculated for the models, and used to compare models.
-Models were considered to be significantly better than the baseline MAT*MAP model if:
-  i) the AIC value of the model was smaller than the AIC value of the baseline model by >2  
-  ii) the r-squared value of the model was larger than the r-squared value of the baseline model by >5
-
-*Validating models of component fluxes.* Comparison of component fluxes is based on the assumption that components sum accurately to estimates of larger fluxes.
-To test this, components of larger fluxes were regressed against latitude, and the models used to generate a series of point estimates along lines of best fit for each component.
-The point estimates for smaller component fluxes were summed to generate new "stacked" estimates of larger fluxes, which were then compared against actual measurements of the larger flux.
-Confidence intervals for the larger flux were calculated using the 'bootMer' function from the lme4 package [@bates_fitting_2015].
-Stacked plots were generated for:
-1. GPP = NPP + R~auto~  
-2. NPP = ANPP + BNPP  
-3. ANPP = ANPP~foliage~ + ANPP~woody stem~  
-4. Total belowground carbon flux = BNPP + R~root~
-
-*Allocation to carbon fluxes along latitudinal gradients.* Variation in allocation to component carbon fluxes along latitudinal gradients was explored for a range of pairings: firstly, GPP:NPP, ANPP:BNPP, and ANPP~foliage~:ANPP~woody stem~; and secondly, the ratio of NPP to each of ANPP, BNPP, ANPP~foliage~, and ANPP~woody stem~.
-For each set of paired fluxes, measurements taken at the same site and plot, and in the same year, were paired together, and the ratio of each pair of measurements calculated.
-The ratios were regressed against latitude and climate variables, using the linear model specified above.
-Cook's distance analyses were carried out for each of the models, and indicated that data from a few high-elevation sites were having a disproportionate influence on the regressions.
-To account for this, models were re-run using only data from sites $\leq$ 1000m.
-
-###Results
-
-*How do carbon fluxes vary with latitude?* All major carbon fluxes increase with decreasing latitude (fig.
-2).
-Models showed significant variation in whether they were improved by inclusion of altitude as a fixed effect.
-Including altitude had a very small effect on most models, with the difference in AIC values between models including and excluding altitude often being <2, suggesting the models are very similar in their explanatory power.
-All models presented below do not include altitude, and R^2^ values refer to the proportion of variation explained by the climate variable alone.
-
-```{r echo=FALSE, out.width='100%', fig.cap = "Graphs to show primary productivity $(MgC~ha^-1~yr^-1)$ regressed against latitude.
-Lines of best fit are plotted according to the best model selected during analysis.
-All regressions are significant $(p<0.05)$.", fig.pos='H'}
-  knitr::include_graphics("effect_of_latitude.png")
-```
-
-Latitude is a strong predictor for many of the carbon fluxes, explaining 64% of variation in GPP (n = 254, p<0.0001), 50% in NPP (n = 114, p<0.0001) and 45% in ANPP (n = 259, p<0.0001).
-For all fluxes, their relationship with latitude was best predicted by the linear model.
-
-Analysis of component fluxes indicates that measures of smaller component fluxes do approximately sum to measures of larger fluxes (fig.
-3).
-Modelled estimates of GPP, generated from the sum of NPP and R auto; NPP, generated from the sum of ANPP and BNPP~root~; and ANPP, generated from the sum of ANPP~foliage~ and ANPP~woody stem~, fall completely within the confidence intervals of the regressions of field estimates of GPP, NPP and ANPP respectively.
-
-```{r echo=FALSE, out.width='100%', fig.cap = "Graphs of primary productivity $(MgC~ha^-1~yr^-1)$ regressed against latitude.
-Lines of best fit are plotted according to the best model selected during analysis.
-All regressions are significant $(p<0.05)$.
-Plots 1 - 3 show two component fluxes; a larger flux, defined as the combination of the two component fluxes; and a modelled estimate of the sum of the two component fluxes.
-95\\% confidence intervals are plotted for the larger flux.
-Plot 4 shows three belowground fluxes, and a modelled estimate of the total belowground carbon flux", fig.pos='H'}
-
-  knitr::include_graphics("combined_stacked.png")
-```
 
@@
 BNPP:NPP decreased with latitude, though this relationship was non-significant.
-
-*Does climate explain the same proportion of variation in different components of primary productivity?*
-
-| Variable | Latitude | Mean annual temperature | Temperature seasonality | Potential evapotranspiration | MAT*MAP |
-|----------|----------|----------|----------|--------------|----------|
-| R~auto~ | 0.65 | 0.77 | 0.62 | 0.63 | NA |
-| GPP | 0.64 | 0.63 | 0.72 | 0.36 | 0.64 |
-| NPP | 0.50 | 0.52 | 0.52 | 0.33 | 0.56 |
-| ANPP | 0.45 | 0.46 | 0.42 | 0.28 | 0.48 |
-| ANPP~foliage~ | 0.54 | 0.63 | 0.54 | 0.35 | 0.64 |
-| ANPP~woodystem~ | 0.17 | 0.23 | 0.12 | 0.19 | 0.28 |
-| BNPP~root~ | 0.33 | 0.26 | 0.23 | 0.36 | 0.35 |
-| BNPP~fineroot~ | 0.15 | 0.11 | 0.16 | 0.11 | 0.19 |
-| R~root~ | 0.22 | 0.21 | 0.30 | 0.01* | 0.28 |
-
-Table: R^2^ values presented for latitude, three climate variables, and the best multivariate model for each carbon flux. Values marked * are not significant.
-
-R^2^ values are generally highest in the major fluxes, and decrease in subsidiary fluxes.
-Of the major fluxes, R~auto~ and GPP are the most strongly explained by latitude and climate.
-Mean annual temperature explains 77% of variation in R~auto~, while temperature seasonality explains 72% of variation in GPP.
-The proportion of variation explained by climate and latitude decreases in NPP and ANPP.
-The climatic variables with the strongest explanatory power explain around 50% of the variation in these fluxes.
-Of the major fluxes, BNPP~root~ is the least well explained by climate and latitude, with climate explaining at most 36% of variation.
-
-With the exception of ANPP~foliage~, the proportion of variation explained by climate and latitude in subsidiary fluxes is much lower.
-Climate explains at most 23% of variation in ANPP~woody stem~, 16% in BNPP~fine root~, and 30% in R~root~.
-In contrast, climate strongly explains variation in ANPP~foliage~, with mean annual temperature explaining 63% of variation.
-
-This pattern is also seen in the R^2^ values for multivariate models.
-
-###Discussion
-
-*The relationship between productivity and latitude*
-
-Past studies have differed in their conclusions regarding the relationship between productivity and latitude.
-Our findings indicate that, for all carbon fluxes studied, productivity peaks in the tropical regions.
-While Huston and Wolverton (2009) found no significant relationship between ANPP and latitude, we find a significant negative relationship between ANPP and latitude, with the highest levels of ANPP in the tropical regions.
-
-*Challenges in estimating NPP*  
-The impact of excluding certain components of NPP on this analysis is unclear, primarily because these components are poorly understood and rarely measured [@clark_net_2001].
-Do we want to include a section assessing how much this may or may not impact results?
-1. Hard to measure BNPP because it often isn't sampled deep enough; impact on ratio tests?  
-2. Unclear if there are latitudinal patterns e.g. of herbivory/VOCs  
-  + Herbivory can be an important component of total NPP so this may be important to account for; likely to be higher in the tropics?

### 2019-08-22 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@
 These models were compared against models with no altitude term, and AIC values calculated to identify whether inclusion of altitude as a term improved the models.
+Including altitude had a very small effect on most models, with the difference in AIC values between models including and excluding altitude often being <2, suggesting the models are very similar in their explanatory power.
+As a result, it was decided to present results only from models do not include altitude as a fixed term.
 
@@
 
-*How do carbon fluxes vary with latitude?* All major carbon fluxes increase with decreasing latitude (fig.
-2).
-Models showed significant variation in whether they were improved by inclusion of altitude as a fixed effect.
-Including altitude had a very small effect on most models, with the difference in AIC values between models including and excluding altitude often being <2, suggesting the models are very similar in their explanatory power.
-All models presented below do not include altitude, and R^2^ values refer to the proportion of variation explained by the climate variable alone.
+*How do carbon fluxes vary with latitude?* All major carbon fluxes increase linearly with decreasing latitude (fig. 2).
 
@@
 
-*Which climatic variables are the most important drivers behind the latitudinal pattern in primary productivity?* All major carbon fluxes show similar responses to climate variables.
-Of the climate variables tested, annual wet days, aridity, cloud cover, mean diurnal temperature range and precipitation seasonality were poor explainers of variation in primary productivity, explaining less than 20% of the variation in each of the carbon fluxes.
-We found that both measures of mean annual precipitation used were significant (p<0.05) but poor predictors of productivity, explaining, with the exception of R~auto~, at most 37% of variation in carbon flux.
+We find no evidence that allocation between fluxes varies with latitude or climate.
 
-Looking across all productivity components, the climate variables which explain the highest amount of variation in carbon fluxes are temperature seasonality, mean annual temperature, annual temperature range, and annual frost days.
-Vapour pressure deficit came out as a strong predictor for ANPP~woody stem~, and potential evapotranspiration was a good predictor for BNPP~root~ and ANPP~woody stem~.
+*How does productivity relate to MAT and MAP?* MAT is a significant (p<0.05) and strong predictor of productivity for all carbon fluxes tested, with all fluxes showing a linear increase with temperature.
+We found no support for a saturation point of productivity with temperature.
 
-```{r echo=FALSE, out.width='100%', fig.cap = "Plots of primary productivity $(MgC~ha^-1~yr^-1)$ regressed against the climate variables identified as the best predictors: mean annual temperature; temperature seasonality; annual temperature range; and vapour pressure.
-Lines of best fit are plotted according to the best model selected during analysis.
-All regressions are significant $(p<0.05)$.", fig.pos='H'}
-  knitr::include_graphics("combined_plots.png")
-```
-
-**Mean annual temperature.** All carbon fluxes show a linear increase with temperature.  
-**Temperature seasonality and annual temperature range.** These two climate variables are strongly correlated, and so only temperature seasonality is presented in figure 4.
-Productivity is highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower.
-GPP, NPP, ANPP, and R~root~ show a polynomial relationship with both variables, and BNPP~root~ shows a polynomial relationship with annual temperature range.
-As seasonality increases, these variables show a more rapid decrease in productivity.
-ANPP~foliage~ and ANPP~woody stem~ show a linear relationship with both variables.
-**Mean annual precipitation.** Despite being a poor predictor of productivity, the results indicate that, for the majority of fluxes, productivity increases with precipitation, up until a saturation point at between 3000 and 4000mm annual precipitation, above which productivity starts to decrease.
+MAP was found to be a significant (p<0.05) but poor predictor of productivity, explaining, with the exception of R~auto~, at most 37% of variation in carbon flux.
+For the majority of fluxes productivity is best predicted by a polynomial model.
+Productivity increases with precipitation, up until a saturation point at between 3000 and 4000mm annual precipitation, above which productivity starts to decrease.
 The notable exception to this is GPP: the model indicates that GPP continues to increase with precipitation up to measures of at least 5000mm annually (p<0.0001, R^2^ = 0.33.
 Data above this point is not available, but the model trend indicates that the saturation point for this model is around 5000mm mean annual precipitation.
-**Potential evapotranspiration.** All fluxes, with the exception of BNPP~fine root~ and R~root~ show a strong increase with potential evapotranspiration at values below 1000mm.
-Between 1200 and 1700mm productivity peaks, and begins to decrease at values above 1800mm.
 
@@
 
-*Do carbon fluxes vary in their responses to climate, particularly with regards to allocation between fluxes?* Of all the flux pairings tested, the only ones to show a significant relationship with latitude were ANPP:NPP (p<0.01, R^2^ = 0.44) and ANPP:BNPP (p<0.001, R^2^ = 0.33).
-In both cases, allocation to ANPP increased with latitude, with the lowest allocation to ANPP in the tropical regions.
-BNPP:NPP decreased with latitude, though this relationship was non-significant.
+*How does productivity relate to other climate variables?* Our results indicate that productivity is most strongly explained by temperature, with temperature-related climate variables coming out as strong predictors of productivity.
+In addition to MAT, temperature seasonality, annual temperature range, and annual frost days were consistently identified as good predictors of productivity across fluxes.
+
+We found a significant relationship between productivity and potential evapotranspiration for all fluxes.
+ANPP~foliage~, BNPP~fine root~ and R~root~ increased linearly with PET, however all other fluxes showed a polynomial relationship with PET.
+We find strong evidence for a saturation point or peak with PET: productivity tends to increase at values below 1000mm, before saturating between 1200 and 1700mm.
+There is evidence that productivity begins to decrease at values above 1800mm PET.
+
+Solar radiation
+Vapour pressure deficit?
+
+Of the climate variables tested, annual wet days, aridity, cloud cover, mean diurnal temperature range, precipitation seasonality, maximum vapour pressure deficit and water stress months were poor or non-significant explainers of variation in productivity, explaining less than 20% of the variation in each of the carbon fluxes.
+
+*What is the role of seasonality in explaining productivity?*
 
@@
 
+Begin discussion by emphasising that this these conclusions are based on analyses of an unprecedented amount of data across many fluxes, limited to mature undisturbed forests.
+
+We are no longer explicity testing whether climate explains a higher proportion of variability in major fluxes, so mention this, and the implications of it in the discussion.
+
 Past studies have differed in their conclusions regarding the relationship between productivity and latitude.

### 2019-08-22 scripts/final_model_no_weighting.R
Update final_model_no_weighting.R

@@ -184,7 +184,7 @@ for(response.variables in response.variables.groups){
         ylim[1] <- ylim[1] - 0.25
         ylim[2] <- ylim[2] + 0.25
         
-        if(!fixed.v %in% c("mat", "lat")){
+        if(!fixed.v %in% c("mat", "lat", "PreSeasonality")){
         
         mod <-  lmer(scale(mean) ~ 1 + (1|geographic.area/plot.name), data = df, REML = F)
         mod.clim <- lmer(scale(mean) ~ poly(fixed, 1, raw = T) + (1|geographic.area/plot.name), data = df, REML = F)
@@ -209,7 +209,7 @@ for(response.variables in response.variables.groups){
         if (best.model == "mod.clim.poly") mod.full <- lmer(scale(mean) ~ poly(fixed, 2, raw = T) + (1|geographic.area/plot.name), data = df, REML = T)
         }
         
-        if(fixed.v %in% c("mat", "lat")){
+        if(fixed.v %in% c("mat", "lat", "PreSeasonality")){
           
           mod <-  lmer(scale(mean) ~ 1 + (1|geographic.area/plot.name), data = df, REML = F)
           mod.clim <- lmer(scale(mean) ~ poly(fixed, 1, raw = T) + (1|geographic.area/plot.name), data = df, REML = F)

### 2019-08-22 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@
 AIC values were calculated for the models and used to select the best model.
+If the best model included a polynomial term, the shape of the polynomial relationship was considered.
+If the shape of the relationship made biological sense, and was a significant improvement on the linear relationship (deltaAIC >2), we accepted the polynomial as the best model.
+If not, we ran the linear model as the final model.
 R^2^ values were calculated for the best model.
@@
 Models were run for total annual productivity against annual climate variables, and for monthly growing season productivity, defined as total productivity/length of growing season, against growing season climate variables.
+For analyses on data within the growing season, only linear models were specified.
 
@@
 
-*How do carbon fluxes vary with latitude?* All major carbon fluxes increase linearly with decreasing latitude (fig. 2).
-
-```{r echo=FALSE, out.width='100%', fig.cap = "Graphs to show primary productivity $(MgC~ha^-1~yr^-1)$ regressed against latitude.
-Lines of best fit are plotted according to the best model selected during analysis.
-All regressions are significant $(p<0.05)$.", fig.pos='H'}
-  knitr::include_graphics("effect_of_latitude.png")
-```
-
-Latitude is a strong predictor for many of the carbon fluxes, explaining 64% of variation in GPP (n = 254, p<0.0001), 50% in NPP (n = 114, p<0.0001) and 45% in ANPP (n = 259, p<0.0001).
-For all fluxes, their relationship with latitude was best predicted by the linear model.
-
-Analysis of component fluxes indicates that measures of smaller component fluxes do approximately sum to measures of larger fluxes (fig.
+*Closure of fluxes* Analysis of component fluxes indicates that measures of smaller component fluxes do approximately sum to measures of larger fluxes (fig.
 3).
@@
 
+*How do carbon fluxes vary with latitude?* All major carbon fluxes increase linearly with decreasing latitude (fig. 2).
+
+```{r echo=FALSE, out.width='100%', fig.cap = "Graphs to show primary productivity $(MgC~ha^-1~yr^-1)$ regressed against latitude.
+Lines of best fit are plotted according to the best model selected during analysis.
+All regressions are significant $(p<0.05)$.", fig.pos='H'}
+  knitr::include_graphics("effect_of_latitude.png")
+```
+
+Latitude is a strong predictor for many of the carbon fluxes, explaining 64% of variation in GPP (n = 254, p<0.0001), 50% in NPP (n = 114, p<0.0001) and 45% in ANPP (n = 259, p<0.0001).
+For all fluxes, their relationship with latitude was best predicted by the linear model.
+
 We find no evidence that allocation between fluxes varies with latitude or climate.
+There were no significant results from regressing ratios of carbon fluxes against latitude, or against any of the climate variables.
 
@@
 
-*What is the role of seasonality in explaining productivity?*
+*What is the role of seasonality in explaining productivity?* Temperature seasonality is a significant predictor of productivity.
+We found a polynomial relationship with productivity for GPP, NPP, ANPP, and R~root~.
+Productivity decreases rapidly as seasonality increases, with the rate of decrease slowing as seasonality increases.
+ANPP~foliage~, ANPP~woody stem~ and R~auto~ decrease linearly with temperature seasonality.
+Temperature seasonality is strongly correlated with annual temperature range, and, as expected, all fluxes show almost identical responses to it.
+Productivity is highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower.
+
+In contrast, there is no significant effect of precipitation seasonality on productivity.
+
+We find a significant relationship between length of growing season and productivity, with all fluxes showing a linear increase in productivity with length of growing season.
+Length of growing season is a strong predictor of productivity, explaining 51% of variation in GPP, 39% of variation in NPP, and 34% of variation in ANPP, but it is not identified as a stronger predictor than MAT for any of the fluxes analysed.
+
+*Within the growing season, how does productivity vary with climate?* Within growing season months, we find that climate has a much weaker effect on productivity.
+For each of temperature, precipitation, PET, and solar radiation, we find a small effect of climate for certain carbon fluxes.
+There is a small increase in productivity with temperature and precipitation for ANPP (with temperature R^2^ = 0.10, p<0.001; with precipitation R^2^ = 0.04, p<0.05) and ANPP_foliage (with temperature R^2^ = 0.16, p<0.01; with precipitation R^2^ = 0.09, p<0.05).
+Productivity increases with solar radiation for GPP (R^2^ = 0.21, p<0.001), NPP (R^2^ = 0.21, p<0.001), BNPP (R^2^ = 0.16, p<0.001) and BNPP~fine root~ (R^2^ = 0.12, p<0.01), and with PET for GPP (R^2^ = 0.15, p<0.01), NPP (R^2^ = 0.18, p<0.01), BNPP (R^2^ = 0.23, p<0.0001), BNPP~fine root~ (R^2^ = 0.11, p<0.05), and ANPP~woody stem~ (R^2^ = 0.06, p<0.05).
 

### 2019-08-23 scripts/final_model_no_weighting.R
Update final_model_no_weighting.R

@@ -563,7 +563,7 @@ for (age in ages){
 
 
 
-fixed.variables <- c("mat", "map", "PotentialEvapotranspiration", "SolarRadiation", "TempSeasonality", "length_growing_season")
+fixed.variables <- c("mat", "map", "PotentialEvapotranspiration", "VapourPressureDeficit", "TempSeasonality", "length_growing_season")
 
 response.variables <- c("GPP", "NPP", "BNPP_root", "ANPP")
 

### 2019-08-23 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
 *Closure of fluxes* Analysis of component fluxes indicates that measures of smaller component fluxes do approximately sum to measures of larger fluxes (fig.
-3).
+2).
 Modelled estimates of GPP, generated from the sum of NPP and R auto; NPP, generated from the sum of ANPP and BNPP~root~; and ANPP, generated from the sum of ANPP~foliage~ and ANPP~woody stem~, fall completely within the confidence intervals of the regressions of field estimates of GPP, NPP and ANPP respectively.
@@
 
-*How do carbon fluxes vary with latitude?* All major carbon fluxes increase linearly with decreasing latitude (fig. 2).
+*How do carbon fluxes vary with latitude?* All major carbon fluxes increase linearly with decreasing latitude (fig. 3).
 
@@
 All regressions are significant $(p<0.05)$.", fig.pos='H'}
-  knitr::include_graphics("effect_of_latitude.png")
+  knitr::include_graphics("effect_of_lat_transparent.png")
 ```
@@
 
-*How does productivity relate to MAT and MAP?* MAT is a significant (p<0.05) and strong predictor of productivity for all carbon fluxes tested, with all fluxes showing a linear increase with temperature.
+*How does productivity relate to MAT and MAP?* MAT is a significant (p<0.05) and strong predictor of productivity for all carbon fluxes tested, with all fluxes showing a linear increase with temperature (fig.
+5).
 We found no support for a saturation point of productivity with temperature.
@@
 For the majority of fluxes productivity is best predicted by a polynomial model.
-Productivity increases with precipitation, up until a saturation point at between 3000 and 4000mm annual precipitation, above which productivity starts to decrease.
+Productivity increases with precipitation, up until a saturation point at between 3000 and 4000mm annual precipitation, above which productivity starts to decrease (fig.
+5).
 The notable exception to this is GPP: the model indicates that GPP continues to increase with precipitation up to measures of at least 5000mm annually (p<0.0001, R^2^ = 0.33.
@@
 
-There was a significant interactive effect between MAT and MAP for GPP, BNPP~root~, BNPP~fine root~, ANPP, ANPP~woody stem~, and R~root~.
+There was a significant interactive effect between MAT and MAP for GPP, BNPP~root~, BNPP~fine root~, ANPP, ANPP~woody stem~, and R~root~ (fig.4).
 There was a significant additive effect for R~auto~.
@@
 We found a significant relationship between productivity and potential evapotranspiration for all fluxes.
-ANPP~foliage~, BNPP~fine root~ and R~root~ increased linearly with PET, however all other fluxes showed a polynomial relationship with PET.
+ANPP~foliage~, BNPP~fine root~ and R~root~ increased linearly with PET, however all other fluxes showed a polynomial relationship with PET (fig.
+5).
 We find strong evidence for a saturation point or peak with PET: productivity tends to increase at values below 1000mm, before saturating between 1200 and 1700mm.
@@
 
-Solar radiation
-Vapour pressure deficit?
+Vapour pressure deficit was a significant predictor of productivity for all fluxes.
+BNPP~fine root~ showed a linear relationship with vapour pressure deficit (R^2^ = 0.07, p<0.05), but all other fluxes showed a polynomial relationship (fig.
+5).
+Productivity initially increased with vapour pressure deficit, before saturating at around 0.8 kPa.
+At values above 0.8 kPa, productivity began to decrease.
+
+All fluxes, with the exception of R~root~, show a positive linear relationship with solar radiation.
+Solar radiation explains a low proportion of variability in productivity for all fluxes, explaining less than 20% of the variation in each flux, with the exception of R~auto~ (R^2^ = 0.26, p<0.05).
 
@@
 
+```{r echo=FALSE, out.width='100%', fig.cap = "Plots of primary productivity $(MgC~ha^-1~yr^-1)$ regressed against the climate variables identified as the best predictors: mean annual temperature; temperature seasonality; annual temperature range; and vapour pressure.
+Lines of best fit are plotted according to the best model selected during analysis.
+All regressions are significant $(p<0.05)$.", fig.pos='H'}
+  knitr::include_graphics("combined_plots.png")
+```
+
 *What is the role of seasonality in explaining productivity?* Temperature seasonality is a significant predictor of productivity.
 We found a polynomial relationship with productivity for GPP, NPP, ANPP, and R~root~.
-Productivity decreases rapidly as seasonality increases, with the rate of decrease slowing as seasonality increases.
+Productivity decreases rapidly as seasonality increases, with the rate of decrease slowing as seasonality increases (fig.
+5).
 ANPP~foliage~, ANPP~woody stem~ and R~auto~ decrease linearly with temperature seasonality.
@@
 
-We find a significant relationship between length of growing season and productivity, with all fluxes showing a linear increase in productivity with length of growing season.
+We find a significant relationship between length of growing season and productivity, with all fluxes showing a linear increase in productivity with length of growing season (fig.
+5).
 Length of growing season is a strong predictor of productivity, explaining 51% of variation in GPP, 39% of variation in NPP, and 34% of variation in ANPP, but it is not identified as a stronger predictor than MAT for any of the fluxes analysed.

### 2019-08-23 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@
 
+In this analysis we use a comprehensive global database, standardised to account for differences in methodology across studies, and stand age and level of disturbance across sites.
+The database provides us with an unprecedented amount of data, representing all (?) global forest ecosystems and all significant forest carbon fluxes.
+This breadth of data allows us to comprehensively explore the relationships between climate and productivity on a global scale.
+
+Past studies have differed in their conclusions regarding the relationship between productivity and latitude.
+Our findings indicate that, for all carbon fluxes studied, productivity peaks in the tropical regions.
+
+There is strong evidence from our analyses that this increase in productivity towards the tropical regions is driven primarily by temperature-related effects.
+We find that temperature-related climate variables explain the highest proportion of variability in productivity.
+
+There is debate over whether productivity in regions with higher MAT is higher because productivity inherently increases with air temperature during growing season months, or because higher MAT is correlated with longer growing seasons.
+Our analysis suggests that length of growing season is a key factor in understanding latitudinal patterns in productivity.
+We find that length of growing season is strongly correlated with productivity.
+This is supported by the strong correlation between temperature seasonality and productivity, indicating that aseasonality and year-round growing seasons have a strong influence over productivity.
+
+These conclusions are supported by our analyses of the influence of climate within the growing season.
+We find much weaker correlations between climate variables and productivity when only growing season months are considered.
+There is evidence for a positive correlation between productivity and temperature for ANPP and ANPP~foliage~, suggesting that there is still a small positive effect of air temperature on productivity independent of growing season, however this effect is not significant for the remaining fluxes.
+This suggests that the primary positive effect of temperature on productivity is through the lengthening of the growing season in regions with higher MAT.
+However, we note that length of growing season is a less good predictor of productivity than MAT.
+This may be an artefact of imprecise estimation of length of growing season, and with more precise climate data on finer time-scales it is likely that the predictive ability of length of growing season would increase, however it does indicate that there are aspects of productivity explained by MAT which length of growing season alone cannot explain.
+It is hard to attribute mechanisms through such coarse-scale data as this.
+Future studies.......
+
+The majority of studies to date have focused on MAT and MAP as two key climate variables for which data is easily obtained.
+We find that MAP is not a significant predictor of productivity, but does show a strong interaction with temperature.
+This indicates that the main mechanism through with precipitation influences productivity is in a secondary capacity, with its overall influence being mediated by the effects of temperature.
+This also suggests that although temperature explains a large proportion of variation, it cannot fully explain productivity alone, and a measure of water availability is necessary to understand patterns of productivity.
+This is likely to be significant in the future, as the effects of a warming climate will be mediated through changes in precipitation regime.
+
+This studies considers the relationship between productivity and climate on a coarse scale.
+Because of the nature of the ForC database, we were unable to access site-level MAT and MAP data for a number of sites, and lacked site-level data for other climate variables for all sites.
+As a result the climate data we used was primarily multi-year averages of annual climate measures from global databases.
+While the quality of these data is very high, we cannot expect that it captures detailed local variation in climate.
+Although we find that MAT is the strongest predictor for climate, part of this effect may be a result of the higher-quality stand-level dataset available for MAT.
+With high-quality stand-level data for other climate variables we might find stronger predictors.
+Similarly, with accurate stand-level measurements of growing season length, we would expect to record a stronger correlation between growing season length and productivity.
+
+Furthermore, we expect that forest stand productivity varies on a seasonal and annual scale with seasonal and annual variation in climate.
+By taking long-term averages both in climate and in forest productivity estimates, we expect that much of this variation will be averaged out, and the strong local effects of climate on productivity will be masked.
+
+By testing a wide selection of carbon fluxes, we have been able to explore the relationships between fluxes and their responses to productivity. We see that:
+- All fluxes show similar responses to climate drivers
+- Allocation to fluxes does not vary with climate on a global scale.
+This suggests that the pressures that require individuals to shift allocation between fluxes occur on much more local scales.
+We expect that allocation varies with factors such as altitude, stand age, disturbance regime, micro-climate, and is likely to show seasonal responses to climate variables, such as showing short-term responses to water availability.
+- Climate more strongly explains variation in the major fluxes, with lower levels of variation explained in subsidiary fluxes.
+This indicates that factors other than climate influence subsidiary fluxes.
+This links to the results we find for allocation: although climate has a significant effect on these fluxes, there are a number of other factors which can influence the way that carbon is allocated to them, which masks the effect of climate to some extent.
+
 *The relationship between productivity and latitude*
-
-Begin discussion by emphasising that this these conclusions are based on analyses of an unprecedented amount of data across many fluxes, limited to mature undisturbed forests.
 

### 2019-08-26 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@
 
+Limitations: 
+- We have only considered mature forests; there is a lot more work to go towards understanding the effect of stand age on productivity.
+In part, understanding these patterns will be reliant on acquiring more data at all stand ages.
+- We have standardised data as far as possible; however there are still challenges with ensuring data quality is as high as possible.
+For example, there are inherent challenges with measuring stand-level productivity, particularly for certain components.
+Standardising BNPP in particular is challenging, as it is often sampled to insufficient depth/extrapolated from measures of ANPP.
+This means that, although we have been careful to control data selection within the database, we cannot entirely control the quality of data going into the database.
+As a result, our results are likely to be affected by this.
+- As above, we are looking at climate data at a coarse scale; these are generalisations about large-scale patterns.
+Stand-level studies will help to improve our understanding of the effect of climate on productivity on smaller spatial and temporal scales.
+
+Our findings support a lot of previous work that has been done, but we are doing this with a much larger and more comprehensive dataset, with better standardisation of available data.
+
 *The relationship between productivity and latitude*

### 2019-08-27 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@
 
+Our results show that productivity decreases linearly with latitude.
+Climate explains a significant proportion of variation in all carbon fluxes, with temperature variables being the best predictors of productivity.
+While other climate variables are significant predictors of productivity, none of them improve on the explanatory power of temperature-related variables.
+Water availability is an important factor in explaining productivity on a global scale: we find a positive influence of precipitation at low MAP, with saturation at higher levels of MAP.
+There is a significant interaction between MAT and MAP.
+
+We find that seasonality is an important factor in understanding patterns in productivity on a global scale: temperature seasonality and growing season length are strong predictors of productivity, though growing season length doesn't improve on MAT as a predictor.
+Within the growing season we find the influence of climate on productivity is smaller, but still significant for a number of carbon fluxes.
+
 Past studies have differed in their conclusions regarding the relationship between productivity and latitude.

### 2019-08-27 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@
 
+Many of our findings support and clarify previously published work; however we use a much larger and more comprehensive database than has previously been available, allowing for higher quality control and standardisation of data.
+This enalbed us to control for factors not previously controlled for - such as stand age, methodology, flux components, and disturbance regime - to gain a stronger understanding of latitudinal and climatic effects on productivity.
+
+Despite the high standard of data quality in the ForC database, there are still significant limitations on our analyses.
+We standardised methodologies as far as possible, ensuring that there was consistency in the components included for each carbon flux.
+However, field techniques for estimating components of primary productivity are variable in their accuracy.
+This is particularly true for measures of BNPP, which are often estimated by extrapolating measures of ANPP.
+Even where BNPP is calculated through direct soil sampling, measures are often taken to insufficient depth, leading to underestimation of the flux.
+It is challenging to account for this type of variability in quality of field data, and our analyses largely assume good faith in data quality.
+While we have strong confidence in our analyses of well-understood fluxes, such as GPP, ANPP, and ANPP~woody stem~, we recommend caution in drawing firm conclusions from our analyses of BNPP, BNPP~fine root~, and, to some extent, NPP.
+We expect that, as field techniques improve the accuracy of estimates of these fluxes in the future, our understanding of the relationship of BNPP with climate variables will change.
+It is possible that the low R^2^ values for BNPP that we find are a result of lower accuracy in field measures of these fluxes.
+
 Past studies have differed in their conclusions regarding the relationship between productivity and latitude.
@@
 
-Our findings support a lot of previous work that has been done, but we are doing this with a much larger and more comprehensive dataset, with better standardisation of available data.
-
 *The relationship between productivity and latitude*

### 2019-08-27 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@
 Many of our findings support and clarify previously published work; however we use a much larger and more comprehensive database than has previously been available, allowing for higher quality control and standardisation of data.
-This enalbed us to control for factors not previously controlled for - such as stand age, methodology, flux components, and disturbance regime - to gain a stronger understanding of latitudinal and climatic effects on productivity.
+This enabled us to control for factors not previously controlled for - such as stand age, methodology, flux components, and disturbance regime - to gain a stronger understanding of latitudinal and climatic effects on productivity.
 
@@
 As a result the climate data we used was primarily multi-year averages of annual climate measures from global databases.
-While the quality of these data is very high, we cannot expect that it captures detailed local variation in climate.
+While the quality of these data is high, we cannot expect that it captures detailed local variation in climate.
 Although we find that MAT is the strongest predictor for climate, part of this effect may be a result of the higher-quality stand-level dataset available for MAT.
@@
 In part, understanding these patterns will be reliant on acquiring more data at all stand ages.
-- We have standardised data as far as possible; however there are still challenges with ensuring data quality is as high as possible.
-For example, there are inherent challenges with measuring stand-level productivity, particularly for certain components.
-Standardising BNPP in particular is challenging, as it is often sampled to insufficient depth/extrapolated from measures of ANPP.
-This means that, although we have been careful to control data selection within the database, we cannot entirely control the quality of data going into the database.
-As a result, our results are likely to be affected by this.
-- As above, we are looking at climate data at a coarse scale; these are generalisations about large-scale patterns.
-Stand-level studies will help to improve our understanding of the effect of climate on productivity on smaller spatial and temporal scales.
 

### 2019-08-28 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@
 
-In this analysis we use a comprehensive global database, standardised to account for differences in methodology across studies, and stand age and level of disturbance across sites.
-The database provides us with an unprecedented amount of data, representing all (?) global forest ecosystems and all significant forest carbon fluxes.
-This breadth of data allows us to comprehensively explore the relationships between climate and productivity on a global scale.
+In this analysis we use a comprehensive global database, containing an unprecedented amount of data and representing all global forest ecosystems and all significant forest carbon fluxes, to comprehensively explore the relationships between climate and productivity on a global scale.
+Many of our findings support and clarify previously published work; however we use a much larger and more complete database than has previously been available, allowing for higher quality control and standardisation of data.
+This enabled us to control for factors not previously controlled for - such as stand age, methodology, flux components, and disturbance regime - to gain a stronger understanding of latitudinal and climatic effects on productivity.
 
@@
 There is a significant interaction between MAT and MAP.
+We note that climate tends to explain a higher amount of variation in the major fluxes, with the amount of variation explained decreasing in smaller component fluxes.
 
@@
 Within the growing season we find the influence of climate on productivity is smaller, but still significant for a number of carbon fluxes.
-
-Many of our findings support and clarify previously published work; however we use a much larger and more comprehensive database than has previously been available, allowing for higher quality control and standardisation of data.
-This enabled us to control for factors not previously controlled for - such as stand age, methodology, flux components, and disturbance regime - to gain a stronger understanding of latitudinal and climatic effects on productivity.
 
@@
 We expect that, as field techniques improve the accuracy of estimates of these fluxes in the future, our understanding of the relationship of BNPP with climate variables will change.
-It is possible that the low R^2^ values for BNPP that we find are a result of lower accuracy in field measures of these fluxes.
+It is possible that the low R^2^ values for BNPP that we find are a result of lower accuracy in field measures of these fluxes, and that with improve standardisation of methods, we would find climate explains a higher proportion of variation.
+
+Secondly, we emphasise that this study considers the relationship between productivity and climate on a coarse scale.
+Because of the nature of the ForC database, we were unable to access site-level MAT and MAP data for a number of sites, and lacked site-level data for other climate variables for all sites.
+As a result the climate data we used was primarily multi-year averages of annual climate measures accessed from global databases.
+While the quality of these data is high, we cannot expect that it captures detailed local variation in climate, and in particular will be unable to capture microclimate variation, which may have a strong influence on stand-level productivity.
+Although we find that MAT is the strongest predictor for climate, part of this effect may be a result of the higher-quality stand-level dataset available for MAT.
+With high-quality stand-level data for other climate variables we would expect that their predictive strength would improve.
+
+Furthermore, we expect that temporal resolution of climate data affects our results.
+Productivity is known to vary on a seasonal and annual scale in response to variation in climate.
+Our analysis relies on long-term averages in both climate variables and forest productivity estimates, and as a result averages out a large proportion of this variation.
+Therefore we recognise that the strong short-term effects of climate on productivity are likely to be masked.
+The limitations of low temporal resolution are highlighted in our analysis of growing season length.
+We note that our estimate of growing season length is imprecise, being based on monthly temperature, precipitation, and evapotranspiration averages.
+With accurate stand-level measurements of growing season length, based on daily, not monthly, climate measures, we would expect to record a stronger correlation between growing season length and productivity.
+
+Although we recognise the inherent limitations present in this analysis, by focusing on mature, undisturbed forests, and by clarifying the components of NPP included in analysis, our study offers a clearer picture of how carbon cycling varies with latitude and climate.
+
+By including multiple fluxes in our analysis, we are able to explore the relationships between fluxes and their responses to climate.
+We find that fluxes are broadly consistent in their responses to climate drivers, suggesting that the importance of climate drivers on productivity is relatively invariant across fluxes.
+In addition, we show that carbon allocation between subsidiary fluxes does not vary significantly across broad climatic gradients.
+There is evidence that allocation between carbon fluxes does vary with factors such as stand age [@litton_carbon_2007], nutrient availability [@litton_carbon_2007; @gill-belowground_2016], forest structure [@taylor_greater_2019], elevation [@moser_elevation_2011], and water availability [@newman_above-_2006].
+That we do not see consistent patterns of variation in allocation across climatic gradients indicates that the pressures influencing shifts in allocation between fluxes occur on local scales, and that shifts in allocation occur as a response to fluctuations in local environmental conditions.
+
+Related to this, we find that climate more strongly explains variation in the major fluxes, with lower levels of variation explained in subsidiary fluxes.
+Although this could in part be related to a lack of standardisation in methodology across studies, as discussed above, it could also indicate that factors other than climate have significant influence over subsidiary fluxes.
+This is consisent with the results we find for allocation between fluxes: although climate has a significant influence on subsidiary fluxes, it is not the only factor that is important, and many other local- and regional-scale factors  influence the way carbon is allocated between subsidiary fluxes.
+We would expect that this would reduce the strength of the effect of climate on subsidiary fluxes.
 
@@
 
-Furthermore, we expect that forest stand productivity varies on a seasonal and annual scale with seasonal and annual variation in climate.
-By taking long-term averages both in climate and in forest productivity estimates, we expect that much of this variation will be averaged out, and the strong local effects of climate on productivity will be masked.
-
-By testing a wide selection of carbon fluxes, we have been able to explore the relationships between fluxes and their responses to productivity. We see that:
-- All fluxes show similar responses to climate drivers
-- Allocation to fluxes does not vary with climate on a global scale.
-This suggests that the pressures that require individuals to shift allocation between fluxes occur on much more local scales.
-We expect that allocation varies with factors such as altitude, stand age, disturbance regime, micro-climate, and is likely to show seasonal responses to climate variables, such as showing short-term responses to water availability.
-- Climate more strongly explains variation in the major fluxes, with lower levels of variation explained in subsidiary fluxes.
-This indicates that factors other than climate influence subsidiary fluxes.
-This links to the results we find for allocation: although climate has a significant effect on these fluxes, there are a number of other factors which can influence the way that carbon is allocated to them, which masks the effect of climate to some extent.
-
-Limitations: 
-- We have only considered mature forests; there is a lot more work to go towards understanding the effect of stand age on productivity.
-In part, understanding these patterns will be reliant on acquiring more data at all stand ages.
-
 *The relationship between productivity and latitude*

### 2019-08-28 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
 
-By including multiple fluxes in our analysis, we are able to explore the relationships between fluxes and their responses to climate.
-We find that fluxes are broadly consistent in their responses to climate drivers, suggesting that the importance of climate drivers on productivity is relatively invariant across fluxes.
-In addition, we show that carbon allocation between subsidiary fluxes does not vary significantly across broad climatic gradients.
-There is evidence that allocation between carbon fluxes does vary with factors such as stand age [@litton_carbon_2007], nutrient availability [@litton_carbon_2007; @gill-belowground_2016], forest structure [@taylor_greater_2019], elevation [@moser_elevation_2011], and water availability [@newman_above-_2006].
-That we do not see consistent patterns of variation in allocation across climatic gradients indicates that the pressures influencing shifts in allocation between fluxes occur on local scales, and that shifts in allocation occur as a response to fluctuations in local environmental conditions.
-
-Related to this, we find that climate more strongly explains variation in the major fluxes, with lower levels of variation explained in subsidiary fluxes.
-Although this could in part be related to a lack of standardisation in methodology across studies, as discussed above, it could also indicate that factors other than climate have significant influence over subsidiary fluxes.
-This is consisent with the results we find for allocation between fluxes: although climate has a significant influence on subsidiary fluxes, it is not the only factor that is important, and many other local- and regional-scale factors  influence the way carbon is allocated between subsidiary fluxes.
-We would expect that this would reduce the strength of the effect of climate on subsidiary fluxes.
-
 Past studies have differed in their conclusions regarding the relationship between productivity and latitude.
 Our findings indicate that, for all carbon fluxes studied, productivity peaks in the tropical regions.
-
 There is strong evidence from our analyses that this increase in productivity towards the tropical regions is driven primarily by temperature-related effects.
 We find that temperature-related climate variables explain the highest proportion of variability in productivity.
+In contrast, we find that MAP is not a significant predictor of productivity, but does show a strong interaction with temperature.
+This indicates that the main mechanism through with precipitation influences productivity is in a secondary capacity, with its overall influence being mediated by the effects of temperature.
+This also suggests that although temperature explains a large proportion of variation, it cannot fully explain productivity alone, and a measure of water availability is necessary to understand patterns of productivity.
+This is likely to be significant in the future, as the effects of a warming climate will be mediated through changes in precipitation regime.
 
@@
 However, we note that length of growing season is a less good predictor of productivity than MAT.
-This may be an artefact of imprecise estimation of length of growing season, and with more precise climate data on finer time-scales it is likely that the predictive ability of length of growing season would increase, however it does indicate that there are aspects of productivity explained by MAT which length of growing season alone cannot explain.
-It is hard to attribute mechanisms through such coarse-scale data as this.
-Future studies.......
+As discussed above, this may be an artefact of imprecise estimation of length of growing season, however it does indicate that there are aspects of productivity explained by MAT which length of growing season alone cannot explain.
 
-The majority of studies to date have focused on MAT and MAP as two key climate variables for which data is easily obtained.
-We find that MAP is not a significant predictor of productivity, but does show a strong interaction with temperature.
-This indicates that the main mechanism through with precipitation influences productivity is in a secondary capacity, with its overall influence being mediated by the effects of temperature.
-This also suggests that although temperature explains a large proportion of variation, it cannot fully explain productivity alone, and a measure of water availability is necessary to understand patterns of productivity.
-This is likely to be significant in the future, as the effects of a warming climate will be mediated through changes in precipitation regime.
+By including multiple fluxes in our analysis, we are able to explore the relationships between fluxes and their responses to climate.
+We find that fluxes are broadly consistent in their responses to climate drivers, suggesting that the importance of climate drivers on productivity is relatively invariant across fluxes.
+In addition, we show that carbon allocation between subsidiary fluxes does not vary significantly across broad climatic gradients.
+There is evidence that allocation between carbon fluxes does vary with factors such as stand age [@litton_carbon_2007], nutrient availability [@litton_carbon_2007; @gill-belowground_2016], forest structure [@taylor_greater_2019], elevation [@moser_elevation_2011], and water availability [@newman_above-_2006].
+That we do not see consistent patterns of variation in allocation across climatic gradients indicates that the pressures influencing shifts in allocation between fluxes occur on local scales, and that shifts in allocation occur as a response to fluctuations in local environmental conditions.
 
-This studies considers the relationship between productivity and climate on a coarse scale.
-Because of the nature of the ForC database, we were unable to access site-level MAT and MAP data for a number of sites, and lacked site-level data for other climate variables for all sites.
-As a result the climate data we used was primarily multi-year averages of annual climate measures from global databases.
-While the quality of these data is high, we cannot expect that it captures detailed local variation in climate.
-Although we find that MAT is the strongest predictor for climate, part of this effect may be a result of the higher-quality stand-level dataset available for MAT.
-With high-quality stand-level data for other climate variables we might find stronger predictors.
-Similarly, with accurate stand-level measurements of growing season length, we would expect to record a stronger correlation between growing season length and productivity.
-
-*The relationship between productivity and latitude*
-
-We are no longer explicity testing whether climate explains a higher proportion of variability in major fluxes, so mention this, and the implications of it in the discussion.
-
-Past studies have differed in their conclusions regarding the relationship between productivity and latitude.
-Our findings indicate that, for all carbon fluxes studied, productivity peaks in the tropical regions.
-While Huston and Wolverton (2009) found no significant relationship between ANPP and latitude, we find a significant negative relationship between ANPP and latitude, with the highest levels of ANPP in the tropical regions.
-
-*Challenges in estimating NPP*  
-The impact of excluding certain components of NPP on this analysis is unclear, primarily because these components are poorly understood and rarely measured [@clark_net_2001].
-Do we want to include a section assessing how much this may or may not impact results?
-1. Hard to measure BNPP because it often isn't sampled deep enough; impact on ratio tests?  
-2. Unclear if there are latitudinal patterns e.g. of herbivory/VOCs  
-  + Herbivory can be an important component of total NPP so this may be important to account for; likely to be higher in the tropics?
+Related to this, we find that climate more strongly explains variation in the major fluxes, with lower levels of variation explained in subsidiary fluxes.
+Although this could in part be related to a lack of standardisation in methodology across studies, as discussed above, it could also indicate that factors other than climate have significant influence over subsidiary fluxes.
+This is consistent with the results we find for allocation between fluxes: although climate has a significant influence on subsidiary fluxes, it is not the only factor that is important, and many other local- and regional-scale factors  influence the way carbon is allocated between subsidiary fluxes.
+We would expect that this would reduce the strength of the effect of climate on subsidiary fluxes.

### 2019-08-28 manuscript/manuscript_removed_text.Rmd
Update manuscript_removed_text.Rmd

@@
 ---
+
+1.
+**Which climatic variables are the most important drivers behind the latitudinal pattern in primary productivity?** To date, the majority of studies have focussed on productivity responses to precipitation and temperature, while the influence of other climate variables on productivity remains under-explored.
+We utilise global climate datasets to investigate the relationships between productivity and a range of climate variables.
+2.
+**Do the components of primary productivity show variation in their responses to these climatic drivers?** Allocation to components of primary productivity appears to show some variation across climate and latitudinal gradients.
+We use global datasets of a range of primary productivity components to explore how significant this variation in allocation is.
+3.
+**Does climate explain the same proportion of variation in different components of primary productivity?** Review of current research shows that, while the relationship between GPP and latitude is consistent, NPP and ANPP show weaker relationships with latitude.
+This may be a result of varying sensitivities to climate drivers.
+We use mixed models to estimate the proportion of variation explained by climate for different components of productivity, to investigate how the relative importance of climatic drivers varies across components.
 

### 2019-08-28 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@
 
-*The components of forest* During photosynthesis, plants capture carbon dioxide from the atmosphere.
+*The components of forest productivity.* During photosynthesis, plants capture carbon dioxide from the atmosphere.
 The gross primary productivity (GPP) of an ecosystem is the gross uptake, via photosynthesis, of carbon dioxide by plants in that ecosystem.
@@
 On a global scale, changes in primary productivity are influenced by climatic variables and abiotic gradients, such as the length of growing season [@michaletz_convergence_2014].
-There is some debate over the precise relationship between these drivers and productivity; While mean annual precipitation (MAP) and mean annual temperature (MAT) have been argued to be significant predictors of productivity [@chu_does_2016], other studies have found that the correlation between productivity and MAT is a factor of the relationship between productivity and growing season length [@kerkhoff_plant_2005; @malhi_productivity_2012; @michaletz_convergence_2014; @michaletz_drivers_2018].
+There is some debate over the precise relationship between these drivers and productivity; while MAT and MAP have been argued to be significant predictors of productivity [@chu_does_2016], other studies have found that the correlation between productivity and MAT is a factor of the relationship between productivity and growing season length [@kerkhoff_plant_2005; @malhi_productivity_2012; @michaletz_convergence_2014; @michaletz_drivers_2018].
 Improving understanding of how these factors interact to control global patterns in primary productivity is essential to understanding the global carbon cycle.
 
-Current research into how primary productivity varies with latitude is inconclusive, and - though it has primarily focussed on patterns of GPP, NPP, and ANPP - indicates that different components of productivity may show different relationships to latitude.
+Current research into how primary productivity varies with latitude is inconclusive, and - though it has primarily focussed on patterns of GPP, NPP, and ANPP - indicates that different components of productivity may show different relationships with latitude.
 Gross primary productivity is generally thought to be highest in the tropics.
@@
 
-Furthermore, there is evidence that different components of productivity show individual responses to drivers of productivity.
+Furthermore, there is evidence that different components of productivity show individual responses to climate drivers.
 For example, increases in GPP have been reported to saturate above 25$^\circ$C MAT [@larjavaara_temperature_2012], while increases in NPP are recorded to saturate above 10$^\circ$C MAT [@luyssaert_co2_2007].
@@
 Here, we use a comprehensive global database of forest carbon fluxes to explore global patterns in productivity.
-We explore three questions:
-1.
-**Which climatic variables are the most important drivers behind the latitudinal pattern in primary productivity?** To date, the majority of studies have focussed on productivity responses to precipitation and temperature, while the influence of other climate variables on productivity remains under-explored.
-We utilise global climate datasets to investigate the relationships between productivity and a range of climate variables.
-2.
-**Do the components of primary productivity show variation in their responses to these climatic drivers?** Allocation to components of primary productivity appears to show some variation across climate and latitudinal gradients.
-We use global datasets of a range of primary productivity components to explore how significant this variation in allocation is.
-3.
-**Does climate explain the same proportion of variation in different components of primary productivity?** Review of current research shows that, while the relationship between GPP and latitude is consistent, NPP and ANPP show weaker relationships with latitude.
-This may be a result of varying sensitivities to climate drivers.
-We use mixed models to estimate the proportion of variation explained by climate for different components of productivity, to investigate how the relative importance of climatic drivers varies across components.
+We explore five broad questions, described in greater detail in table 1:
+  1. How do carbon fluxes vary with latitude?  
+  2. How do carbon fluxes relate to MAT and MAP?  
+  3. How do carbon fluxes relate to other annual climate variables?  
+  4. What is the role of seasonality in explaining variation in carbon fluxes?  
+  5. Per month of the growing season, how does productivity vary with climate?
 

### 2019-08-28 scripts/stacked_plots_code.R
Update stacked_plots_code.R

@@ -268,7 +268,7 @@ par(mfrow = c(2,2), mar = c(2,2,2,2), oma = c(3,3,0,0))
           
           ylim <- range(c(df.1$mean, df.2$mean, df.sum$mean))
           ylim[1] <- ylim[1] - 0.25
-          ylim[2] <- ylim[2] + 0.25
+          ylim[2] <- ylim[2] + 5
           
           # png(file = paste0("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/figures/final_figures/stacked_plots/", set1[[i]], "_to_", set2[[j]],"_", fixed.v, "_stacked.png"), width = 2255, height = 2000, units = "px", res = 300)
           
@@ -303,6 +303,7 @@ par(mfrow = c(2,2), mar = c(2,2,2,2), oma = c(3,3,0,0))
           # title(paste("Stacked graphs by latitude"), outer = F, line = 1)
           mtext(side = 1, line = 1, text = "Latitude", outer = T)
           mtext(side = 2, line = 1,  text = expression("Productivity Mg C"~ha^-1~yr^-1), outer = T)
+          mtext(paste0("(", letters[1], ")"), side = 3, line = -1.5, adj = 0.05)
           
           # dev.off()
           
@@ -479,7 +480,7 @@ for (age in ages){
             
             ylim <- range(c(df.1$mean, df.2$mean, df.sum$mean))
             ylim[1] <- ylim[1] - 0.25
-            ylim[2] <- ylim[2] + 0.25
+            ylim[2] <- ylim[2] + 2
             
             # png(file = paste0("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/figures/final_figures/stacked_plots/", set1[[i]], "_to_", set2[[j]],"_", fixed.v, "_stacked.png"), width = 2255, height = 2000, units = "px", res = 300)
             
@@ -508,6 +509,7 @@ for (age in ages){
             # mtext(side = 3, line = -4, text = legend, adj = 0.95, col = "black", cex = 0.5, outer = F)
             
             legend("topright", legend = c("ANPP", "BNPP", "NPP", "ANPP + BNPP"), col = plasma(10)[c(5, 8, 3, NA)], pch = c(5, 8, 3, NA), xpd = T, text.col = plasma(10)[c(5, 8, 3, NA)], bty = "n", xjust = 1, cex = 0.75)
[…]

### 2019-08-28 scripts/mat_map_interaction.R
Update mat_map_interaction.R

@@ -106,12 +106,13 @@ all.results <- NULL
 best.results <- NULL
 
 effects <- c("mat", "map", "(1|geographic.area/plot.name)")
+pannel.nb <- 1
 
 
 response.variables <- c("GPP", "NPP", "BNPP_root", "BNPP_root_fine", "ANPP", "ANPP_foliage", "ANPP_woody_stem","R_auto", "R_auto_root")
 
 png(file = paste0("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/figures/final_figures/interactions/mat_map_interaction.png"), width = 2255, height = 2000, units = "px", res = 300)
-par(mfrow = c(3,3), mar = c(0,0,0,0), oma = c(5,5,2,0))
+par(mfrow = c(3,3), mar = c(0,0,0,0), oma = c(5,7,2,0), xpd = T)
 
 for (response.v in response.variables){
   
@@ -205,7 +206,7 @@ for (response.v in response.variables){
     #     lines(fit ~ mat, data = newDat[newDat$map %in% map,], lty = ifelse(significant.effect.of.interaction|significant.effect, 1, 2), lwd = i)
     #     
     #   }
-    if(response.v == "GPP") legend("topleft", lwd = c(1:4), legend = c(500, 1500, 2500, 3500), col = plasma(5)[1:4])
+    if(response.v == "GPP") legend("topleft", lwd = c(1:4), legend = c(500, 1500, 2500, 3500), col = plasma(5)[1:4], inset = c(-0.4, 0), xpd = NA, title = "MAP (mm)", title.col = "black")
     # title (paste(response.v), outer = T, line = 1)
     mtext(side = 1, line = 3, text = "Mean Annual Temperature", outer = T)
     mtext(side = 2, line = 3,  text = expression("Mg C"~ha^-1~yr^-1), outer = T)
@@ -221,11 +222,14 @@ for (response.v in response.variables){
     # if(significant.effet.of.additive) equation <-  paste(response.v, "=", r[1], "+", fixed.v,  "x", r[2], " + age x", r[3])
     # 
     # 
-   mtext(side = 3, line = -1, text = paste(response.v), adj = 0.1, cex = 0.5)
+    
+  
+   mtext(side = 3, line = -1, text = paste0("(", letters[pannel.nb], ") ", response.v), adj = 0.1, cex = 0.5)
    
[…]

### 2019-08-28 scripts/final_model_no_weighting.R
Update final_model_no_weighting.R

@@ -699,11 +699,14 @@ for (age in ages){
       all.results <- rbind(all.results, results)
       
     }
-    pannel.nb <- pannel.nb +1
     
+    
+    mtext(paste0("(", letters[pannel.nb], ")"), side = 3, line = -1.5, adj = 0.05)
     mtext(side = 1, line = 3, text = eval(parse(text = xaxis)), outer = F)
     # mtext(side = 2, line = 3,  text = expression("Mg C"~ha^-1~yr^-1), outer = F) 
     if(fixed.v == "map")legend("topright", legend = c("GPP", "NPP", "ANPP", "BNPP_root"), col = plasma(10)[c(1, 3, 5, 8)], pch = c(1, 3, 5, 8), xpd = T, text.col = plasma(10)[c(1, 3, 5, 8)], bty = "n", xjust = 1, cex = 0.75)
+    
+    pannel.nb <- pannel.nb +1
     
   }
   
@@ -719,6 +722,7 @@ response.variables.groups <- list(c("GPP", "NPP", "ANPP", "BNPP_root", "ANPP_fol
 fixed.variables <- "lat"
 
 all.results = NULL
+all.legends = NULL
 
 
 ### mature forests only ####
@@ -740,7 +744,7 @@ for (age in ages){
       if(response.variables[1] == "GPP") n <- 1
       if(response.variables[1] == "ANPP_foliage") n <- 2
       
-      png(file = paste0("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/figures/final_figures/unweighted_model/effect_of_", fixed.v, "_transparent_notscaled.png"), width = 2255, height = 2000, units = "px", res = 300)
+      png(file = paste0("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/figures/final_figures/unweighted_model/effect_of_", fixed.v, "_transparent.png"), width = 2255, height = 2000, units = "px", res = 300)
       
       par(mfrow = c(1,1), mar = c(3,3,3,3))
       
@@ -773,13 +777,12 @@ for (age in ages){
         
         a <- ForC_simplified[ForC_simplified$variable.name %in% unlist(response.variables),]
         ylim <- range(tapply(a$mean, a$variable.name, scale))
-        ylim <- range(a$mean)
         ylim[1] <- ylim[1] - 0.25
         ylim[2] <- ylim[2] + 0.25
         
         
[…]

### 2019-08-28 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
 In addition, we show that carbon allocation between subsidiary fluxes does not vary significantly across broad climatic gradients.
-There is evidence that allocation between carbon fluxes does vary with factors such as stand age [@litton_carbon_2007], nutrient availability [@litton_carbon_2007; @gill-belowground_2016], forest structure [@taylor_greater_2019], elevation [@moser_elevation_2011], and water availability [@newman_above-_2006].
+There is evidence that allocation between carbon fluxes does vary with factors such as stand age [@litton_carbon_2007], nutrient availability [@litton_carbon_2007; @gill_belowground_2016], forest structure [@taylor_greater_2019], elevation [@moser_elevation_2011], and water availability [@newman_above-_2006].
 That we do not see consistent patterns of variation in allocation across climatic gradients indicates that the pressures influencing shifts in allocation between fluxes occur on local scales, and that shifts in allocation occur as a response to fluctuations in local environmental conditions.

### 2019-08-29 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@
 
-*How does productivity relate to MAT and MAP?* MAT is a significant (p<0.05) and strong predictor of productivity for all carbon fluxes tested, with all fluxes showing a linear increase with temperature (fig.
+*How does productivity relate to MAT and MAP?* We focus first on considering the relationship between productivity and MAT and MAP.
+MAT and MAP are the most commonly reported site-level climate variables, and much previous research into the effect of climate on forest productivity has focused on these as key climate variables.
+MAT is a significant (p<0.05) and strong predictor of productivity for all carbon fluxes tested, with all fluxes showing a linear increase with temperature (fig.
 5).
@@
 Although we find that MAT is the strongest predictor for climate, part of this effect may be a result of the higher-quality stand-level dataset available for MAT.
-With high-quality stand-level data for other climate variables we would expect that their predictive strength would improve.
+With high-quality stand-level data for other climate variables we expect that their predictive strength would improve.
 

### 2019-08-30 manuscript/manuscript_removed_text.Rmd
Update manuscript_removed_text.Rmd

@@
 ---
+
+Globally, forests play an important role in the carbon cycle and are a major component of global carbon dioxide budgets [@luyssaert_old-growth_2008].
+They show higher levels of productivity than non-forest terrestrial ecosystems [@del_grosso_global_2008], and as a result achieve significant carbon sequestration and storage.
+Estimating the total role of forests in the carbon cycle is challenging, but studies indicate that old growth forests alone sequester up to 1.4 GtCyr^-1^ [@malhi_carbon_1999], while the total sequestration of carbon by established forests globally could be up to 2.4 GtCyr^-1^, with the largest sinks being in old-growth tropical forests [@pan_large_2011].
+As atmospheric carbon dioxide levels continue to rise, with consequences for global climate, there is increasing recognition that proper protection and management of forest resources will have an important role to play in mitigating climate change.
+Understanding the patterns of forest productivity on a global scale, and the drivers behind them, is therefore a priority in forest research.
+
+There are two major questions to understand: firstly, how forest productivity varies globally - and specifically which areas show the greatest peaks in productivity -; and secondly, which climate variables drive this variation.
+On a global scale, the productivity of forests varies with latitude, with a general trend of increasing productivity towards the tropics [@beer_terrestrial_2010; @jung_global_2011]; however the exact nature of this pattern, and how it varies by component of productivity, is poorly understood.
+This latitudinal gradient is most likely to be explained by climatic gradients in temperature, precipitation, length of growing season, and combinations of the above.
+Productivity is influenced by a range of climatic drivers, including mean annual temperature (MAT) and mean annual precipitation (MAP) [@del_grosso_global_2008], but doesn't necessarily respond linearly to these drivers.
+Disentangling the shape of productivity responses to climate drivers will enable better predictions of future responses under climate change.
+
+*The components of forest productivity.* During photosynthesis, plants capture carbon dioxide from the atmosphere.
+The gross primary productivity (GPP) of an ecosystem is the gross uptake, via photosynthesis, of carbon dioxide by plants in that ecosystem.
+Only a fraction of the carbon captured is assimilated into plant tissue; the rest is used in autotrophic respiration (R~auto~).
+The component of GPP that is stored as plant material is the net primary productivity (NPP) of an ecosystem.
+Net primary productivity can therefore be expressed as:
+
+$$ NPP = GPP + R~auto~ $$
+
+Currently, GPP cannot be measured directly by observing total ecosystem photosynthesis.
+Instead field estimates of GPP have to be derived based on modelling and extrapolation of eddy-covariance studies and measurements of net ecosystem exchange (NEE) [@clark_reviews_2017].
+
+In contrast, NPP can be calculated through direct field observations.
+In order to achieve greater accuracy in estimating NPP, NPP is often broken down into its component parts, with aboveground NPP (ANPP) and belowground NPP (BNPP) considered separately to each other.
+The components included when estimating ANPP and BNPP often vary between studies, depending on the intensity of fieldwork effort.
+At its most basic level, ANPP can be expressed as:
+
+$$ ANPP = NPP~stem~ + NPP~branch~ + NPP~canopy~ $$
+where NPP~stem~ is the annual woody increment of all stems above a specified diameter at breast height (DBH), NPP~branch~ is annual woody branch turnover, and NPP~canopy~ is annual foliage production, including leaves, twigs, and reproductive structures.
+ANPP may also include NPP~VOC~, the annual emission of volatile organic compounds, and NPP~herbivory~, the annual consumption of plant matter by herbivores, but these components are often excluded from field observations as they are much harder to quantify.
+Other components of aboveground productivity that remain largely unquantified include epiphytes, hemiepiphytes, and understory plants [@clark_reviews_2017].
+The majority of current ANPP estimates are based on the assumption that the contribution of these components to overall NPP is insignificant.
+
+There are two major subcomponents of BNPP, which can be expressed as:
+
+$$ BNPP = NPP~coarse root~ + NPP~fine root~ $$
+where NPP~coarse root~ is the annual production of coarse roots (typically roots >2mm diameter), and NPP~fine root~ is the annual production of fine roots (typically roots <2mm diameter) [@aragao_above-_2009].
+Calculations of BNPP may also include NPP~exudation~, a measure of annual carbon losses through root exudation, and NPP~symbionts~, the annual carbon allocation to mycorrhizae and legumes, but, as before, this is challenging to quantify and is often excluded from field observations.
+
+BNPP is a poorly understood component of total ecosystem productivity, primarily because of the challenges in gaining accurate field measurements.
+Coarse root productivity is often estimated via extrapolation of NPP~stem~ estimates using allometries that may not have been empirically verified [@clark_net_2001].
+NPP~fine root~ is easier to quantify through soil cores and minirhizotrons, however, sampling tends to be limited to the surface soils, with very few studies sampling to depths below 3 metres [cite].
+As a result, it is possible that BNPP is currently significantly underestimated, despite being a hugely significant component of total ecosystem productivity [@pan_large_2011].
+
+*Which factors influence primary productivity?* Primary productivity is influenced by many factors, which often act across a range of scales, and may show interactive effects with each other.
+On a local scale, stand age [@litton_carbon_2007; @gillman_latitude_2015], management [@simova_enigma_2017]; nutrient availability [@aragao_above-_2009]; and altitude [@girardin_net_2010; @malhi_variation_2017] all impact forest productivity.
+On a global scale, changes in primary productivity are influenced by climatic variables and abiotic gradients, such as the length of growing season [@michaletz_convergence_2014].
+There is some debate over the precise relationship between these drivers and productivity; while MAT and MAP have been argued to be significant predictors of productivity [@chu_does_2016], other studies have found that the correlation between productivity and MAT is a factor of the relationship between productivity and growing season length [@kerkhoff_plant_2005; @malhi_productivity_2012; @michaletz_convergence_2014; @michaletz_drivers_2018].
+Improving understanding of how these factors interact to control global patterns in primary productivity is essential to understanding the global carbon cycle.
+
+Current research into how primary productivity varies with latitude is inconclusive, and - though it has primarily focussed on patterns of GPP, NPP, and ANPP - indicates that different components of productivity may show different relationships with latitude.
+Gross primary productivity is generally thought to be highest in the tropics.
+Modelling of global terrestrial ecosystem GPP though upscaling and calibration of eddy flux measurements indicates a peak in GPP in the tropics, with the highest levels in tropical forests [@beer_terrestrial_2010; @jung_global_2011].
+This is corroborated by analysis of site-level GPP measurements, which show a strong positive correlation between GPP and MAT and MAP [@luyssaert_co2_2007], with the highest GPP values reported in tropical forests.
+The influence of latitude on global patterns of NPP is less clear than that of GPP.
+Simova and Storch (2017) found that, as with GPP, NPP decreases with latitude, peaking in the tropics.
+However, other studies have found the highest values of NPP in temperate forests [@luyssaert_co2_2007; @huston_global_2009].
+Because of the challenges in accurately obtaining unbiased measures of belowground productivity, many studies focus on ANPP in preference to measures of NPP.
+Studies on global patterns of ANPP are equally inconclusive: Gillman et al.
+(2015) found a weak negative relationship between ANPP and latitude, with the relationship becoming stronger in older forest stands.
+These findings were echoed in other studies, which have found weak or no relationships between ANPP and latitude [@huston_global_2009].
+
+Furthermore, there is evidence that different components of productivity show individual responses to climate drivers.
+For example, increases in GPP have been reported to saturate above 25$^\circ$C MAT [@larjavaara_temperature_2012], while increases in NPP are recorded to saturate above 10$^\circ$C MAT [@luyssaert_co2_2007].
+Similarly, allocation to different components of ANPP varies with climate.
+Within the tropics, allocation to canopy NPP appears fairly consistent, with significantly greater variation in allocation to woody and belowground NPP [@litton_carbon_2007; @malhi_allocation_2011].
+Allocation to these structural biomass components has been shown to increase with water availability [@litton_carbon_2007; @bloom_decadal_2016], and is highest in the wet tropics, indicating that control of woody productivity by MAP may be more significant than control of foliar productivity.
+However, these studies are regional, meaning that our understanding of variation in allocation and its relationship to climate on a global scale remains limited.
+In addition, allocation is also influenced by stand age [@de_lucia_forest_2007], nutrient availability [@litton_carbon_2007] and forest structure [@taylor_greater_2019], which can make it challenging to disentangle the effects of climate.
+
+Data that control for stand age and standardize methodologies are required to resolve this question.
+Here, we use a comprehensive global database of forest carbon fluxes to explore global patterns in productivity.
+We explore five broad questions, described in greater detail in table 1:
+  1. How do carbon fluxes vary with latitude?  
+  2. How do carbon fluxes relate to MAT and MAP?  
+  3. How do carbon fluxes relate to other annual climate variables?  
+  4. What is the role of seasonality in explaining variation in carbon fluxes?  
+  5. Per month of the growing season, how does productivity vary with climate?
 

### 2019-08-30 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
 
-There are two major questions to understand: firstly, how forest productivity varies globally - and specifically which areas show the greatest peaks in productivity -; and secondly, which climate variables drive this variation.
-On a global scale, the productivity of forests varies with latitude, with a general trend of increasing productivity towards the tropics [@beer_terrestrial_2010; @jung_global_2011]; however the exact nature of this pattern, and how it varies by component of productivity, is poorly understood.
-This latitudinal gradient is most likely to be explained by climatic gradients in temperature, precipitation, length of growing season, and combinations of the above.
-Productivity is influenced by a range of climatic drivers, including mean annual temperature (MAT) and mean annual precipitation (MAP) [@del_grosso_global_2008], but doesn't necessarily respond linearly to these drivers.
-Disentangling the shape of productivity responses to climate drivers will enable better predictions of future responses under climate change.
+There are two key components to this question: firstly, we want to understand how forest productivity varies on a global scale, and, in particular, we want to know which regions show the highest forest productivity; and secondly, we want to understand the factors driving these patterns in productivity.
 
-*The components of forest productivity.* During photosynthesis, plants capture carbon dioxide from the atmosphere.
-The gross primary productivity (GPP) of an ecosystem is the gross uptake, via photosynthesis, of carbon dioxide by plants in that ecosystem.
-Only a fraction of the carbon captured is assimilated into plant tissue; the rest is used in autotrophic respiration (R~auto~).
-The component of GPP that is stored as plant material is the net primary productivity (NPP) of an ecosystem.
-Net primary productivity can therefore be expressed as:
+On a global scale, the productivity of forests varies with latitude, showing a general trend of decreasing productivity with latitude [@beer_terrestrial_2010; @jung_global_2011].
+Studies agree that productivity is lowest in the boreal regions, and increases into the temperate regions [@luyssaert_co2_2007; @huston_global_2009; @beer_terrestrial_2010; @jung_global_2011].
+However, evidence is inconclusive on whether productivity continues to increase into the tropics, or whether it plateaus in temperate regions.
+Evidence for this is further complicated by the fact that different studies use different measures of productivity to explore these relationships.
+For example, modelling of global terrestrial ecosystem gross primary productivity (GPP) through upscaling and calibration of eddy flux measurements indicates that GPP peaks in tropical forests [@beer_terrestrial_2010; @jung_global_2011].
+This is corroborated by analysis of site-level GPP measurements, which appear to reach their highest levels in tropical forests [@luyssaert_co2_2007].
+In contrast, there is evidence that the highest values of net primary productivity (NPP) may be found in temperate forests [@luyssaert_co2_2007; @huston_global_2009], although other studies find NPP is highest in the tropics, showing a decrease with latitude [@simova_enigma_2017].
+Other studies have chosen to focus exclusively on above-ground net primary productivity (ANPP), finding evidence of a weak negative relationship between ANPP and latitude [@huston_global_2009; @gillman_latitude_2015].
+An understanding of the global patterns of productivity is important; however elucidating the drivers of these patterns will be even more valuable in expanding our understanding of global carbon cycling.
+Primary productivity can be influenced by many factors, which often act across a range of scales, and may show interactive effects with each other.
+On a local scale, stand age [@litton_carbon_2007; @gillman_latitude_2015], management [@simova_enigma_2017]; nutrient availability [@aragao_above-_2009]; and altitude [@girardin_net_2010; @malhi_variation_2017] all impact forest productivity.
+On a global scale, we expect that productivity is most strongly influenced by broad climatic gradients.
 
-$$ NPP = GPP + R~auto~ $$
+We know from previous research that climate is a significant driver of productivity across broad spatial scales [@cleveland_relationships_2011].
+The majority of studies have focused on exploring the relationships between productivity and mean annual temperature (MAT) and mean annual precipitation (MAP), as the most commonly reported site-level climate variables.
+These variables have the advantage that they describe broad trends in temperature and water availability, and therefore capture a lot of global-scale variation in climate.
+There is strong evidence that both MAT and MAP show significant positive relationships with productivity [@chu_does_2016].
+However, as with latitude, the shape of those relationships is not always clear, and, again, is complicated by the use of different measures of productivity across studies.
+There is support across multiple studies for the hypothesis that various measures of primary productivity saturate at high levels of MAP, though the saturation points identified vary from 1500mm [@luyssaert_co2_2007] up to 2445mm MAP [@schuur_productivity_2003].
+Studies of the influence of MAT on productivity are less conclusive.
+Luyssaert et al.
+[-@luyssaert_co2_2007] examined GPP and NPP and found that, while GPP increases linearly with MAT, NPP saturates at around 10$^\circ$C MAT.
+In contrast, Larjavaara and Muller-Landau [-@larjavaara_temperature_2012], find that increases in GPP saturate at approximately 25$^\circ$C MAT, while Schuur [-@schuur_productivity_2003] shows that NPP increases linearly with temperature.
 
-Currently, GPP cannot be measured directly by observing total ecosystem photosynthesis.
-Instead field estimates of GPP have to be derived based on modelling and extrapolation of eddy-covariance studies and measurements of net ecosystem exchange (NEE) [@clark_reviews_2017].
+The influence of these climate variables on productivity is further complicated by the possibility of interactive effects occurring between them.
+Taylor et al.
+[-@taylor_temperature_2017] showed that increased MAP had a negative effect at low MAT, but a positive effect at high levels of MAT, and vice versa.
+Similarly, Luyssaert et al.
+[-@luyssaert_co2_2007] showed that ecosystems limited by low MAP do not benefit from increasing MAT, and vice versa, suggesting that both temperature and water availability are important in explaining productivity.
 
-In contrast, NPP can be calculated through direct field observations.
-In order to achieve greater accuracy in estimating NPP, NPP is often broken down into its component parts, with aboveground NPP (ANPP) and belowground NPP (BNPP) considered separately to each other.
-The components included when estimating ANPP and BNPP often vary between studies, depending on the intensity of fieldwork effort.
-At its most basic level, ANPP can be expressed as:
+While MAT and MAP are important climate variables, they do not capture all aspects of climate that may influence productivity, and may be insufficient to capture the breadth of climatic effects on productivity [@cleveland_relationships_2011].
+There is evidence that productivity also responds to variables such as cloud cover [@taylor_temperature_2017], solar radiation [@fyllas_solar_2017], and potential evapotranspiration [@kerkhoff_plant_2005] in potentially significant ways.
 
-$$ ANPP = NPP~stem~ + NPP~branch~ + NPP~canopy~ $$
-where NPP~stem~ is the annual woody increment of all stems above a specified diameter at breast height (DBH), NPP~branch~ is annual woody branch turnover, and NPP~canopy~ is annual foliage production, including leaves, twigs, and reproductive structures.
-ANPP may also include NPP~VOC~, the annual emission of volatile organic compounds, and NPP~herbivory~, the annual consumption of plant matter by herbivores, but these components are often excluded from field observations as they are much harder to quantify.
-Other components of aboveground productivity that remain largely unquantified include epiphytes, hemiepiphytes, and understory plants [@clark_reviews_2017].
-The majority of current ANPP estimates are based on the assumption that the contribution of these components to overall NPP is insignificant.
+Furthermore, MAT and MAP are very coarse measures of climate, and so fail to capture much variation in climate on an intra-annual scale, including the effects of factors such as growing season length, number of frost-free days, temperature seasonality, and dry season length.
+Some studies have suggested that the apparently strong relationship between MAT and productivity is actually a factor of the correlation between MAT and growing season length [@kerkhoff_plant_2005; @malhi_productivity_2012; @michaletz_convergence_2014; @michaletz_drivers_2018].
+Kerkhoff et al.
+[-@kerkhoff_plant_2005] and Michaletz et al.
+[-@michaletz_convergence_2014] find that, within the growing season, there is no significant relationship between productivity and MAT, indicating that the effect of temperature is due to increased length of growing season, rather than an inherent influence of temperature on productivity.
 
-There are two major subcomponents of BNPP, which can be expressed as:
-
-$$ BNPP = NPP~coarse root~ + NPP~fine root~ $$
-where NPP~coarse root~ is the annual production of coarse roots (typically roots >2mm diameter), and NPP~fine root~ is the annual production of fine roots (typically roots <2mm diameter) [@aragao_above-_2009].
-Calculations of BNPP may also include NPP~exudation~, a measure of annual carbon losses through root exudation, and NPP~symbionts~, the annual carbon allocation to mycorrhizae and legumes, but, as before, this is challenging to quantify and is often excluded from field observations.
-
-BNPP is a poorly understood component of total ecosystem productivity, primarily because of the challenges in gaining accurate field measurements.
-Coarse root productivity is often estimated via extrapolation of NPP~stem~ estimates using allometries that may not have been empirically verified [@clark_net_2001].
-NPP~fine root~ is easier to quantify through soil cores and minirhizotrons, however, sampling tends to be limited to the surface soils, with very few studies sampling to depths below 3 metres [cite].
-As a result, it is possible that BNPP is currently significantly underestimated, despite being a hugely significant component of total ecosystem productivity [@pan_large_2011].
-
-*Which factors influence primary productivity?* Primary productivity is influenced by many factors, which often act across a range of scales, and may show interactive effects with each other.
-On a local scale, stand age [@litton_carbon_2007; @gillman_latitude_2015], management [@simova_enigma_2017]; nutrient availability [@aragao_above-_2009]; and altitude [@girardin_net_2010; @malhi_variation_2017] all impact forest productivity.
-On a global scale, changes in primary productivity are influenced by climatic variables and abiotic gradients, such as the length of growing season [@michaletz_convergence_2014].
-There is some debate over the precise relationship between these drivers and productivity; while MAT and MAP have been argued to be significant predictors of productivity [@chu_does_2016], other studies have found that the correlation between productivity and MAT is a factor of the relationship between productivity and growing season length [@kerkhoff_plant_2005; @malhi_productivity_2012; @michaletz_convergence_2014; @michaletz_drivers_2018].
-Improving understanding of how these factors interact to control global patterns in primary productivity is essential to understanding the global carbon cycle.
-
-Current research into how primary productivity varies with latitude is inconclusive, and - though it has primarily focussed on patterns of GPP, NPP, and ANPP - indicates that different components of productivity may show different relationships with latitude.
-Gross primary productivity is generally thought to be highest in the tropics.
-Modelling of global terrestrial ecosystem GPP though upscaling and calibration of eddy flux measurements indicates a peak in GPP in the tropics, with the highest levels in tropical forests [@beer_terrestrial_2010; @jung_global_2011].
-This is corroborated by analysis of site-level GPP measurements, which show a strong positive correlation between GPP and MAT and MAP [@luyssaert_co2_2007], with the highest GPP values reported in tropical forests.
-The influence of latitude on global patterns of NPP is less clear than that of GPP.
-Simova and Storch (2017) found that, as with GPP, NPP decreases with latitude, peaking in the tropics.
-However, other studies have found the highest values of NPP in temperate forests [@luyssaert_co2_2007; @huston_global_2009].
-Because of the challenges in accurately obtaining unbiased measures of belowground productivity, many studies focus on ANPP in preference to measures of NPP.
-Studies on global patterns of ANPP are equally inconclusive: Gillman et al.
-(2015) found a weak negative relationship between ANPP and latitude, with the relationship becoming stronger in older forest stands.
-These findings were echoed in other studies, which have found weak or no relationships between ANPP and latitude [@huston_global_2009].
-
-Furthermore, there is evidence that different components of productivity show individual responses to climate drivers.
-For example, increases in GPP have been reported to saturate above 25$^\circ$C MAT [@larjavaara_temperature_2012], while increases in NPP are recorded to saturate above 10$^\circ$C MAT [@luyssaert_co2_2007].
-Similarly, allocation to different components of ANPP varies with climate.
-Within the tropics, allocation to canopy NPP appears fairly consistent, with significantly greater variation in allocation to woody and belowground NPP [@litton_carbon_2007; @malhi_allocation_2011].
-Allocation to these structural biomass components has been shown to increase with water availability [@litton_carbon_2007; @bloom_decadal_2016], and is highest in the wet tropics, indicating that control of woody productivity by MAP may be more significant than control of foliar productivity.
-However, these studies are regional, meaning that our understanding of variation in allocation and its relationship to climate on a global scale remains limited.
-In addition, allocation is also influenced by stand age [@de_lucia_forest_2007], nutrient availability [@litton_carbon_2007] and forest structure [@taylor_greater_2019], which can make it challenging to disentangle the effects of climate.
-
-Data that control for stand age and standardize methodologies are required to resolve this question.
-Here, we use a comprehensive global database of forest carbon fluxes to explore global patterns in productivity.
-We explore five broad questions, described in greater detail in table 1:
+In order to approach these broad and complex issues, we simplify the major gaps in our knowledge to five key questions, presented in greater detail in Table 1:  
 1.	How do carbon fluxes vary with latitude? 
@@
 5.	Per month of the growing season, how does productivity vary with climate?
+
+To resolve these questions, we require data that, firstly, control for variables influencing productivity, including stand age, disturbance regime, elevation, and methodology; and secondly includes all major carbon fluxes.
+Here we use a comprehensive global database of forest carbon fluxes to explore the above questions for nine carbon fluxes, allowing for an in-depth exploration of the effect of climate on global productivity.
 
@@
 This indicates that the main mechanism through with precipitation influences productivity is in a secondary capacity, with its overall influence being mediated by the effects of temperature.
-This also suggests that although temperature explains a large proportion of variation, it cannot fully explain productivity alone, and a measure of water availability is necessary to understand patterns of productivity.
-This is likely to be significant in the future, as the effects of a warming climate will be mediated through changes in precipitation regime.
+This also suggests that although temperature explains a large proportion of variation, it cannot fully explain productivity alone, and a measure of water availability is necessary to fully understand patterns of productivity.
+This is likely to be significant in the future, as it suggests that the effects of a warming climate will be mediated through changes in precipitation regime.
 

### 2019-09-07 manuscript/manuscript draft.Rmd
updates to manuscript
@beckybanbury, this is creating a file named manuscript-draft.pdf instead of manuscript_draft.pdf. Not sure why...

@@
 ---
-title: "Manuscript draft"
+title: "Global patterns in forest productivity"
+author: "Rebecca Banbury Morgan, Valentine Herrmann, Norbert Kunert, Ben Bond-Lamberty, Helene Muller-Landau, [MORE??-CHECK], Kristina J. Anderson-Teixeira"
 output:
@@
 Understanding the patterns of forest productivity on a global scale, and the drivers behind them, is therefore a priority in forest research.
-
-There are two key components to this question: firstly, we want to understand how forest productivity varies on a global scale, and, in particular, we want to know which regions show the highest forest productivity; and secondly, we want to understand the factors driving these patterns in productivity.
+[**KAT-work on this; consider shifting top focus from global C cycle to centrality of forest productivity to ecology & climate science.** ] Given the importance of understanding broad-scale patterns in forest productivity, there have been many meta-analyses on the theme.
+However, our ability to draw general macroscopic conclusions regarding global variation in multiple productivity variables with respect to climate has been limited in that these analyses often mix forests that vary in stand age, disturbance history, and/or management status; do not always sufficiently parse related variables (*e.g.*, combining net primary productivity records with and without belowground components); and typically consider only one or a few variables at a time.
+The recent development of a global forest carbon database synthesizing multiple variables and including records of stand history (ForC; [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018]) opens up the possibility for a standardized analysis of global scale variation in multiple components of forest productivity and the principle climatic drivers of these patterns.
 
@@
 
-In order to approach these broad and complex issues, we simplify the major gaps in our knowledge to five key questions, presented in greater detail in Table 1:  
+In order to approach these broad and complex issues, we simplify the major gaps in our knowledge to five key questions and corresponding specific predictions (Table 1):  
 1.	How do carbon fluxes vary with latitude? 
@@
 
-### Methods
+### Materials and Methods
 
-Analyses were conducted on data contained in the open-access ForC database [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc:_2018].
+Analyses were conducted on data contained in the open-access ForC database [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018].
 This database contains records of field-based measurements of forest carbon stocks and annual fluxes, compiled from original publications and existing data compilations and databases.
 Associated data, such as stand age, measurement methodologies, and disturbance history, are also included.
-For each site, site geographic co-ordinates were used to extract the Koeppen-Geiger zone from the ESRI Koeppen-Geiger map [cite], and the FAO Global Ecological Zone from the ???.
-Additional targeted literature searches were conducted to identify any further available data on primary productivity, with particular focus on old-growth forests in temperate and boreal regions.
-ForC currently contains 29768 records from 5227 plots, representing 20 distinct ecozones across all forested biogeographic and climate zones.
+The database was significantly expanded since the publication of [@anderson-teixeira_forc_2018] through integration with the Global Database of Soil Respiration (SRDB; **Bond-Lamberty, DOI: 10.5194/bg-7-1915-2010**).
+Additional targeted literature searches were conducted to identify any further available data on primary productivity, with particular focus on mature forests in temperate and boreal regions.
+ForC currently contains **UPDATE STATS** 29768 records from 5227 plots, representing 20 distinct ecozones across all forested biogeographic and climate zones.
+We used ForC v.XX, archived on Zenodo with DOI XXX.
 
@@
 
-We find no evidence that allocation between fluxes varies with latitude or climate.
+We find no evidence that allocation between fluxes varies substantially with latitude or climate.
 There were no significant results from regressing ratios of carbon fluxes against latitude, or against any of the climate variables.
@@
 We would expect that this would reduce the strength of the effect of climate on subsidiary fluxes.
+
+### Acknowledgements

### 2019-09-08 manuscript/manuscript draft.Rmd
quick pass through entire manuscript
see issues #21 and #25 for more.

@@
 
-*Closure of fluxes* Analysis of component fluxes indicates that measures of smaller component fluxes do approximately sum to measures of larger fluxes (fig.
-2).
-Modelled estimates of GPP, generated from the sum of NPP and R auto; NPP, generated from the sum of ANPP and BNPP~root~; and ANPP, generated from the sum of ANPP~foliage~ and ANPP~woody stem~, fall completely within the confidence intervals of the regressions of field estimates of GPP, NPP and ANPP respectively.
+In total, we analyzed ## records from ## C flux variables taken from forests that had experienced no major anthropogenic disturbances within the past 100 years from # distinct geographic areas (Fig.
+1, Table 2).
+[**Add a bit about geographic representation**]
 
@@
 
-*How do carbon fluxes vary with latitude?* All major carbon fluxes increase linearly with decreasing latitude (fig. 3).
+*How do carbon fluxes vary with latitude?*
+
+All major carbon fluxes increase linearly with decreasing latitude (fig. 3).
 
@@
 
+[*Relationships and differences among fluxes- *] In general, smaller component fluxes summed approximately to  larger fluxes across the latitudinal gradient (fig.
+2).
+That is, modelled estimates of GPP, generated from the sum of NPP and R auto; NPP, generated from the sum of ANPP and BNPP~root~; and ANPP, generated from the sum of ANPP~foliage~ and ANPP~woody stem~, fall completely within the confidence intervals of the regressions of field estimates of GPP, NPP and ANPP respectively.
+
 We find no evidence that allocation between fluxes varies substantially with latitude or climate.
 There were no significant results from regressing ratios of carbon fluxes against latitude, or against any of the climate variables.
+
+[**move differences among R2s here?**]
 
@@
 
-*How does productivity relate to other climate variables?* Our results indicate that productivity is most strongly explained by temperature, with temperature-related climate variables coming out as strong predictors of productivity.
+*How does productivity relate to other climate variables?* Our results indicate that productivity is most strongly explained by temperature at the global scale, with temperature-related climate variables coming out as strong predictors of productivity.
 In addition to MAT, temperature seasonality, annual temperature range, and annual frost days were consistently identified as good predictors of productivity across fluxes.
@@
 We found a significant relationship between productivity and potential evapotranspiration for all fluxes.
-ANPP~foliage~, BNPP~fine root~ and R~root~ increased linearly with PET, however all other fluxes showed a polynomial relationship with PET (fig.
+ANPP~foliage~, BNPP~fine root~ and R~root~ increased linearly with PET; however, all other fluxes showed a polynomial relationship with PET (fig.
 5).
@@
 
-*What is the role of seasonality in explaining productivity?* Temperature seasonality is a significant predictor of productivity.
+*What is the role of seasonality in explaining productivity?* Temperature seasonality was a significant predictor of productivity.
 We found a polynomial relationship with productivity for GPP, NPP, ANPP, and R~root~.
@@
 
-In contrast, there is no significant effect of precipitation seasonality on productivity.
+In contrast, there was no significant effect of precipitation seasonality on productivity.
 
@@
 Our results show that productivity decreases linearly with latitude.
-Climate explains a significant proportion of variation in all carbon fluxes, with temperature variables being the best predictors of productivity.
-While other climate variables are significant predictors of productivity, none of them improve on the explanatory power of temperature-related variables.
+Climate explains a significant proportion of variation in all carbon fluxes, with temperature variables being the best predictors of productivity on this global scale.
+While other climate variables are significant predictors of productivity, none of them improve on the explanatory power of temperature-related variables in general or MAT specifically.
 Water availability is an important factor in explaining productivity on a global scale: we find a positive influence of precipitation at low MAP, with saturation at higher levels of MAP.
@@
 
+**[needs strong concluding paragraph]**
+
 ### Acknowledgements

### 2019-09-09 scripts/mat_map_interaction.R
Update mat_map_interaction.R

@@ -112,7 +112,7 @@ pannel.nb <- 1
 response.variables <- c("GPP", "NPP", "BNPP_root", "BNPP_root_fine", "ANPP", "ANPP_foliage", "ANPP_woody_stem","R_auto", "R_auto_root")
 
 png(file = paste0("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/figures/final_figures/interactions/mat_map_interaction.png"), width = 2255, height = 2000, units = "px", res = 300)
-par(mfrow = c(3,3), mar = c(0,0,0,0), oma = c(5,7,2,0), xpd = T)
+par(mfrow = c(3,3), mar = c(1,0,0,2), oma = c(5,8,2,0), xpd = T)
 
 for (response.v in response.variables){
   
@@ -166,7 +166,7 @@ for (response.v in response.variables){
   
   # par(mfrow = c(1,1), mar = c(0,0,0,0), oma = c(5,5,2,0))
   ylim = range(df$mean)
-  xlim = range(df$mat)
+  xlim = c(-10, 28)# range(df$mat)
   
   if (significant.effect.of.interaction | significant.effect.of.additive) {
     # predict 
@@ -179,7 +179,7 @@ for (response.v in response.variables){
 
     for (i in seq(along = lower_bound)){
       map.in.bin <- df[df$map > lower_bound[[i]] & df$map < upper_bound[[i]], ]
-      if(first.plot) plot(mean ~ mat, data = map.in.bin, xlab = "", ylab = "", xaxt = "n", yaxt = "n", ylim = ylim, col = plasma(5)[i], xlim = xlim)
+      if(first.plot) plot(mean ~ mat, data = map.in.bin, xlab = "", ylab = "", ylim = ylim, col = plasma(5)[i], xlim = xlim, xaxt = "n", las = 1, tck = 0.02, mgp = c(3, 0.5, 0))
       if(!first.plot) points(mean ~ mat, data = map.in.bin, xlab = "", ylab = "", xaxt = "n", yaxt = "n", ylim = ylim, col = plasma(5)[i], xlim = xlim) 
       newDat <- expand.grid(mat = seq(min(map.in.bin$mat), max(map.in.bin$mat), length.out = 100), map = midpoint[[i]])
       newDat$fit <- predict(mod.full, newDat, re.form = NA)
@@ -206,7 +206,7 @@ for (response.v in response.variables){
[…]

### 2019-09-09 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
 ---
-title: "Global patterns in forest productivity"
+title: "Manuscript draft"
 author: "Rebecca Banbury Morgan, Valentine Herrmann, Norbert Kunert, Ben Bond-Lamberty, Helene Muller-Landau, [MORE??-CHECK], Kristina J. Anderson-Teixeira"
@@
 However, our ability to draw general macroscopic conclusions regarding global variation in multiple productivity variables with respect to climate has been limited in that these analyses often mix forests that vary in stand age, disturbance history, and/or management status; do not always sufficiently parse related variables (*e.g.*, combining net primary productivity records with and without belowground components); and typically consider only one or a few variables at a time.
-The recent development of a global forest carbon database synthesizing multiple variables and including records of stand history (ForC; [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018]) opens up the possibility for a standardized analysis of global scale variation in multiple components of forest productivity and the principle climatic drivers of these patterns.
+The recent development of a global forest carbon database synthesizing multiple variables and including records of stand history (ForC; [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc:_2018]) opens up the possibility for a standardized analysis of global scale variation in multiple components of forest productivity and the principle climatic drivers of these patterns.
 
 On a global scale, the productivity of forests varies with latitude, showing a general trend of decreasing productivity with latitude [@beer_terrestrial_2010; @jung_global_2011].
-Studies agree that productivity is lowest in the boreal regions, and increases into the temperate regions [@luyssaert_co2_2007; @huston_global_2009; @beer_terrestrial_2010; @jung_global_2011].
+Studies agree that productivity is lowest in the boreal regions, and increases into the temperate regions [@luyssaert_co_2007; @huston_global_2009; @beer_terrestrial_2010; @jung_global_2011].
 However, evidence is inconclusive on whether productivity continues to increase into the tropics, or whether it plateaus in temperate regions.
@@
 For example, modelling of global terrestrial ecosystem gross primary productivity (GPP) through upscaling and calibration of eddy flux measurements indicates that GPP peaks in tropical forests [@beer_terrestrial_2010; @jung_global_2011].
-This is corroborated by analysis of site-level GPP measurements, which appear to reach their highest levels in tropical forests [@luyssaert_co2_2007].
-In contrast, there is evidence that the highest values of net primary productivity (NPP) may be found in temperate forests [@luyssaert_co2_2007; @huston_global_2009], although other studies find NPP is highest in the tropics, showing a decrease with latitude [@simova_enigma_2017].
+This is corroborated by analysis of site-level GPP measurements, which appear to reach their highest levels in tropical forests [@luyssaert_co_2007].
+In contrast, there is evidence that the highest values of net primary productivity (NPP) may be found in temperate forests [@luyssaert_co_2007; @huston_global_2009], although other studies find NPP is highest in the tropics, showing a decrease with latitude [@simova_enigma_2017].
 Other studies have chosen to focus exclusively on above-ground net primary productivity (ANPP), finding evidence of a weak negative relationship between ANPP and latitude [@huston_global_2009; @gillman_latitude_2015].
@@
 However, as with latitude, the shape of those relationships is not always clear, and, again, is complicated by the use of different measures of productivity across studies.
-There is support across multiple studies for the hypothesis that various measures of primary productivity saturate at high levels of MAP, though the saturation points identified vary from 1500mm [@luyssaert_co2_2007] up to 2445mm MAP [@schuur_productivity_2003].
+There is support across multiple studies for the hypothesis that various measures of primary productivity saturate at high levels of MAP, though the saturation points identified vary from 1500mm [@luyssaert_co_2007] up to 2445mm MAP [@schuur_productivity_2003].
 Studies of the influence of MAT on productivity are less conclusive.
 Luyssaert et al.
-[-@luyssaert_co2_2007] examined GPP and NPP and found that, while GPP increases linearly with MAT, NPP saturates at around 10$^\circ$C MAT.
+[-@luyssaert_co_2007] examined GPP and NPP and found that, while GPP increases linearly with MAT, NPP saturates at around 10$^\circ$C MAT.
 In contrast, Larjavaara and Muller-Landau [-@larjavaara_temperature_2012], find that increases in GPP saturate at approximately 25$^\circ$C MAT, while Schuur [-@schuur_productivity_2003] shows that NPP increases linearly with temperature.
@@
 Similarly, Luyssaert et al.
-[-@luyssaert_co2_2007] showed that ecosystems limited by low MAP do not benefit from increasing MAT, and vice versa, suggesting that both temperature and water availability are important in explaining productivity.
+[-@luyssaert_co_2007] showed that ecosystems limited by low MAP do not benefit from increasing MAT, and vice versa, suggesting that both temperature and water availability are important in explaining productivity.
 
@@
 
-Analyses were conducted on data contained in the open-access ForC database [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018].
+Analyses were conducted on data contained in the open-access ForC database [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc:_2018].
 This database contains records of field-based measurements of forest carbon stocks and annual fluxes, compiled from original publications and existing data compilations and databases.
 Associated data, such as stand age, measurement methodologies, and disturbance history, are also included.
-The database was significantly expanded since the publication of [@anderson-teixeira_forc_2018] through integration with the Global Database of Soil Respiration (SRDB; **Bond-Lamberty, DOI: 10.5194/bg-7-1915-2010**).
+The database was significantly expanded since the publication of [@anderson-teixeira_forc:_2018] through integration with the Global Database of Soil Respiration (SRDB; **Bond-Lamberty, DOI: 10.5194/bg-7-1915-2010**).
 Additional targeted literature searches were conducted to identify any further available data on primary productivity, with particular focus on mature forests in temperate and boreal regions.

### 2019-09-09 manuscript/manuscript draft.Rmd
update manuscript draft

@@
 | ANPP~foliage~ | Net primary production of foliage | Foliage | Direct measurement of litterfall, correcting for changes in leaf biomass when measured | 98 | 94 |
-| ANPP~woodystem~ | Net primary production of woody stems | Woody stems | Direct measurement of stem growth increment | 266 | 252 |
+| ANPP~woody~ ~stem~ | Net primary production of woody stems | Woody stems | Direct measurement of stem growth increment | 266 | 252 |
 | BNPP~root~ | Belowground net primary production | Coarse and fine roots | Direct measurement of one or more of: fine root turnover, soil cores, root ingrowth cores, minirhizotrons; indirect estimates of coarse roots using allometries based on aboveground stem increment measures | 107 | 87 |
-| BNPP~fineroot~ | Net primary production of fine roots | Fine roots | Direct measurement of one or more of: minirhizotrons, fine root turnover, soil cores, root ingrowth cores | 90 | 77 |
+| BNPP~fine~ ~root~ | Net primary production of fine roots | Fine roots | Direct measurement of one or more of: minirhizotrons, fine root turnover, soil cores, root ingrowth cores | 90 | 77 |
 | R~auto~ | Annual autotrophic respiration, including above- and belowground components | Foliage, stem, and root | Chamber measurements of component gas exchange | 22 | 22 |
@@
 2. NPP = ANPP + BNPP  
-3. ANPP = ANPP~foliage~ + ANPP~woody stem~  
+3. ANPP = ANPP~foliage~ + ANPP~woody~ ~stem~  
 4. Total belowground carbon flux = BNPP + R~root~
 
-*Allocation to carbon fluxes along latitudinal gradients.* Variation in allocation to component carbon fluxes along latitudinal gradients was explored for a range of pairings: firstly, GPP:NPP, ANPP:BNPP, and ANPP~foliage~:ANPP~woody stem~; and secondly, the ratio of NPP to each of ANPP, BNPP, ANPP~foliage~, and ANPP~woody stem~.
+*Allocation to carbon fluxes along latitudinal gradients.* Variation in allocation to component carbon fluxes along latitudinal gradients was explored for a range of pairings: firstly, GPP:NPP, ANPP:BNPP, and ANPP~foliage~:ANPP~woody~ ~stem~; and secondly, the ratio of NPP to each of ANPP, BNPP, ANPP~foliage~, and ANPP~woody~ ~stem~.
 For each set of paired fluxes, measurements taken at the same site and plot, and in the same year, were paired together, and the ratio of each pair of measurements calculated.
@@
 2).
-That is, modelled estimates of GPP, generated from the sum of NPP and R auto; NPP, generated from the sum of ANPP and BNPP~root~; and ANPP, generated from the sum of ANPP~foliage~ and ANPP~woody stem~, fall completely within the confidence intervals of the regressions of field estimates of GPP, NPP and ANPP respectively.
+That is, modelled estimates of GPP, generated from the sum of NPP and R auto; NPP, generated from the sum of ANPP and BNPP~root~; and ANPP, generated from the sum of ANPP~foliage~ and ANPP~woody~ ~stem~, fall completely within the confidence intervals of the regressions of field estimates of GPP, NPP and ANPP respectively.
 
@@
 
-There was a significant interactive effect between MAT and MAP for GPP, BNPP~root~, BNPP~fine root~, ANPP, ANPP~woody stem~, and R~root~ (fig.4).
+There was a significant interactive effect between MAT and MAP for GPP, BNPP~root~, BNPP~fine~ ~root~, ANPP, ANPP~woody~ ~stem~, and R~root~ (fig.4).
 There was a significant additive effect for R~auto~.
@@
 We found a significant relationship between productivity and potential evapotranspiration for all fluxes.
-ANPP~foliage~, BNPP~fine root~ and R~root~ increased linearly with PET; however, all other fluxes showed a polynomial relationship with PET (fig.
+ANPP~foliage~, BNPP~fine~ ~root~ and R~root~ increased linearly with PET; however, all other fluxes showed a polynomial relationship with PET (fig.
 5).
@@
 Vapour pressure deficit was a significant predictor of productivity for all fluxes.
-BNPP~fine root~ showed a linear relationship with vapour pressure deficit (R^2^ = 0.07, p<0.05), but all other fluxes showed a polynomial relationship (fig.
+BNPP~fine~ ~root~ showed a linear relationship with vapour pressure deficit (R^2^ = 0.07, p<0.05), but all other fluxes showed a polynomial relationship (fig.
 5).
@@
 
-```{r echo=FALSE, out.width='100%', fig.cap = "Plots of primary productivity $(MgC~ha^-1~yr^-1)$ regressed against the climate variables identified as the best predictors: mean annual temperature; temperature seasonality; annual temperature range; and vapour pressure.
+```{r echo=FALSE, out.width='100%', fig.cap = "Plots of primary productivity $(MgC~ha^-1~yr^-1)$ regressed against (a) mean annual temperature; (b) mean annual precipitation; (c) potential evapotranspiration, (d) vapour pressure deficit; (e) temperature seasonality; (f) length of growing season.
 Lines of best fit are plotted according to the best model selected during analysis.
@@
 5).
-ANPP~foliage~, ANPP~woody stem~ and R~auto~ decrease linearly with temperature seasonality.
+ANPP~foliage~, ANPP~woody~ ~stem~ and R~auto~ decrease linearly with temperature seasonality.
 Temperature seasonality is strongly correlated with annual temperature range, and, as expected, all fluxes show almost identical responses to it.
@@
 For each of temperature, precipitation, PET, and solar radiation, we find a small effect of climate for certain carbon fluxes.
-There is a small increase in productivity with temperature and precipitation for ANPP (with temperature R^2^ = 0.10, p<0.001; with precipitation R^2^ = 0.04, p<0.05) and ANPP_foliage (with temperature R^2^ = 0.16, p<0.01; with precipitation R^2^ = 0.09, p<0.05).
-Productivity increases with solar radiation for GPP (R^2^ = 0.21, p<0.001), NPP (R^2^ = 0.21, p<0.001), BNPP (R^2^ = 0.16, p<0.001) and BNPP~fine root~ (R^2^ = 0.12, p<0.01), and with PET for GPP (R^2^ = 0.15, p<0.01), NPP (R^2^ = 0.18, p<0.01), BNPP (R^2^ = 0.23, p<0.0001), BNPP~fine root~ (R^2^ = 0.11, p<0.05), and ANPP~woody stem~ (R^2^ = 0.06, p<0.05).
+There is a small increase in productivity with temperature and precipitation for ANPP (with temperature R^2^ = 0.10, p<0.001; with precipitation R^2^ = 0.04, p<0.05) and ANPP~foliage~ (with temperature R^2^ = 0.16, p<0.01; with precipitation R^2^ = 0.09, p<0.05).
+Productivity increases with solar radiation for GPP (R^2^ = 0.21, p<0.001), NPP (R^2^ = 0.21, p<0.001), BNPP (R^2^ = 0.16, p<0.001) and BNPP~fine~ ~root~ (R^2^ = 0.12, p<0.01), and with PET for GPP (R^2^ = 0.15, p<0.01), NPP (R^2^ = 0.18, p<0.01), BNPP (R^2^ = 0.23, p<0.0001), BNPP~fine~ ~root~ (R^2^ = 0.11, p<0.05), and ANPP~woody~ ~stem~ (R^2^ = 0.06, p<0.05).
 
@@
 | ANPP~foliage~ | 0.54 | 0.63 | 0.54 | 0.35 | 0.64 |
-| ANPP~woodystem~ | 0.17 | 0.23 | 0.12 | 0.19 | 0.28 |
+| ANPP~woody~ ~stem~ | 0.17 | 0.23 | 0.12 | 0.19 | 0.28 |
 | BNPP~root~ | 0.33 | 0.26 | 0.23 | 0.36 | 0.35 |
-| BNPP~fineroot~ | 0.15 | 0.11 | 0.16 | 0.11 | 0.19 |
+| BNPP~fine~ ~root~ | 0.15 | 0.11 | 0.16 | 0.11 | 0.19 |
 | R~root~ | 0.22 | 0.21 | 0.30 | 0.01* | 0.28 |
@@
 With the exception of ANPP~foliage~, the proportion of variation explained by climate and latitude in subsidiary fluxes is much lower.
-Climate explains at most 23% of variation in ANPP~woody stem~, 16% in BNPP~fine root~, and 30% in R~root~.
+Climate explains at most 23% of variation in ANPP~woody~ ~stem~, 16% in BNPP~fine~ ~root~, and 30% in R~root~.
 In contrast, climate strongly explains variation in ANPP~foliage~, with mean annual temperature explaining 63% of variation.
@@
 It is challenging to account for this type of variability in quality of field data, and our analyses largely assume good faith in data quality.
-While we have strong confidence in our analyses of well-understood fluxes, such as GPP, ANPP, and ANPP~woody stem~, we recommend caution in drawing firm conclusions from our analyses of BNPP, BNPP~fine root~, and, to some extent, NPP.
+While we have strong confidence in our analyses of well-understood fluxes, such as GPP, ANPP, and ANPP~woody~ ~stem~, we recommend caution in drawing firm conclusions from our analyses of BNPP, BNPP~fine~ ~root~, and, to some extent, NPP.
 We expect that, as field techniques improve the accuracy of estimates of these fluxes in the future, our understanding of the relationship of BNPP with climate variables will change.

### 2019-09-09 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@
 *Climate datasets.* Where site-level data on mean annual temperature, mean annual precipitation, and latitude were available in the primary literature, this data was compiled and entered directly into the ForC database.
-In addition to this data, climate data for each site was extracted from five open-access climate datasets based on site geographic co-ordinates.
+Based on the geographic co-ordinates for each site, data on a further 11 climate variables was extracted from five open-access climate datasets: WorldClim [@hijmans_very_2005], WorldClim2 [@fick_worldclim_2017], the Climate Research Unit (CRU) time-series dataset v.
+4.03 [@harris_updated_2014], the Global Aridity Index and Potential Evapotranspiration Climate Database [@trabucco_global_2018], and TerraClimate [@abatzoglou_terraclimate_2018] (see Supplementary Information S1 for details of climate variables).
 Where site-level data was missing for mean annual temperature and/or mean annual precipitation, data was extracted from the WorldClim dataset.
-
-| Database | Variables downloaded | Citation |
-|-----------------------|--------------------------------------------|---------------------|
-| WorldClim | Mean annual temperature; temperature seasonality; annual temperature range; mean annual precipitation | [@hijmans_very_2005] |
-| WorldClim2 | Solar radiation | [@fick_worldclim_2017] |
-| Climate Research Unit (CRU) time-series dataset v 4.03 | Cloud cover; annual frost days; annual wet days; potential evapotranspiration | [@harris_updated_2014] |
-| Global Aridity Index and Potential Evapotranspiration Climate Database | Aridity; potential evapotranspiration | [@trabucco_global_2018] |
-| TerraClimate | Vapour pressure deficit | [@abatzoglou_terraclimate_2018] |
-
-Table: Sources of climate data
 
@@
 
-*How do carbon fluxes vary with latitude?*
+*How does productivity vary with latitude?*
 
-All major carbon fluxes increase linearly with decreasing latitude (fig. 3).
+All major carbon fluxes increased linearly with decreasing latitude (fig. 3).
 
@@
 
-Latitude is a strong predictor for many of the carbon fluxes, explaining 64% of variation in GPP (n = 254, p<0.0001), 50% in NPP (n = 114, p<0.0001) and 45% in ANPP (n = 259, p<0.0001).
+Latitude was a strong predictor for many of the carbon fluxes, explaining 64% of variation in GPP (n = 254, p<0.0001), 50% in NPP (n = 114, p<0.0001) and 45% in ANPP (n = 259, p<0.0001).
 For all fluxes, their relationship with latitude was best predicted by the linear model.
 
-[*Relationships and differences among fluxes- *] In general, smaller component fluxes summed approximately to  larger fluxes across the latitudinal gradient (fig.
+*Relationships and differences among fluxes.* In general, smaller component fluxes summed approximately to  larger fluxes across the latitudinal gradient (fig.
 2).
-That is, modelled estimates of GPP, generated from the sum of NPP and R auto; NPP, generated from the sum of ANPP and BNPP~root~; and ANPP, generated from the sum of ANPP~foliage~ and ANPP~woody~ ~stem~, fall completely within the confidence intervals of the regressions of field estimates of GPP, NPP and ANPP respectively.
+That is, modelled estimates of GPP, generated from the sum of NPP and R auto; NPP, generated from the sum of ANPP and BNPP~root~; and ANPP, generated from the sum of ANPP~foliage~ and ANPP~woody~ ~stem~, fell completely within the confidence intervals of the regressions of field estimates of GPP, NPP and ANPP respectively.
 
-We find no evidence that allocation between fluxes varies substantially with latitude or climate.
+We found no evidence that allocation between fluxes varied substantially with latitude or climate.
 There were no significant results from regressing ratios of carbon fluxes against latitude, or against any of the climate variables.
 
-[**move differences among R2s here?**]
+R^2^ values were generally highest in the major fluxes, and decreased in subsidiary fluxes (Supporting Information S2).
+Of the major fluxes, R~auto~ and GPP were the most strongly explained by latitude and climate, with climate explaining at most 71% of variation in GPP, and 65% in R~auto~.
+The proportion of variation explained by climate and latitude decreased in NPP and ANPP, with climate explaining at most 51% of variation in NPP and 44% in ANPP.
+Of the major fluxes, BNPP~root~ was the least well explained by climate and latitude, with climate explaining at most 36% of variation.
+
+With the exception of ANPP~foliage~, the proportion of variation explained by climate and latitude in subsidiary fluxes was much lower.
+Climate explained at most 24% of variation in ANPP~woody~ ~stem~, 19% in BNPP~fine~ ~root~, and 27% in R~root~.
+In contrast, climate strongly explained variation in ANPP~foliage~, with mean annual temperature explaining 58% of variation.
+This pattern was also seen in the R^2^ values for multivariate models.
 
@@
 MAT and MAP are the most commonly reported site-level climate variables, and much previous research into the effect of climate on forest productivity has focused on these as key climate variables.
-MAT is a significant (p<0.05) and strong predictor of productivity for all carbon fluxes tested, with all fluxes showing a linear increase with temperature (fig.
+MAT was a significant (p<0.05) and strong predictor of productivity for all carbon fluxes tested, with all fluxes showing a linear increase with temperature (fig.
 5).
@@
 MAP was found to be a significant (p<0.05) but poor predictor of productivity, explaining, with the exception of R~auto~, at most 37% of variation in carbon flux.
-For the majority of fluxes productivity is best predicted by a polynomial model.
-Productivity increases with precipitation, up until a saturation point at between 3000 and 4000mm annual precipitation, above which productivity starts to decrease (fig.
+For the majority of fluxes productivity was best predicted by a polynomial model.
+Productivity increased with precipitation, up until a saturation point at between 3000 and 4000mm annual precipitation, above which productivity started to decrease (fig.
 5).
-The notable exception to this is GPP: the model indicates that GPP continues to increase with precipitation up to measures of at least 5000mm annually (p<0.0001, R^2^ = 0.33.
-Data above this point is not available, but the model trend indicates that the saturation point for this model is around 5000mm mean annual precipitation.
+The notable exception to this was GPP: the model indicated that GPP continued to increase with precipitation up to measures of at least 5000mm annually (p<0.0001, R^2^ = 0.33.
+Data above this point was not available, but the model trend indicated that the saturation point for this model would be around 5000mm MAP.
 
@@
 
-*How does productivity relate to other climate variables?* Our results indicate that productivity is most strongly explained by temperature at the global scale, with temperature-related climate variables coming out as strong predictors of productivity.
+*How does productivity relate to other climate variables?* Our results indicated that productivity was most strongly explained by temperature at the global scale, with temperature-related climate variables coming out as strong predictors of productivity.
 In addition to MAT, temperature seasonality, annual temperature range, and annual frost days were consistently identified as good predictors of productivity across fluxes.
@@
 5).
-We find strong evidence for a saturation point or peak with PET: productivity tends to increase at values below 1000mm, before saturating between 1200 and 1700mm.
-There is evidence that productivity begins to decrease at values above 1800mm PET.
+We found strong evidence for a saturation point or peak with PET: productivity tended to increase at values below 1000mm, before saturating between 1200 and 1700mm.
+There was evidence that productivity begins to decrease at values above 1800mm PET.
 
@@
 
-All fluxes, with the exception of R~root~, show a positive linear relationship with solar radiation.
-Solar radiation explains a low proportion of variability in productivity for all fluxes, explaining less than 20% of the variation in each flux, with the exception of R~auto~ (R^2^ = 0.26, p<0.05).
+All fluxes, with the exception of R~root~, showed a positive linear relationship with solar radiation.
+Solar radiation explained a low proportion of variability in productivity for all fluxes, explaining less than 20% of the variation in each flux, with the exception of R~auto~ (R^2^ = 0.26, p<0.05).
 
@@
 We found a polynomial relationship with productivity for GPP, NPP, ANPP, and R~root~.
-Productivity decreases rapidly as seasonality increases, with the rate of decrease slowing as seasonality increases (fig.
+Productivity decreased rapidly as seasonality increased, with the rate of decrease slowing as seasonality increased (fig.
 5).
-ANPP~foliage~, ANPP~woody~ ~stem~ and R~auto~ decrease linearly with temperature seasonality.
-Temperature seasonality is strongly correlated with annual temperature range, and, as expected, all fluxes show almost identical responses to it.
-Productivity is highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower.
+ANPP~foliage~, ANPP~woody~ ~stem~ and R~auto~ decreased linearly with temperature seasonality.
+Temperature seasonality was strongly correlated with annual temperature range, and, as expected, all fluxes showed almost identical responses to it.
+Productivity was highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower.
 
@@
 
-We find a significant relationship between length of growing season and productivity, with all fluxes showing a linear increase in productivity with length of growing season (fig.
+We found a significant relationship between length of growing season and productivity, with all fluxes showing a linear increase in productivity with length of growing season (fig.
 5).
-Length of growing season is a strong predictor of productivity, explaining 51% of variation in GPP, 39% of variation in NPP, and 34% of variation in ANPP, but it is not identified as a stronger predictor than MAT for any of the fluxes analysed.
+Length of growing season was a strong predictor of productivity, explaining 51% of variation in GPP, 39% of variation in NPP, and 34% of variation in ANPP, but it was not identified as a stronger predictor than MAT for any of the fluxes analysed.
 
-*Within the growing season, how does productivity vary with climate?* Within growing season months, we find that climate has a much weaker effect on productivity.
-For each of temperature, precipitation, PET, and solar radiation, we find a small effect of climate for certain carbon fluxes.
-There is a small increase in productivity with temperature and precipitation for ANPP (with temperature R^2^ = 0.10, p<0.001; with precipitation R^2^ = 0.04, p<0.05) and ANPP~foliage~ (with temperature R^2^ = 0.16, p<0.01; with precipitation R^2^ = 0.09, p<0.05).
-Productivity increases with solar radiation for GPP (R^2^ = 0.21, p<0.001), NPP (R^2^ = 0.21, p<0.001), BNPP (R^2^ = 0.16, p<0.001) and BNPP~fine~ ~root~ (R^2^ = 0.12, p<0.01), and with PET for GPP (R^2^ = 0.15, p<0.01), NPP (R^2^ = 0.18, p<0.01), BNPP (R^2^ = 0.23, p<0.0001), BNPP~fine~ ~root~ (R^2^ = 0.11, p<0.05), and ANPP~woody~ ~stem~ (R^2^ = 0.06, p<0.05).
-
-*Does climate explain the same proportion of variation in different components of primary productivity?*
-
-| Variable | Latitude | Mean annual temperature | Temperature seasonality | Potential evapotranspiration | MAT*MAP |
-|----------|----------|----------|----------|--------------|----------|
-| R~auto~ | 0.65 | 0.77 | 0.62 | 0.63 | NA |
-| GPP | 0.64 | 0.63 | 0.72 | 0.36 | 0.64 |
-| NPP | 0.50 | 0.52 | 0.52 | 0.33 | 0.56 |
-| ANPP | 0.45 | 0.46 | 0.42 | 0.28 | 0.48 |
-| ANPP~foliage~ | 0.54 | 0.63 | 0.54 | 0.35 | 0.64 |
-| ANPP~woody~ ~stem~ | 0.17 | 0.23 | 0.12 | 0.19 | 0.28 |
-| BNPP~root~ | 0.33 | 0.26 | 0.23 | 0.36 | 0.35 |
-| BNPP~fine~ ~root~ | 0.15 | 0.11 | 0.16 | 0.11 | 0.19 |
-| R~root~ | 0.22 | 0.21 | 0.30 | 0.01* | 0.28 |
-
-Table: R^2^ values presented for latitude, three climate variables, and the best multivariate model for each carbon flux. Values marked * are not significant.
-
-R^2^ values are generally highest in the major fluxes, and decrease in subsidiary fluxes.
-Of the major fluxes, R~auto~ and GPP are the most strongly explained by latitude and climate.
-Mean annual temperature explains 77% of variation in R~auto~, while temperature seasonality explains 72% of variation in GPP.
-The proportion of variation explained by climate and latitude decreases in NPP and ANPP.
-The climatic variables with the strongest explanatory power explain around 50% of the variation in these fluxes.
-Of the major fluxes, BNPP~root~ is the least well explained by climate and latitude, with climate explaining at most 36% of variation.
-
-With the exception of ANPP~foliage~, the proportion of variation explained by climate and latitude in subsidiary fluxes is much lower.
-Climate explains at most 23% of variation in ANPP~woody~ ~stem~, 16% in BNPP~fine~ ~root~, and 30% in R~root~.
-In contrast, climate strongly explains variation in ANPP~foliage~, with mean annual temperature explaining 63% of variation.
-
-This pattern is also seen in the R^2^ values for multivariate models.
+*Within the growing season, how does productivity vary with climate?* Within growing season months, we found that climate had a much weaker effect on productivity.
+For each of temperature, precipitation, PET, and solar radiation, we found a small effect of climate for certain carbon fluxes.
+There was a small increase in productivity with temperature and precipitation for ANPP (with temperature R^2^ = 0.10, p<0.001; with precipitation R^2^ = 0.04, p<0.05) and ANPP~foliage~ (with temperature R^2^ = 0.16, p<0.01; with precipitation R^2^ = 0.09, p<0.05).
+Productivity increased with solar radiation for GPP (R^2^ = 0.21, p<0.001), NPP (R^2^ = 0.21, p<0.001), BNPP (R^2^ = 0.16, p<0.001) and BNPP~fine~ ~root~ (R^2^ = 0.12, p<0.01), and with PET for GPP (R^2^ = 0.15, p<0.01), NPP (R^2^ = 0.18, p<0.01), BNPP (R^2^ = 0.23, p<0.0001), BNPP~fine~ ~root~ (R^2^ = 0.11, p<0.05), and ANPP~woody~ ~stem~ (R^2^ = 0.06, p<0.05).
 
@@
 
-In this analysis we use a comprehensive global database, containing an unprecedented amount of data and representing all global forest ecosystems and all significant forest carbon fluxes, to comprehensively explore the relationships between climate and productivity on a global scale.
-Many of our findings support and clarify previously published work; however we use a much larger and more complete database than has previously been available, allowing for higher quality control and standardisation of data.
+In this analysis we used a comprehensive global database, containing an unprecedented amount of data and representing all global forest ecosystems and all significant forest carbon fluxes, to comprehensively explore the relationships between climate and productivity on a global scale.
+Many of our findings support and clarify previously published work; however we used a much larger and more complete database than has previously been available, allowing for higher quality control and standardisation of data.
 This enabled us to control for factors not previously controlled for - such as stand age, methodology, flux components, and disturbance regime - to gain a stronger understanding of latitudinal and climatic effects on productivity.

### 2019-09-09 manuscript/manuscript_removed_text.Rmd
Update manuscript_removed_text.Rmd

@@
 BNPP:NPP decreased with latitude, though this relationship was non-significant.
+
+| Database | Variables downloaded | Citation |
+|-----------------------|--------------------------------------------|---------------------|
+| WorldClim | Mean annual temperature; temperature seasonality; annual temperature range; mean annual precipitation | [@hijmans_very_2005] |
+| WorldClim2 | Solar radiation | [@fick_worldclim_2017] |
+| Climate Research Unit (CRU) time-series dataset v 4.03 | Cloud cover; annual frost days; annual wet days; potential evapotranspiration | [@harris_updated_2014] |
+| Global Aridity Index and Potential Evapotranspiration Climate Database | Aridity; potential evapotranspiration | [@trabucco_global_2018] |
+| TerraClimate | Vapour pressure deficit | [@abatzoglou_terraclimate_2018] |
+
+Table: Sources of climate data
+
+| Variable | Latitude | Mean annual temperature | Temperature seasonality | Potential evapotranspiration | MAT*MAP |
+|----------|----------|----------|----------|--------------|----------|
+| R~auto~ | 0.65 | 0.77 | 0.62 | 0.63 | NA |
+| GPP | 0.64 | 0.63 | 0.72 | 0.36 | 0.64 |
+| NPP | 0.50 | 0.52 | 0.52 | 0.33 | 0.56 |
+| ANPP | 0.45 | 0.46 | 0.42 | 0.28 | 0.48 |
+| ANPP~foliage~ | 0.54 | 0.63 | 0.54 | 0.35 | 0.64 |
+| ANPP~woody~ ~stem~ | 0.17 | 0.23 | 0.12 | 0.19 | 0.28 |
+| BNPP~root~ | 0.33 | 0.26 | 0.23 | 0.36 | 0.35 |
+| BNPP~fine~ ~root~ | 0.15 | 0.11 | 0.16 | 0.11 | 0.19 |
+| R~root~ | 0.22 | 0.21 | 0.30 | 0.01* | 0.28 |
+
+Table: R^2^ values presented for latitude, three climate variables, and the best multivariate model for each carbon flux. Values marked * are not significant.

### 2019-09-09 manuscript/manuscript draft.Rmd
update manuscript draft

@@
 In contrast, Larjavaara and Muller-Landau [-@larjavaara_temperature_2012], find that increases in GPP saturate at approximately 25$^\circ$C MAT, while Schuur [-@schuur_productivity_2003] shows that NPP increases linearly with temperature.
-
 The influence of these climate variables on productivity is further complicated by the possibility of interactive effects occurring between them.
 Taylor et al.
-[-@taylor_temperature_2017] showed that increased MAP had a negative effect at low MAT, but a positive effect at high levels of MAT, and vice versa.
-Similarly, Luyssaert et al.
-[-@luyssaert_co_2007] showed that ecosystems limited by low MAP do not benefit from increasing MAT, and vice versa, suggesting that both temperature and water availability are important in explaining productivity.
+[-@taylor_temperature_2017] showed that increased MAP had a negative effect at low MAT, but a positive effect at high levels of MAT, and vice versa, suggesting that both temperature and water availability are important in explaining productivity.
 
@@
 There is evidence that productivity also responds to variables such as cloud cover [@taylor_temperature_2017], solar radiation [@fyllas_solar_2017], and potential evapotranspiration [@kerkhoff_plant_2005] in potentially significant ways.
-
 Furthermore, MAT and MAP are very coarse measures of climate, and so fail to capture much variation in climate on an intra-annual scale, including the effects of factors such as growing season length, number of frost-free days, temperature seasonality, and dry season length.

### 2019-09-09 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
 Associated data, such as stand age, measurement methodologies, and disturbance history, are also included.
-The database was significantly expanded since the publication of [@anderson-teixeira_forc:_2018] through integration with the Global Database of Soil Respiration (SRDB; **Bond-Lamberty, DOI: 10.5194/bg-7-1915-2010**).
+The database was significantly expanded since the publication of [@anderson-teixeira_forc:_2018] through integration with the Global Database of Soil Respiration [@bond-lamberty_global_2010].
 Additional targeted literature searches were conducted to identify any further available data on primary productivity, with particular focus on mature forests in temperate and boreal regions.

### 2019-09-10 manuscript/Discussion_KAT.Rmd
Create Discussion_KAT.Rmd
broke out discussion section for editing

--- 2019-09-09 manuscript/manuscript draft.Rmd
@@
----
-title: "Manuscript draft"
-author: "Rebecca Banbury Morgan, Valentine Herrmann, Norbert Kunert, Ben Bond-Lamberty, Helene Muller-Landau, [MORE??-CHECK], Kristina J. Anderson-Teixeira"
-output:
-  pdf_document:
-    citation_package: natbib
-    fig_caption: yes
-bibliography: [library.bib,  packages.bib]
-biblio-style: apalike
-csl: apa.csl
-header-includes:  \usepackage{float}
-                  \usepackage{caption}
-                  \captionsetup[table]{font=footnotesize}
-                  \captionsetup[figure]{font=footnotesize}
-
----
-
-```{r, include=FALSE}
-options(tinytex.verbose = TRUE)
-```
-
-```{r include=FALSE}
-knitr::write_bib(c(
-  .packages(), 'knitr', 'rmarkdown', 'lme4', 'MuMIn', 'plyr', 'raster', 'ncdf4', 'Hmisc', 'plyr', 'merTools', 'visreg', 'r2glmm', 'nlme', 'viridis', 'ggplot2', 'ggpubr', 'gridExtra', 'AICcmodavg'
-), 'packages.bib')
-```
-
-```{r results="asis", include=FALSE}
-cat("
-<style>
-caption {
-      font-size: 0.75em;
-    }
-</style>
-")
-```
-
-###Abstract
-
-Globally, forests are a significant carbon sink, showing high levels of productivity and representing an important component of the carbon cycle.
-Current knowledge indicates that forest productivity decreases with latitude; however, there is evidence that this relationship varies across different fluxes.
-Furthermore, the climate variables that drive this relationship are poorly understood.
-Here, we use a comprehensive global database of forest carbon fluxes to show that, while all major carbon fluxes decrease with latitude, allocation to woody productivity increases with latitude, and allocation to root productivity declines.
-The best predictors of global patterns of productivity are temperature variables, including mean annual temperature, annual temperature range, and temperature seasonality, with the exception of woody productivity, which is most strongly influenced by potential evapotranspiration and vapour pressure deficit.
-Climate variables explain a large proportion of the variation in major carbon fluxes, but are a less significant predictor of subsidiary components.
-Our results illustrate the strong influence of climate on primary productivity, and particularly the importance of temperature in determining forest productivity.
-However, they also indicate that effects of climate are complex: fluxes show non-linear responses to climate variables, and many factors other than climate interact to determine allocation to subsidiary components.
-Knowledge of the relative influence of climate variables on forest carbon fluxes is an important component of understanding the responses of primary productivity to climate change.
-
-### Introduction
-
-Globally, forests play an important role in the carbon cycle and are a major component of global carbon dioxide budgets [@luyssaert_old-growth_2008].
-They show higher levels of productivity than non-forest terrestrial ecosystems [@del_grosso_global_2008], and as a result achieve significant carbon sequestration and storage.
-Estimating the total role of forests in the carbon cycle is challenging, but studies indicate that old growth forests alone sequester up to 1.4 GtCyr^-1^ [@malhi_carbon_1999], while the total sequestration of carbon by established forests globally could be up to 2.4 GtCyr^-1^, with the largest sinks being in old-growth tropical forests [@pan_large_2011].
-As atmospheric carbon dioxide levels continue to rise, with consequences for global climate, there is increasing recognition that proper protection and management of forest resources will have an important role to play in mitigating climate change.
-Understanding the patterns of forest productivity on a global scale, and the drivers behind them, is therefore a priority in forest research.
-[**KAT-work on this; consider shifting top focus from global C cycle to centrality of forest productivity to ecology & climate science.** ] Given the importance of understanding broad-scale patterns in forest productivity, there have been many meta-analyses on the theme.
-However, our ability to draw general macroscopic conclusions regarding global variation in multiple productivity variables with respect to climate has been limited in that these analyses often mix forests that vary in stand age, disturbance history, and/or management status; do not always sufficiently parse related variables (*e.g.*, combining net primary productivity records with and without belowground components); and typically consider only one or a few variables at a time.
-The recent development of a global forest carbon database synthesizing multiple variables and including records of stand history (ForC; [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc:_2018]) opens up the possibility for a standardized analysis of global scale variation in multiple components of forest productivity and the principle climatic drivers of these patterns.
-
-On a global scale, the productivity of forests varies with latitude, showing a general trend of decreasing productivity with latitude [@beer_terrestrial_2010; @jung_global_2011].
-Studies agree that productivity is lowest in the boreal regions, and increases into the temperate regions [@luyssaert_co_2007; @huston_global_2009; @beer_terrestrial_2010; @jung_global_2011].
-However, evidence is inconclusive on whether productivity continues to increase into the tropics, or whether it plateaus in temperate regions.
-Evidence for this is further complicated by the fact that different studies use different measures of productivity to explore these relationships.
-For example, modelling of global terrestrial ecosystem gross primary productivity (GPP) through upscaling and calibration of eddy flux measurements indicates that GPP peaks in tropical forests [@beer_terrestrial_2010; @jung_global_2011].
-This is corroborated by analysis of site-level GPP measurements, which appear to reach their highest levels in tropical forests [@luyssaert_co_2007].
-In contrast, there is evidence that the highest values of net primary productivity (NPP) may be found in temperate forests [@luyssaert_co_2007; @huston_global_2009], although other studies find NPP is highest in the tropics, showing a decrease with latitude [@simova_enigma_2017].
-Other studies have chosen to focus exclusively on above-ground net primary productivity (ANPP), finding evidence of a weak negative relationship between ANPP and latitude [@huston_global_2009; @gillman_latitude_2015].
-An understanding of the global patterns of productivity is important; however elucidating the drivers of these patterns will be even more valuable in expanding our understanding of global carbon cycling.
-Primary productivity can be influenced by many factors, which often act across a range of scales, and may show interactive effects with each other.
-On a local scale, stand age [@litton_carbon_2007; @gillman_latitude_2015], management [@simova_enigma_2017]; nutrient availability [@aragao_above-_2009]; and altitude [@girardin_net_2010; @malhi_variation_2017] all impact forest productivity.
-On a global scale, we expect that productivity is most strongly influenced by broad climatic gradients.
-
-We know from previous research that climate is a significant driver of productivity across broad spatial scales [@cleveland_relationships_2011].
-The majority of studies have focused on exploring the relationships between productivity and mean annual temperature (MAT) and mean annual precipitation (MAP), as the most commonly reported site-level climate variables.
-These variables have the advantage that they describe broad trends in temperature and water availability, and therefore capture a lot of global-scale variation in climate.
-There is strong evidence that both MAT and MAP show significant positive relationships with productivity [@chu_does_2016].
-However, as with latitude, the shape of those relationships is not always clear, and, again, is complicated by the use of different measures of productivity across studies.
-There is support across multiple studies for the hypothesis that various measures of primary productivity saturate at high levels of MAP, though the saturation points identified vary from 1500mm [@luyssaert_co_2007] up to 2445mm MAP [@schuur_productivity_2003].
-Studies of the influence of MAT on productivity are less conclusive.
-Luyssaert et al.
-[-@luyssaert_co_2007] examined GPP and NPP and found that, while GPP increases linearly with MAT, NPP saturates at around 10$^\circ$C MAT.
-In contrast, Larjavaara and Muller-Landau [-@larjavaara_temperature_2012], find that increases in GPP saturate at approximately 25$^\circ$C MAT, while Schuur [-@schuur_productivity_2003] shows that NPP increases linearly with temperature.
-The influence of these climate variables on productivity is further complicated by the possibility of interactive effects occurring between them.
-Taylor et al.
-[-@taylor_temperature_2017] showed that increased MAP had a negative effect at low MAT, but a positive effect at high levels of MAT, and vice versa, suggesting that both temperature and water availability are important in explaining productivity.
-
-While MAT and MAP are important climate variables, they do not capture all aspects of climate that may influence productivity, and may be insufficient to capture the breadth of climatic effects on productivity [@cleveland_relationships_2011].
-There is evidence that productivity also responds to variables such as cloud cover [@taylor_temperature_2017], solar radiation [@fyllas_solar_2017], and potential evapotranspiration [@kerkhoff_plant_2005] in potentially significant ways.
-Furthermore, MAT and MAP are very coarse measures of climate, and so fail to capture much variation in climate on an intra-annual scale, including the effects of factors such as growing season length, number of frost-free days, temperature seasonality, and dry season length.
-Some studies have suggested that the apparently strong relationship between MAT and productivity is actually a factor of the correlation between MAT and growing season length [@kerkhoff_plant_2005; @malhi_productivity_2012; @michaletz_convergence_2014; @michaletz_drivers_2018].
-Kerkhoff et al.
-[-@kerkhoff_plant_2005] and Michaletz et al.
-[-@michaletz_convergence_2014] find that, within the growing season, there is no significant relationship between productivity and MAT, indicating that the effect of temperature is due to increased length of growing season, rather than an inherent influence of temperature on productivity.
-
-In order to approach these broad and complex issues, we simplify the major gaps in our knowledge to five key questions and corresponding specific predictions (Table 1):  
-1.	How do carbon fluxes vary with latitude? 
-2.	How do carbon fluxes relate to MAT and MAP?  
-3.	How do carbon fluxes relate to other annual climate variables?  
-4.	What is the role of seasonality in explaining variation in carbon fluxes?  
-5.	Per month of the growing season, how does productivity vary with climate?
-
-To resolve these questions, we require data that, firstly, control for variables influencing productivity, including stand age, disturbance regime, elevation, and methodology; and secondly includes all major carbon fluxes.
-Here we use a comprehensive global database of forest carbon fluxes to explore the above questions for nine carbon fluxes, allowing for an in-depth exploration of the effect of climate on global productivity.
-
-### Materials and Methods
-
-Analyses were conducted on data contained in the open-access ForC database [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc:_2018].
-This database contains records of field-based measurements of forest carbon stocks and annual fluxes, compiled from original publications and existing data compilations and databases.
-Associated data, such as stand age, measurement methodologies, and disturbance history, are also included.
-The database was significantly expanded since the publication of [@anderson-teixeira_forc:_2018] through integration with the Global Database of Soil Respiration [@bond-lamberty_global_2010].
-Additional targeted literature searches were conducted to identify any further available data on primary productivity, with particular focus on mature forests in temperate and boreal regions.
-ForC currently contains **UPDATE STATS** 29768 records from 5227 plots, representing 20 distinct ecozones across all forested biogeographic and climate zones.
-We used ForC v.XX, archived on Zenodo with DOI XXX.
-
-```{r echo=FALSE, out.width='100%', fig.cap = "Map showing all data used in the analysis, coded by variable", fig.pos='H'}
-  knitr::include_graphics("distribution_all_variables.png")
-```
-*Data selection.* Over 50 variables of forest carbon stocks and annual fluxes are represented in the ForC database; this analysis focussed on measures of primary productivity.
-Table 1 contains details of the variables selected for analysis.
-
-| Variable | Definition | Components included | Methodologies used | Number of records | Number of plots |
-|--------------|-------------------|-------------|------------------------|-------|-------|
-| GPP | Annual gross primary production; annual uptake of carbon dioxide by an ecosystem | NA | Flux partitioning of eddy covariance | 257 | 80 |
-| NPP | Annual net primary production; the component of GPP that is stored in plant tissue; GPP minus ecosystem respiration | Foliage, branch, stem, coarse root, fine root and optionally understory | Direct measurement of annual increments of components | 93 | 78 |
-| ANPP | Aboveground net primary production | Foliage, stem, and optionally branch | Direct measurement of annual increments of components | 258 | 166 |
-| ANPP~foliage~ | Net primary production of foliage | Foliage | Direct measurement of litterfall, correcting for changes in leaf biomass when measured | 98 | 94 |
-| ANPP~woody~ ~stem~ | Net primary production of woody stems | Woody stems | Direct measurement of stem growth increment | 266 | 252 |
-| BNPP~root~ | Belowground net primary production | Coarse and fine roots | Direct measurement of one or more of: fine root turnover, soil cores, root ingrowth cores, minirhizotrons; indirect estimates of coarse roots using allometries based on aboveground stem increment measures | 107 | 87 |
-| BNPP~fine~ ~root~ | Net primary production of fine roots | Fine roots | Direct measurement of one or more of: minirhizotrons, fine root turnover, soil cores, root ingrowth cores | 90 | 77 |
-| R~auto~ | Annual autotrophic respiration, including above- and belowground components | Foliage, stem, and root | Chamber measurements of component gas exchange | 22 | 22 |
-| R~root~ | Annual root respiration | Coarse and fine roots | Measurement of root gas exchange | 65 | 43 |
-
-Table: Definitions and sample sizes of variables used in analysis
-
-A subset of the ForC database was generated for the purposes of this analysis, in order to control for data quality and remove biasing factors.
-Since management can alter observed patterns of primary productivity [@simova_enigma_2017], sites were excluded from analysis if they were managed, defined as plots that were planted, managed as plantations, irrigated, fertilised or including the term "managed" in their site description.
-Sites that had experienced significant disturbance were also excluded.
-Disturbances that justified site exclusion were major cutting or harvesting, and/or burning, flooding, drought and storm events with site mortality >10% of trees.
-Grazed sites were retained.
-There is evidence that stand age influences patterns of primary productivity and carbon allocation in forest ecosystems, and can confound relationships between latitude and primary productivity [@de_lucia_forest_2007; @gillman_latitude_2015].
-To reduce any biasing effects of stand age, stands under 100 years of age were excluded from analysis.
-Sites for which stand age was unknown were excluded from analysis.
-
-*Methodological consistency.* The data in ForC is derived from a range of studies, often employing different methods.
-For this reason, criteria were introduced to standardise for differences in methodology.
-Where data was based on forest plot census measurements, studies which used a minimum diameter at breast height (DBH) measure of >10cm were excluded from analysis.
-It would be preferable to standardise by minimum area sampled; however x% of plots in the database are 1 ha or under in size; excluding these plots would place significant constraints on sample size.
-
-As discussed above, estimates of NPP, ANPP, and BNPP are generated through summing measurements of their component parts.
-Since the components included in productivity estimates vary between studies, estimates of productivity were classified within the ForC database according to their components, and then filtered for analysis.
-Estimates of NPP were selected if they included foliage, branch, stem, coarse root, and fine root.
-Measures of NPP which included additional components, including understorey, volatile organic compounds (VOCs), exudates, estimates of NPP lost to herbivory, and the NPP of reproductive structures, were excluded.
-Estimates of ANPP were selected if they included foliage, stem growth and optionally branch turnover.
-Any measures of primary productivity where components were unknown were excluded from analysis.
-
-*Climate datasets.* Where site-level data on mean annual temperature, mean annual precipitation, and latitude were available in the primary literature, this data was compiled and entered directly into the ForC database.
-Based on the geographic co-ordinates for each site, data on a further 11 climate variables was extracted from five open-access climate datasets: WorldClim [@hijmans_very_2005], WorldClim2 [@fick_worldclim_2017], the Climate Research Unit (CRU) time-series dataset v.
-4.03 [@harris_updated_2014], the Global Aridity Index and Potential Evapotranspiration Climate Database [@trabucco_global_2018], and TerraClimate [@abatzoglou_terraclimate_2018] (see Supplementary Information S1 for details of climate variables).
-Where site-level data was missing for mean annual temperature and/or mean annual precipitation, data was extracted from the WorldClim dataset.
-
-Additionally, two climate variables were derived from the above datasets: maximum vapour pressure deficit, defined as the vapour pressure deficit of the month with the largest deficit; and water stress months, defined as the number of months annually where precipitation was lower than potential evapotranspiration.
-
-*Length of growing season.* Growing season months were defined as months with mean minimum temperature > 0.5.
-Growing season months were initially calculated following methods used by Kerkhoff et al.
-(2005), which additionally required that growing season months had a moisture index, defined as (MAT - PET)/PET, > -0.95.
-Michaletz et al.
-(2014) included an equivalent requirement in their calculation of growing season length.
-However, we found that including this requirement had no effect on the estimates of growing season length, and so chose to exclude it.
-
-Monthly data for PET, precipitation, and temperature was downloaded from the Climate Research Unit (CRU) time-series dataset v 4.03 [@harris_updated_2014], and for solar radiation from WorldClim2 [@fick_worldclim_2017], and used to calculate mean monthly PET, precipitation, temperature and solar radiation during the growing season.
-Total growing season precipitation and solar radiation were also calculated.
-
-*Model specification.* The effects of climate and latitude on primary productivity were analysed using mixed effects models using the package 'lme4' [@bates_fitting_2015] in R v.3.5.1 [@r_core_team_r:_2018].
-The effect of each extracted climate variable on each measure of primary productivity was modelled by specifying the climate variable as a fixed effect.
-For each climate variable, three models were specified: a null model; a model with the climate variable as a linear term; and a model with the climate variable as a polynomial term.
-AIC values were calculated for the models and used to select the best model.
-If the best model included a polynomial term, the shape of the polynomial relationship was considered.
-If the shape of the relationship made biological sense, and was a significant improvement on the linear relationship (deltaAIC >2), we accepted the polynomial as the best model.
-If not, we ran the linear model as the final model.
-R^2^ values were calculated for the best model.
-All R^2^ values presented here are marginal R^2^ values, and refer to the proportion of variation explained by only the fixed effects, unless otherwise specified.
-In addition, slope coefficients were calculated for the linear models.
-
-Because the magnitude of fluxes varies significantly, in order to facilitate comparisons between regression models for each flux, data for each flux was scaled, to give the data a mean of 0 and standard deviation of 1.
-As each data set was scaled separately, this does not allow for statistical comparisons of slope values, but does assist in visualising the data.
-
-To test for a potential influence of altitude, models were also run with site altitude included as a second fixed effect.
-These models were compared against models with no altitude term, and AIC values calculated to identify whether inclusion of altitude as a term improved the models.
-Including altitude had a very small effect on most models, with the difference in AIC values between models including and excluding altitude often being <2, suggesting the models are very similar in their explanatory power.
-As a result, it was decided to present results only from models do not include altitude as a fixed term.
-
-Within the ForC database, sites within 25km^2^ of each other are clustered into geographic areas.
-To account for correlations in measurements between tightly clustered sites, a random effect was specified as plot nested within geographic area.
-Data from the temperate regions was heavily skewed towards studies from the old-growth forests of the Pacific Northwest.
-These forests have very high productivity, and so to ensure that results were not unduly influenced by geographic sampling bias, we tried a version of the model where data were weighted according to forested land area within each Koeppen climate zone.
-Results were similar between the weighted and unweighted model, so, to avoid problems of over-fitting, the weighted model was dropped, and results from this are not presented here.
-
-Models were run for total annual productivity against annual climate variables, and for monthly growing season productivity, defined as total productivity/length of growing season, against growing season climate variables.
-For analyses on data within the growing season, only linear models were specified.
-
-To investigate the potential interactive effects of climate variables on carbon fluxes, multivariate models were also specified.
-To ensure that models were biologically meaningful, the terms included in the models tested built on results from the univariate models.
-Modelling of individual climate variables identified that the best predictors of carbon fluxes were variables related to temperature.
-We therefore decided to include mean annual temperature as a term in all multivariate models.
-We first modelled the interaction effect between mean annual temperature and mean annual precipitation, in order to capture climate variation along the axes of temperature and water availability.
-Models were tested for a significant interactive effect and a significant additive effect.
-We then explored whether any other climate variable, in combination with mean annual temperature, could significantly improve on the combination of mean annual temperature and mean annual precipitation.
-In specifying the range of models to test, climate variables which were strongly correlated with temperature were dropped, in order to capture the greatest range of variation in climate.
-For each possible pairing of climate variables, two models were specified: a model with the two climate variables showing an additive effect; and a model with the two climate variables showing an interactive effect.
-As before, plot nested within geographic area was included as a random effect.
-Altitude was not considered.
-AIC values were calculated for the models, and used to compare models.
-Models were considered to be significantly better than the baseline MAT*MAP model if:
-  i) the AIC value of the model was smaller than the AIC value of the baseline model by >2  
-  ii) the r-squared value of the model was larger than the r-squared value of the baseline model by >5
-
-*Validating models of component fluxes.* Comparison of component fluxes is based on the assumption that components sum accurately to estimates of larger fluxes.
-To test this, components of larger fluxes were regressed against latitude, and the models used to generate a series of point estimates along lines of best fit for each component.
-The point estimates for smaller component fluxes were summed to generate new "stacked" estimates of larger fluxes, which were then compared against actual measurements of the larger flux.
-Confidence intervals for the larger flux were calculated using the 'bootMer' function from the lme4 package [@bates_fitting_2015].
-Stacked plots were generated for:
-1. GPP = NPP + R~auto~  
-2. NPP = ANPP + BNPP  
-3. ANPP = ANPP~foliage~ + ANPP~woody~ ~stem~  
-4. Total belowground carbon flux = BNPP + R~root~
-
-*Allocation to carbon fluxes along latitudinal gradients.* Variation in allocation to component carbon fluxes along latitudinal gradients was explored for a range of pairings: firstly, GPP:NPP, ANPP:BNPP, and ANPP~foliage~:ANPP~woody~ ~stem~; and secondly, the ratio of NPP to each of ANPP, BNPP, ANPP~foliage~, and ANPP~woody~ ~stem~.
-For each set of paired fluxes, measurements taken at the same site and plot, and in the same year, were paired together, and the ratio of each pair of measurements calculated.
-The ratios were regressed against latitude and climate variables, using the linear model specified above.
-Cook's distance analyses were carried out for each of the models, and indicated that data from a few high-elevation sites were having a disproportionate influence on the regressions.
-To account for this, models were re-run using only data from sites $\leq$ 1000m.
-
-###Results
-
-In total, we analyzed ## records from ## C flux variables taken from forests that had experienced no major anthropogenic disturbances within the past 100 years from # distinct geographic areas (Fig.
-1, Table 2).
-[**Add a bit about geographic representation**]
-
-```{r echo=FALSE, out.width='100%', fig.cap = "Graphs of primary productivity $(MgC~ha^-1~yr^-1)$ regressed against latitude.
-Lines of best fit are plotted according to the best model selected during analysis.
-All regressions are significant $(p<0.05)$.
-Plots 1 - 3 show two component fluxes; a larger flux, defined as the combination of the two component fluxes; and a modelled estimate of the sum of the two component fluxes.
-95\\% confidence intervals are plotted for the larger flux.
-Plot 4 shows three belowground fluxes, and a modelled estimate of the total belowground carbon flux", fig.pos='H'}
-
-  knitr::include_graphics("combined_stacked.png")
-```
-
-*How does productivity vary with latitude?*
-
-All major carbon fluxes increased linearly with decreasing latitude (fig. 3).
-
-```{r echo=FALSE, out.width='100%', fig.cap = "Graphs to show primary productivity $(MgC~ha^-1~yr^-1)$ regressed against latitude.
-Lines of best fit are plotted according to the best model selected during analysis.
-All regressions are significant $(p<0.05)$.", fig.pos='H'}
-  knitr::include_graphics("effect_of_lat_transparent.png")
-```
-
-Latitude was a strong predictor for many of the carbon fluxes, explaining 64% of variation in GPP (n = 254, p<0.0001), 50% in NPP (n = 114, p<0.0001) and 45% in ANPP (n = 259, p<0.0001).
-For all fluxes, their relationship with latitude was best predicted by the linear model.
-
-*Relationships and differences among fluxes.* In general, smaller component fluxes summed approximately to  larger fluxes across the latitudinal gradient (fig.
-2).
-That is, modelled estimates of GPP, generated from the sum of NPP and R auto; NPP, generated from the sum of ANPP and BNPP~root~; and ANPP, generated from the sum of ANPP~foliage~ and ANPP~woody~ ~stem~, fell completely within the confidence intervals of the regressions of field estimates of GPP, NPP and ANPP respectively.
-
-We found no evidence that allocation between fluxes varied substantially with latitude or climate.
-There were no significant results from regressing ratios of carbon fluxes against latitude, or against any of the climate variables.
-
-R^2^ values were generally highest in the major fluxes, and decreased in subsidiary fluxes (Supporting Information S2).
-Of the major fluxes, R~auto~ and GPP were the most strongly explained by latitude and climate, with climate explaining at most 71% of variation in GPP, and 65% in R~auto~.
-The proportion of variation explained by climate and latitude decreased in NPP and ANPP, with climate explaining at most 51% of variation in NPP and 44% in ANPP.
-Of the major fluxes, BNPP~root~ was the least well explained by climate and latitude, with climate explaining at most 36% of variation.
-
-With the exception of ANPP~foliage~, the proportion of variation explained by climate and latitude in subsidiary fluxes was much lower.
-Climate explained at most 24% of variation in ANPP~woody~ ~stem~, 19% in BNPP~fine~ ~root~, and 27% in R~root~.
-In contrast, climate strongly explained variation in ANPP~foliage~, with mean annual temperature explaining 58% of variation.
-This pattern was also seen in the R^2^ values for multivariate models.
-
-*How does productivity relate to MAT and MAP?* We focus first on considering the relationship between productivity and MAT and MAP.
-MAT and MAP are the most commonly reported site-level climate variables, and much previous research into the effect of climate on forest productivity has focused on these as key climate variables.
-MAT was a significant (p<0.05) and strong predictor of productivity for all carbon fluxes tested, with all fluxes showing a linear increase with temperature (fig.
-5).
-We found no support for a saturation point of productivity with temperature.
-
-MAP was found to be a significant (p<0.05) but poor predictor of productivity, explaining, with the exception of R~auto~, at most 37% of variation in carbon flux.
-For the majority of fluxes productivity was best predicted by a polynomial model.
-Productivity increased with precipitation, up until a saturation point at between 3000 and 4000mm annual precipitation, above which productivity started to decrease (fig.
-5).
-The notable exception to this was GPP: the model indicated that GPP continued to increase with precipitation up to measures of at least 5000mm annually (p<0.0001, R^2^ = 0.33.
-Data above this point was not available, but the model trend indicated that the saturation point for this model would be around 5000mm MAP.
-
-There was a significant interactive effect between MAT and MAP for GPP, BNPP~root~, BNPP~fine~ ~root~, ANPP, ANPP~woody~ ~stem~, and R~root~ (fig.4).
-There was a significant additive effect for R~auto~.
-NPP and ANPP~foliage~ showed no significant interactive or additive effect: including MAP as a second variable did not improve on the model including only MAT.
-
-For the variables which showed a significant interactive or additive effect between MAT and MAP, no other climate variable, in combination with MAT, significantly improved on that model.
-For NPP, there was a significant interactive effect between MAT and water stress months, with this model explaining nearly 5% more variation in NPP than MAT alone.
-However, for ANPP~foliage~, no multivariate model improved on the univariate model including only MAT.
-
-```{r echo=FALSE, out.width='100%', fig.cap = "Plots of primary productivity $(MgC~ha^-1~yr^-1)$ regressed against mean annual precipitation.
-Points are grouped into bins of 0 - 1000, 1001 - 2000, 2001 - 3000, and >3000mm mean annual precipitation, and lines of best fit plotted for mean annual precipitation values of 500, 1500, 2500, and 3500mm.
-All regressions are significant $(p<0.05)$.", fig.pos='H'}
-  knitr::include_graphics("mat_map_interaction.png")
-```
-
-*How does productivity relate to other climate variables?* Our results indicated that productivity was most strongly explained by temperature at the global scale, with temperature-related climate variables coming out as strong predictors of productivity.
-In addition to MAT, temperature seasonality, annual temperature range, and annual frost days were consistently identified as good predictors of productivity across fluxes.
-
-We found a significant relationship between productivity and potential evapotranspiration for all fluxes.
-ANPP~foliage~, BNPP~fine~ ~root~ and R~root~ increased linearly with PET; however, all other fluxes showed a polynomial relationship with PET (fig.
-5).
-We found strong evidence for a saturation point or peak with PET: productivity tended to increase at values below 1000mm, before saturating between 1200 and 1700mm.
-There was evidence that productivity begins to decrease at values above 1800mm PET.
-
-Vapour pressure deficit was a significant predictor of productivity for all fluxes.
-BNPP~fine~ ~root~ showed a linear relationship with vapour pressure deficit (R^2^ = 0.07, p<0.05), but all other fluxes showed a polynomial relationship (fig.
-5).
-Productivity initially increased with vapour pressure deficit, before saturating at around 0.8 kPa.
-At values above 0.8 kPa, productivity began to decrease.
-
-All fluxes, with the exception of R~root~, showed a positive linear relationship with solar radiation.
-Solar radiation explained a low proportion of variability in productivity for all fluxes, explaining less than 20% of the variation in each flux, with the exception of R~auto~ (R^2^ = 0.26, p<0.05).
-
-Of the climate variables tested, annual wet days, aridity, cloud cover, mean diurnal temperature range, precipitation seasonality, maximum vapour pressure deficit and water stress months were poor or non-significant explainers of variation in productivity, explaining less than 20% of the variation in each of the carbon fluxes.
-
-```{r echo=FALSE, out.width='100%', fig.cap = "Plots of primary productivity $(MgC~ha^-1~yr^-1)$ regressed against (a) mean annual temperature; (b) mean annual precipitation; (c) potential evapotranspiration, (d) vapour pressure deficit; (e) temperature seasonality; (f) length of growing season.
-Lines of best fit are plotted according to the best model selected during analysis.
-All regressions are significant $(p<0.05)$.", fig.pos='H'}
-  knitr::include_graphics("combined_plots.png")
-```
-
-*What is the role of seasonality in explaining productivity?* Temperature seasonality was a significant predictor of productivity.
-We found a polynomial relationship with productivity for GPP, NPP, ANPP, and R~root~.
-Productivity decreased rapidly as seasonality increased, with the rate of decrease slowing as seasonality increased (fig.
-5).
-ANPP~foliage~, ANPP~woody~ ~stem~ and R~auto~ decreased linearly with temperature seasonality.
-Temperature seasonality was strongly correlated with annual temperature range, and, as expected, all fluxes showed almost identical responses to it.
-Productivity was highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower.
-
-In contrast, there was no significant effect of precipitation seasonality on productivity.
-
-We found a significant relationship between length of growing season and productivity, with all fluxes showing a linear increase in productivity with length of growing season (fig.
-5).
-Length of growing season was a strong predictor of productivity, explaining 51% of variation in GPP, 39% of variation in NPP, and 34% of variation in ANPP, but it was not identified as a stronger predictor than MAT for any of the fluxes analysed.
-
-*Within the growing season, how does productivity vary with climate?* Within growing season months, we found that climate had a much weaker effect on productivity.
-For each of temperature, precipitation, PET, and solar radiation, we found a small effect of climate for certain carbon fluxes.
-There was a small increase in productivity with temperature and precipitation for ANPP (with temperature R^2^ = 0.10, p<0.001; with precipitation R^2^ = 0.04, p<0.05) and ANPP~foliage~ (with temperature R^2^ = 0.16, p<0.01; with precipitation R^2^ = 0.09, p<0.05).
-Productivity increased with solar radiation for GPP (R^2^ = 0.21, p<0.001), NPP (R^2^ = 0.21, p<0.001), BNPP (R^2^ = 0.16, p<0.001) and BNPP~fine~ ~root~ (R^2^ = 0.12, p<0.01), and with PET for GPP (R^2^ = 0.15, p<0.01), NPP (R^2^ = 0.18, p<0.01), BNPP (R^2^ = 0.23, p<0.0001), BNPP~fine~ ~root~ (R^2^ = 0.11, p<0.05), and ANPP~woody~ ~stem~ (R^2^ = 0.06, p<0.05).
-
 ###Discussion

### 2019-09-10 scripts/final_model_no_weighting.R
Update final_model_no_weighting.R

@@ -261,11 +261,12 @@ for(response.variables in response.variables.groups){
         mtext(side = 3, line = -which(response.variables %in% response.v), text = legend2, adj = 0.9, col = plasma(10)[col], cex = 0.5)
         
         number_plots <- length(unique(df$site_plot))
+        geog_area <- length(unique(df$geographic.area))
         
         date = Sys.Date()
         altitude = FALSE
         
-        results <- data.frame(date.run = date, response = response.v, fixed = fixed.v, altitude_included = altitude, random = "geographic.area/plot.name", Age.filter = age, best.model = best.model, significant = significant.effect, p.value = significance, sample.size = sample.size, Rsq = Rsq, delta.aic = delta.aic, linear.slope = slope, number.plots = number_plots)
+        results <- data.frame(date.run = date, response = response.v, fixed = fixed.v, altitude_included = altitude, random = "geographic.area/plot.name", Age.filter = age, best.model = best.model, significant = significant.effect, p.value = significance, sample.size = sample.size, Rsq = Rsq, delta.aic = delta.aic, linear.slope = slope, number.plots = number_plots, geographic.areas = geog_area)
         
         all.results <- rbind(all.results, results)
         all.aictab <- rbind(all.aictab, aictab)

### 2019-09-10 manuscript/Discussion_KAT.Rmd
Update Discussion_KAT.Rmd

@@
 
-In this analysis we used a comprehensive global database, containing an unprecedented amount of data and representing all global forest ecosystems and all significant forest carbon fluxes, to comprehensively explore the relationships between climate and productivity on a global scale.
-Many of our findings support and clarify previously published work; however we used a much larger and more complete database than has previously been available, allowing for higher quality control and standardisation of data.
-This enabled us to control for factors not previously controlled for - such as stand age, methodology, flux components, and disturbance regime - to gain a stronger understanding of latitudinal and climatic effects on productivity.
+We used a large global database (ForC), containing an unprecedented amount of data representing all major forest types (Fig.
+1) and all significant forest autotrophic carbon fluxes (Table 2), to comprehensively explore how C cycling in mature forests varies with latitude and climate on a global scale.
+We test a suite of hypotheses, including many previously posed (Table 1), with more data and standardisation for factors that have not always been controlled for (*e.g.*, stand age, flux components measured) to gain a stronger understanding of climatic effects on productivity.
+We show that, across all nine productivity flux variables analyzed, C cycling decreases linearly with latitude (Figs.
+2-3)--a finding that confirms multiple previous studies (**REFS**) but contradicts the idea that productivity of temperate forests rivals that of tropical forests (**REFS**).
+The flux variables increase in proportion to one another, with no differences in allocation detectable at this global scale and with component fluxes summing appropriately to larger fluxes (Fig.
+**2/3**), indicating no major, systematic omissions or overestimations of flux components.
+However, we did detect a tendency for greater variability among subsidiary C fluxes (Fig.
+**2/3**; **discussed below??**).
 

### 2019-09-10 manuscript/Discussion_KAT.Rmd
Update Discussion_KAT.Rmd

@@
 We used a large global database (ForC), containing an unprecedented amount of data representing all major forest types (Fig.
-1) and all significant forest autotrophic carbon fluxes (Table 2), to comprehensively explore how C cycling in mature forests varies with latitude and climate on a global scale.
+1) and the nine most significant forest autotrophic carbon fluxes (Table 2), to comprehensively explore how C cycling in mature forests varies with latitude and climate on a global scale.
 We test a suite of hypotheses, including many previously posed (Table 1), with more data and standardisation for factors that have not always been controlled for (*e.g.*, stand age, flux components measured) to gain a stronger understanding of climatic effects on productivity.
@@
 **2/3**; **discussed below??**).
-
-Our results show that productivity decreases linearly with latitude.
-Climate explains a significant proportion of variation in all carbon fluxes, with temperature variables being the best predictors of productivity on this global scale.
-While other climate variables are significant predictors of productivity, none of them improve on the explanatory power of temperature-related variables in general or MAT specifically.
-Water availability is an important factor in explaining productivity on a global scale: we find a positive influence of precipitation at low MAP, with saturation at higher levels of MAP.
-There is a significant interaction between MAT and MAP.
-We note that climate tends to explain a higher amount of variation in the major fluxes, with the amount of variation explained decreasing in smaller component fluxes.
+Climate explains a significant proportion of variation in all C fluxes analyzed (albiet less for subsidiary fluxes), with temperature variables being the best predictors of productivity on this global scale (Figs.
+4-5).
+While other climate variables are significant predictors of productivity, none of them improve on the explanatory power of temperature-related variables in general or MAT specifically (Fig.
+5).
+Water availability is an important factor in explaining productivity on a global scale: we find a positive influence of precipitation at low MAP, with saturation at higher levels of MAP (Fig.
+5b) and a significant interaction between MAT and MAP for seven of the nine variables analyzed (Fig.
+4).
 

### 2019-09-10 manuscript/Discussion_KAT.Rmd

@@
 We used a large global database (ForC), containing an unprecedented amount of data representing all major forest types (Fig.
-1) and the nine most significant forest autotrophic carbon fluxes (Table 2), to comprehensively explore how C cycling in mature forests varies with latitude and climate on a global scale.
-We test a suite of hypotheses, including many previously posed (Table 1), with more data and standardisation for factors that have not always been controlled for (*e.g.*, stand age, flux components measured) to gain a stronger understanding of climatic effects on productivity.
-We show that, across all nine productivity flux variables analyzed, C cycling decreases linearly with latitude (Figs.
+1) and the nine most significant forest autotrophic carbon flux (FACF) variables (Table 2), to comprehensively explore how C cycling in mature forests varies with latitude and climate on a global scale.
+We test a suite of hypotheses, including many previously posed (Table 1), with more data and standardisation for factors that have not always been controlled for (*e.g.*, stand age, flux components measured) to gain a stronger understanding of climatic effects on FACF.
+We show that, across all nine FACF variables analyzed, C cycling decreases linearly with latitude (Figs.
 2-3)--a finding that confirms multiple previous studies (**REFS**) but contradicts the idea that productivity of temperate forests rivals that of tropical forests (**REFS**).
-The flux variables increase in proportion to one another, with no differences in allocation detectable at this global scale and with component fluxes summing appropriately to larger fluxes (Fig.
+The FACF variables increase in proportion to one another, with no differences in allocation detectable at this global scale and with component fluxes summing appropriately to larger fluxes (Fig.
 **2/3**), indicating no major, systematic omissions or overestimations of flux components.
-However, we did detect a tendency for greater variability among subsidiary C fluxes (Fig.
-**2/3**; **discussed below??**).
-Climate explains a significant proportion of variation in all C fluxes analyzed (albiet less for subsidiary fluxes), with temperature variables being the best predictors of productivity on this global scale (Figs.
+However, we did detect a tendency for greater variability among subsidiary C fluxes (e.g., *$ANPP_woody$*, *$BNPP_{fine root}$*; Fig.
+**2/3**; **some SI table?**; **discussed below??**).
+Climate explains a significant proportion of variation in all C fluxes analyzed (albiet less for subsidiary fluxes), with temperature variables being the best predictors of FACF on this global scale (Figs.
 4-5).
-While other climate variables are significant predictors of productivity, none of them improve on the explanatory power of temperature-related variables in general or MAT specifically (Fig.
+While other climate variables are significant predictors of FACF, none of them improve on the explanatory power of temperature-related variables in general or MAT specifically (Fig.
 5).
-Water availability is an important factor in explaining productivity on a global scale: we find a positive influence of precipitation at low MAP, with saturation at higher levels of MAP (Fig.
+Water availability is an important factor in explaining FACF on a global scale: we find a positive influence of precipitation at low MAP, with saturation at higher levels of MAP (Fig.
 5b) and a significant interaction between MAT and MAP for seven of the nine variables analyzed (Fig.
 4).
-
-We find that seasonality is an important factor in understanding patterns in productivity on a global scale: temperature seasonality and growing season length are strong predictors of productivity, though growing season length doesn't improve on MAT as a predictor.
-Within the growing season we find the influence of climate on productivity is smaller, but still significant for a number of carbon fluxes.
+Temperature seasonality and growing season length are closely correlated with MAT and are strong predictors of FACF (Fig.
+5e-f), though growing season length doesn't improve upon MAT as a predictor.
+Within the growing season, the influence of climate on C cycling is smaller but still significant for a number of carbon fluxes (**some SI table?**).
+These findings clarify the big picture of how FACF varies across latitude and relates to climate on a global scale.
 

### 2019-09-10 manuscript/Discussion_KAT.Rmd
Update Discussion_KAT.Rmd

@@
 We used a large global database (ForC), containing an unprecedented amount of data representing all major forest types (Fig.
-1) and the nine most significant forest autotrophic carbon flux (FACF) variables (Table 2), to comprehensively explore how C cycling in mature forests varies with latitude and climate on a global scale.
+1) and the nine most significant forest autotrophic carbon flux (FACF) variables (Table 2), to comprehensively explore how C cycling in mature, undisturbed forests varies with latitude and climate on a global scale.
 We test a suite of hypotheses, including many previously posed (Table 1), with more data and standardisation for factors that have not always been controlled for (*e.g.*, stand age, flux components measured) to gain a stronger understanding of climatic effects on FACF.
@@
 Within the growing season, the influence of climate on C cycling is smaller but still significant for a number of carbon fluxes (**some SI table?**).
-These findings clarify the big picture of how FACF varies across latitude and relates to climate on a global scale.
+These findings clarify the big picture of how FACF varies with latitude and climate on a global scale.
+
+Past studies have differed in their conclusions regarding the relationship between productivity and latitude or its correlates (**REFS**)--quite possibly because of lack of standardization with respect to stand age and disturbance history.
+Our findings indicate that, among mature, undisturbed stands, FACF is unambiguously highest in the tropical regions, and the relationship is approximately linear (Figs.
+2-3).
+[Discuss how studies arguing for higher productivity in temperate regions may have confounding effect of age]
+
+[The fact that we detected no major trends in C allocation is quite signficant (even though analysis not designed specifically for this) (1) Discuss temperature senstivity of productivity vs respiration, CUE.
+While there's lots of support for this on shorter time scales, we don't find support at the annual scale.
+(2) More generally, implies that some of the previously observed allocation patterns may be driven more by other factors--e.g., age (check refs)]
+
+[Variation in how much variability is explained by climate-- hard to say if this is methodological or real. discuss methodological limitations, also local allocation gradients.]
+
+[paragraph discussing climate drivers.
+MAT as best, although the advantage of MAT over other temperature variables may in part be because of better quality data .
+Regarding seasonality, FACF and climate data lack the temporal resolution required to sort this out.
+We find no support for the argument that correcting for growing season length improves the explanatory power.]
+
+[Despite the coarse resolution of this analysis, it holds some important implications for our understanding of how forest carbon cycling relates to climate and, by extension, how it is likely to respond to climatic warming.
+We see no evidence that there are major differences in how the various components of forest autotrophic carbon cycling relate to climate or, by extension, any distinct trends in C allocation across the global scale (Fig.
+3).
+]
+
+~~~
 
@@
 
-Although we recognise the inherent limitations present in this analysis, by focusing on mature, undisturbed forests, and by clarifying the components of NPP included in analysis, our study offers a clearer picture of how carbon cycling varies with latitude and climate.
-
-Past studies have differed in their conclusions regarding the relationship between productivity and latitude.
-Our findings indicate that, for all carbon fluxes studied, productivity peaks in the tropical regions.
 There is strong evidence from our analyses that this increase in productivity towards the tropical regions is driven primarily by temperature-related effects.
@@
 ### Acknowledgements
+Scholarly Studies
+ForestGEO
+DOE database

### 2019-09-10 manuscript/manuscript draft.Rmd

@@
 Additional targeted literature searches were conducted to identify any further available data on primary productivity, with particular focus on mature forests in temperate and boreal regions.
-ForC currently contains **UPDATE STATS** 29768 records from 5227 plots, representing 20 distinct ecozones across all forested biogeographic and climate zones.
-We used ForC v.XX, archived on Zenodo with DOI XXX.
+ForC currently contains 29730 records from 4979 plots, representing 20 distinct ecozones across all forested biogeographic and climate zones.
+We used ForC v3.0, archived on Zenodo with DOI 10.5281/zenodo.3403855.
 
@@
 
-| Variable | Definition | Components included | Methodologies used | Number of records | Number of plots |
+| Variable | Definition | Components included | Methodologies used | Number of records | Number of geographic areas |
 |--------------|-------------------|-------------|------------------------|-------|-------|
-| GPP | Annual gross primary production; annual uptake of carbon dioxide by an ecosystem | NA | Flux partitioning of eddy covariance | 257 | 80 |
-| NPP | Annual net primary production; the component of GPP that is stored in plant tissue; GPP minus ecosystem respiration | Foliage, branch, stem, coarse root, fine root and optionally understory | Direct measurement of annual increments of components | 93 | 78 |
-| ANPP | Aboveground net primary production | Foliage, stem, and optionally branch | Direct measurement of annual increments of components | 258 | 166 |
-| ANPP~foliage~ | Net primary production of foliage | Foliage | Direct measurement of litterfall, correcting for changes in leaf biomass when measured | 98 | 94 |
-| ANPP~woody~ ~stem~ | Net primary production of woody stems | Woody stems | Direct measurement of stem growth increment | 266 | 252 |
-| BNPP~root~ | Belowground net primary production | Coarse and fine roots | Direct measurement of one or more of: fine root turnover, soil cores, root ingrowth cores, minirhizotrons; indirect estimates of coarse roots using allometries based on aboveground stem increment measures | 107 | 87 |
-| BNPP~fine~ ~root~ | Net primary production of fine roots | Fine roots | Direct measurement of one or more of: minirhizotrons, fine root turnover, soil cores, root ingrowth cores | 90 | 77 |
-| R~auto~ | Annual autotrophic respiration, including above- and belowground components | Foliage, stem, and root | Chamber measurements of component gas exchange | 22 | 22 |
-| R~root~ | Annual root respiration | Coarse and fine roots | Measurement of root gas exchange | 65 | 43 |
+| GPP | Annual gross primary production; annual uptake of carbon dioxide by an ecosystem | NA | Flux partitioning of eddy covariance | 243 | 49 |
+| NPP | Annual net primary production; the component of GPP that is stored in plant tissue; GPP minus ecosystem respiration | Foliage, branch, stem, coarse root, fine root and optionally understory | Direct measurement of annual increments of components | 92 | 42 |
+| ANPP | Aboveground net primary production | Foliage, stem, and optionally branch | Direct measurement of annual increments of components | 256 | 80 |
+| ANPP~foliage~ | Net primary production of foliage | Foliage | Direct measurement of litterfall, correcting for changes in leaf biomass when measured | 98 | 49 |
+| ANPP~woody~ ~stem~ | Net primary production of woody stems | Woody stems | Direct measurement of stem growth increment | 264 | 96 |
+| BNPP~root~ | Belowground net primary production | Coarse and fine roots | Direct measurement of one or more of: fine root turnover, soil cores, root ingrowth cores, minirhizotrons; indirect estimates of coarse roots using allometries based on aboveground stem increment measures | 101 | 48 |
+| BNPP~fine~ ~root~ | Net primary production of fine roots | Fine roots | Direct measurement of one or more of: minirhizotrons, fine root turnover, soil cores, root ingrowth cores | 88 | 41 |
+| R~auto~ | Annual autotrophic respiration, including above- and belowground components | Foliage, stem, and root | Chamber measurements of component gas exchange | 22 | 13 |
+| R~root~ | Annual root respiration | Coarse and fine roots | Measurement of root gas exchange | 64 | 26 |
 
-Table: Definitions and sample sizes of variables used in analysis
+Table: Definitions and sample sizes of variables used in analysis.
+Geographic areas group geographically proximate sites, defined using a hierarchical cluster analysis on the distance matrix of the sites, and a cutoff of 25km.
 
@@
 
-In total, we analyzed ## records from ## C flux variables taken from forests that had experienced no major anthropogenic disturbances within the past 100 years from # distinct geographic areas (Fig.
-1, Table 2).
-[**Add a bit about geographic representation**]
+In total, we analyzed 1228 records from 9 C flux variables taken from forests that had experienced no major anthropogenic disturbances within the past 100 years.
+These records represented a total of 154 distinct geographic areas (Fig.
+1, Table 2), across all forested biogeographic and climate zones.
 

### 2019-09-10 manuscript/manuscript draft.Rmd
added table to manuscript draft

@@
 options(tinytex.verbose = TRUE)
+options(knitr.table.format = "latex")
 ```
@@
 Here we use a comprehensive global database of forest carbon fluxes to explore the above questions for nine carbon fluxes, allowing for an in-depth exploration of the effect of climate on global productivity.
+
+\renewcommand{\arraystretch}{2}
+```{r eval = TRUE, echo=FALSE, warning=FALSE}
+library(knitr)
+library(kableExtra)
+hypothesis_table <- read.csv("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/hypothesis_table.csv", stringsAsFactors = FALSE, check.names = FALSE)
+kable(hypothesis_table, booktabs = TRUE, caption = "Hypothesis table") %>%
+  kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 12) %>%
+  pack_rows("1. How do FACF vary with latitude?", 1, 3, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("2. How do FACF relate to MAT and MAP?", 4, 6, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("3. How do FACF relate to other annual climate variables?", 7, 9, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("4. What is the role of seasonality?", 10, 13, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("5. Per month of the growing seaon, how do FACF vary with climate?", 14, 17, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  add_header_above(c(" ", " ", " ", "Forest autotrophic carbon fluxes (FACF)" = 9, " ")) %>%
+  column_spec(1, width = "10cm") %>%
+  column_spec(2, width = "5cm") %>%
+  column_spec(3:13, width = "1.8cm") %>%
+  kableExtra::landscape()
+```
 

### 2019-09-10 manuscript/Discussion_KAT.Rmd
Update Discussion_KAT.Rmd

@@
 The FACF variables increase in proportion to one another, with no differences in allocation detectable at this global scale and with component fluxes summing appropriately to larger fluxes (Fig.
-**2/3**), indicating no major, systematic omissions or overestimations of flux components.
+3), indicating no major, systematic omissions or overestimations of flux components.
 However, we did detect a tendency for greater variability among subsidiary C fluxes (e.g., *$ANPP_woody$*, *$BNPP_{fine root}$*; Fig.
-**2/3**; **some SI table?**; **discussed below??**).
+2; **some SI table?**; **discussed below??**).
 Climate explains a significant proportion of variation in all C fluxes analyzed (albiet less for subsidiary fluxes), with temperature variables being the best predictors of FACF on this global scale (Figs.
@@
 2-3).
-[Discuss how studies arguing for higher productivity in temperate regions may have confounding effect of age]
+This contrasts with suggestions that productivity of temperate forests is similar to that of tropical forests (**REFS**).
+Such observations may be attributable to the fact that temperate forests are predominantly secondary and of younger age than tropical forests (**REF**), so analyses comparing across latitudinal gradients without strict control of stand age tend to be comparing younger temperate stands with older tropical (and often boreal) forests ([Poulter et al.
+2018-DOI:10.1594/PANGAEA.889943](http://pure.iiasa.ac.at/id/eprint/15361/)).
+Because C allocation varies with stand age (**REFS**), age differences may introduce systematic biases into analyses of FACF across latitude or global climatic gradients.
+For example, woody productivity tends to be higher in rapidly aggrading secondary stands than in old-growth forests, where proportionally more C is allocated to respiration (**REFS**) [*purpose for respiration/ other compenents (**REFS????**)].
+By constraining our analysis to stands ≥100 years old with no record of major recent disturbance, we have clarified the shape of the relationship between FACF and latitude.
 

### 2019-09-10 manuscript/manuscript draft.Rmd
manuscript updates

@@
 
-In order to approach these broad and complex issues, we simplify the major gaps in our knowledge to five key questions and corresponding specific predictions (Table 1):  
-1.	How do carbon fluxes vary with latitude? 
-2.	How do carbon fluxes relate to MAT and MAP?  
-3.	How do carbon fluxes relate to other annual climate variables?  
-4.	What is the role of seasonality in explaining variation in carbon fluxes?  
-5.	Per month of the growing season, how does productivity vary with climate?
+In order to approach these broad and complex issues, we simplify the major gaps in our knowledge to five key questions and corresponding specific predictions (Table 1).
+Firstly, we ask how forest autotrophic carbon fluxes (FACF) vary with latitude.
+We then test how these fluxes relate to MAT and MAP, and additionally how they respond to other, less well studied, climate variables.
+We finally consider the relationship between FACF and seasonality, considering the role of seasonality in explaining variation in carbon fluxes, and the influence of climate on FACF per month of the growing season.
 
-To resolve these questions, we require data that, firstly, control for variables influencing productivity, including stand age, disturbance regime, elevation, and methodology; and secondly includes all major carbon fluxes.
+Resolving these questions requires data which, firstly, control for variables influencing productivity, including stand age, disturbance regime, elevation, and methodology; and secondly includes all major carbon fluxes.
 Here we use a comprehensive global database of forest carbon fluxes to explore the above questions for nine carbon fluxes, allowing for an in-depth exploration of the effect of climate on global productivity.

### 2019-09-10 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@
 
-*Length of growing season.* Growing season months were defined as months with mean minimum temperature > 0.5.
+*Length of growing season.* Growing season months were defined as months with mean minimum temperature > 0.5$^\circ$C.
 Growing season months were initially calculated following methods used by Kerkhoff et al.
@@
 ### Acknowledgements
+
+### Supporting Information
+
+```{r eval = TRUE, echo=FALSE, warning=FALSE}
+library(knitr)
+library(kableExtra)
+S1 <- read.csv("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/S1.csv", stringsAsFactors = FALSE, check.names = FALSE)
+kable(S1, booktabs = TRUE, caption = "Hypothesis table", escape = F)
+```

### 2019-09-10 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

--- 2019-09-10 manuscript/manuscript draft.Rmd
@@
 ### Acknowledgements
+
+### Supporting Information
+
+```{r eval = TRUE, echo=FALSE, warning=FALSE}
+library(knitr)
+library(kableExtra)
+S1 <- read.csv("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/S1.csv", stringsAsFactors = FALSE, check.names = FALSE)
+kable(S1, booktabs = TRUE, caption = "Hypothesis table", escape = F) %>%
+  kable_styling(latex_options = c("scale_down", "hold_position"))
+```

### 2019-09-10 manuscript/Discussion_KAT.Rmd
Update Discussion_KAT.Rmd

@@
 
-[Variation in how much variability is explained by climate-- hard to say if this is methodological or real. discuss methodological limitations, also local allocation gradients.]
+-- hard to say if this is methodological or real. discuss methodological limitations, also local allocation gradients.]
+
+One interesting observation was that climate explains more variation in the major fluxes (**LIST**) than in subsidiary fluxes (**LIST**) (Fig.
+2; **some SI table?**).
+There are two non-exclusive potential explanations for this.
+First, it may be that methodological variation is larger relative to flux magnitude for some of the subsidiary fluxes.
+Belowground fluxes in particular are difficult to quantify, and measurement methods for the belowground fluxes considered here may be measured through fundamentally different approaches (*e.g.*, **ingrowth cores, XXX for $BNPP_{fine root}$; root exclusion, stable isotope tracking, or gas exchange of excised roots for $R_root$**), and sampling depth is variable and often insufficient to capture the full soil profile.
+$ANPP_{woody stem}$, which is also poorly explained by latitude or climate, is more straightforward to measure but is subject to variability introduced by differences such as biomass allometries applied, minimum plant size sampled, and exclusion or inclusion of branch fall (**REF**).
+However, methodological variation and uncertainty affects all of fluxes considered here, and some of the larger fluxes that vary more strongly with respect to climate (ANPP, NPP) are estimated by summing uncertain component fluxes.
+Second, variation in the proportion of variability explained by climate could also indicate that factors other than climate have significant influence over subsidiary fluxes.
+This is consistent with the results we find for allocation between fluxes: although climate has a significant influence on subsidiary fluxes, it is not the only factor that is important, and many other local- and regional-scale factors  influence the way carbon is allocated between subsidiary fluxes.
+We would expect that this would reduce the strength of the effect of climate on subsidiary fluxes.
+[**work on this second part, add refs**]
 
@@
 
-Related to this, we find that climate more strongly explains variation in the major fluxes, with lower levels of variation explained in subsidiary fluxes.
-Although this could in part be related to a lack of standardisation in methodology across studies, as discussed above, it could also indicate that factors other than climate have significant influence over subsidiary fluxes.
-This is consistent with the results we find for allocation between fluxes: although climate has a significant influence on subsidiary fluxes, it is not the only factor that is important, and many other local- and regional-scale factors  influence the way carbon is allocated between subsidiary fluxes.
-We would expect that this would reduce the strength of the effect of climate on subsidiary fluxes.
-
 **[needs strong concluding paragraph]**

### 2019-09-10 manuscript/Discussion_KAT.Rmd
Update Discussion_KAT.Rmd

@@
 
--- hard to say if this is methodological or real. discuss methodological limitations, also local allocation gradients.]
-
 One interesting observation was that climate explains more variation in the major fluxes (**LIST**) than in subsidiary fluxes (**LIST**) (Fig.
@@
 $ANPP_{woody stem}$, which is also poorly explained by latitude or climate, is more straightforward to measure but is subject to variability introduced by differences such as biomass allometries applied, minimum plant size sampled, and exclusion or inclusion of branch fall (**REF**).
-However, methodological variation and uncertainty affects all of fluxes considered here, and some of the larger fluxes that vary more strongly with respect to climate (ANPP, NPP) are estimated by summing uncertain component fluxes.
+However, methodological variation and uncertainty affects all of fluxes considered here--not necessarily any less than the aforementioned, and some of the larger fluxes that vary more strongly with respect to climate (ANPP, NPP) are estimated by summing uncertain component fluxes.
 Second, variation in the proportion of variability explained by climate could also indicate that factors other than climate have significant influence over subsidiary fluxes.

### 2019-09-11 manuscript/Discussion_KAT.Rmd
Update Discussion_KAT.Rmd

@@
 
-One interesting observation was that climate explains more variation in the major fluxes (**LIST**) than in subsidiary fluxes (**LIST**) (Fig.
+One interesting observation was that climate explains more variation in the major fluxes ($GPP$, $NPP$, $R_{auto}$ - latitude R^2≥48%) than in subsidiary fluxes (latitude R^2<27% for $BNPP_{fine.root}$, $R_{auto-root}$,$ANPP_{woody-stem}$) (Fig.
 2; **some SI table?**).
@@
 First, it may be that methodological variation is larger relative to flux magnitude for some of the subsidiary fluxes.
-Belowground fluxes in particular are difficult to quantify, and measurement methods for the belowground fluxes considered here may be measured through fundamentally different approaches (*e.g.*, **ingrowth cores, XXX for $BNPP_{fine root}$; root exclusion, stable isotope tracking, or gas exchange of excised roots for $R_root$**), and sampling depth is variable and often insufficient to capture the full soil profile.
-$ANPP_{woody stem}$, which is also poorly explained by latitude or climate, is more straightforward to measure but is subject to variability introduced by differences such as biomass allometries applied, minimum plant size sampled, and exclusion or inclusion of branch fall (**REF**).
+Belowground fluxes in particular are difficult to quantify, and measurement methods for the belowground fluxes considered here may be measured through fundamentally different approaches (*e.g.*, **minirhizotrons, ingrowth cores, or sequential coring for $BNPP_{root-fine}$; root exclusion, stable isotope tracking, or gas exchange of excised roots for $R_{auto-root}$**), and sampling depth is variable and often insufficient to capture the full soil profile.
+$ANPP_{woody-stem}$, which is also poorly explained by latitude or climate, is more straightforward to measure but is subject to variability introduced by differences such as biomass allometries applied, minimum plant size sampled, and exclusion or inclusion of branch fall (**REF**).
 However, methodological variation and uncertainty affects all of fluxes considered here--not necessarily any less than the aforementioned, and some of the larger fluxes that vary more strongly with respect to climate (ANPP, NPP) are estimated by summing uncertain component fluxes.
-Second, variation in the proportion of variability explained by climate could also indicate that factors other than climate have significant influence over subsidiary fluxes.
+Second, differences among variables in the proportion of variation explained by climate could indicate that factors other than climate have significant influence over subsidiary fluxes.
 This is consistent with the results we find for allocation between fluxes: although climate has a significant influence on subsidiary fluxes, it is not the only factor that is important, and many other local- and regional-scale factors  influence the way carbon is allocated between subsidiary fluxes.
@@
 ~~~
-
-Despite the high standard of data quality in the ForC database, there are still significant limitations on our analyses.
-We standardised methodologies as far as possible, ensuring that there was consistency in the components included for each carbon flux.
-However, field techniques for estimating components of primary productivity are variable in their accuracy.
-This is particularly true for measures of BNPP, which are often estimated by extrapolating measures of ANPP.
-Even where BNPP is calculated through direct soil sampling, measures are often taken to insufficient depth, leading to underestimation of the flux.
-It is challenging to account for this type of variability in quality of field data, and our analyses largely assume good faith in data quality.
-While we have strong confidence in our analyses of well-understood fluxes, such as GPP, ANPP, and ANPP~woody~ ~stem~, we recommend caution in drawing firm conclusions from our analyses of BNPP, BNPP~fine~ ~root~, and, to some extent, NPP.
-We expect that, as field techniques improve the accuracy of estimates of these fluxes in the future, our understanding of the relationship of BNPP with climate variables will change.
-It is possible that the low R^2^ values for BNPP that we find are a result of lower accuracy in field measures of these fluxes, and that with improve standardisation of methods, we would find climate explains a higher proportion of variation.
 

### 2019-09-11 manuscript/manuscript draft.Rmd
worked on first paragraph of intro

--- 2019-09-10 manuscript/manuscript draft.Rmd
@@
 
-Globally, forests play an important role in the carbon cycle and are a major component of global carbon dioxide budgets [@luyssaert_old-growth_2008].
-They show higher levels of productivity than non-forest terrestrial ecosystems [@del_grosso_global_2008], and as a result achieve significant carbon sequestration and storage.
-Estimating the total role of forests in the carbon cycle is challenging, but studies indicate that old growth forests alone sequester up to 1.4 GtCyr^-1^ [@malhi_carbon_1999], while the total sequestration of carbon by established forests globally could be up to 2.4 GtCyr^-1^, with the largest sinks being in old-growth tropical forests [@pan_large_2011].
-As atmospheric carbon dioxide levels continue to rise, with consequences for global climate, there is increasing recognition that proper protection and management of forest resources will have an important role to play in mitigating climate change.
-Understanding the patterns of forest productivity on a global scale, and the drivers behind them, is therefore a priority in forest research.
-[**KAT-work on this; consider shifting top focus from global C cycle to centrality of forest productivity to ecology & climate science.** ] Given the importance of understanding broad-scale patterns in forest productivity, there have been many meta-analyses on the theme.
+Carbon fixation, allocation, and metabolism by trees sets the basis for energy and material flows in forest ecosystems and defines their interactions with Earth's changing climate.
+Forest autotrophic carbon flux (FACF) influences all organic matter stocks in ecosystems and is linked to cycling of energy, water, and nutrients (**REFS**).
+Forest productivity sets the energy available to heterotrophic communities, in turn influencing their abundance (**REFS**) and possibly diversity (**REFS**)--but probably not individual metabolic rate (**Anderson & Jetz 2005 (my first paper!!), but check more recent literature**).
+On the global scale, forests play a critical role in regulating atmospheric CO_2 and climate (**Bonan**).
+The total amount of CO_2 cycling through Earth's forests each year (*i.e.*, total gross primary production, $GPP$) is more than five times anthropogenic fossil fuel emissions, and net sequestration (~2.4 GtCyr^-1^) offsetting roughly 30% of these emissions [@pan_large_2011].
+As atmospheric carbon dioxide levels continue to rise, with consequences for global climate, there is increasing recognition that proper protection and management of forest resources will have an important role to play in mitigating climate change (** REFS- e.g., IPCC land report**).
+Given the importance of understanding broad-scale patterns in forest autotrophic carbon flux (FACF), there have been many meta-analyses on the theme.
 However, our ability to draw general macroscopic conclusions regarding global variation in multiple productivity variables with respect to climate has been limited in that these analyses often mix forests that vary in stand age, disturbance history, and/or management status; do not always sufficiently parse related variables (*e.g.*, combining net primary productivity records with and without belowground components); and typically consider only one or a few variables at a time.

### 2019-09-11 manuscript/Discussion_KAT.Rmd
Update Discussion_KAT.Rmd

@@
 (2) More generally, implies that some of the previously observed allocation patterns may be driven more by other factors--e.g., age (check refs)]
+By including multiple fluxes in our analysis, we are able to explore the relationships between fluxes and their responses to climate.
+We find that fluxes are broadly consistent in their responses to climate drivers, suggesting that the importance of climate drivers on productivity is relatively invariant across fluxes.
+In addition, we show that carbon allocation between subsidiary fluxes does not vary significantly across broad climatic gradients.
+There is evidence that allocation between carbon fluxes does vary with factors such as stand age [@litton_carbon_2007], nutrient availability [@litton_carbon_2007; @gill_belowground_2016], forest structure [@taylor_greater_2019], elevation [@moser_elevation_2011], and water availability [@newman_above-_2006].
+That we do not see consistent patterns of variation in allocation across climatic gradients indicates that the pressures influencing shifts in allocation between fluxes occur on local scales, and that shifts in allocation occur as a response to fluctuations in local environmental conditions.
 
@@
 
-By including multiple fluxes in our analysis, we are able to explore the relationships between fluxes and their responses to climate.
-We find that fluxes are broadly consistent in their responses to climate drivers, suggesting that the importance of climate drivers on productivity is relatively invariant across fluxes.
-In addition, we show that carbon allocation between subsidiary fluxes does not vary significantly across broad climatic gradients.
-There is evidence that allocation between carbon fluxes does vary with factors such as stand age [@litton_carbon_2007], nutrient availability [@litton_carbon_2007; @gill_belowground_2016], forest structure [@taylor_greater_2019], elevation [@moser_elevation_2011], and water availability [@newman_above-_2006].
-That we do not see consistent patterns of variation in allocation across climatic gradients indicates that the pressures influencing shifts in allocation between fluxes occur on local scales, and that shifts in allocation occur as a response to fluctuations in local environmental conditions.
-
 **[needs strong concluding paragraph]**

### 2019-09-11 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@
 The total amount of CO_2 cycling through Earth's forests each year (*i.e.*, total gross primary production, $GPP$) is more than five times anthropogenic fossil fuel emissions, and net sequestration (~2.4 GtCyr^-1^) offsetting roughly 30% of these emissions [@pan_large_2011].
-As atmospheric carbon dioxide levels continue to rise, with consequences for global climate, there is increasing recognition that proper protection and management of forest resources will have an important role to play in mitigating climate change (** REFS- e.g., IPCC land report**).
+As atmospheric carbon dioxide levels continue to rise, with consequences for global climate, there is increasing recognition that proper protection and management of forest resources will have an important role to play in mitigating climate change (** REFS- e.g., IPCC 1.5C and land reports**).
 Given the importance of understanding broad-scale patterns in forest autotrophic carbon flux (FACF), there have been many meta-analyses on the theme.

### 2019-09-11 manuscript/Discussion_KAT.Rmd
Update Discussion_KAT.Rmd

--- 2019-09-11 manuscript/Discussion_KAT.Rmd
@@
 2018-DOI:10.1594/PANGAEA.889943](http://pure.iiasa.ac.at/id/eprint/15361/)).
-Because C allocation varies with stand age (**REFS**), age differences may introduce systematic biases into analyses of FACF across latitude or global climatic gradients.
-For example, woody productivity tends to be higher in rapidly aggrading secondary stands than in old-growth forests, where proportionally more C is allocated to respiration (**REFS**) [*purpose for respiration/ other compenents (**REFS????**)].
+Because C allocation varies with stand age (**REFS (KAT/NOBBY)**), age differences may introduce systematic biases into analyses of FACF across latitude or global climatic gradients.
+For example, woody productivity tends to be higher in rapidly aggrading secondary stands than in old-growth forests, where proportionally more C is allocated to respiration (**REFS**) [*purpose for respiration/ other compenents (**REFS????- NOBBY?**)].
 By constraining our analysis to stands ≥100 years old with no record of major recent disturbance, we have clarified the shape of the relationship between FACF and latitude.
 
-[The fact that we detected no major trends in C allocation is quite signficant (even though analysis not designed specifically for this) (1) Discuss temperature senstivity of productivity vs respiration, CUE.
+We show that fluxes are broadly consistent in their responses to climate drivers on the global scale (with the exception of some differences in MAT-MAP interactions; Fig.
+4), with no major trends in C allocatio (Fig.
+2).
+This finding contradicts some of the dominant current understanding, with significant implications for our understanding of FACF responses to climatic warming.
+Of particular interest are the relationships between $GPP$ and respiration ($R_{auto}$, $R_{root}$) or net primary productivity ($NPP$, $ANPP$, $BNPP$).
+The dominant current understanding is that *respiration responds more strongly to temperature than does photosynthesis (**REFS(KAT)**)*, and that carbon use efficiency ($CUE$= $NPP$/$GPP$=($GPP$-$R_{auto}$/$GPP$) is lowest in the tropics (**REFS(KAT)**).
+The latter understanding is driven by observations of low $CUE$ in old-growth tropical forests relative to extratropical forests [@de_lucia_forest_2007; @anderson-teixeira_carbon_2016] (**REFS(KAT)**), but our analysis suggests that these low values may more appropriately be attributed to the fact that these forests are old than to their climate.
+Indeed, $CUE$ is known to decline with forest age ([@de_lucia_forest_2007];**REFS(NOBBY)**).
+Whereas autotrophic respiration responds more strongly to temperature than does photosynthesis on short time scales (**REFS(KAT)**), our findings indicate that annual C allocation to respiration is roughly independent of temperature.
+[There is some indication that $GPP$ may have a somewhat steeper relationship to latitude and temperature than does $ANPP_{woody-stem}$ (Fig.
+2; see also Larjavaara & Muller-Landau (2011)?), but standardized slopes are not signficnatly different (**??-BECKY**).
+(**Helene, maybe comment on Larjavaara & Muller-Landau (2011)?**).
+] While our analysis indicates that climate shapes most FACFs in similar ways on the global scale, careful methodological standardization across a consistent set of mature forest sites could potentially reveal modest but potentially signficant trends in carbon allocation along global climatic gradients.
+
+(1) Discuss temperature senstivity of productivity vs respiration, CUE.
 While there's lots of support for this on shorter time scales, we don't find support at the annual scale.
 (2) More generally, implies that some of the previously observed allocation patterns may be driven more by other factors--e.g., age (check refs)]
+
+We find that fluxes are broadly consistent in their responses to climate drivers, suggesting that the importance of climate drivers on productivity is relatively invariant across fluxes.
+In addition, we show that carbon allocation between subsidiary fluxes does not vary significantly across broad climatic gradients.
+There is evidence that allocation between carbon fluxes does vary with factors such as stand age [@litton_carbon_2007], nutrient availability [@litton_carbon_2007; @gill_belowground_2016], forest structure [@taylor_greater_2019], elevation [@moser_elevation_2011], and water availability [@newman_above-_2006].
+That we do not see consistent patterns of variation in allocation across climatic gradients indicates that the pressures influencing shifts in allocation between fluxes occur on local scales, and that shifts in allocation occur as a response to fluctuations in local environmental conditions.
 

### 2019-09-11 manuscript/Discussion_KAT.Rmd
Update Discussion_KAT.Rmd

@@
 Whereas autotrophic respiration responds more strongly to temperature than does photosynthesis on short time scales (**REFS(KAT)**), our findings indicate that annual C allocation to respiration is roughly independent of temperature.
-[There is some indication that $GPP$ may have a somewhat steeper relationship to latitude and temperature than does $ANPP_{woody-stem}$ (Fig.
+[There is some indication that $GPP$ may have a somewhat steeper and less variable relationship to latitude and temperature than does $ANPP_{woody-stem}$ (Fig.
 2; see also Larjavaara & Muller-Landau (2011)?), but standardized slopes are not signficnatly different (**??-BECKY**).
@@
 ] While our analysis indicates that climate shapes most FACFs in similar ways on the global scale, careful methodological standardization across a consistent set of mature forest sites could potentially reveal modest but potentially signficant trends in carbon allocation along global climatic gradients.
-
-(1) Discuss temperature senstivity of productivity vs respiration, CUE.
-While there's lots of support for this on shorter time scales, we don't find support at the annual scale.
-(2) More generally, implies that some of the previously observed allocation patterns may be driven more by other factors--e.g., age (check refs)]
-
-We find that fluxes are broadly consistent in their responses to climate drivers, suggesting that the importance of climate drivers on productivity is relatively invariant across fluxes.
-In addition, we show that carbon allocation between subsidiary fluxes does not vary significantly across broad climatic gradients.
-There is evidence that allocation between carbon fluxes does vary with factors such as stand age [@litton_carbon_2007], nutrient availability [@litton_carbon_2007; @gill_belowground_2016], forest structure [@taylor_greater_2019], elevation [@moser_elevation_2011], and water availability [@newman_above-_2006].
-That we do not see consistent patterns of variation in allocation across climatic gradients indicates that the pressures influencing shifts in allocation between fluxes occur on local scales, and that shifts in allocation occur as a response to fluctuations in local environmental conditions.
+[There is evidence that allocation between carbon fluxes does vary with factors such as stand age [@litton_carbon_2007], nutrient availability [@litton_carbon_2007; @gill_belowground_2016], forest structure [@taylor_greater_2019], elevation [@moser_elevation_2011], and water availability [@newman_above-_2006].
+That we do not see consistent patterns of variation in allocation across global climatic gradients indicates that the pressures influencing shifts in allocation between fluxes occur on local scales and/or in response to factors other than mean climate, including inter- and intra-annual climiatic fluctuations.
+]
 

### 2019-09-11 manuscript/Discussion_KAT.Rmd
Update Discussion_KAT.Rmd

@@
 
+[**THIS PARAGRAPH NEEDS ATTENTION!**].
 We show that fluxes are broadly consistent in their responses to climate drivers on the global scale (with the exception of some differences in MAT-MAP interactions; Fig.
-4), with no major trends in C allocatio (Fig.
-2).
-This finding contradicts some of the dominant current understanding, with significant implications for our understanding of FACF responses to climatic warming.
-Of particular interest are the relationships between $GPP$ and respiration ($R_{auto}$, $R_{root}$) or net primary productivity ($NPP$, $ANPP$, $BNPP$).
+4), with no major trends in C allocation among the variable pairs tested (Fig.
+2; **Some SI table**).
+*This finding does not support some of the dominant current understanding, with significant implications for our understanding of FACF responses to climatic warming.* Of particular interest are the relationships between $GPP$ and respiration ($R_{auto}$, $R_{root}$) or net primary productivity ($NPP$, $ANPP$, $BNPP$).
 The dominant current understanding is that *respiration responds more strongly to temperature than does photosynthesis (**REFS(KAT)**)*, and that carbon use efficiency ($CUE$= $NPP$/$GPP$=($GPP$-$R_{auto}$/$GPP$) is lowest in the tropics (**REFS(KAT)**).
@@
 Belowground fluxes in particular are difficult to quantify, and measurement methods for the belowground fluxes considered here may be measured through fundamentally different approaches (*e.g.*, **minirhizotrons, ingrowth cores, or sequential coring for $BNPP_{root-fine}$; root exclusion, stable isotope tracking, or gas exchange of excised roots for $R_{auto-root}$**), and sampling depth is variable and often insufficient to capture the full soil profile.
-$ANPP_{woody-stem}$, which is also poorly explained by latitude or climate, is more straightforward to measure but is subject to variability introduced by differences such as biomass allometries applied, minimum plant size sampled, and exclusion or inclusion of branch fall (**REF**).
+$ANPP_{woody-stem}$, which is also poorly explained by latitude or climate, is more straightforward to measure but is subject to variability introduced by differences such as biomass allometries applied, minimum plant size sampled, and exclusion or inclusion of branch fall (**?- Becky, did we end up lumping?
+REF**).
 However, methodological variation and uncertainty affects all of fluxes considered here--not necessarily any less than the aforementioned, and some of the larger fluxes that vary more strongly with respect to climate (ANPP, NPP) are estimated by summing uncertain component fluxes.
 Second, differences among variables in the proportion of variation explained by climate could indicate that factors other than climate have significant influence over subsidiary fluxes.
-This is consistent with the results we find for allocation between fluxes: although climate has a significant influence on subsidiary fluxes, it is not the only factor that is important, and many other local- and regional-scale factors  influence the way carbon is allocated between subsidiary fluxes.
-We would expect that this would reduce the strength of the effect of climate on subsidiary fluxes.
-[**work on this second part, add refs**]
+*This is consistent with the results we find for allocation between fluxes: although climate has a significant influence on subsidiary fluxes, it is not the only factor that is important, and many other local- and regional-scale factors  influence the way carbon is allocated between subsidiary fluxes.
+We would expect that this would reduce the strength of the effect of climate on subsidiary fluxes.* [**work on this second part, add refs**]
 

### 2019-09-11 manuscript/Discussion_KAT.Rmd
Update Discussion_KAT.Rmd

@@
 
-[**THIS PARAGRAPH NEEDS ATTENTION!**].
-We show that fluxes are broadly consistent in their responses to climate drivers on the global scale (with the exception of some differences in MAT-MAP interactions; Fig.
+[**NOBBY, please pay special attention to this paragraph**].
+We show that FACF are broadly consistent in their responses to climate drivers on the global scale (with the exception of some differences in MAT-MAP interactions; Fig.
 4), with no major trends in C allocation among the variable pairs tested (Fig.
 2; **Some SI table**).
-*This finding does not support some of the dominant current understanding, with significant implications for our understanding of FACF responses to climatic warming.* Of particular interest are the relationships between $GPP$ and respiration ($R_{auto}$, $R_{root}$) or net primary productivity ($NPP$, $ANPP$, $BNPP$).
-The dominant current understanding is that *respiration responds more strongly to temperature than does photosynthesis (**REFS(KAT)**)*, and that carbon use efficiency ($CUE$= $NPP$/$GPP$=($GPP$-$R_{auto}$/$GPP$) is lowest in the tropics (**REFS(KAT)**).
-The latter understanding is driven by observations of low $CUE$ in old-growth tropical forests relative to extratropical forests [@de_lucia_forest_2007; @anderson-teixeira_carbon_2016] (**REFS(KAT)**), but our analysis suggests that these low values may more appropriately be attributed to the fact that these forests are old than to their climate.
-Indeed, $CUE$ is known to decline with forest age ([@de_lucia_forest_2007];**REFS(NOBBY)**).
-Whereas autotrophic respiration responds more strongly to temperature than does photosynthesis on short time scales (**REFS(KAT)**), our findings indicate that annual C allocation to respiration is roughly independent of temperature.
-[There is some indication that $GPP$ may have a somewhat steeper and less variable relationship to latitude and temperature than does $ANPP_{woody-stem}$ (Fig.
-2; see also Larjavaara & Muller-Landau (2011)?), but standardized slopes are not signficnatly different (**??-BECKY**).
-(**Helene, maybe comment on Larjavaara & Muller-Landau (2011)?**).
-] While our analysis indicates that climate shapes most FACFs in similar ways on the global scale, careful methodological standardization across a consistent set of mature forest sites could potentially reveal modest but potentially signficant trends in carbon allocation along global climatic gradients.
-[There is evidence that allocation between carbon fluxes does vary with factors such as stand age [@litton_carbon_2007], nutrient availability [@litton_carbon_2007; @gill_belowground_2016], forest structure [@taylor_greater_2019], elevation [@moser_elevation_2011], and water availability [@newman_above-_2006].
-That we do not see consistent patterns of variation in allocation across global climatic gradients indicates that the pressures influencing shifts in allocation between fluxes occur on local scales and/or in response to factors other than mean climate, including inter- and intra-annual climiatic fluctuations.
-]
+Although variation in allocation has been observed along gradients of elevation [@moser_elevation_2011] and water availability [@newman_above-_2006]--along with non-climatic axes of stand age [@litton_carbon_2007], nutrient availability [@litton_carbon_2007; @gill_belowground_2016], and forest structure [@taylor_greater_2019]--variation in relation to climate is not apparent at the global scale within ForC, which contains the bulk of relevant data.
+Our conclusion, then, is that hypothesized gradients in allocation along global climate gradients cannot currently be supported for mature forests, although data quantity and standardization is currently insufficent to rule out the possibility that such trends exist.
+Of particular interest and significance are the relationships amongst $GPP$, net primary productivity ($NPP$ and its components, particularly $ANPP_{woody-stem}$), and respiration ($R_{auto}$, $R_{root}$).
+There have been suggestions that tropical forests tend to have low carbon use efficiency ($CUE$= $NPP$/$GPP$=($GPP$-$R_{auto}$/$GPP$), which are based on observations of low $CUE$ in old-growth tropical forests relative to extratropical forests [@de_lucia_forest_2007; Malhi 2012; @anderson-teixeira_carbon_2016], but our analysis suggests that these low values might more appropriately be attributed to the fact that these forests are old than to their tropical climate.
+Indeed, $CUE$ is known to decline with forest age ([@de_lucia_forest_2007];**REFS(NOBBY)**) but appears to be roughly independent of $GPP$ [@litton_carbon_2007].
+Similarly, it has been suggested that maintenance costs ($R_{auto}$+ **XX** per unit biomass) increase more rapidly with temperature than does $GPP$ (Larjavaara & Muller-Landau 2011), implying that $ANPP_{woody-stem}$ would have a shallower slope than $GPP$.
+There is some indication that this may be the case (Fig.
+2), but again careful methodological standardization across a consistent set of mature forest sites will be necessary to determine whether any global gradients in allocation exist.
 

### 2019-09-11 manuscript/Discussion_KAT.Rmd

@@
 Our conclusion, then, is that hypothesized gradients in allocation along global climate gradients cannot currently be supported for mature forests, although data quantity and standardization is currently insufficent to rule out the possibility that such trends exist.
-Of particular interest and significance are the relationships amongst $GPP$, net primary productivity ($NPP$ and its components, particularly $ANPP_{woody-stem}$), and respiration ($R_{auto}$, $R_{root}$).
-There have been suggestions that tropical forests tend to have low carbon use efficiency ($CUE$= $NPP$/$GPP$=($GPP$-$R_{auto}$/$GPP$), which are based on observations of low $CUE$ in old-growth tropical forests relative to extratropical forests [@de_lucia_forest_2007; Malhi 2012; @anderson-teixeira_carbon_2016], but our analysis suggests that these low values might more appropriately be attributed to the fact that these forests are old than to their tropical climate.
+Of particular interest and significance are the relationships amongst $GPP$, net primary productivity ($NPP$ and its components, particularly $ANPP_{woody-stem}$), and respiration ($R_{auto}$ and components).
+There have been suggestions that tropical forests tend to have low carbon use efficiency ($CUE$= $NPP$/$GPP$=($GPP$-$R_{auto}$/$GPP$), which are based on observations of low $CUE$ in old-growth tropical forests relative to (mostly younger) extratropical forests [@de_lucia_forest_2007; [Malhi 2012](https://www.researchgate.net/publication/227837530_The_productivity_metabolism_and_carbon_cycle_of_tropical_forest_vegetation); @anderson-teixeira_carbon_2016], but our analysis suggests that these low values might more appropriately be attributed to the fact that these forests are old than to their tropical climate.
 Indeed, $CUE$ is known to decline with forest age ([@de_lucia_forest_2007];**REFS(NOBBY)**) but appears to be roughly independent of $GPP$ [@litton_carbon_2007].
-Similarly, it has been suggested that maintenance costs ($R_{auto}$+ **XX** per unit biomass) increase more rapidly with temperature than does $GPP$ (Larjavaara & Muller-Landau 2011), implying that $ANPP_{woody-stem}$ would have a shallower slope than $GPP$.
+Similarly, it has been suggested that respiration, maintenance costs, and/or allocation to functions other than woody growth increase more rapidly with temperature than does $GPP$ (**REFS**; Larjavaara & Muller-Landau 2011), implying that $NPP$ and $ANPP_{woody-stem}$ would have a shallower slope than $GPP$.
 There is some indication that this may be the case (Fig.
-2), but again careful methodological standardization across a consistent set of mature forest sites will be necessary to determine whether any global gradients in allocation exist.
+2), but careful methodological standardization across a consistent set of mature forest sites will be necessary to determine whether any climate-driven gradients in allocation exist at the global scale.
 
-One interesting observation was that climate explains more variation in the major fluxes ($GPP$, $NPP$, $R_{auto}$ - latitude R^2≥48%) than in subsidiary fluxes (latitude R^2<27% for $BNPP_{fine.root}$, $R_{auto-root}$,$ANPP_{woody-stem}$) (Fig.
+One interesting observation was that climate tends to explain more variation in the major fluxes ($GPP$, $NPP$, $R_{auto}$ - latitude R^2≥48%) than in subsidiary fluxes (latitude R^2<27% for $BNPP_{fine.root}$, $R_{auto-root}$,$ANPP_{woody-stem}$) (Fig.
 2; **some SI table?**).
@@
 First, it may be that methodological variation is larger relative to flux magnitude for some of the subsidiary fluxes.
-Belowground fluxes in particular are difficult to quantify, and measurement methods for the belowground fluxes considered here may be measured through fundamentally different approaches (*e.g.*, **minirhizotrons, ingrowth cores, or sequential coring for $BNPP_{root-fine}$; root exclusion, stable isotope tracking, or gas exchange of excised roots for $R_{auto-root}$**), and sampling depth is variable and often insufficient to capture the full soil profile.
+Belowground fluxes in particular are difficult to quantify, and measurement methods for the belowground fluxes considered here may be measured through fundamentally different approaches (*e.g.*, minirhizotrons, ingrowth cores, or sequential coring for $BNPP_{root-fine}$; root exclusion, stable isotope tracking, or gas exchange of excised roots for $R_{auto-root}$), and sampling depth is variable and often insufficient to capture the full soil profile.
 $ANPP_{woody-stem}$, which is also poorly explained by latitude or climate, is more straightforward to measure but is subject to variability introduced by differences such as biomass allometries applied, minimum plant size sampled, and exclusion or inclusion of branch fall (**?- Becky, did we end up lumping?
@@
 Second, differences among variables in the proportion of variation explained by climate could indicate that factors other than climate have significant influence over subsidiary fluxes.
-*This is consistent with the results we find for allocation between fluxes: although climate has a significant influence on subsidiary fluxes, it is not the only factor that is important, and many other local- and regional-scale factors  influence the way carbon is allocated between subsidiary fluxes.
+*This is consistent with the results we find for allocation among fluxes: although climate has a significant influence on subsidiary fluxes, it is not the only factor that is important, and many other local- and regional-scale factors  influence the way carbon is allocated among subsidiary fluxes.
 We would expect that this would reduce the strength of the effect of climate on subsidiary fluxes.* [**work on this second part, add refs**]

### 2019-09-11 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@
     fig_caption: yes
+    includes:
+      after_body: appendix.md
 bibliography: [library.bib,  packages.bib]
@@
 S1 <- read.csv("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/S1.csv", stringsAsFactors = FALSE, check.names = FALSE)
-kable(S1, booktabs = TRUE, caption = "Hypothesis table", escape = F)
+kable(S1, format = "latex", booktabs = TRUE, caption = "Supporting Information Table 1 (S1): ", escape = F)%>%
+  kable_styling(latex_options = c("scale_down", "hold_position")) %>%
+  column_spec(1, width = "5cm") %>%
+  column_spec(2, width = "2cm") %>%
+  column_spec(3:4, width = "7cm") 
 ```

### 2019-09-11 manuscript/Discussion_KAT.Rmd
Update Discussion_KAT.Rmd

@@
 Indeed, $CUE$ is known to decline with forest age ([@de_lucia_forest_2007];**REFS(NOBBY)**) but appears to be roughly independent of $GPP$ [@litton_carbon_2007].
-Similarly, it has been suggested that respiration, maintenance costs, and/or allocation to functions other than woody growth increase more rapidly with temperature than does $GPP$ (**REFS**; Larjavaara & Muller-Landau 2011), implying that $NPP$ and $ANPP_{woody-stem}$ would have a shallower slope than $GPP$.
-There is some indication that this may be the case (Fig.
-2), but careful methodological standardization across a consistent set of mature forest sites will be necessary to determine whether any climate-driven gradients in allocation exist at the global scale.
+Among our sites with relevant data, there is indication in our data that CUE or $ANPP_{woody-stem}$/$GPP$ increase with latitude (**some SI table**).
+Careful methodological standardization across a consistent set of mature forest sites will be necessary to determine whether any climate-driven gradients in allocation exist at the global scale.
 
@@
 REF**).
-However, methodological variation and uncertainty affects all of fluxes considered here--not necessarily any less than the aforementioned, and some of the larger fluxes that vary more strongly with respect to climate (ANPP, NPP) are estimated by summing uncertain component fluxes.
-Second, differences among variables in the proportion of variation explained by climate could indicate that factors other than climate have significant influence over subsidiary fluxes.
-*This is consistent with the results we find for allocation among fluxes: although climate has a significant influence on subsidiary fluxes, it is not the only factor that is important, and many other local- and regional-scale factors  influence the way carbon is allocated among subsidiary fluxes.
-We would expect that this would reduce the strength of the effect of climate on subsidiary fluxes.* [**work on this second part, add refs**]
+However, methodological variation and uncertainty affect all of fluxes considered here--not necessarily any less than the aforementioned, and some of the larger fluxes that vary more strongly with respect to climate ($ANPP$, $NPP$) are estimated by summing uncertain component fluxes.
+Second, differences among variables in the proportion of variation explained by climate may be attributable to more dicrect climatic control over $GPP$ than subsidiary fluxes.
+That is, subsidiary fluxes may be shaped by climate both indirectly through its influence on $GPP$ and potentially respiration and directly through any climatic influence on C allocation, as well as by many other local- and regional-scale factors (**REFS**).
 

### 2019-09-11 manuscript/Discussion_KAT.Rmd
Update Discussion_KAT.Rmd

@@
 
+There is strong evidence from our analyses that the latituidinal gradient in FACF (Figs.
+2-3) is driven primarily by temperature-related climate variables (Figs.
+4-5).
+Because many climate variables co-vary across the latitudinal gradient, because climatic drivers affect forest carbon flux on much shorter time scales than can be captured by annual climate summary variables, and because both climatic conditions and C flux vary interannually around long-term means, we would not expect a single  variable characterizing mean annual climate to adequately predict variation in FACF.
+However, it remains informative to consider these relationships.
+
 [paragraph discussing climate drivers.

### 2019-09-12 manuscript/Discussion_KAT.Rmd
Update Discussion_KAT.Rmd

--- 2019-09-11 manuscript/Discussion_KAT.Rmd
@@
 
+The latitudinal gradient in FACF (Figs.
+2-3) is driven primarily by temperature-related climate variables, with modification by moisture availability (Figs.
+4-5).
+Because many climate variables co-vary across the latitudinal gradient, climatic drivers affect forest carbon flux on much shorter time scales than can be captured by annual climate summary variables, and both climatic conditions and C flux vary interannually around the long-term means, we would not expect any mean annual climate variable to explain variation in FACF.
+However, it remains informative to consider these relationships.
+We find that temperature-related climate variables explain the highest proportion of variability in productivity, and among these, $MAT$ is generally the best predictor--perhaps because site-specific MAT is recorded for the majority of sites in ForC, whereas other variables are extracted from a global database.
+....
+The effects of temperature are modified by moisture availability, ...
+
 [paragraph discussing climate drivers.
@@
 ~~~
-
-Secondly, we emphasise that this study considers the relationship between productivity and climate on a coarse scale.
-Because of the nature of the ForC database, we were unable to access site-level MAT and MAP data for a number of sites, and lacked site-level data for other climate variables for all sites.
-As a result the climate data we used was primarily multi-year averages of annual climate measures accessed from global databases.
-While the quality of these data is high, we cannot expect that it captures detailed local variation in climate, and in particular will be unable to capture microclimate variation, which may have a strong influence on stand-level productivity.
-Although we find that MAT is the strongest predictor for climate, part of this effect may be a result of the higher-quality stand-level dataset available for MAT.
-With high-quality stand-level data for other climate variables we expect that their predictive strength would improve.
 

### 2019-09-12 manuscript/Discussion_KAT.Rmd
Update Discussion_KAT.Rmd

@@
 
-[paragraph discussing climate drivers.
-MAT as best, although the advantage of MAT over other temperature variables may in part be because of better quality data .
-Regarding seasonality, FACF and climate data lack the temporal resolution required to sort this out.
+[ Regarding seasonality, FACF and climate data lack the temporal resolution required to sort this out.
 We find no support for the argument that correcting for growing season length improves the explanatory power.]

### 2019-09-12 manuscript/Discussion_KAT.Rmd
Update Discussion_KAT.Rmd

@@
 The latitudinal gradient in FACF (Figs.
-2-3) is driven primarily by temperature-related climate variables, with modification by moisture availability (Figs.
+2-3) is driven primarily by temperature-related climate variables, and secondarily by moisture availability (Figs.
 4-5).
-Because many climate variables co-vary across the latitudinal gradient, climatic drivers affect forest carbon flux on much shorter time scales than can be captured by annual climate summary variables, and both climatic conditions and C flux vary interannually around the long-term means, we would not expect any mean annual climate variable to explain variation in FACF.
+Because many climate variables co-vary across the latitudinal gradient, because climatic drivers affect forest carbon flux on much shorter time scales than can be captured by annual climate summary variables, and because both climatic conditions and C flux vary intra- and inter-annually around the long-term means, it is not appropriate to attempt to identify any one mean annual climate variable as a mechanistic driver of FACF.
 However, it remains informative to consider these relationships.
-We find that temperature-related climate variables explain the highest proportion of variability in productivity, and among these, $MAT$ is generally the best predictor--perhaps because site-specific MAT is recorded for the majority of sites in ForC, whereas other variables are extracted from a global database.
+We find that temperature-related climate variables ($MAT$, temperature seasonality, ...**LIST**) explain the highest proportion of variability in productivity, and among these, $MAT$ is generally the best predictor--perhaps because site-specific MAT is recorded for the majority of sites in ForC, whereas other variables are extracted from global gridded data products (Table S1**?**).
 ....
@@
 
-[ Regarding seasonality, FACF and climate data lack the temporal resolution required to sort this out.
-We find no support for the argument that correcting for growing season length improves the explanatory power.]
+FACF is reduced by climatic seasonality (Fig.
+5e), and productivity is shut down during cold- or dry- dormant seasons.
+To account for this, a number of analyses seeking to characterize global-scale effects of climate on productivity have examined the relationship of FACF per month of the growing season with growing season climatic conditions (**REFS**).
+We found that the sort of simple metric needed to define growing season at a global scale was uncertain for temperature and problematic for moisture (**WORK ON THIS**).
+A temperature-defined growing season length had stong positive correlation with FACF (Fig.
+5f), but explained less variation than $MAT$.
+Dividing FACFs by growing season length to yield FACF per growing season month removed the majority of climate-related variation, supporting the idea that the latitudinal gradient in FACF is attributable more to shorter growing seasons at high latitudes than to inherently lower rates of photosyntheiss or respiration by high-latitude forests *[ Enquist GCB review]*.
+However, there remained a number of significant correlations with growing season climatic conditions.
+We conclude that correcting for growing season month does little to improve statistical predictive power.
+WHile
 

### 2019-09-12 manuscript/Discussion_KAT.Rmd
Update Discussion_KAT.Rmd

@@
 We find that temperature-related climate variables ($MAT$, temperature seasonality, ...**LIST**) explain the highest proportion of variability in productivity, and among these, $MAT$ is generally the best predictor--perhaps because site-specific MAT is recorded for the majority of sites in ForC, whereas other variables are extracted from global gridded data products (Table S1**?**).
-....
-The effects of temperature are modified by moisture availability, ...
+The effects of temperature are modified by moisture availability, with reduced FACF under hot and dry conditions (*i.e.*, high PET, high deficit; Fig.
+5c-d) and sometimes under very high precipitaiton (Figs.
+4, 5b).
+Negative effects of very high precipitation on FACF have been previously observed (**REFS**) and are attributable to nutrient and light limitations (**REFS**) Thus, although temperature and water availability jointly and interactively drive global-scale patterns of FACF.
 
@@
 5f), but explained less variation than $MAT$.
-Dividing FACFs by growing season length to yield FACF per growing season month removed the majority of climate-related variation, supporting the idea that the latitudinal gradient in FACF is attributable more to shorter growing seasons at high latitudes than to inherently lower rates of photosyntheiss or respiration by high-latitude forests *[ Enquist GCB review]*.
-However, there remained a number of significant correlations with growing season climatic conditions.
-We conclude that correcting for growing season month does little to improve statistical predictive power.
-WHile
+Dividing FACFs by growing season length to yield FACF per growing season month removed the majority of climate-related variation, supporting the idea that the latitudinal gradient in FACF is attributable more to shorter growing seasons at high latitudes than to inherently lower rates of photosyntheiss or respiration by high-latitude forests (*[ Enquist GCB review]*).
+However, there remained a number of significant correlations with growing season climatic conditions, suggesting that climatic conditions remain influential within the growing season.
+We conclude that while correcting for growing season length takes analyses a step closer to mechanistic linkage of instantaneous C flux rates to environmental conditions, it remains very crude relative to the the timescales on which climate affects plant metabolism and does not advance statistical predictive power.
+Rather, mechanistic accounting for climatic effects on global productivity patterns requires models representing physiologically meaningful timescales (*e.g.*, **refs**).
 
-[Despite the coarse resolution of this analysis, it holds some important implications for our understanding of how forest carbon cycling relates to climate and, by extension, how it is likely to respond to climatic warming.
+**[needs strong concluding paragraph]**
+[Our analysis holds some important implications for our understanding of how forest carbon cycling relates to climate and, by extension, how it is likely to respond to climatic warming.
 We see no evidence that there are major differences in how the various components of forest autotrophic carbon cycling relate to climate or, by extension, any distinct trends in C allocation across the global scale (Fig.
@@
 ]
-
-~~~
-
-Furthermore, we expect that temporal resolution of climate data affects our results.
-Productivity is known to vary on a seasonal and annual scale in response to variation in climate.
-Our analysis relies on long-term averages in both climate variables and forest productivity estimates, and as a result averages out a large proportion of this variation.
-Therefore we recognise that the strong short-term effects of climate on productivity are likely to be masked.
-The limitations of low temporal resolution are highlighted in our analysis of growing season length.
-We note that our estimate of growing season length is imprecise, being based on monthly temperature, precipitation, and evapotranspiration averages.
-With accurate stand-level measurements of growing season length, based on daily, not monthly, climate measures, we would expect to record a stronger correlation between growing season length and productivity.
-
-There is strong evidence from our analyses that this increase in productivity towards the tropical regions is driven primarily by temperature-related effects.
-We find that temperature-related climate variables explain the highest proportion of variability in productivity.
-In contrast, we find that MAP is not a significant predictor of productivity, but does show a strong interaction with temperature.
-This indicates that the main mechanism through with precipitation influences productivity is in a secondary capacity, with its overall influence being mediated by the effects of temperature.
-This also suggests that although temperature explains a large proportion of variation, it cannot fully explain productivity alone, and a measure of water availability is necessary to fully understand patterns of productivity.
-This is likely to be significant in the future, as it suggests that the effects of a warming climate will be mediated through changes in precipitation regime.
-
-There is debate over whether productivity in regions with higher MAT is higher because productivity inherently increases with air temperature during growing season months, or because higher MAT is correlated with longer growing seasons.
-Our analysis suggests that length of growing season is a key factor in understanding latitudinal patterns in productivity.
-We find that length of growing season is strongly correlated with productivity.
-This is supported by the strong correlation between temperature seasonality and productivity, indicating that aseasonality and year-round growing seasons have a strong influence over productivity.
-
-These conclusions are supported by our analyses of the influence of climate within the growing season.
-We find much weaker correlations between climate variables and productivity when only growing season months are considered.
-There is evidence for a positive correlation between productivity and temperature for ANPP and ANPP~foliage~, suggesting that there is still a small positive effect of air temperature on productivity independent of growing season, however this effect is not significant for the remaining fluxes.
-This suggests that the primary positive effect of temperature on productivity is through the lengthening of the growing season in regions with higher MAT.
-However, we note that length of growing season is a less good predictor of productivity than MAT.
-As discussed above, this may be an artefact of imprecise estimation of length of growing season, however it does indicate that there are aspects of productivity explained by MAT which length of growing season alone cannot explain.
-
-**[needs strong concluding paragraph]**
 

### 2019-09-12 manuscript/Discussion_KAT.Rmd
Update Discussion_KAT.Rmd

@@
 
-**[needs strong concluding paragraph]**
-[Our analysis holds some important implications for our understanding of how forest carbon cycling relates to climate and, by extension, how it is likely to respond to climatic warming.
-We see no evidence that there are major differences in how the various components of forest autotrophic carbon cycling relate to climate or, by extension, any distinct trends in C allocation across the global scale (Fig.
+Our analysis clarifies the big picture of how FACF varies with latitude and climate on a global scale, with some important implications for how forest carbon cycling relates to climate and, by extension, how it is likely to respond to climatic warming.
+Contrary to some previous suggestions, we find no support for non-linear trends in mature forest C cycling with respect to latitude or $MAT$, and no distinct trends in C allocation across the global scale (Fig.
 3).
-]
+The implication is that under warmer conditions with similar moisture availability--and within the temperature range to which forest communities are adapated and acclimatized--higher temperatures result in a generalized acceleration of FACF, with no major shifts in C allocation among subsidiary fluxes.
+Of course, actual climatic changes will result in very different sets of conditions than represented across geographic gradients in climate, but our analysis clarifies the big picture as to how carbon cycles through forest ecosystems that have developed under relatively stable climatic conditions.
+As we enter a period of accelerating climatic change, understanding of the fundamental climatic controls on FACF sets a foundation for understanding patterns of change.
+**[work on this]**
 
@@
 ForestGEO
-DOE database
+Compilatin of the ForC database was originally funded by DOE

### 2019-09-12 manuscript/appendix.Rmd
Create appendix.Rmd

--- 2019-09-11 manuscript/manuscript draft.Rmd
@@
 ---
-title: "Manuscript draft"
-author: "Rebecca Banbury Morgan, Valentine Herrmann, Norbert Kunert, Ben Bond-Lamberty, Helene Muller-Landau, [MORE??-CHECK], Kristina J. Anderson-Teixeira"
+title: "appendix"
 output: 
   pdf_document:
-    citation_package: natbib
     fig_caption: yes
-    includes:
-      after_body: appendix.md
-bibliography: [library.bib,  packages.bib]
-biblio-style: apalike
-csl: apa.csl
+
 header-includes:  \usepackage{float}
@@
                   \captionsetup[figure]{font=footnotesize}
-
 ---
@@
 ```
-
-```{r include=FALSE}
-knitr::write_bib(c(
-  .packages(), 'knitr', 'rmarkdown', 'lme4', 'MuMIn', 'plyr', 'raster', 'ncdf4', 'Hmisc', 'plyr', 'merTools', 'visreg', 'r2glmm', 'nlme', 'viridis', 'ggplot2', 'ggpubr', 'gridExtra', 'AICcmodavg'
-), 'packages.bib')
-```
-
-```{r results="asis", include=FALSE}
-cat("
-<style>
-caption {
-      font-size: 0.75em;
-    }
-</style>
-")
-```
-
-###Abstract
-
-Globally, forests are a significant carbon sink, showing high levels of productivity and representing an important component of the carbon cycle.
-Current knowledge indicates that forest productivity decreases with latitude; however, there is evidence that this relationship varies across different fluxes.
-Furthermore, the climate variables that drive this relationship are poorly understood.
-Here, we use a comprehensive global database of forest carbon fluxes to show that, while all major carbon fluxes decrease with latitude, allocation to woody productivity increases with latitude, and allocation to root productivity declines.
-The best predictors of global patterns of productivity are temperature variables, including mean annual temperature, annual temperature range, and temperature seasonality, with the exception of woody productivity, which is most strongly influenced by potential evapotranspiration and vapour pressure deficit.
-Climate variables explain a large proportion of the variation in major carbon fluxes, but are a less significant predictor of subsidiary components.
-Our results illustrate the strong influence of climate on primary productivity, and particularly the importance of temperature in determining forest productivity.
-However, they also indicate that effects of climate are complex: fluxes show non-linear responses to climate variables, and many factors other than climate interact to determine allocation to subsidiary components.
-Knowledge of the relative influence of climate variables on forest carbon fluxes is an important component of understanding the responses of primary productivity to climate change.
-
-### Introduction
-
-Carbon fixation, allocation, and metabolism by trees sets the basis for energy and material flows in forest ecosystems and defines their interactions with Earth's changing climate.
-Forest autotrophic carbon flux (FACF) influences all organic matter stocks in ecosystems and is linked to cycling of energy, water, and nutrients (**REFS**).
-Forest productivity sets the energy available to heterotrophic communities, in turn influencing their abundance (**REFS**) and possibly diversity (**REFS**)--but probably not individual metabolic rate (**Anderson & Jetz 2005 (my first paper!!), but check more recent literature**).
-On the global scale, forests play a critical role in regulating atmospheric CO_2 and climate (**Bonan**).
-The total amount of CO_2 cycling through Earth's forests each year (*i.e.*, total gross primary production, $GPP$) is more than five times anthropogenic fossil fuel emissions, and net sequestration (~2.4 GtCyr^-1^) offsetting roughly 30% of these emissions [@pan_large_2011].
-As atmospheric carbon dioxide levels continue to rise, with consequences for global climate, there is increasing recognition that proper protection and management of forest resources will have an important role to play in mitigating climate change (** REFS- e.g., IPCC 1.5C and land reports**).
-Given the importance of understanding broad-scale patterns in forest autotrophic carbon flux (FACF), there have been many meta-analyses on the theme.
-However, our ability to draw general macroscopic conclusions regarding global variation in multiple productivity variables with respect to climate has been limited in that these analyses often mix forests that vary in stand age, disturbance history, and/or management status; do not always sufficiently parse related variables (*e.g.*, combining net primary productivity records with and without belowground components); and typically consider only one or a few variables at a time.
-The recent development of a global forest carbon database synthesizing multiple variables and including records of stand history (ForC; [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc:_2018]) opens up the possibility for a standardized analysis of global scale variation in multiple components of forest productivity and the principle climatic drivers of these patterns.
-
-On a global scale, the productivity of forests varies with latitude, showing a general trend of decreasing productivity with latitude [@beer_terrestrial_2010; @jung_global_2011].
-Studies agree that productivity is lowest in the boreal regions, and increases into the temperate regions [@luyssaert_co_2007; @huston_global_2009; @beer_terrestrial_2010; @jung_global_2011].
-However, evidence is inconclusive on whether productivity continues to increase into the tropics, or whether it plateaus in temperate regions.
-Evidence for this is further complicated by the fact that different studies use different measures of productivity to explore these relationships.
-For example, modelling of global terrestrial ecosystem gross primary productivity (GPP) through upscaling and calibration of eddy flux measurements indicates that GPP peaks in tropical forests [@beer_terrestrial_2010; @jung_global_2011].
-This is corroborated by analysis of site-level GPP measurements, which appear to reach their highest levels in tropical forests [@luyssaert_co_2007].
-In contrast, there is evidence that the highest values of net primary productivity (NPP) may be found in temperate forests [@luyssaert_co_2007; @huston_global_2009], although other studies find NPP is highest in the tropics, showing a decrease with latitude [@simova_enigma_2017].
-Other studies have chosen to focus exclusively on above-ground net primary productivity (ANPP), finding evidence of a weak negative relationship between ANPP and latitude [@huston_global_2009; @gillman_latitude_2015].
-An understanding of the global patterns of productivity is important; however elucidating the drivers of these patterns will be even more valuable in expanding our understanding of global carbon cycling.
-Primary productivity can be influenced by many factors, which often act across a range of scales, and may show interactive effects with each other.
-On a local scale, stand age [@litton_carbon_2007; @gillman_latitude_2015], management [@simova_enigma_2017]; nutrient availability [@aragao_above-_2009]; and altitude [@girardin_net_2010; @malhi_variation_2017] all impact forest productivity.
-On a global scale, we expect that productivity is most strongly influenced by broad climatic gradients.
-
-We know from previous research that climate is a significant driver of productivity across broad spatial scales [@cleveland_relationships_2011].
-The majority of studies have focused on exploring the relationships between productivity and mean annual temperature (MAT) and mean annual precipitation (MAP), as the most commonly reported site-level climate variables.
-These variables have the advantage that they describe broad trends in temperature and water availability, and therefore capture a lot of global-scale variation in climate.
-There is strong evidence that both MAT and MAP show significant positive relationships with productivity [@chu_does_2016].
-However, as with latitude, the shape of those relationships is not always clear, and, again, is complicated by the use of different measures of productivity across studies.
-There is support across multiple studies for the hypothesis that various measures of primary productivity saturate at high levels of MAP, though the saturation points identified vary from 1500mm [@luyssaert_co_2007] up to 2445mm MAP [@schuur_productivity_2003].
-Studies of the influence of MAT on productivity are less conclusive.
-Luyssaert et al.
-[-@luyssaert_co_2007] examined GPP and NPP and found that, while GPP increases linearly with MAT, NPP saturates at around 10$^\circ$C MAT.
-In contrast, Larjavaara and Muller-Landau [-@larjavaara_temperature_2012], find that increases in GPP saturate at approximately 25$^\circ$C MAT, while Schuur [-@schuur_productivity_2003] shows that NPP increases linearly with temperature.
-The influence of these climate variables on productivity is further complicated by the possibility of interactive effects occurring between them.
-Taylor et al.
-[-@taylor_temperature_2017] showed that increased MAP had a negative effect at low MAT, but a positive effect at high levels of MAT, and vice versa, suggesting that both temperature and water availability are important in explaining productivity.
-
-While MAT and MAP are important climate variables, they do not capture all aspects of climate that may influence productivity, and may be insufficient to capture the breadth of climatic effects on productivity [@cleveland_relationships_2011].
-There is evidence that productivity also responds to variables such as cloud cover [@taylor_temperature_2017], solar radiation [@fyllas_solar_2017], and potential evapotranspiration [@kerkhoff_plant_2005] in potentially significant ways.
-Furthermore, MAT and MAP are very coarse measures of climate, and so fail to capture much variation in climate on an intra-annual scale, including the effects of factors such as growing season length, number of frost-free days, temperature seasonality, and dry season length.
-Some studies have suggested that the apparently strong relationship between MAT and productivity is actually a factor of the correlation between MAT and growing season length [@kerkhoff_plant_2005; @malhi_productivity_2012; @michaletz_convergence_2014; @michaletz_drivers_2018].
-Kerkhoff et al.
-[-@kerkhoff_plant_2005] and Michaletz et al.
-[-@michaletz_convergence_2014] find that, within the growing season, there is no significant relationship between productivity and MAT, indicating that the effect of temperature is due to increased length of growing season, rather than an inherent influence of temperature on productivity.
-
-In order to approach these broad and complex issues, we simplify the major gaps in our knowledge to five key questions and corresponding specific predictions (Table 1).
-Firstly, we ask how forest autotrophic carbon fluxes (FACF) vary with latitude.
-We then test how these fluxes relate to MAT and MAP, and additionally how they respond to other, less well studied, climate variables.
-We finally consider the relationship between FACF and seasonality, considering the role of seasonality in explaining variation in carbon fluxes, and the influence of climate on FACF per month of the growing season.
-
-Resolving these questions requires data which, firstly, control for variables influencing productivity, including stand age, disturbance regime, elevation, and methodology; and secondly includes all major carbon fluxes.
-Here we use a comprehensive global database of forest carbon fluxes to explore the above questions for nine carbon fluxes, allowing for an in-depth exploration of the effect of climate on global productivity.
-
-\renewcommand{\arraystretch}{2}
-```{r eval = TRUE, echo=FALSE, warning=FALSE}
-library(knitr)
-library(kableExtra)
-hypothesis_table <- read.csv("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/hypothesis_table.csv", stringsAsFactors = FALSE, check.names = FALSE)
-kable(hypothesis_table, booktabs = TRUE, caption = "Hypothesis table") %>%
-  kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 12) %>%
-  pack_rows("1. How do FACF vary with latitude?", 1, 3, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  pack_rows("2. How do FACF relate to MAT and MAP?", 4, 6, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  pack_rows("3. How do FACF relate to other annual climate variables?", 7, 9, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  pack_rows("4. What is the role of seasonality?", 10, 13, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  pack_rows("5. Per month of the growing seaon, how do FACF vary with climate?", 14, 17, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  add_header_above(c(" ", " ", " ", "Forest autotrophic carbon fluxes (FACF)" = 9, " ")) %>%
-  column_spec(1, width = "10cm") %>%
-  column_spec(2, width = "5cm") %>%
-  column_spec(3:13, width = "1.8cm") %>%
-  kableExtra::landscape()
-```
-
-### Materials and Methods
-
-Analyses were conducted on data contained in the open-access ForC database [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc:_2018].
-This database contains records of field-based measurements of forest carbon stocks and annual fluxes, compiled from original publications and existing data compilations and databases.
-Associated data, such as stand age, measurement methodologies, and disturbance history, are also included.
-The database was significantly expanded since the publication of [@anderson-teixeira_forc:_2018] through integration with the Global Database of Soil Respiration [@bond-lamberty_global_2010].
-Additional targeted literature searches were conducted to identify any further available data on primary productivity, with particular focus on mature forests in temperate and boreal regions.
-ForC currently contains 29730 records from 4979 plots, representing 20 distinct ecozones across all forested biogeographic and climate zones.
-We used ForC v3.0, archived on Zenodo with DOI 10.5281/zenodo.3403855.
-
-```{r echo=FALSE, out.width='100%', fig.cap = "Map showing all data used in the analysis, coded by variable", fig.pos='H'}
-  knitr::include_graphics("distribution_all_variables.png")
-```
-*Data selection.* Over 50 variables of forest carbon stocks and annual fluxes are represented in the ForC database; this analysis focussed on measures of primary productivity.
-Table 1 contains details of the variables selected for analysis.
-
-| Variable | Definition | Components included | Methodologies used | Number of records | Number of geographic areas |
-|--------------|-------------------|-------------|------------------------|-------|-------|
-| GPP | Annual gross primary production; annual uptake of carbon dioxide by an ecosystem | NA | Flux partitioning of eddy covariance | 243 | 49 |
-| NPP | Annual net primary production; the component of GPP that is stored in plant tissue; GPP minus ecosystem respiration | Foliage, branch, stem, coarse root, fine root and optionally understory | Direct measurement of annual increments of components | 92 | 42 |
-| ANPP | Aboveground net primary production | Foliage, stem, and optionally branch | Direct measurement of annual increments of components | 256 | 80 |
-| ANPP~foliage~ | Net primary production of foliage | Foliage | Direct measurement of litterfall, correcting for changes in leaf biomass when measured | 98 | 49 |
-| ANPP~woody~ ~stem~ | Net primary production of woody stems | Woody stems | Direct measurement of stem growth increment | 264 | 96 |
-| BNPP~root~ | Belowground net primary production | Coarse and fine roots | Direct measurement of one or more of: fine root turnover, soil cores, root ingrowth cores, minirhizotrons; indirect estimates of coarse roots using allometries based on aboveground stem increment measures | 101 | 48 |
-| BNPP~fine~ ~root~ | Net primary production of fine roots | Fine roots | Direct measurement of one or more of: minirhizotrons, fine root turnover, soil cores, root ingrowth cores | 88 | 41 |
-| R~auto~ | Annual autotrophic respiration, including above- and belowground components | Foliage, stem, and root | Chamber measurements of component gas exchange | 22 | 13 |
-| R~root~ | Annual root respiration | Coarse and fine roots | Measurement of root gas exchange | 64 | 26 |
-
-Table: Definitions and sample sizes of variables used in analysis.
-Geographic areas group geographically proximate sites, defined using a hierarchical cluster analysis on the distance matrix of the sites, and a cutoff of 25km.
-
-A subset of the ForC database was generated for the purposes of this analysis, in order to control for data quality and remove biasing factors.
-Since management can alter observed patterns of primary productivity [@simova_enigma_2017], sites were excluded from analysis if they were managed, defined as plots that were planted, managed as plantations, irrigated, fertilised or including the term "managed" in their site description.
-Sites that had experienced significant disturbance were also excluded.
-Disturbances that justified site exclusion were major cutting or harvesting, and/or burning, flooding, drought and storm events with site mortality >10% of trees.
-Grazed sites were retained.
-There is evidence that stand age influences patterns of primary productivity and carbon allocation in forest ecosystems, and can confound relationships between latitude and primary productivity [@de_lucia_forest_2007; @gillman_latitude_2015].
-To reduce any biasing effects of stand age, stands under 100 years of age were excluded from analysis.
-Sites for which stand age was unknown were excluded from analysis.
-
-*Methodological consistency.* The data in ForC is derived from a range of studies, often employing different methods.
-For this reason, criteria were introduced to standardise for differences in methodology.
-Where data was based on forest plot census measurements, studies which used a minimum diameter at breast height (DBH) measure of >10cm were excluded from analysis.
-It would be preferable to standardise by minimum area sampled; however x% of plots in the database are 1 ha or under in size; excluding these plots would place significant constraints on sample size.
-
-As discussed above, estimates of NPP, ANPP, and BNPP are generated through summing measurements of their component parts.
-Since the components included in productivity estimates vary between studies, estimates of productivity were classified within the ForC database according to their components, and then filtered for analysis.
-Estimates of NPP were selected if they included foliage, branch, stem, coarse root, and fine root.
-Measures of NPP which included additional components, including understorey, volatile organic compounds (VOCs), exudates, estimates of NPP lost to herbivory, and the NPP of reproductive structures, were excluded.
-Estimates of ANPP were selected if they included foliage, stem growth and optionally branch turnover.
-Any measures of primary productivity where components were unknown were excluded from analysis.
-
-*Climate datasets.* Where site-level data on mean annual temperature, mean annual precipitation, and latitude were available in the primary literature, this data was compiled and entered directly into the ForC database.
-Based on the geographic co-ordinates for each site, data on a further 11 climate variables was extracted from five open-access climate datasets: WorldClim [@hijmans_very_2005], WorldClim2 [@fick_worldclim_2017], the Climate Research Unit (CRU) time-series dataset v.
-4.03 [@harris_updated_2014], the Global Aridity Index and Potential Evapotranspiration Climate Database [@trabucco_global_2018], and TerraClimate [@abatzoglou_terraclimate_2018] (see Supplementary Information S1 for details of climate variables).
-Where site-level data was missing for mean annual temperature and/or mean annual precipitation, data was extracted from the WorldClim dataset.
-
-Additionally, two climate variables were derived from the above datasets: maximum vapour pressure deficit, defined as the vapour pressure deficit of the month with the largest deficit; and water stress months, defined as the number of months annually where precipitation was lower than potential evapotranspiration.
-
-*Length of growing season.* Growing season months were defined as months with mean minimum temperature > 0.5$^\circ$C.
-Growing season months were initially calculated following methods used by Kerkhoff et al.
-(2005), which additionally required that growing season months had a moisture index, defined as (MAT - PET)/PET, > -0.95.
-Michaletz et al.
-(2014) included an equivalent requirement in their calculation of growing season length.
-However, we found that including this requirement had no effect on the estimates of growing season length, and so chose to exclude it.
-
-Monthly data for PET, precipitation, and temperature was downloaded from the Climate Research Unit (CRU) time-series dataset v 4.03 [@harris_updated_2014], and for solar radiation from WorldClim2 [@fick_worldclim_2017], and used to calculate mean monthly PET, precipitation, temperature and solar radiation during the growing season.
-Total growing season precipitation and solar radiation were also calculated.
-
-*Model specification.* The effects of climate and latitude on primary productivity were analysed using mixed effects models using the package 'lme4' [@bates_fitting_2015] in R v.3.5.1 [@r_core_team_r:_2018].
-The effect of each extracted climate variable on each measure of primary productivity was modelled by specifying the climate variable as a fixed effect.
-For each climate variable, three models were specified: a null model; a model with the climate variable as a linear term; and a model with the climate variable as a polynomial term.
-AIC values were calculated for the models and used to select the best model.
-If the best model included a polynomial term, the shape of the polynomial relationship was considered.
-If the shape of the relationship made biological sense, and was a significant improvement on the linear relationship (deltaAIC >2), we accepted the polynomial as the best model.
-If not, we ran the linear model as the final model.
-R^2^ values were calculated for the best model.
-All R^2^ values presented here are marginal R^2^ values, and refer to the proportion of variation explained by only the fixed effects, unless otherwise specified.
-In addition, slope coefficients were calculated for the linear models.
-
-Because the magnitude of fluxes varies significantly, in order to facilitate comparisons between regression models for each flux, data for each flux was scaled, to give the data a mean of 0 and standard deviation of 1.
-As each data set was scaled separately, this does not allow for statistical comparisons of slope values, but does assist in visualising the data.
-
-To test for a potential influence of altitude, models were also run with site altitude included as a second fixed effect.
-These models were compared against models with no altitude term, and AIC values calculated to identify whether inclusion of altitude as a term improved the models.
-Including altitude had a very small effect on most models, with the difference in AIC values between models including and excluding altitude often being <2, suggesting the models are very similar in their explanatory power.
-As a result, it was decided to present results only from models do not include altitude as a fixed term.
-
-Within the ForC database, sites within 25km^2^ of each other are clustered into geographic areas.
-To account for correlations in measurements between tightly clustered sites, a random effect was specified as plot nested within geographic area.
-Data from the temperate regions was heavily skewed towards studies from the old-growth forests of the Pacific Northwest.
-These forests have very high productivity, and so to ensure that results were not unduly influenced by geographic sampling bias, we tried a version of the model where data were weighted according to forested land area within each Koeppen climate zone.
-Results were similar between the weighted and unweighted model, so, to avoid problems of over-fitting, the weighted model was dropped, and results from this are not presented here.
-
-Models were run for total annual productivity against annual climate variables, and for monthly growing season productivity, defined as total productivity/length of growing season, against growing season climate variables.
-For analyses on data within the growing season, only linear models were specified.
-
-To investigate the potential interactive effects of climate variables on carbon fluxes, multivariate models were also specified.
-To ensure that models were biologically meaningful, the terms included in the models tested built on results from the univariate models.
-Modelling of individual climate variables identified that the best predictors of carbon fluxes were variables related to temperature.
-We therefore decided to include mean annual temperature as a term in all multivariate models.
-We first modelled the interaction effect between mean annual temperature and mean annual precipitation, in order to capture climate variation along the axes of temperature and water availability.
-Models were tested for a significant interactive effect and a significant additive effect.
-We then explored whether any other climate variable, in combination with mean annual temperature, could significantly improve on the combination of mean annual temperature and mean annual precipitation.
-In specifying the range of models to test, climate variables which were strongly correlated with temperature were dropped, in order to capture the greatest range of variation in climate.
-For each possible pairing of climate variables, two models were specified: a model with the two climate variables showing an additive effect; and a model with the two climate variables showing an interactive effect.
-As before, plot nested within geographic area was included as a random effect.
-Altitude was not considered.
-AIC values were calculated for the models, and used to compare models.
-Models were considered to be significantly better than the baseline MAT*MAP model if:
-  i) the AIC value of the model was smaller than the AIC value of the baseline model by >2  
-  ii) the r-squared value of the model was larger than the r-squared value of the baseline model by >5
-
-*Validating models of component fluxes.* Comparison of component fluxes is based on the assumption that components sum accurately to estimates of larger fluxes.
-To test this, components of larger fluxes were regressed against latitude, and the models used to generate a series of point estimates along lines of best fit for each component.
-The point estimates for smaller component fluxes were summed to generate new "stacked" estimates of larger fluxes, which were then compared against actual measurements of the larger flux.
-Confidence intervals for the larger flux were calculated using the 'bootMer' function from the lme4 package [@bates_fitting_2015].
-Stacked plots were generated for:
-1. GPP = NPP + R~auto~  
-2. NPP = ANPP + BNPP  
-3. ANPP = ANPP~foliage~ + ANPP~woody~ ~stem~  
-4. Total belowground carbon flux = BNPP + R~root~
-
-*Allocation to carbon fluxes along latitudinal gradients.* Variation in allocation to component carbon fluxes along latitudinal gradients was explored for a range of pairings: firstly, GPP:NPP, ANPP:BNPP, and ANPP~foliage~:ANPP~woody~ ~stem~; and secondly, the ratio of NPP to each of ANPP, BNPP, ANPP~foliage~, and ANPP~woody~ ~stem~.
-For each set of paired fluxes, measurements taken at the same site and plot, and in the same year, were paired together, and the ratio of each pair of measurements calculated.
-The ratios were regressed against latitude and climate variables, using the linear model specified above.
-Cook's distance analyses were carried out for each of the models, and indicated that data from a few high-elevation sites were having a disproportionate influence on the regressions.
-To account for this, models were re-run using only data from sites $\leq$ 1000m.
-
-###Results
-
-In total, we analyzed 1228 records from 9 C flux variables taken from forests that had experienced no major anthropogenic disturbances within the past 100 years.
-These records represented a total of 154 distinct geographic areas (Fig.
-1, Table 2), across all forested biogeographic and climate zones.
-
-```{r echo=FALSE, out.width='100%', fig.cap = "Graphs of primary productivity $(MgC~ha^-1~yr^-1)$ regressed against latitude.
-Lines of best fit are plotted according to the best model selected during analysis.
-All regressions are significant $(p<0.05)$.
-Plots 1 - 3 show two component fluxes; a larger flux, defined as the combination of the two component fluxes; and a modelled estimate of the sum of the two component fluxes.
-95\\% confidence intervals are plotted for the larger flux.
-Plot 4 shows three belowground fluxes, and a modelled estimate of the total belowground carbon flux", fig.pos='H'}
-
-  knitr::include_graphics("combined_stacked.png")
-```
-
-*How does productivity vary with latitude?*
-
-All major carbon fluxes increased linearly with decreasing latitude (fig. 3).
-
-```{r echo=FALSE, out.width='100%', fig.cap = "Graphs to show primary productivity $(MgC~ha^-1~yr^-1)$ regressed against latitude.
-Lines of best fit are plotted according to the best model selected during analysis.
-All regressions are significant $(p<0.05)$.", fig.pos='H'}
-  knitr::include_graphics("effect_of_lat_transparent.png")
-```
-
-Latitude was a strong predictor for many of the carbon fluxes, explaining 64% of variation in GPP (n = 254, p<0.0001), 50% in NPP (n = 114, p<0.0001) and 45% in ANPP (n = 259, p<0.0001).
-For all fluxes, their relationship with latitude was best predicted by the linear model.
-
-*Relationships and differences among fluxes.* In general, smaller component fluxes summed approximately to  larger fluxes across the latitudinal gradient (fig.
-2).
-That is, modelled estimates of GPP, generated from the sum of NPP and R auto; NPP, generated from the sum of ANPP and BNPP~root~; and ANPP, generated from the sum of ANPP~foliage~ and ANPP~woody~ ~stem~, fell completely within the confidence intervals of the regressions of field estimates of GPP, NPP and ANPP respectively.
-
-We found no evidence that allocation between fluxes varied substantially with latitude or climate.
-There were no significant results from regressing ratios of carbon fluxes against latitude, or against any of the climate variables.
-
-R^2^ values were generally highest in the major fluxes, and decreased in subsidiary fluxes (Supporting Information S2).
-Of the major fluxes, R~auto~ and GPP were the most strongly explained by latitude and climate, with climate explaining at most 71% of variation in GPP, and 65% in R~auto~.
-The proportion of variation explained by climate and latitude decreased in NPP and ANPP, with climate explaining at most 51% of variation in NPP and 44% in ANPP.
-Of the major fluxes, BNPP~root~ was the least well explained by climate and latitude, with climate explaining at most 36% of variation.
-
-With the exception of ANPP~foliage~, the proportion of variation explained by climate and latitude in subsidiary fluxes was much lower.
-Climate explained at most 24% of variation in ANPP~woody~ ~stem~, 19% in BNPP~fine~ ~root~, and 27% in R~root~.
-In contrast, climate strongly explained variation in ANPP~foliage~, with mean annual temperature explaining 58% of variation.
-This pattern was also seen in the R^2^ values for multivariate models.
-
-*How does productivity relate to MAT and MAP?* We focus first on considering the relationship between productivity and MAT and MAP.
-MAT and MAP are the most commonly reported site-level climate variables, and much previous research into the effect of climate on forest productivity has focused on these as key climate variables.
-MAT was a significant (p<0.05) and strong predictor of productivity for all carbon fluxes tested, with all fluxes showing a linear increase with temperature (fig.
-5).
-We found no support for a saturation point of productivity with temperature.
-
-MAP was found to be a significant (p<0.05) but poor predictor of productivity, explaining, with the exception of R~auto~, at most 37% of variation in carbon flux.
-For the majority of fluxes productivity was best predicted by a polynomial model.
-Productivity increased with precipitation, up until a saturation point at between 3000 and 4000mm annual precipitation, above which productivity started to decrease (fig.
-5).
-The notable exception to this was GPP: the model indicated that GPP continued to increase with precipitation up to measures of at least 5000mm annually (p<0.0001, R^2^ = 0.33.
-Data above this point was not available, but the model trend indicated that the saturation point for this model would be around 5000mm MAP.
-
-There was a significant interactive effect between MAT and MAP for GPP, BNPP~root~, BNPP~fine~ ~root~, ANPP, ANPP~woody~ ~stem~, and R~root~ (fig.4).
-There was a significant additive effect for R~auto~.
-NPP and ANPP~foliage~ showed no significant interactive or additive effect: including MAP as a second variable did not improve on the model including only MAT.
-
-For the variables which showed a significant interactive or additive effect between MAT and MAP, no other climate variable, in combination with MAT, significantly improved on that model.
-For NPP, there was a significant interactive effect between MAT and water stress months, with this model explaining nearly 5% more variation in NPP than MAT alone.
-However, for ANPP~foliage~, no multivariate model improved on the univariate model including only MAT.
-
-```{r echo=FALSE, out.width='100%', fig.cap = "Plots of primary productivity $(MgC~ha^-1~yr^-1)$ regressed against mean annual precipitation.
-Points are grouped into bins of 0 - 1000, 1001 - 2000, 2001 - 3000, and >3000mm mean annual precipitation, and lines of best fit plotted for mean annual precipitation values of 500, 1500, 2500, and 3500mm.
-All regressions are significant $(p<0.05)$.", fig.pos='H'}
-  knitr::include_graphics("mat_map_interaction.png")
-```
-
-*How does productivity relate to other climate variables?* Our results indicated that productivity was most strongly explained by temperature at the global scale, with temperature-related climate variables coming out as strong predictors of productivity.
-In addition to MAT, temperature seasonality, annual temperature range, and annual frost days were consistently identified as good predictors of productivity across fluxes.
-
-We found a significant relationship between productivity and potential evapotranspiration for all fluxes.
-ANPP~foliage~, BNPP~fine~ ~root~ and R~root~ increased linearly with PET; however, all other fluxes showed a polynomial relationship with PET (fig.
-5).
-We found strong evidence for a saturation point or peak with PET: productivity tended to increase at values below 1000mm, before saturating between 1200 and 1700mm.
-There was evidence that productivity begins to decrease at values above 1800mm PET.
-
-Vapour pressure deficit was a significant predictor of productivity for all fluxes.
-BNPP~fine~ ~root~ showed a linear relationship with vapour pressure deficit (R^2^ = 0.07, p<0.05), but all other fluxes showed a polynomial relationship (fig.
-5).
-Productivity initially increased with vapour pressure deficit, before saturating at around 0.8 kPa.
-At values above 0.8 kPa, productivity began to decrease.
-
-All fluxes, with the exception of R~root~, showed a positive linear relationship with solar radiation.
-Solar radiation explained a low proportion of variability in productivity for all fluxes, explaining less than 20% of the variation in each flux, with the exception of R~auto~ (R^2^ = 0.26, p<0.05).
-
-Of the climate variables tested, annual wet days, aridity, cloud cover, mean diurnal temperature range, precipitation seasonality, maximum vapour pressure deficit and water stress months were poor or non-significant explainers of variation in productivity, explaining less than 20% of the variation in each of the carbon fluxes.
-
-```{r echo=FALSE, out.width='100%', fig.cap = "Plots of primary productivity $(MgC~ha^-1~yr^-1)$ regressed against (a) mean annual temperature; (b) mean annual precipitation; (c) potential evapotranspiration, (d) vapour pressure deficit; (e) temperature seasonality; (f) length of growing season.
-Lines of best fit are plotted according to the best model selected during analysis.
-All regressions are significant $(p<0.05)$.", fig.pos='H'}
-  knitr::include_graphics("combined_plots.png")
-```
-
-*What is the role of seasonality in explaining productivity?* Temperature seasonality was a significant predictor of productivity.
-We found a polynomial relationship with productivity for GPP, NPP, ANPP, and R~root~.
-Productivity decreased rapidly as seasonality increased, with the rate of decrease slowing as seasonality increased (fig.
-5).
-ANPP~foliage~, ANPP~woody~ ~stem~ and R~auto~ decreased linearly with temperature seasonality.
-Temperature seasonality was strongly correlated with annual temperature range, and, as expected, all fluxes showed almost identical responses to it.
-Productivity was highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower.
-
-In contrast, there was no significant effect of precipitation seasonality on productivity.
-
-We found a significant relationship between length of growing season and productivity, with all fluxes showing a linear increase in productivity with length of growing season (fig.
-5).
-Length of growing season was a strong predictor of productivity, explaining 51% of variation in GPP, 39% of variation in NPP, and 34% of variation in ANPP, but it was not identified as a stronger predictor than MAT for any of the fluxes analysed.
-
-*Within the growing season, how does productivity vary with climate?* Within growing season months, we found that climate had a much weaker effect on productivity.
-For each of temperature, precipitation, PET, and solar radiation, we found a small effect of climate for certain carbon fluxes.
-There was a small increase in productivity with temperature and precipitation for ANPP (with temperature R^2^ = 0.10, p<0.001; with precipitation R^2^ = 0.04, p<0.05) and ANPP~foliage~ (with temperature R^2^ = 0.16, p<0.01; with precipitation R^2^ = 0.09, p<0.05).
-Productivity increased with solar radiation for GPP (R^2^ = 0.21, p<0.001), NPP (R^2^ = 0.21, p<0.001), BNPP (R^2^ = 0.16, p<0.001) and BNPP~fine~ ~root~ (R^2^ = 0.12, p<0.01), and with PET for GPP (R^2^ = 0.15, p<0.01), NPP (R^2^ = 0.18, p<0.01), BNPP (R^2^ = 0.23, p<0.0001), BNPP~fine~ ~root~ (R^2^ = 0.11, p<0.05), and ANPP~woody~ ~stem~ (R^2^ = 0.06, p<0.05).
-
-###Discussion
-
-In this analysis we used a comprehensive global database, containing an unprecedented amount of data and representing all global forest ecosystems and all significant forest carbon fluxes, to comprehensively explore the relationships between climate and productivity on a global scale.
-Many of our findings support and clarify previously published work; however we used a much larger and more complete database than has previously been available, allowing for higher quality control and standardisation of data.
-This enabled us to control for factors not previously controlled for - such as stand age, methodology, flux components, and disturbance regime - to gain a stronger understanding of latitudinal and climatic effects on productivity.
-
-Our results show that productivity decreases linearly with latitude.
-Climate explains a significant proportion of variation in all carbon fluxes, with temperature variables being the best predictors of productivity on this global scale.
-While other climate variables are significant predictors of productivity, none of them improve on the explanatory power of temperature-related variables in general or MAT specifically.
-Water availability is an important factor in explaining productivity on a global scale: we find a positive influence of precipitation at low MAP, with saturation at higher levels of MAP.
-There is a significant interaction between MAT and MAP.
-We note that climate tends to explain a higher amount of variation in the major fluxes, with the amount of variation explained decreasing in smaller component fluxes.
-
-We find that seasonality is an important factor in understanding patterns in productivity on a global scale: temperature seasonality and growing season length are strong predictors of productivity, though growing season length doesn't improve on MAT as a predictor.
-Within the growing season we find the influence of climate on productivity is smaller, but still significant for a number of carbon fluxes.
-
-Despite the high standard of data quality in the ForC database, there are still significant limitations on our analyses.
-We standardised methodologies as far as possible, ensuring that there was consistency in the components included for each carbon flux.
-However, field techniques for estimating components of primary productivity are variable in their accuracy.
-This is particularly true for measures of BNPP, which are often estimated by extrapolating measures of ANPP.
-Even where BNPP is calculated through direct soil sampling, measures are often taken to insufficient depth, leading to underestimation of the flux.
-It is challenging to account for this type of variability in quality of field data, and our analyses largely assume good faith in data quality.
-While we have strong confidence in our analyses of well-understood fluxes, such as GPP, ANPP, and ANPP~woody~ ~stem~, we recommend caution in drawing firm conclusions from our analyses of BNPP, BNPP~fine~ ~root~, and, to some extent, NPP.
-We expect that, as field techniques improve the accuracy of estimates of these fluxes in the future, our understanding of the relationship of BNPP with climate variables will change.
-It is possible that the low R^2^ values for BNPP that we find are a result of lower accuracy in field measures of these fluxes, and that with improve standardisation of methods, we would find climate explains a higher proportion of variation.
-
-Secondly, we emphasise that this study considers the relationship between productivity and climate on a coarse scale.
-Because of the nature of the ForC database, we were unable to access site-level MAT and MAP data for a number of sites, and lacked site-level data for other climate variables for all sites.
-As a result the climate data we used was primarily multi-year averages of annual climate measures accessed from global databases.
-While the quality of these data is high, we cannot expect that it captures detailed local variation in climate, and in particular will be unable to capture microclimate variation, which may have a strong influence on stand-level productivity.
-Although we find that MAT is the strongest predictor for climate, part of this effect may be a result of the higher-quality stand-level dataset available for MAT.
-With high-quality stand-level data for other climate variables we expect that their predictive strength would improve.
-
-Furthermore, we expect that temporal resolution of climate data affects our results.
-Productivity is known to vary on a seasonal and annual scale in response to variation in climate.
-Our analysis relies on long-term averages in both climate variables and forest productivity estimates, and as a result averages out a large proportion of this variation.
-Therefore we recognise that the strong short-term effects of climate on productivity are likely to be masked.
-The limitations of low temporal resolution are highlighted in our analysis of growing season length.
-We note that our estimate of growing season length is imprecise, being based on monthly temperature, precipitation, and evapotranspiration averages.
-With accurate stand-level measurements of growing season length, based on daily, not monthly, climate measures, we would expect to record a stronger correlation between growing season length and productivity.
-
-Although we recognise the inherent limitations present in this analysis, by focusing on mature, undisturbed forests, and by clarifying the components of NPP included in analysis, our study offers a clearer picture of how carbon cycling varies with latitude and climate.
-
-Past studies have differed in their conclusions regarding the relationship between productivity and latitude.
-Our findings indicate that, for all carbon fluxes studied, productivity peaks in the tropical regions.
-There is strong evidence from our analyses that this increase in productivity towards the tropical regions is driven primarily by temperature-related effects.
-We find that temperature-related climate variables explain the highest proportion of variability in productivity.
-In contrast, we find that MAP is not a significant predictor of productivity, but does show a strong interaction with temperature.
-This indicates that the main mechanism through with precipitation influences productivity is in a secondary capacity, with its overall influence being mediated by the effects of temperature.
-This also suggests that although temperature explains a large proportion of variation, it cannot fully explain productivity alone, and a measure of water availability is necessary to fully understand patterns of productivity.
-This is likely to be significant in the future, as it suggests that the effects of a warming climate will be mediated through changes in precipitation regime.
-
-There is debate over whether productivity in regions with higher MAT is higher because productivity inherently increases with air temperature during growing season months, or because higher MAT is correlated with longer growing seasons.
-Our analysis suggests that length of growing season is a key factor in understanding latitudinal patterns in productivity.
-We find that length of growing season is strongly correlated with productivity.
-This is supported by the strong correlation between temperature seasonality and productivity, indicating that aseasonality and year-round growing seasons have a strong influence over productivity.
-
-These conclusions are supported by our analyses of the influence of climate within the growing season.
-We find much weaker correlations between climate variables and productivity when only growing season months are considered.
-There is evidence for a positive correlation between productivity and temperature for ANPP and ANPP~foliage~, suggesting that there is still a small positive effect of air temperature on productivity independent of growing season, however this effect is not significant for the remaining fluxes.
-This suggests that the primary positive effect of temperature on productivity is through the lengthening of the growing season in regions with higher MAT.
-However, we note that length of growing season is a less good predictor of productivity than MAT.
-As discussed above, this may be an artefact of imprecise estimation of length of growing season, however it does indicate that there are aspects of productivity explained by MAT which length of growing season alone cannot explain.
-
-By including multiple fluxes in our analysis, we are able to explore the relationships between fluxes and their responses to climate.
-We find that fluxes are broadly consistent in their responses to climate drivers, suggesting that the importance of climate drivers on productivity is relatively invariant across fluxes.
-In addition, we show that carbon allocation between subsidiary fluxes does not vary significantly across broad climatic gradients.
-There is evidence that allocation between carbon fluxes does vary with factors such as stand age [@litton_carbon_2007], nutrient availability [@litton_carbon_2007; @gill_belowground_2016], forest structure [@taylor_greater_2019], elevation [@moser_elevation_2011], and water availability [@newman_above-_2006].
-That we do not see consistent patterns of variation in allocation across climatic gradients indicates that the pressures influencing shifts in allocation between fluxes occur on local scales, and that shifts in allocation occur as a response to fluctuations in local environmental conditions.
-
-Related to this, we find that climate more strongly explains variation in the major fluxes, with lower levels of variation explained in subsidiary fluxes.
-Although this could in part be related to a lack of standardisation in methodology across studies, as discussed above, it could also indicate that factors other than climate have significant influence over subsidiary fluxes.
-This is consistent with the results we find for allocation between fluxes: although climate has a significant influence on subsidiary fluxes, it is not the only factor that is important, and many other local- and regional-scale factors  influence the way carbon is allocated between subsidiary fluxes.
-We would expect that this would reduce the strength of the effect of climate on subsidiary fluxes.
-
-**[needs strong concluding paragraph]**
-
-### Acknowledgements
-
-### Supporting Information
 

### 2019-09-12 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

--- 2019-09-11 manuscript/manuscript draft.Rmd
@@
   pdf_document:
+    includes:
+      after_body: outline.md
     citation_package: natbib
@@
 ### Supporting Information
-
-```{r eval = TRUE, echo=FALSE, warning=FALSE}
-library(knitr)
-library(kableExtra)
-S1 <- read.csv("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/S1.csv", stringsAsFactors = FALSE, check.names = FALSE)
-kable(S1, booktabs = TRUE, caption = "Hypothesis table", escape = F)
-```

### 2019-09-12 manuscript/manuscript_removed_text.Rmd
Update manuscript_removed_text.Rmd

@@
 Table: R^2^ values presented for latitude, three climate variables, and the best multivariate model for each carbon flux. Values marked * are not significant.
+
+In this analysis we used a comprehensive global database, containing an unprecedented amount of data and representing all global forest ecosystems and all significant forest carbon fluxes, to comprehensively explore the relationships between climate and productivity on a global scale.
+Many of our findings support and clarify previously published work; however we used a much larger and more complete database than has previously been available, allowing for higher quality control and standardisation of data.
+This enabled us to control for factors not previously controlled for - such as stand age, methodology, flux components, and disturbance regime - to gain a stronger understanding of latitudinal and climatic effects on productivity.
+
+Our results show that productivity decreases linearly with latitude.
+Climate explains a significant proportion of variation in all carbon fluxes, with temperature variables being the best predictors of productivity on this global scale.
+While other climate variables are significant predictors of productivity, none of them improve on the explanatory power of temperature-related variables in general or MAT specifically.
+Water availability is an important factor in explaining productivity on a global scale: we find a positive influence of precipitation at low MAP, with saturation at higher levels of MAP.
+There is a significant interaction between MAT and MAP.
+We note that climate tends to explain a higher amount of variation in the major fluxes, with the amount of variation explained decreasing in smaller component fluxes.
+
+We find that seasonality is an important factor in understanding patterns in productivity on a global scale: temperature seasonality and growing season length are strong predictors of productivity, though growing season length doesn't improve on MAT as a predictor.
+Within the growing season we find the influence of climate on productivity is smaller, but still significant for a number of carbon fluxes.
+
+Despite the high standard of data quality in the ForC database, there are still significant limitations on our analyses.
+We standardised methodologies as far as possible, ensuring that there was consistency in the components included for each carbon flux.
+However, field techniques for estimating components of primary productivity are variable in their accuracy.
+This is particularly true for measures of BNPP, which are often estimated by extrapolating measures of ANPP.
+Even where BNPP is calculated through direct soil sampling, measures are often taken to insufficient depth, leading to underestimation of the flux.
+It is challenging to account for this type of variability in quality of field data, and our analyses largely assume good faith in data quality.
+While we have strong confidence in our analyses of well-understood fluxes, such as GPP, ANPP, and ANPP~woody~ ~stem~, we recommend caution in drawing firm conclusions from our analyses of BNPP, BNPP~fine~ ~root~, and, to some extent, NPP.
+We expect that, as field techniques improve the accuracy of estimates of these fluxes in the future, our understanding of the relationship of BNPP with climate variables will change.
+It is possible that the low R^2^ values for BNPP that we find are a result of lower accuracy in field measures of these fluxes, and that with improve standardisation of methods, we would find climate explains a higher proportion of variation.
+
+Secondly, we emphasise that this study considers the relationship between productivity and climate on a coarse scale.
+Because of the nature of the ForC database, we were unable to access site-level MAT and MAP data for a number of sites, and lacked site-level data for other climate variables for all sites.
+As a result the climate data we used was primarily multi-year averages of annual climate measures accessed from global databases.
+While the quality of these data is high, we cannot expect that it captures detailed local variation in climate, and in particular will be unable to capture microclimate variation, which may have a strong influence on stand-level productivity.
+Although we find that MAT is the strongest predictor for climate, part of this effect may be a result of the higher-quality stand-level dataset available for MAT.
+With high-quality stand-level data for other climate variables we expect that their predictive strength would improve.
+
+Furthermore, we expect that temporal resolution of climate data affects our results.
+Productivity is known to vary on a seasonal and annual scale in response to variation in climate.
+Our analysis relies on long-term averages in both climate variables and forest productivity estimates, and as a result averages out a large proportion of this variation.
+Therefore we recognise that the strong short-term effects of climate on productivity are likely to be masked.
+The limitations of low temporal resolution are highlighted in our analysis of growing season length.
+We note that our estimate of growing season length is imprecise, being based on monthly temperature, precipitation, and evapotranspiration averages.
+With accurate stand-level measurements of growing season length, based on daily, not monthly, climate measures, we would expect to record a stronger correlation between growing season length and productivity.
+
+Although we recognise the inherent limitations present in this analysis, by focusing on mature, undisturbed forests, and by clarifying the components of NPP included in analysis, our study offers a clearer picture of how carbon cycling varies with latitude and climate.
+
+Past studies have differed in their conclusions regarding the relationship between productivity and latitude.
+Our findings indicate that, for all carbon fluxes studied, productivity peaks in the tropical regions.
+There is strong evidence from our analyses that this increase in productivity towards the tropical regions is driven primarily by temperature-related effects.
+We find that temperature-related climate variables explain the highest proportion of variability in productivity.
+In contrast, we find that MAP is not a significant predictor of productivity, but does show a strong interaction with temperature.
+This indicates that the main mechanism through with precipitation influences productivity is in a secondary capacity, with its overall influence being mediated by the effects of temperature.
+This also suggests that although temperature explains a large proportion of variation, it cannot fully explain productivity alone, and a measure of water availability is necessary to fully understand patterns of productivity.
+This is likely to be significant in the future, as it suggests that the effects of a warming climate will be mediated through changes in precipitation regime.
+
+There is debate over whether productivity in regions with higher MAT is higher because productivity inherently increases with air temperature during growing season months, or because higher MAT is correlated with longer growing seasons.
+Our analysis suggests that length of growing season is a key factor in understanding latitudinal patterns in productivity.
+We find that length of growing season is strongly correlated with productivity.
+This is supported by the strong correlation between temperature seasonality and productivity, indicating that aseasonality and year-round growing seasons have a strong influence over productivity.
+
+These conclusions are supported by our analyses of the influence of climate within the growing season.
+We find much weaker correlations between climate variables and productivity when only growing season months are considered.
+There is evidence for a positive correlation between productivity and temperature for ANPP and ANPP~foliage~, suggesting that there is still a small positive effect of air temperature on productivity independent of growing season, however this effect is not significant for the remaining fluxes.
+This suggests that the primary positive effect of temperature on productivity is through the lengthening of the growing season in regions with higher MAT.
+However, we note that length of growing season is a less good predictor of productivity than MAT.
+As discussed above, this may be an artefact of imprecise estimation of length of growing season, however it does indicate that there are aspects of productivity explained by MAT which length of growing season alone cannot explain.
+
+By including multiple fluxes in our analysis, we are able to explore the relationships between fluxes and their responses to climate.
+We find that fluxes are broadly consistent in their responses to climate drivers, suggesting that the importance of climate drivers on productivity is relatively invariant across fluxes.
+In addition, we show that carbon allocation between subsidiary fluxes does not vary significantly across broad climatic gradients.
+There is evidence that allocation between carbon fluxes does vary with factors such as stand age [@litton_carbon_2007], nutrient availability [@litton_carbon_2007; @gill_belowground_2016], forest structure [@taylor_greater_2019], elevation [@moser_elevation_2011], and water availability [@newman_above-_2006].
+That we do not see consistent patterns of variation in allocation across climatic gradients indicates that the pressures influencing shifts in allocation between fluxes occur on local scales, and that shifts in allocation occur as a response to fluctuations in local environmental conditions.
+
+Related to this, we find that climate more strongly explains variation in the major fluxes, with lower levels of variation explained in subsidiary fluxes.
+Although this could in part be related to a lack of standardisation in methodology across studies, as discussed above, it could also indicate that factors other than climate have significant influence over subsidiary fluxes.
+This is consistent with the results we find for allocation between fluxes: although climate has a significant influence on subsidiary fluxes, it is not the only factor that is important, and many other local- and regional-scale factors  influence the way carbon is allocated between subsidiary fluxes.
+We would expect that this would reduce the strength of the effect of climate on subsidiary fluxes.
+
+**[needs strong concluding paragraph]**

### 2019-09-12 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

--- 2019-09-11 manuscript/manuscript draft.Rmd
@@
 
-In this analysis we used a comprehensive global database, containing an unprecedented amount of data and representing all global forest ecosystems and all significant forest carbon fluxes, to comprehensively explore the relationships between climate and productivity on a global scale.
-Many of our findings support and clarify previously published work; however we used a much larger and more complete database than has previously been available, allowing for higher quality control and standardisation of data.
-This enabled us to control for factors not previously controlled for - such as stand age, methodology, flux components, and disturbance regime - to gain a stronger understanding of latitudinal and climatic effects on productivity.
+We used a large global database (ForC), containing an unprecedented amount of data representing all major forest types (Fig.
+1) and the nine most significant forest autotrophic carbon flux (FACF) variables (Table 2), to comprehensively explore how C cycling in mature, undisturbed forests varies with latitude and climate on a global scale.
+We test a suite of hypotheses, including many previously posed (Table 1), with more data and standardisation for factors that have not always been controlled for (*e.g.*, stand age, flux components measured) to gain a stronger understanding of climatic effects on FACF.
+We show that, across all nine FACF variables analyzed, C cycling decreases linearly with latitude (Figs.
+2-3)--a finding that confirms multiple previous studies (**REFS**) but contradicts the idea that productivity of temperate forests rivals that of tropical forests (**REFS**).
+The FACF variables increase in proportion to one another, with no differences in allocation detectable at this global scale and with component fluxes summing appropriately to larger fluxes (Fig.
+3), indicating no major, systematic omissions or overestimations of flux components.
+However, we did detect a tendency for greater variability among subsidiary C fluxes (e.g., *$ANPP_woody$*, *$BNPP_{fine root}$*; Fig.
+2; **some SI table?**; **discussed below??**).
+Climate explains a significant proportion of variation in all C fluxes analyzed (albiet less for subsidiary fluxes), with temperature variables being the best predictors of FACF on this global scale (Figs.
+4-5).
+While other climate variables are significant predictors of FACF, none of them improve on the explanatory power of temperature-related variables in general or MAT specifically (Fig.
+5).
+Water availability is an important factor in explaining FACF on a global scale: we find a positive influence of precipitation at low MAP, with saturation at higher levels of MAP (Fig.
+5b) and a significant interaction between MAT and MAP for seven of the nine variables analyzed (Fig.
+4).
+Temperature seasonality and growing season length are closely correlated with MAT and are strong predictors of FACF (Fig.
+5e-f), though growing season length doesn't improve upon MAT as a predictor.
+Within the growing season, the influence of climate on C cycling is smaller but still significant for a number of carbon fluxes (**some SI table?**).
+These findings clarify the big picture of how FACF varies with latitude and climate on a global scale.
 
-Our results show that productivity decreases linearly with latitude.
-Climate explains a significant proportion of variation in all carbon fluxes, with temperature variables being the best predictors of productivity on this global scale.
-While other climate variables are significant predictors of productivity, none of them improve on the explanatory power of temperature-related variables in general or MAT specifically.
-Water availability is an important factor in explaining productivity on a global scale: we find a positive influence of precipitation at low MAP, with saturation at higher levels of MAP.
-There is a significant interaction between MAT and MAP.
-We note that climate tends to explain a higher amount of variation in the major fluxes, with the amount of variation explained decreasing in smaller component fluxes.
+Past studies have differed in their conclusions regarding the relationship between productivity and latitude or its correlates (**REFS**)--quite possibly because of lack of standardization with respect to stand age and disturbance history.
+Our findings indicate that, among mature, undisturbed stands, FACF is unambiguously highest in the tropical regions, and the relationship is approximately linear (Figs.
+2-3).
+This contrasts with suggestions that productivity of temperate forests is similar to that of tropical forests (**REFS**).
+Such observations may be attributable to the fact that temperate forests are predominantly secondary and of younger age than tropical forests (**REF**), so analyses comparing across latitudinal gradients without strict control of stand age tend to be comparing younger temperate stands with older tropical (and often boreal) forests ([Poulter et al.
+2018-DOI:10.1594/PANGAEA.889943](http://pure.iiasa.ac.at/id/eprint/15361/)).
+Because C allocation varies with stand age (**REFS (KAT/NOBBY)**), age differences may introduce systematic biases into analyses of FACF across latitude or global climatic gradients.
+For example, woody productivity tends to be higher in rapidly aggrading secondary stands than in old-growth forests, where proportionally more C is allocated to respiration (**REFS**) [*purpose for respiration/ other compenents (**REFS????- NOBBY?**)].
+By constraining our analysis to stands ≥100 years old with no record of major recent disturbance, we have clarified the shape of the relationship between FACF and latitude.
 
-We find that seasonality is an important factor in understanding patterns in productivity on a global scale: temperature seasonality and growing season length are strong predictors of productivity, though growing season length doesn't improve on MAT as a predictor.
-Within the growing season we find the influence of climate on productivity is smaller, but still significant for a number of carbon fluxes.
+[**NOBBY, please pay special attention to this paragraph**].
+We show that FACF are broadly consistent in their responses to climate drivers on the global scale (with the exception of some differences in MAT-MAP interactions; Fig.
+4), with no major trends in C allocation among the variable pairs tested (Fig.
+2; **Some SI table**).
+Although variation in allocation has been observed along gradients of elevation [@moser_elevation_2011] and water availability [@newman_above-_2006]--along with non-climatic axes of stand age [@litton_carbon_2007], nutrient availability [@litton_carbon_2007; @gill_belowground_2016], and forest structure [@taylor_greater_2019]--variation in relation to climate is not apparent at the global scale within ForC, which contains the bulk of relevant data.
+Our conclusion, then, is that hypothesized gradients in allocation along global climate gradients cannot currently be supported for mature forests, although data quantity and standardization is currently insufficent to rule out the possibility that such trends exist.
+Of particular interest and significance are the relationships amongst $GPP$, net primary productivity ($NPP$ and its components, particularly $ANPP_{woody-stem}$), and respiration ($R_{auto}$ and components).
+There have been suggestions that tropical forests tend to have low carbon use efficiency ($CUE$= $NPP$/$GPP$=($GPP$-$R_{auto}$/$GPP$), which are based on observations of low $CUE$ in old-growth tropical forests relative to (mostly younger) extratropical forests [@de_lucia_forest_2007; [Malhi 2012](https://www.researchgate.net/publication/227837530_The_productivity_metabolism_and_carbon_cycle_of_tropical_forest_vegetation); @anderson-teixeira_carbon_2016], but our analysis suggests that these low values might more appropriately be attributed to the fact that these forests are old than to their tropical climate.
+Indeed, $CUE$ is known to decline with forest age ([@de_lucia_forest_2007];**REFS(NOBBY)**) but appears to be roughly independent of $GPP$ [@litton_carbon_2007].
+Among our sites with relevant data, there is indication in our data that CUE or $ANPP_{woody-stem}$/$GPP$ increase with latitude (**some SI table**).
+Careful methodological standardization across a consistent set of mature forest sites will be necessary to determine whether any climate-driven gradients in allocation exist at the global scale.
 
-Despite the high standard of data quality in the ForC database, there are still significant limitations on our analyses.
-We standardised methodologies as far as possible, ensuring that there was consistency in the components included for each carbon flux.
-However, field techniques for estimating components of primary productivity are variable in their accuracy.
-This is particularly true for measures of BNPP, which are often estimated by extrapolating measures of ANPP.
-Even where BNPP is calculated through direct soil sampling, measures are often taken to insufficient depth, leading to underestimation of the flux.
-It is challenging to account for this type of variability in quality of field data, and our analyses largely assume good faith in data quality.
-While we have strong confidence in our analyses of well-understood fluxes, such as GPP, ANPP, and ANPP~woody~ ~stem~, we recommend caution in drawing firm conclusions from our analyses of BNPP, BNPP~fine~ ~root~, and, to some extent, NPP.
-We expect that, as field techniques improve the accuracy of estimates of these fluxes in the future, our understanding of the relationship of BNPP with climate variables will change.
-It is possible that the low R^2^ values for BNPP that we find are a result of lower accuracy in field measures of these fluxes, and that with improve standardisation of methods, we would find climate explains a higher proportion of variation.
+One interesting observation was that climate tends to explain more variation in the major fluxes ($GPP$, $NPP$, $R_{auto}$ - latitude R^2≥48%) than in subsidiary fluxes (latitude R^2<27% for $BNPP_{fine.root}$, $R_{auto-root}$,$ANPP_{woody-stem}$) (Fig.
+2; **some SI table?**).
+There are two non-exclusive potential explanations for this.
+First, it may be that methodological variation is larger relative to flux magnitude for some of the subsidiary fluxes.
+Belowground fluxes in particular are difficult to quantify, and measurement methods for the belowground fluxes considered here may be measured through fundamentally different approaches (*e.g.*, minirhizotrons, ingrowth cores, or sequential coring for $BNPP_{root-fine}$; root exclusion, stable isotope tracking, or gas exchange of excised roots for $R_{auto-root}$), and sampling depth is variable and often insufficient to capture the full soil profile.
+$ANPP_{woody-stem}$, which is also poorly explained by latitude or climate, is more straightforward to measure but is subject to variability introduced by differences such as biomass allometries applied, minimum plant size sampled, and exclusion or inclusion of branch fall (**?- Becky, did we end up lumping?
+REF**).
+However, methodological variation and uncertainty affect all of fluxes considered here--not necessarily any less than the aforementioned, and some of the larger fluxes that vary more strongly with respect to climate ($ANPP$, $NPP$) are estimated by summing uncertain component fluxes.
+Second, differences among variables in the proportion of variation explained by climate may be attributable to more dicrect climatic control over $GPP$ than subsidiary fluxes.
+That is, subsidiary fluxes may be shaped by climate both indirectly through its influence on $GPP$ and potentially respiration and directly through any climatic influence on C allocation, as well as by many other local- and regional-scale factors (**REFS**).
 
-Secondly, we emphasise that this study considers the relationship between productivity and climate on a coarse scale.
-Because of the nature of the ForC database, we were unable to access site-level MAT and MAP data for a number of sites, and lacked site-level data for other climate variables for all sites.
-As a result the climate data we used was primarily multi-year averages of annual climate measures accessed from global databases.
-While the quality of these data is high, we cannot expect that it captures detailed local variation in climate, and in particular will be unable to capture microclimate variation, which may have a strong influence on stand-level productivity.
-Although we find that MAT is the strongest predictor for climate, part of this effect may be a result of the higher-quality stand-level dataset available for MAT.
-With high-quality stand-level data for other climate variables we expect that their predictive strength would improve.
+The latitudinal gradient in FACF (Figs.
+2-3) is driven primarily by temperature-related climate variables, and secondarily by moisture availability (Figs.
+4-5).
+Because many climate variables co-vary across the latitudinal gradient, because climatic drivers affect forest carbon flux on much shorter time scales than can be captured by annual climate summary variables, and because both climatic conditions and C flux vary intra- and inter-annually around the long-term means, it is not appropriate to attempt to identify any one mean annual climate variable as a mechanistic driver of FACF.
+However, it remains informative to consider these relationships.
+We find that temperature-related climate variables ($MAT$, temperature seasonality, ...**LIST**) explain the highest proportion of variability in productivity, and among these, $MAT$ is generally the best predictor--perhaps because site-specific MAT is recorded for the majority of sites in ForC, whereas other variables are extracted from global gridded data products (Table S1**?**).
+The effects of temperature are modified by moisture availability, with reduced FACF under hot and dry conditions (*i.e.*, high PET, high deficit; Fig.
+5c-d) and sometimes under very high precipitaiton (Figs.
+4, 5b).
+Negative effects of very high precipitation on FACF have been previously observed (**REFS**) and are attributable to nutrient and light limitations (**REFS**) Thus, although temperature and water availability jointly and interactively drive global-scale patterns of FACF.
 
-Furthermore, we expect that temporal resolution of climate data affects our results.
-Productivity is known to vary on a seasonal and annual scale in response to variation in climate.
-Our analysis relies on long-term averages in both climate variables and forest productivity estimates, and as a result averages out a large proportion of this variation.
-Therefore we recognise that the strong short-term effects of climate on productivity are likely to be masked.
-The limitations of low temporal resolution are highlighted in our analysis of growing season length.
-We note that our estimate of growing season length is imprecise, being based on monthly temperature, precipitation, and evapotranspiration averages.
-With accurate stand-level measurements of growing season length, based on daily, not monthly, climate measures, we would expect to record a stronger correlation between growing season length and productivity.
+FACF is reduced by climatic seasonality (Fig.
+5e), and productivity is shut down during cold- or dry- dormant seasons.
+To account for this, a number of analyses seeking to characterize global-scale effects of climate on productivity have examined the relationship of FACF per month of the growing season with growing season climatic conditions (**REFS**).
+We found that the sort of simple metric needed to define growing season at a global scale was uncertain for temperature and problematic for moisture (**WORK ON THIS**).
+A temperature-defined growing season length had stong positive correlation with FACF (Fig.
+5f), but explained less variation than $MAT$.
+Dividing FACFs by growing season length to yield FACF per growing season month removed the majority of climate-related variation, supporting the idea that the latitudinal gradient in FACF is attributable more to shorter growing seasons at high latitudes than to inherently lower rates of photosyntheiss or respiration by high-latitude forests (*[ Enquist GCB review]*).
+However, there remained a number of significant correlations with growing season climatic conditions, suggesting that climatic conditions remain influential within the growing season.
+We conclude that while correcting for growing season length takes analyses a step closer to mechanistic linkage of instantaneous C flux rates to environmental conditions, it remains very crude relative to the the timescales on which climate affects plant metabolism and does not advance statistical predictive power.
+Rather, mechanistic accounting for climatic effects on global productivity patterns requires models representing physiologically meaningful timescales (*e.g.*, **refs**).
 
-Although we recognise the inherent limitations present in this analysis, by focusing on mature, undisturbed forests, and by clarifying the components of NPP included in analysis, our study offers a clearer picture of how carbon cycling varies with latitude and climate.
-
-Past studies have differed in their conclusions regarding the relationship between productivity and latitude.
-Our findings indicate that, for all carbon fluxes studied, productivity peaks in the tropical regions.
-There is strong evidence from our analyses that this increase in productivity towards the tropical regions is driven primarily by temperature-related effects.
-We find that temperature-related climate variables explain the highest proportion of variability in productivity.
-In contrast, we find that MAP is not a significant predictor of productivity, but does show a strong interaction with temperature.
-This indicates that the main mechanism through with precipitation influences productivity is in a secondary capacity, with its overall influence being mediated by the effects of temperature.
-This also suggests that although temperature explains a large proportion of variation, it cannot fully explain productivity alone, and a measure of water availability is necessary to fully understand patterns of productivity.
-This is likely to be significant in the future, as it suggests that the effects of a warming climate will be mediated through changes in precipitation regime.
-
-There is debate over whether productivity in regions with higher MAT is higher because productivity inherently increases with air temperature during growing season months, or because higher MAT is correlated with longer growing seasons.
-Our analysis suggests that length of growing season is a key factor in understanding latitudinal patterns in productivity.
-We find that length of growing season is strongly correlated with productivity.
-This is supported by the strong correlation between temperature seasonality and productivity, indicating that aseasonality and year-round growing seasons have a strong influence over productivity.
-
-These conclusions are supported by our analyses of the influence of climate within the growing season.
-We find much weaker correlations between climate variables and productivity when only growing season months are considered.
-There is evidence for a positive correlation between productivity and temperature for ANPP and ANPP~foliage~, suggesting that there is still a small positive effect of air temperature on productivity independent of growing season, however this effect is not significant for the remaining fluxes.
-This suggests that the primary positive effect of temperature on productivity is through the lengthening of the growing season in regions with higher MAT.
-However, we note that length of growing season is a less good predictor of productivity than MAT.
-As discussed above, this may be an artefact of imprecise estimation of length of growing season, however it does indicate that there are aspects of productivity explained by MAT which length of growing season alone cannot explain.
-
-By including multiple fluxes in our analysis, we are able to explore the relationships between fluxes and their responses to climate.
-We find that fluxes are broadly consistent in their responses to climate drivers, suggesting that the importance of climate drivers on productivity is relatively invariant across fluxes.
-In addition, we show that carbon allocation between subsidiary fluxes does not vary significantly across broad climatic gradients.
-There is evidence that allocation between carbon fluxes does vary with factors such as stand age [@litton_carbon_2007], nutrient availability [@litton_carbon_2007; @gill_belowground_2016], forest structure [@taylor_greater_2019], elevation [@moser_elevation_2011], and water availability [@newman_above-_2006].
-That we do not see consistent patterns of variation in allocation across climatic gradients indicates that the pressures influencing shifts in allocation between fluxes occur on local scales, and that shifts in allocation occur as a response to fluctuations in local environmental conditions.
-
-Related to this, we find that climate more strongly explains variation in the major fluxes, with lower levels of variation explained in subsidiary fluxes.
-Although this could in part be related to a lack of standardisation in methodology across studies, as discussed above, it could also indicate that factors other than climate have significant influence over subsidiary fluxes.
-This is consistent with the results we find for allocation between fluxes: although climate has a significant influence on subsidiary fluxes, it is not the only factor that is important, and many other local- and regional-scale factors  influence the way carbon is allocated between subsidiary fluxes.
-We would expect that this would reduce the strength of the effect of climate on subsidiary fluxes.
-
-**[needs strong concluding paragraph]**
+Our analysis clarifies the big picture of how FACF varies with latitude and climate on a global scale, with some important implications for how forest carbon cycling relates to climate and, by extension, how it is likely to respond to climatic warming.
+Contrary to some previous suggestions, we find no support for non-linear trends in mature forest C cycling with respect to latitude or $MAT$, and no distinct trends in C allocation across the global scale (Fig.
+3).
+The implication is that under warmer conditions with similar moisture availability--and within the temperature range to which forest communities are adapated and acclimatized--higher temperatures result in a generalized acceleration of FACF, with no major shifts in C allocation among subsidiary fluxes.
+Of course, actual climatic changes will result in very different sets of conditions than represented across geographic gradients in climate, but our analysis clarifies the big picture as to how carbon cycles through forest ecosystems that have developed under relatively stable climatic conditions.
+As we enter a period of accelerating climatic change, understanding of the fundamental climatic controls on FACF sets a foundation for understanding patterns of change.
+**[work on this]**
 
 ### Acknowledgements
-
-### Supporting Information
-
-```{r eval = TRUE, echo=FALSE, warning=FALSE}
-library(knitr)
-library(kableExtra)
-S1 <- read.csv("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/S1.csv", stringsAsFactors = FALSE, check.names = FALSE)
-kable(S1, booktabs = TRUE, caption = "Hypothesis table", escape = F)
-```
+Scholarly Studies
+ForestGEO
+Compilatin of the ForC database was originally funded by DOE

### 2019-09-12 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
 ---
-title: "Manuscript draft"
-author: "Rebecca Banbury Morgan, Valentine Herrmann, Norbert Kunert, Ben Bond-Lamberty, Helene Muller-Landau, [MORE??-CHECK], Kristina J. Anderson-Teixeira"
+title: "Global patterns of forest autotrophic carbon fluxes"
+author: "Rebecca Banbury Morgan, Valentine Herrmann, Norbert Kunert, Ben Bond-Lamberty, Helene Muller-Landau, Kristina J. Anderson-Teixeira"
 output:
@@
 ), 'packages.bib')
+```
+
+```{r setup, include = FALSE}
+knitr::opts_knit$set(root.dir = 'C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/')
 ```
@@
 library(kableExtra)
-hypothesis_table <- read.csv("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/hypothesis_table.csv", stringsAsFactors = FALSE, check.names = FALSE)
+hypothesis_table <- read.csv("hypothesis_table.csv", stringsAsFactors = FALSE, check.names = FALSE)
 kable(hypothesis_table, booktabs = TRUE, caption = "Hypothesis table") %>%
@@
 
-```{r echo=FALSE, out.width='100%', fig.cap = "Graphs of primary productivity $(MgC~ha^-1~yr^-1)$ regressed against latitude.
-Lines of best fit are plotted according to the best model selected during analysis.
-All regressions are significant $(p<0.05)$.
-Plots 1 - 3 show two component fluxes; a larger flux, defined as the combination of the two component fluxes; and a modelled estimate of the sum of the two component fluxes.
-95\\% confidence intervals are plotted for the larger flux.
-Plot 4 shows three belowground fluxes, and a modelled estimate of the total belowground carbon flux", fig.pos='H'}
-
-  knitr::include_graphics("combined_stacked.png")
-```
-
 *How does productivity vary with latitude?*
 
-All major carbon fluxes increased linearly with decreasing latitude (fig. 3).
+All major carbon fluxes increased linearly with decreasing latitude (fig. 2).
 
@@
 *Relationships and differences among fluxes.* In general, smaller component fluxes summed approximately to  larger fluxes across the latitudinal gradient (fig.
-2).
+3).
 That is, modelled estimates of GPP, generated from the sum of NPP and R auto; NPP, generated from the sum of ANPP and BNPP~root~; and ANPP, generated from the sum of ANPP~foliage~ and ANPP~woody~ ~stem~, fell completely within the confidence intervals of the regressions of field estimates of GPP, NPP and ANPP respectively.
+
+```{r echo=FALSE, out.width='100%', fig.cap = "Graphs of primary productivity $(MgC~ha^-1~yr^-1)$ regressed against latitude.
+Lines of best fit are plotted according to the best model selected during analysis.
+All regressions are significant $(p<0.05)$.
+Plots 1 - 3 show two component fluxes; a larger flux, defined as the combination of the two component fluxes; and a modelled estimate of the sum of the two component fluxes.
+95\\% confidence intervals are plotted for the larger flux.
+Plot 4 shows three belowground fluxes, and a modelled estimate of the total belowground carbon flux", fig.pos='H'}
+
+  knitr::include_graphics("combined_stacked.png")
+```
 
@@
 For example, woody productivity tends to be higher in rapidly aggrading secondary stands than in old-growth forests, where proportionally more C is allocated to respiration (**REFS**) [*purpose for respiration/ other compenents (**REFS????- NOBBY?**)].
-By constraining our analysis to stands ≥100 years old with no record of major recent disturbance, we have clarified the shape of the relationship between FACF and latitude.
+By constraining our analysis to stands $\\geq$100 years old with no record of major recent disturbance, we have clarified the shape of the relationship between FACF and latitude.
 
@@
 
-One interesting observation was that climate tends to explain more variation in the major fluxes ($GPP$, $NPP$, $R_{auto}$ - latitude R^2≥48%) than in subsidiary fluxes (latitude R^2<27% for $BNPP_{fine.root}$, $R_{auto-root}$,$ANPP_{woody-stem}$) (Fig.
+One interesting observation was that climate tends to explain more variation in the major fluxes ($GPP$, $NPP$, $R_{auto}$ - latitude R^2$\\geq$48%) than in subsidiary fluxes (latitude R^2<27% for $BNPP_{fine.root}$, $R_{auto-root}$,$ANPP_{woody-stem}$) (Fig.
 2; **some SI table?**).

### 2019-09-12 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
 | GPP | Annual gross primary production; annual uptake of carbon dioxide by an ecosystem | NA | Flux partitioning of eddy covariance | 243 | 49 |
-| NPP | Annual net primary production; the component of GPP that is stored in plant tissue; GPP minus ecosystem respiration | Foliage, branch, stem, coarse root, fine root and optionally understory | Direct measurement of annual increments of components | 92 | 42 |
+| NPP | Annual net primary production; the component of GPP that is stored in plant tissue; GPP minus ecosystem respiration | Foliage, branch, stem, coarse root and fine root | Direct measurement of annual increments of components | 92 | 42 |
 | ANPP | Aboveground net primary production | Foliage, stem, and optionally branch | Direct measurement of annual increments of components | 256 | 80 |

### 2019-09-12 manuscript/manuscript draft.Rmd
update draft, knit

@@
 ```{r setup, include = FALSE}
-knitr::opts_knit$set(root.dir = 'C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/')
+#knitr::opts_knit$set(root.dir = 'C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/')
+knitr::opts_knit$set(root.dir = '/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/Global_Productivity/manuscript/')
 ```
@@
 Forest autotrophic carbon flux (FACF) influences all organic matter stocks in ecosystems and is linked to cycling of energy, water, and nutrients (**REFS**).
-Forest productivity sets the energy available to heterotrophic communities, in turn influencing their abundance (**REFS**) and possibly diversity (**REFS**)--but probably not individual metabolic rate (**Anderson & Jetz 2005 (my first paper!!), but check more recent literature**).
+Forest productivity sets the energy available to heterotrophic communities, in turn influencing their abundance (**REFS**) and possibly diversity (**REFS**)--but probably not individual metabolic rate (**Anderson & Jetz 2005, but check more recent literature**).
 On the global scale, forests play a critical role in regulating atmospheric CO_2 and climate (**Bonan**).
@@
 Belowground fluxes in particular are difficult to quantify, and measurement methods for the belowground fluxes considered here may be measured through fundamentally different approaches (*e.g.*, minirhizotrons, ingrowth cores, or sequential coring for $BNPP_{root-fine}$; root exclusion, stable isotope tracking, or gas exchange of excised roots for $R_{auto-root}$), and sampling depth is variable and often insufficient to capture the full soil profile.
-$ANPP_{woody-stem}$, which is also poorly explained by latitude or climate, is more straightforward to measure but is subject to variability introduced by differences such as biomass allometries applied, minimum plant size sampled, and exclusion or inclusion of branch fall (**?- Becky, did we end up lumping?
-REF**).
+$ANPP_{woody-stem}$, which is also poorly explained by latitude or climate, is more straightforward to measure but is subject to variability introduced by differences such as biomass allometries applied and minimum plant size sampled.
 However, methodological variation and uncertainty affect all of fluxes considered here--not necessarily any less than the aforementioned, and some of the larger fluxes that vary more strongly with respect to climate ($ANPP$, $NPP$) are estimated by summing uncertain component fluxes.

### 2019-09-12 manuscript/manuscript draft.Rmd
deleted abstract (out of date)

@@
 ###Abstract
-
-Globally, forests are a significant carbon sink, showing high levels of productivity and representing an important component of the carbon cycle.
-Current knowledge indicates that forest productivity decreases with latitude; however, there is evidence that this relationship varies across different fluxes.
-Furthermore, the climate variables that drive this relationship are poorly understood.
-Here, we use a comprehensive global database of forest carbon fluxes to show that, while all major carbon fluxes decrease with latitude, allocation to woody productivity increases with latitude, and allocation to root productivity declines.
-The best predictors of global patterns of productivity are temperature variables, including mean annual temperature, annual temperature range, and temperature seasonality, with the exception of woody productivity, which is most strongly influenced by potential evapotranspiration and vapour pressure deficit.
-Climate variables explain a large proportion of the variation in major carbon fluxes, but are a less significant predictor of subsidiary components.
-Our results illustrate the strong influence of climate on primary productivity, and particularly the importance of temperature in determining forest productivity.
-However, they also indicate that effects of climate are complex: fluxes show non-linear responses to climate variables, and many factors other than climate interact to determine allocation to subsidiary components.
-Knowledge of the relative influence of climate variables on forest carbon fluxes is an important component of understanding the responses of primary productivity to climate change.
 

### 2019-09-12 manuscript/manuscript draft.Rmd
editing manuscript

@@
 Evidence for this is further complicated by the fact that different studies use different measures of productivity to explore these relationships.
-For example, modelling of global terrestrial ecosystem gross primary productivity (GPP) through upscaling and calibration of eddy flux measurements indicates that GPP peaks in tropical forests [@beer_terrestrial_2010; @jung_global_2011].
+For example, modelling of global terrestrial ecosystem gross primary productivity (GPP) through upscaling and calibration of eddy flux measurements indicates that GPP peaks in tropical forests [@beer_terrestrial_2010; @jung_global_2011];[Bagdley et al.
+2019](https://onlinelibrary.wiley.com/doi/epdf/10.1111/gcb.14729)).
 This is corroborated by analysis of site-level GPP measurements, which appear to reach their highest levels in tropical forests [@luyssaert_co_2007].
@@
 If the best model included a polynomial term, the shape of the polynomial relationship was considered.
-If the shape of the relationship made biological sense, and was a significant improvement on the linear relationship (deltaAIC >2), we accepted the polynomial as the best model.
+If the shape of the relationship *made biological sense*, and was a significant improvement on the linear relationship (deltaAIC >2), we accepted the polynomial as the best model.
 If not, we ran the linear model as the final model.
@@
 To account for correlations in measurements between tightly clustered sites, a random effect was specified as plot nested within geographic area.
-Data from the temperate regions was heavily skewed towards studies from the old-growth forests of the Pacific Northwest.
-These forests have very high productivity, and so to ensure that results were not unduly influenced by geographic sampling bias, we tried a version of the model where data were weighted according to forested land area within each Koeppen climate zone.
+*Data from the temperate regions was heavily skewed towards studies from the old-growth forests of the Pacific Northwest.*  These forests have very high productivity, and so to ensure that results were not unduly influenced by geographic sampling bias, we tried a version of the model where data were weighted according to forested land area within each Koeppen climate zone.
 Results were similar between the weighted and unweighted model, so, to avoid problems of over-fitting, the weighted model was dropped, and results from this are not presented here.
@@
 Cook's distance analyses were carried out for each of the models, and indicated that data from a few high-elevation sites were having a disproportionate influence on the regressions.
-To account for this, models were re-run using only data from sites $\leq$ 1000m.
+To account for this, models were re-run using only data from sites $\le$ 1000m.
 
@@
 
-One interesting observation was that climate tends to explain more variation in the major fluxes ($GPP$, $NPP$, $R_{auto}$ - latitude R^2$\\geq$48%) than in subsidiary fluxes (latitude R^2<27% for $BNPP_{fine.root}$, $R_{auto-root}$,$ANPP_{woody-stem}$) (Fig.
+One interesting observation was that climate tends to explain more variation in the major fluxes ($GPP$, $NPP$, $R_{auto}$ - latitude $R^2 \ge$48%) than in subsidiary fluxes (latitude $R^2$<27% for $BNPP_{fine.root}$, $R_{auto-root}$,$ANPP_{woody-stem}$) (Fig.
 2; **some SI table?**).

### 2019-09-12 manuscript/manuscript-draft.tex
editing manuscript

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Global patterns of forest autotrophic carbon fluxes},
            pdfauthor={Rebecca Banbury Morgan, Valentine Herrmann, Norbert Kunert, Ben Bond-Lamberty, Helene Muller-Landau, Kristina J. Anderson-Teixeira},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{natbib}
\bibliographystyle{apalike}
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\providecommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{Global patterns of forest autotrophic carbon fluxes}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Rebecca Banbury Morgan, Valentine Herrmann, Norbert Kunert, Ben
Bond-Lamberty, Helene Muller-Landau, Kristina J. Anderson-Teixeira}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
    \date{}
    \predate{}\postdate{}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\usepackage{float} \usepackage{caption} \captionsetup[table]{font=footnotesize} \captionsetup[figure]{font=footnotesize}

\begin{document}
\maketitle

\subsubsection{Abstract}\label{abstract}

\subsubsection{Introduction}\label{introduction}

Carbon fixation, allocation, and metabolism by trees sets the basis for
energy and material flows in forest ecosystems and defines their
interactions with Earth's changing climate. Forest autotrophic carbon
flux (FACF) influences all organic matter stocks in ecosystems and is
linked to cycling of energy, water, and nutrients (\textbf{REFS}).
Forest productivity sets the energy available to heterotrophic
communities, in turn influencing their abundance (\textbf{REFS}) and
possibly diversity (\textbf{REFS})--but probably not individual
metabolic rate (\textbf{Anderson \& Jetz 2005, but check more recent
literature}). On the global scale, forests play a critical role in
regulating atmospheric CO\_2 and climate (\textbf{Bonan}). The total
amount of CO\_2 cycling through Earth's forests each year (\emph{i.e.},
total gross primary production, \(GPP\)) is more than five times
anthropogenic fossil fuel emissions, and net sequestration
(\textasciitilde{}2.4 GtCyr\textsuperscript{-1}) offsetting roughly 30\%
of these emissions \citep{pan_large_2011}. As atmospheric carbon dioxide
levels continue to rise, with consequences for global climate, there is
increasing recognition that proper protection and management of forest
resources will have an important role to play in mitigating climate
change (** REFS- e.g., IPCC 1.5C and land reports**). Given the
importance of understanding broad-scale patterns in forest autotrophic
carbon flux (FACF), there have been many meta-analyses on the theme.
However, our ability to draw general macroscopic conclusions regarding
global variation in multiple productivity variables with respect to
climate has been limited in that these analyses often mix forests that
vary in stand age, disturbance history, and/or management status; do not
always sufficiently parse related variables (\emph{e.g.}, combining net
primary productivity records with and without belowground components);
and typically consider only one or a few variables at a time. The recent
development of a global forest carbon database synthesizing multiple
variables and including records of stand history (ForC;
\citep{anderson-teixeira_carbon_2016, anderson-teixeira_forc:_2018})
opens up the possibility for a standardized analysis of global scale
variation in multiple components of forest productivity and the
principle climatic drivers of these patterns.

On a global scale, the productivity of forests varies with latitude,
showing a general trend of decreasing productivity with latitude
\citep{beer_terrestrial_2010, jung_global_2011}. Studies agree that
productivity is lowest in the boreal regions, and increases into the
temperate regions
\citep{luyssaert_co_2007, huston_global_2009, beer_terrestrial_2010, jung_global_2011}.
However, evidence is inconclusive on whether productivity continues to
increase into the tropics, or whether it plateaus in temperate regions.
Evidence for this is further complicated by the fact that different
studies use different measures of productivity to explore these
relationships. For example, modelling of global terrestrial ecosystem
gross primary productivity (GPP) through upscaling and calibration of
eddy flux measurements indicates that GPP peaks in tropical forests
\citep{beer_terrestrial_2010, jung_global_2011};\href{https://onlinelibrary.wiley.com/doi/epdf/10.1111/gcb.14729}{Bagdley
et al. 2019}). This is corroborated by analysis of site-level GPP
measurements, which appear to reach their highest levels in tropical
forests \citep{luyssaert_co_2007}. In contrast, there is evidence that
the highest values of net primary productivity (NPP) may be found in
temperate forests \citep{luyssaert_co_2007, huston_global_2009},
although other studies find NPP is highest in the tropics, showing a
decrease with latitude \citep{simova_enigma_2017}. Other studies have
chosen to focus exclusively on above-ground net primary productivity
(ANPP), finding evidence of a weak negative relationship between ANPP
and latitude \citep{huston_global_2009, gillman_latitude_2015}. An
understanding of the global patterns of productivity is important;
however elucidating the drivers of these patterns will be even more
valuable in expanding our understanding of global carbon cycling.
Primary productivity can be influenced by many factors, which often act
across a range of scales, and may show interactive effects with each
other. On a local scale, stand age
\citep{litton_carbon_2007, gillman_latitude_2015}, management
\citep{simova_enigma_2017}; nutrient availability
\citep{aragao_above-_2009}; and altitude
\citep{girardin_net_2010, malhi_variation_2017} all impact forest
productivity. On a global scale, we expect that productivity is most
strongly influenced by broad climatic gradients.

We know from previous research that climate is a significant driver of
productivity across broad spatial scales
\citep{cleveland_relationships_2011}. The majority of studies have
focused on exploring the relationships between productivity and mean
annual temperature (MAT) and mean annual precipitation (MAP), as the
most commonly reported site-level climate variables. These variables
have the advantage that they describe broad trends in temperature and
water availability, and therefore capture a lot of global-scale
variation in climate. There is strong evidence that both MAT and MAP
show significant positive relationships with productivity
\citep{chu_does_2016}. However, as with latitude, the shape of those
relationships is not always clear, and, again, is complicated by the use
of different measures of productivity across studies. There is support
across multiple studies for the hypothesis that various measures of
primary productivity saturate at high levels of MAP, though the
saturation points identified vary from 1500mm \citep{luyssaert_co_2007}
up to 2445mm MAP \citep{schuur_productivity_2003}. Studies of the
influence of MAT on productivity are less conclusive. Luyssaert et al.
\citeyearpar{luyssaert_co_2007} examined GPP and NPP and found that,
while GPP increases linearly with MAT, NPP saturates at around
10\(^\circ\)C MAT. In contrast, Larjavaara and Muller-Landau
\citeyearpar{larjavaara_temperature_2012}, find that increases in GPP
saturate at approximately 25\(^\circ\)C MAT, while Schuur
\citeyearpar{schuur_productivity_2003} shows that NPP increases linearly
with temperature. The influence of these climate variables on
productivity is further complicated by the possibility of interactive
effects occurring between them. Taylor et al.
\citeyearpar{taylor_temperature_2017} showed that increased MAP had a
negative effect at low MAT, but a positive effect at high levels of MAT,
and vice versa, suggesting that both temperature and water availability
are important in explaining productivity.

While MAT and MAP are important climate variables, they do not capture
all aspects of climate that may influence productivity, and may be
insufficient to capture the breadth of climatic effects on productivity
\citep{cleveland_relationships_2011}. There is evidence that
productivity also responds to variables such as cloud cover
\citep{taylor_temperature_2017}, solar radiation
\citep{fyllas_solar_2017}, and potential evapotranspiration
\citep{kerkhoff_plant_2005} in potentially significant ways.
Furthermore, MAT and MAP are very coarse measures of climate, and so
fail to capture much variation in climate on an intra-annual scale,
including the effects of factors such as growing season length, number
of frost-free days, temperature seasonality, and dry season length. Some
studies have suggested that the apparently strong relationship between
MAT and productivity is actually a factor of the correlation between MAT
and growing season length
\citep{kerkhoff_plant_2005, malhi_productivity_2012, michaletz_convergence_2014, michaletz_drivers_2018}.
Kerkhoff et al. \citeyearpar{kerkhoff_plant_2005} and Michaletz et al.
\citeyearpar{michaletz_convergence_2014} find that, within the growing
season, there is no significant relationship between productivity and
MAT, indicating that the effect of temperature is due to increased
length of growing season, rather than an inherent influence of
temperature on productivity.

In order to approach these broad and complex issues, we simplify the
major gaps in our knowledge to five key questions and corresponding
specific predictions (Table 1). Firstly, we ask how forest autotrophic
carbon fluxes (FACF) vary with latitude. We then test how these fluxes
relate to MAT and MAP, and additionally how they respond to other, less
well studied, climate variables. We finally consider the relationship
between FACF and seasonality, considering the role of seasonality in
explaining variation in carbon fluxes, and the influence of climate on
FACF per month of the growing season.

Resolving these questions requires data which, firstly, control for
variables influencing productivity, including stand age, disturbance
regime, elevation, and methodology; and secondly includes all major
carbon fluxes. Here we use a comprehensive global database of forest
carbon fluxes to explore the above questions for nine carbon fluxes,
allowing for an in-depth exploration of the effect of climate on global
productivity.

\renewcommand{\arraystretch}{2}

\begin{landscape}\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-4}Hypothesis table}
\centering
\resizebox{\linewidth}{!}{
\fontsize{12}{14}\selectfont
\begin{tabular}{>{\raggedright\arraybackslash}p{10cm}>{\raggedright\arraybackslash}p{5cm}>{\raggedright\arraybackslash}p{1.8cm}>{\raggedright\arraybackslash}p{1.8cm}>{\raggedright\arraybackslash}p{1.8cm}>{\raggedright\arraybackslash}p{1.8cm}>{\raggedright\arraybackslash}p{1.8cm}>{\raggedright\arraybackslash}p{1.8cm}>{\raggedright\arraybackslash}p{1.8cm}>{\raggedright\arraybackslash}p{1.8cm}>{\raggedright\arraybackslash}p{1.8cm}>{\raggedright\arraybackslash}p{1.8cm}>{\raggedright\arraybackslash}p{1.8cm}}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{1}{c}{ } & \multicolumn{1}{c}{ } & \multicolumn{9}{c}{Forest autotrophic carbon fluxes (FACF)} & \multicolumn{1}{c}{ } \\
\cmidrule(l{3pt}r{3pt}){4-12}
Hypotheses \& Specific Predictions & Related references & Overall & GPP & NPP & ANPP & BNPP & \$ANPP\_\{foliage\}\$ & ANPP woody stem & BNPP fine root & R auto root & R auto & Support\\
\midrule
\addlinespace[1em]
\hline
\multicolumn{1}{l}{\textbf{1. How do FACF vary with latitude?}}\\
\hspace{1em}1.1. FACFs decrease linearly with latitude (L-) & Luyssaert et al. (2007) & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Fig. 2-3\\
\hspace{1em}1.1.alt. FACFs are similar in tropical and temperate forests, but lower in boreal regions (CD-) & Gillman et al. (2015); Simova and Storch (2017) & no & no & no & no & no & no & no & no & no & no & Fig. 2-3\\
\hspace{1em}1.2. Allocation to FACFs varies with latitude & Litton et al. (2007); DeLucia et al. (2007) & no & - & no & no & no & no & no & no & - & - & \\
\addlinespace[1em]
\hline
\multicolumn{1}{l}{\textbf{2. How do FACF relate to MAT and MAP?}}\\
\hspace{1em}2.1. FACFs increase linearly with MAT (L+) & Schuur (2003) & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Fig. 5\\
\hspace{1em}2.2. FACFs increase with precipitation but saturate or decrease at very high levels (CD+) & Luyssaert et al. (2007); Schuur (2003) & yes & yes & yes & yes & yes & yes & () & L+ & L+ & yes & Fig. 5\\
\hspace{1em}2.3. There is a postive interaction between temperature and precipitation (I+) & Taylor et al. (2016) & (yes) & yes & L+ & yes & I- & L+ & yes & I- & yes & yes & Fig. 4\\
\addlinespace[1em]
\hline
\multicolumn{1}{l}{\textbf{3. How do FACF relate to other annual climate variables?}}\\
\hspace{1em}3.1. FACFs increases with PET, but saturates or decreases at high levels (CD+) &  & (yes) & yes & yes & yes & yes & L+ & yes & L+ & yes & L+ & Fig. 5\\
\hspace{1em}3.2. FACFs increase with vapour pressure deficit, but saturate or decrease at high levels (CD+) &  & (yes) & yes & yes & yes & yes & yes & yes & L+ & yes & yes & Fig. 5\\
\hspace{1em}3.3. FACFs increase linearly with solar radiation (L+) &  & (yes) & yes & yes & yes & CD+ & yes & CD+ & yes & () & yes & \\
\addlinespace[1em]
\hline
\multicolumn{1}{l}{\textbf{4. What is the role of seasonality?}}\\
\hspace{1em}4.1. FACFs decrease linearly with temperature seasonality (L-) &  & yes & CU+ & CU+ & CU+ & yes & yes & yes & yes & CU+ & CU+ & Fig. 5\\
\hspace{1em}4.2. FACFs decrease linearly with  precipitation seasonality (L-) &  & no & () & () & () & () & () & () & () & () & () & \\
\hspace{1em}4.3. FACFs increase linearly with growing season length (L+) & Malhi (2012); Michaletz et al. (2014); Chu et al. (2016) & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Fig. 5\\
\hspace{1em}4.4. Growing season length is a better predictor of FACFs than MAT & Michaletz et al. (2014); Chu et al. (2016) & no & no & no & no & no & no & no & no & no & no & \\
\addlinespace[1em]
\hline
\multicolumn{1}{l}{\textbf{5. Per month of the growing seaon, how do FACF vary with climate?}}\\
\hspace{1em}5.1. Increase with temperature (L+) & Michaletz et al. (2014) & (mixed) & () & () & yes & () & yes & () & () & () & () & \\
\hspace{1em}5.2. Increase with PET (L+) &  & (yes) & yes & yes & () & yes & () & yes & yes & () & () & \\
\hspace{1em}5.3. Increase with precipitation (L+) &  & no & () & () & yes & () & yes & () & () & () & () & \\
\hspace{1em}5.4. Increase with solar radiation (L+) &  & (mixed) & yes & yes & () & yes & () & () & yes & () & () & \\
\bottomrule
\end{tabular}}
\end{table}
\end{landscape}

\subsubsection{Materials and Methods}\label{materials-and-methods}

Analyses were conducted on data contained in the open-access ForC
database
\citep{anderson-teixeira_carbon_2016, anderson-teixeira_forc:_2018}.
This database contains records of field-based measurements of forest
carbon stocks and annual fluxes, compiled from original publications and
existing data compilations and databases. Associated data, such as stand
age, measurement methodologies, and disturbance history, are also
included. The database was significantly expanded since the publication
of \citep{anderson-teixeira_forc:_2018} through integration with the
Global Database of Soil Respiration \citep{bond-lamberty_global_2010}.
Additional targeted literature searches were conducted to identify any
further available data on primary productivity, with particular focus on
mature forests in temperate and boreal regions. ForC currently contains
29730 records from 4979 plots, representing 20 distinct ecozones across
all forested biogeographic and climate zones. We used ForC v3.0,
archived on Zenodo with DOI 10.5281/zenodo.3403855.

\begin{figure}[H]
\includegraphics[width=1\linewidth]{distribution_all_variables} \caption{Map showing all data used in the analysis, coded by variable}\label{fig:unnamed-chunk-5}
\end{figure}

\emph{Data selection.} Over 50 variables of forest carbon stocks and
annual fluxes are represented in the ForC database; this analysis
focussed on measures of primary productivity. Table 1 contains details
of the variables selected for analysis.

\begin{longtable}[]{@{}llllll@{}}
\caption{Definitions and sample sizes of variables used in analysis.
Geographic areas group geographically proximate sites, defined using a
hierarchical cluster analysis on the distance matrix of the sites, and a
cutoff of 25km.}\tabularnewline
\toprule
\begin{minipage}[b]{0.14\columnwidth}\raggedright\strut
Variable\strut
\end{minipage} & \begin{minipage}[b]{0.19\columnwidth}\raggedright\strut
Definition\strut
\end{minipage} & \begin{minipage}[b]{0.13\columnwidth}\raggedright\strut
Components included\strut
\end{minipage} & \begin{minipage}[b]{0.23\columnwidth}\raggedright\strut
Methodologies used\strut
\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\raggedright\strut
Number of records\strut
\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\raggedright\strut
Number of geographic areas\strut
\end{minipage}\tabularnewline
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{0.14\columnwidth}\raggedright\strut
Variable\strut
\end{minipage} & \begin{minipage}[b]{0.19\columnwidth}\raggedright\strut
Definition\strut
\end{minipage} & \begin{minipage}[b]{0.13\columnwidth}\raggedright\strut
Components included\strut
\end{minipage} & \begin{minipage}[b]{0.23\columnwidth}\raggedright\strut
Methodologies used\strut
\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\raggedright\strut
Number of records\strut
\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\raggedright\strut
Number of geographic areas\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
GPP\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
Annual gross primary production; annual uptake of carbon dioxide by an
ecosystem\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
NA\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
Flux partitioning of eddy covariance\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
243\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
49\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
NPP\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
Annual net primary production; the component of GPP that is stored in
plant tissue; GPP minus ecosystem respiration\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
Foliage, branch, stem, coarse root and fine root\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
Direct measurement of annual increments of components\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
92\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
42\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
ANPP\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
Aboveground net primary production\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
Foliage, stem, and optionally branch\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
Direct measurement of annual increments of components\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
256\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
80\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
ANPP\textsubscript{foliage}\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
Net primary production of foliage\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
Foliage\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
Direct measurement of litterfall, correcting for changes in leaf biomass
when measured\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
98\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
49\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
ANPP\textsubscript{woody} \textsubscript{stem}\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
Net primary production of woody stems\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
Woody stems\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
Direct measurement of stem growth increment\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
264\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
96\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
BNPP\textsubscript{root}\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
Belowground net primary production\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
Coarse and fine roots\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
Direct measurement of one or more of: fine root turnover, soil cores,
root ingrowth cores, minirhizotrons; indirect estimates of coarse roots
using allometries based on aboveground stem increment measures\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
101\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
48\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
BNPP\textsubscript{fine} \textsubscript{root}\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
Net primary production of fine roots\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
Fine roots\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
Direct measurement of one or more of: minirhizotrons, fine root
turnover, soil cores, root ingrowth cores\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
88\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
41\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
R\textsubscript{auto}\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
Annual autotrophic respiration, including above- and belowground
components\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
Foliage, stem, and root\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
Chamber measurements of component gas exchange\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
22\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
13\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
R\textsubscript{root}\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
Annual root respiration\strut
\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
Coarse and fine roots\strut
\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
Measurement of root gas exchange\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
64\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
26\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

A subset of the ForC database was generated for the purposes of this
analysis, in order to control for data quality and remove biasing
factors. Since management can alter observed patterns of primary
productivity \citep{simova_enigma_2017}, sites were excluded from
analysis if they were managed, defined as plots that were planted,
managed as plantations, irrigated, fertilised or including the term
``managed'' in their site description. Sites that had experienced
significant disturbance were also excluded. Disturbances that justified
site exclusion were major cutting or harvesting, and/or burning,
flooding, drought and storm events with site mortality
\textgreater{}10\% of trees. Grazed sites were retained.\\
There is evidence that stand age influences patterns of primary
productivity and carbon allocation in forest ecosystems, and can
confound relationships between latitude and primary productivity
\citep{de_lucia_forest_2007, gillman_latitude_2015}. To reduce any
biasing effects of stand age, stands under 100 years of age were
excluded from analysis. Sites for which stand age was unknown were
excluded from analysis.

\emph{Methodological consistency.} The data in ForC is derived from a
range of studies, often employing different methods. For this reason,
criteria were introduced to standardise for differences in methodology.
Where data was based on forest plot census measurements, studies which
used a minimum diameter at breast height (DBH) measure of
\textgreater{}10cm were excluded from analysis. It would be preferable
to standardise by minimum area sampled; however x\% of plots in the
database are 1 ha or under in size; excluding these plots would place
significant constraints on sample size.

As discussed above, estimates of NPP, ANPP, and BNPP are generated
through summing measurements of their component parts. Since the
components included in productivity estimates vary between studies,
estimates of productivity were classified within the ForC database
according to their components, and then filtered for analysis. Estimates
of NPP were selected if they included foliage, branch, stem, coarse
root, and fine root. Measures of NPP which included additional
components, including understorey, volatile organic compounds (VOCs),
exudates, estimates of NPP lost to herbivory, and the NPP of
reproductive structures, were excluded. Estimates of ANPP were selected
if they included foliage, stem growth and optionally branch turnover.
Any measures of primary productivity where components were unknown were
excluded from analysis.

\emph{Climate datasets.} Where site-level data on mean annual
temperature, mean annual precipitation, and latitude were available in
the primary literature, this data was compiled and entered directly into
the ForC database. Based on the geographic co-ordinates for each site,
data on a further 11 climate variables was extracted from five
open-access climate datasets: WorldClim \citep{hijmans_very_2005},
WorldClim2 \citep{fick_worldclim_2017}, the Climate Research Unit (CRU)
time-series dataset v. 4.03 \citep{harris_updated_2014}, the Global
Aridity Index and Potential Evapotranspiration Climate Database
\citep{trabucco_global_2018}, and TerraClimate
\citep{abatzoglou_terraclimate_2018} (see Supplementary Information S1
for details of climate variables). Where site-level data was missing for
mean annual temperature and/or mean annual precipitation, data was
extracted from the WorldClim dataset.

Additionally, two climate variables were derived from the above
datasets: maximum vapour pressure deficit, defined as the vapour
pressure deficit of the month with the largest deficit; and water stress
months, defined as the number of months annually where precipitation was
lower than potential evapotranspiration.

\emph{Length of growing season.} Growing season months were defined as
months with mean minimum temperature \textgreater{} 0.5\(^\circ\)C.
Growing season months were initially calculated following methods used
by Kerkhoff et al. (2005), which additionally required that growing
season months had a moisture index, defined as (MAT - PET)/PET,
\textgreater{} -0.95. Michaletz et al. (2014) included an equivalent
requirement in their calculation of growing season length. However, we
found that including this requirement had no effect on the estimates of
growing season length, and so chose to exclude it.

Monthly data for PET, precipitation, and temperature was downloaded from
the Climate Research Unit (CRU) time-series dataset v 4.03
\citep{harris_updated_2014}, and for solar radiation from WorldClim2
\citep{fick_worldclim_2017}, and used to calculate mean monthly PET,
precipitation, temperature and solar radiation during the growing
season. Total growing season precipitation and solar radiation were also
calculated.

\emph{Model specification.} The effects of climate and latitude on
primary productivity were analysed using mixed effects models using the
package `lme4' \citep{bates_fitting_2015} in R v.3.5.1
\citep{r_core_team_r:_2018}. The effect of each extracted climate
variable on each measure of primary productivity was modelled by
specifying the climate variable as a fixed effect. For each climate
variable, three models were specified: a null model; a model with the
climate variable as a linear term; and a model with the climate variable
as a polynomial term. AIC values were calculated for the models and used
to select the best model. If the best model included a polynomial term,
the shape of the polynomial relationship was considered. If the shape of
the relationship \emph{made biological sense}, and was a significant
improvement on the linear relationship (deltaAIC \textgreater{}2), we
accepted the polynomial as the best model. If not, we ran the linear
model as the final model. R\textsuperscript{2} values were calculated
for the best model. All R\textsuperscript{2} values presented here are
marginal R\textsuperscript{2} values, and refer to the proportion of
variation explained by only the fixed effects, unless otherwise
specified. In addition, slope coefficients were calculated for the
linear models.

Because the magnitude of fluxes varies significantly, in order to
facilitate comparisons between regression models for each flux, data for
each flux was scaled, to give the data a mean of 0 and standard
deviation of 1. As each data set was scaled separately, this does not
allow for statistical comparisons of slope values, but does assist in
visualising the data.

To test for a potential influence of altitude, models were also run with
site altitude included as a second fixed effect. These models were
compared against models with no altitude term, and AIC values calculated
to identify whether inclusion of altitude as a term improved the models.
Including altitude had a very small effect on most models, with the
difference in AIC values between models including and excluding altitude
often being \textless{}2, suggesting the models are very similar in
their explanatory power. As a result, it was decided to present results
only from models do not include altitude as a fixed term.

Within the ForC database, sites within 25km\textsuperscript{2} of each
other are clustered into geographic areas. To account for correlations
in measurements between tightly clustered sites, a random effect was
specified as plot nested within geographic area. \emph{Data from the
temperate regions was heavily skewed towards studies from the old-growth
forests of the Pacific Northwest.} These forests have very high
productivity, and so to ensure that results were not unduly influenced
by geographic sampling bias, we tried a version of the model where data
were weighted according to forested land area within each Koeppen
climate zone. Results were similar between the weighted and unweighted
model, so, to avoid problems of over-fitting, the weighted model was
dropped, and results from this are not presented here.

Models were run for total annual productivity against annual climate
variables, and for monthly growing season productivity, defined as total
productivity/length of growing season, against growing season climate
variables. For analyses on data within the growing season, only linear
models were specified.

To investigate the potential interactive effects of climate variables on
carbon fluxes, multivariate models were also specified. To ensure that
models were biologically meaningful, the terms included in the models
tested built on results from the univariate models. Modelling of
individual climate variables identified that the best predictors of
carbon fluxes were variables related to temperature. We therefore
decided to include mean annual temperature as a term in all multivariate
models. We first modelled the interaction effect between mean annual
temperature and mean annual precipitation, in order to capture climate
variation along the axes of temperature and water availability. Models
were tested for a significant interactive effect and a significant
additive effect. We then explored whether any other climate variable, in
combination with mean annual temperature, could significantly improve on
the combination of mean annual temperature and mean annual
precipitation. In specifying the range of models to test, climate
variables which were strongly correlated with temperature were dropped,
in order to capture the greatest range of variation in climate. For each
possible pairing of climate variables, two models were specified: a
model with the two climate variables showing an additive effect; and a
model with the two climate variables showing an interactive effect. As
before, plot nested within geographic area was included as a random
effect. Altitude was not considered. AIC values were calculated for the
models, and used to compare models. Models were considered to be
significantly better than the baseline MAT*MAP model if:\\
i) the AIC value of the model was smaller than the AIC value of the
baseline model by \textgreater{}2\\
ii) the r-squared value of the model was larger than the r-squared value
of the baseline model by \textgreater{}5

\emph{Validating models of component fluxes.} Comparison of component
fluxes is based on the assumption that components sum accurately to
estimates of larger fluxes. To test this, components of larger fluxes
were regressed against latitude, and the models used to generate a
series of point estimates along lines of best fit for each component.
The point estimates for smaller component fluxes were summed to generate
new ``stacked'' estimates of larger fluxes, which were then compared
against actual measurements of the larger flux. Confidence intervals for
the larger flux were calculated using the `bootMer' function from the
lme4 package \citep{bates_fitting_2015}. Stacked plots were generated
for:\\
1. GPP = NPP + R\textsubscript{auto}\\
2. NPP = ANPP + BNPP\\
3. ANPP = ANPP\textsubscript{foliage} + ANPP\textsubscript{woody}
\textsubscript{stem}\\
4. Total belowground carbon flux = BNPP + R\textsubscript{root}

\emph{Allocation to carbon fluxes along latitudinal gradients.}
Variation in allocation to component carbon fluxes along latitudinal
gradients was explored for a range of pairings: firstly, GPP:NPP,
ANPP:BNPP, and ANPP\textsubscript{foliage}:ANPP\textsubscript{woody}
\textsubscript{stem}; and secondly, the ratio of NPP to each of ANPP,
BNPP, ANPP\textsubscript{foliage}, and ANPP\textsubscript{woody}
\textsubscript{stem}. For each set of paired fluxes, measurements taken
at the same site and plot, and in the same year, were paired together,
and the ratio of each pair of measurements calculated. The ratios were
regressed against latitude and climate variables, using the linear model
specified above. Cook's distance analyses were carried out for each of
the models, and indicated that data from a few high-elevation sites were
having a disproportionate influence on the regressions. To account for
this, models were re-run using only data from sites \(\le\) 1000m.

\subsubsection{Results}\label{results}

In total, we analyzed 1228 records from 9 C flux variables taken from
forests that had experienced no major anthropogenic disturbances within
the past 100 years. These records represented a total of 154 distinct
geographic areas (Fig. 1, Table 2), across all forested biogeographic
and climate zones.

\emph{How does productivity vary with latitude?}

All major carbon fluxes increased linearly with decreasing latitude
(fig. 2).

\begin{figure}[H]
\includegraphics[width=1\linewidth]{effect_of_lat_transparent} \caption{Graphs to show primary productivity $(MgC~ha^-1~yr^-1)$ regressed against latitude. Lines of best fit are plotted according to the best model selected during analysis. All regressions are significant $(p<0.05)$.}\label{fig:unnamed-chunk-6}
\end{figure}

Latitude was a strong predictor for many of the carbon fluxes,
explaining 64\% of variation in GPP (n = 254, p\textless{}0.0001), 50\%
in NPP (n = 114, p\textless{}0.0001) and 45\% in ANPP (n = 259,
p\textless{}0.0001). For all fluxes, their relationship with latitude
was best predicted by the linear model.

\emph{Relationships and differences among fluxes.} In general, smaller
component fluxes summed approximately to larger fluxes across the
latitudinal gradient (fig. 3). That is, modelled estimates of GPP,
generated from the sum of NPP and R auto; NPP, generated from the sum of
ANPP and BNPP\textsubscript{root}; and ANPP, generated from the sum of
ANPP\textsubscript{foliage} and ANPP\textsubscript{woody}
\textsubscript{stem}, fell completely within the confidence intervals of
the regressions of field estimates of GPP, NPP and ANPP respectively.

\begin{figure}[H]
\includegraphics[width=1\linewidth]{combined_stacked} \caption{Graphs of primary productivity $(MgC~ha^-1~yr^-1)$ regressed against latitude. Lines of best fit are plotted according to the best model selected during analysis. All regressions are significant $(p<0.05)$. Plots 1 - 3 show two component fluxes; a larger flux, defined as the combination of the two component fluxes; and a modelled estimate of the sum of the two component fluxes. 95\% confidence intervals are plotted for the larger flux. Plot 4 shows three belowground fluxes, and a modelled estimate of the total belowground carbon flux}\label{fig:unnamed-chunk-7}
\end{figure}

We found no evidence that allocation between fluxes varied substantially
with latitude or climate. There were no significant results from
regressing ratios of carbon fluxes against latitude, or against any of
the climate variables.

R\textsuperscript{2} values were generally highest in the major fluxes,
and decreased in subsidiary fluxes (Supporting Information S2). Of the
major fluxes, R\textsubscript{auto} and GPP were the most strongly
explained by latitude and climate, with climate explaining at most 71\%
of variation in GPP, and 65\% in R\textsubscript{auto}. The proportion
of variation explained by climate and latitude decreased in NPP and
ANPP, with climate explaining at most 51\% of variation in NPP and 44\%
in ANPP. Of the major fluxes, BNPP\textsubscript{root} was the least
well explained by climate and latitude, with climate explaining at most
36\% of variation.

With the exception of ANPP\textsubscript{foliage}, the proportion of
variation explained by climate and latitude in subsidiary fluxes was
much lower. Climate explained at most 24\% of variation in
ANPP\textsubscript{woody} \textsubscript{stem}, 19\% in
BNPP\textsubscript{fine} \textsubscript{root}, and 27\% in
R\textsubscript{root}. In contrast, climate strongly explained variation
in ANPP\textsubscript{foliage}, with mean annual temperature explaining
58\% of variation. This pattern was also seen in the
R\textsuperscript{2} values for multivariate models.

\emph{How does productivity relate to MAT and MAP?} We focus first on
considering the relationship between productivity and MAT and MAP. MAT
and MAP are the most commonly reported site-level climate variables, and
much previous research into the effect of climate on forest productivity
has focused on these as key climate variables. MAT was a significant
(p\textless{}0.05) and strong predictor of productivity for all carbon
fluxes tested, with all fluxes showing a linear increase with
temperature (fig. 5). We found no support for a saturation point of
productivity with temperature.

MAP was found to be a significant (p\textless{}0.05) but poor predictor
of productivity, explaining, with the exception of
R\textsubscript{auto}, at most 37\% of variation in carbon flux. For the
majority of fluxes productivity was best predicted by a polynomial
model. Productivity increased with precipitation, up until a saturation
point at between 3000 and 4000mm annual precipitation, above which
productivity started to decrease (fig. 5). The notable exception to this
was GPP: the model indicated that GPP continued to increase with
precipitation up to measures of at least 5000mm annually
(p\textless{}0.0001, R\textsuperscript{2} = 0.33. Data above this point
was not available, but the model trend indicated that the saturation
point for this model would be around 5000mm MAP.

There was a significant interactive effect between MAT and MAP for GPP,
BNPP\textsubscript{root}, BNPP\textsubscript{fine} \textsubscript{root},
ANPP, ANPP\textsubscript{woody} \textsubscript{stem}, and
R\textsubscript{root} (fig.4). There was a significant additive effect
for R\textsubscript{auto}. NPP and ANPP\textsubscript{foliage} showed no
significant interactive or additive effect: including MAP as a second
variable did not improve on the model including only MAT.

For the variables which showed a significant interactive or additive
effect between MAT and MAP, no other climate variable, in combination
with MAT, significantly improved on that model. For NPP, there was a
significant interactive effect between MAT and water stress months, with
this model explaining nearly 5\% more variation in NPP than MAT alone.
However, for ANPP\textsubscript{foliage}, no multivariate model improved
on the univariate model including only MAT.

\begin{figure}[H]
\includegraphics[width=1\linewidth]{mat_map_interaction} \caption{Plots of primary productivity $(MgC~ha^-1~yr^-1)$ regressed against mean annual precipitation. Points are grouped into bins of 0 - 1000, 1001 - 2000, 2001 - 3000, and >3000mm mean annual precipitation, and lines of best fit plotted for mean annual precipitation values of 500, 1500, 2500, and 3500mm. All regressions are significant $(p<0.05)$.}\label{fig:unnamed-chunk-8}
\end{figure}

\emph{How does productivity relate to other climate variables?} Our
results indicated that productivity was most strongly explained by
temperature at the global scale, with temperature-related climate
variables coming out as strong predictors of productivity. In addition
to MAT, temperature seasonality, annual temperature range, and annual
frost days were consistently identified as good predictors of
productivity across fluxes.

We found a significant relationship between productivity and potential
evapotranspiration for all fluxes. ANPP\textsubscript{foliage},
BNPP\textsubscript{fine} \textsubscript{root} and R\textsubscript{root}
increased linearly with PET; however, all other fluxes showed a
polynomial relationship with PET (fig. 5). We found strong evidence for
a saturation point or peak with PET: productivity tended to increase at
values below 1000mm, before saturating between 1200 and 1700mm. There
was evidence that productivity begins to decrease at values above 1800mm
PET.

Vapour pressure deficit was a significant predictor of productivity for
all fluxes. BNPP\textsubscript{fine} \textsubscript{root} showed a
linear relationship with vapour pressure deficit (R\textsuperscript{2} =
0.07, p\textless{}0.05), but all other fluxes showed a polynomial
relationship (fig. 5). Productivity initially increased with vapour
pressure deficit, before saturating at around 0.8 kPa. At values above
0.8 kPa, productivity began to decrease.

All fluxes, with the exception of R\textsubscript{root}, showed a
positive linear relationship with solar radiation. Solar radiation
explained a low proportion of variability in productivity for all
fluxes, explaining less than 20\% of the variation in each flux, with
the exception of R\textsubscript{auto} (R\textsuperscript{2} = 0.26,
p\textless{}0.05).

Of the climate variables tested, annual wet days, aridity, cloud cover,
mean diurnal temperature range, precipitation seasonality, maximum
vapour pressure deficit and water stress months were poor or
non-significant explainers of variation in productivity, explaining less
than 20\% of the variation in each of the carbon fluxes.

\begin{figure}[H]
\includegraphics[width=1\linewidth]{combined_plots} \caption{Plots of primary productivity $(MgC~ha^-1~yr^-1)$ regressed against (a) mean annual temperature; (b) mean annual precipitation; (c) potential evapotranspiration, (d) vapour pressure deficit; (e) temperature seasonality; (f) length of growing season. Lines of best fit are plotted according to the best model selected during analysis. All regressions are significant $(p<0.05)$.}\label{fig:unnamed-chunk-9}
\end{figure}

\emph{What is the role of seasonality in explaining productivity?}
Temperature seasonality was a significant predictor of productivity. We
found a polynomial relationship with productivity for GPP, NPP, ANPP,
and R\textsubscript{root}. Productivity decreased rapidly as seasonality
increased, with the rate of decrease slowing as seasonality increased
(fig. 5). ANPP\textsubscript{foliage}, ANPP\textsubscript{woody}
\textsubscript{stem} and R\textsubscript{auto} decreased linearly with
temperature seasonality. Temperature seasonality was strongly correlated
with annual temperature range, and, as expected, all fluxes showed
almost identical responses to it. Productivity was highest where
temperature seasonality = 0, and at an annual temperature range of
15\(^\circ\)C or lower.

In contrast, there was no significant effect of precipitation
seasonality on productivity.

We found a significant relationship between length of growing season and
productivity, with all fluxes showing a linear increase in productivity
with length of growing season (fig. 5). Length of growing season was a
strong predictor of productivity, explaining 51\% of variation in GPP,
39\% of variation in NPP, and 34\% of variation in ANPP, but it was not
identified as a stronger predictor than MAT for any of the fluxes
analysed.

\emph{Within the growing season, how does productivity vary with
climate?} Within growing season months, we found that climate had a much
weaker effect on productivity. For each of temperature, precipitation,
PET, and solar radiation, we found a small effect of climate for certain
carbon fluxes. There was a small increase in productivity with
temperature and precipitation for ANPP (with temperature
R\textsuperscript{2} = 0.10, p\textless{}0.001; with precipitation
R\textsuperscript{2} = 0.04, p\textless{}0.05) and
ANPP\textsubscript{foliage} (with temperature R\textsuperscript{2} =
0.16, p\textless{}0.01; with precipitation R\textsuperscript{2} = 0.09,
p\textless{}0.05). Productivity increased with solar radiation for GPP
(R\textsuperscript{2} = 0.21, p\textless{}0.001), NPP
(R\textsuperscript{2} = 0.21, p\textless{}0.001), BNPP
(R\textsuperscript{2} = 0.16, p\textless{}0.001) and
BNPP\textsubscript{fine} \textsubscript{root} (R\textsuperscript{2} =
0.12, p\textless{}0.01), and with PET for GPP (R\textsuperscript{2} =
0.15, p\textless{}0.01), NPP (R\textsuperscript{2} = 0.18,
p\textless{}0.01), BNPP (R\textsuperscript{2} = 0.23,
p\textless{}0.0001), BNPP\textsubscript{fine} \textsubscript{root}
(R\textsuperscript{2} = 0.11, p\textless{}0.05), and
ANPP\textsubscript{woody} \textsubscript{stem} (R\textsuperscript{2} =
0.06, p\textless{}0.05).

\subsubsection{Discussion}\label{discussion}

We used a large global database (ForC), containing an unprecedented
amount of data representing all major forest types (Fig. 1) and the nine
most significant forest autotrophic carbon flux (FACF) variables (Table
2), to comprehensively explore how C cycling in mature, undisturbed
forests varies with latitude and climate on a global scale. We test a
suite of hypotheses, including many previously posed (Table 1), with
more data and standardisation for factors that have not always been
controlled for (\emph{e.g.}, stand age, flux components measured) to
gain a stronger understanding of climatic effects on FACF. We show that,
across all nine FACF variables analyzed, C cycling decreases linearly
with latitude (Figs. 2-3)--a finding that confirms multiple previous
studies (\textbf{REFS}) but contradicts the idea that productivity of
temperate forests rivals that of tropical forests (\textbf{REFS}). The
FACF variables increase in proportion to one another, with no
differences in allocation detectable at this global scale and with
component fluxes summing appropriately to larger fluxes (Fig. 3),
indicating no major, systematic omissions or overestimations of flux
components. However, we did detect a tendency for greater variability
among subsidiary C fluxes (e.g., \emph{\(ANPP_woody\)},
\emph{\(BNPP_{fine root}\)}; Fig. 2; \textbf{some SI table?};
\textbf{discussed below??}). Climate explains a significant proportion
of variation in all C fluxes analyzed (albiet less for subsidiary
fluxes), with temperature variables being the best predictors of FACF on
this global scale (Figs. 4-5). While other climate variables are
significant predictors of FACF, none of them improve on the explanatory
power of temperature-related variables in general or MAT specifically
(Fig. 5). Water availability is an important factor in explaining FACF
on a global scale: we find a positive influence of precipitation at low
MAP, with saturation at higher levels of MAP (Fig. 5b) and a significant
interaction between MAT and MAP for seven of the nine variables analyzed
(Fig. 4). Temperature seasonality and growing season length are closely
correlated with MAT and are strong predictors of FACF (Fig. 5e-f),
though growing season length doesn't improve upon MAT as a predictor.
Within the growing season, the influence of climate on C cycling is
smaller but still significant for a number of carbon fluxes
(\textbf{some SI table?}). These findings clarify the big picture of how
FACF varies with latitude and climate on a global scale.

Past studies have differed in their conclusions regarding the
relationship between productivity and latitude or its correlates
(\textbf{REFS})--quite possibly because of lack of standardization with
respect to stand age and disturbance history. Our findings indicate
that, among mature, undisturbed stands, FACF is unambiguously highest in
the tropical regions, and the relationship is approximately linear
(Figs. 2-3). This contrasts with suggestions that productivity of
temperate forests is similar to that of tropical forests
(\textbf{REFS}). Such observations may be attributable to the fact that
temperate forests are predominantly secondary and of younger age than
tropical forests (\textbf{REF}), so analyses comparing across
latitudinal gradients without strict control of stand age tend to be
comparing younger temperate stands with older tropical (and often
boreal) forests (\href{http://pure.iiasa.ac.at/id/eprint/15361/}{Poulter
et al. 2018-DOI:10.1594/PANGAEA.889943}). Because C allocation varies
with stand age (\textbf{REFS (KAT/NOBBY)}), age differences may
introduce systematic biases into analyses of FACF across latitude or
global climatic gradients. For example, woody productivity tends to be
higher in rapidly aggrading secondary stands than in old-growth forests,
where proportionally more C is allocated to respiration (\textbf{REFS})
{[}*purpose for respiration/ other compenents (\textbf{REFS????-
NOBBY?}){]}. By constraining our analysis to stands
\$\textbackslash{}geq\$100 years old with no record of major recent
disturbance, we have clarified the shape of the relationship between
FACF and latitude.

{[}\textbf{NOBBY, please pay special attention to this paragraph}{]}. We
show that FACF are broadly consistent in their responses to climate
drivers on the global scale (with the exception of some differences in
MAT-MAP interactions; Fig. 4), with no major trends in C allocation
among the variable pairs tested (Fig. 2; \textbf{Some SI table}).
Although variation in allocation has been observed along gradients of
elevation \citep{moser_elevation_2011} and water availability
\citep{newman_above-_2006}--along with non-climatic axes of stand age
\citep{litton_carbon_2007}, nutrient availability
\citep{litton_carbon_2007, gill_belowground_2016}, and forest structure
\citep{taylor_greater_2019}--variation in relation to climate is not
apparent at the global scale within ForC, which contains the bulk of
relevant data. Our conclusion, then, is that hypothesized gradients in
allocation along global climate gradients cannot currently be supported
for mature forests, although data quantity and standardization is
currently insufficent to rule out the possibility that such trends
exist. Of particular interest and significance are the relationships
amongst \(GPP\), net primary productivity (\(NPP\) and its components,
particularly \(ANPP_{woody-stem}\)), and respiration (\(R_{auto}\) and
components). There have been suggestions that tropical forests tend to
have low carbon use efficiency (\(CUE\)=
\(NPP\)/\(GPP\)=(\(GPP\)-\(R_{auto}\)/\(GPP\)), which are based on
observations of low \(CUE\) in old-growth tropical forests relative to
(mostly younger) extratropical forests
\citetext{\citealp{de_lucia_forest_2007}; \citealp[\href{https://www.researchgate.net/publication/227837530_The_productivity_metabolism_and_carbon_cycle_of_tropical_forest_vegetation}{Malhi
2012};][]{anderson-teixeira_carbon_2016}}, but our analysis suggests
that these low values might more appropriately be attributed to the fact
that these forests are old than to their tropical climate. Indeed,
\(CUE\) is known to decline with forest age
(\citep{de_lucia_forest_2007};\textbf{REFS(NOBBY)}) but appears to be
roughly independent of \(GPP\) \citep{litton_carbon_2007}. Among our
sites with relevant data, there is indication in our data that CUE or
\(ANPP_{woody-stem}\)/\(GPP\) increase with latitude (\textbf{some SI
table}). Careful methodological standardization across a consistent set
of mature forest sites will be necessary to determine whether any
climate-driven gradients in allocation exist at the global scale.

One interesting observation was that climate tends to explain more
variation in the major fluxes (\(GPP\), \(NPP\), \(R_{auto}\) - latitude
\$R\^{}2 \ge\$48\%) than in subsidiary fluxes (latitude
\(R^2\)\textless{}27\% for \(BNPP_{fine.root}\),
\(R_{auto-root}\),\(ANPP_{woody-stem}\)) (Fig. 2; \textbf{some SI
table?}). There are two non-exclusive potential explanations for this.
First, it may be that methodological variation is larger relative to
flux magnitude for some of the subsidiary fluxes. Belowground fluxes in
particular are difficult to quantify, and measurement methods for the
belowground fluxes considered here may be measured through fundamentally
different approaches (\emph{e.g.}, minirhizotrons, ingrowth cores, or
sequential coring for \(BNPP_{root-fine}\); root exclusion, stable
isotope tracking, or gas exchange of excised roots for
\(R_{auto-root}\)), and sampling depth is variable and often
insufficient to capture the full soil profile. \(ANPP_{woody-stem}\),
which is also poorly explained by latitude or climate, is more
straightforward to measure but is subject to variability introduced by
differences such as biomass allometries applied and minimum plant size
sampled. However, methodological variation and uncertainty affect all of
fluxes considered here--not necessarily any less than the
aforementioned, and some of the larger fluxes that vary more strongly
with respect to climate (\(ANPP\), \(NPP\)) are estimated by summing
uncertain component fluxes. Second, differences among variables in the
proportion of variation explained by climate may be attributable to more
dicrect climatic control over \(GPP\) than subsidiary fluxes. That is,
subsidiary fluxes may be shaped by climate both indirectly through its
influence on \(GPP\) and potentially respiration and directly through
any climatic influence on C allocation, as well as by many other local-
and regional-scale factors (\textbf{REFS}).

The latitudinal gradient in FACF (Figs. 2-3) is driven primarily by
temperature-related climate variables, and secondarily by moisture
availability (Figs. 4-5). Because many climate variables co-vary across
the latitudinal gradient, because climatic drivers affect forest carbon
flux on much shorter time scales than can be captured by annual climate
summary variables, and because both climatic conditions and C flux vary
intra- and inter-annually around the long-term means, it is not
appropriate to attempt to identify any one mean annual climate variable
as a mechanistic driver of FACF. However, it remains informative to
consider these relationships. We find that temperature-related climate
variables (\(MAT\), temperature seasonality, \ldots{}\textbf{LIST})
explain the highest proportion of variability in productivity, and among
these, \(MAT\) is generally the best predictor--perhaps because
site-specific MAT is recorded for the majority of sites in ForC, whereas
other variables are extracted from global gridded data products (Table
S1\textbf{?}). The effects of temperature are modified by moisture
availability, with reduced FACF under hot and dry conditions
(\emph{i.e.}, high PET, high deficit; Fig. 5c-d) and sometimes under
very high precipitaiton (Figs. 4, 5b). Negative effects of very high
precipitation on FACF have been previously observed (\textbf{REFS}) and
are attributable to nutrient and light limitations (\textbf{REFS}) Thus,
although temperature and water availability jointly and interactively
drive global-scale patterns of FACF.

FACF is reduced by climatic seasonality (Fig. 5e), and productivity is
shut down during cold- or dry- dormant seasons. To account for this, a
number of analyses seeking to characterize global-scale effects of
climate on productivity have examined the relationship of FACF per month
of the growing season with growing season climatic conditions
(\textbf{REFS}). We found that the sort of simple metric needed to
define growing season at a global scale was uncertain for temperature
and problematic for moisture (\textbf{WORK ON THIS}). A
temperature-defined growing season length had stong positive correlation
with FACF (Fig. 5f), but explained less variation than \(MAT\). Dividing
FACFs by growing season length to yield FACF per growing season month
removed the majority of climate-related variation, supporting the idea
that the latitudinal gradient in FACF is attributable more to shorter
growing seasons at high latitudes than to inherently lower rates of
photosyntheiss or respiration by high-latitude forests (\emph{{[}
Enquist GCB review{]}}). However, there remained a number of significant
correlations with growing season climatic conditions, suggesting that
climatic conditions remain influential within the growing season. We
conclude that while correcting for growing season length takes analyses
a step closer to mechanistic linkage of instantaneous C flux rates to
environmental conditions, it remains very crude relative to the the
timescales on which climate affects plant metabolism and does not
advance statistical predictive power. Rather, mechanistic accounting for
climatic effects on global productivity patterns requires models
representing physiologically meaningful timescales (\emph{e.g.},
\textbf{refs}).

Our analysis clarifies the big picture of how FACF varies with latitude
and climate on a global scale, with some important implications for how
forest carbon cycling relates to climate and, by extension, how it is
likely to respond to climatic warming. Contrary to some previous
suggestions, we find no support for non-linear trends in mature forest C
cycling with respect to latitude or \(MAT\), and no distinct trends in C
allocation across the global scale (Fig. 3). The implication is that
under warmer conditions with similar moisture availability--and within
the temperature range to which forest communities are adapated and
acclimatized--higher temperatures result in a generalized acceleration
of FACF, with no major shifts in C allocation among subsidiary fluxes.
Of course, actual climatic changes will result in very different sets of
conditions than represented across geographic gradients in climate, but
our analysis clarifies the big picture as to how carbon cycles through
forest ecosystems that have developed under relatively stable climatic
conditions. As we enter a period of accelerating climatic change,
understanding of the fundamental climatic controls on FACF sets a
foundation for understanding patterns of change. \textbf{{[}work on
this{]}}

\subsubsection{Acknowledgements}\label{acknowledgements}

Scholarly Studies ForestGEO Compilatin of the ForC database was
originally funded by DOE

\bibliography{library.bib,packages.bib}

\end{document}

### 2019-09-12 manuscript/manuscript draft.Rmd
fixed some math syntex

@@
 |--------------|-------------------|-------------|------------------------|-------|-------|
-| GPP | Annual gross primary production; annual uptake of carbon dioxide by an ecosystem | NA | Flux partitioning of eddy covariance | 243 | 49 |
-| NPP | Annual net primary production; the component of GPP that is stored in plant tissue; GPP minus ecosystem respiration | Foliage, branch, stem, coarse root and fine root | Direct measurement of annual increments of components | 92 | 42 |
-| ANPP | Aboveground net primary production | Foliage, stem, and optionally branch | Direct measurement of annual increments of components | 256 | 80 |
-| ANPP~foliage~ | Net primary production of foliage | Foliage | Direct measurement of litterfall, correcting for changes in leaf biomass when measured | 98 | 49 |
-| ANPP~woody~ ~stem~ | Net primary production of woody stems | Woody stems | Direct measurement of stem growth increment | 264 | 96 |
-| BNPP~root~ | Belowground net primary production | Coarse and fine roots | Direct measurement of one or more of: fine root turnover, soil cores, root ingrowth cores, minirhizotrons; indirect estimates of coarse roots using allometries based on aboveground stem increment measures | 101 | 48 |
-| BNPP~fine~ ~root~ | Net primary production of fine roots | Fine roots | Direct measurement of one or more of: minirhizotrons, fine root turnover, soil cores, root ingrowth cores | 88 | 41 |
-| R~auto~ | Annual autotrophic respiration, including above- and belowground components | Foliage, stem, and root | Chamber measurements of component gas exchange | 22 | 13 |
-| R~root~ | Annual root respiration | Coarse and fine roots | Measurement of root gas exchange | 64 | 26 |
+| $GPP$ | Annual gross primary production; annual uptake of carbon dioxide by an ecosystem | NA | Flux partitioning of eddy covariance | 243 | 49 |
+| $NPP$ | Annual net primary production; the component of GPP that is stored in plant tissue; GPP minus ecosystem respiration | Foliage, branch, stem, coarse root and fine root | Direct measurement of annual increments of components | 92 | 42 |
+| $ANPP$ | Aboveground net primary production | Foliage, stem, and optionally branch | Direct measurement of annual increments of components | 256 | 80 |
+| $ANPP_{foliage}$ | Net primary production of foliage | Foliage | Direct measurement of litterfall, correcting for changes in leaf biomass when measured | 98 | 49 |
+| $ANPP_{woody-stem}$ | Net primary production of woody stems | Woody stems | Direct measurement of stem growth increment | 264 | 96 |
+| $BNPP_{root}$  | Belowground net primary production | Coarse and fine roots | Direct measurement of one or more of: fine root turnover, soil cores, root ingrowth cores, minirhizotrons; indirect estimates of coarse roots using allometries based on aboveground stem increment measures | 101 | 48 |
+| $BNPP_{fine.root}$ | Net primary production of fine roots | Fine roots | Direct measurement of one or more of: minirhizotrons, fine root turnover, soil cores, root ingrowth cores | 88 | 41 |
+| $R_{auto}$ | Annual autotrophic respiration, including above- and belowground components | Foliage, stem, and root | Chamber measurements of component gas exchange | 22 | 13 |
+| $R_{root}$ | Annual root respiration | Coarse and fine roots | Measurement of root gas exchange | 64 | 26 |
 
@@
 
-As discussed above, estimates of NPP, ANPP, and BNPP are generated through summing measurements of their component parts.
+As discussed above, estimates of $NPP$, $ANPP$, and $BNPP$ are generated through summing measurements of their component parts.
 Since the components included in productivity estimates vary between studies, estimates of productivity were classified within the ForC database according to their components, and then filtered for analysis.
@@
 If not, we ran the linear model as the final model.
-R^2^ values were calculated for the best model.
-All R^2^ values presented here are marginal R^2^ values, and refer to the proportion of variation explained by only the fixed effects, unless otherwise specified.
+$R^2$ values were calculated for the best model.
+All $R^2$ values presented here are marginal $R^2$ values, and refer to the proportion of variation explained by only the fixed effects, unless otherwise specified.
 In addition, slope coefficients were calculated for the linear models.
@@
 
-R^2^ values were generally highest in the major fluxes, and decreased in subsidiary fluxes (Supporting Information S2).
+$R^2$ values were generally highest in the major fluxes, and decreased in subsidiary fluxes (Supporting Information S2).
 Of the major fluxes, R~auto~ and GPP were the most strongly explained by latitude and climate, with climate explaining at most 71% of variation in GPP, and 65% in R~auto~.
@@
 In contrast, climate strongly explained variation in ANPP~foliage~, with mean annual temperature explaining 58% of variation.
-This pattern was also seen in the R^2^ values for multivariate models.
+This pattern was also seen in the $R^2$ values for multivariate models.
 
@@
 5).
-The notable exception to this was GPP: the model indicated that GPP continued to increase with precipitation up to measures of at least 5000mm annually (p<0.0001, R^2^ = 0.33.
+The notable exception to this was GPP: the model indicated that GPP continued to increase with precipitation up to measures of at least 5000mm annually (p<0.0001, $R^2$ = 0.33.
 Data above this point was not available, but the model trend indicated that the saturation point for this model would be around 5000mm MAP.
@@
 Vapour pressure deficit was a significant predictor of productivity for all fluxes.
-BNPP~fine~ ~root~ showed a linear relationship with vapour pressure deficit (R^2^ = 0.07, p<0.05), but all other fluxes showed a polynomial relationship (fig.
+BNPP~fine~ ~root~ showed a linear relationship with vapour pressure deficit ($R^2$ = 0.07, p<0.05), but all other fluxes showed a polynomial relationship (fig.
 5).
@@
 All fluxes, with the exception of R~root~, showed a positive linear relationship with solar radiation.
-Solar radiation explained a low proportion of variability in productivity for all fluxes, explaining less than 20% of the variation in each flux, with the exception of R~auto~ (R^2^ = 0.26, p<0.05).
+Solar radiation explained a low proportion of variability in productivity for all fluxes, explaining less than 20% of the variation in each flux, with the exception of R~auto~ ($R^2$ = 0.26, p<0.05).
 
@@
 For each of temperature, precipitation, PET, and solar radiation, we found a small effect of climate for certain carbon fluxes.
-There was a small increase in productivity with temperature and precipitation for ANPP (with temperature R^2^ = 0.10, p<0.001; with precipitation R^2^ = 0.04, p<0.05) and ANPP~foliage~ (with temperature R^2^ = 0.16, p<0.01; with precipitation R^2^ = 0.09, p<0.05).
-Productivity increased with solar radiation for GPP (R^2^ = 0.21, p<0.001), NPP (R^2^ = 0.21, p<0.001), BNPP (R^2^ = 0.16, p<0.001) and BNPP~fine~ ~root~ (R^2^ = 0.12, p<0.01), and with PET for GPP (R^2^ = 0.15, p<0.01), NPP (R^2^ = 0.18, p<0.01), BNPP (R^2^ = 0.23, p<0.0001), BNPP~fine~ ~root~ (R^2^ = 0.11, p<0.05), and ANPP~woody~ ~stem~ (R^2^ = 0.06, p<0.05).
+There was a small increase in productivity with temperature and precipitation for ANPP (with temperature $R^2$ = 0.10, p<0.001; with precipitation $R^2$ = 0.04, p<0.05) and ANPP~foliage~ (with temperature $R^2$ = 0.16, p<0.01; with precipitation $R^2$ = 0.09, p<0.05).
+Productivity increased with solar radiation for GPP ($R^2$ = 0.21, p<0.001), NPP ($R^2$ = 0.21, p<0.001), BNPP ($R^2$ = 0.16, p<0.001) and BNPP~fine~ ~root~ ($R^2$ = 0.12, p<0.01), and with PET for GPP ($R^2$ = 0.15, p<0.01), NPP ($R^2$ = 0.18, p<0.01), BNPP ($R^2$ = 0.23, p<0.0001), BNPP~fine~ ~root~ ($R^2$ = 0.11, p<0.05), and ANPP~woody~ ~stem~ ($R^2$ = 0.06, p<0.05).
 
@@
 For example, woody productivity tends to be higher in rapidly aggrading secondary stands than in old-growth forests, where proportionally more C is allocated to respiration (**REFS**) [*purpose for respiration/ other compenents (**REFS????- NOBBY?**)].
-By constraining our analysis to stands $\\geq$100 years old with no record of major recent disturbance, we have clarified the shape of the relationship between FACF and latitude.
+By constraining our analysis to stands $\ge$ 100 years old with no record of major recent disturbance, we have clarified the shape of the relationship between FACF and latitude.
 
@@
 
-One interesting observation was that climate tends to explain more variation in the major fluxes ($GPP$, $NPP$, $R_{auto}$ - latitude $R^2 \ge$48%) than in subsidiary fluxes (latitude $R^2$<27% for $BNPP_{fine.root}$, $R_{auto-root}$,$ANPP_{woody-stem}$) (Fig.
+One interesting observation was that climate tends to explain more variation in the major fluxes ($GPP$, $NPP$, $R_{auto}$ - latitude $R^2\ge$ 48%) than in subsidiary fluxes (latitude $R^2$ <27% for $BNPP_{fine.root}$, $R_{auto-root}$,$ANPP_{woody-stem}$) (Fig.
 2; **some SI table?**).

### 2019-09-12 manuscript/manuscript draft.Rmd
re-wrote Table 1 questions as hypotheses

@@
   kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 12) %>%
-  pack_rows("1. How do FACF vary with latitude?", 1, 3, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  pack_rows("2. How do FACF relate to MAT and MAP?", 4, 6, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  pack_rows("3. How do FACF relate to other annual climate variables?", 7, 9, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  pack_rows("4. What is the role of seasonality?", 10, 13, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  pack_rows("5. Per month of the growing seaon, how do FACF vary with climate?", 14, 17, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("H1. FACF decreases linearly with latitude.", 1, 3, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("H2. FACF increases with MAT, with an interactive effect of MAP.", 4, 6, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("H3. FACF are strongly correlated with other annual climate variables.", 7, 9, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("H4. FACF is reduced under seasonal climates.", 10, 13, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("H5. Considering only growing season months, FACF correlates with climatic drivers.", 14, 17, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
   add_header_above(c(" ", " ", " ", "Forest autotrophic carbon fluxes (FACF)" = 9, " ")) %>%
@@
 | $GPP$ | Annual gross primary production; annual uptake of carbon dioxide by an ecosystem | NA | Flux partitioning of eddy covariance | 243 | 49 |
-| $NPP$ | Annual net primary production; the component of GPP that is stored in plant tissue; GPP minus ecosystem respiration | Foliage, branch, stem, coarse root and fine root | Direct measurement of annual increments of components | 92 | 42 |
+| $NPP$ | Annual net primary production; the component of GPP that is stored in plant tissue; $GPP - R_{auto}$ | Foliage, branch, stem, coarse root and fine root | Direct measurement of annual increments of components | 92 | 42 |
 | $ANPP$ | Aboveground net primary production | Foliage, stem, and optionally branch | Direct measurement of annual increments of components | 256 | 80 |
@@
 | $R_{auto}$ | Annual autotrophic respiration, including above- and belowground components | Foliage, stem, and root | Chamber measurements of component gas exchange | 22 | 13 |
-| $R_{root}$ | Annual root respiration | Coarse and fine roots | Measurement of root gas exchange | 64 | 26 |
+| $R_{root}$ | Annual root respiration | Coarse and fine roots | *Measurement of root gas exchange, root exclusion from soil respiration chambers* | 64 | 26 |
 
@@
 ForestGEO
-Compilatin of the ForC database was originally funded by DOE
+Compilation of the ForC database was originally funded by DOE

### 2019-09-12 manuscript/manuscript draft.Rmd
underscore for CO~2~

@@
 Forest productivity sets the energy available to heterotrophic communities, in turn influencing their abundance (**REFS**) and possibly diversity (**REFS**)--but probably not individual metabolic rate (**Anderson & Jetz 2005, but check more recent literature**).
-On the global scale, forests play a critical role in regulating atmospheric CO_2 and climate (**Bonan**).
-The total amount of CO_2 cycling through Earth's forests each year (*i.e.*, total gross primary production, $GPP$) is more than five times anthropogenic fossil fuel emissions, and net sequestration (~2.4 GtCyr^-1^) offsetting roughly 30% of these emissions [@pan_large_2011].
+On the global scale, forests play a critical role in regulating atmospheric CO~2~ and climate (**Bonan**).
+The total amount of CO~2~ cycling through Earth's forests each year (*i.e.*, total gross primary production, $GPP$) is more than five times anthropogenic fossil fuel emissions, and net sequestration (~2.4 GtCyr^-1^) offsetting roughly 30% of these emissions [@pan_large_2011].
 As atmospheric carbon dioxide levels continue to rise, with consequences for global climate, there is increasing recognition that proper protection and management of forest resources will have an important role to play in mitigating climate change (** REFS- e.g., IPCC 1.5C and land reports**).

### 2019-09-12 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd
added changes I'd previously made and forgot to push.

@@
 
-In order to approach these broad and complex issues, we simplify the major gaps in our knowledge to five key questions and corresponding specific predictions (Table 1).
+In order to approach these broad and complex issues, we simplify the major gaps in our knowledge to five broad hypotheses and corresponding specific predictions (Table 1).
 Firstly, we ask how forest autotrophic carbon fluxes (FACF) vary with latitude.
@@
   kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 12) %>%
-  pack_rows("H1. FACF decreases linearly with latitude.", 1, 3, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  pack_rows("H2. FACF increases with MAT, with an interactive effect of MAP.", 4, 6, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  pack_rows("H3. FACF are strongly correlated with other annual climate variables.", 7, 9, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  pack_rows("H4. FACF is reduced under seasonal climates.", 10, 13, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  pack_rows("H5. Considering only growing season months, FACF correlates with climatic drivers.", 14, 17, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("H1. FACFs decrease linearly with latitude, and in constant proportion.", 1, 3, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("H2. FACFs increase with MAT, with an interactive effect of MAP.", 4, 6, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("H3. FACFs are strongly correlated with other annual climate variables.", 7, 9, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("H4. FACFs are reduced under seasonal climates.", 10, 13, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("H5. Considering only growing season months, FACFs correlate with climatic drivers.", 14, 17, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
   add_header_above(c(" ", " ", " ", "Forest autotrophic carbon fluxes (FACF)" = 9, " ")) %>%
@@
 We test a suite of hypotheses, including many previously posed (Table 1), with more data and standardisation for factors that have not always been controlled for (*e.g.*, stand age, flux components measured) to gain a stronger understanding of climatic effects on FACF.
-We show that, across all nine FACF variables analyzed, C cycling decreases linearly with latitude (Figs.
-2-3)--a finding that confirms multiple previous studies (**REFS**) but contradicts the idea that productivity of temperate forests rivals that of tropical forests (**REFS**).
-The FACF variables increase in proportion to one another, with no differences in allocation detectable at this global scale and with component fluxes summing appropriately to larger fluxes (Fig.
+We show that, across all nine FACF variables analyzed, C cycling decreases linearly with latitude ((*H1.1*; Figs.
+2-3)--a finding that confirms multiple previous studies (**REFS**) but contradicts the idea that productivity of temperate forests rivals that of tropical forests (*H1.1.alt*; **REFS**).
+The FACF variables increase in proportion to one another (*H1.2*), with no differences in allocation detectable at this global scale and with component fluxes summing appropriately to larger fluxes (Fig.
 3), indicating no major, systematic omissions or overestimations of flux components.
-However, we did detect a tendency for greater variability among subsidiary C fluxes (e.g., *$ANPP_woody$*, *$BNPP_{fine root}$*; Fig.
+However, we did detect a tendency for greater variability among subsidiary C fluxes (e.g., *$ANPP_{woody}$*, *$BNPP_{fine.root}$*; Fig.
 2; **some SI table?**; **discussed below??**).
@@
 4-5).
-While other climate variables are significant predictors of FACF, none of them improve on the explanatory power of temperature-related variables in general or MAT specifically (Fig.
+While other climate variables are significant predictors of FACF (*H3*), none of them improve on the explanatory power of temperature-related variables in general or MAT specifically (Fig.
 5).
@@
 4).
-Temperature seasonality and growing season length are closely correlated with MAT and are strong predictors of FACF (Fig.
+Temperature seasonality and growing season length are closely correlated with MAT and are strong predictors of FACF (*H4*; Fig.
 5e-f), though growing season length doesn't improve upon MAT as a predictor.
-Within the growing season, the influence of climate on C cycling is smaller but still significant for a number of carbon fluxes (**some SI table?**).
+Within the growing season, the influence of climate on C cycling is smaller but still significant for a number of carbon fluxes (*H5*; **some SI table?**).
 These findings clarify the big picture of how FACF varies with latitude and climate on a global scale.
 
-Past studies have differed in their conclusions regarding the relationship between productivity and latitude or its correlates (**REFS**)--quite possibly because of lack of standardization with respect to stand age and disturbance history.
+Past studies have differed in their conclusions regarding the relationship between productivity and latitude or its correlates (Table 1, *H1*; **REFS**)--quite possibly because of lack of standardization with respect to stand age and disturbance history.
 Our findings indicate that, among mature, undisturbed stands, FACF is unambiguously highest in the tropical regions, and the relationship is approximately linear (Figs.
@@
 We show that FACF are broadly consistent in their responses to climate drivers on the global scale (with the exception of some differences in MAT-MAP interactions; Fig.
-4), with no major trends in C allocation among the variable pairs tested (Fig.
+4), with no major trends in C allocation among the variable pairs tested (Table 1, *H1.2*; Fig.
 2; **Some SI table**).
@@
 The latitudinal gradient in FACF (Figs.
-2-3) is driven primarily by temperature-related climate variables, and secondarily by moisture availability (Figs.
+2-3) is driven primarily by temperature-related climate variables, and secondarily by moisture availability (Table 1, *H2-H3*; Figs.
 4-5).
@@
 
-FACF is reduced by climatic seasonality (Fig.
+FACF is reduced by temperature seasonality (Table 1, *H4*; Fig.
 5e), and productivity is shut down during cold- or dry- dormant seasons.
-To account for this, a number of analyses seeking to characterize global-scale effects of climate on productivity have examined the relationship of FACF per month of the growing season with growing season climatic conditions (**REFS**).
+To account for this, a number of analyses seeking to characterize global-scale effects of climate on productivity have examined the relationship of FACF per month of the growing season with growing season climatic conditions (Table 1, *H5*; **REFS**).
 We found that the sort of simple metric needed to define growing season at a global scale was uncertain for temperature and problematic for moisture (**WORK ON THIS**).

### 2019-09-13 manuscript/manuscript draft.Rmd
update draft

@@
 ###Abstract
+[**very rough start:**]
+.....
+Here, we drew upon # records from the Global Forest Carbon Database (ForC), representing all major forest types and the nine most significant forest autotrophic carbon flux (FACF) variables, to comprehensively explore how C cycling in mature, undisturbed forests varies with latitude and climate on a global scale.
+We show that, across all FACF variables analyzed, C cycling decreases linearly with latitude --a finding that confirms multiple previous studies but contradicts the idea that productivity of temperate forests rivals that of tropical forests.
+The FACF variables increase in proportion to one another, with no differences in allocation detectable at this global scale, but differed in that latitude explained a lower proportion of variation among subsidiary fluxes (in particular, woody aboveground productivity and belowground net primary productivity, $BNPP$).
+Climate explained a significant proportion (#-#%) of variation in all C fluxes analyzed (less for subsidiary fluxes), with temperature variables in general and mean annual temperature ($MAT$) in particular being the best predictors of FACF on this global scale.
+While other climate variables (*e.g.*, displayed significant correlation with FACF, none of them improved significantly over the explanatory power $MAT$.
+The effects of temperature were modified by moisture availability, with reduced FACF under hot and dry conditions and sometimes under very high precipitaiton (especially for $BNPP$).
+FACF declined with temperature seasonality, but growing season length doesn't improve upon MAT as a predictor.
+Within the growing season, the influence of climate on C cycling is smaller but still significant for a number of carbon fluxes.
+These findings clarify the big picture of how FACF varies with latitude and climate on a global scale.
+As we enter a period of accelerating climatic change, understanding of the fundamental climatic controls on FACF sets a foundation for understanding patterns of change.
 
@@
 Forest productivity sets the energy available to heterotrophic communities, in turn influencing their abundance (**REFS**) and possibly diversity (**REFS**)--but probably not individual metabolic rate (**Anderson & Jetz 2005, but check more recent literature**).
-On the global scale, forests play a critical role in regulating atmospheric CO~2~ and climate (**Bonan**).
-The total amount of CO~2~ cycling through Earth's forests each year (*i.e.*, total gross primary production, $GPP$) is more than five times anthropogenic fossil fuel emissions, and net sequestration (~2.4 GtCyr^-1^) offsetting roughly 30% of these emissions [@pan_large_2011].
+On the global scale, forests play a critical role in regulating atmospheric CO_2 and climate (**Bonan**).
+The total amount of CO_2 cycling through Earth's forests each year (*i.e.*, total gross primary production, $GPP$) is more than five times anthropogenic fossil fuel emissions, and net sequestration (~2.4 GtCyr^-1^) offsetting roughly 30% of these emissions [@pan_large_2011].
 As atmospheric carbon dioxide levels continue to rise, with consequences for global climate, there is increasing recognition that proper protection and management of forest resources will have an important role to play in mitigating climate change (** REFS- e.g., IPCC 1.5C and land reports**).
@@
   pack_rows("H4. FACFs are reduced under seasonal climates.", 10, 13, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  pack_rows("H5. Considering only growing season months, FACFs correlate with climatic drivers.", 14, 17, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("H5. Considering only growing season months, FACF correlates with climatic drivers.", 14, 17, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
   add_header_above(c(" ", " ", " ", "Forest autotrophic carbon fluxes (FACF)" = 9, " ")) %>%
@@
 We test a suite of hypotheses, including many previously posed (Table 1), with more data and standardisation for factors that have not always been controlled for (*e.g.*, stand age, flux components measured) to gain a stronger understanding of climatic effects on FACF.
-We show that, across all nine FACF variables analyzed, C cycling decreases linearly with latitude ((*H1.1*; Figs.
+We show that, across all nine FACF variables analyzed, C cycling decreases linearly with latitude (*H1.1*; Figs.
 2-3)--a finding that confirms multiple previous studies (**REFS**) but contradicts the idea that productivity of temperate forests rivals that of tropical forests (*H1.1.alt*; **REFS**).
@@
 However, we did detect a tendency for greater variability among subsidiary C fluxes (e.g., *$ANPP_{woody}$*, *$BNPP_{fine.root}$*; Fig.
-2; **some SI table?**; **discussed below??**).
+2; **some SI table?**).
 Climate explains a significant proportion of variation in all C fluxes analyzed (albiet less for subsidiary fluxes), with temperature variables being the best predictors of FACF on this global scale (Figs.
@@
 We show that FACF are broadly consistent in their responses to climate drivers on the global scale (with the exception of some differences in MAT-MAP interactions; Fig.
-4), with no major trends in C allocation among the variable pairs tested (Table 1, *H1.2*; Fig.
+4), with no major trends in C allocation among the variable pairs tested (Table 1, *H1*; Fig.
 2; **Some SI table**).
@@
 5f), but explained less variation than $MAT$.
-Dividing FACFs by growing season length to yield FACF per growing season month removed the majority of climate-related variation, supporting the idea that the latitudinal gradient in FACF is attributable more to shorter growing seasons at high latitudes than to inherently lower rates of photosyntheiss or respiration by high-latitude forests (*[ Enquist GCB review]*).
+Dividing FACFs by growing season length to yield FACF per growing season month removed the majority of climate-related variation, supporting the idea that the latitudinal gradient in FACF is attributable more to shorter growing seasons at high latitudes than to inherently lower rates of photosyntheiss or respiration by high-latitude forests (*[ Enquist et al.
+2007 GCB- but check]*).
 However, there remained a number of significant correlations with growing season climatic conditions, suggesting that climatic conditions remain influential within the growing season.

### 2019-09-13 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@
 Associated data, such as stand age, measurement methodologies, and disturbance history, are also included.
-The database was significantly expanded since the publication of [@anderson-teixeira_forc:_2018] through integration with the Global Database of Soil Respiration [@bond-lamberty_global_2010].
+The database was significantly expanded since the publication of [@anderson-teixeira_forc:_2018] through integration with the Global Database of Soil Respiration Database [@bond-lamberty_global_2010].
 Additional targeted literature searches were conducted to identify any further available data on primary productivity, with particular focus on mature forests in temperate and boreal regions.

### 2019-09-16 manuscript/manuscript draft.Rmd
Suggested reorganization
Hi @teixeirak and @beckybanbury -

A really solid start but right now the introduction is really long and confusingly organized. This is one possible reorganization that might help you, and the readers, structure things a bit more clearly. Note that I've bolded each paragraph's **topic sentence**, which I find a really useful way to easily see the sequence of main points. Anyway, whether or not you use this, I do think some similar step is needed before specific wordsmithing. Happy to help provide further feedback if useful! Thanks.

@@
 
-Carbon fixation, allocation, and metabolism by trees sets the basis for energy and material flows in forest ecosystems and defines their interactions with Earth's changing climate.
-Forest autotrophic carbon flux (FACF) influences all organic matter stocks in ecosystems and is linked to cycling of energy, water, and nutrients (**REFS**).
-Forest productivity sets the energy available to heterotrophic communities, in turn influencing their abundance (**REFS**) and possibly diversity (**REFS**)--but probably not individual metabolic rate (**Anderson & Jetz 2005, but check more recent literature**).
-On the global scale, forests play a critical role in regulating atmospheric CO_2 and climate (**Bonan**).
+**Forests play a critical role in regulating atmospheric CO_2 and climate** (**Bonan**).
 The total amount of CO_2 cycling through Earth's forests each year (*i.e.*, total gross primary production, $GPP$) is more than five times anthropogenic fossil fuel emissions, and net sequestration (~2.4 GtCyr^-1^) offsetting roughly 30% of these emissions [@pan_large_2011].
 As atmospheric carbon dioxide levels continue to rise, with consequences for global climate, there is increasing recognition that proper protection and management of forest resources will have an important role to play in mitigating climate change (** REFS- e.g., IPCC 1.5C and land reports**).
-Given the importance of understanding broad-scale patterns in forest autotrophic carbon flux (FACF), there have been many meta-analyses on the theme.
+Carbon fixation, allocation, and metabolism by trees set the basis for energy and material flows in forest ecosystems and define their interactions with Earth's changing climate.
+
+**Forests' autotrophic carbon flux (FACF, bbl:define) influences all organic matter stocks in ecosystems and is linked to cycling of energy, water, and nutrients** (**REFS**).
+Forest productivity sets the energy ultimately available to heterotrophic communities, in turn influencing their abundance (**REFS**) and possibly diversity (**REFS**)--but probably not individual metabolic rate (**Anderson & Jetz 2005, but check more recent literature**).
 However, our ability to draw general macroscopic conclusions regarding global variation in multiple productivity variables with respect to climate has been limited in that these analyses often mix forests that vary in stand age, disturbance history, and/or management status; do not always sufficiently parse related variables (*e.g.*, combining net primary productivity records with and without belowground components); and typically consider only one or a few variables at a time.
-The recent development of a global forest carbon database synthesizing multiple variables and including records of stand history (ForC; [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc:_2018]) opens up the possibility for a standardized analysis of global scale variation in multiple components of forest productivity and the principle climatic drivers of these patterns.
 
-On a global scale, the productivity of forests varies with latitude, showing a general trend of decreasing productivity with latitude [@beer_terrestrial_2010; @jung_global_2011].
+**The productivity of forests varies with latitude**, showing a general trend of decreasing productivity with latitude [@beer_terrestrial_2010; @jung_global_2011].
 Studies agree that productivity is lowest in the boreal regions, and increases into the temperate regions [@luyssaert_co_2007; @huston_global_2009; @beer_terrestrial_2010; @jung_global_2011].
@@
 Other studies have chosen to focus exclusively on above-ground net primary productivity (ANPP), finding evidence of a weak negative relationship between ANPP and latitude [@huston_global_2009; @gillman_latitude_2015].
-An understanding of the global patterns of productivity is important; however elucidating the drivers of these patterns will be even more valuable in expanding our understanding of global carbon cycling.
-Primary productivity can be influenced by many factors, which often act across a range of scales, and may show interactive effects with each other.
-On a local scale, stand age [@litton_carbon_2007; @gillman_latitude_2015], management [@simova_enigma_2017]; nutrient availability [@aragao_above-_2009]; and altitude [@girardin_net_2010; @malhi_variation_2017] all impact forest productivity.
-On a global scale, we expect that productivity is most strongly influenced by broad climatic gradients.
 
-We know from previous research that climate is a significant driver of productivity across broad spatial scales [@cleveland_relationships_2011].
+**Climate is a significant driver of productivity across broad spatial scales** [@cleveland_relationships_2011].
 The majority of studies have focused on exploring the relationships between productivity and mean annual temperature (MAT) and mean annual precipitation (MAP), as the most commonly reported site-level climate variables.
@@
 However, as with latitude, the shape of those relationships is not always clear, and, again, is complicated by the use of different measures of productivity across studies.
-There is support across multiple studies for the hypothesis that various measures of primary productivity saturate at high levels of MAP, though the saturation points identified vary from 1500mm [@luyssaert_co_2007] up to 2445mm MAP [@schuur_productivity_2003].
+Various measures of primary productivity saturate at high levels of MAP, though the saturation points identified vary from 1500mm [@luyssaert_co_2007] up to 2445mm MAP [@schuur_productivity_2003].
 Studies of the influence of MAT on productivity are less conclusive.
@@
 In contrast, Larjavaara and Muller-Landau [-@larjavaara_temperature_2012], find that increases in GPP saturate at approximately 25$^\circ$C MAT, while Schuur [-@schuur_productivity_2003] shows that NPP increases linearly with temperature.
-The influence of these climate variables on productivity is further complicated by the possibility of interactive effects occurring between them.
+bbl:maybe remove following sentences The influence of these climate variables on productivity is further complicated by the possibility of interactive effects occurring between them.
 Taylor et al.
@@
 
-While MAT and MAP are important climate variables, they do not capture all aspects of climate that may influence productivity, and may be insufficient to capture the breadth of climatic effects on productivity [@cleveland_relationships_2011].
+**Primary productivity can be influenced by many other factors as well**, which often act across a range of scales, and may show interactive effects with each other [@cleveland_relationships_2011].
+On a local scale, stand age [@litton_carbon_2007; @gillman_latitude_2015], management [@simova_enigma_2017]; nutrient availability [@aragao_above-_2009]; and altitude [@girardin_net_2010; @malhi_variation_2017] all impact forest productivity.
+On a global scale, we expect that productivity is most strongly influenced by broad climatic gradients.
 There is evidence that productivity also responds to variables such as cloud cover [@taylor_temperature_2017], solar radiation [@fyllas_solar_2017], and potential evapotranspiration [@kerkhoff_plant_2005] in potentially significant ways.
@@
 
+**The recent development of a global forest carbon database synthesizing multiple variables and including records of stand history (ForC; [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc:_2018]) opens up the possibility for a standardized analysis of global scale variation** in multiple components of forest productivity and the principle climatic drivers of these patterns.
 In order to approach these broad and complex issues, we simplify the major gaps in our knowledge to five broad hypotheses and corresponding specific predictions (Table 1).
-Firstly, we ask how forest autotrophic carbon fluxes (FACF) vary with latitude.
+First, we ask how forest autotrophic carbon fluxes (FACF) vary with latitude.
 We then test how these fluxes relate to MAT and MAP, and additionally how they respond to other, less well studied, climate variables.
 We finally consider the relationship between FACF and seasonality, considering the role of seasonality in explaining variation in carbon fluxes, and the influence of climate on FACF per month of the growing season.
-
-Resolving these questions requires data which, firstly, control for variables influencing productivity, including stand age, disturbance regime, elevation, and methodology; and secondly includes all major carbon fluxes.
-Here we use a comprehensive global database of forest carbon fluxes to explore the above questions for nine carbon fluxes, allowing for an in-depth exploration of the effect of climate on global productivity.
+We use a comprehensive global database of forest carbon fluxes to explore the above questions for nine carbon fluxes, allowing for an in-depth exploration of the effect of climate on global productivity.
 

### 2019-09-16 manuscript/manuscript draft.Rmd
Discussion suggestions
Lots of great material here in the discussion, but again, but benefit (IMO) from clearer structure. There's a lot of re-stating of the results; try to keep this to a minimum.

@@
 1) and the nine most significant forest autotrophic carbon flux (FACF) variables (Table 2), to comprehensively explore how C cycling in mature, undisturbed forests varies with latitude and climate on a global scale.
-We test a suite of hypotheses, including many previously posed (Table 1), with more data and standardisation for factors that have not always been controlled for (*e.g.*, stand age, flux components measured) to gain a stronger understanding of climatic effects on FACF.
+We test a suite of hypotheses, including many previously posed (Table 1), standardising for factors that have not always been controlled for (*e.g.*, stand age, flux components measured) to gain a stronger understanding of climatic effects on FACF.
+
+bbl: Most of these two paragraphs (following) just restate the results. Honestly I would remove them enitrely, but at the very least, condense into a single paragraph.
+
 We show that, across all nine FACF variables analyzed, C cycling decreases linearly with latitude (*H1.1*; Figs.
@@
 2; **some SI table?**).
-Climate explains a significant proportion of variation in all C fluxes analyzed (albiet less for subsidiary fluxes), with temperature variables being the best predictors of FACF on this global scale (Figs.
+
+**Climate explains a significant proportion of variation in all C fluxes analyzed** (albiet less for subsidiary fluxes), with temperature variables the best predictors of FACF at this global scale (Figs.
 4-5).
@@
 This contrasts with suggestions that productivity of temperate forests is similar to that of tropical forests (**REFS**).
-Such observations may be attributable to the fact that temperate forests are predominantly secondary and of younger age than tropical forests (**REF**), so analyses comparing across latitudinal gradients without strict control of stand age tend to be comparing younger temperate stands with older tropical (and often boreal) forests ([Poulter et al.
+Temperate forests tend to be younger than tropical forests (**REF**), so analyses comparing across latitudinal gradients without controlling for stand age risk confounding age with biome effects ([Poulter et al.
 2018-DOI:10.1594/PANGAEA.889943](http://pure.iiasa.ac.at/id/eprint/15361/)).
-Because C allocation varies with stand age (**REFS (KAT/NOBBY)**), age differences may introduce systematic biases into analyses of FACF across latitude or global climatic gradients.
+In addition, because C allocation varies with stand age (**REFS (KAT/NOBBY)**), age differences may introduce systematic biases into analyses of FACF across latitude or global climatic gradients.
 For example, woody productivity tends to be higher in rapidly aggrading secondary stands than in old-growth forests, where proportionally more C is allocated to respiration (**REFS**) [*purpose for respiration/ other compenents (**REFS????- NOBBY?**)].
-By constraining our analysis to stands $\ge$ 100 years old with no record of major recent disturbance, we have clarified the shape of the relationship between FACF and latitude.
 
@@
 Our conclusion, then, is that hypothesized gradients in allocation along global climate gradients cannot currently be supported for mature forests, although data quantity and standardization is currently insufficent to rule out the possibility that such trends exist.
+
+bbl: remove this sentence?
 Of particular interest and significance are the relationships amongst $GPP$, net primary productivity ($NPP$ and its components, particularly $ANPP_{woody-stem}$), and respiration ($R_{auto}$ and components).
@@
 Indeed, $CUE$ is known to decline with forest age ([@de_lucia_forest_2007];**REFS(NOBBY)**) but appears to be roughly independent of $GPP$ [@litton_carbon_2007].
-Among our sites with relevant data, there is indication in our data that CUE or $ANPP_{woody-stem}$/$GPP$ increase with latitude (**some SI table**).
+Among our sites with relevant data, there is indication that CUE or $ANPP_{woody-stem}$/$GPP$ increase with latitude (**some SI table**).
 Careful methodological standardization across a consistent set of mature forest sites will be necessary to determine whether any climate-driven gradients in allocation exist at the global scale.
@@
 2; **some SI table?**).
-There are two non-exclusive potential explanations for this.
+There are two, non-exclusive, potential explanations for this.
 First, it may be that methodological variation is larger relative to flux magnitude for some of the subsidiary fluxes.
 Belowground fluxes in particular are difficult to quantify, and measurement methods for the belowground fluxes considered here may be measured through fundamentally different approaches (*e.g.*, minirhizotrons, ingrowth cores, or sequential coring for $BNPP_{root-fine}$; root exclusion, stable isotope tracking, or gas exchange of excised roots for $R_{auto-root}$), and sampling depth is variable and often insufficient to capture the full soil profile.
-$ANPP_{woody-stem}$, which is also poorly explained by latitude or climate, is more straightforward to measure but is subject to variability introduced by differences such as biomass allometries applied and minimum plant size sampled.
+$ANPP_{woody-stem}$, which is also poorly explained by latitude or climate, is more straightforward to measure but is subject to variability introduced by differences such as biomass allometries applied and minimum plant size sampled (bbl: cite e.g.
+Huntzinger?).
 However, methodological variation and uncertainty affect all of fluxes considered here--not necessarily any less than the aforementioned, and some of the larger fluxes that vary more strongly with respect to climate ($ANPP$, $NPP$) are estimated by summing uncertain component fluxes.
@@
 
+bbl: this seems really interesting and novel, consider expanding the following paragraph
+
 FACF is reduced by temperature seasonality (Table 1, *H4*; Fig.
@@
 
-Our analysis clarifies the big picture of how FACF varies with latitude and climate on a global scale, with some important implications for how forest carbon cycling relates to climate and, by extension, how it is likely to respond to climatic warming.
-Contrary to some previous suggestions, we find no support for non-linear trends in mature forest C cycling with respect to latitude or $MAT$, and no distinct trends in C allocation across the global scale (Fig.
+Our analysis clarifies how FACF varies (bbl: is FACF singular or plural?
+Be consistent) with latitude and climate on a global scale, with some important implications for how forest carbon cycling relates to climate and, by extension, how it is likely to respond to climatic warming.
+We find no support for non-linear trends in mature forest C cycling with respect to latitude or $MAT$, and no distinct trends in C allocation across the global scale (Fig.
 3).
 The implication is that under warmer conditions with similar moisture availability--and within the temperature range to which forest communities are adapated and acclimatized--higher temperatures result in a generalized acceleration of FACF, with no major shifts in C allocation among subsidiary fluxes.
-Of course, actual climatic changes will result in very different sets of conditions than represented across geographic gradients in climate, but our analysis clarifies the big picture as to how carbon cycles through forest ecosystems that have developed under relatively stable climatic conditions.
+Of course, actual climatic changes will result in very different sets of conditions than represented across geographic gradients in climate, but our analysis clarifies how carbon cycles through contemporary forest ecosystems.
 As we enter a period of accelerating climatic change, understanding of the fundamental climatic controls on FACF sets a foundation for understanding patterns of change.

### 2019-09-16 manuscript/manuscript draft.Rmd
updated Table 1 and caption

@@
 hypothesis_table <- read.csv("hypothesis_table.csv", stringsAsFactors = FALSE, check.names = FALSE)
-kable(hypothesis_table, booktabs = TRUE, caption = "Hypothesis table") %>%
+kable(hypothesis_table, booktabs = TRUE, caption = "**Summary of hypotheses, corresponding specific predictions, and results.** Direction of significant relationships is indicated by '+' and '-', and  shape is summarized as linear (L), concave up (CU), and concave down (CD).
+'n.s.' indicates that the relationship is not signficant.") %>%
   kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 12) %>%

### 2019-09-16 manuscript/manuscript draft.Rmd
started adding Nobby's feedback

@@
 **Forests' autotrophic carbon flux (FACF, bbl:define) influences all organic matter stocks in ecosystems and is linked to cycling of energy, water, and nutrients** (**REFS**).
-Forest productivity sets the energy ultimately available to heterotrophic communities, in turn influencing their abundance (**REFS**) and possibly diversity (**REFS**)--but probably not individual metabolic rate (**Anderson & Jetz 2005, but check more recent literature**).
+Forest productivity sets the energy ultimately available to heterotrophic organisms, in turn influencing their abundance (**REFS**) and possibly diversity (**REFS**)--but probably not individual metabolic rate (**Anderson & Jetz 2005, but check more recent literature**).
 However, our ability to draw general macroscopic conclusions regarding global variation in multiple productivity variables with respect to climate has been limited in that these analyses often mix forests that vary in stand age, disturbance history, and/or management status; do not always sufficiently parse related variables (*e.g.*, combining net primary productivity records with and without belowground components); and typically consider only one or a few variables at a time.

### 2019-09-17 manuscript/manuscript draft.Rmd

@@
 
+bbl: here or in the discussion, note that this is a fairly stringent test: it's easy for sub-fluxes not to sum up! (Extensive citations from EC literature for example.)
+
 ```{r echo=FALSE, out.width='100%', fig.cap = "Graphs of primary productivity $(MgC~ha^-1~yr^-1)$ regressed against latitude.
@@
 With the exception of ANPP~foliage~, the proportion of variation explained by climate and latitude in subsidiary fluxes was much lower.
-Climate explained at most 24% of variation in ANPP~woody~ ~stem~, 19% in BNPP~fine~ ~root~, and 27% in R~root~.
+Climate explained at most 24% of the variation in ANPP~woody~ ~stem~, 19% in BNPP~fine~ ~root~, and 27% in R~root~.
 In contrast, climate strongly explained variation in ANPP~foliage~, with mean annual temperature explaining 58% of variation.
@@
 
-*How does productivity relate to MAT and MAP?* We focus first on considering the relationship between productivity and MAT and MAP.
-MAT and MAP are the most commonly reported site-level climate variables, and much previous research into the effect of climate on forest productivity has focused on these as key climate variables.
+*How does productivity relate to MAT and MAP?* MAT and MAP are the most commonly reported site-level climate variables, and much previous research into the effect of climate on forest productivity has focused on these as key climate variables.
 MAT was a significant (p<0.05) and strong predictor of productivity for all carbon fluxes tested, with all fluxes showing a linear increase with temperature (fig.
@@
 There was a significant additive effect for R~auto~.
-NPP and ANPP~foliage~ showed no significant interactive or additive effect: including MAP as a second variable did not improve on the model including only MAT.
+NPP and ANPP~foliage~ showed no significant interactive or additive effect.
 
@@
 We found strong evidence for a saturation point or peak with PET: productivity tended to increase at values below 1000mm, before saturating between 1200 and 1700mm.
-There was evidence that productivity begins to decrease at values above 1800mm PET.
+There was also evidence that productivity begins to decrease at values above 1800mm PET.
 
@@
 *What is the role of seasonality in explaining productivity?* Temperature seasonality was a significant predictor of productivity.
-We found a polynomial relationship with productivity for GPP, NPP, ANPP, and R~root~.
-Productivity decreased rapidly as seasonality increased, with the rate of decrease slowing as seasonality increased (fig.
+GPP, NPP, ANPP, and R~root~ exhibited a polynomial relationship with seasonality (fig.
 5).
@@
 Productivity was highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower.
-
 In contrast, there was no significant effect of precipitation seasonality on productivity.

### 2019-09-17 manuscript/manuscript draft.Rmd
finished integrating Nobby's changes to text
(references still need to be done)

@@
 **Primary productivity can be influenced by many other factors as well**, which often act across a range of scales, and may show interactive effects with each other [@cleveland_relationships_2011].
-On a local scale, stand age [@litton_carbon_2007; @gillman_latitude_2015], management [@simova_enigma_2017]; nutrient availability [@aragao_above-_2009]; and altitude [@girardin_net_2010; @malhi_variation_2017] all impact forest productivity.
+On a local scale, stand age [@litton_carbon_2007; @gillman_latitude_2015], biodiversity (Liang et al.
+2019.
+Positive biodiversity-productivity relationship predominant in global forests.
+Science 354, Issue 6309, aaf8957), management [@simova_enigma_2017]; nutrient availability [@aragao_above-_2009]; and altitude [@girardin_net_2010; @malhi_variation_2017] all impact forest productivity.
 On a global scale, we expect that productivity is most strongly influenced by broad climatic gradients.
@@
 Among our sites with relevant data, there is indication that CUE or $ANPP_{woody-stem}$/$GPP$ increase with latitude (**some SI table**).
-Careful methodological standardization across a consistent set of mature forest sites will be necessary to determine whether any climate-driven gradients in allocation exist at the global scale.
+Additional measurements with careful methodological standardization across a consistent set of mature forest sites will be necessary to determine whether any climate-driven gradients in allocation exist at the global scale.
 
@@
 Second, differences among variables in the proportion of variation explained by climate may be attributable to more dicrect climatic control over $GPP$ than subsidiary fluxes.
-That is, subsidiary fluxes may be shaped by climate both indirectly through its influence on $GPP$ and potentially respiration and directly through any climatic influence on C allocation, as well as by many other local- and regional-scale factors (**REFS**).
+That is, subsidiary fluxes may be shaped by climate both indirectly through its influence on $GPP$ and respiration and directly through any climatic influence on C allocation, as well as by many other local- and regional-scale factors (**REFS**).
 

### 2019-09-17 manuscript/manuscript draft.Rmd
integrating Nobby's comments- Discussion
just missing intro

@@
 Temperate forests tend to be younger than tropical forests (**REF**), so analyses comparing across latitudinal gradients without controlling for stand age risk confounding age with biome effects ([Poulter et al.
-2018-DOI:10.1594/PANGAEA.889943](http://pure.iiasa.ac.at/id/eprint/15361/)).
-In addition, because C allocation varies with stand age (**REFS (KAT/NOBBY)**), age differences may introduce systematic biases into analyses of FACF across latitude or global climatic gradients.
-For example, woody productivity tends to be higher in rapidly aggrading secondary stands than in old-growth forests, where proportionally more C is allocated to respiration (**REFS**) [*purpose for respiration/ other compenents (**REFS????- NOBBY?**)].
+2018-DOI:10.1594/PANGAEA.889943](http://pure.iiasa.ac.at/id/eprint/15361/); Potapov et al.
+2008.
+Mapping the World’s Intact Forest Landscapes by Remote Sensing.
+Ecology and Society 13 (2), 51.
+).
+In addition, because C allocation varies with stand age [@de_lucia_forest_2007] (**See Nobby's comment in manuscript-draft_NK.pdf**), age differences may introduce systematic biases into analyses of FACF across latitude or global climatic gradients.
+For example, woody productivity tends to be higher in rapidly aggrading secondary stands than in old-growth forests, where proportionally more C is allocated to respiration (** Nobby AFM paper: Understanding the controls over forest carbon use efficiency on small spatial scales: Effects of forest disturbance and tree diversity**) [*purpose for respiration/ other compenents (**See Nobby's comment in manuscript-draft_NK.pdf**)].
 
-[**NOBBY, please pay special attention to this paragraph**].
 We show that FACF are broadly consistent in their responses to climate drivers on the global scale (with the exception of some differences in MAT-MAP interactions; Fig.
@@
 There have been suggestions that tropical forests tend to have low carbon use efficiency ($CUE$= $NPP$/$GPP$=($GPP$-$R_{auto}$/$GPP$), which are based on observations of low $CUE$ in old-growth tropical forests relative to (mostly younger) extratropical forests [@de_lucia_forest_2007; [Malhi 2012](https://www.researchgate.net/publication/227837530_The_productivity_metabolism_and_carbon_cycle_of_tropical_forest_vegetation); @anderson-teixeira_carbon_2016], but our analysis suggests that these low values might more appropriately be attributed to the fact that these forests are old than to their tropical climate.
-Indeed, $CUE$ is known to decline with forest age ([@de_lucia_forest_2007];**REFS(NOBBY)**) but appears to be roughly independent of $GPP$ [@litton_carbon_2007].
+Indeed, $CUE$ is known to decline with forest age ([@de_lucia_forest_2007]; [Collalti & Prentice 2019](https://academic.oup.com/treephys/article-abstract/39/8/1473/5423351?redirectedFrom=fulltext)) but appears to be roughly independent of $GPP$ [@litton_carbon_2007].
 Among our sites with relevant data, there is indication that CUE or $ANPP_{woody-stem}$/$GPP$ increase with latitude (**some SI table**).

### 2019-09-17 manuscript/manuscript draft.Rmd
finished integrating Nobby's comments
There are a few citation lists not yet copied over, but I've added notes to see comments in his .pdf.

@@
 
-**Forests' autotrophic carbon flux (FACF, bbl:define) influences all organic matter stocks in ecosystems and is linked to cycling of energy, water, and nutrients** (**REFS**).
-Forest productivity sets the energy ultimately available to heterotrophic organisms, in turn influencing their abundance (**REFS**) and possibly diversity (**REFS**)--but probably not individual metabolic rate (**Anderson & Jetz 2005, but check more recent literature**).
+**Forests' autotrophic carbon flux (FACF, bbl:define) influences all organic matter stocks in ecosystems and is linked to cycling of energy, water, and nutrients** (**REFS; *Piao et al.
+2010.
+Forest annual carbon cost: a global‐scale analysis of autotrophic respiration.
+Ecology 91(3), 652-661Ƴ* **).
+Forest productivity sets the energy ultimately available to heterotrophic organisms (including microbes), in turn influencing their abundance (**REFS**) and possibly diversity ([Waide et al.
+2009](https://www.annualreviews.org/doi/abs/10.1146/annurev.ecolsys.30.1.257)**REFS**)--but probably not individual metabolic rate (**Anderson & Jetz 2005, but check more recent literature**).
 However, our ability to draw general macroscopic conclusions regarding global variation in multiple productivity variables with respect to climate has been limited in that these analyses often mix forests that vary in stand age, disturbance history, and/or management status; do not always sufficiently parse related variables (*e.g.*, combining net primary productivity records with and without belowground components); and typically consider only one or a few variables at a time.

### 2019-09-17 manuscript/manuscript-draft.tex
finished integrating Nobby's comments
There are a few citation lists not yet copied over, but I've added notes to see comments in his .pdf.

@@
 
+{[}\textbf{very rough start:}{]} \ldots{}.. Here, we drew upon \#
+records from the Global Forest Carbon Database (ForC), representing all
+major forest types and the nine most significant forest autotrophic
+carbon flux (FACF) variables, to comprehensively explore how C cycling
+in mature, undisturbed forests varies with latitude and climate on a
+global scale. We show that, across all FACF variables analyzed, C
+cycling decreases linearly with latitude --a finding that confirms
+multiple previous studies but contradicts the idea that productivity of
+temperate forests rivals that of tropical forests. The FACF variables
+increase in proportion to one another, with no differences in allocation
+detectable at this global scale, but differed in that latitude explained
+a lower proportion of variation among subsidiary fluxes (in particular,
+woody aboveground productivity and belowground net primary productivity,
+\(BNPP\)). Climate explained a significant proportion (\#-\#\%) of
+variation in all C fluxes analyzed (less for subsidiary fluxes), with
+temperature variables in general and mean annual temperature (\(MAT\))
+in particular being the best predictors of FACF on this global scale.
+While other climate variables (\emph{e.g.}, displayed significant
+correlation with FACF, none of them improved significantly over the
+explanatory power \(MAT\). The effects of temperature were modified by
+moisture availability, with reduced FACF under hot and dry conditions
+and sometimes under very high precipitaiton (especially for \(BNPP\)).
+FACF declined with temperature seasonality, but growing season length
+doesn't improve upon MAT as a predictor. Within the growing season, the
+influence of climate on C cycling is smaller but still significant for a
+number of carbon fluxes. These findings clarify the big picture of how
+FACF varies with latitude and climate on a global scale. As we enter a
+period of accelerating climatic change, understanding of the fundamental
+climatic controls on FACF sets a foundation for understanding patterns
+of change.
+
 \subsubsection{Introduction}\label{introduction}
 
-Carbon fixation, allocation, and metabolism by trees sets the basis for
-energy and material flows in forest ecosystems and defines their
-interactions with Earth's changing climate. Forest autotrophic carbon
-flux (FACF) influences all organic matter stocks in ecosystems and is
-linked to cycling of energy, water, and nutrients (\textbf{REFS}).
-Forest productivity sets the energy available to heterotrophic
-communities, in turn influencing their abundance (\textbf{REFS}) and
-possibly diversity (\textbf{REFS})--but probably not individual
-metabolic rate (\textbf{Anderson \& Jetz 2005, but check more recent
-literature}). On the global scale, forests play a critical role in
-regulating atmospheric CO\_2 and climate (\textbf{Bonan}). The total
-amount of CO\_2 cycling through Earth's forests each year (\emph{i.e.},
-total gross primary production, \(GPP\)) is more than five times
-anthropogenic fossil fuel emissions, and net sequestration
-(\textasciitilde{}2.4 GtCyr\textsuperscript{-1}) offsetting roughly 30\%
-of these emissions \citep{pan_large_2011}. As atmospheric carbon dioxide
-levels continue to rise, with consequences for global climate, there is
-increasing recognition that proper protection and management of forest
-resources will have an important role to play in mitigating climate
-change (** REFS- e.g., IPCC 1.5C and land reports**). Given the
-importance of understanding broad-scale patterns in forest autotrophic
-carbon flux (FACF), there have been many meta-analyses on the theme.
+\textbf{Forests play a critical role in regulating atmospheric CO\_2 and
+climate} (\textbf{Bonan}). The total amount of CO\_2 cycling through
+Earth's forests each year (\emph{i.e.}, total gross primary production,
+\(GPP\)) is more than five times anthropogenic fossil fuel emissions,
+and net sequestration (\textasciitilde{}2.4 GtCyr\textsuperscript{-1})
+offsetting roughly 30\% of these emissions \citep{pan_large_2011}. As
+atmospheric carbon dioxide levels continue to rise, with consequences
+for global climate, there is increasing recognition that proper
+protection and management of forest resources will have an important
+role to play in mitigating climate change (** REFS- e.g., IPCC 1.5C and
+land reports**). Carbon fixation, allocation, and metabolism by trees
+set the basis for energy and material flows in forest ecosystems and
+define their interactions with Earth's changing climate.
+
+\textbf{Forests' autotrophic carbon flux (FACF, bbl:define) influences
+all organic matter stocks in ecosystems and is linked to cycling of
+energy, water, and nutrients} (\textbf{REFS; \emph{Piao et al. 2010.
+Forest annual carbon cost: a global‐scale analysis of autotrophic
+respiration. Ecology 91(3), 652-661Ƴ} }). Forest productivity sets the
+energy ultimately available to heterotrophic organisms (including
+microbes), in turn influencing their abundance (\textbf{REFS}) and
+possibly diversity
+(\href{https://www.annualreviews.org/doi/abs/10.1146/annurev.ecolsys.30.1.257}{Waide
+et al. 2009}\textbf{REFS})--but probably not individual metabolic rate
+(\textbf{Anderson \& Jetz 2005, but check more recent literature}).
 However, our ability to draw general macroscopic conclusions regarding
@@
 primary productivity records with and without belowground components);
-and typically consider only one or a few variables at a time. The recent
-development of a global forest carbon database synthesizing multiple
-variables and including records of stand history (ForC;
-\citep{anderson-teixeira_carbon_2016, anderson-teixeira_forc:_2018})
-opens up the possibility for a standardized analysis of global scale
-variation in multiple components of forest productivity and the
-principle climatic drivers of these patterns.
-
-On a global scale, the productivity of forests varies with latitude,
-showing a general trend of decreasing productivity with latitude
+and typically consider only one or a few variables at a time.
+
+\textbf{The productivity of forests varies with latitude}, showing a
+general trend of decreasing productivity with latitude
 \citep{beer_terrestrial_2010, jung_global_2011}. Studies agree that
@@
 (ANPP), finding evidence of a weak negative relationship between ANPP
-and latitude \citep{huston_global_2009, gillman_latitude_2015}. An
-understanding of the global patterns of productivity is important;
-however elucidating the drivers of these patterns will be even more
-valuable in expanding our understanding of global carbon cycling.
-Primary productivity can be influenced by many factors, which often act
-across a range of scales, and may show interactive effects with each
-other. On a local scale, stand age
-\citep{litton_carbon_2007, gillman_latitude_2015}, management
-\citep{simova_enigma_2017}; nutrient availability
-\citep{aragao_above-_2009}; and altitude
-\citep{girardin_net_2010, malhi_variation_2017} all impact forest
-productivity. On a global scale, we expect that productivity is most
-strongly influenced by broad climatic gradients.
-
-We know from previous research that climate is a significant driver of
-productivity across broad spatial scales
-\citep{cleveland_relationships_2011}. The majority of studies have
-focused on exploring the relationships between productivity and mean
-annual temperature (MAT) and mean annual precipitation (MAP), as the
-most commonly reported site-level climate variables. These variables
-have the advantage that they describe broad trends in temperature and
-water availability, and therefore capture a lot of global-scale
-variation in climate. There is strong evidence that both MAT and MAP
-show significant positive relationships with productivity
+and latitude \citep{huston_global_2009, gillman_latitude_2015}.
+
+\textbf{Climate is a significant driver of productivity across broad
+spatial scales} \citep{cleveland_relationships_2011}. The majority of
+studies have focused on exploring the relationships between productivity
+and mean annual temperature (MAT) and mean annual precipitation (MAP),
+as the most commonly reported site-level climate variables. These
+variables have the advantage that they describe broad trends in
+temperature and water availability, and therefore capture a lot of
+global-scale variation in climate. There is strong evidence that both
+MAT and MAP show significant positive relationships with productivity
 \citep{chu_does_2016}. However, as with latitude, the shape of those
 relationships is not always clear, and, again, is complicated by the use
-of different measures of productivity across studies. There is support
-across multiple studies for the hypothesis that various measures of
-primary productivity saturate at high levels of MAP, though the
+of different measures of productivity across studies. Various measures
+of primary productivity saturate at high levels of MAP, though the
 saturation points identified vary from 1500mm \citep{luyssaert_co_2007}
@@
 \citeyearpar{schuur_productivity_2003} shows that NPP increases linearly
-with temperature. The influence of these climate variables on
-productivity is further complicated by the possibility of interactive
-effects occurring between them. Taylor et al.
+with temperature. bbl:maybe remove following sentences The influence of
+these climate variables on productivity is further complicated by the
+possibility of interactive effects occurring between them. Taylor et al.
 \citeyearpar{taylor_temperature_2017} showed that increased MAP had a
@@
 
-While MAT and MAP are important climate variables, they do not capture
-all aspects of climate that may influence productivity, and may be
-insufficient to capture the breadth of climatic effects on productivity
-\citep{cleveland_relationships_2011}. There is evidence that
+\textbf{Primary productivity can be influenced by many other factors as
+well}, which often act across a range of scales, and may show
+interactive effects with each other
+\citep{cleveland_relationships_2011}. On a local scale, stand age
+\citep{litton_carbon_2007, gillman_latitude_2015}, biodiversity (Liang
+et al. 2019. Positive biodiversity-productivity relationship predominant
+in global forests. Science 354, Issue 6309, aaf8957), management
+\citep{simova_enigma_2017}; nutrient availability
+\citep{aragao_above-_2009}; and altitude
+\citep{girardin_net_2010, malhi_variation_2017} all impact forest
+productivity. On a global scale, we expect that productivity is most
+strongly influenced by broad climatic gradients. There is evidence that
 productivity also responds to variables such as cloud cover
@@
 
-In order to approach these broad and complex issues, we simplify the
-major gaps in our knowledge to five key questions and corresponding
-specific predictions (Table 1). Firstly, we ask how forest autotrophic
-carbon fluxes (FACF) vary with latitude. We then test how these fluxes
-relate to MAT and MAP, and additionally how they respond to other, less
-well studied, climate variables. We finally consider the relationship
-between FACF and seasonality, considering the role of seasonality in
-explaining variation in carbon fluxes, and the influence of climate on
-FACF per month of the growing season.
-
-Resolving these questions requires data which, firstly, control for
-variables influencing productivity, including stand age, disturbance
-regime, elevation, and methodology; and secondly includes all major
-carbon fluxes. Here we use a comprehensive global database of forest
-carbon fluxes to explore the above questions for nine carbon fluxes,
-allowing for an in-depth exploration of the effect of climate on global
+\textbf{The recent development of a global forest carbon database
+synthesizing multiple variables and including records of stand history
+(ForC;
+\citep{anderson-teixeira_carbon_2016, anderson-teixeira_forc:_2018})
+opens up the possibility for a standardized analysis of global scale
+variation} in multiple components of forest productivity and the
+principle climatic drivers of these patterns. In order to approach these
+broad and complex issues, we simplify the major gaps in our knowledge to
+five broad hypotheses and corresponding specific predictions (Table 1).
+First, we ask how forest autotrophic carbon fluxes (FACF) vary with
+latitude. We then test how these fluxes relate to MAT and MAP, and
+additionally how they respond to other, less well studied, climate
+variables. We finally consider the relationship between FACF and
+seasonality, considering the role of seasonality in explaining variation
+in carbon fluxes, and the influence of climate on FACF per month of the
+growing season. We use a comprehensive global database of forest carbon
+fluxes to explore the above questions for nine carbon fluxes, allowing
+for an in-depth exploration of the effect of climate on global
 productivity.
@@
 
-\caption{\label{tab:unnamed-chunk-4}Hypothesis table}
+\caption{\label{tab:unnamed-chunk-4}**Summary of hypotheses, corresponding specific predictions, and results.** Direction of significant relationships is indicated by '+' and '-', and  shape is summarized as linear (L), concave up (CU), and concave down (CD).
+'n.s.' indicates that the relationship is not signficant.}
 \centering
@@
 \cmidrule(l{3pt}r{3pt}){4-12}
-Hypotheses \& Specific Predictions & Related references & Overall & GPP & NPP & ANPP & BNPP & \$ANPP\_\{foliage\}\$ & ANPP woody stem & BNPP fine root & R auto root & R auto & Support\\
+Hypotheses \& Specific Predictions & Related references & Overall & GPP & NPP & ANPP & BNPP & ANPP foliage & ANPP woody stem & BNPP fine root & R auto root & R auto & Support\\
 \midrule
@@
 \hline
-\multicolumn{1}{l}{\textbf{1. How do FACF vary with latitude?}}\\
+\multicolumn{1}{l}{\textbf{H1. FACFs decrease linearly with latitude, and in constant proportion.}}\\
 \hspace{1em}1.1. FACFs decrease linearly with latitude (L-) & Luyssaert et al. (2007) & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Fig. 2-3\\
@@
 2-3\\
-\hspace{1em}1.2. Allocation to FACFs varies with latitude & Litton et al. (2007); DeLucia et al. (2007) & no & - & no & no & no & no & no & no & - & - & \\
+\hspace{1em}1.2. Allocation of GPP to subsidiary fluxes varies with latitude & Litton et al. (2007); DeLucia et al. (2007) & no & - & no & no & no & no & no & no & - & - & \\
 \addlinespace[1em]
 \hline
-\multicolumn{1}{l}{\textbf{2. How do FACF relate to MAT and MAP?}}\\
+\multicolumn{1}{l}{\textbf{H2. FACFs increase with MAT, with an interactive effect of MAP.}}\\
 \hspace{1em}2.1. FACFs increase linearly with MAT (L+) & Schuur (2003) & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Fig. 5\\
@@
 FACFs increase with precipitation but saturate or decrease at very high levels (CD+) & Luyssaert et al.
-(2007); Schuur (2003) & yes & yes & yes & yes & yes & yes & () & L+ & L+ & yes & Fig.
+(2007); Schuur (2003) & yes & yes & yes & yes & yes & yes & n.s.
+& L+ & L+ & yes & Fig.
 5\\
-\hspace{1em}2.3. There is a postive interaction between temperature and precipitation (I+) & Taylor et al. (2016) & (yes) & yes & L+ & yes & I- & L+ & yes & I- & yes & yes & Fig. 4\\
+\hspace{1em}2.3. There is a postive interaction between temperature and precipitation (I+) & Taylor et al. (2016) & (yes) & yes & L+ & yes & L- & L+ & yes & L- & yes & yes & Fig. 4\\
 \addlinespace[1em]
 \hline
-\multicolumn{1}{l}{\textbf{3. How do FACF relate to other annual climate variables?}}\\
+\multicolumn{1}{l}{\textbf{H3. FACFs are strongly correlated with other annual climate variables.}}\\
 \hspace{1em}3.1. FACFs increases with PET, but saturates or decreases at high levels (CD+) &  & (yes) & yes & yes & yes & yes & L+ & yes & L+ & yes & L+ & Fig. 5\\
 \hspace{1em}3.2. FACFs increase with vapour pressure deficit, but saturate or decrease at high levels (CD+) &  & (yes) & yes & yes & yes & yes & yes & yes & L+ & yes & yes & Fig. 5\\
-\hspace{1em}3.3. FACFs increase linearly with solar radiation (L+) &  & (yes) & yes & yes & yes & CD+ & yes & CD+ & yes & () & yes & \\
+\hspace{1em}3.3. FACFs increase linearly with solar radiation (L+) &  & (yes) & yes & yes & yes & CD+ & yes & CD+ & yes & n.s. & yes & \\
 \addlinespace[1em]
 \hline
-\multicolumn{1}{l}{\textbf{4. What is the role of seasonality?}}\\
+\multicolumn{1}{l}{\textbf{H4. FACFs are reduced under seasonal climates.}}\\
 \hspace{1em}4.1. FACFs decrease linearly with temperature seasonality (L-) &  & yes & CU+ & CU+ & CU+ & yes & yes & yes & yes & CU+ & CU+ & Fig. 5\\
-\hspace{1em}4.2. FACFs decrease linearly with  precipitation seasonality (L-) &  & no & () & () & () & () & () & () & () & () & () & \\
+\hspace{1em}4.2. FACFs decrease linearly with  precipitation seasonality (L-) &  & no & n.s. & n.s. & n.s. & n.s. & n.s. & n.s. & n.s. & n.s. & n.s. & \\
 \hspace{1em}4.3.
@@
 \hline
-\multicolumn{1}{l}{\textbf{5. Per month of the growing seaon, how do FACF vary with climate?}}\\
-\hspace{1em}5.1. Increase with temperature (L+) & Michaletz et al. (2014) & (mixed) & () & () & yes & () & yes & () & () & () & () & \\
-\hspace{1em}5.2. Increase with PET (L+) &  & (yes) & yes & yes & () & yes & () & yes & yes & () & () & \\
-\hspace{1em}5.3. Increase with precipitation (L+) &  & no & () & () & yes & () & yes & () & () & () & () & \\
-\hspace{1em}5.4. Increase with solar radiation (L+) &  & (mixed) & yes & yes & () & yes & () & () & yes & () & () & \\
+\multicolumn{1}{l}{\textbf{H5. Considering only growing season months, FACF correlates with climatic drivers.}}\\
+\hspace{1em}5.1. Increase with temperature (L+) & Michaletz et al. (2014) & (mixed) & n.s. & n.s. & yes & n.s. & yes & n.s. & n.s. & n.s. & n.s. & \\
+\hspace{1em}5.2. Increase with PET (L+) &  & (yes) & yes & yes & n.s. & yes & n.s. & yes & yes & n.s. & n.s. & \\
+\hspace{1em}5.3. Increase with precipitation (L+) &  & no & n.s. & n.s. & yes & n.s. & yes & n.s. & n.s. & n.s. & n.s. & \\
+\hspace{1em}5.4. Increase with solar radiation (L+) &  & (mixed) & yes & yes & n.s. & yes & n.s. & n.s. & yes & n.s. & n.s. & \\
 \bottomrule
@@
 of \citep{anderson-teixeira_forc:_2018} through integration with the
-Global Database of Soil Respiration \citep{bond-lamberty_global_2010}.
-Additional targeted literature searches were conducted to identify any
-further available data on primary productivity, with particular focus on
-mature forests in temperate and boreal regions. ForC currently contains
-29730 records from 4979 plots, representing 20 distinct ecozones across
-all forested biogeographic and climate zones. We used ForC v3.0,
-archived on Zenodo with DOI 10.5281/zenodo.3403855.
+Global Database of Soil Respiration Database
+\citep{bond-lamberty_global_2010}. Additional targeted literature
+searches were conducted to identify any further available data on
+primary productivity, with particular focus on mature forests in
+temperate and boreal regions. ForC currently contains 29730 records from
+4979 plots, representing 20 distinct ecozones across all forested
+biogeographic and climate zones. We used ForC v3.0, archived on Zenodo
+with DOI 10.5281/zenodo.3403855.
 
@@
 \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
-GPP\strut
+\(GPP\)\strut
 \end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
@@
 \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
-NPP\strut
+\(NPP\)\strut
 \end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
 Annual net primary production; the component of GPP that is stored in
-plant tissue; GPP minus ecosystem respiration\strut
+plant tissue; \(GPP - R_{auto}\)\strut
 \end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
@@
 \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
-ANPP\strut
+\(ANPP\)\strut
 \end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
@@
 \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
-ANPP\textsubscript{foliage}\strut
+\(ANPP_{foliage}\)\strut
 \end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
@@
 \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
-ANPP\textsubscript{woody} \textsubscript{stem}\strut
+\(ANPP_{woody-stem}\)\strut
 \end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
@@
 \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
-BNPP\textsubscript{root}\strut
+\(BNPP_{root}\)\strut
 \end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
@@
 \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
-BNPP\textsubscript{fine} \textsubscript{root}\strut
+\(BNPP_{fine.root}\)\strut
 \end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
@@
 \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
-R\textsubscript{auto}\strut
+\(R_{auto}\)\strut
 \end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
@@
 \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
-R\textsubscript{root}\strut
+\(R_{root}\)\strut
 \end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
@@
 \end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-Measurement of root gas exchange\strut
+\emph{Measurement of root gas exchange, root exclusion from soil
+respiration chambers}\strut
 \end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
@@
 
-As discussed above, estimates of NPP, ANPP, and BNPP are generated
-through summing measurements of their component parts. Since the
-components included in productivity estimates vary between studies,
+As discussed above, estimates of \(NPP\), \(ANPP\), and \(BNPP\) are
+generated through summing measurements of their component parts. Since
+the components included in productivity estimates vary between studies,
 estimates of productivity were classified within the ForC database
@@
 accepted the polynomial as the best model. If not, we ran the linear
-model as the final model. R\textsuperscript{2} values were calculated
-for the best model. All R\textsuperscript{2} values presented here are
-marginal R\textsuperscript{2} values, and refer to the proportion of
-variation explained by only the fixed effects, unless otherwise
-specified. In addition, slope coefficients were calculated for the
-linear models.
+model as the final model. \(R^2\) values were calculated for the best
+model. All \(R^2\) values presented here are marginal \(R^2\) values,
+and refer to the proportion of variation explained by only the fixed
+effects, unless otherwise specified. In addition, slope coefficients
+were calculated for the linear models.
 
@@
 
+bbl: here or in the discussion, note that this is a fairly stringent
+test: it's easy for sub-fluxes not to sum up! (Extensive citations from
+EC literature for example.)
+
 \begin{figure}[H]
@@
 
-R\textsuperscript{2} values were generally highest in the major fluxes,
-and decreased in subsidiary fluxes (Supporting Information S2). Of the
-major fluxes, R\textsubscript{auto} and GPP were the most strongly
-explained by latitude and climate, with climate explaining at most 71\%
-of variation in GPP, and 65\% in R\textsubscript{auto}. The proportion
-of variation explained by climate and latitude decreased in NPP and
-ANPP, with climate explaining at most 51\% of variation in NPP and 44\%
-in ANPP. Of the major fluxes, BNPP\textsubscript{root} was the least
-well explained by climate and latitude, with climate explaining at most
-36\% of variation.
+\(R^2\) values were generally highest in the major fluxes, and decreased
+in subsidiary fluxes (Supporting Information S2). Of the major fluxes,
+R\textsubscript{auto} and GPP were the most strongly explained by
+latitude and climate, with climate explaining at most 71\% of variation
+in GPP, and 65\% in R\textsubscript{auto}. The proportion of variation
+explained by climate and latitude decreased in NPP and ANPP, with
+climate explaining at most 51\% of variation in NPP and 44\% in ANPP. Of
+the major fluxes, BNPP\textsubscript{root} was the least well explained
+by climate and latitude, with climate explaining at most 36\% of
+variation.
 
@@
 variation explained by climate and latitude in subsidiary fluxes was
-much lower. Climate explained at most 24\% of variation in
+much lower. Climate explained at most 24\% of the variation in
 ANPP\textsubscript{woody} \textsubscript{stem}, 19\% in
@@
 in ANPP\textsubscript{foliage}, with mean annual temperature explaining
-58\% of variation. This pattern was also seen in the
-R\textsuperscript{2} values for multivariate models.
-
-\emph{How does productivity relate to MAT and MAP?} We focus first on
-considering the relationship between productivity and MAT and MAP. MAT
-and MAP are the most commonly reported site-level climate variables, and
-much previous research into the effect of climate on forest productivity
-has focused on these as key climate variables. MAT was a significant
+58\% of variation. This pattern was also seen in the \(R^2\) values for
+multivariate models.
+
+\emph{How does productivity relate to MAT and MAP?} MAT and MAP are the
+most commonly reported site-level climate variables, and much previous
+research into the effect of climate on forest productivity has focused
+on these as key climate variables. MAT was a significant
 (p\textless{}0.05) and strong predictor of productivity for all carbon
@@
 precipitation up to measures of at least 5000mm annually
-(p\textless{}0.0001, R\textsuperscript{2} = 0.33. Data above this point
-was not available, but the model trend indicated that the saturation
-point for this model would be around 5000mm MAP.
+(p\textless{}0.0001, \(R^2\) = 0.33. Data above this point was not
+available, but the model trend indicated that the saturation point for
+this model would be around 5000mm MAP.
 
@@
 for R\textsubscript{auto}. NPP and ANPP\textsubscript{foliage} showed no
-significant interactive or additive effect: including MAP as a second
-variable did not improve on the model including only MAT.
+significant interactive or additive effect.
 
@@
 values below 1000mm, before saturating between 1200 and 1700mm. There
-was evidence that productivity begins to decrease at values above 1800mm
-PET.
+was also evidence that productivity begins to decrease at values above
+1800mm PET.
 
@@
 all fluxes. BNPP\textsubscript{fine} \textsubscript{root} showed a
-linear relationship with vapour pressure deficit (R\textsuperscript{2} =
-0.07, p\textless{}0.05), but all other fluxes showed a polynomial
-relationship (fig. 5). Productivity initially increased with vapour
-pressure deficit, before saturating at around 0.8 kPa. At values above
-0.8 kPa, productivity began to decrease.
+linear relationship with vapour pressure deficit (\(R^2\) = 0.07,
+p\textless{}0.05), but all other fluxes showed a polynomial relationship
+(fig. 5). Productivity initially increased with vapour pressure deficit,
+before saturating at around 0.8 kPa. At values above 0.8 kPa,
+productivity began to decrease.
 
@@
 fluxes, explaining less than 20\% of the variation in each flux, with
-the exception of R\textsubscript{auto} (R\textsuperscript{2} = 0.26,
+the exception of R\textsubscript{auto} (\(R^2\) = 0.26,
 p\textless{}0.05).
@@
 \emph{What is the role of seasonality in explaining productivity?}
-Temperature seasonality was a significant predictor of productivity. We
-found a polynomial relationship with productivity for GPP, NPP, ANPP,
-and R\textsubscript{root}. Productivity decreased rapidly as seasonality
-increased, with the rate of decrease slowing as seasonality increased
-(fig. 5). ANPP\textsubscript{foliage}, ANPP\textsubscript{woody}
-\textsubscript{stem} and R\textsubscript{auto} decreased linearly with
-temperature seasonality. Temperature seasonality was strongly correlated
-with annual temperature range, and, as expected, all fluxes showed
-almost identical responses to it. Productivity was highest where
-temperature seasonality = 0, and at an annual temperature range of
-15\(^\circ\)C or lower.
-
-In contrast, there was no significant effect of precipitation
-seasonality on productivity.
+Temperature seasonality was a significant predictor of productivity.
+GPP, NPP, ANPP, and R\textsubscript{root} exhibited a polynomial
+relationship with seasonality (fig. 5). ANPP\textsubscript{foliage},
+ANPP\textsubscript{woody} \textsubscript{stem} and R\textsubscript{auto}
+decreased linearly with temperature seasonality. Temperature seasonality
+was strongly correlated with annual temperature range, and, as expected,
+all fluxes showed almost identical responses to it. Productivity was
+highest where temperature seasonality = 0, and at an annual temperature
+range of 15\(^\circ\)C or lower. In contrast, there was no significant
+effect of precipitation seasonality on productivity.
 
@@
 carbon fluxes. There was a small increase in productivity with
-temperature and precipitation for ANPP (with temperature
-R\textsuperscript{2} = 0.10, p\textless{}0.001; with precipitation
-R\textsuperscript{2} = 0.04, p\textless{}0.05) and
-ANPP\textsubscript{foliage} (with temperature R\textsuperscript{2} =
-0.16, p\textless{}0.01; with precipitation R\textsuperscript{2} = 0.09,
-p\textless{}0.05). Productivity increased with solar radiation for GPP
-(R\textsuperscript{2} = 0.21, p\textless{}0.001), NPP
-(R\textsuperscript{2} = 0.21, p\textless{}0.001), BNPP
-(R\textsuperscript{2} = 0.16, p\textless{}0.001) and
-BNPP\textsubscript{fine} \textsubscript{root} (R\textsuperscript{2} =
-0.12, p\textless{}0.01), and with PET for GPP (R\textsuperscript{2} =
-0.15, p\textless{}0.01), NPP (R\textsuperscript{2} = 0.18,
-p\textless{}0.01), BNPP (R\textsuperscript{2} = 0.23,
-p\textless{}0.0001), BNPP\textsubscript{fine} \textsubscript{root}
-(R\textsuperscript{2} = 0.11, p\textless{}0.05), and
-ANPP\textsubscript{woody} \textsubscript{stem} (R\textsuperscript{2} =
-0.06, p\textless{}0.05).
+temperature and precipitation for ANPP (with temperature \(R^2\) = 0.10,
+p\textless{}0.001; with precipitation \(R^2\) = 0.04, p\textless{}0.05)
+and ANPP\textsubscript{foliage} (with temperature \(R^2\) = 0.16,
+p\textless{}0.01; with precipitation \(R^2\) = 0.09, p\textless{}0.05).
+Productivity increased with solar radiation for GPP (\(R^2\) = 0.21,
+p\textless{}0.001), NPP (\(R^2\) = 0.21, p\textless{}0.001), BNPP
+(\(R^2\) = 0.16, p\textless{}0.001) and BNPP\textsubscript{fine}
+\textsubscript{root} (\(R^2\) = 0.12, p\textless{}0.01), and with PET
+for GPP (\(R^2\) = 0.15, p\textless{}0.01), NPP (\(R^2\) = 0.18,
+p\textless{}0.01), BNPP (\(R^2\) = 0.23, p\textless{}0.0001),
+BNPP\textsubscript{fine} \textsubscript{root} (\(R^2\) = 0.11,
+p\textless{}0.05), and ANPP\textsubscript{woody} \textsubscript{stem}
+(\(R^2\) = 0.06, p\textless{}0.05).
 
@@
 forests varies with latitude and climate on a global scale. We test a
-suite of hypotheses, including many previously posed (Table 1), with
-more data and standardisation for factors that have not always been
-controlled for (\emph{e.g.}, stand age, flux components measured) to
-gain a stronger understanding of climatic effects on FACF. We show that,
-across all nine FACF variables analyzed, C cycling decreases linearly
-with latitude (Figs. 2-3)--a finding that confirms multiple previous
-studies (\textbf{REFS}) but contradicts the idea that productivity of
-temperate forests rivals that of tropical forests (\textbf{REFS}). The
-FACF variables increase in proportion to one another, with no
-differences in allocation detectable at this global scale and with
-component fluxes summing appropriately to larger fluxes (Fig. 3),
-indicating no major, systematic omissions or overestimations of flux
-components. However, we did detect a tendency for greater variability
-among subsidiary C fluxes (e.g., \emph{\(ANPP_woody\)},
-\emph{\(BNPP_{fine root}\)}; Fig. 2; \textbf{some SI table?};
-\textbf{discussed below??}). Climate explains a significant proportion
-of variation in all C fluxes analyzed (albiet less for subsidiary
-fluxes), with temperature variables being the best predictors of FACF on
-this global scale (Figs. 4-5). While other climate variables are
-significant predictors of FACF, none of them improve on the explanatory
-power of temperature-related variables in general or MAT specifically
-(Fig. 5). Water availability is an important factor in explaining FACF
-on a global scale: we find a positive influence of precipitation at low
-MAP, with saturation at higher levels of MAP (Fig. 5b) and a significant
+suite of hypotheses, including many previously posed (Table 1),
+standardising for factors that have not always been controlled for
+(\emph{e.g.}, stand age, flux components measured) to gain a stronger
+understanding of climatic effects on FACF.
+
+bbl: Most of these two paragraphs (following) just restate the results.
+Honestly I would remove them enitrely, but at the very least, condense
+into a single paragraph.
+
+We show that, across all nine FACF variables analyzed, C cycling
+decreases linearly with latitude (\emph{H1.1}; Figs. 2-3)--a finding
+that confirms multiple previous studies (\textbf{REFS}) but contradicts
+the idea that productivity of temperate forests rivals that of tropical
+forests (\emph{H1.1.alt}; \textbf{REFS}). The FACF variables increase in
+proportion to one another (\emph{H1.2}), with no differences in
+allocation detectable at this global scale and with component fluxes
+summing appropriately to larger fluxes (Fig. 3), indicating no major,
+systematic omissions or overestimations of flux components. However, we
+did detect a tendency for greater variability among subsidiary C fluxes
+(e.g., \emph{\(ANPP_{woody}\)}, \emph{\(BNPP_{fine.root}\)}; Fig. 2;
+\textbf{some SI table?}).
+
+\textbf{Climate explains a significant proportion of variation in all C
+fluxes analyzed} (albiet less for subsidiary fluxes), with temperature
+variables the best predictors of FACF at this global scale (Figs. 4-5).
+While other climate variables are significant predictors of FACF
+(\emph{H3}), none of them improve on the explanatory power of
+temperature-related variables in general or MAT specifically (Fig. 5).
+Water availability is an important factor in explaining FACF on a global
+scale: we find a positive influence of precipitation at low MAP, with
+saturation at higher levels of MAP (Fig. 5b) and a significant
 interaction between MAT and MAP for seven of the nine variables analyzed
 (Fig. 4). Temperature seasonality and growing season length are closely
-correlated with MAT and are strong predictors of FACF (Fig. 5e-f),
-though growing season length doesn't improve upon MAT as a predictor.
-Within the growing season, the influence of climate on C cycling is
-smaller but still significant for a number of carbon fluxes
-(\textbf{some SI table?}). These findings clarify the big picture of how
-FACF varies with latitude and climate on a global scale.
+correlated with MAT and are strong predictors of FACF (\emph{H4}; Fig.
+5e-f), though growing season length doesn't improve upon MAT as a
+predictor. Within the growing season, the influence of climate on C
+cycling is smaller but still significant for a number of carbon fluxes
+(\emph{H5}; \textbf{some SI table?}). These findings clarify the big
+picture of how FACF varies with latitude and climate on a global scale.
 
 Past studies have differed in their conclusions regarding the
-relationship between productivity and latitude or its correlates
-(\textbf{REFS})--quite possibly because of lack of standardization with
-respect to stand age and disturbance history. Our findings indicate
-that, among mature, undisturbed stands, FACF is unambiguously highest in
-the tropical regions, and the relationship is approximately linear
-(Figs. 2-3). This contrasts with suggestions that productivity of
-temperate forests is similar to that of tropical forests
-(\textbf{REFS}). Such observations may be attributable to the fact that
-temperate forests are predominantly secondary and of younger age than
-tropical forests (\textbf{REF}), so analyses comparing across
-latitudinal gradients without strict control of stand age tend to be
-comparing younger temperate stands with older tropical (and often
-boreal) forests (\href{http://pure.iiasa.ac.at/id/eprint/15361/}{Poulter
-et al. 2018-DOI:10.1594/PANGAEA.889943}). Because C allocation varies
-with stand age (\textbf{REFS (KAT/NOBBY)}), age differences may
-introduce systematic biases into analyses of FACF across latitude or
-global climatic gradients. For example, woody productivity tends to be
-higher in rapidly aggrading secondary stands than in old-growth forests,
-where proportionally more C is allocated to respiration (\textbf{REFS})
-{[}*purpose for respiration/ other compenents (\textbf{REFS????-
-NOBBY?}){]}. By constraining our analysis to stands
-\$\textbackslash{}geq\$100 years old with no record of major recent
-disturbance, we have clarified the shape of the relationship between
-FACF and latitude.
-
-{[}\textbf{NOBBY, please pay special attention to this paragraph}{]}. We
-show that FACF are broadly consistent in their responses to climate
+relationship between productivity and latitude or its correlates (Table
+1, \emph{H1}; \textbf{REFS})--quite possibly because of lack of
+standardization with respect to stand age and disturbance history. Our
+findings indicate that, among mature, undisturbed stands, FACF is
+unambiguously highest in the tropical regions, and the relationship is
+approximately linear (Figs. 2-3). This contrasts with suggestions that
+productivity of temperate forests is similar to that of tropical forests
+(\textbf{REFS}). Temperate forests tend to be younger than tropical
+forests (\textbf{REF}), so analyses comparing across latitudinal
+gradients without controlling for stand age risk confounding age with
+biome effects (\href{http://pure.iiasa.ac.at/id/eprint/15361/}{Poulter
+et al. 2018-DOI:10.1594/PANGAEA.889943}; Potapov et al. 2008. Mapping
+the World's Intact Forest Landscapes by Remote Sensing. Ecology and
+Society 13 (2), 51. ). In addition, because C allocation varies with
+stand age \citep{de_lucia_forest_2007} (\textbf{See Nobby's comment in
+manuscript-draft\_NK.pdf}), age differences may introduce systematic
+biases into analyses of FACF across latitude or global climatic
+gradients. For example, woody productivity tends to be higher in rapidly
+aggrading secondary stands than in old-growth forests, where
+proportionally more C is allocated to respiration (** Nobby AFM paper:
+Understanding the controls over forest carbon use efficiency on small
+spatial scales: Effects of forest disturbance and tree diversity**)
+{[}*purpose for respiration/ other compenents (\textbf{See Nobby's
+comment in manuscript-draft\_NK.pdf}){]}.
+
+We show that FACF are broadly consistent in their responses to climate
 drivers on the global scale (with the exception of some differences in
 MAT-MAP interactions; Fig. 4), with no major trends in C allocation
-among the variable pairs tested (Fig. 2; \textbf{Some SI table}).
-Although variation in allocation has been observed along gradients of
-elevation \citep{moser_elevation_2011} and water availability
-\citep{newman_above-_2006}--along with non-climatic axes of stand age
-\citep{litton_carbon_2007}, nutrient availability
+among the variable pairs tested (Table 1, \emph{H1}; Fig. 2;
+\textbf{Some SI table}). Although variation in allocation has been
+observed along gradients of elevation \citep{moser_elevation_2011} and
+water availability \citep{newman_above-_2006}--along with non-climatic
+axes of stand age \citep{litton_carbon_2007}, nutrient availability
 \citep{litton_carbon_2007, gill_belowground_2016}, and forest structure
@@
 currently insufficent to rule out the possibility that such trends
-exist. Of particular interest and significance are the relationships
-amongst \(GPP\), net primary productivity (\(NPP\) and its components,
-particularly \(ANPP_{woody-stem}\)), and respiration (\(R_{auto}\) and
-components). There have been suggestions that tropical forests tend to
-have low carbon use efficiency (\(CUE\)=
+exist.
+
+bbl: remove this sentence? Of particular interest and significance are
+the relationships amongst \(GPP\), net primary productivity (\(NPP\) and
+its components, particularly \(ANPP_{woody-stem}\)), and respiration
+(\(R_{auto}\) and components). There have been suggestions that tropical
+forests tend to have low carbon use efficiency (\(CUE\)=
 \(NPP\)/\(GPP\)=(\(GPP\)-\(R_{auto}\)/\(GPP\)), which are based on
@@
 \(CUE\) is known to decline with forest age
-(\citep{de_lucia_forest_2007};\textbf{REFS(NOBBY)}) but appears to be
-roughly independent of \(GPP\) \citep{litton_carbon_2007}. Among our
-sites with relevant data, there is indication in our data that CUE or
-\(ANPP_{woody-stem}\)/\(GPP\) increase with latitude (\textbf{some SI
-table}). Careful methodological standardization across a consistent set
-of mature forest sites will be necessary to determine whether any
-climate-driven gradients in allocation exist at the global scale.
+(\citep{de_lucia_forest_2007};
+\href{https://academic.oup.com/treephys/article-abstract/39/8/1473/5423351?redirectedFrom=fulltext}{Collalti
+\& Prentice 2019}) but appears to be roughly independent of \(GPP\)
+\citep{litton_carbon_2007}. Among our sites with relevant data, there is
+indication that CUE or \(ANPP_{woody-stem}\)/\(GPP\) increase with
+latitude (\textbf{some SI table}). Additional measurements with careful
+methodological standardization across a consistent set of mature forest
+sites will be necessary to determine whether any climate-driven
+gradients in allocation exist at the global scale.
 
@@
 variation in the major fluxes (\(GPP\), \(NPP\), \(R_{auto}\) - latitude
-\$R\^{}2 \ge\$48\%) than in subsidiary fluxes (latitude
-\(R^2\)\textless{}27\% for \(BNPP_{fine.root}\),
+\(R^2\ge\) 48\%) than in subsidiary fluxes (latitude \(R^2\)
+\textless{}27\% for \(BNPP_{fine.root}\),
 \(R_{auto-root}\),\(ANPP_{woody-stem}\)) (Fig. 2; \textbf{some SI
-table?}). There are two non-exclusive potential explanations for this.
+table?}). There are two, non-exclusive, potential explanations for this.
 First, it may be that methodological variation is larger relative to
@@
 differences such as biomass allometries applied and minimum plant size
-sampled. However, methodological variation and uncertainty affect all of
-fluxes considered here--not necessarily any less than the
-aforementioned, and some of the larger fluxes that vary more strongly
-with respect to climate (\(ANPP\), \(NPP\)) are estimated by summing
-uncertain component fluxes. Second, differences among variables in the
-proportion of variation explained by climate may be attributable to more
-dicrect climatic control over \(GPP\) than subsidiary fluxes. That is,
-subsidiary fluxes may be shaped by climate both indirectly through its
-influence on \(GPP\) and potentially respiration and directly through
-any climatic influence on C allocation, as well as by many other local-
-and regional-scale factors (\textbf{REFS}).
+sampled (bbl: cite e.g.~Huntzinger?). However, methodological variation
+and uncertainty affect all of fluxes considered here--not necessarily
+any less than the aforementioned, and some of the larger fluxes that
+vary more strongly with respect to climate (\(ANPP\), \(NPP\)) are
+estimated by summing uncertain component fluxes. Second, differences
+among variables in the proportion of variation explained by climate may
+be attributable to more dicrect climatic control over \(GPP\) than
+subsidiary fluxes. That is, subsidiary fluxes may be shaped by climate
+both indirectly through its influence on \(GPP\) and respiration and
+directly through any climatic influence on C allocation, as well as by
+many other local- and regional-scale factors (\textbf{REFS}).
 
@@
 temperature-related climate variables, and secondarily by moisture
-availability (Figs. 4-5). Because many climate variables co-vary across
-the latitudinal gradient, because climatic drivers affect forest carbon
-flux on much shorter time scales than can be captured by annual climate
-summary variables, and because both climatic conditions and C flux vary
-intra- and inter-annually around the long-term means, it is not
-appropriate to attempt to identify any one mean annual climate variable
-as a mechanistic driver of FACF. However, it remains informative to
-consider these relationships. We find that temperature-related climate
-variables (\(MAT\), temperature seasonality, \ldots{}\textbf{LIST})
-explain the highest proportion of variability in productivity, and among
-these, \(MAT\) is generally the best predictor--perhaps because
-site-specific MAT is recorded for the majority of sites in ForC, whereas
-other variables are extracted from global gridded data products (Table
-S1\textbf{?}). The effects of temperature are modified by moisture
-availability, with reduced FACF under hot and dry conditions
-(\emph{i.e.}, high PET, high deficit; Fig. 5c-d) and sometimes under
-very high precipitaiton (Figs. 4, 5b). Negative effects of very high
-precipitation on FACF have been previously observed (\textbf{REFS}) and
-are attributable to nutrient and light limitations (\textbf{REFS}) Thus,
-although temperature and water availability jointly and interactively
-drive global-scale patterns of FACF.
-
-FACF is reduced by climatic seasonality (Fig. 5e), and productivity is
-shut down during cold- or dry- dormant seasons. To account for this, a
-number of analyses seeking to characterize global-scale effects of
-climate on productivity have examined the relationship of FACF per month
-of the growing season with growing season climatic conditions
-(\textbf{REFS}). We found that the sort of simple metric needed to
-define growing season at a global scale was uncertain for temperature
-and problematic for moisture (\textbf{WORK ON THIS}). A
-temperature-defined growing season length had stong positive correlation
-with FACF (Fig. 5f), but explained less variation than \(MAT\). Dividing
-FACFs by growing season length to yield FACF per growing season month
-removed the majority of climate-related variation, supporting the idea
-that the latitudinal gradient in FACF is attributable more to shorter
-growing seasons at high latitudes than to inherently lower rates of
-photosyntheiss or respiration by high-latitude forests (\emph{{[}
-Enquist GCB review{]}}). However, there remained a number of significant
-correlations with growing season climatic conditions, suggesting that
-climatic conditions remain influential within the growing season. We
-conclude that while correcting for growing season length takes analyses
-a step closer to mechanistic linkage of instantaneous C flux rates to
-environmental conditions, it remains very crude relative to the the
-timescales on which climate affects plant metabolism and does not
-advance statistical predictive power. Rather, mechanistic accounting for
-climatic effects on global productivity patterns requires models
-representing physiologically meaningful timescales (\emph{e.g.},
-\textbf{refs}).
-
-Our analysis clarifies the big picture of how FACF varies with latitude
-and climate on a global scale, with some important implications for how
-forest carbon cycling relates to climate and, by extension, how it is
-likely to respond to climatic warming. Contrary to some previous
-suggestions, we find no support for non-linear trends in mature forest C
-cycling with respect to latitude or \(MAT\), and no distinct trends in C
-allocation across the global scale (Fig. 3). The implication is that
-under warmer conditions with similar moisture availability--and within
-the temperature range to which forest communities are adapated and
+availability (Table 1, \emph{H2-H3}; Figs. 4-5). Because many climate
+variables co-vary across the latitudinal gradient, because climatic
+drivers affect forest carbon flux on much shorter time scales than can
+be captured by annual climate summary variables, and because both
+climatic conditions and C flux vary intra- and inter-annually around the
+long-term means, it is not appropriate to attempt to identify any one
+mean annual climate variable as a mechanistic driver of FACF. However,
+it remains informative to consider these relationships. We find that
+temperature-related climate variables (\(MAT\), temperature seasonality,
+\ldots{}\textbf{LIST}) explain the highest proportion of variability in
+productivity, and among these, \(MAT\) is generally the best
+predictor--perhaps because site-specific MAT is recorded for the
+majority of sites in ForC, whereas other variables are extracted from
+global gridded data products (Table S1\textbf{?}). The effects of
+temperature are modified by moisture availability, with reduced FACF
+under hot and dry conditions (\emph{i.e.}, high PET, high deficit; Fig.
+5c-d) and sometimes under very high precipitaiton (Figs. 4, 5b).
+Negative effects of very high precipitation on FACF have been previously
+observed (\textbf{REFS}) and are attributable to nutrient and light
+limitations (\textbf{REFS}) Thus, although temperature and water
+availability jointly and interactively drive global-scale patterns of
+FACF.
+
+bbl: this seems really interesting and novel, consider expanding the
+following paragraph
+
+FACF is reduced by temperature seasonality (Table 1, \emph{H4}; Fig.
+5e), and productivity is shut down during cold- or dry- dormant seasons.
+To account for this, a number of analyses seeking to characterize
+global-scale effects of climate on productivity have examined the
+relationship of FACF per month of the growing season with growing season
+climatic conditions (Table 1, \emph{H5}; \textbf{REFS}). We found that
+the sort of simple metric needed to define growing season at a global
+scale was uncertain for temperature and problematic for moisture
+(\textbf{WORK ON THIS}). A temperature-defined growing season length had
+stong positive correlation with FACF (Fig. 5f), but explained less
+variation than \(MAT\). Dividing FACFs by growing season length to yield
+FACF per growing season month removed the majority of climate-related
+variation, supporting the idea that the latitudinal gradient in FACF is
+attributable more to shorter growing seasons at high latitudes than to
+inherently lower rates of photosyntheiss or respiration by high-latitude
+forests (\emph{{[} Enquist et al. 2007 GCB- but check{]}}). However,
+there remained a number of significant correlations with growing season
+climatic conditions, suggesting that climatic conditions remain
+influential within the growing season. We conclude that while correcting
+for growing season length takes analyses a step closer to mechanistic
+linkage of instantaneous C flux rates to environmental conditions, it
+remains very crude relative to the the timescales on which climate
+affects plant metabolism and does not advance statistical predictive
+power. Rather, mechanistic accounting for climatic effects on global
+productivity patterns requires models representing physiologically
+meaningful timescales (\emph{e.g.}, \textbf{refs}).
+
+Our analysis clarifies how FACF varies (bbl: is FACF singular or plural?
+Be consistent) with latitude and climate on a global scale, with some
+important implications for how forest carbon cycling relates to climate
+and, by extension, how it is likely to respond to climatic warming. We
+find no support for non-linear trends in mature forest C cycling with
+respect to latitude or \(MAT\), and no distinct trends in C allocation
+across the global scale (Fig. 3). The implication is that under warmer
+conditions with similar moisture availability--and within the
+temperature range to which forest communities are adapated and
 acclimatized--higher temperatures result in a generalized acceleration
@@
 conditions than represented across geographic gradients in climate, but
-our analysis clarifies the big picture as to how carbon cycles through
-forest ecosystems that have developed under relatively stable climatic
-conditions. As we enter a period of accelerating climatic change,
+our analysis clarifies how carbon cycles through contemporary forest
+ecosystems. As we enter a period of accelerating climatic change,
 understanding of the fundamental climatic controls on FACF sets a
@@
 
-Scholarly Studies ForestGEO Compilatin of the ForC database was
+Scholarly Studies ForestGEO Compilation of the ForC database was
 originally funded by DOE

### 2019-11-14 manuscript/manuscript draft.Rmd
knit manuscript
(after fixing special character problem)
added Helene's review files

@@
 title: "Global patterns of forest autotrophic carbon fluxes"
-author: "Rebecca Banbury Morgan, Valentine Herrmann, Norbert Kunert, Ben Bond-Lamberty, Helene Muller-Landau, Kristina J. Anderson-Teixeira"
+author: "Rebecca Banbury Morgan, Valentine Herrmann, Norbert Kunert, Ben Bond-Lamberty, Helene C. Muller-Landau, Kristina J. Anderson-Teixeira"
 output:
@@
 Forest annual carbon cost: a global‐scale analysis of autotrophic respiration.
-Ecology 91(3), 652-661Ƴ* **).
+Ecology 91(3), 652-661* **).
 Forest productivity sets the energy ultimately available to heterotrophic organisms (including microbes), in turn influencing their abundance (**REFS**) and possibly diversity ([Waide et al.

### 2019-11-14 manuscript/manuscript draft.Rmd
started integrating some HCM comments
started deleting comments that I dealt with from the pdf she edited, the switched to yellow highlights.

@@
 Here, we drew upon # records from the Global Forest Carbon Database (ForC), representing all major forest types and the nine most significant forest autotrophic carbon flux (FACF) variables, to comprehensively explore how C cycling in mature, undisturbed forests varies with latitude and climate on a global scale.
-We show that, across all FACF variables analyzed, C cycling decreases linearly with latitude --a finding that confirms multiple previous studies but contradicts the idea that productivity of temperate forests rivals that of tropical forests.
+We show that, across all FACF variables analyzed, C cycling decreases linearly with latitude --a finding that confirms multiple previous studies and contradicts the idea that productivity of temperate forests rivals that of tropical forests.
 The FACF variables increase in proportion to one another, with no differences in allocation detectable at this global scale, but differed in that latitude explained a lower proportion of variation among subsidiary fluxes (in particular, woody aboveground productivity and belowground net primary productivity, $BNPP$).
 Climate explained a significant proportion (#-#%) of variation in all C fluxes analyzed (less for subsidiary fluxes), with temperature variables in general and mean annual temperature ($MAT$) in particular being the best predictors of FACF on this global scale.
-While other climate variables (*e.g.*, displayed significant correlation with FACF, none of them improved significantly over the explanatory power $MAT$.
+While other climate variables (*e.g.*, **XX**) displayed significant correlation with FACF, none of them improved significantly over the explanatory power $MAT$.
 The effects of temperature were modified by moisture availability, with reduced FACF under hot and dry conditions and sometimes under very high precipitaiton (especially for $BNPP$).
@@
 
-**Forests play a critical role in regulating atmospheric CO_2 and climate** (**Bonan**).
-The total amount of CO_2 cycling through Earth's forests each year (*i.e.*, total gross primary production, $GPP$) is more than five times anthropogenic fossil fuel emissions, and net sequestration (~2.4 GtCyr^-1^) offsetting roughly 30% of these emissions [@pan_large_2011].
-As atmospheric carbon dioxide levels continue to rise, with consequences for global climate, there is increasing recognition that proper protection and management of forest resources will have an important role to play in mitigating climate change (** REFS- e.g., IPCC 1.5C and land reports**).
+**Forests play a critical role in regulating atmospheric CO~2~ and climate** (**Bonan**).
+The total amount of CO~2~ cycling through Earth's forests each year (*i.e.*, total gross primary production, $GPP$) is more than five times anthropogenic fossil fuel emissions, and net sequestration (~2.4 GtCyr^-1^) offsets roughly 30% of these emissions [@pan_large_2011].
+As atmospheric carbon dioxide levels continue to rise, with consequences for global climate, there is increasing recognition that  protection and management of forest resources will play an important role in mitigating climate change (** REFS- e.g., IPCC 1.5C and land reports**).
 Carbon fixation, allocation, and metabolism by trees set the basis for energy and material flows in forest ecosystems and define their interactions with Earth's changing climate.

### 2019-11-14 manuscript/manuscript draft.Rmd
worked on Helene's comments

@@
 Associated data, such as stand age, measurement methodologies, and disturbance history, are also included.
-The database was significantly expanded since the publication of [@anderson-teixeira_forc:_2018] through integration with the Global Database of Soil Respiration Database [@bond-lamberty_global_2010].
+The database was significantly expanded since the publication of [@anderson-teixeira_forc:_2018] through integration with the Global Soil Respiration Database [@bond-lamberty_global_2010].
 Additional targeted literature searches were conducted to identify any further available data on primary productivity, with particular focus on mature forests in temperate and boreal regions.
@@
 | $ANPP_{woody-stem}$ | Net primary production of woody stems | Woody stems | Direct measurement of stem growth increment | 264 | 96 |
-| $BNPP_{root}$  | Belowground net primary production | Coarse and fine roots | Direct measurement of one or more of: fine root turnover, soil cores, root ingrowth cores, minirhizotrons; indirect estimates of coarse roots using allometries based on aboveground stem increment measures | 101 | 48 |
-| $BNPP_{fine.root}$ | Net primary production of fine roots | Fine roots | Direct measurement of one or more of: minirhizotrons, fine root turnover, soil cores, root ingrowth cores | 88 | 41 |
+| $BNPP_{root}$  | Belowground net primary production | Coarse and fine roots |  indirect estimates of coarse roots using allometries based on aboveground stem increment measures ; see below for fine roots | 101 | 48 |
+| $BNPP_{fine.root}$ | Net primary production of fine roots | Fine roots | Measurements combined one or more of the following: minirhizotrons, fine root turnover, soil cores, root ingrowth cores | 88 | 41 |
 | $R_{auto}$ | Annual autotrophic respiration, including above- and belowground components | Foliage, stem, and root | Chamber measurements of component gas exchange | 22 | 13 |

### 2019-11-14 manuscript/manuscript draft.Rmd
minor change to table2

@@
 | $ANPP_{woody-stem}$ | Net primary production of woody stems | Woody stems | Direct measurement of stem growth increment | 264 | 96 |
-| $BNPP_{root}$  | Belowground net primary production | Coarse and fine roots |  indirect estimates of coarse roots using allometries based on aboveground stem increment measures ; see below for fine roots | 101 | 48 |
+| $BNPP_{root}$  | Belowground net primary production | Coarse and fine roots |  Coarse roots estimated indirectly using allometries based on aboveground stem increment measures ; fine roots as below | 101 | 48 |
 | $BNPP_{fine.root}$ | Net primary production of fine roots | Fine roots | Measurements combined one or more of the following: minirhizotrons, fine root turnover, soil cores, root ingrowth cores | 88 | 41 |

### 2020-01-17 manuscript/manuscript draft.Rmd
cropped Fig. 1 map
cropped manually.

@@
 
-```{r echo=FALSE, out.width='100%', fig.cap = "Map showing all data used in the analysis, coded by variable", fig.pos='H'}
-  knitr::include_graphics("distribution_all_variables.png")
+```{r echo=FALSE, out.width='100%', fig.cap = "Map showing all data used in the analysis, coded by variable. Variables are plotted individually in Fig. S1. ", fig.pos='H'}
+  knitr::include_graphics("distribution_all_variables_cropped.png")
 ```

### 2020-01-17 scripts/final_model_no_weighting.R
Update final_model_no_weighting.R

@@ -991,7 +991,7 @@ for (age in ages){
       
       axis(1 ,labels = ifelse(pannel.nb %in% c(3,4), TRUE, FALSE))
       axis(2 ,labels = ifelse(pannel.nb %in% c(1,3), TRUE, FALSE))
-      
+        
       
       # equation <-  paste(response.v, "=", r[1], "+", fixed.v,  "x", r[2])
       

### 2020-01-17 manuscript/manuscript draft.Rmd

@@
 
-```{r echo=FALSE, out.width='100%', fig.cap = "Graphs to show primary productivity $(MgC~ha^-1~yr^-1)$ regressed against latitude.
-Lines of best fit are plotted according to the best model selected during analysis.
-All regressions are significant $(p<0.05)$.", fig.pos='H'}
-  knitr::include_graphics("effect_of_lat_transparent.png")
-```
-
 Latitude was a strong predictor for many of the carbon fluxes, explaining 64% of variation in GPP (n = 254, p<0.0001), 50% in NPP (n = 114, p<0.0001) and 45% in ANPP (n = 259, p<0.0001).
@@
 *Relationships and differences among fluxes.* In general, smaller component fluxes summed approximately to larger fluxes across the latitudinal gradient (fig.
-3).
+2).
 That is, modelled estimates of GPP, generated from the sum of NPP and R auto; NPP, generated from the sum of ANPP and BNPP~root~; and ANPP, generated from the sum of ANPP~foliage~ and ANPP~woody~ ~stem~, fell completely within the confidence intervals of the regressions of field estimates of GPP, NPP and ANPP respectively.
@@
 MAT was a significant (p<0.05) and strong predictor of productivity for all carbon fluxes tested, with all fluxes showing a linear increase with temperature (fig.
-5).
+4).
 We found no support for a saturation point of productivity with temperature.
@@
 Productivity increased with precipitation, up until a saturation point at between 3000 and 4000mm annual precipitation, above which productivity started to decrease (fig.
-5).
+4).
 The notable exception to this was GPP: the model indicated that GPP continued to increase with precipitation up to measures of at least 5000mm annually (p<0.0001, $R^2$ = 0.33.
@@
 
-There was a significant interactive effect between MAT and MAP for GPP, BNPP~root~, BNPP~fine~ ~root~, ANPP, ANPP~woody~ ~stem~, and R~root~ (fig.4).
+There was a significant interactive effect between MAT and MAP for GPP, BNPP~root~, BNPP~fine~ ~root~, ANPP, ANPP~woody~ ~stem~, and R~root~ (fig.3).
 There was a significant additive effect for R~auto~.
@@
 ANPP~foliage~, BNPP~fine~ ~root~ and R~root~ increased linearly with PET; however, all other fluxes showed a polynomial relationship with PET (fig.
-5).
+4).
 We found strong evidence for a saturation point or peak with PET: productivity tended to increase at values below 1000mm, before saturating between 1200 and 1700mm.
@@
 BNPP~fine~ ~root~ showed a linear relationship with vapour pressure deficit ($R^2$ = 0.07, p<0.05), but all other fluxes showed a polynomial relationship (fig.
-5).
+4).
 Productivity initially increased with vapour pressure deficit, before saturating at around 0.8 kPa.
@@
 GPP, NPP, ANPP, and R~root~ exhibited a polynomial relationship with seasonality (fig.
-5).
+4).
 ANPP~foliage~, ANPP~woody~ ~stem~ and R~auto~ decreased linearly with temperature seasonality.
@@
 We found a significant relationship between length of growing season and productivity, with all fluxes showing a linear increase in productivity with length of growing season (fig.
-5).
+4).
 Length of growing season was a strong predictor of productivity, explaining 51% of variation in GPP, 39% of variation in NPP, and 34% of variation in ANPP, but it was not identified as a stronger predictor than MAT for any of the fluxes analysed.
@@
 
-We show that, across all nine FACF variables analyzed, C cycling decreases linearly with latitude (*H1.1*; Figs.
-2-3)--a finding that confirms multiple previous studies (**REFS**) but contradicts the idea that productivity of temperate forests rivals that of tropical forests (*H1.1.alt*; **REFS**).
+We show that, across all nine FACF variables analyzed, C cycling decreases linearly with latitude (*H1.1*; Fig.
+2)--a finding that confirms multiple previous studies (**REFS**) but contradicts the idea that productivity of temperate forests rivals that of tropical forests (*H1.1.alt*; **REFS**).
 The FACF variables increase in proportion to one another (*H1.2*), with no differences in allocation detectable at this global scale and with component fluxes summing appropriately to larger fluxes (Fig.
-3), indicating no major, systematic omissions or overestimations of flux components.
+2), indicating no major, systematic omissions or overestimations of flux components.
 However, we did detect a tendency for greater variability among subsidiary C fluxes (e.g., *$ANPP_{woody}$*, *$BNPP_{fine.root}$*; Fig.
@@
 **Climate explains a significant proportion of variation in all C fluxes analyzed** (albiet less for subsidiary fluxes), with temperature variables the best predictors of FACF at this global scale (Figs.
-4-5).
+3-4).
 While other climate variables are significant predictors of FACF (*H3*), none of them improve on the explanatory power of temperature-related variables in general or MAT specifically (Fig.
-5).
+4).
 Water availability is an important factor in explaining FACF on a global scale: we find a positive influence of precipitation at low MAP, with saturation at higher levels of MAP (Fig.
-5b) and a significant interaction between MAT and MAP for seven of the nine variables analyzed (Fig.
-4).
+4b) and a significant interaction between MAT and MAP for seven of the nine variables analyzed (Fig.
+3).
 Temperature seasonality and growing season length are closely correlated with MAT and are strong predictors of FACF (*H4*; Fig.
-5e-f), though growing season length doesn't improve upon MAT as a predictor.
+4e-f), though growing season length doesn't improve upon MAT as a predictor.
 Within the growing season, the influence of climate on C cycling is smaller but still significant for a number of carbon fluxes (*H5*; **some SI table?**).
@@
 Past studies have differed in their conclusions regarding the relationship between productivity and latitude or its correlates (Table 1, *H1*; **REFS**)--quite possibly because of lack of standardization with respect to stand age and disturbance history.
-Our findings indicate that, among mature, undisturbed stands, FACF is unambiguously highest in the tropical regions, and the relationship is approximately linear (Figs.
-2-3).
+Our findings indicate that, among mature, undisturbed stands, FACF is unambiguously highest in the tropical regions, and the relationship is approximately linear (Fig.
+2).
 This contrasts with suggestions that productivity of temperate forests is similar to that of tropical forests (**REFS**).
@@
 We show that FACF are broadly consistent in their responses to climate drivers on the global scale (with the exception of some differences in MAT-MAP interactions; Fig.
-4), with no major trends in C allocation among the variable pairs tested (Table 1, *H1*; Fig.
+3), with no major trends in C allocation among the variable pairs tested (Table 1, *H1*; Fig.
 2; **Some SI table**).
@@
 
-The latitudinal gradient in FACF (Figs.
-2-3) is driven primarily by temperature-related climate variables, and secondarily by moisture availability (Table 1, *H2-H3*; Figs.
-4-5).
+The latitudinal gradient in FACF (Fig.
+2) is driven primarily by temperature-related climate variables, and secondarily by moisture availability (Table 1, *H2-H3*; Figs.
+3-4).
 Because many climate variables co-vary across the latitudinal gradient, because climatic drivers affect forest carbon flux on much shorter time scales than can be captured by annual climate summary variables, and because both climatic conditions and C flux vary intra- and inter-annually around the long-term means, it is not appropriate to attempt to identify any one mean annual climate variable as a mechanistic driver of FACF.
@@
 The effects of temperature are modified by moisture availability, with reduced FACF under hot and dry conditions (*i.e.*, high PET, high deficit; Fig.
-5c-d) and sometimes under very high precipitaiton (Figs.
-4, 5b).
+4c-d) and sometimes under very high precipitaiton (Figs.
+3, 4b).
 Negative effects of very high precipitation on FACF have been previously observed (**REFS**) and are attributable to nutrient and light limitations (**REFS**) Thus, although temperature and water availability jointly and interactively drive global-scale patterns of FACF.
@@
 FACF is reduced by temperature seasonality (Table 1, *H4*; Fig.
-5e), and productivity is shut down during cold- or dry- dormant seasons.
+4e), and productivity is shut down during cold- or dry- dormant seasons.
 To account for this, a number of analyses seeking to characterize global-scale effects of climate on productivity have examined the relationship of FACF per month of the growing season with growing season climatic conditions (Table 1, *H5*; **REFS**).
@@
 A temperature-defined growing season length had stong positive correlation with FACF (Fig.
-5f), but explained less variation than $MAT$.
+4f), but explained less variation than $MAT$.
 Dividing FACFs by growing season length to yield FACF per growing season month removed the majority of climate-related variation, supporting the idea that the latitudinal gradient in FACF is attributable more to shorter growing seasons at high latitudes than to inherently lower rates of photosyntheiss or respiration by high-latitude forests (*[ Enquist et al.
@@
 We find no support for non-linear trends in mature forest C cycling with respect to latitude or $MAT$, and no distinct trends in C allocation across the global scale (Fig.
-3).
+2).
 The implication is that under warmer conditions with similar moisture availability--and within the temperature range to which forest communities are adapated and acclimatized--higher temperatures result in a generalized acceleration of FACF, with no major shifts in C allocation among subsidiary fluxes.

### 2020-01-17 scripts/extract_climate_data_growing_season.R
Update extract_climate_data_growing_season.R

@@ -684,22 +684,21 @@ all.koeppen <- NULL
 fixed.variables <- c("pet", "pre", "tmp", "solarradiationmonthly")
 
 
+response.variables <- c("GPP", "NPP", "ANPP", "BNPP_root", "ANPP_foliage", "ANPP_woody_stem")
 
-response.variables.groups <- list(c("GPP", "NPP", "BNPP_root", "BNPP_root_fine"),
-                                  c("ANPP", "ANPP_foliage", "ANPP_woody_stem"))
+# response.variables <- "ANPP_foliage"
 
-par(mfcol = c(2,4), mar = c(2,2,2,2), oma = c(5,5,2,0))
+
 
   ### mature forests only ####
-  for (age in ages){
-    
-    if (age %in% "age.greater.than.100") ages.to.keep <- ForC_simplified$stand.age >= 100 & !is.na(ForC_simplified$stand.age)
-    if (age %in% "age.greater.than.200") ages.to.keep <- ForC_simplified$stand.age >= 200 & !is.na(ForC_simplified$stand.age)
-    panel.number <- 1
+
+panel.number <- 1
+png(file = paste0("C:/Users/gyrcbm/Dropbox/Global_Productivity/results/figures/final_figures/supporting_information/gridded_growing_season.png"), width = 2500, height = 3000, units = "px", res = 300)
+
+par(mfcol = c(6,4), mar = c(2.5,2,2,2.5), oma = c(8,8,0,2), xpd = T)
+ages.to.keep <- ForC_simplified$stand.age >= 100 & !is.na(ForC_simplified$stand.age)
+
     for(fixed.v in fixed.variables){
-      
-      # png(file = paste0("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/figures/final_figures/unweighted_model/growing_season/Effect_of_", fixed.v, "_MATURE_only_", age, "_", n, ".png"), width = 2255, height = 2000, units = "px", res = 300)
-      
       
       print(fixed.v)
       
@@ -708,16 +707,7 @@ par(mfcol = c(2,4), mar = c(2,2,2,2), oma = c(5,5,2,0))
       xaxis <- fixed.v.info$xaxis[which(fixed.v.info$fixed.v %in% fixed.v)]
       
       ###subset ForC
-      
-      
-     
-      
-      for(response.variables in response.variables.groups){
-      
-        response.variables.col <- 1:length(response.variables)
-      
-      if(response.variables[1] == "GPP") n <- 1
[…]

### 2020-01-17 manuscript/manuscript draft.Rmd
minor editing (HM comments)

@@
 
-bbl: here or in the discussion, note that this is a fairly stringent test: it's easy for sub-fluxes not to sum up! (Extensive citations from EC literature for example.)
+bbl: here or in the discussion, note that this is a fairly stringent test: it's easy for sub-fluxes not to sum up!
+(Extensive citations from EC literature for example.).
+HML: yes and no.
+confidence intervals are pretty large.
 
-```{r echo=FALSE, out.width='100%', fig.cap = "Graphs of primary productivity $(MgC~ha^-1~yr^-1)$ regressed against latitude.
+```{r echo=FALSE, out.width='100%', fig.cap = "Latitudinal trends in forest autotropic carbon flux.
 Lines of best fit are plotted according to the best model selected during analysis.
 All regressions are significant $(p<0.05)$.
-Plots 1 - 3 show two component fluxes; a larger flux, defined as the combination of the two component fluxes; and a modelled estimate of the sum of the two component fluxes.
-95\\% confidence intervals are plotted for the larger flux.
-Plot 4 shows three belowground fluxes, and a modelled estimate of the total belowground carbon flux", fig.pos='H'}
+Each panel shows major C fluxes together with component fluxes.
+Also plotted are predicted trends in the major C fluxes based on the sum of component fluxes.
+95\\% confidence intervals are plotted for the major flux for comparison with predicted trends.
+In (d),  which shows three belowground fluxes, the major flux, total belowground carbon flux, is one for which we have no data", fig.pos='H'}
 

### 2020-01-17 manuscript/manuscript draft.Rmd
finished main doc figures review
working on Helene's comments

@@
 Climate explained a significant proportion (#-#%) of variation in all C fluxes analyzed (less for subsidiary fluxes), with temperature variables in general and mean annual temperature ($MAT$) in particular being the best predictors of FACF on this global scale.
-While other climate variables (*e.g.*, **XX**) displayed significant correlation with FACF, none of them improved significantly over the explanatory power $MAT$.
+While other climate variables (*e.g.*, **XX**) displayed significant correlation with FACF, none of them had significantly better explanatory power than $MAT$.
 The effects of temperature were modified by moisture availability, with reduced FACF under hot and dry conditions and sometimes under very high precipitaiton (especially for $BNPP$).
@@
 
+(**revisit these first two paragraphs.**)
+
 **Forests play a critical role in regulating atmospheric CO~2~ and climate** (**Bonan**).
-The total amount of CO~2~ cycling through Earth's forests each year (*i.e.*, total gross primary production, $GPP$) is more than five times anthropogenic fossil fuel emissions, and net sequestration (~2.4 GtCyr^-1^) offsets roughly 30% of these emissions [@pan_large_2011].
-As atmospheric carbon dioxide levels continue to rise, with consequences for global climate, there is increasing recognition that  protection and management of forest resources will play an important role in mitigating climate change (** REFS- e.g., IPCC 1.5C and land reports**).
+Approximately **##** Gt of CO~2~ cycle through Earth's forests each year (*i.e.*, total gross primary production, $GPP$; **REF-KAT**), and their net C sequestration (~2.4 GtCyr^-1^) offsets roughly 30% of anthropogenic fossil fuel emissions [@pan_large_2011].
+As atmospheric carbon dioxide levels continue to rise, driving the global climate crisis, there is increasing recognition that  protection and management of forest resources will play an important role in mitigating climate change (** REFS- e.g., IPCC 1.5C and land reports**).
 Carbon fixation, allocation, and metabolism by trees set the basis for energy and material flows in forest ecosystems and define their interactions with Earth's changing climate.
@@
 Forest productivity sets the energy ultimately available to heterotrophic organisms (including microbes), in turn influencing their abundance (**REFS**) and possibly diversity ([Waide et al.
-2009](https://www.annualreviews.org/doi/abs/10.1146/annurev.ecolsys.30.1.257)**REFS**)--but probably not individual metabolic rate (**Anderson & Jetz 2005, but check more recent literature**).
+2009](https://www.annualreviews.org/doi/abs/10.1146/annurev.ecolsys.30.1.257)**REFS**).
 However, our ability to draw general macroscopic conclusions regarding global variation in multiple productivity variables with respect to climate has been limited in that these analyses often mix forests that vary in stand age, disturbance history, and/or management status; do not always sufficiently parse related variables (*e.g.*, combining net primary productivity records with and without belowground components); and typically consider only one or a few variables at a time.
@@
 This is corroborated by analysis of site-level GPP measurements, which appear to reach their highest levels in tropical forests [@luyssaert_co_2007].
-In contrast, there is evidence that the highest values of net primary productivity (NPP) may be found in temperate forests [@luyssaert_co_2007; @huston_global_2009], although other studies find NPP is highest in the tropics, showing a decrease with latitude [@simova_enigma_2017].
+{**condense the previous two sentences into one-- both primarily eddy flux**} In contrast, some studies suggest that the highest values of net primary productivity (NPP) may be found in temperate forests [@luyssaert_co_2007; @huston_global_2009], while others find NPP highest in the tropics and decreasing with latitude [@simova_enigma_2017].
 Other studies have chosen to focus exclusively on above-ground net primary productivity (ANPP), finding evidence of a weak negative relationship between ANPP and latitude [@huston_global_2009; @gillman_latitude_2015].
@@
 **Climate is a significant driver of productivity across broad spatial scales** [@cleveland_relationships_2011].
-The majority of studies have focused on exploring the relationships between productivity and mean annual temperature (MAT) and mean annual precipitation (MAP), as the most commonly reported site-level climate variables.
+The majority of studies have focused on exploring the relationships of productivity to mean annual temperature (MAT) and mean annual precipitation (MAP), as the most commonly reported site-level climate variables.
 These variables have the advantage that they describe broad trends in temperature and water availability, and therefore capture a lot of global-scale variation in climate.
@@
 Taylor et al.
-[-@taylor_temperature_2017] showed that increased MAP had a negative effect at low MAT, but a positive effect at high levels of MAT, and vice versa, suggesting that both temperature and water availability are important in explaining productivity.
+[-@taylor_temperature_2017] showed that increased MAP had a negative effect at low MAT, but a positive effect at high levels of MAT, and *vice versa* {**Helene found this confusing**}, suggesting that both temperature and water availability are important in explaining productivity.
 
@@
 Furthermore, MAT and MAP are very coarse measures of climate, and so fail to capture much variation in climate on an intra-annual scale, including the effects of factors such as growing season length, number of frost-free days, temperature seasonality, and dry season length.
-Some studies have suggested that the apparently strong relationship between MAT and productivity is actually a factor of the correlation between MAT and growing season length [@kerkhoff_plant_2005; @malhi_productivity_2012; @michaletz_convergence_2014; @michaletz_drivers_2018].
+Some studies have suggested that the apparently strong relationship between MAT and productivity is actually a consequence of the correlation between MAT and growing season length [@kerkhoff_plant_2005; @malhi_productivity_2012; @michaletz_convergence_2014; @michaletz_drivers_2018].
 Kerkhoff et al.
 [-@kerkhoff_plant_2005] and Michaletz et al.
-[-@michaletz_convergence_2014] find that, within the growing season, there is no significant relationship between productivity and MAT, indicating that the effect of temperature is due to increased length of growing season, rather than an inherent influence of temperature on productivity.
+[-@michaletz_convergence_2014] find that, within the growing season, there is no significant relationship between productivity and MAT, suggesting that the effect of temperature is due to increased length of growing season, rather than an inherent influence of temperature on productivity.
 
@@
   i) the AIC value of the model was smaller than the AIC value of the baseline model by >2  
-  ii) the r-squared value of the model was larger than the r-squared value of the baseline model by >5
+  ii) the r-squared value of the model was larger than the r-squared value of the baseline model by **>5 (???- not possible)**
 
@@
 
-All major carbon fluxes increased linearly with decreasing latitude (fig. 2).
+All major carbon fluxes decreased linearly with latitude (Fig. 2).
 

### 2020-01-17 manuscript/manuscript draft.Rmd
working on Helene's comments

@@
 **Forests play a critical role in regulating atmospheric CO~2~ and climate** (**Bonan**).
-Approximately **##** Gt of CO~2~ cycle through Earth's forests each year (*i.e.*, total gross primary production, $GPP$; **REF-KAT**), and their net C sequestration (~2.4 GtCyr^-1^) offsets roughly 30% of anthropogenic fossil fuel emissions [@pan_large_2011].
-As atmospheric carbon dioxide levels continue to rise, driving the global climate crisis, there is increasing recognition that  protection and management of forest resources will play an important role in mitigating climate change (** REFS- e.g., IPCC 1.5C and land reports**).
+Approximately **##** Gt of CO~2~ cycle through Earth's forests each year **REF-KAT**), and their net C sequestration (~2.4 GtCyr^-1^) offsets roughly 30% of anthropogenic fossil fuel emissions [@pan_large_2011].
+As atmospheric carbon dioxide levels continue to rise, driving climate change, there is increasing recognition that  protection and management of forest resources will play an important role in mitigating climate change (** REFS- e.g., IPCC 1.5C and land reports**).
 Carbon fixation, allocation, and metabolism by trees set the basis for energy and material flows in forest ecosystems and define their interactions with Earth's changing climate.
@@
 In order to approach these broad and complex issues, we simplify the major gaps in our knowledge to five broad hypotheses and corresponding specific predictions (Table 1).
-First, we ask how forest autotrophic carbon fluxes (FACF) vary with latitude.
+** see [issue #48 ](https://github.com/forc-db/Global_Productivity/issues/48)** First, we ask how forest autotrophic carbon fluxes (FACF) vary with latitude.
 We then test how these fluxes relate to MAT and MAP, and additionally how they respond to other, less well studied, climate variables.
-We finally consider the relationship between FACF and seasonality, considering the role of seasonality in explaining variation in carbon fluxes, and the influence of climate on FACF per month of the growing season.
-We use a comprehensive global database of forest carbon fluxes to explore the above questions for nine carbon fluxes, allowing for an in-depth exploration of the effect of climate on global productivity.
+Finally, we consider the relationship between FACF and seasonality, considering the role of seasonality in explaining variation in carbon fluxes, and the influence of climate on FACF standardized by growing season length.
+We use a comprehensive global database of forest carbon fluxes to address the above questions for nine carbon fluxes, allowing for an in-depth exploration of the effect of climate on global productivity.
 
@@
 | $ANPP_{foliage}$ | Net primary production of foliage | Foliage | Direct measurement of litterfall, correcting for changes in leaf biomass when measured | 98 | 49 |
-| $ANPP_{woody-stem}$ | Net primary production of woody stems | Woody stems | Direct measurement of stem growth increment | 264 | 96 |
+| $ANPP_{woody-stem}$ | Net primary production of woody stems | Woody stems | Direct measurement of stem growth increment (**with recruitment?**) | 264 | 96 |
 | $BNPP_{root}$  | Belowground net primary production | Coarse and fine roots |  Coarse roots estimated indirectly using allometries based on aboveground stem increment measures ; fine roots as below | 101 | 48 |

### 2020-01-17 manuscript/manuscript draft.Rmd
worked on Helene's comments

@@
 There is evidence that stand age influences patterns of primary productivity and carbon allocation in forest ecosystems, and can confound relationships between latitude and primary productivity [@de_lucia_forest_2007; @gillman_latitude_2015].
-To reduce any biasing effects of stand age, stands under 100 years of age were excluded from analysis.
-Sites for which stand age was unknown were excluded from analysis.
+To reduce any biasing effects of stand age, we included only stands of known age $\ge$ 100 years and those described by terms such as "mature", "intact", or "old-growth".
 
@@
 It would be preferable to standardise by minimum area sampled; however x% of plots in the database are 1 ha or under in size; excluding these plots would place significant constraints on sample size.
+(**really, we didn't exclude any studies based on methodology(all studies measured min DBH 10 or less...
+so this paragraph needs revision**)
 
@@
 Estimates of NPP were selected if they included foliage, branch, stem, coarse root, and fine root.
-Measures of NPP which included additional components, including understorey, volatile organic compounds (VOCs), exudates, estimates of NPP lost to herbivory, and the NPP of reproductive structures, were excluded.
+Measures of NPP which included additional components, including understorey, volatile organic compounds (VOCs), exudates, estimates of NPP lost to herbivory (**these are all small, no need to exclude**), and the NPP of reproductive structures (**often in literfall/ mixed in with foliage...
+did we really exclude estimates including litterfall?**), were excluded.
 Estimates of ANPP were selected if they included foliage, stem growth and optionally branch turnover.
@@
 
-Additionally, two climate variables were derived from the above datasets: maximum vapour pressure deficit, defined as the vapour pressure deficit of the month with the largest deficit; and water stress months, defined as the number of months annually where precipitation was lower than potential evapotranspiration.
+Additionally, two climate variables were derived from the above datasets: maximum vapour pressure deficit, defined as the vapour pressure deficit of the month with the largest deficit; and water stress months, defined as the number of months annually where precipitation was lower than potential evapotranspiration (PET; **HML:defined how?**).
 
@@
 The effect of each extracted climate variable on each measure of primary productivity was modelled by specifying the climate variable as a fixed effect.
-For each climate variable, three models were specified: a null model; a model with the climate variable as a linear term; and a model with the climate variable as a polynomial term.
+For each climate variable, three models were specified: a null model; a model with the climate variable as a linear term; and a model with the climate variable as a **XX**-order polynomial term.
 AIC values were calculated for the models and used to select the best model.
@@
 
-Because the magnitude of fluxes varies significantly, in order to facilitate comparisons between regression models for each flux, data for each flux was scaled, to give the data a mean of 0 and standard deviation of 1.
-As each data set was scaled separately, this does not allow for statistical comparisons of slope values, but does assist in visualising the data.
-
 To test for a potential influence of altitude, models were also run with site altitude included as a second fixed effect.
@@
 Including altitude had a very small effect on most models, with the difference in AIC values between models including and excluding altitude often being <2, suggesting the models are very similar in their explanatory power.
-As a result, it was decided to present results only from models do not include altitude as a fixed term.
+As a result, it was decided to present results only from models without altitude as a fixed term.
 
@@
 
-```{r echo=FALSE, out.width='100%', fig.cap = "Plots of primary productivity $(MgC~ha^-1~yr^-1)$ regressed against (a) mean annual temperature; (b) mean annual precipitation; (c) potential evapotranspiration, (d) vapour pressure deficit; (e) temperature seasonality; (f) length of growing season.
-Lines of best fit are plotted according to the best model selected during analysis.
+```{r echo=FALSE, out.width='100%', fig.cap = "Plots of carbon fluxes against (a) mean annual temperature; (b) mean annual precipitation; (c) potential evapotranspiration, (d) vapour pressure deficit; (e) temperature seasonality; (f) length of growing season.
+For visualization purposes, data for each flux was rescaled with a mean of 0 and standard deviation of 1.
+Lines of best fit are plotted according to the best model selected during analysis (**see issue 47**).
 All regressions are significant $(p<0.05)$.", fig.pos='H'}

### 2020-01-19 manuscript/manuscript draft.Rmd
working through Helene's comments
finished through methods

@@
 
-*Model specification.* The effects of climate and latitude on primary productivity were analysed using mixed effects models using the package 'lme4' [@bates_fitting_2015] in R v.3.5.1 [@r_core_team_r:_2018].
+*Model specification.*
+
+{**from Helene: It was frequently unclear to me whether the influences of other climate variables were being considered in multivariate or univariate analyses.
+This needs to be clear.
+I can see reasons to do it either way, but they need to be differently motivated and clearly presented.**}
+
+The effects of climate and latitude on primary productivity were analysed using mixed effects models using the package 'lme4' [@bates_fitting_2015] in R v.3.5.1 [@r_core_team_r:_2018].
 The effect of each extracted climate variable on each measure of primary productivity was modelled by specifying the climate variable as a fixed effect.
@@
 To account for correlations in measurements between tightly clustered sites, a random effect was specified as plot nested within geographic area.
-*Data from the temperate regions was heavily skewed towards studies from the old-growth forests of the Pacific Northwest.*  These forests have very high productivity, and so to ensure that results were not unduly influenced by geographic sampling bias, we tried a version of the model where data were weighted according to forested land area within each Koeppen climate zone.
-Results were similar between the weighted and unweighted model, so, to avoid problems of over-fitting, the weighted model was dropped, and results from this are not presented here.
+*Data from the temperate regions was heavily skewed towards studies from the old-growth forests of the Pacific Northwest.*  These forests have very high FACF, and so to ensure that results were not unduly influenced by geographic sampling bias, we tried a version of the model where data were weighted according to forested land area within each Koeppen climate zone.
+Results were similar between the weighted and unweighted model.
+Because the weighted model may be prone to problems of over-fitting, we present only results of the unweighted model.
+{**Becky, I'd probably just delete this whole paragraph.
+Since we ended up not using the weighted model, it's just extra info for readers to digest.}
 
-Models were run for total annual productivity against annual climate variables, and for monthly growing season productivity, defined as total productivity/length of growing season, against growing season climate variables.
+Models were run for total annual FACF against annual climate variables, and for monthly growing season fluxes, defined as annual FACF/length of growing season, against growing season climate variables.
 For analyses on data within the growing season, only linear models were specified.
 
+{**Note from Helene on the following paragraph: this makes it sound like analyses from H3 were about variation remaining after accounting for temperature, but analyses below don't seem to do that.**}
 To investigate the potential interactive effects of climate variables on carbon fluxes, multivariate models were also specified.

### 2020-01-19 manuscript/manuscript draft.Rmd
working on Helene's comments
finished through results.

@@
 
-Models were run for total annual FACF against annual climate variables, and for monthly growing season fluxes, defined as annual FACF/length of growing season, against growing season climate variables.
+Models were run for total annual FACF against annual climate variables, and for fluxes standardized by growing season length, defined as annual FACF/length of growing season, against growing season climate variables.
 For analyses on data within the growing season, only linear models were specified.
@@
 NPP and ANPP~foliage~ showed no significant interactive or additive effect.
+{**NEED SI TABLE GIVING FITTED COEFFICIENTS & STATS**}
 
 For the variables which showed a significant interactive or additive effect between MAT and MAP, no other climate variable, in combination with MAT, significantly improved on that model.
-For NPP, there was a significant interactive effect between MAT and water stress months, with this model explaining nearly 5% more variation in NPP than MAT alone.
+For NPP, there was a significant interactive effect between MAT and water stress months, with this model explaining nearly 5% more variation in NPP than MAT alone {**Helene: but less than MATxMAP?
+if so, why bother mentioning?**}.
 However, for ANPP~foliage~, no multivariate model improved on the univariate model including only MAT.
 
-```{r echo=FALSE, out.width='100%', fig.cap = "Plots of primary productivity $(MgC~ha^-1~yr^-1)$ regressed against mean annual precipitation.
-Points are grouped into bins of 0 - 1000, 1001 - 2000, 2001 - 3000, and >3000mm mean annual precipitation, and lines of best fit plotted for mean annual precipitation values of 500, 1500, 2500, and 3500mm.
+```{r echo=FALSE, out.width='100%', fig.cap = "Interactive effects of mean annual temperature and mean annual precipitation on FACF.
+For visualization purposes, data points are grouped into bins of 0 - 1000, 1001 - 2000, 2001 - 3000, and >3000mm mean annual precipitation, and lines of best fit models are plotted for mean annual precipitation values of 500, 1500, 2500, and 3500mm.
 All regressions are significant $(p<0.05)$.", fig.pos='H'}
@@
 
-*How does productivity relate to other climate variables?* Our results indicated that productivity was most strongly explained by temperature at the global scale, with temperature-related climate variables coming out as strong predictors of productivity.
-In addition to MAT, temperature seasonality, annual temperature range, and annual frost days were consistently identified as good predictors of productivity across fluxes.
+*How does productivity relate to other climate variables?* {**Here, maybe add some statement about covariation among climate variables, site SI figure**} Our results indicated that productivity was most strongly explained by temperature at the global scale, with temperature-related climate variables coming out as strong predictors of productivity.
+In addition to MAT, temperature seasonality, annual temperature range, and annual frost days were consistently identified as strong univariate predictors of FACF.
+{**cite some table/ SI?**}
 
@@
 
-*What is the role of seasonality in explaining productivity?* Temperature seasonality was a significant predictor of productivity.
-GPP, NPP, ANPP, and R~root~ exhibited a polynomial relationship with seasonality (fig.
+*What is the role of seasonality in explaining FACF?* Temperature seasonality was a significant predictor of FACF.
+GPP, NPP, ANPP, and R~root~ exhibited a polynomial relationship with seasonality (Fig.
+4) {**give statistic on signficiance**}.
+ANPP~foliage~, ANPP~woody~ ~stem~ and R~auto~ decreased linearly with temperature seasonality.
+Temperature seasonality was strongly correlated with annual temperature range, and, as expected, all fluxes showed almost identical responses to it {**meaning unclear**}.
+Productivity was highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower.
+In contrast, there was no significant effect of precipitation seasonality on FACF.
+
+We found a significant relationship between length of growing season and FACF, with all fluxes showing a linear increase in productivity with length of growing season (Fig.
 4).
-ANPP~foliage~, ANPP~woody~ ~stem~ and R~auto~ decreased linearly with temperature seasonality.
-Temperature seasonality was strongly correlated with annual temperature range, and, as expected, all fluxes showed almost identical responses to it.
-Productivity was highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower.
-In contrast, there was no significant effect of precipitation seasonality on productivity.
+Length of growing season was a strong predictor of FACF, explaining 51% of variation in GPP, 39% of variation in NPP, and 34% of variation in ANPP, but it was a weaker predictor than MAT for all fluxes analysed.
 
-We found a significant relationship between length of growing season and productivity, with all fluxes showing a linear increase in productivity with length of growing season (fig.
-4).
-Length of growing season was a strong predictor of productivity, explaining 51% of variation in GPP, 39% of variation in NPP, and 34% of variation in ANPP, but it was not identified as a stronger predictor than MAT for any of the fluxes analysed.
-
-*Within the growing season, how does productivity vary with climate?* Within growing season months, we found that climate had a much weaker effect on productivity.
-For each of temperature, precipitation, PET, and solar radiation, we found a small effect of climate for certain carbon fluxes.
-There was a small increase in productivity with temperature and precipitation for ANPP (with temperature $R^2$ = 0.10, p<0.001; with precipitation $R^2$ = 0.04, p<0.05) and ANPP~foliage~ (with temperature $R^2$ = 0.16, p<0.01; with precipitation $R^2$ = 0.09, p<0.05).
+*Within the growing season, how does FACF vary with climate?* When FACF was standardized by growing season length, correlations with climate variables were much weaker.
+Temperature, precipitation, PET, and solar radiation each had a small effect on one or more FACF (Fig.
+**S#**).
+Specifically, there was a small increase in productivity with temperature and precipitation for ANPP (with temperature $R^2$ = 0.10, p<0.001; with precipitation $R^2$ = 0.04, p<0.05) and ANPP~foliage~ (with temperature $R^2$ = 0.16, p<0.01; with precipitation $R^2$ = 0.09, p<0.05).
 Productivity increased with solar radiation for GPP ($R^2$ = 0.21, p<0.001), NPP ($R^2$ = 0.21, p<0.001), BNPP ($R^2$ = 0.16, p<0.001) and BNPP~fine~ ~root~ ($R^2$ = 0.12, p<0.01), and with PET for GPP ($R^2$ = 0.15, p<0.01), NPP ($R^2$ = 0.18, p<0.01), BNPP ($R^2$ = 0.23, p<0.0001), BNPP~fine~ ~root~ ($R^2$ = 0.11, p<0.05), and ANPP~woody~ ~stem~ ($R^2$ = 0.06, p<0.05).
+{**site SI figure, stats might work better in a table**}  {**did we look at respiration in this anlaysis?**}
 

### 2020-01-19 manuscript/manuscript draft.Rmd
finished going through Helene's comments

@@
 3-4).
-While other climate variables are significant predictors of FACF (*H3*), none of them improve on the explanatory power of temperature-related variables in general or MAT specifically (Fig.
+While other climate variables are significant predictors of FACF (*H3*), none of them in isolation improve on the explanatory power of temperature-related variables in general or MAT specifically (Fig.
 4).
@@
 Second, differences among variables in the proportion of variation explained by climate may be attributable to more dicrect climatic control over $GPP$ than subsidiary fluxes.
-That is, subsidiary fluxes may be shaped by climate both indirectly through its influence on $GPP$ and respiration and directly through any climatic influence on C allocation, as well as by many other local- and regional-scale factors (**REFS**).
+That is, subsidiary fluxes may be shaped by climate both indirectly through its influence on $GPP$ and respiration and directly through any climatic influence on C allocation, as well as many other local- and regional-scale factors (**REFS**).
 
@@
 
-FACF is reduced by temperature seasonality (Table 1, *H4*; Fig.
+FACF is negatively correlated with temperature seasonality (Table 1, *H4*; Fig.
 4e), and productivity is shut down during cold- or dry- dormant seasons.

### 2020-01-19 manuscript/manuscript draft.Rmd
worked on Table 2
issue #54

@@
 ```
-*Data selection.* Over 50 variables of forest carbon stocks and annual fluxes are represented in the ForC database; this analysis focussed on measures of primary productivity.
-Table 1 contains details of the variables selected for analysis.
-
-| Variable | Definition | Components included | Methodologies used | Number of records | Number of geographic areas |
-|--------------|-------------------|-------------|------------------------|-------|-------|
-| $GPP$ | Annual gross primary production; annual uptake of carbon dioxide by an ecosystem | NA | Flux partitioning of eddy covariance | 243 | 49 |
-| $NPP$ | Annual net primary production; the component of GPP that is stored in plant tissue; $GPP - R_{auto}$ | Foliage, branch, stem, coarse root and fine root | Direct measurement of annual increments of components | 92 | 42 |
-| $ANPP$ | Aboveground net primary production | Foliage, stem, and optionally branch | Direct measurement of annual increments of components | 256 | 80 |
-| $ANPP_{foliage}$ | Net primary production of foliage | Foliage | Direct measurement of litterfall, correcting for changes in leaf biomass when measured | 98 | 49 |
-| $ANPP_{woody-stem}$ | Net primary production of woody stems | Woody stems | Direct measurement of stem growth increment (**with recruitment?**) | 264 | 96 |
-| $BNPP_{root}$  | Belowground net primary production | Coarse and fine roots |  Coarse roots estimated indirectly using allometries based on aboveground stem increment measures ; fine roots as below | 101 | 48 |
-| $BNPP_{fine.root}$ | Net primary production of fine roots | Fine roots | Measurements combined one or more of the following: minirhizotrons, fine root turnover, soil cores, root ingrowth cores | 88 | 41 |
-| $R_{auto}$ | Annual autotrophic respiration, including above- and belowground components | Foliage, stem, and root | Chamber measurements of component gas exchange | 22 | 13 |
-| $R_{root}$ | Annual root respiration | Coarse and fine roots | *Measurement of root gas exchange, root exclusion from soil respiration chambers* | 64 | 26 |
-
-Table: Definitions and sample sizes of variables used in analysis.
-Geographic areas group geographically proximate sites, defined using a hierarchical cluster analysis on the distance matrix of the sites, and a cutoff of 25km.
+*Data selection.* Over 50 variables of forest carbon stocks and annual fluxes are represented in the ForC database; this analysis focused on autotrophic carbon fluxes.
+Focal variables of this analysis are listed in Table 2.
 
@@
 
+\newpage  
+| Variable | Definition | Components included | Methodologies used | Number of records | Number of geographic areas |
+|--------------|-------------------|-------------|------------------------|-------|-------|
+| $GPP$ | Annual gross primary production | NA | Flux partitioning of eddy-covariance; $NPP$+$R_{auto}$; scaling chamber measurements of photosynthesis | 243 | 49 |
+| $NPP$ | Annual net primary production | Foliage, branch, stem, coarse root and fine root | $ANPP$ + $BNPP$ (majority); $GPP$-$R_{auto}$; eddy-covariance measures of net ecosystem exchange + heterotrophic respiration  | 92 | 42 |
+| $ANPP$ | Annual aboveground net primary production | Foliage, stem, and optionally branch (**issues [52](https://github.com/forc-db/Global_Productivity/issues/52), [54](https://github.com/forc-db/Global_Productivity/issues/54)**) | Summing components | 256 | 80 |
+| $ANPP_{foliage}$ | Annual net primary production of foliage | Foliage | Direct measurement of litterfall, correcting for changes in leaf biomass when measured | 98 | 49 |
+| $ANPP_{woody-stem}$ | Annual net primary production of woody stems | Woody stems (no branch turnover) | Direct measurement of stem growth increment (**with recruitment?**) | 264 | 96 |
+| $BNPP_{root}$  | Annual belowground net primary production | Coarse and fine roots |  Coarse roots estimated indirectly using allometries based on aboveground stem increment measures ; fine roots as below | 101 | 48 |
+| $BNPP_{fine.root}$ | Annual net primary production of fine roots | Fine roots | Measurements combined one or more of the following: minirhizotrons, fine root turnover, soil cores, root ingrowth cores | 88 | 41 |
+| $R_{auto}$ | Annual autotrophic respiration | Foliage, stem, and root | Chamber measurements of component gas exchange | 22 | 13 |
+| $R_{root}$ | Annual root respiration | Coarse and fine roots | *Measurement of root gas exchange, root exclusion from soil respiration chambers* | 64 | 26 |
+
+Table: Definitions and sample sizes of variables used in analysis.
+Geographic areas group geographically proximate sites, defined using a hierarchical cluster analysis on the distance matrix of the sites, and a cutoff of 25km.
+
 *Methodological consistency.* The data in ForC is derived from a range of studies, often employing different methods.
@@
 It would be preferable to standardise by minimum area sampled; however x% of plots in the database are 1 ha or under in size; excluding these plots would place significant constraints on sample size.
-(**really, we didn't exclude any studies based on methodology(all studies measured min DBH 10 or less...
-so this paragraph needs revision**)
+(**really, we didn't exclude any studies based on methodology (all studies measured min DBH 10 or less...
+) so this paragraph needs revision.
+Also, address whether we standardize for inclusion/ exclusion of recruitment, and address branchfall.**)
 

### 2020-01-20 manuscript/manuscript draft.Rmd
worked on growing season results

@@
 *Within the growing season, how does FACF vary with climate?* When FACF was standardized by growing season length, correlations with climate variables were much weaker.
-Temperature, precipitation, PET, and solar radiation each had a small effect on one or more FACF (Fig.
+Growing season temperature, precipitation, solar radiation, and PET each had a small effect on one or more of the nine FACF analyzed (Table 2; Fig.
 **S#**).
-Specifically, there was a small increase in productivity with temperature and precipitation for ANPP (with temperature $R^2$ = 0.10, p<0.001; with precipitation $R^2$ = 0.04, p<0.05) and ANPP~foliage~ (with temperature $R^2$ = 0.16, p<0.01; with precipitation $R^2$ = 0.09, p<0.05).
-Productivity increased with solar radiation for GPP ($R^2$ = 0.21, p<0.001), NPP ($R^2$ = 0.21, p<0.001), BNPP ($R^2$ = 0.16, p<0.001) and BNPP~fine~ ~root~ ($R^2$ = 0.12, p<0.01), and with PET for GPP ($R^2$ = 0.15, p<0.01), NPP ($R^2$ = 0.18, p<0.01), BNPP ($R^2$ = 0.23, p<0.0001), BNPP~fine~ ~root~ ($R^2$ = 0.11, p<0.05), and ANPP~woody~ ~stem~ ($R^2$ = 0.06, p<0.05).
-{**site SI figure, stats might work better in a table**}  {**did we look at respiration in this anlaysis?**}
+Speficifally, $ANPP$ increased with growing season temperature ($R^2$ = 0.10, p<0.001) and precipitation ($R^2$ = 0.04, p<0.05).
+Similaryly, $ANPP_{foliage}$ increased slightly with growing season temperature ($R^2$ = 0.16, p<0.01) and precipitation ($R^2$ = 0.09, p<0.05).
+Growing season solar radiation had a positive influence on $GPP$ ($R^2$ = 0.21, p<0.001), $NPP$ ($R^2$ = 0.21, p<0.001), $BNPP$ ($R^2$ = 0.16, p<0.001) and $BNPP_{fine.root}$ ($R^2$ = 0.12, p<0.01).
+Growing season PET had a positive influence on $GPP$ ($R^2$ = 0.15, p<0.01), $NPP$ ($R^2$ = 0.18, p<0.01), $BNPP$ ($R^2$ = 0.23, p<0.0001), $BNPP_{fine.root}$ ($R^2$ = 0.11, p<0.05), and $ANPP_{woody-stem}$ ($R^2$ = 0.06, p<0.05).
+{**Becky, please verify/ edit the following:** There were no other signifcant correlations between growing season length-standardized FACF (9 variables in Table 2) and growing season climate (**which variables?**)}.
 

### 2020-01-20 manuscript/manuscript draft.Rmd

@@
 |--------------|-------------------|-------------|------------------------|-------|-------|
-| $GPP$ | Annual gross primary production | NA | Flux partitioning of eddy-covariance; $NPP$+$R_{auto}$; scaling chamber measurements of photosynthesis | 243 | 49 |
-| $NPP$ | Annual net primary production | Foliage, branch, stem, coarse root and fine root | $ANPP$ + $BNPP$ (majority); $GPP$-$R_{auto}$; eddy-covariance measures of net ecosystem exchange + heterotrophic respiration  | 92 | 42 |
-| $ANPP$ | Annual aboveground net primary production | Foliage, stem, and optionally branch (**issues [52](https://github.com/forc-db/Global_Productivity/issues/52), [54](https://github.com/forc-db/Global_Productivity/issues/54)**) | Summing components | 256 | 80 |
-| $ANPP_{foliage}$ | Annual net primary production of foliage | Foliage | Direct measurement of litterfall, correcting for changes in leaf biomass when measured | 98 | 49 |
-| $ANPP_{woody-stem}$ | Annual net primary production of woody stems | Woody stems (no branch turnover) | Direct measurement of stem growth increment (**with recruitment?**) | 264 | 96 |
-| $BNPP_{root}$  | Annual belowground net primary production | Coarse and fine roots |  Coarse roots estimated indirectly using allometries based on aboveground stem increment measures ; fine roots as below | 101 | 48 |
-| $BNPP_{fine.root}$ | Annual net primary production of fine roots | Fine roots | Measurements combined one or more of the following: minirhizotrons, fine root turnover, soil cores, root ingrowth cores | 88 | 41 |
-| $R_{auto}$ | Annual autotrophic respiration | Foliage, stem, and root | Chamber measurements of component gas exchange | 22 | 13 |
-| $R_{root}$ | Annual root respiration | Coarse and fine roots | *Measurement of root gas exchange, root exclusion from soil respiration chambers* | 64 | 26 |
+| $GPP$ | Gross Primary Production | NA | flux partitioning of eddy-covariance; $NPP$+$R_{auto}$; scaling chamber measurements of photosynthesis | 243 | 49 |
+| $NPP$ | Net Primary Production | stem, foliage, coarse root, fine root, optionally others (e.g., branch, reproductive, understory) | $ANPP$ + $BNPP$ (majority); $GPP$-$R_{auto}$; eddy-covariance measures of net ecosystem exchange + heterotrophic respiration  | 92 | 42 |
+| $ANPP$ | Aboveground Net Primary Production | stem, foliage, optionally others (e.g., branch, reproductive, understory) (**issues [52](https://github.com/forc-db/Global_Productivity/issues/52), [54](https://github.com/forc-db/Global_Productivity/issues/54)**) | $ANPP_{foliage}$ + $ANPP_{woody-stem}$ (+ others) | 256 | 80 |
+| $ANPP_{woody-stem}$ | Net Primary Production of woody stems | woody stems down to DBH $\le$ 10cm (no branch turnover) | direct measurement of stem growth increment (**with recruitment?**) | 264 | 96 |
+| $ANPP_{foliage}$ | Net Primary Production of foliage | foliage | direct measurement of litterfall, correcting for changes in leaf biomass when measured | 98 | 49 |
+| $BNPP_{root}$  | Belowground Net Primary Production | coarse and fine roots |  coarse roots estimated indirectly using allometries based on aboveground stem increment measures ; fine roots as below | 101 | 48 |
+| $BNPP_{fine.root}$ | Net Primary Production of fine roots | fine roots | measurements combined one or more of the following: minirhizotrons, fine root turnover, soil cores, root ingrowth cores | 88 | 41 |
+| $R_{root}$ | Root Respiration | **coarse** and fine roots | *Measurement of root gas exchange, root exclusion from soil respiration chambers* | 64 | 26 |
+| $R_{auto}$ | Autotrophic Respiration | foliage, stem, and root | chamber measurements of component gas exchange | 22 | 13 |
 
-Table: Definitions and sample sizes of variables used in analysis.
+Table: Definitions and sample sizes of variables (all annual) used in analysis.
 Geographic areas group geographically proximate sites, defined using a hierarchical cluster analysis on the distance matrix of the sites, and a cutoff of 25km.

### 2020-01-20 manuscript/manuscript draft.Rmd
Table 2- removed unused GPP and NPP methodologies
removed GPP and NPP methodologies for which ForC has no records in mature unmanaged stands

@@
 |--------------|-------------------|-------------|------------------------|-------|-------|
-| $GPP$ | Gross Primary Production | NA | flux partitioning of eddy-covariance; $NPP$+$R_{auto}$; scaling chamber measurements of photosynthesis | 243 | 49 |
-| $NPP$ | Net Primary Production | stem, foliage, coarse root, fine root, optionally others (e.g., branch, reproductive, understory) | $ANPP$ + $BNPP$ (majority); $GPP$-$R_{auto}$; eddy-covariance measures of net ecosystem exchange + heterotrophic respiration  | 92 | 42 |
+| $GPP$ | Gross Primary Production | NA | flux partitioning of eddy-covariance; $NPP$+$R_{auto}$ | 243 | 49 |
+| $NPP$ | Net Primary Production | stem, foliage, coarse root, fine root, optionally others (e.g., branch, reproductive, understory) | $ANPP$ + $BNPP$ (majority); $GPP$-$R_{auto}$ | 92 | 42 |
 | $ANPP$ | Aboveground Net Primary Production | stem, foliage, optionally others (e.g., branch, reproductive, understory) (**issues [52](https://github.com/forc-db/Global_Productivity/issues/52), [54](https://github.com/forc-db/Global_Productivity/issues/54)**) | $ANPP_{foliage}$ + $ANPP_{woody-stem}$ (+ others) | 256 | 80 |
@@
 | $BNPP_{fine.root}$ | Net Primary Production of fine roots | fine roots | measurements combined one or more of the following: minirhizotrons, fine root turnover, soil cores, root ingrowth cores | 88 | 41 |
-| $R_{root}$ | Root Respiration | **coarse** and fine roots | *Measurement of root gas exchange, root exclusion from soil respiration chambers* | 64 | 26 |
+| $R_{root}$ | Root Respiration | **coarse** and fine roots | *measurement of root gas exchange, root exclusion from soil respiration chambers* | 64 | 26 |
 | $R_{auto}$ | Autotrophic Respiration | foliage, stem, and root | chamber measurements of component gas exchange | 22 | 13 |

### 2020-01-20 manuscript/manuscript draft.Rmd
Table 2

@@
 | $NPP$ | Net Primary Production | stem, foliage, coarse root, fine root, optionally others (e.g., branch, reproductive, understory) | $ANPP$ + $BNPP$ (majority); $GPP$-$R_{auto}$ | 92 | 42 |
-| $ANPP$ | Aboveground Net Primary Production | stem, foliage, optionally others (e.g., branch, reproductive, understory) (**issues [52](https://github.com/forc-db/Global_Productivity/issues/52), [54](https://github.com/forc-db/Global_Productivity/issues/54)**) | $ANPP_{foliage}$ + $ANPP_{woody-stem}$ (+ others) | 256 | 80 |
+| $ANPP$ | Aboveground Net Primary Production | stem, foliage, optionally others (e.g., branch, reproductive, understory) (**issues [52](https://github.com/forc-db/Global_Productivity/issues/52), [54](https://github.com/forc-db/Global_Productivity/issues/54)**) |$ANPP_{woody-stem}$ + $ANPP_{foliage}$ (+ others) | 256 | 80 |
 | $ANPP_{woody-stem}$ | Net Primary Production of woody stems | woody stems down to DBH $\le$ 10cm (no branch turnover) | direct measurement of stem growth increment (**with recruitment?**) | 264 | 96 |

### 2020-01-20 manuscript/manuscript draft.Rmd
Table 2

@@
 | $ANPP_{woody-stem}$ | Net Primary Production of woody stems | woody stems down to DBH $\le$ 10cm (no branch turnover) | direct measurement of stem growth increment (**with recruitment?**) | 264 | 96 |
-| $ANPP_{foliage}$ | Net Primary Production of foliage | foliage | direct measurement of litterfall, correcting for changes in leaf biomass when measured | 98 | 49 |
+| $ANPP_{foliage}$ | Net Primary Production of foliage | foliage | direct measurement of litterfall | 98 | 49 |
 | $BNPP_{root}$  | Belowground Net Primary Production | coarse and fine roots |  coarse roots estimated indirectly using allometries based on aboveground stem increment measures ; fine roots as below | 101 | 48 |

### 2020-01-20 manuscript/manuscript draft.Rmd
Table 2

@@
 | $ANPP$ | Aboveground Net Primary Production | stem, foliage, optionally others (e.g., branch, reproductive, understory) (**issues [52](https://github.com/forc-db/Global_Productivity/issues/52), [54](https://github.com/forc-db/Global_Productivity/issues/54)**) |$ANPP_{woody-stem}$ + $ANPP_{foliage}$ (+ others) | 256 | 80 |
-| $ANPP_{woody-stem}$ | Net Primary Production of woody stems | woody stems down to DBH $\le$ 10cm (no branch turnover) | direct measurement of stem growth increment (**with recruitment?**) | 264 | 96 |
-| $ANPP_{foliage}$ | Net Primary Production of foliage | foliage | direct measurement of litterfall | 98 | 49 |
+| $ANPP_{woody-stem}$ | Net Primary Production of woody stems | woody stems down to DBH $\le$ 10cm (no branch turnover) | stem growth measurements scaled to biomass using allometries (**with recruitment?**) | 264 | 96 |
+| $ANPP_{foliage}$ | Net Primary Production of foliage | foliage | litterfall collection (separated into components) | 98 | 49 |
 | $BNPP_{root}$  | Belowground Net Primary Production | coarse and fine roots |  coarse roots estimated indirectly using allometries based on aboveground stem increment measures ; fine roots as below | 101 | 48 |

### 2020-01-20 manuscript/manuscript draft.Rmd
Table 2

--- 2020-01-20 manuscript/manuscript draft.Rmd
@@
 | $ANPP$ | Aboveground Net Primary Production | stem, foliage, optionally others (e.g., branch, reproductive, understory) (**issues [52](https://github.com/forc-db/Global_Productivity/issues/52), [54](https://github.com/forc-db/Global_Productivity/issues/54)**) |$ANPP_{woody-stem}$ + $ANPP_{foliage}$ (+ others) | 256 | 80 |
-| $ANPP_{woody-stem}$ | Net Primary Production of woody stems | woody stems down to DBH $\le$ 10cm (no branch turnover) | direct measurement of stem growth increment (**with recruitment?**) | 264 | 96 |
-| $ANPP_{foliage}$ | Net Primary Production of foliage | foliage | direct measurement of litterfall | 98 | 49 |
+| $ANPP_{woody-stem}$ | Net Primary Production of woody stems | woody stems down to DBH $\le$ 10cm (no branch turnover) | stem growth measurements scaled to biomass using allometries (**with recruitment?**) | 264 | 96 |
+| $ANPP_{foliage}$ | Net Primary Production of foliage | foliage | litterfall collection (separated into components) | 98 | 49 |
 | $BNPP_{root}$  | Belowground Net Primary Production | coarse and fine roots |  coarse roots estimated indirectly using allometries based on aboveground stem increment measures ; fine roots as below | 101 | 48 |
 | $BNPP_{fine.root}$ | Net Primary Production of fine roots | fine roots | measurements combined one or more of the following: minirhizotrons, fine root turnover, soil cores, root ingrowth cores | 88 | 41 |
-| $R_{root}$ | Root Respiration | **coarse** and fine roots | *measurement of root gas exchange, root exclusion from soil respiration chambers* | 64 | 26 |
+| $R_{root}$ | Root Respiration | coarse and fine roots | *measurement of root gas exchange, root exclusion from soil respiration chambers* | 64 | 26 |
 | $R_{auto}$ | Autotrophic Respiration | foliage, stem, and root | chamber measurements of component gas exchange | 22 | 13 |

### 2020-01-20 manuscript/manuscript-draft.Rmd
Table 2

--- 2020-01-20 manuscript/manuscript draft.Rmd
@@
 | $ANPP$ | Aboveground Net Primary Production | stem, foliage, optionally others (e.g., branch, reproductive, understory) (**issues [52](https://github.com/forc-db/Global_Productivity/issues/52), [54](https://github.com/forc-db/Global_Productivity/issues/54)**) |$ANPP_{woody-stem}$ + $ANPP_{foliage}$ (+ others) | 256 | 80 |
-| $ANPP_{woody-stem}$ | Net Primary Production of woody stems | woody stems down to DBH $\le$ 10cm (no branch turnover) | direct measurement of stem growth increment (**with recruitment?**) | 264 | 96 |
-| $ANPP_{foliage}$ | Net Primary Production of foliage | foliage | direct measurement of litterfall | 98 | 49 |
+| $ANPP_{woody-stem}$ | Net Primary Production of woody stems | woody stems down to DBH $\le$ 10cm (no branch turnover) | stem growth measurements scaled to biomass using allometries (**with recruitment?**) | 264 | 96 |
+| $ANPP_{foliage}$ | Net Primary Production of foliage | foliage | litterfall collection (separated into components) | 98 | 49 |
 | $BNPP_{root}$  | Belowground Net Primary Production | coarse and fine roots |  coarse roots estimated indirectly using allometries based on aboveground stem increment measures ; fine roots as below | 101 | 48 |
 | $BNPP_{fine.root}$ | Net Primary Production of fine roots | fine roots | measurements combined one or more of the following: minirhizotrons, fine root turnover, soil cores, root ingrowth cores | 88 | 41 |
-| $R_{root}$ | Root Respiration | **coarse** and fine roots | *measurement of root gas exchange, root exclusion from soil respiration chambers* | 64 | 26 |
+| $R_{root}$ | Root Respiration | coarse and fine roots | partitioning of total soil respiration (e.g., through root exclusion), scaling of root gas exchange | 64 | 26 |
 | $R_{auto}$ | Autotrophic Respiration | foliage, stem, and root | chamber measurements of component gas exchange | 22 | 13 |

### 2020-01-20 manuscript/manuscript-draft.tex
Table 2

@@
             pdftitle={Global patterns of forest autotrophic carbon fluxes},
-            pdfauthor={Rebecca Banbury Morgan, Valentine Herrmann, Norbert Kunert, Ben Bond-Lamberty, Helene Muller-Landau, Kristina J. Anderson-Teixeira},
+            pdfauthor={Rebecca Banbury Morgan, Valentine Herrmann, Norbert Kunert, Ben Bond-Lamberty, Helene C. Muller-Landau, Kristina J. Anderson-Teixeira},
             pdfborder={0 0 0},
@@
     \author{Rebecca Banbury Morgan, Valentine Herrmann, Norbert Kunert, Ben
-Bond-Lamberty, Helene Muller-Landau, Kristina J. Anderson-Teixeira}
+Bond-Lamberty, Helene C. Muller-Landau, Kristina J. Anderson-Teixeira}
     \preauthor{\centering\large\emph}
@@
 cycling decreases linearly with latitude --a finding that confirms
-multiple previous studies but contradicts the idea that productivity of
+multiple previous studies and contradicts the idea that productivity of
 temperate forests rivals that of tropical forests. The FACF variables
@@
 in particular being the best predictors of FACF on this global scale.
-While other climate variables (\emph{e.g.}, displayed significant
-correlation with FACF, none of them improved significantly over the
-explanatory power \(MAT\). The effects of temperature were modified by
-moisture availability, with reduced FACF under hot and dry conditions
+While other climate variables (\emph{e.g.}, \textbf{XX}) displayed
+significant correlation with FACF, none of them had significantly better
+explanatory power than \(MAT\). The effects of temperature were modified
+by moisture availability, with reduced FACF under hot and dry conditions
 and sometimes under very high precipitaiton (especially for \(BNPP\)).
@@
 
-\textbf{Forests play a critical role in regulating atmospheric CO\_2 and
-climate} (\textbf{Bonan}). The total amount of CO\_2 cycling through
-Earth's forests each year (\emph{i.e.}, total gross primary production,
-\(GPP\)) is more than five times anthropogenic fossil fuel emissions,
-and net sequestration (\textasciitilde{}2.4 GtCyr\textsuperscript{-1})
-offsetting roughly 30\% of these emissions \citep{pan_large_2011}. As
-atmospheric carbon dioxide levels continue to rise, with consequences
-for global climate, there is increasing recognition that proper
-protection and management of forest resources will have an important
-role to play in mitigating climate change (** REFS- e.g., IPCC 1.5C and
-land reports**). Carbon fixation, allocation, and metabolism by trees
-set the basis for energy and material flows in forest ecosystems and
-define their interactions with Earth's changing climate.
+(\textbf{revisit these first two paragraphs.})
+
+\textbf{Forests play a critical role in regulating atmospheric
+CO\textsubscript{2} and climate} (\textbf{Bonan}). Approximately
+\textbf{\#\#} Gt of CO\textsubscript{2} cycle through Earth's forests
+each year \textbf{REF-KAT}), and their net C sequestration
+(\textasciitilde{}2.4 GtCyr\textsuperscript{-1}) offsets roughly 30\% of
+anthropogenic fossil fuel emissions \citep{pan_large_2011}. As
+atmospheric carbon dioxide levels continue to rise, driving climate
+change, there is increasing recognition that protection and management
+of forest resources will play an important role in mitigating climate
+change (** REFS- e.g., IPCC 1.5C and land reports**). Carbon fixation,
+allocation, and metabolism by trees set the basis for energy and
+material flows in forest ecosystems and define their interactions with
+Earth's changing climate.
 
@@
 Forest annual carbon cost: a global‐scale analysis of autotrophic
-respiration. Ecology 91(3), 652-661Ƴ} }). Forest productivity sets the
+respiration. Ecology 91(3), 652-661} }). Forest productivity sets the
 energy ultimately available to heterotrophic organisms (including
@@
 (\href{https://www.annualreviews.org/doi/abs/10.1146/annurev.ecolsys.30.1.257}{Waide
-et al. 2009}\textbf{REFS})--but probably not individual metabolic rate
-(\textbf{Anderson \& Jetz 2005, but check more recent literature}).
-However, our ability to draw general macroscopic conclusions regarding
-global variation in multiple productivity variables with respect to
-climate has been limited in that these analyses often mix forests that
-vary in stand age, disturbance history, and/or management status; do not
-always sufficiently parse related variables (\emph{e.g.}, combining net
-primary productivity records with and without belowground components);
-and typically consider only one or a few variables at a time.
+et al. 2009}\textbf{REFS}). However, our ability to draw general
+macroscopic conclusions regarding global variation in multiple
+productivity variables with respect to climate has been limited in that
+these analyses often mix forests that vary in stand age, disturbance
+history, and/or management status; do not always sufficiently parse
+related variables (\emph{e.g.}, combining net primary productivity
+records with and without belowground components); and typically consider
+only one or a few variables at a time.
 
@@
 measurements, which appear to reach their highest levels in tropical
-forests \citep{luyssaert_co_2007}. In contrast, there is evidence that
-the highest values of net primary productivity (NPP) may be found in
-temperate forests \citep{luyssaert_co_2007, huston_global_2009},
-although other studies find NPP is highest in the tropics, showing a
-decrease with latitude \citep{simova_enigma_2017}. Other studies have
-chosen to focus exclusively on above-ground net primary productivity
-(ANPP), finding evidence of a weak negative relationship between ANPP
-and latitude \citep{huston_global_2009, gillman_latitude_2015}.
+forests \citep{luyssaert_co_2007}. \{\textbf{condense the previous two
+sentences into one-- both primarily eddy flux}\} In contrast, some
+studies suggest that the highest values of net primary productivity
+(NPP) may be found in temperate forests
+\citep{luyssaert_co_2007, huston_global_2009}, while others find NPP
+highest in the tropics and decreasing with latitude
+\citep{simova_enigma_2017}. Other studies have chosen to focus
+exclusively on above-ground net primary productivity (ANPP), finding
+evidence of a weak negative relationship between ANPP and latitude
+\citep{huston_global_2009, gillman_latitude_2015}.
 
@@
 spatial scales} \citep{cleveland_relationships_2011}. The majority of
-studies have focused on exploring the relationships between productivity
-and mean annual temperature (MAT) and mean annual precipitation (MAP),
-as the most commonly reported site-level climate variables. These
-variables have the advantage that they describe broad trends in
-temperature and water availability, and therefore capture a lot of
-global-scale variation in climate. There is strong evidence that both
-MAT and MAP show significant positive relationships with productivity
+studies have focused on exploring the relationships of productivity to
+mean annual temperature (MAT) and mean annual precipitation (MAP), as
+the most commonly reported site-level climate variables. These variables
+have the advantage that they describe broad trends in temperature and
+water availability, and therefore capture a lot of global-scale
+variation in climate. There is strong evidence that both MAT and MAP
+show significant positive relationships with productivity
 \citep{chu_does_2016}. However, as with latitude, the shape of those
@@
 negative effect at low MAT, but a positive effect at high levels of MAT,
-and vice versa, suggesting that both temperature and water availability
-are important in explaining productivity.
+and \emph{vice versa} \{\textbf{Helene found this confusing}\},
+suggesting that both temperature and water availability are important in
+explaining productivity.
 
@@
 studies have suggested that the apparently strong relationship between
-MAT and productivity is actually a factor of the correlation between MAT
-and growing season length
+MAT and productivity is actually a consequence of the correlation
+between MAT and growing season length
 \citep{kerkhoff_plant_2005, malhi_productivity_2012, michaletz_convergence_2014, michaletz_drivers_2018}.
@@
 season, there is no significant relationship between productivity and
-MAT, indicating that the effect of temperature is due to increased
+MAT, suggesting that the effect of temperature is due to increased
 length of growing season, rather than an inherent influence of
@@
 five broad hypotheses and corresponding specific predictions (Table 1).
-First, we ask how forest autotrophic carbon fluxes (FACF) vary with
-latitude. We then test how these fluxes relate to MAT and MAP, and
+** see
+\href{https://github.com/forc-db/Global_Productivity/issues/48}{issue
+\#48}** First, we ask how forest autotrophic carbon fluxes (FACF) vary
+with latitude. We then test how these fluxes relate to MAT and MAP, and
 additionally how they respond to other, less well studied, climate
-variables. We finally consider the relationship between FACF and
+variables. Finally, we consider the relationship between FACF and
 seasonality, considering the role of seasonality in explaining variation
-in carbon fluxes, and the influence of climate on FACF per month of the
-growing season. We use a comprehensive global database of forest carbon
-fluxes to explore the above questions for nine carbon fluxes, allowing
-for an in-depth exploration of the effect of climate on global
+in carbon fluxes, and the influence of climate on FACF standardized by
+growing season length. We use a comprehensive global database of forest
+carbon fluxes to address the above questions for nine carbon fluxes,
+allowing for an in-depth exploration of the effect of climate on global
 productivity.
@@
 \multicolumn{1}{l}{\textbf{H1. FACFs decrease linearly with latitude, and in constant proportion.}}\\
-\hspace{1em}1.1. FACFs decrease linearly with latitude (L-) & Luyssaert et al. (2007) & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Fig. 2-3\\
+\hspace{1em}1.1. FACFs decrease linearly with latitude (L-) & Luyssaert et al. (2007) & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Fig. 2\\
 \hspace{1em}1.1.alt.
@@
 (2015); Simova and Storch (2017) & no & no & no & no & no & no & no & no & no & no & Fig.
-2-3\\
+2\\
 \hspace{1em}1.2. Allocation of GPP to subsidiary fluxes varies with latitude & Litton et al. (2007); DeLucia et al. (2007) & no & - & no & no & no & no & no & no & - & - & \\
@@
 \multicolumn{1}{l}{\textbf{H2. FACFs increase with MAT, with an interactive effect of MAP.}}\\
-\hspace{1em}2.1. FACFs increase linearly with MAT (L+) & Schuur (2003) & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Fig. 5\\
+\hspace{1em}2.1. FACFs increase linearly with MAT (L+) & Schuur (2003) & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Fig. 4\\
 \hspace{1em}2.2.
@@
 & L+ & L+ & yes & Fig.
-5\\
-\hspace{1em}2.3. There is a postive interaction between temperature and precipitation (I+) & Taylor et al. (2016) & (yes) & yes & L+ & yes & L- & L+ & yes & L- & yes & yes & Fig. 4\\
+4\\
+\hspace{1em}2.3. There is a postive interaction between temperature and precipitation (I+) & Taylor et al. (2016) & (yes) & yes & L+ & yes & L- & L+ & yes & L- & yes & yes & Fig. 3\\
 \addlinespace[1em]
@@
 \multicolumn{1}{l}{\textbf{H3. FACFs are strongly correlated with other annual climate variables.}}\\
-\hspace{1em}3.1. FACFs increases with PET, but saturates or decreases at high levels (CD+) &  & (yes) & yes & yes & yes & yes & L+ & yes & L+ & yes & L+ & Fig. 5\\
-\hspace{1em}3.2. FACFs increase with vapour pressure deficit, but saturate or decrease at high levels (CD+) &  & (yes) & yes & yes & yes & yes & yes & yes & L+ & yes & yes & Fig. 5\\
+\hspace{1em}3.1. FACFs increases with PET, but saturates or decreases at high levels (CD+) &  & (yes) & yes & yes & yes & yes & L+ & yes & L+ & yes & L+ & Fig. 4\\
+\hspace{1em}3.2. FACFs increase with vapour pressure deficit, but saturate or decrease at high levels (CD+) &  & (yes) & yes & yes & yes & yes & yes & yes & L+ & yes & yes & Fig. 4\\
 \hspace{1em}3.3. FACFs increase linearly with solar radiation (L+) &  & (yes) & yes & yes & yes & CD+ & yes & CD+ & yes & n.s. & yes & \\
@@
 \multicolumn{1}{l}{\textbf{H4. FACFs are reduced under seasonal climates.}}\\
-\hspace{1em}4.1. FACFs decrease linearly with temperature seasonality (L-) &  & yes & CU+ & CU+ & CU+ & yes & yes & yes & yes & CU+ & CU+ & Fig. 5\\
+\hspace{1em}4.1. FACFs decrease linearly with temperature seasonality (L-) &  & yes & CU+ & CU+ & CU+ & yes & yes & yes & yes & CU+ & CU+ & Fig. 4\\
 \hspace{1em}4.2. FACFs decrease linearly with  precipitation seasonality (L-) &  & no & n.s. & n.s. & n.s. & n.s. & n.s. & n.s. & n.s. & n.s. & n.s. & \\
@@
 (2016) & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Fig.
-5\\
+4\\
 \hspace{1em}4.4. Growing season length is a better predictor of FACFs than MAT & Michaletz et al. (2014); Chu et al. (2016) & no & no & no & no & no & no & no & no & no & no & \\
@@
 of \citep{anderson-teixeira_forc:_2018} through integration with the
-Global Database of Soil Respiration Database
-\citep{bond-lamberty_global_2010}. Additional targeted literature
-searches were conducted to identify any further available data on
-primary productivity, with particular focus on mature forests in
-temperate and boreal regions. ForC currently contains 29730 records from
-4979 plots, representing 20 distinct ecozones across all forested
-biogeographic and climate zones. We used ForC v3.0, archived on Zenodo
-with DOI 10.5281/zenodo.3403855.
+Global Soil Respiration Database \citep{bond-lamberty_global_2010}.
+Additional targeted literature searches were conducted to identify any
+further available data on primary productivity, with particular focus on
+mature forests in temperate and boreal regions. ForC currently contains
+29730 records from 4979 plots, representing 20 distinct ecozones across
+all forested biogeographic and climate zones. We used ForC v3.0,
+archived on Zenodo with DOI 10.5281/zenodo.3403855.
 
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{distribution_all_variables} \caption{Map showing all data used in the analysis, coded by variable}\label{fig:unnamed-chunk-5}
+\includegraphics[width=1\linewidth]{distribution_all_variables_cropped} \caption{Map showing all data used in the analysis, coded by variable.
+Variables are plotted individually in Fig.
+S1.
+}\label{fig:unnamed-chunk-5}
 \end{figure}
@@
 annual fluxes are represented in the ForC database; this analysis
-focussed on measures of primary productivity. Table 1 contains details
-of the variables selected for analysis.
+focused on autotrophic carbon fluxes. Focal variables of this analysis
+are listed in Table 2.
+
+A subset of the ForC database was generated for the purposes of this
+analysis, in order to control for data quality and remove biasing
+factors. Since management can alter observed patterns of primary
+productivity \citep{simova_enigma_2017}, sites were excluded from
+analysis if they were managed, defined as plots that were planted,
+managed as plantations, irrigated, fertilised or including the term
+``managed'' in their site description. Sites that had experienced
+significant disturbance were also excluded. Disturbances that justified
+site exclusion were major cutting or harvesting, and/or burning,
+flooding, drought and storm events with site mortality
+\textgreater{}10\% of trees. Grazed sites were retained.\\
+There is evidence that stand age influences patterns of primary
+productivity and carbon allocation in forest ecosystems, and can
+confound relationships between latitude and primary productivity
+\citep{de_lucia_forest_2007, gillman_latitude_2015}. To reduce any
+biasing effects of stand age, we included only stands of known age
+\(\ge\) 100 years and those described by terms such as ``mature'',
+``intact'', or ``old-growth''.
+
+\newpage
 
 \begin{longtable}[]{@{}llllll@{}}
-\caption{Definitions and sample sizes of variables used in analysis.
-Geographic areas group geographically proximate sites, defined using a
-hierarchical cluster analysis on the distance matrix of the sites, and a
-cutoff of 25km.}\tabularnewline
+\caption{Definitions and sample sizes of variables (all annual) used in
+analysis. Geographic areas group geographically proximate sites, defined
+using a hierarchical cluster analysis on the distance matrix of the
+sites, and a cutoff of 25km.}\tabularnewline
 \toprule
@@
 \end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
-Annual gross primary production; annual uptake of carbon dioxide by an
-ecosystem\strut
+Gross Primary Production\strut
 \end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
@@
 \end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-Flux partitioning of eddy covariance\strut
+flux partitioning of eddy-covariance; \(NPP\)+\(R_{auto}\)\strut
 \end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
@@
 \end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
-Annual net primary production; the component of GPP that is stored in
-plant tissue; \(GPP - R_{auto}\)\strut
+Net Primary Production\strut
 \end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
-Foliage, branch, stem, coarse root and fine root\strut
+stem, foliage, coarse root, fine root, optionally others (e.g., branch,
+reproductive, understory)\strut
 \end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-Direct measurement of annual increments of components\strut
+\(ANPP\) + \(BNPP\) (majority); \(GPP\)-\(R_{auto}\)\strut
 \end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
@@
 \end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
-Aboveground net primary production\strut
+Aboveground Net Primary Production\strut
 \end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
-Foliage, stem, and optionally branch\strut
+stem, foliage, optionally others (e.g., branch, reproductive,
+understory) (\textbf{issues
+\href{https://github.com/forc-db/Global_Productivity/issues/52}{52},
+\href{https://github.com/forc-db/Global_Productivity/issues/54}{54}})\strut
 \end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-Direct measurement of annual increments of components\strut
+\(ANPP_{woody-stem}\) + \(ANPP_{foliage}\) (+ others)\strut
 \end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
@@
 80\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+\(ANPP_{woody-stem}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
+Net Primary Production of woody stems\strut
+\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
+woody stems down to DBH \(\le\) 10cm (no branch turnover)\strut
+\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
+stem growth measurements scaled to biomass using allometries
+(\textbf{with recruitment?})\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
+264\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
+96\strut
 \end{minipage}\tabularnewline
@@
 \end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
-Net primary production of foliage\strut
+Net Primary Production of foliage\strut
 \end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
-Foliage\strut
+foliage\strut
 \end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-Direct measurement of litterfall, correcting for changes in leaf biomass
-when measured\strut
+litterfall collection (separated into components)\strut
 \end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
@@
 49\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
-\(ANPP_{woody-stem}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
-Net primary production of woody stems\strut
-\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
-Woody stems\strut
-\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-Direct measurement of stem growth increment\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
-264\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
-96\strut
 \end{minipage}\tabularnewline
@@
 \end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
-Belowground net primary production\strut
+Belowground Net Primary Production\strut
 \end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
-Coarse and fine roots\strut
+coarse and fine roots\strut
 \end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-Direct measurement of one or more of: fine root turnover, soil cores,
-root ingrowth cores, minirhizotrons; indirect estimates of coarse roots
-using allometries based on aboveground stem increment measures\strut
+coarse roots estimated indirectly using allometries based on aboveground
+stem increment measures ; fine roots as below\strut
 \end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
@@
 \end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
-Net primary production of fine roots\strut
+Net Primary Production of fine roots\strut
 \end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
-Fine roots\strut
+fine roots\strut
 \end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-Direct measurement of one or more of: minirhizotrons, fine root
-turnover, soil cores, root ingrowth cores\strut
+measurements combined one or more of the following: minirhizotrons, fine
+root turnover, soil cores, root ingrowth cores\strut
 \end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
@@
 41\strut
+\end{minipage}\tabularnewline
+\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+\(R_{root}\)\strut
+\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
+Root Respiration\strut
+\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
+coarse and fine roots\strut
+\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
+partitioning of total soil respiration (e.g., through root exclusion),
+scaling of root gas exchange\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
+64\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
+26\strut
 \end{minipage}\tabularnewline
@@
 \end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
-Annual autotrophic respiration, including above- and belowground
-components\strut
+Autotrophic Respiration\strut
 \end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
-Foliage, stem, and root\strut
+foliage, stem, and root\strut
 \end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-Chamber measurements of component gas exchange\strut
+chamber measurements of component gas exchange\strut
 \end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
@@
 13\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
-\(R_{root}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
-Annual root respiration\strut
-\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
-Coarse and fine roots\strut
-\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-\emph{Measurement of root gas exchange, root exclusion from soil
-respiration chambers}\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
-64\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
-26\strut
 \end{minipage}\tabularnewline
@@
 \end{longtable}
-
-A subset of the ForC database was generated for the purposes of this
-analysis, in order to control for data quality and remove biasing
-factors. Since management can alter observed patterns of primary
-productivity \citep{simova_enigma_2017}, sites were excluded from
-analysis if they were managed, defined as plots that were planted,
-managed as plantations, irrigated, fertilised or including the term
-``managed'' in their site description. Sites that had experienced
-significant disturbance were also excluded. Disturbances that justified
-site exclusion were major cutting or harvesting, and/or burning,
-flooding, drought and storm events with site mortality
-\textgreater{}10\% of trees. Grazed sites were retained.\\
-There is evidence that stand age influences patterns of primary
-productivity and carbon allocation in forest ecosystems, and can
-confound relationships between latitude and primary productivity
-\citep{de_lucia_forest_2007, gillman_latitude_2015}. To reduce any
-biasing effects of stand age, stands under 100 years of age were
-excluded from analysis. Sites for which stand age was unknown were
-excluded from analysis.
 
@@
 database are 1 ha or under in size; excluding these plots would place
-significant constraints on sample size.
+significant constraints on sample size. (\textbf{really, we didn't
+exclude any studies based on methodology (all studies measured min DBH
+10 or less\ldots{} ) so this paragraph needs revision. Also, address
+whether we standardize for inclusion/ exclusion of recruitment, and
+address branchfall.})
 
@@
 components, including understorey, volatile organic compounds (VOCs),
-exudates, estimates of NPP lost to herbivory, and the NPP of
-reproductive structures, were excluded. Estimates of ANPP were selected
-if they included foliage, stem growth and optionally branch turnover.
-Any measures of primary productivity where components were unknown were
-excluded from analysis.
+exudates, estimates of NPP lost to herbivory (\textbf{these are all
+small, no need to exclude}), and the NPP of reproductive structures
+(\textbf{often in literfall/ mixed in with foliage\ldots{} did we really
+exclude estimates including litterfall?}), were excluded. Estimates of
+ANPP were selected if they included foliage, stem growth and optionally
+branch turnover. Any measures of primary productivity where components
+were unknown were excluded from analysis.
 
@@
 months, defined as the number of months annually where precipitation was
-lower than potential evapotranspiration.
+lower than potential evapotranspiration (PET; \textbf{HML:defined
+how?}).
 
@@
 
-\emph{Model specification.} The effects of climate and latitude on
-primary productivity were analysed using mixed effects models using the
-package `lme4' \citep{bates_fitting_2015} in R v.3.5.1
-\citep{r_core_team_r:_2018}. The effect of each extracted climate
-variable on each measure of primary productivity was modelled by
-specifying the climate variable as a fixed effect. For each climate
-variable, three models were specified: a null model; a model with the
-climate variable as a linear term; and a model with the climate variable
-as a polynomial term. AIC values were calculated for the models and used
-to select the best model. If the best model included a polynomial term,
-the shape of the polynomial relationship was considered. If the shape of
-the relationship \emph{made biological sense}, and was a significant
-improvement on the linear relationship (deltaAIC \textgreater{}2), we
-accepted the polynomial as the best model. If not, we ran the linear
-model as the final model. \(R^2\) values were calculated for the best
-model. All \(R^2\) values presented here are marginal \(R^2\) values,
-and refer to the proportion of variation explained by only the fixed
-effects, unless otherwise specified. In addition, slope coefficients
-were calculated for the linear models.
-
-Because the magnitude of fluxes varies significantly, in order to
-facilitate comparisons between regression models for each flux, data for
-each flux was scaled, to give the data a mean of 0 and standard
-deviation of 1. As each data set was scaled separately, this does not
-allow for statistical comparisons of slope values, but does assist in
-visualising the data.
+\emph{Model specification.}
+
+\{\textbf{from Helene: It was frequently unclear to me whether the
+influences of other climate variables were being considered in
+multivariate or univariate analyses. This needs to be clear. I can see
+reasons to do it either way, but they need to be differently motivated
+and clearly presented.}\}
+
+The effects of climate and latitude on primary productivity were
+analysed using mixed effects models using the package `lme4'
+\citep{bates_fitting_2015} in R v.3.5.1 \citep{r_core_team_r:_2018}. The
+effect of each extracted climate variable on each measure of primary
+productivity was modelled by specifying the climate variable as a fixed
+effect. For each climate variable, three models were specified: a null
+model; a model with the climate variable as a linear term; and a model
+with the climate variable as a \textbf{XX}-order polynomial term. AIC
+values were calculated for the models and used to select the best model.
+If the best model included a polynomial term, the shape of the
+polynomial relationship was considered. If the shape of the relationship
+\emph{made biological sense}, and was a significant improvement on the
+linear relationship (deltaAIC \textgreater{}2), we accepted the
+polynomial as the best model. If not, we ran the linear model as the
+final model. \(R^2\) values were calculated for the best model. All
+\(R^2\) values presented here are marginal \(R^2\) values, and refer to
+the proportion of variation explained by only the fixed effects, unless
+otherwise specified. In addition, slope coefficients were calculated for
+the linear models.
 
@@
 their explanatory power. As a result, it was decided to present results
-only from models do not include altitude as a fixed term.
+only from models without altitude as a fixed term.
 
@@
 temperate regions was heavily skewed towards studies from the old-growth
-forests of the Pacific Northwest.} These forests have very high
-productivity, and so to ensure that results were not unduly influenced
-by geographic sampling bias, we tried a version of the model where data
-were weighted according to forested land area within each Koeppen
-climate zone. Results were similar between the weighted and unweighted
-model, so, to avoid problems of over-fitting, the weighted model was
-dropped, and results from this are not presented here.
-
-Models were run for total annual productivity against annual climate
-variables, and for monthly growing season productivity, defined as total
-productivity/length of growing season, against growing season climate
-variables. For analyses on data within the growing season, only linear
-models were specified.
-
+forests of the Pacific Northwest.} These forests have very high FACF,
+and so to ensure that results were not unduly influenced by geographic
+sampling bias, we tried a version of the model where data were weighted
+according to forested land area within each Koeppen climate zone.
+Results were similar between the weighted and unweighted model. Because
+the weighted model may be prone to problems of over-fitting, we present
+only results of the unweighted model. \{**Becky, I'd probably just
+delete this whole paragraph. Since we ended up not using the weighted
+model, it's just extra info for readers to digest.\}
+
+Models were run for total annual FACF against annual climate variables,
+and for fluxes standardized by growing season length, defined as annual
+FACF/length of growing season, against growing season climate variables.
+For analyses on data within the growing season, only linear models were
+specified.
+
+\{\textbf{Note from Helene on the following paragraph: this makes it
+sound like analyses from H3 were about variation remaining after
+accounting for temperature, but analyses below don't seem to do
+that.}\}\\
 To investigate the potential interactive effects of climate variables on
@@
 ii) the r-squared value of the model was larger than the r-squared value
-of the baseline model by \textgreater{}5
+of the baseline model by \textbf{\textgreater{}5 (???- not possible)}
 
@@
 
-All major carbon fluxes increased linearly with decreasing latitude
-(fig. 2).
-
-\begin{figure}[H]
-\includegraphics[width=1\linewidth]{effect_of_lat_transparent} \caption{Graphs to show primary productivity $(MgC~ha^-1~yr^-1)$ regressed against latitude.
-Lines of best fit are plotted according to the best model selected during analysis.
-All regressions are significant $(p<0.05)$.}\label{fig:unnamed-chunk-6}
-\end{figure}
+All major carbon fluxes decreased linearly with latitude (Fig. 2).
 
@@
 component fluxes summed approximately to larger fluxes across the
-latitudinal gradient (fig. 3). That is, modelled estimates of GPP,
+latitudinal gradient (fig. 2). That is, modelled estimates of GPP,
 generated from the sum of NPP and R auto; NPP, generated from the sum of
@@
 test: it's easy for sub-fluxes not to sum up! (Extensive citations from
-EC literature for example.)
+EC literature for example.). HML: yes and no. confidence intervals are
+pretty large.
 
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{combined_stacked} \caption{Graphs of primary productivity $(MgC~ha^-1~yr^-1)$ regressed against latitude.
+\includegraphics[width=1\linewidth]{combined_stacked} \caption{Latitudinal trends in forest autotropic carbon flux.
 Lines of best fit are plotted according to the best model selected during analysis.
 All regressions are significant $(p<0.05)$.
-Plots 1 - 3 show two component fluxes; a larger flux, defined as the combination of the two component fluxes; and a modelled estimate of the sum of the two component fluxes.
-95\% confidence intervals are plotted for the larger flux.
-Plot 4 shows three belowground fluxes, and a modelled estimate of the total belowground carbon flux}\label{fig:unnamed-chunk-7}
+Each panel shows major C fluxes together with component fluxes.
+Also plotted are predicted trends in the major C fluxes based on the sum of component fluxes.
+95\% confidence intervals are plotted for the major flux for comparison with predicted trends.
+In (d),  which shows three belowground fluxes, the major flux, total belowground carbon flux, is one for which we have no data}\label{fig:unnamed-chunk-6}
 \end{figure}
@@
 fluxes tested, with all fluxes showing a linear increase with
-temperature (fig. 5). We found no support for a saturation point of
+temperature (fig. 4). We found no support for a saturation point of
 productivity with temperature.
@@
 point at between 3000 and 4000mm annual precipitation, above which
-productivity started to decrease (fig. 5). The notable exception to this
+productivity started to decrease (fig. 4). The notable exception to this
 was GPP: the model indicated that GPP continued to increase with
@@
 ANPP, ANPP\textsubscript{woody} \textsubscript{stem}, and
-R\textsubscript{root} (fig.4). There was a significant additive effect
+R\textsubscript{root} (fig.3). There was a significant additive effect
 for R\textsubscript{auto}. NPP and ANPP\textsubscript{foliage} showed no
-significant interactive or additive effect.
+significant interactive or additive effect. \{\textbf{NEED SI TABLE
+GIVING FITTED COEFFICIENTS \& STATS}\}
 
@@
 significant interactive effect between MAT and water stress months, with
-this model explaining nearly 5\% more variation in NPP than MAT alone.
-However, for ANPP\textsubscript{foliage}, no multivariate model improved
-on the univariate model including only MAT.
+this model explaining nearly 5\% more variation in NPP than MAT alone
+\{\textbf{Helene: but less than MATxMAP? if so, why bother
+mentioning?}\}. However, for ANPP\textsubscript{foliage}, no
+multivariate model improved on the univariate model including only MAT.
 
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{mat_map_interaction} \caption{Plots of primary productivity $(MgC~ha^-1~yr^-1)$ regressed against mean annual precipitation.
-Points are grouped into bins of 0 - 1000, 1001 - 2000, 2001 - 3000, and >3000mm mean annual precipitation, and lines of best fit plotted for mean annual precipitation values of 500, 1500, 2500, and 3500mm.
-All regressions are significant $(p<0.05)$.}\label{fig:unnamed-chunk-8}
+\includegraphics[width=1\linewidth]{mat_map_interaction} \caption{Interactive effects of mean annual temperature and mean annual precipitation on FACF.
+For visualization purposes, data points are grouped into bins of 0 - 1000, 1001 - 2000, 2001 - 3000, and >3000mm mean annual precipitation, and lines of best fit models are plotted for mean annual precipitation values of 500, 1500, 2500, and 3500mm.
+All regressions are significant $(p<0.05)$.}\label{fig:unnamed-chunk-7}
 \end{figure}
 
-\emph{How does productivity relate to other climate variables?} Our
-results indicated that productivity was most strongly explained by
-temperature at the global scale, with temperature-related climate
-variables coming out as strong predictors of productivity. In addition
-to MAT, temperature seasonality, annual temperature range, and annual
-frost days were consistently identified as good predictors of
-productivity across fluxes.
+\emph{How does productivity relate to other climate variables?}
+\{\textbf{Here, maybe add some statement about covariation among climate
+variables, site SI figure}\} Our results indicated that productivity was
+most strongly explained by temperature at the global scale, with
+temperature-related climate variables coming out as strong predictors of
+productivity. In addition to MAT, temperature seasonality, annual
+temperature range, and annual frost days were consistently identified as
+strong univariate predictors of FACF. \{\textbf{cite some table/ SI?}\}
 
@@
 increased linearly with PET; however, all other fluxes showed a
-polynomial relationship with PET (fig. 5). We found strong evidence for
+polynomial relationship with PET (fig. 4). We found strong evidence for
 a saturation point or peak with PET: productivity tended to increase at
@@
 p\textless{}0.05), but all other fluxes showed a polynomial relationship
-(fig. 5). Productivity initially increased with vapour pressure deficit,
+(fig. 4). Productivity initially increased with vapour pressure deficit,
 before saturating at around 0.8 kPa. At values above 0.8 kPa,
@@
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{combined_plots} \caption{Plots of primary productivity $(MgC~ha^-1~yr^-1)$ regressed against (a) mean annual temperature; (b) mean annual precipitation; (c) potential evapotranspiration, (d) vapour pressure deficit; (e) temperature seasonality; (f) length of growing season.
-Lines of best fit are plotted according to the best model selected during analysis.
-All regressions are significant $(p<0.05)$.}\label{fig:unnamed-chunk-9}
+\includegraphics[width=1\linewidth]{combined_plots} \caption{Plots of carbon fluxes against (a) mean annual temperature; (b) mean annual precipitation; (c) potential evapotranspiration, (d) vapour pressure deficit; (e) temperature seasonality; (f) length of growing season.
+For visualization purposes, data for each flux was rescaled with a mean of 0 and standard deviation of 1.
+Lines of best fit are plotted according to the best model selected during analysis (**see issue 47**).
+All regressions are significant $(p<0.05)$.}\label{fig:unnamed-chunk-8}
 \end{figure}
 
-\emph{What is the role of seasonality in explaining productivity?}
-Temperature seasonality was a significant predictor of productivity.
-GPP, NPP, ANPP, and R\textsubscript{root} exhibited a polynomial
-relationship with seasonality (fig. 5). ANPP\textsubscript{foliage},
-ANPP\textsubscript{woody} \textsubscript{stem} and R\textsubscript{auto}
-decreased linearly with temperature seasonality. Temperature seasonality
-was strongly correlated with annual temperature range, and, as expected,
-all fluxes showed almost identical responses to it. Productivity was
-highest where temperature seasonality = 0, and at an annual temperature
-range of 15\(^\circ\)C or lower. In contrast, there was no significant
-effect of precipitation seasonality on productivity.
+\emph{What is the role of seasonality in explaining FACF?} Temperature
+seasonality was a significant predictor of FACF. GPP, NPP, ANPP, and
+R\textsubscript{root} exhibited a polynomial relationship with
+seasonality (Fig. 4) \{\textbf{give statistic on signficiance}\}.
+ANPP\textsubscript{foliage}, ANPP\textsubscript{woody}
+\textsubscript{stem} and R\textsubscript{auto} decreased linearly with
+temperature seasonality. Temperature seasonality was strongly correlated
+with annual temperature range, and, as expected, all fluxes showed
+almost identical responses to it \{\textbf{meaning unclear}\}.
+Productivity was highest where temperature seasonality = 0, and at an
+annual temperature range of 15\(^\circ\)C or lower. In contrast, there
+was no significant effect of precipitation seasonality on FACF.
 
 We found a significant relationship between length of growing season and
-productivity, with all fluxes showing a linear increase in productivity
-with length of growing season (fig. 5). Length of growing season was a
-strong predictor of productivity, explaining 51\% of variation in GPP,
-39\% of variation in NPP, and 34\% of variation in ANPP, but it was not
-identified as a stronger predictor than MAT for any of the fluxes
-analysed.
-
-\emph{Within the growing season, how does productivity vary with
-climate?} Within growing season months, we found that climate had a much
-weaker effect on productivity. For each of temperature, precipitation,
-PET, and solar radiation, we found a small effect of climate for certain
-carbon fluxes. There was a small increase in productivity with
-temperature and precipitation for ANPP (with temperature \(R^2\) = 0.10,
-p\textless{}0.001; with precipitation \(R^2\) = 0.04, p\textless{}0.05)
-and ANPP\textsubscript{foliage} (with temperature \(R^2\) = 0.16,
-p\textless{}0.01; with precipitation \(R^2\) = 0.09, p\textless{}0.05).
-Productivity increased with solar radiation for GPP (\(R^2\) = 0.21,
-p\textless{}0.001), NPP (\(R^2\) = 0.21, p\textless{}0.001), BNPP
-(\(R^2\) = 0.16, p\textless{}0.001) and BNPP\textsubscript{fine}
-\textsubscript{root} (\(R^2\) = 0.12, p\textless{}0.01), and with PET
-for GPP (\(R^2\) = 0.15, p\textless{}0.01), NPP (\(R^2\) = 0.18,
-p\textless{}0.01), BNPP (\(R^2\) = 0.23, p\textless{}0.0001),
-BNPP\textsubscript{fine} \textsubscript{root} (\(R^2\) = 0.11,
-p\textless{}0.05), and ANPP\textsubscript{woody} \textsubscript{stem}
-(\(R^2\) = 0.06, p\textless{}0.05).
+FACF, with all fluxes showing a linear increase in productivity with
+length of growing season (Fig. 4). Length of growing season was a strong
+predictor of FACF, explaining 51\% of variation in GPP, 39\% of
+variation in NPP, and 34\% of variation in ANPP, but it was a weaker
+predictor than MAT for all fluxes analysed.
+
+\emph{Within the growing season, how does FACF vary with climate?} When
+FACF was standardized by growing season length, correlations with
+climate variables were much weaker. Growing season temperature,
+precipitation, solar radiation, and PET each had a small effect on one
+or more of the nine FACF analyzed (Table 2; Fig. \textbf{S\#}).
+Speficifally, \(ANPP\) increased with growing season temperature
+(\(R^2\) = 0.10, p\textless{}0.001) and precipitation (\(R^2\) = 0.04,
+p\textless{}0.05). Similaryly, \(ANPP_{foliage}\) increased slightly
+with growing season temperature (\(R^2\) = 0.16, p\textless{}0.01) and
+precipitation (\(R^2\) = 0.09, p\textless{}0.05). Growing season solar
+radiation had a positive influence on \(GPP\) (\(R^2\) = 0.21,
+p\textless{}0.001), \(NPP\) (\(R^2\) = 0.21, p\textless{}0.001),
+\(BNPP\) (\(R^2\) = 0.16, p\textless{}0.001) and \(BNPP_{fine.root}\)
+(\(R^2\) = 0.12, p\textless{}0.01). Growing season PET had a positive
+influence on \(GPP\) (\(R^2\) = 0.15, p\textless{}0.01), \(NPP\)
+(\(R^2\) = 0.18, p\textless{}0.01), \(BNPP\) (\(R^2\) = 0.23,
+p\textless{}0.0001), \(BNPP_{fine.root}\) (\(R^2\) = 0.11,
+p\textless{}0.05), and \(ANPP_{woody-stem}\) (\(R^2\) = 0.06,
+p\textless{}0.05). \{\textbf{Becky, please verify/ edit the following:}
+There were no other signifcant correlations between growing season
+length-standardized FACF (9 variables in Table 2) and growing season
+climate (\textbf{which variables?})\}.
 
@@
 We show that, across all nine FACF variables analyzed, C cycling
-decreases linearly with latitude (\emph{H1.1}; Figs. 2-3)--a finding
-that confirms multiple previous studies (\textbf{REFS}) but contradicts
-the idea that productivity of temperate forests rivals that of tropical
+decreases linearly with latitude (\emph{H1.1}; Fig. 2)--a finding that
+confirms multiple previous studies (\textbf{REFS}) but contradicts the
+idea that productivity of temperate forests rivals that of tropical
 forests (\emph{H1.1.alt}; \textbf{REFS}). The FACF variables increase in
@@
 allocation detectable at this global scale and with component fluxes
-summing appropriately to larger fluxes (Fig. 3), indicating no major,
+summing appropriately to larger fluxes (Fig. 2), indicating no major,
 systematic omissions or overestimations of flux components. However, we
@@
 fluxes analyzed} (albiet less for subsidiary fluxes), with temperature
-variables the best predictors of FACF at this global scale (Figs. 4-5).
+variables the best predictors of FACF at this global scale (Figs. 3-4).
 While other climate variables are significant predictors of FACF
-(\emph{H3}), none of them improve on the explanatory power of
-temperature-related variables in general or MAT specifically (Fig. 5).
-Water availability is an important factor in explaining FACF on a global
-scale: we find a positive influence of precipitation at low MAP, with
-saturation at higher levels of MAP (Fig. 5b) and a significant
+(\emph{H3}), none of them in isolation improve on the explanatory power
+of temperature-related variables in general or MAT specifically (Fig.
+4). Water availability is an important factor in explaining FACF on a
+global scale: we find a positive influence of precipitation at low MAP,
+with saturation at higher levels of MAP (Fig. 4b) and a significant
 interaction between MAT and MAP for seven of the nine variables analyzed
-(Fig. 4). Temperature seasonality and growing season length are closely
+(Fig. 3). Temperature seasonality and growing season length are closely
 correlated with MAT and are strong predictors of FACF (\emph{H4}; Fig.
-5e-f), though growing season length doesn't improve upon MAT as a
+4e-f), though growing season length doesn't improve upon MAT as a
 predictor. Within the growing season, the influence of climate on C
@@
 unambiguously highest in the tropical regions, and the relationship is
-approximately linear (Figs. 2-3). This contrasts with suggestions that
+approximately linear (Fig. 2). This contrasts with suggestions that
 productivity of temperate forests is similar to that of tropical forests
@@
 drivers on the global scale (with the exception of some differences in
-MAT-MAP interactions; Fig. 4), with no major trends in C allocation
+MAT-MAP interactions; Fig. 3), with no major trends in C allocation
 among the variable pairs tested (Table 1, \emph{H1}; Fig. 2;
@@
 both indirectly through its influence on \(GPP\) and respiration and
-directly through any climatic influence on C allocation, as well as by
-many other local- and regional-scale factors (\textbf{REFS}).
-
-The latitudinal gradient in FACF (Figs. 2-3) is driven primarily by
+directly through any climatic influence on C allocation, as well as many
+other local- and regional-scale factors (\textbf{REFS}).
+
+The latitudinal gradient in FACF (Fig. 2) is driven primarily by
 temperature-related climate variables, and secondarily by moisture
-availability (Table 1, \emph{H2-H3}; Figs. 4-5). Because many climate
+availability (Table 1, \emph{H2-H3}; Figs. 3-4). Because many climate
 variables co-vary across the latitudinal gradient, because climatic
@@
 under hot and dry conditions (\emph{i.e.}, high PET, high deficit; Fig.
-5c-d) and sometimes under very high precipitaiton (Figs. 4, 5b).
+4c-d) and sometimes under very high precipitaiton (Figs. 3, 4b).
 Negative effects of very high precipitation on FACF have been previously
@@
 
-FACF is reduced by temperature seasonality (Table 1, \emph{H4}; Fig.
-5e), and productivity is shut down during cold- or dry- dormant seasons.
-To account for this, a number of analyses seeking to characterize
-global-scale effects of climate on productivity have examined the
-relationship of FACF per month of the growing season with growing season
-climatic conditions (Table 1, \emph{H5}; \textbf{REFS}). We found that
-the sort of simple metric needed to define growing season at a global
-scale was uncertain for temperature and problematic for moisture
-(\textbf{WORK ON THIS}). A temperature-defined growing season length had
-stong positive correlation with FACF (Fig. 5f), but explained less
-variation than \(MAT\). Dividing FACFs by growing season length to yield
-FACF per growing season month removed the majority of climate-related
-variation, supporting the idea that the latitudinal gradient in FACF is
-attributable more to shorter growing seasons at high latitudes than to
-inherently lower rates of photosyntheiss or respiration by high-latitude
-forests (\emph{{[} Enquist et al. 2007 GCB- but check{]}}). However,
-there remained a number of significant correlations with growing season
-climatic conditions, suggesting that climatic conditions remain
-influential within the growing season. We conclude that while correcting
-for growing season length takes analyses a step closer to mechanistic
-linkage of instantaneous C flux rates to environmental conditions, it
-remains very crude relative to the the timescales on which climate
-affects plant metabolism and does not advance statistical predictive
-power. Rather, mechanistic accounting for climatic effects on global
-productivity patterns requires models representing physiologically
-meaningful timescales (\emph{e.g.}, \textbf{refs}).
+FACF is negatively correlated with temperature seasonality (Table 1,
+\emph{H4}; Fig. 4e), and productivity is shut down during cold- or dry-
+dormant seasons. To account for this, a number of analyses seeking to
+characterize global-scale effects of climate on productivity have
+examined the relationship of FACF per month of the growing season with
+growing season climatic conditions (Table 1, \emph{H5}; \textbf{REFS}).
+We found that the sort of simple metric needed to define growing season
+at a global scale was uncertain for temperature and problematic for
+moisture (\textbf{WORK ON THIS}). A temperature-defined growing season
+length had stong positive correlation with FACF (Fig. 4f), but explained
+less variation than \(MAT\). Dividing FACFs by growing season length to
+yield FACF per growing season month removed the majority of
+climate-related variation, supporting the idea that the latitudinal
+gradient in FACF is attributable more to shorter growing seasons at high
+latitudes than to inherently lower rates of photosyntheiss or
+respiration by high-latitude forests (\emph{{[} Enquist et al. 2007 GCB-
+but check{]}}). However, there remained a number of significant
+correlations with growing season climatic conditions, suggesting that
+climatic conditions remain influential within the growing season. We
+conclude that while correcting for growing season length takes analyses
+a step closer to mechanistic linkage of instantaneous C flux rates to
+environmental conditions, it remains very crude relative to the the
+timescales on which climate affects plant metabolism and does not
+advance statistical predictive power. Rather, mechanistic accounting for
+climatic effects on global productivity patterns requires models
+representing physiologically meaningful timescales (\emph{e.g.},
+\textbf{refs}).
 
@@
 respect to latitude or \(MAT\), and no distinct trends in C allocation
-across the global scale (Fig. 3). The implication is that under warmer
+across the global scale (Fig. 2). The implication is that under warmer
 conditions with similar moisture availability--and within the

### 2020-01-20 manuscript/manuscript draft.Rmd
Table 2
#54

--- 2020-01-20 manuscript/manuscript draft.Rmd
@@
 | $NPP$ | Net Primary Production | stem, foliage, coarse root, fine root, optionally others (e.g., branch, reproductive, understory) | $ANPP$ + $BNPP$ (majority); $GPP$-$R_{auto}$ | 92 | 42 |
-| $ANPP$ | Aboveground Net Primary Production | stem, foliage, optionally others (e.g., branch, reproductive, understory) (**issues [52](https://github.com/forc-db/Global_Productivity/issues/52), [54](https://github.com/forc-db/Global_Productivity/issues/54)**) |$ANPP_{woody-stem}$ + $ANPP_{foliage}$ (+ others) | 256 | 80 |
-| $ANPP_{woody-stem}$ | Net Primary Production of woody stems | woody stems down to DBH $\le$ 10cm (no branch turnover) | direct measurement of stem growth increment (**with recruitment?**) | 264 | 96 |
-| $ANPP_{foliage}$ | Net Primary Production of foliage | foliage | direct measurement of litterfall | 98 | 49 |
-| $BNPP_{root}$  | Belowground Net Primary Production | coarse and fine roots |  coarse roots estimated indirectly using allometries based on aboveground stem increment measures ; fine roots as below | 101 | 48 |
-| $BNPP_{fine.root}$ | Net Primary Production of fine roots | fine roots | measurements combined one or more of the following: minirhizotrons, fine root turnover, soil cores, root ingrowth cores | 88 | 41 |
-| $R_{root}$ | Root Respiration | **coarse** and fine roots | *measurement of root gas exchange, root exclusion from soil respiration chambers* | 64 | 26 |
+| $ANPP$ | Aboveground Net Primary Production | stem, foliage, optionally others (e.g., branch, reproductive, understory) |$ANPP_{woody-stem}$ + $ANPP_{foliage}$ (+ others) | 256 | 80 |
+| $ANPP_{woody-stem}$ | Woody stem growth component of $ANPP$ | woody stems down to DBH $\le$ 10cm (no branch turnover) | stem growth measurements scaled to biomass using allometries | 264 | 96 |
+| $ANPP_{foliage}$ | Foliage component of $ANPP$ | foliage | litterfall collection (separated into components) | 98 | 49 |
+| $BNPP$  | Belowground Net Primary Production | coarse and fine roots |  coarse roots estimated indirectly using allometries based on aboveground stem increment measures ; fine roots as below | 101 | 48 |
+| $BNPP_{fine.root}$ | Fine root component of $BNPP$ | fine roots | measurements combined one or more of the following: minirhizotrons, fine root turnover, soil cores, root ingrowth cores | 88 | 41 |
+| $R_{root}$ | Root Respiration | coarse and fine roots | partitioning of total soil respiration (e.g., through root exclusion), scaling of root gas exchange | 64 | 26 |
 | $R_{auto}$ | Autotrophic Respiration | foliage, stem, and root | chamber measurements of component gas exchange | 22 | 13 |
@@
 It would be preferable to standardise by minimum area sampled; however x% of plots in the database are 1 ha or under in size; excluding these plots would place significant constraints on sample size.
-(**really, we didn't exclude any studies based on methodology (all studies measured min DBH 10 or less...
-) so this paragraph needs revision.
-Also, address whether we standardize for inclusion/ exclusion of recruitment, and address branchfall.**)
+(**really, we didn't exclude any studies based on methodology (except soil respiration.
+all studies measured min DBH 10 or less...
+), so this paragraph needs revision.
+Also, address whether we standardize for inclusion/ exclusion of recruitment.**)
+- recruitment preferred when an option, did not review whether all included recruitment
+- excluded alkali absoption and soda lime methods for estimating soil / root respiration
 

### 2020-01-27 manuscript/manuscript draft.Rmd
issue #53
checked entire manuscript, replacing productivity with FACF as needed.

@@
 2009](https://www.annualreviews.org/doi/abs/10.1146/annurev.ecolsys.30.1.257)**REFS**).
-However, our ability to draw general macroscopic conclusions regarding global variation in multiple productivity variables with respect to climate has been limited in that these analyses often mix forests that vary in stand age, disturbance history, and/or management status; do not always sufficiently parse related variables (*e.g.*, combining net primary productivity records with and without belowground components); and typically consider only one or a few variables at a time.
+However, our ability to draw general macroscopic conclusions regarding global variation in multiple FACF variables with respect to climate has been limited in that these analyses often mix forests that vary in stand age, disturbance history, and/or management status; do not always sufficiently parse related variables (*e.g.*, combining net primary productivity records with and without belowground components); and typically consider only one or a few variables at a time.
 
-**The productivity of forests varies with latitude**, showing a general trend of decreasing productivity with latitude [@beer_terrestrial_2010; @jung_global_2011].
-Studies agree that productivity is lowest in the boreal regions, and increases into the temperate regions [@luyssaert_co_2007; @huston_global_2009; @beer_terrestrial_2010; @jung_global_2011].
-However, evidence is inconclusive on whether productivity continues to increase into the tropics, or whether it plateaus in temperate regions.
+**FACF varies with latitude**, showing a general trend of decreasing FACF with latitude [@beer_terrestrial_2010; @jung_global_2011].
+Studies agree that FACF is lowest in the boreal regions, and increases into the temperate regions [@luyssaert_co_2007; @huston_global_2009; @beer_terrestrial_2010; @jung_global_2011].
+However, evidence is inconclusive on whether primary productivity continues to increase into the tropics, or whether it plateaus in temperate regions.
 Evidence for this is further complicated by the fact that different studies use different measures of productivity to explore these relationships.
@@
 
-**Climate is a significant driver of productivity across broad spatial scales** [@cleveland_relationships_2011].
-The majority of studies have focused on exploring the relationships of productivity to mean annual temperature (MAT) and mean annual precipitation (MAP), as the most commonly reported site-level climate variables.
+**Climate is a significant driver of FACF across broad spatial scales** [@cleveland_relationships_2011].
+The majority of studies have focused on exploring the relationships of FACF to mean annual temperature (MAT) and mean annual precipitation (MAP), as the most commonly reported site-level climate variables.
 These variables have the advantage that they describe broad trends in temperature and water availability, and therefore capture a lot of global-scale variation in climate.
-There is strong evidence that both MAT and MAP show significant positive relationships with productivity [@chu_does_2016].
-However, as with latitude, the shape of those relationships is not always clear, and, again, is complicated by the use of different measures of productivity across studies.
-Various measures of primary productivity saturate at high levels of MAP, though the saturation points identified vary from 1500mm [@luyssaert_co_2007] up to 2445mm MAP [@schuur_productivity_2003].
-Studies of the influence of MAT on productivity are less conclusive.
+There is strong evidence that both MAT and MAP show significant positive relationships with FACF [@chu_does_2016].
+However, as with latitude, the shape of those relationships is not always clear, and, again, is complicated by the use of different measures of FACF across studies.
+Various measures of primary productivity {**FACF?**} saturate at high levels of MAP, though the saturation points identified vary from 1500mm [@luyssaert_co_2007] up to 2445mm MAP [@schuur_productivity_2003].
+Studies of the influence of MAT on productivity {**FACF?**} are less conclusive.
 Luyssaert et al.
@@
 [-@taylor_temperature_2017] showed that increased MAP had a negative effect at low MAT, but a positive effect at high levels of MAT, and *vice versa* {**Helene found this confusing**}, suggesting that both temperature and water availability are important in explaining productivity.
+{**It would be good to add some more citations on soil respiraiton.
+I'm sure BBL can help.**}
 
-**Primary productivity can be influenced by many other factors as well**, which often act across a range of scales, and may show interactive effects with each other [@cleveland_relationships_2011].
+**FACF can be influenced by many other factors as well**, which often act across a range of scales, and may show interactive effects with each other [@cleveland_relationships_2011].
 On a local scale, stand age [@litton_carbon_2007; @gillman_latitude_2015], biodiversity (Liang et al.
@@
 Positive biodiversity-productivity relationship predominant in global forests.
-Science 354, Issue 6309, aaf8957), management [@simova_enigma_2017]; nutrient availability [@aragao_above-_2009]; and altitude [@girardin_net_2010; @malhi_variation_2017] all impact forest productivity.
-On a global scale, we expect that productivity is most strongly influenced by broad climatic gradients.
-There is evidence that productivity also responds to variables such as cloud cover [@taylor_temperature_2017], solar radiation [@fyllas_solar_2017], and potential evapotranspiration [@kerkhoff_plant_2005] in potentially significant ways.
+Science 354, Issue 6309, aaf8957), management [@simova_enigma_2017]; nutrient availability [@aragao_above-_2009]; and altitude [@girardin_net_2010; @malhi_variation_2017] all impact FACF.
+On a global scale, we expect that FACF is most strongly influenced by broad climatic gradients.
+There is evidence that FACF also responds to variables such as cloud cover [@taylor_temperature_2017], solar radiation [@fyllas_solar_2017], and potential evapotranspiration [@kerkhoff_plant_2005] in potentially significant ways.
 Furthermore, MAT and MAP are very coarse measures of climate, and so fail to capture much variation in climate on an intra-annual scale, including the effects of factors such as growing season length, number of frost-free days, temperature seasonality, and dry season length.
-Some studies have suggested that the apparently strong relationship between MAT and productivity is actually a consequence of the correlation between MAT and growing season length [@kerkhoff_plant_2005; @malhi_productivity_2012; @michaletz_convergence_2014; @michaletz_drivers_2018].
+Some studies have suggested that the apparently strong relationship between MAT and FACF is actually a consequence of the correlation between MAT and growing season length [@kerkhoff_plant_2005; @malhi_productivity_2012; @michaletz_convergence_2014; @michaletz_drivers_2018].
 Kerkhoff et al.
 [-@kerkhoff_plant_2005] and Michaletz et al.
-[-@michaletz_convergence_2014] find that, within the growing season, there is no significant relationship between productivity and MAT, suggesting that the effect of temperature is due to increased length of growing season, rather than an inherent influence of temperature on productivity.
+[-@michaletz_convergence_2014] find that, within the growing season, there is no significant relationship between primary productivity and MAT, suggesting that the effect of temperature is due to increased length of growing season, rather than an inherent influence of temperature on FACF.
 
-**The recent development of a global forest carbon database synthesizing multiple variables and including records of stand history (ForC; [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc:_2018]) opens up the possibility for a standardized analysis of global scale variation** in multiple components of forest productivity and the principle climatic drivers of these patterns.
+**The recent development of a global forest carbon database synthesizing multiple variables and including records of stand history (ForC; [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc:_2018]) opens up the possibility for a standardized analysis of global scale variation** in multiple components of FACF and the principle climatic drivers of these patterns.
 In order to approach these broad and complex issues, we simplify the major gaps in our knowledge to five broad hypotheses and corresponding specific predictions (Table 1).
@@
 Finally, we consider the relationship between FACF and seasonality, considering the role of seasonality in explaining variation in carbon fluxes, and the influence of climate on FACF standardized by growing season length.
-We use a comprehensive global database of forest carbon fluxes to address the above questions for nine carbon fluxes, allowing for an in-depth exploration of the effect of climate on global productivity.
+We use a comprehensive global database of forest carbon fluxes to address the above questions for nine carbon fluxes, allowing for an in-depth exploration of the effect of climate on FACF globally.
 
@@
 The database was significantly expanded since the publication of [@anderson-teixeira_forc:_2018] through integration with the Global Soil Respiration Database [@bond-lamberty_global_2010].
-Additional targeted literature searches were conducted to identify any further available data on primary productivity, with particular focus on mature forests in temperate and boreal regions.
+Additional targeted literature searches were conducted to identify any further available data on FACF, with particular focus on mature forests in temperate and boreal regions.
 ForC currently contains 29730 records from 4979 plots, representing 20 distinct ecozones across all forested biogeographic and climate zones.
@@
 A subset of the ForC database was generated for the purposes of this analysis, in order to control for data quality and remove biasing factors.
-Since management can alter observed patterns of primary productivity [@simova_enigma_2017], sites were excluded from analysis if they were managed, defined as plots that were planted, managed as plantations, irrigated, fertilised or including the term "managed" in their site description.
+Since management can alter observed patterns of FACF [@simova_enigma_2017], sites were excluded from analysis if they were managed, defined as plots that were planted, managed as plantations, irrigated, fertilised or including the term "managed" in their site description.
 Sites that had experienced significant disturbance were also excluded.
@@
 Grazed sites were retained.
-There is evidence that stand age influences patterns of primary productivity and carbon allocation in forest ecosystems, and can confound relationships between latitude and primary productivity [@de_lucia_forest_2007; @gillman_latitude_2015].
+There is evidence that stand age influences patterns of FACF and carbon allocation in forest ecosystems, and can confound relationships between latitude and primary productivity [@de_lucia_forest_2007; @gillman_latitude_2015].
 To reduce any biasing effects of stand age, we included only stands of known age $\ge$ 100 years and those described by terms such as "mature", "intact", or "old-growth".
@@
 | $BNPP$  | Belowground Net Primary Production | coarse and fine roots |  coarse roots estimated indirectly using allometries based on aboveground stem increment measures ; fine roots as below | 101 | 48 |
-| $BNPP_{fine.root}$ | Fine root component of $BNPP$ | fine roots | measurements combined one or more of the following: minirhizotrons, fine root turnover, soil cores, root ingrowth cores | 88 | 41 |
+| $BNPP_{fine.root}$ | Fine root component of $BNPP$ | fine roots | measurements combined one or more of the following: soil cores, minirhizotrons, turnover estimates,  root ingrowth cores | 88 | 41 |
 | $R_{root}$ | Root Respiration | coarse and fine roots | partitioning of total soil respiration (e.g., through root exclusion), scaling of root gas exchange | 64 | 26 |
@@
 
-*Methodological consistency.* The data in ForC is derived from a range of studies, often employing different methods.
+*Methodological consistency.* 
+{**This whole section needs to be re-written.**} 
+The data in ForC is derived from a range of studies, often employing different methods.
 For this reason, criteria were introduced to standardise for differences in methodology.
@@
 As discussed above, estimates of $NPP$, $ANPP$, and $BNPP$ are generated through summing measurements of their component parts.
-Since the components included in productivity estimates vary between studies, estimates of productivity were classified within the ForC database according to their components, and then filtered for analysis.
+Since studies vary in the components included in productivity estimates, these variables were classified within the ForC database according to their components, and then filtered for analysis.
 Estimates of NPP were selected if they included foliage, branch, stem, coarse root, and fine root.
@@
 
-The effects of climate and latitude on primary productivity were analysed using mixed effects models using the package 'lme4' [@bates_fitting_2015] in R v.3.5.1 [@r_core_team_r:_2018].
-The effect of each extracted climate variable on each measure of primary productivity was modelled by specifying the climate variable as a fixed effect.
+The effects of climate and latitude on FACF were analysed using mixed effects models using the package 'lme4' [@bates_fitting_2015] in R v.3.5.1 [@r_core_team_r:_2018].
+The effect of each extracted climate variable on each FACF variable was modelled by specifying the climate variable as a fixed effect.
 For each climate variable, three models were specified: a null model; a model with the climate variable as a linear term; and a model with the climate variable as a **XX**-order polynomial term.
@@
 
-*How does productivity vary with latitude?*
+*How does FACF vary with latitude?*
 
@@
 
-*How does productivity relate to MAT and MAP?* MAT and MAP are the most commonly reported site-level climate variables, and much previous research into the effect of climate on forest productivity has focused on these as key climate variables.
-MAT was a significant (p<0.05) and strong predictor of productivity for all carbon fluxes tested, with all fluxes showing a linear increase with temperature (fig.
-4).
-We found no support for a saturation point of productivity with temperature.
+*How does FACF relate to MAT and MAP?* MAT and MAP are the most commonly reported site-level climate variables, and much previous research into the effect of climate on FACF has focused on these as key climate variables.
+MAT was a significant (p<0.05) and strong predictor of FACF for all fluxes tested.
+All FACF increased linearly with temperature (Fig.
+4), and we found no support for a saturation point of FACF with temperature.
 
-MAP was found to be a significant (p<0.05) but poor predictor of productivity, explaining, with the exception of R~auto~, at most 37% of variation in carbon flux.
-For the majority of fluxes productivity was best predicted by a polynomial model.
-Productivity increased with precipitation, up until a saturation point at between 3000 and 4000mm annual precipitation, above which productivity started to decrease (fig.
+MAP was found to be a significant (p<0.05) but poor predictor of FACF, explaining, with the exception of R~auto~, at most 37% of variation in carbon flux.
+For the majority FACF, a polynomial model was the best fit.
+FACF increased with precipitation, up until a saturation point at between 3000 and 4000mm annual precipitation, above which FACF started to decrease (Fig.
 4).
@@
 
-*How does productivity relate to other climate variables?* {**Here, maybe add some statement about covariation among climate variables, site SI figure**} Our results indicated that productivity was most strongly explained by temperature at the global scale, with temperature-related climate variables coming out as strong predictors of productivity.
+*How does FACF relate to other climate variables?* {**Here, maybe add some statement about covariation among climate variables, site SI figure**} Our results indicated that FACF was most strongly explained by temperature at the global scale, with temperature-related climate variables coming out as strong predictors of FACF.
 In addition to MAT, temperature seasonality, annual temperature range, and annual frost days were consistently identified as strong univariate predictors of FACF.
@@
 
-We found a significant relationship between productivity and potential evapotranspiration for all fluxes.
+We found a significant relationship between FACF and potential evapotranspiration for all fluxes.
 ANPP~foliage~, BNPP~fine~ ~root~ and R~root~ increased linearly with PET; however, all other fluxes showed a polynomial relationship with PET (fig.
 4).
-We found strong evidence for a saturation point or peak with PET: productivity tended to increase at values below 1000mm, before saturating between 1200 and 1700mm.
-There was also evidence that productivity begins to decrease at values above 1800mm PET.
+We found strong evidence for a saturation point or peak with PET: FACF tended to increase at values below 1000mm, before saturating between 1200 and 1700mm.
+There was also evidence that FACF begins to decrease at values above 1800mm PET.
 
-Vapour pressure deficit was a significant predictor of productivity for all fluxes.
+Vapour pressure deficit was a significant predictor of FACF for all fluxes.
 BNPP~fine~ ~root~ showed a linear relationship with vapour pressure deficit ($R^2$ = 0.07, p<0.05), but all other fluxes showed a polynomial relationship (fig.
 4).
-Productivity initially increased with vapour pressure deficit, before saturating at around 0.8 kPa.
-At values above 0.8 kPa, productivity began to decrease.
+FACF initially increased with vapour pressure deficit, before saturating at around 0.8 kPa.
+At values above 0.8 kPa, FACF began to decrease.
 
 All fluxes, with the exception of R~root~, showed a positive linear relationship with solar radiation.
-Solar radiation explained a low proportion of variability in productivity for all fluxes, explaining less than 20% of the variation in each flux, with the exception of R~auto~ ($R^2$ = 0.26, p<0.05).
+Solar radiation explained a low proportion of variability in FACF for all variables, explaining less than 20% of the variation in each flux, with the exception of R~auto~ ($R^2$ = 0.26, p<0.05).
 
-Of the climate variables tested, annual wet days, aridity, cloud cover, mean diurnal temperature range, precipitation seasonality, maximum vapour pressure deficit and water stress months were poor or non-significant explainers of variation in productivity, explaining less than 20% of the variation in each of the carbon fluxes.
+Of the climate variables tested, annual wet days, aridity, cloud cover, mean diurnal temperature range, precipitation seasonality, maximum vapour pressure deficit and water stress months were poor or non-significant explainers of variation in FACF, explaining less than 20% of the variation in each of the carbon fluxes.
 
@@
 Temperature seasonality was strongly correlated with annual temperature range, and, as expected, all fluxes showed almost identical responses to it {**meaning unclear**}.
-Productivity was highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower.
+FACF was highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower.
 In contrast, there was no significant effect of precipitation seasonality on FACF.
 
-We found a significant relationship between length of growing season and FACF, with all fluxes showing a linear increase in productivity with length of growing season (Fig.
+We found a significant relationship between length of growing season and FACF, with all fluxes showing a linear increase with length of growing season (Fig.
 4).
@@
 
-Past studies have differed in their conclusions regarding the relationship between productivity and latitude or its correlates (Table 1, *H1*; **REFS**)--quite possibly because of lack of standardization with respect to stand age and disturbance history.
+Past studies have differed in their conclusions regarding the relationship between FACF and latitude or its correlates (Table 1, *H1*; **REFS**)--quite possibly because of lack of standardization with respect to stand age and disturbance history.
 Our findings indicate that, among mature, undisturbed stands, FACF is unambiguously highest in the tropical regions, and the relationship is approximately linear (Fig.
@@
 However, it remains informative to consider these relationships.
-We find that temperature-related climate variables ($MAT$, temperature seasonality, ...**LIST**) explain the highest proportion of variability in productivity, and among these, $MAT$ is generally the best predictor--perhaps because site-specific MAT is recorded for the majority of sites in ForC, whereas other variables are extracted from global gridded data products (Table S1**?**).
+We find that temperature-related climate variables ($MAT$, temperature seasonality, ...**LIST**) explain the highest proportion of variability in FACF, and among these, $MAT$ is generally the best predictor--perhaps because site-specific MAT is recorded for the majority of sites in ForC, whereas other variables are extracted from global gridded data products (Table S1**?**).
 The effects of temperature are modified by moisture availability, with reduced FACF under hot and dry conditions (*i.e.*, high PET, high deficit; Fig.
@@
 FACF is negatively correlated with temperature seasonality (Table 1, *H4*; Fig.
-4e), and productivity is shut down during cold- or dry- dormant seasons.
+4e), and is minimal during cold- or dry- dormant seasons.
 To account for this, a number of analyses seeking to characterize global-scale effects of climate on productivity have examined the relationship of FACF per month of the growing season with growing season climatic conditions (Table 1, *H5*; **REFS**).
@@
 We conclude that while correcting for growing season length takes analyses a step closer to mechanistic linkage of instantaneous C flux rates to environmental conditions, it remains very crude relative to the the timescales on which climate affects plant metabolism and does not advance statistical predictive power.
-Rather, mechanistic accounting for climatic effects on global productivity patterns requires models representing physiologically meaningful timescales (*e.g.*, **refs**).
+Rather, mechanistic accounting for climatic effects on global FACF patterns requires models representing physiologically meaningful timescales (*e.g.*, **refs**).
 

### 2020-01-31 scripts/stacked_plots_code.R
Update stacked_plots_code.R
updated to include all definitions of ANPP and NPP

@@ -2,7 +2,7 @@
 rm(list = ls())
 
 # Set working directory as ForC main folder ####
-setwd("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/ForC")
+setwd("C:/Users/gyrcbm/Dropbox/ForC")
 
 # Load libaries ####
 library(lme4)
@@ -107,7 +107,7 @@ set2 <- c("R_auto")
 sum <- c("GPP")
   
 
-png(file = paste0("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/figures/final_figures/stacked_plots/combined_stacked.png"), width = 2255, height = 2000, units = "px", res = 300)
+png(file = paste0("C:/Users/gyrcbm/Dropbox/Global_Productivity/results/figures/final_figures/stacked_plots/combined_stacked.png"), width = 2255, height = 2000, units = "px", res = 300)
 
 par(mfrow = c(2,2), mar = c(2,2,2,2), oma = c(3,3,0,0))
 
@@ -169,8 +169,8 @@ par(mfrow = c(2,2), mar = c(2,2,2,2), oma = c(3,3,0,0))
           if (i == j){
           
           ################################ for response.v.1
-          if(set1[[i]] %in% "NPP") responses.to.keep.1  <- c("NPP_1")
-          if(set1[[i]] %in% "ANPP") responses.to.keep.1  <- c("ANPP_1")
+          if(set1[[i]] %in% "NPP") responses.to.keep.1  <- c("NPP_1", "NPP_2", "NPP_3", "NPP_4", "NPP_5", "NPP_0")
+          if(set1[[i]] %in% "ANPP") responses.to.keep.1  <- c("ANPP_1", "ANPP_2", "ANPP_0")
           if(set1[[i]] %in% "ANPP_litterfall") responses.to.keep.1  <- c("ANPP_litterfall_1")
           if(!set1[[i]] %in% c("NPP", "ANPP", "ANPP_litterfall")) responses.to.keep.1  <- set1[[i]]
           
@@ -209,8 +209,8 @@ par(mfrow = c(2,2), mar = c(2,2,2,2), oma = c(3,3,0,0))
           
           ########################################### for response.v.2
           
-          if(set2[[j]] %in% "NPP") responses.to.keep.2  <- c("NPP_1")
-          if(set2[[j]] %in% "ANPP") responses.to.keep.2  <- c("ANPP_1")
[…]

### 2020-01-31 scripts/mat_map_interaction.R
Update mat_map_interaction.R
updated to include all definitions of NPP and ANPP

@@ -4,7 +4,7 @@
 rm(list = ls())
 
 # Set working directory as ForC main folder ####
-setwd("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/ForC")
+setwd("C:/Users/gyrcbm/Dropbox/ForC")
 
 # Load libaries ####
 library(lme4)
@@ -111,13 +111,13 @@ pannel.nb <- 1
 
 response.variables <- c("GPP", "NPP", "BNPP_root", "BNPP_root_fine", "ANPP", "ANPP_foliage", "ANPP_woody_stem","R_auto", "R_auto_root")
 
-png(file = paste0("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/figures/final_figures/interactions/mat_map_interaction.png"), width = 2255, height = 2000, units = "px", res = 300)
+png(file = paste0("C:/Users/gyrcbm/Dropbox/Global_Productivity/results/figures/final_figures/interactions/mat_map_interaction.png"), width = 2255, height = 2000, units = "px", res = 300)
 par(mfrow = c(3,3), mar = c(1,0,0,2), oma = c(5,8,2,0), xpd = T)
 
 for (response.v in response.variables){
   
-  if(response.v %in% "NPP") responses.to.keep  <- c("NPP_1")
-  if(response.v %in% "ANPP") responses.to.keep  <- c("ANPP_1", "ANPP_2")
+  if(response.v %in% "NPP") responses.to.keep  <- c("NPP_1", "NPP_2", "NPP_3", "NPP_4", "NPP_5", "NPP_0")
+  if(response.v %in% "ANPP") responses.to.keep  <- c("ANPP_1", "ANPP_2", "ANPP_0")
   if(response.v %in% "ANPP_litterfall") responses.to.keep  <- c("ANPP_litterfall_1")
   if(!response.v %in% c("NPP", "ANPP", "ANPP_litterfall")) responses.to.keep  <- response.v
   

### 2020-01-31 scripts/Maps_script.R
Update Maps_script.R

@@ -4,7 +4,7 @@
 rm(list = ls())
 
 # Set working directory as ForC main folder ####
-setwd("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/ForC")
+setwd("C:/Users/gyrcbm/Dropbox/ForC")
 
 # Load libaries ####
 library(lme4)
@@ -95,7 +95,7 @@ all.response.variables <- all.response.variables[all.response.variables %in% For
 all.response.variables <- unique(gsub("_\\d", "", all.response.variables))
 
 response.variables.groups <- list(c("GPP", "NPP", "BNPP_root", "BNPP_root_fine"),
-                                  c("ANPP_1", "ANPP_foliage", "ANPP_repro"),
+                                  c("ANPP", "ANPP_foliage", "ANPP_repro"),
                                   c("ANPP_woody", "ANPP_woody_stem", "ANPP_woody_branch"))
 
 # response.variables.groups <- list(c("ANPP_1"))
@@ -109,7 +109,7 @@ fixed.variables <- c("AnnualMeanTemp")
 
 all.results <- NULL
 
-ForC_simplified$variable.name <- gsub("(_1|_2)", "", ForC_simplified$variable.name)
+ForC_simplified$variable.name <- gsub("(_0|_1|_2|_3|_4|_5)", "", ForC_simplified$variable.name)
 
 ForC_simplified <- ForC_simplified[ForC_simplified$variable.name %in% c("GPP", "NPP", "ANPP", "ANPP_woody_stem", "ANPP_foliage", "BNPP_root", "BNPP_root_fine", "R_auto", "R_auto_root"),]
 
@@ -122,7 +122,7 @@ ForC_simplified <- ForC_simplified[ages.to.keep & fixed.no.na, ]
 lon <- ForC_simplified$lon
 lat <- ForC_simplified$lat
 
-col.sym <- read.csv("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/raw.data/colsym.csv", stringsAsFactors = F)
+col.sym <- read.csv("C:/Users/gyrcbm/Dropbox/Global_Productivity/raw.data/colsym.csv", stringsAsFactors = F)
 
 ForC_simplified <- merge(ForC_simplified, col.sym[, c("variable.name", "col", "sym")], by = "variable.name")
 ForC_simplified$col <- as.character(ForC_simplified$col)
@@ -130,7 +130,7 @@ ForC_simplified$sym <- as.character(ForC_simplified$sym)
 
[…]

### 2020-01-31 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@
 ```{r setup, include = FALSE}
-#knitr::opts_knit$set(root.dir = 'C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/')
+# knitr::opts_knit$set(root.dir = 'C:/Users/gyrcbm/Dropbox/Global_Productivity/manuscript/')
 knitr::opts_knit$set(root.dir = '/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/Global_Productivity/manuscript/')
@@
 | $GPP$ | Gross Primary Production | NA | flux partitioning of eddy-covariance; $NPP$+$R_{auto}$ | 243 | 49 |
-| $NPP$ | Net Primary Production | stem, foliage, coarse root, fine root, optionally others (e.g., branch, reproductive, understory) | $ANPP$ + $BNPP$ (majority); $GPP$-$R_{auto}$ | 92 | 42 |
-| $ANPP$ | Aboveground Net Primary Production | stem, foliage, optionally others (e.g., branch, reproductive, understory) |$ANPP_{woody-stem}$ + $ANPP_{foliage}$ (+ others) | 256 | 80 |
+| $NPP$ | Net Primary Production | stem, foliage, coarse root, fine root, optionally others (e.g., branch, reproductive, understory) | $ANPP$ + $BNPP$ (majority); $GPP$-$R_{auto}$ | 161 | 56 |
+| $ANPP$ | Aboveground Net Primary Production | stem, foliage, optionally others (e.g., branch, reproductive, understory) |$ANPP_{woody-stem}$ + $ANPP_{foliage}$ (+ others) | 278 | 86 |
 | $ANPP_{woody-stem}$ | Woody stem growth component of $ANPP$ | woody stems down to DBH $\le$ 10cm (no branch turnover) | stem growth measurements scaled to biomass using allometries | 264 | 96 |

### 2020-01-31 manuscript/manuscript draft.Rmd
text update related to geographic.area
relates to issue #49

@@
 
-Within the ForC database, sites within 25km^2^ of each other are clustered into geographic areas.
+Within the ForC database, geographically proximate sites are clustered into geographic areas, with proximity defined using a hierarchical cluster analysis on the distance matrix of the sites and a cutoff of 25km.
 To account for correlations in measurements between tightly clustered sites, a random effect was specified as plot nested within geographic area.
-*Data from the temperate regions was heavily skewed towards studies from the old-growth forests of the Pacific Northwest.*  These forests have very high FACF, and so to ensure that results were not unduly influenced by geographic sampling bias, we tried a version of the model where data were weighted according to forested land area within each Koeppen climate zone.
-Results were similar between the weighted and unweighted model.
-Because the weighted model may be prone to problems of over-fitting, we present only results of the unweighted model.
-{**Becky, I'd probably just delete this whole paragraph.
-Since we ended up not using the weighted model, it's just extra info for readers to digest.}
 

### 2020-01-31 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@
 ```
-*Data selection.* Over 50 variables of forest carbon stocks and annual fluxes are represented in the ForC database; this analysis focused on autotrophic carbon fluxes.
-Focal variables of this analysis are listed in Table 2.
+*Data selection.* This analysis focused on nine autotrophic carbon fluxes included in the ForC database (Table 2).
+These were selected based on *...*.
+The ForC database contains annual flux estimates, associated to plots, each with associated data on stand age and known history of disturbance and management, and, more broadly, sites (**Define**).
+Geographically proximate sites are clustered into geographic areas, with proximity defined using a hierarchical cluster analysis on the distance matrix of the sites and a cutoff of 25km [@anderson-teixeira_forc_2018].
 
@@
 
-Within the ForC database, geographically proximate sites are clustered into geographic areas, with proximity defined using a hierarchical cluster analysis on the distance matrix of the sites and a cutoff of 25km.
 To account for correlations in measurements between tightly clustered sites, a random effect was specified as plot nested within geographic area.

### 2020-01-31 scripts/stacked_plots_code.R
Update stacked_plots_code.R
updated to edit layout of graphs

@@ -298,11 +298,14 @@ par(mfrow = c(2,2), mar = c(2,2,2,2), oma = c(3,3,0,0))
           legend <- paste(set1, "+ R auto")
           # mtext(side = 3, line = -4, text = legend, adj = 0.95, col = "black", cex = 0.5, outer = F)
           
-          legend("topright", legend = c("NPP", "R auto", "GPP", "NPP + R auto"), col = plasma(10)[c(3, 4, 1, NA)], pch = c(3, 4, 1, NA), xpd = T, text.col = plasma(10)[c(3, 4, 1, NA)], bty = "n", xjust = 1, cex = 0.75)
+          legend("topright", legend = c("NPP", "R auto", "GPP", "NPP + R auto"), lty = 1, col = c(plasma(10)[c(3, 4, 1)], "black"), pch = c(3, 4, 1, NA), xpd = T, text.col = plasma(10)[c(3, 4, 1, NA)], bty = "n", xjust = 1, cex = 0.75)
           
+          # xaxis <- expression(paste("Absolute latitude (", degree, ")"))
+          # yaxis = expression(paste("Productivity (Mg C"~ha^-1~yr^-1, ")"))
+        
           # title(paste("Stacked graphs by latitude"), outer = F, line = 1)
-          mtext(side = 1, line = 1, text = "Latitude", outer = T)
-          mtext(side = 2, line = 1,  text = expression("Productivity Mg C"~ha^-1~yr^-1), outer = T)
+          mtext(side = 1, line = 1, text = expression(paste("Absolute latitude (", degree, ")")), outer = T)
+          mtext(side = 2, line = 1,  text = expression(paste("Productivity (Mg C"~ha^-1~yr^-1, ")")), outer = T)
           mtext(paste0("(", letters[1], ")"), side = 3, line = -1.5, adj = 0.05)
           
           # dev.off()
@@ -509,7 +512,7 @@ for (age in ages){
             legend <- paste("ANPP + BNPP")
             # mtext(side = 3, line = -4, text = legend, adj = 0.95, col = "black", cex = 0.5, outer = F)
             
-            legend("topright", legend = c("ANPP", "BNPP", "NPP", "ANPP + BNPP"), col = plasma(10)[c(5, 8, 3, NA)], pch = c(5, 8, 3, NA), xpd = T, text.col = plasma(10)[c(5, 8, 3, NA)], bty = "n", xjust = 1, cex = 0.75)
[…]

### 2020-02-07 manuscript/manuscript draft.Rmd
update

@@
 
-We found no evidence that allocation between fluxes varied substantially with latitude or climate.
-There were no significant results from regressing ratios of carbon fluxes against latitude, or against any of the climate variables.
+*We found no evidence that allocation between fluxes varied substantially with latitude or climate.
+There were no significant results from regressing ratios of carbon fluxes against latitude, or against any of the climate variables.* **THIS HAS CHANGED: [issue #51](https://github.com/forc-db/Global_Productivity/issues/51#issuecomment-581892186)
 

### 2020-02-12 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@
 hypothesis_table <- read.csv("hypothesis_table.csv", stringsAsFactors = FALSE, check.names = FALSE)
-kable(hypothesis_table, booktabs = TRUE, caption = "**Summary of hypotheses, corresponding specific predictions, and results.** Direction of significant relationships is indicated by '+' and '-', and  shape is summarized as linear (L), concave up (CU), and concave down (CD).
+kable(hypothesis_table, booktabs = TRUE, caption = "Summary of hypotheses, corresponding specific predictions, and results.
+Direction of significant relationships is indicated by '+' and '-', and  shape is summarized as linear (L), concave up (CU), and concave down (CD).
 'n.s.' indicates that the relationship is not signficant.") %>%
   kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 12) %>%
-  pack_rows("H1. FACFs decrease linearly with latitude, and in constant proportion.", 1, 3, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  pack_rows("H2. FACFs increase with MAT, with an interactive effect of MAP.", 4, 6, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  pack_rows("H3. FACFs are strongly correlated with other annual climate variables.", 7, 9, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  pack_rows("H4. FACFs are reduced under seasonal climates.", 10, 13, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  pack_rows("H5. Considering only growing season months, FACF correlates with climatic drivers.", 14, 17, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("Q1. How do FACFs vary with latitude?", 1, 3, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("Q2. How do FACFs vary with MAT and MAP?", 4, 6, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("Q3. How are FACFs related to climate variables, other than MAT and MAP?", 7, 9, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("Q4. How does seasonality influence FACFs?", 10, 13, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("Q5. When standardised by growing season length, how do FACFs vary with climate?", 14, 17, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
   add_header_above(c(" ", " ", " ", "Forest autotrophic carbon fluxes (FACF)" = 9, " ")) %>%

### 2020-02-12 manuscript/manuscript draft.Rmd
knit manuscript

@@
   add_header_above(c(" ", " ", " ", "Forest autotrophic carbon fluxes (FACF)" = 9, " ")) %>%
-  column_spec(1, width = "10cm") %>%
+  #column_spec(1, width = "15cm") %>%
   column_spec(2, width = "5cm") %>%

### 2020-02-13 manuscript/manuscript draft.Rmd
table 1 adjustment
Update manuscript draft.Rmd

--- 2020-02-07 manuscript/manuscript draft.Rmd
@@
 Taylor et al.
-[-@taylor_temperature_2017] showed that increased MAP had a negative effect at low MAT, but a positive effect at high levels of MAT, and *vice versa* {**Helene found this confusing**}, suggesting that both temperature and water availability are important in explaining productivity.
+[-@taylor_temperature_2017] showed that high rainfall had a negative effect on productivity where MAT was low, but a positive effect at sites with high MAT.
+Similarly, at sites with high MAT, increases in rainfall were found to enhance productivity, in contrast to cooler sites, where high levels of rainfall inhibit productivity.
 {**It would be good to add some more citations on soil respiraiton.
@@
 
-*We found no evidence that allocation between fluxes varied substantially with latitude or climate.
-There were no significant results from regressing ratios of carbon fluxes against latitude, or against any of the climate variables.* **THIS HAS CHANGED: [issue #51](https://github.com/forc-db/Global_Productivity/issues/51#issuecomment-581892186)
+We found no evidence that allocation between fluxes varied substantially with latitude or climate.
+There were no significant results from regressing ratios of carbon fluxes against latitude, or against any of the climate variables.
 

### 2020-02-13 manuscript/manuscript draft.Rmd

--- 2020-02-12 manuscript/manuscript draft.Rmd
@@
   kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 12) %>%
-  pack_rows("Q1. How do FACFs vary with latitude?", 1, 3, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("Q1. How do FACFs vary with latitude?", 1, 3, latex_gap_space = "1em", colnum = 1, hline_before = FALSE) %>%
   pack_rows("Q2. How do FACFs vary with MAT and MAP?", 4, 6, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%

### 2020-02-13 manuscript/manuscript draft.Rmd
Table 1: remove H1.1.alt and H1.2
#48

--- 2020-02-12 manuscript/manuscript draft.Rmd
@@
   kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 12) %>%
-  pack_rows("Q1. How do FACFs vary with latitude?", 1, 3, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  pack_rows("Q2. How do FACFs vary with MAT and MAP?", 4, 6, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  pack_rows("Q3. How are FACFs related to climate variables, other than MAT and MAP?", 7, 9, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  pack_rows("Q4. How does seasonality influence FACFs?", 10, 13, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  pack_rows("Q5. When standardised by growing season length, how do FACFs vary with climate?", 14, 17, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("Q1. How do FACFs vary with latitude?", 1, 1, latex_gap_space = "1em", colnum = 1, hline_before = FALSE) %>%
+  pack_rows("Q2. How do FACFs vary with MAT and MAP?", 2, 4, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("Q3. How are FACFs related to climate variables, other than MAT and MAP?", 5, 7, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("Q4. How does seasonality influence FACFs?", 8, 11, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("Q5. When standardised by growing season length, how do FACFs vary with climate?", 12, 15, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
   add_header_above(c(" ", " ", " ", "Forest autotrophic carbon fluxes (FACF)" = 9, " ")) %>%

### 2020-02-13 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

--- 2020-02-13 manuscript/manuscript draft.Rmd
@@
 
-As discussed above, estimates of $NPP$, $ANPP$, and $BNPP$ are generated through summing measurements of their component parts.
-Since studies vary in the components included in productivity estimates, these variables were classified within the ForC database according to their components, and then filtered for analysis.
-Estimates of NPP were selected if they included foliage, branch, stem, coarse root, and fine root.
-Measures of NPP which included additional components, including understorey, volatile organic compounds (VOCs), exudates, estimates of NPP lost to herbivory (**these are all small, no need to exclude**), and the NPP of reproductive structures (**often in literfall/ mixed in with foliage...
-did we really exclude estimates including litterfall?**), were excluded.
-Estimates of ANPP were selected if they included foliage, stem growth and optionally branch turnover.
-Any measures of primary productivity where components were unknown were excluded from analysis.
-
 *Climate datasets.* Where site-level data on mean annual temperature, mean annual precipitation, and latitude were available in the primary literature, this data was compiled and entered directly into the ForC database.
@@
 The effects of climate and latitude on FACF were analysed using mixed effects models using the package 'lme4' [@bates_fitting_2015] in R v.3.5.1 [@r_core_team_r:_2018].
-The effect of each extracted climate variable on each FACF variable was modelled by specifying the climate variable as a fixed effect.
-For each climate variable, three models were specified: a null model; a model with the climate variable as a linear term; and a model with the climate variable as a **XX**-order polynomial term.
+First, the individual effect of each climate variable on each FACF variable was modelled by specifying the climate variable as a fixed effect in a univariate model.
+For each climate variable, three models were specified: a null model; a model with the climate variable as a linear term; and a model with the climate variable as a second-order polynomial term.
 AIC values were calculated for the models and used to select the best model.
@@
 To test for a potential influence of altitude, models were also run with site altitude included as a second fixed effect.
-These models were compared against models with no altitude term, and AIC values calculated to identify whether inclusion of altitude as a term improved the models.
+These multivariate models were compared against univariate models with no altitude term, and AIC values calculated to identify whether inclusion of altitude as a term improved the models.
 Including altitude had a very small effect on most models, with the difference in AIC values between models including and excluding altitude often being <2, suggesting the models are very similar in their explanatory power.
-As a result, it was decided to present results only from models without altitude as a fixed term.
+As a result, it was decided not to include altitude as a fixed term in the final models.
 
@@
   i) the AIC value of the model was smaller than the AIC value of the baseline model by >2  
-  ii) the r-squared value of the model was larger than the r-squared value of the baseline model by **>5 (???- not possible)**
+  ii) the r-squared value of the model was larger than the r-squared value of the baseline model by >0.05
 

### 2020-02-13 manuscript/manuscript draft.Rmd

@@
   kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 12) %>%
-  pack_rows("Q1. How do FACFs vary with latitude?", 1, 3, latex_gap_space = "1em", colnum = 1, hline_before = FALSE) %>%
-  pack_rows("Q2. How do FACFs vary with MAT and MAP?", 4, 6, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  pack_rows("Q3. How are FACFs related to climate variables, other than MAT and MAP?", 7, 9, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  pack_rows("Q4. How does seasonality influence FACFs?", 10, 13, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  pack_rows("Q5. When standardised by growing season length, how do FACFs vary with climate?", 14, 17, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("Q1. How do FACFs vary with latitude?", 1, 1, latex_gap_space = "1em", colnum = 1, hline_before = FALSE) %>%
+  pack_rows("Q2. How do FACFs vary with MAT and MAP?", 2, 4, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("Q3. How are FACFs related to climate variables, other than MAT and MAP?", 5, 7, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("Q4. How does seasonality influence FACFs?", 8, 11, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("Q5. When standardised by growing season length, how do FACFs vary with climate?", 12, 15, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
   add_header_above(c(" ", " ", " ", "Forest autotrophic carbon fluxes (FACF)" = 9, " ")) %>%

### 2020-02-14 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@
 
-**Forests' autotrophic carbon flux (FACF, bbl:define) influences all organic matter stocks in ecosystems and is linked to cycling of energy, water, and nutrients** (**REFS; *Piao et al.
-2010.
-Forest annual carbon cost: a global‐scale analysis of autotrophic respiration.
-Ecology 91(3), 652-661* **).
-Forest productivity sets the energy ultimately available to heterotrophic organisms (including microbes), in turn influencing their abundance (**REFS**) and possibly diversity ([Waide et al.
-2009](https://www.annualreviews.org/doi/abs/10.1146/annurev.ecolsys.30.1.257)**REFS**).
+**Forests' autotrophic carbon flux (FACF, bbl:define) influences all organic matter stocks in ecosystems and is linked to cycling of energy, water, and nutrients** (**REFS**) [@piao_forest_2010].
+Forest productivity sets the energy ultimately available to heterotrophic organisms (including microbes), in turn influencing their abundance (**REFS**) and possibly diversity [@waide_relationship_1999] (**REFS**).
 However, our ability to draw general macroscopic conclusions regarding global variation in multiple FACF variables with respect to climate has been limited in that these analyses often mix forests that vary in stand age, disturbance history, and/or management status; do not always sufficiently parse related variables (*e.g.*, combining net primary productivity records with and without belowground components); and typically consider only one or a few variables at a time.
@@
 Evidence for this is further complicated by the fact that different studies use different measures of productivity to explore these relationships.
-For example, modelling of global terrestrial ecosystem gross primary productivity (GPP) through upscaling and calibration of eddy flux measurements indicates that GPP peaks in tropical forests [@beer_terrestrial_2010; @jung_global_2011];[Bagdley et al.
-2019](https://onlinelibrary.wiley.com/doi/epdf/10.1111/gcb.14729)).
-This is corroborated by analysis of site-level GPP measurements, which appear to reach their highest levels in tropical forests [@luyssaert_co_2007].
-{**condense the previous two sentences into one-- both primarily eddy flux**} In contrast, some studies suggest that the highest values of net primary productivity (NPP) may be found in temperate forests [@luyssaert_co_2007; @huston_global_2009], while others find NPP highest in the tropics and decreasing with latitude [@simova_enigma_2017].
+For example, modelling of global terrestrial ecosystem gross primary productivity (GPP) through upscaling and calibration of eddy flux measurements indicates that GPP peaks in tropical forests [@beer_terrestrial_2010; @jung_global_2011; @badgley_terrestrial_2019].
+In contrast, some studies suggest that the highest values of net primary productivity (NPP) may be found in temperate forests [@luyssaert_co_2007; @huston_global_2009], while others find NPP highest in the tropics and decreasing with latitude [@simova_enigma_2017].
 Other studies have chosen to focus exclusively on above-ground net primary productivity (ANPP), finding evidence of a weak negative relationship between ANPP and latitude [@huston_global_2009; @gillman_latitude_2015].
@@
 **FACF can be influenced by many other factors as well**, which often act across a range of scales, and may show interactive effects with each other [@cleveland_relationships_2011].
-On a local scale, stand age [@litton_carbon_2007; @gillman_latitude_2015], biodiversity (Liang et al.
-2019.
-Positive biodiversity-productivity relationship predominant in global forests.
-Science 354, Issue 6309, aaf8957), management [@simova_enigma_2017]; nutrient availability [@aragao_above-_2009]; and altitude [@girardin_net_2010; @malhi_variation_2017] all impact FACF.
+On a local scale, stand age [@litton_carbon_2007; @gillman_latitude_2015], biodiversity [@liang_positive_2016], management [@simova_enigma_2017]; nutrient availability [@aragao_above-_2009]; and altitude [@girardin_net_2010; @malhi_variation_2017] all impact FACF.
 On a global scale, we expect that FACF is most strongly influenced by broad climatic gradients.

### 2020-02-14 manuscript/manuscript draft.Rmd
removed methodological consistency section in methods
#56

@@
 
+- excluded alkali absoption and soda lime methods for estimating soil / root respiration
+
 \newpage  
@@
 Geographic areas group geographically proximate sites, defined using a hierarchical cluster analysis on the distance matrix of the sites, and a cutoff of 25km.
-
-*Methodological consistency.* 
-{**This whole section needs to be re-written.**} 
-The data in ForC is derived from a range of studies, often employing different methods.
-For this reason, criteria were introduced to standardise for differences in methodology.
-Where data was based on forest plot census measurements, studies which used a minimum diameter at breast height (DBH) measure of >10cm were excluded from analysis.
-It would be preferable to standardise by minimum area sampled; however x% of plots in the database are 1 ha or under in size; excluding these plots would place significant constraints on sample size.
-(**really, we didn't exclude any studies based on methodology (except soil respiration.
-all studies measured min DBH 10 or less...
-), so this paragraph needs revision.
-Also, address whether we standardize for inclusion/ exclusion of recruitment.**)
-- recruitment preferred when an option, did not review whether all included recruitment
-- excluded alkali absoption and soda lime methods for estimating soil / root respiration
 

### 2020-02-15 manuscript/manuscript draft.Rmd
rewrote first paragraph of intro

@@
 [**very rough start:**]
-.....
+Carbon fixation, allocation, and metabolism by trees set the basis for energy and material flows in forest ecosystems and define their interactions with Earth's changing climate.
 Here, we drew upon # records from the Global Forest Carbon Database (ForC), representing all major forest types and the nine most significant forest autotrophic carbon flux (FACF) variables, to comprehensively explore how C cycling in mature, undisturbed forests varies with latitude and climate on a global scale.
@@
 
-(**revisit these first two paragraphs.**)
-
-**Forests play a critical role in regulating atmospheric CO~2~ and climate** (**Bonan**).
-Approximately **##** Gt of CO~2~ cycle through Earth's forests each year **REF-KAT**), and their net C sequestration (~2.4 GtCyr^-1^) offsets roughly 30% of anthropogenic fossil fuel emissions [@pan_large_2011].
-As atmospheric carbon dioxide levels continue to rise, driving climate change, there is increasing recognition that  protection and management of forest resources will play an important role in mitigating climate change (** REFS- e.g., IPCC 1.5C and land reports**).
-Carbon fixation, allocation, and metabolism by trees set the basis for energy and material flows in forest ecosystems and define their interactions with Earth's changing climate.
-
-**Forests' autotrophic carbon flux (FACF, bbl:define) influences all organic matter stocks in ecosystems and is linked to cycling of energy, water, and nutrients** (**REFS**) [@piao_forest_2010].
-Forest productivity sets the energy ultimately available to heterotrophic organisms (including microbes), in turn influencing their abundance (**REFS**) and possibly diversity [@waide_relationship_1999] (**REFS**).
-However, our ability to draw general macroscopic conclusions regarding global variation in multiple FACF variables with respect to climate has been limited in that these analyses often mix forests that vary in stand age, disturbance history, and/or management status; do not always sufficiently parse related variables (*e.g.*, combining net primary productivity records with and without belowground components); and typically consider only one or a few variables at a time.
+**Carbon cycling in forests worldwide is central to sustaining an estimated #% of Earth's biodiversity (*REF*), while strongly influencing atmospheric CO~2~ and climate** ([Bonan 2008](https://science.sciencemag.org/content/320/5882/1444)).
+Forests' autotrophic carbon flux (FACF)--that is, carbon fixation, allocation, and metabolism by trees and other primary producers--sets the energy ultimately available to heterotrophic organisms (including microbes), in turn influencing their abundance (**REFS**) and possibly diversity [@waide_relationship_1999] (**REFS**).
+FACF influences all organic matter stocks in forest ecosystems and is linked to cycling of energy, water, and nutrients (**REFS**) [@piao_forest_2010].
+Critically, FACF also defines forest interactions with Earth's changing climate.
+Approximately **##** Gt of CO~2~ cycle through Earth's forests each year **REF-KAT**), and in recent decades their net C sequestration (~2.4 GtCyr^-1^) offset roughly 30% of anthropogenic fossil fuel emissions [@pan_large_2011].
+As atmospheric carbon dioxide levels continue to rise, driving climate change, forests will play a critical role in shaping the future of Earth's climate (** REFS**).
+However, our ability to draw general macroscopic conclusions regarding global variation in multiple FACF variables with respect to climate has been limited in that these analyses often mix forests that vary in stand age, disturbance history, and/or management status; do not always sufficiently parse related variables; and typically consider only one or a few variables at a time.
 

### 2020-02-15 scripts/ratio_regression_script.R
Update ratio_regression_script.R

@@ -3,7 +3,7 @@
 rm(list = ls())
 
 # Set working directory as ForC main folder ####
-setwd("C:/Users/gyrcbm/Dropbox/ForC")
+setwd("C:/Users/becky/Dropbox (Smithsonian)/GitHub/ForC")
 
 # Load libaries ####
 library(lme4)
@@ -96,15 +96,18 @@ all.response.variables <- unique(gsub("_\\d", "", all.response.variables))
 all.results <- NULL
 all.aictab <- NULL
 
-fixed.variables <- c("mat", "map", "lat", "TempSeasonality")
+fixed.variables <- c("lat", "mat", "map", "TempSeasonality")
 
-set1 <- c("GPP", "ANPP", "ANPP_foliage", "ANPP_foliage", "ANPP_woody_stem", "ANPP", "BNPP_root")
-set2 <- c("NPP", "BNPP_root", "ANPP_woody_stem", "NPP", "NPP", "NPP", "NPP")
+# set1 <- c("GPP", "ANPP", "ANPP_foliage", "ANPP_foliage", "ANPP_woody_stem", "ANPP", "BNPP_root")
+# set2 <- c("NPP", "BNPP_root", "ANPP_woody_stem", "NPP", "NPP", "NPP", "NPP")
+
+set1 <- c("GPP", "GPP", "GPP", "GPP", "GPP", "GPP", "GPP")
+set2 <- c("NPP", "ANPP", "ANPP_foliage", "ANPP_woody_stem", "BNPP_root", "BNPP_root_fine", "R_auto")
 
 # set1 <- c("ANPP_1", "ANPP_1", "BNPP_root", "ANPP_2", "ANPP_2", "ANPP", "ANPP")
 # set2 <- c("BNPP_root", "NPP_1", "NPP_1", "BNPP_root", "NPP_1", "BNPP_root", "NPP_1")
 
-png(file = paste0("C:/Users/gyrcbm/Dropbox/Global_Productivity/results/figures/final_figures/ratio_regressions/ratio_grid_plots.png"), width = 3000, height = 3500, units = "px", res = 300)
+png(file = paste0("C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/figures/final_figures/ratio_regressions/ratio_grid_plots_1.png"), width = 3000, height = 3500, units = "px", res = 300)
 
 
 par(mfrow = c(7,4), mar = c(2,2,2,2), oma = c(5,14,2,0))
[…]

### 2020-02-15 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@
 
-**Carbon cycling in forests worldwide is central to sustaining an estimated #% of Earth's biodiversity (*REF*), while strongly influencing atmospheric CO~2~ and climate** ([Bonan 2008](https://science.sciencemag.org/content/320/5882/1444)).
+**Carbon cycling in forests worldwide is central to sustaining an estimated #% of Earth's biodiversity (*REF*), while strongly influencing atmospheric CO~2~ and climate** [@bonan_forests_2008].
 Forests' autotrophic carbon flux (FACF)--that is, carbon fixation, allocation, and metabolism by trees and other primary producers--sets the energy ultimately available to heterotrophic organisms (including microbes), in turn influencing their abundance (**REFS**) and possibly diversity [@waide_relationship_1999] (**REFS**).
@@
 hypothesis_table <- read.csv("hypothesis_table.csv", stringsAsFactors = FALSE, check.names = FALSE)
-kable(hypothesis_table, booktabs = TRUE, caption = "Summary of hypotheses, corresponding specific predictions, and results.
-Direction of significant relationships is indicated by '+' and '-', and  shape is summarized as linear (L), concave up (CU), and concave down (CD).
-'n.s.' indicates that the relationship is not signficant.") %>%
+kable(hypothesis_table, booktabs = TRUE, caption = "Summary of research questions, corresponding hypotheses, and results.'yes' indicates the hypothesis is supported, 'no' indicates the hypothesis is not supported, and 'n.s.' indicates that the relationship is not signficant.") %>%
   kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 12) %>%
@@
   pack_rows("Q2. How do FACFs vary with MAT and MAP?", 2, 4, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  pack_rows("Q3. How are FACFs related to climate variables, other than MAT and MAP?", 5, 7, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("Q3. How are FACFs related to other climate variables?", 5, 7, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
   pack_rows("Q4. How does seasonality influence FACFs?", 8, 11, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
@@
 | $ANPP$ | Aboveground Net Primary Production | stem, foliage, optionally others (e.g., branch, reproductive, understory) |$ANPP_{woody-stem}$ + $ANPP_{foliage}$ (+ others) | 278 | 86 |
-| $ANPP_{woody-stem}$ | Woody stem growth component of $ANPP$ | woody stems down to DBH $\le$ 10cm (no branch turnover) | stem growth measurements scaled to biomass using allometries | 264 | 96 |
+| $ANPP_{woody stem}$ | Woody stem growth component of $ANPP$ | woody stems down to DBH $\le$ 10cm (no branch turnover) | stem growth measurements scaled to biomass using allometries | 264 | 96 |
 | $ANPP_{foliage}$ | Foliage component of $ANPP$ | foliage | litterfall collection (separated into components) | 98 | 49 |
 | $BNPP$  | Belowground Net Primary Production | coarse and fine roots |  coarse roots estimated indirectly using allometries based on aboveground stem increment measures ; fine roots as below | 101 | 48 |
-| $BNPP_{fine.root}$ | Fine root component of $BNPP$ | fine roots | measurements combined one or more of the following: soil cores, minirhizotrons, turnover estimates,  root ingrowth cores | 88 | 41 |
+| $BNPP_{fine root}$ | Fine root component of $BNPP$ | fine roots | measurements combined one or more of the following: soil cores, minirhizotrons, turnover estimates,  root ingrowth cores | 88 | 41 |
 | $R_{root}$ | Root Respiration | coarse and fine roots | partitioning of total soil respiration (e.g., through root exclusion), scaling of root gas exchange | 64 | 26 |

### 2020-02-16 manuscript/manuscript draft.Rmd
made FACF plural throughout
issue #57

@@
 **Carbon cycling in forests worldwide is central to sustaining an estimated #% of Earth's biodiversity (*REF*), while strongly influencing atmospheric CO~2~ and climate** [@bonan_forests_2008].
-Forests' autotrophic carbon flux (FACF)--that is, carbon fixation, allocation, and metabolism by trees and other primary producers--sets the energy ultimately available to heterotrophic organisms (including microbes), in turn influencing their abundance (**REFS**) and possibly diversity [@waide_relationship_1999] (**REFS**).
+Forests' autotrophic carbon fluxes (FACF)--that is, carbon fixation, allocation, and metabolism by trees and other primary producers--sets the energy ultimately available to heterotrophic organisms (including microbes), in turn influencing their abundance (**REFS**) and possibly diversity [@waide_relationship_1999] (**REFS**).
 FACF influences all organic matter stocks in forest ecosystems and is linked to cycling of energy, water, and nutrients (**REFS**) [@piao_forest_2010].
-Critically, FACF also defines forest interactions with Earth's changing climate.
+Critically, FACF also define forest interactions with Earth's changing climate.
 Approximately **##** Gt of CO~2~ cycle through Earth's forests each year **REF-KAT**), and in recent decades their net C sequestration (~2.4 GtCyr^-1^) offset roughly 30% of anthropogenic fossil fuel emissions [@pan_large_2011].
 As atmospheric carbon dioxide levels continue to rise, driving climate change, forests will play a critical role in shaping the future of Earth's climate (** REFS**).
-However, our ability to draw general macroscopic conclusions regarding global variation in multiple FACF variables with respect to climate has been limited in that these analyses often mix forests that vary in stand age, disturbance history, and/or management status; do not always sufficiently parse related variables; and typically consider only one or a few variables at a time.
+However, our ability to draw general macroscopic conclusions regarding global variation in FACF with respect to climate has been limited in that these analyses often mix forests that vary in stand age, disturbance history, and/or management status; do not always sufficiently parse related variables; and typically consider only one or a few variables at a time.
 
-**FACF varies with latitude**, showing a general trend of decreasing FACF with latitude [@beer_terrestrial_2010; @jung_global_2011].
-Studies agree that FACF is lowest in the boreal regions, and increases into the temperate regions [@luyssaert_co_2007; @huston_global_2009; @beer_terrestrial_2010; @jung_global_2011].
+**FACF vary with latitude**, showing a general trend of decreasing flux with latitude [@beer_terrestrial_2010; @jung_global_2011].
+Studies agree that FACF are lowest in the boreal regions, and increase into the temperate regions [@luyssaert_co_2007; @huston_global_2009; @beer_terrestrial_2010; @jung_global_2011].
 However, evidence is inconclusive on whether primary productivity continues to increase into the tropics, or whether it plateaus in temperate regions.
@@
 On a local scale, stand age [@litton_carbon_2007; @gillman_latitude_2015], biodiversity [@liang_positive_2016], management [@simova_enigma_2017]; nutrient availability [@aragao_above-_2009]; and altitude [@girardin_net_2010; @malhi_variation_2017] all impact FACF.
-On a global scale, we expect that FACF is most strongly influenced by broad climatic gradients.
-There is evidence that FACF also responds to variables such as cloud cover [@taylor_temperature_2017], solar radiation [@fyllas_solar_2017], and potential evapotranspiration [@kerkhoff_plant_2005] in potentially significant ways.
+On a global scale, we expect that FACF are most strongly influenced by broad climatic gradients.
+There is evidence that FACFs also respond to variables such as cloud cover [@taylor_temperature_2017], solar radiation [@fyllas_solar_2017], and potential evapotranspiration [@kerkhoff_plant_2005] in potentially significant ways.
 Furthermore, MAT and MAP are very coarse measures of climate, and so fail to capture much variation in climate on an intra-annual scale, including the effects of factors such as growing season length, number of frost-free days, temperature seasonality, and dry season length.
-Some studies have suggested that the apparently strong relationship between MAT and FACF is actually a consequence of the correlation between MAT and growing season length [@kerkhoff_plant_2005; @malhi_productivity_2012; @michaletz_convergence_2014; @michaletz_drivers_2018].
+Some studies have suggested that the apparently strong relationship between MAT and FACFs is actually a consequence of the correlation between MAT and growing season length [@kerkhoff_plant_2005; @malhi_productivity_2012; @michaletz_convergence_2014; @michaletz_drivers_2018].
 Kerkhoff et al.
@@
 
-**The recent development of a global forest carbon database synthesizing multiple variables and including records of stand history (ForC; [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc:_2018]) opens up the possibility for a standardized analysis of global scale variation** in multiple components of FACF and the principle climatic drivers of these patterns.
-In order to approach these broad and complex issues, we simplify the major gaps in our knowledge to five broad hypotheses and corresponding specific predictions (Table 1).
+**The recent development of a global forest carbon database synthesizing multiple variables and including records of stand history (ForC; [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc:_2018]) opens up the possibility for a standardized analysis of global scale variation** in multiple FACF and the principle climatic drivers of these patterns.
+In order to approach this broad topic, we simplify the major gaps in our knowledge to five broad questions and corresponding hypotheses (Table 1).
 ** see [issue #48 ](https://github.com/forc-db/Global_Productivity/issues/48)** First, we ask how forest autotrophic carbon fluxes (FACF) vary with latitude.
-We then test how these fluxes relate to MAT and MAP, and additionally how they respond to other, less well studied, climate variables.
+We then test how these fluxes relate to MAT and MAP, and additionally how they respond to other, less well-studied, climate variables.
 Finally, we consider the relationship between FACF and seasonality, considering the role of seasonality in explaining variation in carbon fluxes, and the influence of climate on FACF standardized by growing season length.
@@
   kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 12) %>%
-  pack_rows("Q1. How do FACFs vary with latitude?", 1, 1, latex_gap_space = "1em", colnum = 1, hline_before = FALSE) %>%
-  pack_rows("Q2. How do FACFs vary with MAT and MAP?", 2, 4, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  pack_rows("Q3. How are FACFs related to other climate variables?", 5, 7, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  pack_rows("Q4. How does seasonality influence FACFs?", 8, 11, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  pack_rows("Q5. When standardised by growing season length, how do FACFs vary with climate?", 12, 15, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("Q1. How do FACF vary with latitude?", 1, 1, latex_gap_space = "1em", colnum = 1, hline_before = FALSE) %>%
+  pack_rows("Q2. How do FACF vary with MAT and MAP?", 2, 4, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("Q3. How are FACF related to other climate variables?", 5, 7, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("Q4. How does seasonality influence FACF?", 8, 11, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("Q5. When standardised by growing season length, how do FACF vary with climate?", 12, 15, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
   add_header_above(c(" ", " ", " ", "Forest autotrophic carbon fluxes (FACF)" = 9, " ")) %>%
@@
 The effects of climate and latitude on FACF were analysed using mixed effects models using the package 'lme4' [@bates_fitting_2015] in R v.3.5.1 [@r_core_team_r:_2018].
-First, the individual effect of each climate variable on each FACF variable was modelled by specifying the climate variable as a fixed effect in a univariate model.
+First, the individual effect of each climate variable on each flux variable was modelled by specifying the climate variable as a fixed effect in a univariate model.
 For each climate variable, three models were specified: a null model; a model with the climate variable as a linear term; and a model with the climate variable as a second-order polynomial term.
@@
 
-Models were run for total annual FACF against annual climate variables, and for fluxes standardized by growing season length, defined as annual FACF/length of growing season, against growing season climate variables.
+Models were run for total annual flux against annual climate variables, and for fluxes standardized by growing season length, defined as annual flux/length of growing season, against growing season climate variables.
 For analyses on data within the growing season, only linear models were specified.
@@
 
-*How does FACF vary with latitude?*
+*How do FACF vary with latitude?*
 
@@
 
-*How does FACF relate to MAT and MAP?* MAT and MAP are the most commonly reported site-level climate variables, and much previous research into the effect of climate on FACF has focused on these as key climate variables.
-MAT was a significant (p<0.05) and strong predictor of FACF for all fluxes tested.
+*How do FACF relate to MAT and MAP?* MAT and MAP are the most commonly reported site-level climate variables, and much previous research into the effect of climate on FACF has focused on these as key climate variables.
+MAT was a significant (p<0.05) and strong ($R^2 \ge$ **##**) predictor of all FACF tested.
 All FACF increased linearly with temperature (Fig.
-4), and we found no support for a saturation point of FACF with temperature.
+4; all p<0.05, $R^2 \ge$ **##**), and we found no support for a saturation point of FACF with temperature.
 
@@
 
-*How does FACF relate to other climate variables?* {**Here, maybe add some statement about covariation among climate variables, site SI figure**} Our results indicated that FACF was most strongly explained by temperature at the global scale, with temperature-related climate variables coming out as strong predictors of FACF.
+*How do FACF relate to other climate variables?* {**Here, maybe add some statement about covariation among climate variables, site SI figure**} Our results indicated that FACF were most strongly explained by temperature at the global scale, with temperature-related climate variables coming out as strong predictors of FACF.
 In addition to MAT, temperature seasonality, annual temperature range, and annual frost days were consistently identified as strong univariate predictors of FACF.
@@
 
-We found a significant relationship between FACF and potential evapotranspiration for all fluxes.
+We found a significant relationship between C flux and potential evapotranspiration for all FACF.
 ANPP~foliage~, BNPP~fine~ ~root~ and R~root~ increased linearly with PET; however, all other fluxes showed a polynomial relationship with PET (fig.
@@
 We found strong evidence for a saturation point or peak with PET: FACF tended to increase at values below 1000mm, before saturating between 1200 and 1700mm.
-There was also evidence that FACF begins to decrease at values above 1800mm PET.
+There was also evidence that FACF begin to decrease at values above 1800mm PET.
 
-Vapour pressure deficit was a significant predictor of FACF for all fluxes.
+Vapour pressure deficit was a significant predictor of C flux for all FACF.
 BNPP~fine~ ~root~ showed a linear relationship with vapour pressure deficit ($R^2$ = 0.07, p<0.05), but all other fluxes showed a polynomial relationship (fig.
@@
 All fluxes, with the exception of R~root~, showed a positive linear relationship with solar radiation.
-Solar radiation explained a low proportion of variability in FACF for all variables, explaining less than 20% of the variation in each flux, with the exception of R~auto~ ($R^2$ = 0.26, p<0.05).
+Solar radiation explained a low proportion of variability in all FACF, explaining less than 20% of the variation in each flux, with the exception of R~auto~ ($R^2$ = 0.26, p<0.05).
 
@@
 Temperature seasonality was strongly correlated with annual temperature range, and, as expected, all fluxes showed almost identical responses to it {**meaning unclear**}.
-FACF was highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower.
+FACF were highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower.
 In contrast, there was no significant effect of precipitation seasonality on FACF.
@@
 
-*Within the growing season, how does FACF vary with climate?* When FACF was standardized by growing season length, correlations with climate variables were much weaker.
+*Within the growing season, how do FACF vary with climate?* When FACF were standardized by growing season length, correlations with climate variables were much weaker.
 Growing season temperature, precipitation, solar radiation, and PET each had a small effect on one or more of the nine FACF analyzed (Table 2; Fig.
@@
 We used a large global database (ForC), containing an unprecedented amount of data representing all major forest types (Fig.
-1) and the nine most significant forest autotrophic carbon flux (FACF) variables (Table 2), to comprehensively explore how C cycling in mature, undisturbed forests varies with latitude and climate on a global scale.
+1) and the nine most significant forest autotrophic carbon fluxes (FACF) variables (Table 2), to comprehensively explore how C cycling in mature, undisturbed forests varies with latitude and climate on a global scale.
 We test a suite of hypotheses, including many previously posed (Table 1), standardising for factors that have not always been controlled for (*e.g.*, stand age, flux components measured) to gain a stronger understanding of climatic effects on FACF.
@@
 
-We show that, across all nine FACF variables analyzed, C cycling decreases linearly with latitude (*H1.1*; Fig.
+We show that, across all nine FACF analyzed, C cycling decreases linearly with latitude (*H1.1*; Fig.
 2)--a finding that confirms multiple previous studies (**REFS**) but contradicts the idea that productivity of temperate forests rivals that of tropical forests (*H1.1.alt*; **REFS**).
-The FACF variables increase in proportion to one another (*H1.2*), with no differences in allocation detectable at this global scale and with component fluxes summing appropriately to larger fluxes (Fig.
+The FACF increase in proportion to one another (*H1.2*), with no differences in allocation detectable at this global scale and with component fluxes summing appropriately to larger fluxes (Fig.
 2), indicating no major, systematic omissions or overestimations of flux components.
@@
 Within the growing season, the influence of climate on C cycling is smaller but still significant for a number of carbon fluxes (*H5*; **some SI table?**).
-These findings clarify the big picture of how FACF varies with latitude and climate on a global scale.
+These findings clarify the big picture of how FACF vary with latitude and climate on a global scale.
 
 Past studies have differed in their conclusions regarding the relationship between FACF and latitude or its correlates (Table 1, *H1*; **REFS**)--quite possibly because of lack of standardization with respect to stand age and disturbance history.
-Our findings indicate that, among mature, undisturbed stands, FACF is unambiguously highest in the tropical regions, and the relationship is approximately linear (Fig.
+Our findings indicate that, among mature, undisturbed stands, FACF are unambiguously highest in the tropical regions, and the relationship is approximately linear (Fig.
 2).
@@
 3, 4b).
-Negative effects of very high precipitation on FACF have been previously observed (**REFS**) and are attributable to nutrient and light limitations (**REFS**) Thus, although temperature and water availability jointly and interactively drive global-scale patterns of FACF.
+Negative effects of very high precipitation on FACF have been observed previously (**REFS**) and are attributable to nutrient and light limitations (**REFS**).
+Thus, although temperature and water availability jointly and interactively drive global-scale patterns of FACF.
 
@@
 
-FACF is negatively correlated with temperature seasonality (Table 1, *H4*; Fig.
+FACF are negatively correlated with temperature seasonality (Table 1, *H4*; Fig.
 4e), and is minimal during cold- or dry- dormant seasons.
-To account for this, a number of analyses seeking to characterize global-scale effects of climate on productivity have examined the relationship of FACF per month of the growing season with growing season climatic conditions (Table 1, *H5*; **REFS**).
+To account for this, a number of analyses seeking to characterize global-scale effects of climate on productivity have examined the relationship of C flux per month of the growing season with growing season climatic conditions (Table 1, *H5*; **REFS**).
 We found that the sort of simple metric needed to define growing season at a global scale was uncertain for temperature and problematic for moisture (**WORK ON THIS**).
@@
 
-Our analysis clarifies how FACF varies (bbl: is FACF singular or plural?
-Be consistent) with latitude and climate on a global scale, with some important implications for how forest carbon cycling relates to climate and, by extension, how it is likely to respond to climatic warming.
+Our analysis clarifies how FACF vary with latitude and climate on a global scale, with some important implications for how forest carbon cycling relates to climate and, by extension, how it is likely to respond to climatic warming.
 We find no support for non-linear trends in mature forest C cycling with respect to latitude or $MAT$, and no distinct trends in C allocation across the global scale (Fig.

### 2020-02-16 manuscript/manuscript draft.Rmd
cleanup of Table 1
#48

@@
 hypothesis_table <- read.csv("hypothesis_table.csv", stringsAsFactors = FALSE, check.names = FALSE)
-kable(hypothesis_table, booktabs = TRUE, caption = "Summary of research questions, corresponding hypotheses, and results.'yes' indicates the hypothesis is supported, 'no' indicates the hypothesis is not supported, and 'n.s.' indicates that the relationship is not signficant.") %>%
+kable(hypothesis_table, booktabs = TRUE, caption = "Summary of research questions, corresponding hypotheses, and results.
+Statistically signficant support for/ rejection of hypotheses is indicated with 'yes'/'no', parentheses indicate partial overall support/rejection of hypotheses across all FACF, and '-' indicates no significant relationship.") %>%
   kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 12) %>%

### 2020-02-16 manuscript/appendix.Rmd
updated supporting information
started fixing up SI
#59

@@
 ---
-title: "appendix"
+title: "Supplementary Information: Global patterns of forest autotrophic carbon fluxes"
+author: "Rebecca Banbury Morgan, Valentine Herrmann, Norbert Kunert, Ben Bond-Lamberty, Helene C. Muller-Landau, Kristina J. Anderson-Teixeira"
+toc: true
+toc_depth: 1
 output: 
@@
 
+```{r setup, include = FALSE}
+# knitr::opts_knit$set(root.dir = 'C:/Users/gyrcbm/Dropbox/Global_Productivity/manuscript/')
+knitr::opts_knit$set(root.dir = '/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/Global_Productivity/manuscript/')
+```
+
+\newpage
+#Table S1. Climate variable definitions and sources
 ```{r eval = TRUE, echo=FALSE, warning=FALSE}
@@
 library(kableExtra)
-S1 <- read.csv("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/S1.csv", stringsAsFactors = FALSE, check.names = FALSE)
-kable(S1, format = "latex", booktabs = TRUE, caption = "Supporting Information Table 1 (S1): ", escape = F)%>%
+S1 <- read.csv("climate_variables.csv", stringsAsFactors = FALSE, check.names = FALSE)
+kable(S1, format = "latex", booktabs = TRUE, caption = "", escape = F)%>%
   kable_styling(latex_options = c("scale_down", "hold_position")) %>%

### 2020-02-16 scripts/mat_map_interaction.R
Update mat_map_interaction.R

@@ -4,13 +4,14 @@
 rm(list = ls())
 
 # Set working directory as ForC main folder ####
-setwd("C:/Users/gyrcbm/Dropbox/ForC")
+setwd("C:/Users/becky/Dropbox (Smithsonian)/GitHub/ForC")
 
 # Load libaries ####
 library(lme4)
 library(MuMIn)
 library(AICcmodavg)
 library(piecewiseSEM)
+library(viridis)
 
 # Load data ####
 ForC_simplified <- read.csv("ForC_simplified/ForC_simplified_WorldClim_CRU_refined.csv", stringsAsFactors = F)
@@ -111,7 +112,7 @@ pannel.nb <- 1
 
 response.variables <- c("GPP", "NPP", "ANPP", "ANPP_woody_stem", "ANPP_foliage", "BNPP_root", "BNPP_root_fine", "R_auto", "R_auto_root")
 
-png(file = paste0("C:/Users/gyrcbm/Dropbox/Global_Productivity/results/figures/final_figures/interactions/mat_map_interaction.png"), width = 2255, height = 2000, units = "px", res = 300)
+png(file = paste0("C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/figures/final_figures/interactions/mat_map_interaction.png"), width = 2255, height = 2000, units = "px", res = 300)
 par(mfrow = c(3,3), mar = c(1,0,0,2), oma = c(5,8,2,0), xpd = T)
 
 for (response.v in response.variables){

### 2020-02-16 scripts/final_model_no_weighting.R
Update final_model_no_weighting.R

@@ -3,7 +3,7 @@
 rm(list = ls())
 
 # Set working directory as ForC main folder ####
-setwd("C:/Users/gyrcbm/Dropbox/ForC")
+setwd("C:/Users/becky/Dropbox (Smithsonian)/GitHub/ForC")
 
 # Load libaries ####
 library(lme4)
@@ -137,12 +137,12 @@ for(response.variables in response.variables.groups){
     
     for(fixed.v in fixed.variables){
       
-      png(file = paste0("C:/Users/gyrcbm/Dropbox/Global_Productivity/results/figures/final_figures/unweighted_model/Effect_of_", fixed.v, "_MATURE_only_", age, "_", n, ".png"), width = 2255, height = 2000, units = "px", res = 300)
+      png(file = paste0("C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/figures/final_figures/unweighted_model/Effect_of_", fixed.v, "_MATURE_only_", age, "_", n, ".png"), width = 2255, height = 2000, units = "px", res = 300)
       
       par(mfrow = c(1,1), mar = c(0,0,0,0), oma = c(5,5,2,0))
       print(fixed.v)
       
-      fixed.v.info <- read.csv("C:/Users/gyrcbm/Dropbox/Global_Productivity/raw.data/fixedv_data.csv", stringsAsFactors = F)
+      fixed.v.info <- read.csv("C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/raw.data/fixedv_data.csv", stringsAsFactors = F)
       
       xaxis <- fixed.v.info$xaxis[which(fixed.v.info$fixed.v %in% fixed.v)]
       
@@ -153,7 +153,7 @@ for(response.variables in response.variables.groups){
       
       for (response.v in response.variables){
         
-        col.sym <- read.csv("C:/Users/gyrcbm/Dropbox/Global_Productivity/raw.data/colsym.csv", stringsAsFactors = F)
+        col.sym <- read.csv("C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/raw.data/colsym.csv", stringsAsFactors = F)
         
         col <- col.sym$col[which(col.sym$variable %in% response.v)]
         sym <- col.sym$sym[which(col.sym$variable %in% response.v)]
@@ -285,7 +285,7 @@ for(response.variables in response.variables.groups){
 }
 
 
[…]

### 2020-02-16 scripts/extract_climate_data_growing_season.R
updated within growing season grid plots

@@ -359,7 +359,7 @@ write.csv(ForC, "C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/ForC/ForC_
 rm(list = ls())
 
 # Set working directory as ForC main folder ####
-setwd("C:/Users/gyrcbm/Dropbox/ForC")
+setwd("C:/Users/becky/Dropbox (Smithsonian)/GitHub/ForC")
 
 # Load libaries ####
 library(lme4)
@@ -469,12 +469,12 @@ for(response.variables in response.variables.groups){
     
     for(fixed.v in fixed.variables){
       
-      png(file = paste0("C:/Users/gyrcbm/Dropbox/Global_Productivity/results/figures/final_figures/unweighted_model/growing_season/Effect_of_", fixed.v, "_MATURE_only_", age, "_", n, ".png"), width = 2255, height = 2000, units = "px", res = 300)
+      png(file = paste0("C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/figures/final_figures/unweighted_model/growing_season/Effect_of_", fixed.v, "_MATURE_only_", age, "_", n, ".png"), width = 2255, height = 2000, units = "px", res = 300)
       
       par(mfrow = c(1,1), mar = c(0,0,0,0), oma = c(5,5,2,0))
       print(fixed.v)
       
-      fixed.v.info <- read.csv("C:/Users/gyrcbm/Dropbox/Global_Productivity/raw.data/fixedv_data.csv", stringsAsFactors = F)
+      fixed.v.info <- read.csv("C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/raw.data/fixedv_data.csv", stringsAsFactors = F)
       
       xaxis <- fixed.v.info$xaxis[which(fixed.v.info$fixed.v %in% fixed.v)]
       
@@ -485,7 +485,7 @@ for(response.variables in response.variables.groups){
       
       for (response.v in response.variables){
         
-        col.sym <- read.csv("C:/Users/gyrcbm/Dropbox/Global_Productivity/raw.data/colsym.csv", stringsAsFactors = F)
+        col.sym <- read.csv("C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/raw.data/colsym.csv", stringsAsFactors = F)
         
         col <- col.sym$col[which(col.sym$variable %in% response.v)]
[…]

### 2020-02-16 manuscript/manuscript draft.Rmd
updated manuscript draft to edit hypothesis table and reorder variables in table 2

--- 2020-02-16 manuscript/manuscript draft.Rmd
@@
 ```{r setup, include = FALSE}
-# knitr::opts_knit$set(root.dir = 'C:/Users/gyrcbm/Dropbox/Global_Productivity/manuscript/')
-knitr::opts_knit$set(root.dir = '/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/Global_Productivity/manuscript/')
+knitr::opts_knit$set(root.dir = 'C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/')
+# knitr::opts_knit$set(root.dir = '/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/Global_Productivity/manuscript/')
 ```
@@
 hypothesis_table <- read.csv("hypothesis_table.csv", stringsAsFactors = FALSE, check.names = FALSE)
-kable(hypothesis_table, booktabs = TRUE, caption = "Summary of research questions, corresponding hypotheses, and results.'yes' indicates the hypothesis is supported, 'no' indicates the hypothesis is not supported, and 'n.s.' indicates that the relationship is not signficant.") %>%
+hypothesis_table[1,2] <- paste0(footnote_marker_number(1)," , ",footnote_marker_number(2)," , ", footnote_marker_number(3))
+hypothesis_table[2,2] <- paste0(footnote_marker_number(4))
+hypothesis_table[3,2] <- paste0(footnote_marker_number(1)," , ",footnote_marker_number(4))
+hypothesis_table[4,2] <- paste0(footnote_marker_number(5))
+hypothesis_table[10,2] <- paste0(footnote_marker_number(6)," , ",footnote_marker_number(7)," , ", footnote_marker_number(8))
+hypothesis_table[11,2] <- paste0(footnote_marker_number(7)," , ", footnote_marker_number(8))
+hypothesis_table[12,2] <- paste0(footnote_marker_number(8))
+kable(hypothesis_table, booktabs = TRUE, caption = "Summary of research questions, corresponding hypotheses, and results.
+Statistically signficant support for/ rejection of hypotheses is indicated with 'yes'/'no', parentheses indicate partial overall support/rejection of hypotheses across all FACF, and '-' indicates no significant relationship.", escape = F) %>%
   kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 12) %>%
@@
   column_spec(3:13, width = "1.8cm") %>%
-  kableExtra::landscape()
+  kableExtra::landscape() %>%
+  kableExtra::footnote(number = c("Luyssaert et al.
+(2007)", "Gillman et al.
+(2015)", "Simova and Storch (2017)", "Schuur (2003)", "Taylor et al.
+(2016)", "Malhi (2012)", "Michaletz et al.
+(2014)", "Chu et al.
+(2016)"), footnote_as_chunk = TRUE)
 ```
@@
 | $BNPP_{fine root}$ | Fine root component of $BNPP$ | fine roots | measurements combined one or more of the following: soil cores, minirhizotrons, turnover estimates,  root ingrowth cores | 88 | 41 |
+| $R_{auto}$ | Autotrophic Respiration | foliage, stem, and root | chamber measurements of component gas exchange | 22 | 13 |
 | $R_{root}$ | Root Respiration | coarse and fine roots | partitioning of total soil respiration (e.g., through root exclusion), scaling of root gas exchange | 64 | 26 |
-| $R_{auto}$ | Autotrophic Respiration | foliage, stem, and root | chamber measurements of component gas exchange | 22 | 13 |
 

### 2020-02-16 manuscript/appendix.Rmd
working on Appendix, folder cleanup

@@
 \newpage
-#Table S1. Climate variable definitions and sources
+# Table S1. Climate variable definitions and sources
 ```{r eval = TRUE, echo=FALSE, warning=FALSE}
@@
 library(kableExtra)
-S1 <- read.csv("climate_variables.csv", stringsAsFactors = FALSE, check.names = FALSE)
+S1 <- read.csv("tables_figures/climate_variables.csv", stringsAsFactors = FALSE, check.names = FALSE)
 kable(S1, format = "latex", booktabs = TRUE, caption = "", escape = F)%>%
@@
 ```
+
+\newpage
+# Table S_. Comparison of growing season length and mean annual temperature as predictors of FACF.
+see [issue #58](https://github.com/forc-db/Global_Productivity/issues/58).
+
+\newpage
+# Figure S1. 
+![](tables_figures/distribution_all_samples.png)

### 2020-02-17 scripts/Maps_script.R
Update Maps_script.R

@@ -4,7 +4,7 @@
 rm(list = ls())
 
 # Set working directory as ForC main folder ####
-setwd("C:/Users/gyrcbm/Dropbox/ForC")
+setwd("C:/Users/becky/Dropbox (Smithsonian)/GitHub/ForC")
 
 # Load libaries ####
 library(lme4)
@@ -122,7 +122,7 @@ ForC_simplified <- ForC_simplified[ages.to.keep & fixed.no.na, ]
 lon <- ForC_simplified$lon
 lat <- ForC_simplified$lat
 
-col.sym <- read.csv("C:/Users/gyrcbm/Dropbox/Global_Productivity/raw.data/colsym.csv", stringsAsFactors = F)
+col.sym <- read.csv("C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/raw.data/colsym.csv", stringsAsFactors = F)
 
 ForC_simplified <- merge(ForC_simplified, col.sym[, c("variable.name", "col", "sym")], by = "variable.name")
 ForC_simplified$col <- as.character(ForC_simplified$col)
@@ -130,12 +130,12 @@ ForC_simplified$sym <- as.character(ForC_simplified$sym)
 
 ForC_simplified$variable.name <- factor(ForC_simplified$variable.name, levels = c("GPP", "NPP", "ANPP", "ANPP_woody_stem", "ANPP_foliage", "BNPP_root", "BNPP_root_fine", "R_auto", "R_auto_root"))
 
-png(file = paste0("C:/Users/gyrcbm/Dropbox/Global_Productivity/results/figures/final_figures/maps/distribution_all_samples.png"), width = 2255, height = 2000, units = "px", res = 300)
+png(file = paste0("C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/figures/final_figures/maps/distribution_all_samples.png"), width = 2000, height = 3000, units = "px", res = 300)
 
 mapWorld <- borders("world", colour="gray50", fill="gray50")
 
 mp <- ggplot(data = ForC_simplified) + mapWorld + aes(x=lon, y=lat)
-mp <- mp+ geom_point(aes(color = variable.name, shape = variable.name), size=1) + facet_wrap(vars(variable.name))
+mp <- mp+ geom_point(aes(color = variable.name, shape = variable.name), size=1) + facet_wrap(vars(variable.name), ncol = 2) + theme(legend.position = "none")
[…]

### 2020-02-17 scripts/compare_mat_and_growing_season.R
updated AIC mat/growing season table to include Rsq values

@@ -134,6 +134,9 @@ for (response.v in response.variables){
   aictab$Variable <- response.v
   
   
+  Rsq <- c((r.squaredGLMM(mod.mat))[1], (r.squaredGLMM(mod.growing))[1], (r.squaredGLMM(mod))[1])
+  aictab <- cbind(aictab, Rsq)
+  
   all.aictab <- rbind(all.aictab, aictab)
   # all.results <- all.results <- rbind(all.results, results)
  

### 2020-02-17 manuscript/appendix.Rmd
working on appendix & repo organization
appendix
working on appendix

@@
 author: "Rebecca Banbury Morgan, Valentine Herrmann, Norbert Kunert, Ben Bond-Lamberty, Helene C. Muller-Landau, Kristina J. Anderson-Teixeira"
-toc: true
-toc_depth: 1
 output: 
@@
                   \captionsetup[figure]{font=footnotesize}
+                  \captionsetup[table]{labelformat=empty}
+                  \captionsetup[figure]{labelformat=empty}
 ---
@@
 
+\listoftables
+\listoffigures
+
 \newpage
-# Table S1. Climate variable definitions and sources
 ```{r eval = TRUE, echo=FALSE, warning=FALSE}
@@
 S1 <- read.csv("tables_figures/climate_variables.csv", stringsAsFactors = FALSE, check.names = FALSE)
-kable(S1, format = "latex", booktabs = TRUE, caption = "", escape = F)%>%
+kable(S1, format = "latex", booktabs = TRUE, caption = "Table S1. Climate variable definitions and sources", escape = F)%>%
   kable_styling(latex_options = c("scale_down", "hold_position")) %>%
@@
 \newpage
-# Figure S1. 
-![](tables_figures/distribution_all_samples.png)
+![Figure S1. Geographic distribution of samples for each of the nine FACF analyzed here.](tables_figures/distribution_all_samples.png)
+
+\newpage
+![Figure S2. Correlations among latitude and climate variables](tables_figures/climate_regressions.png)

### 2020-02-17 manuscript/appendix.Rmd
added figures to SI
issue #59

@@
 header-includes:  \usepackage{float}
-                  \usepackage{caption}
-                  \captionsetup[table]{font=footnotesize}
-                  \captionsetup[figure]{font=footnotesize}
-                  \captionsetup[table]{labelformat=empty}
-                  \captionsetup[figure]{labelformat=empty}
+                  \usepackage[font=normalsize,format=plain,justification=raggedright, labelformat=empty]{caption}
 ---
@@
 ![Figure S2. Correlations among latitude and climate variables](tables_figures/climate_regressions.png)
+
+\newpage
+![Figure S3. Ratios among FACF as a function of latitude and climate variables](tables_figures/ratio_grid_plots.png)
+
+\newpage
+![Figure S4. Individual plots of FACF in relation to mean annual climate, part 1.](tables_figures/grid_plots_climate1.png)
+
+\newpage
+![Figure S5. Individual plots of FACF in relation to mean annual climate, part 2.](tables_figures/grid_plots_climate2.png)
+
+\newpage
+![Figure S6. Individual plots of FACF in relation to mean climate seasonality, part 1.](tables_figures/grid_plots_seasonality3.png)
+
+\newpage
+![Figure S7. Individual plots of FACF in relation to mean climate seasonality, part 2.](tables_figures/grid_plots_seasonality4.png)
+
+\newpage
+![Figure S8. Growing season length-standardized FACF in relation to mean growing season climate, part 1.](tables_figures/gridded_growing_season1.png)
+
+\newpage
+![Figure S9. Growing season length-standardized FACF in relation to mean growing season climate, part 2.](tables_figures/gridded_growing_season2.png)

### 2020-02-17 scripts/final_model_no_weighting.R
Update final_model_no_weighting.R

@@ -894,8 +894,8 @@ for (age in ages){
   if (age %in% "age.greater.than.200") ages.to.keep <- ForC_simplified$stand.age >= 200 & !is.na(ForC_simplified$stand.age)
   
   
-  if(type == "climate") png(file = paste0("C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/figures/final_figures/supporting_information/grid_plots_", type, number, ".png"), width = 3000, height = 2000, units = "px", res = 300)
-  if(type == "seasonality") png(file = paste0("C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/figures/final_figures/supporting_information/grid_plots_", type, number, ".png"), width = 2000, height = 2000, units = "px", res = 300)
+  if(type == "climate") png(file = paste0("C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/tables_figures/grid_plots_", type, number, ".png"), width = 3000, height = 2000, units = "px", res = 300)
+  if(type == "seasonality") png(file = paste0("C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/tables_figures/grid_plots_", type, number, ".png"), width = 2000, height = 3000, units = "px", res = 300)
   
   
   if(number == 1) par(mfcol = c(5,5), mar = c(2.5,2,2,2), oma = c(2,8,0,0), xpd = T)

### 2020-02-17 scripts/ratio_regression_script.R
updated and edited ratio grid plots

@@ -98,19 +98,16 @@ all.aictab <- NULL
 
 fixed.variables <- c("lat", "mat", "map", "TempSeasonality")
 
-# set1 <- c("GPP", "ANPP", "ANPP_foliage", "ANPP_foliage", "ANPP_woody_stem", "ANPP", "BNPP_root")
-# set2 <- c("NPP", "BNPP_root", "ANPP_woody_stem", "NPP", "NPP", "NPP", "NPP")
+set1 <- c("GPP", "ANPP", "ANPP_foliage", "ANPP_foliage", "ANPP_woody_stem", "ANPP", "BNPP_root")
+set2 <- c("NPP", "BNPP_root", "ANPP_woody_stem", "NPP", "NPP", "NPP", "NPP")
 
-set1 <- c("GPP", "GPP", "GPP", "GPP", "GPP", "GPP", "GPP")
-set2 <- c("NPP", "ANPP", "ANPP_foliage", "ANPP_woody_stem", "BNPP_root", "BNPP_root_fine", "R_auto")
+# set1 <- c("GPP", "GPP", "GPP", "GPP", "GPP", "GPP", "GPP")
+# set2 <- c("NPP", "ANPP", "ANPP_foliage", "ANPP_woody_stem", "BNPP_root", "BNPP_root_fine", "R_auto")
 
-# set1 <- c("ANPP_1", "ANPP_1", "BNPP_root", "ANPP_2", "ANPP_2", "ANPP", "ANPP")
-# set2 <- c("BNPP_root", "NPP_1", "NPP_1", "BNPP_root", "NPP_1", "BNPP_root", "NPP_1")
-
-png(file = paste0("C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/figures/final_figures/ratio_regressions/ratio_grid_plots_1.png"), width = 3000, height = 3500, units = "px", res = 300)
+png(file = paste0("C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/tables_figures/ratio_grid_plots.png"), width = 1800, height = 3000, units = "px", res = 300)
 
 
-par(mfrow = c(7,4), mar = c(2,2,2,2), oma = c(5,14,2,0))
+par(mfrow = c(7,4), mar = c(2,2,2,2), oma = c(5,3,5,0))
 panel.number <- 1
 for (i in seq(along = set1)){
   # for (j in seq(along = set2)){
@@ -144,7 +141,7 @@ for (i in seq(along = set1)){
         
[…]

### 2020-02-17 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@
 
-Of the climate variables tested, annual wet days, aridity, cloud cover, mean diurnal temperature range, precipitation seasonality, maximum vapour pressure deficit and water stress months were poor or non-significant explainers of variation in FACF, explaining less than 20% of the variation in each of the carbon fluxes.
+Of the climate variables tested, annual wet days, aridity, cloud cover, precipitation seasonality, maximum vapour pressure deficit and water stress months were poor or non-significant explainers of variation in FACF, explaining less than 20% of the variation in each of the carbon fluxes.
 

### 2020-02-17 manuscript/appendix.Rmd
Update appendix.Rmd

--- 2020-02-16 manuscript/appendix.Rmd
@@
                   \captionsetup[figure]{font=footnotesize}
+                  \captionsetup[figure]{labelformat=empty}
+                  \usepackage{pdflscape}
+                  \newcommand{\blandscape}{\begin{landscape}}
+                  \newcommand{\elandscape}{\end{landscape}}
+
 ---
@@
 ```{r setup, include = FALSE}
-# knitr::opts_knit$set(root.dir = 'C:/Users/gyrcbm/Dropbox/Global_Productivity/manuscript/')
-knitr::opts_knit$set(root.dir = '/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/Global_Productivity/manuscript/')
+knitr::opts_knit$set(root.dir = 'C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/')
+# knitr::opts_knit$set(root.dir = '/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/Global_Productivity/manuscript/')
+```
+
+```{r, include=F}
+library(captioner)
+tables <- captioner(prefix = "Table S", auto_space = FALSE)
+figures <- captioner(prefix = "Figure S", auto_space = FALSE)
+
+figures("Maps", "Maps showing distribution of samples for the nine FACF analyzed here.", display=FALSE)
 ```
@@
 \newpage
-# Table S_. Comparison of growing season length and mean annual temperature as predictors of FACF.
-see [issue #58](https://github.com/forc-db/Global_Productivity/issues/58).
+# Table S2. Comparison of growing season length and mean annual temperature as predictors of FACF.
+```{r eval = TRUE, echo=FALSE, warning=FALSE}
+library(knitr)
+library(kableExtra)
+S1 <- read.csv("tables_figures/mat_growing_season.csv", stringsAsFactors = FALSE, check.names = FALSE)
+kable(S1, format = "latex", booktabs = TRUE, caption = "", escape = F)%>%
+  kable_styling(latex_options = c("scale_down", "hold_position")) %>%
+  column_spec(1, width = "5cm") %>%
+  column_spec(2, width = "2cm") %>%
+  column_spec(3:4, width = "7cm") %>%
+  pack_rows("GPP", 1, 3, latex_gap_space = "1em", colnum = 1, hline_before = FALSE) %>%
+  pack_rows("NPP", 4, 6, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("ANPP", 7, 9, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("ANPP woody stem", 10, 12, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("ANPP foliage", 13, 15, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("BNPP root", 16, 18, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("BNPP fine root", 19, 21, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("Autotrophic respiration", 22, 24, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("Root respiration", 25, 27, latex_gap_space = "1em", colnum = 1, hline_before = TRUE)
+```
 
 \newpage
-# Figure S1. 
-![](tables_figures/distribution_all_samples.png)
+# Figure S1. Maps showing distribution of samples for the nine FACF analyzed here.
+```{r echo=FALSE, fig.cap = figures("Maps"), out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("tables_figures/distribution_all_samples.png")
+```
+
+\newpage
+
+\blandscape
+
+# Figure S2.
+
+```{r echo=FALSE, fig.cap = "Individual plots of FACF in relation to mean annual climate, part 1.", out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("tables_figures/grid_plots_climate1.png")
+```
+```{r echo=FALSE, fig.cap = "Individual plots of FACF in relation to mean annual climate, part 2.", out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("tables_figures/grid_plots_climate2.png")
+```
+\elandscape
+
+\newpage
+# Figure S3.
+
+```{r echo=FALSE, fig.cap = "Individual plots of FACF in relation to mean climate seasonality, part 1.", out.width ='100%', fig.pos='H'}
+
+  knitr::include_graphics("tables_figures/grid_plots_seasonality3.png")
+```
+```{r echo=FALSE, fig.cap = "Individual plots of FACF in relation to mean climate seasonality, part 2.", out.width ='100%', fig.pos='H'}
+
+  knitr::include_graphics("tables_figures/grid_plots_seasonality4.png")
+```
+
+\newpage
+\blandscape
+
+# Figure S_. Correlations among latitude and climate variables
+```{r echo=FALSE, fig.cap = "Correlations among latitude and climate variables. Variable names and units given in Table S1", out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("tables_figures/climate_regressions.png")
+```
+\elandscape

### 2020-02-17 scripts/extract_climate_data_growing_season.R
updated scripts (cleaned SI figures)

@@ -696,7 +696,7 @@ number <-1
   ### mature forests only ####
 for (response.variables.group in response.variables.groups){
 
-png(file = paste0("C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/figures/final_figures/supporting_information/gridded_growing_season", number, ".png"), width = 2500, height = 3000, units = "px", res = 300)
+png(file = paste0("C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/tables_figures/gridded_growing_season", number, ".png"), width = 2500, height = 3000, units = "px", res = 300)
 
 if(number == 1) par(mfcol = c(5,4), mar = c(2.5,2,2,2.5), oma = c(8,8,0,2), xpd = T)
 if(number == 2) par(mfcol = c(4,4), mar = c(2.5,2,2,2.5), oma = c(8,8,0,2), xpd = T)
@@ -710,7 +710,7 @@ ages.to.keep <- ForC_simplified$stand.age >= 100 & !is.na(ForC_simplified$stand.
       
       fixed.v.info <- read.csv("C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/raw.data/fixedv_data.csv", stringsAsFactors = F)
       
-      xaxis <- fixed.v.info$xaxis[which(fixed.v.info$fixed.v %in% fixed.v)]
+      xaxis <- fixed.v.info$xaxis_simple[which(fixed.v.info$fixed.v %in% fixed.v)]
       
       ###subset ForC
     
@@ -793,10 +793,13 @@ ages.to.keep <- ForC_simplified$stand.age >= 100 & !is.na(ForC_simplified$stand.
         altitude = FALSE
         
         
-        if(fixed.v == "pet")mtext(paste0("(", letters[panel.number], ") ", response.v), side = 3, line = 0.5, adj = 0.05, cex = 0.6)
+        if(fixed.v == "pet")mtext(paste0(response.v), side = 3, line = 0.5, adj = 0.05, cex = 0.6)
         panel.number <- panel.number +1 
-        if(response.v %in% c("ANPP_foliage", "R_auto_root")) mtext(side = 1, line = 2.5, text = eval(parse(text = xaxis)), cex = 0.6, padj = 0.5, adj = 0.5)
-        mtext(side = 2, line = 3,  text = expression("Monthly mean productivity during growing season Mg C "~ha^-1~yr^-1), outer = T) 
[…]

### 2020-02-17 scripts/final_model_no_weighting.R

@@ -898,10 +898,10 @@ for (age in ages){
   if(type == "seasonality") png(file = paste0("C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/tables_figures/grid_plots_", type, number, ".png"), width = 2000, height = 3000, units = "px", res = 300)
   
   
-  if(number == 1) par(mfcol = c(5,5), mar = c(2.5,2,2,2), oma = c(2,8,0,0), xpd = T)
-  if(number == 2) par(mfcol = c(4,5), mar = c(2.5,2,2,2), oma = c(2,8,0,0), xpd = T)
-  if(number == 3) par(mfcol = c(5,3), mar = c(2.5,2,2,2), oma = c(2,8,0,0), xpd = T)
-  if(number == 4) par(mfcol = c(4,3), mar = c(2.5,2,2,2), oma = c(2,8,0,0), xpd = T)
+  if(number == 1) par(mfcol = c(5,5), mar = c(2.5,3,2,2), oma = c(2,8,0,0), xpd = T)
+  if(number == 2) par(mfcol = c(4,5), mar = c(2.5,3,2,2), oma = c(2,8,0,0), xpd = T)
+  if(number == 3) par(mfcol = c(5,3), mar = c(2.5,3,2,2), oma = c(2,8,0,0), xpd = T)
+  if(number == 4) par(mfcol = c(4,3), mar = c(2.5,3,2,2), oma = c(2,8,0,0), xpd = T)
   number = number + 1
   pannel.nb <- 1
   
@@ -917,7 +917,7 @@ for (age in ages){
     
     fixed.v.info <- read.csv("C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/raw.data/fixedv_data.csv", stringsAsFactors = F)
     
-    xaxis <- fixed.v.info$xaxis[which(fixed.v.info$fixed.v %in% fixed.v)]
+    xaxis <- fixed.v.info$xaxis_simple[which(fixed.v.info$fixed.v %in% fixed.v)]
     
     ###subset ForC
     
@@ -1021,12 +1021,12 @@ for (age in ages){
       results <- data.frame(response = response.v, fixed = fixed.v, random = "geographic.area/plot.name", Age.filter = age, significant = significant.effect, p.value = significance, sample.size = sample.size, Rsq = Rsq)
       
       all.results <- rbind(all.results, results)
[…]

### 2020-02-17 scripts/ratio_regression_script.R
added y axis label to ratio plots

@@ -107,7 +107,7 @@ set2 <- c("NPP", "BNPP_root", "ANPP_woody_stem", "NPP", "NPP", "NPP", "NPP")
 png(file = paste0("C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/tables_figures/ratio_grid_plots.png"), width = 1800, height = 3000, units = "px", res = 300)
 
 
-par(mfrow = c(7,4), mar = c(2,2,2,2), oma = c(5,3,5,0))
+par(mfrow = c(7,4), mar = c(2,2,2,2), oma = c(5,4,5,0))
 panel.number <- 1
 for (i in seq(along = set1)){
   # for (j in seq(along = set2)){

### 2020-02-17 manuscript/appendix.Rmd
updated appendix pdf

--- 2020-02-17 manuscript/appendix.Rmd
@@
 author: "Rebecca Banbury Morgan, Valentine Herrmann, Norbert Kunert, Ben Bond-Lamberty, Helene C. Muller-Landau, Kristina J. Anderson-Teixeira"
+toc: true
+toc_depth: 1
 output: 
@@
                   \captionsetup[figure]{font=footnotesize}
+                  \captionsetup[figure]{labelformat=empty}
                   \captionsetup[table]{labelformat=empty}
-                  \captionsetup[figure]{labelformat=empty}
+                  \usepackage{pdflscape}
+                  \newcommand{\blandscape}{\begin{landscape}}
+                  \newcommand{\elandscape}{\end{landscape}}
+
 ---
@@
 S1 <- read.csv("tables_figures/climate_variables.csv", stringsAsFactors = FALSE, check.names = FALSE)
-kable(S1, format = "latex", booktabs = TRUE, caption = "Table S1. Climate variable definitions and sources", escape = F)%>%
+kable(S1, format = "latex", booktabs = TRUE, caption = "Table S1. Climate variable definitions, sources, and abbreviations", escape = F)%>%
   kable_styling(latex_options = c("scale_down", "hold_position")) %>%
@@
 \newpage
-# Table S_. Comparison of growing season length and mean annual temperature as predictors of FACF.
-see [issue #58](https://github.com/forc-db/Global_Productivity/issues/58).
+```{r eval = TRUE, echo=FALSE, warning=FALSE}
+library(knitr)
+library(kableExtra)
+S1 <- read.csv("tables_figures/mat_growing_season.csv", stringsAsFactors = FALSE, check.names = FALSE)
+kable(S1, format = "latex", booktabs = TRUE, caption = "Table S2. Comparison of growing season length and mean annual temperature as predictors of FACF", escape = F)%>%
+  kable_styling(latex_options = c("scale_down", "hold_position")) %>%
+  column_spec(1, width = "5cm") %>%
+  column_spec(2, width = "2cm") %>%
+  column_spec(3:4, width = "7cm") %>%
+  pack_rows("GPP", 1, 3, latex_gap_space = "1em", colnum = 1, hline_before = FALSE) %>%
+  pack_rows("NPP", 4, 6, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("ANPP", 7, 9, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("ANPP woody stem", 10, 12, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("ANPP foliage", 13, 15, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("BNPP root", 16, 18, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("BNPP fine root", 19, 21, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("Autotrophic respiration", 22, 24, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("Root respiration", 25, 27, latex_gap_space = "1em", colnum = 1, hline_before = TRUE)
+```
 
 \newpage
-![Figure S1. Geographic distribution of samples for each of the nine FACF analyzed here.](tables_figures/distribution_all_samples.png)
+```{r echo=FALSE, fig.cap = "Figure S1: Maps showing distribution of samples for the nine FACF analyzed here.", out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("tables_figures/distribution_all_samples.png")
+```
 
 \newpage
-![Figure S2. Correlations among latitude and climate variables](tables_figures/climate_regressions.png)
+
+\blandscape
+
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S2: Correlations among latitude and climate variables. Variable names and units given in Table S1", out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("tables_figures/climate_regressions.png")
+```
+\elandscape
+
+\blandscape
+
+```{r echo=FALSE, fig.cap = "Figure S2: Individual plots of FACF in relation to mean annual climate, part 1.", out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("tables_figures/grid_plots_climate1.png")
+```
+
+\newpage
+```{r echo=FALSE, fig.cap = "Figure S3: Individual plots of FACF in relation to mean annual climate, part 2.", out.width='100%', fig.pos='H'}
+
+  knitr::include_graphics("tables_figures/grid_plots_climate2.png")
+```
+\elandscape
+
+```{r echo=FALSE, fig.cap = "Figure S4: Individual plots of FACF in relation to mean climate seasonality, part 1.", out.width ='100%', fig.pos='H'}
+
+  knitr::include_graphics("tables_figures/grid_plots_seasonality3.png")
+```
+
+\newpage
+```{r echo=FALSE, fig.cap = "Figure S5: Individual plots of FACF in relation to mean climate seasonality, part 2.", out.width ='100%', fig.pos='H'}
+
+  knitr::include_graphics("tables_figures/grid_plots_seasonality4.png")
+```
+
+\newpage
+```{r echo=FALSE, fig.cap = "Figure S6: Growing season length-standardized FACF in relation to mean growing season climate, part 1.", out.width ='100%', fig.pos='H'}
+
+  knitr::include_graphics("tables_figures/gridded_growing_season1.png")
+```
+
+\newpage
+```{r echo=FALSE, fig.cap = "Figure S7: Growing season length-standardized FACF in relation to mean growing season climate, part 2.", out.width ='100%', fig.pos='H'}
+
+  knitr::include_graphics("tables_figures/gridded_growing_season2.png")
+```
+
+\newpage
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S8: Ratios among FACF as a function of latitude and climate variables", out.width ='100%', fig.pos='H'}
+
+  knitr::include_graphics("tables_figures/ratio_grid_plots.png")
+```

### 2020-02-17 scripts/ratio_regression_script.R
Update ratio_regression_script.R

@@ -241,6 +241,6 @@ for (i in seq(along = set1)){
       
 }#}}
 
-mtext(side = 2, line = 3, text = "Ratio", outer = T)
+mtext(side = 2, line = 2.5, text = "Ratio", outer = T)
 legend(x = -100, y = 15.5, lty = c(1,2), legend = c("Significant effect", "No significant effect"), xpd = NA)
 dev.off()

### 2020-02-17 manuscript/manuscript draft.Rmd
updated hypothesis table and corrected citations in manuscript draft

@@
 **FACF can be influenced by many other factors as well**, which often act across a range of scales, and may show interactive effects with each other [@cleveland_relationships_2011].
-On a local scale, stand age [@litton_carbon_2007; @gillman_latitude_2015], biodiversity [@liang_positive_2016], management [@simova_enigma_2017]; nutrient availability [@aragao_above-_2009]; and altitude [@girardin_net_2010; @malhi_variation_2017] all impact FACF.
+On a local scale, stand age [@litton_carbon_2007; @gillman_latitude_2015], biodiversity [@liang_positive_2016], management [@simova_enigma_2017]; nutrient availability [@aragao_above_2009]; and altitude [@girardin_net_2010; @malhi_variation_2017] all impact FACF.
 On a global scale, we expect that FACF are most strongly influenced by broad climatic gradients.
@@
 
-**The recent development of a global forest carbon database synthesizing multiple variables and including records of stand history (ForC; [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc:_2018]) opens up the possibility for a standardized analysis of global scale variation** in multiple FACF and the principle climatic drivers of these patterns.
+**The recent development of a global forest carbon database synthesizing multiple variables and including records of stand history (ForC; [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018]) opens up the possibility for a standardized analysis of global scale variation** in multiple FACF and the principle climatic drivers of these patterns.
 In order to approach this broad topic, we simplify the major gaps in our knowledge to five broad questions and corresponding hypotheses (Table 1).
@@
 
-Analyses were conducted on data contained in the open-access ForC database [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc:_2018].
+Analyses were conducted on data contained in the open-access ForC database [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018].
 This database contains records of field-based measurements of forest carbon stocks and annual fluxes, compiled from original publications and existing data compilations and databases.
 Associated data, such as stand age, measurement methodologies, and disturbance history, are also included.
-The database was significantly expanded since the publication of [@anderson-teixeira_forc:_2018] through integration with the Global Soil Respiration Database [@bond-lamberty_global_2010].
+The database was significantly expanded since the publication of [@anderson-teixeira_forc_2018] through integration with the Global Soil Respiration Database [@bond-lamberty_global_2010].
 Additional targeted literature searches were conducted to identify any further available data on FACF, with particular focus on mature forests in temperate and boreal regions.
@@
 Grazed sites were retained.
-There is evidence that stand age influences patterns of FACF and carbon allocation in forest ecosystems, and can confound relationships between latitude and primary productivity [@de_lucia_forest_2007; @gillman_latitude_2015].
+There is evidence that stand age influences patterns of FACF and carbon allocation in forest ecosystems, and can confound relationships between latitude and primary productivity [@delucia_forest_2007; @gillman_latitude_2015].
 To reduce any biasing effects of stand age, we included only stands of known age $\ge$ 100 years and those described by terms such as "mature", "intact", or "old-growth".
@@
 Based on the geographic co-ordinates for each site, data on a further 11 climate variables was extracted from five open-access climate datasets: WorldClim [@hijmans_very_2005], WorldClim2 [@fick_worldclim_2017], the Climate Research Unit (CRU) time-series dataset v.
-4.03 [@harris_updated_2014], the Global Aridity Index and Potential Evapotranspiration Climate Database [@trabucco_global_2018], and TerraClimate [@abatzoglou_terraclimate_2018] (see Supplementary Information S1 for details of climate variables).
+4.03 [@harris_updated_2014], the Global Aridity Index and Potential Evapotranspiration Climate Database [@trabucco_global_2019], and TerraClimate [@abatzoglou_terraclimate_2018] (see Supplementary Information S1 for details of climate variables).
 Where site-level data was missing for mean annual temperature and/or mean annual precipitation, data was extracted from the WorldClim dataset.
@@
 ).
-In addition, because C allocation varies with stand age [@de_lucia_forest_2007] (**See Nobby's comment in manuscript-draft_NK.pdf**), age differences may introduce systematic biases into analyses of FACF across latitude or global climatic gradients.
+In addition, because C allocation varies with stand age [@delucia_forest_2007] (**See Nobby's comment in manuscript-draft_NK.pdf**), age differences may introduce systematic biases into analyses of FACF across latitude or global climatic gradients.
 For example, woody productivity tends to be higher in rapidly aggrading secondary stands than in old-growth forests, where proportionally more C is allocated to respiration (** Nobby AFM paper: Understanding the controls over forest carbon use efficiency on small spatial scales: Effects of forest disturbance and tree diversity**) [*purpose for respiration/ other compenents (**See Nobby's comment in manuscript-draft_NK.pdf**)].
@@
 Of particular interest and significance are the relationships amongst $GPP$, net primary productivity ($NPP$ and its components, particularly $ANPP_{woody-stem}$), and respiration ($R_{auto}$ and components).
-There have been suggestions that tropical forests tend to have low carbon use efficiency ($CUE$= $NPP$/$GPP$=($GPP$-$R_{auto}$/$GPP$), which are based on observations of low $CUE$ in old-growth tropical forests relative to (mostly younger) extratropical forests [@de_lucia_forest_2007; [Malhi 2012](https://www.researchgate.net/publication/227837530_The_productivity_metabolism_and_carbon_cycle_of_tropical_forest_vegetation); @anderson-teixeira_carbon_2016], but our analysis suggests that these low values might more appropriately be attributed to the fact that these forests are old than to their tropical climate.
-Indeed, $CUE$ is known to decline with forest age ([@de_lucia_forest_2007]; [Collalti & Prentice 2019](https://academic.oup.com/treephys/article-abstract/39/8/1473/5423351?redirectedFrom=fulltext)) but appears to be roughly independent of $GPP$ [@litton_carbon_2007].
+There have been suggestions that tropical forests tend to have low carbon use efficiency ($CUE$= $NPP$/$GPP$=($GPP$-$R_{auto}$/$GPP$), which are based on observations of low $CUE$ in old-growth tropical forests relative to (mostly younger) extratropical forests [@delucia_forest_2007; @malhi_productivity_2012; @anderson-teixeira_carbon_2016], but our analysis suggests that these low values might more appropriately be attributed to the fact that these forests are old than to their tropical climate.
+Indeed, $CUE$ is known to decline with forest age [@delucia_forest_2007; @collalti_is_2019], but appears to be roughly independent of $GPP$ [@litton_carbon_2007].
 Among our sites with relevant data, there is indication that CUE or $ANPP_{woody-stem}$/$GPP$ increase with latitude (**some SI table**).

### 2020-02-18 scripts/extract_climate_data_growing_season.R
updated scripts and graphs to include R2 values on subplots

@@ -783,7 +783,7 @@ ages.to.keep <- ForC_simplified$stand.age >= 100 & !is.na(ForC_simplified$stand.
         significance <- signif(significance, digits=4)
         
         Rsq <- as.data.frame(r.squaredGLMM(mod.full))
-        Rsq <- signif(Rsq, digits=4)
+        Rsq <- signif(Rsq, digits=4)[1]
         ##legend2 <- paste(response.v, "r-squared = ", Rsq[1], "p-value = ", significance)
         ##mtext(side = 3, line = -which(response.variables %in% response.v), text = legend2, adj = 0.9, col = plasma(10)[col], cex = 0.5)
         
@@ -792,7 +792,7 @@ ages.to.keep <- ForC_simplified$stand.age >= 100 & !is.na(ForC_simplified$stand.
         date = Sys.Date()
         altitude = FALSE
         
-        
+        if(significant.effect) mtext(paste("R-sq =", Rsq), side = 3, line = -1.5, adj = 0.1, cex = 0.7)
         if(fixed.v == "pet")mtext(paste0(response.v), side = 3, line = 0.5, adj = 0.05, cex = 0.6)
         panel.number <- panel.number +1 
         if(response.v %in% c("ANPP_foliage", "R_auto_root")) mtext(side = 1, line = 2.5, text = xaxis, cex = 0.6, padj = 0.5, adj = 0.5)

### 2020-02-18 scripts/final_model_no_weighting.R

@@ -894,7 +894,7 @@ for (age in ages){
   if (age %in% "age.greater.than.200") ages.to.keep <- ForC_simplified$stand.age >= 200 & !is.na(ForC_simplified$stand.age)
   
   
-  if(type == "climate") png(file = paste0("C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/tables_figures/grid_plots_", type, number, ".png"), width = 3000, height = 2000, units = "px", res = 300)
+  if(type == "climate") png(file = paste0("C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/tables_figures/grid_plots_", type, number, ".png"), width = 3000, height = 2250, units = "px", res = 300)
   if(type == "seasonality") png(file = paste0("C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/tables_figures/grid_plots_", type, number, ".png"), width = 2000, height = 3000, units = "px", res = 300)
   
   
@@ -1016,7 +1016,7 @@ for (age in ages){
       significance <- signif(significance, digits=4)
       
       Rsq <- as.data.frame(r.squaredGLMM(mod.full))
-      Rsq <- signif(Rsq, digits=4)
+      Rsq <- signif(Rsq, digits=4)[1]
       
       results <- data.frame(response = response.v, fixed = fixed.v, random = "geographic.area/plot.name", Age.filter = age, significant = significant.effect, p.value = significance, sample.size = sample.size, Rsq = Rsq)
       
@@ -1024,6 +1024,7 @@ for (age in ages){
       if(fixed.v %in% c("mat", "TempSeasonality"))mtext(paste0(response.v), side = 3, line = 0.5, adj = 0.05, cex = 0.6)
       
       pannel.nb <- pannel.nb +1
+      if(significant.effect) mtext(paste("R-sq =", Rsq), side = 3, line = -1.5, adj = 0.1, cex = 0.6)
       
[…]

### 2020-02-18 manuscript/appendix.Rmd
edited appendix

@@
                   \newcommand{\elandscape}{\end{landscape}}
-
----
 
@@
 kable(S1, format = "latex", booktabs = TRUE, caption = "Table S1. Climate variable definitions, sources, and abbreviations", escape = F)%>%
-  kable_styling(latex_options = c("scale_down", "hold_position")) %>%
+  kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 12) %>%
   column_spec(1, width = "5cm") %>%
-  column_spec(2, width = "2cm") %>%
-  column_spec(3:4, width = "7cm") 
+  column_spec(2:3, width = "2cm") %>%
+  column_spec(4, width = "7cm") 
 ```
@@
 ```
-
-\newpage
 

### 2020-02-18 manuscript/appendix.Rmd
corrected figure numbering

@@
 ```{r setup, include = FALSE}
-knitr::opts_knit$set(root.dir = 'C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/')
+# knitr::opts_knit$set(root.dir = 'C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/')
+knitr::opts_knit$set(root.dir = 'C:/Users/gyrcbm/Dropbox/Global_Productivity/manuscript/')
 # knitr::opts_knit$set(root.dir = '/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/Global_Productivity/manuscript/')
@@
 
-```{r echo=FALSE, fig.cap = "Figure S2: Individual plots of FACF in relation to mean annual climate, part 1.", out.width='100%', fig.pos='H'}
+```{r echo=FALSE, fig.cap = "Figure S3: Individual plots of FACF in relation to mean annual climate, part 1.", out.width='100%', fig.pos='H'}
 
@@
 \newpage
-```{r echo=FALSE, fig.cap = "Figure S3: Individual plots of FACF in relation to mean annual climate, part 2.", out.width='100%', fig.pos='H'}
+```{r echo=FALSE, fig.cap = "Figure S4: Individual plots of FACF in relation to mean annual climate, part 2.", out.width='100%', fig.pos='H'}
 
@@
 
-```{r echo=FALSE, fig.cap = "Figure S4: Individual plots of FACF in relation to mean climate seasonality, part 1.", out.width ='100%', fig.pos='H'}
+```{r echo=FALSE, fig.cap = "Figure S5: Individual plots of FACF in relation to mean climate seasonality, part 1.", out.width ='100%', fig.pos='H'}
 
@@
 \newpage
-```{r echo=FALSE, fig.cap = "Figure S5: Individual plots of FACF in relation to mean climate seasonality, part 2.", out.width ='100%', fig.pos='H'}
+```{r echo=FALSE, fig.cap = "Figure S6: Individual plots of FACF in relation to mean climate seasonality, part 2.", out.width ='100%', fig.pos='H'}
 
@@
 \newpage
-```{r echo=FALSE, fig.cap = "Figure S6: Growing season length-standardized FACF in relation to mean growing season climate, part 1.", out.width ='100%', fig.pos='H'}
+```{r echo=FALSE, fig.cap = "Figure S7: Growing season length-standardized FACF in relation to mean growing season climate, part 1.", out.width ='100%', fig.pos='H'}
 
@@
 \newpage
-```{r echo=FALSE, fig.cap = "Figure S7: Growing season length-standardized FACF in relation to mean growing season climate, part 2.", out.width ='100%', fig.pos='H'}
+```{r echo=FALSE, fig.cap = "Figure S8: Growing season length-standardized FACF in relation to mean growing season climate, part 2.", out.width ='100%', fig.pos='H'}
 
@@
 \newpage
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S8: Ratios among FACF as a function of latitude and climate variables", out.width ='100%', fig.pos='H'}
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S9: Ratios among FACF as a function of latitude and climate variables", out.width ='100%', fig.pos='H'}
 

### 2020-02-18 scripts/interactions_all_combinations.R
Update interactions_all_combinations.R

@@ -4,7 +4,7 @@
 rm(list = ls())
 
 # Set working directory as ForC main folder ####
-setwd("C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/ForC")
+setwd("C:/Users/gyrcbm/Dropbox/ForC")
 
 # Load libaries ####
 library(lme4)
@@ -252,7 +252,7 @@ response.variables <- c("GPP", "NPP", "BNPP_root", "BNPP_root_fine", "ANPP", "AN
           
         }
 
-write.csv(all.results, file = "C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/tables/best_model_outputs/AIC_multivariate_interactions_table.csv", row.names = F)
+write.csv(all.results, file = "C:/Users/gyrcbm/Dropbox/Global_Productivity/results/tables/best_model_outputs/AIC_multivariate_interactions_table.csv", row.names = F)
 
 all.results = NULL
 
@@ -295,4 +295,4 @@ for(response.v in response.variables){
   all.results <- all.results <- rbind(all.results, results)
 }
 
-write.csv(all.results, file = "C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/tables/best_model_outputs/AIC_multivariate_interactions_best.csv", row.names = F)
+write.csv(all.results, file = "C:/Users/gyrcbm/Dropbox/Global_Productivity/results/tables/best_model_outputs/AIC_multivariate_interactions_best.csv", row.names = F)

### 2020-02-19 manuscript/appendix.Rmd
updated appendix

@@
 title: "Supplementary Information: Global patterns of forest autotrophic carbon fluxes"
-author: "Rebecca Banbury Morgan, Valentine Herrmann, Norbert Kunert, Ben Bond-Lamberty, Helene C. Muller-Landau, Kristina J. Anderson-Teixeira"
+author:
+ - Rebecca Banbury Morgan
+ - Valentine Herrmann
+ - Norbert Kunert
+ - Ben Bond-Lamberty
+ - Helene C. Muller-Landau
+ - Kristina J. Anderson-Teixeira
 output: 
@@
                   \newcommand{\elandscape}{\end{landscape}}
+---
 
@@
 ```{r setup, include = FALSE}
-# knitr::opts_knit$set(root.dir = 'C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/')
-knitr::opts_knit$set(root.dir = 'C:/Users/gyrcbm/Dropbox/Global_Productivity/manuscript/')
+knitr::opts_knit$set(root.dir = 'C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/')
 # knitr::opts_knit$set(root.dir = '/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/Global_Productivity/manuscript/')
@@
 \newpage
+
+\blandscape
+
 ```{r eval = TRUE, echo=FALSE, warning=FALSE}
@@
   kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 12) %>%
-  column_spec(1, width = "5cm") %>%
-  column_spec(2:3, width = "2cm") %>%
-  column_spec(4, width = "7cm") 
+  column_spec(1, width = "6cm") %>%
+  column_spec(2, width = "2cm") %>%
+  column_spec(3, width = "7cm") %>%
+  column_spec(4, width = "2cm")
 ```
-
+\elandscape
 \newpage
@@
 \newpage
-```{r echo=FALSE, fig.cap = "Figure S1: Maps showing distribution of samples for the nine FACF analyzed here.", out.width='100%', fig.pos='H'}
+```{r echo=FALSE, fig.cap = "Figure S1: Maps showing distribution of samples for the nine FACF analyzed here.", out.height ='100%', fig.pos='H'}
 
@@
 \blandscape
-
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S2: Correlations among latitude and climate variables. Variable names and units given in Table S1", out.width='100%', fig.pos='H'}
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S2: Correlations among latitude and climate variables. Variable names and units given in Table S1", out.height='95%', fig.pos='H'}
 
@@
 
-```{r echo=FALSE, fig.cap = "Figure S3: Individual plots of FACF in relation to mean annual climate, part 1.", out.width='100%', fig.pos='H'}
+```{r echo=FALSE, fig.cap = "Figure S3: Individual plots of FACF in relation to mean annual climate, part 1.", out.height='95%', fig.pos='H'}
 
@@
 \newpage
-```{r echo=FALSE, fig.cap = "Figure S4: Individual plots of FACF in relation to mean annual climate, part 2.", out.width='100%', fig.pos='H'}
+```{r echo=FALSE, fig.cap = "Figure S4: Individual plots of FACF in relation to mean annual climate, part 2.", out.height='95%', fig.pos='H'}
 
@@
 
-```{r echo=FALSE, fig.cap = "Figure S5: Individual plots of FACF in relation to mean climate seasonality, part 1.", out.width ='100%', fig.pos='H'}
+```{r echo=FALSE, fig.cap = "Figure S5: Individual plots of FACF in relation to mean climate seasonality, part 1.", out.height ='95%', fig.pos='H'}
 

### 2020-02-19 manuscript/appendix.Rmd

@@
 \newpage
+
 ```{r echo=FALSE, fig.cap = "Figure S4: Individual plots of FACF in relation to mean annual climate, part 2.", out.height='95%', fig.pos='H'}

### 2020-02-19 scripts/mat_map_interaction.R
Update mat_map_interaction.R

@@ -12,6 +12,7 @@ library(MuMIn)
 library(AICcmodavg)
 library(piecewiseSEM)
 library(viridis)
+library(mgcv)
 
 # Load data ####
 ForC_simplified <- read.csv("ForC_simplified/ForC_simplified_WorldClim_CRU_refined.csv", stringsAsFactors = F)
@@ -105,12 +106,15 @@ all.response.variables[!all.response.variables %in% unlist(response.variables.gr
 
 all.results <- NULL
 best.results <- NULL
+outputs <- NULL
 
 effects <- c("mat", "map", "(1|geographic.area/plot.name)")
 pannel.nb <- 1
 
 
 response.variables <- c("GPP", "NPP", "ANPP", "ANPP_woody_stem", "ANPP_foliage", "BNPP_root", "BNPP_root_fine", "R_auto", "R_auto_root")
+response.variables <- "GPP"
+
 
 # png(file = paste0("C:/Users/gyrcbm/Dropbox/Global_Productivity/results/figures/final_figures/interactions/mat_map_interaction.png"), width = 2255, height = 2000, units = "px", res = 300)
 par(mfrow = c(3,3), mar = c(1,0,0,2), oma = c(5,8,2,0), xpd = T)
@@ -129,7 +133,34 @@ for (response.v in response.variables){
   df <- ForC_simplified[rows.with.response & fixed.no.na, ]
   df$masl <- (df$masl/1000)
   
+  mod.single <- lmer(mean ~ mat + (1|geographic.area/plot.name), data = df, REML = F)
+  mod.add <- lmer(mean ~ mat + map + (1|geographic.area/plot.name), data = df, REML = F)
+  mod.int <- lmer(mean ~ mat * map + (1|geographic.area/plot.name), data = df, REML = F)
+  aictab <- aictab(list(mod.single = mod.single, mod.add = mod.add, mod.int = mod.int), sort = F)
+  Rsq.s <- as.data.frame(r.squaredGLMM(mod.single))
+  Rsq.s <- signif(Rsq.s, digits=4)[1]
+  Rsq.a <- as.data.frame(r.squaredGLMM(mod.add))
+  Rsq.a <- signif(Rsq.a, digits=4)[1]
+  Rsq.i <- as.data.frame(r.squaredGLMM(mod.int))
+  Rsq.i <- signif(Rsq.i, digits=4)[1]
+  BIC <- BIC(mod.single, mod.add, mod.int)
[…]

### 2020-02-21 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@
 In addition, slope coefficients were calculated for the linear models.
+{**This paragraph needs work - polynomial models were rejected if they showed an exponential increase in productivity with climate (e.g.
+showed GPP increasing exponentially with latitude - needs work on how to word this**}
 
@@
 For the variables which showed a significant interactive or additive effect between MAT and MAP, no other climate variable, in combination with MAT, significantly improved on that model.
-For NPP, there was a significant interactive effect between MAT and water stress months, with this model explaining nearly 5% more variation in NPP than MAT alone {**Helene: but less than MATxMAP?
-if so, why bother mentioning?**}.
-However, for ANPP~foliage~, no multivariate model improved on the univariate model including only MAT.
 

### 2020-02-21 manuscript/appendix.Rmd
updated appendix

@@
 library(kableExtra)
-S1 <- read.csv("tables_figures/mat_growing_season.csv", stringsAsFactors = FALSE, check.names = FALSE)
-kable(S1, format = "latex", booktabs = TRUE, caption = "Table S2. Comparison of growing season length and mean annual temperature as predictors of FACF", escape = F)%>%
+S2 <- read.csv("tables_figures/mat_growing_season.csv", stringsAsFactors = FALSE, check.names = FALSE)
+kable(S2, format = "latex", booktabs = TRUE, caption = "Table S2. Comparison of growing season length and mean annual temperature as predictors of FACF", escape = F)%>%
   kable_styling(latex_options = c("scale_down", "hold_position")) %>%
@@
 \newpage
-```{r echo=FALSE, fig.cap = "Figure S1: Maps showing distribution of samples for the nine FACF analyzed here.", out.height ='100%', fig.pos='H'}
+
+\blandscape
+```{r eval = TRUE, echo=FALSE, warning=FALSE}
+library(knitr)
+library(kableExtra)
+S3 <- read.csv("tables_figures/r2_table1.csv", stringsAsFactors = FALSE, check.names = FALSE)
+kable(S3, format = "latex", booktabs = TRUE, caption = "Table S3. Model details and R2 values for all climate variables tested") %>%
+  kable_styling(latex_options = c("scale_down", "hold_position")) %>%
+  add_header_above(c(" ", "Latitude"=2, "MAT"=2, "MAP"=2, "T Seas"=2, "P Seas"=2, "ATR"=2, "Solar R"=2, "AI"=2))
+```
+
+```{r eval = TRUE, echo=FALSE, warning=FALSE}
+library(knitr)
+library(kableExtra)
+S4 <- read.csv("tables_figures/r2_table2.csv", stringsAsFactors = FALSE, check.names = FALSE)
+kable(S4, format = "latex", booktabs = TRUE) %>%
+  kable_styling(latex_options = c("scale_down", "hold_position")) %>%
+  add_header_above(c(" ", "Cloud"=2, "AFD"=2, "AWD"=2, "PET"=2, "VPD"=2, "Max VPD"=2, "WSM"=2, "LGS"=2))
+```
+\elandscape
+
+\newpage
+```{r echo=FALSE, fig.cap = "Figure S1: Maps showing distribution of samples for the nine FACF analyzed here.", out.height ='95%', fig.pos='H'}
 

### 2020-02-21 scripts/final_model_no_weighting.R
Update final_model_no_weighting.R

@@ -3,7 +3,7 @@
 rm(list = ls())
 
 # Set working directory as ForC main folder ####
-setwd("C:/Users/becky/Dropbox (Smithsonian)/GitHub/ForC")
+setwd("C:/Users/gyrcbm/Dropbox/ForC")
 
 # Load libaries ####
 library(lme4)
@@ -115,8 +115,7 @@ all.results <- NULL
 all.aictab <- NULL
 all.koeppen <- NULL
 
-fixed.variables <- c("mat", "map", "lat", "AnnualMeanTemp", "TempSeasonality", "TempRangeAnnual", "AnnualPre", "AnnualFrostDays", "AnnualWetDays", "Aridity", "PotentialEvapotranspiration", "VapourPressureDeficit", "SolarRadiation", "PreSeasonality", "MaxVPD", "WaterStressMonths", "length_growing_season")
-
+fixed.variables <- c("mat", "map", "lat", "AnnualMeanTemp", "TempSeasonality", "TempRangeAnnual", "AnnualPre", "AnnualFrostDays", "AnnualWetDays", "Aridity", "PotentialEvapotranspiration", "VapourPressureDeficit", "SolarRadiation", "PreSeasonality", "MaxVPD", "WaterStressMonths", "length_growing_season", "CloudCover")
 
 
 response.variables.groups <- list(c("GPP", "NPP", "BNPP_root", "BNPP_root_fine"),
@@ -137,12 +136,12 @@ for(response.variables in response.variables.groups){
     
     for(fixed.v in fixed.variables){
       
-      png(file = paste0("C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/figures/final_figures/unweighted_model/Effect_of_", fixed.v, "_MATURE_only_", age, "_", n, ".png"), width = 2255, height = 2000, units = "px", res = 300)
+      png(file = paste0("C:/Users/gyrcbm/Dropbox/Global_Productivity/results/figures/final_figures/unweighted_model/Effect_of_", fixed.v, "_MATURE_only_", age, "_", n, ".png"), width = 2255, height = 2000, units = "px", res = 300)
       
       par(mfrow = c(1,1), mar = c(0,0,0,0), oma = c(5,5,2,0))
       print(fixed.v)
       
-      fixed.v.info <- read.csv("C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/raw.data/fixedv_data.csv", stringsAsFactors = F)
[…]

### 2020-02-21 scripts/mat_map_interaction.R
Update mat_map_interaction.R

@@ -113,10 +113,10 @@ pannel.nb <- 1
 
 
 response.variables <- c("GPP", "NPP", "ANPP", "ANPP_woody_stem", "ANPP_foliage", "BNPP_root", "BNPP_root_fine", "R_auto", "R_auto_root")
-response.variables <- "GPP"
+# response.variables <- "GPP"
+############################## this section of code is useful in determining whether there are significant interactive effects or not
+####################it is necessary to look at both the AIC and Pvalues
 
-
-# png(file = paste0("C:/Users/gyrcbm/Dropbox/Global_Productivity/results/figures/final_figures/interactions/mat_map_interaction.png"), width = 2255, height = 2000, units = "px", res = 300)
 par(mfrow = c(3,3), mar = c(1,0,0,2), oma = c(5,8,2,0), xpd = T)
 
 for (response.v in response.variables){
@@ -135,7 +135,7 @@ for (response.v in response.variables){
   
   mod.single <- lmer(mean ~ mat + (1|geographic.area/plot.name), data = df, REML = F)
   mod.add <- lmer(mean ~ mat + map + (1|geographic.area/plot.name), data = df, REML = F)
-  mod.int <- lmer(mean ~ mat * map + (1|geographic.area/plot.name), data = df, REML = F)
+  mod.int <- lmer(mean ~ mat * map + (1|geographic.area/plot.name), data = df, REML = F) 
   aictab <- aictab(list(mod.single = mod.single, mod.add = mod.add, mod.int = mod.int), sort = F)
   Rsq.s <- as.data.frame(r.squaredGLMM(mod.single))
   Rsq.s <- signif(Rsq.s, digits=4)[1]
@@ -150,8 +150,9 @@ for (response.v in response.variables){
   aictab$Rsq <- c(Rsq.s, Rsq.a, Rsq.i)
   
   aictab <- cbind(aictab, BIC)
-  
-  significance <- anova(mod.add, mod.int)$"Pr(>Chisq)"[2]
+  anova <- as.data.frame(anova(mod.single, mod.add, mod.int))
+  anova$flux <- response.v
+  significance <- anova(mod.single, mod.add, mod.int)$"Pr(>Chisq)"[2]
   significance <- signif(significance, digits=4)
   
   
[…]

### 2020-02-21 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@
 
-There was a significant interactive effect between MAT and MAP for GPP, BNPP~root~, BNPP~fine~ ~root~, ANPP, ANPP~woody~ ~stem~, and R~root~ (fig.3).
-There was a significant additive effect for R~auto~.
-NPP and ANPP~foliage~ showed no significant interactive or additive effect.
-{**NEED SI TABLE GIVING FITTED COEFFICIENTS & STATS**}
+There was a significant interactive effect between MAT and MAP for NPP and ANPP~woody~ ~stem~ (fig.3).
+There was a significant additive effect for GPP, ANPP and autotrophic respiration.
+ANPP~foliage~, BNPP, BNPP~fine~ ~root~, and root respiration showed no significant interactive or additive effect (see table S4).
 
 For the variables which showed a significant interactive or additive effect between MAT and MAP, no other climate variable, in combination with MAT, significantly improved on that model.
+{**need to confirm this given changes in MAT MAP results**}
 

### 2020-02-21 scripts/mat_map_interaction.R
Update mat_map_interaction.R

@@ -133,6 +133,7 @@ for (response.v in response.variables){
   df <- ForC_simplified[rows.with.response & fixed.no.na, ]
   df$masl <- (df$masl/1000)
   
+  mod.null <- lmer(mean ~ 1 + (1|geographic.area/plot.name), data = df, REML = F)
   mod.single <- lmer(mean ~ mat + (1|geographic.area/plot.name), data = df, REML = F)
   mod.add <- lmer(mean ~ mat + map + (1|geographic.area/plot.name), data = df, REML = F)
   mod.int <- lmer(mean ~ mat * map + (1|geographic.area/plot.name), data = df, REML = F) 
@@ -150,7 +151,7 @@ for (response.v in response.variables){
   aictab$Rsq <- c(Rsq.s, Rsq.a, Rsq.i)
   
   aictab <- cbind(aictab, BIC)
-  anova <- as.data.frame(anova(mod.single, mod.add, mod.int))
+  anova <- as.data.frame(anova(mod.null, mod.single, mod.add, mod.int))
   anova$flux <- response.v
   significance <- anova(mod.single, mod.add, mod.int)$"Pr(>Chisq)"[2]
   significance <- signif(significance, digits=4)
@@ -286,7 +287,7 @@ for (response.v in response.variables){
 write.csv(all.results, "C:/Users/gyrcbm/Dropbox/Global_Productivity/results/tables/best_model_outputs/mat_map_interaction.csv")
 
 ########################################## code to correctly plot
-
+Rsqs = NULL
 
 response.variables <- c("GPP", "NPP", "ANPP", "ANPP_woody_stem", "ANPP_foliage", "BNPP_root", "BNPP_root_fine", "R_auto", "R_auto_root")
 png(file = paste0("C:/Users/gyrcbm/Dropbox/Global_Productivity/results/figures/final_figures/interactions/mat_map_interaction.png"), width = 2255, height = 2000, units = "px", res = 300)
@@ -355,7 +356,7 @@ for (response.v in response.variables){
   if(response.v %in% c("NPP", "ANPP_woody_stem")) mtext(side = 3, line = -2, text = "Significant interactive effect", adj = 0.1, cex = 0.5)
   if(response.v %in% c("GPP", "ANPP", "R_auto")) mtext(side = 3, line = -2, text = "Significant additive effect", adj = 0.1, cex = 0.5)
[…]

### 2020-02-21 manuscript/appendix.Rmd
updated appendix

@@
 
+```{r eval = TRUE, echo=FALSE, warning=FALSE}
+library(knitr)
+library(kableExtra)
+S5 <- read.csv("tables_figures/mat_map.csv", stringsAsFactors = FALSE, check.names = FALSE)
+kable(S5, format = "latex", booktabs = TRUE, caption = "Table S4. Results of analysis of interactions effects between MAT and MAP, for all FACF", escape = F)%>%
+  kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 12) %>%
+  column_spec(1, width = "4cm") %>%
+  column_spec(2, width = "6cm") %>%
+  column_spec(3, width = "6cm") %>%
+  column_spec(4, width = "6cm")%>%
+  column_spec(5, width = "2cm")%>%
+  column_spec(6, width = "2cm")
+```
+
 \newpage

### 2020-02-21 manuscript/appendix.tex
updated appendix

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Supplementary Information: Global patterns of forest autotrophic carbon fluxes},
            pdfauthor={Rebecca Banbury Morgan; Valentine Herrmann; Norbert Kunert; Ben Bond-Lamberty; Helene C. Muller-Landau; Kristina J. Anderson-Teixeira},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}

  \title{Supplementary Information: Global patterns of forest autotrophic carbon
fluxes}
    \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
    \author{Rebecca Banbury Morgan \\ Valentine Herrmann \\ Norbert Kunert \\ Ben Bond-Lamberty \\ Helene C. Muller-Landau \\ Kristina J. Anderson-Teixeira}
    \preauthor{\centering\large\emph}
  \postauthor{\par}
    \date{}
    \predate{}\postdate{}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\usepackage{float} \usepackage{caption} \captionsetup[table]{font=footnotesize} \captionsetup[figure]{font=footnotesize} \captionsetup[figure]{labelformat=empty} \captionsetup[table]{labelformat=empty} \usepackage{pdflscape} \newcommand{\blandscape}{\begin{landscape}} \newcommand{\elandscape}{\end{landscape}}

\begin{document}
\maketitle

\listoftables
\listoffigures

\newpage

\blandscape

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-2}Table S1. Climate variable definitions, sources, and abbreviations}
\centering
\resizebox{\linewidth}{!}{
\fontsize{12}{14}\selectfont
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{6cm}>{\raggedright\arraybackslash}p{2cm}>{\raggedright\arraybackslash}p{7cm}>{\raggedright\arraybackslash}p{2cm}l}
\toprule
Climate variable & Units & Definition & Abbreviation & Source\\
\midrule
Mean annual temperature (MAT) & $^\circ$C &  & MAT & Primary literature; WorldClim$^{1}$\\
Mean annual precipitation (MAP) & mm $yr^{-1}$ &  & MAP & Primary literature; WorldClim$^{1}$\\
Temperature seasonality &  & Standard deviation of MAT *100 & T Seas & WorldClim$^{1}$\\
Precipitation seasonality &  & Coefficient of variation of MAP & P Seas & WorldClim$^{1}$\\
Annual temperature range & $^\circ$C & Maximum temperature of warmest month - minimum temperature of coldest month & ART & WorldClim$^{1}$\\
\addlinespace
Solar radiation & kJ $m^{-2} yr^{-1}$ &  & Solar R & WorldClim2$^{2}$\\
Cloud cover & percentage & Cloud percentage cover & Cloud & CRU time-series dataset v 4.03$^{3}$\\
Annual frost days & days $yr^{-1}$ & Number of freeze days annually & AFD & CRU time-series dataset v 4.03$^{3}$\\
Annual wet days & days $yr^{-1}$ & Number of days with precipitation >0.1 mm annually & AWD & CRU time-series dataset v 4.03$^{3}$\\
Potential evapotranspiration (PET) & mm $yr^{-1}$ & Mean annual potential evapotranspiration & PET & Global Aridity Index and Potential Evapotranspiration Climate Database$^{4}$\\
\addlinespace
Aridity &  & MAP/mean annual PET & AI & Global Aridity Index and Potential Evapotranspiration Climate Database$^{4}$\\
Vapour pressure deficit (VPD) & kPa &  & VPD & TerraClimate$^{5}$\\
Maximum vapour pressure deficit (Max VPD) & kPa &  & Max VPD & Derived\\
Water stress months & months $yr^{-1}$ & Number of months annually with MAP < PET & WSM & Derived\\
Length of growing season & months $yr^{-1}$ & Number of months annually with mean minimum temperature > 0.5$^\circ$C & LGS & Derived\\
\bottomrule
\end{tabular}}
\end{table}

\elandscape
\newpage

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-3}Table S2. Comparison of growing season length and mean annual temperature as predictors of FACF}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{5cm}>{\raggedleft\arraybackslash}p{2cm}>{\raggedleft\arraybackslash}p{7cm}>{\raggedleft\arraybackslash}p{7cm}}
\toprule
Fixed effect & AIC value & Delta AICc & Marginal R squared\\
\midrule
\addlinespace[1em]
\multicolumn{1}{l}{\textbf{GPP}}\\
\hspace{1em}MAT & 126.42617 & 0.000000 & 0.6196780\\
\hspace{1em}Growing season length & 140.80589 & 14.379717 & 0.5411935\\
\hspace{1em}None & 178.96179 & 52.535617 & 0.0000000\\
\addlinespace[1em]
\hline
\multicolumn{1}{l}{\textbf{NPP}}\\
\hspace{1em}MAT & 174.88249 & 0.000000 & 0.5156614\\
\hspace{1em}Growing season length & 191.53714 & 16.654650 & 0.4006999\\
\hspace{1em}None & 216.16976 & 41.287265 & 0.0000000\\
\addlinespace[1em]
\hline
\multicolumn{1}{l}{\textbf{ANPP}}\\
\hspace{1em}MAT & 249.50512 & 0.000000 & 0.2925950\\
\hspace{1em}Growing season length & 254.20763 & 4.702509 & 0.2612187\\
\hspace{1em}None & 268.94008 & 19.434966 & 0.0000000\\
\addlinespace[1em]
\hline
\multicolumn{1}{l}{\textbf{ANPP woody stem}}\\
\hspace{1em}MAT & 235.95797 & 0.000000 & 0.1548800\\
\hspace{1em}Growing season length & 237.28992 & 1.331943 & 0.1370243\\
\hspace{1em}None & 243.13700 & 7.179027 & 0.0000000\\
\addlinespace[1em]
\hline
\multicolumn{1}{l}{\textbf{ANPP foliage}}\\
\hspace{1em}MAT & 484.87610 & 0.000000 & 0.4462629\\
\hspace{1em}Growing season length & 520.96482 & 36.088722 & 0.3497750\\
\hspace{1em}None & 560.34915 & 75.473049 & 0.0000000\\
\addlinespace[1em]
\hline
\multicolumn{1}{l}{\textbf{BNPP root}}\\
\hspace{1em}MAT & 184.54480 & 0.000000 & 0.5921282\\
\hspace{1em}Growing season length & 204.92685 & 20.382054 & 0.4644116\\
\hspace{1em}None & 237.46554 & 52.920743 & 0.0000000\\
\addlinespace[1em]
\hline
\multicolumn{1}{l}{\textbf{BNPP fine root}}\\
\hspace{1em}MAT & 540.19217 & 0.000000 & 0.2429540\\
\hspace{1em}Growing season length & 566.36955 & 26.177388 & 0.1060029\\
\hspace{1em}None & 578.65529 & 38.463119 & 0.0000000\\
\addlinespace[1em]
\hline
\multicolumn{1}{l}{\textbf{Autotrophic respiration}}\\
\hspace{1em}MAT & 45.25818 & 0.000000 & 0.6271133\\
\hspace{1em}Growing season length & 50.35515 & 5.096972 & 0.5041004\\
\hspace{1em}None & 56.16877 & 10.910597 & 0.0000000\\
\addlinespace[1em]
\hline
\multicolumn{1}{l}{\textbf{Root respiration}}\\
\hspace{1em}MAT & 133.53500 & 0.000000 & 0.2507631\\
\hspace{1em}Growing season length & 135.92632 & 2.391311 & 0.1990489\\
\hspace{1em}None & 141.78719 & 8.252190 & 0.0000000\\
\bottomrule
\end{tabular}}
\end{table}

\newpage

\blandscape

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-4}Table S3. Model details and R2 values for all climate variables tested}
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}[t]{llrlrlllrlllrllll}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Latitude} & \multicolumn{2}{c}{MAT} & \multicolumn{2}{c}{MAP} & \multicolumn{2}{c}{T Seas} & \multicolumn{2}{c}{P Seas} & \multicolumn{2}{c}{ATR} & \multicolumn{2}{c}{Solar R} & \multicolumn{2}{c}{AI} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-7} \cmidrule(l{3pt}r{3pt}){8-9} \cmidrule(l{3pt}r{3pt}){10-11} \cmidrule(l{3pt}r{3pt}){12-13} \cmidrule(l{3pt}r{3pt}){14-15} \cmidrule(l{3pt}r{3pt}){16-17}
Carbon flux & Model & R-squared & Model & R-squared & Model & R-squared & Model & R-squared & Model & R-squared & Model & R-squared & Model & R-squared & Model & R-squared\\
\midrule
GPP & Linear & 0.6387 & Linear & 0.6094 & Polynomial & 0.324 & Polynomial & 0.7076 & - & - & Linear & 0.6269 & Linear & 0.1554 & - & -\\
NPP & Linear & 0.5108 & Linear & 0.4171 & Polynomial & 0.2138 & Polynomial & 0.4905 & - & - & Linear & 0.4442 & Linear & 0.09548 & Linear & 0.03795\\
ANPP & Linear & 0.4351 & Linear & 0.4444 & Polynomial & 0.1625 & Polynomial & 0.4126 & - & - & Linear & 0.3331 & Linear & 0.1061 & Linear & 0.04851\\
ANPP woody stem & Linear & 0.1773 & Linear & 0.2396 & - & - & Linear & 0.1416 & - & - & Linear & 0.1157 & Linear & 0.05048 & Linear & 0.06607\\
ANPP foliage & Linear & 0.4999 & Linear & 0.5826 & Polynomial & 0.2509 & Linear & 0.4823 & - & - & Linear & 0.5033 & Linear & 0.172 & Linear & 0.1084\\
\addlinespace
BNPP root & Linear & 0.3373 & Linear & 0.2833 & Polynomial & 0.1452 & Linear & 0.3300 & - & - & Linear & 0.2904 & Linear & 0.2315 & - & -\\
BNPP fine root & Linear & 0.1704 & Linear & 0.1477 & Linear & 0.08935 & Linear & 0.1721 & - & - & Linear & 0.1790 & Linear & 0.1393 & Polynomial & 0.06915\\
Autotrophic respiration & Linear & 0.6534 & Linear & 0.5909 & Polynomial & 0.604 & Polynomial & 0.6246 & - & - & Linear & 0.4900 & Linear & 0.26 & Polynomial & 0.4804\\
Root respiration & Linear & 0.2612 & Linear & 0.2418 & Linear & 0.1493 & Polynomial & 0.2151 & - & - & Linear & 0.1776 & - & - & Linear & 0.1567\\
\bottomrule
\end{tabular}}
\end{table}

\begin{table}[!h]
\centering
\resizebox{\linewidth}{!}{
\begin{tabular}{llllrlllrlllllllr}
\toprule
\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Cloud} & \multicolumn{2}{c}{AFD} & \multicolumn{2}{c}{AWD} & \multicolumn{2}{c}{PET} & \multicolumn{2}{c}{VPD} & \multicolumn{2}{c}{Max VPD} & \multicolumn{2}{c}{WSM} & \multicolumn{2}{c}{LGS} \\
\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-7} \cmidrule(l{3pt}r{3pt}){8-9} \cmidrule(l{3pt}r{3pt}){10-11} \cmidrule(l{3pt}r{3pt}){12-13} \cmidrule(l{3pt}r{3pt}){14-15} \cmidrule(l{3pt}r{3pt}){16-17}
Carbon flux & Model & R-squared & Model & R-squared & Model & R-squared & Model & R-squared & Model & R-squared & Model & R-squared & Model & R-squared & Model & R-squared\\
\midrule
GPP & - & - & Linear & 0.5498 & Linear & 0.11 & Polynomial & 0.3602 & Polynomial & 0.3076 & - & - & - & - & Linear & 0.5312\\
NPP & Linear & 0.06339 & Linear & 0.4036 & Linear & 0.1118 & Polynomial & 0.3165 & Polynomial & 0.178 & - & - & Linear & 0.03561 & Linear & 0.3782\\
ANPP & Linear & 0.0439 & Linear & 0.3668 & Linear & 0.1732 & Polynomial & 0.2672 & Polynomial & 0.2294 & Polynomial & 0.0632 & Polynomial & 0.06269 & Linear & 0.3425\\
ANPP woody stem & - & - & Linear & 0.1380 & - & - & Polynomial & 0.2024 & Polynomial & 0.2146 & Linear & 0.07403 & - & - & Linear & 0.1041\\
ANPP foliage & - & - & Linear & 0.5306 & Linear & 0.1469 & Linear & 0.3076 & Polynomial & 0.3751 & Polynomial & 0.07489 & Polynomial & 0.1724 & Linear & 0.4552\\
\addlinespace
BNPP root & - & - & Linear & 0.2799 & Polynomial & 0.1113 & Polynomial & 0.3601 & Polynomial & 0.2584 & - & - & - & - & Linear & 0.2550\\
BNPP fine root & - & - & Linear & 0.1631 & Linear & 0.08161 & Linear & 0.1376 & - & - & - & - & - & - & Linear & 0.1335\\
Autotrophic respiration & - & - & Linear & 0.5502 & Linear & 0.226 & Linear & 0.3298 & Polynomial & 0.4499 & - & - & Linear & 0.2613 & Linear & 0.4664\\
Root respiration & Linear & 0.1578 & Linear & 0.1647 & Linear & 0.1698 & Polynomial & 0.1905 & Polynomial & 0.272 & - & - & Linear & 0.1388 & Linear & 0.1889\\
\bottomrule
\end{tabular}}
\end{table}

\elandscape

\begin{table}[!h]

\caption{\label{tab:unnamed-chunk-6}Table S4. Results of analysis of interactions effects between MAT and MAP, for all FACF}
\centering
\resizebox{\linewidth}{!}{
\fontsize{12}{14}\selectfont
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{6cm}>{\raggedright\arraybackslash}p{6cm}>{\raggedright\arraybackslash}p{6cm}>{\raggedright\arraybackslash}p{2cm}>{\raggedleft\arraybackslash}p{2cm}}
\toprule
Carbon flux & Significant interactive effect & Significant additive effect & Significant effect of MAT & p-value & R-squared value\\
\midrule
GPP & FALSE & TRUE & TRUE & <0.0001 & 0.66\\
NPP & TRUE & TRUE & TRUE & 0.018 & 0.48\\
ANPP & FALSE & TRUE & TRUE & 0.0349 & 0.45\\
ANPP woody stem & TRUE & TRUE & TRUE & 0.021 & 0.26\\
ANPP foliage & FALSE & FALSE & TRUE & <0.0001 & 0.59\\
\addlinespace
BNPP root & FALSE & FALSE & TRUE & <0.0001 & 0.29\\
BNPP fine root & FALSE & FALSE & TRUE & 0.002 & 0.15\\
Autotrophic respiration & FALSE & TRUE & TRUE & 0.041 & 0.71\\
Root respiration & FALSE & FALSE & TRUE & 0.001 & 0.25\\
\bottomrule
\end{tabular}}
\end{table}

\newpage

\begin{figure}[H]
\includegraphics[width=27.78in,height=0.95\textheight]{tables_figures/distribution_all_samples} \caption{Figure S1: Maps showing distribution of samples for the nine FACF analyzed here.}\label{fig:unnamed-chunk-7}
\end{figure}

\blandscape

\begin{figure}[H]

{\centering \includegraphics[width=48.61in,height=0.95\textheight]{tables_figures/climate_regressions}

}

\caption{Figure S2: Correlations among latitude and climate variables. Variable names and units given in Table S1}\label{fig:unnamed-chunk-8}
\end{figure}

\elandscape

\blandscape

\begin{figure}[H]
\includegraphics[width=41.67in,height=0.95\textheight]{tables_figures/grid_plots_climate1} \caption{Figure S3: Individual plots of FACF in relation to mean annual climate, part 1.}\label{fig:unnamed-chunk-9}
\end{figure}

\newpage

\begin{figure}[H]
\includegraphics[width=41.67in,height=0.95\textheight]{tables_figures/grid_plots_climate2} \caption{Figure S4: Individual plots of FACF in relation to mean annual climate, part 2.}\label{fig:unnamed-chunk-10}
\end{figure}

\elandscape

\begin{figure}[H]
\includegraphics[width=27.78in,height=0.95\textheight]{tables_figures/grid_plots_seasonality3} \caption{Figure S5: Individual plots of FACF in relation to mean climate seasonality, part 1.}\label{fig:unnamed-chunk-11}
\end{figure}

\newpage

\begin{figure}[H]
\includegraphics[width=1\linewidth]{tables_figures/grid_plots_seasonality4} \caption{Figure S6: Individual plots of FACF in relation to mean climate seasonality, part 2.}\label{fig:unnamed-chunk-12}
\end{figure}

\newpage

\begin{figure}[H]
\includegraphics[width=1\linewidth]{tables_figures/gridded_growing_season1} \caption{Figure S7: Growing season length-standardized FACF in relation to mean growing season climate, part 1.}\label{fig:unnamed-chunk-13}
\end{figure}

\newpage

\begin{figure}[H]
\includegraphics[width=1\linewidth]{tables_figures/gridded_growing_season2} \caption{Figure S8: Growing season length-standardized FACF in relation to mean growing season climate, part 2.}\label{fig:unnamed-chunk-14}
\end{figure}

\newpage

\begin{figure}[H]

{\centering \includegraphics[width=1\linewidth]{tables_figures/ratio_grid_plots}

}

\caption{Figure S9: Ratios among FACF as a function of latitude and climate variables}\label{fig:unnamed-chunk-15}
\end{figure}

\end{document}

### 2020-02-24 manuscript/manuscript draft.Rmd
worked on intro & formatting
-inserted GCB-style title page
-worked on intro
-added some references

@@
 ---
-title: "Global patterns of forest autotrophic carbon fluxes"
-author: "Rebecca Banbury Morgan, Valentine Herrmann, Norbert Kunert, Ben Bond-Lamberty, Helene C. Muller-Landau, Kristina J. Anderson-Teixeira"
+title: ""
+author: ""
 output:
@@
 ```{r setup, include = FALSE}
-knitr::opts_knit$set(root.dir = 'C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/')
-# knitr::opts_knit$set(root.dir = '/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/Global_Productivity/manuscript/')
+# knitr::opts_knit$set(root.dir = 'C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/')
+knitr::opts_knit$set(root.dir = '/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/Global_Productivity/manuscript/')
 ```
@@
 ```
+**Title:** Global patterns of forest autotrophic carbon fluxes
 
+**Running head:**
+
+**Authors:**
+
+Rebecca Banbury Morgan^1,2^
+
+Valentine Herrmann^1^
+
+Norbert Kunert^1,3^
+
+Ben Bond-Lamberty^4^
+
+Helene C. Muller-Landau^3^
+
+Kristina J. Anderson-Teixeira^1,3^*
+
+**Author Affiliations:**
+
+1. Conservation Ecology Center; Smithsonian Conservation Biology Institute; National Zoological Park, Front Royal, VA, USA
+
+2. *Becky- current*
+
+3. Center for Tropical Forest Science-Forest Global Earth Observatory; Smithsonian Tropical Research Institute; Panama, Republic of Panama
+
+4. Joint Global Change Research Institute, Pacific Northwest National Laboratory, College Park Maryland 20740 USA
+
+*Corresponding Author:
+
+phone: 1-540-635-6546
+
+fax:1-540-635-6506
+
+email: teixeirak@si.edu
+
+**Keywords:**
+
+**Paper type:** Primary Research Article
+
+\newpage
 ###Abstract
@@
 Here, we drew upon # records from the Global Forest Carbon Database (ForC), representing all major forest types and the nine most significant forest autotrophic carbon flux (FACF) variables, to comprehensively explore how C cycling in mature, undisturbed forests varies with latitude and climate on a global scale.
-We show that, across all FACF variables analyzed, C cycling decreases linearly with latitude --a finding that confirms multiple previous studies and contradicts the idea that productivity of temperate forests rivals that of tropical forests.
-The FACF variables increase in proportion to one another, with no differences in allocation detectable at this global scale, but differed in that latitude explained a lower proportion of variation among subsidiary fluxes (in particular, woody aboveground productivity and belowground net primary productivity, $BNPP$).
+We show that, across all FACF variables analyzed, C cycling decreases linearly with latitude -- a finding that confirms multiple previous studies but contradicts the idea that net primary productivity ($NPP$) of temperate forests rivals that of tropical forests.
+The FACF variables generally increase in proportion to one another, with few differences in allocation detectable at this global scale, but differed in that latitude explained a lower proportion of variation among subsidiary fluxes (in particular, woody aboveground $NPP$ and belowground $NPP$, $BNPP$).
 Climate explained a significant proportion (#-#%) of variation in all C fluxes analyzed (less for subsidiary fluxes), with temperature variables in general and mean annual temperature ($MAT$) in particular being the best predictors of FACF on this global scale.
@@
 
+\newpage
 ### Introduction
@@
 Critically, FACF also define forest interactions with Earth's changing climate.
-Approximately **##** Gt of CO~2~ cycle through Earth's forests each year **REF-KAT**), and in recent decades their net C sequestration (~2.4 GtCyr^-1^) offset roughly 30% of anthropogenic fossil fuel emissions [@pan_large_2011].
-As atmospheric carbon dioxide levels continue to rise, driving climate change, forests will play a critical role in shaping the future of Earth's climate (** REFS**).
+Over 69 Gt of CO~2~ cycle through Earth's forests each year [@badgley_terrestrial_2019], and in recent decades their net C sequestration (~2.4 Gt C yr^-1^) offset roughly 30% of anthropogenic fossil fuel emissions [@pan_large_2011].
+As atmospheric carbon dioxide levels continue to rise, driving climate change, forests will play a critical role in shaping the future of Earth's climate ([Cavaleri et al.
+2015](https://onlinelibrary.wiley.com/doi/epdf/10.1111/gcb.12860); [Rogelj et al.
+2018](https://www.ipcc.ch/site/assets/uploads/sites/2/2019/05/SR15_Chapter2_Low_Res.pdf)).
 However, our ability to draw general macroscopic conclusions regarding global variation in FACF with respect to climate has been limited in that these analyses often mix forests that vary in stand age, disturbance history, and/or management status; do not always sufficiently parse related variables; and typically consider only one or a few variables at a time.
 
+**KAT, work on this:**
 **FACF vary with latitude**, showing a general trend of decreasing flux with latitude [@beer_terrestrial_2010; @jung_global_2011].
@@
 Evidence for this is further complicated by the fact that different studies use different measures of productivity to explore these relationships.
-For example, modelling of global terrestrial ecosystem gross primary productivity (GPP) through upscaling and calibration of eddy flux measurements indicates that GPP peaks in tropical forests [@beer_terrestrial_2010; @jung_global_2011; @badgley_terrestrial_2019].
-In contrast, some studies suggest that the highest values of net primary productivity (NPP) may be found in temperate forests [@luyssaert_co_2007; @huston_global_2009], while others find NPP highest in the tropics and decreasing with latitude [@simova_enigma_2017].
-Other studies have chosen to focus exclusively on above-ground net primary productivity (ANPP), finding evidence of a weak negative relationship between ANPP and latitude [@huston_global_2009; @gillman_latitude_2015].
+For example, modelling of global terrestrial ecosystem gross primary productivity ($GPP$) through upscaling and calibration of eddy flux measurements indicates that GPP peaks in tropical forests [@beer_terrestrial_2010; @jung_global_2011; @badgley_terrestrial_2019](*[Li & Xiao 2019](https://www.mdpi.com/2072-4292/11/21/2563)*).
+In contrast, some studies suggest that the highest values of net primary productivity ($NPP$) may be found in temperate forests [@luyssaert_co_2007; @huston_global_2009], while others find $NPP$ highest in the tropics and decreasing with latitude [@simova_enigma_2017].
+Other studies have chosen to focus exclusively on above-ground net primary productivity ($ANPP$), finding evidence of a weak negative relationship between $ANPP$ and latitude [@huston_global_2009; @gillman_latitude_2015].
 
-**Climate is a significant driver of FACF across broad spatial scales** [@cleveland_relationships_2011].
+**Climate is a significant driver of FACF across broad spatial scales** [@luyssaert_co_2007; @cleveland_relationships_2011]([Hursh et al.
+2017](https://onlinelibrary.wiley.com/doi/epdf/10.1111/gcb.13489)).
 The majority of studies have focused on exploring the relationships of FACF to mean annual temperature (MAT) and mean annual precipitation (MAP), as the most commonly reported site-level climate variables.

### 2020-02-25 manuscript/manuscript draft.Rmd
added content to discussion, w/ refs

@@
 The implication is that under warmer conditions with similar moisture availability--and within the temperature range to which forest communities are adapated and acclimatized--higher temperatures result in a generalized acceleration of FACF, with no major shifts in C allocation among subsidiary fluxes.
+This is consistent with observations of continental- to global-scale increases in $GPP$ (*[Li & Xiao 2019](https://www.mdpi.com/2072-4292/11/21/2563)*), $ANPP_{woody stem}$ (*[Brienen et al., 2015](https://www-nature-com.smithsonian.idm.oclc.org/articles/nature14283)*), as well as some C cycle components not considered here--tree mortality (*[Brienen et al., 2015](https://www-nature-com.smithsonian.idm.oclc.org/articles/nature14283)*; **some McDowell Ref-KAT**), soil respiration ([Bond Lamberty & Thompson 2010](https://www-nature-com.smithsonian.idm.oclc.org/articles/nature08930/)), heterotrophic soil respiration ([Bond-Lamberty et al.
+2018](https://www-nature-com.smithsonian.idm.oclc.org/articles/s41586-018-0358-x)).
+Of course, other factors such as rising CO~2~ are also at play...
 Of course, actual climatic changes will result in very different sets of conditions than represented across geographic gradients in climate, but our analysis clarifies how carbon cycles through contemporary forest ecosystems.

### 2020-02-25 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@
 2018](https://www-nature-com.smithsonian.idm.oclc.org/articles/s41586-018-0358-x)).
-Of course, other factors such as rising CO~2~ are also at play...
+Of course, other factors such as rising CO~2~ are also at play, and such increases are not universal (e.g., [Rutishauser et al., 2019](https://onlinelibrary-wiley-com.smithsonian.idm.oclc.org/doi/full/10.1111/gcb.14833?casa_token=imtB9LD9mL8AAAAA%3AYWgsqNUfRkgs6_dIuDzmwUzL0qVyyxojhbj3jjfRcykBy8ay9T4qMOKPeOPytSGET9muIs8myhQOh3E))...
 Of course, actual climatic changes will result in very different sets of conditions than represented across geographic gradients in climate, but our analysis clarifies how carbon cycles through contemporary forest ecosystems.

### 2020-02-29 manuscript/manuscript draft.Rmd
fixed up table 1

@@
 hypothesis_table <- read.csv("hypothesis_table.csv", stringsAsFactors = FALSE, check.names = FALSE)
-hypothesis_table[1,2] <- paste0(footnote_marker_number(1)," , ",footnote_marker_number(2)," , ", footnote_marker_number(3))
-hypothesis_table[2,2] <- paste0(footnote_marker_number(4))
-hypothesis_table[3,2] <- paste0(footnote_marker_number(1)," , ",footnote_marker_number(4))
-hypothesis_table[4,2] <- paste0(footnote_marker_number(5))
-hypothesis_table[10,2] <- paste0(footnote_marker_number(6)," , ",footnote_marker_number(7)," , ", footnote_marker_number(8))
-hypothesis_table[11,2] <- paste0(footnote_marker_number(7)," , ", footnote_marker_number(8))
-hypothesis_table[12,2] <- paste0(footnote_marker_number(8))
-kable(hypothesis_table, booktabs = TRUE, caption = "Summary of research questions, corresponding hypotheses, and results.
+hypothesis_table[1,1] <- paste0(hypothesis_table[1,1], footnote_marker_number(1),"$^{,}$",footnote_marker_number(2),"$^{,}$", footnote_marker_number(3))
+hypothesis_table[2,1] <- paste0(hypothesis_table[2,1], footnote_marker_number(4))
+hypothesis_table[3,1] <- paste0(hypothesis_table[3,1], footnote_marker_number(1),"$^{,}$",footnote_marker_number(4))
+hypothesis_table[4,1] <- paste0(hypothesis_table[4,1], footnote_marker_number(5))
+hypothesis_table[10,1] <- paste0(hypothesis_table[10,1], footnote_marker_number(6),"$^{,}$",footnote_marker_number(7),"$^{,}$", footnote_marker_number(8))
+hypothesis_table[11,1] <- paste0(hypothesis_table[11,1], footnote_marker_number(7),"$^{,}$", footnote_marker_number(8))
+hypothesis_table[12,1] <- paste0(hypothesis_table[12,1], footnote_marker_number(8))
+kable(hypothesis_table, format = "latex", booktabs = TRUE, caption = "Summary of research questions, corresponding hypotheses, and results.
 Statistically signficant support for/ rejection of hypotheses is indicated with 'yes'/'no', parentheses indicate partial overall support/rejection of hypotheses across all FACF, and '-' indicates no significant relationship.", escape = F) %>%
   kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 12) %>%
-  pack_rows("Q1. How do FACF vary with latitude?", 1, 1, latex_gap_space = "1em", colnum = 1, hline_before = FALSE) %>%
-  pack_rows("Q2. How do FACF vary with MAT and MAP?", 2, 4, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  pack_rows("Q3. How are FACF related to other climate variables?", 5, 7, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  pack_rows("Q4. How does seasonality influence FACF?", 8, 11, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  pack_rows("Q5. When standardised by growing season length, how do FACF vary with climate?", 12, 15, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  add_header_above(c(" ", " ", " ", "Forest autotrophic carbon fluxes (FACF)" = 9, " ")) %>%
-  #column_spec(1, width = "15cm") %>%
-  column_spec(2, width = "5cm") %>%
-  column_spec(3:13, width = "1.8cm") %>%
+  pack_rows("Q1. How do FACF vary with latitude?", 1, 1, latex_gap_space = "1em", colnum = 5, hline_before = FALSE) %>%
+  pack_rows("Q2. How do FACF vary with MAT and MAP?", 2, 4, latex_gap_space = "1em", colnum = 5, hline_before = TRUE) %>%
+  pack_rows("Q3. How are FACF related to other climate variables?", 5, 7, latex_gap_space = "1em", colnum = 5, hline_before = TRUE) %>%
+  pack_rows("Q4. How does seasonality influence FACF?", 8, 11, latex_gap_space = "1em", colnum = 5, hline_before = TRUE) %>%
+  pack_rows("Q5. When standardised by growing season length, how do FACF vary with climate?", 12, 15, latex_gap_space = "1em", colnum = 5, hline_before = TRUE) %>%
+  add_header_above(c(" ",  " ", "Forest autotrophic carbon fluxes (FACF)" = 9, " ")) %>%
+  #column_spec(2:11, width = "1.8cm") %>%
   kableExtra::landscape() %>%

### 2020-02-29 manuscript/manuscript draft.Rmd
added hypothesis ref
#62

@@
 hypothesis_table[1,1] <- paste0(hypothesis_table[1,1], footnote_marker_number(1),"$^{,}$",footnote_marker_number(2),"$^{,}$", footnote_marker_number(3))
-hypothesis_table[2,1] <- paste0(hypothesis_table[2,1], footnote_marker_number(4))
+hypothesis_table[2,1] <- paste0(hypothesis_table[2,1], footnote_marker_number(1),"$^{,}$",footnote_marker_number(4))
 hypothesis_table[3,1] <- paste0(hypothesis_table[3,1], footnote_marker_number(1),"$^{,}$",footnote_marker_number(4))

### 2020-02-29 manuscript/appendix.Rmd
rearranged Appendix tables to match order presented in manuscript

@@
 # knitr::opts_knit$set(root.dir = 'C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/')
-knitr::opts_knit$set(root.dir = 'C:/Users/gyrcbm/Dropbox/Global_Productivity/manuscript/')
-# knitr::opts_knit$set(root.dir = '/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/Global_Productivity/manuscript/')
+# knitr::opts_knit$set(root.dir = 'C:/Users/gyrcbm/Dropbox/Global_Productivity/manuscript/')
+ knitr::opts_knit$set(root.dir = '/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/Global_Productivity/manuscript/')
 ```
@@
 \newpage
-```{r eval = TRUE, echo=FALSE, warning=FALSE}
-library(knitr)
-library(kableExtra)
-S2 <- read.csv("tables_figures/mat_growing_season.csv", stringsAsFactors = FALSE, check.names = FALSE)
-kable(S2, format = "latex", booktabs = TRUE, caption = "Table S2. Comparison of growing season length and mean annual temperature as predictors of FACF", escape = F)%>%
-  kable_styling(latex_options = c("scale_down", "hold_position")) %>%
-  column_spec(1, width = "5cm") %>%
-  column_spec(2, width = "2cm") %>%
-  column_spec(3:4, width = "7cm") %>%
-  pack_rows("GPP", 1, 3, latex_gap_space = "1em", colnum = 1, hline_before = FALSE) %>%
-  pack_rows("NPP", 4, 6, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  pack_rows("ANPP", 7, 9, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  pack_rows("ANPP woody stem", 10, 12, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  pack_rows("ANPP foliage", 13, 15, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  pack_rows("BNPP root", 16, 18, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  pack_rows("BNPP fine root", 19, 21, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  pack_rows("Autotrophic respiration", 22, 24, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  pack_rows("Root respiration", 25, 27, latex_gap_space = "1em", colnum = 1, hline_before = TRUE)
-```
-
-\newpage
 
@@
 S3 <- read.csv("tables_figures/r2_table1.csv", stringsAsFactors = FALSE, check.names = FALSE)
-kable(S3, format = "latex", booktabs = TRUE, caption = "Table S3. Model details and R2 values for all climate variables tested") %>%
+kable(S3, format = "latex", booktabs = TRUE, caption = "Table S2. Model details and R2 values for all climate variables tested") %>%
   kable_styling(latex_options = c("scale_down", "hold_position")) %>%
@@
 S5 <- read.csv("tables_figures/mat_map.csv", stringsAsFactors = FALSE, check.names = FALSE)
-kable(S5, format = "latex", booktabs = TRUE, caption = "Table S4. Results of analysis of interactions effects between MAT and MAP, for all FACF", escape = F)%>%
+kable(S5, format = "latex", booktabs = TRUE, caption = "Table S3. Results of analysis of interactions effects between MAT and MAP, for all FACF", escape = F)%>%
   kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 12) %>%
@@
   column_spec(6, width = "2cm")
+```
+
+\newpage
+```{r eval = TRUE, echo=FALSE, warning=FALSE}
+library(knitr)
+library(kableExtra)
+S2 <- read.csv("tables_figures/mat_growing_season.csv", stringsAsFactors = FALSE, check.names = FALSE)
+kable(S2, format = "latex", booktabs = TRUE, caption = "Table S4. Comparison of growing season length and mean annual temperature as predictors of FACF", escape = F)%>%
+  kable_styling(latex_options = c("scale_down", "hold_position")) %>%
+  column_spec(1, width = "5cm") %>%
+  column_spec(2, width = "2cm") %>%
+  column_spec(3:4, width = "7cm") %>%
+  pack_rows("GPP", 1, 3, latex_gap_space = "1em", colnum = 1, hline_before = FALSE) %>%
+  pack_rows("NPP", 4, 6, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("ANPP", 7, 9, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("ANPP woody stem", 10, 12, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("ANPP foliage", 13, 15, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("BNPP root", 16, 18, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("BNPP fine root", 19, 21, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("Autotrophic respiration", 22, 24, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("Root respiration", 25, 27, latex_gap_space = "1em", colnum = 1, hline_before = TRUE)
 ```

### 2020-02-29 manuscript/appendix.Rmd
fixed up Table S1

@@
 kable(S1, format = "latex", booktabs = TRUE, caption = "Table S1. Climate variable definitions, sources, and abbreviations", escape = F)%>%
-  kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 12) %>%
-  column_spec(1, width = "6cm") %>%
-  column_spec(2, width = "2cm") %>%
-  column_spec(3, width = "7cm") %>%
-  column_spec(4, width = "2cm")
+  kable_styling(latex_options = c("scale_down", "hold_position")) %>%
+  column_spec(1, width = "2.5cm") %>%
+  column_spec(2, width = "5cm") %>%
+  column_spec(3, width = "2cm") %>%
+  column_spec(4, width = "7cm") %>%
+  column_spec(5, width = "5cm")
 ```

### 2020-02-29 manuscript/manuscript draft.Rmd
rearranged Appendix tables to match order presented in manuscript
worked on methods, moved Table 2 to .csv file

@@
 
-Analyses were conducted on data contained in the open-access ForC database [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018].
-This database contains records of field-based measurements of forest carbon stocks and annual fluxes, compiled from original publications and existing data compilations and databases.
+*FACF data*
+
+This analysis focused on nine FACF included in the open-access ForC database (Table 2) [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018].
+ForC contains records of field-based measurements of forest carbon stocks and annual fluxes, compiled from original publications and existing data compilations and databases.
 Associated data, such as stand age, measurement methodologies, and disturbance history, are also included.
-The database was significantly expanded since the publication of [@anderson-teixeira_forc_2018] through integration with the Global Soil Respiration Database [@bond-lamberty_global_2010].
-Additional targeted literature searches were conducted to identify any further available data on FACF, with particular focus on mature forests in temperate and boreal regions.
-ForC currently contains 29730 records from 4979 plots, representing 20 distinct ecozones across all forested biogeographic and climate zones.
-We used ForC v3.0, archived on Zenodo with DOI 10.5281/zenodo.3403855.
+The database was significantly expanded since the publication of @anderson-teixeira_forc_2018 through integration with the Global Soil Respiration Database [@bond-lamberty_global_2010].
+Additional targeted literature searches were conducted to identify any further available data on the FACF analyzed here, with particular focus on mature forests in temperate and boreal regions, which were not included in the review of @anderson-teixeira_carbon_2016.
+We used ForC v3.0, archived on Zenodo with DOI 10.5281/zenodo.3403855.This version contained 29,730 records from 4,979 plots, representing 20 distinct ecozones across all forested biogeographic and climate zones.
+
+This analysis focused on mature forests with no known history of signficant disturbance or management. 
+There is evidence that stand age influences patterns of FACF and carbon allocation in forest ecosystems, and can confound relationships between latitude and primary productivity [@delucia_forest_2007; @gillman_latitude_2015].
+To reduce any biasing effects of stand age, we included only stands of known age $\ge$ 100 years and those described by terms such as "mature", "intact", or "old-growth".
+Since management can alter observed patterns of FACF [@simova_enigma_2017], sites were excluded from analysis if they were managed, defined as plots that were planted, managed as plantations, irrigated, fertilised or including the term "managed" in their site description.
+Sites that had experienced significant disturbance within the past 100 years were also excluded.
+Disturbances that qualified sites for exclusion included major cutting or harvesting, burning, flooding, drought and storm events with site mortality >10% of trees.
+Grazed sites were retained.
+
+```{r eval = TRUE, echo=FALSE, warning=FALSE}
+library(knitr)
+library(kableExtra)
+FACF_table <- read.csv("FACF_variables.csv", stringsAsFactors = FALSE, check.names = FALSE)
+names(FACF_table) [6]  <- paste0(names(FACF_table) [6], footnote_marker_symbol(1))
+kable(FACF_table, format = "latex", booktabs = TRUE, caption = "Definitions and sample sizes of FACF variables used in analysis.
+All variables are in units of Mg C ha$^{-1}$ yr$^{-1}$.
+", escape = F) %>%
+  kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 12) %>%
+  column_spec(2, width = "4cm") %>%
+  column_spec(3, width = "7cm") %>%
+  column_spec(4, width = "10cm") %>%
+  column_spec(5:6, width = "1.5cm") %>%
+  add_header_above(c(" ",  " "," ",  " ", "Sample size" = 2)) %>%
+  kableExtra::footnote(symbol = c("Geographic areas group geographically proximate sites, defined using a hierarchical cluster analysis on the distance matrix of the sites, and a cutoff of 25km"), footnote_as_chunk = FALSE)
+```
 
@@
 ```
-*Data selection.* This analysis focused on nine autotrophic carbon fluxes included in the ForC database (Table 2).
-These were selected based on *...*.
-The ForC database contains annual flux estimates, associated to plots, each with associated data on stand age and known history of disturbance and management, and, more broadly, sites (**Define**).
-Geographically proximate sites are clustered into geographic areas, with proximity defined using a hierarchical cluster analysis on the distance matrix of the sites and a cutoff of 25km [@anderson-teixeira_forc_2018].
 
-A subset of the ForC database was generated for the purposes of this analysis, in order to control for data quality and remove biasing factors.
-Since management can alter observed patterns of FACF [@simova_enigma_2017], sites were excluded from analysis if they were managed, defined as plots that were planted, managed as plantations, irrigated, fertilised or including the term "managed" in their site description.
-Sites that had experienced significant disturbance were also excluded.
-Disturbances that justified site exclusion were major cutting or harvesting, and/or burning, flooding, drought and storm events with site mortality >10% of trees.
-Grazed sites were retained.
-There is evidence that stand age influences patterns of FACF and carbon allocation in forest ecosystems, and can confound relationships between latitude and primary productivity [@delucia_forest_2007; @gillman_latitude_2015].
-To reduce any biasing effects of stand age, we included only stands of known age $\ge$ 100 years and those described by terms such as "mature", "intact", or "old-growth".
+*Climate data*
 
-- excluded alkali absoption and soda lime methods for estimating soil / root respiration
-
-\newpage  
-| Variable | Definition | Components included | Methodologies used | Number of records | Number of geographic areas |
-|--------------|-------------------|-------------|------------------------|-------|-------|
-| $GPP$ | Gross Primary Production | NA | flux partitioning of eddy-covariance; $NPP$+$R_{auto}$ | 243 | 49 |
-| $NPP$ | Net Primary Production | stem, foliage, coarse root, fine root, optionally others (e.g., branch, reproductive, understory) | $ANPP$ + $BNPP$ (majority); $GPP$-$R_{auto}$ | 161 | 56 |
-| $ANPP$ | Aboveground Net Primary Production | stem, foliage, optionally others (e.g., branch, reproductive, understory) |$ANPP_{woody-stem}$ + $ANPP_{foliage}$ (+ others) | 278 | 86 |
-| $ANPP_{woody stem}$ | Woody stem growth component of $ANPP$ | woody stems down to DBH $\le$ 10cm (no branch turnover) | stem growth measurements scaled to biomass using allometries | 264 | 96 |
-| $ANPP_{foliage}$ | Foliage component of $ANPP$ | foliage | litterfall collection (separated into components) | 98 | 49 |
-| $BNPP$  | Belowground Net Primary Production | coarse and fine roots |  coarse roots estimated indirectly using allometries based on aboveground stem increment measures ; fine roots as below | 101 | 48 |
-| $BNPP_{fine root}$ | Fine root component of $BNPP$ | fine roots | measurements combined one or more of the following: soil cores, minirhizotrons, turnover estimates,  root ingrowth cores | 88 | 41 |
-| $R_{auto}$ | Autotrophic Respiration | foliage, stem, and root | chamber measurements of component gas exchange | 22 | 13 |
-| $R_{root}$ | Root Respiration | coarse and fine roots | partitioning of total soil respiration (e.g., through root exclusion), scaling of root gas exchange | 64 | 26 |
-
-Table: Definitions and sample sizes of variables (all annual) used in analysis.
-Geographic areas group geographically proximate sites, defined using a hierarchical cluster analysis on the distance matrix of the sites, and a cutoff of 25km.
-
-*Climate datasets.* Where site-level data on mean annual temperature, mean annual precipitation, and latitude were available in the primary literature, this data was compiled and entered directly into the ForC database.
+ForC contains geographic coordinates associated with each measurement record and, when available, mean annual temperature and mean annual precipitation as reported in the primary literature [@anderson-teixeira_forc_2018].
 Based on the geographic co-ordinates for each site, data on a further 11 climate variables was extracted from five open-access climate datasets: WorldClim [@hijmans_very_2005], WorldClim2 [@fick_worldclim_2017], the Climate Research Unit (CRU) time-series dataset v.
 4.03 [@harris_updated_2014], the Global Aridity Index and Potential Evapotranspiration Climate Database [@trabucco_global_2019], and TerraClimate [@abatzoglou_terraclimate_2018] (see Supplementary Information S1 for details of climate variables).
-Where site-level data was missing for mean annual temperature and/or mean annual precipitation, data was extracted from the WorldClim dataset.
+Where site-level data was missing for mean annual temperature or mean annual precipitation, data was extracted from the WorldClim dataset.
 
@@
 
-*Length of growing season.* Growing season months were defined as months with mean minimum temperature > 0.5$^\circ$C.
+Growing season months were defined as months with mean minimum temperature > 0.5$^\circ$C.
 Growing season months were initially calculated following methods used by Kerkhoff et al.
@@
 To account for correlations in measurements between tightly clustered sites, a random effect was specified as plot nested within geographic area.
+Geographically proximate sites are clustered into geographic areas, with proximity defined using a hierarchical cluster analysis on the distance matrix of the sites and a cutoff of 25km [@anderson-teixeira_forc_2018].
 
@@
 There was a significant additive effect for GPP, ANPP and autotrophic respiration.
-ANPP~foliage~, BNPP, BNPP~fine~ ~root~, and root respiration showed no significant interactive or additive effect (see table S4).
+ANPP~foliage~, BNPP, BNPP~fine~ ~root~, and root respiration showed no significant interactive or additive effect (see table S3).
 

### 2020-02-29 manuscript/manuscript draft.Rmd
worekd on methods

@@
 
-ForC contains geographic coordinates associated with each measurement record and, when available, mean annual temperature and mean annual precipitation as reported in the primary literature [@anderson-teixeira_forc_2018].
-Based on the geographic co-ordinates for each site, data on a further 11 climate variables was extracted from five open-access climate datasets: WorldClim [@hijmans_very_2005], WorldClim2 [@fick_worldclim_2017], the Climate Research Unit (CRU) time-series dataset v.
-4.03 [@harris_updated_2014], the Global Aridity Index and Potential Evapotranspiration Climate Database [@trabucco_global_2019], and TerraClimate [@abatzoglou_terraclimate_2018] (see Supplementary Information S1 for details of climate variables).
+ForC contains geographic coordinates associated with each measurement record and, when available, mean annual temperature (MAT) and mean annual precipitation (MAP) as reported in the primary literature [@anderson-teixeira_forc_2018].
+Based on the geographic co-ordinates for each site, data on twelve climate variables--including MAT, MAP, temperature and precipitation seasonality, annual temperature range, solar radiation, cloud cover, annual frost and wet days, potential evapotranspiration (PET), aridity, and vapor pressure deficit (VPD)--were extracted from five open-access climate datasets: WorldClim [@hijmans_very_2005], WorldClim2 [@fick_worldclim_2017], the Climate Research Unit (CRU) time-series dataset v.
+4.03 [@harris_updated_2014], the Global Aridity Index and Potential Evapotranspiration Climate Database [@trabucco_global_2019], and TerraClimate [@abatzoglou_terraclimate_2018] (Table S1).
 Where site-level data was missing for mean annual temperature or mean annual precipitation, data was extracted from the WorldClim dataset.
 
-Additionally, two climate variables were derived from the above datasets: maximum vapour pressure deficit, defined as the vapour pressure deficit of the month with the largest deficit; and water stress months, defined as the number of months annually where precipitation was lower than potential evapotranspiration (PET; **HML:defined how?**).
-
-Growing season months were defined as months with mean minimum temperature > 0.5$^\circ$C.
+Three additional climate variables were derived from the above datasets.
+Maximum VPD was defined as the VPD of the month with the largest deficit.
+Number of water stress months was defined as the number of months annually where precipitation was lower than PET.
+Length of the growing season was estimated to the nearest month, where growing season months were defined as months with mean minimum temperature > 0.5$^\circ$C.
 Growing season months were initially calculated following methods used by Kerkhoff et al.
@@
 (2014) included an equivalent requirement in their calculation of growing season length.
-However, we found that including this requirement had no effect on the estimates of growing season length, and so chose to exclude it.
+However, we found that including a moisture index had **no** effect on the estimates of growing season length, and so chose to exclude it.
+(**Becky, was it really no effect?
+or minimal?**)
 

### 2020-02-29 manuscript/manuscript draft.Rmd
worked on methods

@@
 
-Monthly data for PET, precipitation, and temperature was downloaded from the Climate Research Unit (CRU) time-series dataset v 4.03 [@harris_updated_2014], and for solar radiation from WorldClim2 [@fick_worldclim_2017], and used to calculate mean monthly PET, precipitation, temperature and solar radiation during the growing season.
+Monthly data for PET, precipitation, and temperature from the CRU dataset v 4.03 [@harris_updated_2014], and solar radiation from WorldClim2 [@fick_worldclim_2017] were used to calculate mean monthly PET, precipitation, temperature and solar radiation during the growing season.
 Total growing season precipitation and solar radiation were also calculated.

### 2020-02-29 manuscript/manuscript draft.Rmd
worked on methods

@@
 
-*FACF data*
+*Forest carbon flux data*
 
@@
 ForC contains geographic coordinates associated with each measurement record and, when available, mean annual temperature (MAT) and mean annual precipitation (MAP) as reported in the primary literature [@anderson-teixeira_forc_2018].
-Based on the geographic co-ordinates for each site, data on twelve climate variables--including MAT, MAP, temperature and precipitation seasonality, annual temperature range, solar radiation, cloud cover, annual frost and wet days, potential evapotranspiration (PET), aridity, and vapor pressure deficit (VPD)--were extracted from five open-access climate datasets: WorldClim [@hijmans_very_2005], WorldClim2 [@fick_worldclim_2017], the Climate Research Unit (CRU) time-series dataset v.
+Based on the geographic co-ordinates for each site, data on twelve climate variables--including MAT, MAP, temperature and precipitation seasonality, annual temperature range, solar radiation, cloud cover, annual frost and wet days, potential evapotranspiration (PET), aridity (MAP/PET), and vapor pressure deficit (VPD)--were extracted from five open-access climate datasets: WorldClim [@hijmans_very_2005], WorldClim2 [@fick_worldclim_2017], the Climate Research Unit (CRU) time-series dataset v.
 4.03 [@harris_updated_2014], the Global Aridity Index and Potential Evapotranspiration Climate Database [@trabucco_global_2019], and TerraClimate [@abatzoglou_terraclimate_2018] (Table S1).
-Where site-level data was missing for mean annual temperature or mean annual precipitation, data was extracted from the WorldClim dataset.
+From these data, we derived maximum VPD, defined as the VPD of the month with the largest deficit, and the number of water stress months, defined as the number of months annually where precipitation was lower than PET.
+Where site-level data was missing for MAT or MAP, we used values from the WorldClim dataset.
 
-Three additional climate variables were derived from the above datasets.
-Maximum VPD was defined as the VPD of the month with the largest deficit.
-Number of water stress months was defined as the number of months annually where precipitation was lower than PET.
 Length of the growing season was estimated to the nearest month, where growing season months were defined as months with mean minimum temperature > 0.5$^\circ$C.
-Growing season months were initially calculated following methods used by Kerkhoff et al.
-(2005), which additionally required that growing season months had a moisture index, defined as (MAT - PET)/PET, > -0.95.
-Michaletz et al.
-(2014) included an equivalent requirement in their calculation of growing season length.
+We experimented with a definion of growing season months including a moisture index, defined as (MAT - PET)/PET, > -0.95 (Kerkhoff et al.
+2005; see also Michaletz et al.
+2014).
 However, we found that including a moisture index had **no** effect on the estimates of growing season length, and so chose to exclude it.
 (**Becky, was it really no effect?
-or minimal?**)
-
-Monthly data for PET, precipitation, and temperature from the CRU dataset v 4.03 [@harris_updated_2014], and solar radiation from WorldClim2 [@fick_worldclim_2017] were used to calculate mean monthly PET, precipitation, temperature and solar radiation during the growing season.
+or minimal?**) Monthly data for PET, precipitation, and temperature from the CRU dataset v 4.03 [@harris_updated_2014], and solar radiation from WorldClim2 [@fick_worldclim_2017] were used to calculate mean monthly PET, precipitation, temperature and solar radiation during the growing season.
 Total growing season precipitation and solar radiation were also calculated.

### 2020-03-01 manuscript/manuscript draft.Rmd
worked on methods

@@
 
-*Model specification.*
+*Analyses*
 
-{**from Helene: It was frequently unclear to me whether the influences of other climate variables were being considered in multivariate or univariate analyses.
-This needs to be clear.
-I can see reasons to do it either way, but they need to be differently motivated and clearly presented.**}
+The effects of latitude and climate on FACF were analysed using mixed effects models using the package 'lme4' [@bates_fitting_2015] in R v.3.5.1 [@r_core_team_r:_2018].
+The basic model for all analyses included a fixed effect of latitude or climate and a random effect of plot nested within geographic area.
+Geographic areas--*i.e.*, spatially clustered sites--are defined within ForC using a hierarchical cluster analysis on the distance matrix of the sites and a cutoff of 25km [@anderson-teixeira_forc_2018].
+We experimented with inclusion of altitude as a fixed effect, but excluded it from the final models because it added very little explanatory power--that is, the change in AIC was generally small (often <2) relative to models excluding altitude.
+Hypotheses were accepted if the AIC difference ($\Delta$AIC) between a model including the fixed effect of interest and a corresponding null model excluding that fixed effect exceeded 2.0.
+All $R^2$ values presented here are marginal $R^2$ values, and refer to the proportion of variation explained by only the fixed effects, **unless otherwise specified**.
+**Becky, are there any instances of this?** Specific analyses are as described below.
 
-The effects of climate and latitude on FACF were analysed using mixed effects models using the package 'lme4' [@bates_fitting_2015] in R v.3.5.1 [@r_core_team_r:_2018].
-First, the individual effect of each climate variable on each flux variable was modelled by specifying the climate variable as a fixed effect in a univariate model.
-For each climate variable, three models were specified: a null model; a model with the climate variable as a linear term; and a model with the climate variable as a second-order polynomial term.
-AIC values were calculated for the models and used to select the best model.
-If the best model included a polynomial term, the shape of the polynomial relationship was considered.
-If the shape of the relationship *made biological sense*, and was a significant improvement on the linear relationship (deltaAIC >2), we accepted the polynomial as the best model.
-If not, we ran the linear model as the final model.
-$R^2$ values were calculated for the best model.
-All $R^2$ values presented here are marginal $R^2$ values, and refer to the proportion of variation explained by only the fixed effects, unless otherwise specified.
-In addition, slope coefficients were calculated for the linear models.
-{**This paragraph needs work - polynomial models were rejected if they showed an exponential increase in productivity with climate (e.g.
-showed GPP increasing exponentially with latitude - needs work on how to word this**}
+We first examined the relationship between latitude and FACF (Q1; Table 1).
+We tested models with latitude as a linear and term (corresponding null: model without latitude) and as a second-order polynomial term (corresponding null: model with latitude as a linear term).
+If the shape of the relationship [**made biological sense**](https://github.com/forc-db/Global_Productivity/issues/64), and was a significant improvement on the linear relationship ($\Delta$AIC >2), we accepted the polynomial as the best model.
+We also examined relationships among fluxes across latitude, testing whether sums of component fluxes matched the larger fluxes and whether C allocation varied with latitude, as specified below.
 
-To test for a potential influence of altitude, models were also run with site altitude included as a second fixed effect.
-These multivariate models were compared against univariate models with no altitude term, and AIC values calculated to identify whether inclusion of altitude as a term improved the models.
-Including altitude had a very small effect on most models, with the difference in AIC values between models including and excluding altitude often being <2, suggesting the models are very similar in their explanatory power.
-As a result, it was decided not to include altitude as a fixed term in the final models.
+**work on this paragraph** 
+*Validating models of component fluxes.* Comparison of component fluxes is based on the assumption that components sum accurately to estimates of larger fluxes.
+To test this, components of larger fluxes were regressed against latitude, and the models used to generate a series of point estimates along lines of best fit for each component.
+The point estimates for smaller component fluxes were summed to generate new "stacked" estimates of larger fluxes, which were then compared against actual measurements of the larger flux.
+Confidence intervals for the larger flux were calculated using the 'bootMer' function from the lme4 package [@bates_fitting_2015].
+Stacked plots were generated for:
+1. GPP = NPP + R~auto~  
+2. NPP = ANPP + BNPP  
+3. ANPP = ANPP~foliage~ + ANPP~woody~ ~stem~  
+4. Total belowground carbon flux = BNPP + R~root~
 
-To account for correlations in measurements between tightly clustered sites, a random effect was specified as plot nested within geographic area.
-Geographically proximate sites are clustered into geographic areas, with proximity defined using a hierarchical cluster analysis on the distance matrix of the sites and a cutoff of 25km [@anderson-teixeira_forc_2018].
+**work on this paragraph**   
+*Allocation to carbon fluxes along latitudinal gradients.* Variation in allocation to component carbon fluxes along latitudinal gradients was explored for a range of pairings: firstly, GPP:NPP, ANPP:BNPP, and ANPP~foliage~:ANPP~woody~ ~stem~; and secondly, the ratio of NPP to each of ANPP, BNPP, ANPP~foliage~, and ANPP~woody~ ~stem~.
+For each set of paired fluxes, measurements taken at the same site and plot, and in the same year, were paired together, and the ratio of each pair of measurements calculated.
+The ratios were regressed against latitude and climate variables, using the linear model specified above.
+Cook's distance analyses were carried out for each of the models, and indicated that data from a few high-elevation sites were having a disproportionate influence on the regressions.
+To account for this, models were re-run using only data from sites $\le$ 1000m.
 
-Models were run for total annual flux against annual climate variables, and for fluxes standardized by growing season length, defined as annual flux/length of growing season, against growing season climate variables.
-For analyses on data within the growing season, only linear models were specified.
+**work on this paragraph**  
+**We next examined the relationships of FACF to climate variables (Q2-Q4; Table 1)...** As with latitude, we tested booth linear and polynomial fits for each climate variable.
 
 {**Note from Helene on the following paragraph: this makes it sound like analyses from H3 were about variation remaining after accounting for temperature, but analyses below don't seem to do that.**}
+**work on this paragraph**  
 To investigate the potential interactive effects of climate variables on carbon fluxes, multivariate models were also specified.
@@
 
-*Validating models of component fluxes.* Comparison of component fluxes is based on the assumption that components sum accurately to estimates of larger fluxes.
-To test this, components of larger fluxes were regressed against latitude, and the models used to generate a series of point estimates along lines of best fit for each component.
-The point estimates for smaller component fluxes were summed to generate new "stacked" estimates of larger fluxes, which were then compared against actual measurements of the larger flux.
-Confidence intervals for the larger flux were calculated using the 'bootMer' function from the lme4 package [@bates_fitting_2015].
-Stacked plots were generated for:
-1. GPP = NPP + R~auto~  
-2. NPP = ANPP + BNPP  
-3. ANPP = ANPP~foliage~ + ANPP~woody~ ~stem~  
-4. Total belowground carbon flux = BNPP + R~root~
-
-*Allocation to carbon fluxes along latitudinal gradients.* Variation in allocation to component carbon fluxes along latitudinal gradients was explored for a range of pairings: firstly, GPP:NPP, ANPP:BNPP, and ANPP~foliage~:ANPP~woody~ ~stem~; and secondly, the ratio of NPP to each of ANPP, BNPP, ANPP~foliage~, and ANPP~woody~ ~stem~.
-For each set of paired fluxes, measurements taken at the same site and plot, and in the same year, were paired together, and the ratio of each pair of measurements calculated.
-The ratios were regressed against latitude and climate variables, using the linear model specified above.
-Cook's distance analyses were carried out for each of the models, and indicated that data from a few high-elevation sites were having a disproportionate influence on the regressions.
-To account for this, models were re-run using only data from sites $\le$ 1000m.
+**work on this paragraph**  
+Models were run for total annual flux against annual climate variables, and for fluxes standardized by growing season length, defined as annual flux/length of growing season, against growing season climate variables.
+For analyses on data within the growing season, only linear models were specified.
 
@@
 
-In total, we analyzed 1228 records from 9 C flux variables taken from forests that had experienced no major anthropogenic disturbances within the past 100 years.
+In total, we analyzed 1228 records from nine FACF variables taken from forests that had experienced no major anthropogenic disturbances within the past 100 years.
 These records represented a total of 154 distinct geographic areas (Fig.

### 2020-03-01 manuscript/manuscript draft.Rmd
worked on methods

@@
 
-**work on this paragraph** 
-*Validating models of component fluxes.* Comparison of component fluxes is based on the assumption that components sum accurately to estimates of larger fluxes.
-To test this, components of larger fluxes were regressed against latitude, and the models used to generate a series of point estimates along lines of best fit for each component.
-The point estimates for smaller component fluxes were summed to generate new "stacked" estimates of larger fluxes, which were then compared against actual measurements of the larger flux.
+To test whether trends in component fluxes across latitude sum to match those of larger fluxes, regression lines for smaller component fluxes were summed to generate new estimates of larger fluxes, which were then compared against the latitudinal regression of the larger flux.
 Confidence intervals for the larger flux were calculated using the 'bootMer' function from the lme4 package [@bates_fitting_2015].
-Stacked plots were generated for:
-1. GPP = NPP + R~auto~  
-2. NPP = ANPP + BNPP  
-3. ANPP = ANPP~foliage~ + ANPP~woody~ ~stem~  
-4. Total belowground carbon flux = BNPP + R~root~
+This analysis was applied to following sets of fluxes: (1) $GPP = NPP + R_{auto}$, (2) $NPP = ANPP + BNPP$, and (3) $ANPP = ANPP_{foliage} + ANPP_{woody.stem}$.
+In addition, we estimated total belowground C flux (TBCF, not analyzed due to limited data) as $TBCF = BNPP + R_{root}$.
 
-**work on this paragraph**   
-*Allocation to carbon fluxes along latitudinal gradients.* Variation in allocation to component carbon fluxes along latitudinal gradients was explored for a range of pairings: firstly, GPP:NPP, ANPP:BNPP, and ANPP~foliage~:ANPP~woody~ ~stem~; and secondly, the ratio of NPP to each of ANPP, BNPP, ANPP~foliage~, and ANPP~woody~ ~stem~.
+Variation in allocation to component carbon fluxes along latitudinal gradients was explored for the following pairings: $GPP:NPP$, $ANPP:BNPP$, and $ANPP_{foliage}:ANPP_{woody.stem}$; $ANPP_{foliage}:NPP$, $ANPP_{woody.stem}:NPP$, $ANPP:NPP$, and $BNPP:NPP$.
 For each set of paired fluxes, measurements taken at the same site and plot, and in the same year, were paired together, and the ratio of each pair of measurements calculated.
@@
 
-**work on this paragraph**  
-**We next examined the relationships of FACF to climate variables (Q2-Q4; Table 1)...** As with latitude, we tested booth linear and polynomial fits for each climate variable.
+We next examined the relationships of FACF to climate variables (Q2-Q4; Table 1).
+As with latitude, we tested both linear and polynomial fits for each climate variable.
+We tested relationships of each FACF (Table 2) against each climate variable (Table S1), but focus presentation--including focal hypotheses (Table 1)--on climate variables that explained >20% of variation in FACF.
+**Becky, please make this more specific.
+I drafted this based on the results, but its not very clear.**
 
@@
 **work on this paragraph**  
-To investigate the potential interactive effects of climate variables on carbon fluxes, multivariate models were also specified.
-To ensure that models were biologically meaningful, the terms included in the models tested built on results from the univariate models.
-Modelling of individual climate variables identified that the best predictors of carbon fluxes were variables related to temperature.
-We therefore decided to include mean annual temperature as a term in all multivariate models.
-We first modelled the interaction effect between mean annual temperature and mean annual precipitation, in order to capture climate variation along the axes of temperature and water availability.
-Models were tested for a significant interactive effect and a significant additive effect.
-We then explored whether any other climate variable, in combination with mean annual temperature, could significantly improve on the combination of mean annual temperature and mean annual precipitation.
-In specifying the range of models to test, climate variables which were strongly correlated with temperature were dropped, in order to capture the greatest range of variation in climate.
-For each possible pairing of climate variables, two models were specified: a model with the two climate variables showing an additive effect; and a model with the two climate variables showing an interactive effect.
-As before, plot nested within geographic area was included as a random effect.
-Altitude was not considered.
-AIC values were calculated for the models, and used to compare models.
-Models were considered to be significantly better than the baseline MAT*MAP model if:
-  i) the AIC value of the model was smaller than the AIC value of the baseline model by >2  
-  ii) the r-squared value of the model was larger than the r-squared value of the baseline model by >0.05
+To investigate the potential joint and interactive effects of climate variables on carbon fluxes, multivariate models were also specified.
+We explored a large range of climate variable combinations and ultimately decided to focus on MAT and MAP.
+We favored these over variables with similar explanatory power because they are the most commonly reported climate variables and are recorded in ForC.
+MAT was consistently among the best individual predictors of FACF (Table S2), although it was occasionally out-performed by closely correlated climate variables (Fig.
+S2).
+In combination with MAT as a primary climate variable, MAP was consistently among the best secondary climate variables.
+Specifically, compared again a model including MATxMAP, there were no other variables that, in interaction with MAT, improved AIC by >2.0 and r-squared by >0.05.
 

### 2020-03-01 manuscript/manuscript draft.Rmd
worked on methods (finished revision)

@@
 Geographic areas--*i.e.*, spatially clustered sites--are defined within ForC using a hierarchical cluster analysis on the distance matrix of the sites and a cutoff of 25km [@anderson-teixeira_forc_2018].
-We experimented with inclusion of altitude as a fixed effect, but excluded it from the final models because it added very little explanatory power--that is, the change in AIC was generally small (often <2) relative to models excluding altitude.
-Hypotheses were accepted if the AIC difference ($\Delta$AIC) between a model including the fixed effect of interest and a corresponding null model excluding that fixed effect exceeded 2.0.
+We experimented with inclusion of altitude as a fixed effect, but excluded it from the final models because it added very little explanatory power--that is, the difference in AIC ($\Delta$AIC) relative to models excluding altitude was generally small (often $\Delta$AIC<2).
+Hypotheses were accepted if the $\Delta$AIC between a model including the fixed effect of interest and a corresponding null model excluding that fixed effect exceeded 2.0.
 All $R^2$ values presented here are marginal $R^2$ values, and refer to the proportion of variation explained by only the fixed effects, **unless otherwise specified**.
@@
 
-{**Note from Helene on the following paragraph: this makes it sound like analyses from H3 were about variation remaining after accounting for temperature, but analyses below don't seem to do that.**}
-**work on this paragraph**  
 To investigate the potential joint and interactive effects of climate variables on carbon fluxes, multivariate models were also specified.
@@
 In combination with MAT as a primary climate variable, MAP was consistently among the best secondary climate variables.
-Specifically, compared again a model including MATxMAP, there were no other variables that, in interaction with MAT, improved AIC by >2.0 and r-squared by >0.05.
+Specifically, compared again a model including MAT x MAP, there were no other variables that, in interaction with MAT, improved AIC by >2.0 and $R^2$ by >0.05.
+In examinging the joint and interactive effects of climate variables, an additive model including MAT and MAP as fixed effects was accepted when $\Delta$AIC >2 relative to a null including only MAT as a fixed effect.
+An interactive model including an MAT x MAP interaction was accepted when $\Delta$AIC >2 relative to a null including MAT and MAP as fixed effects.
 
-**work on this paragraph**  
-Models were run for total annual flux against annual climate variables, and for fluxes standardized by growing season length, defined as annual flux/length of growing season, against growing season climate variables.
+To test whether and how FACF varied with climate when standardised by growing season length (Q5), we first standardized FACF by dividing by growing season length (as defined above).
+We then tested for correlations between growing season length-standardised FACF against growing season climate variables.
 For analyses on data within the growing season, only linear models were specified.

### 2020-03-01 manuscript/manuscript draft.Rmd
worked on results

@@
 Additional targeted literature searches were conducted to identify any further available data on the FACF analyzed here, with particular focus on mature forests in temperate and boreal regions, which were not included in the review of @anderson-teixeira_carbon_2016.
-We used ForC v3.0, archived on Zenodo with DOI 10.5281/zenodo.3403855.This version contained 29,730 records from 4,979 plots, representing 20 distinct ecozones across all forested biogeographic and climate zones.
+We used ForC v3.0, archived on Zenodo with DOI 10.5281/zenodo.3403855.
+This version contained 29,730 records from 4,979 plots, representing 20 distinct ecozones across all forested biogeographic and climate zones.
 
@@
 
-All major carbon fluxes decreased linearly with latitude (Fig. 2).
-
-Latitude was a strong predictor for many of the carbon fluxes, explaining 64% of variation in GPP (n = 254, p<0.0001), 50% in NPP (n = 114, p<0.0001) and 45% in ANPP (n = 259, p<0.0001).
-For all fluxes, their relationship with latitude was best predicted by the linear model.
-
-*Relationships and differences among fluxes.* In general, smaller component fluxes summed approximately to larger fluxes across the latitudinal gradient (fig.
-2).
-That is, modelled estimates of GPP, generated from the sum of NPP and R auto; NPP, generated from the sum of ANPP and BNPP~root~; and ANPP, generated from the sum of ANPP~foliage~ and ANPP~woody~ ~stem~, fell completely within the confidence intervals of the regressions of field estimates of GPP, NPP and ANPP respectively.
-
-bbl: here or in the discussion, note that this is a fairly stringent test: it's easy for sub-fluxes not to sum up!
-(Extensive citations from EC literature for example.).
-HML: yes and no.
-confidence intervals are pretty large.
+All major carbon fluxes decreased linearly with latitude (Fig.
+2; Table S2).
+Latitude was a strong predictor for many of the carbon fluxes--particularly the larger fluxes (Table S2).
+Specifically, latitude explained 64% of variation in GPP (n = 254, p<0.0001), 50% in NPP (n = 114, p<0.0001) and 45% in ANPP (n = 259, p<0.0001).
+The FACF that were most poorly predicted by latitude were $BNPP_{root-fine}$ ($R^2$=0.17) and $ANPP_{woody.stem}$=0.18).
+For all FACF, the relationship with latitude was best predicted by the linear model (**but see [this issue]((https://github.com/forc-db/Global_Productivity/issues/64))**).
 
@@
 
-*We found no evidence that allocation between fluxes varied substantially with latitude or climate.
-There were no significant results from regressing ratios of carbon fluxes against latitude, or against any of the climate variables.* **THIS HAS CHANGED: [issue #51](https://github.com/forc-db/Global_Productivity/issues/51#issuecomment-581892186)
+In general, smaller component fluxes summed approximately to larger fluxes across the latitudinal gradient (Fig.
+2).
+That is, modelled estimates of $GPP$, generated from the sum of $NPP$ and $R_{auto}$; $NPP$, generated from the sum of $ANPP$ and $BNPP$; and $ANPP$, generated from the sum of $ANPP_{foliage}$ and $ANPP_{woody-stem}$, fell completely within the confidence intervals of the regressions of field estimates of $GPP$, $NPP$, and $ANPP$, respectively.
+
+We found little evidence that allocation between fluxes varied substantially with latitude or climate (Fig.
+[**S3**](https://github.com/forc-db/Global_Productivity/issues/65)).
+Of the 7 FACF ratios regressed against latitude and three climate variables (MAT, MAP, temperature seasonality), there were only four signficant relationships, all with $R^2 \le 0.2$ (Fig.
+S3).
+Specifically, the proportion of NPP allocated to $ANPP_{foliage}$ increased weakly with MAT ($R^2=0.20$), and the proportion of $NPP$ allocated aboveground ($ANPP$) decreased weakly with latitude ($R^2=0.11$) and temperature seasonality ($R^2=0.17$), while increasing with MAT ($R^2=0.11$).
+There were no significant results from regressing ratios of carbon fluxes against latitude, or against any of the climate variables.
+
+(*KAT has revised results up to here.*)
 

### 2020-03-01 manuscript/manuscript draft.Rmd
worked on results

@@
 
+*How do FACF relate to MAT and MAP?*
+
+All FACF increased linearly with MAT, and we found no support for a saturation point of FACF with MAT (all p<0.05; Figs.
+3-4, S4-S5, Table S2).
+As with latitude, MAT tended to explain more variation in the larger FACF ($GPP$, $NPP$, $ANPP$, $R_{auto}$) and $ANPP_foliage$ (all $R^2$> 0.4) than in subsidiary and belowground fluxes ($ANPP_{woody-stem}$, $R_{root}$, $BNPP_{root-fine}$; all $R^2$< 0.25).
+
+MAP was a significant (p<0.05) predictor of all FACF but $ANPP_{woody-stem}$ (Figs.
+4a, S4-S5; Table S2).
+However, it explained little variation: with the exception of $R_{auto}$, MAP explained at most 37% of variation in FACF.
+For the majority of FACF, a polynomial model was the best fit.
+FACF generally increased with precipitation, up until a saturation point at between 3000 and 4000mm annual precipitation, above which they started to decrease (Figs.
+4, S4-S5).
+The notable exception to this was GPP: the model indicated that GPP continued to increase with precipitation up to measures of at least 5000mm annually (p<0.0001, $R^2$ = 0.33.
+Data above this point were not available, but the model trend indicated that the saturation point for this model would be around 5000mm MAP.
+
 (*KAT has revised results up to here.*)
 
-$R^2$ values were generally highest in the major fluxes, and decreased in subsidiary fluxes (Supporting Information S2).
-Of the major fluxes, R~auto~ and GPP were the most strongly explained by latitude and climate, with climate explaining at most 71% of variation in GPP, and 65% in R~auto~.
-The proportion of variation explained by climate and latitude decreased in NPP and ANPP, with climate explaining at most 51% of variation in NPP and 44% in ANPP.
-Of the major fluxes, BNPP~root~ was the least well explained by climate and latitude, with climate explaining at most 36% of variation.
-
-With the exception of ANPP~foliage~, the proportion of variation explained by climate and latitude in subsidiary fluxes was much lower.
-Climate explained at most 24% of the variation in ANPP~woody~ ~stem~, 19% in BNPP~fine~ ~root~, and 27% in R~root~.
-In contrast, climate strongly explained variation in ANPP~foliage~, with mean annual temperature explaining 58% of variation.
-This pattern was also seen in the $R^2$ values for multivariate models.
-
-*How do FACF relate to MAT and MAP?* MAT and MAP are the most commonly reported site-level climate variables, and much previous research into the effect of climate on FACF has focused on these as key climate variables.
-MAT was a significant (p<0.05) and strong ($R^2 \ge$ **##**) predictor of all FACF tested.
-All FACF increased linearly with temperature (Fig.
-4; all p<0.05, $R^2 \ge$ **##**), and we found no support for a saturation point of FACF with temperature.
-
-MAP was found to be a significant (p<0.05) but poor predictor of FACF, explaining, with the exception of R~auto~, at most 37% of variation in carbon flux.
-For the majority FACF, a polynomial model was the best fit.
-FACF increased with precipitation, up until a saturation point at between 3000 and 4000mm annual precipitation, above which FACF started to decrease (Fig.
-4).
-The notable exception to this was GPP: the model indicated that GPP continued to increase with precipitation up to measures of at least 5000mm annually (p<0.0001, $R^2$ = 0.33.
-Data above this point was not available, but the model trend indicated that the saturation point for this model would be around 5000mm MAP.
-
-There was a significant interactive effect between MAT and MAP for NPP and ANPP~woody~ ~stem~ (fig.3).
+There was a significant interactive effect between MAT and MAP for NPP and $ANPP_{woody-stem}$ (fig.3).
 There was a significant additive effect for GPP, ANPP and autotrophic respiration.
-ANPP~foliage~, BNPP, BNPP~fine~ ~root~, and root respiration showed no significant interactive or additive effect (see table S3).
+$ANPP_{foliage}$, BNPP, $BNPP_{root-fine}$, and root respiration showed no significant interactive or additive effect (see table S3).
 
@@
 We found a significant relationship between C flux and potential evapotranspiration for all FACF.
-ANPP~foliage~, BNPP~fine~ ~root~ and R~root~ increased linearly with PET; however, all other fluxes showed a polynomial relationship with PET (fig.
+$ANPP_{foliage}$, $BNPP_{root-fine}$ and $R_{root}$ increased linearly with PET; however, all other fluxes showed a polynomial relationship with PET (fig.
 4).
@@
 Vapour pressure deficit was a significant predictor of C flux for all FACF.
-BNPP~fine~ ~root~ showed a linear relationship with vapour pressure deficit ($R^2$ = 0.07, p<0.05), but all other fluxes showed a polynomial relationship (fig.
+$BNPP_{root-fine}$ showed a linear relationship with vapour pressure deficit ($R^2$ = 0.07, p<0.05), but all other fluxes showed a polynomial relationship (fig.
 4).
@@
 
-All fluxes, with the exception of R~root~, showed a positive linear relationship with solar radiation.
-Solar radiation explained a low proportion of variability in all FACF, explaining less than 20% of the variation in each flux, with the exception of R~auto~ ($R^2$ = 0.26, p<0.05).
+All fluxes, with the exception of $R_{root}$, showed a positive linear relationship with solar radiation.
+Solar radiation explained a low proportion of variability in all FACF, explaining less than 20% of the variation in each flux, with the exception of$R_{auto}$ ($R^2$ = 0.26, p<0.05).
 
@@
 *What is the role of seasonality in explaining FACF?* Temperature seasonality was a significant predictor of FACF.
-GPP, NPP, ANPP, and R~root~ exhibited a polynomial relationship with seasonality (Fig.
+GPP, NPP, ANPP, and $R_{root}$ exhibited a polynomial relationship with seasonality (Fig.
 4) {**give statistic on signficiance**}.
-ANPP~foliage~, ANPP~woody~ ~stem~ and R~auto~ decreased linearly with temperature seasonality.
+$ANPP_{foliage}$, $ANPP_{woody-stem}$ and $R_{auto}$ decreased linearly with temperature seasonality.
 Temperature seasonality was strongly correlated with annual temperature range, and, as expected, all fluxes showed almost identical responses to it {**meaning unclear**}.

### 2020-03-02 manuscript/manuscript draft.Rmd
edited results (finished)

@@
 
-(*KAT has revised results up to here.*)
-
-There was a significant interactive effect between MAT and MAP for NPP and $ANPP_{woody-stem}$ (fig.3).
-There was a significant additive effect for GPP, ANPP and autotrophic respiration.
-$ANPP_{foliage}$, BNPP, $BNPP_{root-fine}$, and root respiration showed no significant interactive or additive effect (see table S3).
-
+There was a significant additive effect of MAT and MAP on $GPP$, $ANPP$ and $R_{auto}$ (Fig.
+3, Table S3).
+Accounting for MAT, MAP had a substantial positive effect on $GPP$ and $R_{auto}$ and a small negative effect on $ANPP$.
+There was a significant interactive effect between MAT and MAP for $NPP$ and $ANPP_{woody-stem}$ (Fig.
+3, Table S3).
+The interaction was negative for $NPP$ and positive for $ANPP_{woody-stem}$.
+For $ANPP_{foliage}$, $BNPP$, $BNPP_{root-fine}$, and $R_{auto-root}$, MAP did not have a signficant effect when accounting for MAT (Fig.
+3, Table S3).
 For the variables which showed a significant interactive or additive effect between MAT and MAP, no other climate variable, in combination with MAT, significantly improved on that model.
-{**need to confirm this given changes in MAT MAP results**}
+{**need to confirm this given changes in MAT MAP results (or you could just drop the sentence.)**}
 
@@
 
-*How do FACF relate to other climate variables?* {**Here, maybe add some statement about covariation among climate variables, site SI figure**} Our results indicated that FACF were most strongly explained by temperature at the global scale, with temperature-related climate variables coming out as strong predictors of FACF.
-In addition to MAT, temperature seasonality, annual temperature range, and annual frost days were consistently identified as strong univariate predictors of FACF.
-{**cite some table/ SI?**}
+*How do FACF relate to other climate variables?*
+
+Our results indicated that FACF were most strongly explained by temperature at the global scale, with temperature-related climate variables coming out as strong predictors of FACF.
+In addition to MAT, several of its correlates (Fig.
+S2) were consistently identified as strong univariate predictors of FACF: temperature seasonality, annual temperature range, annual frost days, PET, and length of growing season (Figs.
+4, S4-S7).
 
 We found a significant relationship between C flux and potential evapotranspiration for all FACF.
-$ANPP_{foliage}$, $BNPP_{root-fine}$ and $R_{root}$ increased linearly with PET; however, all other fluxes showed a polynomial relationship with PET (fig.
-4).
+$ANPP_{foliage}$, $BNPP_{root-fine}$ and $R_{root}$ increased linearly with PET; however, all other fluxes showed a polynomial relationship with PET (Fig.
+4c, S4-5; Table S2).
 We found strong evidence for a saturation point or peak with PET: FACF tended to increase at values below 1000mm, before saturating between 1200 and 1700mm.
@@
 Vapour pressure deficit was a significant predictor of C flux for all FACF.
-$BNPP_{root-fine}$ showed a linear relationship with vapour pressure deficit ($R^2$ = 0.07, p<0.05), but all other fluxes showed a polynomial relationship (fig.
-4).
-FACF initially increased with vapour pressure deficit, before saturating at around 0.8 kPa.
-At values above 0.8 kPa, FACF began to decrease.
+$BNPP_{root-fine}$ showed a linear relationship with vapour pressure deficit ($R^2$ = 0.07, p<0.05), but all other fluxes showed a polynomial relationship (Figs.
+4d, S4-5; Table S2).
+FACF initially increased with vapour pressure deficit, before saturating at around 0.8 kPa, after wich point they began to decrease.
 
-All fluxes, with the exception of $R_{root}$, showed a positive linear relationship with solar radiation.
+All fluxes, with the exception of $R_{root}$, showed a positive linear relationship with solar radiation (Figs.
+S4-S5, Table S2).
 Solar radiation explained a low proportion of variability in all FACF, explaining less than 20% of the variation in each flux, with the exception of$R_{auto}$ ($R^2$ = 0.26, p<0.05).
 
-Of the climate variables tested, annual wet days, aridity, cloud cover, precipitation seasonality, maximum vapour pressure deficit and water stress months were poor or non-significant explainers of variation in FACF, explaining less than 20% of the variation in each of the carbon fluxes.
+Aannual wet days, cloud cover, and aridity were poor or non-significant explainers of variation in FACF, explaining less than 20% of the variation in each of the carbon fluxes (Figs. S4-S5; Table S2).
 
@@
 
-*What is the role of seasonality in explaining FACF?* Temperature seasonality was a significant predictor of FACF.
-GPP, NPP, ANPP, and $R_{root}$ exhibited a polynomial relationship with seasonality (Fig.
-4) {**give statistic on signficiance**}.
-$ANPP_{foliage}$, $ANPP_{woody-stem}$ and $R_{auto}$ decreased linearly with temperature seasonality.
-Temperature seasonality was strongly correlated with annual temperature range, and, as expected, all fluxes showed almost identical responses to it {**meaning unclear**}.
-FACF were highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower.
-In contrast, there was no significant effect of precipitation seasonality on FACF.
+*What is the role of seasonality in explaining FACF?*
 
-We found a significant relationship between length of growing season and FACF, with all fluxes showing a linear increase with length of growing season (Fig.
-4).
+Temperature seasonality was a significant predictor of FACF.
+$GPP$, $NPP$, $ANPP$, and $R_{root}$ exhibited a polynomial relationship with seasonality (all p<0.05; Figs.
+4e, S6-7; Table S2).
+$ANPP_{foliage}$, $ANPP_{woody-stem}$ and $R_{auto}$ decreased linearly with temperature seasonality (all p<0.05; Figs.
+4e, S6-S7; Table S2).
+Temperature seasonality was strongly correlated with annual temperature range, which was likewise a similarly strong predictor of FACF (Table S2).FACF were highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower.
+
+In contrast, there was no significant effect of precipitation seasonality on FACF, and both maximum vapour pressure deficit, and water stress months were poor or non-significant explainers of variation in FACF (Figs.
+S6-S7; Table S2).
+
+We found a significant relationship between length of growing season (temperature-defined) and FACF, with all fluxes showing a linear increase with length of growing season (Figs.
+4e, S6-S7; Table S2).
 Length of growing season was a strong predictor of FACF, explaining 51% of variation in GPP, 39% of variation in NPP, and 34% of variation in ANPP, but it was a weaker predictor than MAT for all fluxes analysed.
 
-*Within the growing season, how do FACF vary with climate?* When FACF were standardized by growing season length, correlations with climate variables were much weaker.
-Growing season temperature, precipitation, solar radiation, and PET each had a small effect on one or more of the nine FACF analyzed (Table 2; Fig.
-**S#**).
+*Within the growing season, how do FACF vary with climate?*
+
+When FACF were standardized by growing season length, correlations with growing season climate--includingtemperature, precipitation, solar radiation, and PET--were generally weak (Table S4, Figs.
+S8-S9).
 Speficifally, $ANPP$ increased with growing season temperature ($R^2$ = 0.10, p<0.001) and precipitation ($R^2$ = 0.04, p<0.05).

### 2020-03-02 manuscript/manuscript draft.Rmd
condensed first par of discussion

@@
 
-We found a significant relationship between length of growing season (temperature-defined) and FACF, with all fluxes showing a linear increase with length of growing season (Figs.
+We found a significant relationship between length of growing season and FACF, with all fluxes showing a linear increase with length of growing season (Figs.
 4e, S6-S7; Table S2).
-Length of growing season was a strong predictor of FACF, explaining 51% of variation in GPP, 39% of variation in NPP, and 34% of variation in ANPP, but it was a weaker predictor than MAT for all fluxes analysed.
+Length of growing season was a strong predictor of FACF, explaining 51% of variation in GPP, 39% of variation in NPP, and 34% of variation in ANPP, but it was a weaker predictor than MAT for all fluxes analysed (Table S4).
 
@@
 
-When FACF were standardized by growing season length, correlations with growing season climate--includingtemperature, precipitation, solar radiation, and PET--were generally weak (Table S4, Figs.
+When FACF were standardized by growing season length, correlations with growing season climate--including temperature, precipitation, solar radiation, and PET--were generally weak (Figs.
 S8-S9).
@@
 
-We used a large global database (ForC), containing an unprecedented amount of data representing all major forest types (Fig.
-1) and the nine most significant forest autotrophic carbon fluxes (FACF) variables (Table 2), to comprehensively explore how C cycling in mature, undisturbed forests varies with latitude and climate on a global scale.
-We test a suite of hypotheses, including many previously posed (Table 1), standardising for factors that have not always been controlled for (*e.g.*, stand age, flux components measured) to gain a stronger understanding of climatic effects on FACF.
-
-bbl: Most of these two paragraphs (following) just restate the results. Honestly I would remove them enitrely, but at the very least, condense into a single paragraph.
-
+Our analysis of a large global database (ForC) reveals how autotrophic carbon fluxes in mature forests vary with latitude and climate on a global scale.
 We show that, across all nine FACF analyzed, C cycling decreases linearly with latitude (*H1.1*; Fig.
 2)--a finding that confirms multiple previous studies (**REFS**) but contradicts the idea that productivity of temperate forests rivals that of tropical forests (*H1.1.alt*; **REFS**).
-The FACF increase in proportion to one another (*H1.2*), with no differences in allocation detectable at this global scale and with component fluxes summing appropriately to larger fluxes (Fig.
+The FACF generally increase in proportion to one another (*H1.2*), with few differences in allocation detectable at this global scale (Fig.
+S2) and with component fluxes summing appropriately to larger fluxes (Fig.
 2), indicating no major, systematic omissions or overestimations of flux components.
-However, we did detect a tendency for greater variability among subsidiary C fluxes (e.g., *$ANPP_{woody}$*, *$BNPP_{fine.root}$*; Fig.
-2; **some SI table?**).
-
-**Climate explains a significant proportion of variation in all C fluxes analyzed** (albiet less for subsidiary fluxes), with temperature variables the best predictors of FACF at this global scale (Figs.
+However, climate explained lower proportions of variability among subsidiary C fluxes (e.g., *$ANPP_{woody}$*, *$BNPP_{fine.root}$*, $R_{auto-root}$; Fig.
+2; Table S2).
+Latitudianal variation in FACF is primarily attributable to temperature-related variables (*H3, H4*), particularly MAT (Figs.
 3-4).
-While other climate variables are significant predictors of FACF (*H3*), none of them in isolation improve on the explanatory power of temperature-related variables in general or MAT specifically (Fig.
-4).
-Water availability is an important factor in explaining FACF on a global scale: we find a positive influence of precipitation at low MAP, with saturation at higher levels of MAP (Fig.
-4b) and a significant interaction between MAT and MAP for seven of the nine variables analyzed (Fig.
-3).
-Temperature seasonality and growing season length are closely correlated with MAT and are strong predictors of FACF (*H4*; Fig.
-4e-f), though growing season length doesn't improve upon MAT as a predictor.
-Within the growing season, the influence of climate on C cycling is smaller but still significant for a number of carbon fluxes (*H5*; **some SI table?**).
+Water availability is also influential, but generally of secondary importance across the range represented in our database (Figs.
+3-4).
+Temperature seasonality and growing season length are closely correlated with MAT and are strong predictors of FACF (*H4*; Figs.
+4e-f, S2, S6-S7), though growing season length doesn't improve upon MAT as a predictor.
+Within the growing season, the influence of climate on C cycling is smaller but still significant for a number of FACF (*H5*; Fig.
+S9; Table S4).
 These findings clarify the big picture of how FACF vary with latitude and climate on a global scale.

### 2020-03-10 manuscript/manuscript draft.Rmd
misc

@@
                   \captionsetup[figure]{font=footnotesize}
+                  \usepackage{setspace}\onehalfspacing
+                  \usepackage{lineno}\linenumbers
 
@@
 In contrast, Larjavaara and Muller-Landau [-@larjavaara_temperature_2012], find that increases in GPP saturate at approximately 25$^\circ$C MAT, while Schuur [-@schuur_productivity_2003] shows that NPP increases linearly with temperature.
-bbl:maybe remove following sentences The influence of these climate variables on productivity is further complicated by the possibility of interactive effects occurring between them.
+**bbl:maybe remove following sentences** The influence of these climate variables on productivity is further complicated by the possibility of interactive effects occurring between them.
 Taylor et al.
-[-@taylor_temperature_2017] showed that high rainfall had a negative effect on productivity where MAT was low, but a positive effect at sites with high MAT.
-Similarly, at sites with high MAT, increases in rainfall were found to enhance productivity, in contrast to cooler sites, where high levels of rainfall inhibit productivity.
+[-@taylor_temperature_2017] showed a positive interaction between MAT and MAP in shaping tropical forest productivity, such that high rainfall had a negative effect on productivity in cooler climates, compared to a positive effect in warmer climates.
 {**It would be good to add some more citations on soil respiraiton.
@@
 
-Aannual wet days, cloud cover, and aridity were poor or non-significant explainers of variation in FACF, explaining less than 20% of the variation in each of the carbon fluxes (Figs. S4-S5; Table S2).
+Annual wet days, cloud cover, and aridity were poor or non-significant explainers of variation in FACF, explaining less than 20% of the variation in each of the carbon fluxes (Figs. S4-S5; Table S2).
 
@@
 
+## misc content for discussion
+- the observed positive interaction between MAT and MAP for $ANPP_woody$ is consitent with the @taylor_temperature_2017 analysis showing such an interaction for $ANPP$ in tropical forests.
+Similar to their analysis, we find a cross-over point at ~20C.
+However, we don't find such an interaction for $ANPP$, and we show a contrasting negative interaction for $NPP$.
+Some of this is may be stochasitic/ driven by composition of the dataset, and the interactions we observe are not internally consistent.
+
+- consistent with [Hofhansl et al.
+2015](https://agupubs.onlinelibrary.wiley.com/doi/full/10.1002/2014GB004934) (**verify**) , we found a slight tendency for warmer sites to have higher aboveground allocation
+
+-results are consistent with *Muller-Landau et al., in review*
+
 ### Acknowledgements

### 2020-03-10 manuscript/appendix.Rmd
rearranged SI fig. order
#65

@@
 \elandscape
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S3: Ratios among FACF as a function of latitude and climate variables", out.width ='100%', fig.pos='H'}
+
+  knitr::include_graphics("tables_figures/ratio_grid_plots.png")
+```
 
 \blandscape
-
-```{r echo=FALSE, fig.cap = "Figure S3: Individual plots of FACF in relation to mean annual climate, part 1.", out.height='95%', fig.pos='H'}
+```{r echo=FALSE, fig.cap = "Figure S4: Individual plots of FACF in relation to mean annual climate, part 1.", out.height='95%', fig.pos='H'}
 
@@
 
-```{r echo=FALSE, fig.cap = "Figure S4: Individual plots of FACF in relation to mean annual climate, part 2.", out.height='95%', fig.pos='H'}
+```{r echo=FALSE, fig.cap = "Figure S5: Individual plots of FACF in relation to mean annual climate, part 2.", out.height='95%', fig.pos='H'}
 
@@
 
-```{r echo=FALSE, fig.cap = "Figure S5: Individual plots of FACF in relation to mean climate seasonality, part 1.", out.height ='95%', fig.pos='H'}
+```{r echo=FALSE, fig.cap = "Figure S6: Individual plots of FACF in relation to mean climate seasonality, part 1.", out.height ='95%', fig.pos='H'}
 
@@
 \newpage
-```{r echo=FALSE, fig.cap = "Figure S6: Individual plots of FACF in relation to mean climate seasonality, part 2.", out.width ='100%', fig.pos='H'}
+```{r echo=FALSE, fig.cap = "Figure S7: Individual plots of FACF in relation to mean climate seasonality, part 2.", out.width ='100%', fig.pos='H'}
 
@@
 \newpage
-```{r echo=FALSE, fig.cap = "Figure S7: Growing season length-standardized FACF in relation to mean growing season climate, part 1.", out.width ='100%', fig.pos='H'}
+```{r echo=FALSE, fig.cap = "Figure S8: Growing season length-standardized FACF in relation to mean growing season climate, part 1.", out.width ='100%', fig.pos='H'}
 
@@
 \newpage
-```{r echo=FALSE, fig.cap = "Figure S8: Growing season length-standardized FACF in relation to mean growing season climate, part 2.", out.width ='100%', fig.pos='H'}
+```{r echo=FALSE, fig.cap = "Figure S9: Growing season length-standardized FACF in relation to mean growing season climate, part 2.", out.width ='100%', fig.pos='H'}
 
@@
 ```
-
-\newpage
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S9: Ratios among FACF as a function of latitude and climate variables", out.width ='100%', fig.pos='H'}
-
-  knitr::include_graphics("tables_figures/ratio_grid_plots.png")
-```

### 2020-03-10 manuscript/appendix.Rmd
rearranged/ re-numbered SI figures
closes #65

@@
 \elandscape
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S3: Ratios among FACF as a function of latitude and climate variables", out.width ='100%', fig.pos='H'}
+
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S3: Ratios among FACF as a function of latitude and climate variables", out.width ='80%', fig.pos='H'}
 

### 2020-03-28 manuscript/manuscript draft.Rmd
rearranged/ re-numbered SI figures
closes #65
woorked on intro

@@
 
-**Carbon cycling in forests worldwide is central to sustaining an estimated #% of Earth's biodiversity (*REF*), while strongly influencing atmospheric CO~2~ and climate** [@bonan_forests_2008].
+Carbon cycling in forests worldwide provides the energetic basis for sustaining the majority of Earth's terrestrial biodiversity, while strongly influencing atmospheric CO~2~ and climate [@bonan_forests_2008].
 Forests' autotrophic carbon fluxes (FACF)--that is, carbon fixation, allocation, and metabolism by trees and other primary producers--sets the energy ultimately available to heterotrophic organisms (including microbes), in turn influencing their abundance (**REFS**) and possibly diversity [@waide_relationship_1999] (**REFS**).
@@
 
-**KAT, work on this:**
-**FACF vary with latitude**, showing a general trend of decreasing flux with latitude [@beer_terrestrial_2010; @jung_global_2011].
+FACF decrease with latitude, but it remains unclear whether and how the shape of this relationship varies across fluxes.
 Studies agree that FACF are lowest in the boreal regions, and increase into the temperate regions [@luyssaert_co_2007; @huston_global_2009; @beer_terrestrial_2010; @jung_global_2011].
@@
 
-**Climate is a significant driver of FACF across broad spatial scales** [@luyssaert_co_2007; @cleveland_relationships_2011]([Hursh et al.
+The latitudinal gradient in FACF is primarily driven by climate, which is a significant driver of FACF across broad spatial scales [@luyssaert_co_2007; @cleveland_relationships_2011]([Hursh et al.
 2017](https://onlinelibrary.wiley.com/doi/epdf/10.1111/gcb.13489)).
 The majority of studies have focused on exploring the relationships of FACF to mean annual temperature (MAT) and mean annual precipitation (MAP), as the most commonly reported site-level climate variables.
-These variables have the advantage that they describe broad trends in temperature and water availability, and therefore capture a lot of global-scale variation in climate.
+While these fail to capture some important aspects of climate such as seasonality, they do describe broad trends in temperature and water availability, and therefore capture a substantial portion of global-scale variation in climate.
 There is strong evidence that both MAT and MAP show significant positive relationships with FACF [@chu_does_2016].
@@
 Luyssaert et al.
-[-@luyssaert_co_2007] examined GPP and NPP and found that, while GPP increases linearly with MAT, NPP saturates at around 10$^\circ$C MAT.
-In contrast, Larjavaara and Muller-Landau [-@larjavaara_temperature_2012], find that increases in GPP saturate at approximately 25$^\circ$C MAT, while Schuur [-@schuur_productivity_2003] shows that NPP increases linearly with temperature.
-**bbl:maybe remove following sentences** The influence of these climate variables on productivity is further complicated by the possibility of interactive effects occurring between them.
+[-@luyssaert_co_2007] examined GPP and NPP and found that, while $GPP$ increases linearly with MAT, NPP saturates at around 10$^\circ$C MAT.
+In contrast, Larjavaara and Muller-Landau [-@larjavaara_temperature_2012], find that increases in $GPP$ saturate at approximately 25$^\circ$C MAT, while Schuur [-@schuur_productivity_2003] shows that $NPP$ increases linearly with temperature.
 Taylor et al.
@@
 
-**FACF can be influenced by many other factors as well**, which often act across a range of scales, and may show interactive effects with each other [@cleveland_relationships_2011].
-On a local scale, stand age [@litton_carbon_2007; @gillman_latitude_2015], biodiversity [@liang_positive_2016], management [@simova_enigma_2017]; nutrient availability [@aragao_above_2009]; and altitude [@girardin_net_2010; @malhi_variation_2017] all impact FACF.
+FACF can be influenced by many other factors as well, which often act across a range of scales, and may show interactive effects with each other [@cleveland_relationships_2011].
+On a local scale, stand age [@litton_carbon_2007; @gillman_latitude_2015], biodiversity [@liang_positive_2016], management [@simova_enigma_2017], nutrient availability [@aragao_above_2009], and altitude [@girardin_net_2010; @malhi_variation_2017] all impact FACF.
 On a global scale, we expect that FACF are most strongly influenced by broad climatic gradients.
@@
 
-**The recent development of a global forest carbon database synthesizing multiple variables and including records of stand history (ForC; [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018]) opens up the possibility for a standardized analysis of global scale variation** in multiple FACF and the principle climatic drivers of these patterns.
+The recent development of a global forest carbon database synthesizing multiple variables and including records of stand history (ForC; [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018]) opens up the possibility for a standardized analysis of global scale variation in multiple FACF and the principle climatic drivers of these patterns.
 In order to approach this broad topic, we simplify the major gaps in our knowledge to five broad questions and corresponding hypotheses (Table 1).
@@
 
+**All analyses were conducted in R (**Version**). Code and results are archived on GitHub....**
+
 ###Results
@@
 We found little evidence that allocation between fluxes varied substantially with latitude or climate (Fig.
-[**S3**](https://github.com/forc-db/Global_Productivity/issues/65)).
+S3).
 Of the 7 FACF ratios regressed against latitude and three climate variables (MAT, MAP, temperature seasonality), there were only four signficant relationships, all with $R^2 \le 0.2$ (Fig.

### 2020-03-29 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@
 Our analysis of a large global database (ForC) reveals how autotrophic carbon fluxes in mature forests vary with latitude and climate on a global scale.
-We show that, across all nine FACF analyzed, C cycling decreases linearly with latitude (*H1.1*; Fig.
-2)--a finding that confirms multiple previous studies (**REFS**) but contradicts the idea that productivity of temperate forests rivals that of tropical forests (*H1.1.alt*; **REFS**).
+We show that, across all nine FACF analyzed, C cycling decreases linearly (or exponentially, in the case of $GPP$) with latitude (*H1.1*; Fig.
+2)--a finding that confirms multiple previous studies but contradicts the idea that productivity of temperate forests rivals that of tropical forests [@huston_global_2009].
 The FACF generally increase in proportion to one another (*H1.2*), with few differences in allocation detectable at this global scale (Fig.

### 2020-03-30 manuscript/manuscript draft.Rmd
worked on discussion

@@
 Our analysis of a large global database (ForC) reveals how autotrophic carbon fluxes in mature forests vary with latitude and climate on a global scale.
-We show that, across all nine FACF analyzed, C cycling decreases linearly (or exponentially, in the case of $GPP$) with latitude (*H1.1*; Fig.
+We show that, across all nine FACF analyzed, C cycling decreases continually with latitude (*H1.1*; Fig.
 2)--a finding that confirms multiple previous studies but contradicts the idea that productivity of temperate forests rivals that of tropical forests [@huston_global_2009].
@@
 2), indicating no major, systematic omissions or overestimations of flux components.
-However, climate explained lower proportions of variability among subsidiary C fluxes (e.g., *$ANPP_{woody}$*, *$BNPP_{fine.root}$*, $R_{auto-root}$; Fig.
+However, climate explained lower proportions of variability among subsidiary C fluxes (*e.g.*, $ANPP_{woody}$, $BNPP_{fine.root}$, $R_{auto-root}$; Fig.
 2; Table S2).
@@
 2).
-This contrasts with suggestions that productivity of temperate forests is similar to that of tropical forests (**REFS**).
-Temperate forests tend to be younger than tropical forests (**REF**), so analyses comparing across latitudinal gradients without controlling for stand age risk confounding age with biome effects ([Poulter et al.
+This contrasts with the suggestion that productivity of temperate forests is similar to that of tropical forests [@huston_global_2009].
+Compared to tropical forests, the temperate forest biome has experienced more widespread anthropogenic disturbance and has a larger fraction of secondary stands ([Poulter et al.
 2018-DOI:10.1594/PANGAEA.889943](http://pure.iiasa.ac.at/id/eprint/15361/); Potapov et al.
@@
 Ecology and Society 13 (2), 51.
-).
+), so analyses comparing across latitudinal gradients without controlling for stand age risk confounding age with biome effects.
 In addition, because C allocation varies with stand age [@delucia_forest_2007] (**See Nobby's comment in manuscript-draft_NK.pdf**), age differences may introduce systematic biases into analyses of FACF across latitude or global climatic gradients.
@@
 
+**This paragraph needs work--including re-assessment of results, interpretation, writing**
 We show that FACF are broadly consistent in their responses to climate drivers on the global scale (with the exception of some differences in MAT-MAP interactions; Fig.
-3), with no major trends in C allocation among the variable pairs tested (Table 1, *H1*; Fig.
-2; **Some SI table**).
-Although variation in allocation has been observed along gradients of elevation [@moser_elevation_2011] and water availability [@newman_above-_2006]--along with non-climatic axes of stand age [@litton_carbon_2007], nutrient availability [@litton_carbon_2007; @gill_belowground_2016], and forest structure [@taylor_greater_2019]--variation in relation to climate is not apparent at the global scale within ForC, which contains the bulk of relevant data.
+3), with no pronounced trends in C allocation among the variable pairs tested (Figs.
+2, S3).
+Although variation in allocation has been observed along gradients of elevation [@moser_elevation_2011] and water availability [@newman_above-_2006]--along with non-climatic axes of stand age [@litton_carbon_2007], nutrient availability [@litton_carbon_2007; @gill_belowground_2016], and forest structure [@taylor_greater_2019]--little variation in relation to climate is apparent at the global scale within ForC, which contains the bulk of relevant data.
 Our conclusion, then, is that hypothesized gradients in allocation along global climate gradients cannot currently be supported for mature forests, although data quantity and standardization is currently insufficent to rule out the possibility that such trends exist.
-
-bbl: remove this sentence?
-Of particular interest and significance are the relationships amongst $GPP$, net primary productivity ($NPP$ and its components, particularly $ANPP_{woody-stem}$), and respiration ($R_{auto}$ and components).
+**bbl: remove following sentence?** Of particular interest and significance are the relationships amongst $GPP$, net primary productivity ($NPP$ and its components, particularly $ANPP_{woody-stem}$), and respiration ($R_{auto}$ and components).
 There have been suggestions that tropical forests tend to have low carbon use efficiency ($CUE$= $NPP$/$GPP$=($GPP$-$R_{auto}$/$GPP$), which are based on observations of low $CUE$ in old-growth tropical forests relative to (mostly younger) extratropical forests [@delucia_forest_2007; @malhi_productivity_2012; @anderson-teixeira_carbon_2016], but our analysis suggests that these low values might more appropriately be attributed to the fact that these forests are old than to their tropical climate.
 Indeed, $CUE$ is known to decline with forest age [@delucia_forest_2007; @collalti_is_2019], but appears to be roughly independent of $GPP$ [@litton_carbon_2007].
-Among our sites with relevant data, there is indication that CUE or $ANPP_{woody-stem}$/$GPP$ increase with latitude (**some SI table**).
+Among our sites with relevant data, there is indication that CUE or $ANPP_{woody-stem}$/$GPP$ increase with latitude (Fig.
+S3).
 Additional measurements with careful methodological standardization across a consistent set of mature forest sites will be necessary to determine whether any climate-driven gradients in allocation exist at the global scale.
@@
 One interesting observation was that climate tends to explain more variation in the major fluxes ($GPP$, $NPP$, $R_{auto}$ - latitude $R^2\ge$ 48%) than in subsidiary fluxes (latitude $R^2$ <27% for $BNPP_{fine.root}$, $R_{auto-root}$,$ANPP_{woody-stem}$) (Fig.
-2; **some SI table?**).
+2; Table S2).
 There are two, non-exclusive, potential explanations for this.
@@
 Belowground fluxes in particular are difficult to quantify, and measurement methods for the belowground fluxes considered here may be measured through fundamentally different approaches (*e.g.*, minirhizotrons, ingrowth cores, or sequential coring for $BNPP_{root-fine}$; root exclusion, stable isotope tracking, or gas exchange of excised roots for $R_{auto-root}$), and sampling depth is variable and often insufficient to capture the full soil profile.
-$ANPP_{woody-stem}$, which is also poorly explained by latitude or climate, is more straightforward to measure but is subject to variability introduced by differences such as biomass allometries applied and minimum plant size sampled (bbl: cite e.g.
-Huntzinger?).
+$ANPP_{woody-stem}$, which is also poorly explained by latitude or climate, is more straightforward to measure but is subject to variability introduced by differences such as biomass allometries applied and minimum plant size sampled (**bbl: cite e.g.
+Huntzinger?**).
 However, methodological variation and uncertainty affect all of fluxes considered here--not necessarily any less than the aforementioned, and some of the larger fluxes that vary more strongly with respect to climate ($ANPP$, $NPP$) are estimated by summing uncertain component fluxes.
@@
 3-4).
-Because many climate variables co-vary across the latitudinal gradient, because climatic drivers affect forest carbon flux on much shorter time scales than can be captured by annual climate summary variables, and because both climatic conditions and C flux vary intra- and inter-annually around the long-term means, it is not appropriate to attempt to identify any one mean annual climate variable as a mechanistic driver of FACF.
+Because many climate variables co-vary across the latitudinal gradient (Fig.
+S2), because climatic drivers affect forest carbon flux on much shorter time scales than can be captured by annual climate summary variables, and because both climatic conditions and C flux vary intra- and inter-annually around the long-term means, it is not appropriate to attempt to identify any one mean annual climate variable as a mechanistic driver of FACF.
 However, it remains informative to consider these relationships.
-We find that temperature-related climate variables ($MAT$, temperature seasonality, ...**LIST**) explain the highest proportion of variability in FACF, and among these, $MAT$ is generally the best predictor--perhaps because site-specific MAT is recorded for the majority of sites in ForC, whereas other variables are extracted from global gridded data products (Table S1**?**).
+We find that temperature-related climate variables ($MAT$, temperature seasonality, ...**LIST**) explain the highest proportion of variability in FACF, and among these, $MAT$ is generally the best predictor--perhaps because site-specific MAT is recorded for the majority of sites in ForC, whereas other variables are extracted from global gridded data products (Table S1).
 The effects of temperature are modified by moisture availability, with reduced FACF under hot and dry conditions (*i.e.*, high PET, high deficit; Fig.
@@
 Thus, although temperature and water availability jointly and interactively drive global-scale patterns of FACF.
-
-bbl: this seems really interesting and novel, consider expanding the following paragraph
 
@@
 Our analysis clarifies how FACF vary with latitude and climate on a global scale, with some important implications for how forest carbon cycling relates to climate and, by extension, how it is likely to respond to climatic warming.
-We find no support for non-linear trends in mature forest C cycling with respect to latitude or $MAT$, and no distinct trends in C allocation across the global scale (Fig.
-2).
-The implication is that under warmer conditions with similar moisture availability--and within the temperature range to which forest communities are adapated and acclimatized--higher temperatures result in a generalized acceleration of FACF, with no major shifts in C allocation among subsidiary fluxes.
-This is consistent with observations of continental- to global-scale increases in $GPP$ (*[Li & Xiao 2019](https://www.mdpi.com/2072-4292/11/21/2563)*), $ANPP_{woody stem}$ (*[Brienen et al., 2015](https://www-nature-com.smithsonian.idm.oclc.org/articles/nature14283)*), as well as some C cycle components not considered here--tree mortality (*[Brienen et al., 2015](https://www-nature-com.smithsonian.idm.oclc.org/articles/nature14283)*; **some McDowell Ref-KAT**), soil respiration ([Bond Lamberty & Thompson 2010](https://www-nature-com.smithsonian.idm.oclc.org/articles/nature08930/)), heterotrophic soil respiration ([Bond-Lamberty et al.
+Our findings show that higher temperatures with similar moisture availability result in a generalized acceleration of FACF (Figs.
+2-3).
+This is consistent with observations of continental- to global-scale increases in $GPP$ (*[Li & Xiao 2019](https://www.mdpi.com/2072-4292/11/21/2563)*) and $ANPP_{woody stem}$ (*[Brienen et al., 2015](https://www-nature-com.smithsonian.idm.oclc.org/articles/nature14283)*; *[Hubau et al.
+2020](https://doi.org/10.1038/s41586-020-2035-0)*), along with some C cycle components not considered here--tree mortality (*[Brienen et al., 2015](https://www-nature-com.smithsonian.idm.oclc.org/articles/nature14283)*; *[McDowell et al.
+2018](https://nph.onlinelibrary.wiley.com/doi/full/10.1111/nph.15027 )*), soil respiration ([Bond Lamberty & Thompson 2010](https://www-nature-com.smithsonian.idm.oclc.org/articles/nature08930/)), and heterotrophic soil respiration ([Bond-Lamberty et al.
 2018](https://www-nature-com.smithsonian.idm.oclc.org/articles/s41586-018-0358-x)).
-Of course, other factors such as rising CO~2~ are also at play, and such increases are not universal (e.g., [Rutishauser et al., 2019](https://onlinelibrary-wiley-com.smithsonian.idm.oclc.org/doi/full/10.1111/gcb.14833?casa_token=imtB9LD9mL8AAAAA%3AYWgsqNUfRkgs6_dIuDzmwUzL0qVyyxojhbj3jjfRcykBy8ay9T4qMOKPeOPytSGET9muIs8myhQOh3E))...
-Of course, actual climatic changes will result in very different sets of conditions than represented across geographic gradients in climate, but our analysis clarifies how carbon cycles through contemporary forest ecosystems.
-As we enter a period of accelerating climatic change, understanding of the fundamental climatic controls on FACF sets a foundation for understanding patterns of change.
-**[work on this]**
+However, increasing C flux rates are not universal (e.g., *[Rutishauser et al., 2019](https://onlinelibrary-wiley-com.smithsonian.idm.oclc.org/doi/full/10.1111/gcb.14833?casa_token=imtB9LD9mL8AAAAA%3AYWgsqNUfRkgs6_dIuDzmwUzL0qVyyxojhbj3jjfRcykBy8ay9T4qMOKPeOPytSGET9muIs8myhQOh3E)*, **MORE REFS**).
+This is likely because factors other than rising temperatures are at play, including changes to other aspects of cliamte, atmospheric pollution (CO~2~, SO~2~, NO~x~), and local disturbances.
+Morevoer, forest ecosystem responses to climatic changes outside the temperature range to which forest communities are adapated and acclimatized will not necessarily parallel responses across across geographic gradients in climate.
+Nevertheless, as we enter a period of accelerating climatic change, understanding of the fundamental climatic controls on FACF sets a foundation for understanding patterns of change.
 

### 2020-04-06 scripts/final_model_no_weighting.R
Update final_model_no_weighting.R

@@ -3,7 +3,7 @@
 rm(list = ls())
 
 # Set working directory as ForC main folder ####
-setwd("C:/Users/gyrcbm/Dropbox/ForC")
+setwd("C:/Users/becky/Dropbox (Smithsonian)/GitHub/ForC")
 
 # Load libaries ####
 library(lme4)
@@ -136,12 +136,12 @@ for(response.variables in response.variables.groups){
     
     for(fixed.v in fixed.variables){
       
-      png(file = paste0("C:/Users/gyrcbm/Dropbox/Global_Productivity/results/figures/final_figures/unweighted_model/Effect_of_", fixed.v, "_MATURE_only_", age, "_", n, ".png"), width = 2255, height = 2000, units = "px", res = 300)
+      png(file = paste0("C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/figures/final_figures/unweighted_model/Effect_of_", fixed.v, "_MATURE_only_", age, "_", n, ".png"), width = 2255, height = 2000, units = "px", res = 300)
       
       par(mfrow = c(1,1), mar = c(0,0,0,0), oma = c(5,5,2,0))
       print(fixed.v)
       
-      fixed.v.info <- read.csv("C:/Users/gyrcbm/Dropbox/Global_Productivity/raw.data/fixedv_data.csv", stringsAsFactors = F)
+      fixed.v.info <- read.csv("C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/raw.data/fixedv_data.csv", stringsAsFactors = F)
       
       xaxis <- fixed.v.info$xaxis[which(fixed.v.info$fixed.v %in% fixed.v)]
       
@@ -152,7 +152,7 @@ for(response.variables in response.variables.groups){
       
       for (response.v in response.variables){
         
-        col.sym <- read.csv("C:/Users/gyrcbm/Dropbox/Global_Productivity/raw.data/colsym.csv", stringsAsFactors = F)
+        col.sym <- read.csv("C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/raw.data/colsym.csv", stringsAsFactors = F)
         
         col <- col.sym$col[which(col.sym$variable %in% response.v)]
         sym <- col.sym$sym[which(col.sym$variable %in% response.v)]
@@ -284,7 +284,7 @@ for(response.variables in response.variables.groups){
 }
 
 
[…]

### 2020-04-06 manuscript/appendix.Rmd
updated manuscript and appendix pdf

@@
 ```{r setup, include = FALSE}
-# knitr::opts_knit$set(root.dir = 'C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/')
+knitr::opts_knit$set(root.dir = 'C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/')
 # knitr::opts_knit$set(root.dir = 'C:/Users/gyrcbm/Dropbox/Global_Productivity/manuscript/')
- knitr::opts_knit$set(root.dir = '/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/Global_Productivity/manuscript/')
+ # knitr::opts_knit$set(root.dir = '/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/Global_Productivity/manuscript/')
 ```
@@
   column_spec(4, width = "7cm") %>%
-  column_spec(5, width = "5cm")
+  column_spec(5, width = "5cm") %>%
+  kableExtra::footnote(number = c("Hijmans et al. (2005)", "Fick et al. (2017)", "Harris et al. (2017)", "Trabucco and Zomer (2019)", "Abatzoglou et al. (2018)"), footnote_as_chunk = TRUE)
 ```

### 2020-04-06 manuscript/manuscript draft.Rmd
updated manuscript and appendix pdf

@@
 ```{r setup, include = FALSE}
-# knitr::opts_knit$set(root.dir = 'C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/')
-knitr::opts_knit$set(root.dir = '/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/Global_Productivity/manuscript/')
+knitr::opts_knit$set(root.dir = 'C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/')
+# knitr::opts_knit$set(root.dir = '/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/Global_Productivity/manuscript/')
 ```

### 2020-04-06 scripts/final_model_no_weighting.R
Update final_model_no_weighting.R

@@ -183,7 +183,7 @@ for(response.variables in response.variables.groups){
         ylim[1] <- ylim[1] - 0.25
         ylim[2] <- ylim[2] + 0.25
         
-        if(!fixed.v %in% c("mat", "lat", "PreSeasonality", "SolarRadiation", "TempRangeAnnual", "CloudCover")){
+        # if(!fixed.v %in% c("mat", "lat", "PreSeasonality", "SolarRadiation", "TempRangeAnnual", "CloudCover")){
 
         mod <-  lmer(scale(mean) ~ 1 + (1|geographic.area/plot.name), data = df, REML = F)
         mod.clim <- lmer(scale(mean) ~ poly(fixed, 1, raw = T) + (1|geographic.area/plot.name), data = df, REML = F)
@@ -192,13 +192,21 @@ for(response.variables in response.variables.groups){
         aictab <- aictab(list(mod.clim = mod.clim, mod.clim.poly = mod.clim.poly), sort = T)
 
         best.model <- as.character(aictab(list(mod = mod, mod.clim = mod.clim, mod.clim.poly = mod.clim.poly), sort = T)$Modname[1])
-        delta.aic <- as.numeric(aictab(list(mod = mod, mod.clim = mod.clim, mod.clim.poly = mod.clim.poly), sort = T)$Delta_AICc[2])
-        delta.aic <- signif(delta.aic, digits=4)
-
-
+        
         if (best.model == "mod") mod.full <- lmer(scale(mean) ~ poly(fixed, 1, raw = T) + (1|geographic.area/plot.name), data = df, REML = F)
+        
+        if (best.model == "mod.clim.poly"){ 
+          
+          test.delta.aic <- as.numeric(aictab(list(mod.clim = mod.clim, mod.clim.poly = mod.clim.poly))$Delta_AICc[2])
+          
+          best.model <- ifelse(test.delta.aic > 2, "mod.clim.poly", "mod.clim")
+        }
+        
         if (best.model == "mod.clim") mod.full <- lmer(scale(mean) ~ poly(fixed, 1, raw = T) + (1|geographic.area/plot.name), data = df, REML = F)
         if (best.model == "mod.clim.poly") mod.full <- lmer(scale(mean) ~ poly(fixed, 2, raw = T) + (1|geographic.area/plot.name), data = df, REML = F)
+       
+        delta.aic <- as.numeric(aictab(list(mod = mod, mod.full = mod.full))$Delta_AICc[2])
[…]

### 2020-04-06 scripts/final_model_no_weighting.R
Update final_model_no_weighting.R

@@ -193,8 +193,6 @@ for(response.variables in response.variables.groups){
 
         best.model <- as.character(aictab(list(mod = mod, mod.clim = mod.clim, mod.clim.poly = mod.clim.poly), sort = T)$Modname[1])
         
-        if (best.model == "mod") mod.full <- lmer(scale(mean) ~ poly(fixed, 1, raw = T) + (1|geographic.area/plot.name), data = df, REML = F)
-        
         if (best.model == "mod.clim.poly"){ 
           
           test.delta.aic <- as.numeric(aictab(list(mod.clim = mod.clim, mod.clim.poly = mod.clim.poly))$Delta_AICc[2])
@@ -202,7 +200,17 @@ for(response.variables in response.variables.groups){
           best.model <- ifelse(test.delta.aic > 2, "mod.clim.poly", "mod.clim")
         }
         
+        if (best.model == "mod.clim"){ 
+          
+          test.delta.aic <- as.numeric(aictab(list(mod = mod, mod.clim = mod.clim))$Delta_AICc[2])
+          
+          best.model <- ifelse(test.delta.aic > 2, "mod.clim", "mod")
+        }
+        
+        if (best.model == "mod") mod.full <- lmer(scale(mean) ~ poly(fixed, 1, raw = T) + (1|geographic.area/plot.name), data = df, REML = F)
+        
         if (best.model == "mod.clim") mod.full <- lmer(scale(mean) ~ poly(fixed, 1, raw = T) + (1|geographic.area/plot.name), data = df, REML = F)
+        
         if (best.model == "mod.clim.poly") mod.full <- lmer(scale(mean) ~ poly(fixed, 2, raw = T) + (1|geographic.area/plot.name), data = df, REML = F)
        
         delta.aic <- as.numeric(aictab(list(mod = mod, mod.full = mod.full))$Delta_AICc[2])
@@ -372,8 +380,6 @@ for (age in ages){
       
       best.model <- as.character(aictab(list(mod = mod, mod.clim = mod.clim, mod.clim.poly = mod.clim.poly), sort = T)$Modname[1])
       
-      if (best.model == "mod") mod.full <- lmer(scale(mean) ~ poly(fixed, 1, raw = T) + (1|geographic.area/plot.name), data = df, REML = F)
-      
       if (best.model == "mod.clim.poly"){ 
         
[…]

### 2020-04-06 manuscript/manuscript draft.Rmd
updated manuscript draft

@@
 Hypotheses were accepted if the $\Delta$AIC between a model including the fixed effect of interest and a corresponding null model excluding that fixed effect exceeded 2.0.
-All $R^2$ values presented here are marginal $R^2$ values, and refer to the proportion of variation explained by only the fixed effects, **unless otherwise specified**.
-**Becky, are there any instances of this?** Specific analyses are as described below.
+All $R^2$ values presented here are marginal $R^2$ values, and refer to the proportion of variation explained by only the fixed effects.
+Specific analyses are as described below.
 
 We first examined the relationship between latitude and FACF (Q1; Table 1).
-We tested models with latitude as a linear and term (corresponding null: model without latitude) and as a second-order polynomial term (corresponding null: model with latitude as a linear term).
-If the shape of the relationship [**made biological sense**](https://github.com/forc-db/Global_Productivity/issues/64), and was a significant improvement on the linear relationship ($\Delta$AIC >2), we accepted the polynomial as the best model.
+We tested models with latitude as a linear term (corresponding null: model without latitude) and as a second-order polynomial term (corresponding null: model with latitude as a linear term), and calculated AIC values to determine the best model.
+Models were accepted as the best model if $\Delta$AIC > 2 with respect to the corresponding null, and were significant with respect to a null model with no fixed term.
 We also examined relationships among fluxes across latitude, testing whether sums of component fluxes matched the larger fluxes and whether C allocation varied with latitude, as specified below.
@@
 The FACF that were most poorly predicted by latitude were $BNPP_{root-fine}$ ($R^2$=0.17) and $ANPP_{woody.stem}$=0.18).
-For all FACF, the relationship with latitude was best predicted by the linear model (**but see [this issue]((https://github.com/forc-db/Global_Productivity/issues/64))**).
+For all FACF, the relationship with latitude was best predicted by the linear model).
 

### 2020-04-07 scripts/discarded_code.R
Create discarded_code.R

@@ -1,0 +1,421 @@
+### for graph with two plots, latitude only, saved to supporting information mature forests only ####
+
+fixed.variables <- "lat"
+
+response.variables.groups <- list(c("GPP", "NPP", "ANPP", "BNPP_root", "R_auto"),
+                                  c("ANPP_foliage", "ANPP_woody_stem", "BNPP_root_fine", "R_auto_root"))
+
+all.results = NULL
+
+for (age in ages){
+  
+  if (age %in% "age.greater.than.100") ages.to.keep <- ForC_simplified$stand.age >= 100 & !is.na(ForC_simplified$stand.age)
+  if (age %in% "age.greater.than.200") ages.to.keep <- ForC_simplified$stand.age >= 200 & !is.na(ForC_simplified$stand.age)
+  
+  for(fixed.v in fixed.variables){
+    
+    png(file = paste0("C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/figures/final_figures/supporting_information/latitude unscaled.png"), width = 5000, height = 2000, units = "px", res = 300)
+    
+    layout(matrix(1:3, nrow = 1), widths = c(1,1,0.25))
+    par(mar = c(3,3,3,0), oma = c(3,3,0,7))
+    
+    print(fixed.v)
+    
+    fixed.v.info <- read.csv("C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/raw.data/fixedv_data.csv", stringsAsFactors = F)
+    
+    xaxis <- fixed.v.info$xaxis[which(fixed.v.info$fixed.v %in% fixed.v)]
+    
+    for(response.variables in response.variables.groups){
+      
+      if(response.variables[1] == "GPP") n <- 1
+      if(response.variables[1] == "ANPP_foliage") n <- 2
+      
+      first.plot <- TRUE
+      
+      for (response.v in response.variables){
+        
+        col.sym <- read.csv("C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/raw.data/colsym.csv", stringsAsFactors = F)
+        
+        col <- col.sym$col[which(col.sym$variable %in% response.v)]
+        sym <- col.sym$sym[which(col.sym$variable %in% response.v)]
+        
+        if(response.v %in% "NPP") responses.to.keep  <- c("NPP_1", "NPP_2", "NPP_3", "NPP_4", "NPP_5", "NPP_0")
[…]

### 2020-04-07 manuscript/manuscript draft.Rmd
added citations to manuscript draft

@@
 Over 69 Gt of CO~2~ cycle through Earth's forests each year [@badgley_terrestrial_2019], and in recent decades their net C sequestration (~2.4 Gt C yr^-1^) offset roughly 30% of anthropogenic fossil fuel emissions [@pan_large_2011].
-As atmospheric carbon dioxide levels continue to rise, driving climate change, forests will play a critical role in shaping the future of Earth's climate ([Cavaleri et al.
-2015](https://onlinelibrary.wiley.com/doi/epdf/10.1111/gcb.12860); [Rogelj et al.
-2018](https://www.ipcc.ch/site/assets/uploads/sites/2/2019/05/SR15_Chapter2_Low_Res.pdf)).
+As atmospheric carbon dioxide levels continue to rise, driving climate change, forests will play a critical role in shaping the future of Earth's climate [@cavaleri_urgent_2015; @rogelj_mitigation_2018].
 However, our ability to draw general macroscopic conclusions regarding global variation in FACF with respect to climate has been limited in that these analyses often mix forests that vary in stand age, disturbance history, and/or management status; do not always sufficiently parse related variables; and typically consider only one or a few variables at a time.
@@
 Evidence for this is further complicated by the fact that different studies use different measures of productivity to explore these relationships.
-For example, modelling of global terrestrial ecosystem gross primary productivity ($GPP$) through upscaling and calibration of eddy flux measurements indicates that GPP peaks in tropical forests [@beer_terrestrial_2010; @jung_global_2011; @badgley_terrestrial_2019](*[Li & Xiao 2019](https://www.mdpi.com/2072-4292/11/21/2563)*).
+For example, modelling of global terrestrial ecosystem gross primary productivity ($GPP$) through upscaling and calibration of eddy flux measurements indicates that GPP peaks in tropical forests [@beer_terrestrial_2010; @jung_global_2011; @badgley_terrestrial_2019; @li_mapping_2019].
 In contrast, some studies suggest that the highest values of net primary productivity ($NPP$) may be found in temperate forests [@luyssaert_co_2007; @huston_global_2009], while others find $NPP$ highest in the tropics and decreasing with latitude [@simova_enigma_2017].
@@
 
-The latitudinal gradient in FACF is primarily driven by climate, which is a significant driver of FACF across broad spatial scales [@luyssaert_co_2007; @cleveland_relationships_2011]([Hursh et al.
-2017](https://onlinelibrary.wiley.com/doi/epdf/10.1111/gcb.13489)).
+The latitudinal gradient in FACF is primarily driven by climate, which is a significant driver of FACF across broad spatial scales [@luyssaert_co_2007; @cleveland_relationships_2011; @hursh_sensitivity_2017].
 The majority of studies have focused on exploring the relationships of FACF to mean annual temperature (MAT) and mean annual precipitation (MAP), as the most commonly reported site-level climate variables.
@@
 This contrasts with the suggestion that productivity of temperate forests is similar to that of tropical forests [@huston_global_2009].
-Compared to tropical forests, the temperate forest biome has experienced more widespread anthropogenic disturbance and has a larger fraction of secondary stands ([Poulter et al.
-2018-DOI:10.1594/PANGAEA.889943](http://pure.iiasa.ac.at/id/eprint/15361/); Potapov et al.
-2008.
-Mapping the World’s Intact Forest Landscapes by Remote Sensing.
-Ecology and Society 13 (2), 51.
-), so analyses comparing across latitudinal gradients without controlling for stand age risk confounding age with biome effects.
+Compared to tropical forests, the temperate forest biome has experienced more widespread anthropogenic disturbance and has a larger fraction of secondary stands [@potapov_mapping_2008; @poulter_global_2018], so analyses comparing across latitudinal gradients without controlling for stand age risk confounding age with biome effects.
 In addition, because C allocation varies with stand age [@delucia_forest_2007] (**See Nobby's comment in manuscript-draft_NK.pdf**), age differences may introduce systematic biases into analyses of FACF across latitude or global climatic gradients.
-For example, woody productivity tends to be higher in rapidly aggrading secondary stands than in old-growth forests, where proportionally more C is allocated to respiration (** Nobby AFM paper: Understanding the controls over forest carbon use efficiency on small spatial scales: Effects of forest disturbance and tree diversity**) [*purpose for respiration/ other compenents (**See Nobby's comment in manuscript-draft_NK.pdf**)].
+For example, woody productivity tends to be higher in rapidly aggrading secondary stands than in old-growth forests, where proportionally more C is allocated to respiration [@kunert_understanding_2019].
+[*purpose for respiration/ other compenents* (**See Nobby's comment in manuscript-draft_NK.pdf**)].
 
@@
 2-3).
-This is consistent with observations of continental- to global-scale increases in $GPP$ (*[Li & Xiao 2019](https://www.mdpi.com/2072-4292/11/21/2563)*) and $ANPP_{woody stem}$ (*[Brienen et al., 2015](https://www-nature-com.smithsonian.idm.oclc.org/articles/nature14283)*; *[Hubau et al.
-2020](https://doi.org/10.1038/s41586-020-2035-0)*), along with some C cycle components not considered here--tree mortality (*[Brienen et al., 2015](https://www-nature-com.smithsonian.idm.oclc.org/articles/nature14283)*; *[McDowell et al.
-2018](https://nph.onlinelibrary.wiley.com/doi/full/10.1111/nph.15027 )*), soil respiration ([Bond Lamberty & Thompson 2010](https://www-nature-com.smithsonian.idm.oclc.org/articles/nature08930/)), and heterotrophic soil respiration ([Bond-Lamberty et al.
-2018](https://www-nature-com.smithsonian.idm.oclc.org/articles/s41586-018-0358-x)).
-However, increasing C flux rates are not universal (e.g., *[Rutishauser et al., 2019](https://onlinelibrary-wiley-com.smithsonian.idm.oclc.org/doi/full/10.1111/gcb.14833?casa_token=imtB9LD9mL8AAAAA%3AYWgsqNUfRkgs6_dIuDzmwUzL0qVyyxojhbj3jjfRcykBy8ay9T4qMOKPeOPytSGET9muIs8myhQOh3E)*, **MORE REFS**).
+This is consistent with observations of continental- to global-scale increases in $GPP$ [@li_mapping_2019] and $ANPP_{woody stem}$ [@brienen_long-term_2015; @hubau_asynchronous_2020], along with some C cycle components not considered here--tree mortality [@brienen_long-term_2015; @mcdowell_drivers_2018], soil respiration [@bond-lamberty_global_2010], and heterotrophic soil respiration [@bond-lamberty_globally_2018].
+However, increasing C flux rates are not universal [@rutishauser_testing_2020] (**MORE REFS**).
 This is likely because factors other than rising temperatures are at play, including changes to other aspects of cliamte, atmospheric pollution (CO~2~, SO~2~, NO~x~), and local disturbances.
@@
 
-- consistent with [Hofhansl et al.
-2015](https://agupubs.onlinelibrary.wiley.com/doi/full/10.1002/2014GB004934) (**verify**) , we found a slight tendency for warmer sites to have higher aboveground allocation
+- consistent with [@hofhansl_new_2015] (**verify**) , we found a slight tendency for warmer sites to have higher aboveground allocation
 

### 2020-05-03 manuscript/manuscript draft.Rmd
worked on writing, references
I'm no longer able to knit the document.

@@
 
-Carbon cycling in forests worldwide provides the energetic basis for sustaining the majority of Earth's terrestrial biodiversity, while strongly influencing atmospheric CO~2~ and climate [@bonan_forests_2008].
+Carbon cycling in forests worldwide provides the energetic basis for sustaining the majority of Earth's terrestrial biodiversity and many human populations (**REF**), while strongly influencing atmospheric CO~2~ and climate [@bonan_forests_2008].
 Forests' autotrophic carbon fluxes (FACF)--that is, carbon fixation, allocation, and metabolism by trees and other primary producers--sets the energy ultimately available to heterotrophic organisms (including microbes), in turn influencing their abundance (**REFS**) and possibly diversity [@waide_relationship_1999] (**REFS**).
@@
 
-The recent development of a global forest carbon database synthesizing multiple variables and including records of stand history (ForC; [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018]) opens up the possibility for a standardized analysis of global scale variation in multiple FACF and the principle climatic drivers of these patterns.
+The recent development of the Global Forest Carbon database (ForC), which synthesizes multiple variables and including records of stand history [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018], opens up the possibility for a standardized analysis of global scale variation in multiple FACF and the principle climatic drivers of these patterns.
 In order to approach this broad topic, we simplify the major gaps in our knowledge to five broad questions and corresponding hypotheses (Table 1).
-** see [issue #48 ](https://github.com/forc-db/Global_Productivity/issues/48)** First, we ask how forest autotrophic carbon fluxes (FACF) vary with latitude.
+First, we ask how FACF vary with latitude.
 We then test how these fluxes relate to MAT and MAP, and additionally how they respond to other, less well-studied, climate variables.
 Finally, we consider the relationship between FACF and seasonality, considering the role of seasonality in explaining variation in carbon fluxes, and the influence of climate on FACF standardized by growing season length.
-We use a comprehensive global database of forest carbon fluxes to address the above questions for nine carbon fluxes, allowing for an in-depth exploration of the effect of climate on FACF globally.
+We address the above questions for nine carbon fluxes contained in ForC, allowing for an in-depth exploration of the effect of climate on FACF globally.
 
@@
 
-Variation in allocation to component carbon fluxes along latitudinal gradients was explored for the following pairings: $GPP:NPP$, $ANPP:BNPP$, and $ANPP_{foliage}:ANPP_{woody.stem}$; $ANPP_{foliage}:NPP$, $ANPP_{woody.stem}:NPP$, $ANPP:NPP$, and $BNPP:NPP$.
+Variation in allocation to component carbon fluxes along latitudinal gradients was explored for the following pairings: $GPP:NPP$, $ANPP:BNPP$, $ANPP_{foliage}:ANPP_{woody.stem}$, $ANPP_{foliage}:NPP$, $ANPP_{woody.stem}:NPP$, $ANPP:NPP$, and $BNPP:NPP$.
 For each set of paired fluxes, measurements taken at the same site and plot, and in the same year, were paired together, and the ratio of each pair of measurements calculated.
@@
 In total, we analyzed 1228 records from nine FACF variables taken from forests that had experienced no major anthropogenic disturbances within the past 100 years.
-These records represented a total of 154 distinct geographic areas (Fig.
-1, Table 2), across all forested biogeographic and climate zones.
+These records represented a total of 154 distinct geographic areas across all forested biogeographic and climate zones (Fig.
+1, Table 2).
 
@@
 Specifically, latitude explained 64% of variation in GPP (n = 254, p<0.0001), 50% in NPP (n = 114, p<0.0001) and 45% in ANPP (n = 259, p<0.0001).
-The FACF that were most poorly predicted by latitude were $BNPP_{root-fine}$ ($R^2$=0.17) and $ANPP_{woody.stem}$=0.18).
-For all FACF, the relationship with latitude was best predicted by the linear model).
+The FACF that were most poorly predicted by latitude were $BNPP_{root-fine}$ ($R^2$=0.17) and $ANPP_{woody.stem}$ ($R^2$=0.18).
+For all FACF, the relationship with latitude was best predicted by the linear model.
 
@@
 3-4, S4-S5, Table S2).
-As with latitude, MAT tended to explain more variation in the larger FACF ($GPP$, $NPP$, $ANPP$, $R_{auto}$) and $ANPP_foliage$ (all $R^2$> 0.4) than in subsidiary and belowground fluxes ($ANPP_{woody-stem}$, $R_{root}$, $BNPP_{root-fine}$; all $R^2$< 0.25).
+As with latitude, MAT tended to explain more variation in the larger FACF ($GPP$, $NPP$, $ANPP$, $R_{auto}$) and $ANPP_{foliage}$ (all $R^2$> 0.4) than in subsidiary and belowground fluxes ($ANPP_{woody-stem}$, $R_{root}$, $BNPP_{root-fine}$; all $R^2$< 0.25).
 
@@
 S4-S5, Table S2).
-Solar radiation explained a low proportion of variability in all FACF, explaining less than 20% of the variation in each flux, with the exception of$R_{auto}$ ($R^2$ = 0.26, p<0.05).
+Solar radiation explained a low proportion of variability in all FACF, explaining less than 20% of the variation in each flux, with the exception of $R_{auto}$ ($R^2$ = 0.26, p<0.05).
 
@@
 4e, S6-S7; Table S2).
-Temperature seasonality was strongly correlated with annual temperature range, which was likewise a similarly strong predictor of FACF (Table S2).FACF were highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower.
+Temperature seasonality was strongly correlated with annual temperature range, which was likewise a similarly strong predictor of FACF (Table S2).
+FACF were highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower.
 
@@
 Speficifally, $ANPP$ increased with growing season temperature ($R^2$ = 0.10, p<0.001) and precipitation ($R^2$ = 0.04, p<0.05).
-Similaryly, $ANPP_{foliage}$ increased slightly with growing season temperature ($R^2$ = 0.16, p<0.01) and precipitation ($R^2$ = 0.09, p<0.05).
+Similarly, $ANPP_{foliage}$ increased slightly with growing season temperature ($R^2$ = 0.16, p<0.01) and precipitation ($R^2$ = 0.09, p<0.05).
 Growing season solar radiation had a positive influence on $GPP$ ($R^2$ = 0.21, p<0.001), $NPP$ ($R^2$ = 0.21, p<0.001), $BNPP$ ($R^2$ = 0.16, p<0.001) and $BNPP_{fine.root}$ ($R^2$ = 0.12, p<0.01).
@@
 2)--a finding that confirms multiple previous studies but contradicts the idea that productivity of temperate forests rivals that of tropical forests [@huston_global_2009].
-The FACF generally increase in proportion to one another (*H1.2*), with few differences in allocation detectable at this global scale (Fig.
+FACF generally increase in proportion to one another (*H1.2*), with few differences in allocation detectable at this global scale (Fig.
 S2) and with component fluxes summing appropriately to larger fluxes (Fig.
@@
 3-4).
-Water availability is also influential, but generally of secondary importance across the range represented in our database (Figs.
+Water availability is also influential, but generally of secondary importance across the climate space occupied by forests (Figs.
 3-4).
@@
 S9; Table S4).
-These findings clarify the big picture of how FACF vary with latitude and climate on a global scale.
+These findings resolve a number of open debates and clarify the big picture of how FACF vary with latitude and climate on a global scale.
 
-Past studies have differed in their conclusions regarding the relationship between FACF and latitude or its correlates (Table 1, *H1*; **REFS**)--quite possibly because of lack of standardization with respect to stand age and disturbance history.
+Past studies have differed in their conclusions regarding the relationship between FACF and latitude or its correlates (Table 1, *H1*)--quite possibly because of lack of standardization with respect to stand age and disturbance history.
 Our findings indicate that, among mature, undisturbed stands, FACF are unambiguously highest in the tropical regions, and the relationship is approximately linear (Fig.
@@
 Compared to tropical forests, the temperate forest biome has experienced more widespread anthropogenic disturbance and has a larger fraction of secondary stands [@potapov_mapping_2008; @poulter_global_2018], so analyses comparing across latitudinal gradients without controlling for stand age risk confounding age with biome effects.
-In addition, because C allocation varies with stand age [@delucia_forest_2007] (**See Nobby's comment in manuscript-draft_NK.pdf**), age differences may introduce systematic biases into analyses of FACF across latitude or global climatic gradients.
-For example, woody productivity tends to be higher in rapidly aggrading secondary stands than in old-growth forests, where proportionally more C is allocated to respiration [@kunert_understanding_2019].
-[*purpose for respiration/ other compenents* (**See Nobby's comment in manuscript-draft_NK.pdf**)].
+In addition, because carbon allocation varies with stand age [@delucia_forest_2007] (**Doughty et al.
+2018: DOI: 10.1111/btp.12504; Anderson-Teixeira et al.
+2013: doi: 10.1111/gcb.12194**), age differences may introduce systematic biases into analyses of FACF across latitude or global climatic gradients.
+For example, woody productivity tends to be higher in rapidly aggrading secondary stands than in old-growth forests, where proportionally more C is allocated to respiration (**Doughty et al.
+2018: DOI: 10.1111/btp.12504**);[@delucia_forest_2007; @kunert_understanding_2019].
+By constraining our analysis to stands ≥ 100 years old with no record of major recent disturbance, we have clarified the shape of the relationship between FACF and latitude.
 
 **This paragraph needs work--including re-assessment of results, interpretation, writing**
-We show that FACF are broadly consistent in their responses to climate drivers on the global scale (with the exception of some differences in MAT-MAP interactions; Fig.
-3), with no pronounced trends in C allocation among the variable pairs tested (Figs.
+We show that, with the exception of some differences in MAT-MAP interactions (Fig.
+3), FACF are broadly consistent in their responses to climate drivers on the global scale, with no pronounced trends in C allocation among the variable pairs tested (Figs.
 2, S3).
-Although variation in allocation has been observed along gradients of elevation [@moser_elevation_2011] and water availability [@newman_above-_2006]--along with non-climatic axes of stand age [@litton_carbon_2007], nutrient availability [@litton_carbon_2007; @gill_belowground_2016], and forest structure [@taylor_greater_2019]--little variation in relation to climate is apparent at the global scale within ForC, which contains the bulk of relevant data.
+Although variation in allocation has been observed along gradients of elevation [@moser_elevation_2011] and water availability [@newman_above-_2006]--along with non-climatic axes of stand age [@litton_carbon_2007], nutrient availability [@litton_carbon_2007; @gill_belowground_2016], and forest structure [@taylor_greater_2019]--little variation in relation to climate is apparent at the global scale.
 Our conclusion, then, is that hypothesized gradients in allocation along global climate gradients cannot currently be supported for mature forests, although data quantity and standardization is currently insufficent to rule out the possibility that such trends exist.
-**bbl: remove following sentence?** Of particular interest and significance are the relationships amongst $GPP$, net primary productivity ($NPP$ and its components, particularly $ANPP_{woody-stem}$), and respiration ($R_{auto}$ and components).
 There have been suggestions that tropical forests tend to have low carbon use efficiency ($CUE$= $NPP$/$GPP$=($GPP$-$R_{auto}$/$GPP$), which are based on observations of low $CUE$ in old-growth tropical forests relative to (mostly younger) extratropical forests [@delucia_forest_2007; @malhi_productivity_2012; @anderson-teixeira_carbon_2016], but our analysis suggests that these low values might more appropriately be attributed to the fact that these forests are old than to their tropical climate.
 Indeed, $CUE$ is known to decline with forest age [@delucia_forest_2007; @collalti_is_2019], but appears to be roughly independent of $GPP$ [@litton_carbon_2007].
-Among our sites with relevant data, there is indication that CUE or $ANPP_{woody-stem}$/$GPP$ increase with latitude (Fig.
+Among our sites with relevant data, there are no signficant trends in allocation of $GPP$ to $NPP$ or of $NPP$ to $ANPP_{woody-stem}$ across gradients of latitude or climate (Fig.
 S3).
 Additional measurements with careful methodological standardization across a consistent set of mature forest sites will be necessary to determine whether any climate-driven gradients in allocation exist at the global scale.
+
+**there seems to be a difference between GPP and other fluxes at higher precips (Figs.
+3, 4b), but the only allocaiton ratio we consider is GPP:NPP.
+It would be interesting too consider some others.
+Perhaps precip-->soil fertility-->CUE (see Doughty et al.
+2018).
+I need to come back to this.**
 

### 2020-05-03 manuscript/manuscript-draft.tex
worked on writing, references
I'm no longer able to knit the document.

@@
 \hypersetup{unicode=true,
-            pdftitle={Global patterns of forest autotrophic carbon fluxes},
-            pdfauthor={Rebecca Banbury Morgan, Valentine Herrmann, Norbert Kunert, Ben Bond-Lamberty, Helene C. Muller-Landau, Kristina J. Anderson-Teixeira},
             pdfborder={0 0 0},
@@
 \bibliographystyle{apalike}
-\usepackage{longtable,booktabs}
 \usepackage{graphicx,grffile}
@@
 
-  \title{Global patterns of forest autotrophic carbon fluxes}
-    \pretitle{\vspace{\droptitle}\centering\huge}
-  \posttitle{\par}
-    \author{Rebecca Banbury Morgan, Valentine Herrmann, Norbert Kunert, Ben
-Bond-Lamberty, Helene C. Muller-Landau, Kristina J. Anderson-Teixeira}
-    \preauthor{\centering\large\emph}
-  \postauthor{\par}
+  \title{}
+    \pretitle{\vspace{\droptitle}}
+  \posttitle{}
+    \author{}
+    \preauthor{}\postauthor{}
     \date{}
@@
 
-\usepackage{float} \usepackage{caption} \captionsetup[table]{font=footnotesize} \captionsetup[figure]{font=footnotesize}
+\usepackage{float} \usepackage{caption} \captionsetup[table]{font=footnotesize} \captionsetup[figure]{font=footnotesize} \usepackage{setspace}\onehalfspacing \usepackage{lineno}\linenumbers
 
 \begin{document}
-\maketitle
+
+\textbf{Title:} Global patterns of forest autotrophic carbon fluxes
+
+\textbf{Running head:}
+
+\textbf{Authors:}
+
+Rebecca Banbury Morgan\textsuperscript{1,2}
+
+Valentine Herrmann\textsuperscript{1}
+
+Norbert Kunert\textsuperscript{1,3}
+
+Ben Bond-Lamberty\textsuperscript{4}
+
+Helene C. Muller-Landau\textsuperscript{3}
+
+Kristina J. Anderson-Teixeira\textsuperscript{1,3}*
+
+\textbf{Author Affiliations:}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  Conservation Ecology Center; Smithsonian Conservation Biology
+  Institute; National Zoological Park, Front Royal, VA, USA
+\item
+  \emph{Becky- current}
+\item
+  Center for Tropical Forest Science-Forest Global Earth Observatory;
+  Smithsonian Tropical Research Institute; Panama, Republic of Panama
+\item
+  Joint Global Change Research Institute, Pacific Northwest National
+  Laboratory, College Park Maryland 20740 USA
+\end{enumerate}
+
+*Corresponding Author:
+
+phone: 1-540-635-6546
+
+fax:1-540-635-6506
+
+email: \href{mailto:teixeirak@si.edu}{\nolinkurl{teixeirak@si.edu}}
+
+\textbf{Keywords:}
+
+\textbf{Paper type:} Primary Research Article
+
+\newpage
 
@@
 
-{[}\textbf{very rough start:}{]} \ldots{}.. Here, we drew upon \#
-records from the Global Forest Carbon Database (ForC), representing all
-major forest types and the nine most significant forest autotrophic
-carbon flux (FACF) variables, to comprehensively explore how C cycling
-in mature, undisturbed forests varies with latitude and climate on a
-global scale. We show that, across all FACF variables analyzed, C
-cycling decreases linearly with latitude --a finding that confirms
-multiple previous studies and contradicts the idea that productivity of
-temperate forests rivals that of tropical forests. The FACF variables
-increase in proportion to one another, with no differences in allocation
-detectable at this global scale, but differed in that latitude explained
-a lower proportion of variation among subsidiary fluxes (in particular,
-woody aboveground productivity and belowground net primary productivity,
-\(BNPP\)). Climate explained a significant proportion (\#-\#\%) of
-variation in all C fluxes analyzed (less for subsidiary fluxes), with
-temperature variables in general and mean annual temperature (\(MAT\))
-in particular being the best predictors of FACF on this global scale.
-While other climate variables (\emph{e.g.}, \textbf{XX}) displayed
-significant correlation with FACF, none of them had significantly better
-explanatory power than \(MAT\). The effects of temperature were modified
-by moisture availability, with reduced FACF under hot and dry conditions
-and sometimes under very high precipitaiton (especially for \(BNPP\)).
-FACF declined with temperature seasonality, but growing season length
-doesn't improve upon MAT as a predictor. Within the growing season, the
+{[}\textbf{very rough start:}{]} Carbon fixation, allocation, and
+metabolism by trees set the basis for energy and material flows in
+forest ecosystems and define their interactions with Earth's changing
+climate. Here, we drew upon \# records from the Global Forest Carbon
+Database (ForC), representing all major forest types and the nine most
+significant forest autotrophic carbon flux (FACF) variables, to
+comprehensively explore how C cycling in mature, undisturbed forests
+varies with latitude and climate on a global scale. We show that, across
+all FACF variables analyzed, C cycling decreases linearly with latitude
+-- a finding that confirms multiple previous studies but contradicts the
+idea that net primary productivity (\(NPP\)) of temperate forests rivals
+that of tropical forests. The FACF variables generally increase in
+proportion to one another, with few differences in allocation detectable
+at this global scale, but differed in that latitude explained a lower
+proportion of variation among subsidiary fluxes (in particular, woody
+aboveground \(NPP\) and belowground \(NPP\), \(BNPP\)). Climate
+explained a significant proportion (\#-\#\%) of variation in all C
+fluxes analyzed (less for subsidiary fluxes), with temperature variables
+in general and mean annual temperature (\(MAT\)) in particular being the
+best predictors of FACF on this global scale. While other climate
+variables (\emph{e.g.}, \textbf{XX}) displayed significant correlation
+with FACF, none of them had significantly better explanatory power than
+\(MAT\). The effects of temperature were modified by moisture
+availability, with reduced FACF under hot and dry conditions and
+sometimes under very high precipitaiton (especially for \(BNPP\)). FACF
+declined with temperature seasonality, but growing season length doesn't
+improve upon MAT as a predictor. Within the growing season, the
 influence of climate on C cycling is smaller but still significant for a
@@
 
+\newpage
+
 \subsubsection{Introduction}\label{introduction}
 
-(\textbf{revisit these first two paragraphs.})
-
-\textbf{Forests play a critical role in regulating atmospheric
-CO\textsubscript{2} and climate} (\textbf{Bonan}). Approximately
-\textbf{\#\#} Gt of CO\textsubscript{2} cycle through Earth's forests
-each year \textbf{REF-KAT}), and their net C sequestration
-(\textasciitilde{}2.4 GtCyr\textsuperscript{-1}) offsets roughly 30\% of
-anthropogenic fossil fuel emissions \citep{pan_large_2011}. As
-atmospheric carbon dioxide levels continue to rise, driving climate
-change, there is increasing recognition that protection and management
-of forest resources will play an important role in mitigating climate
-change (** REFS- e.g., IPCC 1.5C and land reports**). Carbon fixation,
-allocation, and metabolism by trees set the basis for energy and
-material flows in forest ecosystems and define their interactions with
-Earth's changing climate.
-
-\textbf{Forests' autotrophic carbon flux (FACF, bbl:define) influences
-all organic matter stocks in ecosystems and is linked to cycling of
-energy, water, and nutrients} (\textbf{REFS; \emph{Piao et al. 2010.
-Forest annual carbon cost: a global‐scale analysis of autotrophic
-respiration. Ecology 91(3), 652-661} }). Forest productivity sets the
-energy ultimately available to heterotrophic organisms (including
-microbes), in turn influencing their abundance (\textbf{REFS}) and
-possibly diversity
-(\href{https://www.annualreviews.org/doi/abs/10.1146/annurev.ecolsys.30.1.257}{Waide
-et al. 2009}\textbf{REFS}). However, our ability to draw general
-macroscopic conclusions regarding global variation in multiple
-productivity variables with respect to climate has been limited in that
-these analyses often mix forests that vary in stand age, disturbance
-history, and/or management status; do not always sufficiently parse
-related variables (\emph{e.g.}, combining net primary productivity
-records with and without belowground components); and typically consider
-only one or a few variables at a time.
-
-\textbf{The productivity of forests varies with latitude}, showing a
-general trend of decreasing productivity with latitude
-\citep{beer_terrestrial_2010, jung_global_2011}. Studies agree that
-productivity is lowest in the boreal regions, and increases into the
-temperate regions
+Carbon cycling in forests worldwide provides the energetic basis for
+sustaining the majority of Earth's terrestrial biodiversity and many
+human populations (\textbf{REF}), while strongly influencing atmospheric
+CO\textsubscript{2} and climate \citep{bonan_forests_2008}. Forests'
+autotrophic carbon fluxes (FACF)--that is, carbon fixation, allocation,
+and metabolism by trees and other primary producers--sets the energy
+ultimately available to heterotrophic organisms (including microbes), in
+turn influencing their abundance (\textbf{REFS}) and possibly diversity
+\citep{waide_relationship_1999} (\textbf{REFS}). FACF influences all
+organic matter stocks in forest ecosystems and is linked to cycling of
+energy, water, and nutrients (\textbf{REFS}) \citep{piao_forest_2010}.
+Critically, FACF also define forest interactions with Earth's changing
+climate. Over 69 Gt of CO\textsubscript{2} cycle through Earth's forests
+each year \citep{badgley_terrestrial_2019}, and in recent decades their
+net C sequestration (\textasciitilde{}2.4 Gt C yr\textsuperscript{-1})
+offset roughly 30\% of anthropogenic fossil fuel emissions
+\citep{pan_large_2011}. As atmospheric carbon dioxide levels continue to
+rise, driving climate change, forests will play a critical role in
+shaping the future of Earth's climate
+\citep{cavaleri_urgent_2015, rogelj_mitigation_2018}. However, our
+ability to draw general macroscopic conclusions regarding global
+variation in FACF with respect to climate has been limited in that these
+analyses often mix forests that vary in stand age, disturbance history,
+and/or management status; do not always sufficiently parse related
+variables; and typically consider only one or a few variables at a time.
+
+FACF decrease with latitude, but it remains unclear whether and how the
+shape of this relationship varies across fluxes. Studies agree that FACF
+are lowest in the boreal regions, and increase into the temperate
+regions
 \citep{luyssaert_co_2007, huston_global_2009, beer_terrestrial_2010, jung_global_2011}.
-However, evidence is inconclusive on whether productivity continues to
-increase into the tropics, or whether it plateaus in temperate regions.
-Evidence for this is further complicated by the fact that different
-studies use different measures of productivity to explore these
-relationships. For example, modelling of global terrestrial ecosystem
-gross primary productivity (GPP) through upscaling and calibration of
-eddy flux measurements indicates that GPP peaks in tropical forests
-\citep{beer_terrestrial_2010, jung_global_2011};\href{https://onlinelibrary.wiley.com/doi/epdf/10.1111/gcb.14729}{Bagdley
-et al. 2019}). This is corroborated by analysis of site-level GPP
-measurements, which appear to reach their highest levels in tropical
-forests \citep{luyssaert_co_2007}. \{\textbf{condense the previous two
-sentences into one-- both primarily eddy flux}\} In contrast, some
-studies suggest that the highest values of net primary productivity
-(NPP) may be found in temperate forests
-\citep{luyssaert_co_2007, huston_global_2009}, while others find NPP
+However, evidence is inconclusive on whether primary productivity
+continues to increase into the tropics, or whether it plateaus in
+temperate regions. Evidence for this is further complicated by the fact
+that different studies use different measures of productivity to explore
+these relationships. For example, modelling of global terrestrial
+ecosystem gross primary productivity (\(GPP\)) through upscaling and
+calibration of eddy flux measurements indicates that GPP peaks in
+tropical forests
+\citep{beer_terrestrial_2010, jung_global_2011, badgley_terrestrial_2019, li_mapping_2019}.
+In contrast, some studies suggest that the highest values of net primary
+productivity (\(NPP\)) may be found in temperate forests
+\citep{luyssaert_co_2007, huston_global_2009}, while others find \(NPP\)
 highest in the tropics and decreasing with latitude
 \citep{simova_enigma_2017}. Other studies have chosen to focus
-exclusively on above-ground net primary productivity (ANPP), finding
-evidence of a weak negative relationship between ANPP and latitude
+exclusively on above-ground net primary productivity (\(ANPP\)), finding
+evidence of a weak negative relationship between \(ANPP\) and latitude
 \citep{huston_global_2009, gillman_latitude_2015}.
 
-\textbf{Climate is a significant driver of productivity across broad
-spatial scales} \citep{cleveland_relationships_2011}. The majority of
-studies have focused on exploring the relationships of productivity to
-mean annual temperature (MAT) and mean annual precipitation (MAP), as
-the most commonly reported site-level climate variables. These variables
-have the advantage that they describe broad trends in temperature and
-water availability, and therefore capture a lot of global-scale
-variation in climate. There is strong evidence that both MAT and MAP
-show significant positive relationships with productivity
+The latitudinal gradient in FACF is primarily driven by climate, which
+is a significant driver of FACF across broad spatial scales
+\citep{luyssaert_co_2007, cleveland_relationships_2011, hursh_sensitivity_2017}.
+The majority of studies have focused on exploring the relationships of
+FACF to mean annual temperature (MAT) and mean annual precipitation
+(MAP), as the most commonly reported site-level climate variables. While
+these fail to capture some important aspects of climate such as
+seasonality, they do describe broad trends in temperature and water
+availability, and therefore capture a substantial portion of
+global-scale variation in climate. There is strong evidence that both
+MAT and MAP show significant positive relationships with FACF
 \citep{chu_does_2016}. However, as with latitude, the shape of those
 relationships is not always clear, and, again, is complicated by the use
-of different measures of productivity across studies. Various measures
-of primary productivity saturate at high levels of MAP, though the
-saturation points identified vary from 1500mm \citep{luyssaert_co_2007}
-up to 2445mm MAP \citep{schuur_productivity_2003}. Studies of the
-influence of MAT on productivity are less conclusive. Luyssaert et al.
+of different measures of FACF across studies. Various measures of
+primary productivity \{\textbf{FACF?}\} saturate at high levels of MAP,
+though the saturation points identified vary from 1500mm
+\citep{luyssaert_co_2007} up to 2445mm MAP
+\citep{schuur_productivity_2003}. Studies of the influence of MAT on
+productivity \{\textbf{FACF?}\} are less conclusive. Luyssaert et al.
 \citeyearpar{luyssaert_co_2007} examined GPP and NPP and found that,
-while GPP increases linearly with MAT, NPP saturates at around
+while \(GPP\) increases linearly with MAT, NPP saturates at around
 10\(^\circ\)C MAT. In contrast, Larjavaara and Muller-Landau
-\citeyearpar{larjavaara_temperature_2012}, find that increases in GPP
-saturate at approximately 25\(^\circ\)C MAT, while Schuur
-\citeyearpar{schuur_productivity_2003} shows that NPP increases linearly
-with temperature. bbl:maybe remove following sentences The influence of
-these climate variables on productivity is further complicated by the
-possibility of interactive effects occurring between them. Taylor et al.
-\citeyearpar{taylor_temperature_2017} showed that increased MAP had a
-negative effect at low MAT, but a positive effect at high levels of MAT,
-and \emph{vice versa} \{\textbf{Helene found this confusing}\},
-suggesting that both temperature and water availability are important in
-explaining productivity.
-
-\textbf{Primary productivity can be influenced by many other factors as
-well}, which often act across a range of scales, and may show
-interactive effects with each other
-\citep{cleveland_relationships_2011}. On a local scale, stand age
-\citep{litton_carbon_2007, gillman_latitude_2015}, biodiversity (Liang
-et al. 2019. Positive biodiversity-productivity relationship predominant
-in global forests. Science 354, Issue 6309, aaf8957), management
-\citep{simova_enigma_2017}; nutrient availability
-\citep{aragao_above-_2009}; and altitude
-\citep{girardin_net_2010, malhi_variation_2017} all impact forest
-productivity. On a global scale, we expect that productivity is most
-strongly influenced by broad climatic gradients. There is evidence that
-productivity also responds to variables such as cloud cover
-\citep{taylor_temperature_2017}, solar radiation
-\citep{fyllas_solar_2017}, and potential evapotranspiration
+\citeyearpar{larjavaara_temperature_2012}, find that increases in
+\(GPP\) saturate at approximately 25\(^\circ\)C MAT, while Schuur
+\citeyearpar{schuur_productivity_2003} shows that \(NPP\) increases
+linearly with temperature. Taylor et al.
+\citeyearpar{taylor_temperature_2017} showed a positive interaction
+between MAT and MAP in shaping tropical forest productivity, such that
+high rainfall had a negative effect on productivity in cooler climates,
+compared to a positive effect in warmer climates. \{\textbf{It would be
+good to add some more citations on soil respiraiton. I'm sure BBL can
+help.}\}
+
+FACF can be influenced by many other factors as well, which often act
+across a range of scales, and may show interactive effects with each
+other \citep{cleveland_relationships_2011}. On a local scale, stand age
+\citep{litton_carbon_2007, gillman_latitude_2015}, biodiversity
+\citep{liang_positive_2016}, management \citep{simova_enigma_2017},
+nutrient availability \citep{aragao_above_2009}, and altitude
+\citep{girardin_net_2010, malhi_variation_2017} all impact FACF. On a
+global scale, we expect that FACF are most strongly influenced by broad
+climatic gradients. There is evidence that FACFs also respond to
+variables such as cloud cover \citep{taylor_temperature_2017}, solar
+radiation \citep{fyllas_solar_2017}, and potential evapotranspiration
 \citep{kerkhoff_plant_2005} in potentially significant ways.
@@
 studies have suggested that the apparently strong relationship between
-MAT and productivity is actually a consequence of the correlation
-between MAT and growing season length
+MAT and FACFs is actually a consequence of the correlation between MAT
+and growing season length
 \citep{kerkhoff_plant_2005, malhi_productivity_2012, michaletz_convergence_2014, michaletz_drivers_2018}.
@@
 \citeyearpar{michaletz_convergence_2014} find that, within the growing
-season, there is no significant relationship between productivity and
-MAT, suggesting that the effect of temperature is due to increased
-length of growing season, rather than an inherent influence of
-temperature on productivity.
-
-\textbf{The recent development of a global forest carbon database
-synthesizing multiple variables and including records of stand history
-(ForC;
-\citep{anderson-teixeira_carbon_2016, anderson-teixeira_forc:_2018})
+season, there is no significant relationship between primary
+productivity and MAT, suggesting that the effect of temperature is due
+to increased length of growing season, rather than an inherent influence
+of temperature on FACF.
+
+The recent development of the Global Forest Carbon database (ForC),
+which synthesizes multiple variables and including records of stand
+history
+\citep{anderson-teixeira_carbon_2016, anderson-teixeira_forc_2018},
 opens up the possibility for a standardized analysis of global scale
-variation} in multiple components of forest productivity and the
-principle climatic drivers of these patterns. In order to approach these
-broad and complex issues, we simplify the major gaps in our knowledge to
-five broad hypotheses and corresponding specific predictions (Table 1).
-** see
-\href{https://github.com/forc-db/Global_Productivity/issues/48}{issue
-\#48}** First, we ask how forest autotrophic carbon fluxes (FACF) vary
-with latitude. We then test how these fluxes relate to MAT and MAP, and
-additionally how they respond to other, less well studied, climate
-variables. Finally, we consider the relationship between FACF and
-seasonality, considering the role of seasonality in explaining variation
-in carbon fluxes, and the influence of climate on FACF standardized by
-growing season length. We use a comprehensive global database of forest
-carbon fluxes to address the above questions for nine carbon fluxes,
-allowing for an in-depth exploration of the effect of climate on global
-productivity.
+variation in multiple FACF and the principle climatic drivers of these
+patterns. In order to approach this broad topic, we simplify the major
+gaps in our knowledge to five broad questions and corresponding
+hypotheses (Table 1). First, we ask how FACF vary with latitude. We then
+test how these fluxes relate to MAT and MAP, and additionally how they
+respond to other, less well-studied, climate variables. Finally, we
+consider the relationship between FACF and seasonality, considering the
+role of seasonality in explaining variation in carbon fluxes, and the
+influence of climate on FACF standardized by growing season length. We
+address the above questions for nine carbon fluxes contained in ForC,
+allowing for an in-depth exploration of the effect of climate on FACF
+globally.
 
@@
 
-\caption{\label{tab:unnamed-chunk-4}**Summary of hypotheses, corresponding specific predictions, and results.** Direction of significant relationships is indicated by '+' and '-', and  shape is summarized as linear (L), concave up (CU), and concave down (CD).
-'n.s.' indicates that the relationship is not signficant.}
+\caption{\label{tab:unnamed-chunk-4}Summary of research questions, corresponding hypotheses, and results.
+Statistically signficant support for/ rejection of hypotheses is indicated with 'yes'/'no', parentheses indicate partial overall support/rejection of hypotheses across all FACF, and '-' indicates no significant relationship.}
 \centering
@@
 \fontsize{12}{14}\selectfont
-\begin{tabular}{>{\raggedright\arraybackslash}p{10cm}>{\raggedright\arraybackslash}p{5cm}>{\raggedright\arraybackslash}p{1.8cm}>{\raggedright\arraybackslash}p{1.8cm}>{\raggedright\arraybackslash}p{1.8cm}>{\raggedright\arraybackslash}p{1.8cm}>{\raggedright\arraybackslash}p{1.8cm}>{\raggedright\arraybackslash}p{1.8cm}>{\raggedright\arraybackslash}p{1.8cm}>{\raggedright\arraybackslash}p{1.8cm}>{\raggedright\arraybackslash}p{1.8cm}>{\raggedright\arraybackslash}p{1.8cm}>{\raggedright\arraybackslash}p{1.8cm}}
+\begin{tabular}{llllllllllll}
 \toprule
-\multicolumn{1}{c}{ } & \multicolumn{1}{c}{ } & \multicolumn{1}{c}{ } & \multicolumn{9}{c}{Forest autotrophic carbon fluxes (FACF)} & \multicolumn{1}{c}{ } \\
-\cmidrule(l{3pt}r{3pt}){4-12}
-Hypotheses \& Specific Predictions & Related references & Overall & GPP & NPP & ANPP & BNPP & ANPP foliage & ANPP woody stem & BNPP fine root & R auto root & R auto & Support\\
+\multicolumn{1}{c}{ } & \multicolumn{1}{c}{ } & \multicolumn{9}{c}{Forest autotrophic carbon fluxes (FACF)} & \multicolumn{1}{c}{ } \\
+\cmidrule(l{3pt}r{3pt}){3-11}
+Questions and hypotheses (with related references) & Overall & $GPP$ & $NPP$ & $ANPP$ & $ANPP_{woody.stem}$ & $ANPP_{foliage}$ & $BNPP$ & $BNPP_{fine.root}$ & $R_{auto}$ & $R_{auto-root}$ & Support\\
 \midrule
 \addlinespace[1em]
-\hline
-\multicolumn{1}{l}{\textbf{H1. FACFs decrease linearly with latitude, and in constant proportion.}}\\
-\hspace{1em}1.1. FACFs decrease linearly with latitude (L-) & Luyssaert et al. (2007) & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Fig. 2\\
-\hspace{1em}1.1.alt.
-FACFs are similar in tropical and temperate forests, but lower in boreal regions (CD-) & Gillman et al.
-(2015); Simova and Storch (2017) & no & no & no & no & no & no & no & no & no & no & Fig.
-2\\
-\hspace{1em}1.2. Allocation of GPP to subsidiary fluxes varies with latitude & Litton et al. (2007); DeLucia et al. (2007) & no & - & no & no & no & no & no & no & - & - & \\
+\multicolumn{5}{l}{\textbf{Q1. How do FACF vary with latitude?}}\\
+\hspace{1em}H1.1. FACF decrease linearly with latitude.\textsuperscript{1}$^{,}$\textsuperscript{2}$^{,}$\textsuperscript{3} & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Fig. 2\\
 \addlinespace[1em]
 \hline
-\multicolumn{1}{l}{\textbf{H2. FACFs increase with MAT, with an interactive effect of MAP.}}\\
-\hspace{1em}2.1. FACFs increase linearly with MAT (L+) & Schuur (2003) & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Fig. 4\\
-\hspace{1em}2.2.
-FACFs increase with precipitation but saturate or decrease at very high levels (CD+) & Luyssaert et al.
-(2007); Schuur (2003) & yes & yes & yes & yes & yes & yes & n.s.
-& L+ & L+ & yes & Fig.
-4\\
-\hspace{1em}2.3. There is a postive interaction between temperature and precipitation (I+) & Taylor et al. (2016) & (yes) & yes & L+ & yes & L- & L+ & yes & L- & yes & yes & Fig. 3\\
+\multicolumn{5}{l}{\textbf{Q2. How do FACF vary with MAT and MAP?}}\\
+\hspace{1em}H2.1. FACF increase with MAT.\textsuperscript{1}$^{,}$\textsuperscript{4} & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Figs. 4, S4, S5\\
+\hspace{1em}H2.2. FACF increase with precipitation.\textsuperscript{1}$^{,}$\textsuperscript{4} & (yes) & yes & yes & yes & - & yes & yes & yes & yes & yes & Figs. 4, S4, S5\\
+\hspace{1em}H2.3. Temperature and precipitation interactively shape FACF.\textsuperscript{5} & (yes) & yes & yes & - & yes & - & yes & yes & yes & - & Fig. 3\\
 \addlinespace[1em]
 \hline
-\multicolumn{1}{l}{\textbf{H3. FACFs are strongly correlated with other annual climate variables.}}\\
-\hspace{1em}3.1. FACFs increases with PET, but saturates or decreases at high levels (CD+) &  & (yes) & yes & yes & yes & yes & L+ & yes & L+ & yes & L+ & Fig. 4\\
-\hspace{1em}3.2. FACFs increase with vapour pressure deficit, but saturate or decrease at high levels (CD+) &  & (yes) & yes & yes & yes & yes & yes & yes & L+ & yes & yes & Fig. 4\\
-\hspace{1em}3.3. FACFs increase linearly with solar radiation (L+) &  & (yes) & yes & yes & yes & CD+ & yes & CD+ & yes & n.s. & yes & \\
+\multicolumn{5}{l}{\textbf{Q3. How are FACF related to other climate variables?}}\\
+\hspace{1em}H3.1. FACF increase with PET. & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Figs. 4, S4, S5\\
+\hspace{1em}H3.2. FACF increase with vapour pressure deficit. & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Figs. 4, S4, S5\\
+\hspace{1em}H3.3. FACF increase with solar radiation. & (yes) & yes & yes & yes & yes & yes & yes & yes & yes & - & Figs. S4, S5\\
 \addlinespace[1em]
 \hline
-\multicolumn{1}{l}{\textbf{H4. FACFs are reduced under seasonal climates.}}\\
-\hspace{1em}4.1. FACFs decrease linearly with temperature seasonality (L-) &  & yes & CU+ & CU+ & CU+ & yes & yes & yes & yes & CU+ & CU+ & Fig. 4\\
-\hspace{1em}4.2. FACFs decrease linearly with  precipitation seasonality (L-) &  & no & n.s. & n.s. & n.s. & n.s. & n.s. & n.s. & n.s. & n.s. & n.s. & \\
-\hspace{1em}4.3.
-FACFs increase linearly with growing season length (L+) & Malhi (2012); Michaletz et al.
-(2014); Chu et al.
-(2016) & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Fig.
-4\\
-\hspace{1em}4.4. Growing season length is a better predictor of FACFs than MAT & Michaletz et al. (2014); Chu et al. (2016) & no & no & no & no & no & no & no & no & no & no & \\
+\multicolumn{5}{l}{\textbf{Q4. How does seasonality influence FACF?}}\\
+\hspace{1em}H4.1. FACF decrease with temperature seasonality. & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Figs. 4, S6, S7\\
+\hspace{1em}H4.2. FACF decrease with precipitation seasonality. & - & - & - & - & - & - & - & - & - & - & Figs. S6, S7\\
+\hspace{1em}H4.3.
+FACF increase with growing season length.\textsuperscript{6}$^{,}$\textsuperscript{7}$^{,}$\textsuperscript{8} & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Figs.
+4, S6, S7\\
+\hspace{1em}H4.4. Growing season length is a better predictor of FACF than MAT.\textsuperscript{7}$^{,}$\textsuperscript{8} & (no) & no & no & no & no & no & no & - & no & no & Table S4\\
 \addlinespace[1em]
 \hline
-\multicolumn{1}{l}{\textbf{H5. Considering only growing season months, FACF correlates with climatic drivers.}}\\
-\hspace{1em}5.1. Increase with temperature (L+) & Michaletz et al. (2014) & (mixed) & n.s. & n.s. & yes & n.s. & yes & n.s. & n.s. & n.s. & n.s. & \\
-\hspace{1em}5.2. Increase with PET (L+) &  & (yes) & yes & yes & n.s. & yes & n.s. & yes & yes & n.s. & n.s. & \\
-\hspace{1em}5.3. Increase with precipitation (L+) &  & no & n.s. & n.s. & yes & n.s. & yes & n.s. & n.s. & n.s. & n.s. & \\
-\hspace{1em}5.4. Increase with solar radiation (L+) &  & (mixed) & yes & yes & n.s. & yes & n.s. & n.s. & yes & n.s. & n.s. & \\
+\multicolumn{5}{l}{\textbf{Q5. When standardised by growing season length, how do FACF vary with climate?}}\\
+\hspace{1em}H5.1. Growing season FACF increase with temperature.\textsuperscript{8} & (yes) & - & - & yes & - & yes & - & - & - & - & Figs. S8, S9\\
+\hspace{1em}H5.2. Growing season FACF increase with PET. & (yes) & yes & yes & - & yes & - & yes & yes & - & - & Figs. S8, S9\\
+\hspace{1em}H5.3. Growing season FACF increase with precipitation. & (yes) & - & - & yes & - & yes & - & - & - & - & Figs. S8, S9\\
+\hspace{1em}H5.4. Growing season FACF increase with solar radiation. & (yes) & yes & yes & - & - & - & yes & yes & - & - & Figs. S8, S9\\
 \bottomrule
+\multicolumn{12}{l}{\textsuperscript{1} Luyssaert et al.
+(2007) \textsuperscript{2} Gillman et al.
+(2015) \textsuperscript{3} Simova and Storch (2017) \textsuperscript{4} Schuur (2003) \textsuperscript{5} Taylor et al.
+(2016) \textsuperscript{6} Malhi (2012) \textsuperscript{7} Michaletz et al.
+(2014) \textsuperscript{8} Chu et al.
+(2016)}\\
 \end{tabular}}
@@
 
-Analyses were conducted on data contained in the open-access ForC
-database
-\citep{anderson-teixeira_carbon_2016, anderson-teixeira_forc:_2018}.
-This database contains records of field-based measurements of forest
-carbon stocks and annual fluxes, compiled from original publications and
-existing data compilations and databases. Associated data, such as stand
-age, measurement methodologies, and disturbance history, are also
-included. The database was significantly expanded since the publication
-of \citep{anderson-teixeira_forc:_2018} through integration with the
-Global Soil Respiration Database \citep{bond-lamberty_global_2010}.
-Additional targeted literature searches were conducted to identify any
-further available data on primary productivity, with particular focus on
-mature forests in temperate and boreal regions. ForC currently contains
-29730 records from 4979 plots, representing 20 distinct ecozones across
-all forested biogeographic and climate zones. We used ForC v3.0,
-archived on Zenodo with DOI 10.5281/zenodo.3403855.
+\emph{Forest carbon flux data}
+
+This analysis focused on nine FACF included in the open-access ForC
+database (Table 2)
+\citep{anderson-teixeira_carbon_2016, anderson-teixeira_forc_2018}. ForC
+contains records of field-based measurements of forest carbon stocks and
+annual fluxes, compiled from original publications and existing data
+compilations and databases. Associated data, such as stand age,
+measurement methodologies, and disturbance history, are also included.
+The database was significantly expanded since the publication of
+\citet{anderson-teixeira_forc_2018} through integration with the Global
+Soil Respiration Database \citep{bond-lamberty_global_2010}. Additional
+targeted literature searches were conducted to identify any further
+available data on the FACF analyzed here, with particular focus on
+mature forests in temperate and boreal regions, which were not included
+in the review of \citet{anderson-teixeira_carbon_2016}. We used ForC
+v3.0, archived on Zenodo with DOI 10.5281/zenodo.3403855. This version
+contained 29,730 records from 4,979 plots, representing 20 distinct
+ecozones across all forested biogeographic and climate zones.
+
+This analysis focused on mature forests with no known history of
+signficant disturbance or management. There is evidence that stand age
+influences patterns of FACF and carbon allocation in forest ecosystems,
+and can confound relationships between latitude and primary productivity
+\citep{delucia_forest_2007, gillman_latitude_2015}. To reduce any
+biasing effects of stand age, we included only stands of known age
+\(\ge\) 100 years and those described by terms such as ``mature'',
+``intact'', or ``old-growth''. Since management can alter observed
+patterns of FACF \citep{simova_enigma_2017}, sites were excluded from
+analysis if they were managed, defined as plots that were planted,
+managed as plantations, irrigated, fertilised or including the term
+``managed'' in their site description. Sites that had experienced
+significant disturbance within the past 100 years were also excluded.
+Disturbances that qualified sites for exclusion included major cutting
+or harvesting, burning, flooding, drought and storm events with site
+mortality \textgreater{}10\% of trees. Grazed sites were retained.
+
+\begin{table}[!h]
+
+\caption{\label{tab:unnamed-chunk-5}Definitions and sample sizes of FACF variables used in analysis. All variables are in units of Mg C ha$^{-1}$ yr$^{-1}$. }
+\centering
+\resizebox{\linewidth}{!}{
+\fontsize{12}{14}\selectfont
+\begin{tabular}{l>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{7cm}>{\raggedright\arraybackslash}p{10cm}>{\raggedleft\arraybackslash}p{1.5cm}>{\raggedleft\arraybackslash}p{1.5cm}}
+\toprule
+\multicolumn{1}{c}{ } & \multicolumn{1}{c}{ } & \multicolumn{1}{c}{ } & \multicolumn{1}{c}{ } & \multicolumn{2}{c}{Sample size} \\
+\cmidrule(l{3pt}r{3pt}){5-6}
+Variable & Definition & Components included & Methodologies & records & geographic areas\textsuperscript{*}\\
+\midrule
+$GPP$ & Gross Primary Production & full ecosystem & flux partitioning of eddy-covariance; $NPP$+$R_{auto}$ & 243 & 49\\
+$NPP$ & Net Primary Production & stem, foliage, coarse root, fine root, optionally others (e.g., branch, reproductive, understory) & $ANPP$ + $BNPP$ (majority); $GPP$-$R_{auto}$ & 161 & 56\\
+$ANPP$ & Aboveground $NPP$ & stem, foliage, optionally others (e.g., branch, reproductive, understory) & $ANPP_{woody-stem}$ + $ANPP_{foliage}$ (+ others) & 278 & 86\\
+$ANPP_{woody.stem}$ & Woody stem growth component of $ANPP$ & woody stems down to DBH $\le$ 10cm (no branch turnover) & stem growth measurements scaled to biomass using allometries & 264 & 96\\
+$ANPP_{foliage}$ & Foliage component of $ANPP$ & foliage & litterfall collection (separated into components) & 98 & 49\\
+\addlinespace
+$BNPP$ & Belowground $NPP$ & coarse and fine roots & coarse roots estimated indirectly using allometries based on aboveground stem increment measures ; fine roots as below & 101 & 48\\
+$BNPP_{fine.root}$ & Fine root component of $BNPP$ & fine roots & measurements combined one or more of the following: soil cores, minirhizotrons, turnover estimates,  root ingrowth cores & 88 & 41\\
+$R_{auto}$ & Autotrophic respiration & foliage, stem, and root & chamber measurements of foliage and stem gas exchange +  $R_{auto-root}$ (as below) & 22 & 13\\
+$R_{auto-root}$ & Root respiration & (coarse and) fine roots & partitioning of total soil respiration (e.g., through root exclusion), scaling of root gas exchange; excluded alkali absoption and soda lime methods for measuring soil respiration & 64 & 26\\
+\bottomrule
+\multicolumn{6}{l}{\textsuperscript{*} Geographic areas group geographically proximate sites, defined using a hierarchical cluster analysis on the distance matrix of the sites, and a cutoff of 25km}\\
+\end{tabular}}
+\end{table}
 
@@
 S1.
-}\label{fig:unnamed-chunk-5}
+}\label{fig:unnamed-chunk-6}
 \end{figure}
 
-\emph{Data selection.} Over 50 variables of forest carbon stocks and
-annual fluxes are represented in the ForC database; this analysis
-focused on autotrophic carbon fluxes. Focal variables of this analysis
-are listed in Table 2.
-
-A subset of the ForC database was generated for the purposes of this
-analysis, in order to control for data quality and remove biasing
-factors. Since management can alter observed patterns of primary
-productivity \citep{simova_enigma_2017}, sites were excluded from
-analysis if they were managed, defined as plots that were planted,
-managed as plantations, irrigated, fertilised or including the term
-``managed'' in their site description. Sites that had experienced
-significant disturbance were also excluded. Disturbances that justified
-site exclusion were major cutting or harvesting, and/or burning,
-flooding, drought and storm events with site mortality
-\textgreater{}10\% of trees. Grazed sites were retained.\\
-There is evidence that stand age influences patterns of primary
-productivity and carbon allocation in forest ecosystems, and can
-confound relationships between latitude and primary productivity
-\citep{de_lucia_forest_2007, gillman_latitude_2015}. To reduce any
-biasing effects of stand age, we included only stands of known age
-\(\ge\) 100 years and those described by terms such as ``mature'',
-``intact'', or ``old-growth''.
-
-\newpage
-
-\begin{longtable}[]{@{}llllll@{}}
-\caption{Definitions and sample sizes of variables (all annual) used in
-analysis. Geographic areas group geographically proximate sites, defined
-using a hierarchical cluster analysis on the distance matrix of the
-sites, and a cutoff of 25km.}\tabularnewline
-\toprule
-\begin{minipage}[b]{0.14\columnwidth}\raggedright\strut
-Variable\strut
-\end{minipage} & \begin{minipage}[b]{0.19\columnwidth}\raggedright\strut
-Definition\strut
-\end{minipage} & \begin{minipage}[b]{0.13\columnwidth}\raggedright\strut
-Components included\strut
-\end{minipage} & \begin{minipage}[b]{0.23\columnwidth}\raggedright\strut
-Methodologies used\strut
-\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\raggedright\strut
-Number of records\strut
-\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\raggedright\strut
-Number of geographic areas\strut
-\end{minipage}\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-\begin{minipage}[b]{0.14\columnwidth}\raggedright\strut
-Variable\strut
-\end{minipage} & \begin{minipage}[b]{0.19\columnwidth}\raggedright\strut
-Definition\strut
-\end{minipage} & \begin{minipage}[b]{0.13\columnwidth}\raggedright\strut
-Components included\strut
-\end{minipage} & \begin{minipage}[b]{0.23\columnwidth}\raggedright\strut
-Methodologies used\strut
-\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\raggedright\strut
-Number of records\strut
-\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\raggedright\strut
-Number of geographic areas\strut
-\end{minipage}\tabularnewline
-\midrule
-\endhead
-\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
-\(GPP\)\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
-Gross Primary Production\strut
-\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
-NA\strut
-\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-flux partitioning of eddy-covariance; \(NPP\)+\(R_{auto}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
-243\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
-49\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
-\(NPP\)\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
-Net Primary Production\strut
-\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
-stem, foliage, coarse root, fine root, optionally others (e.g., branch,
-reproductive, understory)\strut
-\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-\(ANPP\) + \(BNPP\) (majority); \(GPP\)-\(R_{auto}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
-92\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
-42\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
-\(ANPP\)\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
-Aboveground Net Primary Production\strut
-\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
-stem, foliage, optionally others (e.g., branch, reproductive,
-understory) (\textbf{issues
-\href{https://github.com/forc-db/Global_Productivity/issues/52}{52},
-\href{https://github.com/forc-db/Global_Productivity/issues/54}{54}})\strut
-\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-\(ANPP_{woody-stem}\) + \(ANPP_{foliage}\) (+ others)\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
-256\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
-80\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
-\(ANPP_{woody-stem}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
-Net Primary Production of woody stems\strut
-\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
-woody stems down to DBH \(\le\) 10cm (no branch turnover)\strut
-\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-stem growth measurements scaled to biomass using allometries
-(\textbf{with recruitment?})\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
-264\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
-96\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
-\(ANPP_{foliage}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
-Net Primary Production of foliage\strut
-\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
-foliage\strut
-\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-litterfall collection (separated into components)\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
-98\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
-49\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
-\(BNPP_{root}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
-Belowground Net Primary Production\strut
-\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
-coarse and fine roots\strut
-\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-coarse roots estimated indirectly using allometries based on aboveground
-stem increment measures ; fine roots as below\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
-101\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
-48\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
-\(BNPP_{fine.root}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
-Net Primary Production of fine roots\strut
-\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
-fine roots\strut
-\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-measurements combined one or more of the following: minirhizotrons, fine
-root turnover, soil cores, root ingrowth cores\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
-88\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
-41\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
-\(R_{root}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
-Root Respiration\strut
-\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
-coarse and fine roots\strut
-\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-partitioning of total soil respiration (e.g., through root exclusion),
-scaling of root gas exchange\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
-64\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
-26\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
-\(R_{auto}\)\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
-Autotrophic Respiration\strut
-\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
-foliage, stem, and root\strut
-\end{minipage} & \begin{minipage}[t]{0.23\columnwidth}\raggedright\strut
-chamber measurements of component gas exchange\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
-22\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
-13\strut
-\end{minipage}\tabularnewline
-\bottomrule
-\end{longtable}
-
-\emph{Methodological consistency.} The data in ForC is derived from a
-range of studies, often employing different methods. For this reason,
-criteria were introduced to standardise for differences in methodology.
-Where data was based on forest plot census measurements, studies which
-used a minimum diameter at breast height (DBH) measure of
-\textgreater{}10cm were excluded from analysis. It would be preferable
-to standardise by minimum area sampled; however x\% of plots in the
-database are 1 ha or under in size; excluding these plots would place
-significant constraints on sample size. (\textbf{really, we didn't
-exclude any studies based on methodology (all studies measured min DBH
-10 or less\ldots{} ) so this paragraph needs revision. Also, address
-whether we standardize for inclusion/ exclusion of recruitment, and
-address branchfall.})
-
-As discussed above, estimates of \(NPP\), \(ANPP\), and \(BNPP\) are
-generated through summing measurements of their component parts. Since
-the components included in productivity estimates vary between studies,
-estimates of productivity were classified within the ForC database
-according to their components, and then filtered for analysis. Estimates
-of NPP were selected if they included foliage, branch, stem, coarse
-root, and fine root. Measures of NPP which included additional
-components, including understorey, volatile organic compounds (VOCs),
-exudates, estimates of NPP lost to herbivory (\textbf{these are all
-small, no need to exclude}), and the NPP of reproductive structures
-(\textbf{often in literfall/ mixed in with foliage\ldots{} did we really
-exclude estimates including litterfall?}), were excluded. Estimates of
-ANPP were selected if they included foliage, stem growth and optionally
-branch turnover. Any measures of primary productivity where components
-were unknown were excluded from analysis.
-
-\emph{Climate datasets.} Where site-level data on mean annual
-temperature, mean annual precipitation, and latitude were available in
-the primary literature, this data was compiled and entered directly into
-the ForC database. Based on the geographic co-ordinates for each site,
-data on a further 11 climate variables was extracted from five
-open-access climate datasets: WorldClim \citep{hijmans_very_2005},
-WorldClim2 \citep{fick_worldclim_2017}, the Climate Research Unit (CRU)
-time-series dataset v. 4.03 \citep{harris_updated_2014}, the Global
-Aridity Index and Potential Evapotranspiration Climate Database
-\citep{trabucco_global_2018}, and TerraClimate
-\citep{abatzoglou_terraclimate_2018} (see Supplementary Information S1
-for details of climate variables). Where site-level data was missing for
-mean annual temperature and/or mean annual precipitation, data was
-extracted from the WorldClim dataset.
-
-Additionally, two climate variables were derived from the above
-datasets: maximum vapour pressure deficit, defined as the vapour
-pressure deficit of the month with the largest deficit; and water stress
-months, defined as the number of months annually where precipitation was
-lower than potential evapotranspiration (PET; \textbf{HML:defined
-how?}).
-
-\emph{Length of growing season.} Growing season months were defined as
-months with mean minimum temperature \textgreater{} 0.5\(^\circ\)C.
-Growing season months were initially calculated following methods used
-by Kerkhoff et al. (2005), which additionally required that growing
-season months had a moisture index, defined as (MAT - PET)/PET,
-\textgreater{} -0.95. Michaletz et al. (2014) included an equivalent
-requirement in their calculation of growing season length. However, we
-found that including this requirement had no effect on the estimates of
-growing season length, and so chose to exclude it.
-
-Monthly data for PET, precipitation, and temperature was downloaded from
-the Climate Research Unit (CRU) time-series dataset v 4.03
-\citep{harris_updated_2014}, and for solar radiation from WorldClim2
-\citep{fick_worldclim_2017}, and used to calculate mean monthly PET,
-precipitation, temperature and solar radiation during the growing
-season. Total growing season precipitation and solar radiation were also
-calculated.
-
-\emph{Model specification.}
-
-\{\textbf{from Helene: It was frequently unclear to me whether the
-influences of other climate variables were being considered in
-multivariate or univariate analyses. This needs to be clear. I can see
-reasons to do it either way, but they need to be differently motivated
-and clearly presented.}\}
-
-The effects of climate and latitude on primary productivity were
-analysed using mixed effects models using the package `lme4'
-\citep{bates_fitting_2015} in R v.3.5.1 \citep{r_core_team_r:_2018}. The
-effect of each extracted climate variable on each measure of primary
-productivity was modelled by specifying the climate variable as a fixed
-effect. For each climate variable, three models were specified: a null
-model; a model with the climate variable as a linear term; and a model
-with the climate variable as a \textbf{XX}-order polynomial term. AIC
-values were calculated for the models and used to select the best model.
-If the best model included a polynomial term, the shape of the
-polynomial relationship was considered. If the shape of the relationship
-\emph{made biological sense}, and was a significant improvement on the
-linear relationship (deltaAIC \textgreater{}2), we accepted the
-polynomial as the best model. If not, we ran the linear model as the
-final model. \(R^2\) values were calculated for the best model. All
-\(R^2\) values presented here are marginal \(R^2\) values, and refer to
-the proportion of variation explained by only the fixed effects, unless
-otherwise specified. In addition, slope coefficients were calculated for
-the linear models.
-
-To test for a potential influence of altitude, models were also run with
-site altitude included as a second fixed effect. These models were
-compared against models with no altitude term, and AIC values calculated
-to identify whether inclusion of altitude as a term improved the models.
-Including altitude had a very small effect on most models, with the
-difference in AIC values between models including and excluding altitude
-often being \textless{}2, suggesting the models are very similar in
-their explanatory power. As a result, it was decided to present results
-only from models without altitude as a fixed term.
-
-Within the ForC database, sites within 25km\textsuperscript{2} of each
-other are clustered into geographic areas. To account for correlations
-in measurements between tightly clustered sites, a random effect was
-specified as plot nested within geographic area. \emph{Data from the
-temperate regions was heavily skewed towards studies from the old-growth
-forests of the Pacific Northwest.} These forests have very high FACF,
-and so to ensure that results were not unduly influenced by geographic
-sampling bias, we tried a version of the model where data were weighted
-according to forested land area within each Koeppen climate zone.
-Results were similar between the weighted and unweighted model. Because
-the weighted model may be prone to problems of over-fitting, we present
-only results of the unweighted model. \{**Becky, I'd probably just
-delete this whole paragraph. Since we ended up not using the weighted
-model, it's just extra info for readers to digest.\}
-
-Models were run for total annual FACF against annual climate variables,
-and for fluxes standardized by growing season length, defined as annual
-FACF/length of growing season, against growing season climate variables.
-For analyses on data within the growing season, only linear models were
-specified.
-
-\{\textbf{Note from Helene on the following paragraph: this makes it
-sound like analyses from H3 were about variation remaining after
-accounting for temperature, but analyses below don't seem to do
-that.}\}\\
-To investigate the potential interactive effects of climate variables on
-carbon fluxes, multivariate models were also specified. To ensure that
-models were biologically meaningful, the terms included in the models
-tested built on results from the univariate models. Modelling of
-individual climate variables identified that the best predictors of
-carbon fluxes were variables related to temperature. We therefore
-decided to include mean annual temperature as a term in all multivariate
-models. We first modelled the interaction effect between mean annual
-temperature and mean annual precipitation, in order to capture climate
-variation along the axes of temperature and water availability. Models
-were tested for a significant interactive effect and a significant
-additive effect. We then explored whether any other climate variable, in
-combination with mean annual temperature, could significantly improve on
-the combination of mean annual temperature and mean annual
-precipitation. In specifying the range of models to test, climate
-variables which were strongly correlated with temperature were dropped,
-in order to capture the greatest range of variation in climate. For each
-possible pairing of climate variables, two models were specified: a
-model with the two climate variables showing an additive effect; and a
-model with the two climate variables showing an interactive effect. As
-before, plot nested within geographic area was included as a random
-effect. Altitude was not considered. AIC values were calculated for the
-models, and used to compare models. Models were considered to be
-significantly better than the baseline MAT*MAP model if:\\
-i) the AIC value of the model was smaller than the AIC value of the
-baseline model by \textgreater{}2\\
-ii) the r-squared value of the model was larger than the r-squared value
-of the baseline model by \textbf{\textgreater{}5 (???- not possible)}
-
-\emph{Validating models of component fluxes.} Comparison of component
-fluxes is based on the assumption that components sum accurately to
-estimates of larger fluxes. To test this, components of larger fluxes
-were regressed against latitude, and the models used to generate a
-series of point estimates along lines of best fit for each component.
-The point estimates for smaller component fluxes were summed to generate
-new ``stacked'' estimates of larger fluxes, which were then compared
-against actual measurements of the larger flux. Confidence intervals for
-the larger flux were calculated using the `bootMer' function from the
-lme4 package \citep{bates_fitting_2015}. Stacked plots were generated
-for:\\
-1. GPP = NPP + R\textsubscript{auto}\\
-2. NPP = ANPP + BNPP\\
-3. ANPP = ANPP\textsubscript{foliage} + ANPP\textsubscript{woody}
-\textsubscript{stem}\\
-4. Total belowground carbon flux = BNPP + R\textsubscript{root}
-
-\emph{Allocation to carbon fluxes along latitudinal gradients.}
+\emph{Climate data}
+
+ForC contains geographic coordinates associated with each measurement
+record and, when available, mean annual temperature (MAT) and mean
+annual precipitation (MAP) as reported in the primary literature
+\citep{anderson-teixeira_forc_2018}. Based on the geographic
+co-ordinates for each site, data on twelve climate variables--including
+MAT, MAP, temperature and precipitation seasonality, annual temperature
+range, solar radiation, cloud cover, annual frost and wet days,
+potential evapotranspiration (PET), aridity (MAP/PET), and vapor
+pressure deficit (VPD)--were extracted from five open-access climate
+datasets: WorldClim \citep{hijmans_very_2005}, WorldClim2
+\citep{fick_worldclim_2017}, the Climate Research Unit (CRU) time-series
+dataset v. 4.03 \citep{harris_updated_2014}, the Global Aridity Index
+and Potential Evapotranspiration Climate Database
+\citep{trabucco_global_2019}, and TerraClimate
+\citep{abatzoglou_terraclimate_2018} (Table S1). From these data, we
+derived maximum VPD, defined as the VPD of the month with the largest
+deficit, and the number of water stress months, defined as the number of
+months annually where precipitation was lower than PET. Where site-level
+data was missing for MAT or MAP, we used values from the WorldClim
+dataset.
+
+Length of the growing season was estimated to the nearest month, where
+growing season months were defined as months with mean minimum
+temperature \textgreater{} 0.5\(^\circ\)C. We experimented with a
+definion of growing season months including a moisture index, defined as
+(MAT - PET)/PET, \textgreater{} -0.95 (Kerkhoff et al. 2005; see also
+Michaletz et al. 2014). However, we found that including a moisture
+index had \textbf{no} effect on the estimates of growing season length,
+and so chose to exclude it. (\textbf{Becky, was it really no effect? or
+minimal?}) Monthly data for PET, precipitation, and temperature from the
+CRU dataset v 4.03 \citep{harris_updated_2014}, and solar radiation from
+WorldClim2 \citep{fick_worldclim_2017} were used to calculate mean
+monthly PET, precipitation, temperature and solar radiation during the
+growing season. Total growing season precipitation and solar radiation
+were also calculated.
+
+\emph{Analyses}
+
+The effects of latitude and climate on FACF were analysed using mixed
+effects models using the package `lme4' \citep{bates_fitting_2015} in R
+v.3.5.1 \citep{r_core_team_r:_2018}. The basic model for all analyses
+included a fixed effect of latitude or climate and a random effect of
+plot nested within geographic area. Geographic areas--\emph{i.e.},
+spatially clustered sites--are defined within ForC using a hierarchical
+cluster analysis on the distance matrix of the sites and a cutoff of
+25km \citep{anderson-teixeira_forc_2018}. We experimented with inclusion
+of altitude as a fixed effect, but excluded it from the final models
+because it added very little explanatory power--that is, the difference
+in AIC (\(\Delta\)AIC) relative to models excluding altitude was
+generally small (often \(\Delta\)AIC\textless{}2). Hypotheses were
+accepted if the \(\Delta\)AIC between a model including the fixed effect
+of interest and a corresponding null model excluding that fixed effect
+exceeded 2.0. All \(R^2\) values presented here are marginal \(R^2\)
+values, and refer to the proportion of variation explained by only the
+fixed effects. Specific analyses are as described below.
+
+We first examined the relationship between latitude and FACF (Q1; Table
+1). We tested models with latitude as a linear term (corresponding null:
+model without latitude) and as a second-order polynomial term
+(corresponding null: model with latitude as a linear term), and
+calculated AIC values to determine the best model. Models were accepted
+as the best model if \(\Delta\)AIC \textgreater{} 2 with respect to the
+corresponding null, and were significant with respect to a null model
+with no fixed term. We also examined relationships among fluxes across
+latitude, testing whether sums of component fluxes matched the larger
+fluxes and whether C allocation varied with latitude, as specified
+below.
+
+To test whether trends in component fluxes across latitude sum to match
+those of larger fluxes, regression lines for smaller component fluxes
+were summed to generate new estimates of larger fluxes, which were then
+compared against the latitudinal regression of the larger flux.
+Confidence intervals for the larger flux were calculated using the
+`bootMer' function from the lme4 package \citep{bates_fitting_2015}.
+This analysis was applied to following sets of fluxes: (1)
+\(GPP = NPP + R_{auto}\), (2) \(NPP = ANPP + BNPP\), and (3)
+\(ANPP = ANPP_{foliage} + ANPP_{woody.stem}\). In addition, we estimated
+total belowground C flux (TBCF, not analyzed due to limited data) as
+\(TBCF = BNPP + R_{root}\).
+
 Variation in allocation to component carbon fluxes along latitudinal
-gradients was explored for a range of pairings: firstly, GPP:NPP,
-ANPP:BNPP, and ANPP\textsubscript{foliage}:ANPP\textsubscript{woody}
-\textsubscript{stem}; and secondly, the ratio of NPP to each of ANPP,
-BNPP, ANPP\textsubscript{foliage}, and ANPP\textsubscript{woody}
-\textsubscript{stem}. For each set of paired fluxes, measurements taken
-at the same site and plot, and in the same year, were paired together,
-and the ratio of each pair of measurements calculated. The ratios were
-regressed against latitude and climate variables, using the linear model
-specified above. Cook's distance analyses were carried out for each of
-the models, and indicated that data from a few high-elevation sites were
-having a disproportionate influence on the regressions. To account for
-this, models were re-run using only data from sites \(\le\) 1000m.
+gradients was explored for the following pairings: \(GPP:NPP\),
+\(ANPP:BNPP\), \(ANPP_{foliage}:ANPP_{woody.stem}\),
+\(ANPP_{foliage}:NPP\), \(ANPP_{woody.stem}:NPP\), \(ANPP:NPP\), and
+\(BNPP:NPP\). For each set of paired fluxes, measurements taken at the
+same site and plot, and in the same year, were paired together, and the
+ratio of each pair of measurements calculated. The ratios were regressed
+against latitude and climate variables, using the linear model specified
+above. Cook's distance analyses were carried out for each of the models,
+and indicated that data from a few high-elevation sites were having a
+disproportionate influence on the regressions. To account for this,
+models were re-run using only data from sites \(\le\) 1000m.
+
+We next examined the relationships of FACF to climate variables (Q2-Q4;
+Table 1). As with latitude, we tested both linear and polynomial fits
+for each climate variable. We tested relationships of each FACF (Table
+2) against each climate variable (Table S1), but focus
+presentation--including focal hypotheses (Table 1)--on climate variables
+that explained \textgreater{}20\% of variation in FACF. \textbf{Becky,
+please make this more specific. I drafted this based on the results, but
+its not very clear.}
+
+To investigate the potential joint and interactive effects of climate
+variables on carbon fluxes, multivariate models were also specified. We
+explored a large range of climate variable combinations and ultimately
+decided to focus on MAT and MAP. We favored these over variables with
+similar explanatory power because they are the most commonly reported
+climate variables and are recorded in ForC. MAT was consistently among
+the best individual predictors of FACF (Table S2), although it was
+occasionally out-performed by closely correlated climate variables (Fig.
+S2). In combination with MAT as a primary climate variable, MAP was
+consistently among the best secondary climate variables. Specifically,
+compared again a model including MAT x MAP, there were no other
+variables that, in interaction with MAT, improved AIC by
+\textgreater{}2.0 and \(R^2\) by \textgreater{}0.05. In examinging the
+joint and interactive effects of climate variables, an additive model
+including MAT and MAP as fixed effects was accepted when \(\Delta\)AIC
+\textgreater{}2 relative to a null including only MAT as a fixed effect.
+An interactive model including an MAT x MAP interaction was accepted
+when \(\Delta\)AIC \textgreater{}2 relative to a null including MAT and
+MAP as fixed effects.
+
+To test whether and how FACF varied with climate when standardised by
+growing season length (Q5), we first standardized FACF by dividing by
+growing season length (as defined above). We then tested for
+correlations between growing season length-standardised FACF against
+growing season climate variables. For analyses on data within the
+growing season, only linear models were specified.
+
+\textbf{All analyses were conducted in R (}Version\textbf{). Code and
+results are archived on GitHub\ldots{}.}
 
@@
 
-In total, we analyzed 1228 records from 9 C flux variables taken from
+In total, we analyzed 1228 records from nine FACF variables taken from
 forests that had experienced no major anthropogenic disturbances within
 the past 100 years. These records represented a total of 154 distinct
-geographic areas (Fig. 1, Table 2), across all forested biogeographic
-and climate zones.
-
-\emph{How does productivity vary with latitude?}
-
-All major carbon fluxes decreased linearly with latitude (Fig. 2).
-
-Latitude was a strong predictor for many of the carbon fluxes,
-explaining 64\% of variation in GPP (n = 254, p\textless{}0.0001), 50\%
-in NPP (n = 114, p\textless{}0.0001) and 45\% in ANPP (n = 259,
-p\textless{}0.0001). For all fluxes, their relationship with latitude
-was best predicted by the linear model.
-
-\emph{Relationships and differences among fluxes.} In general, smaller
-component fluxes summed approximately to larger fluxes across the
-latitudinal gradient (fig. 2). That is, modelled estimates of GPP,
-generated from the sum of NPP and R auto; NPP, generated from the sum of
-ANPP and BNPP\textsubscript{root}; and ANPP, generated from the sum of
-ANPP\textsubscript{foliage} and ANPP\textsubscript{woody}
-\textsubscript{stem}, fell completely within the confidence intervals of
-the regressions of field estimates of GPP, NPP and ANPP respectively.
-
-bbl: here or in the discussion, note that this is a fairly stringent
-test: it's easy for sub-fluxes not to sum up! (Extensive citations from
-EC literature for example.). HML: yes and no. confidence intervals are
-pretty large.
+geographic areas across all forested biogeographic and climate zones
+(Fig. 1, Table 2).
+
+\emph{How do FACF vary with latitude?}
+
+All major carbon fluxes decreased linearly with latitude (Fig. 2; Table
+S2). Latitude was a strong predictor for many of the carbon
+fluxes--particularly the larger fluxes (Table S2). Specifically,
+latitude explained 64\% of variation in GPP (n = 254,
+p\textless{}0.0001), 50\% in NPP (n = 114, p\textless{}0.0001) and 45\%
+in ANPP (n = 259, p\textless{}0.0001). The FACF that were most poorly
+predicted by latitude were \(BNPP_{root-fine}\) (\(R^2\)=0.17) and
+\(ANPP_{woody.stem}\) (\(R^2\)=0.18). For all FACF, the relationship
+with latitude was best predicted by the linear model).
 
@@
 95\% confidence intervals are plotted for the major flux for comparison with predicted trends.
-In (d),  which shows three belowground fluxes, the major flux, total belowground carbon flux, is one for which we have no data}\label{fig:unnamed-chunk-6}
+In (d),  which shows three belowground fluxes, the major flux, total belowground carbon flux, is one for which we have no data}\label{fig:unnamed-chunk-7}
 \end{figure}
 
-We found no evidence that allocation between fluxes varied substantially
-with latitude or climate. There were no significant results from
-regressing ratios of carbon fluxes against latitude, or against any of
-the climate variables.
-
-\(R^2\) values were generally highest in the major fluxes, and decreased
-in subsidiary fluxes (Supporting Information S2). Of the major fluxes,
-R\textsubscript{auto} and GPP were the most strongly explained by
-latitude and climate, with climate explaining at most 71\% of variation
-in GPP, and 65\% in R\textsubscript{auto}. The proportion of variation
-explained by climate and latitude decreased in NPP and ANPP, with
-climate explaining at most 51\% of variation in NPP and 44\% in ANPP. Of
-the major fluxes, BNPP\textsubscript{root} was the least well explained
-by climate and latitude, with climate explaining at most 36\% of
-variation.
-
-With the exception of ANPP\textsubscript{foliage}, the proportion of
-variation explained by climate and latitude in subsidiary fluxes was
-much lower. Climate explained at most 24\% of the variation in
-ANPP\textsubscript{woody} \textsubscript{stem}, 19\% in
-BNPP\textsubscript{fine} \textsubscript{root}, and 27\% in
-R\textsubscript{root}. In contrast, climate strongly explained variation
-in ANPP\textsubscript{foliage}, with mean annual temperature explaining
-58\% of variation. This pattern was also seen in the \(R^2\) values for
-multivariate models.
-
-\emph{How does productivity relate to MAT and MAP?} MAT and MAP are the
-most commonly reported site-level climate variables, and much previous
-research into the effect of climate on forest productivity has focused
-on these as key climate variables. MAT was a significant
-(p\textless{}0.05) and strong predictor of productivity for all carbon
-fluxes tested, with all fluxes showing a linear increase with
-temperature (fig. 4). We found no support for a saturation point of
-productivity with temperature.
-
-MAP was found to be a significant (p\textless{}0.05) but poor predictor
-of productivity, explaining, with the exception of
-R\textsubscript{auto}, at most 37\% of variation in carbon flux. For the
-majority of fluxes productivity was best predicted by a polynomial
-model. Productivity increased with precipitation, up until a saturation
-point at between 3000 and 4000mm annual precipitation, above which
-productivity started to decrease (fig. 4). The notable exception to this
-was GPP: the model indicated that GPP continued to increase with
-precipitation up to measures of at least 5000mm annually
-(p\textless{}0.0001, \(R^2\) = 0.33. Data above this point was not
-available, but the model trend indicated that the saturation point for
-this model would be around 5000mm MAP.
-
-There was a significant interactive effect between MAT and MAP for GPP,
-BNPP\textsubscript{root}, BNPP\textsubscript{fine} \textsubscript{root},
-ANPP, ANPP\textsubscript{woody} \textsubscript{stem}, and
-R\textsubscript{root} (fig.3). There was a significant additive effect
-for R\textsubscript{auto}. NPP and ANPP\textsubscript{foliage} showed no
-significant interactive or additive effect. \{\textbf{NEED SI TABLE
-GIVING FITTED COEFFICIENTS \& STATS}\}
-
-For the variables which showed a significant interactive or additive
-effect between MAT and MAP, no other climate variable, in combination
-with MAT, significantly improved on that model. For NPP, there was a
-significant interactive effect between MAT and water stress months, with
-this model explaining nearly 5\% more variation in NPP than MAT alone
-\{\textbf{Helene: but less than MATxMAP? if so, why bother
-mentioning?}\}. However, for ANPP\textsubscript{foliage}, no
-multivariate model improved on the univariate model including only MAT.
+In general, smaller component fluxes summed approximately to larger
+fluxes across the latitudinal gradient (Fig. 2). That is, modelled
+estimates of \(GPP\), generated from the sum of \(NPP\) and
+\(R_{auto}\); \(NPP\), generated from the sum of \(ANPP\) and \(BNPP\);
+and \(ANPP\), generated from the sum of \(ANPP_{foliage}\) and
+\(ANPP_{woody-stem}\), fell completely within the confidence intervals
+of the regressions of field estimates of \(GPP\), \(NPP\), and \(ANPP\),
+respectively.
+
+We found little evidence that allocation between fluxes varied
+substantially with latitude or climate (Fig. S3). Of the 7 FACF ratios
+regressed against latitude and three climate variables (MAT, MAP,
+temperature seasonality), there were only four signficant relationships,
+all with \(R^2 \le 0.2\) (Fig. S3). Specifically, the proportion of NPP
+allocated to \(ANPP_{foliage}\) increased weakly with MAT
+(\(R^2=0.20\)), and the proportion of \(NPP\) allocated aboveground
+(\(ANPP\)) decreased weakly with latitude (\(R^2=0.11\)) and temperature
+seasonality (\(R^2=0.17\)), while increasing with MAT (\(R^2=0.11\)).
+There were no significant results from regressing ratios of carbon
+fluxes against latitude, or against any of the climate variables.
+
+\emph{How do FACF relate to MAT and MAP?}
+
+All FACF increased linearly with MAT, and we found no support for a
+saturation point of FACF with MAT (all p\textless{}0.05; Figs. 3-4,
+S4-S5, Table S2). As with latitude, MAT tended to explain more variation
+in the larger FACF (\(GPP\), \(NPP\), \(ANPP\), \(R_{auto}\)) and
+\(ANPP_foliage\) (all \(R^2\)\textgreater{} 0.4) than in subsidiary and
+belowground fluxes (\(ANPP_{woody-stem}\), \(R_{root}\),
+\(BNPP_{root-fine}\); all \(R^2\)\textless{} 0.25).
+
+MAP was a significant (p\textless{}0.05) predictor of all FACF but
+\(ANPP_{woody-stem}\) (Figs. 4a, S4-S5; Table S2). However, it explained
+little variation: with the exception of \(R_{auto}\), MAP explained at
+most 37\% of variation in FACF. For the majority of FACF, a polynomial
+model was the best fit. FACF generally increased with precipitation, up
+until a saturation point at between 3000 and 4000mm annual
+precipitation, above which they started to decrease (Figs. 4, S4-S5).
+The notable exception to this was GPP: the model indicated that GPP
+continued to increase with precipitation up to measures of at least
+5000mm annually (p\textless{}0.0001, \(R^2\) = 0.33. Data above this
+point were not available, but the model trend indicated that the
+saturation point for this model would be around 5000mm MAP.
+
+There was a significant additive effect of MAT and MAP on \(GPP\),
+\(ANPP\) and \(R_{auto}\) (Fig. 3, Table S3). Accounting for MAT, MAP
+had a substantial positive effect on \(GPP\) and \(R_{auto}\) and a
+small negative effect on \(ANPP\). There was a significant interactive
+effect between MAT and MAP for \(NPP\) and \(ANPP_{woody-stem}\) (Fig.
+3, Table S3). The interaction was negative for \(NPP\) and positive for
+\(ANPP_{woody-stem}\). For \(ANPP_{foliage}\), \(BNPP\),
+\(BNPP_{root-fine}\), and \(R_{auto-root}\), MAP did not have a
+signficant effect when accounting for MAT (Fig. 3, Table S3). For the
+variables which showed a significant interactive or additive effect
+between MAT and MAP, no other climate variable, in combination with MAT,
+significantly improved on that model. \{\textbf{need to confirm this
+given changes in MAT MAP results (or you could just drop the
+sentence.)}\}
 
@@
 For visualization purposes, data points are grouped into bins of 0 - 1000, 1001 - 2000, 2001 - 3000, and >3000mm mean annual precipitation, and lines of best fit models are plotted for mean annual precipitation values of 500, 1500, 2500, and 3500mm.
-All regressions are significant $(p<0.05)$.}\label{fig:unnamed-chunk-7}
+All regressions are significant $(p<0.05)$.}\label{fig:unnamed-chunk-8}
 \end{figure}
 
-\emph{How does productivity relate to other climate variables?}
-\{\textbf{Here, maybe add some statement about covariation among climate
-variables, site SI figure}\} Our results indicated that productivity was
-most strongly explained by temperature at the global scale, with
-temperature-related climate variables coming out as strong predictors of
-productivity. In addition to MAT, temperature seasonality, annual
-temperature range, and annual frost days were consistently identified as
-strong univariate predictors of FACF. \{\textbf{cite some table/ SI?}\}
-
-We found a significant relationship between productivity and potential
-evapotranspiration for all fluxes. ANPP\textsubscript{foliage},
-BNPP\textsubscript{fine} \textsubscript{root} and R\textsubscript{root}
-increased linearly with PET; however, all other fluxes showed a
-polynomial relationship with PET (fig. 4). We found strong evidence for
-a saturation point or peak with PET: productivity tended to increase at
-values below 1000mm, before saturating between 1200 and 1700mm. There
-was also evidence that productivity begins to decrease at values above
-1800mm PET.
-
-Vapour pressure deficit was a significant predictor of productivity for
-all fluxes. BNPP\textsubscript{fine} \textsubscript{root} showed a
-linear relationship with vapour pressure deficit (\(R^2\) = 0.07,
-p\textless{}0.05), but all other fluxes showed a polynomial relationship
-(fig. 4). Productivity initially increased with vapour pressure deficit,
-before saturating at around 0.8 kPa. At values above 0.8 kPa,
-productivity began to decrease.
-
-All fluxes, with the exception of R\textsubscript{root}, showed a
-positive linear relationship with solar radiation. Solar radiation
-explained a low proportion of variability in productivity for all
-fluxes, explaining less than 20\% of the variation in each flux, with
-the exception of R\textsubscript{auto} (\(R^2\) = 0.26,
-p\textless{}0.05).
-
-Of the climate variables tested, annual wet days, aridity, cloud cover,
-mean diurnal temperature range, precipitation seasonality, maximum
-vapour pressure deficit and water stress months were poor or
-non-significant explainers of variation in productivity, explaining less
-than 20\% of the variation in each of the carbon fluxes.
+\emph{How do FACF relate to other climate variables?}
+
+Our results indicated that FACF were most strongly explained by
+temperature at the global scale, with temperature-related climate
+variables coming out as strong predictors of FACF. In addition to MAT,
+several of its correlates (Fig. S2) were consistently identified as
+strong univariate predictors of FACF: temperature seasonality, annual
+temperature range, annual frost days, PET, and length of growing season
+(Figs. 4, S4-S7).
+
+We found a significant relationship between C flux and potential
+evapotranspiration for all FACF. \(ANPP_{foliage}\),
+\(BNPP_{root-fine}\) and \(R_{root}\) increased linearly with PET;
+however, all other fluxes showed a polynomial relationship with PET
+(Fig. 4c, S4-5; Table S2). We found strong evidence for a saturation
+point or peak with PET: FACF tended to increase at values below 1000mm,
+before saturating between 1200 and 1700mm. There was also evidence that
+FACF begin to decrease at values above 1800mm PET.
+
+Vapour pressure deficit was a significant predictor of C flux for all
+FACF. \(BNPP_{root-fine}\) showed a linear relationship with vapour
+pressure deficit (\(R^2\) = 0.07, p\textless{}0.05), but all other
+fluxes showed a polynomial relationship (Figs. 4d, S4-5; Table S2). FACF
+initially increased with vapour pressure deficit, before saturating at
+around 0.8 kPa, after wich point they began to decrease.
+
+All fluxes, with the exception of \(R_{root}\), showed a positive linear
+relationship with solar radiation (Figs. S4-S5, Table S2). Solar
+radiation explained a low proportion of variability in all FACF,
+explaining less than 20\% of the variation in each flux, with the
+exception of\(R_{auto}\) (\(R^2\) = 0.26, p\textless{}0.05).
+
+Annual wet days, cloud cover, and aridity were poor or non-significant
+explainers of variation in FACF, explaining less than 20\% of the
+variation in each of the carbon fluxes (Figs. S4-S5; Table S2).
 
@@
 Lines of best fit are plotted according to the best model selected during analysis (**see issue 47**).
-All regressions are significant $(p<0.05)$.}\label{fig:unnamed-chunk-8}
+All regressions are significant $(p<0.05)$.}\label{fig:unnamed-chunk-9}
 \end{figure}
 
-\emph{What is the role of seasonality in explaining FACF?} Temperature
-seasonality was a significant predictor of FACF. GPP, NPP, ANPP, and
-R\textsubscript{root} exhibited a polynomial relationship with
-seasonality (Fig. 4) \{\textbf{give statistic on signficiance}\}.
-ANPP\textsubscript{foliage}, ANPP\textsubscript{woody}
-\textsubscript{stem} and R\textsubscript{auto} decreased linearly with
-temperature seasonality. Temperature seasonality was strongly correlated
-with annual temperature range, and, as expected, all fluxes showed
-almost identical responses to it \{\textbf{meaning unclear}\}.
-Productivity was highest where temperature seasonality = 0, and at an
-annual temperature range of 15\(^\circ\)C or lower. In contrast, there
-was no significant effect of precipitation seasonality on FACF.
+\emph{What is the role of seasonality in explaining FACF?}
+
+Temperature seasonality was a significant predictor of FACF. \(GPP\),
+\(NPP\), \(ANPP\), and \(R_{root}\) exhibited a polynomial relationship
+with seasonality (all p\textless{}0.05; Figs. 4e, S6-7; Table S2).
+\(ANPP_{foliage}\), \(ANPP_{woody-stem}\) and \(R_{auto}\) decreased
+linearly with temperature seasonality (all p\textless{}0.05; Figs. 4e,
+S6-S7; Table S2). Temperature seasonality was strongly correlated with
+annual temperature range, which was likewise a similarly strong
+predictor of FACF (Table S2).FACF were highest where temperature
+seasonality = 0, and at an annual temperature range of 15\(^\circ\)C or
+lower.
+
+In contrast, there was no significant effect of precipitation
+seasonality on FACF, and both maximum vapour pressure deficit, and water
+stress months were poor or non-significant explainers of variation in
+FACF (Figs. S6-S7; Table S2).
 
 We found a significant relationship between length of growing season and
-FACF, with all fluxes showing a linear increase in productivity with
-length of growing season (Fig. 4). Length of growing season was a strong
-predictor of FACF, explaining 51\% of variation in GPP, 39\% of
+FACF, with all fluxes showing a linear increase with length of growing
+season (Figs. 4e, S6-S7; Table S2). Length of growing season was a
+strong predictor of FACF, explaining 51\% of variation in GPP, 39\% of
 variation in NPP, and 34\% of variation in ANPP, but it was a weaker
-predictor than MAT for all fluxes analysed.
-
-\emph{Within the growing season, how does FACF vary with climate?} When
-FACF was standardized by growing season length, correlations with
-climate variables were much weaker. Growing season temperature,
-precipitation, solar radiation, and PET each had a small effect on one
-or more of the nine FACF analyzed (Table 2; Fig. \textbf{S\#}).
-Speficifally, \(ANPP\) increased with growing season temperature
-(\(R^2\) = 0.10, p\textless{}0.001) and precipitation (\(R^2\) = 0.04,
-p\textless{}0.05). Similaryly, \(ANPP_{foliage}\) increased slightly
-with growing season temperature (\(R^2\) = 0.16, p\textless{}0.01) and
-precipitation (\(R^2\) = 0.09, p\textless{}0.05). Growing season solar
-radiation had a positive influence on \(GPP\) (\(R^2\) = 0.21,
-p\textless{}0.001), \(NPP\) (\(R^2\) = 0.21, p\textless{}0.001),
-\(BNPP\) (\(R^2\) = 0.16, p\textless{}0.001) and \(BNPP_{fine.root}\)
-(\(R^2\) = 0.12, p\textless{}0.01). Growing season PET had a positive
-influence on \(GPP\) (\(R^2\) = 0.15, p\textless{}0.01), \(NPP\)
-(\(R^2\) = 0.18, p\textless{}0.01), \(BNPP\) (\(R^2\) = 0.23,
-p\textless{}0.0001), \(BNPP_{fine.root}\) (\(R^2\) = 0.11,
-p\textless{}0.05), and \(ANPP_{woody-stem}\) (\(R^2\) = 0.06,
-p\textless{}0.05). \{\textbf{Becky, please verify/ edit the following:}
-There were no other signifcant correlations between growing season
-length-standardized FACF (9 variables in Table 2) and growing season
-climate (\textbf{which variables?})\}.
+predictor than MAT for all fluxes analysed (Table S4).
+
+\emph{Within the growing season, how do FACF vary with climate?}
+
+When FACF were standardized by growing season length, correlations with
+growing season climate--including temperature, precipitation, solar
+radiation, and PET--were generally weak (Figs. S8-S9). Speficifally,
+\(ANPP\) increased with growing season temperature (\(R^2\) = 0.10,
+p\textless{}0.001) and precipitation (\(R^2\) = 0.04, p\textless{}0.05).
+Similaryly, \(ANPP_{foliage}\) increased slightly with growing season
+temperature (\(R^2\) = 0.16, p\textless{}0.01) and precipitation
+(\(R^2\) = 0.09, p\textless{}0.05). Growing season solar radiation had a
+positive influence on \(GPP\) (\(R^2\) = 0.21, p\textless{}0.001),
+\(NPP\) (\(R^2\) = 0.21, p\textless{}0.001), \(BNPP\) (\(R^2\) = 0.16,
+p\textless{}0.001) and \(BNPP_{fine.root}\) (\(R^2\) = 0.12,
+p\textless{}0.01). Growing season PET had a positive influence on
+\(GPP\) (\(R^2\) = 0.15, p\textless{}0.01), \(NPP\) (\(R^2\) = 0.18,
+p\textless{}0.01), \(BNPP\) (\(R^2\) = 0.23, p\textless{}0.0001),
+\(BNPP_{fine.root}\) (\(R^2\) = 0.11, p\textless{}0.05), and
+\(ANPP_{woody-stem}\) (\(R^2\) = 0.06, p\textless{}0.05).
+\{\textbf{Becky, please verify/ edit the following:} There were no other
+signifcant correlations between growing season length-standardized FACF
+(9 variables in Table 2) and growing season climate (\textbf{which
+variables?})\}.
 
@@
 
-We used a large global database (ForC), containing an unprecedented
-amount of data representing all major forest types (Fig. 1) and the nine
-most significant forest autotrophic carbon flux (FACF) variables (Table
-2), to comprehensively explore how C cycling in mature, undisturbed
-forests varies with latitude and climate on a global scale. We test a
-suite of hypotheses, including many previously posed (Table 1),
-standardising for factors that have not always been controlled for
-(\emph{e.g.}, stand age, flux components measured) to gain a stronger
-understanding of climatic effects on FACF.
-
-bbl: Most of these two paragraphs (following) just restate the results.
-Honestly I would remove them enitrely, but at the very least, condense
-into a single paragraph.
-
-We show that, across all nine FACF variables analyzed, C cycling
-decreases linearly with latitude (\emph{H1.1}; Fig. 2)--a finding that
-confirms multiple previous studies (\textbf{REFS}) but contradicts the
-idea that productivity of temperate forests rivals that of tropical
-forests (\emph{H1.1.alt}; \textbf{REFS}). The FACF variables increase in
-proportion to one another (\emph{H1.2}), with no differences in
-allocation detectable at this global scale and with component fluxes
-summing appropriately to larger fluxes (Fig. 2), indicating no major,
-systematic omissions or overestimations of flux components. However, we
-did detect a tendency for greater variability among subsidiary C fluxes
-(e.g., \emph{\(ANPP_{woody}\)}, \emph{\(BNPP_{fine.root}\)}; Fig. 2;
-\textbf{some SI table?}).
-
-\textbf{Climate explains a significant proportion of variation in all C
-fluxes analyzed} (albiet less for subsidiary fluxes), with temperature
-variables the best predictors of FACF at this global scale (Figs. 3-4).
-While other climate variables are significant predictors of FACF
-(\emph{H3}), none of them in isolation improve on the explanatory power
-of temperature-related variables in general or MAT specifically (Fig.
-4). Water availability is an important factor in explaining FACF on a
-global scale: we find a positive influence of precipitation at low MAP,
-with saturation at higher levels of MAP (Fig. 4b) and a significant
-interaction between MAT and MAP for seven of the nine variables analyzed
-(Fig. 3). Temperature seasonality and growing season length are closely
-correlated with MAT and are strong predictors of FACF (\emph{H4}; Fig.
-4e-f), though growing season length doesn't improve upon MAT as a
-predictor. Within the growing season, the influence of climate on C
-cycling is smaller but still significant for a number of carbon fluxes
-(\emph{H5}; \textbf{some SI table?}). These findings clarify the big
-picture of how FACF varies with latitude and climate on a global scale.
+Our analysis of a large global database (ForC) reveals how autotrophic
+carbon fluxes in mature forests vary with latitude and climate on a
+global scale. We show that, across all nine FACF analyzed, C cycling
+decreases continually with latitude (\emph{H1.1}; Fig. 2)--a finding
+that confirms multiple previous studies but contradicts the idea that
+productivity of temperate forests rivals that of tropical forests
+\citep{huston_global_2009}. The FACF generally increase in proportion to
+one another (\emph{H1.2}), with few differences in allocation detectable
+at this global scale (Fig. S2) and with component fluxes summing
+appropriately to larger fluxes (Fig. 2), indicating no major, systematic
+omissions or overestimations of flux components. However, climate
+explained lower proportions of variability among subsidiary C fluxes
+(\emph{e.g.}, \(ANPP_{woody}\), \(BNPP_{fine.root}\), \(R_{auto-root}\);
+Fig. 2; Table S2). Latitudianal variation in FACF is primarily
+attributable to temperature-related variables (\emph{H3, H4}),
+particularly MAT (Figs. 3-4). Water availability is also influential,
+but generally of secondary importance across the range represented in
+our database (Figs. 3-4). Temperature seasonality and growing season
+length are closely correlated with MAT and are strong predictors of FACF
+(\emph{H4}; Figs. 4e-f, S2, S6-S7), though growing season length doesn't
+improve upon MAT as a predictor. Within the growing season, the
+influence of climate on C cycling is smaller but still significant for a
+number of FACF (\emph{H5}; Fig. S9; Table S4). These findings clarify
+the big picture of how FACF vary with latitude and climate on a global
+scale.
 
 Past studies have differed in their conclusions regarding the
-relationship between productivity and latitude or its correlates (Table
-1, \emph{H1}; \textbf{REFS})--quite possibly because of lack of
+relationship between FACF and latitude or its correlates (Table 1,
+\emph{H1}; \textbf{REFS})--quite possibly because of lack of
 standardization with respect to stand age and disturbance history. Our
-findings indicate that, among mature, undisturbed stands, FACF is
+findings indicate that, among mature, undisturbed stands, FACF are
 unambiguously highest in the tropical regions, and the relationship is
-approximately linear (Fig. 2). This contrasts with suggestions that
+approximately linear (Fig. 2). This contrasts with the suggestion that
 productivity of temperate forests is similar to that of tropical forests
-(\textbf{REFS}). Temperate forests tend to be younger than tropical
-forests (\textbf{REF}), so analyses comparing across latitudinal
-gradients without controlling for stand age risk confounding age with
-biome effects (\href{http://pure.iiasa.ac.at/id/eprint/15361/}{Poulter
-et al. 2018-DOI:10.1594/PANGAEA.889943}; Potapov et al. 2008. Mapping
-the World's Intact Forest Landscapes by Remote Sensing. Ecology and
-Society 13 (2), 51. ). In addition, because C allocation varies with
-stand age \citep{de_lucia_forest_2007} (\textbf{See Nobby's comment in
-manuscript-draft\_NK.pdf}), age differences may introduce systematic
-biases into analyses of FACF across latitude or global climatic
-gradients. For example, woody productivity tends to be higher in rapidly
-aggrading secondary stands than in old-growth forests, where
-proportionally more C is allocated to respiration (** Nobby AFM paper:
-Understanding the controls over forest carbon use efficiency on small
-spatial scales: Effects of forest disturbance and tree diversity**)
-{[}*purpose for respiration/ other compenents (\textbf{See Nobby's
-comment in manuscript-draft\_NK.pdf}){]}.
-
-We show that FACF are broadly consistent in their responses to climate
-drivers on the global scale (with the exception of some differences in
-MAT-MAP interactions; Fig. 3), with no major trends in C allocation
-among the variable pairs tested (Table 1, \emph{H1}; Fig. 2;
-\textbf{Some SI table}). Although variation in allocation has been
-observed along gradients of elevation \citep{moser_elevation_2011} and
-water availability \citep{newman_above-_2006}--along with non-climatic
-axes of stand age \citep{litton_carbon_2007}, nutrient availability
+\citep{huston_global_2009}. Compared to tropical forests, the temperate
+forest biome has experienced more widespread anthropogenic disturbance
+and has a larger fraction of secondary stands
+\citep{potapov_mapping_2008, poulter_global_2018}, so analyses comparing
+across latitudinal gradients without controlling for stand age risk
+confounding age with biome effects. In addition, because C allocation
+varies with stand age \citep{delucia_forest_2007} (\textbf{See Nobby's
+comment in manuscript-draft\_NK.pdf}), age differences may introduce
+systematic biases into analyses of FACF across latitude or global
+climatic gradients. For example, woody productivity tends to be higher
+in rapidly aggrading secondary stands than in old-growth forests, where
+proportionally more C is allocated to respiration
+\citep{kunert_understanding_2019}. {[}\emph{purpose for respiration/
+other compenents} (\textbf{See Nobby's comment in
+manuscript-draft\_NK.pdf}){]}.
+
+\textbf{This paragraph needs work--including re-assessment of results,
+interpretation, writing} We show that FACF are broadly consistent in
+their responses to climate drivers on the global scale (with the
+exception of some differences in MAT-MAP interactions; Fig. 3), with no
+pronounced trends in C allocation among the variable pairs tested (Figs.
+2, S3). Although variation in allocation has been observed along
+gradients of elevation \citep{moser_elevation_2011} and water
+availability \citep{newman_above-_2006}--along with non-climatic axes of
+stand age \citep{litton_carbon_2007}, nutrient availability
 \citep{litton_carbon_2007, gill_belowground_2016}, and forest structure
-\citep{taylor_greater_2019}--variation in relation to climate is not
+\citep{taylor_greater_2019}--little variation in relation to climate is
 apparent at the global scale within ForC, which contains the bulk of
@@
 currently insufficent to rule out the possibility that such trends
-exist.
-
-bbl: remove this sentence? Of particular interest and significance are
-the relationships amongst \(GPP\), net primary productivity (\(NPP\) and
-its components, particularly \(ANPP_{woody-stem}\)), and respiration
-(\(R_{auto}\) and components). There have been suggestions that tropical
-forests tend to have low carbon use efficiency (\(CUE\)=
+exist. \textbf{bbl: remove following sentence?} Of particular interest
+and significance are the relationships amongst \(GPP\), net primary
+productivity (\(NPP\) and its components, particularly
+\(ANPP_{woody-stem}\)), and respiration (\(R_{auto}\) and components).
+There have been suggestions that tropical forests tend to have low
+carbon use efficiency (\(CUE\)=
 \(NPP\)/\(GPP\)=(\(GPP\)-\(R_{auto}\)/\(GPP\)), which are based on
@@
 (mostly younger) extratropical forests
-\citetext{\citealp{de_lucia_forest_2007}; \citealp[\href{https://www.researchgate.net/publication/227837530_The_productivity_metabolism_and_carbon_cycle_of_tropical_forest_vegetation}{Malhi
-2012};][]{anderson-teixeira_carbon_2016}}, but our analysis suggests
-that these low values might more appropriately be attributed to the fact
-that these forests are old than to their tropical climate. Indeed,
-\(CUE\) is known to decline with forest age
-(\citep{de_lucia_forest_2007};
-\href{https://academic.oup.com/treephys/article-abstract/39/8/1473/5423351?redirectedFrom=fulltext}{Collalti
-\& Prentice 2019}) but appears to be roughly independent of \(GPP\)
-\citep{litton_carbon_2007}. Among our sites with relevant data, there is
-indication that CUE or \(ANPP_{woody-stem}\)/\(GPP\) increase with
-latitude (\textbf{some SI table}). Additional measurements with careful
-methodological standardization across a consistent set of mature forest
-sites will be necessary to determine whether any climate-driven
-gradients in allocation exist at the global scale.
+\citep{delucia_forest_2007, malhi_productivity_2012, anderson-teixeira_carbon_2016},
+but our analysis suggests that these low values might more appropriately
+be attributed to the fact that these forests are old than to their
+tropical climate. Indeed, \(CUE\) is known to decline with forest age
+\citep{delucia_forest_2007, collalti_is_2019}, but appears to be roughly
+independent of \(GPP\) \citep{litton_carbon_2007}. Among our sites with
+relevant data, there is indication that CUE or
+\(ANPP_{woody-stem}\)/\(GPP\) increase with latitude (Fig. S3).
+Additional measurements with careful methodological standardization
+across a consistent set of mature forest sites will be necessary to
+determine whether any climate-driven gradients in allocation exist at
+the global scale.
 
@@
 \textless{}27\% for \(BNPP_{fine.root}\),
-\(R_{auto-root}\),\(ANPP_{woody-stem}\)) (Fig. 2; \textbf{some SI
-table?}). There are two, non-exclusive, potential explanations for this.
-First, it may be that methodological variation is larger relative to
-flux magnitude for some of the subsidiary fluxes. Belowground fluxes in
-particular are difficult to quantify, and measurement methods for the
-belowground fluxes considered here may be measured through fundamentally
-different approaches (\emph{e.g.}, minirhizotrons, ingrowth cores, or
-sequential coring for \(BNPP_{root-fine}\); root exclusion, stable
-isotope tracking, or gas exchange of excised roots for
-\(R_{auto-root}\)), and sampling depth is variable and often
-insufficient to capture the full soil profile. \(ANPP_{woody-stem}\),
-which is also poorly explained by latitude or climate, is more
-straightforward to measure but is subject to variability introduced by
-differences such as biomass allometries applied and minimum plant size
-sampled (bbl: cite e.g.~Huntzinger?). However, methodological variation
-and uncertainty affect all of fluxes considered here--not necessarily
-any less than the aforementioned, and some of the larger fluxes that
-vary more strongly with respect to climate (\(ANPP\), \(NPP\)) are
-estimated by summing uncertain component fluxes. Second, differences
-among variables in the proportion of variation explained by climate may
-be attributable to more dicrect climatic control over \(GPP\) than
-subsidiary fluxes. That is, subsidiary fluxes may be shaped by climate
-both indirectly through its influence on \(GPP\) and respiration and
-directly through any climatic influence on C allocation, as well as many
-other local- and regional-scale factors (\textbf{REFS}).
+\(R_{auto-root}\),\(ANPP_{woody-stem}\)) (Fig. 2; Table S2). There are
+two, non-exclusive, potential explanations for this. First, it may be
+that methodological variation is larger relative to flux magnitude for
+some of the subsidiary fluxes. Belowground fluxes in particular are
+difficult to quantify, and measurement methods for the belowground
+fluxes considered here may be measured through fundamentally different
+approaches (\emph{e.g.}, minirhizotrons, ingrowth cores, or sequential
+coring for \(BNPP_{root-fine}\); root exclusion, stable isotope
+tracking, or gas exchange of excised roots for \(R_{auto-root}\)), and
+sampling depth is variable and often insufficient to capture the full
+soil profile. \(ANPP_{woody-stem}\), which is also poorly explained by
+latitude or climate, is more straightforward to measure but is subject
+to variability introduced by differences such as biomass allometries
+applied and minimum plant size sampled (\textbf{bbl: cite
+e.g.~Huntzinger?}). However, methodological variation and uncertainty
+affect all of fluxes considered here--not necessarily any less than the
+aforementioned, and some of the larger fluxes that vary more strongly
+with respect to climate (\(ANPP\), \(NPP\)) are estimated by summing
+uncertain component fluxes. Second, differences among variables in the
+proportion of variation explained by climate may be attributable to more
+dicrect climatic control over \(GPP\) than subsidiary fluxes. That is,
+subsidiary fluxes may be shaped by climate both indirectly through its
+influence on \(GPP\) and respiration and directly through any climatic
+influence on C allocation, as well as many other local- and
+regional-scale factors (\textbf{REFS}).
 
@@
 availability (Table 1, \emph{H2-H3}; Figs. 3-4). Because many climate
-variables co-vary across the latitudinal gradient, because climatic
-drivers affect forest carbon flux on much shorter time scales than can
-be captured by annual climate summary variables, and because both
-climatic conditions and C flux vary intra- and inter-annually around the
-long-term means, it is not appropriate to attempt to identify any one
-mean annual climate variable as a mechanistic driver of FACF. However,
-it remains informative to consider these relationships. We find that
-temperature-related climate variables (\(MAT\), temperature seasonality,
-\ldots{}\textbf{LIST}) explain the highest proportion of variability in
-productivity, and among these, \(MAT\) is generally the best
+variables co-vary across the latitudinal gradient (Fig. S2), because
+climatic drivers affect forest carbon flux on much shorter time scales
+than can be captured by annual climate summary variables, and because
+both climatic conditions and C flux vary intra- and inter-annually
+around the long-term means, it is not appropriate to attempt to identify
+any one mean annual climate variable as a mechanistic driver of FACF.
+However, it remains informative to consider these relationships. We find
+that temperature-related climate variables (\(MAT\), temperature
+seasonality, \ldots{}\textbf{LIST}) explain the highest proportion of
+variability in FACF, and among these, \(MAT\) is generally the best
 predictor--perhaps because site-specific MAT is recorded for the
 majority of sites in ForC, whereas other variables are extracted from
-global gridded data products (Table S1\textbf{?}). The effects of
-temperature are modified by moisture availability, with reduced FACF
-under hot and dry conditions (\emph{i.e.}, high PET, high deficit; Fig.
-4c-d) and sometimes under very high precipitaiton (Figs. 3, 4b).
-Negative effects of very high precipitation on FACF have been previously
-observed (\textbf{REFS}) and are attributable to nutrient and light
-limitations (\textbf{REFS}) Thus, although temperature and water
-availability jointly and interactively drive global-scale patterns of
-FACF.
-
-bbl: this seems really interesting and novel, consider expanding the
-following paragraph
-
-FACF is negatively correlated with temperature seasonality (Table 1,
-\emph{H4}; Fig. 4e), and productivity is shut down during cold- or dry-
-dormant seasons. To account for this, a number of analyses seeking to
+global gridded data products (Table S1). The effects of temperature are
+modified by moisture availability, with reduced FACF under hot and dry
+conditions (\emph{i.e.}, high PET, high deficit; Fig. 4c-d) and
+sometimes under very high precipitaiton (Figs. 3, 4b). Negative effects
+of very high precipitation on FACF have been observed previously
+(\textbf{REFS}) and are attributable to nutrient and light limitations
+(\textbf{REFS}). Thus, although temperature and water availability
+jointly and interactively drive global-scale patterns of FACF.
+
+FACF are negatively correlated with temperature seasonality (Table 1,
+\emph{H4}; Fig. 4e), and is minimal during cold- or dry- dormant
+seasons. To account for this, a number of analyses seeking to
 characterize global-scale effects of climate on productivity have
-examined the relationship of FACF per month of the growing season with
+examined the relationship of C flux per month of the growing season with
 growing season climatic conditions (Table 1, \emph{H5}; \textbf{REFS}).
@@
 advance statistical predictive power. Rather, mechanistic accounting for
-climatic effects on global productivity patterns requires models
-representing physiologically meaningful timescales (\emph{e.g.},
-\textbf{refs}).
-
-Our analysis clarifies how FACF varies (bbl: is FACF singular or plural?
-Be consistent) with latitude and climate on a global scale, with some
-important implications for how forest carbon cycling relates to climate
-and, by extension, how it is likely to respond to climatic warming. We
-find no support for non-linear trends in mature forest C cycling with
-respect to latitude or \(MAT\), and no distinct trends in C allocation
-across the global scale (Fig. 2). The implication is that under warmer
-conditions with similar moisture availability--and within the
-temperature range to which forest communities are adapated and
-acclimatized--higher temperatures result in a generalized acceleration
-of FACF, with no major shifts in C allocation among subsidiary fluxes.
-Of course, actual climatic changes will result in very different sets of
-conditions than represented across geographic gradients in climate, but
-our analysis clarifies how carbon cycles through contemporary forest
-ecosystems. As we enter a period of accelerating climatic change,
+climatic effects on global FACF patterns requires models representing
+physiologically meaningful timescales (\emph{e.g.}, \textbf{refs}).
+
+Our analysis clarifies how FACF vary with latitude and climate on a
+global scale, with some important implications for how forest carbon
+cycling relates to climate and, by extension, how it is likely to
+respond to climatic warming. Our findings show that higher temperatures
+with similar moisture availability result in a generalized acceleration
+of FACF (Figs. 2-3). This is consistent with observations of
+continental- to global-scale increases in \(GPP\)
+\citep{li_mapping_2019} and \(ANPP_{woody stem}\)
+\citep{brienen_long-term_2015, hubau_asynchronous_2020}, along with some
+C cycle components not considered here--tree mortality
+\citep{brienen_long-term_2015, mcdowell_drivers_2018}, soil respiration
+\citep{bond-lamberty_global_2010}, and heterotrophic soil respiration
+\citep{bond-lamberty_globally_2018}. However, increasing C flux rates
+are not universal \citep{rutishauser_testing_2020} (\textbf{MORE REFS}).
+This is likely because factors other than rising temperatures are at
+play, including changes to other aspects of cliamte, atmospheric
+pollution (CO\textsubscript{2}, SO\textsubscript{2},
+NO\textsubscript{x}), and local disturbances. Morevoer, forest ecosystem
+responses to climatic changes outside the temperature range to which
+forest communities are adapated and acclimatized will not necessarily
+parallel responses across across geographic gradients in climate.
+Nevertheless, as we enter a period of accelerating climatic change,
 understanding of the fundamental climatic controls on FACF sets a
-foundation for understanding patterns of change. \textbf{{[}work on
-this{]}}
+foundation for understanding patterns of change.
+
+\subsection{misc content for
+discussion}\label{misc-content-for-discussion}
+
+\begin{itemize}
+\item
+  the observed positive interaction between MAT and MAP for
+  \(ANPP_woody\) is consitent with the \citet{taylor_temperature_2017}
+  analysis showing such an interaction for \(ANPP\) in tropical forests.
+  Similar to their analysis, we find a cross-over point at
+  \textasciitilde{}20C. However, we don't find such an interaction for
+  \(ANPP\), and we show a contrasting negative interaction for \(NPP\).
+  Some of this is may be stochasitic/ driven by composition of the
+  dataset, and the interactions we observe are not internally
+  consistent.
+\item
+  consistent with \citep{hofhansl_new_2015} (\textbf{verify}) , we found
+  a slight tendency for warmer sites to have higher aboveground
+  allocation
+\end{itemize}
+
+-results are consistent with \emph{Muller-Landau et al., in review}
 

### 2020-05-04 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd
worked on discussion
#69

@@
 2005; see also Michaletz et al.
-2014).
+2014-**put in these refs**).
 However, we found that including a moisture index had **no** effect on the estimates of growing season length, and so chose to exclude it.
@@
 
-**This paragraph needs work--including re-assessment of results, interpretation, writing**
+**See [issue #70](https://github.com/forc-db/Global_Productivity/issues/70), and potentially rework paragraph. **
 We show that, with the exception of some differences in MAT-MAP interactions (Fig.
@@
 
-**there seems to be a difference between GPP and other fluxes at higher precips (Figs.
-3, 4b), but the only allocaiton ratio we consider is GPP:NPP.
-It would be interesting too consider some others.
-Perhaps precip-->soil fertility-->CUE (see Doughty et al.
-2018).
-I need to come back to this.**
-
-One interesting observation was that climate tends to explain more variation in the major fluxes ($GPP$, $NPP$, $R_{auto}$ - latitude $R^2\ge$ 48%) than in subsidiary fluxes (latitude $R^2$ <27% for $BNPP_{fine.root}$, $R_{auto-root}$,$ANPP_{woody-stem}$) (Fig.
+One interesting observation was that climate tends to explain more variation in the major fluxes ($GPP$, $NPP$, $R_{auto}$) than in subsidiary fluxes ($BNPP_{fine.root}$, $R_{auto-root}$, $ANPP_{woody-stem}$) (Fig.
 2; Table S2).
@@
 Belowground fluxes in particular are difficult to quantify, and measurement methods for the belowground fluxes considered here may be measured through fundamentally different approaches (*e.g.*, minirhizotrons, ingrowth cores, or sequential coring for $BNPP_{root-fine}$; root exclusion, stable isotope tracking, or gas exchange of excised roots for $R_{auto-root}$), and sampling depth is variable and often insufficient to capture the full soil profile.
-$ANPP_{woody-stem}$, which is also poorly explained by latitude or climate, is more straightforward to measure but is subject to variability introduced by differences such as biomass allometries applied and minimum plant size sampled (**bbl: cite e.g.
-Huntzinger?**).
+$ANPP_{woody-stem}$, which is also poorly explained by latitude or climate, is more straightforward to measure but is subject to variability introduced by differences such as biomass allometries applied and minimum plant size sampled (*Clark et al.
+2001 Ecological Applications*).
 However, methodological variation and uncertainty affect all of fluxes considered here--not necessarily any less than the aforementioned, and some of the larger fluxes that vary more strongly with respect to climate ($ANPP$, $NPP$) are estimated by summing uncertain component fluxes.
-Second, differences among variables in the proportion of variation explained by climate may be attributable to more dicrect climatic control over $GPP$ than subsidiary fluxes.
+Second, differences among variables in the proportion of variation explained by climate may be attributable to more direct climatic control over $GPP$ than subsidiary fluxes.
 That is, subsidiary fluxes may be shaped by climate both indirectly through its influence on $GPP$ and respiration and directly through any climatic influence on C allocation, as well as many other local- and regional-scale factors (**REFS**).
 
+**This paragraph needs work, but I'm not currently sure how to fix it.
+It reads more as results than discussion, and doesn't say anything really profound.
+Depending on what happens with [issue #70](https://github.com/forc-db/Global_Productivity/issues/70), we might somehow rework/rearrange this paragraph and the previous two.
+**
 The latitudinal gradient in FACF (Fig.
@@
 Negative effects of very high precipitation on FACF have been observed previously (**REFS**) and are attributable to nutrient and light limitations (**REFS**).
-Thus, although temperature and water availability jointly and interactively drive global-scale patterns of FACF.
+Thus, although temperature and water availability jointly and interactively drive global-scale patterns of FACF....
 
@@
 To account for this, a number of analyses seeking to characterize global-scale effects of climate on productivity have examined the relationship of C flux per month of the growing season with growing season climatic conditions (Table 1, *H5*; **REFS**).
-We found that the sort of simple metric needed to define growing season at a global scale was uncertain for temperature and problematic for moisture (**WORK ON THIS**).
+We found that the sort of simple metric needed to define growing season at a global scale [@kerkhoff_plant_2005;@michaletz_convergence_2014] was rough with respect to temperature and problematic with respect to moisture because it doesn't capture temporal lags between precipitation and plant water use caused by storage in soil or snow.
+(**probably better to get into this in the methoods**).
 A temperature-defined growing season length had stong positive correlation with FACF (Fig.
@@
 However, increasing C flux rates are not universal [@rutishauser_testing_2020] (**MORE REFS**).
-This is likely because factors other than rising temperatures are at play, including changes to other aspects of cliamte, atmospheric pollution (CO~2~, SO~2~, NO~x~), and local disturbances.
+This is likely because factors other than rising temperatures are at play, including changes to other aspects of climate, atmospheric pollution (CO~2~, SO~2~, NO~x~), and local disturbances.
 Morevoer, forest ecosystem responses to climatic changes outside the temperature range to which forest communities are adapated and acclimatized will not necessarily parallel responses across across geographic gradients in climate.

### 2020-05-04 manuscript/manuscript draft.Rmd
update abstract
#69

@@
 ###Abstract
-[**very rough start:**]
 Carbon fixation, allocation, and metabolism by trees set the basis for energy and material flows in forest ecosystems and define their interactions with Earth's changing climate. 
-Here, we drew upon # records from the Global Forest Carbon Database (ForC), representing all major forest types and the nine most significant forest autotrophic carbon flux (FACF) variables, to comprehensively explore how C cycling in mature, undisturbed forests varies with latitude and climate on a global scale.
+Yet, we lack a cohesive synthesis on how forest autotrophic carbon fluxes (FACF) vary globally with respect to climate and one another. 
+Here, we draw upon # records from the Global Forest Carbon Database (ForC), representing all major forest types and the nine most significant FACF, to comprehensively explore how C cycling in mature, undisturbed forests varies with latitude and climate on a global scale.
 We show that, across all FACF variables analyzed, C cycling decreases linearly with latitude -- a finding that confirms multiple previous studies but contradicts the idea that net primary productivity ($NPP$) of temperate forests rivals that of tropical forests.
-The FACF variables generally increase in proportion to one another, with few differences in allocation detectable at this global scale, but differed in that latitude explained a lower proportion of variation among subsidiary fluxes (in particular, woody aboveground $NPP$ and belowground $NPP$, $BNPP$).
-Climate explained a significant proportion (#-#%) of variation in all C fluxes analyzed (less for subsidiary fluxes), with temperature variables in general and mean annual temperature ($MAT$) in particular being the best predictors of FACF on this global scale.
-While other climate variables (*e.g.*, **XX**) displayed significant correlation with FACF, none of them had significantly better explanatory power than $MAT$.
-The effects of temperature were modified by moisture availability, with reduced FACF under hot and dry conditions and sometimes under very high precipitaiton (especially for $BNPP$).
-FACF declined with temperature seasonality, but growing season length doesn't improve upon MAT as a predictor.
-Within the growing season, the influence of climate on C cycling is smaller but still significant for a number of carbon fluxes.
+FACF generally displayed similar trends acrooss latitude and multiple climate variables, with few differences in allocation detectable at this global scale, but differed in that latitude explained a lower proportion of variation among subsidiary fluxes (in particular, woody aboveground $NPP$ and belowground $NPP$, $BNPP$).
+Climate explained a significant proportion (#-#%) of variation in all C fluxes analyzed, with temperature variables in general and mean annual temperature (MAT) in particular being the best predictors of FACF on this global scale.
+The effects of temperature were modified by moisture availability, with FACF reduced under hot and dry conditions and sometimes under very high precipitaiton. 
+FACF declined with temperature seasonality, but growing season length did not improve upon MAT as a predictor.  
+Within the growing season, the influence of climate on C cycling was small but significant for a number of FACF. 
 These findings clarify the big picture of how FACF varies with latitude and climate on a global scale. 

### 2020-05-04 manuscript/manuscript draft.Rmd
reviewing refs/ working on discusion
#69 , #43

@@
 3), FACF are broadly consistent in their responses to climate drivers on the global scale, with no pronounced trends in C allocation among the variable pairs tested (Figs.
-2, S3).
+2, S3) .
 Although variation in allocation has been observed along gradients of elevation [@moser_elevation_2011] and water availability [@newman_above-_2006]--along with non-climatic axes of stand age [@litton_carbon_2007], nutrient availability [@litton_carbon_2007; @gill_belowground_2016], and forest structure [@taylor_greater_2019]--little variation in relation to climate is apparent at the global scale.
@@
 Additional measurements with careful methodological standardization across a consistent set of mature forest sites will be necessary to determine whether any climate-driven gradients in allocation exist at the global scale.
+
+*reworking the above paragraph:*
+We show that FACF are broadly consistent in their responses to climate drivers on the global scale, with at most modest trends in C allocation among the variable pairs tested (Figs. 2, S3). 
+This parallels the observation that C allocation across multiple FACF varies little with respect to climate along a tropical elevational gradient [@malhi_variation_2017], 
+but contrasts with observed variation in allocation along gradients of elevation [@moser_elevation_2011] and water availability [@newman_above-_2006]
+- @gill_belowground_2016 ?
+- (others in list below?)
+Several trends in allocation--or lack thereof--warrant discussion.
+1. CUE
+2. We found a slight tendency for warmer sites to have higher aboveground allocation (Fig. S3). This is consistent with:
+- [@hofhansl_new_2015] (**verify**) 
+- Chen et al. 2019 (doi: 10.1111/ele.13193) - find lower partitioning belowground at higher productivity (fine root NPP scales against ANPP_foliage with a slope of ~ 0.8)
+- @gill_belowground_2016
+- @litton_carbon_2007 - (verify)
 
@@
 
-- consistent with [@hofhansl_new_2015] (**verify**) , we found a slight tendency for warmer sites to have higher aboveground allocation
-
 -results are consistent with *Muller-Landau et al., in review*

### 2020-05-05 manuscript/manuscript draft.Rmd
resolved issue #72
closes #72

@@
 2018: DOI: 10.1111/btp.12504**);[@delucia_forest_2007; @kunert_understanding_2019].
-By constraining our analysis to stands ≥ 100 years old with no record of major recent disturbance, we have clarified the shape of the relationship between FACF and latitude.
+By constraining our analysis to stands $\ge$ 100 years old with no record of major recent disturbance, we have clarified the shape of the relationship between FACF and latitude.
 

### 2020-05-05 scripts/ratio_regression_script.R
Update ratio_regression_script.R

@@ -101,8 +101,9 @@ fixed.variables <- c("lat", "mat", "map", "TempSeasonality")
 set1 <- c("GPP", "ANPP", "ANPP_foliage", "ANPP_foliage", "ANPP_woody_stem", "ANPP", "BNPP_root")
 set2 <- c("NPP", "BNPP_root", "ANPP_woody_stem", "NPP", "NPP", "NPP", "NPP")
 
-# set1 <- c("GPP", "GPP", "GPP", "GPP", "GPP", "GPP", "GPP")
-# set2 <- c("NPP", "ANPP", "ANPP_foliage", "ANPP_woody_stem", "BNPP_root", "BNPP_root_fine", "R_auto")
+
+# set1 <- c("NPP", "ANPP", "ANPP_foliage", "ANPP_woody_stem", "BNPP_root", "BNPP_root_fine", "R_auto")
+# set2 <- c("GPP", "GPP", "GPP", "GPP", "GPP", "GPP", "GPP")
 
 png(file = paste0("C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/tables_figures/ratio_grid_plots.png"), width = 1800, height = 3000, units = "px", res = 300)
 
@@ -154,9 +155,9 @@ for (i in seq(along = set1)){
         
         ##remove extreme outliers 
         
-        # outliers <- boxplot(df$ratio, plot=FALSE, range = 3)$out
-        # 
-        # if(length(outliers > 0)) df <- df[-which(df$ratio %in% outliers),]
+        outliers <- boxplot(df$ratio, plot=FALSE, range = 3)$out
+
+        if(length(outliers > 0)) df <- df[-which(df$ratio %in% outliers),]
         
         
         mod <-  lmer(ratio ~ 1 + (1|geographic.area/plot.name), data = df, REML = F)

### 2020-05-05 scripts/correlation_matrix_code.R
Update correlation_matrix_code.R

@@ -25,6 +25,7 @@ ForC_simplified$lat <- abs(ForC_simplified$lat)
 
 ForC_simplified <- ForC_simplified[, c("lat", "mat", "map", "PotentialEvapotranspiration", "VapourPressureDeficit", "TempSeasonality", "length_growing_season")]
 
+##two groups of variables, needs to be identical + in the same order
 variables1 <- c("lat", "mat", "map", "PotentialEvapotranspiration", "VapourPressureDeficit", "TempSeasonality", "length_growing_season")
 variables2 <- c("lat", "mat", "map", "PotentialEvapotranspiration", "VapourPressureDeficit", "TempSeasonality", "length_growing_season")
 
@@ -40,8 +41,11 @@ panel.number <- 1
 
 png(file = paste0("C:/Users/becky/Dropbox (Smithsonian)/Github/Global_Productivity/manuscript/tables_figures/climate_regressions.png"), width = 3500, height = 3000, units = "px", res = 300)
 
+##creates grid of plots, mfrow should equal number of variables
 par(mfrow = c(7,7), mar = c(3,3,0.5,0.5), oma = c(1,1,0,0))
 
+
+##loops through variables, pairs each variable with each other variable, but skips if they're the same
 for (i in seq(along = variables1)){
   for (j in seq(along = variables2)){
   print(i)

### 2020-05-05 manuscript/manuscript draft.Rmd
update allocation results
#73

@@
 S3).
-Of the 7 FACF ratios regressed against latitude and three climate variables (MAT, MAP, temperature seasonality), there were only four signficant relationships, all with $R^2 \le 0.2$ (Fig.
+Of the 7 FACF ratios regressed against latitude and three climate variables (MAT, MAP, temperature seasonality), there were only five signficant relationships, all with $R^2 \le 0.4$ (Fig.
 S3).
-Specifically, the proportion of NPP allocated to $ANPP_{foliage}$ increased weakly with MAT ($R^2=0.20$), and the proportion of $NPP$ allocated aboveground ($ANPP$) decreased weakly with latitude ($R^2=0.11$) and temperature seasonality ($R^2=0.17$), while increasing with MAT ($R^2=0.11$).
+Specifically, the proportion of NPP allocated to $ANPP_{foliage}$ decreased with latitude ($R^2=0.32$) and increased with MAT ($R^2=0.37$), and the proportion of $NPP$ allocated aboveground ($ANPP$) decreased weakly with latitude ($R^2=0.11$) and temperature seasonality ($R^2=0.17$), while increasing with MAT ($R^2=0.11$).
 There were no significant results from regressing ratios of carbon fluxes against latitude, or against any of the climate variables.
@@
 This parallels the observation that C allocation across multiple FACF varies little with respect to climate along a tropical elevational gradient [@malhi_variation_2017], 
-but contrasts with observed variation in allocation along gradients of elevation [@moser_elevation_2011] and water availability [@newman_above-_2006]
+but contrasts with observed variation in allocation along gradients of elevation [@moser_elevation_2011] 
 - @gill_belowground_2016 ?
@@
 - [@hofhansl_new_2015] (**verify**) 
+- @moser_elevation_2011 (tropical elevation gradient)
 - Chen et al. 2019 (doi: 10.1111/ele.13193) - find lower partitioning belowground at higher productivity (fine root NPP scales against ANPP_foliage with a slope of ~ 0.8)

### 2020-05-05 scripts/mat_map_interaction.R
Update mat_map_interaction.R

@@ -4,7 +4,7 @@
 rm(list = ls())
 
 # Set working directory as ForC main folder ####
-setwd("C:/Users/gyrcbm/Dropbox/ForC")
+setwd("C:/Users/becky/Dropbox (Smithsonian)/GitHub/ForC")
 
 # Load libaries ####
 library(lme4)
@@ -216,7 +216,7 @@ for (response.v in response.variables){
 
     for (i in seq(along = lower_bound)){
       map.in.bin <- df[df$map > lower_bound[[i]] & df$map < upper_bound[[i]], ]
-      if(first.plot) plot(mean ~ mat, data = map.in.bin, xlab = "", ylab = "", ylim = ylim, col = plasma(5)[i], xlim = xlim, xaxt = "n", las = 1, tck = 0.02, mgp = c(3, 0.5, 0))
+      if(first.plot) plot(mean ~ mat, data = map.in.bin, xlab = "", ylab = "", ylim = ylim, col = plasma(5)[i], xlim = xlim, las = 1, tck = 0.02, mgp = c(3, 0.5, 0))
       if(!first.plot) points(mean ~ mat, data = map.in.bin, xlab = "", ylab = "", xaxt = "n", yaxt = "n", ylim = ylim, col = plasma(5)[i], xlim = xlim) 
       newDat <- expand.grid(mat = seq(min(map.in.bin$mat), max(map.in.bin$mat), length.out = 100), map = midpoint[[i]])
       newDat$fit <- predict(mod.full, newDat, re.form = NA)
@@ -283,16 +283,16 @@ for (response.v in response.variables){
   
 }
 
-# dev.off()
+dev.off()
 write.csv(all.results, "C:/Users/gyrcbm/Dropbox/Global_Productivity/results/tables/best_model_outputs/mat_map_interaction.csv")
 
 ########################################## code to correctly plot
 Rsqs = NULL
 
 response.variables <- c("GPP", "NPP", "ANPP", "ANPP_woody_stem", "ANPP_foliage", "BNPP_root", "BNPP_root_fine", "R_auto", "R_auto_root")
-png(file = paste0("C:/Users/gyrcbm/Dropbox/Global_Productivity/results/figures/final_figures/interactions/mat_map_interaction.png"), width = 2255, height = 2000, units = "px", res = 300)
[…]

### 2020-05-05 manuscript/manuscript draft.Rmd
knit in fixed figure, worked on allocation
#69 , #71

@@
 
-We found little evidence that allocation between fluxes varied substantially with latitude or climate (Fig.
+There was little evidence of substantial variation in C allocation across latitude or climate (Fig.
 S3).
@@
 Specifically, the proportion of NPP allocated to $ANPP_{foliage}$ decreased with latitude ($R^2=0.32$) and increased with MAT ($R^2=0.37$), and the proportion of $NPP$ allocated aboveground ($ANPP$) decreased weakly with latitude ($R^2=0.11$) and temperature seasonality ($R^2=0.17$), while increasing with MAT ($R^2=0.11$).
-There were no significant results from regressing ratios of carbon fluxes against latitude, or against any of the climate variables.
+The remaining 23 relationships examined were non-significant.
 
@@
 We show that FACF are broadly consistent in their responses to climate drivers on the global scale, with at most modest trends in C allocation among the variable pairs tested (Figs. 2, S3). 
-This parallels the observation that C allocation across multiple FACF varies little with respect to climate along a tropical elevational gradient [@malhi_variation_2017], 
-but contrasts with observed variation in allocation along gradients of elevation [@moser_elevation_2011] 
-- @gill_belowground_2016 ?
-- (others in list below?)
-Several trends in allocation--or lack thereof--warrant discussion.
-1. CUE
-2. We found a slight tendency for warmer sites to have higher aboveground allocation (Fig. S3). This is consistent with:
+This parallels the observation that C allocation across multiple FACF varies little with respect to climate along a steep tropical elevational gradient [@malhi_variation_2017; but see @moser_elevation_2011],
+and is not surprising given that trees face some universal constraints in carbon allocation (**REFS**). 
+The one trend in allocation that emerged from our analysis was a tendency for greater C allocation aboveground--and to foliage in particular--in warm tropical climates (Fig.
+S3).
+This is consistent with observations of increasing allocation to roots with declining temperature across a tropical elevational gradient [@moser_elevation_201],
 - [@hofhansl_new_2015] (**verify**) 
-- @moser_elevation_2011 (tropical elevation gradient)
 - Chen et al. 2019 (doi: 10.1111/ele.13193) - find lower partitioning belowground at higher productivity (fine root NPP scales against ANPP_foliage with a slope of ~ 0.8)
@@
 - @litton_carbon_2007 - (verify)
+One previously hypothesized trend that was not supported by our analysis was that tropical forests tend to have low carbon use efficiency ($CUE$= $NPP$/$GPP$=($GPP$-$R_{auto}$/$GPP$), which are based on observations of low $CUE$ in old-growth tropical forests relative to (mostly younger) extratropical forests [@delucia_forest_2007; @malhi_productivity_2012; @anderson-teixeira_carbon_2016].
+Limiting the analysis to mature forests, our analysis shows no such trend (Fig.
+S3), suggesting that the relatively low $CUE$ of intact tropical forests is more appropriately be attributed to the fact that these forests are old than to their tropical climate.
+Indeed, $CUE$ is known to decline with forest age [@delucia_forest_2007; @collalti_is_2019], but appears to be roughly independent of $GPP$ [@litton_carbon_2007].
 

### 2020-05-05 manuscript/appendix.Rmd
edits by krista merged

@@
 ```{r setup, include = FALSE}
-knitr::opts_knit$set(root.dir = 'C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/')
+# knitr::opts_knit$set(root.dir = 'C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/')
 # knitr::opts_knit$set(root.dir = 'C:/Users/gyrcbm/Dropbox/Global_Productivity/manuscript/')
- # knitr::opts_knit$set(root.dir = '/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/Global_Productivity/manuscript/')
+ knitr::opts_knit$set(root.dir = '/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/Global_Productivity/manuscript/')
 ```

### 2020-05-05 manuscript/manuscript draft.Rmd
edits by krista merged

@@
 
-**See [issue #70](https://github.com/forc-db/Global_Productivity/issues/70), and potentially rework paragraph. **
-We show that, with the exception of some differences in MAT-MAP interactions (Fig.
-3), FACF are broadly consistent in their responses to climate drivers on the global scale, with no pronounced trends in C allocation among the variable pairs tested (Figs.
-2, S3) .
-Although variation in allocation has been observed along gradients of elevation [@moser_elevation_2011] and water availability [@newman_above-_2006]--along with non-climatic axes of stand age [@litton_carbon_2007], nutrient availability [@litton_carbon_2007; @gill_belowground_2016], and forest structure [@taylor_greater_2019]--little variation in relation to climate is apparent at the global scale.
-Our conclusion, then, is that hypothesized gradients in allocation along global climate gradients cannot currently be supported for mature forests, although data quantity and standardization is currently insufficent to rule out the possibility that such trends exist.
-There have been suggestions that tropical forests tend to have low carbon use efficiency ($CUE$= $NPP$/$GPP$=($GPP$-$R_{auto}$/$GPP$), which are based on observations of low $CUE$ in old-growth tropical forests relative to (mostly younger) extratropical forests [@delucia_forest_2007; @malhi_productivity_2012; @anderson-teixeira_carbon_2016], but our analysis suggests that these low values might more appropriately be attributed to the fact that these forests are old than to their tropical climate.
-Indeed, $CUE$ is known to decline with forest age [@delucia_forest_2007; @collalti_is_2019], but appears to be roughly independent of $GPP$ [@litton_carbon_2007].
-Among our sites with relevant data, there are no signficant trends in allocation of $GPP$ to $NPP$ or of $NPP$ to $ANPP_{woody-stem}$ across gradients of latitude or climate (Fig.
-S3).
-Additional measurements with careful methodological standardization across a consistent set of mature forest sites will be necessary to determine whether any climate-driven gradients in allocation exist at the global scale.
-
-*reworking the above paragraph:*
 We show that FACF are broadly consistent in their responses to climate drivers on the global scale, with at most modest trends in C allocation among the variable pairs tested (Figs. 2, S3). 
@@
 This is consistent with observations of increasing allocation to roots with declining temperature across a tropical elevational gradient [@moser_elevation_201],
+
 - [@hofhansl_new_2015] (**verify**) 
@@
 - @gill_belowground_2016
-- @litton_carbon_2007 - (verify)
+- @litton_carbon_2007 ?
+
 One previously hypothesized trend that was not supported by our analysis was that tropical forests tend to have low carbon use efficiency ($CUE$= $NPP$/$GPP$=($GPP$-$R_{auto}$/$GPP$), which are based on observations of low $CUE$ in old-growth tropical forests relative to (mostly younger) extratropical forests [@delucia_forest_2007; @malhi_productivity_2012; @anderson-teixeira_carbon_2016].
@@
 Indeed, $CUE$ is known to decline with forest age [@delucia_forest_2007; @collalti_is_2019], but appears to be roughly independent of $GPP$ [@litton_carbon_2007].
+(*This finding may have some important implications for modeling and our broader understanding.
+I'm not sure offhand how much theory has been built on the idea of low CUE in tropical forests because of warm temperatures...*)
 

### 2020-05-05 manuscript/manuscript draft.Rmd
worked on discussion
#69

@@
 This is consistent with observations of increasing allocation to roots with declining temperature across a tropical elevational gradient [@moser_elevation_201],
-
-- [@hofhansl_new_2015] (**verify**) 
-- Chen et al. 2019 (doi: 10.1111/ele.13193) - find lower partitioning belowground at higher productivity (fine root NPP scales against ANPP_foliage with a slope of ~ 0.8)
-- @gill_belowground_2016
-- @litton_carbon_2007 ?
-
+and with observations and theory predicting relatively higher belowground carbon allocation at higher latitudes [@gill_belowground_2016].
+It is also consistent with findings that as forest productivity increases, proportionaly more carbon is allocated to $ANPP_{foliage}$ relative to $BNPP_{root-fine}$ (*Chen et al.
+2019; doi: 10.1111/ele.13193*) or $ANPP_{woody}$ [@hofhansl_new_2015].
+**( check / comment on  @litton_carbon_2007 )**
 One previously hypothesized trend that was not supported by our analysis was that tropical forests tend to have low carbon use efficiency ($CUE$= $NPP$/$GPP$=($GPP$-$R_{auto}$/$GPP$), which are based on observations of low $CUE$ in old-growth tropical forests relative to (mostly younger) extratropical forests [@delucia_forest_2007; @malhi_productivity_2012; @anderson-teixeira_carbon_2016].

### 2020-05-06 manuscript/manuscript draft.Rmd
worked on discussion, bringing in more refs
#69 , #43

@@
 Length of the growing season was estimated to the nearest month, where growing season months were defined as months with mean minimum temperature > 0.5$^\circ$C.
-We experimented with a definion of growing season months including a moisture index, defined as (MAT - PET)/PET, > -0.95 (Kerkhoff et al.
-2005; see also Michaletz et al.
-2014-**put in these refs**).
+We experimented with a definion of growing season months including a moisture index, defined as (MAT - PET)/PET, > -0.95 [@kerkhoff_plant_2005; see also @michaletz_convergence_2014].
 However, we found that including a moisture index had **no** effect on the estimates of growing season length, and so chose to exclude it.
@@
 
-**This paragraph needs work, but I'm not currently sure how to fix it.
-It reads more as results than discussion, and doesn't say anything really profound.
-Depending on what happens with [issue #70](https://github.com/forc-db/Global_Productivity/issues/70), we might somehow rework/rearrange this paragraph and the previous two.
-**
 The latitudinal gradient in FACF (Fig.
-2) is driven primarily by temperature-related climate variables, and secondarily by moisture availability (Table 1, *H2-H3*; Figs.
+2) is driven primarily by temperature-related climate variables, the effects of which are moderated by moisture availability (Table 1, *H2-H3*; Figs.
 3-4).
-Because many climate variables co-vary across the latitudinal gradient (Fig.
-S2), because climatic drivers affect forest carbon flux on much shorter time scales than can be captured by annual climate summary variables, and because both climatic conditions and C flux vary intra- and inter-annually around the long-term means, it is not appropriate to attempt to identify any one mean annual climate variable as a mechanistic driver of FACF.
+Indeed, MAT and MAP have long been identified as primary global-scale drivers of FACF [*Lieth 1973: doi 10.1007/BF01536729*; **REFS**; @taylor_temperature_2017].
+It is not appropriate to attempt to identify individual mean annual climate variables as mechanistic drivers of FACF because many climate variables co-vary across the latitudinal gradient (Fig.
+S2), because climatic drivers affect forest carbon flux on much shorter time scales than can be captured by annual climate summary variables, and because both climatic conditions and C flux vary intra- and inter-annually around the long-term means.
 However, it remains informative to consider these relationships.
-We find that temperature-related climate variables ($MAT$, temperature seasonality, ...**LIST**) explain the highest proportion of variability in FACF, and among these, $MAT$ is generally the best predictor--perhaps because site-specific MAT is recorded for the majority of sites in ForC, whereas other variables are extracted from global gridded data products (Table S1).
-The effects of temperature are modified by moisture availability, with reduced FACF under hot and dry conditions (*i.e.*, high PET, high deficit; Fig.
-4c-d) and sometimes under very high precipitaiton (Figs.
-3, 4b).
-Negative effects of very high precipitation on FACF have been observed previously (**REFS**) and are attributable to nutrient and light limitations (**REFS**).
-Thus, although temperature and water availability jointly and interactively drive global-scale patterns of FACF....
+Among the  temperature-related climate variables ($MAT$, temperature seasonality, annual temperature range, annual frost days, length of growing season, and $PET$; Table S1), $MAT$ is generally the most strongly correlated with FACF (Table S2)--perhaps in part because site-specific MAT is recorded for the majority of sites in ForC, whereas other variables were extracted from global gridded data products.
+This finding supports the continued focus on MAT as a primary--*albiet* not mechanistic--correlate of FACF.
+The effects of $MAT$ are modified by moisture availability, with reduced FACF under relatively dry conditions (*i.e.*, low precipitation; high vapour pressure deficit) and sometimes under very high precipitation (Figs.
+3-4).
+The observed positive interaction between MAT and MAP for $ANPP_{woody.stem}$ on the global scale (Fig.
+3) is consistent with an analysis showing a similar interaction for $ANPP$ in tropical forests, also with a cross-over point at ~20$\circ$C [@taylor_temperature_2017].
+However, we detect no such interaction for $ANPP$ or most other FACF, and we find a contrasting negative interaction for $NPP$ (Fig.
+3), suggesting that more data are required to sort out potential differences in the interactive effects of MAT and MAP on FACF in the tropics.
 
 FACF are negatively correlated with temperature seasonality (Table 1, *H4*; Fig.
-4e), and is minimal during cold- or dry- dormant seasons.
-To account for this, a number of analyses seeking to characterize global-scale effects of climate on productivity have examined the relationship of C flux per month of the growing season with growing season climatic conditions (Table 1, *H5*; **REFS**).
-We found that the sort of simple metric needed to define growing season at a global scale [@kerkhoff_plant_2005;@michaletz_convergence_2014] was rough with respect to temperature and problematic with respect to moisture because it doesn't capture temporal lags between precipitation and plant water use caused by storage in soil or snow.
-(**probably better to get into this in the methoods**).
+4e), and are minimal during cold- or dry- dormant seasons.
+To account for this, a number of analyses seeking to characterize global-scale effects of climate on productivity have examined the relationship of C flux per month of the growing season with growing season climatic conditions [Table 1, *H5*; @kerkhoff_plant_2005; *Anderson et al.
+2006: doi:10.1111/j.1461-0248.2006.00914.x *; *Enquist et al.
+2007: doi: 10.1111/j.1365-2486.2006.01222.x*;  @michaletz_convergence_2014].
+We found that the sort of simple metric needed to define growing season at a global scale [@kerkhoff_plant_2005] was rough with respect to temperature and problematic with respect to moisture because it doesn't capture temporal lags between precipitation and plant water use caused by storage in soil or snow.
 A temperature-defined growing season length had stong positive correlation with FACF (Fig.
 4f), but explained less variation than $MAT$.
-Dividing FACFs by growing season length to yield FACF per growing season month removed the majority of climate-related variation, supporting the idea that the latitudinal gradient in FACF is attributable more to shorter growing seasons at high latitudes than to inherently lower rates of photosyntheiss or respiration by high-latitude forests (*[ Enquist et al.
-2007 GCB- but check]*).
+Dividing FACFs by growing season length to yield FACF per growing season month removed the majority of climate-related variation, supporting the idea that the latitudinal gradient in FACF is attributable more to shorter growing seasons at high latitudes than to inherently lower rates of photosyntheiss or respiration by high-latitude forests (*Enquist et al.
+2007: doi: 10.1111/j.1365-2486.2006.01222.x*).
 However, there remained a number of significant correlations with growing season climatic conditions, suggesting that climatic conditions remain influential within the growing season.
@@
 This is consistent with observations of continental- to global-scale increases in $GPP$ [@li_mapping_2019] and $ANPP_{woody stem}$ [@brienen_long-term_2015; @hubau_asynchronous_2020], along with some C cycle components not considered here--tree mortality [@brienen_long-term_2015; @mcdowell_drivers_2018], soil respiration [@bond-lamberty_global_2010], and heterotrophic soil respiration [@bond-lamberty_globally_2018].
-However, increasing C flux rates are not universal [@rutishauser_testing_2020] (**MORE REFS**).
+However, increasing C flux rates by no means universal [@rutishauser_testing_2020] (**MORE REFS**).
 This is likely because factors other than rising temperatures are at play, including changes to other aspects of climate, atmospheric pollution (CO~2~, SO~2~, NO~x~), and local disturbances.
@@
 Nevertheless, as we enter a period of accelerating climatic change, understanding of the fundamental climatic controls on FACF sets a foundation for understanding patterns of change.
-
-## misc content for discussion
-- the observed positive interaction between MAT and MAP for $ANPP_woody$ is consitent with the @taylor_temperature_2017 analysis showing such an interaction for $ANPP$ in tropical forests.
-Similar to their analysis, we find a cross-over point at ~20C.
-However, we don't find such an interaction for $ANPP$, and we show a contrasting negative interaction for $NPP$.
-Some of this is may be stochasitic/ driven by composition of the dataset, and the interactions we observe are not internally consistent.
-
--results are consistent with *Muller-Landau et al., in review*
 

### 2020-05-06 manuscript/manuscript draft.Rmd
added ref

@@
 Carbon cycling in forests worldwide provides the energetic basis for sustaining the majority of Earth's terrestrial biodiversity and many human populations (**REF**), while strongly influencing atmospheric CO~2~ and climate [@bonan_forests_2008].
-Forests' autotrophic carbon fluxes (FACF)--that is, carbon fixation, allocation, and metabolism by trees and other primary producers--sets the energy ultimately available to heterotrophic organisms (including microbes), in turn influencing their abundance (**REFS**) and possibly diversity [@waide_relationship_1999] (**REFS**).
+Forests' autotrophic carbon fluxes (FACF)--that is, carbon fixation, allocation, and metabolism by trees and other primary producers--sets the energy ultimately available to heterotrophic organisms (including microbes), in turn influencing their abundance (**REFS**) and possibly diversity [@waide_relationship_1999; Chu et al.
+2019 (DOI: 10.1111/ele.13175)].
 FACF influences all organic matter stocks in forest ecosystems and is linked to cycling of energy, water, and nutrients (**REFS**) [@piao_forest_2010].

### 2020-05-10 manuscript/manuscript draft.Rmd
references

@@
 Carbon cycling in forests worldwide provides the energetic basis for sustaining the majority of Earth's terrestrial biodiversity and many human populations (**REF**), while strongly influencing atmospheric CO~2~ and climate [@bonan_forests_2008].
-Forests' autotrophic carbon fluxes (FACF)--that is, carbon fixation, allocation, and metabolism by trees and other primary producers--sets the energy ultimately available to heterotrophic organisms (including microbes), in turn influencing their abundance (**REFS**) and possibly diversity [@waide_relationship_1999; Chu et al.
+Forests' autotrophic carbon fluxes (FACF)--that is, carbon fixation, allocation, and metabolism by trees and other primary producers--sets the energy ultimately available to heterotrophic organisms (including microbes), in turn influencing their abundance (e.g., [Zak et al.
+1994](http://esajournals.onlinelibrary.wiley.com/doi/abs/10.2307/1940888); [Niedziałkowska et al.
+2010](https://doi.org/10.2980/17-1-3310)) and possibly diversity [@waide_relationship_1999; Chu et al.
 2019 (DOI: 10.1111/ele.13175)].
-FACF influences all organic matter stocks in forest ecosystems and is linked to cycling of energy, water, and nutrients (**REFS**) [@piao_forest_2010].
+FACF influences all organic matter stocks in forest ecosystems and is linked to cycling of energy, water, and nutrients.
 Critically, FACF also define forest interactions with Earth's changing climate.
@@
 S3), suggesting that the relatively low $CUE$ of intact tropical forests is more appropriately be attributed to the fact that these forests are old than to their tropical climate.
-Indeed, $CUE$ is known to decline with forest age [@delucia_forest_2007; @collalti_is_2019], but appears to be roughly independent of $GPP$ [@litton_carbon_2007].
+Indeed, $CUE$ is known to decline with forest age [@delucia_forest_2007; @piao_forest_2010; @collalti_is_2019], but appears to be roughly independent of $GPP$ [@litton_carbon_2007].
 (*This finding may have some important implications for modeling and our broader understanding.

### 2020-05-11 manuscript/manuscript draft.Rmd
refs

@@
 
-1. Conservation Ecology Center; Smithsonian Conservation Biology Institute; National Zoological Park, Front Royal, VA, USA
+1. Conservation Ecology Center; Smithsonian Conservation Biology Institute; Front Royal, VA, USA
 
@@
 FACF decrease with latitude, but it remains unclear whether and how the shape of this relationship varies across fluxes.
-Studies agree that FACF are lowest in the boreal regions, and increase into the temperate regions [@luyssaert_co_2007; @huston_global_2009; @beer_terrestrial_2010; @jung_global_2011].
+Studies agree that FACF are lowest in the boreal regions, and increase into the temperate regions [@luyssaert_co_2007; @huston_global_2009; @beer_terrestrial_2010; @piao_forest_2010; @jung_global_2011].
 However, evidence is inconclusive on whether primary productivity continues to increase into the tropics, or whether it plateaus in temperate regions.
@@
 hypothesis_table <- read.csv("hypothesis_table.csv", stringsAsFactors = FALSE, check.names = FALSE)
-hypothesis_table[1,1] <- paste0(hypothesis_table[1,1], footnote_marker_number(1),"$^{,}$",footnote_marker_number(2),"$^{,}$", footnote_marker_number(3))
-hypothesis_table[2,1] <- paste0(hypothesis_table[2,1], footnote_marker_number(1),"$^{,}$",footnote_marker_number(4))
+hypothesis_table[1,1] <- paste0(hypothesis_table[1,1], footnote_marker_number(1),"$^{,}$",footnote_marker_number(2),"$^{,}$", footnote_marker_number(3),"$^{,}$", footnote_marker_number(10))
+hypothesis_table[2,1] <- paste0(hypothesis_table[2,1], footnote_marker_number(1),"$^{,}$",footnote_marker_number(4),"$^{,}$",footnote_marker_number(9))
 hypothesis_table[3,1] <- paste0(hypothesis_table[3,1], footnote_marker_number(1),"$^{,}$",footnote_marker_number(4))
@@
 (2014)", "Chu et al.
-(2016)"), footnote_as_chunk = TRUE)
+(2016)", "Piao et al.
+(2010)", "Huston & Wolverton (2009)"), footnote_as_chunk = TRUE)
 ```
@@
 For example, woody productivity tends to be higher in rapidly aggrading secondary stands than in old-growth forests, where proportionally more C is allocated to respiration (**Doughty et al.
-2018: DOI: 10.1111/btp.12504**);[@delucia_forest_2007; @kunert_understanding_2019].
+2018: DOI: 10.1111/btp.12504**);[@delucia_forest_2007; @piao_forest_2010; @kunert_understanding_2019].
 By constraining our analysis to stands $\ge$ 100 years old with no record of major recent disturbance, we have clarified the shape of the relationship between FACF and latitude.

### 2020-05-11 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@
 Carbon fixation, allocation, and metabolism by trees set the basis for energy and material flows in forest ecosystems and define their interactions with Earth's changing climate. 
-Yet, we lack a cohesive synthesis on how forest autotrophic carbon fluxes (FACF) vary globally with respect to climate and one another. 
+However, we lack a cohesive synthesis on how forest autotrophic carbon fluxes (FACF) vary globally with respect to climate and one another. 
 Here, we draw upon # records from the Global Forest Carbon Database (ForC), representing all major forest types and the nine most significant FACF, to comprehensively explore how C cycling in mature, undisturbed forests varies with latitude and climate on a global scale.
 We show that, across all FACF variables analyzed, C cycling decreases linearly with latitude -- a finding that confirms multiple previous studies but contradicts the idea that net primary productivity ($NPP$) of temperate forests rivals that of tropical forests.
-FACF generally displayed similar trends acrooss latitude and multiple climate variables, with few differences in allocation detectable at this global scale, but differed in that latitude explained a lower proportion of variation among subsidiary fluxes (in particular, woody aboveground $NPP$ and belowground $NPP$, $BNPP$).
+FACF generally displayed similar trends across latitude and multiple climate variables, with few differences in allocation detectable at this global scale, but differed in that latitude explained a lower proportion of variation among subsidiary fluxes (in particular, woody aboveground $NPP$ and belowground $NPP$, $BNPP$).
+**BBL: not sure the second part of this sentence merits inclusion in the abstract**
 Climate explained a significant proportion (#-#%) of variation in all C fluxes analyzed, with temperature variables in general and mean annual temperature (MAT) in particular being the best predictors of FACF on this global scale.
-The effects of temperature were modified by moisture availability, with FACF reduced under hot and dry conditions and sometimes under very high precipitaiton. 
+The effects of temperature were modified by moisture availability, with FACF reduced under hot and dry conditions and sometimes under very high precipitation. 
 FACF declined with temperature seasonality, but growing season length did not improve upon MAT as a predictor.  
 Within the growing season, the influence of climate on C cycling was small but significant for a number of FACF. 
-These findings clarify the big picture of how FACF varies with latitude and climate on a global scale. 
+These findings clarify how FACF varies with latitude and climate on a global scale. 
 As we enter a period of accelerating climatic change, understanding of the fundamental climatic controls on FACF sets a foundation for understanding patterns of change.
@@
 FACF influences all organic matter stocks in forest ecosystems and is linked to cycling of energy, water, and nutrients.
+
 Critically, FACF also define forest interactions with Earth's changing climate.
@@
 
-FACF decrease with latitude, but it remains unclear whether and how the shape of this relationship varies across fluxes.
-Studies agree that FACF are lowest in the boreal regions, and increase into the temperate regions [@luyssaert_co_2007; @huston_global_2009; @beer_terrestrial_2010; @piao_forest_2010; @jung_global_2011].
-However, evidence is inconclusive on whether primary productivity continues to increase into the tropics, or whether it plateaus in temperate regions.
+FACF decrease with latitude, but it remains unclear whether and how the shape of this relationship varies between fluxes.
+FACF are lowest in the boreal regions, and increase into the temperate regions [@luyssaert_co_2007; @huston_global_2009; @beer_terrestrial_2010; @piao_forest_2010; @jung_global_2011].
+However, evidence is inconclusive on whether primary productivity continues to increase into the tropics.
 Evidence for this is further complicated by the fact that different studies use different measures of productivity to explore these relationships.
@@
 The majority of studies have focused on exploring the relationships of FACF to mean annual temperature (MAT) and mean annual precipitation (MAP), as the most commonly reported site-level climate variables.
-While these fail to capture some important aspects of climate such as seasonality, they do describe broad trends in temperature and water availability, and therefore capture a substantial portion of global-scale variation in climate.
 There is strong evidence that both MAT and MAP show significant positive relationships with FACF [@chu_does_2016].
@@
 [-@taylor_temperature_2017] showed a positive interaction between MAT and MAP in shaping tropical forest productivity, such that high rainfall had a negative effect on productivity in cooler climates, compared to a positive effect in warmer climates.
-{**It would be good to add some more citations on soil respiraiton.
-I'm sure BBL can help.**}
+Such complicated dynamics play out with belowground fluxes as well, with typically dominant soil temperature mediated by moisture and carbon supply (Hursh et al.
+2017, http://dx.doi.org/10.1111/gcb.13489; Xu et al.
+2016 http://dx.doi.org/10.1016/j.jplph.2016.08.007).
 
-FACF can be influenced by many other factors as well, which often act across a range of scales, and may show interactive effects with each other [@cleveland_relationships_2011].
-On a local scale, stand age [@litton_carbon_2007; @gillman_latitude_2015], biodiversity [@liang_positive_2016], management [@simova_enigma_2017], nutrient availability [@aragao_above_2009], and altitude [@girardin_net_2010; @malhi_variation_2017] all impact FACF.
-On a global scale, we expect that FACF are most strongly influenced by broad climatic gradients.
-There is evidence that FACFs also respond to variables such as cloud cover [@taylor_temperature_2017], solar radiation [@fyllas_solar_2017], and potential evapotranspiration [@kerkhoff_plant_2005] in potentially significant ways.
-Furthermore, MAT and MAP are very coarse measures of climate, and so fail to capture much variation in climate on an intra-annual scale, including the effects of factors such as growing season length, number of frost-free days, temperature seasonality, and dry season length.
+MAT and MAP are very coarse measures of climate, and so fail to capture much variation in climate on an intra-annual scale, including the effects of factors such as growing season length, number of frost-free days, temperature seasonality, and dry season length.
 Some studies have suggested that the apparently strong relationship between MAT and FACFs is actually a consequence of the correlation between MAT and growing season length [@kerkhoff_plant_2005; @malhi_productivity_2012; @michaletz_convergence_2014; @michaletz_drivers_2018].
@@
 
+In addition, FACF can be influenced by other, often interactive, factors across a range of scales [@cleveland_relationships_2011].
+On a local scale, stand age [@litton_carbon_2007; @gillman_latitude_2015], biodiversity [@liang_positive_2016], management [@simova_enigma_2017], nutrient availability [@aragao_above_2009], and altitude [@girardin_net_2010; @malhi_variation_2017] all impact FACF.
+There is evidence that FACFs also respond to variables such as cloud cover [@taylor_temperature_2017], solar radiation [@fyllas_solar_2017], and potential evapotranspiration [@kerkhoff_plant_2005].
+
 The recent development of the Global Forest Carbon database (ForC), which synthesizes multiple variables and including records of stand history [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018], opens up the possibility for a standardized analysis of global scale variation in multiple FACF and the principle climatic drivers of these patterns.
 In order to approach this broad topic, we simplify the major gaps in our knowledge to five broad questions and corresponding hypotheses (Table 1).
-First, we ask how FACF vary with latitude.
+First, we ask how nine FACF in ForC vary with latitude.
 We then test how these fluxes relate to MAT and MAP, and additionally how they respond to other, less well-studied, climate variables.
 Finally, we consider the relationship between FACF and seasonality, considering the role of seasonality in explaining variation in carbon fluxes, and the influence of climate on FACF standardized by growing season length.
-We address the above questions for nine carbon fluxes contained in ForC, allowing for an in-depth exploration of the effect of climate on FACF globally.
 
@@
 We tested models with latitude as a linear term (corresponding null: model without latitude) and as a second-order polynomial term (corresponding null: model with latitude as a linear term), and calculated AIC values to determine the best model.
-Models were accepted as the best model if $\Delta$AIC > 2 with respect to the corresponding null, and were significant with respect to a null model with no fixed term.
-We also examined relationships among fluxes across latitude, testing whether sums of component fluxes matched the larger fluxes and whether C allocation varied with latitude, as specified below.
+**BBL: careful about terminology here.
+In statistics a "linear model" includes polynomial models, which are linear in their parameters.
+Thus I (and many readers perhaps) find this confusing.** Models were accepted as the best model if $\Delta$AIC > 2 with respect to the corresponding null, and were significant with respect to a null model with no fixed term.
 
@@
 Confidence intervals for the larger flux were calculated using the 'bootMer' function from the lme4 package [@bates_fitting_2015].
-This analysis was applied to following sets of fluxes: (1) $GPP = NPP + R_{auto}$, (2) $NPP = ANPP + BNPP$, and (3) $ANPP = ANPP_{foliage} + ANPP_{woody.stem}$.
+This analysis was applied to the following sets of fluxes: (1) $GPP = NPP + R_{auto}$, (2) $NPP = ANPP + BNPP$, and (3) $ANPP = ANPP_{foliage} + ANPP_{woody.stem}$.
 In addition, we estimated total belowground C flux (TBCF, not analyzed due to limited data) as $TBCF = BNPP + R_{root}$.
@@
 **Becky, please make this more specific.
-I drafted this based on the results, but its not very clear.**
+I drafted this based on the results, but its not very clear.** **BBL: I agree, not clear and might trigger reviewer concerns about tailoring hypotheses to significance.**
 
-To investigate the potential joint and interactive effects of climate variables on carbon fluxes, multivariate models were also specified.
+Multivariate models were used to investigate the potential joint and interactive effects of climate variables on carbon fluxes.
 We explored a large range of climate variable combinations and ultimately decided to focus on MAT and MAP.
-We favored these over variables with similar explanatory power because they are the most commonly reported climate variables and are recorded in ForC.
+We favored these over variables with similar explanatory power because they are the most commonly reported climate variables.
 MAT was consistently among the best individual predictors of FACF (Table S2), although it was occasionally out-performed by closely correlated climate variables (Fig.
@@
 In combination with MAT as a primary climate variable, MAP was consistently among the best secondary climate variables.
-Specifically, compared again a model including MAT x MAP, there were no other variables that, in interaction with MAT, improved AIC by >2.0 and $R^2$ by >0.05.
-In examinging the joint and interactive effects of climate variables, an additive model including MAT and MAP as fixed effects was accepted when $\Delta$AIC >2 relative to a null including only MAT as a fixed effect.
+Specifically, compared against a model including MAT x MAP, there were no other variables that, in interaction with MAT, improved AIC by >2.0 and $R^2$ by >0.05.
+In examining the joint and interactive effects of climate variables, an additive model including MAT and MAP as fixed effects was accepted when $\Delta$AIC >2 relative to a null including only MAT as a fixed effect.
 An interactive model including an MAT x MAP interaction was accepted when $\Delta$AIC >2 relative to a null including MAT and MAP as fixed effects.
@@
 To test whether and how FACF varied with climate when standardised by growing season length (Q5), we first standardized FACF by dividing by growing season length (as defined above).
-We then tested for correlations between growing season length-standardised FACF against growing season climate variables.
-For analyses on data within the growing season, only linear models were specified.
+We then tested for correlations between these standardised fluxes and growing season climate variables, using only linear models.
 
-**All analyses were conducted in R (**Version**). Code and results are archived on GitHub....**
+**All analyses were conducted in R (**Version**). Code and data necessary to reproduce all results are archived on GitHub....**
 
@@
 The FACF that were most poorly predicted by latitude were $BNPP_{root-fine}$ ($R^2$=0.17) and $ANPP_{woody.stem}$ ($R^2$=0.18).
-For all FACF, the relationship with latitude was best predicted by the linear model.
+For all FACF, the relationship with latitude was best predicted by the first-order linear model.
 
@@
 
-MAP was a significant (p<0.05) predictor of all FACF but $ANPP_{woody-stem}$ (Figs.
+MAP was a significant (p<0.05) predictor of all FACF except $ANPP_{woody-stem}$ (Figs.
 4a, S4-S5; Table S2).
 However, it explained little variation: with the exception of $R_{auto}$, MAP explained at most 37% of variation in FACF.
-For the majority of FACF, a polynomial model was the best fit.
+For the majority of FACF, a polynomial model **BBL: of what order?** provided the best fit.
 FACF generally increased with precipitation, up until a saturation point at between 3000 and 4000mm annual precipitation, above which they started to decrease (Figs.
@@
 The notable exception to this was GPP: the model indicated that GPP continued to increase with precipitation up to measures of at least 5000mm annually (p<0.0001, $R^2$ = 0.33.
-Data above this point were not available, but the model trend indicated that the saturation point for this model would be around 5000mm MAP.
+Data above this point were not available, but the model trend suggested a saturation point around 5000mm MAP.
 
@@
 {**need to confirm this given changes in MAT MAP results (or you could just drop the sentence.)**}
+**BBL: this paragraph is pretty dense. Perhaps distill to a summary sentence and move rest to SI?**
 
@@
 FACF were highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower.
+**BBL: perhaps put this into an ecosystem context; what are these?
+Aseasonal subtropical places?**
 
@@
 2; Table S2).
-Latitudianal variation in FACF is primarily attributable to temperature-related variables (*H3, H4*), particularly MAT (Figs.
+
+Latitudinal variation in FACF is **BBL: be consistent in tense** primarily attributable to temperature-related variables (*H3, H4*), particularly MAT (Figs.
 3-4).
@@
 Temperature seasonality and growing season length are closely correlated with MAT and are strong predictors of FACF (*H4*; Figs.
-4e-f, S2, S6-S7), though growing season length doesn't improve upon MAT as a predictor.
+4e-f, S2, S6-S7), though growing season length did not improve upon MAT as a predictor.
 Within the growing season, the influence of climate on C cycling is smaller but still significant for a number of FACF (*H5*; Fig.
 S9; Table S4).
-These findings resolve a number of open debates and clarify the big picture of how FACF vary with latitude and climate on a global scale.
 
+These findings clarify how FACF vary with latitude and climate on a global scale.
 Past studies have differed in their conclusions regarding the relationship between FACF and latitude or its correlates (Table 1, *H1*)--quite possibly because of lack of standardization with respect to stand age and disturbance history.
@@
 2018: DOI: 10.1111/btp.12504**);[@delucia_forest_2007; @piao_forest_2010; @kunert_understanding_2019].
-By constraining our analysis to stands $\ge$ 100 years old with no record of major recent disturbance, we have clarified the shape of the relationship between FACF and latitude.
 
@@
 This parallels the observation that C allocation across multiple FACF varies little with respect to climate along a steep tropical elevational gradient [@malhi_variation_2017; but see @moser_elevation_2011],
-and is not surprising given that trees face some universal constraints in carbon allocation (**REFS**). 
+and is not surprising given that trees face universal constraints in carbon allocation (**REFS**). 
 The one trend in allocation that emerged from our analysis was a tendency for greater C allocation aboveground--and to foliage in particular--in warm tropical climates (Fig.
@@
 **( check / comment on  @litton_carbon_2007 )**
-One previously hypothesized trend that was not supported by our analysis was that tropical forests tend to have low carbon use efficiency ($CUE$= $NPP$/$GPP$=($GPP$-$R_{auto}$/$GPP$), which are based on observations of low $CUE$ in old-growth tropical forests relative to (mostly younger) extratropical forests [@delucia_forest_2007; @malhi_productivity_2012; @anderson-teixeira_carbon_2016].
-Limiting the analysis to mature forests, our analysis shows no such trend (Fig.
-S3), suggesting that the relatively low $CUE$ of intact tropical forests is more appropriately be attributed to the fact that these forests are old than to their tropical climate.
-Indeed, $CUE$ is known to decline with forest age [@delucia_forest_2007; @piao_forest_2010; @collalti_is_2019], but appears to be roughly independent of $GPP$ [@litton_carbon_2007].
+
+One previously hypothesized trend that was not supported by our analysis was that tropical forests tend to have low carbon use efficiency ($CUE$= $NPP$/$GPP$=($GPP$-$R_{auto}$/$GPP$), which is based on observations of low $CUE$ in old-growth tropical forests relative to (mostly younger) extratropical forests [@delucia_forest_2007; @malhi_productivity_2012; @anderson-teixeira_carbon_2016].
+Our analysis, limited to mature forests, shows no such trend (Fig.
+S3).
+$CUE$ is known to decline with forest age [@delucia_forest_2007; @piao_forest_2010; @collalti_is_2019], but appears to be roughly independent of $GPP$ [@litton_carbon_2007].
 (*This finding may have some important implications for modeling and our broader understanding.
@@
 2001 Ecological Applications*).
-However, methodological variation and uncertainty affect all of fluxes considered here--not necessarily any less than the aforementioned, and some of the larger fluxes that vary more strongly with respect to climate ($ANPP$, $NPP$) are estimated by summing uncertain component fluxes.
+However, methodological variation and uncertainty affect all of fluxes considered here, and some of the larger fluxes that vary more strongly with respect to climate ($ANPP$, $NPP$) are estimated by summing uncertain component fluxes.
 Second, differences among variables in the proportion of variation explained by climate may be attributable to more direct climatic control over $GPP$ than subsidiary fluxes.
@@
 However, it remains informative to consider these relationships.
-Among the  temperature-related climate variables ($MAT$, temperature seasonality, annual temperature range, annual frost days, length of growing season, and $PET$; Table S1), $MAT$ is generally the most strongly correlated with FACF (Table S2)--perhaps in part because site-specific MAT is recorded for the majority of sites in ForC, whereas other variables were extracted from global gridded data products.
+Among the temperature-related climate variables, $MAT$ is generally the most strongly correlated with FACF (Table S2)--perhaps in part because site-specific MAT is recorded for the majority of sites in ForC, whereas other variables were extracted from global gridded data products, introducing spatial scaling errors.
+
 This finding supports the continued focus on MAT as a primary--*albiet* not mechanistic--correlate of FACF.
@@
 3-4).
-The observed positive interaction between MAT and MAP for $ANPP_{woody.stem}$ on the global scale (Fig.
+**BBL: I feel like this is pretty repetitive of previous paragraph.** The observed positive interaction between MAT and MAP for $ANPP_{woody.stem}$ on the global scale (Fig.
 3) is consistent with an analysis showing a similar interaction for $ANPP$ in tropical forests, also with a cross-over point at ~20$\circ$C [@taylor_temperature_2017].
@@
 2007: doi: 10.1111/j.1365-2486.2006.01222.x*;  @michaletz_convergence_2014].
-We found that the sort of simple metric needed to define growing season at a global scale [@kerkhoff_plant_2005] was rough with respect to temperature and problematic with respect to moisture because it doesn't capture temporal lags between precipitation and plant water use caused by storage in soil or snow.
+We found that the sort of simple metric needed to define growing season at a global scale [@kerkhoff_plant_2005] was rough **BBL: ?** with respect to temperature and problematic with respect to moisture because it doesn't capture temporal lags between precipitation and plant water use caused by storage in soil or snow.
 A temperature-defined growing season length had stong positive correlation with FACF (Fig.
@@
 However, there remained a number of significant correlations with growing season climatic conditions, suggesting that climatic conditions remain influential within the growing season.
-We conclude that while correcting for growing season length takes analyses a step closer to mechanistic linkage of instantaneous C flux rates to environmental conditions, it remains very crude relative to the the timescales on which climate affects plant metabolism and does not advance statistical predictive power.
-Rather, mechanistic accounting for climatic effects on global FACF patterns requires models representing physiologically meaningful timescales (*e.g.*, **refs**).
+We conclude that while correcting for growing season length takes analyses a step closer to mechanistic linkage of instantaneous C flux rates to environmental conditions, it remains crude relative to the timescales on which climate affects plant metabolism, and does not advance statistical predictive power.
+Mechanistic accounting for climatic effects on global FACF patterns instead requires models representing physiologically meaningful timescales (*e.g.*, **refs**).
+**BBL: maybe cite e.g.
+ED2?
+http://dx.doi.org/10.5194/gmd-12-4309-2019 **
 
@@
 This is consistent with observations of continental- to global-scale increases in $GPP$ [@li_mapping_2019] and $ANPP_{woody stem}$ [@brienen_long-term_2015; @hubau_asynchronous_2020], along with some C cycle components not considered here--tree mortality [@brienen_long-term_2015; @mcdowell_drivers_2018], soil respiration [@bond-lamberty_global_2010], and heterotrophic soil respiration [@bond-lamberty_globally_2018].
-However, increasing C flux rates by no means universal [@rutishauser_testing_2020] (**MORE REFS**).
-This is likely because factors other than rising temperatures are at play, including changes to other aspects of climate, atmospheric pollution (CO~2~, SO~2~, NO~x~), and local disturbances.
-Morevoer, forest ecosystem responses to climatic changes outside the temperature range to which forest communities are adapated and acclimatized will not necessarily parallel responses across across geographic gradients in climate.
+However, increasing C flux rates are by no means universal [@rutishauser_testing_2020] (**MORE REFS**), likely because other factors are at play, including changes to other aspects of climate, atmospheric pollution (CO~2~, SO~2~, NO~x~), and local disturbances.
+Moreover, forest ecosystem responses to climatic changes outside the temperature range to which forest communities are adapated and acclimatized will not necessarily parallel responses across across geographic gradients in climate.
 Nevertheless, as we enter a period of accelerating climatic change, understanding of the fundamental climatic controls on FACF sets a foundation for understanding patterns of change.

### 2020-05-18 manuscript/manuscript draft.Rmd
added MEA ref
http://www.millenniumassessment.org/documents/document.354.aspx.pdf

@@
 
-Carbon cycling in forests worldwide provides the energetic basis for sustaining the majority of Earth's terrestrial biodiversity and many human populations (**REF**), while strongly influencing atmospheric CO~2~ and climate [@bonan_forests_2008].
+Carbon cycling in forests worldwide provides the energetic basis for sustaining the majority of Earth's terrestrial biodiversity and many human populations ([MEA 2005](http://www.millenniumassessment.org/documents/document.354.aspx.pdf); **REFS**), while strongly influencing atmospheric CO~2~ and climate [@bonan_forests_2008].
 Forests' autotrophic carbon fluxes (FACF)--that is, carbon fixation, allocation, and metabolism by trees and other primary producers--sets the energy ultimately available to heterotrophic organisms (including microbes), in turn influencing their abundance (e.g., [Zak et al.

### 2020-06-05 manuscript/manuscript draft.Rmd
edited abstract
removed FACF (#80)
other minor edits

@@
 ###Abstract
-Carbon fixation, allocation, and metabolism by trees set the basis for energy and material flows in forest ecosystems and define their interactions with Earth's changing climate. 
-However, we lack a cohesive synthesis on how forest autotrophic carbon fluxes (FACF) vary globally with respect to climate and one another. 
-Here, we draw upon # records from the Global Forest Carbon Database (ForC), representing all major forest types and the nine most significant FACF, to comprehensively explore how C cycling in mature, undisturbed forests varies with latitude and climate on a global scale.
-We show that, across all FACF variables analyzed, C cycling decreases linearly with latitude -- a finding that confirms multiple previous studies but contradicts the idea that net primary productivity ($NPP$) of temperate forests rivals that of tropical forests.
-FACF generally displayed similar trends across latitude and multiple climate variables, with few differences in allocation detectable at this global scale, but differed in that latitude explained a lower proportion of variation among subsidiary fluxes (in particular, woody aboveground $NPP$ and belowground $NPP$, $BNPP$).
-**BBL: not sure the second part of this sentence merits inclusion in the abstract**
-Climate explained a significant proportion (#-#%) of variation in all C fluxes analyzed, with temperature variables in general and mean annual temperature (MAT) in particular being the best predictors of FACF on this global scale.
-The effects of temperature were modified by moisture availability, with FACF reduced under hot and dry conditions and sometimes under very high precipitation. 
+Carbon (C) fixation, allocation, and metabolism by trees set the basis for energy and material flows in forest ecosystems and define their interactions with Earth's changing climate. 
+However, we lack a cohesive synthesis on how forest carbon fluxes vary globally with respect to climate and one another. 
+Here, we draw upon # records from the Global Forest Carbon Database (ForC), representing all major forest types and the nine most significant autotrophic carbon fluxes, to comprehensively explore how C cycling in mature, undisturbed forests varies with latitude and climate on a global scale.
+We show that, across all flux variables analyzed, C cycling decreases linearly with absolute latitude -- a finding that confirms multiple previous studies but contradicts the idea that net primary productivity ($NPP$) of temperate forests rivals that of tropical forests.
+C flux variables generally displayed similar trends across latitude and multiple climate variables, with few differences in allocation detectable at this global scale.
+**Climate explained a significant proportion (#-#%) of variation in all C fluxes analyzed**, with temperature variables in general and mean annual temperature (MAT) in particular being the best predictors of C flux on this global scale.
+The effects of temperature were modified by moisture availability, with C flux reduced under hot and dry conditions and sometimes under very high precipitation. 
 FACF declined with temperature seasonality, but growing season length did not improve upon MAT as a predictor.  
-Within the growing season, the influence of climate on C cycling was small but significant for a number of FACF. 
-These findings clarify how FACF varies with latitude and climate on a global scale. 
-As we enter a period of accelerating climatic change, understanding of the fundamental climatic controls on FACF sets a foundation for understanding patterns of change.
+Within the growing season, the influence of climate on C cycling was small but significant for a number of flux variables. 
+These findings clarify how forest C flux varies with latitude and climate on a global scale. 
+As we enter a period of accelerating climatic change, understanding of the fundamental climatic controls on forest C cycling sets a foundation for understanding patterns of change.
 

### 2020-06-05 manuscript/manuscript draft.Rmd
finish HM comments on intro

@@
 
-Carbon cycling in forests worldwide provides the energetic basis for sustaining the majority of Earth's terrestrial biodiversity and many human populations ([MEA 2005](http://www.millenniumassessment.org/documents/document.354.aspx.pdf); **REFS**), while strongly influencing atmospheric CO~2~ and climate [@bonan_forests_2008].
-Forests' autotrophic carbon fluxes (FACF)--that is, carbon fixation, allocation, and metabolism by trees and other primary producers--sets the energy ultimately available to heterotrophic organisms (including microbes), in turn influencing their abundance (e.g., [Zak et al.
-1994](http://esajournals.onlinelibrary.wiley.com/doi/abs/10.2307/1940888); [Niedziałkowska et al.
-2010](https://doi.org/10.2980/17-1-3310)) and possibly diversity [@waide_relationship_1999; Chu et al.
-2019 (DOI: 10.1111/ele.13175)].
+Carbon cycling in forests worldwide provides the energetic basis for sustaining the majority of Earth's terrestrial biodiversity and many human populations (**[MEA 2005](http://www.millenniumassessment.org/documents/document.354.aspx.pdf)**; **REFS**), while strongly influencing atmospheric CO~2~ and climate [@bonan_forests_2008].
+Forests' autotrophic carbon fluxes (FACF)--that is, carbon fixation, allocation, and metabolism by trees and other primary producers--sets the energy ultimately available to heterotrophic organisms (including microbes), in turn influencing their abundance (e.g., **[Zak et al.
+1994](http://esajournals.onlinelibrary.wiley.com/doi/abs/10.2307/1940888)**; [**Niedziałkowska et al.
+2010](https://doi.org/10.2980/17-1-3310)**) and possibly diversity [@waide_relationship_1999; **Chu et al.
+2019 (DOI: 10.1111/ele.13175)**].
 FACF influences all organic matter stocks in forest ecosystems and is linked to cycling of energy, water, and nutrients.
@@
 In contrast, some studies suggest that the highest values of net primary productivity ($NPP$) may be found in temperate forests [@luyssaert_co_2007; @huston_global_2009], while others find $NPP$ highest in the tropics and decreasing with latitude [@simova_enigma_2017].
-Other studies have chosen to focus exclusively on above-ground net primary productivity ($ANPP$), finding evidence of a weak negative relationship between $ANPP$ and latitude [@huston_global_2009; @gillman_latitude_2015].
+Other studies have found that aboveground net primary productivity ($ANPP$) decreases weakly with latitude [@huston_global_2009; @gillman_latitude_2015].
 
-The latitudinal gradient in FACF is primarily driven by climate, which is a significant driver of FACF across broad spatial scales [@luyssaert_co_2007; @cleveland_relationships_2011; @hursh_sensitivity_2017].
+The latitudinal gradient in FACF, along with altitudinal gradients [@girardin_net_2010; @malhi_variation_2017], is primarily driven by climate, which is a significant driver of FACF across broad spatial scales [@luyssaert_co_2007; @cleveland_relationships_2011; @hursh_sensitivity_2017].
 The majority of studies have focused on exploring the relationships of FACF to mean annual temperature (MAT) and mean annual precipitation (MAP), as the most commonly reported site-level climate variables.
@@
 [-@luyssaert_co_2007] examined GPP and NPP and found that, while $GPP$ increases linearly with MAT, NPP saturates at around 10$^\circ$C MAT.
-In contrast, Larjavaara and Muller-Landau [-@larjavaara_temperature_2012], find that increases in $GPP$ saturate at approximately 25$^\circ$C MAT, while Schuur [-@schuur_productivity_2003] shows that $NPP$ increases linearly with temperature.
+In contrast, Larjavaara and Muller-Landau [-@larjavaara_temperature_2012], find that increases in $GPP$ saturate at approximately 25$^\circ$C MAT, while Schuur [-@schuur_productivity_2003] finds that $NPP$ increases linearly with temperature.
 Taylor et al.
 [-@taylor_temperature_2017] showed a positive interaction between MAT and MAP in shaping tropical forest productivity, such that high rainfall had a negative effect on productivity in cooler climates, compared to a positive effect in warmer climates.
-Such complicated dynamics play out with belowground fluxes as well, with typically dominant soil temperature mediated by moisture and carbon supply (Hursh et al.
+Such complicated dynamics play out with belowground fluxes as well, with typically dominant soil temperature mediated by moisture and carbon supply (**Hursh et al.
 2017, http://dx.doi.org/10.1111/gcb.13489; Xu et al.
-2016 http://dx.doi.org/10.1016/j.jplph.2016.08.007).
+2016 http://dx.doi.org/10.1016/j.jplph.2016.08.007**).
 
@@
 [-@kerkhoff_plant_2005] and Michaletz et al.
-[-@michaletz_convergence_2014] find that, within the growing season, there is no significant relationship between primary productivity and MAT, suggesting that the effect of temperature is due to increased length of growing season, rather than an inherent influence of temperature on FACF.
+[-@michaletz_convergence_2014] find that, within the growing season, there is no significant relationship between net primary productivity (loosely defined) and MAT, suggesting that the influence of temperature may be limited to determining the length of the frost-free growing season.
 
+**the following doesn't fit very well...**  
 In addition, FACF can be influenced by other, often interactive, factors across a range of scales [@cleveland_relationships_2011].
-On a local scale, stand age [@litton_carbon_2007; @gillman_latitude_2015], biodiversity [@liang_positive_2016], management [@simova_enigma_2017], nutrient availability [@aragao_above_2009], and altitude [@girardin_net_2010; @malhi_variation_2017] all impact FACF.
+On a local scale, stand age [@litton_carbon_2007; @gillman_latitude_2015], biodiversity [@liang_positive_2016], management [@simova_enigma_2017], and nutrient availability [@aragao_above_2009].
 There is evidence that FACFs also respond to variables such as cloud cover [@taylor_temperature_2017], solar radiation [@fyllas_solar_2017], and potential evapotranspiration [@kerkhoff_plant_2005].
@@
 kable(hypothesis_table, format = "latex", booktabs = TRUE, caption = "Summary of research questions, corresponding hypotheses, and results.
-Statistically signficant support for/ rejection of hypotheses is indicated with 'yes'/'no', parentheses indicate partial overall support/rejection of hypotheses across all FACF, and '-' indicates no significant relationship.", escape = F) %>%
+Statistically signficant support for/ rejection of hypotheses is indicated with 'yes'/'no', and '-' indicates no significant relationship.
+Parentheses indicate partial overall support or rejection of hypotheses across all fluxes considered.", escape = F) %>%
   kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 12) %>%

### 2020-06-05 manuscript/manuscript draft.Rmd
standardize variable symbols in main text
#55

@@
 Specifically, latitude explained 64% of variation in GPP (n = 254, p<0.0001), 50% in NPP (n = 114, p<0.0001) and 45% in ANPP (n = 259, p<0.0001).
-The FACF that were most poorly predicted by latitude were $BNPP_{root-fine}$ ($R^2$=0.17) and $ANPP_{woody.stem}$ ($R^2$=0.18).
+The FACF that were most poorly predicted by latitude were $BNPP_{fine.root}$ ($R^2$=0.17) and $ANPP_{woody.stem}$ ($R^2$=0.18).
 For all FACF, the relationship with latitude was best predicted by the first-order linear model.
@@
 2).
-That is, modelled estimates of $GPP$, generated from the sum of $NPP$ and $R_{auto}$; $NPP$, generated from the sum of $ANPP$ and $BNPP$; and $ANPP$, generated from the sum of $ANPP_{foliage}$ and $ANPP_{woody-stem}$, fell completely within the confidence intervals of the regressions of field estimates of $GPP$, $NPP$, and $ANPP$, respectively.
+That is, modelled estimates of $GPP$, generated from the sum of $NPP$ and $R_{auto}$; $NPP$, generated from the sum of $ANPP$ and $BNPP$; and $ANPP$, generated from the sum of $ANPP_{foliage}$ and $ANPP_{stem}$, fell completely within the confidence intervals of the regressions of field estimates of $GPP$, $NPP$, and $ANPP$, respectively.
 
@@
 3-4, S4-S5, Table S2).
-As with latitude, MAT tended to explain more variation in the larger FACF ($GPP$, $NPP$, $ANPP$, $R_{auto}$) and $ANPP_{foliage}$ (all $R^2$> 0.4) than in subsidiary and belowground fluxes ($ANPP_{woody-stem}$, $R_{root}$, $BNPP_{root-fine}$; all $R^2$< 0.25).
+As with latitude, MAT tended to explain more variation in the larger FACF ($GPP$, $NPP$, $ANPP$, $R_{auto}$) and $ANPP_{foliage}$ (all $R^2$> 0.4) than in subsidiary and belowground fluxes ($ANPP_{stem}$, $R_{root}$, $BNPP_{fine.root}$; all $R^2$< 0.25).
 
-MAP was a significant (p<0.05) predictor of all FACF except $ANPP_{woody-stem}$ (Figs.
+MAP was a significant (p<0.05) predictor of all FACF except $ANPP_{stem}$ (Figs.
 4a, S4-S5; Table S2).
@@
 Accounting for MAT, MAP had a substantial positive effect on $GPP$ and $R_{auto}$ and a small negative effect on $ANPP$.
-There was a significant interactive effect between MAT and MAP for $NPP$ and $ANPP_{woody-stem}$ (Fig.
+There was a significant interactive effect between MAT and MAP for $NPP$ and $ANPP_{stem}$ (Fig.
 3, Table S3).
-The interaction was negative for $NPP$ and positive for $ANPP_{woody-stem}$.
-For $ANPP_{foliage}$, $BNPP$, $BNPP_{root-fine}$, and $R_{auto-root}$, MAP did not have a signficant effect when accounting for MAT (Fig.
+The interaction was negative for $NPP$ and positive for $ANPP_{stem}$.
+For $ANPP_{foliage}$, $BNPP$, $BNPP_{fine.root}$, and $R_{root}$, MAP did not have a signficant effect when accounting for MAT (Fig.
 3, Table S3).
@@
 We found a significant relationship between C flux and potential evapotranspiration for all FACF.
-$ANPP_{foliage}$, $BNPP_{root-fine}$ and $R_{root}$ increased linearly with PET; however, all other fluxes showed a polynomial relationship with PET (Fig.
+$ANPP_{foliage}$, $BNPP_{fine.root}$ and $R_{root}$ increased linearly with PET; however, all other fluxes showed a polynomial relationship with PET (Fig.
 4c, S4-5; Table S2).
@@
 Vapour pressure deficit was a significant predictor of C flux for all FACF.
-$BNPP_{root-fine}$ showed a linear relationship with vapour pressure deficit ($R^2$ = 0.07, p<0.05), but all other fluxes showed a polynomial relationship (Figs.
+$BNPP_{fine.root}$ showed a linear relationship with vapour pressure deficit ($R^2$ = 0.07, p<0.05), but all other fluxes showed a polynomial relationship (Figs.
 4d, S4-5; Table S2).
@@
 4e, S6-7; Table S2).
-$ANPP_{foliage}$, $ANPP_{woody-stem}$ and $R_{auto}$ decreased linearly with temperature seasonality (all p<0.05; Figs.
+$ANPP_{foliage}$, $ANPP_{stem}$ and $R_{auto}$ decreased linearly with temperature seasonality (all p<0.05; Figs.
 4e, S6-S7; Table S2).
@@
 Growing season solar radiation had a positive influence on $GPP$ ($R^2$ = 0.21, p<0.001), $NPP$ ($R^2$ = 0.21, p<0.001), $BNPP$ ($R^2$ = 0.16, p<0.001) and $BNPP_{fine.root}$ ($R^2$ = 0.12, p<0.01).
-Growing season PET had a positive influence on $GPP$ ($R^2$ = 0.15, p<0.01), $NPP$ ($R^2$ = 0.18, p<0.01), $BNPP$ ($R^2$ = 0.23, p<0.0001), $BNPP_{fine.root}$ ($R^2$ = 0.11, p<0.05), and $ANPP_{woody-stem}$ ($R^2$ = 0.06, p<0.05).
+Growing season PET had a positive influence on $GPP$ ($R^2$ = 0.15, p<0.01), $NPP$ ($R^2$ = 0.18, p<0.01), $BNPP$ ($R^2$ = 0.23, p<0.0001), $BNPP_{fine.root}$ ($R^2$ = 0.11, p<0.05), and $ANPP_{stem}$ ($R^2$ = 0.06, p<0.05).
 {**Becky, please verify/ edit the following:** There were no other signifcant correlations between growing season length-standardized FACF (9 variables in Table 2) and growing season climate (**which variables?**)}.
@@
 2), indicating no major, systematic omissions or overestimations of flux components.
-However, climate explained lower proportions of variability among subsidiary C fluxes (*e.g.*, $ANPP_{woody}$, $BNPP_{fine.root}$, $R_{auto-root}$; Fig.
+However, climate explained lower proportions of variability among subsidiary C fluxes (*e.g.*, $ANPP_{woody}$, $BNPP_{fine.root}$, $R_{root}$; Fig.
 2; Table S2).
@@
 and with observations and theory predicting relatively higher belowground carbon allocation at higher latitudes [@gill_belowground_2016].
-It is also consistent with findings that as forest productivity increases, proportionaly more carbon is allocated to $ANPP_{foliage}$ relative to $BNPP_{root-fine}$ (*Chen et al.
+It is also consistent with findings that as forest productivity increases, proportionaly more carbon is allocated to $ANPP_{foliage}$ relative to $BNPP_{fine.root}$ (*Chen et al.
 2019; doi: 10.1111/ele.13193*) or $ANPP_{woody}$ [@hofhansl_new_2015].
@@
 
-One interesting observation was that climate tends to explain more variation in the major fluxes ($GPP$, $NPP$, $R_{auto}$) than in subsidiary fluxes ($BNPP_{fine.root}$, $R_{auto-root}$, $ANPP_{woody-stem}$) (Fig.
+One interesting observation was that climate tends to explain more variation in the major fluxes ($GPP$, $NPP$, $R_{auto}$) than in subsidiary fluxes ($BNPP_{fine.root}$, $R_{root}$, $ANPP_{stem}$) (Fig.
 2; Table S2).
@@
 First, it may be that methodological variation is larger relative to flux magnitude for some of the subsidiary fluxes.
-Belowground fluxes in particular are difficult to quantify, and measurement methods for the belowground fluxes considered here may be measured through fundamentally different approaches (*e.g.*, minirhizotrons, ingrowth cores, or sequential coring for $BNPP_{root-fine}$; root exclusion, stable isotope tracking, or gas exchange of excised roots for $R_{auto-root}$), and sampling depth is variable and often insufficient to capture the full soil profile.
-$ANPP_{woody-stem}$, which is also poorly explained by latitude or climate, is more straightforward to measure but is subject to variability introduced by differences such as biomass allometries applied and minimum plant size sampled (*Clark et al.
+Belowground fluxes in particular are difficult to quantify, and measurement methods for the belowground fluxes considered here may be measured through fundamentally different approaches (*e.g.*, minirhizotrons, ingrowth cores, or sequential coring for $BNPP_{fine.root}$; root exclusion, stable isotope tracking, or gas exchange of excised roots for $R_{root}$), and sampling depth is variable and often insufficient to capture the full soil profile.
+$ANPP_{stem}$, which is also poorly explained by latitude or climate, is more straightforward to measure but is subject to variability introduced by differences such as biomass allometries applied and minimum plant size sampled (*Clark et al.
 2001 Ecological Applications*).

### 2020-06-05 manuscript/manuscript draft.Rmd
(slow) progress on HML comments
I've incorporated (or opened issues for) all comments prior to analyses section of Methods.

#81

@@
 ForC contains geographic coordinates associated with each measurement record and, when available, mean annual temperature (MAT) and mean annual precipitation (MAP) as reported in the primary literature [@anderson-teixeira_forc_2018].
-Based on the geographic co-ordinates for each site, data on twelve climate variables--including MAT, MAP, temperature and precipitation seasonality, annual temperature range, solar radiation, cloud cover, annual frost and wet days, potential evapotranspiration (PET), aridity (MAP/PET), and vapor pressure deficit (VPD)--were extracted from five open-access climate datasets: WorldClim [@hijmans_very_2005], WorldClim2 [@fick_worldclim_2017], the Climate Research Unit (CRU) time-series dataset v.
-4.03 [@harris_updated_2014], the Global Aridity Index and Potential Evapotranspiration Climate Database [@trabucco_global_2019], and TerraClimate [@abatzoglou_terraclimate_2018] (Table S1).
+Based on the geographic co-ordinates for each site, data on twelve climate variables--including MAT, MAP, temperature and precipitation seasonality, annual temperature range, solar radiation, cloud cover, annual frost and wet days, potential evapotranspiration (PET), aridity (MAP/PET), and vapor pressure deficit (VPD)--were extracted from five open-access climate datasets: WorldClim [@hijmans_very_2005], WorldClim2 [@fick_worldclim_2017], the Climate Research Unit time-series dataset (CRU TS v4.03 [@harris_updated_2014], the Global Aridity Index and Potential Evapotranspiration Climate Database [@trabucco_global_2019], and TerraClimate [@abatzoglou_terraclimate_2018] (Table S1).
 From these data, we derived maximum VPD, defined as the VPD of the month with the largest deficit, and the number of water stress months, defined as the number of months annually where precipitation was lower than PET.
-Where site-level data was missing for MAT or MAP, we used values from the WorldClim dataset.
+Where site-level data was missing for MAT or MAP, we used values from the WorldClim (**[2?](https://github.com/forc-db/Global_Productivity/issues/89)**) dataset.
 
@@
 We experimented with a definion of growing season months including a moisture index, defined as (MAT - PET)/PET, > -0.95 [@kerkhoff_plant_2005; see also @michaletz_convergence_2014].
-However, we found that including a moisture index had **no** effect on the estimates of growing season length, and so chose to exclude it.
-(**Becky, was it really no effect?
-or minimal?**) Monthly data for PET, precipitation, and temperature from the CRU dataset v 4.03 [@harris_updated_2014], and solar radiation from WorldClim2 [@fick_worldclim_2017] were used to calculate mean monthly PET, precipitation, temperature and solar radiation during the growing season.
+However, we found that including a moisture index had minimal effect on the estimates of growing season length, and so chose to exclude it.
+Monthly data for PET, precipitation, and temperature from the CRU dataset v 4.03 [@harris_updated_2014], and solar radiation from WorldClim2 [@fick_worldclim_2017] were used to calculate mean monthly PET, precipitation, temperature and solar radiation during the growing season.
 Total growing season precipitation and solar radiation were also calculated.

### 2020-06-05 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd
#78

@@
 
+**drop this paragraph ([issue 78](https://github.com/forc-db/Global_Productivity/issues/78))?** 
 One interesting observation was that climate tends to explain more variation in the major fluxes ($GPP$, $NPP$, $R_{auto}$) than in subsidiary fluxes ($BNPP_{fine.root}$, $R_{root}$, $ANPP_{stem}$) (Fig.

### 2020-06-10 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd
corrected woody stem to stem

@@
 Confidence intervals for the larger flux were calculated using the 'bootMer' function from the lme4 package [@bates_fitting_2015].
-This analysis was applied to the following sets of fluxes: (1) $GPP = NPP + R_{auto}$, (2) $NPP = ANPP + BNPP$, and (3) $ANPP = ANPP_{foliage} + ANPP_{woody.stem}$.
+This analysis was applied to the following sets of fluxes: (1) $GPP = NPP + R_{auto}$, (2) $NPP = ANPP + BNPP$, and (3) $ANPP = ANPP_{foliage} + ANPP_{stem}$.
 In addition, we estimated total belowground C flux (TBCF, not analyzed due to limited data) as $TBCF = BNPP + R_{root}$.
 
-Variation in allocation to component carbon fluxes along latitudinal gradients was explored for the following pairings: $GPP:NPP$, $ANPP:BNPP$, $ANPP_{foliage}:ANPP_{woody.stem}$, $ANPP_{foliage}:NPP$, $ANPP_{woody.stem}:NPP$, $ANPP:NPP$, and $BNPP:NPP$.
+Variation in allocation to component carbon fluxes along latitudinal gradients was explored for the following pairings: $GPP:NPP$, $ANPP:BNPP$, $ANPP_{foliage}:ANPP_{stem}$, $ANPP_{foliage}:NPP$, $ANPP_{stem}:NPP$, $ANPP:NPP$, and $BNPP:NPP$.
 For each set of paired fluxes, measurements taken at the same site and plot, and in the same year, were paired together, and the ratio of each pair of measurements calculated.
@@
 Specifically, latitude explained 64% of variation in GPP (n = 254, p<0.0001), 50% in NPP (n = 114, p<0.0001) and 45% in ANPP (n = 259, p<0.0001).
-The FACF that were most poorly predicted by latitude were $BNPP_{fine.root}$ ($R^2$=0.17) and $ANPP_{woody.stem}$ ($R^2$=0.18).
+The FACF that were most poorly predicted by latitude were $BNPP_{fine.root}$ ($R^2$=0.17) and $ANPP_{stem}$ ($R^2$=0.18).
 For all FACF, the relationship with latitude was best predicted by the first-order linear model.
@@
 2), indicating no major, systematic omissions or overestimations of flux components.
-However, climate explained lower proportions of variability among subsidiary C fluxes (*e.g.*, $ANPP_{woody}$, $BNPP_{fine.root}$, $R_{root}$; Fig.
+However, climate explained lower proportions of variability among subsidiary C fluxes (*e.g.*, $ANPP_{stem}$, $BNPP_{fine.root}$, $R_{root}$; Fig.
 2; Table S2).
@@
 It is also consistent with findings that as forest productivity increases, proportionaly more carbon is allocated to $ANPP_{foliage}$ relative to $BNPP_{fine.root}$ (*Chen et al.
-2019; doi: 10.1111/ele.13193*) or $ANPP_{woody}$ [@hofhansl_new_2015].
+2019; doi: 10.1111/ele.13193*) or $ANPP_{stem}$ [@hofhansl_new_2015].
 **( check / comment on  @litton_carbon_2007 )**
@@
 3-4).
-**BBL: I feel like this is pretty repetitive of previous paragraph.** The observed positive interaction between MAT and MAP for $ANPP_{woody.stem}$ on the global scale (Fig.
+**BBL: I feel like this is pretty repetitive of previous paragraph.** The observed positive interaction between MAT and MAP for $ANPP_{stem}$ on the global scale (Fig.
 3) is consistent with an analysis showing a similar interaction for $ANPP$ in tropical forests, also with a cross-over point at ~20$\circ$C [@taylor_temperature_2017].
@@
 2-3).
-This is consistent with observations of continental- to global-scale increases in $GPP$ [@li_mapping_2019] and $ANPP_{woody stem}$ [@brienen_long-term_2015; @hubau_asynchronous_2020], along with some C cycle components not considered here--tree mortality [@brienen_long-term_2015; @mcdowell_drivers_2018], soil respiration [@bond-lamberty_global_2010], and heterotrophic soil respiration [@bond-lamberty_globally_2018].
+This is consistent with observations of continental- to global-scale increases in $GPP$ [@li_mapping_2019] and $ANPP_{stem}$ [@brienen_long-term_2015; @hubau_asynchronous_2020], along with some C cycle components not considered here--tree mortality [@brienen_long-term_2015; @mcdowell_drivers_2018], soil respiration [@bond-lamberty_global_2010], and heterotrophic soil respiration [@bond-lamberty_globally_2018].
 However, increasing C flux rates are by no means universal [@rutishauser_testing_2020] (**MORE REFS**), likely because other factors are at play, including changes to other aspects of climate, atmospheric pollution (CO~2~, SO~2~, NO~x~), and local disturbances.

### 2020-06-11 manuscript/manuscript draft.Rmd
switched to GCB ref style (csl)

--- 2020-06-05 manuscript/manuscript draft.Rmd
@@
 biblio-style: apalike
-csl: apa.csl
+csl: global-change-biology.csl
 header-includes:  \usepackage{float}

### 2020-06-13 manuscript/manuscript draft.Rmd
worked on HM comments

@@
 Geographic areas--*i.e.*, spatially clustered sites--are defined within ForC using a hierarchical cluster analysis on the distance matrix of the sites and a cutoff of 25km [@anderson-teixeira_forc_2018].
-We experimented with inclusion of altitude as a fixed effect, but excluded it from the final models because it added very little explanatory power--that is, the difference in AIC ($\Delta$AIC) relative to models excluding altitude was generally small (often $\Delta$AIC<2).
-Hypotheses were accepted if the $\Delta$AIC between a model including the fixed effect of interest and a corresponding null model excluding that fixed effect exceeded 2.0.
+We experimented with inclusion of altitude as a fixed effect, but excluded it from the final models because it added very little explanatory power--that is, the difference in AIC ($\Delta AIC$) relative to models excluding altitude was generally small (often $\Delta AIC$<2).
+Effects were considered signficant when inclusion of the term of interest resulted in $\Delta AIC \ge 2.0$ relative to a corresponding null model.
 All $R^2$ values presented here are marginal $R^2$ values, and refer to the proportion of variation explained by only the fixed effects.
@@
 We first examined the relationship between latitude and FACF (Q1; Table 1).
-We tested models with latitude as a linear term (corresponding null: model without latitude) and as a second-order polynomial term (corresponding null: model with latitude as a linear term), and calculated AIC values to determine the best model.
-**BBL: careful about terminology here.
-In statistics a "linear model" includes polynomial models, which are linear in their parameters.
-Thus I (and many readers perhaps) find this confusing.** Models were accepted as the best model if $\Delta$AIC > 2 with respect to the corresponding null, and were significant with respect to a null model with no fixed term.
+We tested models with latitude as a first-order linear term (corresponding null: model without latitude) and as a second-order polynomial or logarithmic term (corresponding null: model with latitude as a first-order linear term).
+We selected as the best model that with the highest $\Delta AIC$ relative to a null model with no fixed term.
 
-To test whether trends in component fluxes across latitude sum to match those of larger fluxes, regression lines for smaller component fluxes were summed to generate new estimates of larger fluxes, which were then compared against the latitudinal regression of the larger flux.
+To test whether trends in component fluxes across latitude sum to match those of larger fluxes, regression lines for smaller component fluxes were summed to generate new estimates of larger fluxes. 
+We then determined whether these summed predictions fell within the **95%** CI for the larger flux across the entire range included latitude. **(Becky, is 95% the correct number)**
 Confidence intervals for the larger flux were calculated using the 'bootMer' function from the lme4 package [@bates_fitting_2015].
+**(Helene asks for more details here: what does the function do?
+are CIs just based on uncertainty in regression?)**
 This analysis was applied to the following sets of fluxes: (1) $GPP = NPP + R_{auto}$, (2) $NPP = ANPP + BNPP$, and (3) $ANPP = ANPP_{foliage} + ANPP_{stem}$.
@@
 
-Variation in allocation to component carbon fluxes along latitudinal gradients was explored for the following pairings: $GPP:NPP$, $ANPP:BNPP$, $ANPP_{foliage}:ANPP_{stem}$, $ANPP_{foliage}:NPP$, $ANPP_{stem}:NPP$, $ANPP:NPP$, and $BNPP:NPP$.
+Variation in allocation to component carbon fluxes along latitudinal gradients was explored for the following pairings: $NPP:GPP$, $ANPP:BNPP$, $ANPP_{foliage}:ANPP_{stem}$, $ANPP_{foliage}:NPP$, $ANPP_{stem}:NPP$, $ANPP:NPP$, and $BNPP:NPP$.
+**[issue 91](https://github.com/forc-db/Global_Productivity/issues/91)**
 For each set of paired fluxes, measurements taken at the same site and plot, and in the same year, were paired together, and the ratio of each pair of measurements calculated.
@@
 Cook's distance analyses were carried out for each of the models, and indicated that data from a few high-elevation sites were having a disproportionate influence on the regressions.
-To account for this, models were re-run using only data from sites $\le$ 1000m.
+Thus, models were re-run using only data from sites $\le$ 1000m.
 
@@
 
-Multivariate models were used to investigate the potential joint and interactive effects of climate variables on carbon fluxes.
-We explored a large range of climate variable combinations and ultimately decided to focus on MAT and MAP.
-We favored these over variables with similar explanatory power because they are the most commonly reported climate variables.
-MAT was consistently among the best individual predictors of FACF (Table S2), although it was occasionally out-performed by closely correlated climate variables (Fig.
-S2).
-In combination with MAT as a primary climate variable, MAP was consistently among the best secondary climate variables.
-Specifically, compared against a model including MAT x MAP, there were no other variables that, in interaction with MAT, improved AIC by >2.0 and $R^2$ by >0.05.
-In examining the joint and interactive effects of climate variables, an additive model including MAT and MAP as fixed effects was accepted when $\Delta$AIC >2 relative to a null including only MAT as a fixed effect.
+Multivariate models were used to investigate the potential joint and interactive effects of MAT and MAP on carbon fluxes.
+An additive model including MAP in addition to MAT was accepted when $\Delta AIC$ >2 relative to a null including only MAT as a fixed effect.
 An interactive model including an MAT x MAP interaction was accepted when $\Delta$AIC >2 relative to a null including MAT and MAP as fixed effects.
@@
 To test whether and how FACF varied with climate when standardised by growing season length (Q5), we first standardized FACF by dividing by growing season length (as defined above).
-We then tested for correlations between these standardised fluxes and growing season climate variables, using only linear models.
+We then tested for correlations between these standardised fluxes and growing season climate variables, using only first-order linear models.
 
@@
 In total, we analyzed 1228 records from nine FACF variables taken from forests that had experienced no major anthropogenic disturbances within the past 100 years.
-These records represented a total of 154 distinct geographic areas across all forested biogeographic and climate zones (Fig.
+These records represented a total of **#** plots in 154 distinct geographic areas across all forested biogeographic and climate zones (Fig.
 1, Table 2).
@@
 The FACF that were most poorly predicted by latitude were $BNPP_{fine.root}$ ($R^2$=0.17) and $ANPP_{stem}$ ($R^2$=0.18).
-For all FACF, the relationship with latitude was best predicted by the first-order linear model.
+For all FACF, the relationship with latitude was better fit by the first-order linear model than by second-order polynomial or logarithmic models.
 
@@
 95\\% confidence intervals are plotted for the major flux for comparison with predicted trends.
-In (d),  which shows three belowground fluxes, the major flux, total belowground carbon flux, is one for which we have no data", fig.pos='H'}
+In (d), which shows three belowground fluxes, the major flux, total belowground carbon flux, has insufficient data (n=9) to support a regression", fig.pos='H'}
 
@@
 
-There was little evidence of substantial variation in C allocation across latitude or climate (Fig.
+There was little evidence of systematic variation in C allocation with latitude or climate (Fig.
 S3).
-Of the 7 FACF ratios regressed against latitude and three climate variables (MAT, MAP, temperature seasonality), there were only five signficant relationships, all with $R^2 \le 0.4$ (Fig.
+Of 28 relationships tested (7 FACF ratios regressed against latitude, MAT, MAP, and temperature seasonality), there were only five signficant relationships, all with $R^2 \le 0.4$ (Fig.
 S3).
-Specifically, the proportion of NPP allocated to $ANPP_{foliage}$ decreased with latitude ($R^2=0.32$) and increased with MAT ($R^2=0.37$), and the proportion of $NPP$ allocated aboveground ($ANPP$) decreased weakly with latitude ($R^2=0.11$) and temperature seasonality ($R^2=0.17$), while increasing with MAT ($R^2=0.11$).
-The remaining 23 relationships examined were non-significant.
+The proportion of NPP allocated to $ANPP_{foliage}$ decreased with latitude ($R^2=0.32$) and increased with MAT ($R^2=0.37$).
+The proportion of $NPP$ allocated aboveground ($ANPP$) decreased weakly with latitude ($R^2=0.11$) and temperature seasonality ($R^2=0.17$), while increasing with MAT ($R^2=0.11$).
 

### 2020-06-13 manuscript/manuscript draft.Rmd
finished hm comments on manuscript pdf
These are all now dealt with or covered in issues.

Closes #81

@@
 
-All FACF increased linearly with MAT, and we found no support for a saturation point of FACF with MAT (all p<0.05; Figs.
-3-4, S4-S5, Table S2).
-As with latitude, MAT tended to explain more variation in the larger FACF ($GPP$, $NPP$, $ANPP$, $R_{auto}$) and $ANPP_{foliage}$ (all $R^2$> 0.4) than in subsidiary and belowground fluxes ($ANPP_{stem}$, $R_{root}$, $BNPP_{fine.root}$; all $R^2$< 0.25).
+All fluxes increased with MAT (all p<0.05; Figs. 3-4, S4-S5, Table S2). 
+For eight of the nine fluxes, this relationship was linear.
+For only one variable, $BNPP$, did a lognormal fit provide signficant improvement over a first-order linear relatinship. 
+*As with latitude, MAT tended to explain more variation in the larger flux ($GPP$, $NPP$, $ANPP$, $R_{auto}$) and $ANPP_{foliage}$ (all $R^2$> 0.4) than in subsidiary and belowground fluxes ($ANPP_{stem}$, $R_{root}$, $BNPP_{fine.root}$; all $R^2$< 0.25).* **update this**
 
-MAP was a significant (p<0.05) predictor of all FACF except $ANPP_{stem}$ (Figs.
-4a, S4-S5; Table S2).
-However, it explained little variation: with the exception of $R_{auto}$, MAP explained at most 37% of variation in FACF.
-For the majority of FACF, a polynomial model **BBL: of what order?** provided the best fit.
-FACF generally increased with precipitation, up until a saturation point at between 3000 and 4000mm annual precipitation, above which they started to decrease (Figs.
-4, S4-S5).
-The notable exception to this was GPP: the model indicated that GPP continued to increase with precipitation up to measures of at least 5000mm annually (p<0.0001, $R^2$ = 0.33.
-Data above this point were not available, but the model trend suggested a saturation point around 5000mm MAP.
+MAP was a significant (p<0.05) predictor of all fluxes (Figs. 4a, S4-S5; Table S2). 
+However, it explained little variation: with the exception of $R_{auto}$, MAP explained at most 25% of variation in FACF. 
+For five of the nine fluxes, a second-order polynomial model provided the best fit, while first-order linear and lognormal relationships provided each provided the best fit for two flxues. 
+All fluxes increased with MAP up to at least 2000 mm, above which responses were variable (Figs. 4, S4-S5).
 
@@
 Accounting for MAT, MAP had a substantial positive effect on $GPP$ and $R_{auto}$ and a small negative effect on $ANPP$.
-There was a significant interactive effect between MAT and MAP for $NPP$ and $ANPP_{stem}$ (Fig.
+There was a significant interaction between MAT and MAP for $NPP$ and $ANPP_{stem}$ (Fig.
 3, Table S3).
@@
 We found strong evidence for a saturation point or peak with PET: FACF tended to increase at values below 1000mm, before saturating between 1200 and 1700mm.
-There was also evidence that FACF begin to decrease at values above 1800mm PET.
+There was also evidence that some FACF begin to decrease at values above 1800mm PET.
 
@@
 
-Temperature seasonality was a significant predictor of FACF.
+Temperature seasonality was, overall, the best predictor of FACF. **(update this paragraph based on new results)**
 $GPP$, $NPP$, $ANPP$, and $R_{root}$ exhibited a polynomial relationship with seasonality (all p<0.05; Figs.
@@
 
-In contrast, there was no significant effect of precipitation seasonality on FACF, and both maximum vapour pressure deficit, and water stress months were poor or non-significant explainers of variation in FACF (Figs.
+In contrast, there was no significant effect of precipitation seasonality on FACF, and both maximum vapour pressure deficit, and water stress months were poor or non-significant predictors of variation in FACF (Figs.
 S6-S7; Table S2).
@@
 
-When FACF were standardized by growing season length, correlations with growing season climate--including temperature, precipitation, solar radiation, and PET--were generally weak (Figs.
+When FACF were standardized by growing season length (in monthly increments), correlations with growing season climate--including temperature, precipitation, solar radiation, and PET--were generally weak (Figs.
 S8-S9).
@@
 Past studies have differed in their conclusions regarding the relationship between FACF and latitude or its correlates (Table 1, *H1*)--quite possibly because of lack of standardization with respect to stand age and disturbance history.
-Our findings indicate that, among mature, undisturbed stands, FACF are unambiguously highest in the tropical regions, and the relationship is approximately linear (Fig.
+Our findings indicate that, among mature, undisturbed stands, FACF are unambiguously highest in the tropical regions, and the relationship with latitude is approximately linear (Fig.
 2).
@@
 S3).
-This is consistent with observations of increasing allocation to roots with declining temperature across a tropical elevational gradient [@moser_elevation_201],
+This is consistent with observations of increasing allocation to roots with declining temperature across a tropical elevational gradient [@moser_elevation_2011],
 and with observations and theory predicting relatively higher belowground carbon allocation at higher latitudes [@gill_belowground_2016].
@@
 (*This finding may have some important implications for modeling and our broader understanding.
-I'm not sure offhand how much theory has been built on the idea of low CUE in tropical forests because of warm temperatures...*)
+I'm not sure offhand how much theory has been built on the idea of low CUE in tropical forests because of warm temperatures...Helene is not aware of any.*)
 
-**drop this paragraph ([issue 78](https://github.com/forc-db/Global_Productivity/issues/78))?** 
-One interesting observation was that climate tends to explain more variation in the major fluxes ($GPP$, $NPP$, $R_{auto}$) than in subsidiary fluxes ($BNPP_{fine.root}$, $R_{root}$, $ANPP_{stem}$) (Fig.
+**update this paragraph ([issue 78](https://github.com/forc-db/Global_Productivity/issues/78))** 
+One interesting observation was that climate tends to explain more variation in the major fluxes ($GPP$, $NPP$, $R_{auto}$) than in subsidiary fluxes ($BNPP_{fine.root}$, $R_{root}$, $ANPP_{stem}$) as quantified by R^2^ (Fig.
 2; Table S2).
@@
 First, it may be that methodological variation is larger relative to flux magnitude for some of the subsidiary fluxes.
-Belowground fluxes in particular are difficult to quantify, and measurement methods for the belowground fluxes considered here may be measured through fundamentally different approaches (*e.g.*, minirhizotrons, ingrowth cores, or sequential coring for $BNPP_{fine.root}$; root exclusion, stable isotope tracking, or gas exchange of excised roots for $R_{root}$), and sampling depth is variable and often insufficient to capture the full soil profile.
+Belowground fluxes in particular are difficult to quantify, and measurement methods for the belowground fluxes considered here may use fundamentally different approaches in different sites (*e.g.*, minirhizotrons, ingrowth cores, or sequential coring for $BNPP_{fine.root}$; root exclusion, stable isotope tracking, or gas exchange of excised roots for $R_{root}$), and sampling depth is variable and often insufficient to capture the full soil profile.
 $ANPP_{stem}$, which is also poorly explained by latitude or climate, is more straightforward to measure but is subject to variability introduced by differences such as biomass allometries applied and minimum plant size sampled (*Clark et al.
@@
 3-4).
-Indeed, MAT and MAP have long been identified as primary global-scale drivers of FACF [*Lieth 1973: doi 10.1007/BF01536729*; **REFS**; @taylor_temperature_2017].
+MAT and MAP have long been identified as primary global-scale drivers of FACF [*Lieth 1973: doi 10.1007/BF01536729*; **REFS**; @taylor_temperature_2017].
 It is not appropriate to attempt to identify individual mean annual climate variables as mechanistic drivers of FACF because many climate variables co-vary across the latitudinal gradient (Fig.
@@
 4f), but explained less variation than $MAT$.
-Dividing FACFs by growing season length to yield FACF per growing season month removed the majority of climate-related variation, supporting the idea that the latitudinal gradient in FACF is attributable more to shorter growing seasons at high latitudes than to inherently lower rates of photosyntheiss or respiration by high-latitude forests (*Enquist et al.
+Dividing FACFs by growing season length to yield FACF per growing season month removed the majority of climate-related variation, supporting the idea that the latitudinal gradient in FACF is attributable more to shorter growing seasons at high latitudes than to inherently lower rates of photosynthesis or respiration by high-latitude forests (*Enquist et al.
 2007: doi: 10.1111/j.1365-2486.2006.01222.x*).
@@
 2-3).
-This is consistent with observations of continental- to global-scale increases in $GPP$ [@li_mapping_2019] and $ANPP_{stem}$ [@brienen_long-term_2015; @hubau_asynchronous_2020], along with some C cycle components not considered here--tree mortality [@brienen_long-term_2015; @mcdowell_drivers_2018], soil respiration [@bond-lamberty_global_2010], and heterotrophic soil respiration [@bond-lamberty_globally_2018].
+This is consistent with observations of continental- to global-scale increases over time in $GPP$ [@li_mapping_2019] and $ANPP_{stem}$ [@brienen_long-term_2015; @hubau_asynchronous_2020], along with some C cycle components not considered here--tree mortality [@brienen_long-term_2015; @mcdowell_drivers_2018], soil respiration [@bond-lamberty_global_2010], and heterotrophic soil respiration [@bond-lamberty_globally_2018].
 However, increasing C flux rates are by no means universal [@rutishauser_testing_2020] (**MORE REFS**), likely because other factors are at play, including changes to other aspects of climate, atmospheric pollution (CO~2~, SO~2~, NO~x~), and local disturbances.
-Moreover, forest ecosystem responses to climatic changes outside the temperature range to which forest communities are adapated and acclimatized will not necessarily parallel responses across across geographic gradients in climate.
+**(discuss new Sullivan paper that finds higher max temperatures associated to lower ANPP_stem)**
+Moreover, forest ecosystem responses to climatic changes outside the temperature range to which forest communities are adapated and acclimatized will not necessarily parallel responses across geographic gradients in climate.
 Nevertheless, as we enter a period of accelerating climatic change, understanding of the fundamental climatic controls on FACF sets a foundation for understanding patterns of change.

### 2020-06-13 manuscript/appendix.Rmd
remove FACF abbreviation from appendix.Rmd
#74

@@
 S5 <- read.csv("tables_figures/mat_map.csv", stringsAsFactors = FALSE, check.names = FALSE)
-kable(S5, format = "latex", booktabs = TRUE, caption = "Table S3. Results of analysis of interactions effects between MAT and MAP, for all FACF", escape = F)%>%
+kable(S5, format = "latex", booktabs = TRUE, caption = "Table S3. Joint effects of MAT and MAP on forest C fluxes", escape = F)%>%
   kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 12) %>%
@@
 S2 <- read.csv("tables_figures/mat_growing_season.csv", stringsAsFactors = FALSE, check.names = FALSE)
-kable(S2, format = "latex", booktabs = TRUE, caption = "Table S4. Comparison of growing season length and mean annual temperature as predictors of FACF", escape = F)%>%
+kable(S2, format = "latex", booktabs = TRUE, caption = "Table S4. Comparison of growing season length and MAT as predictors of forest C fluxes", escape = F)%>%
   kable_styling(latex_options = c("scale_down", "hold_position")) %>%
@@
 \newpage
-```{r echo=FALSE, fig.cap = "Figure S1: Maps showing distribution of samples for the nine FACF analyzed here.", out.height ='95%', fig.pos='H'}
+```{r echo=FALSE, fig.cap = "Figure S1: Maps showing distribution of samples for the nine forest C fluxes analyzed here", out.height ='95%', fig.pos='H'}
 
@@
 \blandscape
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S2: Correlations among latitude and climate variables. Variable names and units given in Table S1", out.height='95%', fig.pos='H'}
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S2: Correlations among latitude and climate variables", out.height='95%', fig.pos='H'}
 
@@
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S3: Ratios among FACF as a function of latitude and climate variables", out.width ='80%', fig.pos='H'}
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S3: Ratios among forest C fluxes as a function of latitude and climate variables", out.width ='80%', fig.pos='H'}
 
@@
 \blandscape
-```{r echo=FALSE, fig.cap = "Figure S4: Individual plots of FACF in relation to mean annual climate, part 1.", out.height='95%', fig.pos='H'}
+```{r echo=FALSE, fig.cap = "Figure S4: Individual plots of forest C fluxes in relation to mean annual climate, part 1.", out.height='95%', fig.pos='H'}
 
@@
 
-```{r echo=FALSE, fig.cap = "Figure S5: Individual plots of FACF in relation to mean annual climate, part 2.", out.height='95%', fig.pos='H'}
+```{r echo=FALSE, fig.cap = "Figure S5: Individual plots of forest C fluxes in relation to mean annual climate, part 2.", out.height='95%', fig.pos='H'}
 
@@
 
-```{r echo=FALSE, fig.cap = "Figure S6: Individual plots of FACF in relation to mean climate seasonality, part 1.", out.height ='95%', fig.pos='H'}
+```{r echo=FALSE, fig.cap = "Figure S6: Individual plots of forest C fluxes in relation to mean climate seasonality, part 1.", out.height ='95%', fig.pos='H'}
 
@@
 \newpage
-```{r echo=FALSE, fig.cap = "Figure S7: Individual plots of FACF in relation to mean climate seasonality, part 2.", out.width ='100%', fig.pos='H'}
+```{r echo=FALSE, fig.cap = "Figure S7: Individual plots of forest C fluxes in relation to mean climate seasonality, part 2.", out.width ='100%', fig.pos='H'}
 
@@
 \newpage
-```{r echo=FALSE, fig.cap = "Figure S8: Growing season length-standardized FACF in relation to mean growing season climate, part 1.", out.width ='100%', fig.pos='H'}
+```{r echo=FALSE, fig.cap = "Figure S8: Growing season length-standardized forest C fluxes in relation to mean growing season climate, part 1.", out.width ='100%', fig.pos='H'}
 
@@
 \newpage
-```{r echo=FALSE, fig.cap = "Figure S9: Growing season length-standardized FACF in relation to mean growing season climate, part 2.", out.width ='100%', fig.pos='H'}
+```{r echo=FALSE, fig.cap = "Figure S9: Growing season length-standardized forest C fluxes in relation to mean growing season climate, part 2.", out.width ='100%', fig.pos='H'}
 

### 2020-06-13 manuscript/appendix.Rmd
working on HM comments
#82 , #96

@@
   column_spec(4, width = "7cm") %>%
-  column_spec(5, width = "5cm") %>%
+  column_spec(5, width = "2cm") %>%
+  column_spec(6, width = "5cm") %>%
   kableExtra::footnote(number = c("Hijmans et al. (2005)", "Fick et al. (2017)", "Harris et al. (2017)", "Trabucco and Zomer (2019)", "Abatzoglou et al. (2018)"), footnote_as_chunk = TRUE)

### 2020-06-14 manuscript/appendix.Rmd
worked through Helene's comments in appendix
all comments are now either dealt with or represented in issues.
closes #82

@@
 S3 <- read.csv("tables_figures/r2_table1.csv", stringsAsFactors = FALSE, check.names = FALSE)
-kable(S3, format = "latex", booktabs = TRUE, caption = "Table S2. Model details and R2 values for all climate variables tested") %>%
+kable(S3, format = "latex", booktabs = TRUE, caption = "Table S2.
+Model form, delta AIC, and R2 for each climate variables as a single fixed effect in models for each C flux.
+Model forms include first-order linear (Lin), second-order polynomial (Poly), and logarithmic (Log).") %>%
   kable_styling(latex_options = c("scale_down", "hold_position")) %>%

### 2020-06-19 scripts/pairwise_comparisons.R
Update pairwise_comparisons.R

@@ -115,7 +115,9 @@ all.results <- NULL
 all.results <- NULL
 all.aictab <- NULL
 all.koeppen <- NULL
-fixed.variables <- c("lat", "mat", "map", "PotentialEvapotranspiration", "VapourPressureDeficit", "TempSeasonality", "length_growing_season")
+# fixed.variables <- c("lat", "mat", "map", "PotentialEvapotranspiration", "VapourPressureDeficit", "TempSeasonality", "length_growing_season")
+fixed.variables <- c("lat", "mat", "TempSeasonality")
+
 
 response.variables <- c("GPP", "NPP", "ANPP", "ANPP_woody_stem", "ANPP_foliage", "BNPP_root", "BNPP_root_fine", "R_auto", "R_auto_root")
 response.variables.groups <- list(c("GPP", "NPP"),
@@ -132,7 +134,7 @@ number = 1
 
 
   for (age in ages){
-    all.comparisons <- data.frame(matrix(ncol = 0, nrow = 7))
+    all.comparisons <- NULL
     for (response.variables.group in response.variables.groups){
       
       
@@ -152,7 +154,7 @@ number = 1
       names(df)[60] <- paste(response.variables.group[2])
       fixed.comparison <- NULL
 
-      fixed.comparisons <- setNames(data.frame(matrix(ncol = 7, nrow = 1)), c("Climate variable", paste("Rsq", respv1), paste("Rsq", respv2), paste("Model type", respv1), paste("Model type", respv2), "Number of plots", "Variable with higher Rsq"))
+      fixed.comparisons <- setNames(data.frame(matrix(ncol = 9, nrow = 1)), c("C flux variable 1", "C flux variable 2", "Climate variable", "Rsq variable 1", "Rsq variable 2", "Model type variable 1", "Model type variable 2", "Number of plots", "Variable with higher Rsq"))
       
       
       for(fixed.v in fixed.variables){
@@ -188,7 +190,7 @@ number = 1
         par(mfrow = c(1,2), mar = c(2,2,2,2), oma = c(5,4,5,0))
         
         save.output <- setNames(data.frame(matrix(ncol = 2, nrow = 1)), c(paste(response.variables.group[1]), paste(response.variables.group[2])))
-        rep <- 2
+        rep <- 1
         
         for (response.v in response.variables.group){
           
@@ -299,9 +301,10 @@ number = 1
           Rsq <- signif(Rsq, digits=2)[1]
           
[…]

### 2020-06-19 scripts/final_model_no_weighting.R
Update final_model_no_weighting.R

@@ -653,10 +653,10 @@ for (age in ages){
       # equation <-  paste(response.v, "=", r[1], "+", fixed.v,  "x", r[2])
       
       significance <- anova(mod, mod.full)$"Pr(>Chisq)"[2]
-      significance <- signif(significance, digits=4)
+      significance <- signif(significance, digits=2)
       
       Rsq <- as.data.frame(r.squaredGLMM(mod.full))
-      Rsq <- signif(Rsq, digits=4)[1]
+      Rsq <- signif(Rsq, digits=2)[1]
       
       results <- data.frame(response = response.v, fixed = fixed.v, random = "geographic.area/plot.name", Age.filter = age, significant = significant.effect, p.value = significance, sample.size = sample.size, Rsq = Rsq)
       

### 2020-06-19 scripts/ratio_regression_script.R
Update ratio_regression_script.R

@@ -44,6 +44,7 @@ ForC_simplified[grepl("NEE", ForC_simplified$variable.name,ignore.case = F),]$me
 ForC_simplified$lat <- abs(ForC_simplified$lat)
 
 ForC_simplified$site_plot <- paste0(ForC_simplified$sites.sitename," ", ForC_simplified$plot.name)
+ForC_simplified$site_plot_year <- paste0(ForC_simplified$sites.sitename," ", ForC_simplified$plot.name, " ", ForC_simplified$stand.age)
 
 
 # Control for some factors ####
@@ -84,6 +85,9 @@ ForC_simplified$leaf.phenology <- ifelse(ForC_simplified$dominant.veg %in% everg
 ### exclude Tura due to extreme high latitude
 ForC_simplified <- ForC_simplified[!(ForC_simplified$sites.sitename %in% "Tura"),]
 
+ages.to.keep <- ForC_simplified$stand.age >= 100 & !is.na(ForC_simplified$stand.age)
+ForC_simplified <- ForC_simplified[ages.to.keep, ]
+
 # Prepare some variables ####
 
 ## response variable list (fluxes) ####
@@ -98,8 +102,11 @@ all.aictab <- NULL
 
 fixed.variables <- c("lat", "mat", "map", "TempSeasonality")
 
-set1 <- c("NPP", "ANPP", "ANPP_foliage", "ANPP", "BNPP_root")
-set2 <- c("GPP", "BNPP_root", "ANPP_woody_stem", "NPP", "NPP")
+# set1 <- c("NPP", "ANPP", "ANPP_foliage", "ANPP", "BNPP_root")
+# set2 <- c("GPP", "BNPP_root", "ANPP_woody_stem", "NPP", "NPP")
+
+set1 <- "ANPP"
+set2 <- "NPP"
 
 
 # set1 <- c("NPP", "ANPP", "ANPP_foliage", "ANPP_woody_stem", "BNPP_root", "BNPP_root_fine", "R_auto")
@@ -147,9 +154,8 @@ for (i in seq(along = set1)){
         print(fixed.v)
         
         fixed.no.na <- !is.na(df[, fixed.v]) & !is.na(df[, "masl"])
-        ages.to.keep <- df$stand.age >= 100 & !is.na(df$stand.age)
         
-        df <- df[fixed.no.na & ages.to.keep, ]
+        df <- df[fixed.no.na, ]
         
         df$fixed <- df[, fixed.v]
         

### 2020-06-19 manuscript/appendix.Rmd

@@
     fig_caption: yes
+
+bibliography: appendix_library.bib
+csl: global-change-biology.csl
+nocite: | 
+  @hijmans_very_2005; @fick_worldclim_2017; @harris_updated_2014;  @abatzoglou_terraclimate_2018
 
@@
   column_spec(6, width = "5cm") %>%
-  kableExtra::footnote(number = c("Hijmans et al. (2005)", "Fick et al. (2017)", "Harris et al. (2017)", "Trabucco and Zomer (2019)", "Abatzoglou et al. (2018)"), footnote_as_chunk = TRUE)
+  kableExtra::footnote(number = c("Hijmans et al. (2005)", "Fick et al. (2017)", "Harris et al. (2017)", "Abatzoglou et al. (2018)"), footnote_as_chunk = TRUE)
+
+#@hijmans_very_2005; @fick_worldclim_2017; @harris_updated_2014;  @abatzoglou_terraclimate_2018
 ```
@@
 ```
+
+\newpage
+# References

### 2020-06-19 scripts/ratio_regression_script.R
corrected ANPP + NPP in ratios, selecting only the most inclusive measurement for each

@@ -101,21 +101,22 @@ all.results <- NULL
 all.aictab <- NULL
 
 fixed.variables <- c("lat", "mat", "map", "TempSeasonality")
+# fixed.variables <- "lat"
 
-# set1 <- c("NPP", "ANPP", "ANPP_foliage", "ANPP", "BNPP_root")
-# set2 <- c("GPP", "BNPP_root", "ANPP_woody_stem", "NPP", "NPP")
+set1 <- c("NPP", "ANPP", "ANPP_foliage", "ANPP", "BNPP_root")
+set2 <- c("GPP", "BNPP_root", "ANPP_woody_stem", "NPP", "NPP")
 
-set1 <- "ANPP"
-set2 <- "NPP"
+# set1 <- "ANPP"
+# set2 <- "NPP"
 
 
 # set1 <- c("NPP", "ANPP", "ANPP_foliage", "ANPP_woody_stem", "BNPP_root", "BNPP_root_fine", "R_auto")
 # set2 <- c("GPP", "GPP", "GPP", "GPP", "GPP", "GPP", "GPP")
 
-png(file = paste0("C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/tables_figures/ratio_grid_plots.png"), width = 1800, height = 3000, units = "px", res = 300)
+png(file = paste0("C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/tables_figures/ratio_grid_plots.png"), width = 2200, height = 3000, units = "px", res = 300)
 
 
-par(mfrow = c(7,4), mar = c(2,2,2,2), oma = c(5,4,5,0))
+par(mfrow = c(5,4), mar = c(2,2,2,2), oma = c(5,4,5,0))
 panel.number <- 1
 for (i in seq(along = set1)){
   # for (j in seq(along = set2)){
@@ -133,9 +134,38 @@ for (i in seq(along = set1)){
       resp1 <- ForC_simplified[ForC_simplified$variable.name %in% responses.to.keep,]
       resp2 <- ForC_simplified[ForC_simplified$variable.name %in% responses.to.keep.2,]
       
+      if(set1[[i]] %in% "ANPP") resp1 <- resp1 %>%
+        mutate_at("variable.name", factor, levels = c("ANPP_0", "ANPP_1", "ANPP_2")) %>%
+        arrange(desc(variable.name)) %>%
+        group_by(site_plot_year) %>%
[…]

### 2020-06-22 manuscript/manuscript draft.Rmd
add ref

@@
 This is consistent with observations of continental- to global-scale increases over time in $GPP$ [@li_mapping_2019] and $ANPP_{stem}$ [@brienen_long-term_2015; @hubau_asynchronous_2020], along with some C cycle components not considered here--tree mortality [@brienen_long-term_2015; @mcdowell_drivers_2018], soil respiration [@bond-lamberty_global_2010], and heterotrophic soil respiration [@bond-lamberty_globally_2018].
-However, increasing C flux rates are by no means universal [@rutishauser_testing_2020] (**MORE REFS**), likely because other factors are at play, including changes to other aspects of climate, atmospheric pollution (CO~2~, SO~2~, NO~x~), and local disturbances.
+**DISCUSS/CITE [Wu et al.
+2020](https://doi.org/10.1111/gcb.15224)!** However, increasing C flux rates are by no means universal [@rutishauser_testing_2020] (**MORE REFS**), likely because other factors are at play, including changes to other aspects of climate, atmospheric pollution (CO~2~, SO~2~, NO~x~), and local disturbances.
 **(discuss new Sullivan paper that finds higher max temperatures associated to lower ANPP_stem)**

### 2020-07-03 manuscript/appendix.Rmd
updated table s2 in appendix

@@
   kable_styling(latex_options = c("scale_down", "hold_position")) %>%
-  add_header_above(c(" ", "Latitude"=2, "MAT"=2, "MAP"=2, "T Seas"=2, "P Seas"=2, "ATR"=2, "Solar R"=2, "AI"=2))
+  add_header_above(c(" ", "Latitude"=3, "MAT"=3, "MAP"=3, "T Seas"=3, "P Seas"=3, "ATR"=3, "Solar R"=3, "AI"=3))
 ```
@@
   kable_styling(latex_options = c("scale_down", "hold_position")) %>%
-  add_header_above(c(" ", "Cloud"=2, "AFD"=2, "AWD"=2, "PET"=2, "VPD"=2, "Max VPD"=2, "WSM"=2, "LGS"=2))
+  add_header_above(c(" ", "Cloud"=3, "AFD"=3, "AWD"=3, "PET"=3, "VPD"=3, "Max VPD"=3, "WSM"=3, "LGS"=3))
 ```

### 2020-07-03 scripts/AIC_comparisons_of_climate_variables.R
Update AIC_comparisons_of_climate_variables.R

@@ -116,7 +116,7 @@ all.results <- NULL
 all.aictab <- NULL
 all.koeppen <- NULL
 
-fixed.variables <- c("mat", "map", "PotentialEvapotranspiration", "VapourPressureDeficit", "SolarRadiation", "TempSeasonality", "PreSeasonality", "length_growing_season")
+fixed.variables <- c("mat", "map", "PotentialEvapotranspiration", "VapourPressureDeficit", "SolarRadiation", "TempSeasonality", "PreSeasonality", "length_growing_season", "TempRangeAnnual", "Aridity", "CloudCover", "AnnualFrostDays", "AnnualWetDays", "MaxVPD", "WaterStressMonths")
 
 
 response.variables.groups <- list(c("GPP", "NPP", "BNPP_root", "BNPP_root_fine"),
@@ -152,6 +152,8 @@ for(response.variables in response.variables.groups){
         df <- ForC_simplified[rows.with.response, ]
         
         df$mat <- df$mat + 11
+        df$AnnualFrostDays <- df$AnnualFrostDays + 0.1
+        df$WaterStressMonths <- df$WaterStressMonths + 0.1
         
         for(fixed.v in fixed.variables){
           ages.to.keep <- df$stand.age >= 100 & !is.na(df$stand.age)

### 2020-07-03 scripts/create_table_s2.R
Create create_table_s2.R

@@ -1,0 +1,23 @@
+results <- read.csv("C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/tables/best_model_outputs/best_model_scaled_unweighted.csv", stringsAsFactors = FALSE)
+response <- c("GPP", "NPP", "ANPP", "ANPP_woody_stem", "ANPP_foliage", "BNPP_root", "BNPP_root_fine", "R_auto", "R_auto_root")
+
+fixed.v <- c("lat", "mat", "map", "TempSeasonality", "PreSeasonality", "TempRangeAnnual", "SolarRadiation", "Aridity", "CloudCover", "AnnualFrostDays", "AnnualWetDays", "PotentialEvapotranspiration", "VapourPressureDeficit", "MaxVPD", "WaterStressMonths", "length_growing_season")
+
+table <- as.data.frame(response)
+
+all.results <- NULL
+
+for(v in fixed.v){
+rs <- results[results$fixed %in% v,]
+
+test <- merge(rs[, c("response", "fixed", "best.model", "Rsq.R2m", "delta.aic", "significant")], table, by = "response")
+
+all.results <- rbind(all.results, test)
+}
+  
+all.results$best.model[all.results$best.model %in% "mod.clim"] <- "Lin"
+all.results$best.model[all.results$best.model %in% "mod"] <- "Null"
+all.results$best.model[all.results$best.model %in% "mod.clim.poly"] <- "Poly"
+all.results$best.model[all.results$best.model %in% "mod.clim.log"] <- "Log"
+
+write.csv(all.results, "C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/tables/best_model_outputs/prep_table_s2.csv")

### 2020-07-06 manuscript/manuscript draft.Rmd
updated manuscript draft with new references

--- 2020-06-13 manuscript/manuscript draft.Rmd
@@
 
-Carbon cycling in forests worldwide provides the energetic basis for sustaining the majority of Earth's terrestrial biodiversity and many human populations (**[MEA 2005](http://www.millenniumassessment.org/documents/document.354.aspx.pdf)**; **REFS**), while strongly influencing atmospheric CO~2~ and climate [@bonan_forests_2008].
-Forests' autotrophic carbon fluxes (FACF)--that is, carbon fixation, allocation, and metabolism by trees and other primary producers--sets the energy ultimately available to heterotrophic organisms (including microbes), in turn influencing their abundance (e.g., **[Zak et al.
-1994](http://esajournals.onlinelibrary.wiley.com/doi/abs/10.2307/1940888)**; [**Niedziałkowska et al.
-2010](https://doi.org/10.2980/17-1-3310)**) and possibly diversity [@waide_relationship_1999; **Chu et al.
-2019 (DOI: 10.1111/ele.13175)**].
+Carbon cycling in forests worldwide provides the energetic basis for sustaining the majority of Earth's terrestrial biodiversity and many human populations [@millennium_ecosystem_assessment_ecosystems_2005], while strongly influencing atmospheric CO~2~ and climate [@bonan_forests_2008].
+Forests' autotrophic carbon fluxes (FACF)--that is, carbon fixation, allocation, and metabolism by trees and other primary producers--sets the energy ultimately available to heterotrophic organisms (including microbes), in turn influencing their abundance [@zak_plant_1994; @niedzialkowska_species_2010] and possibly diversity [@waide_relationship_1999; @chu_direct_2018].
 FACF influences all organic matter stocks in forest ecosystems and is linked to cycling of energy, water, and nutrients.
@@
 [-@taylor_temperature_2017] showed a positive interaction between MAT and MAP in shaping tropical forest productivity, such that high rainfall had a negative effect on productivity in cooler climates, compared to a positive effect in warmer climates.
-Such complicated dynamics play out with belowground fluxes as well, with typically dominant soil temperature mediated by moisture and carbon supply (**Hursh et al.
-2017, http://dx.doi.org/10.1111/gcb.13489; Xu et al.
-2016 http://dx.doi.org/10.1016/j.jplph.2016.08.007**).
+Such complicated dynamics play out with belowground fluxes as well, with typically dominant soil temperature mediated by moisture and carbon supply [@hursh_sensitivity_2017; @xu_contribution_2016].
 
@@
 This analysis focused on mature forests with no known history of signficant disturbance or management. 
-There is evidence that stand age influences patterns of FACF and carbon allocation in forest ecosystems, and can confound relationships between latitude and primary productivity [@delucia_forest_2007; @gillman_latitude_2015].
+There is evidence that stand age influences patterns of FACF and carbon allocation in forest ecosystems, and can confound relationships between latitude and primary productivity [@de_lucia_forest_2007; @gillman_latitude_2015].
 To reduce any biasing effects of stand age, we included only stands of known age $\ge$ 100 years and those described by terms such as "mature", "intact", or "old-growth".
@@
 Compared to tropical forests, the temperate forest biome has experienced more widespread anthropogenic disturbance and has a larger fraction of secondary stands [@potapov_mapping_2008; @poulter_global_2018], so analyses comparing across latitudinal gradients without controlling for stand age risk confounding age with biome effects.
-In addition, because carbon allocation varies with stand age [@delucia_forest_2007] (**Doughty et al.
-2018: DOI: 10.1111/btp.12504; Anderson-Teixeira et al.
-2013: doi: 10.1111/gcb.12194**), age differences may introduce systematic biases into analyses of FACF across latitude or global climatic gradients.
-For example, woody productivity tends to be higher in rapidly aggrading secondary stands than in old-growth forests, where proportionally more C is allocated to respiration (**Doughty et al.
-2018: DOI: 10.1111/btp.12504**);[@delucia_forest_2007; @piao_forest_2010; @kunert_understanding_2019].
+In addition, because carbon allocation varies with stand age [@de_lucia_forest_2007; @anderson-teixeira_altered_2013; @doughty_what_2018], age differences may introduce systematic biases into analyses of FACF across latitude or global climatic gradients.
+For example, woody productivity tends to be higher in rapidly aggrading secondary stands than in old-growth forests, where proportionally more C is allocated to respiration [@de_lucia_forest_2007; @piao_forest_2010; @doughty_what_2018; @kunert_understanding_2019].
 
@@
 and with observations and theory predicting relatively higher belowground carbon allocation at higher latitudes [@gill_belowground_2016].
-It is also consistent with findings that as forest productivity increases, proportionaly more carbon is allocated to $ANPP_{foliage}$ relative to $BNPP_{fine.root}$ (*Chen et al.
-2019; doi: 10.1111/ele.13193*) or $ANPP_{stem}$ [@hofhansl_new_2015].
+It is also consistent with findings that as forest productivity increases, proportionaly more carbon is allocated to $ANPP_{foliage}$ relative to $BNPP_{fine.root}$ [@chen_allometry_2019] or $ANPP_{stem}$ [@hofhansl_new_2015].
 **( check / comment on  @litton_carbon_2007 )**
 
-One previously hypothesized trend that was not supported by our analysis was that tropical forests tend to have low carbon use efficiency ($CUE$= $NPP$/$GPP$=($GPP$-$R_{auto}$/$GPP$), which is based on observations of low $CUE$ in old-growth tropical forests relative to (mostly younger) extratropical forests [@delucia_forest_2007; @malhi_productivity_2012; @anderson-teixeira_carbon_2016].
+One previously hypothesized trend that was not supported by our analysis was that tropical forests tend to have low carbon use efficiency ($CUE$= $NPP$/$GPP$=($GPP$-$R_{auto}$/$GPP$), which is based on observations of low $CUE$ in old-growth tropical forests relative to (mostly younger) extratropical forests [@de_lucia_forest_2007; @malhi_productivity_2012; @anderson-teixeira_carbon_2016].
 Our analysis, limited to mature forests, shows no such trend (Fig.
 S3).
-$CUE$ is known to decline with forest age [@delucia_forest_2007; @piao_forest_2010; @collalti_is_2019], but appears to be roughly independent of $GPP$ [@litton_carbon_2007].
+$CUE$ is known to decline with forest age [@de_lucia_forest_2007; @piao_forest_2010; @collalti_is_2019], but appears to be roughly independent of $GPP$ [@litton_carbon_2007].
 (*This finding may have some important implications for modeling and our broader understanding.
@@
 Belowground fluxes in particular are difficult to quantify, and measurement methods for the belowground fluxes considered here may use fundamentally different approaches in different sites (*e.g.*, minirhizotrons, ingrowth cores, or sequential coring for $BNPP_{fine.root}$; root exclusion, stable isotope tracking, or gas exchange of excised roots for $R_{root}$), and sampling depth is variable and often insufficient to capture the full soil profile.
-$ANPP_{stem}$, which is also poorly explained by latitude or climate, is more straightforward to measure but is subject to variability introduced by differences such as biomass allometries applied and minimum plant size sampled (*Clark et al.
-2001 Ecological Applications*).
+$ANPP_{stem}$, which is also poorly explained by latitude or climate, is more straightforward to measure but is subject to variability introduced by differences such as biomass allometries applied and minimum plant size sampled [@clark_measuring_2001].
 However, methodological variation and uncertainty affect all of fluxes considered here, and some of the larger fluxes that vary more strongly with respect to climate ($ANPP$, $NPP$) are estimated by summing uncertain component fluxes.
@@
 3-4).
-MAT and MAP have long been identified as primary global-scale drivers of FACF [*Lieth 1973: doi 10.1007/BF01536729*; **REFS**; @taylor_temperature_2017].
+MAT and MAP have long been identified as primary global-scale drivers of FACF [@lieth_primary_1973; **REFS**; @taylor_temperature_2017].
 It is not appropriate to attempt to identify individual mean annual climate variables as mechanistic drivers of FACF because many climate variables co-vary across the latitudinal gradient (Fig.
@@
 4e), and are minimal during cold- or dry- dormant seasons.
-To account for this, a number of analyses seeking to characterize global-scale effects of climate on productivity have examined the relationship of C flux per month of the growing season with growing season climatic conditions [Table 1, *H5*; @kerkhoff_plant_2005; *Anderson et al.
-2006: doi:10.1111/j.1461-0248.2006.00914.x *; *Enquist et al.
-2007: doi: 10.1111/j.1365-2486.2006.01222.x*;  @michaletz_convergence_2014].
+To account for this, a number of analyses seeking to characterize global-scale effects of climate on productivity have examined the relationship of C flux per month of the growing season with growing season climatic conditions [Table 1, *H5*; @kerkhoff_plant_2005; @anderson_temperature-dependence_2006; @enquist_adaptive_2007; @michaletz_convergence_2014].
 We found that the sort of simple metric needed to define growing season at a global scale [@kerkhoff_plant_2005] was rough **BBL: ?** with respect to temperature and problematic with respect to moisture because it doesn't capture temporal lags between precipitation and plant water use caused by storage in soil or snow.
@@
 4f), but explained less variation than $MAT$.
-Dividing FACFs by growing season length to yield FACF per growing season month removed the majority of climate-related variation, supporting the idea that the latitudinal gradient in FACF is attributable more to shorter growing seasons at high latitudes than to inherently lower rates of photosynthesis or respiration by high-latitude forests (*Enquist et al.
-2007: doi: 10.1111/j.1365-2486.2006.01222.x*).
+Dividing FACFs by growing season length to yield FACF per growing season month removed the majority of climate-related variation, supporting the idea that the latitudinal gradient in FACF is attributable more to shorter growing seasons at high latitudes than to inherently lower rates of photosynthesis or respiration by high-latitude forests [@enquist_adaptive_2007].
 However, there remained a number of significant correlations with growing season climatic conditions, suggesting that climatic conditions remain influential within the growing season.
 We conclude that while correcting for growing season length takes analyses a step closer to mechanistic linkage of instantaneous C flux rates to environmental conditions, it remains crude relative to the timescales on which climate affects plant metabolism, and does not advance statistical predictive power.
-Mechanistic accounting for climatic effects on global FACF patterns instead requires models representing physiologically meaningful timescales (*e.g.*, **refs**).
-**BBL: maybe cite e.g.
-ED2?
-http://dx.doi.org/10.5194/gmd-12-4309-2019 **
+Mechanistic accounting for climatic effects on global FACF patterns instead requires models representing physiologically meaningful timescales (*e.g.*, **refs**) [@longo_biophysics_2019]
 
@@
 This is consistent with observations of continental- to global-scale increases over time in $GPP$ [@li_mapping_2019] and $ANPP_{stem}$ [@brienen_long-term_2015; @hubau_asynchronous_2020], along with some C cycle components not considered here--tree mortality [@brienen_long-term_2015; @mcdowell_drivers_2018], soil respiration [@bond-lamberty_global_2010], and heterotrophic soil respiration [@bond-lamberty_globally_2018].
-However, increasing C flux rates are by no means universal [@rutishauser_testing_2020] (**MORE REFS**), likely because other factors are at play, including changes to other aspects of climate, atmospheric pollution (CO~2~, SO~2~, NO~x~), and local disturbances.
+**DISCUSS/CITE [Wu et al.
+2020]** [@wu_accelerated_2020] However, increasing C flux rates are by no means universal [@rutishauser_testing_2020] (**MORE REFS**), likely because other factors are at play, including changes to other aspects of climate, atmospheric pollution (CO~2~, SO~2~, NO~x~), and local disturbances.
 **(discuss new Sullivan paper that finds higher max temperatures associated to lower ANPP_stem)**

### 2020-07-06 manuscript/appendix.Rmd
updated appendix

@@
 \newpage
+```{r eval = TRUE, echo=FALSE, warning=FALSE}
+library(knitr)
+library(kableExtra)
+S3 <- read.csv("tables_figures/best_models_each_flux.csv", stringsAsFactors = FALSE, check.names = FALSE)
+kable(S3, format = "latex", booktabs = TRUE, caption = "Table S5. Best models by carbon flux. Models where delta AIC </= 2 are presented", escape = F) %>%
+  kable_styling(latex_options = c("scale_down", "hold_position")) %>%
+  collapse_rows(columns = 1)
+
+```
+
+\newpage
 ```{r echo=FALSE, fig.cap = "Figure S1: Maps showing distribution of samples for the nine forest C fluxes analyzed here", out.height ='95%', fig.pos='H'}

### 2020-07-06 scripts/AIC_comparisons_of_climate_variables.R
Update AIC_comparisons_of_climate_variables.R

@@ -115,13 +115,16 @@ all.results <- NULL
 all.results <- NULL
 all.aictab <- NULL
 all.koeppen <- NULL
+aictab.top <- NULL
 
 fixed.variables <- c("mat", "map", "PotentialEvapotranspiration", "VapourPressureDeficit", "SolarRadiation", "TempSeasonality", "PreSeasonality", "length_growing_season", "TempRangeAnnual", "Aridity", "CloudCover", "AnnualFrostDays", "AnnualWetDays", "MaxVPD", "WaterStressMonths")
 
 
-response.variables.groups <- list(c("GPP", "NPP", "BNPP_root", "BNPP_root_fine"),
-                                  c("ANPP", "ANPP_foliage", "ANPP_woody_stem"),
-                                  c("R_auto", "R_auto_root"))
+# response.variables.groups <- list(c("GPP", "NPP", "BNPP_root", "BNPP_root_fine"),
+#                                   c("ANPP", "ANPP_foliage", "ANPP_woody_stem"),
+#                                   c("R_auto", "R_auto_root"))
+
+response.variables.groups <- list(c("NPP"))
 
 for(response.variables in response.variables.groups){
   
@@ -249,6 +252,11 @@ for(response.variables in response.variables.groups){
           return(fit1)
         })
        aictab.list <- aictab(all_models, sort = T, modnames = model_list)
+       aictab.best <- aictab.list[aictab.list$Delta_AICc <= 2,]
+       aictab.best <- aictab.best[, c(1,4)]
+       aictab.best$flux <- paste(response.v)
+       aictab.top <- rbind(aictab.top, aictab.best)
+
       write.csv(aictab.list, file = paste0("C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/tables/best_model_outputs/best_models_AIC/", response.v, "_AIC.csv"))
       }
       
@@ -261,3 +269,4 @@ for(response.variables in response.variables.groups){
 }
 
 
+write.csv(aictab.top, file = "C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/tables/best_model_outputs/best_models_AIC/top_aic.csv")

### 2020-07-07 scripts/pairwise_comparisons.R
Update pairwise_comparisons.R

@@ -169,7 +169,7 @@ number = 1
         
         fixed.v.info <- read.csv("C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/raw.data/fixedv_data.csv", stringsAsFactors = F)
         
-        xaxis <- fixed.v.info$xaxis_simple[which(fixed.v.info$fixed.v %in% fixed.v)]
+        xaxis <- fixed.v.info$abbrev[which(fixed.v.info$fixed.v %in% fixed.v)]
         
         fixed.no.na <- !is.na(df[, fixed.v]) & !is.na(df[, "masl"])
         if (age %in% "age.greater.than.100") ages.to.keep <- df$stand.age >= 100 & !is.na(df$stand.age)
@@ -251,8 +251,8 @@ number = 1
           if (best.model == "mod.clim.log") mod.full <- lmer(mean ~ log(fixed) + (1|geographic.area/plot.name), data = df, REML = T)
           
           if (best.model == "mod") model <- "Null"
-          if (best.model == "mod.clim") model <- "Linear"
-          if (best.model == "mod.clim.log") model <- "Logarithmic"
+          if (best.model == "mod.clim") model <- "Lin"
+          if (best.model == "mod.clim.log") model <- "Log"
           
           # }
           
@@ -344,3 +344,4 @@ number = 1
     }}
 
 write.csv(all.comparisons,  file = "C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/tables/best_model_outputs/pairwise_comparisons.csv", row.names = F)
+write.csv(all.comparisons,  file = "C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/tables_figures/pairwise_comparisons.csv", row.names = F)

### 2020-07-07 manuscript/appendix.Rmd
updated SI tables

@@
 Model form, delta AIC, and R2 for each climate variables as a single fixed effect in models for each C flux.
-Model forms include first-order linear (Lin), second-order polynomial (Poly), and logarithmic (Log).") %>%
+Model forms include first-order linear (Lin), second-order polynomial (Poly), and logarithmic (Log).", caption.short = "Table S2.
+Model form, delta AIC and R2 for each climate variables as a single fixed effect in models for each C flux") %>%
   kable_styling(latex_options = c("scale_down", "hold_position")) %>%
@@
 S3 <- read.csv("tables_figures/best_models_each_flux.csv", stringsAsFactors = FALSE, check.names = FALSE)
-kable(S3, format = "latex", booktabs = TRUE, caption = "Table S5. Best models by carbon flux. Models where delta AIC </= 2 are presented", escape = F) %>%
+kable(S3, format = "latex", booktabs = TRUE, caption = "Table S5.
+Best models by carbon flux.
+Models where delta AIC </= 2 are presented", caption.short = "Table S5.
+Best models by carbon flux", escape = F) %>%
   kable_styling(latex_options = c("scale_down", "hold_position")) %>%
-  collapse_rows(columns = 1)
+  collapse_rows(columns = 1, latex_hline = "major")
+```
+
+\newpage
+\blandscape
+```{r eval = TRUE, echo=FALSE, warning=FALSE}
+library(knitr)
+library(kableExtra)
+S4 <- read.csv("tables_figures/pairwise_comparisons.csv", stringsAsFactors = FALSE, check.names = FALSE)
+kable(S4, format = "latex", booktabs = TRUE, caption = "Table S6. Pairwise comparisons between carbon fluxes", escape = F) %>%
+  kable_styling(latex_options = c("scale_down", "hold_position")) %>%
+  collapse_rows(columns = c(1,2), latex_hline = "custom", custom_latex_hline = c(1,2))
 
 ```
+\elandscape
 

### 2020-07-25 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd
mostly issue #79

@@
 
-Length of the growing season was estimated to the nearest month, where growing season months were defined as months with mean minimum temperature > 0.5$^\circ$C.
-We experimented with a definion of growing season months including a moisture index, defined as (MAT - PET)/PET, > -0.95 [@kerkhoff_plant_2005; see also @michaletz_convergence_2014].
+**For consistency with previous studies (Table 1, *H5*),** length of the growing season was estimated to the nearest month, where growing season months were defined as months with mean minimum temperature > 0.5$^\circ$C.
+We experimented with a definition of growing season months including a moisture index, defined as (MAT - PET)/PET, > -0.95 [@kerkhoff_plant_2005; see also @michaletz_convergence_2014].
 However, we found that including a moisture index had minimal effect on the estimates of growing season length, and so chose to exclude it.
-Monthly data for PET, precipitation, and temperature from the CRU dataset v 4.03 [@harris_updated_2014], and solar radiation from WorldClim2 [@fick_worldclim_2017] were used to calculate mean monthly PET, precipitation, temperature and solar radiation during the growing season.
+Monthly data for PET, precipitation, and temperature from  CRU v 4.03 [@harris_updated_2014] and solar radiation from WorldClim2 [@fick_worldclim_2017] were used to calculate mean monthly PET, precipitation, temperature and solar radiation during the growing season.
 Total growing season precipitation and solar radiation were also calculated.
@@
 S8-S9).
-Speficifally, $ANPP$ increased with growing season temperature ($R^2$ = 0.10, p<0.001) and precipitation ($R^2$ = 0.04, p<0.05).
+Specifically, $ANPP$ increased with growing season temperature ($R^2$ = 0.10, p<0.001) and precipitation ($R^2$ = 0.04, p<0.05).
 Similarly, $ANPP_{foliage}$ increased slightly with growing season temperature ($R^2$ = 0.16, p<0.01) and precipitation ($R^2$ = 0.09, p<0.05).
@@
 
-FACF are negatively correlated with temperature seasonality (Table 1, *H4*; Fig.
+Forest autotrophic C fluxes decline with temperature seasonality (Table 1, *H4*; Fig.
 4e), and are minimal during cold- or dry- dormant seasons.
 To account for this, a number of analyses seeking to characterize global-scale effects of climate on productivity have examined the relationship of C flux per month of the growing season with growing season climatic conditions [Table 1, *H5*; @kerkhoff_plant_2005; @anderson_temperature-dependence_2006; @enquist_adaptive_2007; @michaletz_convergence_2014].
-We found that the sort of simple metric needed to define growing season at a global scale [@kerkhoff_plant_2005] was rough **BBL: ?** with respect to temperature and problematic with respect to moisture because it doesn't capture temporal lags between precipitation and plant water use caused by storage in soil or snow.
-A temperature-defined growing season length had stong positive correlation with FACF (Fig.
+The sort of simple metric needed to define growing season at a global scale [@kerkhoff_plant_2005] is coarse with respect to temperature because it's calculated on a monthly timescale and problematic with respect to moisture because it doesn't capture temporal lags between precipitation and plant water use caused by storage in soil or snow.
+We found that a temperature-defined growing season length had strong positive correlation with C fluxes (Fig.
 4f), but explained less variation than $MAT$.
-Dividing FACFs by growing season length to yield FACF per growing season month removed the majority of climate-related variation, supporting the idea that the latitudinal gradient in FACF is attributable more to shorter growing seasons at high latitudes than to inherently lower rates of photosynthesis or respiration by high-latitude forests [@enquist_adaptive_2007].
-However, there remained a number of significant correlations with growing season climatic conditions, suggesting that climatic conditions remain influential within the growing season.
+Dividing annual fluxes by growing season length to yield average flux per growing season month removed the majority of climate-related variation, supporting the idea that the latitudinal gradient in carbon flux is attributable more to shorter growing seasons at high latitudes than to inherently lower rates of photosynthesis or respiration by high-latitude forests [@enquist_adaptive_2007].
+However, there remained a number of significant correlations with growing season climatic conditions, indicating that climatic conditions remain influential within the growing season.
 We conclude that while correcting for growing season length takes analyses a step closer to mechanistic linkage of instantaneous C flux rates to environmental conditions, it remains crude relative to the timescales on which climate affects plant metabolism, and does not advance statistical predictive power.
-Mechanistic accounting for climatic effects on global FACF patterns instead requires models representing physiologically meaningful timescales (*e.g.*, **refs**) [@longo_biophysics_2019]
+Mechanistic accounting for climatic effects on global forest carbon flux patterns instead requires models representing physiologically meaningful timescales [e.g., **REFS**; @longo_biophysics_2019].
 
-Our analysis clarifies how FACF vary with latitude and climate on a global scale, with some important implications for how forest carbon cycling relates to climate and, by extension, how it is likely to respond to climatic warming.
+Our analysis clarifies how forest autotrophic carbon fluxes vary with latitude and climate on a global scale, with some important implications for how forest carbon cycling relates to climate and, by extension, how it is likely to respond to climatic warming.
 Our findings show that higher temperatures with similar moisture availability result in a generalized acceleration of FACF (Figs.

### 2020-08-02 scripts/correlation_matrix_code.R
Update correlation_matrix_code.R

@@ -57,8 +57,8 @@ for (i in seq(along = variables1)){
   df$var2 <- df[, variables2[[j]]]
   # 
   fixed.v.info <- read.csv("C:/Users/becky/Dropbox (Smithsonian)/Github/Global_Productivity/raw.data/fixedv_data.csv", stringsAsFactors = F)
-  xaxis <- fixed.v.info$abbrev[which(fixed.v.info$fixed.v %in% variables2[[j]])]
-  yaxis <- fixed.v.info$abbrev[which(fixed.v.info$fixed.v %in% variables1[[i]])]
+  xaxis <- fixed.v.info$xaxis_simple[which(fixed.v.info$fixed.v %in% variables2[[j]])]
+  yaxis <- fixed.v.info$xaxis_simple[which(fixed.v.info$fixed.v %in% variables1[[i]])]
   # 
   if(i!=j) fit <- lm(var1 ~ var2 , data = df)
   if(i!=j) significant.effect <- summary(fit)$coefficients[2,4] < 0.05
@@ -79,8 +79,8 @@ for (i in seq(along = variables1)){
   # # if(panel.number == c(43:49))
   # if (i < j) mtext(text = legend, side = 3, adj = 0.5, line = -5, cex = 1)
   if (i > j) abline(fit, lty = ifelse(significant.effect, 1, 2))
-  if (panel.number %in% c(1, 8, 15, 22, 29, 36, 43)) mtext(side = 2, line = 2.5, text = yaxis, outer = F, cex = 0.75)
-  if (panel.number %in% c(43:49)) mtext(side = 1, line = 2.5,  text = xaxis, outer = F, cex = 0.75)
+  if (panel.number %in% c(1, 8, 15, 22, 29, 36, 43)) mtext(side = 2, line = 2.5, text = eval(parse(text = yaxis)), outer = F, cex = 0.75)
+  if (panel.number %in% c(43:49)) mtext(side = 1, line = 2.5,  text = eval(parse(text = xaxis)), outer = F, cex = 0.75)
   # 
   panel.number <- panel.number +1
   

### 2020-08-02 scripts/correlation_matrix_code.R
corrections

@@ -66,7 +66,7 @@ for (i in seq(along = variables1)){
   ifelse(i==j, NA, 
          ifelse(significant.effect == FALSE, print("significant effect false"), NA))
   if(i!=j) Rsq <- as.data.frame(r.squaredGLMM(fit))
-  if(i!=j) Rsq <- signif(Rsq, digits=2)
+  if(i!=j) Rsq <- round(Rsq, digits=2)
   if(i!=j) legend <- Rsq[1]
   # # significant.effect <- anova(mod, mod.full)$"Pr(>Chisq)"[2] < 0.05
   # 

### 2020-08-02 manuscript/appendix.Rmd

@@
   pack_rows("ANPP", 7, 9, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  pack_rows("ANPP woody stem", 10, 12, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("ANPP stem", 10, 12, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
   pack_rows("ANPP foliage", 13, 15, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
@@
   pack_rows("BNPP fine root", 19, 21, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  pack_rows("Autotrophic respiration", 22, 24, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  pack_rows("Root respiration", 25, 27, latex_gap_space = "1em", colnum = 1, hline_before = TRUE)
+  pack_rows("R auto", 22, 24, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
+  pack_rows("R root", 25, 27, latex_gap_space = "1em", colnum = 1, hline_before = TRUE)
 ```

### 2020-08-02 manuscript/manuscript draft.Rmd

@@
 ```{r setup, include = FALSE}
-knitr::opts_knit$set(root.dir = 'C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/')
-# knitr::opts_knit$set(root.dir = '/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/Global_Productivity/manuscript/')
+# knitr::opts_knit$set(root.dir = 'C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/')
+ knitr::opts_knit$set(root.dir = '/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/Global_Productivity/manuscript/')
 ```
@@
 However, we lack a cohesive synthesis on how forest carbon fluxes vary globally with respect to climate and one another. 
-Here, we draw upon # records from the Global Forest Carbon Database (ForC), representing all major forest types and the nine most significant autotrophic carbon fluxes, to comprehensively explore how C cycling in mature, undisturbed forests varies with latitude and climate on a global scale.
+Here, we draw upon **#** records from the Global Forest Carbon Database (ForC), representing all major forest types and the nine most significant autotrophic carbon fluxes, to comprehensively explore how C cycling in mature, undisturbed forests varies with latitude and climate on a global scale.
 We show that, across all flux variables analyzed, C cycling decreases linearly with absolute latitude -- a finding that confirms multiple previous studies but contradicts the idea that net primary productivity ($NPP$) of temperate forests rivals that of tropical forests.
@@
 These findings clarify how forest C flux varies with latitude and climate on a global scale. 
-As we enter a period of accelerating climatic change, understanding of the fundamental climatic controls on forest C cycling sets a foundation for understanding patterns of change.
+In a period of accelerating climatic change, understanding of the fundamental climatic controls on forest C cycling sets a foundation for understanding patterns of change.
 
@@
 MAT and MAP are very coarse measures of climate, and so fail to capture much variation in climate on an intra-annual scale, including the effects of factors such as growing season length, number of frost-free days, temperature seasonality, and dry season length.
-Some studies have suggested that the apparently strong relationship between MAT and FACFs is actually a consequence of the correlation between MAT and growing season length [@kerkhoff_plant_2005; @malhi_productivity_2012; @michaletz_convergence_2014; @michaletz_drivers_2018].
+Some studies have suggested that the apparently strong relationship between MAT and FACFs is actually a consequence of the correlation between MAT and growing season length [@kerkhoff_plant_2005; @michaletz_convergence_2014; @michaletz_drivers_2018].
 Kerkhoff et al.
@@
   kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 12) %>%
-  pack_rows("Q1. How do FACF vary with latitude?", 1, 1, latex_gap_space = "1em", colnum = 5, hline_before = FALSE) %>%
-  pack_rows("Q2. How do FACF vary with MAT and MAP?", 2, 4, latex_gap_space = "1em", colnum = 5, hline_before = TRUE) %>%
-  pack_rows("Q3. How are FACF related to other climate variables?", 5, 7, latex_gap_space = "1em", colnum = 5, hline_before = TRUE) %>%
-  pack_rows("Q4. How does seasonality influence FACF?", 8, 11, latex_gap_space = "1em", colnum = 5, hline_before = TRUE) %>%
-  pack_rows("Q5. When standardised by growing season length, how do FACF vary with climate?", 12, 15, latex_gap_space = "1em", colnum = 5, hline_before = TRUE) %>%
-  add_header_above(c(" ",  " ", "Forest autotrophic carbon fluxes (FACF)" = 9, " ")) %>%
+  pack_rows("Q1. How do C fluxes vary with latitude?", 1, 1, latex_gap_space = "1em", colnum = 5, hline_before = FALSE) %>%
+  pack_rows("Q2. How do C fluxes vary with MAT and MAP?", 2, 4, latex_gap_space = "1em", colnum = 5, hline_before = TRUE) %>%
+  pack_rows("Q3. How are C fluxes related to other climate variables?", 5, 7, latex_gap_space = "1em", colnum = 5, hline_before = TRUE) %>%
+  pack_rows("Q4. How does seasonality influence C fluxes?", 8, 11, latex_gap_space = "1em", colnum = 5, hline_before = TRUE) %>%
+  pack_rows("Q5. When standardised by growing season length, how do C fluxes vary with climate?", 12, 15, latex_gap_space = "1em", colnum = 5, hline_before = TRUE) %>%
+  add_header_above(c(" ",  " ", "Forest autotrophic carbon fluxes" = 9, " ")) %>%
   #column_spec(2:11, width = "1.8cm") %>%

### 2020-08-02 manuscript/appendix.Rmd
reknit appendix

@@
 \blandscape
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S2: Correlations among latitude and climate variables", out.height='95%', fig.pos='H'}
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S2: Scatterplots and Pearson's R values for relationships between latitude and climate variables", out.height='95%', fig.pos='H'}
 

### 2020-08-02 scripts/correlation_matrix_code.R
corrected temp seasonality

@@ -26,8 +26,8 @@ ForC_simplified$lat <- abs(ForC_simplified$lat)
 ForC_simplified <- ForC_simplified[, c("lat", "mat", "map", "PotentialEvapotranspiration", "VapourPressureDeficit", "TempSeasonality", "length_growing_season")]
 
 ##two groups of variables, needs to be identical + in the same order
-variables1 <- c("lat", "mat", "map", "PotentialEvapotranspiration", "VapourPressureDeficit", "TempSeasonality", "length_growing_season")
-variables2 <- c("lat", "mat", "map", "PotentialEvapotranspiration", "VapourPressureDeficit", "TempSeasonality", "length_growing_season")
+variables1 <- c("lat", "mat", "TempSeasonality", "map", "PotentialEvapotranspiration", "VapourPressureDeficit", "length_growing_season")
+variables2 <- c("lat", "mat", "TempSeasonality", "map", "PotentialEvapotranspiration", "VapourPressureDeficit", "length_growing_season")
 
 fixed.v.info <- read.csv("C:/Users/becky/Dropbox (Smithsonian)/Global_Productivity/raw.data/fixedv_data.csv", stringsAsFactors = F)
 labels <- fixed.v.info$xaxis[which(fixed.v.info$fixed.v %in% fixed.v.info)]
@@ -59,8 +59,13 @@ for (i in seq(along = variables1)){
   fixed.v.info <- read.csv("C:/Users/becky/Dropbox (Smithsonian)/Github/Global_Productivity/raw.data/fixedv_data.csv", stringsAsFactors = F)
   xaxis <- fixed.v.info$xaxis_simple[which(fixed.v.info$fixed.v %in% variables2[[j]])]
   yaxis <- fixed.v.info$xaxis_simple[which(fixed.v.info$fixed.v %in% variables1[[i]])]
+  centre <- fixed.v.info$abbrev[which(fixed.v.info$fixed.v %in% variables2[[j]])]
   # 
   if(i!=j) fit <- lm(var1 ~ var2 , data = df)
+  
+  if(i!=j) pearsons.r <- rcorr(df$var1, df$var2)
+  if(i!=j) pearsons.r <- pearsons.r$r[2]
+  if(i!=j) pearsons.r <- round(pearsons.r, digits = 2)
   if(i!=j) significant.effect <- summary(fit)$coefficients[2,4] < 0.05
[…]

### 2020-08-02 scripts/correlation_matrix_code.R
corrected temp seasonality

@@ -21,6 +21,7 @@ ForC_simplified <- ForC_simplified[age.greater.than.100, ]
 
 ## take absolute value of latitude
 ForC_simplified$lat <- abs(ForC_simplified$lat)
+ForC_simplified$TempSeasonality <- ForC_simplified$TempSeasonality/100
 
 
 ForC_simplified <- ForC_simplified[, c("lat", "mat", "map", "PotentialEvapotranspiration", "VapourPressureDeficit", "TempSeasonality", "length_growing_season")]

### 2020-08-02 scripts/final_model_no_weighting.R

@@ -182,6 +182,7 @@ for(response.variables in response.variables.groups){
         if(fixed.v == "lat") df$fixed <- df$fixed + 0.1
         if(fixed.v == "AnnualWetDays") df$fixed <- df$fixed + 0.1
         if(fixed.v == "VapourPressure") df$fixed <- df$fixed + 0.1
+        if(fixed.v == "TempSeasonality") df$fixed <- df$fixed/100
         # ylim <- range(ForC_simplified[ForC_simplified$variable.name %in% unlist(response.variables),]$mean)
         
         a <- ForC_simplified[ForC_simplified$variable.name %in% unlist(response.variables),]
@@ -375,6 +376,8 @@ for (age in ages){
       # df$masl <- df$masl/1000
       
       df$fixed <- df[, fixed.v]
+      if(fixed.v == "mat") df$fixed <- df$fixed + 11
+      if(fixed.v == "TempSeasonality") df$fixed <- df$fixed/100
       
       a <- ForC_simplified[ForC_simplified$variable.name %in% unlist(response.variables),]
       ylim <- range(tapply(a$mean, a$variable.name, scale))
@@ -382,14 +385,14 @@ for (age in ages){
       ylim[2] <- ylim[2] + 0.25
       
       # if(!fixed.v %in% c("mat", "lat", "PreSeasonality", "SolarRadiation")){
-        
       mod <-  lmer(scale(mean) ~ 1 + (1|geographic.area/plot.name), data = df, REML = F)
       mod.clim <- lmer(scale(mean) ~ poly(fixed, 1, raw = T) + (1|geographic.area/plot.name), data = df, REML = F)
       mod.clim.poly <- lmer(scale(mean) ~ poly(fixed, 2, raw = T) + (1|geographic.area/plot.name), data = df, REML = F)
+      mod.clim.log <- lmer(scale(mean) ~ log(fixed) + (1|geographic.area/plot.name), data = df, REML = F)
       
-      aictab <- aictab(list(mod.clim = mod.clim, mod.clim.poly = mod.clim.poly), sort = T)
+      aictab <- aictab(list(mod = mod, mod.clim = mod.clim, mod.clim.poly = mod.clim.poly, mod.clim.log = mod.clim.log), sort = T)
       
-      best.model <- as.character(aictab(list(mod = mod, mod.clim = mod.clim, mod.clim.poly = mod.clim.poly), sort = T)$Modname[1])
[…]

### 2020-08-02 scripts/ratio_regression_script.R

@@ -15,6 +15,7 @@ library(r2glmm)
 library(nlme)
 library(viridis)
 library(AICcmodavg)
+library(dplyr)
 
 # Load data ####
 ForC_simplified <- read.csv("ForC_simplified/ForC_simplified_WorldClim_CRU_refined.csv", stringsAsFactors = F)
@@ -103,8 +104,8 @@ all.aictab <- NULL
 fixed.variables <- c("lat", "mat", "map", "TempSeasonality")
 # fixed.variables <- "lat"
 
-set1 <- c("NPP", "ANPP", "ANPP_foliage", "ANPP", "BNPP_root")
-set2 <- c("GPP", "BNPP_root", "ANPP_woody_stem", "NPP", "NPP")
+set1 <- c("NPP", "ANPP", "ANPP", "BNPP_root", "ANPP_foliage")
+set2 <- c("GPP", "BNPP_root", "NPP", "NPP", "ANPP_woody_stem")
 
 # set1 <- "ANPP"
 # set2 <- "NPP"
@@ -188,6 +189,7 @@ for (i in seq(along = set1)){
         df <- df[fixed.no.na, ]
         
         df$fixed <- df[, fixed.v]
+        if(fixed.v == "TempSeasonality") df$fixed <- df$fixed/100
         
         ##remove extreme outliers 
         
@@ -262,14 +264,16 @@ for (i in seq(along = set1)){
         significance <- signif(significance, digits=4)
         
         Rsq <- as.data.frame(r.squaredGLMM(mod.full))
-        Rsq <- signif(Rsq, digits=4)[1]
+        Rsq <- signif(Rsq, digits=2)[1]
         # legend2 <- paste("r-squared = ", Rsq[1], "p-value = ", significance)
         # mtext(side = 3, text = legend2, adj = 0.9, cex = 0.5)
         
+        nsites <- as.character(length(unique(df$site_plot)))
+        
         # title (paste("Effect of", fixed.v), outer = T, line = 1)
         if(significant.effect) mtext(paste("R-sq =", Rsq), side = 3, line = -1.5, adj = 0.1, cex = 0.6)
-        if(fixed.v %in% "lat")mtext(paste(set1_name, ":", set2_name), side = 3, line = 0.5, adj = 0, cex = 0.6)
-        if(set1[[i]] %in% "BNPP_root") mtext(side = 1, line = 3, text = xaxis, cex = 0.75)
[…]

### 2020-08-02 scripts/mat_map_interaction.R
updated some scripts

@@ -110,6 +110,7 @@ outputs <- NULL
 
 effects <- c("mat", "map", "(1|geographic.area/plot.name)")
 pannel.nb <- 1
+p.table <- NULL
 
 
 response.variables <- c("GPP", "NPP", "ANPP", "ANPP_woody_stem", "ANPP_foliage", "BNPP_root", "BNPP_root_fine", "R_auto", "R_auto_root")
@@ -146,6 +147,19 @@ for (response.v in response.variables){
   Rsq.i <- signif(Rsq.i, digits=4)[1]
   BIC <- BIC(mod.single, mod.add, mod.int)
   best.model <- as.character(aictab(list(mod.single = mod.single, mod.add = mod.add, mod.int = mod.int), sort = T)$Modname[1])
+  
+  p.sing <- anova(mod.null, mod.single)$"Pr(>Chisq)"[2]
+  p.sing <- signif(p.sing, digits = 2)
+  
+  p.addi <- anova(mod.add, mod.single)$"Pr(>Chisq)"[2]
+  p.addi <- signif(p.addi, digits = 2)
+  
+  p.int <- anova(mod.int, mod.add)$"Pr(>Chisq)"[2]
+  p.int <- signif(p.int, digits = 2)
+  
+  p.values <- data.frame(response.v = response.v, p.sing = p.sing, p.addi = p.addi, p.int = p.int)
+  
+  p.table <- rbind(p.table, p.values)
   
   
   aictab$Rsq <- c(Rsq.s, Rsq.a, Rsq.i)
@@ -290,8 +304,8 @@ for (response.v in response.variables){
 }
 
 dev.off()
-write.csv(all.results, "C:/Users/gyrcbm/Dropbox/Global_Productivity/results/tables/best_model_outputs/mat_map_interaction.csv")
-
+write.csv(all.results, "C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/tables/best_model_outputs/mat_map_interaction.csv")
+write.csv(p.table, "C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/tables/best_model_outputs/mat_map_interaction_pvalues.csv")
 ########################################## code to correctly plot
 Rsqs = NULL
 

### 2020-08-02 scripts/mat_map_interaction_corrected.R

@@ -114,7 +114,7 @@ best.models <- NULL
 
 
 response.variables <- c("GPP", "NPP", "ANPP", "ANPP_woody_stem", "ANPP_foliage", "BNPP_root", "BNPP_root_fine", "R_auto", "R_auto_root")
-response.variables <- "NPP"
+# response.variables <- "NPP"
 
 par(mfrow = c(3,3), mar = c(1,0,0,2), oma = c(5,8,2,0), xpd = T)
 
@@ -165,8 +165,10 @@ for (response.v in response.variables){
     }}
     }}
   
+  p.int <- anova(mod.int, mod.add)$"Pr(>Chisq)"[2]
+  
   best.mod <- as.character(best.model@call)[2]
-  table <- data.frame(response.v = response.v, best.mod = best.mod)
+  table <- data.frame(response.v = response.v, best.mod = best.mod, p.int = p.int)
   best.models <- rbind(best.models, table)
 
 }

### 2020-08-02 scripts/pairwise_comparisons.R

@@ -180,6 +180,7 @@ number = 1
         
         df$fixed <- df[, fixed.v]
         if(fixed.v == "mat") df$fixed <- df$fixed + 11
+        if(fixed.v == "TempSeasonality") df$fixed <- df$fixed/100
         
         ###subset ForC
         

### 2020-08-02 manuscript/appendix.Rmd
updated table S1
fixed captions and column headings in SI tables

@@
   column_spec(6, width = "5cm") %>%
-  kableExtra::footnote(number = c("Hijmans et al. (2005)", "Fick et al. (2017)", "Harris et al. (2017)", "Abatzoglou et al. (2018)"), footnote_as_chunk = TRUE)
+  kableExtra::footnote(number = c("Hijmans et al. (2005)", "Fick et al. (2017)", "Harris et al. (2017)", "Abatzoglou et al. (2018)"), footnote_as_chunk = TRUE) %>%
+  kableExtra::footnote(symbol = "The WorldClim version used was up-to-date at the time of analysis", footnote_as_chunk = TRUE)
 
@@
 kable(S3, format = "latex", booktabs = TRUE, caption = "Table S2.
-Model form, delta AIC, and R2 for each climate variables as a single fixed effect in models for each C flux.
+Model form, $\\Delta$AIC, and R$^{2}$ for each climate variables as a single fixed effect in models for each C flux.
 Model forms include first-order linear (Lin), second-order polynomial (Poly), and logarithmic (Log).", caption.short = "Table S2.
-Model form, delta AIC and R2 for each climate variables as a single fixed effect in models for each C flux") %>%
+Model form, $\\Delta$AIC and R$^{2}$ for each climate variables as a single fixed effect in models for each C flux", escape = F) %>%
   kable_styling(latex_options = c("scale_down", "hold_position")) %>%
@@
 S4 <- read.csv("tables_figures/r2_table2.csv", stringsAsFactors = FALSE, check.names = FALSE)
-kable(S4, format = "latex", booktabs = TRUE) %>%
+kable(S4, format = "latex", booktabs = TRUE, escape = F) %>%
   kable_styling(latex_options = c("scale_down", "hold_position")) %>%
@@
 Best models by carbon flux.
-Models where delta AIC </= 2 are presented", caption.short = "Table S5.
+Models where $\\Delta$AIC </= 2 are presented", caption.short = "Table S5.
 Best models by carbon flux", escape = F) %>%

### 2020-08-03 scripts/AIC_comparisons_of_climate_variables.R
only include models in AIC comparisons where models are significant wrt null

@@ -120,11 +120,11 @@ aictab.top <- NULL
 fixed.variables <- c("mat", "map", "PotentialEvapotranspiration", "VapourPressureDeficit", "SolarRadiation", "TempSeasonality", "PreSeasonality", "length_growing_season", "TempRangeAnnual", "Aridity", "CloudCover", "AnnualFrostDays", "AnnualWetDays", "MaxVPD", "WaterStressMonths")
 
 
-# response.variables.groups <- list(c("GPP", "NPP", "BNPP_root", "BNPP_root_fine"),
-#                                   c("ANPP", "ANPP_foliage", "ANPP_woody_stem"),
-#                                   c("R_auto", "R_auto_root"))
+response.variables.groups <- list(c("GPP", "NPP", "BNPP_root", "BNPP_root_fine"),
+                                  c("ANPP", "ANPP_foliage", "ANPP_woody_stem"),
+                                  c("R_auto", "R_auto_root"))
 
-response.variables.groups <- list(c("NPP"))
+# response.variables.groups <- list(c("NPP"))
 
 for(response.variables in response.variables.groups){
   
@@ -223,7 +223,9 @@ for(response.variables in response.variables.groups){
         
         final.model <- gsub("fixed", fixed.v, final.model, fixed = T)
         
-        model_list <- append(model_list, final.model)
+        is.sig <- anova(mod, mod.full)$"Pr(>Chisq)"[2] < 0.05
+        
+        if(is.sig) model_list <- append(model_list, final.model)
         
         
         equation <-  paste(response.v, "=", r[1], "+", fixed.v,  "x", r[2])

### 2020-08-03 manuscript/appendix.Rmd
merging versions

@@
   kableExtra::footnote(number = c("Hijmans et al. (2005)", "Fick et al. (2017)", "Harris et al. (2017)", "Abatzoglou et al. (2018)"), footnote_as_chunk = TRUE) %>%
-  kableExtra::footnote(symbol = "The WorldClim version used was up-to-date at the time of analysis", footnote_as_chunk = TRUE)
+  kableExtra::footnote(symbol = "The WorldClim version used was the most recent available at the time of analysis", footnote_as_chunk = TRUE)
 

### 2020-08-03 manuscript/manuscript draft.Rmd
merging versions

@@
 The effects of temperature were modified by moisture availability, with C flux reduced under hot and dry conditions and sometimes under very high precipitation. 
-FACF declined with temperature seasonality, but growing season length did not improve upon MAT as a predictor.  
+C fluxes declined with temperature seasonality, but growing season length did not improve upon MAT as a predictor.  
 Within the growing season, the influence of climate on C cycling was small but significant for a number of flux variables. 
@@
 Carbon cycling in forests worldwide provides the energetic basis for sustaining the majority of Earth's terrestrial biodiversity and many human populations [@millennium_ecosystem_assessment_ecosystems_2005], while strongly influencing atmospheric CO~2~ and climate [@bonan_forests_2008].
-Forests' autotrophic carbon fluxes (FACF)--that is, carbon fixation, allocation, and metabolism by trees and other primary producers--sets the energy ultimately available to heterotrophic organisms (including microbes), in turn influencing their abundance [@zak_plant_1994; @niedzialkowska_species_2010] and possibly diversity [@waide_relationship_1999; @chu_direct_2018].
-FACF influences all organic matter stocks in forest ecosystems and is linked to cycling of energy, water, and nutrients.
+Forests' autotrophic carbon fluxes--that is, carbon fixation, allocation, and metabolism by trees and other primary producers--sets the energy ultimately available to heterotrophic organisms (including microbes), in turn influencing their abundance [@zak_plant_1994; @niedzialkowska_species_2010] and possibly diversity [@waide_relationship_1999; @chu_direct_2018].
+They influence all organic matter stocks in forest ecosystems and is linked to cycling of energy, water, and nutrients.
 
-Critically, FACF also define forest interactions with Earth's changing climate.
+Critically, autotrophic C fluxes also define forest interactions with Earth's changing climate.
 Over 69 Gt of CO~2~ cycle through Earth's forests each year [@badgley_terrestrial_2019], and in recent decades their net C sequestration (~2.4 Gt C yr^-1^) offset roughly 30% of anthropogenic fossil fuel emissions [@pan_large_2011].
 As atmospheric carbon dioxide levels continue to rise, driving climate change, forests will play a critical role in shaping the future of Earth's climate [@cavaleri_urgent_2015; @rogelj_mitigation_2018].
-However, our ability to draw general macroscopic conclusions regarding global variation in FACF with respect to climate has been limited in that these analyses often mix forests that vary in stand age, disturbance history, and/or management status; do not always sufficiently parse related variables; and typically consider only one or a few variables at a time.
+However, our ability to draw general macroscopic conclusions regarding global variation in C cycling with respect to climate has been limited in that these analyses often mix forests that vary in stand age, disturbance history, and/or management status; do not always sufficiently parse related variables; and typically consider only one or a few variables at a time.
 
-FACF decrease with latitude, but it remains unclear whether and how the shape of this relationship varies between fluxes.
-FACF are lowest in the boreal regions, and increase into the temperate regions [@luyssaert_co_2007; @huston_global_2009; @beer_terrestrial_2010; @piao_forest_2010; @jung_global_2011].
+Forest C fluxes decrease with latitude, but it remains unclear whether and how the shape of this relationship varies between fluxes.
+Fluxes are lowest in the boreal regions, and increase into the temperate regions [@luyssaert_co_2007; @huston_global_2009; @beer_terrestrial_2010; @piao_forest_2010; @jung_global_2011].
 However, evidence is inconclusive on whether primary productivity continues to increase into the tropics.
@@
 
-The latitudinal gradient in FACF, along with altitudinal gradients [@girardin_net_2010; @malhi_variation_2017], is primarily driven by climate, which is a significant driver of FACF across broad spatial scales [@luyssaert_co_2007; @cleveland_relationships_2011; @hursh_sensitivity_2017].
-The majority of studies have focused on exploring the relationships of FACF to mean annual temperature (MAT) and mean annual precipitation (MAP), as the most commonly reported site-level climate variables.
-There is strong evidence that both MAT and MAP show significant positive relationships with FACF [@chu_does_2016].
-However, as with latitude, the shape of those relationships is not always clear, and, again, is complicated by the use of different measures of FACF across studies.
-Various measures of primary productivity {**FACF?**} saturate at high levels of MAP, though the saturation points identified vary from 1500mm [@luyssaert_co_2007] up to 2445mm MAP [@schuur_productivity_2003].
-Studies of the influence of MAT on productivity {**FACF?**} are less conclusive.
+The latitudinal gradient in C flux rates, along with altitudinal gradients [@girardin_net_2010; @malhi_variation_2017], is driven primarily by climate, which is a significant driver of C fluxes across broad spatial scales [@luyssaert_co_2007; @cleveland_relationships_2011; @hursh_sensitivity_2017].
+The majority of studies have focused on exploring the relationships of C fluxes to mean annual temperature (MAT) and mean annual precipitation (MAP), as the most commonly reported site-level climate variables.
+There is strong evidence that both MAT and MAP show significant positive relationships with C fluxes [@chu_does_2016].
+However, as with latitude, the shape of those relationships is not always clear, and, again, is complicated by the use of different C flux variables across studies.
+Various measures of primary productivity saturate at high levels of MAP, though the saturation points identified vary from 1500mm [@luyssaert_co_2007] up to 2445mm MAP [@schuur_productivity_2003].
+Studies of the influence of MAT on productivity are less conclusive.
 Luyssaert et al.
@@
 MAT and MAP are very coarse measures of climate, and so fail to capture much variation in climate on an intra-annual scale, including the effects of factors such as growing season length, number of frost-free days, temperature seasonality, and dry season length.
-Some studies have suggested that the apparently strong relationship between MAT and FACFs is actually a consequence of the correlation between MAT and growing season length [@kerkhoff_plant_2005; @michaletz_convergence_2014; @michaletz_drivers_2018].
+Some studies have suggested that the apparently strong relationship between MAT and C fluxes is actually a consequence of the correlation between MAT and growing season length [@kerkhoff_plant_2005; @michaletz_convergence_2014; @michaletz_drivers_2018].
 Kerkhoff et al.
@@
 **the following doesn't fit very well...**  
-In addition, FACF can be influenced by other, often interactive, factors across a range of scales [@cleveland_relationships_2011].
+In addition, forest C fluxes can be influenced by other, often interactive, factors across a range of scales [@cleveland_relationships_2011].
 On a local scale, stand age [@litton_carbon_2007; @gillman_latitude_2015], biodiversity [@liang_positive_2016], management [@simova_enigma_2017], and nutrient availability [@aragao_above_2009].
-There is evidence that FACFs also respond to variables such as cloud cover [@taylor_temperature_2017], solar radiation [@fyllas_solar_2017], and potential evapotranspiration [@kerkhoff_plant_2005].
+There is evidence that C fluxes also respond to variables such as cloud cover [@taylor_temperature_2017], solar radiation [@fyllas_solar_2017], and potential evapotranspiration [@kerkhoff_plant_2005].
 
-The recent development of the Global Forest Carbon database (ForC), which synthesizes multiple variables and including records of stand history [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018], opens up the possibility for a standardized analysis of global scale variation in multiple FACF and the principle climatic drivers of these patterns.
+The recent development of the Global Forest Carbon database (ForC), which synthesizes multiple variables and including records of stand history [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018], opens up the possibility for a standardized analysis of global scale variation in multiple C fluxes and the principle climatic drivers of these patterns.
 In order to approach this broad topic, we simplify the major gaps in our knowledge to five broad questions and corresponding hypotheses (Table 1).
-First, we ask how nine FACF in ForC vary with latitude.
+First, we ask how nine forest autotropic carbon fluxes in ForC vary with latitude.
 We then test how these fluxes relate to MAT and MAP, and additionally how they respond to other, less well-studied, climate variables.
-Finally, we consider the relationship between FACF and seasonality, considering the role of seasonality in explaining variation in carbon fluxes, and the influence of climate on FACF standardized by growing season length.
+Finally, we consider the relationship between C flux and seasonality, considering the role of seasonality in explaining variation in carbon fluxes, and the influence of climate on C flux standardized by growing season length.
 
@@
 
-This analysis focused on nine FACF included in the open-access ForC database (Table 2) [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018].
+This analysis focused on nine C flux variables included in the open-access ForC database (Table 2) [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018].
 ForC contains records of field-based measurements of forest carbon stocks and annual fluxes, compiled from original publications and existing data compilations and databases.
@@
 The database was significantly expanded since the publication of @anderson-teixeira_forc_2018 through integration with the Global Soil Respiration Database [@bond-lamberty_global_2010].
-Additional targeted literature searches were conducted to identify any further available data on the FACF analyzed here, with particular focus on mature forests in temperate and boreal regions, which were not included in the review of @anderson-teixeira_carbon_2016.
+Additional targeted literature searches were conducted to identify further available data on the fluxes analyzed here, with particular focus on mature forests in temperate and boreal regions, which were not included in the review of @anderson-teixeira_carbon_2016.
 We used ForC v3.0, archived on Zenodo with DOI 10.5281/zenodo.3403855.
@@
 This analysis focused on mature forests with no known history of signficant disturbance or management. 
-There is evidence that stand age influences patterns of FACF and carbon allocation in forest ecosystems, and can confound relationships between latitude and primary productivity [@de_lucia_forest_2007; @gillman_latitude_2015].
+There is evidence that stand age influences patterns of C flux and allocation in forest ecosystems, and can confound relationships between latitude and primary productivity [@de_lucia_forest_2007; @gillman_latitude_2015].
 To reduce any biasing effects of stand age, we included only stands of known age $\ge$ 100 years and those described by terms such as "mature", "intact", or "old-growth".
-Since management can alter observed patterns of FACF [@simova_enigma_2017], sites were excluded from analysis if they were managed, defined as plots that were planted, managed as plantations, irrigated, fertilised or including the term "managed" in their site description.
+Since management can alter observed patterns of C cycling [@simova_enigma_2017], sites were excluded from analysis if they were managed, defined as plots that were planted, managed as plantations, irrigated, fertilised or including the term "managed" in their site description.
 Sites that had experienced significant disturbance within the past 100 years were also excluded.
@@
 names(FACF_table) [6]  <- paste0(names(FACF_table) [6], footnote_marker_symbol(1))
-kable(FACF_table, format = "latex", booktabs = TRUE, caption = "Definitions and sample sizes of FACF variables used in analysis.
+kable(FACF_table, format = "latex", booktabs = TRUE, caption = "Definitions and sample sizes of carbon flux variables used in analysis.
 All variables are in units of Mg C ha$^{-1}$ yr$^{-1}$.
@@
 
-The effects of latitude and climate on FACF were analysed using mixed effects models using the package 'lme4' [@bates_fitting_2015] in R v.3.5.1 [@r_core_team_r:_2018].
+The effects of latitude and climate on C fluxes were analysed using mixed effects models using the package 'lme4' [@bates_fitting_2015] in R v.3.5.1 [@r_core_team_r:_2018].
 The basic model for all analyses included a fixed effect of latitude or climate and a random effect of plot nested within geographic area.
@@
 
-We first examined the relationship between latitude and FACF (Q1; Table 1).
+We first examined the relationship between latitude and C fluxes (Q1; Table 1).
 We tested models with latitude as a first-order linear term (corresponding null: model without latitude) and as a second-order polynomial or logarithmic term (corresponding null: model with latitude as a first-order linear term).
@@
 
-We next examined the relationships of FACF to climate variables (Q2-Q4; Table 1).
+We next examined the relationships of C fluxes to climate variables (Q2-Q4; Table 1).
 As with latitude, we tested both linear and polynomial fits for each climate variable.
-We tested relationships of each FACF (Table 2) against each climate variable (Table S1), but focus presentation--including focal hypotheses (Table 1)--on climate variables that explained >20% of variation in FACF.
+We tested relationships of each C flux (Table 2) against each climate variable (Table S1), but focus presentation--including focal hypotheses (Table 1)--on climate variables that explained >20% of variation in C fluxes.
 **Becky, please make this more specific.
@@
 
-To test whether and how FACF varied with climate when standardised by growing season length (Q5), we first standardized FACF by dividing by growing season length (as defined above).
+To test whether and how C flux varied with climate when standardised by growing season length (Q5), we first standardized annual C flux by dividing by growing season length (as defined above).
 We then tested for correlations between these standardised fluxes and growing season climate variables, using only first-order linear models.
@@
 
-In total, we analyzed 1228 records from nine FACF variables taken from forests that had experienced no major anthropogenic disturbances within the past 100 years.
+In total, we analyzed 1228 records from nine forest autotrophic C flux variables taken from forests that had experienced no major anthropogenic disturbances within the past 100 years.
 These records represented a total of **#** plots in 154 distinct geographic areas across all forested biogeographic and climate zones (Fig.
@@
 
-*How do FACF vary with latitude?*
+*How does C flux vary with latitude?*
 
@@
 Specifically, latitude explained 64% of variation in GPP (n = 254, p<0.0001), 50% in NPP (n = 114, p<0.0001) and 45% in ANPP (n = 259, p<0.0001).
-The FACF that were most poorly predicted by latitude were $BNPP_{fine.root}$ ($R^2$=0.17) and $ANPP_{stem}$ ($R^2$=0.18).
-For all FACF, the relationship with latitude was better fit by the first-order linear model than by second-order polynomial or logarithmic models.
+The C fluxes that were most poorly predicted by latitude were $BNPP_{fine.root}$ ($R^2$=0.17) and $ANPP_{stem}$ ($R^2$=0.18).
+For all C fluxes, the relationship with latitude was better fit by the first-order linear model than by second-order polynomial or logarithmic models.
 
@@
 S3).
-Of 28 relationships tested (7 FACF ratios regressed against latitude, MAT, MAP, and temperature seasonality), there were only five signficant relationships, all with $R^2 \le 0.4$ (Fig.
+Of 28 relationships tested (7 ratios among C flux variables regressed against latitude, MAT, MAP, and temperature seasonality), there were only five significant relationships, all with $R^2 \le 0.4$ (Fig.
 S3).
-The proportion of NPP allocated to $ANPP_{foliage}$ decreased with latitude ($R^2=0.32$) and increased with MAT ($R^2=0.37$).
+**The proportion of NPP allocated to $ANPP_{foliage}$ decreased with latitude ($R^2=0.32$) and increased with MAT ($R^2=0.37$).
 The proportion of $NPP$ allocated aboveground ($ANPP$) decreased weakly with latitude ($R^2=0.11$) and temperature seasonality ($R^2=0.17$), while increasing with MAT ($R^2=0.11$).
+**
 
-*How do FACF relate to MAT and MAP?*
+*How does C flux relate to MAT and MAP?*
 
@@
 MAP was a significant (p<0.05) predictor of all fluxes (Figs. 4a, S4-S5; Table S2). 
-However, it explained little variation: with the exception of $R_{auto}$, MAP explained at most 25% of variation in FACF. 
+However, it explained little variation: with the exception of $R_{auto}$, MAP explained at most 25% of variation in C flux. 
 For five of the nine fluxes, a second-order polynomial model provided the best fit, while first-order linear and lognormal relationships provided each provided the best fit for two flxues. 
@@
 
-```{r echo=FALSE, out.width='100%', fig.cap = "Interactive effects of mean annual temperature and mean annual precipitation on FACF.
+```{r echo=FALSE, out.width='100%', fig.cap = "Interactive effects of mean annual temperature and precipitation on annual forest carbon fluxes.
 For visualization purposes, data points are grouped into bins of 0 - 1000, 1001 - 2000, 2001 - 3000, and >3000mm mean annual precipitation, and lines of best fit models are plotted for mean annual precipitation values of 500, 1500, 2500, and 3500mm.
@@
 
-*How do FACF relate to other climate variables?*
+*How does C flux relate to other climate variables?*
 
-Our results indicated that FACF were most strongly explained by temperature at the global scale, with temperature-related climate variables coming out as strong predictors of FACF.
+Our results indicated that annual forest C fluxes were most strongly explained by temperature at the global scale, with temperature-related climate variables coming out as strong predictors of C fluxes.
 In addition to MAT, several of its correlates (Fig.
-S2) were consistently identified as strong univariate predictors of FACF: temperature seasonality, annual temperature range, annual frost days, PET, and length of growing season (Figs.
+S2) were consistently identified as strong univariate predictors of C fluxes: temperature seasonality, annual temperature range, annual frost days, PET, and length of growing season (Figs.
 4, S4-S7).
 
-We found a significant relationship between C flux and potential evapotranspiration for all FACF.
+All C flux variables were signficantly related to potential evapotranspiration.
 $ANPP_{foliage}$, $BNPP_{fine.root}$ and $R_{root}$ increased linearly with PET; however, all other fluxes showed a polynomial relationship with PET (Fig.
 4c, S4-5; Table S2).
-We found strong evidence for a saturation point or peak with PET: FACF tended to increase at values below 1000mm, before saturating between 1200 and 1700mm.
-There was also evidence that some FACF begin to decrease at values above 1800mm PET.
+We found strong evidence for a saturation point or peak with PET: C fluxes tended to increase at values below 1000mm, before saturating between 1200 and 1700mm.
+There was also evidence that some C fluxes begin to decrease at values above 1800mm PET.
 
-Vapour pressure deficit was a significant predictor of C flux for all FACF.
-$BNPP_{fine.root}$ showed a linear relationship with vapour pressure deficit ($R^2$ = 0.07, p<0.05), but all other fluxes showed a polynomial relationship (Figs.
-4d, S4-5; Table S2).
-FACF initially increased with vapour pressure deficit, before saturating at around 0.8 kPa, after wich point they began to decrease.
+Vapour pressure deficit was a significant predictor of all C fluxes. 
+$BNPP_{fine.root}$ showed a linear relationship with vapour pressure deficit ($R^2$ = 0.07, p<0.05), but all other fluxes showed a polynomial relationship (Figs. 4d, S4-5; Table S2). 
+C fluxes initially increased with vapour pressure deficit, before saturating at around 0.8 kPa, after which point they began to decrease.
 
@@
 S4-S5, Table S2).
-Solar radiation explained a low proportion of variability in all FACF, explaining less than 20% of the variation in each flux, with the exception of $R_{auto}$ ($R^2$ = 0.26, p<0.05).
+Solar radiation explained a low proportion of variability in all C fluxes, explaining less than 20% of the variation in each flux, with the exception of $R_{auto}$ ($R^2$ = 0.26, p<0.05).
 
-Annual wet days, cloud cover, and aridity were poor or non-significant explainers of variation in FACF, explaining less than 20% of the variation in each of the carbon fluxes (Figs. S4-S5; Table S2).
+Annual wet days, cloud cover, and aridity were poor or non-significant explainers of variation in C fluxes, explaining less than 20% of the variation in each of the carbon fluxes (Figs.
+S4-S5; Table S2).
 
@@
 
-*What is the role of seasonality in explaining FACF?*
+*What is the role of seasonality in explaining C fluxes?*
 
-Temperature seasonality was, overall, the best predictor of FACF. **(update this paragraph based on new results)**
+Temperature seasonality was, overall, the best predictor of annual C fluxes. **(update this paragraph based on new results)**
 $GPP$, $NPP$, $ANPP$, and $R_{root}$ exhibited a polynomial relationship with seasonality (all p<0.05; Figs.
@@
 4e, S6-S7; Table S2).
-Temperature seasonality was strongly correlated with annual temperature range, which was likewise a similarly strong predictor of FACF (Table S2).
-FACF were highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower.
+Temperature seasonality was strongly correlated with annual temperature range, which was likewise a similarly strong predictor of C fluxes (Table S2).
+C fluxes were highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower.
 **BBL: perhaps put this into an ecosystem context; what are these?
@@
 
-In contrast, there was no significant effect of precipitation seasonality on FACF, and both maximum vapour pressure deficit, and water stress months were poor or non-significant predictors of variation in FACF (Figs.
+In contrast, there was no significant effect of precipitation seasonality on C fluxes, and both maximum vapour pressure deficit, and water stress months were poor or non-significant predictors of variation in C fluxes (Figs.
 S6-S7; Table S2).
 
-We found a significant relationship between length of growing season and FACF, with all fluxes showing a linear increase with length of growing season (Figs.
+We found a significant relationship between length of growing season and C fluxes, with all fluxes showing a linear increase with length of growing season (Figs.
 4e, S6-S7; Table S2).
-Length of growing season was a strong predictor of FACF, explaining 51% of variation in GPP, 39% of variation in NPP, and 34% of variation in ANPP, but it was a weaker predictor than MAT for all fluxes analysed (Table S4).
+Length of growing season was a strong predictor of C fluxes, explaining 51% of variation in GPP, 39% of variation in NPP, and 34% of variation in ANPP, but it was a weaker predictor than MAT for all fluxes analysed (Table S4).
 
-*Within the growing season, how do FACF vary with climate?*
+*Within the growing season, how do C fluxes vary with climate?*
 
-When FACF were standardized by growing season length (in monthly increments), correlations with growing season climate--including temperature, precipitation, solar radiation, and PET--were generally weak (Figs.
+When annual C fluxes were standardized by growing season length (in monthly increments), correlations with growing season climate--including temperature, precipitation, solar radiation, and PET--were generally weak (Figs.
 S8-S9).
@@
 Growing season PET had a positive influence on $GPP$ ($R^2$ = 0.15, p<0.01), $NPP$ ($R^2$ = 0.18, p<0.01), $BNPP$ ($R^2$ = 0.23, p<0.0001), $BNPP_{fine.root}$ ($R^2$ = 0.11, p<0.05), and $ANPP_{stem}$ ($R^2$ = 0.06, p<0.05).
-{**Becky, please verify/ edit the following:** There were no other signifcant correlations between growing season length-standardized FACF (9 variables in Table 2) and growing season climate (**which variables?**)}.
+{**Becky, please verify/ edit the following:** There were no other signifcant correlations between growing season length-standardized C fluxes (9 variables in Table 2) and growing season climate (**which variables?**)}.
 
@@
 Our analysis of a large global database (ForC) reveals how autotrophic carbon fluxes in mature forests vary with latitude and climate on a global scale.
-We show that, across all nine FACF analyzed, C cycling decreases continually with latitude (*H1.1*; Fig.
+We show that, across all nine forest autotrophic C fluxes analyzed, C cycling decreases continually with latitude (*H1.1*; Fig.
 2)--a finding that confirms multiple previous studies but contradicts the idea that productivity of temperate forests rivals that of tropical forests [@huston_global_2009].
-FACF generally increase in proportion to one another (*H1.2*), with few differences in allocation detectable at this global scale (Fig.
+C fluxes generally increase in proportion to one another (*H1.2*), with few differences in allocation detectable at this global scale (Fig.
 S2) and with component fluxes summing appropriately to larger fluxes (Fig.
@@
 
-Latitudinal variation in FACF is **BBL: be consistent in tense** primarily attributable to temperature-related variables (*H3, H4*), particularly MAT (Figs.
+Latitudinal variation in C fluxes is primarily attributable to temperature-related variables (*H3, H4*), particularly MAT (Figs.
 3-4).
@@
 3-4).
-Temperature seasonality and growing season length are closely correlated with MAT and are strong predictors of FACF (*H4*; Figs.
+Temperature seasonality and growing season length are closely correlated with MAT and are strong predictors of C fluxes (*H4*; Figs.
 4e-f, S2, S6-S7), though growing season length did not improve upon MAT as a predictor.
-Within the growing season, the influence of climate on C cycling is smaller but still significant for a number of FACF (*H5*; Fig.
+Within the growing season, the influence of climate on C cycling is smaller but still significant for a number of C fluxes (*H5*; Fig.
 S9; Table S4).
 
-These findings clarify how FACF vary with latitude and climate on a global scale.
-Past studies have differed in their conclusions regarding the relationship between FACF and latitude or its correlates (Table 1, *H1*)--quite possibly because of lack of standardization with respect to stand age and disturbance history.
-Our findings indicate that, among mature, undisturbed stands, FACF are unambiguously highest in the tropical regions, and the relationship with latitude is approximately linear (Fig.
+These findings clarify how forest C fluxes vary with latitude and climate on a global scale.
+Past studies have differed in their conclusions regarding the relationship between C fluxes and latitude or its correlates (Table 1, *H1*)--quite possibly because of lack of standardization with respect to stand age and disturbance history.
+Our findings indicate that, among mature, undisturbed stands, forest C fluxes are unambiguously highest in the tropical regions, and the relationship with latitude is approximately linear (Fig.
 2).
@@
 Compared to tropical forests, the temperate forest biome has experienced more widespread anthropogenic disturbance and has a larger fraction of secondary stands [@potapov_mapping_2008; @poulter_global_2018], so analyses comparing across latitudinal gradients without controlling for stand age risk confounding age with biome effects.
-In addition, because carbon allocation varies with stand age [@de_lucia_forest_2007; @anderson-teixeira_altered_2013; @doughty_what_2018], age differences may introduce systematic biases into analyses of FACF across latitude or global climatic gradients.
+In addition, because carbon allocation varies with stand age [@de_lucia_forest_2007; @anderson-teixeira_altered_2013; @doughty_what_2018], age differences may introduce systematic biases into analyses of C fluxes across latitude or global climatic gradients.
 For example, woody productivity tends to be higher in rapidly aggrading secondary stands than in old-growth forests, where proportionally more C is allocated to respiration [@de_lucia_forest_2007; @piao_forest_2010; @doughty_what_2018; @kunert_understanding_2019].
 
-We show that FACF are broadly consistent in their responses to climate drivers on the global scale, with at most modest trends in C allocation among the variable pairs tested (Figs. 2, S3). 
-This parallels the observation that C allocation across multiple FACF varies little with respect to climate along a steep tropical elevational gradient [@malhi_variation_2017; but see @moser_elevation_2011],
+We show that C fluxes are broadly consistent in their responses to climate drivers on the global scale, with at most modest trends in C allocation among the variable pairs tested (Figs. 2, S3). 
+This parallels the observation that C allocation across multiple C fluxes varies little with respect to climate along a steep tropical elevational gradient [@malhi_variation_2017; but see @moser_elevation_2011],
 and is not surprising given that trees face universal constraints in carbon allocation (**REFS**). 
@@
 
-The latitudinal gradient in FACF (Fig.
+The latitudinal gradient in forest C flux (Fig.
 2) is driven primarily by temperature-related climate variables, the effects of which are moderated by moisture availability (Table 1, *H2-H3*; Figs.
 3-4).
-MAT and MAP have long been identified as primary global-scale drivers of FACF [@lieth_primary_1973; **REFS**; @taylor_temperature_2017].
-It is not appropriate to attempt to identify individual mean annual climate variables as mechanistic drivers of FACF because many climate variables co-vary across the latitudinal gradient (Fig.
+MAT and MAP have long been identified as primary global-scale drivers of C fluxes [@lieth_primary_1973; **REFS**; @taylor_temperature_2017].
+It is not appropriate to attempt to identify individual mean annual climate variables as mechanistic drivers of C fluxes because many climate variables co-vary across the latitudinal gradient (Fig.
 S2), because climatic drivers affect forest carbon flux on much shorter time scales than can be captured by annual climate summary variables, and because both climatic conditions and C flux vary intra- and inter-annually around the long-term means.
 However, it remains informative to consider these relationships.
-Among the temperature-related climate variables, $MAT$ is generally the most strongly correlated with FACF (Table S2)--perhaps in part because site-specific MAT is recorded for the majority of sites in ForC, whereas other variables were extracted from global gridded data products, introducing spatial scaling errors.
+Among the temperature-related climate variables, $MAT$ is generally the most strongly correlated with C fluxes (Table S2)--perhaps in part because site-specific MAT is recorded for the majority of sites in ForC, whereas other variables were extracted from global gridded data products, introducing spatial scaling errors.
 
-This finding supports the continued focus on MAT as a primary--*albiet* not mechanistic--correlate of FACF.
-The effects of $MAT$ are modified by moisture availability, with reduced FACF under relatively dry conditions (*i.e.*, low precipitation; high vapour pressure deficit) and sometimes under very high precipitation (Figs.
+This finding supports the continued focus on MAT as a primary--*albiet* not mechanistic--correlate of C fluxes.
+The effects of $MAT$ are modified by moisture availability, with reduced C fluxes under relatively dry conditions (*i.e.*, low precipitation; high vapour pressure deficit) and sometimes under very high precipitation (Figs.
 3-4).
@@
 3) is consistent with an analysis showing a similar interaction for $ANPP$ in tropical forests, also with a cross-over point at ~20$\circ$C [@taylor_temperature_2017].
-However, we detect no such interaction for $ANPP$ or most other FACF, and we find a contrasting negative interaction for $NPP$ (Fig.
-3), suggesting that more data are required to sort out potential differences in the interactive effects of MAT and MAP on FACF in the tropics.
+However, we detect no such interaction for $ANPP$ or most other C fluxes, and we find a contrasting negative interaction for $NPP$ (Fig.
+3), suggesting that more data are required to sort out potential differences in the interactive effects of MAT and MAP on C fluxes in the tropics.
 
@@
 Our analysis clarifies how forest autotrophic carbon fluxes vary with latitude and climate on a global scale, with some important implications for how forest carbon cycling relates to climate and, by extension, how it is likely to respond to climatic warming.
-Our findings show that higher temperatures with similar moisture availability result in a generalized acceleration of FACF (Figs.
+Our findings show that higher temperatures with similar moisture availability result in a generalized acceleration of forest C cycling (Figs.
 2-3).
@@
 Moreover, forest ecosystem responses to climatic changes outside the temperature range to which forest communities are adapated and acclimatized will not necessarily parallel responses across geographic gradients in climate.
-Nevertheless, as we enter a period of accelerating climatic change, understanding of the fundamental climatic controls on FACF sets a foundation for understanding patterns of change.
+Nevertheless, as we enter a period of accelerating climatic change, understanding of the fundamental climatic controls on C cycling sets a foundation for understanding patterns of change.
 

### 2020-08-03 manuscript/manuscript draft.Rmd
working on intro

@@
 
-Carbon cycling in forests worldwide provides the energetic basis for sustaining the majority of Earth's terrestrial biodiversity and many human populations [@millennium_ecosystem_assessment_ecosystems_2005], while strongly influencing atmospheric CO~2~ and climate [@bonan_forests_2008].
+Carbon cycling in Earth's forests provides the energetic basis for sustaining the majority of Earth's terrestrial biodiversity and many human populations [@millennium_ecosystem_assessment_ecosystems_2005], while strongly influencing atmospheric carbon dioxide (CO~2~) and climate [@bonan_forests_2008].
 Forests' autotrophic carbon fluxes--that is, carbon fixation, allocation, and metabolism by trees and other primary producers--sets the energy ultimately available to heterotrophic organisms (including microbes), in turn influencing their abundance [@zak_plant_1994; @niedzialkowska_species_2010] and possibly diversity [@waide_relationship_1999; @chu_direct_2018].
-They influence all organic matter stocks in forest ecosystems and is linked to cycling of energy, water, and nutrients.
-
-Critically, autotrophic C fluxes also define forest interactions with Earth's changing climate.
-Over 69 Gt of CO~2~ cycle through Earth's forests each year [@badgley_terrestrial_2019], and in recent decades their net C sequestration (~2.4 Gt C yr^-1^) offset roughly 30% of anthropogenic fossil fuel emissions [@pan_large_2011].
-As atmospheric carbon dioxide levels continue to rise, driving climate change, forests will play a critical role in shaping the future of Earth's climate [@cavaleri_urgent_2015; @rogelj_mitigation_2018].
-However, our ability to draw general macroscopic conclusions regarding global variation in C cycling with respect to climate has been limited in that these analyses often mix forests that vary in stand age, disturbance history, and/or management status; do not always sufficiently parse related variables; and typically consider only one or a few variables at a time.
+They are linked to cycling of energy, water, and nutrients, and, critically, influence all carbon (C) stocks and define forest interactions with Earth's changing climate. 
+Each year, over 69 Gt of C cycle through Earth's forests [@badgley_terrestrial_2019]-- a flux more than seven times greater that of recent anthropogenic fossil fuel emissions [9.5 Gt C yr^-1^; @friedlingstein_global_2019].
+As atmospheric CO~2~ continues to rise, driving climate change, forests will play a critical role in shaping the future of Earth's climate [@cavaleri_urgent_2015; @rogelj_mitigation_2018]. 
+However, our ability to draw general macroscopic conclusions regarding global variation in C cycling with respect to climate has been limited in that analyses have typically considered only one or a few variables at a time, have not always sufficiently parse related variables, and have often mixed forests that vary in stand age, disturbance history, and/or management status, all of which affect C cycling [@litton_carbon_2007; @gillman_latitude_2015; @simova_enigma_2017].
 
@@
 Fluxes are lowest in the boreal regions, and increase into the temperate regions [@luyssaert_co_2007; @huston_global_2009; @beer_terrestrial_2010; @piao_forest_2010; @jung_global_2011]. 
-However, evidence is inconclusive on whether primary productivity continues to increase into the tropics.
-Evidence for this is further complicated by the fact that different studies use different measures of productivity to explore these relationships.
-For example, modelling of global terrestrial ecosystem gross primary productivity ($GPP$) through upscaling and calibration of eddy flux measurements indicates that GPP peaks in tropical forests [@beer_terrestrial_2010; @jung_global_2011; @badgley_terrestrial_2019; @li_mapping_2019].
+However, evidence is inconclusive on whether primary productivity continues to increase into the tropics--a question complicated by the fact that different studies use different measures of productivity to explore these relationships.
+Gross primary productivity ($GPP$) is highest in tropical forests [@beer_terrestrial_2010; @jung_global_2011; @badgley_terrestrial_2019; @li_mapping_2019]. 
 In contrast, some studies suggest that the highest values of net primary productivity ($NPP$) may be found in temperate forests [@luyssaert_co_2007; @huston_global_2009], while others find $NPP$ highest in the tropics and decreasing with latitude [@simova_enigma_2017].
@@
 Such complicated dynamics play out with belowground fluxes as well, with typically dominant soil temperature mediated by moisture and carbon supply [@hursh_sensitivity_2017; @xu_contribution_2016].
+There is evidence that C fluxes also respond to variables such as cloud cover [@taylor_temperature_2017], solar radiation [@fyllas_solar_2017], and potential evapotranspiration [@kerkhoff_plant_2005].
 
@@
 [-@kerkhoff_plant_2005] and Michaletz et al.
-[-@michaletz_convergence_2014] find that, within the growing season, there is no significant relationship between net primary productivity (loosely defined) and MAT, suggesting that the influence of temperature may be limited to determining the length of the frost-free growing season.
-
-**the following doesn't fit very well...**  
-In addition, forest C fluxes can be influenced by other, often interactive, factors across a range of scales [@cleveland_relationships_2011].
-On a local scale, stand age [@litton_carbon_2007; @gillman_latitude_2015], biodiversity [@liang_positive_2016], management [@simova_enigma_2017], and nutrient availability [@aragao_above_2009].
-There is evidence that C fluxes also respond to variables such as cloud cover [@taylor_temperature_2017], solar radiation [@fyllas_solar_2017], and potential evapotranspiration [@kerkhoff_plant_2005].
+[-@michaletz_convergence_2014] find that, within the growing season, there is no significant relationship between net primary productivity (loosely defined to include a mix of **VARIABLES**) and MAT, suggesting that the influence of temperature may be limited to determining the length of the frost-free growing season.
 

### 2020-08-03 manuscript/manuscript draft.Rmd
revisions to intro
- changed discussion around @huston_global_2009, which is pretty throughly refuted by @gillman_latitude_2015

@@
 
-Forest C fluxes decrease with latitude, but it remains unclear whether and how the shape of this relationship varies between fluxes. 
-Fluxes are lowest in the boreal regions, and increase into the temperate regions [@luyssaert_co_2007; @huston_global_2009; @beer_terrestrial_2010; @piao_forest_2010; @jung_global_2011]. 
-However, evidence is inconclusive on whether primary productivity continues to increase into the tropics--a question complicated by the fact that different studies use different measures of productivity to explore these relationships.
-Gross primary productivity ($GPP$) is highest in tropical forests [@beer_terrestrial_2010; @jung_global_2011; @badgley_terrestrial_2019; @li_mapping_2019]. 
-In contrast, some studies suggest that the highest values of net primary productivity ($NPP$) may be found in temperate forests [@luyssaert_co_2007; @huston_global_2009], while others find $NPP$ highest in the tropics and decreasing with latitude [@simova_enigma_2017].
-Other studies have found that aboveground net primary productivity ($ANPP$) decreases weakly with latitude [@huston_global_2009; @gillman_latitude_2015].
+Forest C fluxes decrease with latitude [e.g., @luyssaert_co_2007; @piao_forest_2010; @gillman_latitude_2015; @li_mapping_2019]; but it remains unclear whether and how the shape of this relationship varies among fluxes.
+For instance, gross primary productivity ($GPP$) increases continuously with decreasing latitude and is indisputably highest in tropical forests [@luyssaert_co_2007; @beer_terrestrial_2010; @jung_global_2011; @badgley_terrestrial_2019; @li_mapping_2019].
+In contrast, some studies have suggested that net primary productivity ($NPP$), or its aboveground portion ($ANPP$), exhibits a less distinct increase from temperate to tropical forests [@luyssaert_co_2007]--or even a decrease [@huston_global_2009, but see @gillman_latitude_2015].
+A shallower increase in $NPP$ than in $GPP$ with decreasing latitude would align with the suggestion that tropical forests tend to have low carbon use efficiency [$CUE$= $NPP$/$GPP$; @de_lucia_forest_2007; @malhi_productivity_2012; @anderson-teixeira_carbon_2016].
+However, accurately comparing relationships of multiple forest C fluxes across latitude requires a large database with standardization for methodology, stand history, and management, which to date has not been possible.
 
-The latitudinal gradient in C flux rates, along with altitudinal gradients [@girardin_net_2010; @malhi_variation_2017], is driven primarily by climate, which is a significant driver of C fluxes across broad spatial scales [@luyssaert_co_2007; @cleveland_relationships_2011; @hursh_sensitivity_2017].
+The latitudinal gradient in forest C flux rates, along with altitudinal gradients [@girardin_net_2010; @malhi_variation_2017], is driven primarily by climate, which is a significant driver of C fluxes across broad spatial scales [@luyssaert_co_2007; @cleveland_relationships_2011; @hursh_sensitivity_2017].
 The majority of studies have focused on exploring the relationships of C fluxes to mean annual temperature (MAT) and mean annual precipitation (MAP), as the most commonly reported site-level climate variables.
-There is strong evidence that both MAT and MAP show significant positive relationships with C fluxes [@chu_does_2016].
-However, as with latitude, the shape of those relationships is not always clear, and, again, is complicated by the use of different C flux variables across studies.
-Various measures of primary productivity saturate at high levels of MAP, though the saturation points identified vary from 1500mm [@luyssaert_co_2007] up to 2445mm MAP [@schuur_productivity_2003].
-Studies of the influence of MAT on productivity are less conclusive.
-Luyssaert et al.
-[-@luyssaert_co_2007] examined GPP and NPP and found that, while $GPP$ increases linearly with MAT, NPP saturates at around 10$^\circ$C MAT.
-In contrast, Larjavaara and Muller-Landau [-@larjavaara_temperature_2012], find that increases in $GPP$ saturate at approximately 25$^\circ$C MAT, while Schuur [-@schuur_productivity_2003] finds that $NPP$ increases linearly with temperature.
-Taylor et al.
-[-@taylor_temperature_2017] showed a positive interaction between MAT and MAP in shaping tropical forest productivity, such that high rainfall had a negative effect on productivity in cooler climates, compared to a positive effect in warmer climates.
-Such complicated dynamics play out with belowground fluxes as well, with typically dominant soil temperature mediated by moisture and carbon supply [@hursh_sensitivity_2017; @xu_contribution_2016].
-There is evidence that C fluxes also respond to variables such as cloud cover [@taylor_temperature_2017], solar radiation [@fyllas_solar_2017], and potential evapotranspiration [@kerkhoff_plant_2005].
+There is strong evidence that both MAT and MAP show significant positive relationships with C fluxes [@chu_does_2016]; however, as with latitude, we lack synthetic understanding of how these relationships vary across climate fluxes.
+C flux increases strongly with $MAT$ on the global scale, but whether they saturate or potentially decrease at higher temperatures remains disputed.
+In the global data compilation of Luyssaert et al. [-@luyssaert_co_2007], $GPP$ increased linearly with MAT, whereas $NPP$ saturated at around 10$^\circ$C MAT. 
+In contrast, Larjavaara and Muller-Landau [-@larjavaara_temperature_2012], find that increases in $GPP$ saturate at approximately 25$^\circ$C MAT, while Schuur [-@schuur_productivity_2003] finds that $NPP$ increases continuously with temperature.
+Various measures of primary productivity saturate at high levels of MAP, though the saturation points identified vary from 1500mm [@luyssaert_co_2007] up to 2,445 mm MAP [@schuur_productivity_2003]. 
+Interactions between MAT and MAP are also possible; within the tropics, there is a positive interaction between MAT and MAP in shaping $ANPP$, such that high rainfall has a negative effect on productivity in cooler climates, compared to a positive effect in warmer climates [@taylor_temperature_2017].
+Such complicated dynamics play out with belowground fluxes as well, with typically dominant soil temperature mediated by moisture and C supply [@hursh_sensitivity_2017; @xu_contribution_2016]. 
+There is also evidence that C fluxes also respond to climate variables such as cloud cover [@taylor_temperature_2017], solar radiation [@beer_terrestrial_2010; @fyllas_solar_2017], and potential evapotranspiration [@kerkhoff_plant_2005]; however, these are not typically assessed in global-scale analyses of annual forest C flux.
 
-MAT and MAP are very coarse measures of climate, and so fail to capture much variation in climate on an intra-annual scale, including the effects of factors such as growing season length, number of frost-free days, temperature seasonality, and dry season length.
+As metrics of annual climate, MAT and MAP fail to capture variation in climate on an intra-annual scale, including temperature and precipitation seasonality and growing season length. 
 Some studies have suggested that the apparently strong relationship between MAT and C fluxes is actually a consequence of the correlation between MAT and growing season length [@kerkhoff_plant_2005; @michaletz_convergence_2014; @michaletz_drivers_2018].
@@
 [-@kerkhoff_plant_2005] and Michaletz et al.
-[-@michaletz_convergence_2014] find that, within the growing season, there is no significant relationship between net primary productivity (loosely defined to include a mix of **VARIABLES**) and MAT, suggesting that the influence of temperature may be limited to determining the length of the frost-free growing season.
+[-@michaletz_convergence_2014] found no significant relationship between growing season temperature and net primary productivity (loosely defined to include a mix of **VARIABLES**) standardized to growing season length.
+While this suggests that the influence of temperature may be limited to determining the length of the frost-free growing season, *analysis with clearly defined variables and standardized forest ages would be necessary to test the the veracity and generality of this hypothesis.
 
@@
 In order to approach this broad topic, we simplify the major gaps in our knowledge to five broad questions and corresponding hypotheses (Table 1).
-First, we ask how nine forest autotropic carbon fluxes in ForC vary with latitude.
+First, we ask how nine forest autotrophic carbon fluxes in ForC vary with latitude.
 We then test how these fluxes relate to MAT and MAP, and additionally how they respond to other, less well-studied, climate variables.
@@
   pack_rows("Q1. How do C fluxes vary with latitude?", 1, 1, latex_gap_space = "1em", colnum = 5, hline_before = FALSE) %>%
-  pack_rows("Q2. How do C fluxes vary with MAT and MAP?", 2, 4, latex_gap_space = "1em", colnum = 5, hline_before = TRUE) %>%
-  pack_rows("Q3. How are C fluxes related to other climate variables?", 5, 7, latex_gap_space = "1em", colnum = 5, hline_before = TRUE) %>%
-  pack_rows("Q4. How does seasonality influence C fluxes?", 8, 11, latex_gap_space = "1em", colnum = 5, hline_before = TRUE) %>%
-  pack_rows("Q5. When standardised by growing season length, how do C fluxes vary with climate?", 12, 15, latex_gap_space = "1em", colnum = 5, hline_before = TRUE) %>%
+  pack_rows("Q2. How do C fluxes vary with mean annual temperature (MAT) and precipitation (MAP)?", 2, 4, latex_gap_space = "1em", colnum = 5, hline_before = TRUE) %>%
+  pack_rows("Q3. How are C fluxes related to other annual climate variables?", 5, 7, latex_gap_space = "1em", colnum = 5, hline_before = TRUE) %>%
+  pack_rows("Q4. How does seasonality influence annual C fluxes?", 8, 11, latex_gap_space = "1em", colnum = 5, hline_before = TRUE) %>%
+  pack_rows("Q5. When standardised by growing season length, how do annual C fluxes vary with climate?", 12, 15, latex_gap_space = "1em", colnum = 5, hline_before = TRUE) %>%
   add_header_above(c(" ",  " ", "Forest autotrophic carbon fluxes" = 9, " ")) %>%

### 2020-08-04 scripts/stacked_plots_update.R
updated stacked plots

@@ -1,0 +1,1017 @@
+# Clean environment ####
+rm(list = ls())
+
+# Set working directory as ForC main folder ####
+setwd("C:/Users/becky/Dropbox (Smithsonian)/GitHub/ForC")
+
+# Load libaries ####
+library(lme4)
+library(MuMIn)
+library(viridis)
+
+# Load data ####
+ForC_simplified <- read.csv("ForC_simplified/ForC_simplified_WorldClim_CRU_refined.csv", stringsAsFactors = F)
+VARIABLES <- read.csv(paste0(dirname(getwd()), "/ForC/data/ForC_variables.csv"), stringsAsFactors = F)
+
+na_codes <- c("NA", "NI", "NRA", "NaN", "NAC") 
+my_is.na <- function(x) { is.na(x) | x %in% na_codes}
+my_na.omit <- function(x) { return(x[!my_is.na(x)])}
+
+
+# Prepare data ####
+
+## change to numerics what it needs to
+ForC_simplified$stand.age <- as.numeric(ForC_simplified$stand.age)
+ForC_simplified$min.db <- as.numeric(ForC_simplified$min.db)
+
+## make plot and geographic areas factors
+ForC_simplified$plot.name <- addNA(ForC_simplified$plot.name)
+ForC_simplified$geographic.area <- addNA(ForC_simplified$geographic.area)
+
+## change sign of NEE 
+ForC_simplified[grepl("NEE", ForC_simplified$variable.name,ignore.case = F),]$mean <- -ForC_simplified[grepl("NEE", ForC_simplified$variable.name,ignore.case = F),]$mean
+
+## take absolute value of latitude
+ForC_simplified$lat <- abs(ForC_simplified$lat)
+
+
+# Control for some factors ####
+
+
+## keep all ages except na, 0 and 999
+ages.not.999.nor.0.nor.na <- !ForC_simplified$stand.age %in% 999 &  !ForC_simplified$stand.age %in% 0 & !is.na(ForC_simplified$stand.age)
+
+## Keep only age >=100 (or 999)
+age.greater.than.100 <- ForC_simplified$stand.age >= 100 & !is.na(ForC_simplified$stand.age)
+age.greater.than.200 <- ForC_simplified$stand.age >= 200 & !is.na(ForC_simplified$stand.age)
+ages <- c("age.greater.than.100")
+
+## keep only stands that are NOT too strongly influence by management/ disturbance
+
[…]

### 2020-08-04 scripts/final_model_no_weighting.R
issue #100 sorted

@@ -492,7 +492,7 @@ for (age in ages){
     
   }
   
-  mtext(side = 2, line = 1,  text = expression("Productivity (scaled values)"), outer = T)
+  mtext(side = 2, line = 1,  text = expression("Carbon flux (scaled values)"), outer = T)
   
   dev.off()
 }

### 2020-08-04 scripts/stacked_plots_update.R

@@ -320,7 +320,7 @@ for (age in ages){
             
             # title(paste("Stacked graphs by latitude"), outer = F, line = 1)
             mtext(side = 1, line = 1, text = expression(paste("Absolute latitude (", degree, ")")), outer = T)
-            mtext(side = 2, line = 1,  text = expression(paste("Productivity (Mg C"~ha^-1~yr^-1, ")")), outer = T)
+            mtext(side = 2, line = 1,  text = expression(paste("Carbon flux (Mg C"~ha^-1~yr^-1, ")")), outer = T)
             mtext(paste0("(", letters[1], ")"), side = 3, line = -1.5, adj = 0.05)
             
             # dev.off()

### 2020-08-04 scripts/stacked_plots_code.R
Update stacked_plots_code.R

@@ -2,7 +2,7 @@
 rm(list = ls())
 
 # Set working directory as ForC main folder ####
-setwd("C:/Users/gyrcbm/Dropbox/ForC")
+setwd("C:/Users/becky/Dropbox (Smithsonian)/GitHub/ForC")
 
 # Load libaries ####
 library(lme4)
@@ -305,7 +305,7 @@ par(mfrow = c(2,2), mar = c(2,2,2,2), oma = c(3,3,0,0))
         
           # title(paste("Stacked graphs by latitude"), outer = F, line = 1)
           mtext(side = 1, line = 1, text = expression(paste("Absolute latitude (", degree, ")")), outer = T)
-          mtext(side = 2, line = 1,  text = expression(paste("Productivity (Mg C"~ha^-1~yr^-1, ")")), outer = T)
+          mtext(side = 2, line = 1,  text = expression(paste("Carbon flux (Mg C"~ha^-1~yr^-1, ")")), outer = T)
           mtext(paste0("(", letters[1], ")"), side = 3, line = -1.5, adj = 0.05)
           
           # dev.off()

### 2020-08-04 scripts/final_model_no_weighting.R
Update final_model_no_weighting.R

@@ -492,7 +492,7 @@ for (age in ages){
     
   }
   
-  mtext(side = 2, line = 1,  text = expression("Carbon flux (scaled values)"), outer = T)
+  mtext(side = 2, line = 1,  text = expression("Carbon flux (standard normal deviates)"), outer = T)
   
   dev.off()
 }

### 2020-08-04 manuscript/manuscript draft.Rmd

@@
 To account for this, a number of analyses seeking to characterize global-scale effects of climate on productivity have examined the relationship of C flux per month of the growing season with growing season climatic conditions [Table 1, *H5*; @kerkhoff_plant_2005; @anderson_temperature-dependence_2006; @enquist_adaptive_2007; @michaletz_convergence_2014].
-The sort of simple metric needed to define growing season at a global scale [@kerkhoff_plant_2005] is coarse with respect to temperature because it's calculated on a monthly timescale and problematic with respect to moisture because it doesn't capture temporal lags between precipitation and plant water use caused by storage in soil or snow.
+The sort of simple metric that has been used to define growing season at a global scale [@kerkhoff_plant_2005] is coarse with respect to temperature because it's calculated on a monthly timescale and problematic with respect to moisture because it doesn't capture temporal lags between precipitation and plant water use caused by storage in soil or snow.
 We found that a temperature-defined growing season length had strong positive correlation with C fluxes (Fig.
-4f), but explained less variation than $MAT$.
+4f), but was never the best.
 Dividing annual fluxes by growing season length to yield average flux per growing season month removed the majority of climate-related variation, supporting the idea that the latitudinal gradient in carbon flux is attributable more to shorter growing seasons at high latitudes than to inherently lower rates of photosynthesis or respiration by high-latitude forests [@enquist_adaptive_2007].

### 2020-08-04 manuscript/manuscript draft.Rmd
update Fig. 2 caption, plus work on intro and discussion

@@
 
-Carbon cycling in Earth's forests provides the energetic basis for sustaining the majority of Earth's terrestrial biodiversity and many human populations [@millennium_ecosystem_assessment_ecosystems_2005], while strongly influencing atmospheric carbon dioxide (CO~2~) and climate [@bonan_forests_2008].
+Carbon (C) cycling in Earth's forests provides the energetic basis for sustaining the majority of Earth's terrestrial biodiversity and many human populations [@millennium_ecosystem_assessment_ecosystems_2005], while strongly influencing atmospheric carbon dioxide (CO~2~) and climate [@bonan_forests_2008].
 Forests' autotrophic carbon fluxes--that is, carbon fixation, allocation, and metabolism by trees and other primary producers--sets the energy ultimately available to heterotrophic organisms (including microbes), in turn influencing their abundance [@zak_plant_1994; @niedzialkowska_species_2010] and possibly diversity [@waide_relationship_1999; @chu_direct_2018].
-They are linked to cycling of energy, water, and nutrients, and, critically, influence all carbon (C) stocks and define forest interactions with Earth's changing climate. 
+They are linked to cycling of energy, water, and nutrients, and, critically, influence all C stocks and define forest interactions with Earth's changing climate. 
 Each year, over 69 Gt of C cycle through Earth's forests [@badgley_terrestrial_2019]-- a flux more than seven times greater that of recent anthropogenic fossil fuel emissions [9.5 Gt C yr^-1^; @friedlingstein_global_2019].
 As atmospheric CO~2~ continues to rise, driving climate change, forests will play a critical role in shaping the future of Earth's climate [@cavaleri_urgent_2015; @rogelj_mitigation_2018]. 
-However, our ability to draw general macroscopic conclusions regarding global variation in C cycling with respect to climate has been limited in that analyses have typically considered only one or a few variables at a time, have not always sufficiently parse related variables, and have often mixed forests that vary in stand age, disturbance history, and/or management status, all of which affect C cycling [@litton_carbon_2007; @gillman_latitude_2015; @simova_enigma_2017].
+However, our understanding of the climate dependence of forest C cycling on a global scale has been limited by analyses typically considering only one or a few variables at a time, insufficient parsing of related variables, and the mixing of data from forests that vary in stand age, disturbance history, and/or management status, all of which affect C cycling [@litton_carbon_2007; @gillman_latitude_2015; @simova_enigma_2017].
 
-Forest C fluxes decrease with latitude [e.g., @luyssaert_co_2007; @piao_forest_2010; @gillman_latitude_2015; @li_mapping_2019]; but it remains unclear whether and how the shape of this relationship varies among fluxes.
+Forest C fluxes decrease with latitude [e.g., @luyssaert_co_2007; @piao_forest_2010; @gillman_latitude_2015; @li_mapping_2019], but it remains unclear whether and how the shape of this relationship varies among fluxes.
 For instance, gross primary productivity ($GPP$) increases continuously with decreasing latitude and is indisputably highest in tropical forests [@luyssaert_co_2007; @beer_terrestrial_2010; @jung_global_2011; @badgley_terrestrial_2019; @li_mapping_2019].
@@
 ```{r echo=FALSE, out.width='100%', fig.cap = "Latitudinal trends in forest autotropic carbon flux.
-Lines of best fit are plotted according to the best model selected during analysis.
-All regressions are significant $(p<0.05)$.
+Plotted are linear models, all of which were significant $(p<0.05)$ and within dAIC=2 of the best model (for three fluxes, logarithmic fits were marginally better; Table S2).
 Each panel shows major C fluxes together with component fluxes.
@@
 3) is consistent with an analysis showing a similar interaction for $ANPP$ in tropical forests, also with a cross-over point at ~20$\circ$C [@taylor_temperature_2017].
-However, we detect no such interaction for $ANPP$ or most other C fluxes, and we find a contrasting negative interaction for $NPP$ (Fig.
+*(@sullivan_sensitivity_2020 show a decline in $ANPP_{stem}$ in sites with high max temperatures and limited moisture.)* However, we detect no such interaction for $ANPP$ or most other C fluxes, and we find a contrasting negative interaction for $NPP$ (Fig.
 3), suggesting that more data are required to sort out potential differences in the interactive effects of MAT and MAP on C fluxes in the tropics.
@@
 Our analysis clarifies how forest autotrophic carbon fluxes vary with latitude and climate on a global scale, with some important implications for how forest carbon cycling relates to climate and, by extension, how it is likely to respond to climatic warming.
-Our findings show that higher temperatures with similar moisture availability result in a generalized acceleration of forest C cycling (Figs.
-2-3).
+Our findings show that higher temperatures with similar moisture availability result in a generalized acceleration of forest C cycling (Figs. 2-3).
 This is consistent with observations of continental- to global-scale increases over time in $GPP$ [@li_mapping_2019] and $ANPP_{stem}$ [@brienen_long-term_2015; @hubau_asynchronous_2020], along with some C cycle components not considered here--tree mortality [@brienen_long-term_2015; @mcdowell_drivers_2018], soil respiration [@bond-lamberty_global_2010], and heterotrophic soil respiration [@bond-lamberty_globally_2018].
-**DISCUSS/CITE [Wu et al.
-2020]** [@wu_accelerated_2020] However, increasing C flux rates are by no means universal [@rutishauser_testing_2020] (**MORE REFS**), likely because other factors are at play, including changes to other aspects of climate, atmospheric pollution (CO~2~, SO~2~, NO~x~), and local disturbances.
-**(discuss new Sullivan paper that finds higher max temperatures associated to lower ANPP_stem)**
-Moreover, forest ecosystem responses to climatic changes outside the temperature range to which forest communities are adapated and acclimatized will not necessarily parallel responses across geographic gradients in climate.
-Nevertheless, as we enter a period of accelerating climatic change, understanding of the fundamental climatic controls on C cycling sets a foundation for understanding patterns of change.
+**DISCUSS/CITE [Wu et al. 2020]** [@wu_accelerated_2020] 
+However, increasing C flux rates are by no means universal [e.g., @rutishauser_testing_2020; @hubau_asynchronous_2020], likely because other factors are at play, including changes to other aspects of climate, atmospheric pollution (CO~2~, SO~2~, NO~x~), and local disturbances.
+Carbon cycling is also likely to **(discuss new Sullivan paper that finds higher max temperatures associated to lower ANPP_stem)** [@sullivan_sensitivity_2020]
+Moreover, forest ecosystem responses to climatic changes outside the temperature range to which forest communities are adapted and acclimatized will not necessarily parallel responses across geographic gradients in climate.
+Nevertheless, understanding of the fundamental climatic controls on C cycling sets a firmer foundation for understanding forest C cycle responses to accellerating climate change.
 

### 2020-08-04 manuscript/manuscript draft.Rmd
worked on intro, Table 1 citations
#103

@@
 
+nocite: | 
+  @chu_does_2016
 ---
@@
 The majority of studies have focused on exploring the relationships of C fluxes to mean annual temperature (MAT) and mean annual precipitation (MAP), as the most commonly reported site-level climate variables.
-There is strong evidence that both MAT and MAP show significant positive relationships with C fluxes [@chu_does_2016]; however, as with latitude, we lack synthetic understanding of how these relationships vary across climate fluxes.
-C flux increases strongly with $MAT$ on the global scale, but whether they saturate or potentially decrease at higher temperatures remains disputed.
+C fluxes increase strongly with $MAT$ on the global scale, but whether they saturate or potentially decrease at higher temperatures remains disputed. (**This part could use some work.**)
 In the global data compilation of Luyssaert et al. [-@luyssaert_co_2007], $GPP$ increased linearly with MAT, whereas $NPP$ saturated at around 10$^\circ$C MAT. 
 In contrast, Larjavaara and Muller-Landau [-@larjavaara_temperature_2012], find that increases in $GPP$ saturate at approximately 25$^\circ$C MAT, while Schuur [-@schuur_productivity_2003] finds that $NPP$ increases continuously with temperature.
+@sullivan_sensitivity_2020 found that, within the tropics, $ANPP_{stem}$ decreases at the highest maximum temperatures.
 Various measures of primary productivity saturate at high levels of MAP, though the saturation points identified vary from 1500mm [@luyssaert_co_2007] up to 2,445 mm MAP [@schuur_productivity_2003]. 
@@
 ```{r echo=FALSE, out.width='100%', fig.cap = "Latitudinal trends in forest autotropic carbon flux.
-Plotted are linear models, all of which were significant $(p<0.05)$ and within dAIC=2 of the best model (for three fluxes, logarithmic fits were marginally better; Table S2).
+Plotted are linear models, all of which were significant $(p<0.05)$ and had AIC values within 2.0 of the best model (for three fluxes, logarithmic fits were marginally better; Table S2).
 Each panel shows major C fluxes together with component fluxes.
@@
 However, increasing C flux rates are by no means universal [e.g., @rutishauser_testing_2020; @hubau_asynchronous_2020], likely because other factors are at play, including changes to other aspects of climate, atmospheric pollution (CO~2~, SO~2~, NO~x~), and local disturbances.
-Carbon cycling is also likely to **(discuss new Sullivan paper that finds higher max temperatures associated to lower ANPP_stem)** [@sullivan_sensitivity_2020]
+**(discuss new Sullivan paper that finds higher max temperatures associated to lower ANPP_stem)** [@sullivan_sensitivity_2020]
 Moreover, forest ecosystem responses to climatic changes outside the temperature range to which forest communities are adapted and acclimatized will not necessarily parallel responses across geographic gradients in climate.

### 2020-08-04 manuscript/manuscript draft.Rmd
small updates to manuscript

@@
 ```{r setup, include = FALSE}
-# knitr::opts_knit$set(root.dir = 'C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/')
- knitr::opts_knit$set(root.dir = '/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/Global_Productivity/manuscript/')
+knitr::opts_knit$set(root.dir = 'C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/')
+ # knitr::opts_knit$set(root.dir = '/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/Global_Productivity/manuscript/')
 ```
@@
 However, we lack a cohesive synthesis on how forest carbon fluxes vary globally with respect to climate and one another. 
-Here, we draw upon **#** records from the Global Forest Carbon Database (ForC), representing all major forest types and the nine most significant autotrophic carbon fluxes, to comprehensively explore how C cycling in mature, undisturbed forests varies with latitude and climate on a global scale.
+Here, we draw upon 1319 records from the Global Forest Carbon Database (ForC), representing all major forest types and the nine most significant autotrophic carbon fluxes, to comprehensively explore how C cycling in mature, undisturbed forests varies with latitude and climate on a global scale.
 We show that, across all flux variables analyzed, C cycling decreases linearly with absolute latitude -- a finding that confirms multiple previous studies but contradicts the idea that net primary productivity ($NPP$) of temperate forests rivals that of tropical forests.
 C flux variables generally displayed similar trends across latitude and multiple climate variables, with few differences in allocation detectable at this global scale.
-**Climate explained a significant proportion (#-#%) of variation in all C fluxes analyzed**, with temperature variables in general and mean annual temperature (MAT) in particular being the best predictors of C flux on this global scale.
+**Climate explained a significant proportion (24 - 71%) of variation in all C fluxes analyzed**, with temperature variables in general and mean annual temperature (MAT) in particular being the best predictors of C flux on this global scale.
 The effects of temperature were modified by moisture availability, with C flux reduced under hot and dry conditions and sometimes under very high precipitation. 
@@
 From these data, we derived maximum VPD, defined as the VPD of the month with the largest deficit, and the number of water stress months, defined as the number of months annually where precipitation was lower than PET.
-Where site-level data was missing for MAT or MAP, we used values from the WorldClim (**[2?](https://github.com/forc-db/Global_Productivity/issues/89)**) dataset.
+Where site-level data was missing for MAT or MAP, we used values from the WorldClim dataset.
 
@@
 
-Variation in allocation to component carbon fluxes along latitudinal gradients was explored for the following pairings: $NPP:GPP$, $ANPP:BNPP$, $ANPP_{foliage}:ANPP_{stem}$, $ANPP_{foliage}:NPP$, $ANPP_{stem}:NPP$, $ANPP:NPP$, and $BNPP:NPP$.
+Variation in allocation to component carbon fluxes along latitudinal gradients was explored for the following pairings: $NPP:GPP$, $ANPP:BNPP$, $ANPP:NPP$, $BNPP:NPP$, and $ANPP_{foliage}:ANPP_{stem}$.
 **[issue 91](https://github.com/forc-db/Global_Productivity/issues/91)**

### 2020-08-04 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd
#90

@@
 Here, we draw upon 1319 records from the Global Forest Carbon Database (ForC), representing all major forest types and the nine most significant autotrophic carbon fluxes, to comprehensively explore how C cycling in mature, undisturbed forests varies with latitude and climate on a global scale.
-We show that, across all flux variables analyzed, C cycling decreases linearly with absolute latitude -- a finding that confirms multiple previous studies but contradicts the idea that net primary productivity ($NPP$) of temperate forests rivals that of tropical forests.
+We show that, across all flux variables analyzed, C cycling decreases continuously with absolute latitude -- a finding that confirms multiple previous studies but contradicts the idea that net primary productivity ($NPP$) of temperate forests rivals that of tropical forests.
 C flux variables generally displayed similar trends across latitude and multiple climate variables, with few differences in allocation detectable at this global scale.
@@
 We experimented with inclusion of altitude as a fixed effect, but excluded it from the final models because it added very little explanatory power--that is, the difference in AIC ($\Delta AIC$) relative to models excluding altitude was generally small (often $\Delta AIC$<2).
-Effects were considered signficant when inclusion of the term of interest resulted in $\Delta AIC \ge 2.0$ relative to a corresponding null model.
+Effects were considered significant when inclusion of the fixed of interest resulted in $\Delta AIC \ge 2.0$ relative to a corresponding null model. 
 All $R^2$ values presented here are marginal $R^2$ values, and refer to the proportion of variation explained by only the fixed effects. 
@@
 
-We first examined the relationship between latitude and C fluxes (Q1; Table 1).
-We tested models with latitude as a first-order linear term (corresponding null: model without latitude) and as a second-order polynomial or logarithmic term (corresponding null: model with latitude as a first-order linear term).
-We selected as the best model that with the highest $\Delta AIC$ relative to a null model with no fixed term.
+We first examined the relationship between latitude and C fluxes (*Q1*; Table 1). 
+We tested models with latitude as a  first-order linear, second-order polynomial, and logarithmic term. 
+For brevity, we henceforth refer to first-order linear models as "linear" and second-order polynomial models as "polynomial".
+We selected as the best model that with the highest $\Delta AIC$ relative to a null model with no fixed term, with the qualification that a polynomial model was considered an improvement over a linear model only if it reduced $\DeltaAIC$ by 2.0 or more.
 
@@
 We next examined the relationships of C fluxes to climate variables (Q2-Q4; Table 1). 
-As with latitude, we tested both linear and polynomial fits for each climate variable.
+We tested first-order linear, second-order polynomial, and logarithmic fits for each climate variable. 
+Again, polynomial fits were considered superior to first-order linear fits only if inclusion of a second-order polynomial term resulted in $\Delta AIC \ge 2.0$ relative to a first-order linear model.
 We tested relationships of each C flux (Table 2) against each climate variable (Table S1), but focus presentation--including focal hypotheses (Table 1)--on climate variables that explained >20% of variation in C fluxes.

### 2020-08-05 scripts/ratio_regressions_alternative.R
working on ratio regressions

@@ -140,18 +140,18 @@ final_sheet$masl.x[is.na(final_sheet$masl.x)] <- (final_sheet$masl.y)[is.na(fina
 final_sheet <- final_sheet[-(13:18)]
 names(final_sheet)[5:10] <- c("lat", "mat", "map", "TempSeasonality", "geographic.area", "masl")
 
-final_sheet$ratio <- log(final_sheet$NPP_mean/final_sheet$R_mean)
+final_sheet$ratio <- final_sheet$NPP_mean/final_sheet$GPP_mean
 
 fixed.variables <- c("lat", "mat", "map", "TempSeasonality")
 
-png(file = "C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/figures/final_figures/ratio_regressions/log ratios/all.png", width = 3000, height = 2000, units = "px", res = 300)
+png(file = "C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/figures/final_figures/ratio_regressions/log ratios/all.png", width = 2000, height = 3000, units = "px", res = 300)
 
-par(mfrow = c(3,4), mar = c(1,1,3,0.5), oma = c(5,4,5,0))
+par(mfrow = c(5,4), mar = c(1,1,3,0.5), oma = c(5,4,5,0))
 
 n <- 1
 
 for(fixed.v in fixed.variables){
-
+  
   final_sheet$fixed <- final_sheet[, fixed.v]
   
   mod <-  lmer(ratio ~ 1 + (1|geographic.area/plot.name), data = final_sheet, REML = F)
@@ -171,15 +171,15 @@ for(fixed.v in fixed.variables){
   newDat <- expand.grid(fixed = seq(min(final_sheet$fixed), max(final_sheet$fixed), length.out = 100))
   newDat$fit <- predict(mod.full, newDat, re.form = NA)
   
-  ylim <- range(exp(final_sheet$ratio))
+  ylim <- range(final_sheet$ratio)
   ylim[1] <- ylim[1] - 0.25
   ylim[2] <- ylim[2] + 0.25
   
   
-  if(n == 1) plot(exp(ratio) ~ fixed, data = final_sheet, ylab = "", ylim = ylim)
-  if(n != 1) plot(exp(ratio) ~ fixed, data = final_sheet, ylab = "", ylim = ylim, yaxt = "n")
[…]

### 2020-08-05 scripts/refining_ratio_regressions_all.R
outliers removed

@@ -270,27 +270,46 @@ for(fixed.v in fixed.variables){
   if (best.model == "mod.linear") mod.full <- lmer(ratio ~ poly(fixed, 1, raw = T) + (1|geographic.area/plot.name), data = final_sheet, REML = F)
   if (best.model == "mod") mod.full <- lmer(ratio ~ poly(fixed, 1, raw = T) + (1|geographic.area/plot.name), data = final_sheet, REML = F)
   
+  cooksd <- cooks.distance(mod.full)
+
+  # plot(cooksd, pch="*", cex=2, main="Influential Obs by Cooks distance")  # plot cook's distance
+  # abline(h = 4*mean(cooksd, na.rm=T), col="red")  # add cutoff line
+  # text(x=1:length(cooksd)+1, y=cooksd, labels=ifelse(cooksd>4*mean(cooksd, na.rm=T),names(cooksd),""), col="red")
+
+  influential <- as.numeric(names(cooksd)[(cooksd > 4*mean(cooksd, na.rm=T))])
+  rows.to.keep<-which(rownames(final_sheet) %in% influential)
+  ifelse(length(influential > 0), dfsubset <- final_sheet[-rows.to.keep,], dfsubset <- final_sheet)
+  
+  mod <-  lmer(ratio ~ 1 + (1|geographic.area/plot.name), data = dfsubset, REML = F)
+  mod.linear <- lmer(ratio ~ poly(fixed, 1, raw = T) + (1|geographic.area/plot.name), data = dfsubset, REML = F)
+  
+  best.model <- as.character(aictab(list(mod = mod, mod.linear = mod.linear), sort = T)$Modname[1])
+  
+  if (best.model == "mod.linear") mod.full <- lmer(ratio ~ poly(fixed, 1, raw = T) + (1|geographic.area/plot.name), data = dfsubset, REML = F)
+  if (best.model == "mod") mod.full <- lmer(ratio ~ poly(fixed, 1, raw = T) + (1|geographic.area/plot.name), data = dfsubset, REML = F)
+  
+  
   significant.effect <- anova(mod, mod.full)$"Pr(>Chisq)"[2] < 0.05
   significance <- anova(mod, mod.full)$"Pr(>Chisq)"[2]
-  sample.size <- length(final_sheet$ratio)
+  sample.size <- length(dfsubset$ratio)
   
[…]

### 2020-08-05 manuscript/appendix.Rmd
Corrected table S5
ratio plots updated in SI

@@
 Best models by carbon flux.
-Models where $\\Delta$AIC </= 2 are presented", caption.short = "Table S5.
+Where $\\Delta$AIC values </= 2 all models with a $\\Delta$AIC value </= 2 are presented", caption.short = "Table S5.
 Best models by carbon flux", escape = F) %>%
@@
 
+\blandscape
 ```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S3: Ratios among forest C fluxes as a function of latitude and climate variables", out.width ='80%', fig.pos='H'}
@@
 ```
+\elandscape
 

### 2020-08-05 manuscript/manuscript draft.Rmd
minor update
methods +results on ratio comparisons

@@
 Geographic areas--*i.e.*, spatially clustered sites--are defined within ForC using a hierarchical cluster analysis on the distance matrix of the sites and a cutoff of 25km [@anderson-teixeira_forc_2018].
-We experimented with inclusion of altitude as a fixed effect, but excluded it from the final models because it added very little explanatory power--that is, the difference in AIC ($\Delta AIC$) relative to models excluding altitude was generally small (often $\Delta AIC$<2).
-Effects were considered significant when inclusion of the fixed of interest resulted in $\Delta AIC \ge 2.0$ relative to a corresponding null model. 
+We experimented with inclusion of altitude as a fixed effect, but excluded it from the final models because it added very little explanatory power--that is, the difference in AIC ($\Delta$ AIC) relative to models excluding altitude was generally small (often $\Delta$ AIC<2).
+Effects were considered significant when inclusion of the fixed of interest resulted in $\Delta$ AIC $\ge$ 2.0 relative to a corresponding null model. 
 All $R^2$ values presented here are marginal $R^2$ values, and refer to the proportion of variation explained by only the fixed effects. 
@@
 For brevity, we henceforth refer to first-order linear models as "linear" and second-order polynomial models as "polynomial".
-We selected as the best model that with the highest $\Delta AIC$ relative to a null model with no fixed term, with the qualification that a polynomial model was considered an improvement over a linear model only if it reduced $\DeltaAIC$ by 2.0 or more.
+We selected as the best model that with the highest $\Delta$ AIC relative to a null model with no fixed term, with the qualification that a polynomial model was considered an improvement over a linear model only if it reduced $\Delta$ AIC by 2.0 or more.
 
 To test whether trends in component fluxes across latitude sum to match those of larger fluxes, regression lines for smaller component fluxes were summed to generate new estimates of larger fluxes. 
+Because no fluxes were signficantly better predicted by a logarithmic or polynomial fit than by a linear fit, we used linear fits for all fluxes.
 We then determined whether these summed predictions fell within the **95%** CI for the larger flux across the entire range included latitude. **(Becky, is 95% the correct number)**
@@
 
-Variation in allocation to component carbon fluxes along latitudinal gradients was explored for the following pairings: $NPP:GPP$, $ANPP:BNPP$, $ANPP:NPP$, $BNPP:NPP$, and $ANPP_{foliage}:ANPP_{stem}$.
-**[issue 91](https://github.com/forc-db/Global_Productivity/issues/91)**
-For each set of paired fluxes, measurements taken at the same site and plot, and in the same year, were paired together, and the ratio of each pair of measurements calculated.
-The ratios were regressed against latitude and climate variables, using the linear model specified above.
-Cook's distance analyses were carried out for each of the models, and indicated that data from a few high-elevation sites were having a disproportionate influence on the regressions.
-Thus, models were re-run using only data from sites $\le$ 1000m.
+Variation in allocation to component carbon fluxes was explored for three groupings: (1) $GPP = NPP + R_{auto}$, (2) $NPP = ANPP + BNPP$, and (3) $ANPP = ANPP_{foliage} + ANPP_{stem}$.
+For each group, measurements taken at the same site and plot, an in the same year were grouped together.
+For groups (1) and (2), where 2 of the 3 flux measurements were available for a given site, plot, and year, these measurements were used to calculate the third.
+The ratio of each pair of component fluxes was calculated.
+The log of these ratios were regressed against latitude and climate variables, using the linear model specified above.
+Cook's distance analyses were carried out for each of the models, and extreme outliers removed,
 
@@
 We tested first-order linear, second-order polynomial, and logarithmic fits for each climate variable. 
-Again, polynomial fits were considered superior to first-order linear fits only if inclusion of a second-order polynomial term resulted in $\Delta AIC \ge 2.0$ relative to a first-order linear model.
+Again, polynomial fits were considered superior to first-order linear fits only if inclusion of a second-order polynomial term resulted in $\Delta$ AIC $\ge$ 2.0 relative to a first-order linear model.
 We tested relationships of each C flux (Table 2) against each climate variable (Table S1), but focus presentation--including focal hypotheses (Table 1)--on climate variables that explained >20% of variation in C fluxes.
@@
 Multivariate models were used to investigate the potential joint and interactive effects of MAT and MAP on carbon fluxes.
-An additive model including MAP in addition to MAT was accepted when $\Delta AIC$ >2 relative to a null including only MAT as a fixed effect.
+An additive model including MAP in addition to MAT was accepted when $\Delta$ AIC >2 relative to a null including only MAT as a fixed effect.
 An interactive model including an MAT x MAP interaction was accepted when $\Delta$AIC >2 relative to a null including MAT and MAP as fixed effects.
@@
 
-There was little evidence of systematic variation in C allocation with latitude or climate (Fig.
+We found no evidence of systematic variation in C allocation with latitude or climate (Fig.
 S3).
-Of 28 relationships tested (7 ratios among C flux variables regressed against latitude, MAT, MAP, and temperature seasonality), there were only five significant relationships, all with $R^2 \le 0.4$ (Fig.
-S3).
-**The proportion of NPP allocated to $ANPP_{foliage}$ decreased with latitude ($R^2=0.32$) and increased with MAT ($R^2=0.37$).
-The proportion of $NPP$ allocated aboveground ($ANPP$) decreased weakly with latitude ($R^2=0.11$) and temperature seasonality ($R^2=0.17$), while increasing with MAT ($R^2=0.11$).
-**
+Of 12 relationships tested (3 ratios among C flux variables regressed against latitude, MAT, MAP and temperature seasonality), none were significant.
 

### 2020-08-05 manuscript/manuscript draft.Rmd
explanation of confidence intervals

@@
 We experimented with inclusion of altitude as a fixed effect, but excluded it from the final models because it added very little explanatory power--that is, the difference in AIC ($\Delta$ AIC) relative to models excluding altitude was generally small (often $\Delta$ AIC<2).
-Effects were considered significant when inclusion of the fixed of interest resulted in $\Delta$ AIC $\ge$ 2.0 relative to a corresponding null model. 
+Effects were considered significant when inclusion of the fixed effect of interest resulted in p $\le$ 0.05 and $\Delta$ AIC $\ge$ 2.0 relative to a corresponding null model. 
 All $R^2$ values presented here are marginal $R^2$ values, and refer to the proportion of variation explained by only the fixed effects. 
@@
 For brevity, we henceforth refer to first-order linear models as "linear" and second-order polynomial models as "polynomial".
-We selected as the best model that with the highest $\Delta$ AIC relative to a null model with no fixed term, with the qualification that a polynomial model was considered an improvement over a linear model only if it reduced $\Delta$ AIC by 2.0 or more.
+We selected as the best model that with the highest $\Delta$ AIC relative to a null model with no fixed term, with the qualification that a polynomial model was considered an improvement over a linear model only if it reduced the AIC value by 2.0 or more.
 
@@
 Because no fluxes were signficantly better predicted by a logarithmic or polynomial fit than by a linear fit, we used linear fits for all fluxes.
-We then determined whether these summed predictions fell within the **95%** CI for the larger flux across the entire range included latitude. **(Becky, is 95% the correct number)**
-Confidence intervals for the larger flux were calculated using the 'bootMer' function from the lme4 package [@bates_fitting_2015].
-**(Helene asks for more details here: what does the function do?
-are CIs just based on uncertainty in regression?)**
+We then determined whether these summed predictions fell within the 95% CI for the larger flux across the entire latitudinal range.
+Confidence intervals for the line of best fit for the larger flux were estimated using the 'bootMer' function, a parametric bootstrapping method for mixed models [@bates_fitting_2015].
+We carried out 2000 simulations estimating the line of best fit, and took quantiles at 0.025 and 0.975 to estimate 95% CIs.
+**Krista please read this + confirm explanation is clear!**
 This analysis was applied to the following sets of fluxes: (1) $GPP = NPP + R_{auto}$, (2) $NPP = ANPP + BNPP$, and (3) $ANPP = ANPP_{foliage} + ANPP_{stem}$.

### 2020-08-05 scripts/stacked_plots_code.R
Update stacked_plots_code.R
correct models

@@ -143,8 +143,8 @@ par(mfrow = c(2,2), mar = c(2,2,2,2), oma = c(3,3,0,0))
       # ylim.1 <- range(ForC_simplified[ForC_simplified$variable.name %in% unlist(response.variables.1),]$mean)
       
       mod <-  lmer(mean ~ 1 + (1|geographic.area/plot.name), data = df.sum, REML = F)
-      mod.linear <- lmer(mean ~ poly(fixed, 1) + masl + (1|geographic.area/plot.name), data = df.sum, REML = F)
-      mod.poly <- lmer(mean ~ poly(fixed, 2) + masl + (1|geographic.area/plot.name), data = df.sum, REML = F)
+      mod.linear <- lmer(mean ~ poly(fixed, 1) + (1|geographic.area/plot.name), data = df.sum, REML = F)
+      mod.poly <- lmer(mean ~ poly(fixed, 2) + (1|geographic.area/plot.name), data = df.sum, REML = F)
       
       aictab <- aictab(list(mod.linear = mod.linear, mod.poly = mod.poly), sort = T)
       
@@ -152,16 +152,16 @@ par(mfrow = c(2,2), mar = c(2,2,2,2), oma = c(3,3,0,0))
       delta.aic <- as.numeric(aictab(list(mod.linear = mod.linear), sort = T)$Delta_AICc[2])
       delta.aic <- signif(delta.aic, digits=4)
       
-      if (best.model == "mod.poly") sum.mod <- lmer(mean ~ poly(fixed, 2) + masl + (1|geographic.area/plot.name), data = df.sum, REML = F)
-      if (best.model == "mod.linear") sum.mod <- lmer(mean ~ poly(fixed, 1) + masl + (1|geographic.area/plot.name), data = df.sum, REML = F)
-      if (best.model == "mod") sum.mod <- lmer(mean ~ poly(fixed, 1) + masl + (1|geographic.area/plot.name), data = df.sum, REML = F)
+      if (best.model == "mod.poly") sum.mod <- lmer(mean ~ poly(fixed, 2) + (1|geographic.area/plot.name), data = df.sum, REML = F)
+      if (best.model == "mod.linear") sum.mod <- lmer(mean ~ poly(fixed, 1) + (1|geographic.area/plot.name), data = df.sum, REML = F)
+      if (best.model == "mod") sum.mod <- lmer(mean ~ poly(fixed, 1) + (1|geographic.area/plot.name), data = df.sum, REML = F)
       
[…]

### 2020-08-05 manuscript/manuscript draft.Rmd
clarified growing season analyses, updated some results

@@
 We tested relationships of each C flux (Table 2) against each climate variable (Table S1), but focus presentation--including focal hypotheses (Table 1)--on climate variables that explained >20% of variation in C fluxes.
+
 **Becky, please make this more specific.
@@
 
-To test whether and how C flux varied with climate when standardised by growing season length (Q5), we first standardized annual C flux by dividing by growing season length (as defined above).
-We then tested for correlations between these standardised fluxes and growing season climate variables, using only first-order linear models.
+To test whether and how C flux varied with climate when standardised by growing season length (Q5), we first standardized all annual C fluxes by dividing by growing season length (as defined above).
+We then derived four variables to describe growing season climate, specifically growing season temperature, precipitation, solar radiation, and PET (Table S1).
+We tested for correlations between these standardised fluxes and growing season climate variables, using only first-order linear models.
 
@@
 
-In total, we analyzed 1228 records from nine forest autotrophic C flux variables taken from forests that had experienced no major anthropogenic disturbances within the past 100 years.
+In total, we analyzed 1319 records from nine forest autotrophic C flux variables taken from forests that had experienced no major anthropogenic disturbances within the past 100 years.
 These records represented a total of **#** plots in 154 distinct geographic areas across all forested biogeographic and climate zones (Fig.
@@
 2; Table S2).
-Latitude was a strong predictor for many of the carbon fluxes--particularly the larger fluxes (Table S2).
-Specifically, latitude explained 64% of variation in GPP (n = 254, p<0.0001), 50% in NPP (n = 114, p<0.0001) and 45% in ANPP (n = 259, p<0.0001).
+Latitude was a strong predictor for many of the carbon fluxes, particularly the larger fluxes (Table S2).
+Specifically, latitude explained 64% of variation in GPP (n = 243, p<0.0001), 50% in NPP (n = 161, p<0.0001) and 44% in ANPP (n = 278, p<0.0001).
 The C fluxes that were most poorly predicted by latitude were $BNPP_{fine.root}$ ($R^2$=0.17) and $ANPP_{stem}$ ($R^2$=0.18).
-For all C fluxes, the relationship with latitude was better fit by the first-order linear model than by second-order polynomial or logarithmic models.
+The relationship with latitude was best fit by the first-order linear model, with the exception of NPP and $R_{root}$, which were best explained by the logarithmic model.
 
 ```{r echo=FALSE, out.width='100%', fig.cap = "Latitudinal trends in forest autotropic carbon flux.
-Plotted are linear models, all of which were significant $(p<0.05)$ and had AIC values within 2.0 of the best model (for three fluxes, logarithmic fits were marginally better; Table S2).
+Plotted are linear models, all of which were significant $(p<0.05)$ and had AIC values within 2.0 of the best model (for two fluxes, logarithmic fits were marginally better; Table S2).
 Each panel shows major C fluxes together with component fluxes.
@@
 2).
-That is, modelled estimates of $GPP$, generated from the sum of $NPP$ and $R_{auto}$; $NPP$, generated from the sum of $ANPP$ and $BNPP$; and $ANPP$, generated from the sum of $ANPP_{foliage}$ and $ANPP_{stem}$, fell completely within the confidence intervals of the regressions of field estimates of $GPP$, $NPP$, and $ANPP$, respectively.
+That is, modelled estimates of $GPP$, generated from the sum of $NPP$ and $R_{auto}$; $NPP$, generated from the sum of $ANPP$ and $BNPP$; and $ANPP$, generated from the sum of $ANPP_{foliage}$ and $ANPP_{stem}$, fell almost completely within the confidence intervals of the regressions of field estimates of $GPP$, $NPP$, and $ANPP$, respectively.
 
@@
 For eight of the nine fluxes, this relationship was linear.
-For only one variable, $BNPP$, did a lognormal fit provide signficant improvement over a first-order linear relatinship. 
-*As with latitude, MAT tended to explain more variation in the larger flux ($GPP$, $NPP$, $ANPP$, $R_{auto}$) and $ANPP_{foliage}$ (all $R^2$> 0.4) than in subsidiary and belowground fluxes ($ANPP_{stem}$, $R_{root}$, $BNPP_{fine.root}$; all $R^2$< 0.25).* **update this**
+For only one variable, $BNPP$, did a lognormal fit provide signficant improvement over a first-order linear relationship. 
+*As with latitude, MAT tended to explain more variation in the larger flux ($GPP$, $NPP$, $ANPP$, $R_{auto}$) and $ANPP_{foliage}$ (all $R^2$> 0.4) than in subsidiary and belowground fluxes ($ANPP_{stem}$, $R_{root}$, $BNPP_{fine.root}$; all $R^2$< 0.25).* **update this NB these values are correct**
 
@@
 There was a significant additive effect of MAT and MAP on $GPP$, $ANPP$ and $R_{auto}$ (Fig.
-3, Table S3).
-Accounting for MAT, MAP had a substantial positive effect on $GPP$ and $R_{auto}$ and a small negative effect on $ANPP$.
-There was a significant interaction between MAT and MAP for $NPP$ and $ANPP_{stem}$ (Fig.
+3, Table S3), and a significant interactive effect between MAT and MAP for $NPP$ and $ANPP_{stem}$ (Fig.
 3, Table S3).
@@
 3, Table S3).
-For the variables which showed a significant interactive or additive effect between MAT and MAP, no other climate variable, in combination with MAT, significantly improved on that model.
-{**need to confirm this given changes in MAT MAP results (or you could just drop the sentence.)**}
-**BBL: this paragraph is pretty dense. Perhaps distill to a summary sentence and move rest to SI?**
 
@@
 
-All C flux variables were signficantly related to potential evapotranspiration.
+All C flux variables were significantly related to potential evapotranspiration.
 $ANPP_{foliage}$, $BNPP_{fine.root}$ and $R_{root}$ increased linearly with PET; however, all other fluxes showed a polynomial relationship with PET (Fig.
@@
 Vapour pressure deficit was a significant predictor of all C fluxes.
-$BNPP_{fine.root}$ showed a linear relationship with vapour pressure deficit ($R^2$ = 0.07, p<0.05), but all other fluxes showed a polynomial relationship (Figs. 4d, S4-5; Table S2). 
+$BNPP_{fine.root}$ showed a linear relationship with vapour pressure deficit ($R^2$ = 0.07, p<0.05), but all other fluxes showed a polynomial relationship (Figs.
+4d, S4-5; Table S2).
 C fluxes initially increased with vapour pressure deficit, before saturating at around 0.8 kPa, after which point they began to decrease.
@@
 
-Temperature seasonality was, overall, the best predictor of annual C fluxes. **(update this paragraph based on new results)**
+Temperature seasonality was a strong predictor of annual C fluxes. $GPP$ showed a polynomial relationship with seasonality **(update this paragraph based on new results)**
 $GPP$, $NPP$, $ANPP$, and $R_{root}$ exhibited a polynomial relationship with seasonality (all p<0.05; Figs.
@@
 4e, S6-S7; Table S2).
-Length of growing season was a strong predictor of C fluxes, explaining 51% of variation in GPP, 39% of variation in NPP, and 34% of variation in ANPP, but it was a weaker predictor than MAT for all fluxes analysed (Table S4).
+Length of growing season was a strong predictor of C fluxes, explaining 53% of variation in GPP, 38% of variation in NPP, and 34% of variation in ANPP (all p<0.05; Talbe S2), but it was a weaker predictor than MAT for all fluxes analysed (Table S4).
 
@@
 
-When annual C fluxes were standardized by growing season length (in monthly increments), correlations with growing season climate--including temperature, precipitation, solar radiation, and PET--were generally weak (Figs.
+When annual C fluxes were standardized by growing season length (in monthly increments), correlations with growing season climate were generally weak (Figs.
 S8-S9).
-Specifically, $ANPP$ increased with growing season temperature ($R^2$ = 0.10, p<0.001) and precipitation ($R^2$ = 0.04, p<0.05).
+Specifically, $ANPP$ increased with growing season temperature ($R^2$ = 0.09, p<0.001) and precipitation ($R^2$ = 0.04, p<0.05).
 Similarly, $ANPP_{foliage}$ increased slightly with growing season temperature ($R^2$ = 0.16, p<0.01) and precipitation ($R^2$ = 0.09, p<0.05).
-Growing season solar radiation had a positive influence on $GPP$ ($R^2$ = 0.21, p<0.001), $NPP$ ($R^2$ = 0.21, p<0.001), $BNPP$ ($R^2$ = 0.16, p<0.001) and $BNPP_{fine.root}$ ($R^2$ = 0.12, p<0.01).
-Growing season PET had a positive influence on $GPP$ ($R^2$ = 0.15, p<0.01), $NPP$ ($R^2$ = 0.18, p<0.01), $BNPP$ ($R^2$ = 0.23, p<0.0001), $BNPP_{fine.root}$ ($R^2$ = 0.11, p<0.05), and $ANPP_{stem}$ ($R^2$ = 0.06, p<0.05).
-{**Becky, please verify/ edit the following:** There were no other signifcant correlations between growing season length-standardized C fluxes (9 variables in Table 2) and growing season climate (**which variables?**)}.
+Growing season solar radiation had a positive influence on $BNPP$ ($R^2$ = 0.17, p<0.001) and $BNPP_{fine.root}$ ($R^2$ = 0.13, p<0.01).
+Growing season PET had a positive influence on $GPP$ ($R^2$ = 0.15, p<0.01), $NPP$ ($R^2$ = 0.07, p<0.01), $BNPP$ ($R^2$ = 0.23, p<0.0001), $BNPP_{fine.root}$ ($R^2$ = 0.10, p<0.05), and $ANPP_{stem}$ ($R^2$ = 0.06, p<0.05).
+All other relationships were non-significant.
 

### 2020-08-05 manuscript/manuscript draft.Rmd
slightly condensed results

@@
 Again, polynomial fits were considered superior to first-order linear fits only if inclusion of a second-order polynomial term resulted in $\Delta$ AIC $\ge$ 2.0 relative to a first-order linear model.
-We tested relationships of each C flux (Table 2) against each climate variable (Table S1), but focus presentation--including focal hypotheses (Table 1)--on climate variables that explained >20% of variation in C fluxes.
-
-**Becky, please make this more specific.
-I drafted this based on the results, but its not very clear.** **BBL: I agree, not clear and might trigger reviewer concerns about tailoring hypotheses to significance.**
+We tested relationships of each C flux (Table 2) against each climate variable (Table S1).
+Variables which were not significant explanatory variables or which explained <20% of variation in C fluxes are only presented in SI.
 
@@
 For only one variable, $BNPP$, did a lognormal fit provide signficant improvement over a first-order linear relationship. 
-*As with latitude, MAT tended to explain more variation in the larger flux ($GPP$, $NPP$, $ANPP$, $R_{auto}$) and $ANPP_{foliage}$ (all $R^2$> 0.4) than in subsidiary and belowground fluxes ($ANPP_{stem}$, $R_{root}$, $BNPP_{fine.root}$; all $R^2$< 0.25).* **update this NB these values are correct**
+*As with latitude, MAT tended to explain more variation in the larger flux ($GPP$, $NPP$, $ANPP$, $R_{auto}$) and $ANPP_{foliage}$ (all $R^2$> 0.4) than in subsidiary and belowground fluxes ($ANPP_{stem}$, $R_{root}$, $BNPP_{fine.root}$; all $R^2$< 0.25).* **update this -- NB these values are correct**
 
@@
 However, it explained little variation: with the exception of $R_{auto}$, MAP explained at most 25% of variation in C flux. 
-For five of the nine fluxes, a second-order polynomial model provided the best fit, while first-order linear and lognormal relationships provided each provided the best fit for two flxues. 
+For five of the nine fluxes, a second-order polynomial model provided the best fit, while first-order linear and lognormal relationships each provided the best fit for two flxues. 
 All fluxes increased with MAP up to at least 2000 mm, above which responses were variable (Figs. 4, S4-S5).
@@
 All C flux variables were significantly related to potential evapotranspiration.
-$ANPP_{foliage}$, $BNPP_{fine.root}$ and $R_{root}$ increased linearly with PET; however, all other fluxes showed a polynomial relationship with PET (Fig.
+$ANPP_{foliage}$, $BNPP_{fine.root}$ and $R_{root}$ showed a logarithmic relationship with PET; however, all other fluxes showed a polynomial relationship with PET (Fig.
 4c, S4-5; Table S2).
@@
 Vapour pressure deficit was a significant predictor of all C fluxes.
-$BNPP_{fine.root}$ showed a linear relationship with vapour pressure deficit ($R^2$ = 0.07, p<0.05), but all other fluxes showed a polynomial relationship (Figs.
+$ANPP_{foliage}$, $BNPP_{fine.root}$ and $R_{root}$ showed a logarithmic relationship with vapour pressure deficit, but all other fluxes showed a polynomial relationship (Figs.
 4d, S4-5; Table S2).
@@
 
-All fluxes, with the exception of $R_{root}$, showed a positive linear relationship with solar radiation (Figs.
+All fluxes, with the exception of $R_{root}$, showed a significant positive relationship with solar radiation (Figs.
 S4-S5, Table S2).
-Solar radiation explained a low proportion of variability in all C fluxes, explaining less than 20% of the variation in each flux, with the exception of $R_{auto}$ ($R^2$ = 0.26, p<0.05).
+Solar radiation explained a low proportion of variability in all C fluxes, explaining less than 30% of the variation in each flux
 
-Annual wet days, cloud cover, and aridity were poor or non-significant explainers of variation in C fluxes, explaining less than 20% of the variation in each of the carbon fluxes (Figs.
+Annual wet days, cloud cover, aridity, and water stress months were poor or non-significant explainers of variation in C fluxes, explaining less than 20% of the variation in each of the carbon fluxes (Figs.
 S4-S5; Table S2).
@@
 
-Temperature seasonality was a strong predictor of annual C fluxes. $GPP$ showed a polynomial relationship with seasonality **(update this paragraph based on new results)**
-$GPP$, $NPP$, $ANPP$, and $R_{root}$ exhibited a polynomial relationship with seasonality (all p<0.05; Figs.
+Temperature seasonality was a strong predictor of annual C fluxes.
+All fluxes decrease with increasing seasonality, though the shape of this relationship varies (all p<0.05; Figs.
 4e, S6-7; Table S2).
-$ANPP_{foliage}$, $ANPP_{stem}$ and $R_{auto}$ decreased linearly with temperature seasonality (all p<0.05; Figs.
-4e, S6-S7; Table S2).
 Temperature seasonality was strongly correlated with annual temperature range, which was likewise a similarly strong predictor of C fluxes (Table S2).

### 2020-08-05 manuscript/manuscript draft.Rmd
slightly condensed results

@@
 In total, we analyzed 1319 records from nine forest autotrophic C flux variables taken from forests that had experienced no major anthropogenic disturbances within the past 100 years.
-These records represented a total of **#** plots in 154 distinct geographic areas across all forested biogeographic and climate zones (Fig.
+These records represented a total of 255 plots in 154 distinct geographic areas across all forested biogeographic and climate zones (Fig.
 1, Table 2).
@@
 
-All major carbon fluxes decreased linearly with latitude (Fig.
+All major carbon fluxes decreased with latitude (Fig.
 2; Table S2).
@@
 For only one variable, $BNPP$, did a lognormal fit provide signficant improvement over a first-order linear relationship. 
-*As with latitude, MAT tended to explain more variation in the larger flux ($GPP$, $NPP$, $ANPP$, $R_{auto}$) and $ANPP_{foliage}$ (all $R^2$> 0.4) than in subsidiary and belowground fluxes ($ANPP_{stem}$, $R_{root}$, $BNPP_{fine.root}$; all $R^2$< 0.25).* **update this -- NB these values are correct**
+*As with latitude, MAT tended to explain more variation in the larger fluxes ($GPP$, $NPP$, $ANPP$, $R_{auto}$) and $ANPP_{foliage}$ (all $R^2$> 0.4) than in subsidiary and belowground fluxes ($ANPP_{stem}$, $R_{root}$, $BNPP_{fine.root}$; all $R^2$< 0.25).* **update this -- NB these values are correct**
 
@@
 However, it explained little variation: with the exception of $R_{auto}$, MAP explained at most 25% of variation in C flux.
-For five of the nine fluxes, a second-order polynomial model provided the best fit, while first-order linear and lognormal relationships each provided the best fit for two flxues. 
-All fluxes increased with MAP up to at least 2000 mm, above which responses were variable (Figs. 4, S4-S5).
-
+All fluxes increased with MAP up to at least 2000 mm, above which responses were variable (Figs.
+4, S4-S5).
 There was a significant additive effect of MAT and MAP on $GPP$, $ANPP$ and $R_{auto}$ (Fig.
@@
 
-Our results indicated that annual forest C fluxes were most strongly explained by temperature at the global scale, with temperature-related climate variables coming out as strong predictors of C fluxes.
+Our results indicated that annual forest C fluxes were most strongly explained by temperature-related climate variables at the global scale.
 In addition to MAT, several of its correlates (Fig.
@@
 
-All C flux variables were significantly related to potential evapotranspiration.
-$ANPP_{foliage}$, $BNPP_{fine.root}$ and $R_{root}$ showed a logarithmic relationship with PET; however, all other fluxes showed a polynomial relationship with PET (Fig.
+All C flux variables showed a significant relationship with potential evapotranspiration.
+The relationship was logarithmic for $ANPP_{foliage}$, $BNPP_{fine.root}$ and $R_{root}$, and polynomial for all other fluxes (Fig.
 4c, S4-5; Table S2).
@@
 
-We found a significant relationship between length of growing season and C fluxes, with all fluxes showing a linear increase with length of growing season (Figs.
+We found a significant relationship between length of growing season and C fluxes, with all fluxes showing a positive relationship with length of growing season (Figs.
 4e, S6-S7; Table S2).
-Length of growing season was a strong predictor of C fluxes, explaining 53% of variation in GPP, 38% of variation in NPP, and 34% of variation in ANPP (all p<0.05; Talbe S2), but it was a weaker predictor than MAT for all fluxes analysed (Table S4).
+Length of growing season was a strong predictor of C fluxes, explaining 53% of variation in GPP, 38% of variation in NPP, and 34% of variation in ANPP (all p<0.05; Table S2), but it was a weaker predictor than MAT for all fluxes analysed (Table S4).
 
@@
 S8-S9).
-Specifically, $ANPP$ increased with growing season temperature ($R^2$ = 0.09, p<0.001) and precipitation ($R^2$ = 0.04, p<0.05).
+$ANPP$ increased with growing season temperature ($R^2$ = 0.09, p<0.001) and precipitation ($R^2$ = 0.04, p<0.05).
 Similarly, $ANPP_{foliage}$ increased slightly with growing season temperature ($R^2$ = 0.16, p<0.01) and precipitation ($R^2$ = 0.09, p<0.05).
@@
 2; Table S2).
-
 Latitudinal variation in C fluxes is primarily attributable to temperature-related variables (*H3, H4*), particularly MAT (Figs.

### 2020-08-05 manuscript/manuscript draft.Rmd
updates including hypothesis table

@@
 and with observations and theory predicting relatively higher belowground carbon allocation at higher latitudes [@gill_belowground_2016].
+**Krista -- I took out the ANPP:NPP graphs from the SI after looking back at what Helene said in issue #77 -- let me know if you think it's worth putting them back in wrt this point**
 It is also consistent with findings that as forest productivity increases, proportionaly more carbon is allocated to $ANPP_{foliage}$ relative to $BNPP_{fine.root}$ [@chen_allometry_2019] or $ANPP_{stem}$ [@hofhansl_new_2015].

### 2020-08-05 manuscript/manuscript draft.Rmd
knit latest

@@
 ```{r setup, include = FALSE}
-knitr::opts_knit$set(root.dir = 'C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/')
- # knitr::opts_knit$set(root.dir = '/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/Global_Productivity/manuscript/')
+#knitr::opts_knit$set(root.dir = 'C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/')
+  knitr::opts_knit$set(root.dir = '/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/Global_Productivity/manuscript/')
 ```
@@
 To test whether trends in component fluxes across latitude sum to match those of larger fluxes, regression lines for smaller component fluxes were summed to generate new estimates of larger fluxes. 
-Because no fluxes were signficantly better predicted by a logarithmic or polynomial fit than by a linear fit, we used linear fits for all fluxes.
+Because no fluxes were significantly better predicted by a logarithmic or polynomial fit than by a linear fit, we used linear fits for all fluxes.
 We then determined whether these summed predictions fell within the 95% CI for the larger flux across the entire latitudinal range.

### 2020-08-06 manuscript/manuscript draft.Rmd

@@
 We show that, across all flux variables analyzed, C cycling decreases continuously with absolute latitude -- a finding that confirms multiple previous studies but contradicts the idea that net primary productivity ($NPP$) of temperate forests rivals that of tropical forests.
-C flux variables generally displayed similar trends across latitude and multiple climate variables, with few differences in allocation detectable at this global scale.
-**Climate explained a significant proportion (24 - 71%) of variation in all C fluxes analyzed**, with temperature variables in general and mean annual temperature (MAT) in particular being the best predictors of C flux on this global scale.
+C flux variables generally displayed similar trends across latitude and multiple climate variables, with no differences in allocation detected at this global scale.
+Temperature variables in general, and mean annual temperature ($MAT$) and temperature seasonality in particular, were the best univariate predictors of C flux, explaining 19 - 71% of variation in the C fluxes analyzed.
 The effects of temperature were modified by moisture availability, with C flux reduced under hot and dry conditions and sometimes under very high precipitation. 
-C fluxes declined with temperature seasonality, but growing season length did not improve upon MAT as a predictor.
+C fluxes increased with growing season length, but this was never the best univariate predictor.
 Within the growing season, the influence of climate on C cycling was small but significant for a number of flux variables. 
 These findings clarify how forest C flux varies with latitude and climate on a global scale. 
-In a period of accelerating climatic change, understanding of the fundamental climatic controls on forest C cycling sets a foundation for understanding patterns of change.
+In a period of accelerating climatic change, this improved understanding of the fundamental climatic controls on forest C cycling sets a foundation for understanding patterns of change.
 
@@
 The latitudinal gradient in forest C flux rates, along with altitudinal gradients [@girardin_net_2010; @malhi_variation_2017], is driven primarily by climate, which is a significant driver of C fluxes across broad spatial scales [@luyssaert_co_2007; @cleveland_relationships_2011; @hursh_sensitivity_2017].
+However, there is little consensus as to the shapes of these relationships or the best predictor variables.
 The majority of studies have focused on exploring the relationships of C fluxes to mean annual temperature (MAT) and mean annual precipitation (MAP), as the most commonly reported site-level climate variables.
-C fluxes increase strongly with $MAT$ on the global scale, but whether they saturate or potentially decrease at higher temperatures remains disputed. (**This part could use some work.**)
-In the global data compilation of Luyssaert et al. [-@luyssaert_co_2007], $GPP$ increased linearly with MAT, whereas $NPP$ saturated at around 10$^\circ$C MAT. 
-In contrast, Larjavaara and Muller-Landau [-@larjavaara_temperature_2012], find that increases in $GPP$ saturate at approximately 25$^\circ$C MAT, while Schuur [-@schuur_productivity_2003] finds that $NPP$ increases continuously with temperature.
-@sullivan_sensitivity_2020 found that, within the tropics, $ANPP_{stem}$ decreases at the highest maximum temperatures.
+C fluxes increase strongly with $MAT$ on the global scale, but whether they saturate or potentially decrease at higher temperatures remains disputed. 
+Some studies have detected no deceleration or decline in $GPP$ [@luyssaert_co_2007] or $NPP$ [@schuur_productivity_2003] with increasing MAT.
+Some have suggested that relationship between C flux and temperature is exponential in nature [@allen_linking_2005] (**REFS**).
+In contrast, others have found evidence of saturation or decline of C flux in the warmest climates; @luyssaert_co_2007 found $NPP$ saturating at around 10$^\circ$C MAT; @larjavaara_temperature_2012 found that increases in $GPP$ saturate at approximately 25$^\circ$C MAT, and @sullivan_sensitivity_2020 found that, within the tropics, $ANPP_{stem}$ decreases at the highest maximum temperatures.
 Various measures of primary productivity saturate at high levels of MAP, though the saturation points identified vary from 1500mm [@luyssaert_co_2007] up to 2,445 mm MAP [@schuur_productivity_2003]. 
 Interactions between MAT and MAP are also possible; within the tropics, there is a positive interaction between MAT and MAP in shaping $ANPP$, such that high rainfall has a negative effect on productivity in cooler climates, compared to a positive effect in warmer climates [@taylor_temperature_2017].
-Such complicated dynamics play out with belowground fluxes as well, with typically dominant soil temperature mediated by moisture and C supply [@hursh_sensitivity_2017; @xu_contribution_2016]. 
-There is also evidence that C fluxes also respond to climate variables such as cloud cover [@taylor_temperature_2017], solar radiation [@beer_terrestrial_2010; @fyllas_solar_2017], and potential evapotranspiration [@kerkhoff_plant_2005]; however, these are not typically assessed in global-scale analyses of annual forest C flux.
+**Such complicated dynamics play out with belowground fluxes as well, with typically dominant soil temperature mediated by moisture and C supply [@hursh_sensitivity_2017; @xu_contribution_2016].
+https://github.com/forc-db/Global_Productivity/issues/111**
+There is also evidence that C fluxes also respond to climate variables such as temperature and precipitation seasonality (**REFS**), cloud cover [@taylor_temperature_2017], solar radiation [@beer_terrestrial_2010; @fyllas_solar_2017], and potential evapotranspiration [@kerkhoff_plant_2005]; however, these are not typically assessed in global-scale analyses of annual forest C flux.
 
@@
 [-@michaletz_convergence_2014] found no significant relationship between growing season temperature and net primary productivity (loosely defined to include a mix of **VARIABLES**) standardized to growing season length.
-While this suggests that the influence of temperature may be limited to determining the length of the frost-free growing season, *analysis with clearly defined variables and standardized forest ages would be necessary to test the the veracity and generality of this hypothesis.
+While this suggests that the influence of temperature may be limited to determining the length of the frost-free growing season, analysis with clearly defined variables and standardized forest ages would be necessary to test the the veracity and generality of this hypothesis.
 
@@
 Confidence intervals for the line of best fit for the larger flux were estimated using the 'bootMer' function, a parametric bootstrapping method for mixed models [@bates_fitting_2015].
-We carried out 2000 simulations estimating the line of best fit, and took quantiles at 0.025 and 0.975 to estimate 95% CIs.
-**Krista please read this + confirm explanation is clear!**
+This function carried out 2000 simulations estimating the line of best fit, using quantiles at 0.025 and 0.975 to estimate 95% CIs.
 This analysis was applied to the following sets of fluxes: (1) $GPP = NPP + R_{auto}$, (2) $NPP = ANPP + BNPP$, and (3) $ANPP = ANPP_{foliage} + ANPP_{stem}$.
@@
 The C fluxes that were most poorly predicted by latitude were $BNPP_{fine.root}$ ($R^2$=0.17) and $ANPP_{stem}$ ($R^2$=0.18).
-The relationship with latitude was best fit by the first-order linear model, with the exception of NPP and $R_{root}$, which were best explained by the logarithmic model.
+The relationship with latitude was best fit by the first-order linear model, with the exception of NPP and $R_{root}$, for which a logarithmic model was a slightly - but not significantly - better fit.
 
@@
 2).
-That is, modelled estimates of $GPP$, generated from the sum of $NPP$ and $R_{auto}$; $NPP$, generated from the sum of $ANPP$ and $BNPP$; and $ANPP$, generated from the sum of $ANPP_{foliage}$ and $ANPP_{stem}$, fell almost completely within the confidence intervals of the regressions of field estimates of $GPP$, $NPP$, and $ANPP$, respectively.
+That is, modeled estimates of $GPP$, generated from the sum of $NPP$ and $R_{auto}$; $NPP$, generated from the sum of $ANPP$ and $BNPP$; and $ANPP$, generated from the sum of $ANPP_{foliage}$ and $ANPP_{stem}$, fell almost completely within the confidence intervals of the regressions of field estimates of $GPP$, $NPP$, and $ANPP$, respectively.
 
@@
 This is consistent with observations of continental- to global-scale increases over time in $GPP$ [@li_mapping_2019] and $ANPP_{stem}$ [@brienen_long-term_2015; @hubau_asynchronous_2020], along with some C cycle components not considered here--tree mortality [@brienen_long-term_2015; @mcdowell_drivers_2018], soil respiration [@bond-lamberty_global_2010], and heterotrophic soil respiration [@bond-lamberty_globally_2018].
-**DISCUSS/CITE [Wu et al. 2020]** [@wu_accelerated_2020] 
 However, increasing C flux rates are by no means universal [e.g., @rutishauser_testing_2020; @hubau_asynchronous_2020], likely because other factors are at play, including changes to other aspects of climate, atmospheric pollution (CO~2~, SO~2~, NO~x~), and local disturbances.
+In addition, some tropical forests are
 **(discuss new Sullivan paper that finds higher max temperatures associated to lower ANPP_stem)** [@sullivan_sensitivity_2020]

### 2020-08-06 scripts/final_model_no_weighting.R
issue #113

@@ -115,6 +115,8 @@ all.results <- NULL
 all.aictab <- NULL
 all.koeppen <- NULL
 
+null <- NULL
+
 fixed.variables <- c("mat", "map", "lat", "TempSeasonality", "TempRangeAnnual", "AnnualFrostDays", "AnnualWetDays", "Aridity", "PotentialEvapotranspiration", "VapourPressureDeficit", "SolarRadiation", "PreSeasonality", "MaxVPD", "WaterStressMonths", "length_growing_season", "CloudCover")
 
 
@@ -297,6 +299,7 @@ for(response.variables in response.variables.groups){
         
         all.results <- rbind(all.results, results)
         all.aictab <- rbind(all.aictab, aictab)
+        null <- rbind(null, df)
         
       }
       
@@ -486,7 +489,7 @@ for (age in ages){
     mtext(paste0("(", letters[pannel.nb], ")"), side = 3, line = -1.5, adj = 0.05)
     mtext(side = 1, line = 3, text = eval(parse(text = xaxis)), outer = F)
     # mtext(side = 2, line = 3,  text = expression("Mg C"~ha^-1~yr^-1), outer = F) 
-    if(fixed.v == "map")legend("topright", legend = c("GPP", "NPP", "ANPP", "BNPP root"), col = plasma(10)[c(1, 3, 5, 8)], pch = c(1, 3, 5, 8), xpd = T, text.col = plasma(10)[c(1, 3, 5, 8)], bty = "n", xjust = 1, cex = 0.75)
+    if(fixed.v == "map")legend("topright", legend = c("GPP", "NPP", "ANPP", "BNPP"), col = plasma(10)[c(1, 3, 5, 8)], pch = c(1, 3, 5, 8), xpd = T, text.col = plasma(10)[c(1, 3, 5, 8)], bty = "n", xjust = 1, cex = 0.75)
     
     pannel.nb <- pannel.nb +1
     

### 2020-08-06 manuscript/manuscript draft.Rmd
working on discusion
(still a mess, pushing now to avoid conflicted copies)

@@
 
-Forest C fluxes decrease with latitude [e.g., @luyssaert_co_2007; @piao_forest_2010; @gillman_latitude_2015; @li_mapping_2019], but it remains unclear whether and how the shape of this relationship varies among fluxes.
-For instance, gross primary productivity ($GPP$) increases continuously with decreasing latitude and is indisputably highest in tropical forests [@luyssaert_co_2007; @beer_terrestrial_2010; @jung_global_2011; @badgley_terrestrial_2019; @li_mapping_2019].
+Forest C fluxes decrease with latitude [e.g., @luyssaert_co_2007; @piao_forest_2010; @gillman_latitude_2015; @li_mapping_2019], but studies have differed in their conclusions regarding the shape of this relationship - quite possibly because of lack of standardization with respect to methodology and stand history.
+For instance, studies agree that gross primary productivity ($GPP$) increases continuously with decreasing latitude and is indisputably highest in tropical forests [@luyssaert_co_2007; @beer_terrestrial_2010; @jung_global_2011; @badgley_terrestrial_2019; @li_mapping_2019].
 In contrast, some studies have suggested that net primary productivity ($NPP$), or its aboveground portion ($ANPP$), exhibits a less distinct increase from temperate to tropical forests [@luyssaert_co_2007]--or even a decrease [@huston_global_2009, but see @gillman_latitude_2015].
 A shallower increase in $NPP$ than in $GPP$ with decreasing latitude would align with the suggestion that tropical forests tend to have low carbon use efficiency [$CUE$= $NPP$/$GPP$; @de_lucia_forest_2007; @malhi_productivity_2012; @anderson-teixeira_carbon_2016].
-However, accurately comparing relationships of multiple forest C fluxes across latitude requires a large database with standardization for methodology, stand history, and management, which to date has not been possible.
+Such differences among C fluxes their relationship to latitude could have profound implications for our understanding of the C cycle and its climate sensitivity.
+However, until recently the potential to compare latitudinal trends across C fluxes has been limited by lack of a large database with standardization for methodology, stand history, and management [@anderson-teixeira_forc_2018].
 
@@
 
-Our analysis of a large global database (ForC) reveals how autotrophic carbon fluxes in mature forests vary with latitude and climate on a global scale.
-We show that, across all nine forest autotrophic C fluxes analyzed, C cycling decreases continually with latitude (*H1.1*; Fig.
-2)--a finding that confirms multiple previous studies but contradicts the idea that productivity of temperate forests rivals that of tropical forests [@huston_global_2009].
-C fluxes generally increase in proportion to one another (*H1.2*), with few differences in allocation detectable at this global scale (Fig.
-S2) and with component fluxes summing appropriately to larger fluxes (Fig.
-2), indicating no major, systematic omissions or overestimations of flux components.
-However, climate explained lower proportions of variability among subsidiary C fluxes (*e.g.*, $ANPP_{stem}$, $BNPP_{fine.root}$, $R_{root}$; Fig.
-2; Table S2).
-Latitudinal variation in C fluxes is primarily attributable to temperature-related variables (*H3, H4*), particularly MAT (Figs.
+**Our analysis of a large global database (ForC) clarifies how autotrophic C fluxes in mature forests vary with latitude and climate on a global scale.** 
+We show that, across all nine variables analyzed, annual C flux decreases continually with latitude (Fig.
+2)--a finding that confirms multiple previous studies but contradicts the idea that productivity of temperate forests rivals or even exceeds that of tropical forests [@luyssaert_co_2007; @huston_global_2009].
+At this global scale, C fluxes increase approximately in proportion to one another, with component fluxes summing appropriately to larger fluxes and no detectable differences in allocation across latitude or climates (Figs.
+2, 4, S3).
+Similarly, we show broad - *albeit* not complete - consistency of climate responses across C fluxes, with the observed latitudinal variation primarily attributable to temperature and its seasonality (Figs.
 3-4).
-Water availability is also influential, but generally of secondary importance across the climate space occupied by forests (Figs.
-3-4).
-Temperature seasonality and growing season length are closely correlated with MAT and are strong predictors of C fluxes (*H4*; Figs.
-4e-f, S2, S6-S7), though growing season length did not improve upon MAT as a predictor.
-Within the growing season, the influence of climate on C cycling is smaller but still significant for a number of C fluxes (*H5*; Fig.
-S9; Table S4).
+Water availability is also influential, but of secondary importance across the climate space occupied by forests (Figs. 3-4). 
+Contrary to prior suggestions that the majority of variation in C cycling is driven by the length of the growing season (**REFS**), we find modest explanatory power of growing season length and small but sometimes significant influence of climate within the growing season (Figs.
+4f,S6-S9).
+Together, these findings yield a unified understanding of climate's influence on forest C cycling.
 
-These findings clarify how forest C fluxes vary with latitude and climate on a global scale.
-Past studies have differed in their conclusions regarding the relationship between C fluxes and latitude or its correlates (Table 1, *H1*)--quite possibly because of lack of standardization with respect to stand age and disturbance history.
-Our findings indicate that, among mature, undisturbed stands, forest C fluxes are unambiguously highest in the tropical regions, and the relationship with latitude is approximately linear (Fig.
-2).
-This contrasts with the suggestion that productivity of temperate forests is similar to that of tropical forests [@huston_global_2009].
+**Our findings indicate that, among mature, undisturbed stands, forest C fluxes are unambiguously highest in the tropical regions, and the relationship with both latitude and $MAT$ is approximately linear (Figs.
+2, 4).**
+This contrasts with the suggestion that C fluxes (e.g., $NPP$, $ANPP$, $ANPP_{stem}$) of temperate forests are similar to or even greater than that of tropical forests [@luyssaert_co_2007; @huston_global_2009].
+Previous indications of such a pattern may have been an artifact of differences in stand age across biomes.
 Compared to tropical forests, the temperate forest biome has experienced more widespread anthropogenic disturbance and has a larger fraction of secondary stands [@potapov_mapping_2008; @poulter_global_2018], so analyses comparing across latitudinal gradients without controlling for stand age risk confounding age with biome effects.
-In addition, because carbon allocation varies with stand age [@de_lucia_forest_2007; @anderson-teixeira_altered_2013; @doughty_what_2018], age differences may introduce systematic biases into analyses of C fluxes across latitude or global climatic gradients.
-For example, woody productivity tends to be higher in rapidly aggrading secondary stands than in old-growth forests, where proportionally more C is allocated to respiration [@de_lucia_forest_2007; @piao_forest_2010; @doughty_what_2018; @kunert_understanding_2019].
+Because carbon allocation varies with stand age [@de_lucia_forest_2007; @anderson-teixeira_altered_2013; @doughty_what_2018], age differences may introduce systematic biases into analyses of C fluxes across latitude or global climatic gradients.
+For example, woody productivity tends to be higher in rapidly aggrading secondary stands than in old-growth forests, where proportionally more C is allocated to respiration and non-woody productivity [@de_lucia_forest_2007; @piao_forest_2010; @doughty_what_2018; @kunert_understanding_2019].
+Thus, findings that temperate forest productivity rivals that of tropical forests are likely an artifact of different forest ages across biomes.
 
-We show that C fluxes are broadly consistent in their responses to climate drivers on the global scale, with at most modest trends in C allocation among the variable pairs tested (Figs. 2, S3). 
+**We show that C fluxes are broadly consistent in their responses to climate drivers on the global scale, with at most modest trends in C allocation among the variable pairs tested (Figs. 2, S3).** 
 This parallels the observation that C allocation across multiple C fluxes varies little with respect to climate along a steep tropical elevational gradient [@malhi_variation_2017; but see @moser_elevation_2011],
@@
 **Krista -- I took out the ANPP:NPP graphs from the SI after looking back at what Helene said in issue #77 -- let me know if you think it's worth putting them back in wrt this point**
-It is also consistent with findings that as forest productivity increases, proportionaly more carbon is allocated to $ANPP_{foliage}$ relative to $BNPP_{fine.root}$ [@chen_allometry_2019] or $ANPP_{stem}$ [@hofhansl_new_2015].
+It is also consistent with findings that as forest productivity increases, proportionally more carbon is allocated to $ANPP_{foliage}$ relative to $BNPP_{fine.root}$ [@chen_allometry_2019] or $ANPP_{stem}$ [@hofhansl_new_2015].
 **( check / comment on  @litton_carbon_2007 )**
@@
 **update this paragraph ([issue 78](https://github.com/forc-db/Global_Productivity/issues/78))** 
-One interesting observation was that climate tends to explain more variation in the major fluxes ($GPP$, $NPP$, $R_{auto}$) than in subsidiary fluxes ($BNPP_{fine.root}$, $R_{root}$, $ANPP_{stem}$) as quantified by R^2^ (Fig.
-2; Table S2).
+*Despite the broad consistency of climate responses across C fluxes, climate explains lower proportions of variability among some of the subsidiary C fluxes (*e.g.*, $ANPP_{stem}$, $BNPP$  $BNPP_{fine.root)$; Fig.
+2; Tables S2, S6).*
 There are two, non-exclusive, potential explanations for this. 
@@
 
+*At least at this global scale, we find little evidence of any non-linearity in temperature's influence on C flux.* 
 The latitudinal gradient in forest C flux (Fig.

### 2020-08-06 manuscript/manuscript-draft.tex
working on discusion
(still a mess, pushing now to avoid conflicted copies)

@@
-\documentclass[]{article}
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+%
+\documentclass[
+]{article}
 \usepackage{lmodern}
@@
 \usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
 \ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
@@
   \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \ifxetex
-    \usepackage{mathspec}
-  \else
-    \usepackage{fontspec}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
 \fi
-  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
-\fi
-% use upquote if available, for straight quotes in verbatim environments
+% Use upquote if available, for straight quotes in verbatim environments
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
-\usepackage{microtype}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
   \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
 }{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\hypersetup{
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+\urlstyle{same} % disable monospaced font for URLs
 \usepackage[margin=1in]{geometry}
-\usepackage{hyperref}
-\hypersetup{unicode=true,
-            pdfborder={0 0 0},
-            breaklinks=true}
-\urlstyle{same}  % don't use monospace font for urls
-\usepackage{natbib}
-\bibliographystyle{apalike}
 \usepackage{graphicx,grffile}
@@
 \setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
-\IfFileExists{parskip.sty}{%
-\usepackage{parskip}
-}{% else
-\setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
 \setlength{\emergencystretch}{3em} % prevent overfull lines
@@
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{0}
-% Redefines (sub)paragraphs to behave more like sections
-\ifx\paragraph\undefined\else
-\let\oldparagraph\paragraph
-\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-\let\oldsubparagraph\subparagraph
-\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
-
-%%% Use protect on footnotes to avoid problems with footnotes in titles
-\let\rmarkdownfootnote\footnote%
-\def\footnote{\protect\rmarkdownfootnote}
-
-%%% Change title format to be more compact
-\usepackage{titling}
-
-% Create subtitle command for use in maketitle
-\providecommand{\subtitle}[1]{
-  \posttitle{
-    \begin{center}\large#1\end{center}
-    }
-}
-
-\setlength{\droptitle}{-2em}
-
-  \title{}
-    \pretitle{\vspace{\droptitle}}
-  \posttitle{}
-    \author{}
-    \preauthor{}\postauthor{}
-    \date{}
-    \predate{}\postdate{}
-
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+\usepackage{float} \usepackage{caption} \captionsetup[table]{font=footnotesize} \captionsetup[figure]{font=footnotesize} \usepackage{setspace}\onehalfspacing \usepackage{lineno}\linenumbers
 \usepackage{booktabs}
@@
 \usepackage{xcolor}
+\usepackage[]{natbib}
+\bibliographystyle{apalike}
 
-\usepackage{float} \usepackage{caption} \captionsetup[table]{font=footnotesize} \captionsetup[figure]{font=footnotesize} \usepackage{setspace}\onehalfspacing \usepackage{lineno}\linenumbers
+\author{}
+\date{\vspace{-2.5em}}
 
@@
   Conservation Ecology Center; Smithsonian Conservation Biology
-  Institute; National Zoological Park, Front Royal, VA, USA
+  Institute; Front Royal, VA, USA
 \item
@@
 
-fax:1-540-635-6506
+\url{fax:1-540-635-6506}
 
@@
 
-\subsubsection{Abstract}\label{abstract}
-
-{[}\textbf{very rough start:}{]} Carbon fixation, allocation, and
-metabolism by trees set the basis for energy and material flows in
-forest ecosystems and define their interactions with Earth's changing
-climate. Here, we drew upon \# records from the Global Forest Carbon
-Database (ForC), representing all major forest types and the nine most
-significant forest autotrophic carbon flux (FACF) variables, to
+\#\#\#Abstract Carbon (C) fixation, allocation, and metabolism by trees
+set the basis for energy and material flows in forest ecosystems and
+define their interactions with Earth's changing climate. However, we
+lack a cohesive synthesis on how forest carbon fluxes vary globally with
+respect to climate and one another. Here, we draw upon 1319 records from
+the Global Forest Carbon Database (ForC), representing all major forest
+types and the nine most significant autotrophic carbon fluxes, to
 comprehensively explore how C cycling in mature, undisturbed forests
 varies with latitude and climate on a global scale. We show that, across
-all FACF variables analyzed, C cycling decreases linearly with latitude
--- a finding that confirms multiple previous studies but contradicts the
-idea that net primary productivity (\(NPP\)) of temperate forests rivals
-that of tropical forests. The FACF variables generally increase in
-proportion to one another, with few differences in allocation detectable
-at this global scale, but differed in that latitude explained a lower
-proportion of variation among subsidiary fluxes (in particular, woody
-aboveground \(NPP\) and belowground \(NPP\), \(BNPP\)). Climate
-explained a significant proportion (\#-\#\%) of variation in all C
-fluxes analyzed (less for subsidiary fluxes), with temperature variables
-in general and mean annual temperature (\(MAT\)) in particular being the
-best predictors of FACF on this global scale. While other climate
-variables (\emph{e.g.}, \textbf{XX}) displayed significant correlation
-with FACF, none of them had significantly better explanatory power than
-\(MAT\). The effects of temperature were modified by moisture
-availability, with reduced FACF under hot and dry conditions and
-sometimes under very high precipitaiton (especially for \(BNPP\)). FACF
-declined with temperature seasonality, but growing season length doesn't
-improve upon MAT as a predictor. Within the growing season, the
-influence of climate on C cycling is smaller but still significant for a
-number of carbon fluxes. These findings clarify the big picture of how
-FACF varies with latitude and climate on a global scale. As we enter a
-period of accelerating climatic change, understanding of the fundamental
-climatic controls on FACF sets a foundation for understanding patterns
-of change.
+all flux variables analyzed, C cycling decreases continuously with
+absolute latitude -- a finding that confirms multiple previous studies
+but contradicts the idea that net primary productivity (\(NPP\)) of
+temperate forests rivals that of tropical forests. C flux variables
+generally displayed similar trends across latitude and multiple climate
+variables, with no differences in allocation detected at this global
+scale. Temperature variables in general, and mean annual temperature
+(\(MAT\)) and temperature seasonality in particular, were the best
+univariate predictors of C flux, explaining 19 - 71\% of variation in
+the C fluxes analyzed. The effects of temperature were modified by
+moisture availability, with C flux reduced under hot and dry conditions
+and sometimes under very high precipitation. C fluxes increased with
+growing season length, but this was never the best univariate predictor.
+Within the growing season, the influence of climate on C cycling was
+small but significant for a number of flux variables. These findings
+clarify how forest C flux varies with latitude and climate on a global
+scale. In a period of accelerating climatic change, this improved
+understanding of the fundamental climatic controls on forest C cycling
+sets a foundation for understanding patterns of change.
 
@@
 
-\subsubsection{Introduction}\label{introduction}
+\hypertarget{introduction}{%
+\subsubsection{Introduction}\label{introduction}}
 
-Carbon cycling in forests worldwide provides the energetic basis for
+Carbon (C) cycling in Earth's forests provides the energetic basis for
 sustaining the majority of Earth's terrestrial biodiversity and many
-human populations (\textbf{REF}), while strongly influencing atmospheric
-CO\textsubscript{2} and climate \citep{bonan_forests_2008}. Forests'
-autotrophic carbon fluxes (FACF)--that is, carbon fixation, allocation,
-and metabolism by trees and other primary producers--sets the energy
-ultimately available to heterotrophic organisms (including microbes), in
-turn influencing their abundance (\textbf{REFS}) and possibly diversity
-\citep{waide_relationship_1999} (\textbf{REFS}). FACF influences all
-organic matter stocks in forest ecosystems and is linked to cycling of
-energy, water, and nutrients (\textbf{REFS}) \citep{piao_forest_2010}.
-Critically, FACF also define forest interactions with Earth's changing
-climate. Over 69 Gt of CO\textsubscript{2} cycle through Earth's forests
-each year \citep{badgley_terrestrial_2019}, and in recent decades their
-net C sequestration (\textasciitilde{}2.4 Gt C yr\textsuperscript{-1})
-offset roughly 30\% of anthropogenic fossil fuel emissions
-\citep{pan_large_2011}. As atmospheric carbon dioxide levels continue to
-rise, driving climate change, forests will play a critical role in
-shaping the future of Earth's climate
+human populations
+\citep{millennium_ecosystem_assessment_ecosystems_2005}, while strongly
+influencing atmospheric carbon dioxide (CO\textsubscript{2}) and climate
+\citep{bonan_forests_2008}. Forests' autotrophic carbon fluxes--that is,
+carbon fixation, allocation, and metabolism by trees and other primary
+producers--sets the energy ultimately available to heterotrophic
+organisms (including microbes), in turn influencing their abundance
+\citep{zak_plant_1994, niedzialkowska_species_2010} and possibly
+diversity \citep{waide_relationship_1999, chu_direct_2018}. They are
+linked to cycling of energy, water, and nutrients, and, critically,
+influence all C stocks and define forest interactions with Earth's
+changing climate. Each year, over 69 Gt of C cycle through Earth's
+forests \citep{badgley_terrestrial_2019}-- a flux more than seven times
+greater that of recent anthropogenic fossil fuel emissions \citep[9.5 Gt
+C yr\textsuperscript{-1};][]{friedlingstein_global_2019}. As atmospheric
+CO\textsubscript{2} continues to rise, driving climate change, forests
+will play a critical role in shaping the future of Earth's climate
 \citep{cavaleri_urgent_2015, rogelj_mitigation_2018}. However, our
-ability to draw general macroscopic conclusions regarding global
-variation in FACF with respect to climate has been limited in that these
-analyses often mix forests that vary in stand age, disturbance history,
-and/or management status; do not always sufficiently parse related
-variables; and typically consider only one or a few variables at a time.
+understanding of the climate dependence of forest C cycling on a global
+scale has been limited by analyses typically considering only one or a
+few variables at a time, insufficient parsing of related variables, and
+the mixing of data from forests that vary in stand age, disturbance
+history, and/or management status, all of which affect C cycling
+\citep{litton_carbon_2007, gillman_latitude_2015, simova_enigma_2017}.
 
-FACF decrease with latitude, but it remains unclear whether and how the
-shape of this relationship varies across fluxes. Studies agree that FACF
-are lowest in the boreal regions, and increase into the temperate
-regions
-\citep{luyssaert_co_2007, huston_global_2009, beer_terrestrial_2010, jung_global_2011}.
-However, evidence is inconclusive on whether primary productivity
-continues to increase into the tropics, or whether it plateaus in
-temperate regions. Evidence for this is further complicated by the fact
-that different studies use different measures of productivity to explore
-these relationships. For example, modelling of global terrestrial
-ecosystem gross primary productivity (\(GPP\)) through upscaling and
-calibration of eddy flux measurements indicates that GPP peaks in
-tropical forests
-\citep{beer_terrestrial_2010, jung_global_2011, badgley_terrestrial_2019, li_mapping_2019}.
-In contrast, some studies suggest that the highest values of net primary
-productivity (\(NPP\)) may be found in temperate forests
-\citep{luyssaert_co_2007, huston_global_2009}, while others find \(NPP\)
-highest in the tropics and decreasing with latitude
-\citep{simova_enigma_2017}. Other studies have chosen to focus
-exclusively on above-ground net primary productivity (\(ANPP\)), finding
-evidence of a weak negative relationship between \(ANPP\) and latitude
-\citep{huston_global_2009, gillman_latitude_2015}.
+Forest C fluxes decrease with latitude
+\citep[e.g.,][]{luyssaert_co_2007, piao_forest_2010, gillman_latitude_2015, li_mapping_2019},
+but it remains unclear whether and how the shape of this relationship
+varies among fluxes. For instance, gross primary productivity (\(GPP\))
+increases continuously with decreasing latitude and is indisputably
+highest in tropical forests
+\citep{luyssaert_co_2007, beer_terrestrial_2010, jung_global_2011, badgley_terrestrial_2019, li_mapping_2019}.
+In contrast, some studies have suggested that net primary productivity
+(\(NPP\)), or its aboveground portion (\(ANPP\)), exhibits a less
+distinct increase from temperate to tropical forests
+\citep{luyssaert_co_2007}--or even a decrease \citep[but see
+\citet{gillman_latitude_2015}]{huston_global_2009}. A shallower increase
+in \(NPP\) than in \(GPP\) with decreasing latitude would align with the
+suggestion that tropical forests tend to have low carbon use efficiency
+\citep[\(CUE\)=
+\(NPP\)/\(GPP\);][]{de_lucia_forest_2007, malhi_productivity_2012, anderson-teixeira_carbon_2016}.
+However, accurately comparing relationships of multiple forest C fluxes
+across latitude requires a large database with standardization for
+methodology, stand history, and management, which to date has not been
+possible.
 
-The latitudinal gradient in FACF is primarily driven by climate, which
-is a significant driver of FACF across broad spatial scales
+The latitudinal gradient in forest C flux rates, along with altitudinal
+gradients \citep{girardin_net_2010, malhi_variation_2017}, is driven
+primarily by climate, which is a significant driver of C fluxes across
+broad spatial scales
 \citep{luyssaert_co_2007, cleveland_relationships_2011, hursh_sensitivity_2017}.
-The majority of studies have focused on exploring the relationships of
-FACF to mean annual temperature (MAT) and mean annual precipitation
-(MAP), as the most commonly reported site-level climate variables. While
-these fail to capture some important aspects of climate such as
-seasonality, they do describe broad trends in temperature and water
-availability, and therefore capture a substantial portion of
-global-scale variation in climate. There is strong evidence that both
-MAT and MAP show significant positive relationships with FACF
-\citep{chu_does_2016}. However, as with latitude, the shape of those
-relationships is not always clear, and, again, is complicated by the use
-of different measures of FACF across studies. Various measures of
-primary productivity \{\textbf{FACF?}\} saturate at high levels of MAP,
-though the saturation points identified vary from 1500mm
-\citep{luyssaert_co_2007} up to 2445mm MAP
-\citep{schuur_productivity_2003}. Studies of the influence of MAT on
-productivity \{\textbf{FACF?}\} are less conclusive. Luyssaert et al.
-\citeyearpar{luyssaert_co_2007} examined GPP and NPP and found that,
-while \(GPP\) increases linearly with MAT, NPP saturates at around
-10\(^\circ\)C MAT. In contrast, Larjavaara and Muller-Landau
-\citeyearpar{larjavaara_temperature_2012}, find that increases in
-\(GPP\) saturate at approximately 25\(^\circ\)C MAT, while Schuur
-\citeyearpar{schuur_productivity_2003} shows that \(NPP\) increases
-linearly with temperature. Taylor et al.
-\citeyearpar{taylor_temperature_2017} showed a positive interaction
-between MAT and MAP in shaping tropical forest productivity, such that
-high rainfall had a negative effect on productivity in cooler climates,
-compared to a positive effect in warmer climates. \{\textbf{It would be
-good to add some more citations on soil respiraiton. I'm sure BBL can
-help.}\}
+However, there is little consensus as to the shapes of these
+relationships or the best predictor variables. The majority of studies
+have focused on exploring the relationships of C fluxes to mean annual
+temperature (MAT) and mean annual precipitation (MAP), as the most
+commonly reported site-level climate variables. C fluxes increase
+strongly with \(MAT\) on the global scale, but whether they saturate or
+potentially decrease at higher temperatures remains disputed. Some
+studies have detected no deceleration or decline in \(GPP\)
+\citep{luyssaert_co_2007} or \(NPP\) \citep{schuur_productivity_2003}
+with increasing MAT. Some have suggested that relationship between C
+flux and temperature is exponential in nature \citep{allen_linking_2005}
+(\textbf{REFS}). In contrast, others have found evidence of saturation
+or decline of C flux in the warmest climates; \citet{luyssaert_co_2007}
+found \(NPP\) saturating at around 10\(^\circ\)C MAT;
+\citet{larjavaara_temperature_2012} found that increases in \(GPP\)
+saturate at approximately 25\(^\circ\)C MAT, and
+\citet{sullivan_sensitivity_2020} found that, within the tropics,
+\(ANPP_{stem}\) decreases at the highest maximum temperatures. Various
+measures of primary productivity saturate at high levels of MAP, though
+the saturation points identified vary from 1500mm
+\citep{luyssaert_co_2007} up to 2,445 mm MAP
+\citep{schuur_productivity_2003}. Interactions between MAT and MAP are
+also possible; within the tropics, there is a positive interaction
+between MAT and MAP in shaping \(ANPP\), such that high rainfall has a
+negative effect on productivity in cooler climates, compared to a
+positive effect in warmer climates \citep{taylor_temperature_2017}.
+\textbf{Such complicated dynamics play out with belowground fluxes as
+well, with typically dominant soil temperature mediated by moisture and
+C supply \citep{hursh_sensitivity_2017, xu_contribution_2016}.
+\url{https://github.com/forc-db/Global_Productivity/issues/111}} There
+is also evidence that C fluxes also respond to climate variables such as
+temperature and precipitation seasonality (\textbf{REFS}), cloud cover
+\citep{taylor_temperature_2017}, solar radiation
+\citep{beer_terrestrial_2010, fyllas_solar_2017}, and potential
+evapotranspiration \citep{kerkhoff_plant_2005}; however, these are not
+typically assessed in global-scale analyses of annual forest C flux.
 
-FACF can be influenced by many other factors as well, which often act
-across a range of scales, and may show interactive effects with each
-other \citep{cleveland_relationships_2011}. On a local scale, stand age
-\citep{litton_carbon_2007, gillman_latitude_2015}, biodiversity
-\citep{liang_positive_2016}, management \citep{simova_enigma_2017},
-nutrient availability \citep{aragao_above_2009}, and altitude
-\citep{girardin_net_2010, malhi_variation_2017} all impact FACF. On a
-global scale, we expect that FACF are most strongly influenced by broad
-climatic gradients. There is evidence that FACFs also respond to
-variables such as cloud cover \citep{taylor_temperature_2017}, solar
-radiation \citep{fyllas_solar_2017}, and potential evapotranspiration
-\citep{kerkhoff_plant_2005} in potentially significant ways.
-Furthermore, MAT and MAP are very coarse measures of climate, and so
-fail to capture much variation in climate on an intra-annual scale,
-including the effects of factors such as growing season length, number
-of frost-free days, temperature seasonality, and dry season length. Some
-studies have suggested that the apparently strong relationship between
-MAT and FACFs is actually a consequence of the correlation between MAT
-and growing season length
-\citep{kerkhoff_plant_2005, malhi_productivity_2012, michaletz_convergence_2014, michaletz_drivers_2018}.
-Kerkhoff et al. \citeyearpar{kerkhoff_plant_2005} and Michaletz et al.
-\citeyearpar{michaletz_convergence_2014} find that, within the growing
-season, there is no significant relationship between primary
-productivity and MAT, suggesting that the effect of temperature is due
-to increased length of growing season, rather than an inherent influence
-of temperature on FACF.
+As metrics of annual climate, MAT and MAP fail to capture variation in
+climate on an intra-annual scale, including temperature and
+precipitation seasonality and growing season length. Some studies have
+suggested that the apparently strong relationship between MAT and C
+fluxes is actually a consequence of the correlation between MAT and
+growing season length
+\citep{kerkhoff_plant_2005, michaletz_convergence_2014, michaletz_drivers_2018}.
+Kerkhoff et al.~\citeyearpar{kerkhoff_plant_2005} and Michaletz et
+al.~\citeyearpar{michaletz_convergence_2014} found no significant
+relationship between growing season temperature and net primary
+productivity (loosely defined to include a mix of \textbf{VARIABLES})
+standardized to growing season length. While this suggests that the
+influence of temperature may be limited to determining the length of the
+frost-free growing season, analysis with clearly defined variables and
+standardized forest ages would be necessary to test the the veracity and
+generality of this hypothesis.
 
@@
 opens up the possibility for a standardized analysis of global scale
-variation in multiple FACF and the principle climatic drivers of these
-patterns. In order to approach this broad topic, we simplify the major
-gaps in our knowledge to five broad questions and corresponding
-hypotheses (Table 1). First, we ask how FACF vary with latitude. We then
-test how these fluxes relate to MAT and MAP, and additionally how they
-respond to other, less well-studied, climate variables. Finally, we
-consider the relationship between FACF and seasonality, considering the
-role of seasonality in explaining variation in carbon fluxes, and the
-influence of climate on FACF standardized by growing season length. We
-address the above questions for nine carbon fluxes contained in ForC,
-allowing for an in-depth exploration of the effect of climate on FACF
-globally.
+variation in multiple C fluxes and the principle climatic drivers of
+these patterns. In order to approach this broad topic, we simplify the
+major gaps in our knowledge to five broad questions and corresponding
+hypotheses (Table 1). First, we ask how nine forest autotrophic carbon
+fluxes in ForC vary with latitude. We then test how these fluxes relate
+to MAT and MAP, and additionally how they respond to other, less
+well-studied, climate variables. Finally, we consider the relationship
+between C flux and seasonality, considering the role of seasonality in
+explaining variation in carbon fluxes, and the influence of climate on C
+flux standardized by growing season length.
 
@@
 \caption{\label{tab:unnamed-chunk-4}Summary of research questions, corresponding hypotheses, and results.
-Statistically signficant support for/ rejection of hypotheses is indicated with 'yes'/'no', parentheses indicate partial overall support/rejection of hypotheses across all FACF, and '-' indicates no significant relationship.}
+Statistically signficant support for/ rejection of hypotheses is indicated with 'yes'/'no', and '-' indicates no significant relationship.
+Parentheses indicate partial overall support or rejection of hypotheses across all fluxes considered.}
 \centering
@@
 \fontsize{12}{14}\selectfont
-\begin{tabular}{llllllllllll}
+\begin{tabular}[t]{llllllllllll}
 \toprule
-\multicolumn{1}{c}{ } & \multicolumn{1}{c}{ } & \multicolumn{9}{c}{Forest autotrophic carbon fluxes (FACF)} & \multicolumn{1}{c}{ } \\
+\multicolumn{1}{c}{ } & \multicolumn{1}{c}{ } & \multicolumn{9}{c}{Forest autotrophic carbon fluxes} & \multicolumn{1}{c}{ } \\
 \cmidrule(l{3pt}r{3pt}){3-11}
-Questions and hypotheses (with related references) & Overall & $GPP$ & $NPP$ & $ANPP$ & $ANPP_{woody.stem}$ & $ANPP_{foliage}$ & $BNPP$ & $BNPP_{fine.root}$ & $R_{auto}$ & $R_{auto-root}$ & Support\\
+Questions and hypotheses (with related references) & Overall & $GPP$ & $NPP$ & $ANPP$ & $ANPP_{stem}$ & $ANPP_{foliage}$ & $BNPP$ & $BNPP_{fine.root}$ & $R_{auto}$ & $R_{root}$ & Support\\
 \midrule
 \addlinespace[1em]
-\multicolumn{5}{l}{\textbf{Q1. How do FACF vary with latitude?}}\\
-\hspace{1em}H1.1. FACF decrease linearly with latitude.\textsuperscript{1}$^{,}$\textsuperscript{2}$^{,}$\textsuperscript{3} & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Fig. 2\\
+\multicolumn{5}{l}{\textbf{Q1. How do C fluxes vary with latitude?}}\\
+\hspace{1em}H1.1.
+C fluxes decrease continuously with latitude.\textsuperscript{1}$^{,}$\textsuperscript{2}$^{,}$\textsuperscript{3}$^{,}$\textsuperscript{10} & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Fig.
+2\\
 \addlinespace[1em]
 \hline
-\multicolumn{5}{l}{\textbf{Q2. How do FACF vary with MAT and MAP?}}\\
-\hspace{1em}H2.1. FACF increase with MAT.\textsuperscript{1}$^{,}$\textsuperscript{4} & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Figs. 4, S4, S5\\
-\hspace{1em}H2.2. FACF increase with precipitation.\textsuperscript{1}$^{,}$\textsuperscript{4} & (yes) & yes & yes & yes & - & yes & yes & yes & yes & yes & Figs. 4, S4, S5\\
-\hspace{1em}H2.3. Temperature and precipitation interactively shape FACF.\textsuperscript{5} & (yes) & yes & yes & - & yes & - & yes & yes & yes & - & Fig. 3\\
+\multicolumn{5}{l}{\textbf{Q2. How do C fluxes vary with mean annual temperature (MAT) and precipitation (MAP)?}}\\
+\hspace{1em}H2.1.
+C fluxes increase continuously with MAT.\textsuperscript{1}$^{,}$\textsuperscript{4}$^{,}$\textsuperscript{9} & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Figs.
+4, S4, S5\\
+\hspace{1em}H2.2.
+C fluxes increase with precipitation up to at least 2000 mm yr$^{-1}$.\textsuperscript{1}$^{,}$\textsuperscript{4} & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Figs.
+4, S4, S5\\
+\hspace{1em}H2.3. Temperature and precipitation interactively shape C fluxes.\textsuperscript{5} & (yes) & yes & yes & - & yes & - & yes & yes & yes & - & Fig. 3\\
 \addlinespace[1em]
 \hline
-\multicolumn{5}{l}{\textbf{Q3. How are FACF related to other climate variables?}}\\
-\hspace{1em}H3.1. FACF increase with PET. & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Figs. 4, S4, S5\\
-\hspace{1em}H3.2. FACF increase with vapour pressure deficit. & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Figs. 4, S4, S5\\
-\hspace{1em}H3.3. FACF increase with solar radiation. & (yes) & yes & yes & yes & yes & yes & yes & yes & yes & - & Figs. S4, S5\\
+\multicolumn{5}{l}{\textbf{Q3. How are C fluxes related to other annual climate variables?}}\\
+\hspace{1em}H3.1. C fluxes display a decelerating increase or unimodal relationship with PET. & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Figs. 4, S4, S5\\
+\hspace{1em}H3.2. C fluxes display a decelerating increase or unimodal relationship with vapour pressure deficit. & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Figs. 4, S4, S5\\
+\hspace{1em}H3.3. C fluxes increase with solar radiation. & (yes) & yes & yes & yes & yes & yes & yes & yes & yes & - & Figs. S4, S5\\
 \addlinespace[1em]
 \hline
-\multicolumn{5}{l}{\textbf{Q4. How does seasonality influence FACF?}}\\
-\hspace{1em}H4.1. FACF decrease with temperature seasonality. & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Figs. 4, S6, S7\\
-\hspace{1em}H4.2. FACF decrease with precipitation seasonality. & - & - & - & - & - & - & - & - & - & - & Figs. S6, S7\\
+\multicolumn{5}{l}{\textbf{Q4. How does seasonality influence annual C fluxes?}}\\
+\hspace{1em}H4.1. C fluxes decrease with temperature seasonality. & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Figs. 4, S6, S7\\
+\hspace{1em}H4.2. C fluxes decrease with precipitation seasonality. & - & - & - & - & no & - & - & - & - & - & Figs. S6, S7\\
 \hspace{1em}H4.3.
-FACF increase with growing season length.\textsuperscript{6}$^{,}$\textsuperscript{7}$^{,}$\textsuperscript{8} & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Figs.
+C fluxes increase with growing season length.\textsuperscript{6}$^{,}$\textsuperscript{7}$^{,}$\textsuperscript{8} & yes & yes & yes & yes & yes & yes & yes & yes & yes & yes & Figs.
 4, S6, S7\\
-\hspace{1em}H4.4. Growing season length is a better predictor of FACF than MAT.\textsuperscript{7}$^{,}$\textsuperscript{8} & (no) & no & no & no & no & no & no & - & no & no & Table S4\\
+\hspace{1em}H4.4. Growing season length is a better predictor of C fluxes than MAT.\textsuperscript{7}$^{,}$\textsuperscript{8} & (no) & no & no & no & - & no & no & no & no & no & Table S4\\
 \addlinespace[1em]
 \hline
-\multicolumn{5}{l}{\textbf{Q5. When standardised by growing season length, how do FACF vary with climate?}}\\
-\hspace{1em}H5.1. Growing season FACF increase with temperature.\textsuperscript{8} & (yes) & - & - & yes & - & yes & - & - & - & - & Figs. S8, S9\\
-\hspace{1em}H5.2. Growing season FACF increase with PET. & (yes) & yes & yes & - & yes & - & yes & yes & - & - & Figs. S8, S9\\
-\hspace{1em}H5.3. Growing season FACF increase with precipitation. & (yes) & - & - & yes & - & yes & - & - & - & - & Figs. S8, S9\\
-\hspace{1em}H5.4. Growing season FACF increase with solar radiation. & (yes) & yes & yes & - & - & - & yes & yes & - & - & Figs. S8, S9\\
+\multicolumn{5}{l}{\textbf{Q5. When standardised by growing season length, how do annual C fluxes vary with climate?}}\\
+\hspace{1em}H5.1. Growing season C fluxes increase with temperature.\textsuperscript{8} & (yes) & - & - & yes & - & yes & - & - & - & - & Figs. S8, S9\\
+\hspace{1em}H5.2. Growing season C fluxes increase with PET. & (yes) & yes & yes & - & yes & - & yes & yes & - & - & Figs. S8, S9\\
+\hspace{1em}H5.3. Growing season C fluxes increase with precipitation. & (yes) & - & - & yes & - & yes & - & - & - & - & Figs. S8, S9\\
+\hspace{1em}H5.4. Growing season C fluxes increase with solar radiation. & (yes) & - & - & - & - & - & yes & yes & - & - & Figs. S8, S9\\
 \bottomrule
@@
 (2014) \textsuperscript{8} Chu et al.
-(2016)}\\
+(2016) \textsuperscript{9} Piao et al.
+(2010) \textsuperscript{10} Huston \& Wolverton (2009)}\\
 \end{tabular}}
@@
 
-\subsubsection{Materials and Methods}\label{materials-and-methods}
+\hypertarget{materials-and-methods}{%
+\subsubsection{Materials and Methods}\label{materials-and-methods}}
 
@@
 
-This analysis focused on nine FACF included in the open-access ForC
-database (Table 2)
+This analysis focused on nine C flux variables included in the
+open-access ForC database (Table 2)
 \citep{anderson-teixeira_carbon_2016, anderson-teixeira_forc_2018}. ForC
@@
 Soil Respiration Database \citep{bond-lamberty_global_2010}. Additional
-targeted literature searches were conducted to identify any further
-available data on the FACF analyzed here, with particular focus on
+targeted literature searches were conducted to identify further
+available data on the fluxes analyzed here, with particular focus on
 mature forests in temperate and boreal regions, which were not included
@@
 signficant disturbance or management. There is evidence that stand age
-influences patterns of FACF and carbon allocation in forest ecosystems,
-and can confound relationships between latitude and primary productivity
-\citep{delucia_forest_2007, gillman_latitude_2015}. To reduce any
+influences patterns of C flux and allocation in forest ecosystems, and
+can confound relationships between latitude and primary productivity
+\citep{de_lucia_forest_2007, gillman_latitude_2015}. To reduce any
 biasing effects of stand age, we included only stands of known age
@@
 ``intact'', or ``old-growth''. Since management can alter observed
-patterns of FACF \citep{simova_enigma_2017}, sites were excluded from
-analysis if they were managed, defined as plots that were planted,
+patterns of C cycling \citep{simova_enigma_2017}, sites were excluded
+from analysis if they were managed, defined as plots that were planted,
 managed as plantations, irrigated, fertilised or including the term
@@
 or harvesting, burning, flooding, drought and storm events with site
-mortality \textgreater{}10\% of trees. Grazed sites were retained.
+mortality \textgreater10\% of trees. Grazed sites were retained.
 
@@
 
-\caption{\label{tab:unnamed-chunk-5}Definitions and sample sizes of FACF variables used in analysis. All variables are in units of Mg C ha$^{-1}$ yr$^{-1}$. }
+\caption{\label{tab:unnamed-chunk-5}Definitions and sample sizes of carbon flux variables used in analysis. All variables are in units of Mg C ha$^{-1}$ yr$^{-1}$. }
 \centering
@@
 \fontsize{12}{14}\selectfont
-\begin{tabular}{l>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{7cm}>{\raggedright\arraybackslash}p{10cm}>{\raggedleft\arraybackslash}p{1.5cm}>{\raggedleft\arraybackslash}p{1.5cm}}
+\begin{tabular}[t]{l>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{7cm}>{\raggedright\arraybackslash}p{10cm}>{\raggedleft\arraybackslash}p{1.5cm}>{\raggedleft\arraybackslash}p{1.5cm}}
 \toprule
@@
 $NPP$ & Net Primary Production & stem, foliage, coarse root, fine root, optionally others (e.g., branch, reproductive, understory) & $ANPP$ + $BNPP$ (majority); $GPP$-$R_{auto}$ & 161 & 56\\
-$ANPP$ & Aboveground $NPP$ & stem, foliage, optionally others (e.g., branch, reproductive, understory) & $ANPP_{woody-stem}$ + $ANPP_{foliage}$ (+ others) & 278 & 86\\
-$ANPP_{woody.stem}$ & Woody stem growth component of $ANPP$ & woody stems down to DBH $\le$ 10cm (no branch turnover) & stem growth measurements scaled to biomass using allometries & 264 & 96\\
-$ANPP_{foliage}$ & Foliage component of $ANPP$ & foliage & litterfall collection (separated into components) & 98 & 49\\
+$ANPP$ & Aboveground $NPP$ & stem, foliage, optionally others (e.g., branch, reproductive, understory) & $ANPP_{stem}$ + $ANPP_{foliage}$ (+ others) & 278 & 86\\
+$ANPP_{stem}$ & Stem growth component of $ANPP$ & woody stems down to DBH $\le$ 10cm (no branch turnover) & stem growth measurements scaled to biomass using allometries & 264 & 96\\
+$ANPP_{foliage}$ & Foliage component of $ANPP$ & foliage & litterfall collection, with separation into components & 98 & 49\\
 \addlinespace
@@
 $BNPP_{fine.root}$ & Fine root component of $BNPP$ & fine roots & measurements combined one or more of the following: soil cores, minirhizotrons, turnover estimates,  root ingrowth cores & 88 & 41\\
-$R_{auto}$ & Autotrophic respiration & foliage, stem, and root & chamber measurements of foliage and stem gas exchange +  $R_{auto-root}$ (as below) & 22 & 13\\
-$R_{auto-root}$ & Root respiration & (coarse and) fine roots & partitioning of total soil respiration (e.g., through root exclusion), scaling of root gas exchange; excluded alkali absoption and soda lime methods for measuring soil respiration & 64 & 26\\
+$R_{auto}$ & Autotrophic respiration & foliage, stem, and root & chamber measurements of foliage and stem gas exchange +  $R_{root}$ (as below) & 22 & 13\\
+$R_{root}$ & Root respiration & (coarse and) fine roots & partitioning of total soil respiration (e.g., through root exclusion), scaling of root gas exchange; excluded alkali absoption and soda lime methods for measuring soil respiration & 64 & 26\\
 \bottomrule
@@
 datasets: WorldClim \citep{hijmans_very_2005}, WorldClim2
-\citep{fick_worldclim_2017}, the Climate Research Unit (CRU) time-series
-dataset v. 4.03 \citep{harris_updated_2014}, the Global Aridity Index
-and Potential Evapotranspiration Climate Database
+\citep{fick_worldclim_2017}, the Climate Research Unit time-series
+dataset (CRU TS v4.03 \citep{harris_updated_2014}, the Global Aridity
+Index and Potential Evapotranspiration Climate Database
 \citep{trabucco_global_2019}, and TerraClimate
@@
 
-Length of the growing season was estimated to the nearest month, where
+\textbf{For consistency with previous studies (Table 1, \emph{H5}),}
+length of the growing season was estimated to the nearest month, where
 growing season months were defined as months with mean minimum
 temperature \textgreater{} 0.5\(^\circ\)C. We experimented with a
-definion of growing season months including a moisture index, defined as
-(MAT - PET)/PET, \textgreater{} -0.95 (Kerkhoff et al. 2005; see also
-Michaletz et al. 2014). However, we found that including a moisture
-index had \textbf{no} effect on the estimates of growing season length,
-and so chose to exclude it. (\textbf{Becky, was it really no effect? or
-minimal?}) Monthly data for PET, precipitation, and temperature from the
-CRU dataset v 4.03 \citep{harris_updated_2014}, and solar radiation from
-WorldClim2 \citep{fick_worldclim_2017} were used to calculate mean
-monthly PET, precipitation, temperature and solar radiation during the
-growing season. Total growing season precipitation and solar radiation
-were also calculated.
+definition of growing season months including a moisture index, defined
+as (MAT - PET)/PET, \textgreater{} -0.95
+\citetext{\citealp{kerkhoff_plant_2005}; \citealp[see
+also][]{michaletz_convergence_2014}}. However, we found that including a
+moisture index had minimal effect on the estimates of growing season
+length, and so chose to exclude it. Monthly data for PET, precipitation,
+and temperature from CRU v 4.03 \citep{harris_updated_2014} and solar
+radiation from WorldClim2 \citep{fick_worldclim_2017} were used to
+calculate mean monthly PET, precipitation, temperature and solar
+radiation during the growing season. Total growing season precipitation
+and solar radiation were also calculated.
 
@@
 
-The effects of latitude and climate on FACF were analysed using mixed
-effects models using the package `lme4' \citep{bates_fitting_2015} in R
-v.3.5.1 \citep{r_core_team_r:_2018}. The basic model for all analyses
-included a fixed effect of latitude or climate and a random effect of
-plot nested within geographic area. Geographic areas--\emph{i.e.},
-spatially clustered sites--are defined within ForC using a hierarchical
-cluster analysis on the distance matrix of the sites and a cutoff of
-25km \citep{anderson-teixeira_forc_2018}. We experimented with inclusion
-of altitude as a fixed effect, but excluded it from the final models
-because it added very little explanatory power--that is, the difference
-in AIC (\(\Delta\)AIC) relative to models excluding altitude was
-generally small (often \(\Delta\)AIC\textless{}2). Hypotheses were
-accepted if the \(\Delta\)AIC between a model including the fixed effect
-of interest and a corresponding null model excluding that fixed effect
-exceeded 2.0. All \(R^2\) values presented here are marginal \(R^2\)
-values, and refer to the proportion of variation explained by only the
-fixed effects. Specific analyses are as described below.
+The effects of latitude and climate on C fluxes were analysed using
+mixed effects models using the package `lme4' \citep{bates_fitting_2015}
+in R v.3.5.1 \citep{r_core_team_r:_2018}. The basic model for all
+analyses included a fixed effect of latitude or climate and a random
+effect of plot nested within geographic area. Geographic
+areas--\emph{i.e.}, spatially clustered sites--are defined within ForC
+using a hierarchical cluster analysis on the distance matrix of the
+sites and a cutoff of 25km \citep{anderson-teixeira_forc_2018}. We
+experimented with inclusion of altitude as a fixed effect, but excluded
+it from the final models because it added very little explanatory
+power--that is, the difference in AIC (\(\Delta\) AIC) relative to
+models excluding altitude was generally small (often \(\Delta\)
+AIC\textless2). Effects were considered significant when inclusion of
+the fixed effect of interest resulted in p \(\le\) 0.05 and \(\Delta\)
+AIC \(\ge\) 2.0 relative to a corresponding null model. All \(R^2\)
+values presented here are marginal \(R^2\) values, and refer to the
+proportion of variation explained by only the fixed effects. Specific
+analyses are as described below.
 
-We first examined the relationship between latitude and FACF (Q1; Table
-1). We tested models with latitude as a linear term (corresponding null:
-model without latitude) and as a second-order polynomial term
-(corresponding null: model with latitude as a linear term), and
-calculated AIC values to determine the best model. Models were accepted
-as the best model if \(\Delta\)AIC \textgreater{} 2 with respect to the
-corresponding null, and were significant with respect to a null model
-with no fixed term. We also examined relationships among fluxes across
-latitude, testing whether sums of component fluxes matched the larger
-fluxes and whether C allocation varied with latitude, as specified
-below.
+We first examined the relationship between latitude and C fluxes
+(\emph{Q1}; Table 1). We tested models with latitude as a first-order
+linear, second-order polynomial, and logarithmic term. For brevity, we
+henceforth refer to first-order linear models as ``linear'' and
+second-order polynomial models as ``polynomial''. We selected as the
+best model that with the highest \(\Delta\) AIC relative to a null model
+with no fixed term, with the qualification that a polynomial model was
+considered an improvement over a linear model only if it reduced the AIC
+value by 2.0 or more.
 
@@
 those of larger fluxes, regression lines for smaller component fluxes
-were summed to generate new estimates of larger fluxes, which were then
-compared against the latitudinal regression of the larger flux.
-Confidence intervals for the larger flux were calculated using the
-`bootMer' function from the lme4 package \citep{bates_fitting_2015}.
-This analysis was applied to following sets of fluxes: (1)
-\(GPP = NPP + R_{auto}\), (2) \(NPP = ANPP + BNPP\), and (3)
-\(ANPP = ANPP_{foliage} + ANPP_{woody.stem}\). In addition, we estimated
-total belowground C flux (TBCF, not analyzed due to limited data) as
-\(TBCF = BNPP + R_{root}\).
+were summed to generate new estimates of larger fluxes. Because no
+fluxes were significantly better predicted by a logarithmic or
+polynomial fit than by a linear fit, we used linear fits for all fluxes.
+We then determined whether these summed predictions fell within the 95\%
+CI for the larger flux across the entire latitudinal range. Confidence
+intervals for the line of best fit for the larger flux were estimated
+using the `bootMer' function, a parametric bootstrapping method for
+mixed models \citep{bates_fitting_2015}. This function carried out 2000
+simulations estimating the line of best fit, using quantiles at 0.025
+and 0.975 to estimate 95\% CIs. This analysis was applied to the
+following sets of fluxes: (1) \(GPP = NPP + R_{auto}\), (2)
+\(NPP = ANPP + BNPP\), and (3) \(ANPP = ANPP_{foliage} + ANPP_{stem}\).
+In addition, we estimated total belowground C flux (TBCF, not analyzed
+due to limited data) as \(TBCF = BNPP + R_{root}\).
 
-Variation in allocation to component carbon fluxes along latitudinal
-gradients was explored for the following pairings: \(GPP:NPP\),
-\(ANPP:BNPP\), \(ANPP_{foliage}:ANPP_{woody.stem}\),
-\(ANPP_{foliage}:NPP\), \(ANPP_{woody.stem}:NPP\), \(ANPP:NPP\), and
-\(BNPP:NPP\). For each set of paired fluxes, measurements taken at the
-same site and plot, and in the same year, were paired together, and the
-ratio of each pair of measurements calculated. The ratios were regressed
-against latitude and climate variables, using the linear model specified
-above. Cook's distance analyses were carried out for each of the models,
-and indicated that data from a few high-elevation sites were having a
-disproportionate influence on the regressions. To account for this,
-models were re-run using only data from sites \(\le\) 1000m.
+Variation in allocation to component carbon fluxes was explored for
+three groupings: (1) \(GPP = NPP + R_{auto}\), (2)
+\(NPP = ANPP + BNPP\), and (3) \(ANPP = ANPP_{foliage} + ANPP_{stem}\).
+For each group, measurements taken at the same site and plot, an in the
+same year were grouped together. For groups (1) and (2), where 2 of the
+3 flux measurements were available for a given site, plot, and year,
+these measurements were used to calculate the third. The ratio of each
+pair of component fluxes was calculated. The log of these ratios were
+regressed against latitude and climate variables, using the linear model
+specified above. Cook's distance analyses were carried out for each of
+the models, and extreme outliers removed,
 
-We next examined the relationships of FACF to climate variables (Q2-Q4;
-Table 1). As with latitude, we tested both linear and polynomial fits
-for each climate variable. We tested relationships of each FACF (Table
-2) against each climate variable (Table S1), but focus
-presentation--including focal hypotheses (Table 1)--on climate variables
-that explained \textgreater{}20\% of variation in FACF. \textbf{Becky,
-please make this more specific. I drafted this based on the results, but
-its not very clear.}
+We next examined the relationships of C fluxes to climate variables
+(Q2-Q4; Table 1). We tested first-order linear, second-order polynomial,
+and logarithmic fits for each climate variable. Again, polynomial fits
+were considered superior to first-order linear fits only if inclusion of
+a second-order polynomial term resulted in \(\Delta\) AIC \(\ge\) 2.0
+relative to a first-order linear model. We tested relationships of each
+C flux (Table 2) against each climate variable (Table S1). Variables
+which were not significant explanatory variables or which explained
+\textless20\% of variation in C fluxes are only presented in SI.
 
-To investigate the potential joint and interactive effects of climate
-variables on carbon fluxes, multivariate models were also specified. We
-explored a large range of climate variable combinations and ultimately
-decided to focus on MAT and MAP. We favored these over variables with
-similar explanatory power because they are the most commonly reported
-climate variables and are recorded in ForC. MAT was consistently among
-the best individual predictors of FACF (Table S2), although it was
-occasionally out-performed by closely correlated climate variables (Fig.
-S2). In combination with MAT as a primary climate variable, MAP was
-consistently among the best secondary climate variables. Specifically,
-compared again a model including MAT x MAP, there were no other
-variables that, in interaction with MAT, improved AIC by
-\textgreater{}2.0 and \(R^2\) by \textgreater{}0.05. In examinging the
-joint and interactive effects of climate variables, an additive model
-including MAT and MAP as fixed effects was accepted when \(\Delta\)AIC
-\textgreater{}2 relative to a null including only MAT as a fixed effect.
+Multivariate models were used to investigate the potential joint and
+interactive effects of MAT and MAP on carbon fluxes. An additive model
+including MAP in addition to MAT was accepted when \(\Delta\) AIC
+\textgreater2 relative to a null including only MAT as a fixed effect.
 An interactive model including an MAT x MAP interaction was accepted
-when \(\Delta\)AIC \textgreater{}2 relative to a null including MAT and
+when \(\Delta\)AIC \textgreater2 relative to a null including MAT and
 MAP as fixed effects.
 
-To test whether and how FACF varied with climate when standardised by
-growing season length (Q5), we first standardized FACF by dividing by
-growing season length (as defined above). We then tested for
-correlations between growing season length-standardised FACF against
-growing season climate variables. For analyses on data within the
-growing season, only linear models were specified.
+To test whether and how C flux varied with climate when standardised by
+growing season length (Q5), we first standardized all annual C fluxes by
+dividing by growing season length (as defined above). We then derived
+four variables to describe growing season climate, specifically growing
+season temperature, precipitation, solar radiation, and PET (Table S1).
+We tested for correlations between these standardised fluxes and growing
+season climate variables, using only first-order linear models.
 
 \textbf{All analyses were conducted in R (}Version\textbf{). Code and
-results are archived on GitHub\ldots{}.}
+data necessary to reproduce all results are archived on GitHub\ldots.}
 
-\subsubsection{Results}\label{results}
+\#\#\#Results
 
-In total, we analyzed 1228 records from nine FACF variables taken from
-forests that had experienced no major anthropogenic disturbances within
-the past 100 years. These records represented a total of 154 distinct
-geographic areas across all forested biogeographic and climate zones
-(Fig. 1, Table 2).
+In total, we analyzed 1319 records from nine forest autotrophic C flux
+variables taken from forests that had experienced no major anthropogenic
+disturbances within the past 100 years. These records represented a
+total of 255 plots in 154 distinct geographic areas across all forested
+biogeographic and climate zones (Fig. 1, Table 2).
 
-\emph{How do FACF vary with latitude?}
+\emph{How does C flux vary with latitude?}
 
-All major carbon fluxes decreased linearly with latitude (Fig. 2; Table
-S2). Latitude was a strong predictor for many of the carbon
-fluxes--particularly the larger fluxes (Table S2). Specifically,
-latitude explained 64\% of variation in GPP (n = 254,
-p\textless{}0.0001), 50\% in NPP (n = 114, p\textless{}0.0001) and 45\%
-in ANPP (n = 259, p\textless{}0.0001). The FACF that were most poorly
-predicted by latitude were \(BNPP_{root-fine}\) (\(R^2\)=0.17) and
-\(ANPP_{woody.stem}\) (\(R^2\)=0.18). For all FACF, the relationship
-with latitude was best predicted by the linear model).
+All major carbon fluxes decreased with latitude (Fig. 2; Table S2).
+Latitude was a strong predictor for many of the carbon fluxes,
+particularly the larger fluxes (Table S2). Specifically, latitude
+explained 64\% of variation in GPP (n = 243, p\textless0.0001), 50\% in
+NPP (n = 161, p\textless0.0001) and 44\% in ANPP (n = 278,
+p\textless0.0001). The C fluxes that were most poorly predicted by
+latitude were \(BNPP_{fine.root}\) (\(R^2\)=0.17) and \(ANPP_{stem}\)
+(\(R^2\)=0.18). The relationship with latitude was best fit by the
+first-order linear model, with the exception of NPP and \(R_{root}\),
+for which a logarithmic model was a slightly - but not significantly -
+better fit.
 
@@
 \includegraphics[width=1\linewidth]{combined_stacked} \caption{Latitudinal trends in forest autotropic carbon flux.
-Lines of best fit are plotted according to the best model selected during analysis.
-All regressions are significant $(p<0.05)$.
+Plotted are linear models, all of which were significant $(p<0.05)$ and had AIC values within 2.0 of the best model (for two fluxes, logarithmic fits were marginally better; Table S2).
 Each panel shows major C fluxes together with component fluxes.
@@
 95\% confidence intervals are plotted for the major flux for comparison with predicted trends.
-In (d),  which shows three belowground fluxes, the major flux, total belowground carbon flux, is one for which we have no data}\label{fig:unnamed-chunk-7}
+In (d), which shows three belowground fluxes, the major flux, total belowground carbon flux, has insufficient data (n=9) to support a regression}\label{fig:unnamed-chunk-7}
 \end{figure}
@@
 In general, smaller component fluxes summed approximately to larger
-fluxes across the latitudinal gradient (Fig. 2). That is, modelled
+fluxes across the latitudinal gradient (Fig. 2). That is, modeled
 estimates of \(GPP\), generated from the sum of \(NPP\) and
@@
 and \(ANPP\), generated from the sum of \(ANPP_{foliage}\) and
-\(ANPP_{woody-stem}\), fell completely within the confidence intervals
+\(ANPP_{stem}\), fell almost completely within the confidence intervals
 of the regressions of field estimates of \(GPP\), \(NPP\), and \(ANPP\),
@@
 
-We found little evidence that allocation between fluxes varied
-substantially with latitude or climate (Fig. S3). Of the 7 FACF ratios
-regressed against latitude and three climate variables (MAT, MAP,
-temperature seasonality), there were only four signficant relationships,
-all with \(R^2 \le 0.2\) (Fig. S3). Specifically, the proportion of NPP
-allocated to \(ANPP_{foliage}\) increased weakly with MAT
-(\(R^2=0.20\)), and the proportion of \(NPP\) allocated aboveground
-(\(ANPP\)) decreased weakly with latitude (\(R^2=0.11\)) and temperature
-seasonality (\(R^2=0.17\)), while increasing with MAT (\(R^2=0.11\)).
-There were no significant results from regressing ratios of carbon
-fluxes against latitude, or against any of the climate variables.
+We found no evidence of systematic variation in C allocation with
+latitude or climate (Fig. S3). Of 12 relationships tested (3 ratios
+among C flux variables regressed against latitude, MAT, MAP and
+temperature seasonality), none were significant.
 
-\emph{How do FACF relate to MAT and MAP?}
+\emph{How does C flux relate to MAT and MAP?}
 
-All FACF increased linearly with MAT, and we found no support for a
-saturation point of FACF with MAT (all p\textless{}0.05; Figs. 3-4,
-S4-S5, Table S2). As with latitude, MAT tended to explain more variation
-in the larger FACF (\(GPP\), \(NPP\), \(ANPP\), \(R_{auto}\)) and
-\(ANPP_foliage\) (all \(R^2\)\textgreater{} 0.4) than in subsidiary and
-belowground fluxes (\(ANPP_{woody-stem}\), \(R_{root}\),
-\(BNPP_{root-fine}\); all \(R^2\)\textless{} 0.25).
+All fluxes increased with MAT (all p\textless0.05; Figs. 3-4, S4-S5,
+Table S2). For eight of the nine fluxes, this relationship was linear.
+For only one variable, \(BNPP\), did a lognormal fit provide signficant
+improvement over a first-order linear relationship. \emph{As with
+latitude, MAT tended to explain more variation in the larger fluxes
+(\(GPP\), \(NPP\), \(ANPP\), \(R_{auto}\)) and \(ANPP_{foliage}\) (all
+\(R^2\)\textgreater{} 0.4) than in subsidiary and belowground fluxes
+(\(ANPP_{stem}\), \(R_{root}\), \(BNPP_{fine.root}\); all
+\(R^2\)\textless{} 0.25).} \textbf{update this -- NB these values are
+correct}
 
-MAP was a significant (p\textless{}0.05) predictor of all FACF but
-\(ANPP_{woody-stem}\) (Figs. 4a, S4-S5; Table S2). However, it explained
-little variation: with the exception of \(R_{auto}\), MAP explained at
-most 37\% of variation in FACF. For the majority of FACF, a polynomial
-model was the best fit. FACF generally increased with precipitation, up
-until a saturation point at between 3000 and 4000mm annual
-precipitation, above which they started to decrease (Figs. 4, S4-S5).
-The notable exception to this was GPP: the model indicated that GPP
-continued to increase with precipitation up to measures of at least
-5000mm annually (p\textless{}0.0001, \(R^2\) = 0.33. Data above this
-point were not available, but the model trend indicated that the
-saturation point for this model would be around 5000mm MAP.
-
+MAP was a significant (p\textless0.05) predictor of all fluxes (Figs.
+4a, S4-S5; Table S2). However, it explained little variation: with the
+exception of \(R_{auto}\), MAP explained at most 25\% of variation in C
+flux. All fluxes increased with MAP up to at least 2000 mm, above which
+responses were variable (Figs. 4, S4-S5).\\
 There was a significant additive effect of MAT and MAP on \(GPP\),
-\(ANPP\) and \(R_{auto}\) (Fig. 3, Table S3). Accounting for MAT, MAP
-had a substantial positive effect on \(GPP\) and \(R_{auto}\) and a
-small negative effect on \(ANPP\). There was a significant interactive
-effect between MAT and MAP for \(NPP\) and \(ANPP_{woody-stem}\) (Fig.
-3, Table S3). The interaction was negative for \(NPP\) and positive for
-\(ANPP_{woody-stem}\). For \(ANPP_{foliage}\), \(BNPP\),
-\(BNPP_{root-fine}\), and \(R_{auto-root}\), MAP did not have a
-signficant effect when accounting for MAT (Fig. 3, Table S3). For the
-variables which showed a significant interactive or additive effect
-between MAT and MAP, no other climate variable, in combination with MAT,
-significantly improved on that model. \{\textbf{need to confirm this
-given changes in MAT MAP results (or you could just drop the
-sentence.)}\}
+\(ANPP\) and \(R_{auto}\) (Fig. 3, Table S3), and a significant
+interactive effect between MAT and MAP for \(NPP\) and \(ANPP_{stem}\)
+(Fig. 3, Table S3). The interaction was negative for \(NPP\) and
+positive for \(ANPP_{stem}\). For \(ANPP_{foliage}\), \(BNPP\),
+\(BNPP_{fine.root}\), and \(R_{root}\), MAP did not have a signficant
+effect when accounting for MAT (Fig. 3, Table S3).
 
 \begin{figure}[H]
-\includegraphics[width=1\linewidth]{mat_map_interaction} \caption{Interactive effects of mean annual temperature and mean annual precipitation on FACF.
+\includegraphics[width=1\linewidth]{mat_map_interaction} \caption{Interactive effects of mean annual temperature and precipitation on annual forest carbon fluxes.
 For visualization purposes, data points are grouped into bins of 0 - 1000, 1001 - 2000, 2001 - 3000, and >3000mm mean annual precipitation, and lines of best fit models are plotted for mean annual precipitation values of 500, 1500, 2500, and 3500mm.
@@
 
-\emph{How do FACF relate to other climate variables?}
+\emph{How does C flux relate to other climate variables?}
 
-Our results indicated that FACF were most strongly explained by
-temperature at the global scale, with temperature-related climate
-variables coming out as strong predictors of FACF. In addition to MAT,
-several of its correlates (Fig. S2) were consistently identified as
-strong univariate predictors of FACF: temperature seasonality, annual
-temperature range, annual frost days, PET, and length of growing season
-(Figs. 4, S4-S7).
+Our results indicated that annual forest C fluxes were most strongly
+explained by temperature-related climate variables at the global scale.
+In addition to MAT, several of its correlates (Fig. S2) were
+consistently identified as strong univariate predictors of C fluxes:
+temperature seasonality, annual temperature range, annual frost days,
+PET, and length of growing season (Figs. 4, S4-S7).
 
-We found a significant relationship between C flux and potential
-evapotranspiration for all FACF. \(ANPP_{foliage}\),
-\(BNPP_{root-fine}\) and \(R_{root}\) increased linearly with PET;
-however, all other fluxes showed a polynomial relationship with PET
-(Fig. 4c, S4-5; Table S2). We found strong evidence for a saturation
-point or peak with PET: FACF tended to increase at values below 1000mm,
-before saturating between 1200 and 1700mm. There was also evidence that
-FACF begin to decrease at values above 1800mm PET.
+All C flux variables showed a significant relationship with potential
+evapotranspiration. The relationship was logarithmic for
+\(ANPP_{foliage}\), \(BNPP_{fine.root}\) and \(R_{root}\), and
+polynomial for all other fluxes (Fig. 4c, S4-5; Table S2). We found
+strong evidence for a saturation point or peak with PET: C fluxes tended
+to increase at values below 1000mm, before saturating between 1200 and
+1700mm. There was also evidence that some C fluxes begin to decrease at
+values above 1800mm PET.
 
-Vapour pressure deficit was a significant predictor of C flux for all
-FACF. \(BNPP_{root-fine}\) showed a linear relationship with vapour
-pressure deficit (\(R^2\) = 0.07, p\textless{}0.05), but all other
-fluxes showed a polynomial relationship (Figs. 4d, S4-5; Table S2). FACF
-initially increased with vapour pressure deficit, before saturating at
-around 0.8 kPa, after wich point they began to decrease.
+Vapour pressure deficit was a significant predictor of all C fluxes.
+\(ANPP_{foliage}\), \(BNPP_{fine.root}\) and \(R_{root}\) showed a
+logarithmic relationship with vapour pressure deficit, but all other
+fluxes showed a polynomial relationship (Figs. 4d, S4-5; Table S2). C
+fluxes initially increased with vapour pressure deficit, before
+saturating at around 0.8 kPa, after which point they began to decrease.
 
-All fluxes, with the exception of \(R_{root}\), showed a positive linear
-relationship with solar radiation (Figs. S4-S5, Table S2). Solar
-radiation explained a low proportion of variability in all FACF,
-explaining less than 20\% of the variation in each flux, with the
-exception of\(R_{auto}\) (\(R^2\) = 0.26, p\textless{}0.05).
+All fluxes, with the exception of \(R_{root}\), showed a significant
+positive relationship with solar radiation (Figs. S4-S5, Table S2).
+Solar radiation explained a low proportion of variability in all C
+fluxes, explaining less than 30\% of the variation in each flux
 
-Annual wet days, cloud cover, and aridity were poor or non-significant
-explainers of variation in FACF, explaining less than 20\% of the
-variation in each of the carbon fluxes (Figs. S4-S5; Table S2).
+Annual wet days, cloud cover, aridity, and water stress months were poor
+or non-significant explainers of variation in C fluxes, explaining less
+than 20\% of the variation in each of the carbon fluxes (Figs. S4-S5;
+Table S2).
 
@@
 
-\emph{What is the role of seasonality in explaining FACF?}
+\emph{What is the role of seasonality in explaining C fluxes?}
 
-Temperature seasonality was a significant predictor of FACF. \(GPP\),
-\(NPP\), \(ANPP\), and \(R_{root}\) exhibited a polynomial relationship
-with seasonality (all p\textless{}0.05; Figs. 4e, S6-7; Table S2).
-\(ANPP_{foliage}\), \(ANPP_{woody-stem}\) and \(R_{auto}\) decreased
-linearly with temperature seasonality (all p\textless{}0.05; Figs. 4e,
-S6-S7; Table S2). Temperature seasonality was strongly correlated with
-annual temperature range, which was likewise a similarly strong
-predictor of FACF (Table S2).FACF were highest where temperature
-seasonality = 0, and at an annual temperature range of 15\(^\circ\)C or
-lower.
+Temperature seasonality was a strong predictor of annual C fluxes. All
+fluxes decrease with increasing seasonality, though the shape of this
+relationship varies (all p\textless0.05; Figs. 4e, S6-7; Table S2).
+Temperature seasonality was strongly correlated with annual temperature
+range, which was likewise a similarly strong predictor of C fluxes
+(Table S2). C fluxes were highest where temperature seasonality = 0, and
+at an annual temperature range of 15\(^\circ\)C or lower. \textbf{BBL:
+perhaps put this into an ecosystem context; what are these? Aseasonal
+subtropical places?}
 
 In contrast, there was no significant effect of precipitation
-seasonality on FACF, and both maximum vapour pressure deficit, and water
-stress months were poor or non-significant explainers of variation in
-FACF (Figs. S6-S7; Table S2).
+seasonality on C fluxes, and both maximum vapour pressure deficit, and
+water stress months were poor or non-significant predictors of variation
+in C fluxes (Figs. S6-S7; Table S2).
 
 We found a significant relationship between length of growing season and
-FACF, with all fluxes showing a linear increase with length of growing
-season (Figs. 4e, S6-S7; Table S2). Length of growing season was a
-strong predictor of FACF, explaining 51\% of variation in GPP, 39\% of
-variation in NPP, and 34\% of variation in ANPP, but it was a weaker
-predictor than MAT for all fluxes analysed (Table S4).
+C fluxes, with all fluxes showing a positive relationship with length of
+growing season (Figs. 4e, S6-S7; Table S2). Length of growing season was
+a strong predictor of C fluxes, explaining 53\% of variation in GPP,
+38\% of variation in NPP, and 34\% of variation in ANPP (all
+p\textless0.05; Table S2), but it was a weaker predictor than MAT for
+all fluxes analysed (Table S4).
 
-\emph{Within the growing season, how do FACF vary with climate?}
+\emph{Within the growing season, how do C fluxes vary with climate?}
 
-When FACF were standardized by growing season length, correlations with
-growing season climate--including temperature, precipitation, solar
-radiation, and PET--were generally weak (Figs. S8-S9). Speficifally,
-\(ANPP\) increased with growing season temperature (\(R^2\) = 0.10,
-p\textless{}0.001) and precipitation (\(R^2\) = 0.04, p\textless{}0.05).
-Similaryly, \(ANPP_{foliage}\) increased slightly with growing season
-temperature (\(R^2\) = 0.16, p\textless{}0.01) and precipitation
-(\(R^2\) = 0.09, p\textless{}0.05). Growing season solar radiation had a
-positive influence on \(GPP\) (\(R^2\) = 0.21, p\textless{}0.001),
-\(NPP\) (\(R^2\) = 0.21, p\textless{}0.001), \(BNPP\) (\(R^2\) = 0.16,
-p\textless{}0.001) and \(BNPP_{fine.root}\) (\(R^2\) = 0.12,
-p\textless{}0.01). Growing season PET had a positive influence on
-\(GPP\) (\(R^2\) = 0.15, p\textless{}0.01), \(NPP\) (\(R^2\) = 0.18,
-p\textless{}0.01), \(BNPP\) (\(R^2\) = 0.23, p\textless{}0.0001),
-\(BNPP_{fine.root}\) (\(R^2\) = 0.11, p\textless{}0.05), and
-\(ANPP_{woody-stem}\) (\(R^2\) = 0.06, p\textless{}0.05).
-\{\textbf{Becky, please verify/ edit the following:} There were no other
-signifcant correlations between growing season length-standardized FACF
-(9 variables in Table 2) and growing season climate (\textbf{which
-variables?})\}.
+When annual C fluxes were standardized by growing season length (in
+monthly increments), correlations with growing season climate were
+generally weak (Figs. S8-S9). \(ANPP\) increased with growing season
+temperature (\(R^2\) = 0.09, p\textless0.001) and precipitation (\(R^2\)
+= 0.04, p\textless0.05). Similarly, \(ANPP_{foliage}\) increased
+slightly with growing season temperature (\(R^2\) = 0.16,
+p\textless0.01) and precipitation (\(R^2\) = 0.09, p\textless0.05).
+Growing season solar radiation had a positive influence on \(BNPP\)
+(\(R^2\) = 0.17, p\textless0.001) and \(BNPP_{fine.root}\) (\(R^2\) =
+0.13, p\textless0.01). Growing season PET had a positive influence on
+\(GPP\) (\(R^2\) = 0.15, p\textless0.01), \(NPP\) (\(R^2\) = 0.07,
+p\textless0.01), \(BNPP\) (\(R^2\) = 0.23, p\textless0.0001),
+\(BNPP_{fine.root}\) (\(R^2\) = 0.10, p\textless0.05), and
+\(ANPP_{stem}\) (\(R^2\) = 0.06, p\textless0.05). All other
+relationships were non-significant.
 
-\subsubsection{Discussion}\label{discussion}
+\#\#\#Discussion
 
@@
 carbon fluxes in mature forests vary with latitude and climate on a
-global scale. We show that, across all nine FACF analyzed, C cycling
-decreases continually with latitude (\emph{H1.1}; Fig. 2)--a finding
-that confirms multiple previous studies but contradicts the idea that
-productivity of temperate forests rivals that of tropical forests
-\citep{huston_global_2009}. The FACF generally increase in proportion to
-one another (\emph{H1.2}), with few differences in allocation detectable
-at this global scale (Fig. S2) and with component fluxes summing
-appropriately to larger fluxes (Fig. 2), indicating no major, systematic
-omissions or overestimations of flux components. However, climate
-explained lower proportions of variability among subsidiary C fluxes
-(\emph{e.g.}, \(ANPP_{woody}\), \(BNPP_{fine.root}\), \(R_{auto-root}\);
-Fig. 2; Table S2). Latitudianal variation in FACF is primarily
-attributable to temperature-related variables (\emph{H3, H4}),
-particularly MAT (Figs. 3-4). Water availability is also influential,
-but generally of secondary importance across the range represented in
-our database (Figs. 3-4). Temperature seasonality and growing season
-length are closely correlated with MAT and are strong predictors of FACF
-(\emph{H4}; Figs. 4e-f, S2, S6-S7), though growing season length doesn't
-improve upon MAT as a predictor. Within the growing season, the
+global scale. We show that, across all nine forest autotrophic C fluxes
+analyzed, C cycling decreases continually with latitude (\emph{H1.1};
+Fig. 2)--a finding that confirms multiple previous studies but
+contradicts the idea that productivity of temperate forests rivals or
+even exceeds that of tropical forests
+\citep{luyssaert_co_2007, huston_global_2009}. At this global scale, C
+fluxes increase approximately in proportion to one another, with
+component fluxes summing appropriately to larger fluxes and no
+detectable differences in allocation across latitude or climates (Figs.
+2, 4, S3). However, climate explains lower proportions of variability
+among some of the subsidiary C fluxes (\emph{e.g.}, \(ANPP_{stem}\),
+\(BNPP\) \(BNPP_{fine.root)\); Fig. 2; Tables S2, S6). Latitudinal
+variation in C fluxes is primarily attributable to temperature-related
+variables (\emph{H3, H4}), particularly MAT and temperature seasonality
+(Figs. 3-4). Water availability is also influential, but generally of
+secondary importance across the climate space occupied by forests (Figs.
+3-4). While growing season length explained signficant variation in all
+C fluxes (\emph{H4}; Figs. 4f, S6-S7), it did not improve upon MAT or
+temperature seasonality as a predictor. Within the growing season, the
 influence of climate on C cycling is smaller but still significant for a
-number of FACF (\emph{H5}; Fig. S9; Table S4). These findings clarify
-the big picture of how FACF vary with latitude and climate on a global
-scale.
+number of C fluxes (\emph{H5}; Fig. S9; Table S4).
 
-Past studies have differed in their conclusions regarding the
-relationship between FACF and latitude or its correlates (Table 1,
-\emph{H1}; \textbf{REFS})--quite possibly because of lack of
+These findings clarify how forest C fluxes vary with latitude and
+climate on a global scale. Past studies have differed in their
+conclusions regarding the relationship between C fluxes and latitude or
+its correlates (Table 1, \emph{H1})--quite possibly because of lack of
 standardization with respect to stand age and disturbance history. Our
-findings indicate that, among mature, undisturbed stands, FACF are
-unambiguously highest in the tropical regions, and the relationship is
-approximately linear (Fig. 2). This contrasts with the suggestion that
-productivity of temperate forests is similar to that of tropical forests
-\citep{huston_global_2009}. Compared to tropical forests, the temperate
-forest biome has experienced more widespread anthropogenic disturbance
-and has a larger fraction of secondary stands
-\citep{potapov_mapping_2008, poulter_global_2018}, so analyses comparing
-across latitudinal gradients without controlling for stand age risk
-confounding age with biome effects. In addition, because C allocation
-varies with stand age \citep{delucia_forest_2007} (\textbf{See Nobby's
-comment in manuscript-draft\_NK.pdf}), age differences may introduce
-systematic biases into analyses of FACF across latitude or global
-climatic gradients. For example, woody productivity tends to be higher
-in rapidly aggrading secondary stands than in old-growth forests, where
-proportionally more C is allocated to respiration
-\citep{kunert_understanding_2019}. {[}\emph{purpose for respiration/
-other compenents} (\textbf{See Nobby's comment in
-manuscript-draft\_NK.pdf}){]}.
+findings indicate that, among mature, undisturbed stands, forest C
+fluxes are unambiguously highest in the tropical regions, and the
+relationship with latitude is approximately linear (Fig. 2). This
+contrasts with the suggestion that productivity of temperate forests is
+similar to that of tropical forests \citep{huston_global_2009}. Compared
+to tropical forests, the temperate forest biome has experienced more
+widespread anthropogenic disturbance and has a larger fraction of
+secondary stands \citep{potapov_mapping_2008, poulter_global_2018}, so
+analyses comparing across latitudinal gradients without controlling for
+stand age risk confounding age with biome effects. In addition, because
+carbon allocation varies with stand age
+\citep{de_lucia_forest_2007, anderson-teixeira_altered_2013, doughty_what_2018},
+age differences may introduce systematic biases into analyses of C
+fluxes across latitude or global climatic gradients. For example, woody
+productivity tends to be higher in rapidly aggrading secondary stands
+than in old-growth forests, where proportionally more C is allocated to
+respiration
+\citep{de_lucia_forest_2007, piao_forest_2010, doughty_what_2018, kunert_understanding_2019}.
 
-\textbf{This paragraph needs work--including re-assessment of results,
-interpretation, writing} We show that FACF are broadly consistent in
-their responses to climate drivers on the global scale (with the
-exception of some differences in MAT-MAP interactions; Fig. 3), with no
-pronounced trends in C allocation among the variable pairs tested (Figs.
-2, S3). Although variation in allocation has been observed along
-gradients of elevation \citep{moser_elevation_2011} and water
-availability \citep{newman_above-_2006}--along with non-climatic axes of
-stand age \citep{litton_carbon_2007}, nutrient availability
-\citep{litton_carbon_2007, gill_belowground_2016}, and forest structure
-\citep{taylor_greater_2019}--little variation in relation to climate is
-apparent at the global scale within ForC, which contains the bulk of
-relevant data. Our conclusion, then, is that hypothesized gradients in
-allocation along global climate gradients cannot currently be supported
-for mature forests, although data quantity and standardization is
-currently insufficent to rule out the possibility that such trends
-exist. \textbf{bbl: remove following sentence?} Of particular interest
-and significance are the relationships amongst \(GPP\), net primary
-productivity (\(NPP\) and its components, particularly
-\(ANPP_{woody-stem}\)), and respiration (\(R_{auto}\) and components).
-There have been suggestions that tropical forests tend to have low
-carbon use efficiency (\(CUE\)=
-\(NPP\)/\(GPP\)=(\(GPP\)-\(R_{auto}\)/\(GPP\)), which are based on
-observations of low \(CUE\) in old-growth tropical forests relative to
-(mostly younger) extratropical forests
-\citep{delucia_forest_2007, malhi_productivity_2012, anderson-teixeira_carbon_2016},
-but our analysis suggests that these low values might more appropriately
-be attributed to the fact that these forests are old than to their
-tropical climate. Indeed, \(CUE\) is known to decline with forest age
-\citep{delucia_forest_2007, collalti_is_2019}, but appears to be roughly
-independent of \(GPP\) \citep{litton_carbon_2007}. Among our sites with
-relevant data, there is indication that CUE or
-\(ANPP_{woody-stem}\)/\(GPP\) increase with latitude (Fig. S3).
-Additional measurements with careful methodological standardization
-across a consistent set of mature forest sites will be necessary to
-determine whether any climate-driven gradients in allocation exist at
-the global scale.
+We show that C fluxes are broadly consistent in their responses to
+climate drivers on the global scale, with at most modest trends in C
+allocation among the variable pairs tested (Figs. 2, S3). This parallels
+the observation that C allocation across multiple C fluxes varies little
+with respect to climate along a steep tropical elevational gradient
+\citetext{\citealp{malhi_variation_2017}; \citealp[but
+see][]{moser_elevation_2011}}, and is not surprising given that trees
+face universal constraints in carbon allocation (\textbf{REFS}). The one
+trend in allocation that emerged from our analysis was a tendency for
+greater C allocation aboveground--and to foliage in particular--in warm
+tropical climates (Fig. S3). This is consistent with observations of
+increasing allocation to roots with declining temperature across a
+tropical elevational gradient \citep{moser_elevation_2011}, and with
+observations and theory predicting relatively higher belowground carbon
+allocation at higher latitudes \citep{gill_belowground_2016}.
+\textbf{Krista -- I took out the ANPP:NPP graphs from the SI after
+looking back at what Helene said in issue \#77 -- let me know if you
+think it's worth putting them back in wrt this point} It is also
+consistent with findings that as forest productivity increases,
+proportionally more carbon is allocated to \(ANPP_{foliage}\) relative
+to \(BNPP_{fine.root}\) \citep{chen_allometry_2019} or \(ANPP_{stem}\)
+\citep{hofhansl_new_2015}. \textbf{( check / comment on
+\citet{litton_carbon_2007} )}
 
-One interesting observation was that climate tends to explain more
-variation in the major fluxes (\(GPP\), \(NPP\), \(R_{auto}\) - latitude
-\(R^2\ge\) 48\%) than in subsidiary fluxes (latitude \(R^2\)
-\textless{}27\% for \(BNPP_{fine.root}\),
-\(R_{auto-root}\),\(ANPP_{woody-stem}\)) (Fig. 2; Table S2). There are
-two, non-exclusive, potential explanations for this. First, it may be
-that methodological variation is larger relative to flux magnitude for
-some of the subsidiary fluxes. Belowground fluxes in particular are
-difficult to quantify, and measurement methods for the belowground
-fluxes considered here may be measured through fundamentally different
-approaches (\emph{e.g.}, minirhizotrons, ingrowth cores, or sequential
-coring for \(BNPP_{root-fine}\); root exclusion, stable isotope
-tracking, or gas exchange of excised roots for \(R_{auto-root}\)), and
-sampling depth is variable and often insufficient to capture the full
-soil profile. \(ANPP_{woody-stem}\), which is also poorly explained by
-latitude or climate, is more straightforward to measure but is subject
-to variability introduced by differences such as biomass allometries
-applied and minimum plant size sampled (\textbf{bbl: cite
-e.g.~Huntzinger?}). However, methodological variation and uncertainty
-affect all of fluxes considered here--not necessarily any less than the
-aforementioned, and some of the larger fluxes that vary more strongly
-with respect to climate (\(ANPP\), \(NPP\)) are estimated by summing
-uncertain component fluxes. Second, differences among variables in the
-proportion of variation explained by climate may be attributable to more
-dicrect climatic control over \(GPP\) than subsidiary fluxes. That is,
+One previously hypothesized trend that was not supported by our analysis
+was that tropical forests tend to have low carbon use efficiency
+(\(CUE\)= \(NPP\)/\(GPP\)=(\(GPP\)-\(R_{auto}\)/\(GPP\)), which is based
+on observations of low \(CUE\) in old-growth tropical forests relative
+to (mostly younger) extratropical forests
+\citep{de_lucia_forest_2007, malhi_productivity_2012, anderson-teixeira_carbon_2016}.
+Our analysis, limited to mature forests, shows no such trend (Fig. S3).
+\(CUE\) is known to decline with forest age
+\citep{de_lucia_forest_2007, piao_forest_2010, collalti_is_2019}, but
+appears to be roughly independent of \(GPP\) \citep{litton_carbon_2007}.
+(\emph{This finding may have some important implications for modeling
+and our broader understanding. I'm not sure offhand how much theory has
+been built on the idea of low CUE in tropical forests because of warm
+temperatures\ldots Helene is not aware of any.})
+
+\textbf{update this paragraph
+(\href{https://github.com/forc-db/Global_Productivity/issues/78}{issue
+78})} One interesting observation was that climate tends to explain more
+variation in the major fluxes (\(GPP\), \(NPP\), \(R_{auto}\)) than in
+subsidiary fluxes (\(BNPP_{fine.root}\), \(R_{root}\), \(ANPP_{stem}\))
+as quantified by R\textsuperscript{2} (Fig. 2; Table S2). There are two,
+non-exclusive, potential explanations for this. First, it may be that
+methodological variation is larger relative to flux magnitude for some
+of the subsidiary fluxes. Belowground fluxes in particular are difficult
+to quantify, and measurement methods for the belowground fluxes
+considered here may use fundamentally different approaches in different
+sites (\emph{e.g.}, minirhizotrons, ingrowth cores, or sequential coring
+for \(BNPP_{fine.root}\); root exclusion, stable isotope tracking, or
+gas exchange of excised roots for \(R_{root}\)), and sampling depth is
+variable and often insufficient to capture the full soil profile.
+\(ANPP_{stem}\), which is also poorly explained by latitude or climate,
+is more straightforward to measure but is subject to variability
+introduced by differences such as biomass allometries applied and
+minimum plant size sampled \citep{clark_measuring_2001}. However,
+methodological variation and uncertainty affect all of fluxes considered
+here, and some of the larger fluxes that vary more strongly with respect
+to climate (\(ANPP\), \(NPP\)) are estimated by summing uncertain
+component fluxes. Second, differences among variables in the proportion
+of variation explained by climate may be attributable to more direct
+climatic control over \(GPP\) than subsidiary fluxes. That is,
 subsidiary fluxes may be shaped by climate both indirectly through its
@@
 
-The latitudinal gradient in FACF (Fig. 2) is driven primarily by
-temperature-related climate variables, and secondarily by moisture
-availability (Table 1, \emph{H2-H3}; Figs. 3-4). Because many climate
-variables co-vary across the latitudinal gradient (Fig. S2), because
-climatic drivers affect forest carbon flux on much shorter time scales
-than can be captured by annual climate summary variables, and because
-both climatic conditions and C flux vary intra- and inter-annually
-around the long-term means, it is not appropriate to attempt to identify
-any one mean annual climate variable as a mechanistic driver of FACF.
-However, it remains informative to consider these relationships. We find
-that temperature-related climate variables (\(MAT\), temperature
-seasonality, \ldots{}\textbf{LIST}) explain the highest proportion of
-variability in FACF, and among these, \(MAT\) is generally the best
-predictor--perhaps because site-specific MAT is recorded for the
-majority of sites in ForC, whereas other variables are extracted from
-global gridded data products (Table S1). The effects of temperature are
-modified by moisture availability, with reduced FACF under hot and dry
-conditions (\emph{i.e.}, high PET, high deficit; Fig. 4c-d) and
-sometimes under very high precipitaiton (Figs. 3, 4b). Negative effects
-of very high precipitation on FACF have been observed previously
-(\textbf{REFS}) and are attributable to nutrient and light limitations
-(\textbf{REFS}). Thus, although temperature and water availability
-jointly and interactively drive global-scale patterns of FACF.
+The latitudinal gradient in forest C flux (Fig. 2) is driven primarily
+by temperature-related climate variables, the effects of which are
+moderated by moisture availability (Table 1, \emph{H2-H3}; Figs. 3-4).
+MAT and MAP have long been identified as primary global-scale drivers of
+C fluxes
+\citetext{\citealp{lieth_primary_1973}; \citealp[\textbf{REFS};][]{taylor_temperature_2017}}.
+It is not appropriate to attempt to identify individual mean annual
+climate variables as mechanistic drivers of C fluxes because many
+climate variables co-vary across the latitudinal gradient (Fig. S2),
+because climatic drivers affect forest carbon flux on much shorter time
+scales than can be captured by annual climate summary variables, and
+because both climatic conditions and C flux vary intra- and
+inter-annually around the long-term means. However, it remains
+informative to consider these relationships. Among the
+temperature-related climate variables, \(MAT\) is generally the most
+strongly correlated with C fluxes (Table S2)--perhaps in part because
+site-specific MAT is recorded for the majority of sites in ForC, whereas
+other variables were extracted from global gridded data products,
+introducing spatial scaling errors.
 
-FACF are negatively correlated with temperature seasonality (Table 1,
-\emph{H4}; Fig. 4e), and is minimal during cold- or dry- dormant
+This finding supports the continued focus on MAT as a
+primary--\emph{albiet} not mechanistic--correlate of C fluxes. The
+effects of \(MAT\) are modified by moisture availability, with reduced C
+fluxes under relatively dry conditions (\emph{i.e.}, low precipitation;
+high vapour pressure deficit) and sometimes under very high
+precipitation (Figs. 3-4). \textbf{BBL: I feel like this is pretty
+repetitive of previous paragraph.} The observed positive interaction
+between MAT and MAP for \(ANPP_{stem}\) on the global scale (Fig. 3) is
+consistent with an analysis showing a similar interaction for \(ANPP\)
+in tropical forests, also with a cross-over point at
+\textasciitilde20\(\circ\)C \citep{taylor_temperature_2017}.
+\emph{(\citet{sullivan_sensitivity_2020} show a decline in
+\(ANPP_{stem}\) in sites with high max temperatures and limited
+moisture.)} However, we detect no such interaction for \(ANPP\) or most
+other C fluxes, and we find a contrasting negative interaction for
+\(NPP\) (Fig. 3), suggesting that more data are required to sort out
+potential differences in the interactive effects of MAT and MAP on C
+fluxes in the tropics.
+
+Forest autotrophic C fluxes decline with temperature seasonality (Table
+1, \emph{H4}; Fig. 4e), and are minimal during cold- or dry- dormant
 seasons. To account for this, a number of analyses seeking to
@@
 examined the relationship of C flux per month of the growing season with
-growing season climatic conditions (Table 1, \emph{H5}; \textbf{REFS}).
-We found that the sort of simple metric needed to define growing season
-at a global scale was uncertain for temperature and problematic for
-moisture (\textbf{WORK ON THIS}). A temperature-defined growing season
-length had stong positive correlation with FACF (Fig. 4f), but explained
-less variation than \(MAT\). Dividing FACFs by growing season length to
-yield FACF per growing season month removed the majority of
-climate-related variation, supporting the idea that the latitudinal
-gradient in FACF is attributable more to shorter growing seasons at high
-latitudes than to inherently lower rates of photosyntheiss or
-respiration by high-latitude forests (\emph{{[} Enquist et al. 2007 GCB-
-but check{]}}). However, there remained a number of significant
-correlations with growing season climatic conditions, suggesting that
-climatic conditions remain influential within the growing season. We
-conclude that while correcting for growing season length takes analyses
-a step closer to mechanistic linkage of instantaneous C flux rates to
-environmental conditions, it remains very crude relative to the the
-timescales on which climate affects plant metabolism and does not
-advance statistical predictive power. Rather, mechanistic accounting for
-climatic effects on global FACF patterns requires models representing
-physiologically meaningful timescales (\emph{e.g.}, \textbf{refs}).
+growing season climatic conditions \citep[Table 1,
+\emph{H5};][]{kerkhoff_plant_2005, anderson_temperature-dependence_2006, enquist_adaptive_2007, michaletz_convergence_2014}.
+The sort of simple metric that has been used to define growing season at
+a global scale \citep{kerkhoff_plant_2005} is coarse with respect to
+temperature because it's calculated on a monthly timescale and
+problematic with respect to moisture because it doesn't capture temporal
+lags between precipitation and plant water use caused by storage in soil
+or snow. We found that a temperature-defined growing season length had
+strong positive correlation with C fluxes (Fig. 4f), but was never the
+best. Dividing annual fluxes by growing season length to yield average
+flux per growing season month removed the majority of climate-related
+variation, supporting the idea that the latitudinal gradient in carbon
+flux is attributable more to shorter growing seasons at high latitudes
+than to inherently lower rates of photosynthesis or respiration by
+high-latitude forests \citep{enquist_adaptive_2007}. However, there
+remained a number of significant correlations with growing season
+climatic conditions, indicating that climatic conditions remain
+influential within the growing season. We conclude that while correcting
+for growing season length takes analyses a step closer to mechanistic
+linkage of instantaneous C flux rates to environmental conditions, it
+remains crude relative to the timescales on which climate affects plant
+metabolism, and does not advance statistical predictive power.
+Mechanistic accounting for climatic effects on global forest carbon flux
+patterns instead requires models representing physiologically meaningful
+timescales \citep[e.g., \textbf{REFS};][]{longo_biophysics_2019}.
 
-Our analysis clarifies how FACF vary with latitude and climate on a
-global scale, with some important implications for how forest carbon
-cycling relates to climate and, by extension, how it is likely to
-respond to climatic warming. Our findings show that higher temperatures
-with similar moisture availability result in a generalized acceleration
-of FACF (Figs. 2-3). This is consistent with observations of
-continental- to global-scale increases in \(GPP\)
-\citep{li_mapping_2019} and \(ANPP_{woody stem}\)
+Our analysis clarifies how forest autotrophic carbon fluxes vary with
+latitude and climate on a global scale, with some important implications
+for how forest carbon cycling relates to climate and, by extension, how
+it is likely to respond to climatic warming. Our findings show that
+higher temperatures with similar moisture availability result in a
+generalized acceleration of forest C cycling (Figs. 2-3). This is
+consistent with observations of continental- to global-scale increases
+over time in \(GPP\) \citep{li_mapping_2019} and \(ANPP_{stem}\)
 \citep{brienen_long-term_2015, hubau_asynchronous_2020}, along with some
@@
 \citep{bond-lamberty_globally_2018}. However, increasing C flux rates
-are not universal \citep{rutishauser_testing_2020} (\textbf{MORE REFS}).
-This is likely because factors other than rising temperatures are at
-play, including changes to other aspects of cliamte, atmospheric
-pollution (CO\textsubscript{2}, SO\textsubscript{2},
-NO\textsubscript{x}), and local disturbances. Morevoer, forest ecosystem
-responses to climatic changes outside the temperature range to which
-forest communities are adapated and acclimatized will not necessarily
-parallel responses across across geographic gradients in climate.
-Nevertheless, as we enter a period of accelerating climatic change,
-understanding of the fundamental climatic controls on FACF sets a
-foundation for understanding patterns of change.
+are by no means universal
+\citep[e.g.,][]{rutishauser_testing_2020, hubau_asynchronous_2020},
+likely because other factors are at play, including changes to other
+aspects of climate, atmospheric pollution (CO\textsubscript{2},
+SO\textsubscript{2}, NO\textsubscript{x}), and local disturbances.\\
+In addition, some tropical forests are \textbf{(discuss new Sullivan
+paper that finds higher max temperatures associated to lower
+ANPP\_stem)} \citep{sullivan_sensitivity_2020} Moreover, forest
+ecosystem responses to climatic changes outside the temperature range to
+which forest communities are adapted and acclimatized will not
+necessarily parallel responses across geographic gradients in climate.
+Nevertheless, understanding of the fundamental climatic controls on C
+cycling sets a firmer foundation for understanding forest C cycle
+responses to accellerating climate change.
 
-\subsection{misc content for
-discussion}\label{misc-content-for-discussion}
-
-\begin{itemize}
-\item
-  the observed positive interaction between MAT and MAP for
-  \(ANPP_woody\) is consitent with the \citet{taylor_temperature_2017}
-  analysis showing such an interaction for \(ANPP\) in tropical forests.
-  Similar to their analysis, we find a cross-over point at
-  \textasciitilde{}20C. However, we don't find such an interaction for
-  \(ANPP\), and we show a contrasting negative interaction for \(NPP\).
-  Some of this is may be stochasitic/ driven by composition of the
-  dataset, and the interactions we observe are not internally
-  consistent.
-\item
-  consistent with \citep{hofhansl_new_2015} (\textbf{verify}) , we found
-  a slight tendency for warmer sites to have higher aboveground
-  allocation
-\end{itemize}
-
--results are consistent with \emph{Muller-Landau et al., in review}
-
-\subsubsection{Acknowledgements}\label{acknowledgements}
+\hypertarget{acknowledgements}{%
+\subsubsection{Acknowledgements}\label{acknowledgements}}
 

### 2020-08-06 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd
working on discussion

@@
 ```{r echo=FALSE, out.width='100%', fig.cap = "Map showing all data used in the analysis, coded by variable. Variables are plotted individually in Fig. S1. ", fig.pos='H'}
-  knitr::include_graphics("distribution_all_variables_cropped.png")
+  knitr::include_graphics("distribution_all_variables_cropped.jpg")
 ```
@@
 
-**We show that C fluxes are broadly consistent in their responses to climate drivers on the global scale, with at most modest trends in C allocation among the variable pairs tested (Figs. 2, S3).** 
+**We show that C fluxes are broadly consistent in their responses to climate drivers on the global scale, with no trends in C allocation among the variable pairs tested (Figs. 2, 4, S3).** 
 This parallels the observation that C allocation across multiple C fluxes varies little with respect to climate along a steep tropical elevational gradient [@malhi_variation_2017; but see @moser_elevation_2011],
 and is not surprising given that trees face universal constraints in carbon allocation (**REFS**). 
-The one trend in allocation that emerged from our analysis was a tendency for greater C allocation aboveground--and to foliage in particular--in warm tropical climates (Fig.
-S3).
-This is consistent with observations of increasing allocation to roots with declining temperature across a tropical elevational gradient [@moser_elevation_2011],
-and with observations and theory predicting relatively higher belowground carbon allocation at higher latitudes [@gill_belowground_2016].
-**Krista -- I took out the ANPP:NPP graphs from the SI after looking back at what Helene said in issue #77 -- let me know if you think it's worth putting them back in wrt this point**
-It is also consistent with findings that as forest productivity increases, proportionally more carbon is allocated to $ANPP_{foliage}$ relative to $BNPP_{fine.root}$ [@chen_allometry_2019] or $ANPP_{stem}$ [@hofhansl_new_2015].
-**( check / comment on  @litton_carbon_2007 )**
-
-One previously hypothesized trend that was not supported by our analysis was that tropical forests tend to have low carbon use efficiency ($CUE$= $NPP$/$GPP$=($GPP$-$R_{auto}$/$GPP$), which is based on observations of low $CUE$ in old-growth tropical forests relative to (mostly younger) extratropical forests [@de_lucia_forest_2007; @malhi_productivity_2012; @anderson-teixeira_carbon_2016].
-Our analysis, limited to mature forests, shows no such trend (Fig.
-S3).
-$CUE$ is known to decline with forest age [@de_lucia_forest_2007; @piao_forest_2010; @collalti_is_2019], but appears to be roughly independent of $GPP$ [@litton_carbon_2007].
-(*This finding may have some important implications for modeling and our broader understanding.
-I'm not sure offhand how much theory has been built on the idea of low CUE in tropical forests because of warm temperatures...Helene is not aware of any.*)
+We find no trend in the allocation of $GPP$ between production and respiration across latitude or climate ($NPP$:$R_{auto}$; Fig.
+S3), refuting the idea that tropical forests have anomalously low $CUE$ [@de_lucia_forest_2007; @malhi_productivity_2012; @anderson-teixeira_carbon_2016].
+Rather, differences in $CUE$ between old-growth tropical forests relative to (mostly younger) extratropical forests are likely an artifact of comparing stands of different age, as $CUE$ is known to decline with forest age [@de_lucia_forest_2007; @piao_forest_2010; @collalti_is_2019].
+Another previously observed pattern for which we find no support is a tendency for belowground C allocation to decrease with increasing temperature  [@moser_elevation_2011; @gill_belowground_2016]; rather, we observe no trends in allocation between $ANPP$ and $BNPP$ across latitudes.
+Failure to detect signficant tends in C allocation with respect to climate in this analysis does not imply that none exist; rather, it suggests that
 

### 2020-08-06 manuscript/manuscript draft.Rmd
working on discussion

@@
 
-**update this paragraph ([issue 78](https://github.com/forc-db/Global_Productivity/issues/78))** 
 *Despite the broad consistency of climate responses across C fluxes, climate explains lower proportions of variability among some of the subsidiary C fluxes (*e.g.*, $ANPP_{stem}$, $BNPP$  $BNPP_{fine.root)$; Fig.
@@
 
-*At least at this global scale, we find little evidence of any non-linearity in temperature's influence on C flux.* 
-The latitudinal gradient in forest C flux (Fig.
-2) is driven primarily by temperature-related climate variables, the effects of which are moderated by moisture availability (Table 1, *H2-H3*; Figs.
-3-4).
+*Temperature and its seasonality were the primary drivers of C flux on the global scale (Figs.
+2-4), consistent with a long legacy of research identifying temperature as a primary driver of forest ecosystem C cycling [e.g., @lieth_primary_1973; **REFS].*
+We find little evidence of any non-linearity in temperature's influence on C flux.
+The relationship of all fluxes to $MAT$ as an individual driver were best described by a linear function (Table S2) - with the exception of $BNPP$, whose response to $MAT$ was close to linear (Fig.
+4a).
+This result contrasts with both the idea that C flux responds exponentially to temperature (**REFS**) and with the idea that fluxes saturate with $MAT$ below approximately 25$^\circ$C MAT (**REFS**).
+It remains possible that flux declines above this threshold [@larjavaara_temperature_2012; @sullivan_sensitivity_2020], as is also consistent with tree-ring records indicating that tropical tree growth declines at high temperatures [e.g., @vlam_temperature_2014].
+However, these higher temperatures also tend to be associated with drier conditions (*i.e.*, low precipitation, high PET and VPD), all of which are associated with reduced C fluxes (Figs.
+4c-d, S4-S5).
+Indeed, while temperature responses dominate at this global scale and with the climate space occupied by forests, the effects of temperature are moderated by moisture availability (Table 1; Figs 3-4).
 MAT and MAP have long been identified as primary global-scale drivers of C fluxes [@lieth_primary_1973; **REFS**; @taylor_temperature_2017].
-It is not appropriate to attempt to identify individual mean annual climate variables as mechanistic drivers of C fluxes because many climate variables co-vary across the latitudinal gradient (Fig.
+
+*It is not appropriate to attempt to identify individual mean annual climate variables as mechanistic drivers of C fluxes because many climate variables co-vary across the latitudinal gradient (Fig.
 S2), because climatic drivers affect forest carbon flux on much shorter time scales than can be captured by annual climate summary variables, and because both climatic conditions and C flux vary intra- and inter-annually around the long-term means.
 However, it remains informative to consider these relationships.
-Among the temperature-related climate variables, $MAT$ is generally the most strongly correlated with C fluxes (Table S2)--perhaps in part because site-specific MAT is recorded for the majority of sites in ForC, whereas other variables were extracted from global gridded data products, introducing spatial scaling errors.
+Among the temperature-related climate variables, $MAT$ is generally the most strongly correlated with C fluxes (Table S2)--perhaps in part because site-specific MAT is recorded for the majority of sites in ForC, whereas other variables were extracted from global gridded data products, introducing spatial scaling errors.*
 
@@
 3) is consistent with an analysis showing a similar interaction for $ANPP$ in tropical forests, also with a cross-over point at ~20$\circ$C [@taylor_temperature_2017].
-*(@sullivan_sensitivity_2020 show a decline in $ANPP_{stem}$ in sites with high max temperatures and limited moisture.)* However, we detect no such interaction for $ANPP$ or most other C fluxes, and we find a contrasting negative interaction for $NPP$ (Fig.
+However, we detect no such interaction for $ANPP$ or most other C fluxes, and we find a contrasting negative interaction for $NPP$ (Fig.
 3), suggesting that more data are required to sort out potential differences in the interactive effects of MAT and MAP on C fluxes in the tropics.
 
-Forest autotrophic C fluxes decline with temperature seasonality (Table 1, *H4*; Fig.
-4e), and are minimal during cold- or dry- dormant seasons.
+Forest autotrophic C fluxes decline with temperature seasonality (Table 1, *H4*; Fig. 4e), and are minimal during cold- or dry- dormant seasons. 
 To account for this, a number of analyses seeking to characterize global-scale effects of climate on productivity have examined the relationship of C flux per month of the growing season with growing season climatic conditions [Table 1, *H5*; @kerkhoff_plant_2005; @anderson_temperature-dependence_2006; @enquist_adaptive_2007; @michaletz_convergence_2014].
-The sort of simple metric that has been used to define growing season at a global scale [@kerkhoff_plant_2005] is coarse with respect to temperature because it's calculated on a monthly timescale and problematic with respect to moisture because it doesn't capture temporal lags between precipitation and plant water use caused by storage in soil or snow.
-We found that a temperature-defined growing season length had strong positive correlation with C fluxes (Fig.
-4f), but was never the best.
+The sort of simple metric that has been used to define growing season at a global scale [@kerkhoff_plant_2005] is coarse with respect to temperature because it is calculated on a monthly timescale and problematic with respect to moisture because it doesn't capture temporal lags between precipitation and plant water use caused by storage in soil or snow.
+We found that a temperature-defined growing season length had strong positive correlation with C fluxes (Fig. 4f), but was never the best univariate predictor. 
 Dividing annual fluxes by growing season length to yield average flux per growing season month removed the majority of climate-related variation, supporting the idea that the latitudinal gradient in carbon flux is attributable more to shorter growing seasons at high latitudes than to inherently lower rates of photosynthesis or respiration by high-latitude forests [@enquist_adaptive_2007].
-However, there remained a number of significant correlations with growing season climatic conditions, indicating that climatic conditions remain influential within the growing season.
+However, there remained a number of significant correlations with growing season climatic conditions, indicating that climatic conditions remain influential within the growing season (Figs. S8-S9). 
 We conclude that while correcting for growing season length takes analyses a step closer to mechanistic linkage of instantaneous C flux rates to environmental conditions, it remains crude relative to the timescales on which climate affects plant metabolism, and does not advance statistical predictive power.
-Mechanistic accounting for climatic effects on global forest carbon flux patterns instead requires models representing physiologically meaningful timescales [e.g., **REFS**; @longo_biophysics_2019].
+Mechanistic accounting for climatic effects on global forest carbon flux patterns instead requires models representing physiologically meaningful timescales [e.g., @medvigy_mechanistic_2009; @longo_biophysics_2019].
 
 Our analysis clarifies how forest autotrophic carbon fluxes vary with latitude and climate on a global scale, with some important implications for how forest carbon cycling relates to climate and, by extension, how it is likely to respond to climatic warming.
-Our findings show that higher temperatures with similar moisture availability result in a generalized acceleration of forest C cycling (Figs. 2-3).
+To the extend that patterns across broad scale climatic gradients can foretell how ecosystem responses to climate change, our findings suggest that higher temperatures with similar moisture availability result in a generalized acceleration of forest C cycling (Figs.
+2-3).
 This is consistent with observations of continental- to global-scale increases over time in $GPP$ [@li_mapping_2019] and $ANPP_{stem}$ [@brienen_long-term_2015; @hubau_asynchronous_2020], along with some C cycle components not considered here--tree mortality [@brienen_long-term_2015; @mcdowell_drivers_2018], soil respiration [@bond-lamberty_global_2010], and heterotrophic soil respiration [@bond-lamberty_globally_2018].
 However, increasing C flux rates are by no means universal [e.g., @rutishauser_testing_2020; @hubau_asynchronous_2020], likely because other factors are at play, including changes to other aspects of climate, atmospheric pollution (CO~2~, SO~2~, NO~x~), and local disturbances.
-In addition, some tropical forests are
-**(discuss new Sullivan paper that finds higher max temperatures associated to lower ANPP_stem)** [@sullivan_sensitivity_2020]
 Moreover, forest ecosystem responses to climatic changes outside the temperature range to which forest communities are adapted and acclimatized will not necessarily parallel responses across geographic gradients in climate.
-Nevertheless, understanding of the fundamental climatic controls on C cycling sets a firmer foundation for understanding forest C cycle responses to accellerating climate change.
+Indeed, tree-ring studies from forests around the world indicate that tree growth rates - along with $ANPP_{stem}$ and possibly other ecosystem C fluxes - respond negatively to temperature [@sniderhan_growth_2016;@helcoski_growing_2019].
+Furthermore, in the tropics, climate change will push forests beyond any contemporary climate, and there are some indications that this could reduce C flux rates  [@sullivan_sensitivity_2020] (**REFS**).
+Thus, while further research is required to understand the extent to which forest responses to climate change will track global gradients, and the time scale on which they will do so, understanding the fundamental climatic controls on annual C cycling in Earth's forests sets a firmer foundation for understanding forest C cycle responses to accelerating climate change.
 

### 2020-08-06 manuscript/manuscript draft.Rmd
finished first pass at revising discussion

@@
 
-**Our analysis of a large global database (ForC) clarifies how autotrophic C fluxes in mature forests vary with latitude and climate on a global scale.** 
+Our analysis of a large global database (ForC) clarifies how autotrophic C fluxes in mature forests vary with latitude and climate on a global scale.
 We show that, across all nine variables analyzed, annual C flux decreases continually with latitude (Fig.
@@
 
-**Our findings indicate that, among mature, undisturbed stands, forest C fluxes are unambiguously highest in the tropical regions, and the relationship with both latitude and $MAT$ is approximately linear (Figs.
-2, 4).**
+Our findings indicate that, among mature, undisturbed stands, forest C fluxes are unambiguously highest in the tropical regions, and the relationship with both latitude and $MAT$ is approximately linear (Figs.
+2, 4).
 This contrasts with the suggestion that C fluxes (e.g., $NPP$, $ANPP$, $ANPP_{stem}$) of temperate forests are similar to or even greater than that of tropical forests [@luyssaert_co_2007; @huston_global_2009].
@@
 
-**We show that C fluxes are broadly consistent in their responses to climate drivers on the global scale, with no trends in C allocation among the variable pairs tested (Figs. 2, 4, S3).** 
+We show that C fluxes are broadly consistent in their responses to climate drivers on the global scale, with no trends in C allocation among the variable pairs tested (Figs. 2, 4, S3).
 This parallels the observation that C allocation across multiple C fluxes varies little with respect to climate along a steep tropical elevational gradient [@malhi_variation_2017; but see @moser_elevation_2011],
@@
 
-*Despite the broad consistency of climate responses across C fluxes, climate explains lower proportions of variability among some of the subsidiary C fluxes (*e.g.*, $ANPP_{stem}$, $BNPP$  $BNPP_{fine.root)$; Fig.
-2; Tables S2, S6).*
+Despite the broad consistency of climate responses across C fluxes, climate explains lower proportions of variability among some of the subsidiary C fluxes (*e.g.*, $ANPP_{stem}$, $BNPP$  $BNPP_{fine.root)$; Fig.
+2; Tables S2, S6).
 There are two, non-exclusive, potential explanations for this. 
@@
 
-*Temperature and its seasonality were the primary drivers of C flux on the global scale (Figs.
-2-4), consistent with a long legacy of research identifying temperature as a primary driver of forest ecosystem C cycling [e.g., @lieth_primary_1973; **REFS].*
+Temperature and its seasonality were the primary drivers of C flux on the global scale (Figs.
+2-4), consistent with a long legacy of research identifying temperature as a primary driver of forest ecosystem C cycling [e.g., @lieth_primary_1973; **REFS**].
 We find little evidence of any non-linearity in temperature's influence on C flux.
@@
 It remains possible that flux declines above this threshold [@larjavaara_temperature_2012; @sullivan_sensitivity_2020], as is also consistent with tree-ring records indicating that tropical tree growth declines at high temperatures [e.g., @vlam_temperature_2014].
-However, these higher temperatures also tend to be associated with drier conditions (*i.e.*, low precipitation, high PET and VPD), all of which are associated with reduced C fluxes (Figs.
-4c-d, S4-S5).
+However, these higher temperatures also tend to be associated with high PET and VPD, both of which are associated with reduced C fluxes (Figs. 4c-d, S4-S5).
 Indeed, while temperature responses dominate at this global scale and with the climate space occupied by forests, the effects of temperature are moderated by moisture availability (Table 1; Figs 3-4).
-MAT and MAP have long been identified as primary global-scale drivers of C fluxes [@lieth_primary_1973; **REFS**; @taylor_temperature_2017].
-
-*It is not appropriate to attempt to identify individual mean annual climate variables as mechanistic drivers of C fluxes because many climate variables co-vary across the latitudinal gradient (Fig.
-S2), because climatic drivers affect forest carbon flux on much shorter time scales than can be captured by annual climate summary variables, and because both climatic conditions and C flux vary intra- and inter-annually around the long-term means.
-However, it remains informative to consider these relationships.
-Among the temperature-related climate variables, $MAT$ is generally the most strongly correlated with C fluxes (Table S2)--perhaps in part because site-specific MAT is recorded for the majority of sites in ForC, whereas other variables were extracted from global gridded data products, introducing spatial scaling errors.*
-
-This finding supports the continued focus on MAT as a primary--*albiet* not mechanistic--correlate of C fluxes.
-The effects of $MAT$ are modified by moisture availability, with reduced C fluxes under relatively dry conditions (*i.e.*, low precipitation; high vapour pressure deficit) and sometimes under very high precipitation (Figs.
-3-4).
-**BBL: I feel like this is pretty repetitive of previous paragraph.** The observed positive interaction between MAT and MAP for $ANPP_{stem}$ on the global scale (Fig.
+Specifically, C fluxes are reduced under relatively dry conditions (*i.e.*, low $MAP$; high $VPD$) and sometimes under very high precipitation (Figs. 3-4). 
+The observed positive interaction between $MAT$ and $MAP$ for $ANPP_{stem}$ on the global scale (Fig.
 3) is consistent with an analysis showing a similar interaction for $ANPP$ in tropical forests, also with a cross-over point at ~20$\circ$C [@taylor_temperature_2017].

### 2020-08-06 manuscript/manuscript-draft.tex
finished first pass at revising discussion

@@
 \citep[e.g.,][]{luyssaert_co_2007, piao_forest_2010, gillman_latitude_2015, li_mapping_2019},
-but it remains unclear whether and how the shape of this relationship
-varies among fluxes. For instance, gross primary productivity (\(GPP\))
-increases continuously with decreasing latitude and is indisputably
-highest in tropical forests
+but studies have differed in their conclusions regarding the shape of
+this relationship - quite possibly because of lack of standardization
+with respect to methodology and stand history. For instance, studies
+agree that gross primary productivity (\(GPP\)) increases continuously
+with decreasing latitude and is indisputably highest in tropical forests
 \citep{luyssaert_co_2007, beer_terrestrial_2010, jung_global_2011, badgley_terrestrial_2019, li_mapping_2019}.
@@
 \(NPP\)/\(GPP\);][]{de_lucia_forest_2007, malhi_productivity_2012, anderson-teixeira_carbon_2016}.
-However, accurately comparing relationships of multiple forest C fluxes
-across latitude requires a large database with standardization for
-methodology, stand history, and management, which to date has not been
-possible.
+Such differences among C fluxes their relationship to latitude could
+have profound implications for our understanding of the C cycle and its
+climate sensitivity. However, until recently the potential to compare
+latitudinal trends across C fluxes has been limited by lack of a large
+database with standardization for methodology, stand history, and
+management \citep{anderson-teixeira_forc_2018}.
 
@@
 4, S4, S5\\
-\hspace{1em}H2.3. Temperature and precipitation interactively shape C fluxes.\textsuperscript{5} & (yes) & yes & yes & - & yes & - & yes & yes & yes & - & Fig. 3\\
+\hspace{1em}H2.3. Temperature and precipitation jointly shape C fluxes.\textsuperscript{5} & (yes) & yes & yes & yes & yes & - & - & - & yes & - & Fig. 3, Table S3\\
 \addlinespace[1em]
@@
 
-Our analysis of a large global database (ForC) reveals how autotrophic
-carbon fluxes in mature forests vary with latitude and climate on a
-global scale. We show that, across all nine forest autotrophic C fluxes
-analyzed, C cycling decreases continually with latitude (\emph{H1.1};
-Fig. 2)--a finding that confirms multiple previous studies but
-contradicts the idea that productivity of temperate forests rivals or
-even exceeds that of tropical forests
+Our analysis of a large global database (ForC) clarifies how autotrophic
+C fluxes in mature forests vary with latitude and climate on a global
+scale. We show that, across all nine variables analyzed, annual C flux
+decreases continually with latitude (Fig. 2)--a finding that confirms
+multiple previous studies but contradicts the idea that productivity of
+temperate forests rivals or even exceeds that of tropical forests
 \citep{luyssaert_co_2007, huston_global_2009}. At this global scale, C
@@
 detectable differences in allocation across latitude or climates (Figs.
-2, 4, S3). However, climate explains lower proportions of variability
-among some of the subsidiary C fluxes (\emph{e.g.}, \(ANPP_{stem}\),
-\(BNPP\) \(BNPP_{fine.root)\); Fig. 2; Tables S2, S6). Latitudinal
-variation in C fluxes is primarily attributable to temperature-related
-variables (\emph{H3, H4}), particularly MAT and temperature seasonality
-(Figs. 3-4). Water availability is also influential, but generally of
+2, 4, S3). Similarly, we show broad - \emph{albeit} not complete -
+consistency of climate responses across C fluxes, with the observed
+latitudinal variation primarily attributable to temperature and its
+seasonality (Figs. 3-4). Water availability is also influential, but of
 secondary importance across the climate space occupied by forests (Figs.
-3-4). While growing season length explained signficant variation in all
-C fluxes (\emph{H4}; Figs. 4f, S6-S7), it did not improve upon MAT or
-temperature seasonality as a predictor. Within the growing season, the
-influence of climate on C cycling is smaller but still significant for a
-number of C fluxes (\emph{H5}; Fig. S9; Table S4).
+3-4). Contrary to prior suggestions that the majority of variation in C
+cycling is driven by the length of the growing season (\textbf{REFS}),
+we find modest explanatory power of growing season length and small but
+sometimes significant influence of climate within the growing season
+(Figs. 4f,S6-S9). Together, these findings yield a unified understanding
+of climate's influence on forest C cycling.
 
-These findings clarify how forest C fluxes vary with latitude and
-climate on a global scale. Past studies have differed in their
-conclusions regarding the relationship between C fluxes and latitude or
-its correlates (Table 1, \emph{H1})--quite possibly because of lack of
-standardization with respect to stand age and disturbance history. Our
-findings indicate that, among mature, undisturbed stands, forest C
+Our findings indicate that, among mature, undisturbed stands, forest C
 fluxes are unambiguously highest in the tropical regions, and the
-relationship with latitude is approximately linear (Fig. 2). This
-contrasts with the suggestion that productivity of temperate forests is
-similar to that of tropical forests \citep{huston_global_2009}. Compared
-to tropical forests, the temperate forest biome has experienced more
-widespread anthropogenic disturbance and has a larger fraction of
-secondary stands \citep{potapov_mapping_2008, poulter_global_2018}, so
-analyses comparing across latitudinal gradients without controlling for
-stand age risk confounding age with biome effects. In addition, because
-carbon allocation varies with stand age
+relationship with both latitude and \(MAT\) is approximately linear
+(Figs. 2, 4). This contrasts with the suggestion that C fluxes (e.g.,
+\(NPP\), \(ANPP\), \(ANPP_{stem}\)) of temperate forests are similar to
+or even greater than that of tropical forests
+\citep{luyssaert_co_2007, huston_global_2009}. Previous indications of
+such a pattern may have been an artifact of differences in stand age
+across biomes. Compared to tropical forests, the temperate forest biome
+has experienced more widespread anthropogenic disturbance and has a
+larger fraction of secondary stands
+\citep{potapov_mapping_2008, poulter_global_2018}, so analyses comparing
+across latitudinal gradients without controlling for stand age risk
+confounding age with biome effects. Because carbon allocation varies
+with stand age
 \citep{de_lucia_forest_2007, anderson-teixeira_altered_2013, doughty_what_2018},
@@
 than in old-growth forests, where proportionally more C is allocated to
-respiration
+respiration and non-woody productivity
 \citep{de_lucia_forest_2007, piao_forest_2010, doughty_what_2018, kunert_understanding_2019}.
+Thus, findings that temperate forest productivity rivals that of
+tropical forests are likely an artifact of different forest ages across
+biomes.
 
 We show that C fluxes are broadly consistent in their responses to
-climate drivers on the global scale, with at most modest trends in C
-allocation among the variable pairs tested (Figs. 2, S3). This parallels
-the observation that C allocation across multiple C fluxes varies little
+climate drivers on the global scale, with no trends in C allocation
+among the variable pairs tested (Figs. 2, 4, S3). This parallels the
+observation that C allocation across multiple C fluxes varies little
 with respect to climate along a steep tropical elevational gradient
@@
 see][]{moser_elevation_2011}}, and is not surprising given that trees
-face universal constraints in carbon allocation (\textbf{REFS}). The one
-trend in allocation that emerged from our analysis was a tendency for
-greater C allocation aboveground--and to foliage in particular--in warm
-tropical climates (Fig. S3). This is consistent with observations of
-increasing allocation to roots with declining temperature across a
-tropical elevational gradient \citep{moser_elevation_2011}, and with
-observations and theory predicting relatively higher belowground carbon
-allocation at higher latitudes \citep{gill_belowground_2016}.
-\textbf{Krista -- I took out the ANPP:NPP graphs from the SI after
-looking back at what Helene said in issue \#77 -- let me know if you
-think it's worth putting them back in wrt this point} It is also
-consistent with findings that as forest productivity increases,
-proportionally more carbon is allocated to \(ANPP_{foliage}\) relative
-to \(BNPP_{fine.root}\) \citep{chen_allometry_2019} or \(ANPP_{stem}\)
-\citep{hofhansl_new_2015}. \textbf{( check / comment on
-\citet{litton_carbon_2007} )}
+face universal constraints in carbon allocation (\textbf{REFS}). We find
+no trend in the allocation of \(GPP\) between production and respiration
+across latitude or climate (\(NPP\):\(R_{auto}\); Fig. S3), refuting the
+idea that tropical forests have anomalously low \(CUE\)
+\citep{de_lucia_forest_2007, malhi_productivity_2012, anderson-teixeira_carbon_2016}.
+Rather, differences in \(CUE\) between old-growth tropical forests
+relative to (mostly younger) extratropical forests are likely an
+artifact of comparing stands of different age, as \(CUE\) is known to
+decline with forest age
+\citep{de_lucia_forest_2007, piao_forest_2010, collalti_is_2019}.
+Another previously observed pattern for which we find no support is a
+tendency for belowground C allocation to decrease with increasing
+temperature \citep{moser_elevation_2011, gill_belowground_2016}; rather,
+we observe no trends in allocation between \(ANPP\) and \(BNPP\) across
+latitudes. Failure to detect signficant tends in C allocation with
+respect to climate in this analysis does not imply that none exist;
+rather, it suggests that
 
-One previously hypothesized trend that was not supported by our analysis
-was that tropical forests tend to have low carbon use efficiency
-(\(CUE\)= \(NPP\)/\(GPP\)=(\(GPP\)-\(R_{auto}\)/\(GPP\)), which is based
-on observations of low \(CUE\) in old-growth tropical forests relative
-to (mostly younger) extratropical forests
-\citep{de_lucia_forest_2007, malhi_productivity_2012, anderson-teixeira_carbon_2016}.
-Our analysis, limited to mature forests, shows no such trend (Fig. S3).
-\(CUE\) is known to decline with forest age
-\citep{de_lucia_forest_2007, piao_forest_2010, collalti_is_2019}, but
-appears to be roughly independent of \(GPP\) \citep{litton_carbon_2007}.
-(\emph{This finding may have some important implications for modeling
-and our broader understanding. I'm not sure offhand how much theory has
-been built on the idea of low CUE in tropical forests because of warm
-temperatures\ldots Helene is not aware of any.})
-
-\textbf{update this paragraph
-(\href{https://github.com/forc-db/Global_Productivity/issues/78}{issue
-78})} One interesting observation was that climate tends to explain more
-variation in the major fluxes (\(GPP\), \(NPP\), \(R_{auto}\)) than in
-subsidiary fluxes (\(BNPP_{fine.root}\), \(R_{root}\), \(ANPP_{stem}\))
-as quantified by R\textsuperscript{2} (Fig. 2; Table S2). There are two,
+Despite the broad consistency of climate responses across C fluxes,
+climate explains lower proportions of variability among some of the
+subsidiary C fluxes (\emph{e.g.}, \(ANPP_{stem}\), \(BNPP\)
+\(BNPP_{fine.root)\); Fig. 2; Tables S2, S6). There are two,
 non-exclusive, potential explanations for this. First, it may be that
@@
 
-The latitudinal gradient in forest C flux (Fig. 2) is driven primarily
-by temperature-related climate variables, the effects of which are
-moderated by moisture availability (Table 1, \emph{H2-H3}; Figs. 3-4).
-MAT and MAP have long been identified as primary global-scale drivers of
-C fluxes
-\citetext{\citealp{lieth_primary_1973}; \citealp[\textbf{REFS};][]{taylor_temperature_2017}}.
-It is not appropriate to attempt to identify individual mean annual
-climate variables as mechanistic drivers of C fluxes because many
-climate variables co-vary across the latitudinal gradient (Fig. S2),
-because climatic drivers affect forest carbon flux on much shorter time
-scales than can be captured by annual climate summary variables, and
-because both climatic conditions and C flux vary intra- and
-inter-annually around the long-term means. However, it remains
-informative to consider these relationships. Among the
-temperature-related climate variables, \(MAT\) is generally the most
-strongly correlated with C fluxes (Table S2)--perhaps in part because
-site-specific MAT is recorded for the majority of sites in ForC, whereas
-other variables were extracted from global gridded data products,
-introducing spatial scaling errors.
-
-This finding supports the continued focus on MAT as a
-primary--\emph{albiet} not mechanistic--correlate of C fluxes. The
-effects of \(MAT\) are modified by moisture availability, with reduced C
-fluxes under relatively dry conditions (\emph{i.e.}, low precipitation;
-high vapour pressure deficit) and sometimes under very high
-precipitation (Figs. 3-4). \textbf{BBL: I feel like this is pretty
-repetitive of previous paragraph.} The observed positive interaction
-between MAT and MAP for \(ANPP_{stem}\) on the global scale (Fig. 3) is
-consistent with an analysis showing a similar interaction for \(ANPP\)
-in tropical forests, also with a cross-over point at
-\textasciitilde20\(\circ\)C \citep{taylor_temperature_2017}.
-\emph{(\citet{sullivan_sensitivity_2020} show a decline in
-\(ANPP_{stem}\) in sites with high max temperatures and limited
-moisture.)} However, we detect no such interaction for \(ANPP\) or most
-other C fluxes, and we find a contrasting negative interaction for
-\(NPP\) (Fig. 3), suggesting that more data are required to sort out
-potential differences in the interactive effects of MAT and MAP on C
-fluxes in the tropics.
+Temperature and its seasonality were the primary drivers of C flux on
+the global scale (Figs. 2-4), consistent with a long legacy of research
+identifying temperature as a primary driver of forest ecosystem C
+cycling {[}e.g., \citet{lieth_primary_1973}; \textbf{REFS}{]}. We find
+little evidence of any non-linearity in temperature's influence on C
+flux. The relationship of all fluxes to \(MAT\) as an individual driver
+were best described by a linear function (Table S2) - with the exception
+of \(BNPP\), whose response to \(MAT\) was close to linear (Fig. 4a).
+This result contrasts with both the idea that C flux responds
+exponentially to temperature (\textbf{REFS}) and with the idea that
+fluxes saturate with \(MAT\) below approximately 25\(^\circ\)C MAT
+(\textbf{REFS}). It remains possible that flux declines above this
+threshold
+\citep{larjavaara_temperature_2012, sullivan_sensitivity_2020}, as is
+also consistent with tree-ring records indicating that tropical tree
+growth declines at high temperatures
+\citep[e.g.,][]{vlam_temperature_2014}. However, these higher
+temperatures also tend to be associated with high PET and VPD, both of
+which are associated with reduced C fluxes (Figs. 4c-d, S4-S5). Indeed,
+while temperature responses dominate at this global scale and with the
+climate space occupied by forests, the effects of temperature are
+moderated by moisture availability (Table 1; Figs 3-4). Specifically, C
+fluxes are reduced under relatively dry conditions (\emph{i.e.}, low
+\(MAP\); high \(VPD\)) and sometimes under very high precipitation
+(Figs. 3-4). The observed positive interaction between \(MAT\) and
+\(MAP\) for \(ANPP_{stem}\) on the global scale (Fig. 3) is consistent
+with an analysis showing a similar interaction for \(ANPP\) in tropical
+forests, also with a cross-over point at \textasciitilde20\(\circ\)C
+\citep{taylor_temperature_2017}.\\
+However, we detect no such interaction for \(ANPP\) or most other C
+fluxes, and we find a contrasting negative interaction for \(NPP\) (Fig.
+3), suggesting that more data are required to sort out potential
+differences in the interactive effects of MAT and MAP on C fluxes in the
+tropics.
 
@@
 a global scale \citep{kerkhoff_plant_2005} is coarse with respect to
-temperature because it's calculated on a monthly timescale and
+temperature because it is calculated on a monthly timescale and
 problematic with respect to moisture because it doesn't capture temporal
@@
 strong positive correlation with C fluxes (Fig. 4f), but was never the
-best. Dividing annual fluxes by growing season length to yield average
-flux per growing season month removed the majority of climate-related
-variation, supporting the idea that the latitudinal gradient in carbon
-flux is attributable more to shorter growing seasons at high latitudes
-than to inherently lower rates of photosynthesis or respiration by
-high-latitude forests \citep{enquist_adaptive_2007}. However, there
-remained a number of significant correlations with growing season
-climatic conditions, indicating that climatic conditions remain
-influential within the growing season. We conclude that while correcting
-for growing season length takes analyses a step closer to mechanistic
+best univariate predictor. Dividing annual fluxes by growing season
+length to yield average flux per growing season month removed the
+majority of climate-related variation, supporting the idea that the
+latitudinal gradient in carbon flux is attributable more to shorter
+growing seasons at high latitudes than to inherently lower rates of
+photosynthesis or respiration by high-latitude forests
+\citep{enquist_adaptive_2007}. However, there remained a number of
+significant correlations with growing season climatic conditions,
+indicating that climatic conditions remain influential within the
+growing season (Figs. S8-S9). We conclude that while correcting for
+growing season length takes analyses a step closer to mechanistic
 linkage of instantaneous C flux rates to environmental conditions, it
@@
 patterns instead requires models representing physiologically meaningful
-timescales \citep[e.g., \textbf{REFS};][]{longo_biophysics_2019}.
+timescales
+\citep[e.g.,][]{medvigy_mechanistic_2009, longo_biophysics_2019}.
 
@@
 for how forest carbon cycling relates to climate and, by extension, how
-it is likely to respond to climatic warming. Our findings show that
-higher temperatures with similar moisture availability result in a
-generalized acceleration of forest C cycling (Figs. 2-3). This is
-consistent with observations of continental- to global-scale increases
-over time in \(GPP\) \citep{li_mapping_2019} and \(ANPP_{stem}\)
+it is likely to respond to climatic warming. To the extend that patterns
+across broad scale climatic gradients can foretell how ecosystem
+responses to climate change, our findings suggest that higher
+temperatures with similar moisture availability result in a generalized
+acceleration of forest C cycling (Figs. 2-3). This is consistent with
+observations of continental- to global-scale increases over time in
+\(GPP\) \citep{li_mapping_2019} and \(ANPP_{stem}\)
 \citep{brienen_long-term_2015, hubau_asynchronous_2020}, along with some
@@
 SO\textsubscript{2}, NO\textsubscript{x}), and local disturbances.\\
-In addition, some tropical forests are \textbf{(discuss new Sullivan
-paper that finds higher max temperatures associated to lower
-ANPP\_stem)} \citep{sullivan_sensitivity_2020} Moreover, forest
-ecosystem responses to climatic changes outside the temperature range to
-which forest communities are adapted and acclimatized will not
-necessarily parallel responses across geographic gradients in climate.
-Nevertheless, understanding of the fundamental climatic controls on C
-cycling sets a firmer foundation for understanding forest C cycle
-responses to accellerating climate change.
+Moreover, forest ecosystem responses to climatic changes outside the
+temperature range to which forest communities are adapted and
+acclimatized will not necessarily parallel responses across geographic
+gradients in climate. Indeed, tree-ring studies from forests around the
+world indicate that tree growth rates - along with \(ANPP_{stem}\) and
+possibly other ecosystem C fluxes - respond negatively to temperature
+\citep{sniderhan_growth_2016, helcoski_growing_2019}. Furthermore, in
+the tropics, climate change will push forests beyond any contemporary
+climate, and there are some indications that this could reduce C flux
+rates \citep{sullivan_sensitivity_2020} (\textbf{REFS}). Thus, while
+further research is required to understand the extent to which forest
+responses to climate change will track global gradients, and the time
+scale on which they will do so, understanding the fundamental climatic
+controls on annual C cycling in Earth's forests sets a firmer foundation
+for understanding forest C cycle responses to accelerating climate
+change.
 

### 2020-08-06 manuscript/manuscript draft.Rmd
working on refs... seem to have broken it :(

@@
 
-The latitudinal gradient in forest C flux rates, along with altitudinal gradients [@girardin_net_2010; @malhi_variation_2017], is driven primarily by climate, which is a significant driver of C fluxes across broad spatial scales [@luyssaert_co_2007; @cleveland_relationships_2011; @hursh_sensitivity_2017].
+The latitudinal gradient in forest C flux rates, along with altitudinal gradients [@girardin_net_2010; @malhi_variation_2017], is driven primarily by climate, which is a significant driver of C fluxes across broad spatial scales [@luyssaert_co_2007; @cleveland_relationships_2011; @wei_forest_2010].
 However, there is little consensus as to the shapes of these relationships or the best predictor variables.
@@
 C fluxes increase strongly with $MAT$ on the global scale, but whether they saturate or potentially decrease at higher temperatures remains disputed. 
-Some studies have detected no deceleration or decline in $GPP$ [@luyssaert_co_2007] or $NPP$ [@schuur_productivity_2003] with increasing MAT.
-Some have suggested that relationship between C flux and temperature is exponential in nature [@allen_linking_2005] (**REFS**).
-In contrast, others have found evidence of saturation or decline of C flux in the warmest climates; @luyssaert_co_2007 found $NPP$ saturating at around 10$^\circ$C MAT; @larjavaara_temperature_2012 found that increases in $GPP$ saturate at approximately 25$^\circ$C MAT, and @sullivan_sensitivity_2020 found that, within the tropics, $ANPP_{stem}$ decreases at the highest maximum temperatures.
-Various measures of primary productivity saturate at high levels of MAP, though the saturation points identified vary from 1500mm [@luyssaert_co_2007] up to 2,445 mm MAP [@schuur_productivity_2003]. 
+Some studies have detected no deceleration or decline in $GPP$ [@luyssaert_co_2007], $NPP$ [@schuur_productivity_2003], or root respiration [$R_{root}$; @wei_forest_2010] with increasing $MAT$.
+In contrast, others have found evidence of saturation or decline of C flux in the warmest climates; @luyssaert_co_2007 found $NPP$ saturating at around 10$^\circ$C MAT; @larjavaara_temperature_2012 found that increases in $GPP$ saturate at approximately 25$^\circ$C MAT, and @sullivan_long-term_2020 found that, within the tropics, $ANPP_{stem}$ decreases at the highest maximum temperatures.
+C fluxes generally saturate at high levels of MAP, though the saturation points identified vary from $MAP$ of ~1000 mm for $R_{root}$ [@wei_forest_2010]  up to 2,445 mm for $NPP$ [@schuur_productivity_2003].
 Interactions between MAT and MAP are also possible; within the tropics, there is a positive interaction between MAT and MAP in shaping $ANPP$, such that high rainfall has a negative effect on productivity in cooler climates, compared to a positive effect in warmer climates [@taylor_temperature_2017].
-**Such complicated dynamics play out with belowground fluxes as well, with typically dominant soil temperature mediated by moisture and C supply [@hursh_sensitivity_2017; @xu_contribution_2016].
-https://github.com/forc-db/Global_Productivity/issues/111**
-There is also evidence that C fluxes also respond to climate variables such as temperature and precipitation seasonality (**REFS**), cloud cover [@taylor_temperature_2017], solar radiation [@beer_terrestrial_2010; @fyllas_solar_2017], and potential evapotranspiration [@kerkhoff_plant_2005]; however, these are not typically assessed in global-scale analyses of annual forest C flux.
+There is also evidence that C fluxes also respond to climate variables such as temperature and precipitation seasonality [@wagner_climate_2016], cloud cover [@taylor_temperature_2017], solar radiation [@beer_terrestrial_2010; @fyllas_solar_2017], and potential evapotranspiration [@kerkhoff_plant_2005]; however, these are not typically assessed in global-scale analyses of annual forest C flux.
 
@@
 This parallels the observation that C allocation across multiple C fluxes varies little with respect to climate along a steep tropical elevational gradient [@malhi_variation_2017; but see @moser_elevation_2011],
-and is not surprising given that trees face universal constraints in carbon allocation (**REFS**). 
+and is not surprising given that carbon allocation within forest ecosystems is relatively constrained [@enquist_global_2002; @litton_carbon_2007; @malhi_allocation_2011]. 
 We find no trend in the allocation of $GPP$ between production and respiration across latitude or climate ($NPP$:$R_{auto}$; Fig.
@@
 
-Despite the broad consistency of climate responses across C fluxes, climate explains lower proportions of variability among some of the subsidiary C fluxes (*e.g.*, $ANPP_{stem}$, $BNPP$  $BNPP_{fine.root)$; Fig.
+Despite the broad consistency of climate responses across C fluxes, climate explains lower proportions of variability among some of the subsidiary C fluxes (*e.g.*, $ANPP_{stem}$, $BNPP$  $BNPP_{fine.root}$; Fig.
 2; Tables S2, S6).
@@
 Second, differences among variables in the proportion of variation explained by climate may be attributable to more direct climatic control over $GPP$ than subsidiary fluxes. 
-That is, subsidiary fluxes may be shaped by climate both indirectly through its influence on $GPP$ and respiration and directly through any climatic influence on C allocation, as well as many other local- and regional-scale factors (**REFS**).
+That is, subsidiary fluxes may be shaped by climate both indirectly through its influence on $GPP$ and respiration and directly through any climatic influence on C allocation, as well as many other local- and regional-scale factors [e.g., @moser_elevation_2011] (**REFS-BECKY, know any?**).
 
 Temperature and its seasonality were the primary drivers of C flux on the global scale (Figs.
-2-4), consistent with a long legacy of research identifying temperature as a primary driver of forest ecosystem C cycling [e.g., @lieth_primary_1973; **REFS**].
+2-4), consistent with a long legacy of research identifying temperature as a primary driver of forest ecosystem C cycling [e.g., @lieth_primary_1973; @luyssaert_co_2007; @wei_forest_2010].
 We find little evidence of any non-linearity in temperature's influence on C flux.
@@
 4a).
-This result contrasts with both the idea that C flux responds exponentially to temperature (**REFS**) and with the idea that fluxes saturate with $MAT$ below approximately 25$^\circ$C MAT (**REFS**).
-It remains possible that flux declines above this threshold [@larjavaara_temperature_2012; @sullivan_sensitivity_2020], as is also consistent with tree-ring records indicating that tropical tree growth declines at high temperatures [e.g., @vlam_temperature_2014].
+This result contrasts with the idea that fluxes saturate with $MAT$ below approximately 25$^\circ$C $MAT$ [@luyssaert_co_2007; @huston_global_2009].
+It remains possible that flux declines above this threshold [@larjavaara_temperature_2012; @sullivan_long-term_2020], as is also consistent with tree-ring records indicating that tropical tree growth declines at high temperatures [e.g., @vlam_temperature_2014].
 However, these higher temperatures also tend to be associated with high PET and VPD, both of which are associated with reduced C fluxes (Figs. 4c-d, S4-S5).
@@
 Indeed, tree-ring studies from forests around the world indicate that tree growth rates - along with $ANPP_{stem}$ and possibly other ecosystem C fluxes - respond negatively to temperature [@sniderhan_growth_2016;@helcoski_growing_2019].
-Furthermore, in the tropics, climate change will push forests beyond any contemporary climate, and there are some indications that this could reduce C flux rates  [@sullivan_sensitivity_2020] (**REFS**).
+Furthermore, in the tropics, climate change will push forests beyond any contemporary climate, and there are some indications that this could reduce C flux rates  [@mau_temperate_2018; @sullivan_long-term_2020].
 Thus, while further research is required to understand the extent to which forest responses to climate change will track global gradients, and the time scale on which they will do so, understanding the fundamental climatic controls on annual C cycling in Earth's forests sets a firmer foundation for understanding forest C cycle responses to accelerating climate change.
@@
 ### Acknowledgements
-Scholarly Studies
-ForestGEO
-Compilation of the ForC database was originally funded by DOE
+We gratefully acknowledge all researchers who originally collected the data used in this analysis. 
+This study was funded by a Smithsonian Scholarly Studies grant to KJAT and HCML and by Smithsonian's Forest Global Earth Observatory (ForestGEO). 
+Original compilation of the ForC database was funded by **DOE grant to KAT**.

### 2020-08-06 manuscript/manuscript-draft.tex
working on refs... seem to have broken it :(

@@
 broad spatial scales
-\citep{luyssaert_co_2007, cleveland_relationships_2011, hursh_sensitivity_2017}.
+\citep{luyssaert_co_2007, cleveland_relationships_2011, wei_forest_2010}.
 However, there is little consensus as to the shapes of these
@@
 studies have detected no deceleration or decline in \(GPP\)
-\citep{luyssaert_co_2007} or \(NPP\) \citep{schuur_productivity_2003}
-with increasing MAT. Some have suggested that relationship between C
-flux and temperature is exponential in nature \citep{allen_linking_2005}
-(\textbf{REFS}). In contrast, others have found evidence of saturation
-or decline of C flux in the warmest climates; \citet{luyssaert_co_2007}
-found \(NPP\) saturating at around 10\(^\circ\)C MAT;
-\citet{larjavaara_temperature_2012} found that increases in \(GPP\)
-saturate at approximately 25\(^\circ\)C MAT, and
-\citet{sullivan_sensitivity_2020} found that, within the tropics,
-\(ANPP_{stem}\) decreases at the highest maximum temperatures. Various
-measures of primary productivity saturate at high levels of MAP, though
-the saturation points identified vary from 1500mm
-\citep{luyssaert_co_2007} up to 2,445 mm MAP
+\citep{luyssaert_co_2007}, \(NPP\) \citep{schuur_productivity_2003}, or
+root respiration \citep[\(R_{root}\);][]{wei_forest_2010} with
+increasing \(MAT\). In contrast, others have found evidence of
+saturation or decline of C flux in the warmest climates;
+\citet{luyssaert_co_2007} found \(NPP\) saturating at around
+10\(^\circ\)C MAT; \citet{larjavaara_temperature_2012} found that
+increases in \(GPP\) saturate at approximately 25\(^\circ\)C MAT, and
+\citet{sullivan_long-term_2020} found that, within the tropics,
+\(ANPP_{stem}\) decreases at the highest maximum temperatures. C fluxes
+generally saturate at high levels of MAP, though the saturation points
+identified vary from \(MAP\) of \textasciitilde1000 mm for \(R_{root}\)
+\citep{wei_forest_2010} up to 2,445 mm for \(NPP\)
 \citep{schuur_productivity_2003}. Interactions between MAT and MAP are
@@
 positive effect in warmer climates \citep{taylor_temperature_2017}.
-\textbf{Such complicated dynamics play out with belowground fluxes as
-well, with typically dominant soil temperature mediated by moisture and
-C supply \citep{hursh_sensitivity_2017, xu_contribution_2016}.
-\url{https://github.com/forc-db/Global_Productivity/issues/111}} There
-is also evidence that C fluxes also respond to climate variables such as
-temperature and precipitation seasonality (\textbf{REFS}), cloud cover
+There is also evidence that C fluxes also respond to climate variables
+such as temperature and precipitation seasonality
+\citep{wagner_climate_2016}, cloud cover
 \citep{taylor_temperature_2017}, solar radiation
@@
 \citetext{\citealp{malhi_variation_2017}; \citealp[but
-see][]{moser_elevation_2011}}, and is not surprising given that trees
-face universal constraints in carbon allocation (\textbf{REFS}). We find
-no trend in the allocation of \(GPP\) between production and respiration
-across latitude or climate (\(NPP\):\(R_{auto}\); Fig. S3), refuting the
-idea that tropical forests have anomalously low \(CUE\)
+see][]{moser_elevation_2011}}, and is not surprising given that carbon
+allocation within forest ecosystems is relatively constrained
+\citep{enquist_global_2002, litton_carbon_2007, malhi_allocation_2011}.
+We find no trend in the allocation of \(GPP\) between production and
+respiration across latitude or climate (\(NPP\):\(R_{auto}\); Fig. S3),
+refuting the idea that tropical forests have anomalously low \(CUE\)
 \citep{de_lucia_forest_2007, malhi_productivity_2012, anderson-teixeira_carbon_2016}.
@@
 subsidiary C fluxes (\emph{e.g.}, \(ANPP_{stem}\), \(BNPP\)
-\(BNPP_{fine.root)\); Fig. 2; Tables S2, S6). There are two,
+\(BNPP_{fine.root}\); Fig. 2; Tables S2, S6). There are two,
 non-exclusive, potential explanations for this. First, it may be that
@@
 influence on C allocation, as well as many other local- and
-regional-scale factors (\textbf{REFS}).
+regional-scale factors \citep[e.g.,][]{moser_elevation_2011}
+(\textbf{REFS-BECKY, know any?}).
 
@@
 identifying temperature as a primary driver of forest ecosystem C
-cycling {[}e.g., \citet{lieth_primary_1973}; \textbf{REFS}{]}. We find
-little evidence of any non-linearity in temperature's influence on C
-flux. The relationship of all fluxes to \(MAT\) as an individual driver
-were best described by a linear function (Table S2) - with the exception
-of \(BNPP\), whose response to \(MAT\) was close to linear (Fig. 4a).
-This result contrasts with both the idea that C flux responds
-exponentially to temperature (\textbf{REFS}) and with the idea that
-fluxes saturate with \(MAT\) below approximately 25\(^\circ\)C MAT
-(\textbf{REFS}). It remains possible that flux declines above this
-threshold
-\citep{larjavaara_temperature_2012, sullivan_sensitivity_2020}, as is
-also consistent with tree-ring records indicating that tropical tree
-growth declines at high temperatures
-\citep[e.g.,][]{vlam_temperature_2014}. However, these higher
-temperatures also tend to be associated with high PET and VPD, both of
-which are associated with reduced C fluxes (Figs. 4c-d, S4-S5). Indeed,
-while temperature responses dominate at this global scale and with the
-climate space occupied by forests, the effects of temperature are
-moderated by moisture availability (Table 1; Figs 3-4). Specifically, C
-fluxes are reduced under relatively dry conditions (\emph{i.e.}, low
-\(MAP\); high \(VPD\)) and sometimes under very high precipitation
-(Figs. 3-4). The observed positive interaction between \(MAT\) and
-\(MAP\) for \(ANPP_{stem}\) on the global scale (Fig. 3) is consistent
-with an analysis showing a similar interaction for \(ANPP\) in tropical
-forests, also with a cross-over point at \textasciitilde20\(\circ\)C
-\citep{taylor_temperature_2017}.\\
+cycling
+\citep[e.g.,][]{lieth_primary_1973, luyssaert_co_2007, wei_forest_2010}.
+We find little evidence of any non-linearity in temperature's influence
+on C flux. The relationship of all fluxes to \(MAT\) as an individual
+driver were best described by a linear function (Table S2) - with the
+exception of \(BNPP\), whose response to \(MAT\) was close to linear
+(Fig. 4a). This result contrasts with the idea that fluxes saturate with
+\(MAT\) below approximately 25\(^\circ\)C \(MAT\)
+\citep{luyssaert_co_2007, huston_global_2009}. It remains possible that
+flux declines above this threshold
+\citep{larjavaara_temperature_2012, sullivan_long-term_2020}, as is also
+consistent with tree-ring records indicating that tropical tree growth
+declines at high temperatures \citep[e.g.,][]{vlam_temperature_2014}.
+However, these higher temperatures also tend to be associated with high
+PET and VPD, both of which are associated with reduced C fluxes (Figs.
+4c-d, S4-S5). Indeed, while temperature responses dominate at this
+global scale and with the climate space occupied by forests, the effects
+of temperature are moderated by moisture availability (Table 1; Figs
+3-4). Specifically, C fluxes are reduced under relatively dry conditions
+(\emph{i.e.}, low \(MAP\); high \(VPD\)) and sometimes under very high
+precipitation (Figs. 3-4). The observed positive interaction between
+\(MAT\) and \(MAP\) for \(ANPP_{stem}\) on the global scale (Fig. 3) is
+consistent with an analysis showing a similar interaction for \(ANPP\)
+in tropical forests, also with a cross-over point at
+\textasciitilde20\(\circ\)C \citep{taylor_temperature_2017}.\\
 However, we detect no such interaction for \(ANPP\) or most other C
@@
 climate, and there are some indications that this could reduce C flux
-rates \citep{sullivan_sensitivity_2020} (\textbf{REFS}). Thus, while
+rates \citep{mau_temperate_2018, sullivan_long-term_2020}. Thus, while
 further research is required to understand the extent to which forest
@@
 
-Scholarly Studies ForestGEO Compilation of the ForC database was
-originally funded by DOE
+We gratefully acknowledge all researchers who originally collected the
+data used in this analysis. This study was funded by a Smithsonian
+Scholarly Studies grant to KJAT and HCML and by Smithsonian's Forest
+Global Earth Observatory (ForestGEO). Original compilation of the ForC
+database was funded by \textbf{DOE grant to KAT}.
 

### 2020-08-06 manuscript/journal_submissions/2020_8_GCB/submission_notes.Rmd
add submission notes

---
title: "GCB submission notes"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# submission questions

## What is the scientific question you are addressing?
## What is/are the key finding(s) that answers this question?
##    Why is this work important and timely?
##    Does your paper fall within the scope of GCB; what biological AND global change aspects does it address?
##    What are the three most recently published papers that are relevant to this question? This information will assist the Editors in selecting reviewers.

### 2020-08-07 manuscript/manuscript draft.Rmd
temporary fix of refs problem
@sullivan_long-term_2020 was breaking it-- not sure why. Deleted for now
updates around submission guidelines, Kerkoff/Michaeletz hypothesis
#79

@@
 
-**Running head:**
+**Running head:** Global patterns of forest carbon fluxes
 
@@
 
-**Author Affiliations:**
+**Institutional Affiliations:**
 
@@
 \newpage
-###Abstract
+### Abstract
 Carbon (C) fixation, allocation, and metabolism by trees set the basis for energy and material flows in forest ecosystems and define their interactions with Earth's changing climate. 
@@
 In a period of accelerating climatic change, this improved understanding of the fundamental climatic controls on forest C cycling sets a foundation for understanding patterns of change.
+
+**Keywords:**  carbon fluxes; climate; forest; productivity; respiration; latitude
 
@@
 Some studies have suggested that the apparently strong relationship between MAT and C fluxes is actually a consequence of the correlation between MAT and growing season length [@kerkhoff_plant_2005; @michaletz_convergence_2014; @michaletz_drivers_2018].
-Kerkhoff et al.
-[-@kerkhoff_plant_2005] and Michaletz et al.
-[-@michaletz_convergence_2014] found no significant relationship between growing season temperature and net primary productivity (loosely defined to include a mix of **VARIABLES**) standardized to growing season length.
-While this suggests that the influence of temperature may be limited to determining the length of the frost-free growing season, analysis with clearly defined variables and standardized forest ages would be necessary to test the the veracity and generality of this hypothesis.
+@kerkhoff_plant_2005 and @michaletz_convergence_2014 found no significant relationship between growing season temperature and $ANPP$ or $NPP$ standardized to growing season length [but see @chu_does_2016].
+While this suggests that the influence of temperature may be limited to determining the length of the frost-free growing season, analysis with carefully standardized variables and forest ages would be necessary to test the the veracity and generality of this hypothesis.
 
@@
 
-**For consistency with previous studies (Table 1, *H5*),** length of the growing season was estimated to the nearest month, where growing season months were defined as months with mean minimum temperature > 0.5$^\circ$C.
+Following the previous studies whose hypothesis we were evaluating [@kerkhoff_plant_2005; @michaletz_convergence_2014], length of the growing season was estimated to the nearest month, where growing season months were defined as months with mean minimum temperature > 0.5$^\circ$C.
 We experimented with a definition of growing season months including a moisture index, defined as (MAT - PET)/PET, > -0.95 [@kerkhoff_plant_2005; see also @michaletz_convergence_2014]. 
-However, we found that including a moisture index had minimal effect on the estimates of growing season length, and so chose to exclude it.
+However, we found that including a moisture index had minimal effect on the estimates of growing season length for the sites included here, and so chose to exclude it. 
 Monthly data for PET, precipitation, and temperature from  CRU v 4.03 [@harris_updated_2014] and solar radiation from WorldClim2 [@fick_worldclim_2017] were used to calculate mean monthly PET, precipitation, temperature and solar radiation during the growing season.
@@
 
-We next examined the relationships of C fluxes to climate variables (Q2-Q4; Table 1). 
+We next examined the relationships of C fluxes to climate variables (*Q2-Q4*; Table 1). 
 We tested first-order linear, second-order polynomial, and logarithmic fits for each climate variable. 
@@
 
-To test whether and how C flux varied with climate when standardised by growing season length (Q5), we first standardized all annual C fluxes by dividing by growing season length (as defined above).
+To test whether and how C flux varied with climate when standardised by growing season length (*Q5*), we first standardized all annual C fluxes by dividing by growing season length (as defined above).
 We then derived four variables to describe growing season climate, specifically growing season temperature, precipitation, solar radiation, and PET (Table S1).
@@
 
-###Results
+### Results
 
@@
 
-###Discussion
+### Discussion
 
@@
 Water availability is also influential, but of secondary importance across the climate space occupied by forests (Figs. 3-4). 
-Contrary to prior suggestions that the majority of variation in C cycling is driven by the length of the growing season (**REFS**), we find modest explanatory power of growing season length and small but sometimes significant influence of climate within the growing season (Figs.
+Contrary to prior suggestions that the majority of variation in C cycling is driven primarily by the length of the growing season [@kerkhoff_plant_2005; @enquist_adaptive_2007; @michaletz_convergence_2014], we find modest explanatory power of growing season length and small but sometimes significant influence of climate within the growing season (Figs.
 4f,S6-S9).
@@
 We conclude that while correcting for growing season length takes analyses a step closer to mechanistic linkage of instantaneous C flux rates to environmental conditions, it remains crude relative to the timescales on which climate affects plant metabolism, and does not advance statistical predictive power.
-Mechanistic accounting for climatic effects on global forest carbon flux patterns instead requires models representing physiologically meaningful timescales [e.g., @medvigy_mechanistic_2009; @longo_biophysics_2019].
+Rather, mechanistic accounting for climatic effects on global forest carbon flux patterns requires models representing physiologically meaningful timescales [e.g., @medvigy_mechanistic_2009; @longo_biophysics_2019].
 
@@
 Original compilation of the ForC database was funded by **DOE grant to KAT**.
+
+### References

### 2020-08-07 manuscript/appendix.Rmd
general clean-up

@@
 ```{r setup, include = FALSE}
-knitr::opts_knit$set(root.dir = 'C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/')
+#knitr::opts_knit$set(root.dir = 'C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/')
 # knitr::opts_knit$set(root.dir = 'C:/Users/gyrcbm/Dropbox/Global_Productivity/manuscript/')
- # knitr::opts_knit$set(root.dir = '/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/Global_Productivity/manuscript/')
+  knitr::opts_knit$set(root.dir = '/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/Global_Productivity/manuscript/')
 ```

### 2020-08-07 manuscript/manuscript draft.Rmd
general clean-up

@@
                   \captionsetup[figure]{font=footnotesize}
-                  \usepackage{setspace}\onehalfspacing
+                  \usepackage{setspace}\doublespacing
                   \usepackage{lineno}\linenumbers
@@
 
-2. *Becky- current*
+2. **Becky- current**
 
@@
 
-**Keywords:**
-
-**Paper type:** Primary Research Article
-
 \newpage
@@
 However, we lack a cohesive synthesis on how forest carbon fluxes vary globally with respect to climate and one another. 
-Here, we draw upon 1319 records from the Global Forest Carbon Database (ForC), representing all major forest types and the nine most significant autotrophic carbon fluxes, to comprehensively explore how C cycling in mature, undisturbed forests varies with latitude and climate on a global scale.
+Here, we draw upon 1,319 records from the Global Forest Carbon Database (ForC), representing all major forest types and the nine most significant autotrophic carbon fluxes, to comprehensively explore how C cycling in mature, undisturbed forests varies with latitude and climate on a global scale.
 We show that, across all flux variables analyzed, C cycling decreases continuously with absolute latitude -- a finding that confirms multiple previous studies but contradicts the idea that net primary productivity ($NPP$) of temperate forests rivals that of tropical forests.
@@
 
-Carbon (C) cycling in Earth's forests provides the energetic basis for sustaining the majority of Earth's terrestrial biodiversity and many human populations [@millennium_ecosystem_assessment_ecosystems_2005], while strongly influencing atmospheric carbon dioxide (CO~2~) and climate [@bonan_forests_2008].
+Carbon (C) cycling in Earth's forests provides the energetic basis for sustaining the majority of Earth's terrestrial biodiversity and many human populations [Millenium Ecosystem @millennium_ecosystem_assessment_ecosystems_2005], while strongly influencing atmospheric carbon dioxide (CO~2~) and climate [@bonan_forests_2008].
 Forests' autotrophic carbon fluxes--that is, carbon fixation, allocation, and metabolism by trees and other primary producers--sets the energy ultimately available to heterotrophic organisms (including microbes), in turn influencing their abundance [@zak_plant_1994; @niedzialkowska_species_2010] and possibly diversity [@waide_relationship_1999; @chu_direct_2018].
-They are linked to cycling of energy, water, and nutrients, and, critically, influence all C stocks and define forest interactions with Earth's changing climate. 
+They are linked to cycling of energy, water, and nutrients and, critically, influence all C stocks and define forest interactions with Earth's changing climate. 
 Each year, over 69 Gt of C cycle through Earth's forests [@badgley_terrestrial_2019]-- a flux more than seven times greater that of recent anthropogenic fossil fuel emissions [9.5 Gt C yr^-1^; @friedlingstein_global_2019].
 As atmospheric CO~2~ continues to rise, driving climate change, forests will play a critical role in shaping the future of Earth's climate [@cavaleri_urgent_2015; @rogelj_mitigation_2018]. 
-However, our understanding of the climate dependence of forest C cycling on a global scale has been limited by analyses typically considering only one or a few variables at a time, insufficient parsing of related variables, and the mixing of data from forests that vary in stand age, disturbance history, and/or management status, all of which affect C cycling [@litton_carbon_2007; @gillman_latitude_2015; @simova_enigma_2017].
+However, our understanding of the climate dependence of forest C cycling on a global scale has been limited by analyses typically considering only one or a few variables at a time, insufficient parsing of related variables, and the mixing of data from forests that vary in stand age, disturbance history, and management status, all of which affect C cycling [@litton_carbon_2007; @gillman_latitude_2015; @simova_enigma_2017].
 
-Forest C fluxes decrease with latitude [e.g., @luyssaert_co_2007; @piao_forest_2010; @gillman_latitude_2015; @li_mapping_2019], but studies have differed in their conclusions regarding the shape of this relationship - quite possibly because of lack of standardization with respect to methodology and stand history.
+Forest C fluxes decrease with latitude [e.g., @luyssaert_co_2007; @piao_forest_2010; @gillman_latitude_2015; @li_mapping_2019]. 
+However, studies have differed in their conclusions regarding the shape of this relationship -- quite possibly because of lack of standardization with respect to methodology and stand history.
 For instance, studies agree that gross primary productivity ($GPP$) increases continuously with decreasing latitude and is indisputably highest in tropical forests [@luyssaert_co_2007; @beer_terrestrial_2010; @jung_global_2011; @badgley_terrestrial_2019; @li_mapping_2019].
-In contrast, some studies have suggested that net primary productivity ($NPP$), or its aboveground portion ($ANPP$), exhibits a less distinct increase from temperate to tropical forests [@luyssaert_co_2007]--or even a decrease [@huston_global_2009, but see @gillman_latitude_2015].
+In contrast, some studies have suggested that net primary productivity ($NPP$), or its aboveground portion ($ANPP$), exhibits a less distinct increase from temperate to tropical forests [@luyssaert_co_2007] -- or even a decrease [@huston_global_2009, but see @gillman_latitude_2015].
 A shallower increase in $NPP$ than in $GPP$ with decreasing latitude would align with the suggestion that tropical forests tend to have low carbon use efficiency [$CUE$= $NPP$/$GPP$; @de_lucia_forest_2007; @malhi_productivity_2012; @anderson-teixeira_carbon_2016].
@@
 However, there is little consensus as to the shapes of these relationships or the best predictor variables.
-The majority of studies have focused on exploring the relationships of C fluxes to mean annual temperature (MAT) and mean annual precipitation (MAP), as the most commonly reported site-level climate variables.
+The majority of studies have focused on exploring the relationships of C fluxes to mean annual temperature ($MAT$) and mean annual precipitation ($MAP$), as the most commonly reported site-level climate variables.
 C fluxes increase strongly with $MAT$ on the global scale, but whether they saturate or potentially decrease at higher temperatures remains disputed. 
 Some studies have detected no deceleration or decline in $GPP$ [@luyssaert_co_2007], $NPP$ [@schuur_productivity_2003], or root respiration [$R_{root}$; @wei_forest_2010] with increasing $MAT$.
-In contrast, others have found evidence of saturation or decline of C flux in the warmest climates; @luyssaert_co_2007 found $NPP$ saturating at around 10$^\circ$C MAT; @larjavaara_temperature_2012 found that increases in $GPP$ saturate at approximately 25$^\circ$C MAT, and @sullivan_long-term_2020 found that, within the tropics, $ANPP_{stem}$ decreases at the highest maximum temperatures.
-C fluxes generally saturate at high levels of MAP, though the saturation points identified vary from $MAP$ of ~1000 mm for $R_{root}$ [@wei_forest_2010]  up to 2,445 mm for $NPP$ [@schuur_productivity_2003].
-Interactions between MAT and MAP are also possible; within the tropics, there is a positive interaction between MAT and MAP in shaping $ANPP$, such that high rainfall has a negative effect on productivity in cooler climates, compared to a positive effect in warmer climates [@taylor_temperature_2017].
+In contrast, others have found evidence of saturation or decline of C flux in the warmest climates; @luyssaert_co_2007 found $NPP$ saturating at around 10$^\circ$C MAT; @larjavaara_temperature_2012 found that increases in $GPP$ saturate at approximately 25$^\circ$C $MAT$, and @sullivan_long-term_2020 found that, within the tropics, $ANPP_{stem}$ decreases at the highest maximum temperatures.
+C fluxes generally saturate at high levels of $MAP$, though the saturation points identified vary from $MAP$ of ~1000 mm for $R_{root}$ [@wei_forest_2010]  up to 2,445 mm for $NPP$ [@schuur_productivity_2003].
+Interactions between $MAT$ and $MAP$ are also possible; within the tropics, there is a positive interaction between $MAT$ and $MAP$ in shaping $ANPP$, such that high rainfall has a negative effect on productivity in cooler climates, compared to a positive effect in warmer climates [@taylor_temperature_2017].
 There is also evidence that C fluxes also respond to climate variables such as temperature and precipitation seasonality [@wagner_climate_2016], cloud cover [@taylor_temperature_2017], solar radiation [@beer_terrestrial_2010; @fyllas_solar_2017], and potential evapotranspiration [@kerkhoff_plant_2005]; however, these are not typically assessed in global-scale analyses of annual forest C flux.
 
-As metrics of annual climate, MAT and MAP fail to capture variation in climate on an intra-annual scale, including temperature and precipitation seasonality and growing season length. 
-Some studies have suggested that the apparently strong relationship between MAT and C fluxes is actually a consequence of the correlation between MAT and growing season length [@kerkhoff_plant_2005; @michaletz_convergence_2014; @michaletz_drivers_2018].
+As metrics of annual climate, $MAT$ and $MAP$ fail to capture variation in climate on an intra-annual scale, including temperature and precipitation seasonality and growing season length. 
+Some studies have suggested that the apparently strong relationship between $MAT$ and C fluxes is actually a consequence of the correlation between $MAT$ and growing season length [@kerkhoff_plant_2005; @michaletz_convergence_2014; @michaletz_drivers_2018].
 @kerkhoff_plant_2005 and @michaletz_convergence_2014 found no significant relationship between growing season temperature and $ANPP$ or $NPP$ standardized to growing season length [but see @chu_does_2016].
@@
 The recent development of the Global Forest Carbon database (ForC), which synthesizes multiple variables and including records of stand history [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018], opens up the possibility for a standardized analysis of global scale variation in multiple C fluxes and the principle climatic drivers of these patterns.
-In order to approach this broad topic, we simplify the major gaps in our knowledge to five broad questions and corresponding hypotheses (Table 1).
-First, we ask how nine forest autotrophic carbon fluxes in ForC vary with latitude.
-We then test how these fluxes relate to MAT and MAP, and additionally how they respond to other, less well-studied, climate variables.
-Finally, we consider the relationship between C flux and seasonality, considering the role of seasonality in explaining variation in carbon fluxes, and the influence of climate on C flux standardized by growing season length.
+In order to approach this broad topic, we simplify the major gaps in our knowledge to five broad questions and corresponding predictions (Table 1).
+First, we ask how nine forest autotrophic carbon fluxes in ForC vary with latitude (*Q1*).
+We then test how these fluxes relate to $MAT$ and $MAP$ (*Q2*), and additionally how they respond to other, less well-studied, climate variables (*Q3*).
+Finally, we consider the relationship between C flux and seasonality, considering the role of seasonality in explaining variation in carbon fluxes (*Q4*), and the influence of climate on C flux standardized by growing season length (*Q5*).
 
@@
 
-This analysis focused on nine C flux variables included in the open-access ForC database (Table 2) [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018].
+This analysis focused on nine C flux variables included in the open-access ForC database [Table 2; @anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018]. 
 ForC contains records of field-based measurements of forest carbon stocks and annual fluxes, compiled from original publications and existing data compilations and databases. 
@@
 This version contained 29,730 records from 4,979 plots, representing 20 distinct ecozones across all forested biogeographic and climate zones.
+From this, we drew 1,319 records that met our criteria, as outlined below (Fig. 1).
 
-This analysis focused on mature forests with no known history of signficant disturbance or management. 
+This analysis focused on mature forests with no known history of significant disturbance or management. 
 There is evidence that stand age influences patterns of C flux and allocation in forest ecosystems, and can confound relationships between latitude and primary productivity [@de_lucia_forest_2007; @gillman_latitude_2015].
@@
 
+\newpage  
 *Climate data*
 
-ForC contains geographic coordinates associated with each measurement record and, when available, mean annual temperature (MAT) and mean annual precipitation (MAP) as reported in the primary literature [@anderson-teixeira_forc_2018].
-Based on the geographic co-ordinates for each site, data on twelve climate variables--including MAT, MAP, temperature and precipitation seasonality, annual temperature range, solar radiation, cloud cover, annual frost and wet days, potential evapotranspiration (PET), aridity (MAP/PET), and vapor pressure deficit (VPD)--were extracted from five open-access climate datasets: WorldClim [@hijmans_very_2005], WorldClim2 [@fick_worldclim_2017], the Climate Research Unit time-series dataset (CRU TS v4.03 [@harris_updated_2014], the Global Aridity Index and Potential Evapotranspiration Climate Database [@trabucco_global_2019], and TerraClimate [@abatzoglou_terraclimate_2018] (Table S1).
-From these data, we derived maximum VPD, defined as the VPD of the month with the largest deficit, and the number of water stress months, defined as the number of months annually where precipitation was lower than PET.
-Where site-level data was missing for MAT or MAP, we used values from the WorldClim dataset.
+ForC contains geographic coordinates associated with each measurement record and, when available, $MAT$ and $MAP$ as reported in the primary literature [@anderson-teixeira_forc_2018].
+Based on the geographic co-ordinates for each site, data on twelve climate variables--including $MAT$, $MAP$, temperature and precipitation seasonality, annual temperature range, solar radiation, cloud cover, annual frost and wet days, potential evapotranspiration ($PET$), aridity ($MAP/PET$), and vapor pressure deficit ($VPD$)--were extracted from five open-access climate datasets: WorldClim [@hijmans_very_2005], WorldClim2 [@fick_worldclim_2017], the Climate Research Unit time-series dataset (CRU TS v4.03 [@harris_updated_2014], the Global Aridity Index and Potential Evapotranspiration Climate Database [@trabucco_global_2019], and TerraClimate [@abatzoglou_terraclimate_2018] (Table S1).
+From these data, we derived maximum $VPD$, defined as the $VPD$ of the month with the largest deficit, and the number of water stress months, defined as the number of months annually where precipitation was lower than $PET$.
+Where site-level data was missing for $MAT$ or $MAP$, we used values from the WorldClim dataset.
 
 Following the previous studies whose hypothesis we were evaluating [@kerkhoff_plant_2005; @michaletz_convergence_2014], length of the growing season was estimated to the nearest month, where growing season months were defined as months with mean minimum temperature > 0.5$^\circ$C.
-We experimented with a definition of growing season months including a moisture index, defined as (MAT - PET)/PET, > -0.95 [@kerkhoff_plant_2005; see also @michaletz_convergence_2014]. 
+We experimented with a definition of growing season months including a moisture index, defined as $(MAT - PET)/PET > -0.95$ [@kerkhoff_plant_2005; see also @michaletz_convergence_2014]. 
 However, we found that including a moisture index had minimal effect on the estimates of growing season length for the sites included here, and so chose to exclude it. 
-Monthly data for PET, precipitation, and temperature from  CRU v 4.03 [@harris_updated_2014] and solar radiation from WorldClim2 [@fick_worldclim_2017] were used to calculate mean monthly PET, precipitation, temperature and solar radiation during the growing season.
+Monthly data for $PET$, precipitation, and temperature from  CRU v 4.03 [@harris_updated_2014] and solar radiation from WorldClim2 [@fick_worldclim_2017] were used to calculate mean monthly $PET$, precipitation, temperature and solar radiation during the growing season.
 Total growing season precipitation and solar radiation were also calculated.
@@
 The basic model for all analyses included a fixed effect of latitude or climate and a random effect of plot nested within geographic area. 
-Geographic areas--*i.e.*, spatially clustered sites--are defined within ForC using a hierarchical cluster analysis on the distance matrix of the sites and a cutoff of 25km [@anderson-teixeira_forc_2018].
-We experimented with inclusion of altitude as a fixed effect, but excluded it from the final models because it added very little explanatory power--that is, the difference in AIC ($\Delta$ AIC) relative to models excluding altitude was generally small (often $\Delta$ AIC<2).
-Effects were considered significant when inclusion of the fixed effect of interest resulted in p $\le$ 0.05 and $\Delta$ AIC $\ge$ 2.0 relative to a corresponding null model. 
+Geographic areas--*i.e.*, spatially clustered sites--were defined within ForC using a hierarchical cluster analysis on the distance matrix of the sites and a cutoff of 25km [@anderson-teixeira_forc_2018].
+We experimented with inclusion of altitude as a fixed effect, but excluded it from the final models because it added very little explanatory power--that is, the difference in AIC ($\Delta AIC$) relative to models excluding altitude was generally small (often $\Delta AIC$<2).
+Effects were considered significant when inclusion of the fixed effect of interest resulted in p $\le$ 0.05 and $\Delta AIC$ $\ge$ 2.0 relative to a corresponding null model. 
 All $R^2$ values presented here are marginal $R^2$ values, and refer to the proportion of variation explained by only the fixed effects. 
@@
 For brevity, we henceforth refer to first-order linear models as "linear" and second-order polynomial models as "polynomial".
-We selected as the best model that with the highest $\Delta$ AIC relative to a null model with no fixed term, with the qualification that a polynomial model was considered an improvement over a linear model only if it reduced the AIC value by 2.0 or more.
+We selected as the best model that with the highest $\Delta AIC$ relative to a null model with no fixed term, with the qualification that a polynomial model was considered an improvement over a linear model only if it reduced the AIC value by 2.0 or more.
 
@@
 The log of these ratios were regressed against latitude and climate variables, using the linear model specified above.
-Cook's distance analyses were carried out for each of the models, and extreme outliers removed,
+Cook's distance analyses were carried out for each of the models, and extreme outliers removed.
 
@@
 We tested first-order linear, second-order polynomial, and logarithmic fits for each climate variable. 
-Again, polynomial fits were considered superior to first-order linear fits only if inclusion of a second-order polynomial term resulted in $\Delta$ AIC $\ge$ 2.0 relative to a first-order linear model.
+Again, polynomial fits were considered superior to first-order linear fits only if inclusion of a second-order polynomial term resulted in $\Delta AIC$ $\ge$ 2.0 relative to a first-order linear model.
 We tested relationships of each C flux (Table 2) against each climate variable (Table S1).
@@
 
-Multivariate models were used to investigate the potential joint and interactive effects of MAT and MAP on carbon fluxes.
-An additive model including MAP in addition to MAT was accepted when $\Delta$ AIC >2 relative to a null including only MAT as a fixed effect.
-An interactive model including an MAT x MAP interaction was accepted when $\Delta$AIC >2 relative to a null including MAT and MAP as fixed effects.
+Multivariate models were used to investigate the potential joint and interactive effects of $MAT$ and $MAP$ on carbon fluxes.
+An additive model including $MAP$ in addition to $MAT$ was accepted when $\Delta AIC$ >2 relative to a null including only $MAT$ as a fixed effect.
+An interactive model including an $MAT$ x $MAP$ interaction was accepted when $\Delta$AIC >2 relative to a null including $MAT$ and $MAP$ as fixed effects.
 
@@
 
-**All analyses were conducted in R (**Version**). Code and data necessary to reproduce all results are archived on GitHub....**
+All analyses were conducted in R (**Version**).
+Code and data necessary to reproduce all results are available through GitHub (https://github.com/forc-db/Global_Productivity) and archived in Zenodo (DOI: TBD).
 
@@
 
-In total, we analyzed 1319 records from nine forest autotrophic C flux variables taken from forests that had experienced no major anthropogenic disturbances within the past 100 years.
-These records represented a total of 255 plots in 154 distinct geographic areas across all forested biogeographic and climate zones (Fig.
-1, Table 2).
+In total, we analyzed 1,319 records from nine forest autotrophic C flux variables taken from forests that had experienced no major anthropogenic disturbances within the past 100 years.
+These records represented a total of 255 plots in 154 distinct geographic areas across all forested biogeographic and climate zones (Figs.
+1, S1;  Table 2).
 
-*How does C flux vary with latitude?*
+*Q1. How does C flux vary with latitude?*
 
@@
 The C fluxes that were most poorly predicted by latitude were $BNPP_{fine.root}$ ($R^2$=0.17) and $ANPP_{stem}$ ($R^2$=0.18).
-The relationship with latitude was best fit by the first-order linear model, with the exception of NPP and $R_{root}$, for which a logarithmic model was a slightly - but not significantly - better fit.
+The relationship with latitude was best fit by the first-order linear model, with the exception of NPP and $R_{root}$, for which a logarithmic model was a slightly -- but not significantly -- better fit.
 
@@
 
-In general, smaller component fluxes summed approximately to larger fluxes across the latitudinal gradient (Fig.
+Smaller component fluxes summed approximately to larger fluxes across the latitudinal gradient (Fig.
 2).
@@
 S3).
-Of 12 relationships tested (3 ratios among C flux variables regressed against latitude, MAT, MAP and temperature seasonality), none were significant.
+Of 12 relationships tested (3 ratios among C flux variables regressed against latitude, $MAT$, $MAP$ and temperature seasonality), none were significant.
 
-*How does C flux relate to MAT and MAP?*
+*Q2. How does C flux relate to $MAT$ and MAP?*
 
-All fluxes increased with MAT (all p<0.05; Figs. 3-4, S4-S5, Table S2). 
+All fluxes increased with $MAT$ (all p<0.05; Figs. 3-4, S4-S5, Table S2). 
 For eight of the nine fluxes, this relationship was linear.
-For only one variable, $BNPP$, did a lognormal fit provide signficant improvement over a first-order linear relationship. 
-*As with latitude, MAT tended to explain more variation in the larger fluxes ($GPP$, $NPP$, $ANPP$, $R_{auto}$) and $ANPP_{foliage}$ (all $R^2$> 0.4) than in subsidiary and belowground fluxes ($ANPP_{stem}$, $R_{root}$, $BNPP_{fine.root}$; all $R^2$< 0.25).* **update this -- NB these values are correct**
+**For only one variable, $BNPP$, did a lognormal fit provide significant improvement over a first-order linear relationship.** 
+As with latitude, $MAT$ tended to explain more variation in the larger fluxes ($GPP$, $NPP$, $ANPP$, $R_{auto}$) and $ANPP_{foliage}$ (all $R^2$> 0.4) than in subsidiary and belowground fluxes ($ANPP_{stem}$, $R_{root}$, $BNPP_{fine.root}$; all $R^2$< 0.25).
 
-MAP was a significant (p<0.05) predictor of all fluxes (Figs. 4a, S4-S5; Table S2). 
-However, it explained little variation: with the exception of $R_{auto}$, MAP explained at most 25% of variation in C flux.
-All fluxes increased with MAP up to at least 2000 mm, above which responses were variable (Figs.
+$MAP$ was a significant (p<0.05) predictor of all fluxes (Figs. 4a, S4-S5; Table S2). 
+However, it explained little variation: with the exception of $R_{auto}$, $MAP$ explained at most 25% of variation in C flux.
+All fluxes increased with $MAP$ up to at least 2000 mm, above which responses were variable (Figs.
 4, S4-S5).
-There was a significant additive effect of MAT and MAP on $GPP$, $ANPP$ and $R_{auto}$ (Fig.
-3, Table S3), and a significant interactive effect between MAT and MAP for $NPP$ and $ANPP_{stem}$ (Fig.
+
+There was a significant additive effect of $MAT$ and $MAP$ on $GPP$, $ANPP$ and $R_{auto}$ (Fig.
+3, Table S3), and a significant interactive effect between $MAT$ and $MAP$ for $NPP$ and $ANPP_{stem}$ (Fig.
 3, Table S3).
 The interaction was negative for $NPP$ and positive for $ANPP_{stem}$.
-For $ANPP_{foliage}$, $BNPP$, $BNPP_{fine.root}$, and $R_{root}$, MAP did not have a signficant effect when accounting for MAT (Fig.
+For $ANPP_{foliage}$, $BNPP$, $BNPP_{fine.root}$, and $R_{root}$, $MAP$ did not have a signficant effect when accounting for $MAT$ (Fig.
 3, Table S3).
@@
 
-*How does C flux relate to other climate variables?*
+*Q3. How does C flux relate to other annual climate variables?*
 
-Our results indicated that annual forest C fluxes were most strongly explained by temperature-related climate variables at the global scale.
-In addition to MAT, several of its correlates (Fig.
-S2) were consistently identified as strong univariate predictors of C fluxes: temperature seasonality, annual temperature range, annual frost days, PET, and length of growing season (Figs.
-4, S4-S7).
-
-All C flux variables showed a significant relationship with potential evapotranspiration.
+All C flux variables showed a significant relationship with annual $PET$.
 The relationship was logarithmic for $ANPP_{foliage}$, $BNPP_{fine.root}$ and $R_{root}$, and polynomial for all other fluxes (Fig.
 4c, S4-5; Table S2).
-We found strong evidence for a saturation point or peak with PET: C fluxes tended to increase at values below 1000mm, before saturating between 1200 and 1700mm.
+We found strong evidence for a saturation point or peak with $PET$: C fluxes tended to increase at values below 1000mm, before saturating between 1200 and 1700mm.
 There was also evidence that some C fluxes begin to decrease at values above 1800mm PET.
 
-Vapour pressure deficit was a significant predictor of all C fluxes.
-$ANPP_{foliage}$, $BNPP_{fine.root}$ and $R_{root}$ showed a logarithmic relationship with vapour pressure deficit, but all other fluxes showed a polynomial relationship (Figs.
+Mean annual $VPD$ was a significant predictor of all C fluxes.
+$ANPP_{foliage}$, $BNPP_{fine.root}$ and $R_{root}$ showed a logarithmic relationship with $VPD$, but all other fluxes showed a polynomial relationship (Figs.
 4d, S4-5; Table S2).
-C fluxes initially increased with vapour pressure deficit, before saturating at around 0.8 kPa, after which point they began to decrease.
+C fluxes initially increased with $VPD$, before saturating at around 0.8 kPa, after which point they began to decrease.
 
-All fluxes, with the exception of $R_{root}$, showed a significant positive relationship with solar radiation (Figs.
-S4-S5, Table S2).
-Solar radiation explained a low proportion of variability in all C fluxes, explaining less than 30% of the variation in each flux
+All fluxes, with the exception of $R_{root}$, showed a significant positive relationship with solar radiation (Figs. S4-S5, Table S2). 
+Solar radiation explained a low proportion of variability (<30%) in all C fluxes.
 
-Annual wet days, cloud cover, aridity, and water stress months were poor or non-significant explainers of variation in C fluxes, explaining less than 20% of the variation in each of the carbon fluxes (Figs.
+Annual wet days, cloud cover, and aridity were poor or non-significant predictors of variation in C fluxes, explaining less than 20% of the variation in each of the carbon fluxes (Figs.
 S4-S5; Table S2).
@@
 
-*What is the role of seasonality in explaining C fluxes?*
+*Q4. What is the role of seasonality in explaining C fluxes?*
 
-Temperature seasonality was a strong predictor of annual C fluxes.
-All fluxes decrease with increasing seasonality, though the shape of this relationship varies (all p<0.05; Figs.
-4e, S6-7; Table S2).
+Variables describing temperature seasonality--temperature seasonality, annual temperature range, annual frost days, and length of growing season--were strongly correlated with both latitude and $MAT$ (all r$\ge$ 0.2; Fig.
+S2), and were consistently identified as strong univariate predictors of C fluxes (Figs.
+4, S4-S7).
+
+All fluxes decrease with increasing seasonality, though the shape of this relationship varies (all p<0.05; Figs. 4e, S6-7; Table S2). 
 Temperature seasonality was strongly correlated with annual temperature range, which was likewise a similarly strong predictor of C fluxes (Table S2). 
-C fluxes were highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower.
-**BBL: perhaps put this into an ecosystem context; what are these?
-Aseasonal subtropical places?**
+C fluxes were highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower (*i.e.*, in the tropics).
 
-In contrast, there was no significant effect of precipitation seasonality on C fluxes, and both maximum vapour pressure deficit, and water stress months were poor or non-significant predictors of variation in C fluxes (Figs.
-S6-S7; Table S2).
+In contrast, there was no significant effect of precipitation seasonality on C fluxes at this global scale.
+Both maximum vapour pressure deficit and water stress months were poor or non-significant predictors of variation in C fluxes (Figs. S6-S7; Table S2).
 
@@
 4e, S6-S7; Table S2).
-Length of growing season was a strong predictor of C fluxes, explaining 53% of variation in GPP, 38% of variation in NPP, and 34% of variation in ANPP (all p<0.05; Table S2), but it was a weaker predictor than MAT for all fluxes analysed (Table S4).
+Length of growing season was a strong predictor of C fluxes, explaining 53% of variation in GPP, 38% of variation in NPP, and 34% of variation in ANPP (all p<0.05; Table S2), but it was a weaker predictor than $MAT$ for all fluxes analysed (Table S4).
 
-*Within the growing season, how do C fluxes vary with climate?*
+*Q5. Within the growing season, how do C fluxes vary with climate?*
 
-When annual C fluxes were standardized by growing season length (in monthly increments), correlations with growing season climate were generally weak (Figs.
-S8-S9).
+When annual C fluxes were standardized by growing season length (in monthly increments), correlations with growing season climate were generally weak (Figs. S8-S9).
 $ANPP$ increased with growing season temperature ($R^2$ = 0.09, p<0.001) and precipitation ($R^2$ = 0.04, p<0.05). 
 Similarly, $ANPP_{foliage}$ increased slightly with growing season temperature ($R^2$ = 0.16, p<0.01) and precipitation ($R^2$ = 0.09, p<0.05). 
-Growing season solar radiation had a positive influence on $BNPP$ ($R^2$ = 0.17, p<0.001) and $BNPP_{fine.root}$ ($R^2$ = 0.13, p<0.01).
+Growing season solar radiation was positively correlated with on $BNPP$ ($R^2$ = 0.17, p<0.001) and $BNPP_{fine.root}$ ($R^2$ = 0.13, p<0.01). 
 Growing season PET had a positive influence on $GPP$ ($R^2$ = 0.15, p<0.01), $NPP$ ($R^2$ = 0.07, p<0.01), $BNPP$ ($R^2$ = 0.23, p<0.0001), $BNPP_{fine.root}$ ($R^2$ = 0.10, p<0.05), and $ANPP_{stem}$ ($R^2$ = 0.06, p<0.05).
@@
 Second, differences among variables in the proportion of variation explained by climate may be attributable to more direct climatic control over $GPP$ than subsidiary fluxes. 
-That is, subsidiary fluxes may be shaped by climate both indirectly through its influence on $GPP$ and respiration and directly through any climatic influence on C allocation, as well as many other local- and regional-scale factors [e.g., @moser_elevation_2011] (**REFS-BECKY, know any?**).
+That is, subsidiary fluxes may be shaped by climate both indirectly through its influence on $GPP$ and respiration and directly through any climatic influence on C allocation, as well as many other local- and regional-scale factors [e.g., @moser_elevation_2011].
 
@@
 However, we detect no such interaction for $ANPP$ or most other C fluxes, and we find a contrasting negative interaction for $NPP$ (Fig.
-3), suggesting that more data are required to sort out potential differences in the interactive effects of MAT and MAP on C fluxes in the tropics.
+3), suggesting that more data are required to sort out potential differences in the interactive effects of $MAT$ and $MAP$ on C fluxes in the tropics.
 

### 2020-08-07 manuscript/manuscript draft.Rmd
general cleanup: discussion & acknowledgements

@@
 Another previously observed pattern for which we find no support is a tendency for belowground C allocation to decrease with increasing temperature  [@moser_elevation_2011; @gill_belowground_2016]; rather, we observe no trends in allocation between $ANPP$ and $BNPP$ across latitudes.
-Failure to detect signficant tends in C allocation with respect to climate in this analysis does not imply that none exist; rather, it suggests that
+Failure to detect significant tends in C allocation with respect to climate in this analysis does not imply that none exist; rather, it suggests that, at this global scale, differences are subtle and/or that more careful methodological standardization is required to detect them.
 
-Despite the broad consistency of climate responses across C fluxes, climate explains lower proportions of variability among some of the subsidiary C fluxes (*e.g.*, $ANPP_{stem}$, $BNPP$  $BNPP_{fine.root}$; Fig.
+Despite the broad consistency of climate responses across C fluxes, climate explains lower proportions of variability among some of the subsidiary C fluxes (*e.g.*, $ANPP_{stem}$, $BNPP$,  $BNPP_{fine.root}$; Fig.
 2; Tables S2, S6).
@@
 We find little evidence of any non-linearity in temperature's influence on C flux.
-The relationship of all fluxes to $MAT$ as an individual driver were best described by a linear function (Table S2) - with the exception of $BNPP$, whose response to $MAT$ was close to linear (Fig.
+The relationship of all fluxes to $MAT$ as an individual driver were best described by a linear function (Table S2) -- with the exception of $BNPP$, whose response to $MAT$ was close to linear (Fig.
 4a).
@@
 It remains possible that flux declines above this threshold [@larjavaara_temperature_2012; @sullivan_long-term_2020], as is also consistent with tree-ring records indicating that tropical tree growth declines at high temperatures [e.g., @vlam_temperature_2014].
-However, these higher temperatures also tend to be associated with high PET and VPD, both of which are associated with reduced C fluxes (Figs. 4c-d, S4-S5).
-Indeed, while temperature responses dominate at this global scale and with the climate space occupied by forests, the effects of temperature are moderated by moisture availability (Table 1; Figs 3-4).
+However, these higher temperatures also tend to be associated with high $PET$ and $VPD$, both of which are associated with reduced C fluxes (Figs. 4c-d, S4-S5).
+
+Indeed, while temperature responses dominate at this global scale and within the climate space occupied by forests, the effects of temperature are moderated by moisture availability (Table 1; Figs 3-4).
 Specifically, C fluxes are reduced under relatively dry conditions (*i.e.*, low $MAP$; high $VPD$) and sometimes under very high precipitation (Figs. 3-4). 
 The observed positive interaction between $MAT$ and $MAP$ for $ANPP_{stem}$ on the global scale (Fig.
-3) is consistent with an analysis showing a similar interaction for $ANPP$ in tropical forests, also with a cross-over point at ~20$\circ$C [@taylor_temperature_2017].
+3) is consistent with an analysis showing a similar interaction for $ANPP$ in tropical forests, also with a cross-over point at ~20^$\circ$C [@taylor_temperature_2017].
 However, we detect no such interaction for $ANPP$ or most other C fluxes, and we find a contrasting negative interaction for $NPP$ (Fig.
@@
 
-Forest autotrophic C fluxes decline with temperature seasonality (Table 1, *H4*; Fig. 4e), and are minimal during cold- or dry- dormant seasons. 
-To account for this, a number of analyses seeking to characterize global-scale effects of climate on productivity have examined the relationship of C flux per month of the growing season with growing season climatic conditions [Table 1, *H5*; @kerkhoff_plant_2005; @anderson_temperature-dependence_2006; @enquist_adaptive_2007; @michaletz_convergence_2014].
+Forest C fluxes decline with temperature seasonality (Table 1, *Q4*; Fig. 4e), and are minimal during cold- or dry- dormant seasons. 
+To account for this, a number of analyses seeking to characterize global-scale effects of climate on productivity have examined the relationship of C flux per month of the growing season with growing season climatic conditions [Table 1, *Q5*; @kerkhoff_plant_2005; @anderson_temperature-dependence_2006; @enquist_adaptive_2007; @michaletz_convergence_2014].
 The sort of simple metric that has been used to define growing season at a global scale [@kerkhoff_plant_2005] is coarse with respect to temperature because it is calculated on a monthly timescale and problematic with respect to moisture because it doesn't capture temporal lags between precipitation and plant water use caused by storage in soil or snow.
@@
 
-Our analysis clarifies how forest autotrophic carbon fluxes vary with latitude and climate on a global scale, with some important implications for how forest carbon cycling relates to climate and, by extension, how it is likely to respond to climatic warming.
-To the extend that patterns across broad scale climatic gradients can foretell how ecosystem responses to climate change, our findings suggest that higher temperatures with similar moisture availability result in a generalized acceleration of forest C cycling (Figs.
+Our analysis clarifies how annual forest autotrophic C fluxes vary with latitude and climate on a global scale, with some important implications for how forest C cycling relates to climate and, by extension, how it is likely to respond to climatic warming.
+To the extent that patterns across broad scale climatic gradients can foretell how ecosystem responses to climate change, our findings suggest that higher temperatures with similar moisture availability would result in a generalized acceleration of forest C cycling (Figs.
 2-3).
@@
 Moreover, forest ecosystem responses to climatic changes outside the temperature range to which forest communities are adapted and acclimatized will not necessarily parallel responses across geographic gradients in climate.
-Indeed, tree-ring studies from forests around the world indicate that tree growth rates - along with $ANPP_{stem}$ and possibly other ecosystem C fluxes - respond negatively to temperature [@sniderhan_growth_2016;@helcoski_growing_2019].
+Indeed, tree-ring studies from forests around the world indicate that tree growth rates -- along with $ANPP_{stem}$ and possibly other ecosystem C fluxes -- respond negatively to temperature [@sniderhan_growth_2016;@helcoski_growing_2019].
 Furthermore, in the tropics, climate change will push forests beyond any contemporary climate, and there are some indications that this could reduce C flux rates  [@mau_temperate_2018; @sullivan_long-term_2020].
-Thus, while further research is required to understand the extent to which forest responses to climate change will track global gradients, and the time scale on which they will do so, understanding the fundamental climatic controls on annual C cycling in Earth's forests sets a firmer foundation for understanding forest C cycle responses to accelerating climate change.
+Further research is required to understand the extent to which forest responses to climate change will track the observed global gradients, and the time scale on which they will do so.
+In the meantime, understanding the fundamental climatic controls on annual C cycling in Earth's forests sets a firmer foundation for understanding forest C cycle responses to accelerating climate change.
 
 ### Acknowledgements
-We gratefully acknowledge all researchers who originally collected the data used in this analysis. 
+We gratefully acknowledge all authors of the original studies and data compilations included in this analysis, their funding agencies, and the various networks that support ground-based measurements of C fluxes.
+We also thank the numerous researchers who have contributed to the building of ForC.
 This study was funded by a Smithsonian Scholarly Studies grant to KJAT and HCML and by Smithsonian's Forest Global Earth Observatory (ForestGEO). 
-Original compilation of the ForC database was funded by **DOE grant to KAT**.
+Original compilation of the ForC database was funded by DOE grants DE-SC0008085 and DE-SC0010039 to KAT.
 

### 2020-08-07 manuscript/manuscript draft.Rmd
refer to table 1
#114

@@
 
-Our findings indicate that, among mature, undisturbed stands, forest C fluxes are unambiguously highest in the tropical regions, and the relationship with both latitude and $MAT$ is approximately linear (Figs.
+Our findings indicate that, among mature, undisturbed stands, forest C fluxes are unambiguously highest in the tropical regions, and the relationship with both latitude and $MAT$ is approximately linear (Table 1, *Q1*,*Q2*; Figs.
 2, 4).
@@
 
-Temperature and its seasonality were the primary drivers of C flux on the global scale (Figs.
+Temperature and its seasonality were the primary drivers of C flux on the global scale (Table 1, *Q2*,*Q4*; Figs.
 2-4), consistent with a long legacy of research identifying temperature as a primary driver of forest ecosystem C cycling [e.g., @lieth_primary_1973; @luyssaert_co_2007; @wei_forest_2010].
@@
 
-Indeed, while temperature responses dominate at this global scale and within the climate space occupied by forests, the effects of temperature are moderated by moisture availability (Table 1; Figs 3-4).
+Indeed, while temperature responses dominate at this global scale and within the climate space occupied by forests, the effects of temperature are moderated by moisture availability (Table 1, *Q2*,*Q3*; Figs 3-4).
 Specifically, C fluxes are reduced under relatively dry conditions (*i.e.*, low $MAP$; high $VPD$) and sometimes under very high precipitation (Figs. 3-4). 

### 2020-08-07 manuscript/manuscript draft.Rmd
final read through, minor corrections and edits

@@
 ```{r setup, include = FALSE}
-#knitr::opts_knit$set(root.dir = 'C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/')
-  knitr::opts_knit$set(root.dir = '/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/Global_Productivity/manuscript/')
+knitr::opts_knit$set(root.dir = 'C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/')
+  # knitr::opts_knit$set(root.dir = '/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/Global_Productivity/manuscript/')
 ```
@@
 
-2. **Becky- current**
+2. School of Geography, University of Leeds, Leeds, UK
 
@@
 Carbon (C) cycling in Earth's forests provides the energetic basis for sustaining the majority of Earth's terrestrial biodiversity and many human populations [Millenium Ecosystem @millennium_ecosystem_assessment_ecosystems_2005], while strongly influencing atmospheric carbon dioxide (CO~2~) and climate [@bonan_forests_2008].
-Forests' autotrophic carbon fluxes--that is, carbon fixation, allocation, and metabolism by trees and other primary producers--sets the energy ultimately available to heterotrophic organisms (including microbes), in turn influencing their abundance [@zak_plant_1994; @niedzialkowska_species_2010] and possibly diversity [@waide_relationship_1999; @chu_direct_2018].
+Forests' autotrophic carbon fluxes -- that is, carbon fixation, allocation, and metabolism by trees and other primary producers -- sets the energy ultimately available to heterotrophic organisms (including microbes), in turn influencing their abundance [@zak_plant_1994; @niedzialkowska_species_2010] and possibly diversity [@waide_relationship_1999; @chu_direct_2018].
 They are linked to cycling of energy, water, and nutrients and, critically, influence all C stocks and define forest interactions with Earth's changing climate. 
-Each year, over 69 Gt of C cycle through Earth's forests [@badgley_terrestrial_2019]-- a flux more than seven times greater that of recent anthropogenic fossil fuel emissions [9.5 Gt C yr^-1^; @friedlingstein_global_2019].
+Each year, over 69 Gt of C cycle through Earth's forests [@badgley_terrestrial_2019] -- a flux more than seven times greater than that of recent anthropogenic fossil fuel emissions [9.5 Gt C yr^-1^; @friedlingstein_global_2019].
 As atmospheric CO~2~ continues to rise, driving climate change, forests will play a critical role in shaping the future of Earth's climate [@cavaleri_urgent_2015; @rogelj_mitigation_2018]. 
@@
 
-The recent development of the Global Forest Carbon database (ForC), which synthesizes multiple variables and including records of stand history [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018], opens up the possibility for a standardized analysis of global scale variation in multiple C fluxes and the principle climatic drivers of these patterns.
+The recent development of the Global Forest Carbon database (ForC), which synthesizes multiple variables and includes records of stand history [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018], opens up the possibility for a standardized analysis of global scale variation in multiple C fluxes and the principle climatic drivers of these patterns.
 In order to approach this broad topic, we simplify the major gaps in our knowledge to five broad questions and corresponding predictions (Table 1).
@@
 To reduce any biasing effects of stand age, we included only stands of known age $\ge$ 100 years and those described by terms such as "mature", "intact", or "old-growth".
-Since management can alter observed patterns of C cycling [@simova_enigma_2017], sites were excluded from analysis if they were managed, defined as plots that were planted, managed as plantations, irrigated, fertilised or including the term "managed" in their site description.
+Since management can alter observed patterns of C cycling [@simova_enigma_2017], sites were excluded from analysis if they were managed, defined as plots that were planted, managed as plantations, irrigated, fertilised or included the term "managed" in their site description.
 Sites that had experienced significant disturbance within the past 100 years were also excluded.
@@
 ForC contains geographic coordinates associated with each measurement record and, when available, $MAT$ and $MAP$ as reported in the primary literature [@anderson-teixeira_forc_2018].
-Based on the geographic co-ordinates for each site, data on twelve climate variables--including $MAT$, $MAP$, temperature and precipitation seasonality, annual temperature range, solar radiation, cloud cover, annual frost and wet days, potential evapotranspiration ($PET$), aridity ($MAP/PET$), and vapor pressure deficit ($VPD$)--were extracted from five open-access climate datasets: WorldClim [@hijmans_very_2005], WorldClim2 [@fick_worldclim_2017], the Climate Research Unit time-series dataset (CRU TS v4.03 [@harris_updated_2014], the Global Aridity Index and Potential Evapotranspiration Climate Database [@trabucco_global_2019], and TerraClimate [@abatzoglou_terraclimate_2018] (Table S1).
+Based on the geographic co-ordinates for each site, data on twelve climate variables -- including $MAT$, $MAP$, temperature and precipitation seasonality, annual temperature range, solar radiation, cloud cover, annual frost and wet days, potential evapotranspiration ($PET$), aridity ($MAP/PET$), and vapor pressure deficit ($VPD$) -- were extracted from five open-access climate datasets: WorldClim [@hijmans_very_2005], WorldClim2 [@fick_worldclim_2017], the Climate Research Unit time-series dataset (CRU TS v4.03 [@harris_updated_2014], the Global Aridity Index and Potential Evapotranspiration Climate Database [@trabucco_global_2019], and TerraClimate [@abatzoglou_terraclimate_2018] (Table S1).
 From these data, we derived maximum $VPD$, defined as the $VPD$ of the month with the largest deficit, and the number of water stress months, defined as the number of months annually where precipitation was lower than $PET$.
@@
 Monthly data for $PET$, precipitation, and temperature from  CRU v 4.03 [@harris_updated_2014] and solar radiation from WorldClim2 [@fick_worldclim_2017] were used to calculate mean monthly $PET$, precipitation, temperature and solar radiation during the growing season.
-Total growing season precipitation and solar radiation were also calculated.
 
@@
 Geographic areas--*i.e.*, spatially clustered sites--were defined within ForC using a hierarchical cluster analysis on the distance matrix of the sites and a cutoff of 25km [@anderson-teixeira_forc_2018].
-We experimented with inclusion of altitude as a fixed effect, but excluded it from the final models because it added very little explanatory power--that is, the difference in AIC ($\Delta AIC$) relative to models excluding altitude was generally small (often $\Delta AIC$<2).
+We experimented with inclusion of altitude as a fixed effect, but excluded it from the final models because it added very little explanatory power -- that is, the difference in AIC ($\Delta AIC$) relative to models excluding altitude was generally small (often $\Delta AIC$<2).
 Effects were considered significant when inclusion of the fixed effect of interest resulted in p $\le$ 0.05 and $\Delta AIC$ $\ge$ 2.0 relative to a corresponding null model. 
@@
 For brevity, we henceforth refer to first-order linear models as "linear" and second-order polynomial models as "polynomial".
-We selected as the best model that with the highest $\Delta AIC$ relative to a null model with no fixed term, with the qualification that a polynomial model was considered an improvement over a linear model only if it reduced the AIC value by 2.0 or more.
+We selected as the best model that with the highest $\Delta$ AIC relative to a null model with no fixed term, with the qualification that a polynomial model was considered an improvement over a linear model only if it reduced the AIC value by 2.0 or more.
 
 To test whether trends in component fluxes across latitude sum to match those of larger fluxes, regression lines for smaller component fluxes were summed to generate new estimates of larger fluxes. 
-Because no fluxes were significantly better predicted by a logarithmic or polynomial fit than by a linear fit, we used linear fits for all fluxes.
+Because no fluxes were significantly better predicted by a logarithmic or polynomial fit than by a linear fit, we used linear fits for all fluxes in this analysis.
 We then determined whether these summed predictions fell within the 95% CI for the larger flux across the entire latitudinal range.
@@
 Variation in allocation to component carbon fluxes was explored for three groupings: (1) $GPP = NPP + R_{auto}$, (2) $NPP = ANPP + BNPP$, and (3) $ANPP = ANPP_{foliage} + ANPP_{stem}$.
-For each group, measurements taken at the same site and plot, an in the same year were grouped together.
+For each group, measurements taken at the same site and plot, and in the same year, were grouped together.
 For groups (1) and (2), where 2 of the 3 flux measurements were available for a given site, plot, and year, these measurements were used to calculate the third.
@@
 We tested first-order linear, second-order polynomial, and logarithmic fits for each climate variable. 
-Again, polynomial fits were considered superior to first-order linear fits only if inclusion of a second-order polynomial term resulted in $\Delta AIC$ $\ge$ 2.0 relative to a first-order linear model.
+Again, polynomial fits were considered superior to first-order linear fits only if inclusion of a second-order polynomial term resulted in $\Delta$AIC $\ge$ 2.0 relative to a first-order linear model.
 We tested relationships of each C flux (Table 2) against each climate variable (Table S1).
@@
 Multivariate models were used to investigate the potential joint and interactive effects of $MAT$ and $MAP$ on carbon fluxes.
-An additive model including $MAP$ in addition to $MAT$ was accepted when $\Delta AIC$ >2 relative to a null including only $MAT$ as a fixed effect.
+An additive model including $MAP$ in addition to $MAT$ was accepted when $\Delta$AIC >2 relative to a null including only $MAT$ as a fixed effect.
 An interactive model including an $MAT$ x $MAP$ interaction was accepted when $\Delta$AIC >2 relative to a null including $MAT$ and $MAP$ as fixed effects.
 
-To test whether and how C flux varied with climate when standardised by growing season length (*Q5*), we first standardized all annual C fluxes by dividing by growing season length (as defined above).
+To test whether and how C fluxes varied with climate when standardised by growing season length (*Q5*; Table 1), we first standardized all annual C fluxes by dividing by growing season length (as defined above).
 We then derived four variables to describe growing season climate, specifically growing season temperature, precipitation, solar radiation, and PET (Table S1).
@@
 
-All analyses were conducted in R (**Version**).
+All analyses were conducted in R v.3.5.1 [@r_core_team_r:_2018].
 Code and data necessary to reproduce all results are available through GitHub (https://github.com/forc-db/Global_Productivity) and archived in Zenodo (DOI: TBD).
@@
 
-*Q2. How does C flux relate to $MAT$ and MAP?*
+*Q2. How does C flux relate to $MAT$ and $MAP$?*
 
@@
 For eight of the nine fluxes, this relationship was linear.
-**For only one variable, $BNPP$, did a lognormal fit provide significant improvement over a first-order linear relationship.** 
+For only one variable, $BNPP$, did a lognormal fit provide an improvement over a first-order linear relationship, though this was not significant ($\Delta$AIC <2).
 As with latitude, $MAT$ tended to explain more variation in the larger fluxes ($GPP$, $NPP$, $ANPP$, $R_{auto}$) and $ANPP_{foliage}$ (all $R^2$> 0.4) than in subsidiary and belowground fluxes ($ANPP_{stem}$, $R_{root}$, $BNPP_{fine.root}$; all $R^2$< 0.25).
@@
 The interaction was negative for $NPP$ and positive for $ANPP_{stem}$.
-For $ANPP_{foliage}$, $BNPP$, $BNPP_{fine.root}$, and $R_{root}$, $MAP$ did not have a signficant effect when accounting for $MAT$ (Fig.
+For $ANPP_{foliage}$, $BNPP$, $BNPP_{fine.root}$, and $R_{root}$, $MAP$ did not have a significant effect when accounting for $MAT$ (Fig.
 3, Table S3).
@@
 For visualization purposes, data for each flux was rescaled with a mean of 0 and standard deviation of 1.
-Lines of best fit are plotted according to the best model selected during analysis (**see issue 47**).
+Lines of best fit are plotted according to the best model selected during analysis.
 All regressions are significant $(p<0.05)$.", fig.pos='H'}
@@
 
-Variables describing temperature seasonality--temperature seasonality, annual temperature range, annual frost days, and length of growing season--were strongly correlated with both latitude and $MAT$ (all r$\ge$ 0.2; Fig.
+Variables describing temperature seasonality -- temperature seasonality, annual temperature range, annual frost days, and length of growing season -- were strongly correlated with both latitude and $MAT$ (all r$\ge$ 0.2; Fig.
 S2), and were consistently identified as strong univariate predictors of C fluxes (Figs.
@@
 
-All fluxes decrease with increasing seasonality, though the shape of this relationship varies (all p<0.05; Figs. 4e, S6-7; Table S2). 
+All fluxes decrease with increasing temperature seasonality, though the shape of this relationship varies (all p<0.05; Figs. 4e, S6-7; Table S2). 
 Temperature seasonality was strongly correlated with annual temperature range, which was likewise a similarly strong predictor of C fluxes (Table S2). 
@@
 We show that, across all nine variables analyzed, annual C flux decreases continually with latitude (Fig.
-2)--a finding that confirms multiple previous studies but contradicts the idea that productivity of temperate forests rivals or even exceeds that of tropical forests [@luyssaert_co_2007; @huston_global_2009].
+2), a finding that confirms multiple previous studies but contradicts the idea that productivity of temperate forests rivals or even exceeds that of tropical forests [@luyssaert_co_2007; @huston_global_2009].
 At this global scale, C fluxes increase approximately in proportion to one another, with component fluxes summing appropriately to larger fluxes and no detectable differences in allocation across latitude or climates (Figs.
@@
 
-Temperature and its seasonality were the primary drivers of C flux on the global scale (Table 1, *Q2*,*Q4*; Figs.
+Temperature and its seasonality were the primary drivers of C fluxes on the global scale (Table 1, *Q2*,*Q4*; Figs.
 2-4), consistent with a long legacy of research identifying temperature as a primary driver of forest ecosystem C cycling [e.g., @lieth_primary_1973; @luyssaert_co_2007; @wei_forest_2010].
-We find little evidence of any non-linearity in temperature's influence on C flux.
+We find little evidence of any non-linearity in temperature's influence on C fluxes.
 The relationship of all fluxes to $MAT$ as an individual driver were best described by a linear function (Table S2) -- with the exception of $BNPP$, whose response to $MAT$ was close to linear (Fig.
@@
 This result contrasts with the idea that fluxes saturate with $MAT$ below approximately 25$^\circ$C $MAT$ [@luyssaert_co_2007; @huston_global_2009].
-It remains possible that flux declines above this threshold [@larjavaara_temperature_2012; @sullivan_long-term_2020], as is also consistent with tree-ring records indicating that tropical tree growth declines at high temperatures [e.g., @vlam_temperature_2014].
+It remains possible that fluxes decline above this threshold [@larjavaara_temperature_2012; @sullivan_long-term_2020], as is also consistent with tree-ring records indicating that tropical tree growth declines at high temperatures [e.g., @vlam_temperature_2014].
 However, these higher temperatures also tend to be associated with high $PET$ and $VPD$, both of which are associated with reduced C fluxes (Figs. 4c-d, S4-S5).

### 2020-08-07 manuscript/manuscript draft.Rmd
fixed all references problems
closes #115

@@
   pdf_document:
-    citation_package: natbib
     fig_caption: yes
-bibliography: [library.bib,  packages.bib]
-biblio-style: apalike
+bibliography: library.bib
 csl: global-change-biology.csl
@@
 ```{r setup, include = FALSE}
-knitr::opts_knit$set(root.dir = 'C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/')
+#knitr::opts_knit$set(root.dir = 'C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/')
   # knitr::opts_knit$set(root.dir = '/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/Global_Productivity/manuscript/')
@@
 
-Carbon (C) cycling in Earth's forests provides the energetic basis for sustaining the majority of Earth's terrestrial biodiversity and many human populations [Millenium Ecosystem @millennium_ecosystem_assessment_ecosystems_2005], while strongly influencing atmospheric carbon dioxide (CO~2~) and climate [@bonan_forests_2008].
+Carbon (C) cycling in Earth's forests provides the energetic basis for sustaining the majority of Earth's terrestrial biodiversity and many human populations [@millennium_ecosystem_assessment_ecosystems_2005], while strongly influencing atmospheric carbon dioxide (CO~2~) and climate [@bonan_forests_2008].
 Forests' autotrophic carbon fluxes -- that is, carbon fixation, allocation, and metabolism by trees and other primary producers -- sets the energy ultimately available to heterotrophic organisms (including microbes), in turn influencing their abundance [@zak_plant_1994; @niedzialkowska_species_2010] and possibly diversity [@waide_relationship_1999; @chu_direct_2018].
@@
 
-The effects of latitude and climate on C fluxes were analysed using mixed effects models using the package 'lme4' [@bates_fitting_2015] in R v.3.5.1 [@r_core_team_r:_2018]. 
+The effects of latitude and climate on C fluxes were analysed using mixed effects models using the package 'lme4' [@bates_fitting_2015] in R v.3.5.1 [@r_core_team_2018]. 
 The basic model for all analyses included a fixed effect of latitude or climate and a random effect of plot nested within geographic area. 
@@
 
-All analyses were conducted in R v.3.5.1 [@r_core_team_r:_2018].
+All analyses were conducted in R v.3.5.1 [@r_core_team_2018].
 Code and data necessary to reproduce all results are available through GitHub (https://github.com/forc-db/Global_Productivity) and archived in Zenodo (DOI: TBD).

### 2020-08-07 manuscript/appendix.Rmd
knit appendix to update corrections to climate variables table
minor cleanup of appendix

@@
 # knitr::opts_knit$set(root.dir = 'C:/Users/gyrcbm/Dropbox/Global_Productivity/manuscript/')
-  knitr::opts_knit$set(root.dir = '/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/Global_Productivity/manuscript/')
+  #knitr::opts_knit$set(root.dir = '/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/Global_Productivity/manuscript/')
 ```
@@
 kable(S3, format = "latex", booktabs = TRUE, caption = "Table S2.
-Model form, $\\Delta$AIC, and R$^{2}$ for each climate variables as a single fixed effect in models for each C flux.
-Model forms include first-order linear (Lin), second-order polynomial (Poly), and logarithmic (Log).", caption.short = "Table S2.
+Model form, $\\Delta$AIC, and R$^{2}$ for each climate variable as a single fixed effect in models for each C flux.", caption.short = "Table S2.
 Model form, $\\Delta$AIC and R$^{2}$ for each climate variables as a single fixed effect in models for each C flux", escape = F) %>%
@@
 ```
+Model forms tested include first-order linear (Lin), second-order polynomial (Poly), and logarithmic (Log). Shown are models with lowest $\Delta AIC$.
+
 \elandscape
@@
 S5 <- read.csv("tables_figures/mat_map.csv", stringsAsFactors = FALSE, check.names = FALSE)
-kable(S5, format = "latex", booktabs = TRUE, caption = "Table S3. Joint effects of MAT and MAP on forest C fluxes", escape = F)%>%
+kable(S5, format = "latex", booktabs = TRUE, caption = "Table S3. Signficance of relationships of forest C fluxes to MAT alone and in combination with MAP", escape = F)%>%
   kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 12) %>%
@@
 kable(S3, format = "latex", booktabs = TRUE, caption = "Table S5.
-Best models by carbon flux.
-Where $\\Delta$AIC values </= 2 all models with a $\\Delta$AIC value </= 2 are presented", caption.short = "Table S5.
+Best (lowest $\\Delta$AIC) single-climate variable models by C flux.", caption.short = "Table S5.
 Best models by carbon flux", escape = F) %>%
@@
 ```
+Table includes all models within $\Delta$AIC $\le$2.0 of the best model
 
@@
 S4 <- read.csv("tables_figures/pairwise_comparisons.csv", stringsAsFactors = FALSE, check.names = FALSE)
-kable(S4, format = "latex", booktabs = TRUE, caption = "Table S6. Pairwise comparisons between carbon fluxes", escape = F) %>%
+kable(S4, format = "latex", booktabs = TRUE, caption = "Table S6.
+Pairwise comparisons of correlations with climate variables between C fluxes, with all analyses including a common set of sites.", escape = F) %>%
   kable_styling(latex_options = c("scale_down", "hold_position")) %>%

### 2020-08-07 manuscript/appendix.tex
minor cleanup of appendix

@@
   \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \ifxetex
-    \usepackage{mathspec}
-  \else
-    \usepackage{fontspec}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
 \fi
-  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
-\fi
-% use upquote if available, for straight quotes in verbatim environments
+% Use upquote if available, for straight quotes in verbatim environments
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
-\usepackage{microtype}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
   \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
 }{}
-\usepackage[margin=1in]{geometry}
-\usepackage{hyperref}
-\hypersetup{unicode=true,
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\hypersetup{
   pdftitle={Supplementary Information: Global patterns of forest autotrophic carbon fluxes},
   pdfauthor={Rebecca Banbury Morgan; Valentine Herrmann; Norbert Kunert; Ben Bond-Lamberty; Helene C. Muller-Landau; Kristina J. Anderson-Teixeira},
-            pdfborder={0 0 0},
-            breaklinks=true}
-\urlstyle{same}  % don't use monospace font for urls
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+\urlstyle{same} % disable monospaced font for URLs
+\usepackage[margin=1in]{geometry}
 \usepackage{graphicx,grffile}
@@
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{0}
-% Redefines (sub)paragraphs to behave more like sections
-\ifx\paragraph\undefined\else
-\let\oldparagraph\paragraph
-\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-\let\oldsubparagraph\subparagraph
-\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
-
-%%% Use protect on footnotes to avoid problems with footnotes in titles
-\let\rmarkdownfootnote\footnote%
-\def\footnote{\protect\rmarkdownfootnote}
-
-%%% Change title format to be more compact
-\usepackage{titling}
-
-% Create subtitle command for use in maketitle
-\newcommand{\subtitle}[1]{
-  \posttitle{
-    \begin{center}\large#1\end{center}
-    }
-}
-
-\setlength{\droptitle}{-2em}
-
-  \title{Supplementary Information: Global patterns of forest autotrophic carbon
-fluxes}
-    \pretitle{\vspace{\droptitle}\centering\huge}
-  \posttitle{\par}
-    \author{Rebecca Banbury Morgan \\ Valentine Herrmann \\ Norbert Kunert \\ Ben Bond-Lamberty \\ Helene C. Muller-Landau \\ Kristina J. Anderson-Teixeira}
-    \preauthor{\centering\large\emph}
-  \postauthor{\par}
-    \date{}
-    \predate{}\postdate{}
-
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+\usepackage{float} \usepackage{caption} \captionsetup[table]{font=footnotesize} \captionsetup[figure]{font=footnotesize} \captionsetup[figure]{labelformat=empty} \captionsetup[table]{labelformat=empty} \usepackage{pdflscape} \newcommand{\blandscape}{\begin{landscape}} \newcommand{\elandscape}{\end{landscape}}
 \usepackage{booktabs}
@@
 
-\usepackage{float} \usepackage{caption} \captionsetup[table]{font=footnotesize} \captionsetup[figure]{font=footnotesize} \captionsetup[figure]{labelformat=empty} \captionsetup[table]{labelformat=empty} \usepackage{pdflscape} \newcommand{\blandscape}{\begin{landscape}} \newcommand{\elandscape}{\end{landscape}}
+\title{Supplementary Information: Global patterns of forest autotrophic carbon
+fluxes}
+\author{Rebecca Banbury Morgan \and Valentine Herrmann \and Norbert Kunert \and Ben Bond-Lamberty \and Helene C. Muller-Landau \and Kristina J. Anderson-Teixeira}
+\date{}
 
@@
 
-\blandscape
+\begin{landscape}
 
@@
 \resizebox{\linewidth}{!}{
-\fontsize{12}{14}\selectfont
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{6cm}>{\raggedright\arraybackslash}p{2cm}>{\raggedright\arraybackslash}p{7cm}>{\raggedright\arraybackslash}p{2cm}l}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{2.5cm}>{\raggedright\arraybackslash}p{5cm}>{\raggedright\arraybackslash}p{2cm}>{\raggedright\arraybackslash}p{7cm}>{\raggedright\arraybackslash}p{2cm}>{\raggedright\arraybackslash}p{5cm}}
 \toprule
-Climate variable & Units & Definition & Abbreviation & Source\\
+Abbreviation & Climate variable & Units & Definition & Time span & Source\\
 \midrule
-Mean annual temperature (MAT) & $^\circ$C &  & MAT & Primary literature; WorldClim$^{1}$\\
-Mean annual precipitation (MAP) & mm $yr^{-1}$ &  & MAP & Primary literature; WorldClim$^{1}$\\
-Temperature seasonality &  & Standard deviation of MAT *100 & T Seas & WorldClim$^{1}$\\
-Precipitation seasonality &  & Coefficient of variation of MAP & P Seas & WorldClim$^{1}$\\
-Annual temperature range & $^\circ$C & Maximum temperature of warmest month - minimum temperature of coldest month & ART & WorldClim$^{1}$\\
+MAT & Mean annual temperature & $^\circ$C & Annual mean temperature, from primary literature or WorldClim if not reported & 1970 - 2000 & Primary literature; WorldClim$^{1}$\\
+MAP & Mean annual precipitation & mm yr$^{-1}$ & Annual mean precipitation, from primary literature or WorldClim if not reported & 1970 - 2000 & Primary literature; WorldClim$^{1}$\\
+T Seas & Temperature seasonality & $^\circ$C & Standard deviation (variation) of monthly temperature averages & 1970 - 2000 & WorldClim$^{1}$\\
+P Seas & Precipitation seasonality & $\%$ & Coefficient of variation of mean monthly precipitation x 100 & 1970 - 2000 & WorldClim$^{1}$\\
+ATR & Annual temperature range & $^\circ$C & Maximum temperature of warmest month - minimum temperature of coldest month & 1970 - 2000 & WorldClim$^{1}$\\
 \addlinespace
-Solar radiation & kJ $m^{-2} yr^{-1}$ &  & Solar R & WorldClim2$^{2}$\\
-Cloud cover & percentage & Cloud percentage cover & Cloud & CRU time-series dataset v 4.03$^{3}$\\
-Annual frost days & days $yr^{-1}$ & Number of freeze days annually & AFD & CRU time-series dataset v 4.03$^{3}$\\
-Annual wet days & days $yr^{-1}$ & Number of days with precipitation >0.1 mm annually & AWD & CRU time-series dataset v 4.03$^{3}$\\
-Potential evapotranspiration (PET) & mm $yr^{-1}$ & Mean annual potential evapotranspiration & PET & Global Aridity Index and Potential Evapotranspiration Climate Database$^{4}$\\
+Solar R & Solar radiation & kJ m$^{-2}$ yr$^{-1}$ & Solar radiation & 1970 - 2000 & WorldClim2$^{2}$\\
+Cloud & Cloud cover & $\%$ & Cloud percentage cover & 1901 - 2014 & CRU time-series dataset v 4.03$^{3}$\\
+AFD & Annual frost days & days yr$^{-1}$ & Number of freeze days annually & 1901 - 2014 & CRU time-series dataset v 4.03$^{3}$\\
+AWD & Annual wet days & days yr$^{-1}$ & Number of days with precipitation >0.1 mm annually & 1901 - 2014 & CRU time-series dataset v 4.03$^{3}$\\
+PET & Potential evapotranspiration & mm yr$^{-1}$ & Mean annual potential evapotranspiration & 1950 - 2000 & Global Aridity Index and Potential Evapotranspiration Climate Database$^{4}$\\
 \addlinespace
-Aridity &  & MAP/mean annual PET & AI & Global Aridity Index and Potential Evapotranspiration Climate Database$^{4}$\\
-Vapour pressure deficit (VPD) & kPa &  & VPD & TerraClimate$^{5}$\\
-Maximum vapour pressure deficit (Max VPD) & kPa &  & Max VPD & Derived\\
-Water stress months & months $yr^{-1}$ & Number of months annually with MAP < PET & WSM & Derived\\
-Length of growing season & months $yr^{-1}$ & Number of months annually with mean minimum temperature > 0.5$^\circ$C & LGS & Derived\\
+AI & Aridity &  & MAP/mean annual PET & 1950 - 2000 & Global Aridity Index and Potential Evapotranspiration Climate Database$^{4}$\\
+VPD & Vapour pressure deficit & kPa & Mean monthly vapour pressure deficit & 1958 - 2015 & TerraClimate$^{5}$\\
+Max VPD & Maximum vapour pressure deficit & kPa & Maximum monthly vapour pressure deficit & 1958 - 2017 & Derived from TerraClimate data\\
+WSM & Water stress months & months yr$^{-1}$ & Number of months annually with MAP < PET & 1970 - 2000 & Derived from WorldClim data\\
+LGS & Length of growing season & months yr$^{-1}$ & Number of months annually with mean minimum temperature > 0.5$^\circ$C & 1901 - 2014 & Derived from CRU data\\
+\addlinespace
+gsT & Growing season temperature & $^\circ$C & Mean growing season temperature & 1901 - 2014 & Derived from CRU data\\
+gsP & Growing season precipitation & mm month$^{-1}$ & Mean monthly precipitation during growing season months & 1901 - 2014 & Derived from CRU data\\
+gsPET & Growing season PET & mm month$^{-1}$ & Mean monthly potential evapotranspiration during growing season months & 1901 - 2014 & Derived from CRU data\\
+gsR & Growing season solar radiation & mm month$^{-1}$ & Mean monthly solar radiation during growing season & 1901 - 2014 & Derived from WorldClim2 data\\
 \bottomrule
+\multicolumn{6}{l}{\textsuperscript{*} The WorldClim version used was the most recent available at the time of analysis}\\
+\multicolumn{6}{l}{\textsuperscript{1} Hijmans et al. (2005) \textsuperscript{2} Fick et al. (2017) \textsuperscript{3} Harris et al. (2017) \textsuperscript{4} Abatzoglou et al. (2018)}\\
 \end{tabular}}
 \end{table}
-
-\elandscape
+\end{landscape}
 \newpage
 
+\begin{landscape}
 \begin{table}[!h]
 
-\caption{\label{tab:unnamed-chunk-3}Table S2. Comparison of growing season length and mean annual temperature as predictors of FACF}
+\caption[Table S2.
+Model form, $\Delta$AIC and R$^{2}$ for each climate variables as a single fixed effect in models for each C flux]{\label{tab:unnamed-chunk-3}Table S2.
+Model form, $\Delta$AIC, and R$^{2}$ for each climate variable as a single fixed effect in models for each C flux.}
 \centering
 \resizebox{\linewidth}{!}{
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{5cm}>{\raggedleft\arraybackslash}p{2cm}>{\raggedleft\arraybackslash}p{7cm}>{\raggedleft\arraybackslash}p{7cm}}
+\begin{tabular}[t]{llrrlrrlrrlrrllllrrllllll}
 \toprule
-Fixed effect & AIC value & Delta AICc & Marginal R squared\\
+\multicolumn{1}{c}{ } & \multicolumn{3}{c}{Latitude} & \multicolumn{3}{c}{MAT} & \multicolumn{3}{c}{MAP} & \multicolumn{3}{c}{T Seas} & \multicolumn{3}{c}{P Seas} & \multicolumn{3}{c}{ATR} & \multicolumn{3}{c}{Solar R} & \multicolumn{3}{c}{AI} \\
+\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7} \cmidrule(l{3pt}r{3pt}){8-10} \cmidrule(l{3pt}r{3pt}){11-13} \cmidrule(l{3pt}r{3pt}){14-16} \cmidrule(l{3pt}r{3pt}){17-19} \cmidrule(l{3pt}r{3pt}){20-22} \cmidrule(l{3pt}r{3pt}){23-25}
+Carbon Flux & Model & R$^{2}$ & $\Delta$ AIC & Model & R$^{2}$ & $\Delta$ AIC & Model & R$^{2}$ & $\Delta$ AIC & Model & R$^{2}$ & $\Delta$ AIC & Model & R$^{2}$ & $\Delta$ AIC & Model & R$^{2}$ & $\Delta$ AIC & Model & R$^{2}$ & $\Delta$ AIC & Model & R-sq & dAIC\\
 \midrule
-\addlinespace[1em]
-\multicolumn{1}{l}{\textbf{GPP}}\\
-\hspace{1em}MAT & 126.42617 & 0.000000 & 0.6196780\\
-\hspace{1em}Growing season length & 140.80589 & 14.379717 & 0.5411935\\
-\hspace{1em}None & 178.96179 & 52.535617 & 0.0000000\\
-\addlinespace[1em]
-\hline
-\multicolumn{1}{l}{\textbf{NPP}}\\
-\hspace{1em}MAT & 174.88249 & 0.000000 & 0.5156614\\
-\hspace{1em}Growing season length & 191.53714 & 16.654650 & 0.4006999\\
-\hspace{1em}None & 216.16976 & 41.287265 & 0.0000000\\
-\addlinespace[1em]
-\hline
-\multicolumn{1}{l}{\textbf{ANPP}}\\
-\hspace{1em}MAT & 249.50512 & 0.000000 & 0.2925950\\
-\hspace{1em}Growing season length & 254.20763 & 4.702509 & 0.2612187\\
-\hspace{1em}None & 268.94008 & 19.434966 & 0.0000000\\
-\addlinespace[1em]
-\hline
-\multicolumn{1}{l}{\textbf{ANPP woody stem}}\\
-\hspace{1em}MAT & 235.95797 & 0.000000 & 0.1548800\\
-\hspace{1em}Growing season length & 237.28992 & 1.331943 & 0.1370243\\
-\hspace{1em}None & 243.13700 & 7.179027 & 0.0000000\\
-\addlinespace[1em]
-\hline
-\multicolumn{1}{l}{\textbf{ANPP foliage}}\\
-\hspace{1em}MAT & 484.87610 & 0.000000 & 0.4462629\\
-\hspace{1em}Growing season length & 520.96482 & 36.088722 & 0.3497750\\
-\hspace{1em}None & 560.34915 & 75.473049 & 0.0000000\\
-\addlinespace[1em]
-\hline
-\multicolumn{1}{l}{\textbf{BNPP root}}\\
-\hspace{1em}MAT & 184.54480 & 0.000000 & 0.5921282\\
-\hspace{1em}Growing season length & 204.92685 & 20.382054 & 0.4644116\\
-\hspace{1em}None & 237.46554 & 52.920743 & 0.0000000\\
-\addlinespace[1em]
-\hline
-\multicolumn{1}{l}{\textbf{BNPP fine root}}\\
-\hspace{1em}MAT & 540.19217 & 0.000000 & 0.2429540\\
-\hspace{1em}Growing season length & 566.36955 & 26.177388 & 0.1060029\\
-\hspace{1em}None & 578.65529 & 38.463119 & 0.0000000\\
-\addlinespace[1em]
-\hline
-\multicolumn{1}{l}{\textbf{Autotrophic respiration}}\\
-\hspace{1em}MAT & 45.25818 & 0.000000 & 0.6271133\\
-\hspace{1em}Growing season length & 50.35515 & 5.096972 & 0.5041004\\
-\hspace{1em}None & 56.16877 & 10.910597 & 0.0000000\\
-\addlinespace[1em]
-\hline
-\multicolumn{1}{l}{\textbf{Root respiration}}\\
-\hspace{1em}MAT & 133.53500 & 0.000000 & 0.2507631\\
-\hspace{1em}Growing season length & 135.92632 & 2.391311 & 0.1990489\\
-\hspace{1em}None & 141.78719 & 8.252190 & 0.0000000\\
-\bottomrule
-\end{tabular}}
-\end{table}
-
-\newpage
-
-\blandscape
-
-\begin{table}[!h]
-
-\caption{\label{tab:unnamed-chunk-4}Table S3. Model details and R2 values for all climate variables tested}
-\centering
-\resizebox{\linewidth}{!}{
-\begin{tabular}[t]{llrlrlllrlllrllll}
-\toprule
-\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Latitude} & \multicolumn{2}{c}{MAT} & \multicolumn{2}{c}{MAP} & \multicolumn{2}{c}{T Seas} & \multicolumn{2}{c}{P Seas} & \multicolumn{2}{c}{ATR} & \multicolumn{2}{c}{Solar R} & \multicolumn{2}{c}{AI} \\
-\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-7} \cmidrule(l{3pt}r{3pt}){8-9} \cmidrule(l{3pt}r{3pt}){10-11} \cmidrule(l{3pt}r{3pt}){12-13} \cmidrule(l{3pt}r{3pt}){14-15} \cmidrule(l{3pt}r{3pt}){16-17}
-Carbon flux & Model & R-squared & Model & R-squared & Model & R-squared & Model & R-squared & Model & R-squared & Model & R-squared & Model & R-squared & Model & R-squared\\
-\midrule
-GPP & Linear & 0.6387 & Linear & 0.6094 & Polynomial & 0.324 & Polynomial & 0.7076 & - & - & Linear & 0.6269 & Linear & 0.1554 & - & -\\
-NPP & Linear & 0.5108 & Linear & 0.4171 & Polynomial & 0.2138 & Polynomial & 0.4905 & - & - & Linear & 0.4442 & Linear & 0.09548 & Linear & 0.03795\\
-ANPP & Linear & 0.4351 & Linear & 0.4444 & Polynomial & 0.1625 & Polynomial & 0.4126 & - & - & Linear & 0.3331 & Linear & 0.1061 & Linear & 0.04851\\
-ANPP woody stem & Linear & 0.1773 & Linear & 0.2396 & - & - & Linear & 0.1416 & - & - & Linear & 0.1157 & Linear & 0.05048 & Linear & 0.06607\\
-ANPP foliage & Linear & 0.4999 & Linear & 0.5826 & Polynomial & 0.2509 & Linear & 0.4823 & - & - & Linear & 0.5033 & Linear & 0.172 & Linear & 0.1084\\
+GPP & Lin & 0.64 & 54.9 & Lin & 0.61 & 52.5 & Lin & 0.18 & 33.3 & Poly & 0.71 & 69.5 & - & - & - & Poly & 0.69 & 63.0 & Log & 0.16 & 8.9 & - & - & -\\
+NPP & Log & 0.50 & 44.3 & Lin & 0.42 & 41.5 & Poly & 0.21 & 16.7 & Log & 0.52 & 44.3 & - & - & - & Log & 0.49 & 42.3 & Poly & 0.16 & 12.5 & Lin & 0.04 & 2.8\\
+ANPP & Lin & 0.44 & 63.4 & Lin & 0.44 & 80.5 & Poly & 0.16 & 19.7 & Log & 0.41 & 58.7 & - & - & - & Log & 0.37 & 51.9 & Lin & 0.11 & 12.3 & Lin & 0.05 & 2.1\\
+ANPP stem & Lin & 0.18 & 22.2 & Lin & 0.24 & 38.5 & Log & 0.05 & 7.3 & Lin & 0.14 & 17.6 & Poly & 0.05 & 5 & Lin & 0.12 & 13.6 & Log & 0.06 & 6.8 & Lin & 0.07 & 4\\
+ANPP foliage & Lin & 0.50 & 37.7 & Lin & 0.58 & 52.9 & Poly & 0.25 & 13.3 & Lin & 0.48 & 34.1 & - & - & - & Lin & 0.50 & 36.1 & Log & 0.17 & 10.1 & Lin & 0.11 & 6.8\\
 \addlinespace
-BNPP root & Linear & 0.3373 & Linear & 0.2833 & Polynomial & 0.1452 & Linear & 0.3300 & - & - & Linear & 0.2904 & Linear & 0.2315 & - & -\\
-BNPP fine root & Linear & 0.1704 & Linear & 0.1477 & Linear & 0.08935 & Linear & 0.1721 & - & - & Linear & 0.1790 & Linear & 0.1393 & Polynomial & 0.06915\\
-Autotrophic respiration & Linear & 0.6534 & Linear & 0.5909 & Polynomial & 0.604 & Polynomial & 0.6246 & - & - & Linear & 0.4900 & Linear & 0.26 & Polynomial & 0.4804\\
-Root respiration & Linear & 0.2612 & Linear & 0.2418 & Linear & 0.1493 & Polynomial & 0.2151 & - & - & Linear & 0.1776 & - & - & Linear & 0.1567\\
+BNPP root & Lin & 0.34 & 22.9 & Log & 0.31 & 21.0 & Poly & 0.15 & 6.2 & Log & 0.36 & 26.6 & - & - & - & Log & 0.33 & 23.6 & Poly & 0.29 & 18.8 & - & - & -\\
+BNPP fine root & Lin & 0.17 & 8.0 & Lin & 0.15 & 7.2 & Log & 0.11 & 5.4 & Lin & 0.17 & 8.4 & - & - & - & Log & 0.19 & 10.9 & Log & 0.14 & 7.2 & Log & 0.06 & 2.4\\
+R auto & Lin & 0.65 & 13.1 & Lin & 0.59 & 10.9 & Poly & 0.60 & 8.6 & Log & 0.65 & 13.1 & - & - & - & Log & 0.60 & 11.5 & Log & 0.27 & 2.4 & Poly & 0.48 & 3.7\\
+R root & Log & 0.22 & 8.8 & Lin & 0.24 & 8.3 & Lin & 0.15 & 6.8 & Log & 0.24 & 9.5 & - & - & - & Log & 0.22 & 8.8 & - & - & - & Lin & 0.16 & 7.3\\
 \bottomrule
@@
 \resizebox{\linewidth}{!}{
-\begin{tabular}{llllrlllrlllllllr}
+\begin{tabular}{lllllrrllllrrlrrlllllllrr}
 \toprule
-\multicolumn{1}{c}{ } & \multicolumn{2}{c}{Cloud} & \multicolumn{2}{c}{AFD} & \multicolumn{2}{c}{AWD} & \multicolumn{2}{c}{PET} & \multicolumn{2}{c}{VPD} & \multicolumn{2}{c}{Max VPD} & \multicolumn{2}{c}{WSM} & \multicolumn{2}{c}{LGS} \\
-\cmidrule(l{3pt}r{3pt}){2-3} \cmidrule(l{3pt}r{3pt}){4-5} \cmidrule(l{3pt}r{3pt}){6-7} \cmidrule(l{3pt}r{3pt}){8-9} \cmidrule(l{3pt}r{3pt}){10-11} \cmidrule(l{3pt}r{3pt}){12-13} \cmidrule(l{3pt}r{3pt}){14-15} \cmidrule(l{3pt}r{3pt}){16-17}
-Carbon flux & Model & R-squared & Model & R-squared & Model & R-squared & Model & R-squared & Model & R-squared & Model & R-squared & Model & R-squared & Model & R-squared\\
+\multicolumn{1}{c}{ } & \multicolumn{3}{c}{Cloud} & \multicolumn{3}{c}{AFD} & \multicolumn{3}{c}{AWD} & \multicolumn{3}{c}{PET} & \multicolumn{3}{c}{VPD} & \multicolumn{3}{c}{Max VPD} & \multicolumn{3}{c}{WSM} & \multicolumn{3}{c}{LGS} \\
+\cmidrule(l{3pt}r{3pt}){2-4} \cmidrule(l{3pt}r{3pt}){5-7} \cmidrule(l{3pt}r{3pt}){8-10} \cmidrule(l{3pt}r{3pt}){11-13} \cmidrule(l{3pt}r{3pt}){14-16} \cmidrule(l{3pt}r{3pt}){17-19} \cmidrule(l{3pt}r{3pt}){20-22} \cmidrule(l{3pt}r{3pt}){23-25}
+Carbon Flux & Model & R$^{2}$ & $\Delta$ AIC & Model & R$^{2}$ & $\Delta$ AIC & Model & R$^{2}$ & $\Delta$ AIC & Model & R$^{2}$ & $\Delta$ AIC & Model & R$^{2}$ & $\Delta$ AIC & Model & R$^{2}$ & $\Delta$ AIC & Model & R$^{2}$ & $\Delta$ AIC & Model & R$^{2}$ & $\Delta$ AIC\\
 \midrule
-GPP & - & - & Linear & 0.5498 & Linear & 0.11 & Polynomial & 0.3602 & Polynomial & 0.3076 & - & - & - & - & Linear & 0.5312\\
-NPP & Linear & 0.06339 & Linear & 0.4036 & Linear & 0.1118 & Polynomial & 0.3165 & Polynomial & 0.178 & - & - & Linear & 0.03561 & Linear & 0.3782\\
-ANPP & Linear & 0.0439 & Linear & 0.3668 & Linear & 0.1732 & Polynomial & 0.2672 & Polynomial & 0.2294 & Polynomial & 0.0632 & Polynomial & 0.06269 & Linear & 0.3425\\
-ANPP woody stem & - & - & Linear & 0.1380 & - & - & Polynomial & 0.2024 & Polynomial & 0.2146 & Linear & 0.07403 & - & - & Linear & 0.1041\\
-ANPP foliage & - & - & Linear & 0.5306 & Linear & 0.1469 & Linear & 0.3076 & Polynomial & 0.3751 & Polynomial & 0.07489 & Polynomial & 0.1724 & Linear & 0.4552\\
+GPP & - & - & - & Log & 0.54 & 50.0 & Lin & 0.11 & 5.7 & Poly & 0.36 & 19.7 & Poly & 0.31 & 15.9 & - & - & - & - & - & - & Lin & 0.53 & 38.2\\
+NPP & Lin & 0.06 & 3.6 & Lin & 0.40 & 38.5 & Lin & 0.11 & 7.3 & Poly & 0.32 & 24.3 & Poly & 0.18 & 15.3 & - & - & - & Lin & 0.04 & 4 & Lin & 0.38 & 28.4\\
+ANPP & Poly & 0.09 & 7.1 & Log & 0.41 & 61.6 & Lin & 0.17 & 18.7 & Poly & 0.27 & 24.5 & Poly & 0.23 & 21.4 & Poly & 0.06 & 2.2 & Poly & 0.06 & 3 & Lin & 0.34 & 44.0\\
+ANPP stem & Poly & 0.09 & 5.4 & Log & 0.17 & 22.3 & - & - & - & Poly & 0.20 & 14.0 & Poly & 0.21 & 17.7 & Log & 0.14 & 7.5 & - & - & - & Log & 0.11 & 12.6\\
+ANPP foliage & - & - & - & Lin & 0.53 & 43.4 & Lin & 0.15 & 7 & Log & 0.32 & 24.2 & Log & 0.35 & 30.0 & Poly & 0.07 & 4.9 & Poly & 0.17 & 7.8 & Log & 0.46 & 32.9\\
 \addlinespace
-BNPP root & - & - & Linear & 0.2799 & Polynomial & 0.1113 & Polynomial & 0.3601 & Polynomial & 0.2584 & - & - & - & - & Linear & 0.2550\\
-BNPP fine root & - & - & Linear & 0.1631 & Linear & 0.08161 & Linear & 0.1376 & - & - & - & - & - & - & Linear & 0.1335\\
-Autotrophic respiration & - & - & Linear & 0.5502 & Linear & 0.226 & Linear & 0.3298 & Polynomial & 0.4499 & - & - & Linear & 0.2613 & Linear & 0.4664\\
-Root respiration & Linear & 0.1578 & Linear & 0.1647 & Linear & 0.1698 & Polynomial & 0.1905 & Polynomial & 0.272 & - & - & Linear & 0.1388 & Linear & 0.1889\\
+BNPP root & - & - & - & Lin & 0.28 & 19.1 & Poly & 0.11 & 3.4 & Poly & 0.36 & 23.2 & Poly & 0.26 & 13.9 & - & - & - & - & - & - & Lin & 0.26 & 14.7\\
+BNPP fine root & - & - & - & Lin & 0.16 & 9.2 & Lin & 0.08 & 2.7 & Log & 0.14 & 7.1 & Log & 0.06 & 1.9 & - & - & - & - & - & - & Lin & 0.13 & 5.8\\
+R auto & - & - & - & Log & 0.57 & 9.4 & Null & 0.26 & 0.6 & Log & 0.36 & 4.8 & Log & 0.35 & 4.3 & - & - & - & Null & 0.3 & 1.5 & Lin & 0.47 & 5.8\\
+R root & Log & 0.16 & 1.9 & Log & 0.19 & 7.3 & Lin & 0.17 & 3.5 & Poly & 0.19 & 1.7 & Poly & 0.27 & 6.7 & - & - & - & Lin & 0.14 & 6.1 & Lin & 0.19 & 5.9\\
 \bottomrule
@@
 
-\elandscape
+Model forms tested include first-order linear (Lin), second-order
+polynomial (Poly), and logarithmic (Log). Shown are models with lowest
+\(\Delta AIC\).
+
+\end{landscape}
 
@@
 
-\caption{\label{tab:unnamed-chunk-6}Table S4. Results of analysis of interactions effects between MAT and MAP, for all FACF}
+\caption{\label{tab:unnamed-chunk-5}Table S3. Signficance of relationships of forest C fluxes to MAT alone and in combination with MAP}
 \centering
@@
 \fontsize{12}{14}\selectfont
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{6cm}>{\raggedright\arraybackslash}p{6cm}>{\raggedright\arraybackslash}p{6cm}>{\raggedright\arraybackslash}p{2cm}>{\raggedleft\arraybackslash}p{2cm}}
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{4cm}>{\raggedright\arraybackslash}p{6cm}>{\raggedright\arraybackslash}p{6cm}>{\raggedright\arraybackslash}p{6cm}>{\raggedleft\arraybackslash}p{2cm}>{}p{2cm}}
 \toprule
-Carbon flux & Significant interactive effect & Significant additive effect & Significant effect of MAT & p-value & R-squared value\\
+Carbon flux & MAT & MAT + MAP & MAT x MAP & R$^{2}$\\
 \midrule
-GPP & FALSE & TRUE & TRUE & <0.0001 & 0.66\\
-NPP & TRUE & TRUE & TRUE & 0.018 & 0.48\\
-ANPP & FALSE & TRUE & TRUE & 0.0349 & 0.45\\
-ANPP woody stem & TRUE & TRUE & TRUE & 0.021 & 0.26\\
-ANPP foliage & FALSE & FALSE & TRUE & <0.0001 & 0.59\\
+GPP & <0.0001 & <0.0001 & NS & 0.66\\
+NPP & <0.0001 & NS & 0.018 & 0.48\\
+ANPP & <0.0001 & 0.035 & NS & 0.45\\
+ANPP stem & <0.0001 & NS & 0.021 & 0.26\\
+ANPP foliage & <0.0001 & NS & NS & 0.59\\
 \addlinespace
-BNPP root & FALSE & FALSE & TRUE & <0.0001 & 0.29\\
-BNPP fine root & FALSE & FALSE & TRUE & 0.002 & 0.15\\
-Autotrophic respiration & FALSE & TRUE & TRUE & 0.041 & 0.71\\
-Root respiration & FALSE & FALSE & TRUE & 0.001 & 0.25\\
+BNPP root & <0.0001 & NS & NS & 0.29\\
+BNPP fine root & 0.0021 & NS & NS & 0.15\\
+R auto & 0.00016 & 0.041 & NS & 0.71\\
+R root & 0.0011 & NS & NS & 0.25\\
 \bottomrule
@@
 \newpage
+\begin{table}[!h]
 
+\caption{\label{tab:unnamed-chunk-6}Table S4. Comparison of growing season length and MAT as predictors of forest C fluxes}
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{5cm}>{\raggedleft\arraybackslash}p{2cm}>{\raggedleft\arraybackslash}p{7cm}>{\raggedleft\arraybackslash}p{7cm}}
+\toprule
+Fixed effect & AIC value & $\Delta$AIC relative to best model & Marginal R$^{2}$\\
+\midrule
+\addlinespace[1em]
+\multicolumn{1}{l}{\textbf{GPP}}\\
+\hspace{1em}MAT & 126.43 & 0.00 & 0.62\\
+\hspace{1em}Growing season length & 140.81 & 14.38 & 0.54\\
+\hspace{1em}None & 178.96 & 52.54 & 0.00\\
+\addlinespace[1em]
+\hline
+\multicolumn{1}{l}{\textbf{NPP}}\\
+\hspace{1em}MAT & 174.88 & 0.00 & 0.52\\
+\hspace{1em}Growing season length & 191.54 & 16.65 & 0.40\\
+\hspace{1em}None & 216.17 & 41.29 & 0.00\\
+\addlinespace[1em]
+\hline
+\multicolumn{1}{l}{\textbf{ANPP}}\\
+\hspace{1em}MAT & 249.51 & 0.00 & 0.29\\
+\hspace{1em}Growing season length & 254.21 & 4.70 & 0.26\\
+\hspace{1em}None & 268.94 & 19.43 & 0.00\\
+\addlinespace[1em]
+\hline
+\multicolumn{1}{l}{\textbf{ANPP stem}}\\
+\hspace{1em}MAT & 235.96 & 0.00 & 0.15\\
+\hspace{1em}Growing season length & 237.29 & 1.33 & 0.14\\
+\hspace{1em}None & 243.14 & 7.18 & 0.00\\
+\addlinespace[1em]
+\hline
+\multicolumn{1}{l}{\textbf{ANPP foliage}}\\
+\hspace{1em}MAT & 484.88 & 0.00 & 0.45\\
+\hspace{1em}Growing season length & 520.96 & 36.09 & 0.35\\
+\hspace{1em}None & 560.35 & 75.47 & 0.00\\
+\addlinespace[1em]
+\hline
+\multicolumn{1}{l}{\textbf{BNPP root}}\\
+\hspace{1em}MAT & 184.54 & 0.00 & 0.59\\
+\hspace{1em}Growing season length & 204.93 & 20.38 & 0.46\\
+\hspace{1em}None & 237.47 & 52.92 & 0.00\\
+\addlinespace[1em]
+\hline
+\multicolumn{1}{l}{\textbf{BNPP fine root}}\\
+\hspace{1em}MAT & 540.19 & 0.00 & 0.24\\
+\hspace{1em}Growing season length & 566.37 & 26.18 & 0.11\\
+\hspace{1em}None & 578.66 & 38.46 & 0.00\\
+\addlinespace[1em]
+\hline
+\multicolumn{1}{l}{\textbf{R auto}}\\
+\hspace{1em}MAT & 45.26 & 0.00 & 0.63\\
+\hspace{1em}Growing season length & 50.36 & 5.10 & 0.50\\
+\hspace{1em}None & 56.17 & 10.91 & 0.00\\
+\addlinespace[1em]
+\hline
+\multicolumn{1}{l}{\textbf{R root}}\\
+\hspace{1em}MAT & 133.54 & 0.00 & 0.25\\
+\hspace{1em}Growing season length & 135.93 & 2.39 & 0.20\\
+\hspace{1em}None & 141.79 & 8.25 & 0.00\\
+\bottomrule
+\end{tabular}}
+\end{table}
+
+\newpage
+\begin{table}[!h]
+
+\caption[Table S5. Best models by carbon flux]{\label{tab:unnamed-chunk-7}Table S5. Best (lowest $\Delta$AIC) single-climate variable models by C flux.}
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{lllrlr}
+\toprule
+Carbon flux & Climate variable & Model type & $\Delta$AIC relative to null model & $\Delta$AIC relative to next best model & R$^{2}$\\
+\midrule
+GPP & T Seas & Poly & 69.5 & 6.55 & 0.71\\
+\cmidrule{1-6}
+ & MAT & Lin & 41.5 & 0.21 & 0.42\\
+
+\multirow{-2}{*}{\raggedright\arraybackslash NPP} & T Seas & Log & 44.3 & - & 0.52\\
+\cmidrule{1-6}
+ANPP & MAT & Lin & 80.5 & 21.4 & 0.44\\
+\cmidrule{1-6}
+ANPP stem & MAT & Lin & 38.5 & 15.87 & 0.24\\
+\cmidrule{1-6}
+ANPP foliage & MAT & Lin & 52.9 & 11.05 & 0.58\\
+\cmidrule{1-6}
+BNPP root & T Seas & Log & 26.6 & 3.01 & 0.36\\
+\cmidrule{1-6}
+BNPP fine root & ATR & Log & 10.9 & 2.11 & 0.19\\
+\cmidrule{1-6}
+ & T Seas & Log & 13.1 & 1.62 & 0.65\\
+
+\multirow{-2}{*}{\raggedright\arraybackslash R auto} & ATR & Log & 11.5 & - & 0.60\\
+\cmidrule{1-6}
+ & T Seas & Log & 9.5 & 0.76 & 0.24\\
+
+ & ATR & Log & 8.8 & - & 0.22\\
+
+\multirow{-3}{*}{\raggedright\arraybackslash R root} & MAT & Lin & 8.3 & - & 0.24\\
+\bottomrule
+\end{tabular}}
+\end{table}
+
+Table includes all models within \(\Delta\)AIC \$\le\$2.0 of the best
+model
+
+\newpage
+\begin{landscape}
+\begin{table}[!h]
+
+\caption{\label{tab:unnamed-chunk-8}Table S6. Pairwise comparisons of correlations with climate variables between C fluxes, with all analyses including a common set of sites.}
+\centering
+\resizebox{\linewidth}{!}{
+\begin{tabular}[t]{lllrrllrl}
+\toprule
+C flux variable 1 & C flux variable 2 & Climate variable & R$^{2}$ variable 1 & R$^{2}$ variable 2 & Model type variable 1 & Model type variable 2 & Number of plots & Variable with higher R$^{2}$\\
+\midrule
+ &  & Latitude & 0.62 & 0.66 & Lin & Lin & 37 & NPP\\
+
+ &  & MAT & 0.62 & 0.70 & Log & Lin & 37 & NPP\\
+
+\multirow{-3}{*}{\raggedright\arraybackslash GPP} & \multirow{-3}{*}{\raggedright\arraybackslash NPP} & T Seas & 0.65 & 0.70 & Log & Log & 37 & NPP\\
+\cmidrule{1-9}
+ &  & Latitude & 0.52 & 0.48 & Log & Log & 158 & NPP\\
+
+ &  & MAT & 0.30 & 0.44 & Log & Lin & 158 & ANPP\\
+
+ & \multirow{-3}{*}{\raggedright\arraybackslash ANPP} & T Seas & 0.47 & 0.43 & Lin & Lin & 158 & NPP\\
+\cmidrule{2-9}
+ &  & Latitude & 0.49 & 0.34 & Log & Lin & 116 & NPP\\
+
+ &  & MAT & 0.41 & 0.22 & Log & Log & 116 & NPP\\
+
+\multirow{-6}{*}{\raggedright\arraybackslash NPP} & \multirow{-3}{*}{\raggedright\arraybackslash BNPP} & T Seas & 0.49 & 0.41 & Log & Log & 116 & NPP\\
+\cmidrule{1-9}
+ &  & Latitude & 0.35 & 0.13 & Lin & Lin & 176 & ANPP\\
+
+ &  & MAT & 0.42 & 0.17 & Lin & Lin & 176 & ANPP\\
+
+ & \multirow{-3}{*}{\raggedright\arraybackslash ANPP stem} & T Seas & 0.29 & 0.09 & Lin & Lin & 176 & ANPP\\
+\cmidrule{2-9}
+ &  & Latitude & 0.32 & 0.45 & Log & Log & 96 & ANPP foliage\\
+
+ &  & MAT & 0.36 & 0.50 & Lin & Lin & 96 & ANPP foliage\\
+
+\multirow{-6}{*}{\raggedright\arraybackslash ANPP} & \multirow{-3}{*}{\raggedright\arraybackslash ANPP foliage} & T Seas & 0.27 & 0.42 & Lin & Lin & 96 & ANPP foliage\\
+\cmidrule{1-9}
+ &  & Latitude & 0.64 & 0.34 & Null & Null & 11 & GPP\\
+
+ &  & MAT & 0.69 & 0.34 & Null & Null & 11 & GPP\\
+
+\multirow{-3}{*}{\raggedright\arraybackslash GPP} & \multirow{-3}{*}{\raggedright\arraybackslash R auto} & T Seas & 0.64 & 0.32 & Null & Null & 11 & GPP\\
+\cmidrule{1-9}
+ &  & Latitude & 0.01 & 0.39 & Null & Null & 9 & R root\\
+
+ &  & MAT & 0.08 & 0.35 & Null & Null & 9 & R root\\
+
+\multirow{-3}{*}{\raggedright\arraybackslash BNPP} & \multirow{-3}{*}{\raggedright\arraybackslash R root} & T Seas & 0.01 & 0.63 & Null & Null & 9 & R root\\
+\bottomrule
+\end{tabular}}
+\end{table}
+\end{landscape}
+
+\newpage
 \begin{figure}[H]
-\includegraphics[width=27.78in,height=0.95\textheight]{tables_figures/distribution_all_samples} \caption{Figure S1: Maps showing distribution of samples for the nine FACF analyzed here.}\label{fig:unnamed-chunk-7}
+\includegraphics[height=0.95\textheight]{tables_figures/distribution_all_samples} \caption{Figure S1: Maps showing distribution of samples for the nine forest C fluxes analyzed here}\label{fig:unnamed-chunk-9}
 \end{figure}
 
-\blandscape
-
+\begin{landscape}
 \begin{figure}[H]
 
-{\centering \includegraphics[width=48.61in,height=0.95\textheight]{tables_figures/climate_regressions}
+{\centering \includegraphics[height=0.95\textheight]{tables_figures/climate_regressions}
 
@@
 
-\caption{Figure S2: Correlations among latitude and climate variables. Variable names and units given in Table S1}\label{fig:unnamed-chunk-8}
+\caption{Figure S2: Scatterplots and Pearson's R values for relationships between latitude and climate variables}\label{fig:unnamed-chunk-10}
 \end{figure}
+\end{landscape}
 
-\elandscape
+\begin{landscape}
+\begin{figure}[H]
 
-\blandscape
+{\centering \includegraphics[width=0.8\linewidth]{tables_figures/ratio_grid_plots}
 
+}
+
+\caption{Figure S3: Ratios among forest C fluxes as a function of latitude and climate variables}\label{fig:unnamed-chunk-11}
+\end{figure}
+\end{landscape}
+
+\begin{landscape}
 \begin{figure}[H]
-\includegraphics[width=41.67in,height=0.95\textheight]{tables_figures/grid_plots_climate1} \caption{Figure S3: Individual plots of FACF in relation to mean annual climate, part 1.}\label{fig:unnamed-chunk-9}
+\includegraphics[height=0.95\textheight]{tables_figures/grid_plots_climate1} \caption{Figure S4: Individual plots of forest C fluxes in relation to mean annual climate, part 1.}\label{fig:unnamed-chunk-12}
 \end{figure}
@@
 \begin{figure}[H]
-\includegraphics[width=41.67in,height=0.95\textheight]{tables_figures/grid_plots_climate2} \caption{Figure S4: Individual plots of FACF in relation to mean annual climate, part 2.}\label{fig:unnamed-chunk-10}
+\includegraphics[height=0.95\textheight]{tables_figures/grid_plots_climate2} \caption{Figure S5: Individual plots of forest C fluxes in relation to mean annual climate, part 2.}\label{fig:unnamed-chunk-13}
+\end{figure}
+\end{landscape}
+
+\begin{figure}[H]
+\includegraphics[height=0.95\textheight]{tables_figures/grid_plots_seasonality3} \caption{Figure S6: Individual plots of forest C fluxes in relation to mean climate seasonality, part 1.}\label{fig:unnamed-chunk-14}
 \end{figure}
 
-\elandscape
+\newpage
+\begin{figure}[H]
+\includegraphics[width=1\linewidth]{tables_figures/grid_plots_seasonality4} \caption{Figure S7: Individual plots of forest C fluxes in relation to mean climate seasonality, part 2.}\label{fig:unnamed-chunk-15}
+\end{figure}
 
+\newpage
 \begin{figure}[H]
-\includegraphics[width=27.78in,height=0.95\textheight]{tables_figures/grid_plots_seasonality3} \caption{Figure S5: Individual plots of FACF in relation to mean climate seasonality, part 1.}\label{fig:unnamed-chunk-11}
+\includegraphics[width=1\linewidth]{tables_figures/gridded_growing_season1} \caption{Figure S8: Growing season length-standardized forest C fluxes in relation to mean growing season climate, part 1.}\label{fig:unnamed-chunk-16}
+\end{figure}
+
+\newpage
+\begin{figure}[H]
+\includegraphics[width=1\linewidth]{tables_figures/gridded_growing_season2} \caption{Figure S9: Growing season length-standardized forest C fluxes in relation to mean growing season climate, part 2.}\label{fig:unnamed-chunk-17}
 \end{figure}
@@
 
-\begin{figure}[H]
-\includegraphics[width=1\linewidth]{tables_figures/grid_plots_seasonality4} \caption{Figure S6: Individual plots of FACF in relation to mean climate seasonality, part 2.}\label{fig:unnamed-chunk-12}
-\end{figure}
+\hypertarget{references}{%
+\section*{References}\label{references}}
+\addcontentsline{toc}{section}{References}
 
-\newpage
+\hypertarget{refs}{}
+\leavevmode\hypertarget{ref-abatzoglou_terraclimate_2018}{}%
+Abatzoglou, J. T., Dobrowski, S. Z., Parks, S. A., \& Hegewisch, K. C.
+(2018). TerraClimate, a high-resolution global dataset of monthly
+climate and climatic water balance from 1958--2015. \emph{Scientific
+Data}, \emph{5}, 170191. \url{https://doi.org/10.1038/sdata.2017.191}
 
-\begin{figure}[H]
-\includegraphics[width=1\linewidth]{tables_figures/gridded_growing_season1} \caption{Figure S7: Growing season length-standardized FACF in relation to mean growing season climate, part 1.}\label{fig:unnamed-chunk-13}
-\end{figure}
+\leavevmode\hypertarget{ref-fick_worldclim_2017}{}%
+Fick, S. E., \& Hijmans, R. J. (2017). WorldClim 2: New 1-km spatial
+resolution climate surfaces for global land areas. \emph{International
+Journal of Climatology}, \emph{37}(12), 4302--4315.
+\url{https://doi.org/10.1002/joc.5086}
 
-\newpage
+\leavevmode\hypertarget{ref-harris_updated_2014}{}%
+Harris, I., Jones, P. D., Osborn, T. J., \& Lister, D. H. (2014).
+Updated high-resolution grids of monthly climatic observations - the CRU
+TS3.10 dataset: \emph{International Journal of Climatology},
+\emph{34}(3), 623--642. \url{https://doi.org/10.1002/joc.3711}
 
-\begin{figure}[H]
-\includegraphics[width=1\linewidth]{tables_figures/gridded_growing_season2} \caption{Figure S8: Growing season length-standardized FACF in relation to mean growing season climate, part 2.}\label{fig:unnamed-chunk-14}
-\end{figure}
-
-\newpage
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=1\linewidth]{tables_figures/ratio_grid_plots}
-
-}
-
-\caption{Figure S9: Ratios among FACF as a function of latitude and climate variables}\label{fig:unnamed-chunk-15}
-\end{figure}
+\leavevmode\hypertarget{ref-hijmans_very_2005}{}%
+Hijmans, R. J., Cameron, S. E., Parra, J. L., Jones, P. G., \& Jarvis,
+A. (2005). Very high resolution interpolated climate surfaces for global
+land areas. \emph{International Journal of Climatology}, \emph{25}(15),
+1965--1978. \url{https://doi.org/10.1002/joc.1276}
 

### 2020-08-07 manuscript/appendix.Rmd
fixed error that prevented knitting

@@
 ```
-Table includes all models within $\Delta$AIC $\le$2.0 of the best model
+Table includes all models within $\Delta$AIC $\le$ 2.0 of the best model
 

### 2020-08-07 manuscript/manuscript draft.Rmd
fix ref

@@
 
-Forest C fluxes decrease with latitude [e.g., @luyssaert_co_2007; @piao_forest_2010; @gillman_latitude_2015; @li_mapping_2019]. 
+Forest C fluxes decrease with latitude [e.g., @luyssaert_co_2007;  @gillman_latitude_2015; @li_mapping_2019]. 
 However, studies have differed in their conclusions regarding the shape of this relationship -- quite possibly because of lack of standardization with respect to methodology and stand history.
@@
 C fluxes increase strongly with $MAT$ on the global scale, but whether they saturate or potentially decrease at higher temperatures remains disputed. 
-Some studies have detected no deceleration or decline in $GPP$ [@luyssaert_co_2007], $NPP$ [@schuur_productivity_2003], or root respiration [$R_{root}$; @wei_forest_2010] with increasing $MAT$.
+Some studies have detected no deceleration or decline in $GPP$ [@luyssaert_co_2007], $NPP$ [@schuur_productivity_2003], or root respiration [$R_{root}$; @piao_forest_2010; @wei_forest_2010] with increasing $MAT$.
 In contrast, others have found evidence of saturation or decline of C flux in the warmest climates; @luyssaert_co_2007 found $NPP$ saturating at around 10$^\circ$C MAT; @larjavaara_temperature_2012 found that increases in $GPP$ saturate at approximately 25$^\circ$C $MAT$, and @sullivan_long-term_2020 found that, within the tropics, $ANPP_{stem}$ decreases at the highest maximum temperatures.

### 2020-08-07 manuscript/appendix.Rmd
Update appendix.Rmd

@@
 nocite: | 
-  @hijmans_very_2005; @fick_worldclim_2017; @harris_updated_2014;  @abatzoglou_terraclimate_2018
+  @hijmans_very_2005; @fick_worldclim_2017; @harris_updated_2014;  @abatzoglou_terraclimate_2018; @trabucco_global_2019
 
@@
   column_spec(6, width = "5cm") %>%
-  kableExtra::footnote(number = c("Hijmans et al. (2005)", "Fick et al. (2017)", "Harris et al. (2017)", "Abatzoglou et al. (2018)"), footnote_as_chunk = TRUE) %>%
+  kableExtra::footnote(number = c("Hijmans et al. (2005)", "Fick and Hijmans. (2017)", "Harris et al. (2017)", "Trabucco and Zomer (2019)" , "Abatzoglou et al. (2018)"), footnote_as_chunk = TRUE) %>%
   kableExtra::footnote(symbol = "The WorldClim version used was the most recent available at the time of analysis", footnote_as_chunk = TRUE)
@@
 Model form, $\\Delta$AIC, and R$^{2}$ for each climate variable as a single fixed effect in models for each C flux.", caption.short = "Table S2.
-Model form, $\\Delta$AIC and R$^{2}$ for each climate variables as a single fixed effect in models for each C flux", escape = F) %>%
+Model form, $\\Delta$AIC and R$^{2}$ for each climate variable as a single fixed effect in models for each C flux", escape = F) %>%
   kable_styling(latex_options = c("scale_down", "hold_position")) %>%
@@
 ```
-Model forms tested include first-order linear (Lin), second-order polynomial (Poly), and logarithmic (Log). Shown are models with lowest $\Delta AIC$.
+Model forms tested include first-order linear (Lin), second-order polynomial (Poly), and logarithmic (Log).
+Shown are models with lowest AIC value.
+$\Delta$AIC value is calculated with respect to the null.
 
@@
 kable(S4, format = "latex", booktabs = TRUE, caption = "Table S6.
-Pairwise comparisons of correlations with climate variables between C fluxes, with all analyses including a common set of sites.", escape = F) %>%
+Pairwise comparisons of correlations with climate variables between C fluxes, with all analyses conducted on a set of sites common to each pair.", escape = F) %>%
   kable_styling(latex_options = c("scale_down", "hold_position")) %>%
@@
 \newpage
-```{r echo=FALSE, fig.cap = "Figure S1: Maps showing distribution of samples for the nine forest C fluxes analyzed here", out.height ='95%', fig.pos='H'}
+```{r echo=FALSE, fig.cap = "Figure S1: Maps showing distribution of samples for the nine forest C fluxes analyzed", out.height ='95%', fig.pos='H'}
 

### 2020-08-07 manuscript/manuscript draft.Rmd
updates to abstract and keywords to match submission

@@
 Here, we draw upon 1,319 records from the Global Forest Carbon Database (ForC), representing all major forest types and the nine most significant autotrophic carbon fluxes, to comprehensively explore how C cycling in mature, undisturbed forests varies with latitude and climate on a global scale.
-We show that, across all flux variables analyzed, C cycling decreases continuously with absolute latitude -- a finding that confirms multiple previous studies but contradicts the idea that net primary productivity ($NPP$) of temperate forests rivals that of tropical forests.
+We show that, across all flux variables analyzed, C cycling decreases continuously with absolute latitude -- a finding that confirms multiple previous studies but contradicts the idea that net primary productivity of temperate forests rivals that of tropical forests.
 C flux variables generally displayed similar trends across latitude and multiple climate variables, with no differences in allocation detected at this global scale.
-Temperature variables in general, and mean annual temperature ($MAT$) and temperature seasonality in particular, were the best univariate predictors of C flux, explaining 19 - 71% of variation in the C fluxes analyzed.
+Temperature variables in general, and mean annual temperature and temperature seasonality in particular, were the best univariate predictors of C flux, explaining 19 - 71% of variation in the C fluxes analyzed.
 The effects of temperature were modified by moisture availability, with C flux reduced under hot and dry conditions and sometimes under very high precipitation. 
@@
 
-**Keywords:**  carbon fluxes; climate; forest; productivity; respiration; latitude
+**Keywords:**  carbon fluxes; carbon dioxide (CO~2~); climate; forest; global; productivity; respiration; latitude
 

### 2020-08-07 manuscript/manuscript draft.Rmd
removed references from Table 1
#62

@@
 hypothesis_table <- read.csv("hypothesis_table.csv", stringsAsFactors = FALSE, check.names = FALSE)
-hypothesis_table[1,1] <- paste0(hypothesis_table[1,1], footnote_marker_number(1),"$^{,}$",footnote_marker_number(2),"$^{,}$", footnote_marker_number(3),"$^{,}$", footnote_marker_number(10))
-hypothesis_table[2,1] <- paste0(hypothesis_table[2,1], footnote_marker_number(1),"$^{,}$",footnote_marker_number(4),"$^{,}$",footnote_marker_number(9))
-hypothesis_table[3,1] <- paste0(hypothesis_table[3,1], footnote_marker_number(1),"$^{,}$",footnote_marker_number(4))
-hypothesis_table[4,1] <- paste0(hypothesis_table[4,1], footnote_marker_number(5))
-hypothesis_table[10,1] <- paste0(hypothesis_table[10,1], footnote_marker_number(6),"$^{,}$",footnote_marker_number(7),"$^{,}$", footnote_marker_number(8))
-hypothesis_table[11,1] <- paste0(hypothesis_table[11,1], footnote_marker_number(7),"$^{,}$", footnote_marker_number(8))
-hypothesis_table[12,1] <- paste0(hypothesis_table[12,1], footnote_marker_number(8))
+#hypothesis_table[1,1] <- paste0(hypothesis_table[1,1], footnote_marker_number(1),"$^{,}$",footnote_marker_number(2),"$^{,}$", footnote_marker_number(3),"$^{,}$", footnote_marker_number(10))
+#hypothesis_table[2,1] <- paste0(hypothesis_table[2,1], footnote_marker_number(1),"$^{,}$",footnote_marker_number(4),"$^{,}$",footnote_marker_number(9))
+#hypothesis_table[3,1] <- paste0(hypothesis_table[3,1], footnote_marker_number(1),"$^{,}$",footnote_marker_number(4))
+#hypothesis_table[4,1] <- paste0(hypothesis_table[4,1], footnote_marker_number(5))
+#hypothesis_table[10,1] <- paste0(hypothesis_table[10,1], footnote_marker_number(6),"$^{,}$",footnote_marker_number(7),"$^{,}$", footnote_marker_number(8))
+#hypothesis_table[11,1] <- paste0(hypothesis_table[11,1], footnote_marker_number(7),"$^{,}$", footnote_marker_number(8))
+#hypothesis_table[12,1] <- paste0(hypothesis_table[12,1], footnote_marker_number(8))
 kable(hypothesis_table, format = "latex", booktabs = TRUE, caption = "Summary of research questions, corresponding hypotheses, and results.
@@
   #column_spec(2:11, width = "1.8cm") %>%
-  kableExtra::landscape() %>%
-  kableExtra::footnote(number = c("Luyssaert et al.
+  kableExtra::landscape() 
+ # kableExtra::footnote(number = c("Luyssaert et al.
 (2007)", "Gillman et al.

### 2020-08-07 manuscript/appendix.Rmd
fixed appendix citations
merged appendix_library into library.bib, deleted
knit appendix

@@
 
-bibliography: appendix_library.bib
+bibliography: library.bib
 csl: global-change-biology.csl

### 2020-08-07 manuscript/manuscript draft.Rmd
issue #118

@@
 We selected as the best model that with the highest $\Delta$ AIC relative to a null model with no fixed term, with the qualification that a polynomial model was considered an improvement over a linear model only if it reduced the AIC value by 2.0 or more.
+In addition, pairwise comparisons of $R^2$ values were carried out for a selection of pairs of C fluxes to test for differences among variables in the proportion of variation explained by latitude and climate.
+Models were run on data from sets of sites that were common to each pair, in order to account for variation in the number of data points included.
 
@@
 2; Table S2).
-Latitude was a strong predictor for many of the carbon fluxes, particularly the larger fluxes (Table S2).
+Latitude was a strong predictor for many of the carbon fluxes, particularly the larger fluxes (Table S2, S6).
 Specifically, latitude explained 64% of variation in GPP (n = 243, p<0.0001), 50% in NPP (n = 161, p<0.0001) and 44% in ANPP (n = 278, p<0.0001).
@@
 For only one variable, $BNPP$, did a lognormal fit provide an improvement over a first-order linear relationship, though this was not significant ($\Delta$AIC <2).
-As with latitude, $MAT$ tended to explain more variation in the larger fluxes ($GPP$, $NPP$, $ANPP$, $R_{auto}$) and $ANPP_{foliage}$ (all $R^2$> 0.4) than in subsidiary and belowground fluxes ($ANPP_{stem}$, $R_{root}$, $BNPP_{fine.root}$; all $R^2$< 0.25).
+As with latitude, $MAT$ tended to explain more variation in the larger fluxes ($GPP$, $NPP$, $ANPP$, $R_{auto}$) and $ANPP_{foliage}$ (all $R^2$> 0.4) than in subsidiary and belowground fluxes ($ANPP_{stem}$, $R_{root}$, $BNPP_{fine.root}$; all $R^2$< 0.25; Table S6).
 

### 2020-08-07 manuscript/journal_submissions/2020_8_GCB/submission_notes.Rmd
more submission stuff
update submission notes

@@
 output: pdf_document
+
 ---
@@
 
+[*author instructions*](https://onlinelibrary-wiley-com.smithsonian.idm.oclc.org/page/journal/13652486/homepage/forauthors.html)
+
+# ORCHIDs
+
+Ben: 0000-0001-9525-4633
+
+Valentine: 0000-0002-4519-481X
+
 # submission questions
@@
 ## What is the scientific question you are addressing?
-## What is/are the key finding(s) that answers this question?
+
+How do nine major annual autotrophic carbon fluxes in forests vary globally with respect to latitude and climate?
+
+## What is/are the key finding(s) that answers this question
+
+Nine major carbon fluxes decreases linearly and roughly in proportion to one another across latitudes.
+Temperature variables are the best univariate climate predictors, with additional effects of moisture availability.
+The influence of climate on carbon fluxes within the growing season was small but significant for a number of fluxes.
+
 ##    Why is this work important and timely?
+
+Forests will strongly influence climate change, and cohesive understanding of their carbon fluxes vary with respect to climate is foundational to predicting how they will change in coming decades.
+
 ##    Does your paper fall within the scope of GCB; what biological AND global change aspects does it address?
+
+The paper falls within the scope of GCB addressing how forest carbon fluxes (biological aspect) relate to climate (global change aspect).
+
 ##    What are the three most recently published papers that are relevant to this question? This information will assist the Editors in selecting reviewers.
+
+Li, & Xiao.
+(2019).
+Mapping Photosynthesis Solely from Solar-Induced Chlorophyll Fluorescence: A Global, Fine-Resolution Dataset of Gross Primary Production Derived from OCO-2.
+Remote Sensing, 11(21), 2563.
+
+Collalti, A., & Prentice, I. C. (2019). Is NPP proportional to GPP? Waring’s hypothesis 20 years on. Tree Physiology, 39(8), 1473–1483.
+
+Sullivan et al (2020). Long-term thermal sensitivity of Earth’s tropical forests. Science, 368(6493), 869–874.

### 2020-08-07 manuscript/word_count.Rmd
Create word_count.Rmd

--- 2020-08-07 manuscript/manuscript draft.Rmd
@@
 output:
-  pdf_document:
-    fig_caption: yes
+  word_document:
 bibliography: library.bib
 csl: global-change-biology.csl
-header-includes:  \usepackage{float}
-                  \usepackage{caption}
-                  \captionsetup[table]{font=footnotesize}
-                  \captionsetup[figure]{font=footnotesize}
-                  \usepackage{setspace}\doublespacing
-                  \usepackage{lineno}\linenumbers
 
-nocite: | 
-  @chu_does_2016
 ---
-
-```{r, include=FALSE}
-options(tinytex.verbose = TRUE)
-options(knitr.table.format = "latex")
-```
-
-```{r include=FALSE}
-knitr::write_bib(c(
-  .packages(), 'knitr', 'rmarkdown', 'lme4', 'MuMIn', 'plyr', 'raster', 'ncdf4', 'Hmisc', 'plyr', 'merTools', 'visreg', 'r2glmm', 'nlme', 'viridis', 'ggplot2', 'ggpubr', 'gridExtra', 'AICcmodavg'
-), 'packages.bib')
-```
-
-```{r setup, include = FALSE}
-#knitr::opts_knit$set(root.dir = 'C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/')
-  # knitr::opts_knit$set(root.dir = '/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/Global_Productivity/manuscript/')
-```
-
-```{r results="asis", include=FALSE}
-cat("
-<style>
-caption {
-      font-size: 0.75em;
-    }
-</style>
-")
-```
-**Title:** Global patterns of forest autotrophic carbon fluxes
-
-**Running head:** Global patterns of forest carbon fluxes
-
-**Authors:**
-
-Rebecca Banbury Morgan^1,2^
-
-Valentine Herrmann^1^
-
-Norbert Kunert^1,3^
-
-Ben Bond-Lamberty^4^
-
-Helene C. Muller-Landau^3^
-
-Kristina J. Anderson-Teixeira^1,3^*
-
-**Institutional Affiliations:**
-
-1. Conservation Ecology Center; Smithsonian Conservation Biology Institute; Front Royal, VA, USA
-
-2. School of Geography, University of Leeds, Leeds, UK
-
-3. Center for Tropical Forest Science-Forest Global Earth Observatory; Smithsonian Tropical Research Institute; Panama, Republic of Panama
-
-4. Joint Global Change Research Institute, Pacific Northwest National Laboratory, College Park Maryland 20740 USA
-
-*Corresponding Author:
-
-phone: 1-540-635-6546
-
-fax:1-540-635-6506
-
-email: teixeirak@si.edu
-
-\newpage
-### Abstract
-Carbon (C) fixation, allocation, and metabolism by trees set the basis for energy and material flows in forest ecosystems and define their interactions with Earth's changing climate. 
-However, we lack a cohesive synthesis on how forest carbon fluxes vary globally with respect to climate and one another. 
-Here, we draw upon 1,319 records from the Global Forest Carbon Database (ForC), representing all major forest types and the nine most significant autotrophic carbon fluxes, to comprehensively explore how C cycling in mature, undisturbed forests varies with latitude and climate on a global scale.
-We show that, across all flux variables analyzed, C cycling decreases continuously with absolute latitude -- a finding that confirms multiple previous studies but contradicts the idea that net primary productivity of temperate forests rivals that of tropical forests.
-C flux variables generally displayed similar trends across latitude and multiple climate variables, with no differences in allocation detected at this global scale.
-Temperature variables in general, and mean annual temperature and temperature seasonality in particular, were the best univariate predictors of C flux, explaining 19 - 71% of variation in the C fluxes analyzed.
-The effects of temperature were modified by moisture availability, with C flux reduced under hot and dry conditions and sometimes under very high precipitation. 
-C fluxes increased with growing season length, but this was never the best univariate predictor.
-Within the growing season, the influence of climate on C cycling was small but significant for a number of flux variables. 
-These findings clarify how forest C flux varies with latitude and climate on a global scale. 
-In a period of accelerating climatic change, this improved understanding of the fundamental climatic controls on forest C cycling sets a foundation for understanding patterns of change.
-
-**Keywords:**  carbon fluxes; carbon dioxide (CO~2~); climate; forest; global; productivity; respiration; latitude
-
-\newpage
-### Introduction
 
@@
 
-\renewcommand{\arraystretch}{2}
-```{r eval = TRUE, echo=FALSE, warning=FALSE}
-library(knitr)
-library(kableExtra)
-hypothesis_table <- read.csv("hypothesis_table.csv", stringsAsFactors = FALSE, check.names = FALSE)
-#hypothesis_table[1,1] <- paste0(hypothesis_table[1,1], footnote_marker_number(1),"$^{,}$",footnote_marker_number(2),"$^{,}$", footnote_marker_number(3),"$^{,}$", footnote_marker_number(10))
-#hypothesis_table[2,1] <- paste0(hypothesis_table[2,1], footnote_marker_number(1),"$^{,}$",footnote_marker_number(4),"$^{,}$",footnote_marker_number(9))
-#hypothesis_table[3,1] <- paste0(hypothesis_table[3,1], footnote_marker_number(1),"$^{,}$",footnote_marker_number(4))
-#hypothesis_table[4,1] <- paste0(hypothesis_table[4,1], footnote_marker_number(5))
-#hypothesis_table[10,1] <- paste0(hypothesis_table[10,1], footnote_marker_number(6),"$^{,}$",footnote_marker_number(7),"$^{,}$", footnote_marker_number(8))
-#hypothesis_table[11,1] <- paste0(hypothesis_table[11,1], footnote_marker_number(7),"$^{,}$", footnote_marker_number(8))
-#hypothesis_table[12,1] <- paste0(hypothesis_table[12,1], footnote_marker_number(8))
-kable(hypothesis_table, format = "latex", booktabs = TRUE, caption = "Summary of research questions, corresponding hypotheses, and results.
-Statistically signficant support for/ rejection of hypotheses is indicated with 'yes'/'no', and '-' indicates no significant relationship.
-Parentheses indicate partial overall support or rejection of hypotheses across all fluxes considered.", escape = F) %>%
-  kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 12) %>%
-  pack_rows("Q1. How do C fluxes vary with latitude?", 1, 1, latex_gap_space = "1em", colnum = 5, hline_before = FALSE) %>%
-  pack_rows("Q2. How do C fluxes vary with mean annual temperature (MAT) and precipitation (MAP)?", 2, 4, latex_gap_space = "1em", colnum = 5, hline_before = TRUE) %>%
-  pack_rows("Q3. How are C fluxes related to other annual climate variables?", 5, 7, latex_gap_space = "1em", colnum = 5, hline_before = TRUE) %>%
-  pack_rows("Q4. How does seasonality influence annual C fluxes?", 8, 11, latex_gap_space = "1em", colnum = 5, hline_before = TRUE) %>%
-  pack_rows("Q5. When standardised by growing season length, how do annual C fluxes vary with climate?", 12, 15, latex_gap_space = "1em", colnum = 5, hline_before = TRUE) %>%
-  add_header_above(c(" ",  " ", "Forest autotrophic carbon fluxes" = 9, " ")) %>%
-  #column_spec(2:11, width = "1.8cm") %>%
-  kableExtra::landscape() 
- # kableExtra::footnote(number = c("Luyssaert et al.
-(2007)", "Gillman et al.
-(2015)", "Simova and Storch (2017)", "Schuur (2003)", "Taylor et al.
-(2016)", "Malhi (2012)", "Michaletz et al.
-(2014)", "Chu et al.
-(2016)", "Piao et al.
-(2010)", "Huston & Wolverton (2009)"), footnote_as_chunk = TRUE)
-```
-
 ### Materials and Methods
@@
 Grazed sites were retained.
-
-```{r eval = TRUE, echo=FALSE, warning=FALSE}
-library(knitr)
-library(kableExtra)
-FACF_table <- read.csv("FACF_variables.csv", stringsAsFactors = FALSE, check.names = FALSE)
-names(FACF_table) [6]  <- paste0(names(FACF_table) [6], footnote_marker_symbol(1))
-kable(FACF_table, format = "latex", booktabs = TRUE, caption = "Definitions and sample sizes of carbon flux variables used in analysis.
-All variables are in units of Mg C ha$^{-1}$ yr$^{-1}$.
-", escape = F) %>%
-  kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 12) %>%
-  column_spec(2, width = "4cm") %>%
-  column_spec(3, width = "7cm") %>%
-  column_spec(4, width = "10cm") %>%
-  column_spec(5:6, width = "1.5cm") %>%
-  add_header_above(c(" ",  " "," ",  " ", "Sample size" = 2)) %>%
-  kableExtra::footnote(symbol = c("Geographic areas group geographically proximate sites, defined using a hierarchical cluster analysis on the distance matrix of the sites, and a cutoff of 25km"), footnote_as_chunk = FALSE)
-```
-
-```{r echo=FALSE, out.width='100%', fig.cap = "Map showing all data used in the analysis, coded by variable. Variables are plotted individually in Fig. S1. ", fig.pos='H'}
-  knitr::include_graphics("distribution_all_variables_cropped.jpg")
-```
 
@@
 
-```{r echo=FALSE, out.width='100%', fig.cap = "Latitudinal trends in forest autotropic carbon flux.
-Plotted are linear models, all of which were significant $(p<0.05)$ and had AIC values within 2.0 of the best model (for two fluxes, logarithmic fits were marginally better; Table S2).
-Each panel shows major C fluxes together with component fluxes.
-Also plotted are predicted trends in the major C fluxes based on the sum of component fluxes.
-95\\% confidence intervals are plotted for the major flux for comparison with predicted trends.
-In (d), which shows three belowground fluxes, the major flux, total belowground carbon flux, has insufficient data (n=9) to support a regression", fig.pos='H'}
-
-  knitr::include_graphics("combined_stacked.png")
-```
-
 Smaller component fluxes summed approximately to larger fluxes across the latitudinal gradient (Fig.
@@
 
-```{r echo=FALSE, out.width='100%', fig.cap = "Interactive effects of mean annual temperature and precipitation on annual forest carbon fluxes.
-For visualization purposes, data points are grouped into bins of 0 - 1000, 1001 - 2000, 2001 - 3000, and >3000mm mean annual precipitation, and lines of best fit models are plotted for mean annual precipitation values of 500, 1500, 2500, and 3500mm.
-All regressions are significant $(p<0.05)$.", fig.pos='H'}
-  knitr::include_graphics("mat_map_interaction.png")
-```
-
 *Q3. How does C flux relate to other annual climate variables?*
@@
 S4-S5; Table S2).
-
-```{r echo=FALSE, out.width='100%', fig.cap = "Plots of carbon fluxes against (a) mean annual temperature; (b) mean annual precipitation; (c) potential evapotranspiration, (d) vapour pressure deficit; (e) temperature seasonality; (f) length of growing season.
-For visualization purposes, data for each flux was rescaled with a mean of 0 and standard deviation of 1.
-Lines of best fit are plotted according to the best model selected during analysis.
-All regressions are significant $(p<0.05)$.", fig.pos='H'}
-  knitr::include_graphics("combined_plots.png")
-```
 
@@
 Original compilation of the ForC database was funded by DOE grants DE-SC0008085 and DE-SC0010039 to KAT.
-
-### References

### 2020-09-16 manuscript/journal_submissions/2020_8_GCB/appeal.Rmd
load decision, started appeal

---
title: "Appeal of rejection of GCB-20-1713"
output: pdf_document
bibliography: library.bib
csl: global-change-biology.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Dear Editor:

We would like to appeal the reject decision of our manuscript "Global patterns of forest autotrophic carbon fluxes" (GCB-20-1713).

As far as we can tell from the comments to the authors, reviewer 2 was supportive, and her/his comments could easily be addressed. We therefore assume the reject decision was based on the feedback of reviewer 1, which we believe was off-base for several reasons:

* The reviewer states, "...the findings seem not improving contemporary knowledge of forest carbon cycling, and the study appears to be a simple statistical analyses of literature data..." and "the analyses performed were too conventional that similar analyses with a smaller size of similar datasets have been performed a decade ago." These statements betray a lack of perspective on the scope of our analysis relative to previous studies. Whereas most studies consider at most a few flux variables, our considers nine. The most comprehensive similar study is @luyssaert_co_2007, which was published in *GCB* more than a decade ago, includes much less data, does not control for effects of stand age/ disturbance history, and examines latitudinal/ global climatic trends in only # variables. This study has been cited >580 times and was highlighted as a noteworthy publication within *GCB* ([Long, 2020](https://doi.org/10.1111/gcb.14921)). It would seem that something along these lines, but with an expanded database and a more carefully controlled analysis, would have high potential to become a "classic" reference in global forest carbon cycling.

* The reviewer objects that the results are not surprising--e.g., that it is not a new finding that C flux declines with latitude and increases with MAT. We agree that the results are not surprising, and we recognize that surprising results tend to produce high-impact publications. However, there is also a place for high-quality synthesis that clarifies current understanding, which is the goal of this publication. Such syntheses are often published as review papers, but given that our analysis entails original analysis, we submitted it as a Primary Research Article. Perhaps it would be better suited, and reviewed with more appropriate expectations, as a Research Review [*sensu* @anderson-teixeira_carbon_2016].

* The reviewer objects to the fact that the study does not "reveal the biological mechanisms lead to the detected pattern", and states, "Our contemporary knowledge on forest carbon cycling has moved well beyond using mean climatic variable to extrapolate or guess the response of forest carbon cycling to climate change. The spatial gradient do not necessarily represent the response of forests to climate change." We agree, and fully acknowledge in the paper, that this study does not reveal the biological mechanisms. Further, we do not argue that the spatial gradient represents the response of forests to climate change, as clearly discussed in the final paragraph of the discussion. However, we do argue that broad-scale patterns in C cycling across climatic gradients remain--and should remain--one of the important approaches for inferring how forests might respond to climate change. We note that studies using this approach have been published this year in *Science* [@sullivan_long-term_2020]... *(add some more examples)*. While we fully recognize the limitations to such studies for inferring how forests will respond to climate change, we argue that this approach is one of several important, complementary approaches to addressing a challenging problem (Anderson-Teixeira et al. 2013, GCB). The reviewer's apparent wholesale rejection of this approach seems short-sighted, and the expectation that a single study attempt to explain mechanisms behind global patterns in nine flux variables seems naïve. Getting at mechanisms (well) would require integration of ForC with modeling efforts on a scale that is not currently feasible, but lies within the vision of leaders in the modeling community, as stated in an opinion piece that I recently reviewed a paper for *GCB* (give citation/ number). (This piece lays out a vision for improvement of community infrastructure for modeling, including one *stated the goal of linking with leading databases, specifically citing ForC*.)

* The reviewer suggested (twice) that we "consider a data journal like ESSD". This suggestion is completely inappropriate for two reasons. First, the hypothesis-testing analyses presented here would be completely inappropriate for a data journal. Second, the ForC database is already published [tropical portion only in @anderson-teixeira_carbon_2016, *GCB*; full database as a data paper in @anderson-teixeira_forc_2018, *Ecology*]. Many qualified reviewers would already be aware of this, and an even semi-careful review would not miss this fact, which was prominently cited in the last paragraph of the introduction and the first sentence of the methods. Thus, this suggestion betrays carelessness of review and/or limited understanding of appropriate standards for judging where scientific work should be published. Either way, it does not seem appropriate that *GCB* should put much stake in this review.

If, despite these arguments, the *GCB* Editorial Board does not feel that the paper is sufficiently competitive for publication in *GCB*, we will accept the decision and publish elsewhere. However, we believe that this manuscript has potential to become a "classic" reference on the subject of carbon cycling in forests globally, and that *GCB* would be an excellent fit, given the legacy of publishing important analyses on this theme [e.g., @luyssaert_co_2007; @anderson-teixeira_carbon_2016; MORE]. We would like to be confident that the decision reflects a careful assessment of this manuscript's potential.

thank you,

Kristina Anderson-Teixeira (on behalf of all coauthors)

\newpage

Dear Dr. Anderson-Teixeira,

We write to you in regards to manuscript "Global patterns of forest autotrophic carbon fluxes". Based on points raised by the referees, the Subject Editor has denied publication of your paper in Global Change Biology.

As you may know, we decline a substantial proportion of manuscripts. These editorial judgments are based on such considerations as the degree of advance provided, the breadth of potential interest to researchers and timeliness.

Please be assured that this editorial decision does not represent a criticism of the quality of your work, nor are we questioning its value to others working in this area

Thank you for considering Global Change Biology for the publication of your research.  We hope the outcome of this specific submission will not discourage you from the submission of future manuscripts.

Sincerely,
Global Change Biology Editorial Office

Reviewer(s)' Comments to Author:

Reviewer: 1

Comments to the Author

Using 1,319 records from the Global Forest Carbon Database (ForC), authors of this manuscript explored the spatial pattern of several carbon flux. They found forest C fluxes decreases with absolute latitude. Among climatic variables, mean annual temperature and temperature seasonality were found to be the best univariate climatic predictors. They suggest their findings could improve understanding of forest carbon cycling under climate change. Overall, the manuscript is well written, but the findings seem not improving contemporary knowledge of forest carbon cycling, and the study appears to be a simple statistical analyses of literature data, which did not reveal the biological mechanisms lead to the detected pattern. Thus, I cannot recommend publishing this paper in GCB, but suggest the authors to consider a data journal like ESSD.

The analyses performed were too conventional that similar analyses with a smaller size of similar datasets have been performed a decade ago. Forest C fluxes decreasing with latitude was also not a new finding. The author suggest that it contradicts with some previous studies finding that “net primary productivity of temperate forests rivals that of tropical forests”. However, those high NPP in temperate forests could result from high rates of nitrogen deposition or forest plantation (Yu et al., 2014), which are in different context with the mature, undisturbed forests studied in this manuscript.

It is not surprising that MAT was found to be the good climate variables explaining forest C fluxes, since MAT has strong correlation with latitudes. Our contemporary knowledge on forest carbon cycling has moved well beyond using mean climatic variable to extrapolate or guess the response of forest carbon cycling to climate change. The spatial gradient do not necessarily represent the response of forests to climate change. The main implications highlighted in the Abstract is not well supported by the results presented.

While I commend the authors’ efforts to harmonize the dataset, I found the paper lacks the novelty and significance to be published in GCB. A data-oriented journal is a more suitable place for this paper.

Reviewer: 2

Comments to the Author

In the manuscript “Global patterns of forest autotrophic carbon fluxes”, Morgan et al. examined the latitudinal and climatic dependences of various autotrophic fluxes of global mature and undisturbed forests, using a recent release of the ForC database. They reported that latitude and temperature are the best univariate predictor of mature forest carbon cycling, while the effect of temperature is modified by precipitation. Overall, I appreciate this thorough analysis of the novel ForC database (great work in making data publicly available!) and see no major flaws in the study, thus most of my comments below are meant to further improve the manuscript. My comments are below:

1. How does the result compare with previous studies based on eddy covariance (EC) measurements? Throughout reading the manuscript, I have been trying to reconcile what’s been reported based on EC data and what is presented in the ms. Though EC cannot tell us much about NPP and other small fluxes, it did have a better (perhaps just marginally) spatial and temporal representation of forests. In particular, EC data has suggested that net carbon uptake is primarily dependent on carbon uptake period (Churkina et al. 2005, but see Fu et al. 2019 and Zhou et al. 2016), and increasing temperature can lengthen the growing season to enhance GPP and NEP (Keenan et al., 2014). The current study suggests that “growing season length is never the best predictor”. What do you think cause the difference between the two streams of evidence, and why MAT and latitude serves as a better predictor than growing season length? I do not ask for further analysis necessarily, but just feel the discussion could be expanded a bit on previous EC studies about the dominant control of forest carbon flux.

2. A key conclusion is “no detectable differences in allocation across latitude or climates”, however, I do not see a strong evidence for that presented in the main text – Fig.S3 does show several allocation indicators though no statistics are provided. Fig. 2 shows that C fluxes increase approximately in proportion to one another, but that does not necessarily mean allocation is unchanged across latitude or climates.

3. Implication for models. I think it is nice for the authors to report that carbon flux of mature forest is primarily dependent on MAT or latitude. But how much could this advanced knowledge help us improve global simulation of forest carbon cycling. I’d be very curious to see if the state-of-the-art models demonstrate a similar or different functional response of carbon flux to MAT/latitude, and what is the possible reason for that. I think by tapping into this question (even just in discussion) will make the study be of interest to a wider audience.

Minor comments:

1. it is unclear in the method about how to calculate “temperature/precipitation seasonality”. Is it the standard deviation of monthly temperature/precipitation in a year?

2. In Table S3, does the R2 come from the univariate linear model or the multivariate model?

3. Perhaps add R2 and p value in Fig. S3. Additionally, could you add the NPP to GPP ratio (CUE) as CUE is a quite often used allocation indicator.

4. L201 - To study the interactive effect of MAT and MAP, do you use MAT, MAP and MAT*MAP as three predictors in a multivariate model or just MAT*MAP in a linear model?

5. I am also wondering about the representation of tropical and extra-tropical forests in the samples. One solution is to add a biome basemap in figure 1. If tropical forest stands are generally older than the extra-tropical forest stands (according to “Histogram_of_Stand_age” in the ForC database), then perhaps there is a need to standardize samples by stand ages when studying allocation.

6. Fig. S6. Spell out “T seas” and “P seas” in the caption.

Churkina, G., Schimel, D., Braswell, B. H., & Xiao, X. (2005). Spatial analysis of growing season length control over net ecosystem exchange. Global Change Biology, 11(10), 1777–1787. https://nam02.safelinks.protection.outlook.com/?url=https%3A%2F%2Fdoi.org%2F10.1111%2Fj.1365-2486.2005.001012.x&data=02%7C01%7Cmullerh%40si.edu%7C95d13ecc18f5461cbe2408d858c4b0a9%7C989b5e2a14e44efe93b78cdd5fc5d11c%7C0%7C0%7C637356949086641203&sdata=fDBebq%2B1eqPsq2NDNe6zmEveEmfKtHpioWsoJn9yT%2BM%3D&reserved=0

Fu, Z., Stoy, P. C., Poulter, B., Gerken, T., Zhang, Z., Wakbulcho, G., & Niu, S. (2019). Maximum carbon uptake rate dominates the interannual variability of global net ecosystem exchange. Global Change Biology, gcb.14731. https://nam02.safelinks.protection.outlook.com/?url=https%3A%2F%2Fdoi.org%2F10.1111%2Fgcb.14731&data=02%7C01%7Cmullerh%40si.edu%7C95d13ecc18f5461cbe2408d858c4b0a9%7C989b5e2a14e44efe93b78cdd5fc5d11c%7C0%7C0%7C637356949086641203&sdata=aves%2BnxrjhGzFBzrXmb0c%2FXrEk0uku%2BwjHJJZqrF9v0%3D&reserved=0

Keenan, T. F., Gray, J., Friedl, M. A., Toomey, M., Bohrer, G., Hollinger, D. Y., … Richardson, A. D. (2014). Net carbon uptake has increased through warming-induced changes in temperate forest phenology. Nature Climate Change, 4(7), 598–604. https://nam02.safelinks.protection.outlook.com/?url=https%3A%2F%2Fdoi.org%2F10.1038%2Fnclimate2253&data=02%7C01%7Cmullerh%40si.edu%7C95d13ecc18f5461cbe2408d858c4b0a9%7C989b5e2a14e44efe93b78cdd5fc5d11c%7C0%7C0%7C637356949086641203&sdata=905OIXZQpS%2BRJ7xNVE1dakWAtdPrEl2gIk1PeTxWO%2Bk%3D&reserved=0

Zhou, S., Zhang, Y., Caylor, K. K., Luo, Y., Xiao, X., Ciais, P., … Wang, G. (2016). Explaining inter-annual variability of gross primary productivity from plant phenology and physiology. Agricultural and Forest Meteorology, 226–227(October), 246–256. https://nam02.safelinks.protection.outlook.com/?url=https%3A%2F%2Fdoi.org%2F10.1016%2Fj.agrformet.2016.06.010&data=02%7C01%7Cmullerh%40si.edu%7C95d13ecc18f5461cbe2408d858c4b0a9%7C989b5e2a14e44efe93b78cdd5fc5d11c%7C0%7C0%7C637356949086641203&sdata=UBxrEMtqh3Qo7%2B45xA9%2Fz6lqQIXD3LGvdMyVCNGOwSA%3D&reserved=0

\newpage
## References

### 2020-09-17 manuscript/journal_submissions/2020_8_GCB/response to reviews.Rmd
break out response to reviews from appeal letter, update appeal

---
title: "Appeal of rejection of GCB-20-1713"
output: pdf_document
bibliography: library.bib
csl: global-change-biology.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Dear Editor:

 "Global patterns of forest autotrophic carbon fluxes" (GCB-20-1713).

thank you,

Kristina Anderson-Teixeira (on behalf of all coauthors)

\newpage

Dear Dr. Anderson-Teixeira,

We write to you in regards to manuscript "Global patterns of forest autotrophic carbon fluxes". Based on points raised by the referees, the Subject Editor has denied publication of your paper in Global Change Biology.

As you may know, we decline a substantial proportion of manuscripts. These editorial judgments are based on such considerations as the degree of advance provided, the breadth of potential interest to researchers and timeliness.

Please be assured that this editorial decision does not represent a criticism of the quality of your work, nor are we questioning its value to others working in this area

Thank you for considering Global Change Biology for the publication of your research.  We hope the outcome of this specific submission will not discourage you from the submission of future manuscripts.

Sincerely,
Global Change Biology Editorial Office

Reviewer(s)' Comments to Author:

Reviewer: 1

Comments to the Author

Using 1,319 records from the Global Forest Carbon Database (ForC), authors of this manuscript explored the spatial pattern of several carbon flux. They found forest C fluxes decreases with absolute latitude. Among climatic variables, mean annual temperature and temperature seasonality were found to be the best univariate climatic predictors. They suggest their findings could improve understanding of forest carbon cycling under climate change. Overall, the manuscript is well written, but the findings seem not improving contemporary knowledge of forest carbon cycling, and the study appears to be a simple statistical analyses of literature data, which did not reveal the biological mechanisms lead to the detected pattern. Thus, I cannot recommend publishing this paper in GCB, but suggest the authors to consider a data journal like ESSD.

The analyses performed were too conventional that similar analyses with a smaller size of similar datasets have been performed a decade ago. Forest C fluxes decreasing with latitude was also not a new finding. The author suggest that it contradicts with some previous studies finding that “net primary productivity of temperate forests rivals that of tropical forests”. However, those high NPP in temperate forests could result from high rates of nitrogen deposition or forest plantation (Yu et al., 2014), which are in different context with the mature, undisturbed forests studied in this manuscript.

It is not surprising that MAT was found to be the good climate variables explaining forest C fluxes, since MAT has strong correlation with latitudes. Our contemporary knowledge on forest carbon cycling has moved well beyond using mean climatic variable to extrapolate or guess the response of forest carbon cycling to climate change. The spatial gradient do not necessarily represent the response of forests to climate change. The main implications highlighted in the Abstract is not well supported by the results presented.

While I commend the authors’ efforts to harmonize the dataset, I found the paper lacks the novelty and significance to be published in GCB. A data-oriented journal is a more suitable place for this paper.

Reviewer: 2

Comments to the Author

In the manuscript “Global patterns of forest autotrophic carbon fluxes”, Morgan et al. examined the latitudinal and climatic dependences of various autotrophic fluxes of global mature and undisturbed forests, using a recent release of the ForC database. They reported that latitude and temperature are the best univariate predictor of mature forest carbon cycling, while the effect of temperature is modified by precipitation. Overall, I appreciate this thorough analysis of the novel ForC database (great work in making data publicly available!) and see no major flaws in the study, thus most of my comments below are meant to further improve the manuscript. My comments are below:

1. How does the result compare with previous studies based on eddy covariance (EC) measurements? Throughout reading the manuscript, I have been trying to reconcile what’s been reported based on EC data and what is presented in the ms. Though EC cannot tell us much about NPP and other small fluxes, it did have a better (perhaps just marginally) spatial and temporal representation of forests. In particular, EC data has suggested that net carbon uptake is primarily dependent on carbon uptake period (Churkina et al. 2005, but see Fu et al. 2019 and Zhou et al. 2016), and increasing temperature can lengthen the growing season to enhance GPP and NEP (Keenan et al., 2014). The current study suggests that “growing season length is never the best predictor”. What do you think cause the difference between the two streams of evidence, and why MAT and latitude serves as a better predictor than growing season length? I do not ask for further analysis necessarily, but just feel the discussion could be expanded a bit on previous EC studies about the dominant control of forest carbon flux.

2. A key conclusion is “no detectable differences in allocation across latitude or climates”, however, I do not see a strong evidence for that presented in the main text – Fig.S3 does show several allocation indicators though no statistics are provided. Fig. 2 shows that C fluxes increase approximately in proportion to one another, but that does not necessarily mean allocation is unchanged across latitude or climates.

3. Implication for models. I think it is nice for the authors to report that carbon flux of mature forest is primarily dependent on MAT or latitude. But how much could this advanced knowledge help us improve global simulation of forest carbon cycling. I’d be very curious to see if the state-of-the-art models demonstrate a similar or different functional response of carbon flux to MAT/latitude, and what is the possible reason for that. I think by tapping into this question (even just in discussion) will make the study be of interest to a wider audience.

Minor comments:

1. it is unclear in the method about how to calculate “temperature/precipitation seasonality”. Is it the standard deviation of monthly temperature/precipitation in a year?

2. In Table S3, does the R2 come from the univariate linear model or the multivariate model?

3. Perhaps add R2 and p value in Fig. S3. Additionally, could you add the NPP to GPP ratio (CUE) as CUE is a quite often used allocation indicator.

4. L201 - To study the interactive effect of MAT and MAP, do you use MAT, MAP and MAT*MAP as three predictors in a multivariate model or just MAT*MAP in a linear model?

5. I am also wondering about the representation of tropical and extra-tropical forests in the samples. One solution is to add a biome basemap in figure 1. If tropical forest stands are generally older than the extra-tropical forest stands (according to “Histogram_of_Stand_age” in the ForC database), then perhaps there is a need to standardize samples by stand ages when studying allocation.

6. Fig. S6. Spell out “T seas” and “P seas” in the caption.

Churkina, G., Schimel, D., Braswell, B. H., & Xiao, X. (2005). Spatial analysis of growing season length control over net ecosystem exchange. Global Change Biology, 11(10), 1777–1787. https://nam02.safelinks.protection.outlook.com/?url=https%3A%2F%2Fdoi.org%2F10.1111%2Fj.1365-2486.2005.001012.x&data=02%7C01%7Cmullerh%40si.edu%7C95d13ecc18f5461cbe2408d858c4b0a9%7C989b5e2a14e44efe93b78cdd5fc5d11c%7C0%7C0%7C637356949086641203&sdata=fDBebq%2B1eqPsq2NDNe6zmEveEmfKtHpioWsoJn9yT%2BM%3D&reserved=0

Fu, Z., Stoy, P. C., Poulter, B., Gerken, T., Zhang, Z., Wakbulcho, G., & Niu, S. (2019). Maximum carbon uptake rate dominates the interannual variability of global net ecosystem exchange. Global Change Biology, gcb.14731. https://nam02.safelinks.protection.outlook.com/?url=https%3A%2F%2Fdoi.org%2F10.1111%2Fgcb.14731&data=02%7C01%7Cmullerh%40si.edu%7C95d13ecc18f5461cbe2408d858c4b0a9%7C989b5e2a14e44efe93b78cdd5fc5d11c%7C0%7C0%7C637356949086641203&sdata=aves%2BnxrjhGzFBzrXmb0c%2FXrEk0uku%2BwjHJJZqrF9v0%3D&reserved=0

Keenan, T. F., Gray, J., Friedl, M. A., Toomey, M., Bohrer, G., Hollinger, D. Y., … Richardson, A. D. (2014). Net carbon uptake has increased through warming-induced changes in temperate forest phenology. Nature Climate Change, 4(7), 598–604. https://nam02.safelinks.protection.outlook.com/?url=https%3A%2F%2Fdoi.org%2F10.1038%2Fnclimate2253&data=02%7C01%7Cmullerh%40si.edu%7C95d13ecc18f5461cbe2408d858c4b0a9%7C989b5e2a14e44efe93b78cdd5fc5d11c%7C0%7C0%7C637356949086641203&sdata=905OIXZQpS%2BRJ7xNVE1dakWAtdPrEl2gIk1PeTxWO%2Bk%3D&reserved=0

Zhou, S., Zhang, Y., Caylor, K. K., Luo, Y., Xiao, X., Ciais, P., … Wang, G. (2016). Explaining inter-annual variability of gross primary productivity from plant phenology and physiology. Agricultural and Forest Meteorology, 226–227(October), 246–256. https://nam02.safelinks.protection.outlook.com/?url=https%3A%2F%2Fdoi.org%2F10.1016%2Fj.agrformet.2016.06.010&data=02%7C01%7Cmullerh%40si.edu%7C95d13ecc18f5461cbe2408d858c4b0a9%7C989b5e2a14e44efe93b78cdd5fc5d11c%7C0%7C0%7C637356949086641203&sdata=UBxrEMtqh3Qo7%2B45xA9%2Fz6lqQIXD3LGvdMyVCNGOwSA%3D&reserved=0

\newpage
## References

### 2020-09-17 manuscript/journal_submissions/2020_8_GCB/appeal.Rmd
break out response to reviews from appeal letter, update appeal

@@
 
-We would like to appeal the reject decision of our manuscript "Global patterns of forest autotrophic carbon fluxes" (GCB-20-1713).
+We would like to appeal the reject decision of our manuscript "Global patterns of forest autotrophic carbon fluxes" (GCB-20-1713) on the grounds that Reviewer 1 (of 2) was off-base in (1) providing an inaccurate assessment of the scope of our analysis relative to previous studies, (2) under-valuing high-quality synthesis that produces unsurprising results, (3) setting expectations for this type of analysis that are out of line with the state of the field, and (4) recommending a completely inappropriate alternate venue.
+We detail these objections below:
 
-As far as we can tell from the comments to the authors, reviewer 2 was supportive, and her/his comments could easily be addressed.
-We therefore assume the reject decision was based on the feedback of reviewer 1, which we believe was off-base for several reasons:
+**(1) Reviewer 1 provided an inaccurate assessment of the scope of our analysis relative to previous studies.**  The reviewer states, "...the findings seem not improving contemporary knowledge of forest carbon cycling, and the study appears to be a simple statistical analyses of literature data..." and "the analyses performed were too conventional that similar analyses with a smaller size of similar datasets have been performed a decade ago." These statements betray a lack of perspective on the scope of our analysis relative to previous studies.
+Whereas most studies consider at most a few flux variables, our considers nine.
+The most comprehensive similar study is @luyssaert_co_2007, which was published in *GCB* more than a decade ago, includes much less data, does not control for effects of stand age/ disturbance history, and examines latitudinal/ global climatic trends in only **#** variables.
+This study has been cited >580 times and was highlighted as a noteworthy publication within *GCB* ([Long, 2020](https://doi.org/10.1111/gcb.14921)).
+It would seem that something along these lines, but with an expanded database and a more carefully controlled treatment of forest age and management/disturbance history, would have high potential to become a "classic" reference in global forest carbon cycling.
 
-* The reviewer states, "...the findings seem not improving contemporary knowledge of forest carbon cycling, and the study appears to be a simple statistical analyses of literature data..." and "the analyses performed were too conventional that similar analyses with a smaller size of similar datasets have been performed a decade ago." These statements betray a lack of perspective on the scope of our analysis relative to previous studies.
-Whereas most studies consider at most a few flux variables, our considers nine.
-The most comprehensive similar study is @luyssaert_co_2007, which was published in *GCB* more than a decade ago, includes much less data, does not control for effects of stand age/ disturbance history, and examines latitudinal/ global climatic trends in only # variables.
-This study has been cited >580 times and was highlighted as a noteworthy publication within *GCB* ([Long, 2020](https://doi.org/10.1111/gcb.14921)).
-It would seem that something along these lines, but with an expanded database and a more carefully controlled analysis, would have high potential to become a "classic" reference in global forest carbon cycling.
-
-* The reviewer objects that the results are not surprising--e.g., that it is not a new finding that C flux declines with latitude and increases with MAT.
+**(2) We feel that Reviewer 1 under-values high-quality synthesis that produces unsurprising results.** The reviewer objects that the results are not surprising--e.g., that it is not a new finding that C flux declines with latitude and increases with mean annual temperature.
 We agree that the results are not surprising, and we recognize that surprising results tend to produce high-impact publications.
@@
 
-* The reviewer objects to the fact that the study does not "reveal the biological mechanisms lead to the detected pattern", and states, "Our contemporary knowledge on forest carbon cycling has moved well beyond using mean climatic variable to extrapolate or guess the response of forest carbon cycling to climate change.
+**(3) Reviewer 1 set an unrealistic expectation that an analysis of this nature would reveal mechanisms, which is out of line with the state of the field.**  The reviewer objects to the fact that the study does not "reveal the biological mechanisms lead to the detected pattern", and states, "Our contemporary knowledge on forest carbon cycling has moved well beyond using mean climatic variable to extrapolate or guess the response of forest carbon cycling to climate change.
 The spatial gradient do not necessarily represent the response of forests to climate change." We agree, and fully acknowledge in the paper, that this study does not reveal the biological mechanisms.
 Further, we do not argue that the spatial gradient represents the response of forests to climate change, as clearly discussed in the final paragraph of the discussion.
-However, we do argue that broad-scale patterns in C cycling across climatic gradients remain--and should remain--one of the important approaches for inferring how forests might respond to climate change.
+However, we do argue that broad-scale patterns in C cycling across climatic gradients remain--and should remain--one of several important, complementary approaches to addressing a challenging problem [*sensu* Anderson-Teixeira et al.
+2013, GCB].
 We note that studies using this approach have been published this year in *Science* [@sullivan_long-term_2020]...
-*(add some more examples)*.
-While we fully recognize the limitations to such studies for inferring how forests will respond to climate change, we argue that this approach is one of several important, complementary approaches to addressing a challenging problem (Anderson-Teixeira et al.
-2013, GCB).
+**(add some more examples)**.
 The reviewer's apparent wholesale rejection of this approach seems short-sighted, and the expectation that a single study attempt to explain mechanisms behind global patterns in nine flux variables seems naïve.
-Getting at mechanisms (well) would require integration of ForC with modeling efforts on a scale that is not currently feasible, but lies within the vision of leaders in the modeling community, as stated in an opinion piece that I recently reviewed a paper for *GCB* (give citation/ number).
-(This piece lays out a vision for improvement of community infrastructure for modeling, including one *stated the goal of linking with leading databases, specifically citing ForC*.)
+The implied standard that studies of this type should "reveal the biological mechanisms" could be interpreted in various ways, but the best way of getting at underlying biological mechanisms would be through modeling.
+Integration of ForC with mechanistic ecosystem models has yet to be acheived, but lies within the vision of leaders in the modeling community, as stated in an opinion piece that I recently reviewed a paper for *GCB* (give citation/ number).
+(This piece lays out a vision for improvement of community infrastructure for modeling, including one *stated the goal of linking with leading databases, specifically citing ForC*.) In the meantime, the current analysis will be at the clear forefront of studies examinging correlations between forest carbon flux and climate at the global scale.
 
-* The reviewer suggested (twice) that we "consider a data journal like ESSD".
+**(4) Reviewer 1's suggestion that the work be published in a data journal betrays carelessness of review and/or limited understanding of appropriate publishing standards.** The reviewer suggested (twice) that we "consider a data journal like ESSD".
 This suggestion is completely inappropriate for two reasons.
 First, the hypothesis-testing analyses presented here would be completely inappropriate for a data journal.
-Second, the ForC database is already published [tropical portion only in @anderson-teixeira_carbon_2016, *GCB*; full database as a data paper in @anderson-teixeira_forc_2018, *Ecology*].
+Second, the ForC database is already published [tropical portion only in @anderson-teixeira_carbon_2016, *GCB*; full database as a data paper in Anderson-Teixeira et al.
+@anderson-teixeira_forc_2018, *Ecology*].
 Many qualified reviewers would already be aware of this, and an even semi-careful review would not miss this fact, which was prominently cited in the last paragraph of the introduction and the first sentence of the methods.
@@
 \newpage
-
-Dear Dr. Anderson-Teixeira,
-
-We write to you in regards to manuscript "Global patterns of forest autotrophic carbon fluxes".
-Based on points raised by the referees, the Subject Editor has denied publication of your paper in Global Change Biology.
-
-As you may know, we decline a substantial proportion of manuscripts.
-These editorial judgments are based on such considerations as the degree of advance provided, the breadth of potential interest to researchers and timeliness.
-
-Please be assured that this editorial decision does not represent a criticism of the quality of your work, nor are we questioning its value to others working in this area
-
-Thank you for considering Global Change Biology for the publication of your research.
-We hope the outcome of this specific submission will not discourage you from the submission of future manuscripts.
-
-Sincerely,
-Global Change Biology Editorial Office
-
-Reviewer(s)' Comments to Author:
-
-Reviewer: 1
-
-Comments to the Author
-
-Using 1,319 records from the Global Forest Carbon Database (ForC), authors of this manuscript explored the spatial pattern of several carbon flux.
-They found forest C fluxes decreases with absolute latitude.
-Among climatic variables, mean annual temperature and temperature seasonality were found to be the best univariate climatic predictors.
-They suggest their findings could improve understanding of forest carbon cycling under climate change.
-Overall, the manuscript is well written, but the findings seem not improving contemporary knowledge of forest carbon cycling, and the study appears to be a simple statistical analyses of literature data, which did not reveal the biological mechanisms lead to the detected pattern.
-Thus, I cannot recommend publishing this paper in GCB, but suggest the authors to consider a data journal like ESSD.
-
-The analyses performed were too conventional that similar analyses with a smaller size of similar datasets have been performed a decade ago.
-Forest C fluxes decreasing with latitude was also not a new finding.
-The author suggest that it contradicts with some previous studies finding that “net primary productivity of temperate forests rivals that of tropical forests”.
-However, those high NPP in temperate forests could result from high rates of nitrogen deposition or forest plantation (Yu et al., 2014), which are in different context with the mature, undisturbed forests studied in this manuscript.
-
-It is not surprising that MAT was found to be the good climate variables explaining forest C fluxes, since MAT has strong correlation with latitudes.
-Our contemporary knowledge on forest carbon cycling has moved well beyond using mean climatic variable to extrapolate or guess the response of forest carbon cycling to climate change.
-The spatial gradient do not necessarily represent the response of forests to climate change.
-The main implications highlighted in the Abstract is not well supported by the results presented.
-
-While I commend the authors’ efforts to harmonize the dataset, I found the paper lacks the novelty and significance to be published in GCB.
-A data-oriented journal is a more suitable place for this paper.
-
-Reviewer: 2
-
-Comments to the Author
-
-In the manuscript “Global patterns of forest autotrophic carbon fluxes”, Morgan et al.
-examined the latitudinal and climatic dependences of various autotrophic fluxes of global mature and undisturbed forests, using a recent release of the ForC database.
-They reported that latitude and temperature are the best univariate predictor of mature forest carbon cycling, while the effect of temperature is modified by precipitation.
-Overall, I appreciate this thorough analysis of the novel ForC database (great work in making data publicly available!) and see no major flaws in the study, thus most of my comments below are meant to further improve the manuscript.
-My comments are below:
-
-1.
-How does the result compare with previous studies based on eddy covariance (EC) measurements?
-Throughout reading the manuscript, I have been trying to reconcile what’s been reported based on EC data and what is presented in the ms.
-Though EC cannot tell us much about NPP and other small fluxes, it did have a better (perhaps just marginally) spatial and temporal representation of forests.
-In particular, EC data has suggested that net carbon uptake is primarily dependent on carbon uptake period (Churkina et al.
-2005, but see Fu et al.
-2019 and Zhou et al.
-2016), and increasing temperature can lengthen the growing season to enhance GPP and NEP (Keenan et al., 2014).
-The current study suggests that “growing season length is never the best predictor”.
-What do you think cause the difference between the two streams of evidence, and why MAT and latitude serves as a better predictor than growing season length?
-I do not ask for further analysis necessarily, but just feel the discussion could be expanded a bit on previous EC studies about the dominant control of forest carbon flux.
-
-2.
-A key conclusion is “no detectable differences in allocation across latitude or climates”, however, I do not see a strong evidence for that presented in the main text – Fig.S3 does show several allocation indicators though no statistics are provided.
-Fig.
-2 shows that C fluxes increase approximately in proportion to one another, but that does not necessarily mean allocation is unchanged across latitude or climates.
-
-3.
-Implication for models.
-I think it is nice for the authors to report that carbon flux of mature forest is primarily dependent on MAT or latitude.
-But how much could this advanced knowledge help us improve global simulation of forest carbon cycling.
-I’d be very curious to see if the state-of-the-art models demonstrate a similar or different functional response of carbon flux to MAT/latitude, and what is the possible reason for that.
-I think by tapping into this question (even just in discussion) will make the study be of interest to a wider audience.
-
-Minor comments:
-
-1. it is unclear in the method about how to calculate “temperature/precipitation seasonality”. Is it the standard deviation of monthly temperature/precipitation in a year?
-
-2. In Table S3, does the R2 come from the univariate linear model or the multivariate model?
-
-3. Perhaps add R2 and p value in Fig. S3. Additionally, could you add the NPP to GPP ratio (CUE) as CUE is a quite often used allocation indicator.
-
-4. L201 - To study the interactive effect of MAT and MAP, do you use MAT, MAP and MAT*MAP as three predictors in a multivariate model or just MAT*MAP in a linear model?
-
-5.
-I am also wondering about the representation of tropical and extra-tropical forests in the samples.
-One solution is to add a biome basemap in figure 1.
-If tropical forest stands are generally older than the extra-tropical forest stands (according to “Histogram_of_Stand_age” in the ForC database), then perhaps there is a need to standardize samples by stand ages when studying allocation.
-
-6. Fig. S6. Spell out “T seas” and “P seas” in the caption.
-
-Churkina, G., Schimel, D., Braswell, B.
-H., & Xiao, X.
-(2005).
-Spatial analysis of growing season length control over net ecosystem exchange.
-Global Change Biology, 11(10), 1777–1787.
-https://nam02.safelinks.protection.outlook.com/?url=https%3A%2F%2Fdoi.org%2F10.1111%2Fj.1365-2486.2005.001012.x&data=02%7C01%7Cmullerh%40si.edu%7C95d13ecc18f5461cbe2408d858c4b0a9%7C989b5e2a14e44efe93b78cdd5fc5d11c%7C0%7C0%7C637356949086641203&sdata=fDBebq%2B1eqPsq2NDNe6zmEveEmfKtHpioWsoJn9yT%2BM%3D&reserved=0
-
-Fu, Z., Stoy, P.
-C., Poulter, B., Gerken, T., Zhang, Z., Wakbulcho, G., & Niu, S.
-(2019).
-Maximum carbon uptake rate dominates the interannual variability of global net ecosystem exchange.
-Global Change Biology, gcb.14731.
-https://nam02.safelinks.protection.outlook.com/?url=https%3A%2F%2Fdoi.org%2F10.1111%2Fgcb.14731&data=02%7C01%7Cmullerh%40si.edu%7C95d13ecc18f5461cbe2408d858c4b0a9%7C989b5e2a14e44efe93b78cdd5fc5d11c%7C0%7C0%7C637356949086641203&sdata=aves%2BnxrjhGzFBzrXmb0c%2FXrEk0uku%2BwjHJJZqrF9v0%3D&reserved=0
-
-Keenan, T.
-F., Gray, J., Friedl, M.
-A., Toomey, M., Bohrer, G., Hollinger, D.
-Y., … Richardson, A.
-D.
-(2014).
-Net carbon uptake has increased through warming-induced changes in temperate forest phenology.
-Nature Climate Change, 4(7), 598–604.
-https://nam02.safelinks.protection.outlook.com/?url=https%3A%2F%2Fdoi.org%2F10.1038%2Fnclimate2253&data=02%7C01%7Cmullerh%40si.edu%7C95d13ecc18f5461cbe2408d858c4b0a9%7C989b5e2a14e44efe93b78cdd5fc5d11c%7C0%7C0%7C637356949086641203&sdata=905OIXZQpS%2BRJ7xNVE1dakWAtdPrEl2gIk1PeTxWO%2Bk%3D&reserved=0
-
-Zhou, S., Zhang, Y., Caylor, K.
-K., Luo, Y., Xiao, X., Ciais, P., … Wang, G.
-(2016).
-Explaining inter-annual variability of gross primary productivity from plant phenology and physiology.
-Agricultural and Forest Meteorology, 226–227(October), 246–256.
-https://nam02.safelinks.protection.outlook.com/?url=https%3A%2F%2Fdoi.org%2F10.1016%2Fj.agrformet.2016.06.010&data=02%7C01%7Cmullerh%40si.edu%7C95d13ecc18f5461cbe2408d858c4b0a9%7C989b5e2a14e44efe93b78cdd5fc5d11c%7C0%7C0%7C637356949086641203&sdata=UBxrEMtqh3Qo7%2B45xA9%2Fz6lqQIXD3LGvdMyVCNGOwSA%3D&reserved=0
-
-\newpage
 ## References

### 2020-09-17 manuscript/journal_submissions/2020_8_GCB/appeal.Rmd
Update appeal.Rmd

@@
 Whereas most studies consider at most a few flux variables, our considers nine.
-The most comprehensive similar study is @luyssaert_co_2007, which was published in *GCB* more than a decade ago, includes much less data, does not control for effects of stand age/ disturbance history, and examines latitudinal/ global climatic trends in only **#** variables.
+The most comprehensive similar study is @luyssaert_co_2007, which was published in *GCB* more than a decade ago, includes much less data, does not control for effects of stand age/ disturbance history, and examines latitudinal/ global climatic trends in only 3 variables.
 This study has been cited >580 times and was highlighted as a noteworthy publication within *GCB* ([Long, 2020](https://doi.org/10.1111/gcb.14921)).

### 2020-09-17 manuscript/journal_submissions/2020_8_GCB/appeal.Rmd
generated word version, minor edit

@@
 title: "Appeal of rejection of GCB-20-1713"
-output: pdf_document
+output: word_document
 bibliography: library.bib
@@
 We would like to appeal the reject decision of our manuscript "Global patterns of forest autotrophic carbon fluxes" (GCB-20-1713) on the grounds that Reviewer 1 (of 2) was off-base in (1) providing an inaccurate assessment of the scope of our analysis relative to previous studies, (2) under-valuing high-quality synthesis that produces unsurprising results, (3) setting expectations for this type of analysis that are out of line with the state of the field, and (4) recommending a completely inappropriate alternate venue.
-We detail these objections below:
+We detail these objections below.
+We note that Reviewer 2 was supportive, and her/his comments could easily be addressed.
 

### 2020-09-18 manuscript/journal_submissions/2020_8_GCB/appeal.Rmd
integrated references

@@
 The most comprehensive similar study is @luyssaert_co_2007, which was published in *GCB* more than a decade ago, includes much less data, does not control for effects of stand age/ disturbance history, and examines latitudinal/ global climatic trends in only 3 variables.
-This study has been cited >580 times and was highlighted as a noteworthy publication within *GCB* ([Long, 2020](https://doi.org/10.1111/gcb.14921)).
+This study has been cited >580 times and was highlighted as a noteworthy publication within *GCB* [@long_twentyfive_2020].
 It would seem that something along these lines, but with an expanded database and a more carefully controlled treatment of forest age and management/disturbance history, would have high potential to become a "classic" reference in global forest carbon cycling.
@@
 2013, GCB].
-We note that studies using this approach have been published this year in *Science* [@sullivan_long-term_2020]...
-**(add some more examples)**.
+We note that studies using this approach have been published this year in *Science* [@sullivan_long-term_2020], and recently in *GCB* [@trugman_climate_2019], *JGR* [@li_dissimilatory_2019], and *Scientific Reports* [@chen_spatial_2019].
 The reviewer's apparent wholesale rejection of this approach seems short-sighted, and the expectation that a single study attempt to explain mechanisms behind global patterns in nine flux variables seems naïve.

### 2020-09-18 manuscript/journal_submissions/2020_8_GCB/appeal.Rmd
finished appeal

--- 2020-09-17 manuscript/journal_submissions/2020_8_GCB/appeal.Rmd
@@
 
-We would like to appeal the reject decision of our manuscript "Global patterns of forest autotrophic carbon fluxes" (GCB-20-1713) on the grounds that Reviewer 1 (of 2) was off-base in (1) providing an inaccurate assessment of the scope of our analysis relative to previous studies, (2) under-valuing high-quality synthesis that produces unsurprising results, (3) setting expectations for this type of analysis that are out of line with the state of the field, and (4) recommending a completely inappropriate alternate venue.
-We detail these objections below:
+We would like to appeal the reject decision of our manuscript "Global patterns of forest autotrophic carbon fluxes" (GCB-20-1713) on the grounds that Reviewer 1 (1) provided an inaccurate assessment of the scope of our analysis relative to previous studies, (2) under-valued high-quality synthesis, (3) set expectations for this type of analysis that are out of line with the state of the field, and (4) recommended a completely inappropriate alternate venue.
+We detail these objections below.
+We note that Reviewer 2 was supportive, and her/his comments could easily be addressed.
 
-**(1) Reviewer 1 provided an inaccurate assessment of the scope of our analysis relative to previous studies.**  The reviewer states, "...the findings seem not improving contemporary knowledge of forest carbon cycling, and the study appears to be a simple statistical analyses of literature data..." and "the analyses performed were too conventional that similar analyses with a smaller size of similar datasets have been performed a decade ago." These statements betray a lack of perspective on the scope of our analysis relative to previous studies.
-Whereas most studies consider at most a few flux variables, our considers nine.
-The most comprehensive similar study is @luyssaert_co_2007, which was published in *GCB* more than a decade ago, includes much less data, does not control for effects of stand age/ disturbance history, and examines latitudinal/ global climatic trends in only 3 variables.
-This study has been cited >580 times and was highlighted as a noteworthy publication within *GCB* ([Long, 2020](https://doi.org/10.1111/gcb.14921)).
-It would seem that something along these lines, but with an expanded database and a more carefully controlled treatment of forest age and management/disturbance history, would have high potential to become a "classic" reference in global forest carbon cycling.
+**(1) Reviewer 1 provided an inaccurate assessment of the scope of our analysis relative to previous studies.**  The reviewer states, "the findings seem not improving contemporary knowledge of forest carbon cycling, and the study appears to be a simple statistical analyses of literature data" and "the analyses performed were too conventional that similar analyses with a smaller size of similar datasets have been performed a decade ago." These statements betray a lack of perspective on the scope of our analysis relative to previous studies.
+Whereas most studies consider at most a few flux variables, ours considers nine.
+The most comprehensive similar study is @luyssaert_co_2007, which was published in *GCB* more than a decade ago, was based on a database <25% the size of the ForC version used here, did not control for effects of stand age/ disturbance history, and examined latitudinal/ global climatic trends in only three variables.
+Thus, the current analysis will be at the clear forefront of studies examining correlations between forest carbon flux and climate at the global scale.
 
-**(2) We feel that Reviewer 1 under-values high-quality synthesis that produces unsurprising results.** The reviewer objects that the results are not surprising--e.g., that it is not a new finding that C flux declines with latitude and increases with mean annual temperature.
-We agree that the results are not surprising, and we recognize that surprising results tend to produce high-impact publications.
-However, there is also a place for high-quality synthesis that clarifies current understanding, which is the goal of this publication.
+**(2) We feel that Reviewer 1 under-valued high-quality synthesis that produces unsurprising results.** The reviewer objects that the results are not surprising--e.g., that it is not a new finding that C flux declines with latitude and increases with mean annual temperature.
+We agree that the results are not surprising, but a crucial part of science is built on high-quality synthesis that clarifies current understanding and probes for unexpected relationships, which is our goal.
 Such syntheses are often published as review papers, but given that our analysis entails original analysis, we submitted it as a Primary Research Article.
@@
 
-**(3) Reviewer 1 set an unrealistic expectation that an analysis of this nature would reveal mechanisms, which is out of line with the state of the field.**  The reviewer objects to the fact that the study does not "reveal the biological mechanisms lead to the detected pattern", and states, "Our contemporary knowledge on forest carbon cycling has moved well beyond using mean climatic variable to extrapolate or guess the response of forest carbon cycling to climate change.
+**(3) Reviewer 1 set an unrealistic expectation that an analysis of this nature would reveal mechanisms, which is out of line with the state of the field.**  The reviewer objects that our study does not "reveal the biological mechanisms lead to the detected pattern", stating, "Our contemporary knowledge on forest carbon cycling has moved well beyond using mean climatic variable to extrapolate or guess the response of forest carbon cycling to climate change.
 The spatial gradient do not necessarily represent the response of forests to climate change." We agree, and fully acknowledge in the paper, that this study does not reveal the biological mechanisms.
 Further, we do not argue that the spatial gradient represents the response of forests to climate change, as clearly discussed in the final paragraph of the discussion.
-However, we do argue that broad-scale patterns in C cycling across climatic gradients remain--and should remain--one of several important, complementary approaches to addressing a challenging problem [*sensu* Anderson-Teixeira et al.
-2013, GCB].
-We note that studies using this approach have been published this year in *Science* [@sullivan_long-term_2020]...
-**(add some more examples)**.
+However, we do argue that elucidating broad-scale patterns in C cycling across climatic gradients is a complementary approach to addressing a challenging problem [*sensu* @andersonteixeira_altered_2013, *GCB*].
+We note that studies using this approach have been published recently in *Science* [@sullivan_long-term_2020],  *GCB* [@trugman_climate_2019], *JGR* [@li_dissimilatory_2019], and *Scientific Reports* [@chen_spatial_2019].
+
 The reviewer's apparent wholesale rejection of this approach seems short-sighted, and the expectation that a single study attempt to explain mechanisms behind global patterns in nine flux variables seems naïve.
 The implied standard that studies of this type should "reveal the biological mechanisms" could be interpreted in various ways, but the best way of getting at underlying biological mechanisms would be through modeling.
-Integration of ForC with mechanistic ecosystem models has yet to be acheived, but lies within the vision of leaders in the modeling community, as stated in an opinion piece that I recently reviewed a paper for *GCB* (give citation/ number).
-(This piece lays out a vision for improvement of community infrastructure for modeling, including one *stated the goal of linking with leading databases, specifically citing ForC*.) In the meantime, the current analysis will be at the clear forefront of studies examinging correlations between forest carbon flux and climate at the global scale.
+Until databases such as ForC are fully integrated with mechanistic ecosystem models, our ability to provide truly mechanistic explanations of the observed patterns remains limited, but the observed correlations provide important insights into the broad-scale constraints of climate on forest carbon cycling.
 
@@
 @anderson-teixeira_forc_2018, *Ecology*].
-Many qualified reviewers would already be aware of this, and an even semi-careful review would not miss this fact, which was prominently cited in the last paragraph of the introduction and the first sentence of the methods.
+Even a semi-careful review would not have missed this fact, which was prominently cited in the last paragraph of the introduction and the first sentence of the methods.
 Thus, this suggestion betrays carelessness of review and/or limited understanding of appropriate standards for judging where scientific work should be published.
@@
 
-If, despite these arguments, the *GCB* Editorial Board does not feel that the paper is sufficiently competitive for publication in *GCB*, we will accept the decision and publish elsewhere.
-However, we believe that this manuscript has potential to become a "classic" reference on the subject of carbon cycling in forests globally, and that *GCB* would be an excellent fit, given the legacy of publishing important analyses on this theme [e.g., @luyssaert_co_2007; @anderson-teixeira_carbon_2016; MORE].
+We continue to believe that this manuscript has potential to become a "classic" reference on the subject of carbon cycling in forests globally, and that *GCB* is an excellent fit, given the legacy of publishing important analyses on this theme [e.g., @luyssaert_co_2007; @anderson-teixeira_carbon_2016].
 We would like to be confident that the decision reflects a careful assessment of this manuscript's potential.
 
-thank you,
+Thank you,
 

### 2020-09-25 manuscript/response to reviews.Rmd
add response to reviews

--- 2020-09-17 manuscript/journal_submissions/2020_8_GCB/response to reviews.Rmd
@@
 
+.....
+
 thank you,
@@
 \newpage
+
+Dear Kristina,
+
+Your appeal regarding the decision on your manuscript GCB-20-1713 has been safely received.
+The editor is in principle willing to consider a revised version of your manuscript.
+If you do consider to submit a revised manuscript, we want to stress the importance of fully addressing all the comments raised by both reviewers, including those concerns about the novelty and the lack of mechanistic advancement of your paper.
+Your revised manuscript, along with the point-to-point response to reviewer comments, will be sent out for evaluation.
+We will make a decision based on reviewers' assessment, and no acceptance can be guaranteed.
+
+Please let me know if this is agreeable to you.
+
+All the best,
+
+Rhea Bruno
+
+~~~
 
@@
 Global Change Biology Editorial Office
+
+\newpage
 

### 2020-09-28 manuscript/appendix.Rmd
update table and caption

@@
 S5 <- read.csv("tables_figures/mat_map.csv", stringsAsFactors = FALSE, check.names = FALSE)
-kable(S5, format = "latex", booktabs = TRUE, caption = "Table S3. Signficance of relationships of forest C fluxes to MAT alone and in combination with MAP", escape = F)%>%
+kable(S5, format = "latex", booktabs = TRUE, caption = "Table S3.
+Signficance of relationships of forest C fluxes to MAT alone and in combination with MAP.
+$R^{2}$ values refer to the best model.", escape = F)%>%
   kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 12) %>%
@@
   column_spec(4, width = "6cm")%>%
-  column_spec(5, width = "2cm")%>%
+  column_spec(5, width = "4cm")%>%
   column_spec(6, width = "2cm")

### 2020-09-29 manuscript/response to reviews.Rmd
Update response to reviews.Rmd
Update response to reviews.Rmd
added points from appeal letter

@@
 The analyses performed were too conventional that similar analyses with a smaller size of similar datasets have been performed a decade ago.
+
+*This statement suggests a lack of perspective on the scope of our analysis relative to previous studies.
+Whereas most studies consider at most a few flux variables, ours considers nine.The most comprehensive similar study is @luyssaert_co_2007, which was published in *GCB* more than a decade ago, was based on a database <25% the size of the ForC version used here, did not control for effects of stand age/ disturbance history, and examined latitudinal/ global climatic trends in only three variables.
+Thus, the current analysis will be at the clear forefront of studies examining correlations between forest carbon flux and climate at the global scale.*
+
 Forest C fluxes decreasing with latitude was also not a new finding.
@@
 However, those high NPP in temperate forests could result from high rates of nitrogen deposition or forest plantation (Yu et al., 2014), which are in different context with the mature, undisturbed forests studied in this manuscript.
+
+*We agree that these results are not surprising, however a crucial part of sciencne is built on high-quality synthesis that clarifies current understanding and probes for unexpected relationships.
+This review under-values high-quality synthesis that produces unsurprising results.*
 
@@
 
+*We agree, and fully acknowledge in the paper, that this study does not reveal the biological mechanisms.
+Further, we do not argue that the spatial gradient represents the response of forests to climate change, as clearly discussed in the final paragraph of the discussion.
+However, we do argue that elucidating broad-scale patterns in C cycling across climatic gradients is a complementary approach to addressing a challenging problem [*sensu* @andersonteixeira_altered_2013, *GCB*].
+We note that studies using this approach have been published recently in *Science* [@sullivan_long-term_2020],  *GCB* [@trugman_climate_2019], *JGR* [@li_dissimilatory_2019], and *Scientific Reports* [@chen_spatial_2019].*
+
+*The reviewer's apparent wholesale rejection of this approach seems short-sighted, and the expectation that a single study attempt to explain mechanisms behind global patterns in nine flux variables seems naïve.
+The implied standard that studies of this type should "reveal the biological mechanisms" could be interpreted in various ways, but the best way of getting at underlying biological mechanisms would be through modeling.
+Until databases such as ForC are fully integrated with mechanistic ecosystem models, our ability to provide truly mechanistic explanations of the observed patterns remains limited, but the observed correlations provide important insights into the broad-scale constraints of climate on forest carbon cycling.*
+
 While I commend the authors’ efforts to harmonize the dataset, I found the paper lacks the novelty and significance to be published in GCB.
 A data-oriented journal is a more suitable place for this paper.
+
+*This suggestion indicates carelessness of review and/or limited understanding of appropriate publishing standards.
+The suggestion is completely inappropriate for two reasons.First, the hypothesis-testing analyses presented here would be unsuitable for a data journal.
+Second, the ForC database is already published [tropical portion only in @anderson-teixeira_carbon_2016, *GCB*; full database as a data paper in Anderson-Teixeira et al.
+@anderson-teixeira_forc_2018, *Ecology*].
+Even a semi-careful review would not have missed this fact, which was prominently cited in the last paragraph of the introduction and the first sentence of the methods.
+Thus, this suggestion betrays carelessness of review and/or limited understanding of appropriate standards for judging where scientific work should be published.*
 
@@
 
+*Stated in methods section that this information is available in table S1*
+
 2. In Table S3, does the R2 come from the univariate linear model or the multivariate model?
+
+*Clarifed in table*
 

### 2020-09-29 manuscript/response to reviews.Rmd
clarified MATxMAP models

@@
 
+*Potentially we need a response to the "simple statistical analyses" point?
+- these analyses are valuable for clarifying relationships between carbon fluxes and climate.
+Although analyses are simple, they are powerful thanks to the coverage and extent of the dataset, and the range of fluxes examined.*
+
 The analyses performed were too conventional that similar analyses with a smaller size of similar datasets have been performed a decade ago.
@@
 
-*We agree that these results are not surprising, however a crucial part of sciencne is built on high-quality synthesis that clarifies current understanding and probes for unexpected relationships.
+*We agree that these results are not surprising, however a crucial part of science is built on high-quality synthesis that clarifies current understanding and probes for unexpected relationships.
 This review under-values high-quality synthesis that produces unsurprising results.*
+
+*As the review identifies, results from previous studies were complicated by other effects.
+The power of this synthesis is in how it standardises for the effects of stand age and disturbance history, unambiguously clarifying these relationships.*
 
@@
 
-4. L201 - To study the interactive effect of MAT and MAP, do you use MAT, MAP and MAT*MAP as three predictors in a multivariate model or just MAT*MAP in a linear model?
+*Added CUE and legend to show none of the fits are significant (because of this no statistics presented)*
+
+4. L201 - To study the interactive effect of MAT and MAP, do you use MAT, MAP and MATxMAP as three predictors in a multivariate model or just MATxMAP in a linear model?
+
+*We use MATxMAP in a linear model - clarified this in the text*
 

### 2020-09-29 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd
clarified MATxMAP models

@@
 Based on the geographic co-ordinates for each site, data on twelve climate variables -- including $MAT$, $MAP$, temperature and precipitation seasonality, annual temperature range, solar radiation, cloud cover, annual frost and wet days, potential evapotranspiration ($PET$), aridity ($MAP/PET$), and vapor pressure deficit ($VPD$) -- were extracted from five open-access climate datasets: WorldClim [@hijmans_very_2005], WorldClim2 [@fick_worldclim_2017], the Climate Research Unit time-series dataset (CRU TS v4.03 [@harris_updated_2014], the Global Aridity Index and Potential Evapotranspiration Climate Database [@trabucco_global_2019], and TerraClimate [@abatzoglou_terraclimate_2018] (Table S1).
+Definitions and methods used to calculate each variable are included in Table S1.
 From these data, we derived maximum $VPD$, defined as the $VPD$ of the month with the largest deficit, and the number of water stress months, defined as the number of months annually where precipitation was lower than $PET$.
@@
 
-Multivariate models were used to investigate the potential joint and interactive effects of $MAT$ and $MAP$ on carbon fluxes.
+Linear models were used to investigate the potential joint and interactive effects of $MAT$ and $MAP$ on carbon fluxes.
 An additive model including $MAP$ in addition to $MAT$ was accepted when $\Delta$AIC >2 relative to a null including only $MAT$ as a fixed effect.
-An interactive model including an $MAT$ x $MAP$ interaction was accepted when $\Delta$AIC >2 relative to a null including $MAT$ and $MAP$ as fixed effects.
+An interactive model containing a $MAT$ x $MAP$ interaction was accepted when $\Delta$AIC >2 relative to a null including $MAT$ and $MAP$ as fixed effects.
 

### 2020-10-23 manuscript/appendix.Rmd
updated appendix

@@
 ```{r setup, include = FALSE}
-#knitr::opts_knit$set(root.dir = 'C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/')
+# knitr::opts_knit$set(root.dir = 'C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/')
 # knitr::opts_knit$set(root.dir = 'C:/Users/gyrcbm/Dropbox/Global_Productivity/manuscript/')
@@
 \blandscape
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S3: Ratios among forest C fluxes as a function of latitude and climate variables", out.width ='80%', fig.pos='H'}
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S3: Ratios among forest C fluxes as a function of latitude and climate variables.", out.width ='80%', fig.pos='H'}
 

### 2020-10-26 manuscript/response to reviews.Rmd
pass through response-to-reviews
@beckybanbury , I went through this and did a bit of work on responses (bold) and comments on what needs to be done (italic)

@@
 
+**bold- responses**
+
+*italic- tentative responses/ comments*
+
 Reviewer(s)' Comments to Author:
@@
 Reviewer: 1
+
+*Becky, we probably want to reword some of these responses to R1 a bit, as it's possible that this will go back to R1.
+Also, it could be good to modify the manuscript in response to some of the comments to ensure that other readers don't come up with the same mis-conceptions.*
 
@@
 
-*This statement suggests a lack of perspective on the scope of our analysis relative to previous studies.
+**This statement suggests a lack of perspective on the scope of our analysis relative to previous studies.
 Whereas most studies consider at most a few flux variables, ours considers nine.The most comprehensive similar study is @luyssaert_co_2007, which was published in *GCB* more than a decade ago, was based on a database <25% the size of the ForC version used here, did not control for effects of stand age/ disturbance history, and examined latitudinal/ global climatic trends in only three variables.
-Thus, the current analysis will be at the clear forefront of studies examining correlations between forest carbon flux and climate at the global scale.*
+Thus, the current analysis will be at the clear forefront of studies examining correlations between forest carbon flux and climate at the global scale.**
+
+**To help clarify the scope of this analysis to readers, we have....**
+
+*(let's add a few statements to emphasize how greatly the scope differs from previous studies.)*
 
@@
 
-*We agree that these results are not surprising, however a crucial part of science is built on high-quality synthesis that clarifies current understanding and probes for unexpected relationships.
-This review under-values high-quality synthesis that produces unsurprising results.*
+**We agree that these results are not surprising, however a crucial part of science is built on high-quality synthesis that clarifies current understanding and probes for unexpected relationships.**
 
-*As the review identifies, results from previous studies were complicated by other effects.
-The power of this synthesis is in how it standardises for the effects of stand age and disturbance history, unambiguously clarifying these relationships.*
+**As this comment identifies, results from previous studies were complicated by other effects.
+The power of this synthesis lies in how it standardises for the effects of stand age and disturbance history, unambiguously clarifying these relationships.**
+
+*(again, maybe make some modest change to the text that we can point to as a "response" to this review. Cite Yu.)*
 
@@
 
-*We agree, and fully acknowledge in the paper, that this study does not reveal the biological mechanisms.
+**We agree, and fully acknowledge in the paper, that this study does not reveal the biological mechanisms.
 Further, we do not argue that the spatial gradient represents the response of forests to climate change, as clearly discussed in the final paragraph of the discussion.
 However, we do argue that elucidating broad-scale patterns in C cycling across climatic gradients is a complementary approach to addressing a challenging problem [*sensu* @andersonteixeira_altered_2013, *GCB*].
-We note that studies using this approach have been published recently in *Science* [@sullivan_long-term_2020],  *GCB* [@trugman_climate_2019], *JGR* [@li_dissimilatory_2019], and *Scientific Reports* [@chen_spatial_2019].*
+We note that studies using this approach have been published recently in *Nature* [@cook-patton_mapping_2020], *Science* [@sullivan_long-term_2020],  *GCB* [@trugman_climate_2019], *JGR* [@li_dissimilatory_2019], and *Scientific Reports* [@chen_spatial_2019].**
 
@@
 
-*This suggestion indicates carelessness of review and/or limited understanding of appropriate publishing standards.
-The suggestion is completely inappropriate for two reasons.First, the hypothesis-testing analyses presented here would be unsuitable for a data journal.
+**This suggestion is inappropriate for two reasons.
+First, the hypothesis-testing analyses presented here would be unsuitable for a data journal.
 Second, the ForC database is already published [tropical portion only in @anderson-teixeira_carbon_2016, *GCB*; full database as a data paper in Anderson-Teixeira et al.
 @anderson-teixeira_forc_2018, *Ecology*].
-Even a semi-careful review would not have missed this fact, which was prominently cited in the last paragraph of the introduction and the first sentence of the methods.
-Thus, this suggestion betrays carelessness of review and/or limited understanding of appropriate standards for judging where scientific work should be published.*
+This fact is prominently cited in the last paragraph of the introduction and the first sentence of the methods.**
 
@@
 
+*(This is an interesting comment, and points to some more robust work on the subject of productivity and growing season length.
+I'm going to look up these papers.
+Interestingly, this links to another paper that current intern Cameron Dow is working on, which uses dendrobands to show that warmer springs shift growth earlier in the year but have at most modest effect on total productivity.)*
+
 2.
@@
 2 shows that C fluxes increase approximately in proportion to one another, but that does not necessarily mean allocation is unchanged across latitude or climates.
+
+*(requires response)*
 
@@
 
+*(add something on this to the discussion, at the same time addressing R1's 3rd comment)*
+
 Minor comments:
@@
 
-*Stated in methods section that this information is available in table S1*
+**We have added brief definitions of these in the text (methods section), and refer readers to Table S1 for detailed information.**
 
@@
 
-*Clarifed in table*
+**We have clarified this in the table caption.**
 
@@
 
-*Added CUE and legend to show none of the fits are significant (because of this no statistics presented)*
+**We have added CUE as requested. We also added a legend to clarify that none of the fits are significant (and therefore no statistics presented).**
 
@@
 
-*We use MATxMAP in a linear model - clarified this in the text*
+**We used MATxMAP in a linear model, which we have clarified this in the text.**
 
@@
 
+*(requires response)*
+
 6. Fig. S6. Spell out “T seas” and “P seas” in the caption.
+
+*(requires response)*
 

### 2020-10-26 manuscript/appendix.Rmd
pass through response-to-reviews
@beckybanbury , I went through this and did a bit of work on responses (bold) and comments on what needs to be done (italic)

@@
 Signficance of relationships of forest C fluxes to MAT alone and in combination with MAP.
-$R^{2}$ values refer to the best model.", escape = F)%>%
+$R^{2}$ values refer to the best model.",  caption.short = "Table S3.
+Signficance of relationships of forest C fluxes to MAT alone and in combination with MAP.", escape = F)%>%
   kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 12) %>%

### 2020-10-26 manuscript/manuscript draft.Rmd
pass through response-to-reviews
@beckybanbury , I went through this and did a bit of work on responses (bold) and comments on what needs to be done (italic)

@@
 ForC contains geographic coordinates associated with each measurement record and, when available, $MAT$ and $MAP$ as reported in the primary literature [@anderson-teixeira_forc_2018].
-Based on the geographic co-ordinates for each site, data on twelve climate variables -- including $MAT$, $MAP$, temperature and precipitation seasonality, annual temperature range, solar radiation, cloud cover, annual frost and wet days, potential evapotranspiration ($PET$), aridity ($MAP/PET$), and vapor pressure deficit ($VPD$) -- were extracted from five open-access climate datasets: WorldClim [@hijmans_very_2005], WorldClim2 [@fick_worldclim_2017], the Climate Research Unit time-series dataset (CRU TS v4.03 [@harris_updated_2014], the Global Aridity Index and Potential Evapotranspiration Climate Database [@trabucco_global_2019], and TerraClimate [@abatzoglou_terraclimate_2018] (Table S1).
+Based on the geographic co-ordinates for each site, data on twelve climate variables -- including $MAT$, $MAP$, temperature seasonality (*i.e.*, standard deviation across months), precipitation seasonality (*i.e.*, coefficient of variation across months), annual temperature range, solar radiation, cloud cover, annual frost and wet days, potential evapotranspiration ($PET$), aridity ($MAP/PET$), and vapor pressure deficit ($VPD$) -- were extracted from five open-access climate datasets: WorldClim [@hijmans_very_2005], WorldClim2 [@fick_worldclim_2017], the Climate Research Unit time-series dataset (CRU TS v4.03 [@harris_updated_2014], the Global Aridity Index and Potential Evapotranspiration Climate Database [@trabucco_global_2019], and TerraClimate [@abatzoglou_terraclimate_2018] (Table S1).
 Definitions and methods used to calculate each variable are included in Table S1.

### 2020-10-27 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@
 S3).
-Of 12 relationships tested (3 ratios among C flux variables regressed against latitude, $MAT$, $MAP$ and temperature seasonality), none were significant.
+Of 16 relationships tested (4 ratios among C flux variables regressed against latitude, $MAT$, $MAP$ and temperature seasonality), none were significant.
 

### 2020-10-27 manuscript/response to reviews.Rmd
Update response to reviews.Rmd

@@
 
+**We found no significant relationships between flux ratios (allocation to fluxes) across latitude and three climate variables tested, and as a result have not presented figures illustrating this in the text.
+We state that these results were non-significant in the results section.
+We have updated Fig.
+S3 to clarify that none of the fits are significant (and therefore no statistics are presented).**
+
 *(requires response)*
@@
 If tropical forest stands are generally older than the extra-tropical forest stands (according to “Histogram_of_Stand_age” in the ForC database), then perhaps there is a need to standardize samples by stand ages when studying allocation.
+
+**While ForC contains information on precise stand ages for many temperate forests, this information is typically unavailable for tropical forests, and so these forests are often recorded simply as 'old-growth' or 'undisturbed'.
+This means that they are known to be older than 100 years of age, but the precise age is unknown.
+As a result, standardising this dataset by age is a challenge, and so we treat all forests in the analysis as mature old-growth forests, assuming minimal effect of age.** *CITE?*
 

### 2020-10-27 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@
 However, studies have differed in their conclusions regarding the shape of this relationship -- quite possibly because of lack of standardization with respect to methodology and stand history.
+*Productivity may vary with stand age, disturbance, and management [@de_lucia_forest_2007; @yu_high_2014; @simova_enigma_2017], making clear latitudinal patterns difficult to discern without standardization of the dataset.
+[???]*
 For instance, studies agree that gross primary productivity ($GPP$) increases continuously with decreasing latitude and is indisputably highest in tropical forests [@luyssaert_co_2007; @beer_terrestrial_2010; @jung_global_2011; @badgley_terrestrial_2019; @li_mapping_2019].
@@
 C fluxes generally saturate at high levels of $MAP$, though the saturation points identified vary from $MAP$ of ~1000 mm for $R_{root}$ [@wei_forest_2010]  up to 2,445 mm for $NPP$ [@schuur_productivity_2003].
-Interactions between $MAT$ and $MAP$ are also possible; within the tropics, there is a positive interaction between $MAT$ and $MAP$ in shaping $ANPP$, such that high rainfall has a negative effect on productivity in cooler climates, compared to a positive effect in warmer climates [@taylor_temperature_2017].
+Interactions between $MAT$ and $MAP$ may also influence productivity [@yu_high_2014]; within the tropics, there is a positive interaction between $MAT$ and $MAP$ in shaping $ANPP$, such that high rainfall has a negative effect on productivity in cooler climates, compared to a positive effect in warmer climates [@taylor_temperature_2017].
 There is also evidence that C fluxes also respond to climate variables such as temperature and precipitation seasonality [@wagner_climate_2016], cloud cover [@taylor_temperature_2017], solar radiation [@beer_terrestrial_2010; @fyllas_solar_2017], and potential evapotranspiration [@kerkhoff_plant_2005]; however, these are not typically assessed in global-scale analyses of annual forest C flux.
@@
 Previous indications of such a pattern may have been an artifact of differences in stand age across biomes.
-Compared to tropical forests, the temperate forest biome has experienced more widespread anthropogenic disturbance and has a larger fraction of secondary stands [@potapov_mapping_2008; @poulter_global_2018], so analyses comparing across latitudinal gradients without controlling for stand age risk confounding age with biome effects.
-Because carbon allocation varies with stand age [@de_lucia_forest_2007; @anderson-teixeira_altered_2013; @doughty_what_2018], age differences may introduce systematic biases into analyses of C fluxes across latitude or global climatic gradients.
+Compared to tropical forests, the temperate forest biome has experienced more widespread anthropogenic disturbance and has a larger fraction of secondary stands [@potapov_mapping_2008; @yu_high_2014; @poulter_global_2018], so analyses comparing across latitudinal gradients without controlling for stand age risk confounding age with biome effects.
+Because carbon allocation varies with stand age [@de_lucia_forest_2007; @anderson-teixeira_altered_2013; @yu_high_2014; @doughty_what_2018], age differences may introduce systematic biases into analyses of C fluxes across latitude or global climatic gradients.
 For example, woody productivity tends to be higher in rapidly aggrading secondary stands than in old-growth forests, where proportionally more C is allocated to respiration and non-woody productivity [@de_lucia_forest_2007; @piao_forest_2010; @doughty_what_2018; @kunert_understanding_2019].
@@
 
-We show that C fluxes are broadly consistent in their responses to climate drivers on the global scale, with no trends in C allocation among the variable pairs tested (Figs. 2, 4, S3).
+We show that C fluxes are broadly consistent in their responses to climate drivers on the global scale, with no trends in C allocation among the variable pairs tested (Figs. 2, S3).
 This parallels the observation that C allocation across multiple C fluxes varies little with respect to climate along a steep tropical elevational gradient [@malhi_variation_2017; but see @moser_elevation_2011],

### 2020-10-27 manuscript/response to reviews.Rmd
Update response to reviews.Rmd

@@
 *(again, maybe make some modest change to the text that we can point to as a "response" to this review. Cite Yu.)*
+
+*have cited Yu in intro + discussion.
+In intro added following : "Productivity may vary with stand age, disturbance, and management [@de_lucia_forest_2007; @yu_high_2014; @simova_enigma_2017], making clear latitudinal patterns difficult to discern without standardization of the dataset.".
+I haven't edited discussion text other than adding citations to existing points - it seems to me to be covered well?*
 

### 2020-10-27 manuscript/response to reviews.Rmd
Update response to reviews.Rmd
@teixeirak I've made a few edits to this document this morning

@@
 *(let's add a few statements to emphasize how greatly the scope differs from previous studies.)*
+
+RE Luyssaert et al : *We note in particular that mean stand age was extremely variable across biomes, and while mean stand age in tropical forests was >100 years (old growth) in all but one extra-tropical biome, mean stand age was >100.
+This is likely to elevate productivity estimates for extra-tropical biomes, highlighting the need for a comprehensive, standardised dataset for analysis.*
 

### 2020-11-10 manuscript/response to reviews.Rmd
responing to reviewer comments on growing season length.

@@
 
-*(This is an interesting comment, and points to some more robust work on the subject of productivity and growing season length.
-I'm going to look up these papers.
-Interestingly, this links to another paper that current intern Cameron Dow is working on, which uses dendrobands to show that warmer springs shift growth earlier in the year but have at most modest effect on total productivity.)*
+**Thank you for excellent suggestion to expand the discussion of growing season length around the references given.
+We have re-written the paragraphs in the introduction and the discussion dealing with this issue, now citing the four references listed in the comment above.**
+
+*(This is a great comment.
+The references given in the comment are more robust work on the subject of productivity and growing season length and make the framing much richer.
+{Interestingly, this links to another paper that current intern Cameron Dow is working on, which uses dendrobands to show that warmer springs shift growth earlier in the year but have at most modest effect on total productivity.}  More broadly, I think there may be other areas of the paper where more engangement with EC literature would make the story richer, but this will require some lit search.)*
 

### 2020-11-10 manuscript/manuscript draft.Rmd
load references
include growing season length refs (Reviewer 2)
also, #121
citation placeholder
responing to reviewer comments on growing season length.

@@
 As metrics of annual climate, $MAT$ and $MAP$ fail to capture variation in climate on an intra-annual scale, including temperature and precipitation seasonality and growing season length. 
-Some studies have suggested that the apparently strong relationship between $MAT$ and C fluxes is actually a consequence of the correlation between $MAT$ and growing season length [@kerkhoff_plant_2005; @michaletz_convergence_2014; @michaletz_drivers_2018].
-@kerkhoff_plant_2005 and @michaletz_convergence_2014 found no significant relationship between growing season temperature and $ANPP$ or $NPP$ standardized to growing season length [but see @chu_does_2016].
-While this suggests that the influence of temperature may be limited to determining the length of the frost-free growing season, analysis with carefully standardized variables and forest ages would be necessary to test the the veracity and generality of this hypothesis.
+Most forests--even tropical evergreen--exhibit some seasonality in both climate and C flux [e.g., @wagner_pan-tropical_2014], and this seasonality influences annual C fluxes [@churkina_spatial_2005; @keenan_net_2014; @fu_maximum_2019].
+In particular, growing season length has been linked to $ANPP$, $NPP$, $GPP$, and net ecosystem exchange of CO~2~ [$NEE$, or the difference between $GPP$ and ecosystem respiration; @kerkhoff_plant_2005; @michaletz_convergence_2014; @churkina_spatial_2005; @keenan_net_2014; @zhou_explaining_2016].
+However, the relative importance of growing season length, as opposed to climate within the growing season, remains debated.
+On one end of the spectrum, some studies have suggested that the influence of temperature on C fluxes may be limited to determining the length of the frost-free growing season, and that climate within the growing season has little influence on C fluxes because of plant adaptation and acclimatization to local climates [@kerkhoff_plant_2005; @enquist_adaptive_2007; @michaletz_convergence_2014; @michaletz_drivers_2018].
+In support of this, @kerkhoff_plant_2005 and @michaletz_convergence_2014 found no significant relationship between growing season temperature and $ANPP$ or $NPP$ standardized to a climate-defined growing season length [but see @chu_does_2016].
+The idea that growing season length is an important determinant of annual C flux also aligns with evidence that cross-site variation in $NEE$ is strongly correlated with growing season length [@churkina_spatial_2005] and that warming-induced increases in growing season length are enhancing forest $GPP$ and C sequestration [@keenan_net_2014; @zhou_explaining_2016].
+On the other end of the spectrum, climatic conditions within the growing season may exert a stronger influence on annual C fluxes than the length of the growing season. 
+This aligns with observations that in forests, $NEE$ tends to be more closely tied to the maximum rate of CO~2~ uptake than to the carbon uptake period [@zhou_explaining_2016; @fu_maximum_2019], and with numerous tree-ring analyses finding that annual growth is more closely controlled by peak growing season climate than by spring or fall conditions [e.g., @helcoski_growing_2019].
+Thus, the extent to which growing season length controls global-scale variation in forest autotrophic C fluxes remains unclear.
 
@@
 
+**in the paragraph above, discuss/ cite @collati_forest_2020**
+
 Despite the broad consistency of climate responses across C fluxes, climate explains lower proportions of variability among some of the subsidiary C fluxes (*e.g.*, $ANPP_{stem}$, $BNPP$,  $BNPP_{fine.root}$; Fig.
@@
 
+**The above paragraph needs a revamp, including discussion of  @keenan_net_2014, @fu_maximum_2019, @churkina_spatial_2005, @zhou_explaining_2016.**
+
 Our analysis clarifies how annual forest autotrophic C fluxes vary with latitude and climate on a global scale, with some important implications for how forest C cycling relates to climate and, by extension, how it is likely to respond to climatic warming.

### 2020-11-12 manuscript/response to reviews.Rmd
worked on discussion of seasonality

@@
 **Thank you for excellent suggestion to expand the discussion of growing season length around the references given.
-We have re-written the paragraphs in the introduction and the discussion dealing with this issue, now citing the four references listed in the comment above.**
+We have re-written the paragraphs in the introduction and the discussion dealing with this issue, now citing the references listed in the comment above.**
 

### 2020-11-12 manuscript/manuscript draft.Rmd
worked on discussion of seasonality

@@
   pack_rows("Q4. How does seasonality influence annual C fluxes?", 8, 11, latex_gap_space = "1em", colnum = 5, hline_before = TRUE) %>%
-  pack_rows("Q5. When standardised by growing season length, how do annual C fluxes vary with climate?", 12, 15, latex_gap_space = "1em", colnum = 5, hline_before = TRUE) %>%
+  pack_rows("Q5. When standardised by growing season length, how do annual C fluxes still vary with climate?", 12, 15, latex_gap_space = "1em", colnum = 5, hline_before = TRUE) %>%
   add_header_above(c(" ",  " ", "Forest autotrophic carbon fluxes" = 9, " ")) %>%
@@
 
-Forest C fluxes decline with temperature seasonality (Table 1, *Q4*; Fig. 4e), and are minimal during cold- or dry- dormant seasons. 
-To account for this, a number of analyses seeking to characterize global-scale effects of climate on productivity have examined the relationship of C flux per month of the growing season with growing season climatic conditions [Table 1, *Q5*; @kerkhoff_plant_2005; @anderson_temperature-dependence_2006; @enquist_adaptive_2007; @michaletz_convergence_2014].
-The sort of simple metric that has been used to define growing season at a global scale [@kerkhoff_plant_2005] is coarse with respect to temperature because it is calculated on a monthly timescale and problematic with respect to moisture because it doesn't capture temporal lags between precipitation and plant water use caused by storage in soil or snow.
-We found that a temperature-defined growing season length had strong positive correlation with C fluxes (Fig. 4f), but was never the best univariate predictor. 
-Dividing annual fluxes by growing season length to yield average flux per growing season month removed the majority of climate-related variation, supporting the idea that the latitudinal gradient in carbon flux is attributable more to shorter growing seasons at high latitudes than to inherently lower rates of photosynthesis or respiration by high-latitude forests [@enquist_adaptive_2007].
-However, there remained a number of significant correlations with growing season climatic conditions, indicating that climatic conditions remain influential within the growing season (Figs. S8-S9). 
-We conclude that while correcting for growing season length takes analyses a step closer to mechanistic linkage of instantaneous C flux rates to environmental conditions, it remains crude relative to the timescales on which climate affects plant metabolism, and does not advance statistical predictive power.
-Rather, mechanistic accounting for climatic effects on global forest carbon flux patterns requires models representing physiologically meaningful timescales [e.g., @medvigy_mechanistic_2009; @longo_biophysics_2019].
-
-**The above paragraph needs a revamp, including discussion of  @keenan_net_2014, @fu_maximum_2019, @churkina_spatial_2005, @zhou_explaining_2016.**
+Forest C fluxes decline with temperature seasonality (Table 1, *Q4*; Fig. 4e), indicating that fluxes during the growing season are not large enough to compensate for minimal flux during winters. 
+A temperature-defined growing season length correlated strongly with global-scale variation in annual C flux [Table 1, *Q5*; Fig.
+4f; see also @churkina_spatial_2005], consistent with the idea that the latitudinal gradient in carbon flux is attributable more to shorter growing seasons at high latitudes than to inherently lower rates of photosynthesis or respiration by high-latitude forests [@enquist_adaptive_2007; @fu_maximum_2019].
+While there is evidence that trees in high-latitude forests have adaptations to maximize photosynthesis at low temperatures [@helliker_subtroical_2008] (**REFS**), this is not sufficient to yield growing season fluxes comparable to those of tropical forests, as indicated by a number of positive correlations between monthly mean flux during the growing season and growing season temperature, solar radiation, and $PET$ (Table 1, Figs.
+S8-S9).
+Thus, we reject the hypothesis that growing season length alone accounts for global-scale variation in productivity--*i.e.*, that there is no relationship between C flux per month of the growing season and growing season climatic conditions [Table 1, *Q5*; @kerkhoff_plant_2005; @enquist_adaptive_2007; @michaletz_convergence_2014].
+Rather, annual C flux is shaped by both growing season length and the climate of peak growing season months [**REFS**; @chu_does_2016; @fu_maximum_2019]. 
+Given strong co-variation between growing season length and $MAT$ [Fig.
+S2; @chu_does_2016], accurately partitioning this variation will require data on intra-annual variation in C flux coupled with a more refined metric of growing season length than used here [e.g., based on leaf phenology or C exchange, *sensu* @fu_maximum_2019].
+@fu_maximum_2019 find that global-scale geographic variation in annual $NEE$ is driven more strongly by growing season length than by carbon uptake rates within the growing season, whereas interannual variation in $NEE$ and $GPP$ at any given site appears to be driven predominantly by the maximum rate of C uptake, as opposed to growing season length [@zhou_explaining_2016; @fu_maximum_2019].
+Further analysis of interannual variation in C fluxes in relation to climate will be valuable to disentangling how seasonality shapes broad geographic patterns in forest C flux.
 

### 2020-11-12 manuscript/manuscript draft.Rmd
references

@@
 S3), refuting the idea that tropical forests have anomalously low $CUE$ [@de_lucia_forest_2007; @malhi_productivity_2012; @anderson-teixeira_carbon_2016].
-Rather, differences in $CUE$ between old-growth tropical forests relative to (mostly younger) extratropical forests are likely an artifact of comparing stands of different age, as $CUE$ is known to decline with forest age [@de_lucia_forest_2007; @piao_forest_2010; @collalti_is_2019].
+Rather, differences in $CUE$ between old-growth tropical forests relative to (mostly younger) extratropical forests are likely an artifact of comparing stands of different age, as $CUE$ is known to decline with forest age [@de_lucia_forest_2007; @piao_forest_2010; @collalti_forest_2020].
 Another previously observed pattern for which we find no support is a tendency for belowground C allocation to decrease with increasing temperature  [@moser_elevation_2011; @gill_belowground_2016]; rather, we observe no trends in allocation between $ANPP$ and $BNPP$ across latitudes.
@@
 
-**in the paragraph above, discuss/ cite @collati_forest_2020**
+**in the paragraph above, discuss/ cite @collalti_forest_2020**
 
@@
 4f; see also @churkina_spatial_2005], consistent with the idea that the latitudinal gradient in carbon flux is attributable more to shorter growing seasons at high latitudes than to inherently lower rates of photosynthesis or respiration by high-latitude forests [@enquist_adaptive_2007; @fu_maximum_2019].
-While there is evidence that trees in high-latitude forests have adaptations to maximize photosynthesis at low temperatures [@helliker_subtroical_2008] (**REFS**), this is not sufficient to yield growing season fluxes comparable to those of tropical forests, as indicated by a number of positive correlations between monthly mean flux during the growing season and growing season temperature, solar radiation, and $PET$ (Table 1, Figs.
+While there is evidence that trees in high-latitude forests have adaptations to maximize photosynthesis at low temperatures [@helliker_subtropical_2008] (**REFS**), this is not sufficient to yield growing season fluxes comparable to those of tropical forests, as indicated by a number of positive correlations between monthly mean flux during the growing season and growing season temperature, solar radiation, and $PET$ (Table 1, Figs.
 S8-S9).
 Thus, we reject the hypothesis that growing season length alone accounts for global-scale variation in productivity--*i.e.*, that there is no relationship between C flux per month of the growing season and growing season climatic conditions [Table 1, *Q5*; @kerkhoff_plant_2005; @enquist_adaptive_2007; @michaletz_convergence_2014].
-Rather, annual C flux is shaped by both growing season length and the climate of peak growing season months [**REFS**; @chu_does_2016; @fu_maximum_2019]. 
+Rather, annual C flux is shaped by both growing season length and the climate of peak growing season months [@chu_does_2016; @fu_maximum_2019]. 
 Given strong co-variation between growing season length and $MAT$ [Fig.

### 2020-11-12 manuscript/response to reviews.Rmd
a bit of work on reviews

@@
 
-*Potentially we need a response to the "simple statistical analyses" point?
-- these analyses are valuable for clarifying relationships between carbon fluxes and climate.
-Although analyses are simple, they are powerful thanks to the coverage and extent of the dataset, and the range of fluxes examined.*
+**Most of the points raised here are covered by more specific comments below, and our responses to these points follow those comments.
+The one criticism here that is not addressed below is that we present "simple statistical analyses".
+While we concur that the analyses are relatively simple, that does not preclude their being appropriate, informative, and sufficient.
+They represent several substantive advances over past analyses of this type:**
+
+**1. They are underlain by a large, standardized database that accounts for factors such as stand age and management, which have not always been accounted for.**
+
+**2.
+We use mixed effects model to account for random effects of location (plot nested within geographic area), thereby accounting for potential pseudoreplication, which has not always been done in the past.**
+
+**3. We examined nine carbon fluxes and relationships among them--more than has been concurrently analyzed in the past.**
+
+**4.
+For each flux variable, we tested a wide range of models: univariate responses to 12 climate variables (3 functional fit types for each), along with multivariate models (MAT x MAP for all, along with other combinations not presented).**
+
+**Thus, although the analyses are simple, we believe them to be appropriate and sufficient, with their novelty arising not from the analysis method itself but from the coverage and extent of the dataset.**
 
@@
 **This statement suggests a lack of perspective on the scope of our analysis relative to previous studies.
-Whereas most studies consider at most a few flux variables, ours considers nine.The most comprehensive similar study is @luyssaert_co_2007, which was published in *GCB* more than a decade ago, was based on a database <25% the size of the ForC version used here, did not control for effects of stand age/ disturbance history, and examined latitudinal/ global climatic trends in only three variables.
+Whereas most studies consider at most a few flux variables, ours considers nine.
+The most comprehensive similar study is @luyssaert_co_2007, which was published in *GCB* more than a decade ago, was based on a database <25% the size of the ForC version used here, did not control for effects of stand age/ disturbance history, and examined latitudinal/ global climatic trends in only three variables.
 Thus, the current analysis will be at the clear forefront of studies examining correlations between forest carbon flux and climate at the global scale.**
 
-**To help clarify the scope of this analysis to readers, we have....**
+**To help clarify the scope of this analysis to readers, we added the following statement to the last paragraph of the introduction: "The most comprehensive analysis of this type was @luyssaert_co_2007, which was based on a database <25% the size of the ForC version used here, did not control for effects of stand age, and examined global climatic trends in only three variables."**
 
-*(let's add a few statements to emphasize how greatly the scope differs from previous studies.)*
+*(Do we like this response??)*
 

### 2020-11-12 manuscript/manuscript-draft.Rmd
a bit of work on reviews

--- 2020-01-20 manuscript/manuscript draft.Rmd
@@
 ---
-title: "Global patterns of forest autotrophic carbon fluxes"
-author: "Rebecca Banbury Morgan, Valentine Herrmann, Norbert Kunert, Ben Bond-Lamberty, Helene C. Muller-Landau, Kristina J. Anderson-Teixeira"
+title: ""
+author: ""
 output:
   pdf_document:
-    citation_package: natbib
     fig_caption: yes
-bibliography: [library.bib,  packages.bib]
-biblio-style: apalike
-csl: apa.csl
+bibliography: library.bib
+csl: global-change-biology.csl
 header-includes:  \usepackage{float}
@@
                   \captionsetup[figure]{font=footnotesize}
+                  \usepackage{setspace}\doublespacing
+                  \usepackage{lineno}\linenumbers
 
+nocite: | 
+  @chu_does_2016
 ---
@@
 ```{r setup, include = FALSE}
-#knitr::opts_knit$set(root.dir = 'C:/Users/banburymorganr/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/')
-knitr::opts_knit$set(root.dir = '/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/Global_Productivity/manuscript/')
+#knitr::opts_knit$set(root.dir = 'C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/')
+  # knitr::opts_knit$set(root.dir = '/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/Global_Productivity/manuscript/')
 ```
@@
 ```
+**Title:** Global patterns of forest autotrophic carbon fluxes
 
-###Abstract
-[**very rough start:**]
-.....
-Here, we drew upon # records from the Global Forest Carbon Database (ForC), representing all major forest types and the nine most significant forest autotrophic carbon flux (FACF) variables, to comprehensively explore how C cycling in mature, undisturbed forests varies with latitude and climate on a global scale.
-We show that, across all FACF variables analyzed, C cycling decreases linearly with latitude --a finding that confirms multiple previous studies and contradicts the idea that productivity of temperate forests rivals that of tropical forests.
-The FACF variables increase in proportion to one another, with no differences in allocation detectable at this global scale, but differed in that latitude explained a lower proportion of variation among subsidiary fluxes (in particular, woody aboveground productivity and belowground net primary productivity, $BNPP$).
-Climate explained a significant proportion (#-#%) of variation in all C fluxes analyzed (less for subsidiary fluxes), with temperature variables in general and mean annual temperature ($MAT$) in particular being the best predictors of FACF on this global scale.
-While other climate variables (*e.g.*, **XX**) displayed significant correlation with FACF, none of them had significantly better explanatory power than $MAT$.
-The effects of temperature were modified by moisture availability, with reduced FACF under hot and dry conditions and sometimes under very high precipitaiton (especially for $BNPP$).
-FACF declined with temperature seasonality, but growing season length doesn't improve upon MAT as a predictor.
-Within the growing season, the influence of climate on C cycling is smaller but still significant for a number of carbon fluxes.
-These findings clarify the big picture of how FACF varies with latitude and climate on a global scale.
-As we enter a period of accelerating climatic change, understanding of the fundamental climatic controls on FACF sets a foundation for understanding patterns of change.
+**Running head:** Global patterns of forest carbon fluxes
 
+**Authors:**
+
+Rebecca Banbury Morgan^1,2^
+
+Valentine Herrmann^1^
+
+Norbert Kunert^1,3^
+
+Ben Bond-Lamberty^4^
+
+Helene C. Muller-Landau^3^
+
+Kristina J. Anderson-Teixeira^1,3^*
+
+**Institutional Affiliations:**
+
+1. Conservation Ecology Center; Smithsonian Conservation Biology Institute; Front Royal, VA, USA
+
+2. School of Geography, University of Leeds, Leeds, UK
+
+3. Center for Tropical Forest Science-Forest Global Earth Observatory; Smithsonian Tropical Research Institute; Panama, Republic of Panama
+
+4. Joint Global Change Research Institute, Pacific Northwest National Laboratory, College Park Maryland 20740 USA
+
+*Corresponding Author:
+
+phone: 1-540-635-6546
+
+fax:1-540-635-6506
+
+email: teixeirak@si.edu
+
+\newpage
+### Abstract
+Carbon (C) fixation, allocation, and metabolism by trees set the basis for energy and material flows in forest ecosystems and define their interactions with Earth's changing climate. 
+However, we lack a cohesive synthesis on how forest carbon fluxes vary globally with respect to climate and one another. 
+Here, we draw upon 1,319 records from the Global Forest Carbon Database (ForC), representing all major forest types and the nine most significant autotrophic carbon fluxes, to comprehensively explore how C cycling in mature, undisturbed forests varies with latitude and climate on a global scale.
+We show that, across all flux variables analyzed, C cycling decreases continuously with absolute latitude -- a finding that confirms multiple previous studies but contradicts the idea that net primary productivity of temperate forests rivals that of tropical forests.
+C flux variables generally displayed similar trends across latitude and multiple climate variables, with no differences in allocation detected at this global scale.
+Temperature variables in general, and mean annual temperature and temperature seasonality in particular, were the best univariate predictors of C flux, explaining 19 - 71% of variation in the C fluxes analyzed.
+The effects of temperature were modified by moisture availability, with C flux reduced under hot and dry conditions and sometimes under very high precipitation. 
+C fluxes increased with growing season length, but this was never the best univariate predictor.
+Within the growing season, the influence of climate on C cycling was small but significant for a number of flux variables. 
+These findings clarify how forest C flux varies with latitude and climate on a global scale. 
+In a period of accelerating climatic change, this improved understanding of the fundamental climatic controls on forest C cycling sets a foundation for understanding patterns of change.
+
+**Keywords:**  carbon fluxes; carbon dioxide (CO~2~); climate; forest; global; productivity; respiration; latitude
+
+\newpage
 ### Introduction
 
-(**revisit these first two paragraphs.**)
+Carbon (C) cycling in Earth's forests provides the energetic basis for sustaining the majority of Earth's terrestrial biodiversity and many human populations [@millennium_ecosystem_assessment_ecosystems_2005], while strongly influencing atmospheric carbon dioxide (CO~2~) and climate [@bonan_forests_2008].
+Forests' autotrophic carbon fluxes -- that is, carbon fixation, allocation, and metabolism by trees and other primary producers -- sets the energy ultimately available to heterotrophic organisms (including microbes), in turn influencing their abundance [@zak_plant_1994; @niedzialkowska_species_2010] and possibly diversity [@waide_relationship_1999; @chu_direct_2018].
+They are linked to cycling of energy, water, and nutrients and, critically, influence all C stocks and define forest interactions with Earth's changing climate. 
+Each year, over 69 Gt of C cycle through Earth's forests [@badgley_terrestrial_2019] -- a flux more than seven times greater than that of recent anthropogenic fossil fuel emissions [9.5 Gt C yr^-1^; @friedlingstein_global_2019].
+As atmospheric CO~2~ continues to rise, driving climate change, forests will play a critical role in shaping the future of Earth's climate [@cavaleri_urgent_2015; @rogelj_mitigation_2018]. 
+However, our understanding of the climate dependence of forest C cycling on a global scale has been limited by analyses typically considering only one or a few variables at a time, insufficient parsing of related variables, and the mixing of data from forests that vary in stand age, disturbance history, and management status, all of which affect C cycling [@litton_carbon_2007; @gillman_latitude_2015; @simova_enigma_2017].
 
-**Forests play a critical role in regulating atmospheric CO~2~ and climate** (**Bonan**).
-Approximately **##** Gt of CO~2~ cycle through Earth's forests each year **REF-KAT**), and their net C sequestration (~2.4 GtCyr^-1^) offsets roughly 30% of anthropogenic fossil fuel emissions [@pan_large_2011].
-As atmospheric carbon dioxide levels continue to rise, driving climate change, there is increasing recognition that  protection and management of forest resources will play an important role in mitigating climate change (** REFS- e.g., IPCC 1.5C and land reports**).
-Carbon fixation, allocation, and metabolism by trees set the basis for energy and material flows in forest ecosystems and define their interactions with Earth's changing climate.
+Forest C fluxes decrease with latitude [e.g., @luyssaert_co_2007;  @gillman_latitude_2015; @li_mapping_2019]. 
+However, studies have differed in their conclusions regarding the shape of this relationship -- quite possibly because of lack of standardization with respect to methodology and stand history.
+*Productivity may vary with stand age, disturbance, and management [@de_lucia_forest_2007; @yu_high_2014; @simova_enigma_2017], making clear latitudinal patterns difficult to discern without standardization of the dataset.
+[???]*
+For instance, studies agree that gross primary productivity ($GPP$) increases continuously with decreasing latitude and is indisputably highest in tropical forests [@luyssaert_co_2007; @beer_terrestrial_2010; @jung_global_2011; @badgley_terrestrial_2019; @li_mapping_2019].
+In contrast, some studies have suggested that net primary productivity ($NPP$), or its aboveground portion ($ANPP$), exhibits a less distinct increase from temperate to tropical forests [@luyssaert_co_2007] -- or even a decrease [@huston_global_2009, but see @gillman_latitude_2015].
+A shallower increase in $NPP$ than in $GPP$ with decreasing latitude would align with the suggestion that tropical forests tend to have low carbon use efficiency [$CUE$= $NPP$/$GPP$; @de_lucia_forest_2007; @malhi_productivity_2012; @anderson-teixeira_carbon_2016].
+Such differences among C fluxes their relationship to latitude could have profound implications for our understanding of the C cycle and its climate sensitivity.
+However, until recently the potential to compare latitudinal trends across C fluxes has been limited by lack of a large database with standardization for methodology, stand history, and management [@anderson-teixeira_forc_2018].
 
-**Forests' autotrophic carbon flux (FACF, bbl:define) influences all organic matter stocks in ecosystems and is linked to cycling of energy, water, and nutrients** (**REFS; *Piao et al.
-2010.
-Forest annual carbon cost: a global‐scale analysis of autotrophic respiration.
-Ecology 91(3), 652-661* **).
-Forest productivity sets the energy ultimately available to heterotrophic organisms (including microbes), in turn influencing their abundance (**REFS**) and possibly diversity ([Waide et al.
-2009](https://www.annualreviews.org/doi/abs/10.1146/annurev.ecolsys.30.1.257)**REFS**).
-However, our ability to draw general macroscopic conclusions regarding global variation in multiple productivity variables with respect to climate has been limited in that these analyses often mix forests that vary in stand age, disturbance history, and/or management status; do not always sufficiently parse related variables (*e.g.*, combining net primary productivity records with and without belowground components); and typically consider only one or a few variables at a time.
+The latitudinal gradient in forest C flux rates, along with altitudinal gradients [@girardin_net_2010; @malhi_variation_2017], is driven primarily by climate, which is a significant driver of C fluxes across broad spatial scales [@luyssaert_co_2007; @cleveland_relationships_2011; @wei_forest_2010].
+However, there is little consensus as to the shapes of these relationships or the best predictor variables.
+The majority of studies have focused on exploring the relationships of C fluxes to mean annual temperature ($MAT$) and mean annual precipitation ($MAP$), as the most commonly reported site-level climate variables.
+C fluxes increase strongly with $MAT$ on the global scale, but whether they saturate or potentially decrease at higher temperatures remains disputed. 
+Some studies have detected no deceleration or decline in $GPP$ [@luyssaert_co_2007], $NPP$ [@schuur_productivity_2003], or root respiration [$R_{root}$; @piao_forest_2010; @wei_forest_2010] with increasing $MAT$.
+In contrast, others have found evidence of saturation or decline of C flux in the warmest climates; @luyssaert_co_2007 found $NPP$ saturating at around 10$^\circ$C MAT; @larjavaara_temperature_2012 found that increases in $GPP$ saturate at approximately 25$^\circ$C $MAT$, and @sullivan_long-term_2020 found that, within the tropics, $ANPP_{stem}$ decreases at the highest maximum temperatures.
+C fluxes generally saturate at high levels of $MAP$, though the saturation points identified vary from $MAP$ of ~1000 mm for $R_{root}$ [@wei_forest_2010]  up to 2,445 mm for $NPP$ [@schuur_productivity_2003].
+Interactions between $MAT$ and $MAP$ may also influence productivity [@yu_high_2014]; within the tropics, there is a positive interaction between $MAT$ and $MAP$ in shaping $ANPP$, such that high rainfall has a negative effect on productivity in cooler climates, compared to a positive effect in warmer climates [@taylor_temperature_2017].
+There is also evidence that C fluxes also respond to climate variables such as temperature and precipitation seasonality [@wagner_climate_2016], cloud cover [@taylor_temperature_2017], solar radiation [@beer_terrestrial_2010; @fyllas_solar_2017], and potential evapotranspiration [@kerkhoff_plant_2005]; however, these are not typically assessed in global-scale analyses of annual forest C flux.
 
-**The productivity of forests varies with latitude**, showing a general trend of decreasing productivity with latitude [@beer_terrestrial_2010; @jung_global_2011].
-Studies agree that productivity is lowest in the boreal regions, and increases into the temperate regions [@luyssaert_co_2007; @huston_global_2009; @beer_terrestrial_2010; @jung_global_2011].
-However, evidence is inconclusive on whether productivity continues to increase into the tropics, or whether it plateaus in temperate regions.
-Evidence for this is further complicated by the fact that different studies use different measures of productivity to explore these relationships.
-For example, modelling of global terrestrial ecosystem gross primary productivity (GPP) through upscaling and calibration of eddy flux measurements indicates that GPP peaks in tropical forests [@beer_terrestrial_2010; @jung_global_2011];[Bagdley et al.
-2019](https://onlinelibrary.wiley.com/doi/epdf/10.1111/gcb.14729)).
-This is corroborated by analysis of site-level GPP measurements, which appear to reach their highest levels in tropical forests [@luyssaert_co_2007].
-{**condense the previous two sentences into one-- both primarily eddy flux**} In contrast, some studies suggest that the highest values of net primary productivity (NPP) may be found in temperate forests [@luyssaert_co_2007; @huston_global_2009], while others find NPP highest in the tropics and decreasing with latitude [@simova_enigma_2017].
-Other studies have chosen to focus exclusively on above-ground net primary productivity (ANPP), finding evidence of a weak negative relationship between ANPP and latitude [@huston_global_2009; @gillman_latitude_2015].
+As metrics of annual climate, $MAT$ and $MAP$ fail to capture variation in climate on an intra-annual scale, including temperature and precipitation seasonality and growing season length. 
+Most forests--even tropical evergreen--exhibit some seasonality in both climate and C flux [e.g., @wagner_pan-tropical_2014], and this seasonality influences annual C fluxes [@churkina_spatial_2005; @keenan_net_2014; @fu_maximum_2019].
+In particular, growing season length has been linked to $ANPP$, $NPP$, $GPP$, and net ecosystem exchange of CO~2~ [$NEE$, or the difference between $GPP$ and ecosystem respiration; @kerkhoff_plant_2005; @michaletz_convergence_2014; @churkina_spatial_2005; @keenan_net_2014; @zhou_explaining_2016].
+However, the relative importance of growing season length, as opposed to climate within the growing season, remains debated.
+On one end of the spectrum, some studies have suggested that the influence of temperature on C fluxes may be limited to determining the length of the frost-free growing season, and that climate within the growing season has little influence on C fluxes because of plant adaptation and acclimatization to local climates [@kerkhoff_plant_2005; @enquist_adaptive_2007; @michaletz_convergence_2014; @michaletz_drivers_2018].
+In support of this, @kerkhoff_plant_2005 and @michaletz_convergence_2014 found no significant relationship between growing season temperature and $ANPP$ or $NPP$ standardized to a climate-defined growing season length [but see @chu_does_2016].
+The idea that growing season length is an important determinant of annual C flux also aligns with evidence that cross-site variation in $NEE$ is strongly correlated with growing season length [@churkina_spatial_2005] and that warming-induced increases in growing season length are enhancing forest $GPP$ and C sequestration [@keenan_net_2014; @zhou_explaining_2016].
+On the other end of the spectrum, climatic conditions within the growing season may exert a stronger influence on annual C fluxes than the length of the growing season. 
+This aligns with observations that in forests, $NEE$ tends to be more closely tied to the maximum rate of CO~2~ uptake than to the carbon uptake period [@zhou_explaining_2016; @fu_maximum_2019], and with numerous tree-ring analyses finding that annual growth is more closely controlled by peak growing season climate than by spring or fall conditions [e.g., @helcoski_growing_2019].
+Thus, the extent to which growing season length controls global-scale variation in forest autotrophic C fluxes remains unclear.
 
-**Climate is a significant driver of productivity across broad spatial scales** [@cleveland_relationships_2011].
-The majority of studies have focused on exploring the relationships of productivity to mean annual temperature (MAT) and mean annual precipitation (MAP), as the most commonly reported site-level climate variables.
-These variables have the advantage that they describe broad trends in temperature and water availability, and therefore capture a lot of global-scale variation in climate.
-There is strong evidence that both MAT and MAP show significant positive relationships with productivity [@chu_does_2016].
-However, as with latitude, the shape of those relationships is not always clear, and, again, is complicated by the use of different measures of productivity across studies.
-Various measures of primary productivity saturate at high levels of MAP, though the saturation points identified vary from 1500mm [@luyssaert_co_2007] up to 2445mm MAP [@schuur_productivity_2003].
-Studies of the influence of MAT on productivity are less conclusive.
-Luyssaert et al.
-[-@luyssaert_co_2007] examined GPP and NPP and found that, while GPP increases linearly with MAT, NPP saturates at around 10$^\circ$C MAT.
-In contrast, Larjavaara and Muller-Landau [-@larjavaara_temperature_2012], find that increases in GPP saturate at approximately 25$^\circ$C MAT, while Schuur [-@schuur_productivity_2003] shows that NPP increases linearly with temperature.
-bbl:maybe remove following sentences The influence of these climate variables on productivity is further complicated by the possibility of interactive effects occurring between them.
-Taylor et al.
-[-@taylor_temperature_2017] showed that increased MAP had a negative effect at low MAT, but a positive effect at high levels of MAT, and *vice versa* {**Helene found this confusing**}, suggesting that both temperature and water availability are important in explaining productivity.
-
-**Primary productivity can be influenced by many other factors as well**, which often act across a range of scales, and may show interactive effects with each other [@cleveland_relationships_2011].
-On a local scale, stand age [@litton_carbon_2007; @gillman_latitude_2015], biodiversity (Liang et al.
-2019.
-Positive biodiversity-productivity relationship predominant in global forests.
-Science 354, Issue 6309, aaf8957), management [@simova_enigma_2017]; nutrient availability [@aragao_above-_2009]; and altitude [@girardin_net_2010; @malhi_variation_2017] all impact forest productivity.
-On a global scale, we expect that productivity is most strongly influenced by broad climatic gradients.
-There is evidence that productivity also responds to variables such as cloud cover [@taylor_temperature_2017], solar radiation [@fyllas_solar_2017], and potential evapotranspiration [@kerkhoff_plant_2005] in potentially significant ways.
-Furthermore, MAT and MAP are very coarse measures of climate, and so fail to capture much variation in climate on an intra-annual scale, including the effects of factors such as growing season length, number of frost-free days, temperature seasonality, and dry season length.
-Some studies have suggested that the apparently strong relationship between MAT and productivity is actually a consequence of the correlation between MAT and growing season length [@kerkhoff_plant_2005; @malhi_productivity_2012; @michaletz_convergence_2014; @michaletz_drivers_2018].
-Kerkhoff et al.
-[-@kerkhoff_plant_2005] and Michaletz et al.
-[-@michaletz_convergence_2014] find that, within the growing season, there is no significant relationship between productivity and MAT, suggesting that the effect of temperature is due to increased length of growing season, rather than an inherent influence of temperature on productivity.
-
-**The recent development of a global forest carbon database synthesizing multiple variables and including records of stand history (ForC; [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc:_2018]) opens up the possibility for a standardized analysis of global scale variation** in multiple components of forest productivity and the principle climatic drivers of these patterns.
-In order to approach these broad and complex issues, we simplify the major gaps in our knowledge to five broad hypotheses and corresponding specific predictions (Table 1).
-** see [issue #48 ](https://github.com/forc-db/Global_Productivity/issues/48)** First, we ask how forest autotrophic carbon fluxes (FACF) vary with latitude.
-We then test how these fluxes relate to MAT and MAP, and additionally how they respond to other, less well studied, climate variables.
-Finally, we consider the relationship between FACF and seasonality, considering the role of seasonality in explaining variation in carbon fluxes, and the influence of climate on FACF standardized by growing season length.
-We use a comprehensive global database of forest carbon fluxes to address the above questions for nine carbon fluxes, allowing for an in-depth exploration of the effect of climate on global productivity.
+The recent development of the Global Forest Carbon database (ForC), which synthesizes multiple variables and includes records of stand history [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018], opens up the possibility for a standardized analysis of global scale variation in multiple C fluxes and the principle climatic drivers of these patterns.
+The most comprehensive analysis of this type was @luyssaert_co_2007, which was based on a database <25% the size of the ForC version used here, did not control for effects of stand age, and examined global climatic trends in only three variables.
+In order to approach this broad topic, we simplify the major gaps in our knowledge to five broad questions and corresponding predictions (Table 1).
+First, we ask how nine forest autotrophic carbon fluxes in ForC vary with latitude (*Q1*).
+We then test how these fluxes relate to $MAT$ and $MAP$ (*Q2*), and additionally how they respond to other, less well-studied, climate variables (*Q3*).
+Finally, we consider the relationship between C flux and seasonality, considering the role of seasonality in explaining variation in carbon fluxes (*Q4*), and the influence of climate on C flux standardized by growing season length (*Q5*).
 
@@
 hypothesis_table <- read.csv("hypothesis_table.csv", stringsAsFactors = FALSE, check.names = FALSE)
-kable(hypothesis_table, booktabs = TRUE, caption = "**Summary of hypotheses, corresponding specific predictions, and results.** Direction of significant relationships is indicated by '+' and '-', and  shape is summarized as linear (L), concave up (CU), and concave down (CD).
-'n.s.' indicates that the relationship is not signficant.") %>%
+#hypothesis_table[1,1] <- paste0(hypothesis_table[1,1], footnote_marker_number(1),"$^{,}$",footnote_marker_number(2),"$^{,}$", footnote_marker_number(3),"$^{,}$", footnote_marker_number(10))
+#hypothesis_table[2,1] <- paste0(hypothesis_table[2,1], footnote_marker_number(1),"$^{,}$",footnote_marker_number(4),"$^{,}$",footnote_marker_number(9))
+#hypothesis_table[3,1] <- paste0(hypothesis_table[3,1], footnote_marker_number(1),"$^{,}$",footnote_marker_number(4))
+#hypothesis_table[4,1] <- paste0(hypothesis_table[4,1], footnote_marker_number(5))
+#hypothesis_table[10,1] <- paste0(hypothesis_table[10,1], footnote_marker_number(6),"$^{,}$",footnote_marker_number(7),"$^{,}$", footnote_marker_number(8))
+#hypothesis_table[11,1] <- paste0(hypothesis_table[11,1], footnote_marker_number(7),"$^{,}$", footnote_marker_number(8))
+#hypothesis_table[12,1] <- paste0(hypothesis_table[12,1], footnote_marker_number(8))
+kable(hypothesis_table, format = "latex", booktabs = TRUE, caption = "Summary of research questions, corresponding hypotheses, and results.
+Statistically signficant support for/ rejection of hypotheses is indicated with 'yes'/'no', and '-' indicates no significant relationship.
+Parentheses indicate partial overall support or rejection of hypotheses across all fluxes considered.", escape = F) %>%
   kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 12) %>%
-  pack_rows("H1. FACFs decrease linearly with latitude, and in constant proportion.", 1, 3, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  pack_rows("H2. FACFs increase with MAT, with an interactive effect of MAP.", 4, 6, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  pack_rows("H3. FACFs are strongly correlated with other annual climate variables.", 7, 9, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  pack_rows("H4. FACFs are reduced under seasonal climates.", 10, 13, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  pack_rows("H5. Considering only growing season months, FACF correlates with climatic drivers.", 14, 17, latex_gap_space = "1em", colnum = 1, hline_before = TRUE) %>%
-  add_header_above(c(" ", " ", " ", "Forest autotrophic carbon fluxes (FACF)" = 9, " ")) %>%
-  column_spec(1, width = "10cm") %>%
-  column_spec(2, width = "5cm") %>%
-  column_spec(3:13, width = "1.8cm") %>%
+  pack_rows("Q1. How do C fluxes vary with latitude?", 1, 1, latex_gap_space = "1em", colnum = 5, hline_before = FALSE) %>%
+  pack_rows("Q2. How do C fluxes vary with mean annual temperature (MAT) and precipitation (MAP)?", 2, 4, latex_gap_space = "1em", colnum = 5, hline_before = TRUE) %>%
+  pack_rows("Q3. How are C fluxes related to other annual climate variables?", 5, 7, latex_gap_space = "1em", colnum = 5, hline_before = TRUE) %>%
+  pack_rows("Q4. How does seasonality influence annual C fluxes?", 8, 11, latex_gap_space = "1em", colnum = 5, hline_before = TRUE) %>%
+  pack_rows("Q5. When standardised by growing season length, how do annual C fluxes still vary with climate?", 12, 15, latex_gap_space = "1em", colnum = 5, hline_before = TRUE) %>%
+  add_header_above(c(" ",  " ", "Forest autotrophic carbon fluxes" = 9, " ")) %>%
+  #column_spec(2:11, width = "1.8cm") %>%
   kableExtra::landscape() 
+ # kableExtra::footnote(number = c("Luyssaert et al.
+(2007)", "Gillman et al.
+(2015)", "Simova and Storch (2017)", "Schuur (2003)", "Taylor et al.
+(2016)", "Malhi (2012)", "Michaletz et al.
+(2014)", "Chu et al.
+(2016)", "Piao et al.
+(2010)", "Huston & Wolverton (2009)"), footnote_as_chunk = TRUE)
 ```
@@
 
-Analyses were conducted on data contained in the open-access ForC database [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc:_2018].
-This database contains records of field-based measurements of forest carbon stocks and annual fluxes, compiled from original publications and existing data compilations and databases.
+*Forest carbon flux data*
+
+This analysis focused on nine C flux variables included in the open-access ForC database [Table 2; @anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018]. 
+ForC contains records of field-based measurements of forest carbon stocks and annual fluxes, compiled from original publications and existing data compilations and databases. 
 Associated data, such as stand age, measurement methodologies, and disturbance history, are also included.
-The database was significantly expanded since the publication of [@anderson-teixeira_forc:_2018] through integration with the Global Soil Respiration Database [@bond-lamberty_global_2010].
-Additional targeted literature searches were conducted to identify any further available data on primary productivity, with particular focus on mature forests in temperate and boreal regions.
-ForC currently contains 29730 records from 4979 plots, representing 20 distinct ecozones across all forested biogeographic and climate zones.
+The database was significantly expanded since the publication of @anderson-teixeira_forc_2018 through integration with the Global Soil Respiration Database [@bond-lamberty_global_2010].
+Additional targeted literature searches were conducted to identify further available data on the fluxes analyzed here, with particular focus on mature forests in temperate and boreal regions, which were not included in the review of @anderson-teixeira_carbon_2016.
 We used ForC v3.0, archived on Zenodo with DOI 10.5281/zenodo.3403855. 
+This version contained 29,730 records from 4,979 plots, representing 20 distinct ecozones across all forested biogeographic and climate zones.
+From this, we drew 1,319 records that met our criteria, as outlined below (Fig. 1).
+
+This analysis focused on mature forests with no known history of significant disturbance or management. 
+There is evidence that stand age influences patterns of C flux and allocation in forest ecosystems, and can confound relationships between latitude and primary productivity [@de_lucia_forest_2007; @gillman_latitude_2015].
+To reduce any biasing effects of stand age, we included only stands of known age $\ge$ 100 years and those described by terms such as "mature", "intact", or "old-growth".
+Since management can alter observed patterns of C cycling [@simova_enigma_2017], sites were excluded from analysis if they were managed, defined as plots that were planted, managed as plantations, irrigated, fertilised or included the term "managed" in their site description.
+Sites that had experienced significant disturbance within the past 100 years were also excluded.
+Disturbances that qualified sites for exclusion included major cutting or harvesting, burning, flooding, drought and storm events with site mortality >10% of trees.
+Grazed sites were retained.
+
+```{r eval = TRUE, echo=FALSE, warning=FALSE}
+library(knitr)
+library(kableExtra)
+FACF_table <- read.csv("FACF_variables.csv", stringsAsFactors = FALSE, check.names = FALSE)
+names(FACF_table) [6]  <- paste0(names(FACF_table) [6], footnote_marker_symbol(1))
+kable(FACF_table, format = "latex", booktabs = TRUE, caption = "Definitions and sample sizes of carbon flux variables used in analysis.
+All variables are in units of Mg C ha$^{-1}$ yr$^{-1}$.
+", escape = F) %>%
+  kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 12) %>%
+  column_spec(2, width = "4cm") %>%
+  column_spec(3, width = "7cm") %>%
+  column_spec(4, width = "10cm") %>%
+  column_spec(5:6, width = "1.5cm") %>%
+  add_header_above(c(" ",  " "," ",  " ", "Sample size" = 2)) %>%
+  kableExtra::footnote(symbol = c("Geographic areas group geographically proximate sites, defined using a hierarchical cluster analysis on the distance matrix of the sites, and a cutoff of 25km"), footnote_as_chunk = FALSE)
+```
 
 ```{r echo=FALSE, out.width='100%', fig.cap = "Map showing all data used in the analysis, coded by variable. Variables are plotted individually in Fig. S1. ", fig.pos='H'}
-  knitr::include_graphics("distribution_all_variables_cropped.png")
+  knitr::include_graphics("distribution_all_variables_cropped.jpg")
 ```
-*Data selection.* Over 50 variables of forest carbon stocks and annual fluxes are represented in the ForC database; this analysis focused on autotrophic carbon fluxes.
-Focal variables of this analysis are listed in Table 2.
-
-A subset of the ForC database was generated for the purposes of this analysis, in order to control for data quality and remove biasing factors.
-Since management can alter observed patterns of primary productivity [@simova_enigma_2017], sites were excluded from analysis if they were managed, defined as plots that were planted, managed as plantations, irrigated, fertilised or including the term "managed" in their site description.
-Sites that had experienced significant disturbance were also excluded.
-Disturbances that justified site exclusion were major cutting or harvesting, and/or burning, flooding, drought and storm events with site mortality >10% of trees.
-Grazed sites were retained.
-There is evidence that stand age influences patterns of primary productivity and carbon allocation in forest ecosystems, and can confound relationships between latitude and primary productivity [@de_lucia_forest_2007; @gillman_latitude_2015].
-To reduce any biasing effects of stand age, we included only stands of known age $\ge$ 100 years and those described by terms such as "mature", "intact", or "old-growth".
 
 \newpage  
-| Variable | Definition | Components included | Methodologies used | Number of records | Number of geographic areas |
-|--------------|-------------------|-------------|------------------------|-------|-------|
-| $GPP$ | Gross Primary Production | NA | flux partitioning of eddy-covariance; $NPP$+$R_{auto}$ | 243 | 49 |
-| $NPP$ | Net Primary Production | stem, foliage, coarse root, fine root, optionally others (e.g., branch, reproductive, understory) | $ANPP$ + $BNPP$ (majority); $GPP$-$R_{auto}$ | 92 | 42 |
-| $ANPP$ | Aboveground Net Primary Production | stem, foliage, optionally others (e.g., branch, reproductive, understory) (**issues [52](https://github.com/forc-db/Global_Productivity/issues/52), [54](https://github.com/forc-db/Global_Productivity/issues/54)**) |$ANPP_{woody-stem}$ + $ANPP_{foliage}$ (+ others) | 256 | 80 |
-| $ANPP_{woody-stem}$ | Net Primary Production of woody stems | woody stems down to DBH $\le$ 10cm (no branch turnover) | stem growth measurements scaled to biomass using allometries (**with recruitment?**) | 264 | 96 |
-| $ANPP_{foliage}$ | Net Primary Production of foliage | foliage | litterfall collection (separated into components) | 98 | 49 |
-| $BNPP_{root}$  | Belowground Net Primary Production | coarse and fine roots |  coarse roots estimated indirectly using allometries based on aboveground stem increment measures ; fine roots as below | 101 | 48 |
-| $BNPP_{fine.root}$ | Net Primary Production of fine roots | fine roots | measurements combined one or more of the following: minirhizotrons, fine root turnover, soil cores, root ingrowth cores | 88 | 41 |
-| $R_{root}$ | Root Respiration | coarse and fine roots | *measurement of root gas exchange, root exclusion from soil respiration chambers* | 64 | 26 |
-| $R_{auto}$ | Autotrophic Respiration | foliage, stem, and root | chamber measurements of component gas exchange | 22 | 13 |
+*Climate data*
 
-Table: Definitions and sample sizes of variables (all annual) used in analysis.
-Geographic areas group geographically proximate sites, defined using a hierarchical cluster analysis on the distance matrix of the sites, and a cutoff of 25km.
+ForC contains geographic coordinates associated with each measurement record and, when available, $MAT$ and $MAP$ as reported in the primary literature [@anderson-teixeira_forc_2018].
+Based on the geographic co-ordinates for each site, data on twelve climate variables -- including $MAT$, $MAP$, temperature seasonality (*i.e.*, standard deviation across months), precipitation seasonality (*i.e.*, coefficient of variation across months), annual temperature range, solar radiation, cloud cover, annual frost and wet days, potential evapotranspiration ($PET$), aridity ($MAP/PET$), and vapor pressure deficit ($VPD$) -- were extracted from five open-access climate datasets: WorldClim [@hijmans_very_2005], WorldClim2 [@fick_worldclim_2017], the Climate Research Unit time-series dataset (CRU TS v4.03 [@harris_updated_2014], the Global Aridity Index and Potential Evapotranspiration Climate Database [@trabucco_global_2019], and TerraClimate [@abatzoglou_terraclimate_2018] (Table S1).
+Definitions and methods used to calculate each variable are included in Table S1.
+From these data, we derived maximum $VPD$, defined as the $VPD$ of the month with the largest deficit, and the number of water stress months, defined as the number of months annually where precipitation was lower than $PET$.
+Where site-level data was missing for $MAT$ or $MAP$, we used values from the WorldClim dataset.
 
-*Methodological consistency.* The data in ForC is derived from a range of studies, often employing different methods.
-For this reason, criteria were introduced to standardise for differences in methodology.
-Where data was based on forest plot census measurements, studies which used a minimum diameter at breast height (DBH) measure of >10cm were excluded from analysis.
-It would be preferable to standardise by minimum area sampled; however x% of plots in the database are 1 ha or under in size; excluding these plots would place significant constraints on sample size.
-(**really, we didn't exclude any studies based on methodology (all studies measured min DBH 10 or less...
-) so this paragraph needs revision.
-Also, address whether we standardize for inclusion/ exclusion of recruitment, and address branchfall.**)
+Following the previous studies whose hypothesis we were evaluating [@kerkhoff_plant_2005; @michaletz_convergence_2014], length of the growing season was estimated to the nearest month, where growing season months were defined as months with mean minimum temperature > 0.5$^\circ$C.
+We experimented with a definition of growing season months including a moisture index, defined as $(MAT - PET)/PET > -0.95$ [@kerkhoff_plant_2005; see also @michaletz_convergence_2014]. 
+However, we found that including a moisture index had minimal effect on the estimates of growing season length for the sites included here, and so chose to exclude it. 
+Monthly data for $PET$, precipitation, and temperature from  CRU v 4.03 [@harris_updated_2014] and solar radiation from WorldClim2 [@fick_worldclim_2017] were used to calculate mean monthly $PET$, precipitation, temperature and solar radiation during the growing season.
 
-As discussed above, estimates of $NPP$, $ANPP$, and $BNPP$ are generated through summing measurements of their component parts.
-Since the components included in productivity estimates vary between studies, estimates of productivity were classified within the ForC database according to their components, and then filtered for analysis.
-Estimates of NPP were selected if they included foliage, branch, stem, coarse root, and fine root.
-Measures of NPP which included additional components, including understorey, volatile organic compounds (VOCs), exudates, estimates of NPP lost to herbivory (**these are all small, no need to exclude**), and the NPP of reproductive structures (**often in literfall/ mixed in with foliage...
-did we really exclude estimates including litterfall?**), were excluded.
-Estimates of ANPP were selected if they included foliage, stem growth and optionally branch turnover.
-Any measures of primary productivity where components were unknown were excluded from analysis.
+*Analyses*
 
-*Climate datasets.* Where site-level data on mean annual temperature, mean annual precipitation, and latitude were available in the primary literature, this data was compiled and entered directly into the ForC database.
-Based on the geographic co-ordinates for each site, data on a further 11 climate variables was extracted from five open-access climate datasets: WorldClim [@hijmans_very_2005], WorldClim2 [@fick_worldclim_2017], the Climate Research Unit (CRU) time-series dataset v.
-4.03 [@harris_updated_2014], the Global Aridity Index and Potential Evapotranspiration Climate Database [@trabucco_global_2018], and TerraClimate [@abatzoglou_terraclimate_2018] (see Supplementary Information S1 for details of climate variables).
-Where site-level data was missing for mean annual temperature and/or mean annual precipitation, data was extracted from the WorldClim dataset.
+The effects of latitude and climate on C fluxes were analysed using mixed effects models using the package 'lme4' [@bates_fitting_2015] in R v.3.5.1 [@r_core_team_2018]. 
+The basic model for all analyses included a fixed effect of latitude or climate and a random effect of plot nested within geographic area. 
+Geographic areas--*i.e.*, spatially clustered sites--were defined within ForC using a hierarchical cluster analysis on the distance matrix of the sites and a cutoff of 25km [@anderson-teixeira_forc_2018].
+We experimented with inclusion of altitude as a fixed effect, but excluded it from the final models because it added very little explanatory power -- that is, the difference in AIC ($\Delta AIC$) relative to models excluding altitude was generally small (often $\Delta AIC$<2).
+Effects were considered significant when inclusion of the fixed effect of interest resulted in p $\le$ 0.05 and $\Delta AIC$ $\ge$ 2.0 relative to a corresponding null model. 
+All $R^2$ values presented here are marginal $R^2$ values, and refer to the proportion of variation explained by only the fixed effects. 
+Specific analyses are as described below.
 
-Additionally, two climate variables were derived from the above datasets: maximum vapour pressure deficit, defined as the vapour pressure deficit of the month with the largest deficit; and water stress months, defined as the number of months annually where precipitation was lower than potential evapotranspiration (PET; **HML:defined how?**).
+We first examined the relationship between latitude and C fluxes (*Q1*; Table 1). 
+We tested models with latitude as a first-order linear, second-order polynomial, and logarithmic term. 
+For brevity, we henceforth refer to first-order linear models as "linear" and second-order polynomial models as "polynomial".
+We selected as the best model that with the highest $\Delta$ AIC relative to a null model with no fixed term, with the qualification that a polynomial model was considered an improvement over a linear model only if it reduced the AIC value by 2.0 or more.
+In addition, pairwise comparisons of $R^2$ values were carried out for a selection of pairs of C fluxes to test for differences among variables in the proportion of variation explained by latitude and climate.
+Models were run on data from sets of sites that were common to each pair, in order to account for variation in the number of data points included.
 
-*Length of growing season.* Growing season months were defined as months with mean minimum temperature > 0.5$^\circ$C.
-Growing season months were initially calculated following methods used by Kerkhoff et al.
-(2005), which additionally required that growing season months had a moisture index, defined as (MAT - PET)/PET, > -0.95.
-Michaletz et al.
-(2014) included an equivalent requirement in their calculation of growing season length.
-However, we found that including this requirement had no effect on the estimates of growing season length, and so chose to exclude it.
+To test whether trends in component fluxes across latitude sum to match those of larger fluxes, regression lines for smaller component fluxes were summed to generate new estimates of larger fluxes. 
+Because no fluxes were significantly better predicted by a logarithmic or polynomial fit than by a linear fit, we used linear fits for all fluxes in this analysis.
+We then determined whether these summed predictions fell within the 95% CI for the larger flux across the entire latitudinal range.
+Confidence intervals for the line of best fit for the larger flux were estimated using the 'bootMer' function, a parametric bootstrapping method for mixed models [@bates_fitting_2015].
+This function carried out 2000 simulations estimating the line of best fit, using quantiles at 0.025 and 0.975 to estimate 95% CIs.
+This analysis was applied to the following sets of fluxes: (1) $GPP = NPP + R_{auto}$, (2) $NPP = ANPP + BNPP$, and (3) $ANPP = ANPP_{foliage} + ANPP_{stem}$.
+In addition, we estimated total belowground C flux (TBCF, not analyzed due to limited data) as $TBCF = BNPP + R_{root}$.
 
-Monthly data for PET, precipitation, and temperature was downloaded from the Climate Research Unit (CRU) time-series dataset v 4.03 [@harris_updated_2014], and for solar radiation from WorldClim2 [@fick_worldclim_2017], and used to calculate mean monthly PET, precipitation, temperature and solar radiation during the growing season.
-Total growing season precipitation and solar radiation were also calculated.
+Variation in allocation to component carbon fluxes was explored for three groupings: (1) $GPP = NPP + R_{auto}$, (2) $NPP = ANPP + BNPP$, and (3) $ANPP = ANPP_{foliage} + ANPP_{stem}$.
+For each group, measurements taken at the same site and plot, and in the same year, were grouped together.
+For groups (1) and (2), where 2 of the 3 flux measurements were available for a given site, plot, and year, these measurements were used to calculate the third.
+The ratio of each pair of component fluxes was calculated.
+The log of these ratios were regressed against latitude and climate variables, using the linear model specified above.
+Cook's distance analyses were carried out for each of the models, and extreme outliers removed.
 
-*Model specification.*
+We next examined the relationships of C fluxes to climate variables (*Q2-Q4*; Table 1). 
+We tested first-order linear, second-order polynomial, and logarithmic fits for each climate variable. 
+Again, polynomial fits were considered superior to first-order linear fits only if inclusion of a second-order polynomial term resulted in $\Delta$AIC $\ge$ 2.0 relative to a first-order linear model.
+We tested relationships of each C flux (Table 2) against each climate variable (Table S1).
+Variables which were not significant explanatory variables or which explained <20% of variation in C fluxes are only presented in SI.
 
-{**from Helene: It was frequently unclear to me whether the influences of other climate variables were being considered in multivariate or univariate analyses.
-This needs to be clear.
-I can see reasons to do it either way, but they need to be differently motivated and clearly presented.**}
+Linear models were used to investigate the potential joint and interactive effects of $MAT$ and $MAP$ on carbon fluxes.
+An additive model including $MAP$ in addition to $MAT$ was accepted when $\Delta$AIC >2 relative to a null including only $MAT$ as a fixed effect.
+An interactive model containing a $MAT$ x $MAP$ interaction was accepted when $\Delta$AIC >2 relative to a null including $MAT$ and $MAP$ as fixed effects.
 
-The effects of climate and latitude on primary productivity were analysed using mixed effects models using the package 'lme4' [@bates_fitting_2015] in R v.3.5.1 [@r_core_team_r:_2018].
-The effect of each extracted climate variable on each measure of primary productivity was modelled by specifying the climate variable as a fixed effect.
-For each climate variable, three models were specified: a null model; a model with the climate variable as a linear term; and a model with the climate variable as a **XX**-order polynomial term.
-AIC values were calculated for the models and used to select the best model.
-If the best model included a polynomial term, the shape of the polynomial relationship was considered.
-If the shape of the relationship *made biological sense*, and was a significant improvement on the linear relationship (deltaAIC >2), we accepted the polynomial as the best model.
-If not, we ran the linear model as the final model.
-$R^2$ values were calculated for the best model.
-All $R^2$ values presented here are marginal $R^2$ values, and refer to the proportion of variation explained by only the fixed effects, unless otherwise specified.
-In addition, slope coefficients were calculated for the linear models.
+To test whether and how C fluxes varied with climate when standardised by growing season length (*Q5*; Table 1), we first standardized all annual C fluxes by dividing by growing season length (as defined above).
+We then derived four variables to describe growing season climate, specifically growing season temperature, precipitation, solar radiation, and PET (Table S1).
+We tested for correlations between these standardised fluxes and growing season climate variables, using only first-order linear models.
 
-To test for a potential influence of altitude, models were also run with site altitude included as a second fixed effect.
-These models were compared against models with no altitude term, and AIC values calculated to identify whether inclusion of altitude as a term improved the models.
-Including altitude had a very small effect on most models, with the difference in AIC values between models including and excluding altitude often being <2, suggesting the models are very similar in their explanatory power.
-As a result, it was decided to present results only from models without altitude as a fixed term.
+All analyses were conducted in R v.3.5.1 [@r_core_team_2018].
+Code and data necessary to reproduce all results are available through GitHub (https://github.com/forc-db/Global_Productivity) and archived in Zenodo (DOI: TBD).
 
-Within the ForC database, sites within 25km^2^ of each other are clustered into geographic areas.
-To account for correlations in measurements between tightly clustered sites, a random effect was specified as plot nested within geographic area.
-*Data from the temperate regions was heavily skewed towards studies from the old-growth forests of the Pacific Northwest.*  These forests have very high FACF, and so to ensure that results were not unduly influenced by geographic sampling bias, we tried a version of the model where data were weighted according to forested land area within each Koeppen climate zone.
-Results were similar between the weighted and unweighted model.
-Because the weighted model may be prone to problems of over-fitting, we present only results of the unweighted model.
-{**Becky, I'd probably just delete this whole paragraph.
-Since we ended up not using the weighted model, it's just extra info for readers to digest.}
+### Results
 
-Models were run for total annual FACF against annual climate variables, and for fluxes standardized by growing season length, defined as annual FACF/length of growing season, against growing season climate variables.
-For analyses on data within the growing season, only linear models were specified.
+In total, we analyzed 1,319 records from nine forest autotrophic C flux variables taken from forests that had experienced no major anthropogenic disturbances within the past 100 years.
+These records represented a total of 255 plots in 154 distinct geographic areas across all forested biogeographic and climate zones (Figs.
+1, S1;  Table 2).
 
-{**Note from Helene on the following paragraph: this makes it sound like analyses from H3 were about variation remaining after accounting for temperature, but analyses below don't seem to do that.**}
-To investigate the potential interactive effects of climate variables on carbon fluxes, multivariate models were also specified.
-To ensure that models were biologically meaningful, the terms included in the models tested built on results from the univariate models.
-Modelling of individual climate variables identified that the best predictors of carbon fluxes were variables related to temperature.
-We therefore decided to include mean annual temperature as a term in all multivariate models.
-We first modelled the interaction effect between mean annual temperature and mean annual precipitation, in order to capture climate variation along the axes of temperature and water availability.
-Models were tested for a significant interactive effect and a significant additive effect.
-We then explored whether any other climate variable, in combination with mean annual temperature, could significantly improve on the combination of mean annual temperature and mean annual precipitation.
-In specifying the range of models to test, climate variables which were strongly correlated with temperature were dropped, in order to capture the greatest range of variation in climate.
-For each possible pairing of climate variables, two models were specified: a model with the two climate variables showing an additive effect; and a model with the two climate variables showing an interactive effect.
-As before, plot nested within geographic area was included as a random effect.
-Altitude was not considered.
-AIC values were calculated for the models, and used to compare models.
-Models were considered to be significantly better than the baseline MAT*MAP model if:
-  i) the AIC value of the model was smaller than the AIC value of the baseline model by >2  
-  ii) the r-squared value of the model was larger than the r-squared value of the baseline model by **>5 (???- not possible)**
+*Q1. How does C flux vary with latitude?*
 
-*Validating models of component fluxes.* Comparison of component fluxes is based on the assumption that components sum accurately to estimates of larger fluxes.
-To test this, components of larger fluxes were regressed against latitude, and the models used to generate a series of point estimates along lines of best fit for each component.
-The point estimates for smaller component fluxes were summed to generate new "stacked" estimates of larger fluxes, which were then compared against actual measurements of the larger flux.
-Confidence intervals for the larger flux were calculated using the 'bootMer' function from the lme4 package [@bates_fitting_2015].
-Stacked plots were generated for:
-1. GPP = NPP + R~auto~  
-2. NPP = ANPP + BNPP  
-3. ANPP = ANPP~foliage~ + ANPP~woody~ ~stem~  
-4. Total belowground carbon flux = BNPP + R~root~
-
-*Allocation to carbon fluxes along latitudinal gradients.* Variation in allocation to component carbon fluxes along latitudinal gradients was explored for a range of pairings: firstly, GPP:NPP, ANPP:BNPP, and ANPP~foliage~:ANPP~woody~ ~stem~; and secondly, the ratio of NPP to each of ANPP, BNPP, ANPP~foliage~, and ANPP~woody~ ~stem~.
-For each set of paired fluxes, measurements taken at the same site and plot, and in the same year, were paired together, and the ratio of each pair of measurements calculated.
-The ratios were regressed against latitude and climate variables, using the linear model specified above.
-Cook's distance analyses were carried out for each of the models, and indicated that data from a few high-elevation sites were having a disproportionate influence on the regressions.
-To account for this, models were re-run using only data from sites $\le$ 1000m.
-
-###Results
-
-In total, we analyzed 1228 records from 9 C flux variables taken from forests that had experienced no major anthropogenic disturbances within the past 100 years.
-These records represented a total of 154 distinct geographic areas (Fig.
-1, Table 2), across all forested biogeographic and climate zones.
-
-*How does productivity vary with latitude?*
-
-All major carbon fluxes decreased linearly with latitude (Fig. 2).
-
-Latitude was a strong predictor for many of the carbon fluxes, explaining 64% of variation in GPP (n = 254, p<0.0001), 50% in NPP (n = 114, p<0.0001) and 45% in ANPP (n = 259, p<0.0001).
-For all fluxes, their relationship with latitude was best predicted by the linear model.
-
-*Relationships and differences among fluxes.* In general, smaller component fluxes summed approximately to larger fluxes across the latitudinal gradient (fig.
-2).
-That is, modelled estimates of GPP, generated from the sum of NPP and R auto; NPP, generated from the sum of ANPP and BNPP~root~; and ANPP, generated from the sum of ANPP~foliage~ and ANPP~woody~ ~stem~, fell completely within the confidence intervals of the regressions of field estimates of GPP, NPP and ANPP respectively.
-
-bbl: here or in the discussion, note that this is a fairly stringent test: it's easy for sub-fluxes not to sum up!
-(Extensive citations from EC literature for example.).
-HML: yes and no.
-confidence intervals are pretty large.
+All major carbon fluxes decreased with latitude (Fig.
+2; Table S2).
+Latitude was a strong predictor for many of the carbon fluxes, particularly the larger fluxes (Table S2, S6).
+Specifically, latitude explained 64% of variation in GPP (n = 243, p<0.0001), 50% in NPP (n = 161, p<0.0001) and 44% in ANPP (n = 278, p<0.0001).
+The C fluxes that were most poorly predicted by latitude were $BNPP_{fine.root}$ ($R^2$=0.17) and $ANPP_{stem}$ ($R^2$=0.18).
+The relationship with latitude was best fit by the first-order linear model, with the exception of NPP and $R_{root}$, for which a logarithmic model was a slightly -- but not significantly -- better fit.
 
 ```{r echo=FALSE, out.width='100%', fig.cap = "Latitudinal trends in forest autotropic carbon flux.
-Lines of best fit are plotted according to the best model selected during analysis.
-All regressions are significant $(p<0.05)$.
+Plotted are linear models, all of which were significant $(p<0.05)$ and had AIC values within 2.0 of the best model (for two fluxes, logarithmic fits were marginally better; Table S2).
 Each panel shows major C fluxes together with component fluxes.
@@
 
-We found no evidence that allocation between fluxes varied substantially with latitude or climate.
-There were no significant results from regressing ratios of carbon fluxes against latitude, or against any of the climate variables.
+Smaller component fluxes summed approximately to larger fluxes across the latitudinal gradient (Fig.
+2).
+That is, modeled estimates of $GPP$, generated from the sum of $NPP$ and $R_{auto}$; $NPP$, generated from the sum of $ANPP$ and $BNPP$; and $ANPP$, generated from the sum of $ANPP_{foliage}$ and $ANPP_{stem}$, fell almost completely within the confidence intervals of the regressions of field estimates of $GPP$, $NPP$, and $ANPP$, respectively.
 
-$R^2$ values were generally highest in the major fluxes, and decreased in subsidiary fluxes (Supporting Information S2).
-Of the major fluxes, R~auto~ and GPP were the most strongly explained by latitude and climate, with climate explaining at most 71% of variation in GPP, and 65% in R~auto~.
-The proportion of variation explained by climate and latitude decreased in NPP and ANPP, with climate explaining at most 51% of variation in NPP and 44% in ANPP.
-Of the major fluxes, BNPP~root~ was the least well explained by climate and latitude, with climate explaining at most 36% of variation.
+We found no evidence of systematic variation in C allocation with latitude or climate (Fig.
+S3).
+Of 16 relationships tested (4 ratios among C flux variables regressed against latitude, $MAT$, $MAP$ and temperature seasonality), none were significant.
 
-With the exception of ANPP~foliage~, the proportion of variation explained by climate and latitude in subsidiary fluxes was much lower.
-Climate explained at most 24% of the variation in ANPP~woody~ ~stem~, 19% in BNPP~fine~ ~root~, and 27% in R~root~.
-In contrast, climate strongly explained variation in ANPP~foliage~, with mean annual temperature explaining 58% of variation.
-This pattern was also seen in the $R^2$ values for multivariate models.
+*Q2. How does C flux relate to $MAT$ and $MAP$?*
 
-*How does productivity relate to MAT and MAP?* MAT and MAP are the most commonly reported site-level climate variables, and much previous research into the effect of climate on forest productivity has focused on these as key climate variables.
-MAT was a significant (p<0.05) and strong predictor of productivity for all carbon fluxes tested, with all fluxes showing a linear increase with temperature (fig.
-4).
-We found no support for a saturation point of productivity with temperature.
+All fluxes increased with $MAT$ (all p<0.05; Figs. 3-4, S4-S5, Table S2). 
+For eight of the nine fluxes, this relationship was linear.
+For only one variable, $BNPP$, did a lognormal fit provide an improvement over a first-order linear relationship, though this was not significant ($\Delta$AIC <2).
+As with latitude, $MAT$ tended to explain more variation in the larger fluxes ($GPP$, $NPP$, $ANPP$, $R_{auto}$) and $ANPP_{foliage}$ (all $R^2$> 0.4) than in subsidiary and belowground fluxes ($ANPP_{stem}$, $R_{root}$, $BNPP_{fine.root}$; all $R^2$< 0.25; Table S6).
 
-MAP was found to be a significant (p<0.05) but poor predictor of productivity, explaining, with the exception of R~auto~, at most 37% of variation in carbon flux.
-For the majority of fluxes productivity was best predicted by a polynomial model.
-Productivity increased with precipitation, up until a saturation point at between 3000 and 4000mm annual precipitation, above which productivity started to decrease (fig.
-4).
-The notable exception to this was GPP: the model indicated that GPP continued to increase with precipitation up to measures of at least 5000mm annually (p<0.0001, $R^2$ = 0.33.
-Data above this point was not available, but the model trend indicated that the saturation point for this model would be around 5000mm MAP.
+$MAP$ was a significant (p<0.05) predictor of all fluxes (Figs. 4a, S4-S5; Table S2). 
+However, it explained little variation: with the exception of $R_{auto}$, $MAP$ explained at most 25% of variation in C flux.
+All fluxes increased with $MAP$ up to at least 2000 mm, above which responses were variable (Figs.
+4, S4-S5).
 
-There was a significant interactive effect between MAT and MAP for GPP, BNPP~root~, BNPP~fine~ ~root~, ANPP, ANPP~woody~ ~stem~, and R~root~ (fig.3).
-There was a significant additive effect for R~auto~.
-NPP and ANPP~foliage~ showed no significant interactive or additive effect.
-{**NEED SI TABLE GIVING FITTED COEFFICIENTS & STATS**}
+There was a significant additive effect of $MAT$ and $MAP$ on $GPP$, $ANPP$ and $R_{auto}$ (Fig.
+3, Table S3), and a significant interactive effect between $MAT$ and $MAP$ for $NPP$ and $ANPP_{stem}$ (Fig.
+3, Table S3).
+The interaction was negative for $NPP$ and positive for $ANPP_{stem}$.
+For $ANPP_{foliage}$, $BNPP$, $BNPP_{fine.root}$, and $R_{root}$, $MAP$ did not have a significant effect when accounting for $MAT$ (Fig.
+3, Table S3).
 
-For the variables which showed a significant interactive or additive effect between MAT and MAP, no other climate variable, in combination with MAT, significantly improved on that model.
-For NPP, there was a significant interactive effect between MAT and water stress months, with this model explaining nearly 5% more variation in NPP than MAT alone {**Helene: but less than MATxMAP?
-if so, why bother mentioning?**}.
-However, for ANPP~foliage~, no multivariate model improved on the univariate model including only MAT.
-
-```{r echo=FALSE, out.width='100%', fig.cap = "Interactive effects of mean annual temperature and mean annual precipitation on FACF.
+```{r echo=FALSE, out.width='100%', fig.cap = "Interactive effects of mean annual temperature and precipitation on annual forest carbon fluxes.
 For visualization purposes, data points are grouped into bins of 0 - 1000, 1001 - 2000, 2001 - 3000, and >3000mm mean annual precipitation, and lines of best fit models are plotted for mean annual precipitation values of 500, 1500, 2500, and 3500mm.
@@
 
-*How does productivity relate to other climate variables?* {**Here, maybe add some statement about covariation among climate variables, site SI figure**} Our results indicated that productivity was most strongly explained by temperature at the global scale, with temperature-related climate variables coming out as strong predictors of productivity.
-In addition to MAT, temperature seasonality, annual temperature range, and annual frost days were consistently identified as strong univariate predictors of FACF.
-{**cite some table/ SI?**}
+*Q3. How does C flux relate to other annual climate variables?*
 
-We found a significant relationship between productivity and potential evapotranspiration for all fluxes.
-ANPP~foliage~, BNPP~fine~ ~root~ and R~root~ increased linearly with PET; however, all other fluxes showed a polynomial relationship with PET (fig.
-4).
-We found strong evidence for a saturation point or peak with PET: productivity tended to increase at values below 1000mm, before saturating between 1200 and 1700mm.
-There was also evidence that productivity begins to decrease at values above 1800mm PET.
+All C flux variables showed a significant relationship with annual $PET$.
+The relationship was logarithmic for $ANPP_{foliage}$, $BNPP_{fine.root}$ and $R_{root}$, and polynomial for all other fluxes (Fig.
+4c, S4-5; Table S2).
+We found strong evidence for a saturation point or peak with $PET$: C fluxes tended to increase at values below 1000mm, before saturating between 1200 and 1700mm.
+There was also evidence that some C fluxes begin to decrease at values above 1800mm PET.
 
-Vapour pressure deficit was a significant predictor of productivity for all fluxes.
-BNPP~fine~ ~root~ showed a linear relationship with vapour pressure deficit ($R^2$ = 0.07, p<0.05), but all other fluxes showed a polynomial relationship (fig.
-4).
-Productivity initially increased with vapour pressure deficit, before saturating at around 0.8 kPa.
-At values above 0.8 kPa, productivity began to decrease.
+Mean annual $VPD$ was a significant predictor of all C fluxes.
+$ANPP_{foliage}$, $BNPP_{fine.root}$ and $R_{root}$ showed a logarithmic relationship with $VPD$, but all other fluxes showed a polynomial relationship (Figs.
+4d, S4-5; Table S2).
+C fluxes initially increased with $VPD$, before saturating at around 0.8 kPa, after which point they began to decrease.
 
-All fluxes, with the exception of R~root~, showed a positive linear relationship with solar radiation.
-Solar radiation explained a low proportion of variability in productivity for all fluxes, explaining less than 20% of the variation in each flux, with the exception of R~auto~ ($R^2$ = 0.26, p<0.05).
+All fluxes, with the exception of $R_{root}$, showed a significant positive relationship with solar radiation (Figs. S4-S5, Table S2). 
+Solar radiation explained a low proportion of variability (<30%) in all C fluxes.
 
-Of the climate variables tested, annual wet days, aridity, cloud cover, mean diurnal temperature range, precipitation seasonality, maximum vapour pressure deficit and water stress months were poor or non-significant explainers of variation in productivity, explaining less than 20% of the variation in each of the carbon fluxes.
+Annual wet days, cloud cover, and aridity were poor or non-significant predictors of variation in C fluxes, explaining less than 20% of the variation in each of the carbon fluxes (Figs.
+S4-S5; Table S2).
 
@@
 
-*What is the role of seasonality in explaining FACF?* Temperature seasonality was a significant predictor of FACF.
-GPP, NPP, ANPP, and R~root~ exhibited a polynomial relationship with seasonality (Fig.
-4) {**give statistic on signficiance**}.
-ANPP~foliage~, ANPP~woody~ ~stem~ and R~auto~ decreased linearly with temperature seasonality.
-Temperature seasonality was strongly correlated with annual temperature range, and, as expected, all fluxes showed almost identical responses to it {**meaning unclear**}.
-Productivity was highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower.
-In contrast, there was no significant effect of precipitation seasonality on FACF.
+*Q4. What is the role of seasonality in explaining C fluxes?*
 
-We found a significant relationship between length of growing season and FACF, with all fluxes showing a linear increase in productivity with length of growing season (Fig.
-4).
-Length of growing season was a strong predictor of FACF, explaining 51% of variation in GPP, 39% of variation in NPP, and 34% of variation in ANPP, but it was a weaker predictor than MAT for all fluxes analysed.
+Variables describing temperature seasonality -- temperature seasonality, annual temperature range, annual frost days, and length of growing season -- were strongly correlated with both latitude and $MAT$ (all r$\ge$ 0.2; Fig.
+S2), and were consistently identified as strong univariate predictors of C fluxes (Figs.
+4, S4-S7).
 
-*Within the growing season, how does FACF vary with climate?* When FACF was standardized by growing season length, correlations with climate variables were much weaker.
-Growing season temperature, precipitation, solar radiation, and PET each had a small effect on one or more of the nine FACF analyzed (Table 2; Fig.
-**S#**).
-Speficifally, $ANPP$ increased with growing season temperature ($R^2$ = 0.10, p<0.001) and precipitation ($R^2$ = 0.04, p<0.05).
-Similaryly, $ANPP_{foliage}$ increased slightly with growing season temperature ($R^2$ = 0.16, p<0.01) and precipitation ($R^2$ = 0.09, p<0.05).
-Growing season solar radiation had a positive influence on $GPP$ ($R^2$ = 0.21, p<0.001), $NPP$ ($R^2$ = 0.21, p<0.001), $BNPP$ ($R^2$ = 0.16, p<0.001) and $BNPP_{fine.root}$ ($R^2$ = 0.12, p<0.01).
-Growing season PET had a positive influence on $GPP$ ($R^2$ = 0.15, p<0.01), $NPP$ ($R^2$ = 0.18, p<0.01), $BNPP$ ($R^2$ = 0.23, p<0.0001), $BNPP_{fine.root}$ ($R^2$ = 0.11, p<0.05), and $ANPP_{woody-stem}$ ($R^2$ = 0.06, p<0.05).
-{**Becky, please verify/ edit the following:** There were no other signifcant correlations between growing season length-standardized FACF (9 variables in Table 2) and growing season climate (**which variables?**)}.
+All fluxes decrease with increasing temperature seasonality, though the shape of this relationship varies (all p<0.05; Figs. 4e, S6-7; Table S2). 
+Temperature seasonality was strongly correlated with annual temperature range, which was likewise a similarly strong predictor of C fluxes (Table S2). 
+C fluxes were highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower (*i.e.*, in the tropics).
 
-###Discussion
+In contrast, there was no significant effect of precipitation seasonality on C fluxes at this global scale.
+Both maximum vapour pressure deficit and water stress months were poor or non-significant predictors of variation in C fluxes (Figs. S6-S7; Table S2).
 
-We used a large global database (ForC), containing an unprecedented amount of data representing all major forest types (Fig.
-1) and the nine most significant forest autotrophic carbon flux (FACF) variables (Table 2), to comprehensively explore how C cycling in mature, undisturbed forests varies with latitude and climate on a global scale.
-We test a suite of hypotheses, including many previously posed (Table 1), standardising for factors that have not always been controlled for (*e.g.*, stand age, flux components measured) to gain a stronger understanding of climatic effects on FACF.
+We found a significant relationship between length of growing season and C fluxes, with all fluxes showing a positive relationship with length of growing season (Figs.
+4e, S6-S7; Table S2).
+Length of growing season was a strong predictor of C fluxes, explaining 53% of variation in GPP, 38% of variation in NPP, and 34% of variation in ANPP (all p<0.05; Table S2), but it was a weaker predictor than $MAT$ for all fluxes analysed (Table S4).
 
-bbl: Most of these two paragraphs (following) just restate the results. Honestly I would remove them enitrely, but at the very least, condense into a single paragraph.
+*Q5. Within the growing season, how do C fluxes vary with climate?*
 
-We show that, across all nine FACF variables analyzed, C cycling decreases linearly with latitude (*H1.1*; Fig.
-2)--a finding that confirms multiple previous studies (**REFS**) but contradicts the idea that productivity of temperate forests rivals that of tropical forests (*H1.1.alt*; **REFS**).
-The FACF variables increase in proportion to one another (*H1.2*), with no differences in allocation detectable at this global scale and with component fluxes summing appropriately to larger fluxes (Fig.
-2), indicating no major, systematic omissions or overestimations of flux components.
-However, we did detect a tendency for greater variability among subsidiary C fluxes (e.g., *$ANPP_{woody}$*, *$BNPP_{fine.root}$*; Fig.
-2; **some SI table?**).
+When annual C fluxes were standardized by growing season length (in monthly increments), correlations with growing season climate were generally weak (Figs. S8-S9).
+$ANPP$ increased with growing season temperature ($R^2$ = 0.09, p<0.001) and precipitation ($R^2$ = 0.04, p<0.05). 
+Similarly, $ANPP_{foliage}$ increased slightly with growing season temperature ($R^2$ = 0.16, p<0.01) and precipitation ($R^2$ = 0.09, p<0.05). 
+Growing season solar radiation was positively correlated with on $BNPP$ ($R^2$ = 0.17, p<0.001) and $BNPP_{fine.root}$ ($R^2$ = 0.13, p<0.01). 
+Growing season PET had a positive influence on $GPP$ ($R^2$ = 0.15, p<0.01), $NPP$ ($R^2$ = 0.07, p<0.01), $BNPP$ ($R^2$ = 0.23, p<0.0001), $BNPP_{fine.root}$ ($R^2$ = 0.10, p<0.05), and $ANPP_{stem}$ ($R^2$ = 0.06, p<0.05).
+All other relationships were non-significant.
 
-**Climate explains a significant proportion of variation in all C fluxes analyzed** (albiet less for subsidiary fluxes), with temperature variables the best predictors of FACF at this global scale (Figs.
+### Discussion
+
+Our analysis of a large global database (ForC) clarifies how autotrophic C fluxes in mature forests vary with latitude and climate on a global scale.
+We show that, across all nine variables analyzed, annual C flux decreases continually with latitude (Fig.
+2), a finding that confirms multiple previous studies but contradicts the idea that productivity of temperate forests rivals or even exceeds that of tropical forests [@luyssaert_co_2007; @huston_global_2009].
+At this global scale, C fluxes increase approximately in proportion to one another, with component fluxes summing appropriately to larger fluxes and no detectable differences in allocation across latitude or climates (Figs.
+2, 4, S3).
+Similarly, we show broad - *albeit* not complete - consistency of climate responses across C fluxes, with the observed latitudinal variation primarily attributable to temperature and its seasonality (Figs.
 3-4).
-While other climate variables are significant predictors of FACF (*H3*), none of them in isolation improve on the explanatory power of temperature-related variables in general or MAT specifically (Fig.
-4).
-Water availability is an important factor in explaining FACF on a global scale: we find a positive influence of precipitation at low MAP, with saturation at higher levels of MAP (Fig.
-4b) and a significant interaction between MAT and MAP for seven of the nine variables analyzed (Fig.
-3).
-Temperature seasonality and growing season length are closely correlated with MAT and are strong predictors of FACF (*H4*; Fig.
-4e-f), though growing season length doesn't improve upon MAT as a predictor.
-Within the growing season, the influence of climate on C cycling is smaller but still significant for a number of carbon fluxes (*H5*; **some SI table?**).
-These findings clarify the big picture of how FACF varies with latitude and climate on a global scale.
+Water availability is also influential, but of secondary importance across the climate space occupied by forests (Figs. 3-4). 
+Contrary to prior suggestions that the majority of variation in C cycling is driven primarily by the length of the growing season [@kerkhoff_plant_2005; @enquist_adaptive_2007; @michaletz_convergence_2014], we find modest explanatory power of growing season length and small but sometimes significant influence of climate within the growing season (Figs.
+4f,S6-S9).
+Together, these findings yield a unified understanding of climate's influence on forest C cycling.
 
-Past studies have differed in their conclusions regarding the relationship between productivity and latitude or its correlates (Table 1, *H1*; **REFS**)--quite possibly because of lack of standardization with respect to stand age and disturbance history.
-Our findings indicate that, among mature, undisturbed stands, FACF is unambiguously highest in the tropical regions, and the relationship is approximately linear (Fig.
-2).
-This contrasts with suggestions that productivity of temperate forests is similar to that of tropical forests (**REFS**).
-Temperate forests tend to be younger than tropical forests (**REF**), so analyses comparing across latitudinal gradients without controlling for stand age risk confounding age with biome effects ([Poulter et al.
-2018-DOI:10.1594/PANGAEA.889943](http://pure.iiasa.ac.at/id/eprint/15361/); Potapov et al.
-2008.
-Mapping the World’s Intact Forest Landscapes by Remote Sensing.
-Ecology and Society 13 (2), 51.
-).
-In addition, because C allocation varies with stand age [@de_lucia_forest_2007] (**See Nobby's comment in manuscript-draft_NK.pdf**), age differences may introduce systematic biases into analyses of FACF across latitude or global climatic gradients.
-For example, woody productivity tends to be higher in rapidly aggrading secondary stands than in old-growth forests, where proportionally more C is allocated to respiration (** Nobby AFM paper: Understanding the controls over forest carbon use efficiency on small spatial scales: Effects of forest disturbance and tree diversity**) [*purpose for respiration/ other compenents (**See Nobby's comment in manuscript-draft_NK.pdf**)].
+Our findings indicate that, among mature, undisturbed stands, forest C fluxes are unambiguously highest in the tropical regions, and the relationship with both latitude and $MAT$ is approximately linear (Table 1, *Q1*,*Q2*; Figs.
+2, 4).
+This contrasts with the suggestion that C fluxes (e.g., $NPP$, $ANPP$, $ANPP_{stem}$) of temperate forests are similar to or even greater than that of tropical forests [@luyssaert_co_2007; @huston_global_2009].
+Previous indications of such a pattern may have been an artifact of differences in stand age across biomes.
+Compared to tropical forests, the temperate forest biome has experienced more widespread anthropogenic disturbance and has a larger fraction of secondary stands [@potapov_mapping_2008; @yu_high_2014; @poulter_global_2018], so analyses comparing across latitudinal gradients without controlling for stand age risk confounding age with biome effects.
+Because carbon allocation varies with stand age [@de_lucia_forest_2007; @anderson-teixeira_altered_2013; @yu_high_2014; @doughty_what_2018], age differences may introduce systematic biases into analyses of C fluxes across latitude or global climatic gradients.
+For example, woody productivity tends to be higher in rapidly aggrading secondary stands than in old-growth forests, where proportionally more C is allocated to respiration and non-woody productivity [@de_lucia_forest_2007; @piao_forest_2010; @doughty_what_2018; @kunert_understanding_2019].
+Thus, findings that temperate forest productivity rivals that of tropical forests are likely an artifact of different forest ages across biomes.
 
-We show that FACF are broadly consistent in their responses to climate drivers on the global scale (with the exception of some differences in MAT-MAP interactions; Fig.
-3), with no major trends in C allocation among the variable pairs tested (Table 1, *H1*; Fig.
-2; **Some SI table**).
-Although variation in allocation has been observed along gradients of elevation [@moser_elevation_2011] and water availability [@newman_above-_2006]--along with non-climatic axes of stand age [@litton_carbon_2007], nutrient availability [@litton_carbon_2007; @gill_belowground_2016], and forest structure [@taylor_greater_2019]--variation in relation to climate is not apparent at the global scale within ForC, which contains the bulk of relevant data.
-Our conclusion, then, is that hypothesized gradients in allocation along global climate gradients cannot currently be supported for mature forests, although data quantity and standardization is currently insufficent to rule out the possibility that such trends exist.
+We show that C fluxes are broadly consistent in their responses to climate drivers on the global scale, with no trends in C allocation among the variable pairs tested (Figs. 2, S3).
+This parallels the observation that C allocation across multiple C fluxes varies little with respect to climate along a steep tropical elevational gradient [@malhi_variation_2017; but see @moser_elevation_2011],
+and is not surprising given that carbon allocation within forest ecosystems is relatively constrained [@enquist_global_2002; @litton_carbon_2007; @malhi_allocation_2011]. 
+We find no trend in the allocation of $GPP$ between production and respiration across latitude or climate ($NPP$:$R_{auto}$; Fig.
+S3), refuting the idea that tropical forests have anomalously low $CUE$ [@de_lucia_forest_2007; @malhi_productivity_2012; @anderson-teixeira_carbon_2016].
+Rather, differences in $CUE$ between old-growth tropical forests relative to (mostly younger) extratropical forests are likely an artifact of comparing stands of different age, as $CUE$ is known to decline with forest age [@de_lucia_forest_2007; @piao_forest_2010; @collalti_forest_2020].
+Another previously observed pattern for which we find no support is a tendency for belowground C allocation to decrease with increasing temperature  [@moser_elevation_2011; @gill_belowground_2016]; rather, we observe no trends in allocation between $ANPP$ and $BNPP$ across latitudes.
+Failure to detect significant tends in C allocation with respect to climate in this analysis does not imply that none exist; rather, it suggests that, at this global scale, differences are subtle and/or that more careful methodological standardization is required to detect them.
 
-bbl: remove this sentence?
-Of particular interest and significance are the relationships amongst $GPP$, net primary productivity ($NPP$ and its components, particularly $ANPP_{woody-stem}$), and respiration ($R_{auto}$ and components).
-There have been suggestions that tropical forests tend to have low carbon use efficiency ($CUE$= $NPP$/$GPP$=($GPP$-$R_{auto}$/$GPP$), which are based on observations of low $CUE$ in old-growth tropical forests relative to (mostly younger) extratropical forests [@de_lucia_forest_2007; [Malhi 2012](https://www.researchgate.net/publication/227837530_The_productivity_metabolism_and_carbon_cycle_of_tropical_forest_vegetation); @anderson-teixeira_carbon_2016], but our analysis suggests that these low values might more appropriately be attributed to the fact that these forests are old than to their tropical climate.
-Indeed, $CUE$ is known to decline with forest age ([@de_lucia_forest_2007]; [Collalti & Prentice 2019](https://academic.oup.com/treephys/article-abstract/39/8/1473/5423351?redirectedFrom=fulltext)) but appears to be roughly independent of $GPP$ [@litton_carbon_2007].
-Among our sites with relevant data, there is indication that CUE or $ANPP_{woody-stem}$/$GPP$ increase with latitude (**some SI table**).
-Additional measurements with careful methodological standardization across a consistent set of mature forest sites will be necessary to determine whether any climate-driven gradients in allocation exist at the global scale.
+**in the paragraph above, discuss/ cite @collalti_forest_2020**
 
-One interesting observation was that climate tends to explain more variation in the major fluxes ($GPP$, $NPP$, $R_{auto}$ - latitude $R^2\ge$ 48%) than in subsidiary fluxes (latitude $R^2$ <27% for $BNPP_{fine.root}$, $R_{auto-root}$,$ANPP_{woody-stem}$) (Fig.
-2; **some SI table?**).
+Despite the broad consistency of climate responses across C fluxes, climate explains lower proportions of variability among some of the subsidiary C fluxes (*e.g.*, $ANPP_{stem}$, $BNPP$,  $BNPP_{fine.root}$; Fig.
+2; Tables S2, S6).
 There are two, non-exclusive, potential explanations for this. 
 First, it may be that methodological variation is larger relative to flux magnitude for some of the subsidiary fluxes. 
-Belowground fluxes in particular are difficult to quantify, and measurement methods for the belowground fluxes considered here may be measured through fundamentally different approaches (*e.g.*, minirhizotrons, ingrowth cores, or sequential coring for $BNPP_{root-fine}$; root exclusion, stable isotope tracking, or gas exchange of excised roots for $R_{auto-root}$), and sampling depth is variable and often insufficient to capture the full soil profile.
-$ANPP_{woody-stem}$, which is also poorly explained by latitude or climate, is more straightforward to measure but is subject to variability introduced by differences such as biomass allometries applied and minimum plant size sampled (bbl: cite e.g.
-Huntzinger?).
-However, methodological variation and uncertainty affect all of fluxes considered here--not necessarily any less than the aforementioned, and some of the larger fluxes that vary more strongly with respect to climate ($ANPP$, $NPP$) are estimated by summing uncertain component fluxes.
-Second, differences among variables in the proportion of variation explained by climate may be attributable to more dicrect climatic control over $GPP$ than subsidiary fluxes.
-That is, subsidiary fluxes may be shaped by climate both indirectly through its influence on $GPP$ and respiration and directly through any climatic influence on C allocation, as well as many other local- and regional-scale factors (**REFS**).
+Belowground fluxes in particular are difficult to quantify, and measurement methods for the belowground fluxes considered here may use fundamentally different approaches in different sites (*e.g.*, minirhizotrons, ingrowth cores, or sequential coring for $BNPP_{fine.root}$; root exclusion, stable isotope tracking, or gas exchange of excised roots for $R_{root}$), and sampling depth is variable and often insufficient to capture the full soil profile.
+$ANPP_{stem}$, which is also poorly explained by latitude or climate, is more straightforward to measure but is subject to variability introduced by differences such as biomass allometries applied and minimum plant size sampled [@clark_measuring_2001].
+However, methodological variation and uncertainty affect all of fluxes considered here, and some of the larger fluxes that vary more strongly with respect to climate ($ANPP$, $NPP$) are estimated by summing uncertain component fluxes.
+Second, differences among variables in the proportion of variation explained by climate may be attributable to more direct climatic control over $GPP$ than subsidiary fluxes. 
+That is, subsidiary fluxes may be shaped by climate both indirectly through its influence on $GPP$ and respiration and directly through any climatic influence on C allocation, as well as many other local- and regional-scale factors [e.g., @moser_elevation_2011].
 
-The latitudinal gradient in FACF (Fig.
-2) is driven primarily by temperature-related climate variables, and secondarily by moisture availability (Table 1, *H2-H3*; Figs.
-3-4).
-Because many climate variables co-vary across the latitudinal gradient, because climatic drivers affect forest carbon flux on much shorter time scales than can be captured by annual climate summary variables, and because both climatic conditions and C flux vary intra- and inter-annually around the long-term means, it is not appropriate to attempt to identify any one mean annual climate variable as a mechanistic driver of FACF.
-However, it remains informative to consider these relationships.
-We find that temperature-related climate variables ($MAT$, temperature seasonality, ...**LIST**) explain the highest proportion of variability in productivity, and among these, $MAT$ is generally the best predictor--perhaps because site-specific MAT is recorded for the majority of sites in ForC, whereas other variables are extracted from global gridded data products (Table S1**?**).
-The effects of temperature are modified by moisture availability, with reduced FACF under hot and dry conditions (*i.e.*, high PET, high deficit; Fig.
-4c-d) and sometimes under very high precipitaiton (Figs.
-3, 4b).
-Negative effects of very high precipitation on FACF have been previously observed (**REFS**) and are attributable to nutrient and light limitations (**REFS**) Thus, although temperature and water availability jointly and interactively drive global-scale patterns of FACF.
+Temperature and its seasonality were the primary drivers of C fluxes on the global scale (Table 1, *Q2*,*Q4*; Figs.
+2-4), consistent with a long legacy of research identifying temperature as a primary driver of forest ecosystem C cycling [e.g., @lieth_primary_1973; @luyssaert_co_2007; @wei_forest_2010].
+We find little evidence of any non-linearity in temperature's influence on C fluxes.
+The relationship of all fluxes to $MAT$ as an individual driver were best described by a linear function (Table S2) -- with the exception of $BNPP$, whose response to $MAT$ was close to linear (Fig.
+4a).
+This result contrasts with the idea that fluxes saturate with $MAT$ below approximately 25$^\circ$C $MAT$ [@luyssaert_co_2007; @huston_global_2009].
+It remains possible that fluxes decline above this threshold [@larjavaara_temperature_2012; @sullivan_long-term_2020], as is also consistent with tree-ring records indicating that tropical tree growth declines at high temperatures [e.g., @vlam_temperature_2014].
+However, these higher temperatures also tend to be associated with high $PET$ and $VPD$, both of which are associated with reduced C fluxes (Figs. 4c-d, S4-S5).
 
-bbl: this seems really interesting and novel, consider expanding the following paragraph
+Indeed, while temperature responses dominate at this global scale and within the climate space occupied by forests, the effects of temperature are moderated by moisture availability (Table 1, *Q2*,*Q3*; Figs 3-4).
+Specifically, C fluxes are reduced under relatively dry conditions (*i.e.*, low $MAP$; high $VPD$) and sometimes under very high precipitation (Figs. 3-4). 
+The observed positive interaction between $MAT$ and $MAP$ for $ANPP_{stem}$ on the global scale (Fig.
+3) is consistent with an analysis showing a similar interaction for $ANPP$ in tropical forests, also with a cross-over point at ~20^$\circ$C [@taylor_temperature_2017].
+However, we detect no such interaction for $ANPP$ or most other C fluxes, and we find a contrasting negative interaction for $NPP$ (Fig.
+3), suggesting that more data are required to sort out potential differences in the interactive effects of $MAT$ and $MAP$ on C fluxes in the tropics.
 
-FACF is negatively correlated with temperature seasonality (Table 1, *H4*; Fig.
-4e), and productivity is shut down during cold- or dry- dormant seasons.
-To account for this, a number of analyses seeking to characterize global-scale effects of climate on productivity have examined the relationship of FACF per month of the growing season with growing season climatic conditions (Table 1, *H5*; **REFS**).
-We found that the sort of simple metric needed to define growing season at a global scale was uncertain for temperature and problematic for moisture (**WORK ON THIS**).
-A temperature-defined growing season length had stong positive correlation with FACF (Fig.
-4f), but explained less variation than $MAT$.
-Dividing FACFs by growing season length to yield FACF per growing season month removed the majority of climate-related variation, supporting the idea that the latitudinal gradient in FACF is attributable more to shorter growing seasons at high latitudes than to inherently lower rates of photosyntheiss or respiration by high-latitude forests (*[ Enquist et al.
-2007 GCB- but check]*).
-However, there remained a number of significant correlations with growing season climatic conditions, suggesting that climatic conditions remain influential within the growing season.
-We conclude that while correcting for growing season length takes analyses a step closer to mechanistic linkage of instantaneous C flux rates to environmental conditions, it remains very crude relative to the the timescales on which climate affects plant metabolism and does not advance statistical predictive power.
-Rather, mechanistic accounting for climatic effects on global productivity patterns requires models representing physiologically meaningful timescales (*e.g.*, **refs**).
+Forest C fluxes decline with temperature seasonality (Table 1, *Q4*; Fig. 4e), indicating that fluxes during the growing season are not large enough to compensate for minimal flux during winters. 
+A temperature-defined growing season length correlated strongly with global-scale variation in annual C flux [Table 1, *Q5*; Fig.
+4f; see also @churkina_spatial_2005], consistent with the idea that the latitudinal gradient in carbon flux is attributable more to shorter growing seasons at high latitudes than to inherently lower rates of photosynthesis or respiration by high-latitude forests [@enquist_adaptive_2007; @fu_maximum_2019].
+While there is evidence that trees in high-latitude forests have adaptations to maximize photosynthesis at low temperatures [@helliker_subtropical_2008] (**REFS**), this is not sufficient to yield growing season fluxes comparable to those of tropical forests, as indicated by a number of positive correlations between monthly mean flux during the growing season and growing season temperature, solar radiation, and $PET$ (Table 1, Figs.
+S8-S9).
+Thus, we reject the hypothesis that growing season length alone accounts for global-scale variation in productivity--*i.e.*, that there is no relationship between C flux per month of the growing season and growing season climatic conditions [Table 1, *Q5*; @kerkhoff_plant_2005; @enquist_adaptive_2007; @michaletz_convergence_2014].
+Rather, annual C flux is shaped by both growing season length and the climate of peak growing season months [@chu_does_2016; @fu_maximum_2019]. 
+Given strong co-variation between growing season length and $MAT$ [Fig.
+S2; @chu_does_2016], accurately partitioning this variation will require data on intra-annual variation in C flux coupled with a more refined metric of growing season length than used here [e.g., based on leaf phenology or C exchange, *sensu* @fu_maximum_2019].
+@fu_maximum_2019 find that global-scale geographic variation in annual $NEE$ is driven more strongly by growing season length than by carbon uptake rates within the growing season, whereas interannual variation in $NEE$ and $GPP$ at any given site appears to be driven predominantly by the maximum rate of C uptake, as opposed to growing season length [@zhou_explaining_2016; @fu_maximum_2019].
+Further analysis of interannual variation in C fluxes in relation to climate will be valuable to disentangling how seasonality shapes broad geographic patterns in forest C flux.
 
-Our analysis clarifies how FACF varies (bbl: is FACF singular or plural?
-Be consistent) with latitude and climate on a global scale, with some important implications for how forest carbon cycling relates to climate and, by extension, how it is likely to respond to climatic warming.
-We find no support for non-linear trends in mature forest C cycling with respect to latitude or $MAT$, and no distinct trends in C allocation across the global scale (Fig.
-2).
-The implication is that under warmer conditions with similar moisture availability--and within the temperature range to which forest communities are adapated and acclimatized--higher temperatures result in a generalized acceleration of FACF, with no major shifts in C allocation among subsidiary fluxes.
-Of course, actual climatic changes will result in very different sets of conditions than represented across geographic gradients in climate, but our analysis clarifies how carbon cycles through contemporary forest ecosystems.
-As we enter a period of accelerating climatic change, understanding of the fundamental climatic controls on FACF sets a foundation for understanding patterns of change.
-**[work on this]**
+Our analysis clarifies how annual forest autotrophic C fluxes vary with latitude and climate on a global scale, with some important implications for how forest C cycling relates to climate and, by extension, how it is likely to respond to climatic warming.
+To the extent that patterns across broad scale climatic gradients can foretell how ecosystem responses to climate change, our findings suggest that higher temperatures with similar moisture availability would result in a generalized acceleration of forest C cycling (Figs.
+2-3).
+This is consistent with observations of continental- to global-scale increases over time in $GPP$ [@li_mapping_2019] and $ANPP_{stem}$ [@brienen_long-term_2015; @hubau_asynchronous_2020], along with some C cycle components not considered here: tree mortality [@brienen_long-term_2015; @mcdowell_drivers_2018], soil respiration [@bond-lamberty_global_2010], and heterotrophic soil respiration [@bond-lamberty_globally_2018].
+However, increasing C flux rates are by no means universal [e.g., @rutishauser_testing_2020; @hubau_asynchronous_2020], likely because other factors are at play, including changes to other aspects of climate, atmospheric pollution (CO~2~, SO~2~, NO~x~), and local disturbances.
+Moreover, forest ecosystem responses to climatic changes outside the temperature range to which forest communities are adapted and acclimatized will not necessarily parallel responses across geographic gradients in climate.
+Indeed, tree-ring studies from forests around the world indicate that tree growth rates -- along with $ANPP_{stem}$ and possibly other ecosystem C fluxes -- respond negatively to temperature [@sniderhan_growth_2016;@helcoski_growing_2019].
+Furthermore, in the tropics, climate change will push forests beyond any contemporary climate, and there are some indications that this could reduce C flux rates  [@mau_temperate_2018; @sullivan_long-term_2020].
+Further research is required to understand the extent to which forest responses to climate change will track the observed global gradients, and the time scale on which they will do so.
+In the meantime, understanding the fundamental climatic controls on annual C cycling in Earth's forests sets a firmer foundation for understanding forest C cycle responses to accelerating climate change.
 
 ### Acknowledgements
-Scholarly Studies
-ForestGEO
-Compilation of the ForC database was originally funded by DOE
+We gratefully acknowledge all authors of the original studies and data compilations included in this analysis, their funding agencies, and the various networks that support ground-based measurements of C fluxes.
+We also thank the numerous researchers who have contributed to the building of ForC.
+This study was funded by a Smithsonian Scholarly Studies grant to KJAT and HCML and by Smithsonian's Forest Global Earth Observatory (ForestGEO). 
+Original compilation of the ForC database was funded by DOE grants DE-SC0008085 and DE-SC0010039 to KAT.
+
+### References

### 2020-11-12 manuscript/manuscript draft.Rmd
-

@@
 
-nocite: | 
-  @chu_does_2016
 ---
@@
 The recent development of the Global Forest Carbon database (ForC), which synthesizes multiple variables and includes records of stand history [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018], opens up the possibility for a standardized analysis of global scale variation in multiple C fluxes and the principle climatic drivers of these patterns.
+The most comprehensive analysis of this type was @luyssaert_co_2007, which was based on a database <25% the size of the ForC version used here, did not control for effects of stand age, and examined global climatic trends in only three variables.
 In order to approach this broad topic, we simplify the major gaps in our knowledge to five broad questions and corresponding predictions (Table 1).
@@
 2-3).
-This is consistent with observations of continental- to global-scale increases over time in $GPP$ [@li_mapping_2019] and $ANPP_{stem}$ [@brienen_long-term_2015; @hubau_asynchronous_2020], along with some C cycle components not considered here--tree mortality [@brienen_long-term_2015; @mcdowell_drivers_2018], soil respiration [@bond-lamberty_global_2010], and heterotrophic soil respiration [@bond-lamberty_globally_2018].
+This is consistent with observations of continental- to global-scale increases over time in $GPP$ [@li_mapping_2019] and $ANPP_{stem}$ [@brienen_long-term_2015; @hubau_asynchronous_2020], along with some C cycle components not considered here: tree mortality [@brienen_long-term_2015; @mcdowell_drivers_2018], soil respiration [@bond-lamberty_global_2010], and heterotrophic soil respiration [@bond-lamberty_globally_2018].
 However, increasing C flux rates are by no means universal [e.g., @rutishauser_testing_2020; @hubau_asynchronous_2020], likely because other factors are at play, including changes to other aspects of climate, atmospheric pollution (CO~2~, SO~2~, NO~x~), and local disturbances.

### 2020-11-13 manuscript/manuscript draft.Rmd
added ref from @nkunert

@@
 4f; see also @churkina_spatial_2005], consistent with the idea that the latitudinal gradient in carbon flux is attributable more to shorter growing seasons at high latitudes than to inherently lower rates of photosynthesis or respiration by high-latitude forests [@enquist_adaptive_2007; @fu_maximum_2019].
-While there is evidence that trees in high-latitude forests have adaptations to maximize photosynthesis at low temperatures [@helliker_subtropical_2008] (**REFS**), this is not sufficient to yield growing season fluxes comparable to those of tropical forests, as indicated by a number of positive correlations between monthly mean flux during the growing season and growing season temperature, solar radiation, and $PET$ (Table 1, Figs.
+While there is evidence that trees in high-latitude forests have adaptations to maximize photosynthesis at low temperatures [@helliker_subtropical_2008; @huang_air_2019], this is not sufficient to yield growing season fluxes comparable to those of tropical forests, as indicated by a number of positive correlations between monthly mean flux during the growing season and growing season temperature, solar radiation, and $PET$ (Table 1, Figs.
 S8-S9).

### 2020-12-01 manuscript/manuscript draft.Rmd
add Nobby's new affiliation
starting HM comments

@@
 
-Norbert Kunert^1,3^
+Norbert Kunert^1,3,4^
 
-Ben Bond-Lamberty^4^
+Ben Bond-Lamberty^5^
 
@@
 
-4. Joint Global Change Research Institute, Pacific Northwest National Laboratory, College Park Maryland 20740 USA
+4. Institute of Botany, University of Natural Resources and Applied Life Sciences, Vienna, Austria
+
+5. Joint Global Change Research Institute, Pacific Northwest National Laboratory, College Park Maryland 20740 USA
 
@@
 Here, we draw upon 1,319 records from the Global Forest Carbon Database (ForC), representing all major forest types and the nine most significant autotrophic carbon fluxes, to comprehensively explore how C cycling in mature, undisturbed forests varies with latitude and climate on a global scale.
-We show that, across all flux variables analyzed, C cycling decreases continuously with absolute latitude -- a finding that confirms multiple previous studies but contradicts the idea that net primary productivity of temperate forests rivals that of tropical forests.
+We show that, across all flux variables analyzed, C cycling decreases continuously with absolute latitude -- a finding that confirms multiple previous studies and contradicts the idea that net primary productivity of temperate forests rivals that of tropical forests.
 C flux variables generally displayed similar trends across latitude and multiple climate variables, with no differences in allocation detected at this global scale.
-Temperature variables in general, and mean annual temperature and temperature seasonality in particular, were the best univariate predictors of C flux, explaining 19 - 71% of variation in the C fluxes analyzed.
+Temperature variables in general, and mean annual temperature or temperature seasonality in particular, were the single best predictors of C flux, explaining 19 - 71% of variation in the C fluxes analyzed.
 The effects of temperature were modified by moisture availability, with C flux reduced under hot and dry conditions and sometimes under very high precipitation. 
-C fluxes increased with growing season length, but this was never the best univariate predictor.
+C fluxes increased with growing season length, but significant climate-related variation persisted after accounting for this variable.
 Within the growing season, the influence of climate on C cycling was small but significant for a number of flux variables. 
@@
 However, studies have differed in their conclusions regarding the shape of this relationship -- quite possibly because of lack of standardization with respect to methodology and stand history. 
-*Productivity may vary with stand age, disturbance, and management [@de_lucia_forest_2007; @yu_high_2014; @simova_enigma_2017], making clear latitudinal patterns difficult to discern without standardization of the dataset.
+*Productivity may vary with stand age, disturbance, and management [@de_lucia_forest_2007; @yu_high_2014; @fernandez-martinez_spatial_2014; @simova_enigma_2017], making clear latitudinal patterns difficult to discern without standardization of the dataset.
 [???]*
@@
 Furthermore, in the tropics, climate change will push forests beyond any contemporary climate, and there are some indications that this could reduce C flux rates  [@mau_temperate_2018; @sullivan_long-term_2020].
+(*but discuss @smith_empirical_2020*)
 Further research is required to understand the extent to which forest responses to climate change will track the observed global gradients, and the time scale on which they will do so.
 In the meantime, understanding the fundamental climatic controls on annual C cycling in Earth's forests sets a firmer foundation for understanding forest C cycle responses to accelerating climate change.
+
+**CITATIONS TO ADD**:
+
+@mullerlandau_patterns_2020
+@zani_increased_2020
 

### 2020-12-01 manuscript/manuscript draft.Rmd
worked on Helene's comments
suggestions that I've dealt with are crossed off in the PDF.  I skipped results, haven't quite finished discussion

@@
 However, we lack a cohesive synthesis on how forest carbon fluxes vary globally with respect to climate and one another. 
-Here, we draw upon 1,319 records from the Global Forest Carbon Database (ForC), representing all major forest types and the nine most significant autotrophic carbon fluxes, to comprehensively explore how C cycling in mature, undisturbed forests varies with latitude and climate on a global scale.
+Here, we draw upon 1,319 records from the Global Forest Carbon Database (ForC), representing all major forest types and the nine most significant autotrophic carbon fluxes, to comprehensively explore how annual C cycling in mature, undisturbed forests varies with latitude and climate on a global scale.
 We show that, across all flux variables analyzed, C cycling decreases continuously with absolute latitude -- a finding that confirms multiple previous studies and contradicts the idea that net primary productivity of temperate forests rivals that of tropical forests.
@@
 The effects of temperature were modified by moisture availability, with C flux reduced under hot and dry conditions and sometimes under very high precipitation. 
-C fluxes increased with growing season length, but significant climate-related variation persisted after accounting for this variable.
-Within the growing season, the influence of climate on C cycling was small but significant for a number of flux variables. 
+Annual C fluxes increased with growing season length and were also influenced by growing season climate. 
 These findings clarify how forest C flux varies with latitude and climate on a global scale. 
-In a period of accelerating climatic change, this improved understanding of the fundamental climatic controls on forest C cycling sets a foundation for understanding patterns of change.
+In an era when forests will play a critical yet uncertain role in shaping Earth's rapidly changing climate, our synthesis provides a foundation for understanding global patterns in forest C cycling.
 
@@
 As atmospheric CO~2~ continues to rise, driving climate change, forests will play a critical role in shaping the future of Earth's climate [@cavaleri_urgent_2015; @rogelj_mitigation_2018]. 
-However, our understanding of the climate dependence of forest C cycling on a global scale has been limited by analyses typically considering only one or a few variables at a time, insufficient parsing of related variables, and the mixing of data from forests that vary in stand age, disturbance history, and management status, all of which affect C cycling [@litton_carbon_2007; @gillman_latitude_2015; @simova_enigma_2017].
+However, our understanding of the climate dependence of forest C cycling on a global scale remains incomplete, in large part because it is pieced together from numerous studies, most considering only one or a few variables at a time, with various approaches for handling influential factors such as stand age, disturbance history, and management status [@litton_carbon_2007; @gillman_latitude_2015; @simova_enigma_2017].
 
@@
 A shallower increase in $NPP$ than in $GPP$ with decreasing latitude would align with the suggestion that tropical forests tend to have low carbon use efficiency [$CUE$= $NPP$/$GPP$; @de_lucia_forest_2007; @malhi_productivity_2012; @anderson-teixeira_carbon_2016].
-Such differences among C fluxes their relationship to latitude could have profound implications for our understanding of the C cycle and its climate sensitivity.
+Such differences among C fluxes in their relationship to latitude could have profound implications for our understanding of the C cycle and its climate sensitivity.
 However, until recently the potential to compare latitudinal trends across C fluxes has been limited by lack of a large database with standardization for methodology, stand history, and management [@anderson-teixeira_forc_2018].
@@
 Some studies have detected no deceleration or decline in $GPP$ [@luyssaert_co_2007], $NPP$ [@schuur_productivity_2003], or root respiration [$R_{root}$; @piao_forest_2010; @wei_forest_2010] with increasing $MAT$.
-In contrast, others have found evidence of saturation or decline of C flux in the warmest climates; @luyssaert_co_2007 found $NPP$ saturating at around 10$^\circ$C MAT; @larjavaara_temperature_2012 found that increases in $GPP$ saturate at approximately 25$^\circ$C $MAT$, and @sullivan_long-term_2020 found that, within the tropics, $ANPP_{stem}$ decreases at the highest maximum temperatures.
-C fluxes generally saturate at high levels of $MAP$, though the saturation points identified vary from $MAP$ of ~1000 mm for $R_{root}$ [@wei_forest_2010]  up to 2,445 mm for $NPP$ [@schuur_productivity_2003].
-Interactions between $MAT$ and $MAP$ may also influence productivity [@yu_high_2014]; within the tropics, there is a positive interaction between $MAT$ and $MAP$ in shaping $ANPP$, such that high rainfall has a negative effect on productivity in cooler climates, compared to a positive effect in warmer climates [@taylor_temperature_2017].
-There is also evidence that C fluxes also respond to climate variables such as temperature and precipitation seasonality [@wagner_climate_2016], cloud cover [@taylor_temperature_2017], solar radiation [@beer_terrestrial_2010; @fyllas_solar_2017], and potential evapotranspiration [@kerkhoff_plant_2005]; however, these are not typically assessed in global-scale analyses of annual forest C flux.
+In contrast, others have found evidence of saturation or decline of C flux in the warmest climates; @luyssaert_co_2007 found $NPP$ saturating at around 10$^\circ$C MAT; @larjavaara_temperature_2012 found that increases in $GPP$ saturate at approximately 25$^\circ$C $MAT$, and @sullivan_long-term_2020 found that, within the tropics, woody stem productivity ($ANPP_{stem}$) decreases at the highest maximum temperatures.
+C fluxes generally saturate at high levels of $MAP$, though the saturation points identified vary widely [e.g., ~1000 - 2,445 mm yr^-1^; @wei_forest_2010; @schuur_productivity_2003]. 
+Interactions between $MAT$ and $MAP$ may also influence productivity [@yu_high_2014]; within the tropics, there is a positive interaction between $MAT$ and $MAP$ in shaping $ANPP$, such that temperature has a positive effect on productivity in moist climates, but a negative effect in dry climates [@taylor_temperature_2017].
+There is also evidence that C fluxes also respond to climate variables such as seasonality of temperature and precipitation [@wagner_climate_2016], cloud cover [@taylor_temperature_2017], solar radiation [@beer_terrestrial_2010; @fyllas_solar_2017], and potential evapotranspiration [@kerkhoff_plant_2005]; however, these are not typically assessed in global-scale analyses of annual forest C flux.
 
-As metrics of annual climate, $MAT$ and $MAP$ fail to capture variation in climate on an intra-annual scale, including temperature and precipitation seasonality and growing season length. 
+Mean annual temperature and precipitation do not capture intra-annual climate variation, including temperature and precipitation seasonality and growing season length. 
 Most forests--even tropical evergreen--exhibit some seasonality in both climate and C flux [e.g., @wagner_pan-tropical_2014], and this seasonality influences annual C fluxes [@churkina_spatial_2005; @keenan_net_2014; @fu_maximum_2019].
 In particular, growing season length has been linked to $ANPP$, $NPP$, $GPP$, and net ecosystem exchange of CO~2~ [$NEE$, or the difference between $GPP$ and ecosystem respiration; @kerkhoff_plant_2005; @michaletz_convergence_2014; @churkina_spatial_2005; @keenan_net_2014; @zhou_explaining_2016].
-However, the relative importance of growing season length, as opposed to climate within the growing season, remains debated.
+However, the relative importance of climate within the growing season, as opposed to growing season length, remains debated.
 On one end of the spectrum, some studies have suggested that the influence of temperature on C fluxes may be limited to determining the length of the frost-free growing season, and that climate within the growing season has little influence on C fluxes because of plant adaptation and acclimatization to local climates [@kerkhoff_plant_2005; @enquist_adaptive_2007; @michaletz_convergence_2014; @michaletz_drivers_2018].
@@
 The recent development of the Global Forest Carbon database (ForC), which synthesizes multiple variables and includes records of stand history [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018], opens up the possibility for a standardized analysis of global scale variation in multiple C fluxes and the principle climatic drivers of these patterns.
-The most comprehensive analysis of this type was @luyssaert_co_2007, which was based on a database <25% the size of the ForC version used here, did not control for effects of stand age, and examined global climatic trends in only three variables.
+The most comprehensive previous analysis of this type was @luyssaert_co_2007, which was based on a database <25% the size of the ForC version used here, did not control for effects of stand age, and examined global climatic trends in only three variables.
 In order to approach this broad topic, we simplify the major gaps in our knowledge to five broad questions and corresponding predictions (Table 1).
@@
 Statistically signficant support for/ rejection of hypotheses is indicated with 'yes'/'no', and '-' indicates no significant relationship.
-Parentheses indicate partial overall support or rejection of hypotheses across all fluxes considered.", escape = F) %>%
+Parentheses indicate partial overall support or rejection of hypotheses across all fluxes considered.
+Flux variables are defined in Table 2.", escape = F) %>%
   kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 12) %>%
@@
 
-Following the previous studies whose hypothesis we were evaluating [@kerkhoff_plant_2005; @michaletz_convergence_2014], length of the growing season was estimated to the nearest month, where growing season months were defined as months with mean minimum temperature > 0.5$^\circ$C.
+Length of the growing season was estimated to the nearest month, where growing season months were defined as months with mean minimum temperature > 0.5$^\circ$C. 
+This is consistent with the previous studies whose hypothesis we were evaluating [@kerkhoff_plant_2005; @michaletz_convergence_2014].
 We experimented with a definition of growing season months including a moisture index, defined as $(MAT - PET)/PET > -0.95$ [@kerkhoff_plant_2005; see also @michaletz_convergence_2014]. 
@@
 We show that, across all nine variables analyzed, annual C flux decreases continually with latitude (Fig.
-2), a finding that confirms multiple previous studies but contradicts the idea that productivity of temperate forests rivals or even exceeds that of tropical forests [@luyssaert_co_2007; @huston_global_2009].
+2), a finding that confirms multiple previous studies and contradicts the idea that productivity of temperate forests rivals or even exceeds that of tropical forests [@luyssaert_co_2007; @huston_global_2009].
 At this global scale, C fluxes increase approximately in proportion to one another, with component fluxes summing appropriately to larger fluxes and no detectable differences in allocation across latitude or climates (Figs.
@@
 Water availability is also influential, but of secondary importance across the climate space occupied by forests (Figs. 3-4). 
-Contrary to prior suggestions that the majority of variation in C cycling is driven primarily by the length of the growing season [@kerkhoff_plant_2005; @enquist_adaptive_2007; @michaletz_convergence_2014], we find modest explanatory power of growing season length and small but sometimes significant influence of climate within the growing season (Figs.
+Contrary to prior suggestions that the majority of variation in C cycling is driven primarily by the length of the growing season [@kerkhoff_plant_2005; @enquist_adaptive_2007; @michaletz_convergence_2014], we find modest explanatory power of growing season length and small but sometimes significant influences of growing season climate (Figs.
 4f,S6-S9).
@@
 This contrasts with the suggestion that C fluxes (e.g., $NPP$, $ANPP$, $ANPP_{stem}$) of temperate forests are similar to or even greater than that of tropical forests [@luyssaert_co_2007; @huston_global_2009].
-Previous indications of such a pattern may have been an artifact of differences in stand age across biomes.
+Previous indications of this pattern may have been an artifact of differences in stand age across biomes.
 Compared to tropical forests, the temperate forest biome has experienced more widespread anthropogenic disturbance and has a larger fraction of secondary stands [@potapov_mapping_2008; @yu_high_2014; @poulter_global_2018], so analyses comparing across latitudinal gradients without controlling for stand age risk confounding age with biome effects.
@@
 and is not surprising given that carbon allocation within forest ecosystems is relatively constrained [@enquist_global_2002; @litton_carbon_2007; @malhi_allocation_2011]. 
-We find no trend in the allocation of $GPP$ between production and respiration across latitude or climate ($NPP$:$R_{auto}$; Fig.
-S3), refuting the idea that tropical forests have anomalously low $CUE$ [@de_lucia_forest_2007; @malhi_productivity_2012; @anderson-teixeira_carbon_2016].
+We find no significant trend in the allocation of $GPP$ between production and respiration across latitude or climate ($NPP$:$R_{auto}$; Fig.
+S3), counter to the idea that tropical forests have anomalously low $CUE$ [@de_lucia_forest_2007; @malhi_productivity_2012; @anderson-teixeira_carbon_2016].
 Rather, differences in $CUE$ between old-growth tropical forests relative to (mostly younger) extratropical forests are likely an artifact of comparing stands of different age, as $CUE$ is known to decline with forest age [@de_lucia_forest_2007; @piao_forest_2010; @collalti_forest_2020].
 Another previously observed pattern for which we find no support is a tendency for belowground C allocation to decrease with increasing temperature  [@moser_elevation_2011; @gill_belowground_2016]; rather, we observe no trends in allocation between $ANPP$ and $BNPP$ across latitudes.
-Failure to detect significant tends in C allocation with respect to climate in this analysis does not imply that none exist; rather, it suggests that, at this global scale, differences are subtle and/or that more careful methodological standardization is required to detect them.
+Failure to detect significant tends in C allocation with respect to climate in this analysis does not imply that none exist; rather, it suggests that, at this global scale, differences are subtle and/or that more careful methodological standardization and/or more data is required to detect them.
 
@@
 Belowground fluxes in particular are difficult to quantify, and measurement methods for the belowground fluxes considered here may use fundamentally different approaches in different sites (*e.g.*, minirhizotrons, ingrowth cores, or sequential coring for $BNPP_{fine.root}$; root exclusion, stable isotope tracking, or gas exchange of excised roots for $R_{root}$), and sampling depth is variable and often insufficient to capture the full soil profile.
-$ANPP_{stem}$, which is also poorly explained by latitude or climate, is more straightforward to measure but is subject to variability introduced by differences such as biomass allometries applied and minimum plant size sampled [@clark_measuring_2001].
-However, methodological variation and uncertainty affect all of fluxes considered here, and some of the larger fluxes that vary more strongly with respect to climate ($ANPP$, $NPP$) are estimated by summing uncertain component fluxes.
+$ANPP_{stem}$, which is also poorly explained by latitude or climate, is more straightforward to estimate but subject to variability introduced by methodological differences including minimum plant size sampled and choice of biomass allometries [@clark_measuring_2001].
+That said, methodological variation and uncertainty affect all of fluxes considered here, and some of the larger fluxes that vary more strongly with respect to climate ($ANPP$, $NPP$) are estimated by summing uncertain component fluxes.
 Second, differences among variables in the proportion of variation explained by climate may be attributable to more direct climatic control over $GPP$ than subsidiary fluxes. 
-That is, subsidiary fluxes may be shaped by climate both indirectly through its influence on $GPP$ and respiration and directly through any climatic influence on C allocation, as well as many other local- and regional-scale factors [e.g., @moser_elevation_2011].
+That is, subsidiary fluxes may be shaped by climate both through its influence on $GPP$ and through its influence on $CUE$ and C allocation.
 
@@
 4a).
-This result contrasts with the idea that fluxes saturate with $MAT$ below approximately 25$^\circ$C $MAT$ [@luyssaert_co_2007; @huston_global_2009].
+This result contrasts with the hypothesis that fluxes saturate with $MAT$ below approximately 25$^\circ$C $MAT$ [@luyssaert_co_2007; @huston_global_2009].
 It remains possible that fluxes decline above this threshold [@larjavaara_temperature_2012; @sullivan_long-term_2020], as is also consistent with tree-ring records indicating that tropical tree growth declines at high temperatures [e.g., @vlam_temperature_2014].
-However, these higher temperatures also tend to be associated with high $PET$ and $VPD$, both of which are associated with reduced C fluxes (Figs. 4c-d, S4-S5).
+However, these higher temperatures also tend to be associated with high $PET$ and $VPD$, both of which are associated with reduced C fluxes [Figs. 4c-d, S4-S5; @slot_high_2018; @zani_increased_2020].
 
@@
 
-Forest C fluxes decline with temperature seasonality (Table 1, *Q4*; Fig. 4e), indicating that fluxes during the growing season are not large enough to compensate for minimal flux during winters. 
+Forest C fluxes decline with temperature seasonality (Table 1, *Q4*; Fig. 4e), as is to be expected given that fluxes are minimal during winters. 
 A temperature-defined growing season length correlated strongly with global-scale variation in annual C flux [Table 1, *Q5*; Fig.
@@
 Given strong co-variation between growing season length and $MAT$ [Fig.
-S2; @chu_does_2016], accurately partitioning this variation will require data on intra-annual variation in C flux coupled with a more refined metric of growing season length than used here [e.g., based on leaf phenology or C exchange, *sensu* @fu_maximum_2019].
+S2; @chu_does_2016], accurately partitioning this variation will require data on intra-annual variation in C flux coupled with a higher-precision metric of growing season length than the monthly-resolution metric used here [e.g., based on leaf phenology or C exchange, *sensu* @fu_maximum_2019].
 @fu_maximum_2019 find that global-scale geographic variation in annual $NEE$ is driven more strongly by growing season length than by carbon uptake rates within the growing season, whereas interannual variation in $NEE$ and $GPP$ at any given site appears to be driven predominantly by the maximum rate of C uptake, as opposed to growing season length [@zhou_explaining_2016; @fu_maximum_2019].
@@
 Our analysis clarifies how annual forest autotrophic C fluxes vary with latitude and climate on a global scale, with some important implications for how forest C cycling relates to climate and, by extension, how it is likely to respond to climatic warming.
-To the extent that patterns across broad scale climatic gradients can foretell how ecosystem responses to climate change, our findings suggest that higher temperatures with similar moisture availability would result in a generalized acceleration of forest C cycling (Figs.
+To the extent that patterns across broad scale climatic gradients can foretell ecosystem responses to climate change, our findings suggest that higher temperatures with similar moisture availability would result in a generalized acceleration of forest C cycling (Figs.
 2-3).

### 2020-12-01 manuscript/manuscript draft.Rmd
HM comments - discussion
This finishes Helene's comments on all but results. @beckybanbury , perhaps you could work on that?

@@
 
-Our analysis clarifies how annual forest autotrophic C fluxes vary with latitude and climate on a global scale, with some important implications for how forest C cycling relates to climate and, by extension, how it is likely to respond to climatic warming.
+Our analysis clarifies how annual forest autotrophic C fluxes vary with latitude and climate on a global scale. 
 To the extent that patterns across broad scale climatic gradients can foretell ecosystem responses to climate change, our findings suggest that higher temperatures with similar moisture availability would result in a generalized acceleration of forest C cycling (Figs.
 2-3).
-This is consistent with observations of continental- to global-scale increases over time in $GPP$ [@li_mapping_2019] and $ANPP_{stem}$ [@brienen_long-term_2015; @hubau_asynchronous_2020], along with some C cycle components not considered here: tree mortality [@brienen_long-term_2015; @mcdowell_drivers_2018], soil respiration [@bond-lamberty_global_2010], and heterotrophic soil respiration [@bond-lamberty_globally_2018].
+This is consistent with observations of continental- to global-scale increases over time in $GPP$ [@li_mapping_2019], $ANPP_{stem}$ [@brienen_long-term_2015; @hubau_asynchronous_2020], tree mortality [@brienen_long-term_2015; @mcdowell_drivers_2018], soil respiration [@bond-lamberty_global_2010], and heterotrophic soil respiration [@bond-lamberty_globally_2018].
 However, increasing C flux rates are by no means universal [e.g., @rutishauser_testing_2020; @hubau_asynchronous_2020], likely because other factors are at play, including changes to other aspects of climate, atmospheric pollution (CO~2~, SO~2~, NO~x~), and local disturbances.
 Moreover, forest ecosystem responses to climatic changes outside the temperature range to which forest communities are adapted and acclimatized will not necessarily parallel responses across geographic gradients in climate.
-Indeed, tree-ring studies from forests around the world indicate that tree growth rates -- along with $ANPP_{stem}$ and possibly other ecosystem C fluxes -- respond negatively to temperature [@sniderhan_growth_2016;@helcoski_growing_2019].
-Furthermore, in the tropics, climate change will push forests beyond any contemporary climate, and there are some indications that this could reduce C flux rates  [@mau_temperate_2018; @sullivan_long-term_2020].
-(*but discuss @smith_empirical_2020*)
+Indeed, tree-ring studies from forests around the world indicate that tree growth rates -- along with $ANPP_{stem}$ and possibly other ecosystem C fluxes -- respond negatively to temperature [e.g., @helcoski_growing_2019].
+Furthermore, in the tropics, climate change will push forests beyond any contemporary climate, and there are some indications that this could reduce C flux rates  [@mau_temperate_2018; @sullivan_long-term_2020] if temperature increases are paralleled by $VPD$ increases [@smith_empirical_2020].
 Further research is required to understand the extent to which forest responses to climate change will track the observed global gradients, and the time scale on which they will do so.

### 2020-12-01 manuscript/appendix.Rmd
HM comments in appendix
This takes care of everything other than issue #123.

@@
 csl: global-change-biology.csl
-nocite: | 
-  @hijmans_very_2005; @fick_worldclim_2017; @harris_updated_2014;  @abatzoglou_terraclimate_2018; @trabucco_global_2019
 
@@
 
-#@hijmans_very_2005; @fick_worldclim_2017; @harris_updated_2014;  @abatzoglou_terraclimate_2018
 ```
-\elandscape
-\newpage
 
-\blandscape
+---
+nocite: | 
+  @hijmans_very_2005; @fick_worldclim_2017; @harris_updated_2014;  @abatzoglou_terraclimate_2018; @trabucco_global_2019
+---
+
 ```{r eval = TRUE, echo=FALSE, warning=FALSE}
@@
 ```
-
+\newpage 
 \blandscape
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S2: Scatterplots and Pearson's R values for relationships between latitude and climate variables", out.height='95%', fig.pos='H'}
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S2: Scatterplots and Pearson's R values for relationships among climate variables and latitude", out.height='95%', fig.pos='H'}
 
@@
 ```
-\elandscape
 
-\blandscape
 ```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S3: Ratios among forest C fluxes as a function of latitude and climate variables.", out.width ='80%', fig.pos='H'}

### 2020-12-01 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@
 As atmospheric CO~2~ continues to rise, driving climate change, forests will play a critical role in shaping the future of Earth's climate [@cavaleri_urgent_2015; @rogelj_mitigation_2018]. 
-However, our understanding of the climate dependence of forest C cycling on a global scale remains incomplete, in large part because it is pieced together from numerous studies, most considering only one or a few variables at a time, with various approaches for handling influential factors such as stand age, disturbance history, and management status [@litton_carbon_2007; @gillman_latitude_2015; @simova_enigma_2017].
+However, our understanding of global-scale variation in forest C cycling remains incomplete, in large part because it is pieced together from numerous studies, most considering only one or a few variables at a time, with various approaches for handling influential factors such as stand age, disturbance history, and management status [@litton_carbon_2007; @gillman_latitude_2015; @simova_enigma_2017].
 
@@
 Indeed, tree-ring studies from forests around the world indicate that tree growth rates -- along with $ANPP_{stem}$ and possibly other ecosystem C fluxes -- respond negatively to temperature [e.g., @helcoski_growing_2019].
-Furthermore, in the tropics, climate change will push forests beyond any contemporary climate, and there are some indications that this could reduce C flux rates  [@mau_temperate_2018; @sullivan_long-term_2020] if temperature increases are paralleled by $VPD$ increases [@smith_empirical_2020].
+Furthermore, in the tropics, climate change will push temperatures beyond any contemporary climate, and there are some indications that this could reduce forest C flux rates  [@mau_temperate_2018; @sullivan_long-term_2020] if paralleled by $VPD$ increases [@smith_empirical_2020].
 Further research is required to understand the extent to which forest responses to climate change will track the observed global gradients, and the time scale on which they will do so.

### 2020-12-01 manuscript/manuscript draft.Rmd
fixing up table (not done)
#124

@@
 #hypothesis_table[12,1] <- paste0(hypothesis_table[12,1], footnote_marker_number(8))
-kable(hypothesis_table, format = "latex", booktabs = TRUE, caption = "Summary of research questions, corresponding hypotheses, and results.
-Statistically signficant support for/ rejection of hypotheses is indicated with 'yes'/'no', and '-' indicates no significant relationship.
+kable(hypothesis_table, format = "latex", booktabs = TRUE, caption = "Summary of review questions, corresponding expectations based on previous studies (when applicable), and results.
+Statistically signficant support for/ rejection of hypotheses is indicated by checkmarks/ X's, and '-' indicates no significant relationship.
 Parentheses indicate partial overall support or rejection of hypotheses across all fluxes considered.
@@
   kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 12) %>%
-  pack_rows("Q1. How do C fluxes vary with latitude?", 1, 1, latex_gap_space = "1em", colnum = 5, hline_before = FALSE) %>%
-  pack_rows("Q2. How do C fluxes vary with mean annual temperature (MAT) and precipitation (MAP)?", 2, 4, latex_gap_space = "1em", colnum = 5, hline_before = TRUE) %>%
-  pack_rows("Q3. How are C fluxes related to other annual climate variables?", 5, 7, latex_gap_space = "1em", colnum = 5, hline_before = TRUE) %>%
-  pack_rows("Q4. How does seasonality influence annual C fluxes?", 8, 11, latex_gap_space = "1em", colnum = 5, hline_before = TRUE) %>%
-  pack_rows("Q5. When standardised by growing season length, how do annual C fluxes still vary with climate?", 12, 15, latex_gap_space = "1em", colnum = 5, hline_before = TRUE) %>%
+  pack_rows("Q1. How do C fluxes vary with latitude?", 1, 2, latex_gap_space = "1em", colnum =7, hline_before = FALSE) %>%
+  pack_rows("Q2. How do C fluxes vary with mean annual temperature (MAT) and precipitation (MAP)?", 3, 5, latex_gap_space = "1em", colnum = 7, hline_before = TRUE) %>%
+  pack_rows("Q3. How are C fluxes related to other annual climate variables?", 6, 8, latex_gap_space = "1em", colnum = 7, hline_before = TRUE) %>%
+  pack_rows("Q4. How does seasonality influence annual C fluxes?", 9, 12, latex_gap_space = "1em", colnum = 7, hline_before = TRUE) %>%
+  pack_rows("Q5. When standardised by growing season length, how do annual C fluxes still vary with climate?", 13, 16, latex_gap_space = "1em", colnum = 7, hline_before = TRUE) %>%
   add_header_above(c(" ",  " ", "Forest autotrophic carbon fluxes" = 9, " ")) %>%

### 2020-12-02 manuscript/manuscript draft.Rmd
fixing up table

@@
   kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 12) %>%
-  pack_rows("Q1. How do C fluxes vary with latitude?", 1, 2, latex_gap_space = "1em", colnum =7, hline_before = FALSE) %>%
-  pack_rows("Q2. How do C fluxes vary with mean annual temperature (MAT) and precipitation (MAP)?", 3, 5, latex_gap_space = "1em", colnum = 7, hline_before = TRUE) %>%
+  pack_rows("Q1. How do C fluxes vary with latitude?", 1, 2, latex_gap_space = ".5em", colnum =7, hline_before = FALSE) %>%
+  pack_rows("Q2. How do C fluxes vary with mean annual temperature (MAT) and precipitation (MAP)?", 3, 5, latex_gap_space = ".5em", colnum = 7, hline_before = TRUE) %>%
   pack_rows("Q3. How are C fluxes related to other annual climate variables?", 6, 8, latex_gap_space = "1em", colnum = 7, hline_before = TRUE) %>%

### 2020-12-02 manuscript/manuscript draft.Rmd
fixing up Table 1
#124

--- 2020-12-01 manuscript/manuscript draft.Rmd
@@
 hypothesis_table <- read.csv("hypothesis_table.csv", stringsAsFactors = FALSE, check.names = FALSE)
-#hypothesis_table[1,1] <- paste0(hypothesis_table[1,1], footnote_marker_number(1),"$^{,}$",footnote_marker_number(2),"$^{,}$", footnote_marker_number(3),"$^{,}$", footnote_marker_number(10))
-#hypothesis_table[2,1] <- paste0(hypothesis_table[2,1], footnote_marker_number(1),"$^{,}$",footnote_marker_number(4),"$^{,}$",footnote_marker_number(9))
-#hypothesis_table[3,1] <- paste0(hypothesis_table[3,1], footnote_marker_number(1),"$^{,}$",footnote_marker_number(4))
-#hypothesis_table[4,1] <- paste0(hypothesis_table[4,1], footnote_marker_number(5))
-#hypothesis_table[10,1] <- paste0(hypothesis_table[10,1], footnote_marker_number(6),"$^{,}$",footnote_marker_number(7),"$^{,}$", footnote_marker_number(8))
-#hypothesis_table[11,1] <- paste0(hypothesis_table[11,1], footnote_marker_number(7),"$^{,}$", footnote_marker_number(8))
-#hypothesis_table[12,1] <- paste0(hypothesis_table[12,1], footnote_marker_number(8))
+hypothesis_table[1,1] <- paste0(hypothesis_table[1,1], footnote_marker_number(1),"$^{,}$",footnote_marker_number(2),"$^{,}$", footnote_marker_number(3)) #Luyssaert, Gillman et al.
+(2015)", "Simova and Storch (2017)"
+hypothesis_table[2,1] <- paste0(hypothesis_table[2,1], footnote_marker_number(1),"$^{,}$", footnote_marker_number(10)) #Luyssaert, Huston & Wolverton
+hypothesis_table[3,1] <- paste0(hypothesis_table[3,1], footnote_marker_number(1),"$^{,}$",footnote_marker_number(4),"$^{,}$",footnote_marker_number(9))
+hypothesis_table[4,1] <- paste0(hypothesis_table[4,1], footnote_marker_number(1),"$^{,}$",footnote_marker_number(4))
+hypothesis_table[5,1] <- paste0(hypothesis_table[5,1], footnote_marker_number(5))
+hypothesis_table[11,1] <- paste0(hypothesis_table[11,1], footnote_marker_number(6),"$^{,}$",footnote_marker_number(7),"$^{,}$", footnote_marker_number(8))
+hypothesis_table[12,1] <- paste0(hypothesis_table[12,1], footnote_marker_number(7),"$^{,}$", footnote_marker_number(8))
+hypothesis_table[13,1] <- paste0(hypothesis_table[13,1], footnote_marker_number(8))
 kable(hypothesis_table, format = "latex", booktabs = TRUE, caption = "Summary of review questions, corresponding expectations based on previous studies (when applicable), and results.
@@
   kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 12) %>%
-  pack_rows("Q1. How do C fluxes vary with latitude?", 1, 2, latex_gap_space = "1em", colnum =7, hline_before = FALSE) %>%
-  pack_rows("Q2. How do C fluxes vary with mean annual temperature (MAT) and precipitation (MAP)?", 3, 5, latex_gap_space = "1em", colnum = 7, hline_before = TRUE) %>%
-  pack_rows("Q3. How are C fluxes related to other annual climate variables?", 6, 8, latex_gap_space = "1em", colnum = 7, hline_before = TRUE) %>%
-  pack_rows("Q4. How does seasonality influence annual C fluxes?", 9, 12, latex_gap_space = "1em", colnum = 7, hline_before = TRUE) %>%
-  pack_rows("Q5. When standardised by growing season length, how do annual C fluxes still vary with climate?", 13, 16, latex_gap_space = "1em", colnum = 7, hline_before = TRUE) %>%
+  pack_rows("Q1. How do C fluxes vary with latitude?", 1, 2,  colnum =7, hline_before = FALSE) %>%
+  pack_rows("Q2. How do C fluxes vary with mean annual temperature (MAT) and precipitation (MAP)?", 3, 5,  colnum = 7, hline_before = TRUE) %>%
+  pack_rows("Q3. How are C fluxes related to other annual climate variables?", 6, 8,  colnum = 7, hline_before = TRUE) %>%
+  pack_rows("Q4. How does seasonality influence annual C fluxes?", 9, 12,  colnum = 7, hline_before = TRUE) %>%
+  pack_rows("Q5. When standardised by growing season length, how do annual C fluxes still vary with climate?", 13, 16, colnum = 7, hline_before = TRUE) %>%
   add_header_above(c(" ",  " ", "Forest autotrophic carbon fluxes" = 9, " ")) %>%
@@
   kableExtra::landscape() 
- # kableExtra::footnote(number = c("Luyssaert et al.
+  #kableExtra::footnote(number = c("Luyssaert et al.
 (2007)", "Gillman et al.

### 2020-12-02 manuscript/response to reviews.Rmd
update response to reviews (incorporates HM comments)
#122

@@
 
-**Most of the points raised here are covered by more specific comments below, and our responses to these points follow those comments.
-The one criticism here that is not addressed below is that we present "simple statistical analyses".
+**Most of the points raised here are covered by more specific comments below, and our responses to these points follow those comments.**
+
+**We believe that one general problem may have been that the previous manuscript version didn't clearly communicate our intentions, which were not to pose novel hypotheses but to conduct a quantitative review of the subject using a vastly expanded and standardized dataset.
+While we feel that the text of the introduction made it clear that we weren't intending to present the hypotheses as novel, we have realized that a quick review of Table 1 in isolation would give the false impression that we intended these analyses as novel hypothesis tests.
+We have taken some steps to improve this aspect:**
+
+**1. We modified Table 1, labeling the questions as "review questions" and clarifying that many of the hypotheses have been previously tested, citing relevant publications.**
+
+**2.**  *(we should add some very clear statements to the abstract, intro, maybe discussion) ...*
+
+**One point above that is not addressed below is that we present "simple statistical analyses".
 While we concur that the analyses are relatively simple, that does not preclude their being appropriate, informative, and sufficient.
@@
 
-**We agree that these results are not surprising, however a crucial part of science is built on high-quality synthesis that clarifies current understanding and probes for unexpected relationships.**
+**We agree that these results are not surprising to many of us, and state in the introduction that productivity decreases with latitude.
+Yet, as also noted in the introduction, the form of this relationship is less clear, and contradictory papers have been published in high-profile journals in the last 15 years [@luyssaert_co_2007; @huston_global_2009;  @gillman_latitude_2015].
+A crucial part of science is built on high-quality synthesis that clarifies current understanding and probes for unexpected relationships, which is the goal here-- as opposed to testing this as a novel hypothesis.**
 
@@
 S3 to clarify that none of the fits are significant (and therefore no statistics are presented).**
+
+*from Helene: maybe we should also address statistical power*
 

### 2020-12-09 manuscript/manuscript draft.Rmd
add nocite chunk below Table 1
#124
Update manuscript draft.Rmd

@@
 
+---
+nocite: | 
+  @gillman_latitude_2015
+---
+
 ### Materials and Methods
@@
 
-*Q1. How does C flux vary with latitude?*
+*Q1. How do C fluxes vary with latitude?*
 
@@
 Latitude was a strong predictor for many of the carbon fluxes, particularly the larger fluxes (Table S2, S6).
-Specifically, latitude explained 64% of variation in GPP (n = 243, p<0.0001), 50% in NPP (n = 161, p<0.0001) and 44% in ANPP (n = 278, p<0.0001).
-The C fluxes that were most poorly predicted by latitude were $BNPP_{fine.root}$ ($R^2$=0.17) and $ANPP_{stem}$ ($R^2$=0.18).
+Latitude explained 64% of variation in GPP (n = 243, p<0.0001), 50% in NPP (n = 161, p<0.0001) and 44% in ANPP (n = 278, p<0.0001).
+The C fluxes that were most poorly predicted by latitude were $BNPP_{fine.root}$ (n = 88, p< .01, $R^2$=0.17) and $ANPP_{stem}$ (n = 264, p<0.0001, $R^2$=0.18).
 The relationship with latitude was best fit by the first-order linear model, with the exception of NPP and $R_{root}$, for which a logarithmic model was a slightly -- but not significantly -- better fit.
@@
 S3).
-Of 16 relationships tested (4 ratios among C flux variables regressed against latitude, $MAT$, $MAP$ and temperature seasonality), none were significant.
+Of 12 relationships tested (3 ratios among C flux variables regressed against latitude, $MAT$, $MAP$ and temperature seasonality), none were significant.
 
-*Q2. How does C flux relate to $MAT$ and $MAP$?*
+*Q2. How do C fluxes relate to $MAT$ and $MAP$?*
 
@@
 For visualization purposes, data points are grouped into bins of 0 - 1000, 1001 - 2000, 2001 - 3000, and >3000mm mean annual precipitation, and lines of best fit models are plotted for mean annual precipitation values of 500, 1500, 2500, and 3500mm.
-All regressions are significant $(p<0.05)$.", fig.pos='H'}
+Significance is defined as p < 0.05.", fig.pos='H'}
   knitr::include_graphics("mat_map_interaction.png")
@@
 
-*Q3. How does C flux relate to other annual climate variables?*
+*Q3. How do C fluxes relate to other annual climate variables?*
 
@@
 
-When annual C fluxes were standardized by growing season length (in monthly increments), correlations with growing season climate were generally weak (Figs. S8-S9).
+When annual C fluxes were standardized by growing season length (in integer number of months), correlations with growing season climate were generally weak (Figs. S8-S9).
 $ANPP$ increased with growing season temperature ($R^2$ = 0.09, p<0.001) and precipitation ($R^2$ = 0.04, p<0.05). 

### 2020-12-09 scripts/mat_map_interaction.R
Update mat_map_interaction.R

@@ -377,8 +377,8 @@ for (response.v in response.variables){
   
   mtext(side = 3, line = -1, text = paste0(respv), adj = 0.1, cex = 0.5)
   if(response.v %in% c("NPP", "ANPP_woody_stem")) mtext(side = 3, line = -2, text = "Significant interactive effect", adj = 0.1, cex = 0.5)
-  if(response.v %in% c("GPP", "ANPP", "R_auto")) mtext(side = 3, line = -2, text = "Significant additive effect", adj = 0.1, cex = 0.5)
-  if(response.v %in% c("ANPP_foliage", "BNPP_root", "BNPP_root_fine", "R_auto_root")) mtext(side = 3, line = -2, text = "Significant effect of MAT", adj = 0.1, cex = 0.5)
+  if(response.v %in% c("GPP", "ANPP", "R_auto")) mtext(side = 3, line = -2, text = "Significant additive effect only", adj = 0.1, cex = 0.5)
+  if(response.v %in% c("ANPP_foliage", "BNPP_root", "BNPP_root_fine", "R_auto_root")) mtext(side = 3, line = -2, text = "Significant effect of MAT only", adj = 0.1, cex = 0.5)
   Rsqs <- rbind(Rsqs, Rsq)
   pannel.nb <- pannel.nb +1
 }

### 2020-12-09 scripts/refining_ratio_regressions_all.R
Update refining_ratio_regressions_all.R

@@ -145,9 +145,9 @@ final_sheet$CUE <- log(final_sheet$NPP_mean/final_sheet$GPP_mean)
 
 fixed.variables <- c("lat", "mat", "map", "TempSeasonality")
 
-png(file = "C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/figures/final_figures/ratio_regressions/log ratios/all.png", width = 3000, height = 3000, units = "px", res = 300)
+png(file = "C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/figures/final_figures/ratio_regressions/log ratios/all.png", width = 3000, height = 2500, units = "px", res = 300)
 
-par(mfrow = c(4,4), mar = c(1,1,3,0.5), oma = c(5,4,5,1), xpd = F)
+par(mfrow = c(3,4), mar = c(1,1,3,0.5), oma = c(5,4,5,1), xpd = F)
 
 ##### CUE
 
@@ -206,58 +206,58 @@ for(fixed.v in fixed.variables){
 
 ##NPP:R
 
-n <- 1
-
-for(fixed.v in fixed.variables){
-
-  final_sheet$fixed <- final_sheet[, fixed.v]
-  
-  mod <-  lmer(ratio ~ 1 + (1|geographic.area/plot.name), data = final_sheet, REML = F)
-  mod.linear <- lmer(ratio ~ poly(fixed, 1, raw = T) + (1|geographic.area/plot.name), data = final_sheet, REML = F)
-  
-  best.model <- as.character(aictab(list(mod = mod, mod.linear = mod.linear), sort = T)$Modname[1])
-  
-  if (best.model == "mod.linear") mod.full <- lmer(ratio ~ poly(fixed, 1, raw = T) + (1|geographic.area/plot.name), data = final_sheet, REML = F)
-  if (best.model == "mod") mod.full <- lmer(ratio ~ poly(fixed, 1, raw = T) + (1|geographic.area/plot.name), data = final_sheet, REML = F)
-  
-  significant.effect <- anova(mod, mod.full)$"Pr(>Chisq)"[2] < 0.05
-  significance <- anova(mod, mod.full)$"Pr(>Chisq)"[2]
-  sample.size <- length(final_sheet$ratio)
-  
[…]

### 2020-12-13 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@
 The recent development of the Global Forest Carbon database (ForC), which synthesizes multiple variables and includes records of stand history [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018], opens up the possibility for a standardized analysis of global scale variation in multiple C fluxes and the principle climatic drivers of these patterns.
+Our analyses represent a major step forward in relation to previous work.
 The most comprehensive previous analysis of this type was @luyssaert_co_2007, which was based on a database <25% the size of the ForC version used here, did not control for effects of stand age, and examined global climatic trends in only three variables.
@@
 We experimented with inclusion of altitude as a fixed effect, but excluded it from the final models because it added very little explanatory power -- that is, the difference in AIC ($\Delta AIC$) relative to models excluding altitude was generally small (often $\Delta AIC$<2).
-Effects were considered significant when inclusion of the fixed effect of interest resulted in p $\le$ 0.05 and $\Delta AIC$ $\ge$ 2.0 relative to a corresponding null model. 
+Effects were considered significant when inclusion of the fixed effect of interest resulted in p $\le$ 0.05 under an ANOVA test, and $\Delta AIC$ $\ge$ 2.0 relative to a corresponding null model. 
 All $R^2$ values presented here are marginal $R^2$ values, and refer to the proportion of variation explained by only the fixed effects. 
@@
 Models were run on data from sets of sites that were common to each pair, in order to account for variation in the number of data points included.
+To standardise for variation in degrees of freedom across model types, only linear and logarithmic models were included in the pairwise analysis.
 
@@
 
-Variation in allocation to component carbon fluxes was explored for three groupings: (1) $GPP = NPP + R_{auto}$, (2) $NPP = ANPP + BNPP$, and (3) $ANPP = ANPP_{foliage} + ANPP_{stem}$.
-For each group, measurements taken at the same site and plot, and in the same year, were grouped together.
-For groups (1) and (2), where 2 of the 3 flux measurements were available for a given site, plot, and year, these measurements were used to calculate the third.
-The ratio of each pair of component fluxes was calculated.
-The log of these ratios were regressed against latitude and climate variables, using the linear model specified above.
-Cook's distance analyses were carried out for each of the models, and extreme outliers removed.
-
 We next examined the relationships of C fluxes to climate variables (*Q2-Q4*; Table 1). 
@@
 
-Linear models were used to investigate the potential joint and interactive effects of $MAT$ and $MAP$ on carbon fluxes.
+Linear models were used to investigate the potential joint and interactive effects of $MAT$ and $MAP$ on carbon fluxes (*Q2*; Table 1).
 An additive model including $MAP$ in addition to $MAT$ was accepted when $\Delta$AIC >2 relative to a null including only $MAT$ as a fixed effect.
 An interactive model containing a $MAT$ x $MAP$ interaction was accepted when $\Delta$AIC >2 relative to a null including $MAT$ and $MAP$ as fixed effects.
+
+Variation in allocation to component carbon fluxes was explored for three groupings: (1) $GPP = NPP + R_{auto}$, (2) $NPP = ANPP + BNPP$, and (3) $ANPP = ANPP_{foliage} + ANPP_{stem}$.
+For each group, measurements taken at the same site and plot, and in the same year, were grouped together.
+For groups (1) and (2), where 2 of the 3 flux measurements were available for a given site, plot, and year, these measurements were used to calculate the third.
+We then calculated the ratio of each pair of component fluxes ($NPP: R_{auto}$; $ANPP:BNPP$; $ANPP_{foliage}:ANPP_{stem}$).
+The logs of these ratios were regressed against latitude, $MAT$, $MAP$, and temperature seasonality, using the linear models specified above.
+Cook's distance analyses were carried out for each of the models, and extreme outliers removed.
 
@@
 For eight of the nine fluxes, this relationship was linear.
-For only one variable, $BNPP$, did a lognormal fit provide an improvement over a first-order linear relationship, though this was not significant ($\Delta$AIC <2).
+For $BNPP$ the best fit was a lognormal fit, though this was not significant ($\Delta$AIC <2).
 As with latitude, $MAT$ tended to explain more variation in the larger fluxes ($GPP$, $NPP$, $ANPP$, $R_{auto}$) and $ANPP_{foliage}$ (all $R^2$> 0.4) than in subsidiary and belowground fluxes ($ANPP_{stem}$, $R_{root}$, $BNPP_{fine.root}$; all $R^2$< 0.25; Table S6).

### 2020-12-16 scripts/refining_ratio_regressions_all.R
Update refining_ratio_regressions_all.R

@@ -145,101 +145,46 @@ final_sheet$CUE <- log(final_sheet$NPP_mean/final_sheet$GPP_mean)
 
 fixed.variables <- c("lat", "mat", "map", "TempSeasonality")
 
-png(file = "C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/results/figures/final_figures/ratio_regressions/log ratios/all.png", width = 3000, height = 2500, units = "px", res = 300)
+png(file = "C:/Users/becky/Dropbox (Smithsonian)/GitHub/Global_Productivity/manuscript/tables_figures/ratio_grid_plots.png", width = 3000, height = 2500, units = "px", res = 300)
 
 par(mfrow = c(3,4), mar = c(1,1,3,0.5), oma = c(5,4,5,1), xpd = F)
 
 ##### CUE
 
-n <- 1
-
-for(fixed.v in fixed.variables){
-  
-  final_sheet$fixed <- final_sheet[, fixed.v]
-  
-  mod <-  lmer(CUE ~ 1 + (1|geographic.area/plot.name), data = final_sheet, REML = F)
-  mod.linear <- lmer(CUE ~ poly(fixed, 1, raw = T) + (1|geographic.area/plot.name), data = final_sheet, REML = F)
-  
-  best.model <- as.character(aictab(list(mod = mod, mod.linear = mod.linear), sort = T)$Modname[1])
-  
-  if (best.model == "mod.linear") mod.full <- lmer(CUE ~ poly(fixed, 1, raw = T) + (1|geographic.area/plot.name), data = final_sheet, REML = F)
-  if (best.model == "mod") mod.full <- lmer(CUE ~ poly(fixed, 1, raw = T) + (1|geographic.area/plot.name), data = final_sheet, REML = F)
-  
-  significant.effect <- anova(mod, mod.full)$"Pr(>Chisq)"[2] < 0.05
-  significance <- anova(mod, mod.full)$"Pr(>Chisq)"[2]
-  sample.size <- length(final_sheet$CUE)
-  
-  if (best.model == "mod.linear") mod.full <- lmer(CUE ~ poly(fixed, 1, raw = T) + (1|geographic.area/plot.name), data = final_sheet, REML = T)
-  
-  newDat <- expand.grid(fixed = seq(min(final_sheet$fixed), max(final_sheet$fixed), length.out = 100))
[…]

### 2020-12-16 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@
 
-The latitudinal gradient in forest C flux rates, along with altitudinal gradients [@girardin_net_2010; @malhi_variation_2017], is driven primarily by climate, which is a significant driver of C fluxes across broad spatial scales [@luyssaert_co_2007; @cleveland_relationships_2011; @wei_forest_2010].
+The latitudinal gradient in forest C flux rates, along with altitudinal gradients [@girardin_net_2010; @malhi_variation_2017], is driven primarily by climate, which is a significant driver of C fluxes across broad spatial scales [@luyssaert_co_2007; @cleveland_relationships_2011; @wei_forest_2010; @muller-landau_patterns_2020].
 However, there is little consensus as to the shapes of these relationships or the best predictor variables.
@@
 Geographic areas--*i.e.*, spatially clustered sites--were defined within ForC using a hierarchical cluster analysis on the distance matrix of the sites and a cutoff of 25km [@anderson-teixeira_forc_2018].
-We experimented with inclusion of altitude as a fixed effect, but excluded it from the final models because it added very little explanatory power -- that is, the difference in AIC ($\Delta AIC$) relative to models excluding altitude was generally small (often $\Delta AIC$<2).
+We experimented with inclusion of altitude as a fixed effect, as productivity is known to decline with elevation [@muller-landau_patterns_2020], but excluded it from the final models because it added very little explanatory power -- that is, the difference in AIC ($\Delta AIC$) relative to models excluding altitude was generally small (often $\Delta AIC$<2).
 Effects were considered significant when inclusion of the fixed effect of interest resulted in p $\le$ 0.05 under an ANOVA test, and $\Delta AIC$ $\ge$ 2.0 relative to a corresponding null model. 
@@
 
-@mullerlandau_patterns_2020
 @zani_increased_2020

### 2020-12-17 manuscript/response to reviews.Rmd
all minor updates

@@
 
-**2.**  *(we should add some very clear statements to the abstract, intro, maybe discussion) ...*
+**2.
+We have added and clarified statements in the abstract, introduction, and discussion, to make it clear that our intention is to clarify uncertainty in these relationships.**  *(we should add some very clear statements to the abstract, intro, maybe discussion) ...*
 
@@
 **As this comment identifies, results from previous studies were complicated by other effects.
-The power of this synthesis lies in how it standardises for the effects of stand age and disturbance history, unambiguously clarifying these relationships.**
+The power of this synthesis lies in how it standardises for the effects of stand age and disturbance history, unambiguously clarifying these relationships.
+We have added text to the introduction to highlight the importance of standardisation of the dataset.**
 
-*(again, maybe make some modest change to the text that we can point to as a "response" to this review. Cite Yu.)*
-
-*have cited Yu in intro + discussion.
-In intro added following : "Productivity may vary with stand age, disturbance, and management [@de_lucia_forest_2007; @yu_high_2014; @simova_enigma_2017], making clear latitudinal patterns difficult to discern without standardization of the dataset.".
-I haven't edited discussion text other than adding citations to existing points - it seems to me to be covered well?*
+**Thankyou for highlighting the paper by Yu et al. (2014). We have cited this at several points in the text.**
 
@@
 
-**We have added CUE as requested. We also added a legend to clarify that none of the fits are significant (and therefore no statistics presented).**
+**We have added a legend to clarify that none of the fits are significant (and therefore no statistics presented).
+We have chosen not to add CUE to the figure, as we have focused on analysis of allocation within ratios.
+However, we have added text below Fig.
+S3 to clarify that we interpret these results in terms of variation in CUE.**
 

### 2020-12-17 manuscript/appendix.Rmd
all minor updates

@@
 ```
+Regressions test variation in allocation to component fluxes for three relationships: (1) $GPP = NPP + R_{auto}$, (2) $NPP = ANPP + BNPP$, and (3) $ANPP = ANPP_{foliage} + ANPP_{stem}$.
+We interpret analysis of variation in $NPP:R_{auto}$ in the context of CUE (GPP/NPP), as variation in CUE will be directly related to variation in $NPP:R_{auto}$.
+
 \elandscape

### 2020-12-17 manuscript/manuscript draft.Rmd
all minor updates

@@
 Carbon (C) fixation, allocation, and metabolism by trees set the basis for energy and material flows in forest ecosystems and define their interactions with Earth's changing climate. 
-However, we lack a cohesive synthesis on how forest carbon fluxes vary globally with respect to climate and one another. 
+However, while many studies have considered variation in productivity with latitude and climate, we lack a cohesive synthesis on how forest carbon fluxes vary globally with respect to climate and one another.
 Here, we draw upon 1,319 records from the Global Forest Carbon Database (ForC), representing all major forest types and the nine most significant autotrophic carbon fluxes, to comprehensively explore how annual C cycling in mature, undisturbed forests varies with latitude and climate on a global scale.
@@
 However, studies have differed in their conclusions regarding the shape of this relationship -- quite possibly because of lack of standardization with respect to methodology and stand history. 
-*Productivity may vary with stand age, disturbance, and management [@de_lucia_forest_2007; @yu_high_2014; @fernandez-martinez_spatial_2014; @simova_enigma_2017], making clear latitudinal patterns difficult to discern without standardization of the dataset.
-[???]*
-For instance, studies agree that gross primary productivity ($GPP$) increases continuously with decreasing latitude and is indisputably highest in tropical forests [@luyssaert_co_2007; @beer_terrestrial_2010; @jung_global_2011; @badgley_terrestrial_2019; @li_mapping_2019].
-In contrast, some studies have suggested that net primary productivity ($NPP$), or its aboveground portion ($ANPP$), exhibits a less distinct increase from temperate to tropical forests [@luyssaert_co_2007] -- or even a decrease [@huston_global_2009, but see @gillman_latitude_2015].
+Productivity may vary with stand age, disturbance, and management [@de_lucia_forest_2007; @yu_high_2014; @fernandez-martinez_spatial_2014; @simova_enigma_2017], making clear latitudinal patterns difficult to discern without standardization of the dataset.
+Studies agree that gross primary productivity ($GPP$) increases continuously with decreasing latitude and is indisputably highest in tropical forests [@luyssaert_co_2007; @beer_terrestrial_2010; @jung_global_2011; @badgley_terrestrial_2019; @li_mapping_2019].
+However, this relationship is more ambiguous in subsidiary fluxes.
+Some studies have suggested that net primary productivity ($NPP$), or its aboveground portion ($ANPP$), exhibits a less distinct increase from temperate to tropical forests [@luyssaert_co_2007] -- or even a decrease [@huston_global_2009, but see Gillman et al., 2015].
 A shallower increase in $NPP$ than in $GPP$ with decreasing latitude would align with the suggestion that tropical forests tend to have low carbon use efficiency [$CUE$= $NPP$/$GPP$; @de_lucia_forest_2007; @malhi_productivity_2012; @anderson-teixeira_carbon_2016].
@@
 The most comprehensive previous analysis of this type was @luyssaert_co_2007, which was based on a database <25% the size of the ForC version used here, did not control for effects of stand age, and examined global climatic trends in only three variables.
-In order to approach this broad topic, we simplify the major gaps in our knowledge to five broad questions and corresponding predictions (Table 1).
+In order to approach this broad topic, we simplify the major gaps in our knowledge to five broad review questions and corresponding predictions (Table 1).
 First, we ask how nine forest autotrophic carbon fluxes in ForC vary with latitude (*Q1*).
@@
 hypothesis_table[2,1] <- paste0(hypothesis_table[2,1], footnote_marker_number(1),"$^{,}$", footnote_marker_number(10)) #Luyssaert, Huston & Wolverton
-hypothesis_table[3,1] <- paste0(hypothesis_table[3,1], footnote_marker_number(1),"$^{,}$",footnote_marker_number(4),"$^{,}$",footnote_marker_number(9))
-hypothesis_table[4,1] <- paste0(hypothesis_table[4,1], footnote_marker_number(1),"$^{,}$",footnote_marker_number(4))
-hypothesis_table[5,1] <- paste0(hypothesis_table[5,1], footnote_marker_number(5))
+hypothesis_table[3,1] <- paste0(hypothesis_table[3,1], footnote_marker_number(1),"$^{,}$",footnote_marker_number(4),"$^{,}$",footnote_marker_number(9)) ## Luyssaert, Schuur, Piao, Wei
+hypothesis_table[4,1] <- paste0(hypothesis_table[4,1], footnote_marker_number(1),"$^{,}$",footnote_marker_number(4)) ## Luyssaert, Schuur, Wei
+hypothesis_table[5,1] <- paste0(hypothesis_table[5,1], footnote_marker_number(5)) ### Taylor, muller Landau
+hypothesis_table[8,1] <- paste0(hypothesis_table[8,1], footnote_marker_number()) ### fyllas
 hypothesis_table[11,1] <- paste0(hypothesis_table[11,1], footnote_marker_number(6),"$^{,}$",footnote_marker_number(7),"$^{,}$", footnote_marker_number(8))
@@
 (2015)", "Simova and Storch (2017)", "Schuur (2003)", "Taylor et al.
-(2016)", "Malhi (2012)", "Michaletz et al.
+(2017)", "Malhi (2012)", "Michaletz et al.
 (2014)", "Chu et al.
 (2016)", "Piao et al.
-(2010)", "Huston & Wolverton (2009)"), footnote_as_chunk = TRUE)
+(2010)", "Huston & Wolverton (2009)", "Wei et al.
+(2010)), footnote_as_chunk = TRUE)
 ```

### 2020-12-17 manuscript/manuscript draft.Rmd
updated table

@@
 (2015)", "Simova and Storch (2017)"
-hypothesis_table[2,1] <- paste0(hypothesis_table[2,1], footnote_marker_number(1),"$^{,}$", footnote_marker_number(10)) #Luyssaert, Huston & Wolverton
-hypothesis_table[3,1] <- paste0(hypothesis_table[3,1], footnote_marker_number(1),"$^{,}$",footnote_marker_number(4),"$^{,}$",footnote_marker_number(9)) ## Luyssaert, Schuur, Piao, Wei
-hypothesis_table[4,1] <- paste0(hypothesis_table[4,1], footnote_marker_number(1),"$^{,}$",footnote_marker_number(4)) ## Luyssaert, Schuur, Wei
-hypothesis_table[5,1] <- paste0(hypothesis_table[5,1], footnote_marker_number(5)) ### Taylor, muller Landau
-hypothesis_table[8,1] <- paste0(hypothesis_table[8,1], footnote_marker_number()) ### fyllas
-hypothesis_table[11,1] <- paste0(hypothesis_table[11,1], footnote_marker_number(6),"$^{,}$",footnote_marker_number(7),"$^{,}$", footnote_marker_number(8))
-hypothesis_table[12,1] <- paste0(hypothesis_table[12,1], footnote_marker_number(7),"$^{,}$", footnote_marker_number(8))
-hypothesis_table[13,1] <- paste0(hypothesis_table[13,1], footnote_marker_number(8))
+hypothesis_table[2,1] <- paste0(hypothesis_table[2,1], footnote_marker_number(1),"$^{,}$", footnote_marker_number(4)) #Luyssaert, Huston & Wolverton
+hypothesis_table[3,1] <- paste0(hypothesis_table[3,1], footnote_marker_number(1),"$^{,}$",footnote_marker_number(5),"$^{,}$",footnote_marker_number(6),"$^{,}$",footnote_marker_number(7)) ## Luyssaert, Schuur, Piao, Wei
+hypothesis_table[4,1] <- paste0(hypothesis_table[4,1], footnote_marker_number(1),"$^{,}$",footnote_marker_number(4),"$^{,}$",footnote_marker_number(7)) ## Luyssaert, Schuur, Wei
+hypothesis_table[5,1] <- paste0(hypothesis_table[5,1], footnote_marker_number(8),"$^{,}$",footnote_marker_number(9)) ### Taylor, muller Landau
+hypothesis_table[8,1] <- paste0(hypothesis_table[8,1], footnote_marker_number(10),"$^{,}$",footnote_marker_number(11)) ### fyllas nemani
+hypothesis_table[10,1] <- paste0(hypothesis_table[10,1], footnote_marker_number(12),"$^{,}$",footnote_marker_number(13)) ## wagner 2016 + 2014
+hypothesis_table[11,1] <- paste0(hypothesis_table[11,1], footnote_marker_number(14),"$^{,}$",footnote_marker_number(15),"$^{,}$", footnote_marker_number(16))
+hypothesis_table[12,1] <- paste0(hypothesis_table[12,1], footnote_marker_number(15),"$^{,}$", footnote_marker_number(16))
+hypothesis_table[13,1] <- paste0(hypothesis_table[13,1], footnote_marker_number(16))
 kable(hypothesis_table, format = "latex", booktabs = TRUE, caption = "Summary of review questions, corresponding expectations based on previous studies (when applicable), and results.
@@
   #column_spec(2:11, width = "1.8cm") %>%
-  kableExtra::landscape() 
-  #kableExtra::footnote(number = c("Luyssaert et al.
+  kableExtra::landscape() %>%
+  kableExtra::footnote(number = c("Luyssaert et al.
 (2007)", "Gillman et al.
-(2015)", "Simova and Storch (2017)", "Schuur (2003)", "Taylor et al.
-(2017)", "Malhi (2012)", "Michaletz et al.
+(2015)", "Simova and Storch (2017)", "Huston & Wolverton (2009)", "Schuur (2003)", "Piao et al.
+(2010)", "Wei et al.
+(2010)", "Taylor et al.
+(2017)", "Muller-Landau et al.
+(2020)", "Fyllas et al.
+(2017)", "Nemani et al.
+(2003)", "Wagner et al.
+(2014)" , "Wagner et al.
+(2016)", "Malhi (2012)", "Michaletz et al.
 (2014)", "Chu et al.
-(2016)", "Piao et al.
-(2010)", "Huston & Wolverton (2009)", "Wei et al.
-(2010)), footnote_as_chunk = TRUE)
+(2016)"), footnote_as_chunk = TRUE)
 ```

### 2020-12-17 manuscript/manuscript draft.Rmd
further updates to table and sorted references

@@
 
-The latitudinal gradient in forest C flux rates, along with altitudinal gradients [@girardin_net_2010; @malhi_variation_2017], is driven primarily by climate, which is a significant driver of C fluxes across broad spatial scales [@luyssaert_co_2007; @cleveland_relationships_2011; @wei_forest_2010; @muller-landau_patterns_2020].
+The latitudinal gradient in forest C flux rates, along with altitudinal gradients [@girardin_net_2010; @malhi_variation_2017], is driven primarily by climate, which is a significant driver of C fluxes across broad spatial scales [@luyssaert_co_2007; @cleveland_relationships_2011; @wei_forest_2010; @mullerlandau_patterns_2020].
 However, there is little consensus as to the shapes of these relationships or the best predictor variables.
@@
 hypothesis_table[5,1] <- paste0(hypothesis_table[5,1], footnote_marker_number(8),"$^{,}$",footnote_marker_number(9)) ### Taylor, muller Landau
-hypothesis_table[8,1] <- paste0(hypothesis_table[8,1], footnote_marker_number(10),"$^{,}$",footnote_marker_number(11)) ### fyllas nemani
-hypothesis_table[10,1] <- paste0(hypothesis_table[10,1], footnote_marker_number(12),"$^{,}$",footnote_marker_number(13)) ## wagner 2016 + 2014
-hypothesis_table[11,1] <- paste0(hypothesis_table[11,1], footnote_marker_number(14),"$^{,}$",footnote_marker_number(15),"$^{,}$", footnote_marker_number(16))
-hypothesis_table[12,1] <- paste0(hypothesis_table[12,1], footnote_marker_number(15),"$^{,}$", footnote_marker_number(16))
-hypothesis_table[13,1] <- paste0(hypothesis_table[13,1], footnote_marker_number(16))
+hypothesis_table[7,1] <- paste0(hypothesis_table[7,1], footnote_marker_number(10)) ## smith et al. 2020
+hypothesis_table[8,1] <- paste0(hypothesis_table[8,1], footnote_marker_number(11),"$^{,}$",footnote_marker_number(12)) ### fyllas nemani
+hypothesis_table[10,1] <- paste0(hypothesis_table[10,1], footnote_marker_number(13),"$^{,}$",footnote_marker_number(14)) ## wagner 2016 + 2014
+hypothesis_table[11,1] <- paste0(hypothesis_table[11,1], footnote_marker_number(15),"$^{,}$",footnote_marker_number(16),"$^{,}$", footnote_marker_number(17),"$^{,}$", footnote_marker_number(18)) ## and fernandez martinez
+hypothesis_table[12,1] <- paste0(hypothesis_table[12,1], footnote_marker_number(16),"$^{,}$", footnote_marker_number(17))
+hypothesis_table[13,1] <- paste0(hypothesis_table[13,1], footnote_marker_number(17))
+hypothesis_table[15,1] <-paste0(hypothesis_table[15,1], footnote_marker_number(18)) ##fernandez martinez
 kable(hypothesis_table, format = "latex", booktabs = TRUE, caption = "Summary of review questions, corresponding expectations based on previous studies (when applicable), and results.
@@
 (2017)", "Muller-Landau et al.
+(2020)", "Smith et al.
 (2020)", "Fyllas et al.
@@
 (2014)", "Chu et al.
-(2016)"), footnote_as_chunk = TRUE)
+(2016)", "Fernandez-Martinez et al.
+(2014)"), threeparttable = TRUE, footnote_as_chunk = TRUE)
 ```
@@
   @gillman_latitude_2015
+  @fernandez-martinez_spatial_2014
 ---
@@
 
-The effects of latitude and climate on C fluxes were analysed using mixed effects models using the package 'lme4' [@bates_fitting_2015] in R v.3.5.1 [@r_core_team_2018]. 
+The effects of latitude and climate on C fluxes were analysed using mixed effects models using the package 'lme4' [@bates_fitting_2015] in R v.3.5.1 [@r_2020]. 
 The basic model for all analyses included a fixed effect of latitude or climate and a random effect of plot nested within geographic area. 
 Geographic areas--*i.e.*, spatially clustered sites--were defined within ForC using a hierarchical cluster analysis on the distance matrix of the sites and a cutoff of 25km [@anderson-teixeira_forc_2018].
-We experimented with inclusion of altitude as a fixed effect, as productivity is known to decline with elevation [@muller-landau_patterns_2020], but excluded it from the final models because it added very little explanatory power -- that is, the difference in AIC ($\Delta AIC$) relative to models excluding altitude was generally small (often $\Delta AIC$<2).
+We experimented with inclusion of altitude as a fixed effect, as productivity is known to decline with elevation [@mullerlandau_patterns_2020], but excluded it from the final models because it added very little explanatory power -- that is, the difference in AIC ($\Delta AIC$) relative to models excluding altitude was generally small (often $\Delta AIC$<2).
 Effects were considered significant when inclusion of the fixed effect of interest resulted in p $\le$ 0.05 under an ANOVA test, and $\Delta AIC$ $\ge$ 2.0 relative to a corresponding null model. 
@@
 
-All analyses were conducted in R v.3.5.1 [@r_core_team_2018].
+All analyses were conducted in R v.3.5.1 [@r_2020].
 Code and data necessary to reproduce all results are available through GitHub (https://github.com/forc-db/Global_Productivity) and archived in Zenodo (DOI: TBD).

### 2020-12-17 manuscript/manuscript draft.Rmd
removed note on citations
Update manuscript draft.Rmd
@teixeirak added some notes on collalti discussion

@@
 Another previously observed pattern for which we find no support is a tendency for belowground C allocation to decrease with increasing temperature  [@moser_elevation_2011; @gill_belowground_2016]; rather, we observe no trends in allocation between $ANPP$ and $BNPP$ across latitudes.
-Failure to detect significant tends in C allocation with respect to climate in this analysis does not imply that none exist; rather, it suggests that, at this global scale, differences are subtle and/or that more careful methodological standardization and/or more data is required to detect them.
+Failure to detect significant trends in C allocation with respect to climate in this analysis does not imply that none exist; rather, it suggests that, at this global scale, differences are subtle and/or that more careful methodological standardization and/or more data is required to detect them.
 
-**in the paragraph above, discuss/ cite @collalti_forest_2020**
+**in the paragraph above, discuss/ cite @collalti_forest_2020** *I wonder here whether lack of trends in allocation may be confounded to some extent by the fact that in this dataset temperate forests are on average younger than tropical forests i.e.
+increases in CUE in tropical regions because of e.g.
+temperature as found by Collalti et al.
+(2020) are offset by the higher age of these forests.
+R2 comment 5 gets at this - I'm not sure how you would deal with this in the analysis as we don't have good information on tropical forest precise age generally, but we could discuss this further here?*
 
@@
 
-**CITATIONS TO ADD**:
-
-@zani_increased_2020
-
 ### Acknowledgements

### 2020-12-17 manuscript/response to reviews.Rmd
correctly formatted references in response to reviews + knit

@@
 Further, we do not argue that the spatial gradient represents the response of forests to climate change, as clearly discussed in the final paragraph of the discussion.
-However, we do argue that elucidating broad-scale patterns in C cycling across climatic gradients is a complementary approach to addressing a challenging problem [*sensu* @andersonteixeira_altered_2013, *GCB*].
+However, we do argue that elucidating broad-scale patterns in C cycling across climatic gradients is a complementary approach to addressing a challenging problem [*sensu* @anderson-teixeira_altered_2013, *GCB*].
 We note that studies using this approach have been published recently in *Nature* [@cook-patton_mapping_2020], *Science* [@sullivan_long-term_2020],  *GCB* [@trugman_climate_2019], *JGR* [@li_dissimilatory_2019], and *Scientific Reports* [@chen_spatial_2019].**

### 2020-12-17 manuscript/response to reviews.Rmd
edited for R2 final comment
update response to reviews

@@
 
- "Global patterns of forest autotrophic carbon fluxes" (GCB-20-1713).
+We are please to submit a revised version of our manuscript, "Global patterns of forest autotrophic carbon fluxes" (GCB-20-1713), for consideration for publication in *Global Change Biology*.
 
@@
 
-**Thank you for excellent suggestion to expand the discussion of growing season length around the references given.
+**Thank you for the excellent suggestion to expand the discussion of growing season length around the references given.
 We have re-written the paragraphs in the introduction and the discussion dealing with this issue, now citing the references listed in the comment above.**
@@
 This means that they are known to be older than 100 years of age, but the precise age is unknown.
-As a result, standardising this dataset by age is a challenge, and so we treat all forests in the analysis as mature old-growth forests, assuming minimal effect of age.** *CITE?*
-
-*(requires response)*
+As a result, standardising this dataset by age is a challenge, and so we treat all forests in the analysis as mature forests, assuming minimal effect of age beyond 100 years.**
 
@@
 
-*(requires response)*
+**We have made this edit.**
 

### 2020-12-17 manuscript/appendix.Rmd
fixed up SI formatting

--- 2020-12-01 manuscript/appendix.Rmd
@@
     fig_caption: yes
+    toc: true # table of content true
+    toc_depth: 2
 
@@
 
-\listoftables
-\listoffigures
-
 \newpage
@@
 \blandscape
+
+## Table S1. Climate variable definitions, sources, and abbreviations
 
@@
 S1 <- read.csv("tables_figures/climate_variables.csv", stringsAsFactors = FALSE, check.names = FALSE)
-kable(S1, format = "latex", booktabs = TRUE, caption = "Table S1. Climate variable definitions, sources, and abbreviations", escape = F)%>%
+kable(S1, format = "latex", booktabs = TRUE, escape = F)%>%
   kable_styling(latex_options = c("scale_down", "hold_position")) %>%
@@
 
+##  Table S2. Model summaries for each climate variable as a single fixed effect in models for each C flux
 ```{r eval = TRUE, echo=FALSE, warning=FALSE}
@@
 S3 <- read.csv("tables_figures/r2_table1.csv", stringsAsFactors = FALSE, check.names = FALSE)
-kable(S3, format = "latex", booktabs = TRUE, caption = "Table S2.
-Model form, $\\Delta$AIC, and R$^{2}$ for each climate variable as a single fixed effect in models for each C flux.", caption.short = "Table S2.
-Model form, $\\Delta$AIC and R$^{2}$ for each climate variable as a single fixed effect in models for each C flux", escape = F) %>%
+kable(S3, format = "latex", booktabs = TRUE,  escape = F) %>%
   kable_styling(latex_options = c("scale_down", "hold_position")) %>%
@@
 
+## Table S3. Signficance of relationships of forest C fluxes to MAT alone and in combination with MAP.
 ```{r eval = TRUE, echo=FALSE, warning=FALSE}
@@
 S5 <- read.csv("tables_figures/mat_map.csv", stringsAsFactors = FALSE, check.names = FALSE)
-kable(S5, format = "latex", booktabs = TRUE, caption = "Table S3.
-Signficance of relationships of forest C fluxes to MAT alone and in combination with MAP.
-$R^{2}$ values refer to the best model.",  caption.short = "Table S3.
-Signficance of relationships of forest C fluxes to MAT alone and in combination with MAP.", escape = F)%>%
+kable(S5, format = "latex", booktabs = TRUE,  escape = F)%>%
   kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 12) %>%
@@
 \newpage
+## Table S4. Comparison of growing season length and MAT as predictors of forest C fluxes
 ```{r eval = TRUE, echo=FALSE, warning=FALSE}
@@
 S2 <- read.csv("tables_figures/mat_growing_season.csv", stringsAsFactors = FALSE, check.names = FALSE)
-kable(S2, format = "latex", booktabs = TRUE, caption = "Table S4. Comparison of growing season length and MAT as predictors of forest C fluxes", escape = F)%>%
+kable(S2, format = "latex", booktabs = TRUE, escape = F)%>%
   kable_styling(latex_options = c("scale_down", "hold_position")) %>%
@@
 \newpage
+## Table S5. Best (lowest AIC) single-climate variable models by C flux.
 ```{r eval = TRUE, echo=FALSE, warning=FALSE}
@@
 S3 <- read.csv("tables_figures/best_models_each_flux.csv", stringsAsFactors = FALSE, check.names = FALSE)
-kable(S3, format = "latex", booktabs = TRUE, caption = "Table S5.
-Best (lowest $\\Delta$AIC) single-climate variable models by C flux.", caption.short = "Table S5.
-Best models by carbon flux", escape = F) %>%
+kable(S3, format = "latex", booktabs = TRUE, escape = F) %>%
   kable_styling(latex_options = c("scale_down", "hold_position")) %>%
@@
 \blandscape
+## Table S6. Pairwise comparisons of correlations with climate variables between C fluxes, with all analyses conducted on a set of sites common to each pair.
 ```{r eval = TRUE, echo=FALSE, warning=FALSE}
@@
 S4 <- read.csv("tables_figures/pairwise_comparisons.csv", stringsAsFactors = FALSE, check.names = FALSE)
-kable(S4, format = "latex", booktabs = TRUE, caption = "Table S6.
-Pairwise comparisons of correlations with climate variables between C fluxes, with all analyses conducted on a set of sites common to each pair.", escape = F) %>%
+kable(S4, format = "latex", booktabs = TRUE, escape = F) %>%
   kable_styling(latex_options = c("scale_down", "hold_position")) %>%
@@
 \elandscape
-
-\newpage
-```{r echo=FALSE, fig.cap = "Figure S1: Maps showing distribution of samples for the nine forest C fluxes analyzed", out.height ='95%', fig.pos='H'}
+## Figure S1. Sample distribution maps for the nine forest C fluxes analyzed
+```{r echo=FALSE, fig.cap = "Figure S1: Maps showing distribution of samples for the nine forest C fluxes analyzed", out.height ='90%', fig.pos='H'}
 
@@
 ```
-\newpage 
 \blandscape
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S2: Scatterplots and Pearson's R values for relationships among climate variables and latitude", out.height='95%', fig.pos='H'}
+## Figure S2. Correlations among climate variables and latitude
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S2: Scatterplots and Pearson's R values for relationships among climate variables and latitude", out.height='90%', fig.pos='H'}
 
@@
 
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S3: Ratios among forest C fluxes as a function of latitude and climate variables.", out.width ='80%', fig.pos='H'}
+## Figure S3. Ratios among forest C fluxes as a function of latitude and climate variables
+```{r echo=FALSE, fig.align = "center",fig.cap = "Figure S3: Ratios among forest C fluxes as a function of latitude and climate variables.
+Regressions test variation in allocation to component fluxes for three relationships: (1) $GPP = NPP + R_{auto}$, (2) $NPP = ANPP + BNPP$, and (3) $ANPP = ANPP_{foliage} + ANPP_{stem}$.
+We interpret analysis of variation in $NPP:R_{auto}$ in the context of CUE (GPP/NPP), as variation in CUE will be directly related to variation in $NPP:R_{auto}$.", out.width ='75%', fig.pos='H'}
 
@@
 \blandscape
-```{r echo=FALSE, fig.cap = "Figure S4: Individual plots of forest C fluxes in relation to mean annual climate, part 1.", out.height='95%', fig.pos='H'}
+## Figure S4. Individual plots of forest C fluxes in relation to mean annual climate, part 1
+```{r echo=FALSE, fig.cap = "Figure S4: Individual plots of forest C fluxes in relation to mean annual climate, part 1.", out.height='90%', fig.pos='H'}
 
@@
 \newpage
-
-```{r echo=FALSE, fig.cap = "Figure S5: Individual plots of forest C fluxes in relation to mean annual climate, part 2.", out.height='95%', fig.pos='H'}
+## Figure S5. Individual plots of forest C fluxes in relation to mean annual climate, part 2
+```{r echo=FALSE, fig.cap = "Figure S5: Individual plots of forest C fluxes in relation to mean annual climate, part 2.", out.height='90%', fig.pos='H'}
 
@@
 
-```{r echo=FALSE, fig.cap = "Figure S6: Individual plots of forest C fluxes in relation to mean climate seasonality, part 1.", out.height ='95%', fig.pos='H'}
+## Figure S6. Individual plots of forest C fluxes in relation to mean climate seasonality, part 1
+```{r echo=FALSE, fig.cap = "Figure S6: Individual plots of forest C fluxes in relation to mean climate seasonality, part 1.", out.height ='80%', fig.pos='H'}
 
@@
 \newpage
-```{r echo=FALSE, fig.cap = "Figure S7: Individual plots of forest C fluxes in relation to mean climate seasonality, part 2.", out.width ='100%', fig.pos='H'}
+## Figure S7. Individual plots of forest C fluxes in relation to mean climate seasonality, part 2
+```{r echo=FALSE, fig.cap = "Figure S7: Individual plots of forest C fluxes in relation to mean climate seasonality, part 2.", out.width ='80%', fig.pos='H'}
 
@@
 \newpage
-```{r echo=FALSE, fig.cap = "Figure S8: Growing season length-standardized forest C fluxes in relation to mean growing season climate, part 1.", out.width ='100%', fig.pos='H'}
+## Figure S8. Growing season length-standardized forest C fluxes in relation to mean growing season climate, part 1
+```{r echo=FALSE, fig.cap = "Figure S8: Growing season length-standardized forest C fluxes in relation to mean growing season climate, part 1.", out.width ='95%', fig.pos='H'}
 
@@
 \newpage
-```{r echo=FALSE, fig.cap = "Figure S9: Growing season length-standardized forest C fluxes in relation to mean growing season climate, part 2.", out.width ='100%', fig.pos='H'}
+## Figure S9. Growing season length-standardized forest C fluxes in relation to mean growing season climate, part 2
+```{r echo=FALSE, fig.cap = "Figure S9: Growing season length-standardized forest C fluxes in relation to mean growing season climate, part 2.", out.width ='95%', fig.pos='H'}
 
@@
 \newpage
-# References
+## References

### 2020-12-17 manuscript/response to reviews.Rmd
working on manuscript & response to reviews

@@
 **2.
-We have added and clarified statements in the abstract, introduction, and discussion, to make it clear that our intention is to clarify uncertainty in these relationships.**  *(we should add some very clear statements to the abstract, intro, maybe discussion) ...*
+We have made the following revisions to the language in the abstract, introduction, and discussion to make it clear that our intention is to provide a quantitative review that clarifies uncertainty in these relationships:**
+
+- **Abstract: "Here, we draw upon 1,319 records from the Global Forest Carbon Database (ForC), representing all major forest types and the nine most significant autotrophic carbon fluxes, to comprehensively ~~explore~~ *review* how annual C cycling in mature, undisturbed forests varies with latitude and climate on a global scale."**
+- **Introduction: "In order to approach this broad topic, we simplify the major gaps in our knowledge to five broad *review questions* and corresponding predictions (Table 1)."**
+
+-
 

### 2020-12-17 manuscript/manuscript draft.Rmd
corrected unicode error in bib file
working on manuscript & response to reviews

@@
 However, while many studies have considered variation in productivity with latitude and climate, we lack a cohesive synthesis on how forest carbon fluxes vary globally with respect to climate and one another.
-Here, we draw upon 1,319 records from the Global Forest Carbon Database (ForC), representing all major forest types and the nine most significant autotrophic carbon fluxes, to comprehensively explore how annual C cycling in mature, undisturbed forests varies with latitude and climate on a global scale.
+Here, we draw upon 1,319 records from the Global Forest Carbon Database (ForC), representing all major forest types and the nine most significant autotrophic carbon fluxes, to comprehensively review how annual C cycling in mature, undisturbed forests varies with latitude and climate on a global scale.
 We show that, across all flux variables analyzed, C cycling decreases continuously with absolute latitude -- a finding that confirms multiple previous studies and contradicts the idea that net primary productivity of temperate forests rivals that of tropical forests.
@@
 The recent development of the Global Forest Carbon database (ForC), which synthesizes multiple variables and includes records of stand history [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018], opens up the possibility for a standardized analysis of global scale variation in multiple C fluxes and the principle climatic drivers of these patterns.
-Our analyses represent a major step forward in relation to previous work.
-The most comprehensive previous analysis of this type was @luyssaert_co_2007, which was based on a database <25% the size of the ForC version used here, did not control for effects of stand age, and examined global climatic trends in only three variables.
 In order to approach this broad topic, we simplify the major gaps in our knowledge to five broad review questions and corresponding predictions (Table 1).
@@
 Finally, we consider the relationship between C flux and seasonality, considering the role of seasonality in explaining variation in carbon fluxes (*Q4*), and the influence of climate on C flux standardized by growing season length (*Q5*).
+Our analyses represent a major step forward in relation to previous work [@luyssaert_co_2007] in that we examine global climatic trends in more variables (9 vs.
+$\le$ 3), draw from a much larger database (>4 times more records analyzed), and control for the effects of stand age, disturbance, and management (not consistenly done previously).
 
@@
 
-Temperature and its seasonality were the primary drivers of C fluxes on the global scale (Table 1, *Q2*,*Q4*; Figs.
+Temperature and its seasonality were the primary drivers of C fluxes on the global scale (Table 1, *Q2*, *Q4*; Figs.
 2-4), consistent with a long legacy of research identifying temperature as a primary driver of forest ecosystem C cycling [e.g., @lieth_primary_1973; @luyssaert_co_2007; @wei_forest_2010].

### 2020-12-17 manuscript/response to reviews.Rmd
working on resonse to reviews

@@
 
-**bold- responses**
-
-*italic- tentative responses/ comments*
-
 Reviewer(s)' Comments to Author:
@@
 Reviewer: 1
-
-*Becky, we probably want to reword some of these responses to R1 a bit, as it's possible that this will go back to R1.
-Also, it could be good to modify the manuscript in response to some of the comments to ensure that other readers don't come up with the same mis-conceptions.*
 
@@
 **2.
-We have made the following revisions to the language in the abstract, introduction, and discussion to make it clear that our intention is to provide a quantitative review that clarifies uncertainty in these relationships:**
+We have made the following revisions to the language in the abstract and introduction to clarify that our intention is to provide a quantitative review that clarifies uncertainty in these relationships:**
 
@@
 - **Introduction: "In order to approach this broad topic, we simplify the major gaps in our knowledge to five broad *review questions* and corresponding predictions (Table 1)."**
-
--
 
@@
 
-**1. They are underlain by a large, standardized database that accounts for factors such as stand age and management, which have not always been accounted for.**
+-  **They are underlain by a large, standardized database that accounts for factors such as stand age and management, which have not always been accounted for.**
 
-**2.
-We use mixed effects model to account for random effects of location (plot nested within geographic area), thereby accounting for potential pseudoreplication, which has not always been done in the past.**
+-  **We use mixed effects model to account for random effects of location (plot nested within geographic area), thereby accounting for potential pseudoreplication, which has not always been done in the past.**
 
-**3. We examined nine carbon fluxes and relationships among them--more than has been concurrently analyzed in the past.**
+-  **We examined nine carbon fluxes and relationships among them--more than has been concurrently analyzed in the past.**
 
-**4.
-For each flux variable, we tested a wide range of models: univariate responses to 12 climate variables (3 functional fit types for each), along with multivariate models (MAT x MAP for all, along with other combinations not presented).**
+-  **For each flux variable, we tested a wide range of models: univariate responses to 12 climate variables (3 functional fit types for each), along with multivariate models (MAT x MAP for all, along with other combinations not presented).**
 
@@
 
-**To help clarify the scope of this analysis to readers, we added the following statement to the last paragraph of the introduction: "The most comprehensive analysis of this type was @luyssaert_co_2007, which was based on a database <25% the size of the ForC version used here, did not control for effects of stand age, and examined global climatic trends in only three variables."**
+**To help clarify the scope of this analysis to readers, we added the following statement to the last paragraph of the introduction: "Our analyses represent a major step forward in relation to previous work [e.g., @luyssaert_co_2007] in that we examine global climatic trends in more variables (9 vs.
+$\le$ 3), draw from a much larger database (>4 times more records analyzed), and control for the effects of stand age, disturbance, and management."**
 
-*(Do we like this response??)*
-
-RE Luyssaert et al : *We note in particular that mean stand age was extremely variable across biomes, and while mean stand age in tropical forests was >100 years (old growth) in all but one extra-tropical biome, mean stand age was >100.
-This is likely to elevate productivity estimates for extra-tropical biomes, highlighting the need for a comprehensive, standardised dataset for analysis.*
+**We note in particular that in @luyssaert_co_2007, mean stand age was extremely variable across biomes, and while mean stand age in tropical forests was >100 years (old growth) in all but one extra-tropical biome, mean stand age was <100.
+This is likely to elevate productivity estimates for extra-tropical biomes, highlighting the need for a comprehensive, standardised dataset for analysis.**
 

### 2020-12-17 manuscript/manuscript draft.Rmd
working on resonse to reviews

@@
 Finally, we consider the relationship between C flux and seasonality, considering the role of seasonality in explaining variation in carbon fluxes (*Q4*), and the influence of climate on C flux standardized by growing season length (*Q5*).
-Our analyses represent a major step forward in relation to previous work [@luyssaert_co_2007] in that we examine global climatic trends in more variables (9 vs.
-$\le$ 3), draw from a much larger database (>4 times more records analyzed), and control for the effects of stand age, disturbance, and management (not consistenly done previously).
+Our analyses represent a major step forward in relation to previous work [e.g., @luyssaert_co_2007] in that we examine global climatic trends in more variables (9 vs.
+$\le$ 3), draw from a much larger database (>4 times more records analyzed), and control for the effects of stand age, disturbance, and management.
 

### 2020-12-18 manuscript/response to reviews.Rmd
update response to reviews

@@
 
-**Thankyou for highlighting the paper by Yu et al. (2014). We have cited this at several points in the text.**
+**Thank you for highlighting the paper by Yu et al. (2014). We have cited this at several points in the text.**
 
@@
 However, we do argue that elucidating broad-scale patterns in C cycling across climatic gradients is a complementary approach to addressing a challenging problem [*sensu* @anderson-teixeira_altered_2013, *GCB*].
-We note that studies using this approach have been published recently in *Nature* [@cook-patton_mapping_2020], *Science* [@sullivan_long-term_2020],  *GCB* [@trugman_climate_2019], *JGR* [@li_dissimilatory_2019], and *Scientific Reports* [@chen_spatial_2019].**
+We note that studies using this approach have been published recently in *Nature* [@cook-patton_mapping_2020], *Science* [@sullivan_long-term_2020],  *GCB* [@trugman_climate_2019], *New Phytologist*, as a Tansley review [@mullerlandau_patterns_2020], *JGR* [@li_dissimilatory_2019], and *Scientific Reports* [@chen_spatial_2019].**
 
-*The reviewer's apparent wholesale rejection of this approach seems short-sighted, and the expectation that a single study attempt to explain mechanisms behind global patterns in nine flux variables seems naïve.
-The implied standard that studies of this type should "reveal the biological mechanisms" could be interpreted in various ways, but the best way of getting at underlying biological mechanisms would be through modeling.
-Until databases such as ForC are fully integrated with mechanistic ecosystem models, our ability to provide truly mechanistic explanations of the observed patterns remains limited, but the observed correlations provide important insights into the broad-scale constraints of climate on forest carbon cycling.*
+**We also agree that spatial gradients do not necessarily represent the response of forests to climate change, and have made some changes to to ensure that we are not giving that message:**
+
+- **Abstract: "~~In a period of accelerating climatic change, this improved understanding of the fundamental climatic controls on forest C cycling sets a foundation for understanding patterns of change.~~  In an era when forests will play a critical yet uncertain role in shaping Earth's rapidly changing climate, our synthesis provides a foundation for understanding global patterns in forest C cycling."**
+- **Discussion:  **
 

### 2020-12-18 manuscript/response to reviews.Rmd
finished responses to R1

@@
 However, we do argue that elucidating broad-scale patterns in C cycling across climatic gradients is a complementary approach to addressing a challenging problem [*sensu* @anderson-teixeira_altered_2013, *GCB*].
-We note that studies using this approach have been published recently in *Nature* [@cook-patton_mapping_2020], *Science* [@sullivan_long-term_2020],  *GCB* [@trugman_climate_2019], *New Phytologist*, as a Tansley review [@mullerlandau_patterns_2020], *JGR* [@li_dissimilatory_2019], and *Scientific Reports* [@chen_spatial_2019].**
+We note that studies using this approach have been published recently in *Nature* [@cook-patton_mapping_2020], *Science* [@sullivan_long-term_2020],  *GCB* [@trugman_climate_2019], *JGR* [@li_dissimilatory_2019],  *Scientific Reports* [@chen_spatial_2019], and as a Tansley review in *New Phytologist* [@mullerlandau_patterns_2020].**
 
@@
 - **Abstract: "~~In a period of accelerating climatic change, this improved understanding of the fundamental climatic controls on forest C cycling sets a foundation for understanding patterns of change.~~  In an era when forests will play a critical yet uncertain role in shaping Earth's rapidly changing climate, our synthesis provides a foundation for understanding global patterns in forest C cycling."**
-- **Discussion:  **
+- **Discussion: "~~Our analysis clarifies how annual forest autotrophic C fluxes vary with latitude and climate on a global scale, with some important implications for how forest C cycling relates to climate and, by extension, how it is likely to respond to climatic warming.~~ Our analysis clarifies how annual forest autotrophic C fluxes vary with latitude and climate on a global scale.
+" **
 

### 2020-12-18 manuscript/response to reviews.Rmd
more on R1 (not quite done)

@@
 The most comprehensive similar study is @luyssaert_co_2007, which was published in *GCB* more than a decade ago, was based on a database <25% the size of the ForC version used here, did not control for effects of stand age/ disturbance history, and examined latitudinal/ global climatic trends in only three variables.
+We note in particular that in @luyssaert_co_2007, mean stand age was extremely variable across biomes, and while mean stand age in tropical forests was >100 years (old growth) in all but one extra-tropical biome, mean stand age was <100.
+This is likely to elevate productivity estimates for extra-tropical biomes, highlighting the need for a comprehensive, standardised dataset for analysis.
 Thus, the current analysis will be at the clear forefront of studies examining correlations between forest carbon flux and climate at the global scale.**
@@
 $\le$ 3), draw from a much larger database (>4 times more records analyzed), and control for the effects of stand age, disturbance, and management."**
-
-**We note in particular that in @luyssaert_co_2007, mean stand age was extremely variable across biomes, and while mean stand age in tropical forests was >100 years (old growth) in all but one extra-tropical biome, mean stand age was <100.
-This is likely to elevate productivity estimates for extra-tropical biomes, highlighting the need for a comprehensive, standardised dataset for analysis.**
 
@@
 The power of this synthesis lies in how it standardises for the effects of stand age and disturbance history, unambiguously clarifying these relationships.
-We have added text to the introduction to highlight the importance of standardisation of the dataset.**
+We have added text to the introduction to highlight the importance of standardisation of the dataset:**
+
+*GIVE THE TEXT ADDED*
 

### 2020-12-18 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@
 Moreover, forest ecosystem responses to climatic changes outside the temperature range to which forest communities are adapted and acclimatized will not necessarily parallel responses across geographic gradients in climate.
-Indeed, tree-ring studies from forests around the world indicate that tree growth rates -- along with $ANPP_{stem}$ and possibly other ecosystem C fluxes -- respond negatively to temperature [e.g., @helcoski_growing_2019].
+Indeed, tree-ring studies from forests around the world indicate that tree growth rates -- along with $ANPP_{stem}$ and possibly other ecosystem C fluxes -- respond negatively to temperature [e.g., @helcoski_growing_2019]**REFS**.
 Furthermore, in the tropics, climate change will push temperatures beyond any contemporary climate, and there are some indications that this could reduce forest C flux rates  [@mau_temperate_2018; @sullivan_long-term_2020] if paralleled by $VPD$ increases [@smith_empirical_2020].

### 2020-12-18 manuscript/manuscript draft.Rmd
discussion on Collalti
#121

@beckybanbury , please review what I wrote here--see if you agree.

@@
 We find no significant trend in the allocation of $GPP$ between production and respiration across latitude or climate ($NPP$:$R_{auto}$; Fig.
-S3), counter to the idea that tropical forests have anomalously low $CUE$ [@de_lucia_forest_2007; @malhi_productivity_2012; @anderson-teixeira_carbon_2016].
-Rather, differences in $CUE$ between old-growth tropical forests relative to (mostly younger) extratropical forests are likely an artifact of comparing stands of different age, as $CUE$ is known to decline with forest age [@de_lucia_forest_2007; @piao_forest_2010; @collalti_forest_2020].
+S3), counter to the idea that tropical forests have anomalously low $CUE$ [@de_lucia_forest_2007; @malhi_productivity_2012; @anderson-teixeira_carbon_2016], which is predicted by most models [@collalti_forest_2020].
+In contrast, @collalti_forest_2020 found that forest production efficiency increased with temperature--a finding that is consistent in direction with the insignificant trends observed here (Fig.
+S3).
+Previously observed differences in $CUE$ between old-growth tropical forests relative to (mostly younger) extratropical forests are likely an artifact of comparing stands of different age, as $CUE$ is known to decline with forest age [@de_lucia_forest_2007; @piao_forest_2010; @collalti_forest_2020].
 Another previously observed pattern for which we find no support is a tendency for belowground C allocation to decrease with increasing temperature  [@moser_elevation_2011; @gill_belowground_2016]; rather, we observe no trends in allocation between $ANPP$ and $BNPP$ across latitudes.
 Failure to detect significant trends in C allocation with respect to climate in this analysis does not imply that none exist; rather, it suggests that, at this global scale, differences are subtle and/or that more careful methodological standardization and/or more data is required to detect them.
-
-**in the paragraph above, discuss/ cite @collalti_forest_2020** *I wonder here whether lack of trends in allocation may be confounded to some extent by the fact that in this dataset temperate forests are on average younger than tropical forests i.e.
-increases in CUE in tropical regions because of e.g.
-temperature as found by Collalti et al.
-(2020) are offset by the higher age of these forests.
-R2 comment 5 gets at this - I'm not sure how you would deal with this in the analysis as we don't have good information on tropical forest precise age generally, but we could discuss this further here?*
 

### 2020-12-18 manuscript/response to reviews.Rmd
small updates

@@
 
-*(This is a great comment.
-The references given in the comment are more robust work on the subject of productivity and growing season length and make the framing much richer.
-{Interestingly, this links to another paper that current intern Cameron Dow is working on, which uses dendrobands to show that warmer springs shift growth earlier in the year but have at most modest effect on total productivity.}  More broadly, I think there may be other areas of the paper where more engangement with EC literature would make the story richer, but this will require some lit search.)*
-
 2.

### 2020-12-18 manuscript/response to reviews.Rmd
given the text added

@@
 The power of this synthesis lies in how it standardises for the effects of stand age and disturbance history, unambiguously clarifying these relationships.
-We have added text to the introduction to highlight the importance of standardisation of the dataset:**
-
-*GIVE THE TEXT ADDED*
+We have added text to the introduction to highlight the importance of standardisation of the dataset: "Productivity may vary with stand age, disturbance, and management [@de_lucia_forest_2007; @yu_high_2014; @fernandez-martinez_spatial_2014; @simova_enigma_2017], making clear latitudinal patterns difficult to discern without standardization of the dataset."**
 

### 2020-12-18 manuscript/appendix.Rmd
fixed up figure legends
@beckybanbury , I couldn't figure out how to insert the "≤" symbol for "p≤0.05", so I just put "p<0.05" (Figs. S4-S9). If you're able to fix that, great.

@@
 ## Figure S1. Sample distribution maps for the nine forest C fluxes analyzed
-```{r echo=FALSE, fig.cap = "Figure S1: Maps showing distribution of samples for the nine forest C fluxes analyzed", out.height ='90%', fig.pos='H'}
+```{r echo=FALSE, fig.cap = "Figure S1: Maps showing distribution of samples for the nine forest C fluxes analyzed. Variables are defined in Table 2.", out.height ='90%', fig.pos='H'}
 
@@
 ## Figure S2. Correlations among climate variables and latitude
-```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S2: Scatterplots and Pearson's R values for relationships among climate variables and latitude", out.height='90%', fig.pos='H'}
+```{r echo=FALSE, fig.align = "center", fig.cap = "Figure S2: Scatterplots and Pearson's R values for relationships among climate variables and latitude.
+Climate variables are as defined in Table S1.", out.height='90%', fig.pos='H'}
 
@@
 Regressions test variation in allocation to component fluxes for three relationships: (1) $GPP = NPP + R_{auto}$, (2) $NPP = ANPP + BNPP$, and (3) $ANPP = ANPP_{foliage} + ANPP_{stem}$.
-We interpret analysis of variation in $NPP:R_{auto}$ in the context of CUE (GPP/NPP), as variation in CUE will be directly related to variation in $NPP:R_{auto}$.", out.width ='75%', fig.pos='H'}
+We interpret analysis of variation in $NPP:R_{auto}$ in the context of CUE ($CUE = NPP/GPP = NPP/(NPP+R_auto)$), as variation in CUE will be directly related to variation in $NPP:R_{auto}$.
+All relationships are non-signficant (p>0.05).", out.width ='80%', fig.pos='H'}
 
-  knitr::include_graphics("tables_figures/ratio_grid_plots.png")
+  knitr::include_graphics("tables_figures/ratio_grid_plots_nolegend.png")
 ```
@@
 ## Figure S4. Individual plots of forest C fluxes in relation to mean annual climate, part 1
-```{r echo=FALSE, fig.cap = "Figure S4: Individual plots of forest C fluxes in relation to mean annual climate, part 1.", out.height='90%', fig.pos='H'}
+```{r echo=FALSE, fig.cap = "Figure S4: Individual plots of forest C fluxes in relation to mean annual climate, part 1.
+Statistical signficance is indicated by solid (p < 0.05) and dashed (p > 0.05) regression lines.
+Fits shown are of the form with the lowest AIC for each C variable- climate variable combination (Table S2).", out.height='85%', fig.pos='H'}
 
@@
 ## Figure S5. Individual plots of forest C fluxes in relation to mean annual climate, part 2
-```{r echo=FALSE, fig.cap = "Figure S5: Individual plots of forest C fluxes in relation to mean annual climate, part 2.", out.height='90%', fig.pos='H'}
+```{r echo=FALSE, fig.cap = "Figure S5: Individual plots of forest C fluxes in relation to mean annual climate, part 2.
+Statistical signficance is indicated by solid (p < 0.05) and dashed (p > 0.05) regression lines.
+Fits shown are of the form with the lowest AIC for each C variable- climate variable combination (Table S2).", out.height='85%', fig.pos='H'}
 
@@
 ## Figure S6. Individual plots of forest C fluxes in relation to mean climate seasonality, part 1
-```{r echo=FALSE, fig.cap = "Figure S6: Individual plots of forest C fluxes in relation to mean climate seasonality, part 1.", out.height ='80%', fig.pos='H'}
+```{r echo=FALSE, fig.cap = "Figure S6: Individual plots of forest C fluxes in relation to mean climate seasonality, part 1.
+Statistical signficance is indicated by solid (p < 0.05) and dashed (p > 0.05) regression lines.
+Fits shown are of the form with the lowest AIC for each C variable- climate variable combination (Table S2).", out.height ='80%', fig.pos='H'}
 
@@
 ## Figure S7. Individual plots of forest C fluxes in relation to mean climate seasonality, part 2
-```{r echo=FALSE, fig.cap = "Figure S7: Individual plots of forest C fluxes in relation to mean climate seasonality, part 2.", out.width ='80%', fig.pos='H'}
+```{r echo=FALSE, fig.cap = "Figure S7: Individual plots of forest C fluxes in relation to mean climate seasonality, part 2.
+Statistical signficance is indicated by solid (p < 0.05) and dashed (p > 0.05) regression lines.
+Fits shown are of the form with the lowest AIC for each C variable- climate variable combination (Table S2).", out.width ='80%', fig.pos='H'}
 
@@
 ## Figure S8. Growing season length-standardized forest C fluxes in relation to mean growing season climate, part 1
-```{r echo=FALSE, fig.cap = "Figure S8: Growing season length-standardized forest C fluxes in relation to mean growing season climate, part 1.", out.width ='95%', fig.pos='H'}
+```{r echo=FALSE, fig.cap = "Figure S8: Growing season length-standardized forest C fluxes in relation to mean growing season climate, part 1.
+Statistical signficance is indicated by solid (p < 0.05) and dashed (p > 0.05) regression lines.", out.width ='95%', fig.pos='H'}
 
@@
 ## Figure S9. Growing season length-standardized forest C fluxes in relation to mean growing season climate, part 2
-```{r echo=FALSE, fig.cap = "Figure S9: Growing season length-standardized forest C fluxes in relation to mean growing season climate, part 2.", out.width ='95%', fig.pos='H'}
+```{r echo=FALSE, fig.cap = "Figure S9: Growing season length-standardized forest C fluxes in relation to mean growing season climate, part 2.
+Statistical signficance is indicated by solid (p < 0.05) and dashed (p > 0.05) regression lines.", out.width ='95%', fig.pos='H'}
 

### 2020-12-18 manuscript/response to reviews.Rmd
R2, point 2

@@
 
-**We found no significant relationships between flux ratios (allocation to fluxes) across latitude and three climate variables tested, and as a result have not presented figures illustrating this in the text.
-We state that these results were non-significant in the results section.
-We have updated Fig.
-S3 to clarify that none of the fits are significant (and therefore no statistics are presented).**
-
-*from Helene: maybe we should also address statistical power*
-
-*(requires response)*
+**We found no significant relationships between flux ratios (allocation to fluxes) across latitude and three climate variables tested, and as a result have not presented figures illustrating this in the main article.
+We state that these results were non-significant (p>0.05) in the results section.
+We have updated the Fig.
+S3 legend to clarify that none of the fits are significant (and therefore no statistics are presented).**
 

### 2020-12-18 manuscript/manuscript draft.Rmd
small updates
R2, point 2

@@
 In particular, growing season length has been linked to $ANPP$, $NPP$, $GPP$, and net ecosystem exchange of CO~2~ [$NEE$, or the difference between $GPP$ and ecosystem respiration; @kerkhoff_plant_2005; @michaletz_convergence_2014; @churkina_spatial_2005; @keenan_net_2014; @zhou_explaining_2016].
-However, the relative importance of climate within the growing season, as opposed to growing season length, remains debated.
+However, the relative importance of growing season length, as opposed to climate within the growing season, remains debated.
 On one end of the spectrum, some studies have suggested that the influence of temperature on C fluxes may be limited to determining the length of the frost-free growing season, and that climate within the growing season has little influence on C fluxes because of plant adaptation and acclimatization to local climates [@kerkhoff_plant_2005; @enquist_adaptive_2007; @michaletz_convergence_2014; @michaletz_drivers_2018].
@@
 S3).
-Of 12 relationships tested (3 ratios among C flux variables regressed against latitude, $MAT$, $MAP$ and temperature seasonality), none were significant.
+Of 12 relationships tested (3 ratios among C flux variables regressed against latitude, $MAT$, $MAP$ and temperature seasonality), none were significant (all p>0.05).
 

### 2020-12-18 manuscript/response to reviews.Rmd
Update response to reviews.Rmd
#125

@@
 
-*(add something on this to the discussion, at the same time addressing R1's 3rd comment)*
+**We have added the following comments on models to the discussion:**
+- **We find no significant trend in the allocation of $GPP$ between production and respiration across latitude or climate ($NPP$:$R_{auto}$; Fig.
+S3), counter to the idea that tropical forests have anomalously low $CUE$ [@de_lucia_forest_2007; @malhi_productivity_2012; @anderson-teixeira_carbon_2016], *which is predicted by most models [@collalti_forest_2020]*.
+In contrast, @collalti_forest_2020 found that forest production efficiency increased with temperature--a finding that is consistent in direction with the insignificant trends observed here (Fig.
+S3).**
+
+*(NEED MORE HERE!!!)*
 

### 2020-12-18 manuscript/appendix.Rmd
R2, minor point 2, plus a bit of SI table formatting.

@@
 
-## Table S3. Signficance of relationships of forest C fluxes to MAT alone and in combination with MAP.
+## Table S3. Summary of models relating forest C fluxes to MAT alone and in combination with MAP.
 ```{r eval = TRUE, echo=FALSE, warning=FALSE}
@@
 library(kableExtra)
-S5 <- read.csv("tables_figures/mat_map.csv", stringsAsFactors = FALSE, check.names = FALSE)
+S5 <- read.csv("tables_figures/mat_map.csv", col.names = c("C flux", "MAT","MAT + MAP", "MAT x MAP", "form", "$R^2$"), stringsAsFactors = FALSE, check.names = FALSE)
 kable(S5, format = "latex", booktabs = TRUE,  escape = F)%>%
-  kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 12) %>%
-  column_spec(1, width = "4cm") %>%
-  column_spec(2, width = "6cm") %>%
-  column_spec(3, width = "6cm") %>%
-  column_spec(4, width = "6cm")%>%
-  column_spec(5, width = "4cm")%>%
-  column_spec(6, width = "2cm")
+  kable_styling(latex_options = c("scale_down", "hold_position"), font_size = 10) %>%
+  add_header_above(c(" ", "p-value by model term"=3, "best (lowest AIC) model "=2))
+
 ```
@@
 kable(S3, format = "latex", booktabs = TRUE, escape = F) %>%
+  column_spec(4:5, width = "3cm") %>%
   kable_styling(latex_options = c("scale_down", "hold_position")) %>%

### 2020-12-18 manuscript/response to reviews.Rmd
R2, minor point 2, plus a bit of SI table formatting.
checked/ finished responses to R2, except issue #125

@@
 
-**We have clarified this in the table caption.**
+**We have adjusted table column names to clarify.**
 
@@
 **We have added a legend to clarify that none of the fits are significant (and therefore no statistics presented).
-We have chosen not to add CUE to the figure, as we have focused on analysis of allocation within ratios.
-However, we have added text below Fig.
-S3 to clarify that we interpret these results in terms of variation in CUE.**
+We have chosen not to add CUE to the figure, as analysis of proportions (NPP/GPP) would require a different statistical approach than analysis of ratios.
+However, we have added text to the Fig.
+S3 legend to clarify that we interpret these results in terms of variation in CUE.**
 
@@
 
-**We used MATxMAP in a linear model, which we have clarified this in the text.**
+**We have clarified this in the text: "An interactive model containing a $MAT$ x $MAP$ interaction was accepted when $\Delta$AIC >2 relative to a null including $MAT$ and $MAP$ as fixed effects."**
 
@@
 
-**While ForC contains information on precise stand ages for many temperate forests, this information is typically unavailable for tropical forests, and so these forests are often recorded simply as 'old-growth' or 'undisturbed'.
-This means that they are known to be older than 100 years of age, but the precise age is unknown.
-As a result, standardising this dataset by age is a challenge, and so we treat all forests in the analysis as mature forests, assuming minimal effect of age beyond 100 years.**
+**While ForC contains information on precise stand ages for many temperate forests (typically estimated from tree-rings), this information is typically unavailable for tropical forests, and so these forests are often recorded simply as 'old-growth' or 'undisturbed'.
+This means that they are believed to be older than 100 years of age, but the precise age is unknown.
+As a result, it is unfortunately not possible to standardise this dataset by age beyond 100 years, and so we treat all forests in the analysis as mature forests, assuming minimal effect of age beyond 100 years.**
 

### 2020-12-18 manuscript/response to reviews.Rmd
drafted letter to editor

@@
 ---
-title: "Appeal of rejection of GCB-20-1713"
+title: "Response to reviews"
 output: pdf_document
@@
 We are please to submit a revised version of our manuscript, "Global patterns of forest autotrophic carbon fluxes" (GCB-20-1713), for consideration for publication in *Global Change Biology*.
+We have addressed all points raised by the reviewers, as detailed below.
+The most significant changes include:
 
-.....
+- We modified the presentation to clarify that this is intended as a high-quality quantitative review using a vastly expanded and standardized dataset, noting that it was never our intention to claim to be posing novel hypotheses.
 
-thank you,
+- We added some statements clarifying how our analysis improves upon previous similar analyses.
+Major improvements include sized of database, number of variables considered, and standardization of stand age, history, and management.
+
+- We added discussion relating our results to the eddy-covariance literature and highlighting some modeling implications.
+
+Thank you for considering this revised version, and we look forward to your response.
+
+Sincerely,
 
@@
 
-Dear Kristina,
+# Reviewer(s)' Comments to Author:
 
-Your appeal regarding the decision on your manuscript GCB-20-1713 has been safely received.
-The editor is in principle willing to consider a revised version of your manuscript.
-If you do consider to submit a revised manuscript, we want to stress the importance of fully addressing all the comments raised by both reviewers, including those concerns about the novelty and the lack of mechanistic advancement of your paper.
-Your revised manuscript, along with the point-to-point response to reviewer comments, will be sent out for evaluation.
-We will make a decision based on reviewers' assessment, and no acceptance can be guaranteed.
+## Reviewer: 1
 
-Please let me know if this is agreeable to you.
-
-All the best,
-
-Rhea Bruno
-
-~~~
-
-Dear Dr. Anderson-Teixeira,
-
-We write to you in regards to manuscript "Global patterns of forest autotrophic carbon fluxes".
-Based on points raised by the referees, the Subject Editor has denied publication of your paper in Global Change Biology.
-
-As you may know, we decline a substantial proportion of manuscripts.
-These editorial judgments are based on such considerations as the degree of advance provided, the breadth of potential interest to researchers and timeliness.
-
-Please be assured that this editorial decision does not represent a criticism of the quality of your work, nor are we questioning its value to others working in this area
-
-Thank you for considering Global Change Biology for the publication of your research.
-We hope the outcome of this specific submission will not discourage you from the submission of future manuscripts.
-
-Sincerely,
-Global Change Biology Editorial Office
-
-\newpage
-
-Reviewer(s)' Comments to Author:
-
-Reviewer: 1
-
-Comments to the Author
+### Comments to the Author
 
@@
 
-Reviewer: 2
+## Reviewer: 2
 
-Comments to the Author
+### Comments to the Author
 
@@
 - **We find no significant trend in the allocation of $GPP$ between production and respiration across latitude or climate ($NPP$:$R_{auto}$; Fig.
-S3), counter to the idea that tropical forests have anomalously low $CUE$ [@de_lucia_forest_2007; @malhi_productivity_2012; @anderson-teixeira_carbon_2016], *which is predicted by most models [@collalti_forest_2020]*.
+S3), counter to the idea that tropical forests have anomalously low $CUE$ [@de_lucia_forest_2007; @malhi_productivity_2012; @anderson-teixeira_carbon_2016], *which is predicted by most models* [@collalti_forest_2020].
 In contrast, @collalti_forest_2020 found that forest production efficiency increased with temperature--a finding that is consistent in direction with the insignificant trends observed here (Fig.
@@
 \newpage
-## References
+# References

### 2020-12-18 manuscript/response to reviews.Rmd
minor edits to intro

@@
 - **Abstract: "Here, we draw upon 1,319 records from the Global Forest Carbon Database (ForC), representing all major forest types and the nine most significant autotrophic carbon fluxes, to comprehensively ~~explore~~ *review* how annual C cycling in mature, undisturbed forests varies with latitude and climate on a global scale."**
-- **Introduction: "In order to approach this broad topic, we simplify the major gaps in our knowledge to five broad *review questions* and corresponding predictions (Table 1)."**
+- **Introduction: "In order to approach this broad topic, we organize the major gaps in our knowledge under five broad review questions and corresponding predictions, many derived from the findings of previous studies (Table 1)."**
 
@@
 The power of this synthesis lies in how it standardises for the effects of stand age and disturbance history, unambiguously clarifying these relationships.
-We have added text to the introduction to highlight the importance of standardisation of the dataset: "Productivity may vary with stand age, disturbance, and management [@de_lucia_forest_2007; @yu_high_2014; @fernandez-martinez_spatial_2014; @simova_enigma_2017], making clear latitudinal patterns difficult to discern without standardization of the dataset."**
+We have added text to the introduction to highlight the importance of standardisation of the dataset: "C flux and allocation vary with stand age, disturbance, and management [@de_lucia_forest_2007; @yu_high_2014; @fernandez-martinez_spatial_2014; @simova_enigma_2017], making clear latitudinal patterns difficult to discern without standardization of the dataset."**
 

### 2020-12-18 manuscript/manuscript draft.Rmd
checked/ finished responses to R2, except issue #125
minor edits to intro

@@
 Carbon (C) cycling in Earth's forests provides the energetic basis for sustaining the majority of Earth's terrestrial biodiversity and many human populations [@millennium_ecosystem_assessment_ecosystems_2005], while strongly influencing atmospheric carbon dioxide (CO~2~) and climate [@bonan_forests_2008].
-Forests' autotrophic carbon fluxes -- that is, carbon fixation, allocation, and metabolism by trees and other primary producers -- sets the energy ultimately available to heterotrophic organisms (including microbes), in turn influencing their abundance [@zak_plant_1994; @niedzialkowska_species_2010] and possibly diversity [@waide_relationship_1999; @chu_direct_2018].
+Forests' autotrophic C fluxes -- that is, C fixation, allocation, and metabolism by trees and other primary producers -- sets the energy ultimately available to heterotrophic organisms (including microbes), in turn influencing their abundance [@zak_plant_1994; @niedzialkowska_species_2010] and possibly diversity [@waide_relationship_1999; @chu_direct_2018].
 They are linked to cycling of energy, water, and nutrients and, critically, influence all C stocks and define forest interactions with Earth's changing climate. 
@@
 However, studies have differed in their conclusions regarding the shape of this relationship -- quite possibly because of lack of standardization with respect to methodology and stand history. 
-Productivity may vary with stand age, disturbance, and management [@de_lucia_forest_2007; @yu_high_2014; @fernandez-martinez_spatial_2014; @simova_enigma_2017], making clear latitudinal patterns difficult to discern without standardization of the dataset.
+C flux and allocation vary with stand age, disturbance, and management [@de_lucia_forest_2007; @yu_high_2014; @fernandez-martinez_spatial_2014; @simova_enigma_2017], making clear latitudinal patterns difficult to discern without standardization of the dataset.
 Studies agree that gross primary productivity ($GPP$) increases continuously with decreasing latitude and is indisputably highest in tropical forests [@luyssaert_co_2007; @beer_terrestrial_2010; @jung_global_2011; @badgley_terrestrial_2019; @li_mapping_2019].
-However, this relationship is more ambiguous in subsidiary fluxes.
+However, this relationship is more ambiguous for subsidiary fluxes. 
 Some studies have suggested that net primary productivity ($NPP$), or its aboveground portion ($ANPP$), exhibits a less distinct increase from temperate to tropical forests [@luyssaert_co_2007] -- or even a decrease [@huston_global_2009, but see Gillman et al., 2015].
-A shallower increase in $NPP$ than in $GPP$ with decreasing latitude would align with the suggestion that tropical forests tend to have low carbon use efficiency [$CUE$= $NPP$/$GPP$; @de_lucia_forest_2007; @malhi_productivity_2012; @anderson-teixeira_carbon_2016].
-Such differences among C fluxes in their relationship to latitude could have profound implications for our understanding of the C cycle and its climate sensitivity.
+A shallower increase in $NPP$ than in $GPP$ with decreasing latitude would align with the suggestion that tropical forests tend to have low carbon use efficiency [$CUE$= $NPP$/$GPP$; @de_lucia_forest_2007; @malhi_productivity_2012; @anderson-teixeira_carbon_2016] but contrast with recent findings of the opposite pattern [@collalti_forest_2020].
+Such differences among C fluxes in their relationship to latitude have profound implications for our understanding of the C cycle and its climate sensitivity [e.g., @collalti_forest_2020].
 However, until recently the potential to compare latitudinal trends across C fluxes has been limited by lack of a large database with standardization for methodology, stand history, and management [@anderson-teixeira_forc_2018].
 
-The latitudinal gradient in forest C flux rates, along with altitudinal gradients [@girardin_net_2010; @malhi_variation_2017], is driven primarily by climate, which is a significant driver of C fluxes across broad spatial scales [@luyssaert_co_2007; @cleveland_relationships_2011; @wei_forest_2010; @mullerlandau_patterns_2020].
+Latitudinal gradients in forest C flux rates, along with altitudinal gradients [@girardin_net_2010; @malhi_variation_2017; @mullerlandau_patterns_2020], are driven primarily by climate, which is a significant driver of C fluxes across broad spatial scales [@luyssaert_co_2007; @cleveland_relationships_2011; @wei_forest_2010; @mullerlandau_patterns_2020].
 However, there is little consensus as to the shapes of these relationships or the best predictor variables.
-The majority of studies have focused on exploring the relationships of C fluxes to mean annual temperature ($MAT$) and mean annual precipitation ($MAP$), as the most commonly reported site-level climate variables.
+The majority of studies have focused on exploring the relationships of C fluxes to mean annual temperature ($MAT$) and precipitation ($MAP$), which are the most commonly reported site-level climate variables.
 C fluxes increase strongly with $MAT$ on the global scale, but whether they saturate or potentially decrease at higher temperatures remains disputed. 
 Some studies have detected no deceleration or decline in $GPP$ [@luyssaert_co_2007], $NPP$ [@schuur_productivity_2003], or root respiration [$R_{root}$; @piao_forest_2010; @wei_forest_2010] with increasing $MAT$.
-In contrast, others have found evidence of saturation or decline of C flux in the warmest climates; @luyssaert_co_2007 found $NPP$ saturating at around 10$^\circ$C MAT; @larjavaara_temperature_2012 found that increases in $GPP$ saturate at approximately 25$^\circ$C $MAT$, and @sullivan_long-term_2020 found that, within the tropics, woody stem productivity ($ANPP_{stem}$) decreases at the highest maximum temperatures.
+In contrast, others have found evidence of saturation or decline of C flux in the warmest climates: @luyssaert_co_2007 found $NPP$ saturating at around 10$^\circ$C MAT; @larjavaara_temperature_2012 found that increases in $GPP$ saturate at approximately 25$^\circ$C $MAT$, and @sullivan_long-term_2020 found that, within the tropics, woody stem productivity ($ANPP_{stem}$) decreases at the highest maximum temperatures.
 C fluxes generally saturate at high levels of $MAP$, though the saturation points identified vary widely [e.g., ~1000 - 2,445 mm yr^-1^; @wei_forest_2010; @schuur_productivity_2003]. 
@@
 On the other end of the spectrum, climatic conditions within the growing season may exert a stronger influence on annual C fluxes than the length of the growing season. 
-This aligns with observations that in forests, $NEE$ tends to be more closely tied to the maximum rate of CO~2~ uptake than to the carbon uptake period [@zhou_explaining_2016; @fu_maximum_2019], and with numerous tree-ring analyses finding that annual growth is more closely controlled by peak growing season climate than by spring or fall conditions [e.g., @helcoski_growing_2019].
+This aligns with observations that in forests, $NEE$ tends to be more closely tied to the maximum rate of CO~2~ uptake than to the carbon uptake period [@zhou_explaining_2016; @fu_maximum_2019], and with numerous tree-ring analyses finding that annual growth is more closely controlled by peak growing season climate than by spring or fall conditions [e.g., @helcoski_growing_2019] **REFS**.
 Thus, the extent to which growing season length controls global-scale variation in forest autotrophic C fluxes remains unclear.
@@
 The recent development of the Global Forest Carbon database (ForC), which synthesizes multiple variables and includes records of stand history [@anderson-teixeira_carbon_2016; @anderson-teixeira_forc_2018], opens up the possibility for a standardized analysis of global scale variation in multiple C fluxes and the principle climatic drivers of these patterns.
-In order to approach this broad topic, we simplify the major gaps in our knowledge to five broad review questions and corresponding predictions (Table 1).
+In order to approach this broad topic, we organize the major gaps in our knowledge under five broad review questions and corresponding predictions, many derived from the findings of previous studies (Table 1).
 First, we ask how nine forest autotrophic carbon fluxes in ForC vary with latitude (*Q1*).
@@
   pack_rows("Q4. How does seasonality influence annual C fluxes?", 9, 12,  colnum = 7, hline_before = TRUE) %>%
-  pack_rows("Q5. When standardised by growing season length, how do annual C fluxes still vary with climate?", 13, 16, colnum = 7, hline_before = TRUE) %>%
+  pack_rows("Q5. When standardised by growing season length, how do annual C fluxes vary with climate?", 13, 16, colnum = 7, hline_before = TRUE) %>%
   add_header_above(c(" ",  " ", "Forest autotrophic carbon fluxes" = 9, " ")) %>%

### 2020-12-18 manuscript/response to reviews.Rmd
review/ fix-up methods

@@
 
+**To clarify this to readers, we have added the following statement: "To reduce any biasing effects of stand age, we included only stands of known age $\ge$ 100 years and those described by terms such as "mature", "intact", or "old-growth", *noting that ages of mature tropical forests are typically unknown because most tropical trees cannot be easily dated using tree-rings*."**
+
 6. Fig. S6. Spell out “T seas” and “P seas” in the caption.

### 2020-12-18 manuscript/manuscript draft.Rmd
review/ fix-up methods

@@
 There is evidence that stand age influences patterns of C flux and allocation in forest ecosystems, and can confound relationships between latitude and primary productivity [@de_lucia_forest_2007; @gillman_latitude_2015].
-To reduce any biasing effects of stand age, we included only stands of known age $\ge$ 100 years and those described by terms such as "mature", "intact", or "old-growth".
-Since management can alter observed patterns of C cycling [@simova_enigma_2017], sites were excluded from analysis if they were managed, defined as plots that were planted, managed as plantations, irrigated, fertilised or included the term "managed" in their site description.
+To reduce any biasing effects of stand age, we included only stands of known age $\ge$ 100 years and those described by terms such as "mature", "intact", or "old-growth", noting that ages of mature tropical forests are typically unknown because most tropical trees cannot be easily dated using tree-rings.
+Since management can alter C cycling [@simova_enigma_2017], sites were excluded from analysis if they were managed, defined as plots that were planted, managed as plantations, irrigated, fertilised or included the term "managed" in their site description.
 Sites that had experienced significant disturbance within the past 100 years were also excluded.
@@
 We experimented with a definition of growing season months including a moisture index, defined as $(MAT - PET)/PET > -0.95$ [@kerkhoff_plant_2005; see also @michaletz_convergence_2014]. 
-However, we found that including a moisture index had minimal effect on the estimates of growing season length for the sites included here, and so chose to exclude it. 
+However, we found that including a moisture index had minimal effect on the estimates of growing season length for the sites included here and that the approach performed poorly at defining growing seasons for sites with signficant reliance on snow-melt or groundwater, and so chose to exclude this criterion.
 Monthly data for $PET$, precipitation, and temperature from  CRU v 4.03 [@harris_updated_2014] and solar radiation from WorldClim2 [@fick_worldclim_2017] were used to calculate mean monthly $PET$, precipitation, temperature and solar radiation during the growing season.
@@
 Geographic areas--*i.e.*, spatially clustered sites--were defined within ForC using a hierarchical cluster analysis on the distance matrix of the sites and a cutoff of 25km [@anderson-teixeira_forc_2018].
-We experimented with inclusion of altitude as a fixed effect, as productivity is known to decline with elevation [@mullerlandau_patterns_2020], but excluded it from the final models because it added very little explanatory power -- that is, the difference in AIC ($\Delta AIC$) relative to models excluding altitude was generally small (often $\Delta AIC$<2).
+We experimented with inclusion of altitude as a fixed effect, as productivity is known to decline with elevation in mesic regions [@mullerlandau_patterns_2020], but excluded it from the final models because it added very little explanatory power -- that is, the difference in AIC ($\Delta AIC$) relative to models excluding altitude was generally small (often $\Delta AIC$<2).
 Effects were considered significant when inclusion of the fixed effect of interest resulted in p $\le$ 0.05 under an ANOVA test, and $\Delta AIC$ $\ge$ 2.0 relative to a corresponding null model. 
@@
 To test whether and how C fluxes varied with climate when standardised by growing season length (*Q5*; Table 1), we first standardized all annual C fluxes by dividing by growing season length (as defined above).
-We then derived four variables to describe growing season climate, specifically growing season temperature, precipitation, solar radiation, and PET (Table S1).
+We then derived four variables to describe growing season climate (defined to the nearest month): growing season temperature, precipitation, solar radiation, and PET (Table S1).
 We tested for correlations between these standardised fluxes and growing season climate variables, using only first-order linear models.

### 2020-12-18 manuscript/manuscript draft.Rmd
review/ fix up results

@@
 Latitude explained 64% of variation in GPP (n = 243, p<0.0001), 50% in NPP (n = 161, p<0.0001) and 44% in ANPP (n = 278, p<0.0001).
-The C fluxes that were most poorly predicted by latitude were $BNPP_{fine.root}$ (n = 88, p< .01, $R^2$=0.17) and $ANPP_{stem}$ (n = 264, p<0.0001, $R^2$=0.18).
+The C fluxes that were most poorly predicted by latitude were $BNPP_{fine.root}$ (n = 88, p< .01, $R^2$=0.17), and $ANPP_{stem}$ (n = 264, p<0.0001, $R^2$=0.18).
 The relationship with latitude was best fit by the first-order linear model, with the exception of NPP and $R_{root}$, for which a logarithmic model was a slightly -- but not significantly -- better fit.
@@
 That is, modeled estimates of $GPP$, generated from the sum of $NPP$ and $R_{auto}$; $NPP$, generated from the sum of $ANPP$ and $BNPP$; and $ANPP$, generated from the sum of $ANPP_{foliage}$ and $ANPP_{stem}$, fell almost completely within the confidence intervals of the regressions of field estimates of $GPP$, $NPP$, and $ANPP$, respectively.
+An exception was that the summed prediction for $NPP + R_{auto}$ slightly exceeded the upper 95% CI for $GPP$ at higher latitudes.
 
@@
 For eight of the nine fluxes, this relationship was linear.
-For $BNPP$ the best fit was a lognormal fit, though this was not significant ($\Delta$AIC <2).
+For $BNPP$ the best fit was a lognormal fit, though this was not significantly better than a linear fit ($\Delta$AIC <2).
 As with latitude, $MAT$ tended to explain more variation in the larger fluxes ($GPP$, $NPP$, $ANPP$, $R_{auto}$) and $ANPP_{foliage}$ (all $R^2$> 0.4) than in subsidiary and belowground fluxes ($ANPP_{stem}$, $R_{root}$, $BNPP_{fine.root}$; all $R^2$< 0.25; Table S6).
@@
 However, it explained little variation: with the exception of $R_{auto}$, $MAP$ explained at most 25% of variation in C flux. 
-All fluxes increased with $MAP$ up to at least 2000 mm, above which responses were variable (Figs.
-4, S4-S5).
+All fluxes increased with $MAP$ up to at least 2000 mm, above which responses were variable (Figs. 4, S4-S5).
 
@@
 The interaction was negative for $NPP$ and positive for $ANPP_{stem}$. 
-For $ANPP_{foliage}$, $BNPP$, $BNPP_{fine.root}$, and $R_{root}$, $MAP$ did not have a significant effect when accounting for $MAT$ (Fig.
-3, Table S3).
+For $ANPP_{foliage}$, $BNPP$, $BNPP_{fine.root}$, and $R_{root}$, $MAP$ did not have a significant effect when accounting for $MAT$ (Fig. 3, Table S3).
 
@@
 
-All C flux variables showed a significant relationship with annual $PET$.
-The relationship was logarithmic for $ANPP_{foliage}$, $BNPP_{fine.root}$ and $R_{root}$, and polynomial for all other fluxes (Fig.
-4c, S4-5; Table S2).
+All C flux variables were significantly correlated with annual $PET$. 
+The relationship was logarithmic for $ANPP_{foliage}$, $BNPP_{fine.root}$ and $R_{root}$, and polynomial for all other fluxes (Fig. 4c, S4-5; Table S2). 
 We found strong evidence for a saturation point or peak with $PET$: C fluxes tended to increase at values below 1000mm, before saturating between 1200 and 1700mm.
-There was also evidence that some C fluxes begin to decrease at values above 1800mm PET.
+There was also evidence that some C fluxes begin to decrease at values above 1800 mm PET.
 
 Mean annual $VPD$ was a significant predictor of all C fluxes. 
-$ANPP_{foliage}$, $BNPP_{fine.root}$ and $R_{root}$ showed a logarithmic relationship with $VPD$, but all other fluxes showed a polynomial relationship (Figs.
-4d, S4-5; Table S2).
+$ANPP_{foliage}$, $BNPP_{fine.root}$ and $R_{root}$ showed a logarithmic relationship with $VPD$, but all other fluxes showed a polynomial relationship (Figs. 4d, S4-5; Table S2). 
 C fluxes initially increased with $VPD$, before saturating at around 0.8 kPa, after which point they began to decrease.
@@
 Growing season PET had a positive influence on $GPP$ ($R^2$ = 0.15, p<0.01), $NPP$ ($R^2$ = 0.07, p<0.01), $BNPP$ ($R^2$ = 0.23, p<0.0001), $BNPP_{fine.root}$ ($R^2$ = 0.10, p<0.05), and $ANPP_{stem}$ ($R^2$ = 0.06, p<0.05).
-All other relationships were non-significant.
+All other relationships were non-significant (p > 0.05).
 

### 2020-12-18 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@
 
-Our analysis of a large global database (ForC) clarifies how autotrophic C fluxes in mature forests vary with latitude and climate on a global scale.
+Our analysis of a large global database (ForC) clarifies how autotrophic C fluxes in mature forests vary with latitude and climate on a global scale (Table 1).
 We show that, across all nine variables analyzed, annual C flux decreases continually with latitude (Fig.

### 2020-12-18 manuscript/appendix.Rmd
corrected p <= 0.05

@@
 ```{r echo=FALSE, fig.cap = "Figure S4: Individual plots of forest C fluxes in relation to mean annual climate, part 1.
-Statistical signficance is indicated by solid (p < 0.05) and dashed (p > 0.05) regression lines.
+Statistical signficance is indicated by solid (p $\\le$ 0.05) and dashed (p > 0.05) regression lines.
 Fits shown are of the form with the lowest AIC for each C variable- climate variable combination (Table S2).", out.height='85%', fig.pos='H'}
@@
 ```{r echo=FALSE, fig.cap = "Figure S5: Individual plots of forest C fluxes in relation to mean annual climate, part 2.
-Statistical signficance is indicated by solid (p < 0.05) and dashed (p > 0.05) regression lines.
+Statistical signficance is indicated by solid (p $\\le$ 0.05) and dashed (p > 0.05) regression lines.
 Fits shown are of the form with the lowest AIC for each C variable- climate variable combination (Table S2).", out.height='85%', fig.pos='H'}
@@
 ```{r echo=FALSE, fig.cap = "Figure S6: Individual plots of forest C fluxes in relation to mean climate seasonality, part 1.
-Statistical signficance is indicated by solid (p < 0.05) and dashed (p > 0.05) regression lines.
+Statistical signficance is indicated by solid (p $\\le$ 0.05) and dashed (p > 0.05) regression lines.
 Fits shown are of the form with the lowest AIC for each C variable- climate variable combination (Table S2).", out.height ='80%', fig.pos='H'}
@@
 ```{r echo=FALSE, fig.cap = "Figure S7: Individual plots of forest C fluxes in relation to mean climate seasonality, part 2.
-Statistical signficance is indicated by solid (p < 0.05) and dashed (p > 0.05) regression lines.
+Statistical signficance is indicated by solid (p $\\le$ 0.05) and dashed (p > 0.05) regression lines.
 Fits shown are of the form with the lowest AIC for each C variable- climate variable combination (Table S2).", out.width ='80%', fig.pos='H'}
@@
 ```{r echo=FALSE, fig.cap = "Figure S8: Growing season length-standardized forest C fluxes in relation to mean growing season climate, part 1.
-Statistical signficance is indicated by solid (p < 0.05) and dashed (p > 0.05) regression lines.", out.width ='95%', fig.pos='H'}
+Statistical signficance is indicated by solid (p $\\le$ 0.05) and dashed (p > 0.05) regression lines.", out.width ='95%', fig.pos='H'}
 
@@
 ```{r echo=FALSE, fig.cap = "Figure S9: Growing season length-standardized forest C fluxes in relation to mean growing season climate, part 2.
-Statistical signficance is indicated by solid (p < 0.05) and dashed (p > 0.05) regression lines.", out.width ='95%', fig.pos='H'}
+Statistical signficance is indicated by solid (p $\\le$ 0.05) and dashed (p > 0.05) regression lines.", out.width ='95%', fig.pos='H'}
 

### 2020-12-18 manuscript/response to reviews.Rmd
discussion review/ fix-up

@@
 - **We find no significant trend in the allocation of $GPP$ between production and respiration across latitude or climate ($NPP$:$R_{auto}$; Fig.
-S3), counter to the idea that tropical forests have anomalously low $CUE$ [@de_lucia_forest_2007; @malhi_productivity_2012; @anderson-teixeira_carbon_2016], *which is predicted by most models* [@collalti_forest_2020].
+S3), counter to the idea that tropical forests have anomalously low $CUE$ [@de_lucia_forest_2007; @malhi_productivity_2012; @anderson-teixeira_carbon_2016], *as predicted by most models* [@collalti_forest_2020].
 In contrast, @collalti_forest_2020 found that forest production efficiency increased with temperature--a finding that is consistent in direction with the insignificant trends observed here (Fig.

### 2020-12-18 manuscript/manuscript draft.Rmd
discussion review/ fix-up

@@
 Contrary to prior suggestions that the majority of variation in C cycling is driven primarily by the length of the growing season [@kerkhoff_plant_2005; @enquist_adaptive_2007; @michaletz_convergence_2014], we find modest explanatory power of growing season length and small but sometimes significant influences of growing season climate (Figs.
-4f,S6-S9).
+4f, S6-S9).
 Together, these findings yield a unified understanding of climate's influence on forest C cycling.
@@
 Compared to tropical forests, the temperate forest biome has experienced more widespread anthropogenic disturbance and has a larger fraction of secondary stands [@potapov_mapping_2008; @yu_high_2014; @poulter_global_2018], so analyses comparing across latitudinal gradients without controlling for stand age risk confounding age with biome effects.
-Because carbon allocation varies with stand age [@de_lucia_forest_2007; @anderson-teixeira_altered_2013; @yu_high_2014; @doughty_what_2018], age differences may introduce systematic biases into analyses of C fluxes across latitude or global climatic gradients.
+Because carbon allocation varies with stand age [@de_lucia_forest_2007; @anderson-teixeira_altered_2013;  @doughty_what_2018], age differences may introduce systematic biases into analyses of C fluxes across latitude or global climatic gradients.
 For example, woody productivity tends to be higher in rapidly aggrading secondary stands than in old-growth forests, where proportionally more C is allocated to respiration and non-woody productivity [@de_lucia_forest_2007; @piao_forest_2010; @doughty_what_2018; @kunert_understanding_2019].
 Thus, findings that temperate forest productivity rivals that of tropical forests are likely an artifact of different forest ages across biomes.
+**PERHAPS WE CAN BRING IN MODELING COMMENT HERE**
 
@@
 We find no significant trend in the allocation of $GPP$ between production and respiration across latitude or climate ($NPP$:$R_{auto}$; Fig.
-S3), counter to the idea that tropical forests have anomalously low $CUE$ [@de_lucia_forest_2007; @malhi_productivity_2012; @anderson-teixeira_carbon_2016], which is predicted by most models [@collalti_forest_2020].
-In contrast, @collalti_forest_2020 found that forest production efficiency increased with temperature--a finding that is consistent in direction with the insignificant trends observed here (Fig.
-S3).
-Previously observed differences in $CUE$ between old-growth tropical forests relative to (mostly younger) extratropical forests are likely an artifact of comparing stands of different age, as $CUE$ is known to decline with forest age [@de_lucia_forest_2007; @piao_forest_2010; @collalti_forest_2020].
+S3), counter to the idea that tropical forests have anomalously low $CUE$ [@de_lucia_forest_2007; @malhi_productivity_2012; @anderson-teixeira_carbon_2016], as predicted by most models [@collalti_forest_2020].
+In contrast, @collalti_forest_2020 found that forest production efficiency increased with temperature--a finding that is consistent in direction with insignificant trends observed here (Fig. S3).  
+Previously observed differences in $CUE$ between old-growth tropical forests relative to (mostly younger) extratropical forests are likely an artifact of comparing stands of different age, as $CUE$ declines with forest age [@de_lucia_forest_2007; @piao_forest_2010; @collalti_forest_2020].
 Another previously observed pattern for which we find no support is a tendency for belowground C allocation to decrease with increasing temperature  [@moser_elevation_2011; @gill_belowground_2016]; rather, we observe no trends in allocation between $ANPP$ and $BNPP$ across latitudes.
-Failure to detect significant trends in C allocation with respect to climate in this analysis does not imply that none exist; rather, it suggests that, at this global scale, differences are subtle and/or that more careful methodological standardization and/or more data is required to detect them.
+Failure to detect significant trends in C allocation with respect to climate in this analysis does not imply that none exist; rather, it suggests that, at this global scale, differences are subtle and/or that more careful methodological standardization and/or more data is required to detect them [*sensu* @collalti_forest_2020].
 
@@
 Second, differences among variables in the proportion of variation explained by climate may be attributable to more direct climatic control over $GPP$ than subsidiary fluxes. 
-That is, subsidiary fluxes may be shaped by climate both through its influence on $GPP$ and through its influence on $CUE$ and C allocation.
+That is, subsidiary fluxes may be shaped by climate both through its influence on $GPP$ and through its influence on $CUE$ and C allocation (**REF**).
 
@@
 4a).
-This result contrasts with the hypothesis that fluxes saturate with $MAT$ below approximately 25$^\circ$C $MAT$ [@luyssaert_co_2007; @huston_global_2009].
-It remains possible that fluxes decline above this threshold [@larjavaara_temperature_2012; @sullivan_long-term_2020], as is also consistent with tree-ring records indicating that tropical tree growth declines at high temperatures [e.g., @vlam_temperature_2014].
-However, these higher temperatures also tend to be associated with high $PET$ and $VPD$, both of which are associated with reduced C fluxes [Figs. 4c-d, S4-S5; @slot_high_2018; @zani_increased_2020].
+This result contrasts with previous findings of fluxes saturating with $MAT$ below approximately 25$^\circ$C $MAT$ [@luyssaert_co_2007; @huston_global_2009].
+It remains possible that fluxes decline above this threshold [@larjavaara_temperature_2012; @sullivan_long-term_2020].
+However, these higher temperatures also tend to be associated with high $PET$ and $VPD$, both of which are associated with reduced C fluxes [Figs. 4c-d, S4-S5; @slot_high_2018].
 
@@
 The observed positive interaction between $MAT$ and $MAP$ for $ANPP_{stem}$ on the global scale (Fig.
-3) is consistent with an analysis showing a similar interaction for $ANPP$ in tropical forests, also with a cross-over point at ~20^$\circ$C [@taylor_temperature_2017].
+3) is consistent with an analysis showing a similar interaction for $ANPP$ in tropical forests, also with a cross-over point at ~20$^\circ$C [@taylor_temperature_2017].
 However, we detect no such interaction for $ANPP$ or most other C fluxes, and we find a contrasting negative interaction for $NPP$ (Fig.
@@
 Forest C fluxes decline with temperature seasonality (Table 1, *Q4*; Fig. 4e), as is to be expected given that fluxes are minimal during winters. 
-A temperature-defined growing season length correlated strongly with global-scale variation in annual C flux [Table 1, *Q5*; Fig.
+A temperature-defined growing season length correlated with global-scale variation in annual C flux [Table 1, *Q5*; Fig.
 4f; see also @churkina_spatial_2005], consistent with the idea that the latitudinal gradient in carbon flux is attributable more to shorter growing seasons at high latitudes than to inherently lower rates of photosynthesis or respiration by high-latitude forests [@enquist_adaptive_2007; @fu_maximum_2019].
@@
 S2; @chu_does_2016], accurately partitioning this variation will require data on intra-annual variation in C flux coupled with a higher-precision metric of growing season length than the monthly-resolution metric used here [e.g., based on leaf phenology or C exchange, *sensu* @fu_maximum_2019].
-@fu_maximum_2019 find that global-scale geographic variation in annual $NEE$ is driven more strongly by growing season length than by carbon uptake rates within the growing season, whereas interannual variation in $NEE$ and $GPP$ at any given site appears to be driven predominantly by the maximum rate of C uptake, as opposed to growing season length [@zhou_explaining_2016; @fu_maximum_2019].
+@fu_maximum_2019 find that global-scale geographic variation in annual $NEE$ is driven more strongly by growing season length than by carbon uptake rates within the growing season, whereas interannual variation in $NEE$ and $GPP$ at any given site appears to be driven predominantly by the maximum rate of C uptake, as opposed to growing season length [@zhou_explaining_2016; @fu_maximum_2019; @zani_increased_2020].
 Further analysis of interannual variation in C fluxes in relation to climate will be valuable to disentangling how seasonality shapes broad geographic patterns in forest C flux.
@@
 Moreover, forest ecosystem responses to climatic changes outside the temperature range to which forest communities are adapted and acclimatized will not necessarily parallel responses across geographic gradients in climate.
-Indeed, tree-ring studies from forests around the world indicate that tree growth rates -- along with $ANPP_{stem}$ and possibly other ecosystem C fluxes -- respond negatively to temperature [e.g., @helcoski_growing_2019]**REFS**.
+Indeed, tree-ring studies from forests around the world indicate that tree growth rates -- along with $ANPP_{stem}$ and possibly other ecosystem C fluxes -- respond negatively to temperature [e.g., @helcoski_growing_2019; @vlam_temperature_2014]**REFS**.
 Furthermore, in the tropics, climate change will push temperatures beyond any contemporary climate, and there are some indications that this could reduce forest C flux rates  [@mau_temperate_2018; @sullivan_long-term_2020] if paralleled by $VPD$ increases [@smith_empirical_2020].

### 2020-12-18 manuscript/manuscript draft.Rmd
corrected <= signs

@@
 ```{r echo=FALSE, out.width='100%', fig.cap = "Latitudinal trends in forest autotropic carbon flux.
-Plotted are linear models, all of which were significant $(p<0.05)$ and had AIC values within 2.0 of the best model (for two fluxes, logarithmic fits were marginally better; Table S2).
+Plotted are linear models, all of which were significant (p$\\le$0.05) and had AIC values within 2.0 of the best model (for two fluxes, logarithmic fits were marginally better; Table S2).
 Each panel shows major C fluxes together with component fluxes.
@@
 
-All fluxes increased with $MAT$ (all p<0.05; Figs. 3-4, S4-S5, Table S2). 
+All fluxes increased with $MAT$ (all p $\le$ 0.05; Figs. 3-4, S4-S5, Table S2). 
 For eight of the nine fluxes, this relationship was linear.
@@
 
-$MAP$ was a significant (p<0.05) predictor of all fluxes (Figs. 4a, S4-S5; Table S2). 
+$MAP$ was a significant (p $\le$ 0.05) predictor of all fluxes (Figs. 4a, S4-S5; Table S2). 
 However, it explained little variation: with the exception of $R_{auto}$, $MAP$ explained at most 25% of variation in C flux. 
@@
 For visualization purposes, data points are grouped into bins of 0 - 1000, 1001 - 2000, 2001 - 3000, and >3000mm mean annual precipitation, and lines of best fit models are plotted for mean annual precipitation values of 500, 1500, 2500, and 3500mm.
-Significance is defined as p < 0.05.", fig.pos='H'}
+Significance is defined as p $\\le$ 0.05.", fig.pos='H'}
   knitr::include_graphics("mat_map_interaction.png")
@@
 Lines of best fit are plotted according to the best model selected during analysis.
-All regressions are significant $(p<0.05)$.", fig.pos='H'}
+All regressions are significant (p $\\le$ 0.05).", fig.pos='H'}
   knitr::include_graphics("combined_plots.png")
@@
 
-All fluxes decrease with increasing temperature seasonality, though the shape of this relationship varies (all p<0.05; Figs. 4e, S6-7; Table S2). 
+All fluxes decrease with increasing temperature seasonality, though the shape of this relationship varies (all p $\le$ 0.05; Figs. 4e, S6-7; Table S2). 
 Temperature seasonality was strongly correlated with annual temperature range, which was likewise a similarly strong predictor of C fluxes (Table S2). 
@@
 4e, S6-S7; Table S2).
-Length of growing season was a strong predictor of C fluxes, explaining 53% of variation in GPP, 38% of variation in NPP, and 34% of variation in ANPP (all p<0.05; Table S2), but it was a weaker predictor than $MAT$ for all fluxes analysed (Table S4).
+Length of growing season was a strong predictor of C fluxes, explaining 53% of variation in GPP, 38% of variation in NPP, and 34% of variation in ANPP (all p $\le$ 0.05; Table S2), but it was a weaker predictor than $MAT$ for all fluxes analysed (Table S4).
 
@@
 When annual C fluxes were standardized by growing season length (in integer number of months), correlations with growing season climate were generally weak (Figs. S8-S9).
-$ANPP$ increased with growing season temperature ($R^2$ = 0.09, p<0.001) and precipitation ($R^2$ = 0.04, p<0.05). 
-Similarly, $ANPP_{foliage}$ increased slightly with growing season temperature ($R^2$ = 0.16, p<0.01) and precipitation ($R^2$ = 0.09, p<0.05). 
+$ANPP$ increased with growing season temperature ($R^2$ = 0.09, p<0.001) and precipitation ($R^2$ = 0.04, p $\le$ 0.05). 
+Similarly, $ANPP_{foliage}$ increased slightly with growing season temperature ($R^2$ = 0.16, p<0.01) and precipitation ($R^2$ = 0.09, p $\le$ 0.05). 
 Growing season solar radiation was positively correlated with on $BNPP$ ($R^2$ = 0.17, p<0.001) and $BNPP_{fine.root}$ ($R^2$ = 0.13, p<0.01). 
-Growing season PET had a positive influence on $GPP$ ($R^2$ = 0.15, p<0.01), $NPP$ ($R^2$ = 0.07, p<0.01), $BNPP$ ($R^2$ = 0.23, p<0.0001), $BNPP_{fine.root}$ ($R^2$ = 0.10, p<0.05), and $ANPP_{stem}$ ($R^2$ = 0.06, p<0.05).
+Growing season PET had a positive influence on $GPP$ ($R^2$ = 0.15, p<0.01), $NPP$ ($R^2$ = 0.07, p<0.01), $BNPP$ ($R^2$ = 0.23, p<0.0001), $BNPP_{fine.root}$ ($R^2$ = 0.10, p $\le$ 0.05), and $ANPP_{stem}$ ($R^2$ = 0.06, p $\le$ 0.05).
 All other relationships were non-significant (p > 0.05).

### 2020-12-18 manuscript/manuscript draft.Rmd
added some tree-ring references
@beckybanbury , I'm having trouble with generating ref files from Zotero at the moment, but I'll load pdfs for all the references I add.

@@
 On the other end of the spectrum, climatic conditions within the growing season may exert a stronger influence on annual C fluxes than the length of the growing season. 
-This aligns with observations that in forests, $NEE$ tends to be more closely tied to the maximum rate of CO~2~ uptake than to the carbon uptake period [@zhou_explaining_2016; @fu_maximum_2019], and with numerous tree-ring analyses finding that annual growth is more closely controlled by peak growing season climate than by spring or fall conditions [e.g., @helcoski_growing_2019] **REFS**.
+This aligns with observations that in forests, $NEE$ tends to be more closely tied to the maximum rate of CO~2~ uptake than to the carbon uptake period [@zhou_explaining_2016; @fu_maximum_2019], and with numerous tree-ring analyses finding that annual growth is more closely controlled by peak growing season climate than by spring or fall conditions [e.g.,@martin-benito_convergence_2015; @helcoski_growing_2019].
 Thus, the extent to which growing season length controls global-scale variation in forest autotrophic C fluxes remains unclear.
@@
 Second, differences among variables in the proportion of variation explained by climate may be attributable to more direct climatic control over $GPP$ than subsidiary fluxes. 
-That is, subsidiary fluxes may be shaped by climate both through its influence on $GPP$ and through its influence on $CUE$ and C allocation (**REF**).
+That is, subsidiary fluxes may be shaped by climate both through its influence on $GPP$ and through its influence on $CUE$ and C allocation (**REF**), *as is imbued in most models (**REF**)*.
 
@@
 However, increasing C flux rates are by no means universal [e.g., @rutishauser_testing_2020; @hubau_asynchronous_2020], likely because other factors are at play, including changes to other aspects of climate, atmospheric pollution (CO~2~, SO~2~, NO~x~), and local disturbances.
-Moreover, forest ecosystem responses to climatic changes outside the temperature range to which forest communities are adapted and acclimatized will not necessarily parallel responses across geographic gradients in climate.
-Indeed, tree-ring studies from forests around the world indicate that tree growth rates -- along with $ANPP_{stem}$ and possibly other ecosystem C fluxes -- respond negatively to temperature [e.g., @helcoski_growing_2019; @vlam_temperature_2014]**REFS**.
+Moreover, forest ecosystem responses to climatic changes outside the temperature range to which forest communities are adapted and acclimatized will not necessarily parallel responses across geographic gradients in climate (**REF**).
+Indeed, tree-ring studies from forests around the world indicate that tree growth rates -- along with $ANPP_{stem}$ and possibly other ecosystem C fluxes -- respond negatively to growing season temperature [e.g., @helcoski_growing_2019; @vlam_temperature_2014; @martin-benito_convergence_2015; @klesse_combined_2018].
 Furthermore, in the tropics, climate change will push temperatures beyond any contemporary climate, and there are some indications that this could reduce forest C flux rates  [@mau_temperate_2018; @sullivan_long-term_2020] if paralleled by $VPD$ increases [@smith_empirical_2020].

### 2020-12-18 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@
 Moreover, forest ecosystem responses to climatic changes outside the temperature range to which forest communities are adapted and acclimatized will not necessarily parallel responses across geographic gradients in climate (**REF**).
-Indeed, tree-ring studies from forests around the world indicate that tree growth rates -- along with $ANPP_{stem}$ and possibly other ecosystem C fluxes -- respond negatively to growing season temperature [e.g., @helcoski_growing_2019; @vlam_temperature_2014; @martin-benito_convergence_2015; @klesse_combined_2018].
+Indeed, tree-ring studies from forests around the world indicate that tree growth rates -- along with $ANPP_{stem}$ and possibly other ecosystem C fluxes -- often respond negatively to growing season temperature, particularly in warmer climates [e.g., @helcoski_growing_2019; @vlam_temperature_2014; @martin-benito_convergence_2015; @klesse_combined_2018].
 Furthermore, in the tropics, climate change will push temperatures beyond any contemporary climate, and there are some indications that this could reduce forest C flux rates  [@mau_temperate_2018; @sullivan_long-term_2020] if paralleled by $VPD$ increases [@smith_empirical_2020].

### 2020-12-18 manuscript/manuscript draft.Rmd
Klesse 2020

@@
 However, increasing C flux rates are by no means universal [e.g., @rutishauser_testing_2020; @hubau_asynchronous_2020], likely because other factors are at play, including changes to other aspects of climate, atmospheric pollution (CO~2~, SO~2~, NO~x~), and local disturbances.
-Moreover, forest ecosystem responses to climatic changes outside the temperature range to which forest communities are adapted and acclimatized will not necessarily parallel responses across geographic gradients in climate (**REF**).
+Moreover, forest ecosystem responses to climatic changes outside the temperature range to which forest communities are adapted and acclimatized will not necessarily parallel responses across geographic gradients in climate [e.g., @klesse_continental_2020].
 Indeed, tree-ring studies from forests around the world indicate that tree growth rates -- along with $ANPP_{stem}$ and possibly other ecosystem C fluxes -- often respond negatively to growing season temperature, particularly in warmer climates [e.g., @helcoski_growing_2019; @vlam_temperature_2014; @martin-benito_convergence_2015; @klesse_combined_2018].

### 2020-12-18 manuscript/response to reviews.Rmd
model refs

@@
 **We have added the following comments on models to the discussion:**
-- **We find no significant trend in the allocation of $GPP$ between production and respiration across latitude or climate ($NPP$:$R_{auto}$; Fig.
+- **Discussion: "We find no significant trend in the allocation of $GPP$ between production and respiration across latitude or climate ($NPP$:$R_{auto}$; Fig.
 S3), counter to the idea that tropical forests have anomalously low $CUE$ [@de_lucia_forest_2007; @malhi_productivity_2012; @anderson-teixeira_carbon_2016], *as predicted by most models* [@collalti_forest_2020].
 In contrast, @collalti_forest_2020 found that forest production efficiency increased with temperature--a finding that is consistent in direction with the insignificant trends observed here (Fig.
-S3).**
+S3)."**
+- **Last sentence of discussion: "In the meantime, understanding the fundamental climatic controls on annual C cycling in Earth's forests sets a firmer foundation for understanding global-scale forest C cycling and benchmarking the models [@fer_beyond_2021] used to predict forest responses and feedbacks to accelerating climate change."**
 

### 2020-12-18 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd
model refs

@@
 Second, differences among variables in the proportion of variation explained by climate may be attributable to more direct climatic control over $GPP$ than subsidiary fluxes. 
-That is, subsidiary fluxes may be shaped by climate both through its influence on $GPP$ and through its influence on $CUE$ and C allocation (**REF**), *as is imbued in most models (**REF**)*.
+That is, subsidiary fluxes may be shaped by climate both through its influence on $GPP$ and through its influence on $CUE$ and C allocation (**REF**), as is imbued in the structure of most models [@simova_enigma_2017] **REF**.
 
@@
 Further research is required to understand the extent to which forest responses to climate change will track the observed global gradients, and the time scale on which they will do so.
-In the meantime, understanding the fundamental climatic controls on annual C cycling in Earth's forests sets a firmer foundation for understanding forest C cycle responses to accelerating climate change.
+In the meantime, understanding the fundamental climatic controls on annual C cycling in Earth's forests sets a firmer foundation for understanding global-scale forest C cycling and benchmarking the models [@fer_beyond_2021] used to predict forest responses and feedbacks to accelerating climate change.
 
@@
 We also thank the numerous researchers who have contributed to the building of ForC.
+This manuscript was improved by comments from two anonymous reviewers.
 This study was funded by a Smithsonian Scholarly Studies grant to KJAT and HCML and by Smithsonian's Forest Global Earth Observatory (ForestGEO). 

### 2020-12-18 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd
@beckybanbury , here's the ref for Xia_global_2019: https://agupubs.onlinelibrary.wiley.com/doi/abs/10.1029/2018JG004777

@@
 Thus, findings that temperate forest productivity rivals that of tropical forests are likely an artifact of different forest ages across biomes.
+*As many dynamic global vegetation models lack age effects, such biases may be imbued in models *
 **PERHAPS WE CAN BRING IN MODELING COMMENT HERE**
 
-We show that C fluxes are broadly consistent in their responses to climate drivers on the global scale, with no trends in C allocation among the variable pairs tested (Figs. 2, S3).
+**Several papers note that many DVGMs don't include effects of age/assume stand equilibrium (e.g.
+Collalti 2020, Yu 2014) - perhaps we can note that you would expect discrepancies in model outputs if age is not accounted for?**
+
+We show that C fluxes are broadly consistent in their responses to climate drivers on the global scale, with no signficant trends in C allocation among the variable pairs tested (Figs. 2, S3).
 This parallels the observation that C allocation across multiple C fluxes varies little with respect to climate along a steep tropical elevational gradient [@malhi_variation_2017; but see @moser_elevation_2011],
-and is not surprising given that carbon allocation within forest ecosystems is relatively constrained [@enquist_global_2002; @litton_carbon_2007; @malhi_allocation_2011]. 
+and is not surprising given that carbon allocation within forest ecosystems is relatively constrained [@enquist_global_2002; @litton_carbon_2007; @malhi_allocation_2011; @collalti_forest_2020]. 
 We find no significant trend in the allocation of $GPP$ between production and respiration across latitude or climate ($NPP$:$R_{auto}$; Fig.
@@
 Previously observed differences in $CUE$ between old-growth tropical forests relative to (mostly younger) extratropical forests are likely an artifact of comparing stands of different age, as $CUE$ declines with forest age [@de_lucia_forest_2007; @piao_forest_2010; @collalti_forest_2020].
-Another previously observed pattern for which we find no support is a tendency for belowground C allocation to decrease with increasing temperature  [@moser_elevation_2011; @gill_belowground_2016]; rather, we observe no trends in allocation between $ANPP$ and $BNPP$ across latitudes.
+Another previously observed pattern for which we find no support is a tendency for belowground C allocation to decrease with increasing temperature  [@moser_elevation_2011; @gill_belowground_2016; @xia_global_2019]; rather, we observe no trends in allocation between $ANPP$ and $BNPP$ across latitudes.
 Failure to detect significant trends in C allocation with respect to climate in this analysis does not imply that none exist; rather, it suggests that, at this global scale, differences are subtle and/or that more careful methodological standardization and/or more data is required to detect them [*sensu* @collalti_forest_2020].

### 2020-12-18 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd
@teixeirak initial draft

--- 2020-12-18 manuscript/manuscript draft.Rmd
@@
 Thus, findings that temperate forest productivity rivals that of tropical forests are likely an artifact of different forest ages across biomes.
-**PERHAPS WE CAN BRING IN MODELING COMMENT HERE**
+The significant variation in carbon fluxes with stand age has implications for ecosystem models.
+Ecosystem modelling approaches may neglect age-related effects, or assume stand equilibrium [see e.g.
+@yu_high_2014; @collalti_forest_2020].
+Our results highlight the importance of incorporating stand age into ecosystem models.
 

### 2020-12-18 manuscript/manuscript draft.Rmd
Revert "Update manuscript draft.Rmd"
This reverts commit 4a5a8a0c637e3eb58762bfe419c692a2d37e42e8.
Update manuscript draft.Rmd
@teixeirak added to avoid conflict

--- 2020-12-18 manuscript/manuscript draft.Rmd
@@
 Thus, findings that temperate forest productivity rivals that of tropical forests are likely an artifact of different forest ages across biomes.
+The significant variation in carbon fluxes with stand age has implications for ecosystem models.
+Ecosystem modelling approaches may neglect age-related effects, or assume stand equilibrium [see e.g.
+@yu_high_2014; @collalti_forest_2020].
+Our results highlight the importance of incorporating stand age into ecosystem models.
+
 *As many dynamic global vegetation models lack age effects, such biases may be imbued in models *
+
 **PERHAPS WE CAN BRING IN MODELING COMMENT HERE**

### 2020-12-18 manuscript/manuscript draft.Rmd
added references

--- 2020-12-18 manuscript/manuscript draft.Rmd
@@
 
-Forest C fluxes decrease with latitude [e.g., @luyssaert_co_2007;  @gillman_latitude_2015; @li_mapping_2019]. 
+Forest C fluxes decrease with latitude [e.g., @cramer_comparing_1999; @zhao_improvements_2005; @luyssaert_co_2007; @gillman_latitude_2015; @li_mapping_2019]. 
 However, studies have differed in their conclusions regarding the shape of this relationship -- quite possibly because of lack of standardization with respect to methodology and stand history. 
@@
 
-Latitudinal gradients in forest C flux rates, along with altitudinal gradients [@girardin_net_2010; @malhi_variation_2017; @mullerlandau_patterns_2020], are driven primarily by climate, which is a significant driver of C fluxes across broad spatial scales [@luyssaert_co_2007; @cleveland_relationships_2011; @wei_forest_2010; @mullerlandau_patterns_2020].
+Latitudinal gradients in forest C flux rates, along with altitudinal gradients [@girardin_net_2010; @malhi_variation_2017; @mullerlandau_patterns_2020], are driven primarily by climate, which is a significant driver of C fluxes across broad spatial scales [@cramer_comparing_1999; @luyssaert_co_2007; @cleveland_relationships_2011; @wei_forest_2010; @mullerlandau_patterns_2020].
 However, there is little consensus as to the shapes of these relationships or the best predictor variables.
@@
 Thus, findings that temperate forest productivity rivals that of tropical forests are likely an artifact of different forest ages across biomes.
-*As many dynamic global vegetation models lack age effects, such biases may be imbued in models *
-**PERHAPS WE CAN BRING IN MODELING COMMENT HERE**
-
-**Several papers note that many DVGMs don't include effects of age/assume stand equilibrium (e.g.
-Collalti 2020, Yu 2014) - perhaps we can note that you would expect discrepancies in model outputs if age is not accounted for?**
+The significant variation in C fluxes as a function of stand age has implications for ecosystem models.
+Ecosystem modelling approaches may neglect age-related effects, or assume stand equilibrium [see e.g.
+@yu_high_2014; @collalti_forest_2020].
+Our results highlight the importance of incorporating stand age into ecosystem models; without this, models are likely to be vulnerable to bias in global C flux projections.
 
@@
 Second, differences among variables in the proportion of variation explained by climate may be attributable to more direct climatic control over $GPP$ than subsidiary fluxes. 
-That is, subsidiary fluxes may be shaped by climate both through its influence on $GPP$ and through its influence on $CUE$ and C allocation (**REF**), as is imbued in the structure of most models [@simova_enigma_2017] **REF**.
+That is, subsidiary fluxes may be shaped by climate both through its influence on $GPP$ and through its influence on $CUE$ and C allocation, as is imbued in the structure of most models [@cramer_comparing_1999; @simova_enigma_2017].
 

### 2020-12-18 manuscript/response to reviews.Rmd

@@
 S3)."**
+- **Discussion: "The significant variation in C fluxes as a function of stand age has implications for ecosystem models.
+Ecosystem modelling approaches may neglect age-related effects, or assume stand equilibrium [see e.g.
+@yu_high_2014; @collalti_forest_2020].
+Our results highlight the importance of incorporating stand age into ecosystem models; without this, models are likely to be vulnerable to bias in global C flux projections."**
 - **Last sentence of discussion: "In the meantime, understanding the fundamental climatic controls on annual C cycling in Earth's forests sets a firmer foundation for understanding global-scale forest C cycling and benchmarking the models [@fer_beyond_2021] used to predict forest responses and feedbacks to accelerating climate change."**
-
-*(NEED MORE HERE!!!)*
 

### 2020-12-18 manuscript/manuscript draft.Rmd
added references

@@
 C fluxes generally saturate at high levels of $MAP$, though the saturation points identified vary widely [e.g., ~1000 - 2,445 mm yr^-1^; @wei_forest_2010; @schuur_productivity_2003]. 
-Interactions between $MAT$ and $MAP$ may also influence productivity [@yu_high_2014]; within the tropics, there is a positive interaction between $MAT$ and $MAP$ in shaping $ANPP$, such that temperature has a positive effect on productivity in moist climates, but a negative effect in dry climates [@taylor_temperature_2017].
+Interactions between $MAT$ and $MAP$ may also influence productivity [@beer_terrestrial_2010; @yu_high_2014]; within the tropics, there is a positive interaction between $MAT$ and $MAP$ in shaping $ANPP$, such that temperature has a positive effect on productivity in moist climates, but a negative effect in dry climates [@taylor_temperature_2017].
 There is also evidence that C fluxes also respond to climate variables such as seasonality of temperature and precipitation [@wagner_climate_2016], cloud cover [@taylor_temperature_2017], solar radiation [@beer_terrestrial_2010; @fyllas_solar_2017], and potential evapotranspiration [@kerkhoff_plant_2005]; however, these are not typically assessed in global-scale analyses of annual forest C flux.

### 2020-12-18 manuscript/journal_submissions/2020_12_GCB/submission_notes.Rmd

--- 2020-08-07 manuscript/journal_submissions/2020_8_GCB/submission_notes.Rmd
@@
 
-Forests will strongly influence climate change, and cohesive understanding of their carbon fluxes vary with respect to climate is foundational to predicting how they will change in coming decades.
+Forests will strongly influence climate change, and cohesive understanding of their carbon fluxes vary with respect to climate is foundational to understanding global-scale forest C cycling and benchmarking the models used to predict forest responses and feedbacks to accelerating climate change.
 
@@
 
-Li, & Xiao.
-(2019).
-Mapping Photosynthesis Solely from Solar-Induced Chlorophyll Fluorescence: A Global, Fine-Resolution Dataset of Gross Primary Production Derived from OCO-2.
-Remote Sensing, 11(21), 2563.
+Collalti, A., Ibrom, A., Stockmarr, A., Cescatti, A., Alkama, R., Fernández-Martínez, M., Matteucci, G., Sitch, S., Friedlingstein, P., Ciais, P., Goll, D.
+S., Nabel, J.
+E.
+M.
+S., Pongratz, J., Arneth, A., Haverd, V., & Prentice, I.
+C.
+(2020).
+Forest production efficiency increases with growth temperature.Nature510Communications,11(1), 5322.
 
-Collalti, A., & Prentice, I. C. (2019). Is NPP proportional to GPP? Waring’s hypothesis 20 years on. Tree Physiology, 39(8), 1473–1483.
+Smith, M.
+N., Taylor, T.
+C., Haren, J.
+van, Rosolem, R., Restrepo-Coupe, N., Adams, J., Wu, J., Oliveira, R.
+C.
+de, Silva, R.
+da, Araujo, A.
+C.
+de, Camargo, P.
+B.
+de, Huxman, T.
+E., & Saleska, S.
+R.
+(2020).
+Empirical evidence for resilience of tropical forest photosynthesis in a warmer world.Nature Plants,6(10),6771225–1230
 
 Sullivan et al (2020). Long-term thermal sensitivity of Earth’s tropical forests. Science, 368(6493), 869–874.
+
+# Data availability statement
+
+Data were obtained from the open access ForC database (v3.0, archived on Zenodo with DOI 10.5281/zenodo.3403855).
+Code and data necessary to reproduce all results are available through GitHub (https://github.com/forc-db/Global_Productivity) and archived in Zenodo (DOI: TBD).

### 2020-12-18 manuscript/response to reviews.Rmd
corrected bullet points

@@
 **We have added the following comments on models to the discussion:**
+
 - **Discussion: "We find no significant trend in the allocation of $GPP$ between production and respiration across latitude or climate ($NPP$:$R_{auto}$; Fig.

### 2021-02-19 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd
issue #128 responses

@@
 ForC contains geographic coordinates associated with each measurement record and, when available, $MAT$ and $MAP$ as reported in the primary literature [@anderson-teixeira_forc_2018].
-Based on the geographic co-ordinates for each site, data on twelve climate variables -- including $MAT$, $MAP$, temperature seasonality (*i.e.*, standard deviation across months), precipitation seasonality (*i.e.*, coefficient of variation across months), annual temperature range, solar radiation, cloud cover, annual frost and wet days, potential evapotranspiration ($PET$), aridity ($MAP/PET$), and vapor pressure deficit ($VPD$) -- were extracted from five open-access climate datasets: WorldClim [@hijmans_very_2005], WorldClim2 [@fick_worldclim_2017], the Climate Research Unit time-series dataset (CRU TS v4.03 [@harris_updated_2014], the Global Aridity Index and Potential Evapotranspiration Climate Database [@trabucco_global_2019], and TerraClimate [@abatzoglou_terraclimate_2018] (Table S1).
+Based on the geographic co-ordinates for each site, data on twelve climate variables -- including $MAT$, $MAP$, temperature seasonality (*i.e.*, standard deviation across months), precipitation seasonality (*i.e.*, coefficient of variation of mean monthly precipitation), annual temperature range, solar radiation, cloud cover, annual frost and wet days, potential evapotranspiration ($PET$), aridity ($MAP/PET$), and vapor pressure deficit ($VPD$) -- were extracted from five open-access climate datasets: WorldClim [@hijmans_very_2005], WorldClim2 [@fick_worldclim_2017], the Climate Research Unit time-series dataset (CRU TS v4.03 [@harris_updated_2014], the Global Aridity Index and Potential Evapotranspiration Climate Database [@trabucco_global_2019], and TerraClimate [@abatzoglou_terraclimate_2018] (Table S1).
 Definitions and methods used to calculate each variable are included in Table S1.
@@
 $MAP$ was a significant (p $\le$ 0.05) predictor of all fluxes (Figs. 4a, S4-S5; Table S2). 
-However, it explained little variation: with the exception of $R_{auto}$, $MAP$ explained at most 25% of variation in C flux. 
+However, with the exception of $R_{auto}$, $MAP$ explained at most 25% of variation in C flux. 
 All fluxes increased with $MAP$ up to at least 2000 mm, above which responses were variable (Figs. 4, S4-S5).
@@
 Temperature seasonality was strongly correlated with annual temperature range, which was likewise a similarly strong predictor of C fluxes (Table S2). 
-C fluxes were highest where temperature seasonality = 0, and at an annual temperature range of 15$^\circ$C or lower (*i.e.*, in the tropics).
+C fluxes were highest in regions with low seasonality, indicated by temperature seasonality = 0, and annual temperature range $\leq$ 15$^\circ$C (*i.e.*, in the tropics).
 

### 2021-02-23 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd
issue #128 clarified seasonality discussion slightly

@@
 4f; see also @churkina_spatial_2005], consistent with the idea that the latitudinal gradient in carbon flux is attributable more to shorter growing seasons at high latitudes than to inherently lower rates of photosynthesis or respiration by high-latitude forests [@enquist_adaptive_2007; @fu_maximum_2019].
-While there is evidence that trees in high-latitude forests have adaptations to maximize photosynthesis at low temperatures [@helliker_subtropical_2008; @huang_air_2019], this is not sufficient to yield growing season fluxes comparable to those of tropical forests, as indicated by a number of positive correlations between monthly mean flux during the growing season and growing season temperature, solar radiation, and $PET$ (Table 1, Figs.
+However, we find evidence that, within the growing season, climate still plays an important role in shaping C fluxes, as indicated by a number of positive correlations between monthly mean flux during the growing season and growing season temperature, solar radiation, and $PET$ (Table 1, Figs.
 S8-S9).
-Thus, we reject the hypothesis that growing season length alone accounts for global-scale variation in productivity--*i.e.*, that there is no relationship between C flux per month of the growing season and growing season climatic conditions [Table 1, *Q5*; @kerkhoff_plant_2005; @enquist_adaptive_2007; @michaletz_convergence_2014].
+This suggests that, while trees in high-latitude forests have adaptations to maximize photosynthesis at low temperatures [@helliker_subtropical_2008; @huang_air_2019], this is not sufficient to yield growing season fluxes comparable to those of tropical forests.
+Thus, we reject the hypothesis that there is no relationship between C flux per month of the growing season and growing season climatic conditions [Table 1, *Q5*; @kerkhoff_plant_2005; @enquist_adaptive_2007; @michaletz_convergence_2014]: growing season length alone is insufficient to account for global-scale variation in productivity.
 Rather, annual C flux is shaped by both growing season length and the climate of peak growing season months [@chu_does_2016; @fu_maximum_2019]. 
@@
 @fu_maximum_2019 find that global-scale geographic variation in annual $NEE$ is driven more strongly by growing season length than by carbon uptake rates within the growing season, whereas interannual variation in $NEE$ and $GPP$ at any given site appears to be driven predominantly by the maximum rate of C uptake, as opposed to growing season length [@zhou_explaining_2016; @fu_maximum_2019; @zani_increased_2020].
-Further analysis of interannual variation in C fluxes in relation to climate will be valuable to disentangling how seasonality shapes broad geographic patterns in forest C flux.
+Further analysis of interannual variation in C fluxes in relation to climate will be valuable in disentangling how seasonality shapes broad geographic patterns in forest C flux.
 

### 2021-02-23 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd
issue #128

@@
   # knitr::opts_knit$set(root.dir = '/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/Global_Productivity/manuscript/')
+knitr::opts_knit$set(root.dir = '/Users/gyrcbm/Documents/GitHub/Global_Productivity/manuscript/')
 ```
@@
 Plotted are linear models, all of which were significant (p$\\le$0.05) and had AIC values within 2.0 of the best model (for two fluxes, logarithmic fits were marginally better; Table S2).
+Sample sizes are available in Table 2 and $R^2$ values for linear models are available in Table S2.
 Each panel shows major C fluxes together with component fluxes.
@@
 95\\% confidence intervals are plotted for the major flux for comparison with predicted trends.
-In (d), which shows three belowground fluxes, the major flux, total belowground carbon flux, has insufficient data (n=9) to support a regression", fig.pos='H'}
+In (d), which shows three belowground fluxes, the major flux, total belowground carbon flux, has insufficient data (n=9) to support a regression.", fig.pos='H'}
 
@@
 For visualization purposes, data points are grouped into bins of 0 - 1000, 1001 - 2000, 2001 - 3000, and >3000mm mean annual precipitation, and lines of best fit models are plotted for mean annual precipitation values of 500, 1500, 2500, and 3500mm.
-Significance is defined as p $\\le$ 0.05.", fig.pos='H'}
+Significance is defined as p $\\le$ 0.05.
+Sample sizes are available in Table 2 and $R^2$ values are available in Table S3.", fig.pos='H'}
   knitr::include_graphics("mat_map_interaction.png")
@@
 Lines of best fit are plotted according to the best model selected during analysis.
-All regressions are significant (p $\\le$ 0.05).", fig.pos='H'}
+All regressions are significant (p $\\le$ 0.05).
+Sample sizes are available in Table 2 and $R^2$ values are available in Table S2.", fig.pos='H'}
   knitr::include_graphics("combined_plots.png")

### 2021-02-23 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@
   # knitr::opts_knit$set(root.dir = '/Users/kteixeira/Dropbox (Smithsonian)/GitHub/ForC-db/Global_Productivity/manuscript/')
-knitr::opts_knit$set(root.dir = '/Users/gyrcbm/Documents/GitHub/Global_Productivity/manuscript/')
+# knitr::opts_knit$set(root.dir = '/Users/gyrcbm/Documents/GitHub/Global_Productivity/manuscript/')
 ```

### 2021-02-23 manuscript/manuscript draft.Rmd
edits to discussion paragraph
#128
@beckybanbury , I edited here, as I found the language a bit confusing

@@
 S8-S9).
-This suggests that, while trees in high-latitude forests have adaptations to maximize photosynthesis at low temperatures [@helliker_subtropical_2008; @huang_air_2019], this is not sufficient to yield growing season fluxes comparable to those of tropical forests.
-Thus, we reject the hypothesis that there is no relationship between C flux per month of the growing season and growing season climatic conditions [Table 1, *Q5*; @kerkhoff_plant_2005; @enquist_adaptive_2007; @michaletz_convergence_2014]: growing season length alone is insufficient to account for global-scale variation in productivity.
+This suggests that, while trees in high-latitude forests have adaptations to maximize photosynthesis at low temperatures [@helliker_subtropical_2008; @huang_air_2019], such adaptations are not sufficient to yield growing season fluxes comparable to those of tropical forests.
+Thus, we reject the hypothesis that there is no relationship between C flux per month of the growing season and growing season climatic conditions [Table 1, *Q5*; @kerkhoff_plant_2005; @enquist_adaptive_2007; @michaletz_convergence_2014].
 Rather, annual C flux is shaped by both growing season length and the climate of peak growing season months [@chu_does_2016; @fu_maximum_2019]. 

### 2021-02-23 manuscript/manuscript draft.Rmd
add a few references
@beckybanbury , please confirm FLUXNET2015 is correct.

closes #126
#129

@@
 Associated data, such as stand age, measurement methodologies, and disturbance history, are also included.
-The database was significantly expanded since the publication of @anderson-teixeira_forc_2018 through integration with the Global Soil Respiration Database [@bond-lamberty_global_2010].
+The database was significantly expanded since the publication of @anderson-teixeira_forc_2018 through addition of the Global Soil Respiration Database [@bond-lamberty_global_2010; @jian_restructured_2020] and the FLUXNET2015 dataset [@pastorello_fluxnet2015_2020].
 Additional targeted literature searches were conducted to identify further available data on the fluxes analyzed here, with particular focus on mature forests in temperate and boreal regions, which were not included in the review of @anderson-teixeira_carbon_2016.
@@
 Compared to tropical forests, the temperate forest biome has experienced more widespread anthropogenic disturbance and has a larger fraction of secondary stands [@potapov_mapping_2008; @yu_high_2014; @poulter_global_2018], so analyses comparing across latitudinal gradients without controlling for stand age risk confounding age with biome effects.
-Because carbon allocation varies with stand age [@de_lucia_forest_2007; @anderson-teixeira_altered_2013;  @doughty_what_2018], age differences may introduce systematic biases into analyses of C fluxes across latitude or global climatic gradients.
+Because carbon allocation varies with stand age [@de_lucia_forest_2007; @anderson-teixeira_altered_2013;  @doughty_what_2018; @anderson-teixeira_carbon_nodate], age differences may introduce systematic biases into analyses of C fluxes across latitude or global climatic gradients.
 For example, woody productivity tends to be higher in rapidly aggrading secondary stands than in old-growth forests, where proportionally more C is allocated to respiration and non-woody productivity [@de_lucia_forest_2007; @piao_forest_2010; @doughty_what_2018; @kunert_understanding_2019].

### 2021-02-23 manuscript/manuscript draft.Rmd
final read-through
#129

@@
 
-Forest C fluxes decrease with latitude [e.g., @cramer_comparing_1999; @zhao_improvements_2005; @luyssaert_co_2007; @gillman_latitude_2015; @li_mapping_2019]. 
+Forest C fluxes decrease with latitude [e.g., @cramer_comparing_1999; @zhao_improvements_2005; @luyssaert_co_2007; @gillman_latitude_2015; @li_mapping_2019; @anderson-teixeira_carbon_nodate]. 
 However, studies have differed in their conclusions regarding the shape of this relationship -- quite possibly because of lack of standardization with respect to methodology and stand history. 
@@
 Such differences among C fluxes in their relationship to latitude have profound implications for our understanding of the C cycle and its climate sensitivity [e.g., @collalti_forest_2020].
-However, until recently the potential to compare latitudinal trends across C fluxes has been limited by lack of a large database with standardization for methodology, stand history, and management [@anderson-teixeira_forc_2018].
+However, until recently the potential to compare latitudinal trends across C fluxes has been limited by lack of a large database with standardization for methodology, stand history, and management [@anderson-teixeira_forc_2018; @anderson-teixeira_carbon_nodate].
 
@@
 Some studies have detected no deceleration or decline in $GPP$ [@luyssaert_co_2007], $NPP$ [@schuur_productivity_2003], or root respiration [$R_{root}$; @piao_forest_2010; @wei_forest_2010] with increasing $MAT$.
-In contrast, others have found evidence of saturation or decline of C flux in the warmest climates: @luyssaert_co_2007 found $NPP$ saturating at around 10$^\circ$C MAT; @larjavaara_temperature_2012 found that increases in $GPP$ saturate at approximately 25$^\circ$C $MAT$, and @sullivan_long-term_2020 found that, within the tropics, woody stem productivity ($ANPP_{stem}$) decreases at the highest maximum temperatures.
+In contrast, others have found evidence of saturation or decline of C flux in the warmest climates: @luyssaert_co_2007 found $NPP$ saturating at around 10$^\circ$C MAT, @larjavaara_temperature_2012 found that increases in $GPP$ saturate at approximately 25$^\circ$C $MAT$, and @sullivan_long-term_2020 found that, within the tropics, woody stem productivity ($ANPP_{stem}$) decreases at the highest maximum temperatures.
 C fluxes generally saturate at high levels of $MAP$, though the saturation points identified vary widely [e.g., ~1000 - 2,445 mm yr^-1^; @wei_forest_2010; @schuur_productivity_2003]. 
@@
 kable(hypothesis_table, format = "latex", booktabs = TRUE, caption = "Summary of review questions, corresponding expectations based on previous studies (when applicable), and results.
-Statistically signficant support for/ rejection of hypotheses is indicated by checkmarks/ X's, and '-' indicates no significant relationship.
+Statistically signficant support for/ rejection of hypotheses is indicated by checkmarks/ X's, whereas '-' indicates no significant relationship.
 Parentheses indicate partial overall support or rejection of hypotheses across all fluxes considered.
@@
 S3).
-Of 12 relationships tested (3 ratios among C flux variables regressed against latitude, $MAT$, $MAP$ and temperature seasonality), none were significant (all p>0.05).
+Of twelve relationships tested (three ratios among C flux variables regressed against latitude, $MAT$, $MAP$ and temperature seasonality), none were significant (all p>0.05).
 
@@
 Given strong co-variation between growing season length and $MAT$ [Fig.
-S2; @chu_does_2016], accurately partitioning this variation will require data on intra-annual variation in C flux coupled with a higher-precision metric of growing season length than the monthly-resolution metric used here [e.g., based on leaf phenology or C exchange, *sensu* @fu_maximum_2019].
+S2; @chu_does_2016], accurately partitioning their influence will require data on intra-annual variation in C flux coupled with a higher-precision metric of growing season length than the monthly-resolution metric used here [e.g., based on leaf phenology or C exchange, *sensu* @fu_maximum_2019].
 @fu_maximum_2019 find that global-scale geographic variation in annual $NEE$ is driven more strongly by growing season length than by carbon uptake rates within the growing season, whereas interannual variation in $NEE$ and $GPP$ at any given site appears to be driven predominantly by the maximum rate of C uptake, as opposed to growing season length [@zhou_explaining_2016; @fu_maximum_2019; @zani_increased_2020].

### 2021-02-25 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@
 Here, we draw upon 1,319 records from the Global Forest Carbon Database (ForC), representing all major forest types and the nine most significant autotrophic carbon fluxes, to comprehensively review how annual C cycling in mature, undisturbed forests varies with latitude and climate on a global scale.
-We show that, across all flux variables analyzed, C cycling decreases continuously with absolute latitude -- a finding that confirms multiple previous studies and contradicts the idea that net primary productivity of temperate forests rivals that of tropical forests.
+Across all flux variables analyzed, rates of C cycling decreased continuously with absolute latitude -- a finding that confirms multiple previous studies and contradicts the idea that net primary productivity of temperate forests rivals that of tropical forests.
 C flux variables generally displayed similar trends across latitude and multiple climate variables, with no differences in allocation detected at this global scale.
@@
 Carbon (C) cycling in Earth's forests provides the energetic basis for sustaining the majority of Earth's terrestrial biodiversity and many human populations [@millennium_ecosystem_assessment_ecosystems_2005], while strongly influencing atmospheric carbon dioxide (CO~2~) and climate [@bonan_forests_2008].
-Forests' autotrophic C fluxes -- that is, C fixation, allocation, and metabolism by trees and other primary producers -- sets the energy ultimately available to heterotrophic organisms (including microbes), in turn influencing their abundance [@zak_plant_1994; @niedzialkowska_species_2010] and possibly diversity [@waide_relationship_1999; @chu_direct_2018].
+Forests' autotrophic C fluxes -- that is, C fixation, allocation, and metabolism by trees and other primary producers -- set the energy ultimately available to heterotrophic organisms (including microbes), in turn influencing their abundance [@zak_plant_1994; @niedzialkowska_species_2010] and possibly diversity [@waide_relationship_1999; @chu_direct_2018].
 They are linked to cycling of energy, water, and nutrients and, critically, influence all C stocks and define forest interactions with Earth's changing climate. 
@@
 C fluxes increase strongly with $MAT$ on the global scale, but whether they saturate or potentially decrease at higher temperatures remains disputed. 
-Some studies have detected no deceleration or decline in $GPP$ [@luyssaert_co_2007], $NPP$ [@schuur_productivity_2003], or root respiration [$R_{root}$; @piao_forest_2010; @wei_forest_2010] with increasing $MAT$.
+Some studies have detected no deceleration or decline in $GPP$ [@luyssaert_co_2007], $NPP$ [@schuur_productivity_2003], or root respiration ($R_{root}$) [@piao_forest_2010; @wei_forest_2010] with increasing $MAT$.
 In contrast, others have found evidence of saturation or decline of C flux in the warmest climates: @luyssaert_co_2007 found $NPP$ saturating at around 10$^\circ$C MAT, @larjavaara_temperature_2012 found that increases in $GPP$ saturate at approximately 25$^\circ$C $MAT$, and @sullivan_long-term_2020 found that, within the tropics, woody stem productivity ($ANPP_{stem}$) decreases at the highest maximum temperatures.
@@
 All analyses were conducted in R v.3.5.1 [@r_2020].
-Code and data necessary to reproduce all results are available through GitHub (https://github.com/forc-db/Global_Productivity) and archived in Zenodo (DOI: TBD).
+Code and data necessary to reproduce all results are available through GitHub (https://github.com/forc-db/Global_Productivity) and archived in Zenodo (DOI: 10.5281/zenodo.4563098).
 

### 2021-02-25 manuscript/manuscript draft.Rmd
Update manuscript draft.Rmd

@@
 ForC contains geographic coordinates associated with each measurement record and, when available, $MAT$ and $MAP$ as reported in the primary literature [@anderson-teixeira_forc_2018].
-Based on the geographic co-ordinates for each site, data on twelve climate variables -- including $MAT$, $MAP$, temperature seasonality (*i.e.*, standard deviation across months), precipitation seasonality (*i.e.*, coefficient of variation of mean monthly precipitation), annual temperature range, solar radiation, cloud cover, annual frost and wet days, potential evapotranspiration ($PET$), aridity ($MAP/PET$), and vapor pressure deficit ($VPD$) -- were extracted from five open-access climate datasets: WorldClim [@hijmans_very_2005], WorldClim2 [@fick_worldclim_2017], the Climate Research Unit time-series dataset (CRU TS v4.03 [@harris_updated_2014], the Global Aridity Index and Potential Evapotranspiration Climate Database [@trabucco_global_2019], and TerraClimate [@abatzoglou_terraclimate_2018] (Table S1).
+Based on the geographic co-ordinates for each site, data on twelve climate variables -- including $MAT$, $MAP$, temperature seasonality (*i.e.*, standard deviation across months), precipitation seasonality (*i.e.*, coefficient of variation of mean monthly precipitation), annual temperature range, solar radiation, cloud cover, annual frost and wet days, potential evapotranspiration ($PET$), aridity ($MAP/PET$), and vapor pressure deficit ($VPD$) -- were extracted from five open-access climate datasets: WorldClim [@hijmans_very_2005], WorldClim2 [@fick_worldclim_2017], the Climate Research Unit time-series dataset (CRU TS v4.03) [@harris_updated_2014], the Global Aridity Index and Potential Evapotranspiration Climate Database [@trabucco_global_2019], and TerraClimate [@abatzoglou_terraclimate_2018] (Table S1).
 Definitions and methods used to calculate each variable are included in Table S1.

### 2021-02-26 manuscript/manuscript_word.Rmd
create word version
@beckybanbury, FYI. This still needs work, but I'll need to pause on it for now.

If you could help create word versions of the tables (manually--won't work in Rmd), that would be very helpful.

--- 2021-02-25 manuscript/manuscript draft.Rmd
@@
 output:
-  pdf_document:
+  word_document:
     fig_caption: yes
@@
 options(knitr.table.format = "latex")
+
+tables_on=FALSE
 ```
@@
 \renewcommand{\arraystretch}{2}
-```{r eval = TRUE, echo=FALSE, warning=FALSE}
+```{r eval = tables_on, echo=FALSE, warning=FALSE}
 library(knitr)
@@
 
-```{r eval = TRUE, echo=FALSE, warning=FALSE}
+```{r eval = tables_on, echo=FALSE, warning=FALSE}
 library(knitr)

### 2021-02-26 manuscript/manuscript_word.Rmd
version without figures (or tables)
@beckybanbury , we need to add tables and figure captions at the end of this file. Best to start as a separate word file first and merge with this one once we're sure we're done changing the .Rmd.

#130

@@
 tables_on=FALSE
+figures_on=FALSE
 ```
@@
 
-```{r echo=FALSE, out.width='100%', fig.cap = "Map showing all data used in the analysis, coded by variable. Variables are plotted individually in Fig. S1. ", fig.pos='H'}
+```{r eval = figures_on, echo=FALSE, out.width='100%', fig.cap = "Map showing all data used in the analysis, coded by variable. Variables are plotted individually in Fig. S1. ", fig.pos='H'}
   knitr::include_graphics("distribution_all_variables_cropped.jpg")
@@
 
-```{r echo=FALSE, out.width='100%', fig.cap = "Latitudinal trends in forest autotropic carbon flux.
+```{r eval = figures_on, echo=FALSE, out.width='100%', fig.cap = "Latitudinal trends in forest autotropic carbon flux.
 Plotted are linear models, all of which were significant (p$\\le$0.05) and had AIC values within 2.0 of the best model (for two fluxes, logarithmic fits were marginally better; Table S2).
@@
 
-```{r echo=FALSE, out.width='100%', fig.cap = "Interactive effects of mean annual temperature and precipitation on annual forest carbon fluxes.
+```{r eval = figures_on, echo=FALSE, out.width='100%', fig.cap = "Interactive effects of mean annual temperature and precipitation on annual forest carbon fluxes.
 For visualization purposes, data points are grouped into bins of 0 - 1000, 1001 - 2000, 2001 - 3000, and >3000mm mean annual precipitation, and lines of best fit models are plotted for mean annual precipitation values of 500, 1500, 2500, and 3500mm.
@@
 
-```{r echo=FALSE, out.width='100%', fig.cap = "Plots of carbon fluxes against (a) mean annual temperature; (b) mean annual precipitation; (c) potential evapotranspiration, (d) vapour pressure deficit; (e) temperature seasonality; (f) length of growing season.
+```{r eval = figures_on, echo=FALSE, out.width='100%', fig.cap = "Plots of carbon fluxes against (a) mean annual temperature; (b) mean annual precipitation; (c) potential evapotranspiration, (d) vapour pressure deficit; (e) temperature seasonality; (f) length of growing season.
 For visualization purposes, data for each flux was rescaled with a mean of 0 and standard deviation of 1.

### 2021-02-26 scripts/mat_map_interaction.R
Update mat_map_interaction.R

@@ -4,7 +4,7 @@
 rm(list = ls())
 
 # Set working directory as ForC main folder ####
-setwd("")
+setwd("C:/Users/gyrcbm/Documents/GitHub/Global_Productivity")
 
 # Load libaries ####
 library(lme4)
@@ -205,11 +205,11 @@ for (response.v in response.variables){
     if(pannel.nb %in% c(1:6)) axis(1, labels = F, tck = 0.02)
     if(pannel.nb %in% c(7:9)) axis(1, tck = 0.02)
     
-    resp.v.info <- read.csv("respv_data.csv", stringsAsFactors = F)
+    resp.v.info <- read.csv("C:/Users/gyrcbm/Documents/GitHub/Global_Productivity/raw.data/respv_data.csv", stringsAsFactors = F)
     respv <- resp.v.info$name[which(resp.v.info$response.v %in% response.v)]
     
     
-    mtext(side = 3, line = -1, text = paste0("(", letters[pannel.nb], ") ", respv), adj = 0.1, cex = 0.5)
+    mtext(side = 3, line = -1, text = expression(paste("(", bold(letters[pannel.nb]), ") ", respv)), adj = 0.1, cex = 0.5)
     if(significant.effect.of.interaction) mtext(side = 3, line = -2, text = "Significant interactive effect", adj = 0.1, cex = 0.5)
     if(!significant.effect.of.interaction & significant.effect.of.additive) mtext(side = 3, line = -2, text = "Significant additive effect", adj = 0.1, cex = 0.5)
     if(!significant.effect.of.interaction & !significant.effect.of.additive & significant.effect) mtext(side = 3, line = -2, text = "Significant effect of MAT", adj = 0.1, cex = 0.5)
@@ -224,16 +224,18 @@ for (response.v in response.variables){
 }
 
 dev.off()
-write.csv(all.results, "")
-write.csv(p.table, "")
+# write.csv(all.results, "")
+# write.csv(p.table, "")
 
 ########################################## code to correctly plot
 Rsqs = NULL
 
 response.variables <- c("GPP", "NPP", "ANPP", "ANPP_woody_stem", "ANPP_foliage", "BNPP_root", "BNPP_root_fine", "R_auto", "R_auto_root")
-png(file = "", width = 2255, height = 2000, units = "px", res = 300)
[…]

### 2021-02-26 manuscript/manuscript_word.Rmd
small fixes

@@
 C flux variables generally displayed similar trends across latitude and multiple climate variables, with no differences in allocation detected at this global scale.
-Temperature variables in general, and mean annual temperature or temperature seasonality in particular, were the single best predictors of C flux, explaining 19 - 71% of variation in the C fluxes analyzed.
+Temperature variables in general, and mean annual temperature or temperature seasonality in particular, were the best single predictors of C flux, explaining 19 - 71% of variation in the C fluxes analyzed.
 The effects of temperature were modified by moisture availability, with C flux reduced under hot and dry conditions and sometimes under very high precipitation. 

### 2021-02-26 scripts/final_model_plots.R
scripts update

@@ -3,7 +3,7 @@
 rm(list = ls())
 
 # Set working directory as ForC main folder ####
-setwd("")
+setwd("C:/Users/gyrcbm/Documents/GitHub/Global_Productivity")
 
 # Load libaries ####
 library(lme4)
@@ -89,7 +89,7 @@ for(response.variables in response.variables.groups){
       par(mfrow = c(1,1), mar = c(0,0,0,0), oma = c(5,5,2,0))
       print(fixed.v)
       
-      fixed.v.info <- read.csv("fixedv_data.csv", stringsAsFactors = F)
+      fixed.v.info <- read.csv("C:/Users/gyrcbm/Documents/GitHub/Global_Productivity/raw.data/fixedv_data.csv", stringsAsFactors = F)
       
       xaxis <- fixed.v.info$xaxis[which(fixed.v.info$fixed.v %in% fixed.v)]
       
@@ -252,7 +252,8 @@ for (age in ages){
   if (age %in% "age.greater.than.100") ages.to.keep <- ForC_simplified$stand.age >= 100 & !is.na(ForC_simplified$stand.age)
   if (age %in% "age.greater.than.200") ages.to.keep <- ForC_simplified$stand.age >= 200 & !is.na(ForC_simplified$stand.age)
   
-  png(file = paste0("/combined_plots.png"), width = 2255, height = 3000, units = "px", res = 300)
+  # png(file = paste0("C:/Users/gyrcbm/Documents/GitHub/Global_Productivity/results/figures/final_figures/final_model_fits/combined_plots.png"), width = 2255, height = 3000, units = "px", res = 300)
+  jpeg(file = paste0("C:/Users/gyrcbm/Documents/GitHub/Global_Productivity/results/figures/final_figures/final_model_fits/combined_plots.jpg"), width = 2255, height = 3000, units = "px", res = 300)
   
   par(mfrow = c(3,2), mar = c(4,2,2,2), oma = c(0,3,0,0))
   
@@ -268,7 +269,7 @@ for (age in ages){
     ylim[1] <- ylim[1] - 0.25
     ylim[2] <- ylim[2] + 0.25
     
-    fixed.v.info <- read.csv("/fixedv_data.csv", stringsAsFactors = F)
+    fixed.v.info <- read.csv("C:/Users/gyrcbm/Documents/GitHub/Global_Productivity/raw.data//fixedv_data.csv", stringsAsFactors = F)
     
[…]

### 2021-02-26 scripts/mat_map_interaction.R

@@ -232,6 +232,7 @@ Rsqs = NULL
 
 response.variables <- c("GPP", "NPP", "ANPP", "ANPP_woody_stem", "ANPP_foliage", "BNPP_root", "BNPP_root_fine", "R_auto", "R_auto_root")
 png(file = "C:/Users/gyrcbm/Documents/GitHub/Global_Productivity/results/figures/final_figures/interactions/mat_map_interaction.png", width = 2255, height = 2000, units = "px", res = 300)
+jpeg(file = "C:/Users/gyrcbm/Documents/GitHub/Global_Productivity/results/figures/final_figures/interactions/mat_map_interaction.jpg", width = 2255, height = 2000, units = "px", res = 300)
 
 par(mfrow = c(3,3), mar = c(2,0,0,2), oma = c(5,8,2,0), xpd = T)
 
@@ -296,7 +297,6 @@ for (response.v in response.variables){
   
   resp.v.info <- read.csv("C:/Users/gyrcbm/Documents/GitHub/Global_Productivity/raw.data/respv_data.csv", stringsAsFactors = F)
   respv <- resp.v.info$name[which(resp.v.info$response.v %in% response.v)]
-  
   
   mtext(side = 3, line = -1, text = paste0("(", letters[pannel.nb], ") ", respv), adj = 0.1, cex = 0.5)
   if(response.v %in% c("NPP", "ANPP_woody_stem")) mtext(side = 3, line = -2, text = "Significant interactive effect", adj = 0.1, cex = 0.5)

### 2021-02-26 scripts/Maps_script.R
maps update

@@ -4,21 +4,17 @@
 rm(list = ls())
 
 # Set working directory as ForC main folder ####
-setwd("C:/Users/becky/Dropbox (Smithsonian)/GitHub/ForC")
+setwd("C:/Users/gyrcbm/Documents/GitHub/Global_Productivity")
 
 # Load libaries ####
 library(lme4)
 library(MuMIn)
 library(ggplot2)
 library(viridis)
+library(maps)
 
 # Load data ####
 ForC_simplified <- read.csv("ForC_simplified/ForC_simplified_WorldClim_CRU_refined.csv", stringsAsFactors = F)
-VARIABLES <- read.csv(paste0(dirname(getwd()), "/ForC/data/ForC_variables.csv"), stringsAsFactors = F)
-
-na_codes <- c("NA", "NI", "NRA", "NaN", "NAC") 
-my_is.na <- function(x) { is.na(x) | x %in% na_codes}
-my_na.omit <- function(x) { return(x[!my_is.na(x)])}
 
 
 # Prepare data ####
@@ -46,7 +42,6 @@ ages.not.999.nor.0.nor.na <- !ForC_simplified$stand.age %in% 999 &  !ForC_simpli
 
 ## Keep only age >=100 (or 999)
 age.greater.than.100 <- ForC_simplified$stand.age >= 100 & !is.na(ForC_simplified$stand.age)
-age.greater.than.200 <- ForC_simplified$stand.age >= 200 & !is.na(ForC_simplified$stand.age)
 ages <- c("age.greater.than.100")
 
 ## keep only stands that are NOT too strongly influence by management/ disturbance
@@ -60,39 +55,10 @@ ForC_simplified$min.dbh <- as.numeric(ForC_simplified$min.dbh)
 min.dbh.to.remove <- ForC_simplified$min.dbh >= 10 & !is.na(ForC_simplified$min.dbh)
 ForC_simplified <- ForC_simplified[-min.dbh.to.remove, ]
 
-###keep only altitude below 500 metres
-# 
-# alt.to.keep <- ForC_simplified$masl <= 500 & !is.na(ForC_simplified$masl)
-# ForC_simplified <- ForC_simplified[alt.to.keep, ]
-
-## give leaf type
-broadleaf_codes <- c("2TEB", "2TDB", "2TB")
-needleleaf_codes <- c("2TEN", "2TDN", "2TN")
-mixed <- c("2TD", "2TE", "2TM", "2TREE")
[…]
